### 2017-01-17 Manuscripts/cogsci-2017/outline.tex
fresh start...

\documentclass[10pt,letterpaper]{article}

\usepackage{cogsci}
\usepackage{pslatex}
\usepackage[natbibapa]{apacite}

% set up PGF
\usepackage{pgfplots}
\pgfplotsset{compat=1.13}
\newcommand\inputpgf[2]{{
\let\pgfimageWithoutPath\pgfimage
\renewcommand{\pgfimage}[2][]{\pgfimageWithoutPath[##1]{#1/##2}}
\input{#1/#2}
}}

\title{PACKER: An Exemplar Model of Category Generation}
 %JLA: I like an exemplar model of category generation better, the only issue is there already is an exemplar model. Need to figure out how to get PACKER or something like it as the acronym...
 % 		is that Jern & Kemp 2013 named their own exemplar model for category generation  ("copy-and-tweak"). We need to implement something like it...
 % OTHER IDEAS:
 %  Contrast effects in category generation
 %

%JLA: Is it possible to have the single line author with two names since our affiliations are identical?
\author{
{\large \bf Nolan Conaway (nconaway@wisc.edu)} \\
Department of Psychology, 1202 W. Johnson Street \\
Madison, WI 53706 USA
  \AND 
{\large \bf Joseph L. Austerweil (austerweil@wisc.edu)} \\
Department of Psychology, 1202 W. Johnson Street \\
Madison, WI 53706 USA
}

\begin{document}

\maketitle

\begin{abstract}
FILLER

\textbf{Keywords:} 
Categorization; etc
\end{abstract}

\section{Introduction}

\begin{itemize}
\item One of our most intriguing cognitive capabilities is the ability to generate new ideas and concepts.
\item Most of the formal work on concepts and categorization has focused on category learning
\item Work in the field has ... [standard passage] ... However, the creative use of conceptual knowledge has been has scarcely been the subject of scientific inquiry.
\item \textit{Briefly} review Ward's contribution. Use words like `foundational', `canonical', etc. Mention main psychological principle(s) Ward discovered about how people create new categories (copy and tweak?). Need to cite an formal work in this line. Looks like Ward or Barsalou may have made one (see pg. 93 of Jern \& Kemp, 13)
\item Jern \& Kemp as pioneering contemporary research on formal models of category generation. Main psychological principle 
\item \textbf{Our angle}: Jern \& Kemp capture one aspect of generating new categories: they should be informed by statistics of features across categories related to the new category (potentially also including that the new category should have good category coherence). However, it misses a major aspect of generating new categories -- They should be different from the other related categories. We define a novel exemplar model for category generation, the PACKER model, which creates new categories by balancing two constraints: (1) new categories should be different from the contrast category and (2) new categories should have be internally coherent.
%\item \textbf{Our angle}: Jern \& Kemp explain the copying-structure aspect of generation, but fails to explain how it is otherwise constrained -- by what already exists, and what can possibly exist.
\item \textit{Need to make this clear}: we are not suggesting Jern \& Kemp have failed to capture a core aspect of generation, but that there are other properties that can be explained through one of the field's oldest traditions.

\end{itemize}

\section{Previous work}
\begin{itemize}
\item Review some of Ward's work. Use words like `foundational', `canonical', etc. Mention empirical findings, how they were shown, and that they held across children and adults
\item Ward not enough. Not just copy and tweak. People use distributional information across (multiple) related categories to generate new categories. Summarize Jern \& Kemp's empirical results from Expts 4 and 5.
\item Copy-and-tweak exemplar model 
\item Jern \& Kemp's hierarchical Bayesian model
\end{itemize}

\section{PACKER: An Exemplar Model}
\begin{itemize}
\item Standard passage: ``exemplar models have enjoyed broad success in explaining classification learning... '' (add more domains that exemplar models have been applied to). Here we test a novel exemplar model of generation. 
\item Model performance is constrained by two forces -- categories are generated such that same-category examples should be similar to one another, and opposite category examples should not be similar to one another.
\item Mathematical description, make sure to describe influence of parameters on performance.
\end{itemize}

\subsection{Qualitative (less mathematical description of) predictions}
\begin{itemize}
\item Category \textit{structure} affects performance only insofar as it constrains inter-exemplar similarities. 
\item Category \textit{location} in the space is crucial because it constrains remaining possible locations.
\item \textbf{Maybe}: show positive space, negative space, and combination of spaces to describe model performance.
\end{itemize}

\section{Behavioral Experiment}
\begin{itemize}
\item Testing the model in a simplified domain, following paradigm set by Jern \& Kemp.
\item Goal is to determine if category \textit{structure} is the only determinant of generation. Category structure is held constant, location in space is manipulated.
\item Description of Jern \& Kemp and copy-and-tweak model predictions.
\item Depiction of our exemplar model predictions. \textit{Ideas}: Representative simulations, aggregate heat-map, probability spaces following 0 and 1 items generated. %i like aggregate heat-map of representative simulations that demonstrates how qualitatively different behavior can come out of model from different parameter values (and have some generated exemplars)
\end{itemize}

\subsection{Participants \& Materials}
\subsection{Procedure}
\subsection{Results}

\begin{itemize}
\item Following predictions of our exemplar model, the main interest was the use of space above and below the contrast category. [If-then logic]...
\item Obvious effects of category location: do people tend to place the new category somewhere else? See Figure \ref{fig:middle-bottom-yranges}, Table \ref{table:subset-table}.
\item The more interesting issue: does the remaining space constrain what types of categories people come up with? 
\item Analysis is tricky because vertical range is not normally distributed (see Figure \ref{fig:middle-bottom-yranges}). Report parametric / nonparametric stats.
\item Following core prediction of exemplar model, fewer participants used the top and bottom of the space when assigned the Bottom category, report 2x2x2 contingency table stats.
\item However, we also observed a great deal of individual differences in the form of ... [report types of categories, maybe show samples?]
%JLA: Yeah representative participants from each category and approximate size
\end{itemize}

\section{Simulations}
\section{Discussion}
\section{Acknowledgments}

\clearpage

\begin{figure*}
    \begin{center}
    \inputpgf{figs/}{example-prob-spaces.pgf}
    \caption{Exemplar generation following exposure to one member of category `A' and one member of category `B'. Areas in which generation is not likely are shaded white, high probability areas are shaded blue. \textit{Left}: Model generation gradient when only considering the contrast category (A). \textit{Center}: Model generation gradient when only considering the targeted category (B). \textit{Right}: The combined spaces.  }
    \label{fig:example-prob-spaces}
    \end{center}
\end{figure*}

\begin{figure}
    \begin{center}
    \input{figs/middle-bottom-conditions.pgf}
    \caption{Experiment conditions.}
    \label{fig:middle-bottom-conditions}
    \end{center}
\end{figure}

\begin{figure*}
    \begin{center}
    \input{figs/middle-bottom-yranges.pgf}
    \caption{Behavioral results. Each line shows the minimum and maximum value of a generated category along the Y (vertical) axis. Dots along each line represent the positions of individual exemplars in the category, and each participant's category is shown on a separate line. Participants are sorted by overall Y axis range, and then by condition.}
    \label{fig:middle-bottom-yranges}
    \end{center}
\end{figure*}

\begin{table}[!ht]
\begin{center} 
\caption{Behavioral results.} 
\label{table:subset-table} 
\vskip 0.12in
\begin{tabular}{ l r r}
    \textbf{Middle}         & Used top rows & No top rows \\
    \hline
    Used bottom rows        &  31 & 18  \\
    No bottom rows          &  11 &  1  \\
    \\
    \textbf{Bottom}         & Used top rows & No top rows \\
    \hline
    Used bottom rows        & 22 & 8 \\
    No bottom rows          & 29 & 2 \\
\end{tabular}
\end{center} 
\end{table}

% \bibliographystyle{apacite}
% \setlength{\bibleftmargin}{.025in}
% \setlength{\bibindent}{-\bibleftmargin}
% \bibliography{references}

\end{document}

### 2017-01-17 Manuscripts/model-descriptions/copy-and-tweak/copy-and-tweak.tex
fresh start...

%!TEX output_directory = latex_out/

\documentclass[12pt]{article}
\usepackage[letterpaper, margin=1in, headheight=15pt]{geometry}
\usepackage{setspace, amsmath}

\begin{document}

\section{The Exemplar Copy-And-Tweak Model}

This model assumed that people represent categories as a collection of stored observations. When prompted to generate new examples, they copy and randomly tweak one of the stored examples. Thus, the model's generation is a two-part process:

\begin{enumerate}
  \item Select a source example from memory.
  \item Generate an example that is similar-but-not-too-similar to the source.
\end{enumerate}

More formally, let $x$ be a $j$-exemplar by $k$-feature matrix, corresponding to every stored exemplar. The probability that a source example $z$ is selected is given by:

\begin{equation}
  p(z) = exp(f_j) / \sum_j{ exp(f_j) }
\end{equation}

$f_j$ is a vector of values specifying each item's relative chance of being selected. In theory, $f_j$ may be set arbitrarily, but in practice it is sensible to differentiate between members of the target and contrast categories. For example, $f_j = -\gamma$ for members of contrast categories, and $f_j = \gamma$ for members of the target category. The resulting free parameter $\gamma$ ($>0$) thus controls the degree favoritism for members of the target category, with larger values producing larger differentials.

After a source exemplar $z$ is selected, similarity between candidate generation options $y$ is computed:

\begin{equation}
  s(y,z) = exp( -c \sum_k{|y_k - z_k|w_k})
\end{equation}

Where $c$ and $w_k$ are the standard specificity and attention weights discussed elsewhere. The probability a candidate stimulus will be generated is given by:

\begin{equation}
  p(y|z) \propto  
  \begin{cases}
     exp(\theta s(y,z)), & \text{if } s(y,z) \leq \phi \\
    0, & \text{otherwise}
  \end{cases}
  \label{eq:generation-choice}
\end{equation}

where $\theta$ is a response determinism parameter. $\phi$ is a similarity tolerance parameter ($0 \leq \phi \leq 1$) specifying the amount of dissimilarity tolerated between $y$ and $z$. Larger values allow for more similar examples to be generated. The final choice is made using a normalization of all possible values of $p(y|z)$.

\subsection*{The Practical Implementation}

In practice, however, it is useful to be able to generate model predictions more deterministically (i.e., without respect to any given source exemplar $z$). Generalizing things a little, the overall probability that an item $y$ will be generated is the sum of its generation probabilities given every possible source example $z$, $p(y|z)$, weighted by the probability each source will be selected, $p(z)$:

\begin{equation}
  p(y) = \sum_z{p(z)p(y|z) }
\end{equation}

In this way, source-independent probabilities can be obtained without monte-carlo simulation. First, the probability of each source $p(z)$ is computed (which is also required for simulation). Then, the pairwise similarity $s(y,z)$ between each source and each generation candidate is computed. Then the generation probability $p(y|z)$ is computed as per Equation \ref{eq:generation-choice}.  Then $p(y)$ is computed as above.

\end{document}

### 2017-01-17 Manuscripts/model-descriptions/jern-kemp-conjugate/jern-kemp-conjugate.tex
fresh start...

%!TEX output_directory = latex_out/

\documentclass[12pt]{article}
\usepackage[letterpaper, margin=1in, headheight=15pt]{geometry}
\usepackage{amsmath,fancyvrb}

\begin{document}
\VerbatimFootnotes

\section*{Updated Version of the Jern \& Kemp (2013) Model}

Working backwards, what is needed for simulation of exemplar generation is the probability of generating a stimulus $x$ given exposure to members of the target category $x_b$:
\
\begin{equation}
p(x | x_b) = ?
\end{equation}
\
where $x_b$ may be empty. Jern \& Kemp's model achieves this using a generative process. Category members ($x_a$ or $x_b$; more generally written as $x_y$) are assumed to have been generated using an underlying category distribution (specifically, multivariate normal):
\
\begin{equation}
  x_y \sim Normal(\mu_{y}, \Sigma_{y})
\end{equation}
\
$p(x | x_y)$ is proportional to the candidate's density under the target category's distribution. Thus, obtaining the category distribution parameters $(\mu_{y}, \Sigma_{y})$ is key for generation.

In Jern \& Kemp's original model, category distribution parameters are inferred exclusively from exposure to the category's members. To better simulate our paradigm, category distribution parameters (specifically, the $\Sigma_{y}$ parameter) in our implementation are also affected by a higher level domain distribution (specifically, Inverse-Wishart), from which $\Sigma_{y}$ are assumed to be generated:

\begin{equation}
  \Sigma_{y} \sim Wishart^{-1}(\Sigma_D, df_D)
\end{equation}
\
This document describes how we have obtained the relevant distributions.

\subsection*{Computing $\mu_y$}
Because Jern \& Kemp assume that the location of $x_y$ is independent of contrast categories, the category $\mu_{y}$ can be inferred exclusively based on exposure to the category's members. Assuming $(\mu_y, \Sigma_y)$ are Normal-Inverse-Wishart distributed (unknown mean, unknown variance):
\
\begin{equation}
  \mu_y = \dfrac
    {\kappa\mu_{0} + n_y \bar{x_y}}
    {\kappa + n_y}
    \label{eq:category_mus}
\end{equation}
\ 
where:
\begin{itemize}
    \setlength\itemsep{-0.5em}
    \item $\mu_{0}$ is the prior mean along $p$ dimensions. Here we set it to the middle of the space.
    \item $\kappa$ is a scalar hyper-parameter, roughly weighting the importance of $\mu_{0}$. $\kappa$ must be greater than zero.
    \item $n_y$ is the number of observations in $x_y$
    \item $\bar{x_y}$ is the sample mean along $p$ dimensions
\end{itemize}

In the case of a populated class, $\mu_{y}$ ends up lying somewhere between $\mu_{0}$ and $\bar{x_y}$, depending on $\kappa_{0}$ and $n_y$. In the case of an empty class, $n_y = 0$, thus Equation \ref{eq:category_mus} reduces to $\mu_{y} = \mu_{0}$. Because we set $\mu_0$ to the center of the space, this outcome is the same as if we had integrated over all possible $\mu_y$.

\subsection*{Computing $\Sigma_D$}

Unlike $\mu_y$, $\Sigma_y$ cannot be computed considering only the members of category $y$. Instead, $\Sigma_y$ is influenced both by the distribution of $x_y$ and by members of other categories through $\Sigma_D$.

$\Sigma_D$ is inferred based on the observed (empirical) category covariances $C_y$. We assume these covariances to be Wishart-distributed, and so $\Sigma_D$ can be computed as:

\
\begin{equation}
    \Sigma_D = \Sigma_0 + \sum_{y}{C_y}
\end{equation}
\
$\Sigma_{0}$ is a $p$-by-$p$ prior covariance matrix. We use a $p$-dimensional identity matrix $I_p$ multiplied element-wise against a free parameter, $\gamma$, controlling the amount of variance assumed by the prior:
\
\begin{equation}
    \Sigma_0 =  I_p\gamma
\end{equation}

\subsubsection*{Note}
In actuality, we ended up doing a bunch of matrix inverting that I do not understand:
\begin{verbatim}
Domain_Sigma = np.linalg.inv(Prior_Sigma)
for y in ncategories:
  Domain_Sigma += np.linalg.inv(Category_Sigmas[:,:,y])
Domain_Sigma = np.linalg.inv(Domain_Sigma)
\end{verbatim}
This is definitely not in the math, and I think it has something to do with Wishart vs. Inverse-Wishart. I think it also had something to do with the issue that its weird to just take the sum of the covariance matrices: the domain ends up assuming much larger covariances compared to any of the categories. We've never nailed it down. \textbf{Joe}, do you know what this is about?

\subsection*{Computing $\Sigma_y$}

In previous versions, we sampled $\Sigma_y \sim Wishart^{-1}(\Sigma_D, df_D)$, with which we sampled $x \sim Normal(\mu_y, \Sigma_y)$. That was the approach taken by Jern \& Kemp. Sampling $x$, however, is not the primary goal of the model -- the main purpose of the model is to predict $p(x | x_b)$, and from those data we can sample $x$.

We could monte-carlo sample $x$ as described above to obtain $p(x | x_b)$. But this is computationally intensive, as each trial would need a separate sampler. Instead, it would be better to develop a method to infer $\Sigma_y$ without sampling.

Assuming $(\mu_y, \Sigma_y)$ are Normal-Inverse-Wishart distributed, $\Sigma_y$ can be computed as:

\begin{equation}
  \Sigma_y = [\Sigma_D\nu + C_y +
    \dfrac
    {\kappa n_y}
    {\kappa + n_y}
    (\bar{x_y}-\mu_y)(\bar{x_y}-\mu_y)^T
  ] (\nu + n_y)^{-1}
  \label{eq:Sigma_y}
\end{equation}
\
$\kappa$, $\bar{x_y}$, $C_y$, $n_y$, $\mu_0$,  are the same values as described above. $\nu$ is an additional free parameter, weighting the importance of $\Sigma_{D}$. $\nu$ must be greater than $p-1$. When $x_b$ is empty, Equation \ref{eq:Sigma_y} reduces to $\Sigma_y = \Sigma_D$.

\subsection*{Computing response probabilities $p(x | x_b)$}

$x$ are assumed to be drawn from a the distribution given by $Normal(\mu_{y}, \Sigma_{y})$. Thus,

\begin{equation}
  p(x) \propto Normal(x | \mu_{b}, \Sigma_{b})
\end{equation}

In practice, $p(x)$ is computed by first obtaining the relative density of every possible generation candidate $x_i$ under the category distribution. The end probability is a normalization of these values:

\begin{equation}
  p(x) = \dfrac
    {exp( \theta Normal(x | \mu_{b}, \Sigma_{b}))}
    {\sum_i exp( \theta Normal(x_i |\mu_{b}, \Sigma_{b}))} 
\end{equation}
\
where $\theta$ is a response determinism parameter.

\subsection*{Description of free parameters}

\begin{itemize}
    \setlength\itemsep{-0.5em}
    \item $\kappa$. Scalar, $\kappa>0$. Weights the importance of $\mu_{0}$ in inferring category $\mu_{y}$.
    \item $\gamma$. Scalar, $\gamma>0$. Weights the importance of $\Sigma_{0}$ in inferring the domain $\Sigma_{D}$.
    \item $\nu$. Scalar, $\nu > p-1$. Weights the importance of $\Sigma_{D}$ in inferring the domain $\Sigma_{y}$.
    \item $\theta$. Scalar, $\theta > 0$. Response determinism parameter.
\end{itemize}

\end{document}

### 2017-01-17 Manuscripts/model-descriptions/packer/packer.tex
fresh start...

%!TEX output_directory = latex_out/

\documentclass[12pt]{article}
\usepackage[letterpaper, margin=1in, headheight=15pt]{geometry}
\usepackage{amsmath, listings}

\begin{document}

\section{The PACKER Model: Producing Alike or Contrasting Knowledge with Exemplar Representations}

This model assumes that people represent categories as exemplars in a multidimensional space, and that generation is constrained by both similarity to members of the category being generated \textit{and} dissimilarity to members of other categories. The model assumes people generate categories that are dissimilar to known categories and have strong within-class similarity.

First, let's assume exemplars $x$ are represented as points within an $k$-dimensional space. When prompted to generate the first member of a new class, participants seek out stimuli that have low similarity to all assigned examples.

Distance between exemplars $X_i$ and $X_j$ is computed using the city-block metric:

\begin{equation}
  d(x_i,x_j) = \sum_{k}{|x_{ik} - x_{jk}|}w_k
\end{equation}

Where $w_k$ is a vector of feature weights, corresponding to the importance of each feature in the similarity computation. The attention weights are of interest in explaining individual differences in generation strategy, and they may have some relation to the relative dimensional variance of existing categories, but for the general case they can be uniform.

The similarity between exemplars $x_i$ and $x_j$ is then computed using Shepard's law:

\begin{equation}
  s(x_i,x_j) = exp( {-c d(x_i,x_j)} )
\end{equation}

Where $c$ is a free parameter setting the specificity of examples. Larger values of $c$ produce more \textit{specific} exemplars -- only stimuli very close to the exemplar will be viewed as similar. Due to the nature of the exponential function, similarity values will lie in range $0-1$, with a value of $1$ indicating an exact match between $x_i$ and $x_j$.

When prompted to make a generation decision, participants are thought to consider both similarity to examples from other categories as well as similarity to examples in the target category. More formally, the summed similarity $ss$ between item $x_i$ and the model's stored exemplars $x_j$ can be computed as:

\begin{equation}
  ss(x_i, x_j) = \sum_j{f_j s(x_i, x_j)}
\end{equation}

Where $f_j$ is a function specifying each stored example's degree of contribution toward generation. Although $f_j$ may be set arbitrarily, for our purposes it is relevant to use the class assignment. For known members of contrast categories, $f_j = -\phi$. For known members of the target category, $f_j = \gamma$. $\phi$ and $\gamma$ are free parameters ($\geq 0$) that control the contribution of same- and opposite- category members to generation. Larger values for either parameter produce greater consideration of either constraint. Values of 0 for either parameter result in no effect for that type of similarity.

This arrangement has a number of desirable properties. When $\phi = \gamma$, the similarity to contrast categories is effectively subtracted from the similarity to the target category -- negative values for $ss(x, x_j)$ thus indicate more contrast category similarity, and positive values indicate more target category similarity. Because $\phi$ and $\gamma$ are parameterized, this trade off can be managed explicitly.

The probability that a given item $x$ will be generated given the model's memory $x_j$ is computed using relative summed similarity values across all generation candidates $x_i$:

\begin{equation}
p(x|x_j) = \dfrac
    { exp( { \theta ss(x, x_j) } ) }
    { \sum_i{ exp( { \theta ss(x_i, x_j) } ) } }
\end{equation}

Where $\theta$ ($\geq 0$) is a free parameter controlling overall response determinism. Larger values of $\theta$ produce greater adherence to the relative probabilities of the choices (items with larger probabilities become more likely), and the special case of $\theta = 0$ produces random responding. Participants are not allowed to generate the same example twice, thus $p_x$ for these items is manually set to 0.

\end{document}

### 2017-01-17 Experiments/middle-bottom/analysis/plot.conditions.py

@@ -1,0 +1,53 @@
+import sqlite3, os, sys
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+
+sys.path.insert(0, "../../../Modules/") # generate-categories/Modules
+import utils
+
+
+
+# 72 73 74 75 76 77 78 79 80
+# 63 64 65 66 67 68 69 70 71
+# 54 55 56 57 58 59 60 61 62
+# 45 46 47 48 49 50 51 52 53
+# 36 37 38 39 40 41 42 43 44
+# 27 28 29 30 31 32 33 34 35
+# 18 19 20 21 22 23 24 25 26
+#  9 10 11 12 13 14 15 16 17
+#  0  1  2  3  4  5  6  7  8
+values = np.linspace(-1, 1, 9)
+alphas = dict(
+	Middle = [30, 32, 48, 50 ],
+	Bottom = [12, 14, 30, 32],
+)
+
+stimuli = np.fliplr(utils.cartesian([values, values]))
+textsettings = dict(
+		verticalalignment='top', 
+		horizontalalignment='left',
+		fontsize = 11.0)
+
+
+
+f, ax= plt.subplots(1,2, figsize=(3, 1.5))
+counter = 0
+for k, a  in alphas.items():
+	h = ax[counter]
+	utils.plotclasses(h, stimuli, a, [])		
+	h.text(-1.1,1.1, k, **textsettings)
+	h.axis([-1.2, 1.2, -1.2, 1.2])
+	[i.set_linewidth(0.75) for i in h.spines.itervalues()]
+
+	counter += 1
+# plt.show()
+f.savefig('conditions.png', bbox_inches='tight', transparent=False)
+
+
+import os, matplotlib
+os.environ["PATH"] += os.pathsep + '/Library/TeX/texbin/'
+opts = {'pgf.texsystem': 'pdflatex'}
+matplotlib.rcParams.update(opts)
+f.savefig('../../../Manuscripts/cogsci-2017/figs/middle-bottom-conditions.pgf',
+	bbox_inches='tight', transparent=False)

### 2017-01-17 Experiments/middle-bottom/analysis/plot.heatmaps.py

@@ -1,0 +1,45 @@
+import sqlite3, sys
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+
+np.set_printoptions(precision = 2)
+sys.path.insert(0, "../../../Modules/")
+import utils
+
+
+
+con = sqlite3.connect('../data/experiment.db')
+infodf = pd.read_sql_query("SELECT * from participants", con)
+stimuli = pd.read_sql_query("SELECT * from stimuli", con).as_matrix()
+alphas = pd.read_sql_query("SELECT * from alphas", con)
+generation = pd.read_sql_query("SELECT * from generation", con)
+con.close()
+
+f, ax = plt.subplots(1,2, figsize = (4,2))
+counter = 0
+for i, rows in infodf.groupby('condition'):
+
+	pids = rows.participant
+	betas = generation.loc[generation.participant.isin(pids), 'stimulus']
+	
+	counts = np.array([sum(betas==j) for j in range(stimuli.shape[0])])
+	ps = counts / float(sum(counts))
+	g = utils.gradientroll(ps,'roll')[:,:,0]
+
+	print(g)
+	print(i, np.max(g))
+
+	h = ax[counter]
+	utils.plotgradient(h, g, stimuli[alphas[i],:], [], clim = (0.0, 0.15))
+	counter += 1
+
+
+import os, matplotlib
+os.environ["PATH"] += os.pathsep + '/Library/TeX/texbin/'
+opts = {'pgf.texsystem': 'pdflatex'}
+matplotlib.rcParams.update(opts)
+f.savefig('../../../Manuscripts/cogsci-2017/figs/behavioral-heatmaps.pgf',	bbox_inches='tight')
+
+f.savefig('heatmaps.png', bbox_inches='tight', transparent=False)
+

### 2017-01-17 Experiments/middle-bottom/analysis/top.bottom.subset.py

@@ -1,0 +1,29 @@
+import sqlite3
+import pandas as pd
+
+# import data
+con = sqlite3.connect('../data/experiment.db')
+info = pd.read_sql_query("SELECT * from participants", con)
+stats = pd.read_sql_query("SELECT * from betastats", con)
+con.close()
+
+stats = pd.merge(stats, info[['participant', 'condition']], on = 'participant')
+
+print 'Raw counts for table:'
+for condition, rows in stats.groupby('condition'):
+	print condition
+	g = rows.groupby(['top_used', 'bottom_used']).size().reset_index()
+	print pd.pivot_table(g, index = 'top_used', columns = 'bottom_used')
+
+
+from scipy.stats import fisher_exact
+
+for i in ['top_used', 'bottom_used', 'top_and_bottom', 'bottom_only', 'top_only']:
+	g = stats.groupby(['condition', i]).size().reset_index()
+	c = pd.pivot_table(g, index = 'condition', columns = i, fill_value=0)
+	pval = fisher_exact(c)[1]
+	pval = round(pval,4)
+	print 'Fishers exact test comparing groups on: ' + i + '\t\tp = ' + str(pval)
+	# print c
+
+

### 2017-01-17 Experiments/middle-bottom/analysis/util/plot.py

@@ -1,0 +1,63 @@
+# some plotting helpers
+from util.funcs import *
+
+
+def plotclasses(h, stimuli, alphas, betas):
+	textsettings = dict(
+		verticalalignment='center', 
+		horizontalalignment='center',
+		fontsize = 11.0)
+
+	h.axis(np.array([-1, 1, -1, 1])*1.2)
+	for i in alphas:
+		x, y = stimuli[i,0], stimuli[i,1]
+		h.text(x,y,'A',color = 'r', **textsettings)
+
+	for i in betas:
+		x, y = stimuli[i,0], stimuli[i,1]
+		h.text(x,y,'B',color = 'b', **textsettings)
+
+	h.get_xaxis().set_visible(False)
+	h.get_yaxis().set_visible(False)
+	h.set_aspect('equal', adjustable='box')
+
+
+def plotgradient(h, G, alphas, stimuli, clim = (), cmap = 'Blues'):
+	"""
+	Plot a gradient using matplotlib.
+	 - h is the handle to the axis
+	 - G is the matrix being plotted, 
+	 - alphas are the known members of the alpha category
+	 - clim (optional) defines the limits of the colormap.
+	"""
+
+	# generate clims if not provided
+	if not clim:
+		clim = (np.min(G), np.max(G))
+
+	# make sure G is 2D
+	if G.ndim > 2:
+		raise Exception("G has too many dimensions. Size: " + str(G.shape))
+
+	# plot gradient
+	h.imshow(np.flipud(G), clim = clim, 
+		origin='lower', interpolation="nearest", cmap = cmap)
+
+	# rescale stimulus coordinates
+	scstimuli = gradientspace(stimuli, G.shape[0])
+
+	# show annotations
+	textsettings = dict(
+		verticalalignment='center', 
+		horizontalalignment='center',
+		fontsize = 11.0)
+
+	coords = scstimuli[alphas,:]
+	for j in range(coords.shape[0]):
+		h.text(coords[j,0],coords[j,1], 'A',
+			color = 'red', **textsettings)
+
+	h.get_xaxis().set_visible(False)
+	h.get_yaxis().set_visible(False)
+	h.set_aspect('equal', adjustable='box')
+	h.axis([-0.5, G.shape[1]-0.5, -0.5, G.shape[0]-0.5])

### 2017-01-17 Modules/models/__init__.py

@@ -1,0 +1,4 @@
+import Optimize
+from CopyTweak import CopyTweak
+from Packer import Packer
+from ConjugateJK13 import ConjugateJK13

### 2017-01-17 Modules/utils/__init__.py

@@ -1,0 +1,2 @@
+from utils import *
+from plot import *

### 2017-01-18 Manuscripts/model-descriptions/jern-kemp-conjugate/jern-kemp-conjugate.tex
update

@@
 \
-$p(x | x_y)$ is proportional to the candidate's density under the target category's distribution. Thus, obtaining the category distribution parameters $(\mu_{y}, \Sigma_{y})$ is key for generation.
-
-In Jern \& Kemp's original model, category distribution parameters are inferred exclusively from exposure to the category's members.
-To better simulate our paradigm, category distribution parameters (specifically, the $\Sigma_{y}$ parameter) in our implementation are also affected by a higher level domain distribution (specifically, Inverse-Wishart), from which $\Sigma_{y}$ are assumed to be generated:
-
-\begin{equation}
-  \Sigma_{y} \sim Wishart^{-1}(\Sigma_D, df_D)
-\end{equation}
-\
-This document describes how we have obtained the relevant distributions.
+$p(x | x_y)$ is proportional to the candidate's density under the target category's distribution.
+Thus, obtaining the category distribution parameters $(\mu_{y}, \Sigma_{y})$ is key for generation.
+This document describes how we compute these variables in a conjugate model.
 
 \subsection*{Computing $\mu_y$}
-Because Jern \& Kemp assume that the location of $x_y$ is independent of contrast categories, the category $\mu_{y}$ can be inferred exclusively based on exposure to the category's members.
 Assuming $(\mu_y, \Sigma_y)$ are Normal-Inverse-Wishart distributed (unknown mean, unknown variance):
@@
 
-\subsubsection*{Note}
-In actuality, we ended up doing a bunch of matrix inverting that I do not understand:
-\begin{verbatim}
-Domain_Sigma = np.linalg.inv(Prior_Sigma)
-for y in ncategories:
-  Domain_Sigma += np.linalg.inv(Category_Sigmas[:,:,y])
-Domain_Sigma = np.linalg.inv(Domain_Sigma)
-\end{verbatim}
-This is definitely not in the math, and I think it has something to do with Wishart vs.
-Inverse-Wishart.
-I think it also had something to do with the issue that its weird to just take the sum of the covariance matrices: the domain ends up assuming much larger covariances compared to any of the categories.
-We've never nailed it down.
-\textbf{Joe}, do you know what this is about?
+Thus, categories are assumed to have some degree of variance along each feature (specified by $\gamma$), but not are assumed to possess feature-feature correlations.
 
 \subsection*{Computing $\Sigma_y$}
-
-In previous versions, we sampled $\Sigma_y \sim Wishart^{-1}(\Sigma_D, df_D)$, with which we sampled $x \sim Normal(\mu_y, \Sigma_y)$.
-That was the approach taken by Jern \& Kemp.
-Sampling $x$, however, is not the primary goal of the model -- the main purpose of the model is to predict $p(x | x_b)$, and from those data we can sample $x$.
-
-We could monte-carlo sample $x$ as described above to obtain $p(x | x_b)$.
-But this is computationally intensive, as each trial would need a separate sampler.
-Instead, it would be better to develop a method to infer $\Sigma_y$ without sampling.
 

### 2017-01-18 Experiments/middle-bottom/modeling/get-best-params.py

@@ -1,0 +1,36 @@
+import sys, pickle
+import pandas as pd
+import numpy as np
+
+import scipy.optimize as op
+
+# add modeling module
+sys.path.insert(0, "../../../Modules/") # generate-categories/Modules
+from models import CopyTweak, Packer, ConjugateJK13, Optimize
+import utils
+
+# get data from pickle
+with open( "data.pickle", "rb" ) as f:
+	trials, stimuli = pickle.load( f )
+
+
+# options for the optimization routine
+options = dict(
+	method = 'Nelder-Mead',
+	options = dict(maxiter = 500, disp = False),
+	tol = 0.01,
+) 
+
+for o in [CopyTweak, Packer, ConjugateJK13]:
+	inits = np.random.uniform(low = 1.0001, high = 2.0, size = (4,)).tolist()
+
+	# edit for model specific ranges
+	if o == CopyTweak:
+		inits[1] = inits[1] - 1.5 # within pref could be positive or negative
+		inits[2] = inits[2] - 1.0 # tolerance is [0 1]
+
+	if o == ConjugateJK13:
+		inits[0] = inits[0] - 1.0 # categiry mean bias should be smallish
+		inits[2] = inits[2] - 1.0 # domain variance assumption should be smallish
+
+	Optimize.hillclimber(o, inits, trials, stimuli, options)

### 2017-01-18 Modules/models/Optimize.py

@@ -12,8 +12,8 @@ def _callback_fun_(xk):
 	"""
 	Function executed at each step of the hill-climber
 	"""
-	print '\t[' + ', '.join([str(round(i,4)) for i in xk]) + ']'
-	# print '\b.',
+	# print '\t[' + ', '.join([str(round(i,4)) for i in xk]) + ']'
+	print '\b.',
 	sys.stdout.flush()
 
 
@@ -39,8 +39,12 @@ def costfun(params, model_obj, dataset, stimuli):
 		predictions = obj.get_generation_ps(stimuli, 1)
 		ps[i] = predictions[trial['response']]
 
+	# check for NaN
 	if np.any(np.isnan(ps)):
 		raise Exception('You got nan probabilities. Sorry :-(')
+
+	# remove zeros to prevent infs
+	ps[ps<1e-308] = 1e-308
 
 	lps = np.log(ps)
 	loglike = np.sum(lps)

### 2017-01-20 Manuscripts/cogsci-2017/paper.tex
started the cogsci proceedings

\documentclass[10pt,letterpaper]{article}

\usepackage{cogsci}
\usepackage{pslatex}
\usepackage{amsmath}
\usepackage[natbibapa]{apacite}

% set up PGF
\usepackage{pgfplots}
\pgfplotsset{compat=1.13}
\newcommand\inputpgf[2]{{
\let\pgfimageWithoutPath\pgfimage
\renewcommand{\pgfimage}[2][]{\pgfimageWithoutPath[##1]{#1/##2}}
\input{#1/#2}
}}

\title{PACKER: An Exemplar Model of Category Generation}
 %JLA: I like an exemplar model of category generation better, the only issue is there already is an exemplar model. Need to figure out how to get PACKER or something like it as the acronym...
 % 		is that Jern & Kemp 2013 named their own exemplar model for category generation  ("copy-and-tweak"). We need to implement something like it...

\author{
{ \large \bf Nolan Conaway (nconaway@wisc.edu) } \\
{ \large \bf Joseph L. Austerweil (austerweil@wisc.edu) } \\
Department of Psychology, 1202 W. Johnson Street \\
Madison, WI 53706 USA
}

\begin{document}

\maketitle

\begin{abstract}
FILLER

\textbf{Keywords:} 
Categorization; etc
\end{abstract}

\section{Introduction}

One of the most intriguing capabilities of human cognition is the ability to creatively generate new ideas and concepts. % Some statement about how often people use concepts creatively, prevalence, etc
The creative use of knowledge has, however, scarcely been the subject of scientific inquiry in the field of categorization. Most research on concepts and categorization has conformed to a traditional artificial classification learning paradigm \citep{kurtz2015human}, wherein the primary use of category knowledge is to discriminate between two or more experimenter-defined classes.

In one of the earliest explorations on the creative use of conceptual knowledge, Ward \& colleagues \citep{marsh1999inadvertent,smith1993constraining,ward2002role,ward1994structured} provided clear evidence of the role of prior knowledge in concept generation. In a canonical study, \citet{ward1994structured} asked participants to draw and describe novel species of plants and animals that might exist on other planets. Generation was strongly constrained by prior knowledge of Earth plants and animals: people generated alien species using the same structural forms found on Earth (e.g., eyes, legs, wings), exemplars tended to obey the same feature correlations observed on Earth (e.g., feathers tended to co-occur with wings), and exemplars drawn from the same species were less variable than animals drawn from different species (as is the case on Earth).

% early formal models, intro copy and tweak. [joe: did you ever get those papers?]

In a more recent report, \citet{jern2013probabilistic} developed a hierarchical model capable of explaining the influence of prior knowledge observed by \citet{ward1994structured}. In their model, categories are represented as distributions in multidimensional space, where known exemplars are assumed to have been generated from the underlying category distribution. Category distributions are in turn thought to be generated by an underlying domain distribution, representing common distributional characteristics of known categories (e.g., feature variance, feature correlations). When asked to generate exemplars from a novel category, people are thought to sample a new category from the domain distribution, and then sample exemplars from the category distribution. Thus, the generated exemplars tend to obey the same distributional properties of known categories.

\citet{jern2013probabilistic} reported several experiments in support of their model's predictions. In their paradigm, participants are first exposed to members from a set of known categories within a $k$-dimensional feature space (in their experiments, $k=3$). After observing the assigned examples, participants are asked to generate exemplars from a new category within the domain. Participants presented with $k$ scales that can be used to adjust the feature values of the generated stimulus, and are given unlimited time to create each example. As in the classic \cite{ward1994structured} experiment, \citet{jern2013probabilistic} found that people generated categories possessing the same feature correlations as the known categories in the domain.

\section{PACKER: An Exemplar Model}

The work presented by Ward \& Colleagues, as well as that of \citet{jern2013probabilistic}, undoubtedly captures a core principle of the creative use of conceptual knowledge: category generation is informed by the distributional characteristics of known concepts in the target domain. However, this work fails to describe the inverse: how do generated categories \textit{differ} from known categories? In this paper, we introduce a novel exemplar-based approach to category generation, the PACKER model (\textit{Producing Alike and Contrasting Knowledge using Exemplar Representations}), which creates new categories by balancing two constraints: (1) new categories should be different from known categories (minimizing between-class similarity), and (2) new categories should be internally coherent (maximizing within-class similarity).

% Is this evident? "we are not suggesting Jern \& Kemp have failed to capture a core aspect of generation, but that there are other properties that can be explained through one of the field's oldest traditions."

The PACKER model is in many senses an extension of the highly influential Generalized Context Model of category learning (GCM; \citealp{nosofsky1984choice}). The model assumes that each category is represented by a collection of exemplars within in a $k$-dimensional stimulus space, and that generation is constrained by both similarity to members of the category being generated as well as similarity to members of other categories.

As in the GCM, the similarity between two examples, $s(x_i, x_j)$, is defined as an inverse-exponential function of distance:
\begin{equation}
  s(x_i,x_j) = exp( -c \sum_{k}{|x_{ik} - x_{jk}|}w_k ) 
\end{equation}
where $w_k$ is a vector of attention parameters ($\sum_k{w_k} = 1$), weighting to the importance of each feature in the similarity computation. The $c$ parameter controls the specificity of examples: larger values of $c$ produce more \textit{specific} representations (exemplars that do not generalize broadly). The attention weights are of interest in explaining individual differences in generation strategy (discussed below), but for our formal simulations they are set uniformly.

When prompted to generate a new example, the model considers both the summed similarity to examples from other categories as well as the summed similarity to examples in the target category. More formally, the summed similarity $ss$ between generation candidate $y$ and the model's stored exemplars $x_j$ can be computed as:
\begin{equation}
  ss(y, x_j) = \sum_j{f_j s(y, x_j)}
\end{equation}
where $f_j$ is a function specifying each example's degree of contribution toward generation. Although $f_j$ may be set arbitrarily, in PACKER it is set according to class assignment. For known members of contrast categories, $f_j = -\phi$. For known members of the target category, $f_j = \gamma$. $\phi$ and $\gamma$ are thus free parameters ($\geq 0$) controlling the contribution of target- and contrast-category similarity to generation. Larger values for either parameter produce greater consideration of either type of similarity. When $\phi = \gamma$, similarity to contrast categories is effectively subtracted from similarity to the target category. Negative values for $ss(y, x_j)$ would then indicate that $y$ is more similar to members of contrast categories, and positive values indicate $y$ is more similar to members of the target category.

% NOTE: Maybe we should use a different term for summed similarity? Or maybe just indicate specifically how this deviates from classification in the GCM.

The probability that a given candidate $y$ will be generated is evaluated using the \citet{luce1977choice} choice axiom. Candidates with larger summed similarity are more likely to be generated compared to candidates with smaller summed similarity:
\begin{equation}
p(y) = \dfrac
    { exp( { \theta \cdot ss(y, x_j) } ) }
    { \sum_i{ exp( { \theta \cdot ss(y_i, x_j) } ) } }
\end{equation}
where $\theta$ ($\geq 0$) is a free parameter controlling overall response determinism.

\subsubsection{Summary.}
The PACKER model described above explains exemplar and category generation through one of the field's oldest traditions: exemplar representations. The model proposes that people generate categories by maximizing within-category similarity and minimizing between-category similarity. Unlike the model proposed by \citet{jern2013probabilistic}, the distributional properties of known categories affects performance only insofar as it constrains inter-exemplar similarities. Instead, the \textit{location} of known categories in the domain is crucial because it constrains remaining possible locations for new categories. See Figure \ref{fig:example-prob-spaces} for a depiction of the model's behavior.

\begin{figure*}
    \begin{center}
    \inputpgf{figs/}{example-prob-spaces.pgf}
    \caption{Predicted generation of a category `B' example, following exposure to one member of category `A' and one member of category `B'. Areas in which generation is not likely are shaded white; high probability areas are shaded blue. \textit{Left}: Predictions given $\{\phi = 1$, $\gamma = 0\}$ (contrast influence only). \textit{Center}: Predictions given $\{\phi = 0$, $\gamma = 1\}$ (target influence only).  \textit{Right}: Predictions given $\{\phi = 1$, $\gamma = 1\}$ (both constraints considered).  }
    \label{fig:example-prob-spaces}
    \end{center}
\end{figure*}

\section{Behavioral Experiment}

The behavioral experiment described below was designed to test a core prediction of the PACKER model: that the \textit{location} of contrast categories (as opposed to their structure), influences generation. The experiment follows the paradigm set by \citet{jern2013probabilistic}. First, participants are exposed to members of a known category (`Alpha', or `A'), and are then asked to generate exemplars belonging to a new category (`Beta', or `B').

Because the goal of the study was to determine if the structure of known categories structure is the only determinant of generation, our experiment manipulated only the location of the Alpha category within the domain (the structure of the category was held constant). See Figure \ref{fig:middle-bottom-conditions} for a depiction of the experimental conditions. In both condition, members of the Alpha category are tightly clustered, with equal variance on both features and no correlation between features. The conditions differ only in the Y-Axis position of the Alpha category: in the `Middle' condition the Alpha category is placed in the center of the space, in the `Bottom' condition the Alpha category is placed in the bottom-center of the space.

By consequence, the \citet{jern2013probabilistic} model predicts that there should be no difference in the structure of generated categories, as the distributional characteristics of the domain do not differ between the two conditions. Conversely, the PACKER model can predict strong between-condition differences. % What are the predictions---

\begin{figure}
    \begin{center}
    \input{figs/middle-bottom-conditions.pgf}
    \caption{Experiment conditions.}
    \label{fig:middle-bottom-conditions}
    \end{center}
\end{figure}

\subsection{Participants \& Materials}
\subsection{Procedure}
\subsection{Results}

\begin{itemize}
\item Following predictions of our exemplar model, the main interest was the use of space above and below the contrast category. [If-then logic]...
\item Obvious effects of category location: do people tend to place the new category somewhere else? See Figure \ref{fig:middle-bottom-yranges}, Table \ref{table:subset-table}.
\item The more interesting issue: does the remaining space constrain what types of categories people come up with? 
\item Analysis is tricky because vertical range is not normally distributed (see Figure \ref{fig:middle-bottom-yranges}). Report parametric / nonparametric stats.
\item Following core prediction of exemplar model, fewer participants used the top and bottom of the space when assigned the Bottom category, report 2x2x2 contingency table stats.
\item However, we also observed a great deal of individual differences in the form of ... [report types of categories, maybe show samples?]
%JLA: Yeah representative participants from each category and approximate size
\end{itemize}

\section{Simulations}
\section{Discussion}
\section{Acknowledgments}

\begin{figure*}
    \begin{center}
    \input{figs/middle-bottom-yranges.pgf}
    \caption{Behavioral results. Each line shows the minimum and maximum value of a generated category along the Y (vertical) axis. Dots along each line represent the positions of individual exemplars in the category, and each participant's category is shown on a separate line. Participants are sorted by overall Y axis range, and then by condition.}
    \label{fig:middle-bottom-yranges}
    \end{center}
\end{figure*}

\begin{table}[!ht]
\begin{center} 
\caption{Behavioral results.} 
\label{table:subset-table} 
\vskip 0.12in
\begin{tabular}{ l r r}
    \textbf{Middle}         & Used top rows & No top rows \\
    \hline
    Used bottom rows        &  31 & 18  \\
    No bottom rows          &  11 &  1  \\
    \\
    \textbf{Bottom}         & Used top rows & No top rows \\
    \hline
    Used bottom rows        & 22 & 8 \\
    No bottom rows          & 29 & 2 \\
\end{tabular}
\end{center} 
\end{table}

\bibliographystyle{apacite}
\setlength{\bibleftmargin}{.025in}
\setlength{\bibindent}{-\bibleftmargin}
\bibliography{references}

\end{document}

### 2017-01-20 Manuscripts/cogsci-2017/paper.tex
added required sentence to acknowledgements
update to proceedings

@@
 \begin{equation}
-  ss(y, x_j) = \sum_j{f_j s(y, x_j)}
+  ss(y, x) = \sum_j{\hat{f_j} s(y, x_j)}
 \end{equation}
-where $f_j$ is a function specifying each example's degree of contribution toward generation.
-Although $f_j$ may be set arbitrarily, in PACKER it is set according to class assignment.
-For known members of contrast categories, $f_j = -\phi$.
-For known members of the target category, $f_j = \gamma$.
-$\phi$ and $\gamma$ are thus free parameters ($\geq 0$) controlling the contribution of target- and contrast-category similarity to generation.
-Larger values for either parameter produce greater consideration of either type of similarity.
-When $\phi = \gamma$, similarity to contrast categories is effectively subtracted from similarity to the target category.
-Negative values for $ss(y, x_j)$ would then indicate that $y$ is more similar to members of contrast categories, and positive values indicate $y$ is more similar to members of the target category.
+where $\hat{f_j}$ is a function specifying each example's degree of contribution toward generation.
+Although $\hat{f_j}$ may be set arbitrarily, in PACKER it is set according to class assignment.
+For known members of contrast categories, $\hat{f_j} = \phi$.
+For known members of the target category, $\hat{f_j} = \gamma$.
+$\phi$ and $\gamma$ are free parameters ($-\infty \leq \phi, \gamma \leq \infty$) controlling the contribution of target- and contrast-category similarity to generation.
+Larger absolute values for either parameter produce greater consideration of that type of similarity, with values of 0 producing no effect.
+Negative values produce a `repelling' effect (exemplars are less likely to be generated nearby).
+Conversely, positive values produce a `pulling' effect (exemplars are more likely to be generated nearby).
 
-% NOTE: Maybe we should use a different term for summed similarity? Or maybe just indicate specifically how this deviates from classification in the GCM.
+Because the model's main proposals are that new categories should be different from existing categories, and exemplars belonging to the same category should be similar to one another, PACKER is commonly simulated under the constraints that $\phi < 0$, $\gamma > 0$, thus, similarity to contrast categories is effectively subtracted from similarity to the target category.
+When $-\phi = \gamma$, negative summed similarities indicate that $y$ is more similar to members of contrast categories, and positive values indicate $y$ is more similar to members of the target category.
 
@@
 p(y) = \dfrac
-    { exp( { \theta \cdot ss(y, x_j) } ) }
-    { \sum_i{ exp( { \theta \cdot ss(y_i, x_j) } ) } }
+    { exp( { \theta \cdot ss(y, x) } ) }
+    { \sum_i{ exp( { \theta \cdot ss(y_i, x) } ) } }
 \end{equation}
 where $\theta$ ($\geq 0$) is a free parameter controlling overall response determinism.
-
-\subsubsection{Summary.}
-The PACKER model described above explains exemplar and category generation through one of the field's oldest traditions: exemplar representations.
-The model proposes that people generate categories by maximizing within-category similarity and minimizing between-category similarity.
-Unlike the model proposed by \citet{jern2013probabilistic}, the distributional properties of known categories affects performance only insofar as it constrains inter-exemplar similarities.
-Instead, the \textit{location} of known categories in the domain is crucial because it constrains remaining possible locations for new categories.
-See Figure \ref{fig:example-prob-spaces} for a depiction of the model's behavior.
 
@@
 Areas in which generation is not likely are shaded white; high probability areas are shaded blue.
-\textit{Left}: Predictions given $\{\phi = 1$, $\gamma = 0\}$ (contrast influence only).
+\textit{Left}: Predictions given $\{\phi = -1$, $\gamma = 0\}$ (contrast influence only).
 \textit{Center}: Predictions given $\{\phi = 0$, $\gamma = 1\}$ (target influence only).
-\textit{Right}: Predictions given $\{\phi = 1$, $\gamma = 1\}$ (both constraints considered).
+\textit{Right}: Predictions given $\{\phi = -1$, $\gamma = 1\}$ (both constraints considered).
 }
@@
 
+\subsubsection{Summary.}
+The PACKER model described above explains exemplar and category generation through exemplar similarity (one of the field's oldest traditions).
+The model proposes that people generate categories by maximizing within-category similarity and minimizing between-category similarity.
+Unlike the model proposed by \citet{jern2013probabilistic}, the distributional properties of known categories affects performance only insofar as it constrains inter-exemplar similarities.
+Instead, the \textit{location} of known categories in the domain is crucial because it constrains remaining possible locations for new categories.
+See Figure \ref{fig:example-prob-spaces} for a depiction of the model's behavior.
+
 \section{Behavioral Experiment}
@@
 The behavioral experiment described below was designed to test a core prediction of the PACKER model: that the \textit{location} of contrast categories (as opposed to their structure), influences generation.
-The experiment follows the paradigm set by \citet{jern2013probabilistic}.
-First, participants are exposed to members of a known category (`Alpha', or `A'), and are then asked to generate exemplars belonging to a new category (`Beta', or `B').
-
-Because the goal of the study was to determine if the structure of known categories structure is the only determinant of generation, our experiment manipulated only the location of the Alpha category within the domain (the structure of the category was held constant).
-See Figure \ref{fig:middle-bottom-conditions} for a depiction of the experimental conditions.
-In both condition, members of the Alpha category are tightly clustered, with equal variance on both features and no correlation between features.
-The conditions differ only in the Y-Axis position of the Alpha category: in the `Middle' condition the Alpha category is placed in the center of the space, in the `Bottom' condition the Alpha category is placed in the bottom-center of the space.
-
-By consequence, the \citet{jern2013probabilistic} model predicts that there should be no difference in the structure of generated categories, as the distributional characteristics of the domain do not differ between the two conditions.
-Conversely, the PACKER model can predict strong between-condition differences.
-% What are the predictions---
+The experiment follows the paradigm set by \citet{jern2013probabilistic}: first, participants are exposed to members of a known category (`Alpha', or `A'), and are then asked to generate exemplars belonging to a new category (`Beta', or `B').
 
@@
 
+Because the goal of the study was to determine if the structure of known categories is the only determinant of generation, our experiment manipulated only the location of the Alpha category within the domain (the structure of the category was held constant).
+See Figure \ref{fig:middle-bottom-conditions} for a depiction of the experimental conditions.
+In both conditions, members of the Alpha category are tightly clustered, with equal variance on both features and no correlation between features.
+Our manipulation is fairly `weak' --  the conditions have only a slight difference in the Y-Axis position of the Alpha category: in the `Middle' condition the Alpha category is placed in the center of the space, in the `Bottom' condition the Alpha category is placed in the bottom-center of the space.
+
+By consequence, the \citet{jern2013probabilistic} model predicts that there should be no difference in the generated categories, as the distributional characteristics of the domain do not differ between the two conditions.
+However, the PACKER model can predict strong between-condition differences.
+Assuming exemplars are generated such that each category has strong within-class similarity and weak between-class similarity, top areas of the stimulus space should be used for generation more frequently in the Bottom condition (compared to Middle), as  stimuli in this area possess weaker similarity to members of the Alpha category.
+Conversely, the bottom areas of the space should be less frequently used for generation in the Bottom condition (compared to Middle), as stimuli in this area possess greater similarity to the Alpha category.
+
 \subsection{Participants \& Materials}
+
+We recruited 122 participants from Amazon Mechanical Turk.
+All participants were located in the United States.
+61 Participants were assigned to the `Middle' condition, 61 were assigned to the `Bottom' condition.
+Stimuli were squares varying in color (RGB 25--230) and size (3.0--5.8cm), see Figure \ref{fig:stimuli-samples} for samples.
+The assignment of perceptual features (color, size) to axes of the domain space (x, y) was counterbalanced across participants.
+
+\begin{figure}
+    \begin{center}
+    \input{figs/stimuli-samples.pgf}
+    \caption{Sample stimuli (not drawn to scale).}
+    \label{fig:stimuli-samples}
+    \end{center}
+\end{figure}
+
 \subsection{Procedure}
+
+Participants began the experiment with a short training phase (3 blocks of 4 trials), where they iteratively observed exemplars belonging to the `Alpha' category.
+Participants were instructed to learn as much as they can about the Alpha category, and that they would answer a series of test questions afterwards.
+On each trial, a single exemplar from the Alpha category was presented, and participants were given as much time as they desired before moving on.
+
+Following the training phase, participants were asked to generate four examples belonging to another category called `Beta'.
+Participants were instructed that members of the Beta category could be quite similar or different depending on what they think makes the most sense for the category, but that they were not allowed to make the same example twice.
+
+As in the \citet{jern2013probabilistic} experiments, generation was completed using a sliding-scale interface.
+Participants were presented with two scales that could be used to change the color and size of the generated example.
+An on-screen preview of the example updated whenever one of the features was changed.
+Participants could generate any example along an evenly-spaced 9x9 grid, and they could change the example as much as they wanted before moving on.
+Neither the members of the Alpha category nor the previously generated Beta examples were visible during generation.
+
+% % Should we bother reporting this? we never even looked at the data.
+% Following generation participants completed a generalization phase where they were asked to classify 81 stimuli sampled across the domain space (along a 9x9 grid).
+On each trial, a single stimulus was presented and participants were asked to classify the item into either the Alpha or Beta category.
+The members of the Alpha and Beta categories were presented during the generalization phase, but feedback was not provided on responses.
+
 \subsection{Results}
 
-\begin{itemize}
-\item Following predictions of our exemplar model, the main interest was the use of space above and below the contrast category. [If-then logic]...
-\item Obvious effects of category location: do people tend to place the new category somewhere else? See Figure \ref{fig:middle-bottom-yranges}, Table \ref{table:subset-table}.
-\item The more interesting issue: does the remaining space constrain what types of categories people come up with? 
-\item Analysis is tricky because vertical range is not normally distributed (see Figure \ref{fig:middle-bottom-yranges}). Report parametric / nonparametric stats.
-\item Following core prediction of exemplar model, fewer participants used the top and bottom of the space when assigned the Bottom category, report 2x2x2 contingency table stats.
-\item However, we also observed a great deal of individual differences in the form of ... [report types of categories, maybe show samples?]
-%JLA: Yeah representative participants from each category and approximate size
-\end{itemize}
+Because the conditions differ only in their location along the Y (vertical) axis (see Figure \ref{fig:middle-bottom-conditions}), our main interest is in generation of Beta examples above and below the contrast category.
 
-\section{Simulations}
-\section{Discussion}
-\section{Acknowledgments}
+Table \ref{table:subset-table} contains the number of participants in each condition who generated at least one example possessing one of the
 
-\begin{figure*}
-    \begin{center}
-    \input{figs/middle-bottom-yranges.pgf}
-    \caption{Behavioral results.
-Each line shows the minimum and maximum value of a generated category along the Y (vertical) axis.
-Dots along each line represent the positions of individual exemplars in the category, and each participant's category is shown on a separate line.
-Participants are sorted by overall Y axis range, and then by condition.}
-    \label{fig:middle-bottom-yranges}
-    \end{center}
-\end{figure*}
+Fisher's Exact Tests reveal that a greater number of participants assigned to the Middle condition generated at least one Beta example in the bottom rows of the space, $p < 0.001$.
+Similarly, a greater number of Middle participants generated all four Beta examples in the bottom rows of the space, $p = 0.003$.
+Participants in the Bottom condition were more marginally more likely to generate at least one Beta example in the top rows of the space, $p = 0.088$, and they were significantly more likely to generate all four examples in the top rows of the space, $p = 0.017$.
+The conditions did not differ in the number of participants who placed at least one Beta example in the top rows \textit{and} at least one Beta example in the bottom rows, $p = 0.14$.
 
@@
 
+\begin{figure*}
+    \begin{center}
+    \input{figs/middle-bottom-yranges.pgf}
+    \caption{Behavioral results.
+Each line shows the minimum and maximum value of a generated category along the Y (vertical) axis.
+Dots along each line represent the positions of individual exemplars in the category, and each participant's category is shown on a separate line.
+Participants are sorted by overall Y axis range, and then by condition.}
+    \label{fig:middle-bottom-yranges}
+    \end{center}
+\end{figure*}
+
+\subsection{Summary}
+
+\section{Simulations}
+\section{Discussion}
+\section{Acknowledgments}
+Support for this research was provided by the Office of the Vice Chancellor for Research and Graduate Education at the University of Wisconsin - Madison with funding from the Wisconsin Alumni Research Foundation.
+We would like to thanks Kenneth J.
+Kurtz for helpful comments on this work.
+% any other thanks?
+
 \bibliographystyle{apacite}

### 2017-01-20 Experiments/middle-bottom/modeling/positive-negative-spaces.py

@@ -17,9 +17,9 @@ cats = [A,B]
 # params for PACKER
 c, phi = 1.0, 3.0
 
-pos =     [c, 0.0, 1.0, phi]
-neg =     [c, 1.0, 0.0, phi]
-pos_neg = [c, 1.0, 1.0, phi]
+pos =     [c,  0.0, 1.0, phi]
+neg =     [c, -1.0, 0.0, phi]
+pos_neg = [c, -1.0, 1.0, phi]
 
 prob_spaces = {
     'Target Influence': Packer(cats,pos),
@@ -38,7 +38,7 @@ for k in ['Contrast Influence', 'Target Influence', 'Combination']:
     print max(ps)
 
     g = utils.gradientroll(ps, 'roll')[:,:,0]
-    utils.plotgradient(h, g, A, B, cmap = 'PuBu', beta_col = 'w')
+    utils.plotgradient(h, g, A, B, cmap = 'Blues', beta_col = 'w')
     h.set_title(k, fontsize = 11)
 
     counter += 1
@@ -49,4 +49,4 @@ import os, matplotlib
 os.environ["PATH"] += os.pathsep + '/Library/TeX/texbin/'
 opts = {'pgf.texsystem': 'pdflatex'}
 matplotlib.rcParams.update(opts)
-# f.savefig('../../../Manuscripts/cogsci-2017/figs/example-prob-spaces.pgf', bbox_inches='tight')
+f.savefig('../../../Manuscripts/cogsci-2017/figs/example-prob-spaces.pgf', bbox_inches='tight')

### 2017-01-21 Manuscripts/cogsci-2017/paper.tex
updates to figs / manuscript

@@
+%!TEX output_directory = latex_out/
+
 \documentclass[10pt,letterpaper]{article}
@@
 % Some statement about how often people use concepts creatively, prevalence, etc
-The creative use of knowledge has, however, scarcely been the subject of scientific inquiry in the field of categorization.
-Most research on concepts and categorization has conformed to a traditional artificial classification learning paradigm \citep{kurtz2015human}, wherein the primary use of category knowledge is to discriminate between two or more experimenter-defined classes.
+The creative use of knowledge has, however, infrequently been the subject of scientific inquiry in the field of categorization.
+Most research on concepts and categorization has instead conformed to a traditional artificial classification learning paradigm \citep{kurtz2015human}, wherein the primary use of category knowledge is to discriminate between two or more experimenter-defined classes.
 
@@
 After observing the assigned examples, participants are asked to generate exemplars from a new category within the domain.
-Participants presented with $k$ scales that can be used to adjust the feature values of the generated stimulus, and are given unlimited time to create each example.
+Participants presented with a set of scales scales that can be used to adjust the feature values of the generated stimulus, and are given unlimited time to create each example.
 As in the classic \cite{ward1994structured} experiment, \citet{jern2013probabilistic} found that people generated categories possessing the same feature correlations as the known categories in the domain.
@@
 
-% Is this evident?
-"we are not suggesting Jern \& Kemp have failed to capture a core aspect of generation, but that there are other properties that can be explained through one of the field's oldest traditions."
+% Is this evident?: "we are not suggesting Jern \& Kemp have failed to capture a core aspect of generation, but that there are other properties that can be explained through one of the field's oldest traditions."
 
-The PACKER model is in many senses an extension of the highly influential Generalized Context Model of category learning (GCM; \citealp{nosofsky1984choice}).
-The model assumes that each category is represented by a collection of exemplars within in a $k$-dimensional stimulus space, and that generation is constrained by both similarity to members of the category being generated as well as similarity to members of other categories.
+The PACKER model is an extension of the highly influential Generalized Context Model of category learning (GCM; \citealp{nosofsky1984choice}).
+The model assumes that each category is represented by a collection of exemplars within in a $k$-dimensional psychological space, and that generation is constrained by both similarity to members of the target category (the category in which a stimulus is being generated) as well as similarity to members of other categories.
 
@@
 When prompted to generate a new example, the model considers both the summed similarity to examples from other categories as well as the summed similarity to examples in the target category.
-More formally, the summed similarity $ss$ between generation candidate $y$ and the model's stored exemplars $x_j$ can be computed as:
+More formally, the summed similarity $ss$ between generation candidate $y$ and the model's stored exemplars $x$ can be computed as:
 \begin{equation}
@@
 
+\subsubsection{Summary.}
+
+The PACKER model described above explains exemplar and category generation under the traditional exemplar-similarity approach that has guided much of the field's theoretical and empirical work for over three decades.
+The model proposes that people generate categories by maximizing within-category similarity and minimizing between-category similarity.
+See Figure \ref{fig:example-prob-spaces} for a depiction of the model's behavior.
+
 \begin{figure*}
@@
 
-\subsubsection{Summary.}
-The PACKER model described above explains exemplar and category generation through exemplar similarity (one of the field's oldest traditions).
-The model proposes that people generate categories by maximizing within-category similarity and minimizing between-category similarity.
-Unlike the model proposed by \citet{jern2013probabilistic}, the distributional properties of known categories affects performance only insofar as it constrains inter-exemplar similarities.
+Although PACKER appeals to the well-studied exemplar framework, it remains unique as an account of conceptual generation because, unlike the accounts put forward by \citet{jern2013probabilistic} and Ward \& colleagues, it specifies precisely how generated categories should \textit{differ} from known classes in the domain.
+In this paper, we explore whether an exemplar-similarity approach can offer insight into the creative use of conceptual knowledge.
+In particular, we focus on a key difference between PACKER and previous accounts: unlike the model proposed by \citeauthor{jern2013probabilistic}, the distributional properties of known categories affects performance in PACKER only insofar as it constrains inter-exemplar similarities.
 Instead, the \textit{location} of known categories in the domain is crucial because it constrains remaining possible locations for new categories.
-See Figure \ref{fig:example-prob-spaces} for a depiction of the model's behavior.
 
@@
 
-The behavioral experiment described below was designed to test a core prediction of the PACKER model: that the \textit{location} of contrast categories (as opposed to their structure), influences generation.
+The behavioral experiment described below was designed to test the prediction made that the \textit{location} of contrast categories (as opposed to their structure), influences generation.
+As such, our experiment manipulates only the location of the known categories within the domain (the structure of the categories is held constant).
+
 The experiment follows the paradigm set by \citet{jern2013probabilistic}: first, participants are exposed to members of a known category (`Alpha', or `A'), and are then asked to generate exemplars belonging to a new category (`Beta', or `B').
+In both conditions, members of the Alpha category are tightly clustered, with equal variance on both features and no correlation between features.
+See Figure \ref{fig:middle-bottom-conditions} for a depiction of the experimental conditions.
+Our manipulation is fairly `weak': the conditions have only a slight difference in the Y-Axis position of the Alpha category.
+In the `Middle' condition the Alpha category is placed in the center of the space, in the `Bottom' condition the Alpha category is placed in the bottom-center of the space.
 
@@
     \input{figs/middle-bottom-conditions.pgf}
-    \caption{Experiment conditions.}
+    \caption{Conditions tested in the behavioral experiment.}
     \label{fig:middle-bottom-conditions}
@@
 
-Because the goal of the study was to determine if the structure of known categories is the only determinant of generation, our experiment manipulated only the location of the Alpha category within the domain (the structure of the category was held constant).
-See Figure \ref{fig:middle-bottom-conditions} for a depiction of the experimental conditions.
-In both conditions, members of the Alpha category are tightly clustered, with equal variance on both features and no correlation between features.
-Our manipulation is fairly `weak' --  the conditions have only a slight difference in the Y-Axis position of the Alpha category: in the `Middle' condition the Alpha category is placed in the center of the space, in the `Bottom' condition the Alpha category is placed in the bottom-center of the space.
-
-By consequence, the \citet{jern2013probabilistic} model predicts that there should be no difference in the generated categories, as the distributional characteristics of the domain do not differ between the two conditions.
-However, the PACKER model can predict strong between-condition differences.
-Assuming exemplars are generated such that each category has strong within-class similarity and weak between-class similarity, top areas of the stimulus space should be used for generation more frequently in the Bottom condition (compared to Middle), as  stimuli in this area possess weaker similarity to members of the Alpha category.
-Conversely, the bottom areas of the space should be less frequently used for generation in the Bottom condition (compared to Middle), as stimuli in this area possess greater similarity to the Alpha category.
+Because the distributional characteristics of the domain do not differ between the Middle and Bottom conditions, the \citet{jern2013probabilistic} model predicts that there should be no difference in the categories generated within each domain.
+However, the PACKER model proposes that exemplars are generated such that each category has strong within-class similarity and weak between-class similarity.
+By consequence, the upper areas of the stimulus space should be used for generation more frequently in the Bottom condition compared to Middle, as stimuli in this area possess weaker similarity to members of the Alpha category.
+Conversely, the bottom areas of the space should be less frequently used for generation in the Bottom condition compared to Middle, as stimuli in this area possess greater similarity to the Alpha category.
 
@@
 
-Because the conditions differ only in their location along the Y (vertical) axis (see Figure \ref{fig:middle-bottom-conditions}), our main interest is in generation of Beta examples above and below the contrast category.
+Because the conditions differ only in their location along the vertical axis (see Figure \ref{fig:middle-bottom-conditions}), our main interest is in the generation of Beta examples above and below the contrast category.
+A highly detailed picture of these data is shown in Figure \ref{fig:middle-bottom-yranges}, wherein each participant's generated category is depicted in terms of its use of the vertical axis.
+As is evident in Figure \ref{fig:middle-bottom-yranges}, we observed a great deal of variance in generation strategy: whereas some participants generated all four Beta examples within a narrow range of the Y axis, others generated Beta examples along a fairly wide range.
 
-Table \ref{table:subset-table} contains the number of participants in each condition who generated at least one example possessing one of the
+\begin{figure*}
+    \begin{center}
+    \input{figs/middle-bottom-yranges.pgf}
+    \caption{Behavioral results.
+Each line shows the minimum and maximum value of a generated category along the Y (vertical) axis.
+Dots along each line represent the positions of individual exemplars in the category, and each participant's category is shown on a separate line.
+Participants are sorted by overall Y axis range, and then by condition.}
+    \label{fig:middle-bottom-yranges}
+    \end{center}
+\end{figure*}
 
-Fisher's Exact Tests reveal that a greater number of participants assigned to the Middle condition generated at least one Beta example in the bottom rows of the space, $p < 0.001$.
-Similarly, a greater number of Middle participants generated all four Beta examples in the bottom rows of the space, $p = 0.003$.
-Participants in the Bottom condition were more marginally more likely to generate at least one Beta example in the top rows of the space, $p = 0.088$, and they were significantly more likely to generate all four examples in the top rows of the space, $p = 0.017$.
-The conditions did not differ in the number of participants who placed at least one Beta example in the top rows \textit{and} at least one Beta example in the bottom rows, $p = 0.14$.
+To obtain a more abstracted view of our results, we determined the number of participants in each condition who placed at least one Beta example within the top `rows' of the space (one of the top two possible Y axis values), as well as the number of participants who placed at least one example in one of the bottom rows.
+The resulting contingency tables data are shown in Table \ref{table:subset-table}.
 
-\begin{table}[!ht]
+\begin{table}
 \begin{center} 
@@
 
-\begin{figure*}
-    \begin{center}
-    \input{figs/middle-bottom-yranges.pgf}
-    \caption{Behavioral results.
-Each line shows the minimum and maximum value of a generated category along the Y (vertical) axis.
-Dots along each line represent the positions of individual exemplars in the category, and each participant's category is shown on a separate line.
-Participants are sorted by overall Y axis range, and then by condition.}
-    \label{fig:middle-bottom-yranges}
-    \end{center}
-\end{figure*}
+Fisher's Exact Tests reveal that a greater number of participants assigned to the Middle condition generated at least one Beta example in the bottom rows of the space, $p < 0.001$.
+Participants in the Bottom condition were more marginally more likely to generate at least one Beta example in the top rows of the space, $p = 0.088$.
+The conditions did not differ in the number of participants who placed at least one Beta example in the top rows \textit{and} at least one Beta example in the bottom rows, $p = 0.14$.
+Finally, a greater number of Middle participants generated all four Beta examples in the bottom rows of the space, $p = 0.003$, and a greater number of Bottom participants generated all four examples in the top rows of the space, $p = 0.017$.
 
@@
 Support for this research was provided by the Office of the Vice Chancellor for Research and Graduate Education at the University of Wisconsin - Madison with funding from the Wisconsin Alumni Research Foundation.
-We would like to thanks Kenneth J.
+We would like to thank Kenneth J.
 Kurtz for helpful comments on this work.

### 2017-01-21 Experiments/middle-bottom/analysis/plot.heatmaps.py

@@ -16,9 +16,12 @@ alphas = pd.read_sql_query("SELECT * from alphas", con)
 generation = pd.read_sql_query("SELECT * from generation", con)
 con.close()
 
+from scipy.ndimage.filters import gaussian_filter
+
+clim = (0.0, 0.10)
+
 f, ax = plt.subplots(1,2, figsize = (4,2))
-counter = 0
-for i, rows in infodf.groupby('condition'):
+for i, (K, rows) in enumerate(infodf.groupby('condition')):
 
 	pids = rows.participant
 	betas = generation.loc[generation.participant.isin(pids), 'stimulus']
@@ -27,13 +30,22 @@ for i, rows in infodf.groupby('condition'):
 	ps = counts / float(sum(counts))
 	g = utils.gradientroll(ps,'roll')[:,:,0]
 
+	# g = gaussian_filter(g, 0.8)
+
 	print(g)
-	print(i, np.max(g))
+	print(K, np.max(g))
 
-	h = ax[counter]
-	utils.plotgradient(h, g, stimuli[alphas[i],:], [], clim = (0.0, 0.15))
-	counter += 1
+	h = ax[i]
+	im = utils.plotgradient(h, g, stimuli[alphas[K],:], [], clim = clim)
+	h.set_title(K, fontsize = 12)
 
+
+# add colorbar
+f.subplots_adjust(right=0.8)
+cbar = f.add_axes([0.83, 0.225, 0.03, 0.54])
+f.colorbar(im, cax=cbar, ticks = clim)
+cbar.set_yticklabels(['Lowest\nProbability', 'Greatest\nProbability'])
+cbar.tick_params(length = 0)
 
 import os, matplotlib
 os.environ["PATH"] += os.pathsep + '/Library/TeX/texbin/'

### 2017-01-21 Experiments/middle-bottom/analysis/plot.yrange.data.py

@@ -60,7 +60,7 @@ def plotlines(h, data, stats):
 	h.legend(loc="center left", fontsize = 11, frameon=False)
 
 
-fh = plt.figure(figsize = (7,2))
+fh = plt.figure(figsize = (8,2))
 plotlines(fh.gca(), df, stats)
 [i.set_linewidth(1.0) for i in fh.gca().spines.itervalues()]
 

### 2017-01-21 Experiments/middle-bottom/modeling/plot.range.x.location.py

@@ -39,7 +39,7 @@ model_param_pairs = [
         )],
     [Packer, dict(
         specificity = 0.562922970884,
-        between = 1.76500997943,
+        between = -1.76500997943,
         within = 1.55628620461,
         determinism = 1.99990124401,
         )],
@@ -51,7 +51,7 @@ model_param_pairs = [
         )],
 ]
 
-model_obj, params = model_param_pairs[2]
+model_obj, params = model_param_pairs[1]
 nsamples = 20
 
 # get simulated data
@@ -62,7 +62,6 @@ for i, row in stats.groupby('participant'):
 
 	# get weights
 	rs = np.array(row[['xrange','yrange']])[0]
-	rs = 1.0 / (rs + 1.0/9.0)
 	params['wts'] = rs / float(np.sum(rs))
 	# params['wts'] = np.array([0.5, 0.5])
 
@@ -83,42 +82,45 @@ simulated.loc[pd.isnull(simulated['var']),'var'] = 1.0
 simulated['size'] /= nsamples
 
 
-scatter_settings = dict(s = 185, alpha = 1.0, marker = 's', 
+scatter_settings = dict(s = 185, marker = 's', 
 			edgecolors = 'gray', linewidth = 0.5, cmap = 'PuOr',
-			vmin = -2, vmax = 2)
+			vmin = -2.0, vmax = 2.0)
 
-from scipy.interpolate import interp2d
+
+from scipy.ndimage.filters import gaussian_filter
+
 
 # plotting
 f, ax = plt.subplots(2,2,figsize = (5,5))
 for colnum, c in enumerate(pd.unique(info.condition)):
 	A = stimuli[alphas[c],:]
+	
 	for j, data in enumerate([observed, simulated]):
 		h = ax[j][colnum]
 		df = data.loc[data.condition == c]
 
 		# get x/y pos of examples
-		nums = df.stimulus.as_matrix().astype(int)
-		x = stimuli[nums,0]
-		y = stimuli[nums,1]
+		x, y = stimuli[:,0], stimuli[:,1]
 	
-		# compute color
-		n = df['size'].as_matrix()
-		sumx = df['mean'].as_matrix() * n
-		sig = df['var'].as_matrix()
-		sig[sig==0] = 0.01
-		vals  = (0.0/sig +  sumx / sig) / (1.0/sig + n/sig)
+		# compute colors
+		vals = np.zeros(nstimuli)
[…]

### 2017-01-21 Modules/utils/plot.py

@@ -46,7 +46,7 @@ def plotgradient(h, G, alphas, betas,
 		raise Exception("G has too many dimensions. Size: " + str(G.shape))
 
 	# plot gradient
-	h.imshow(np.flipud(G), 
+	im = h.imshow(np.flipud(G), 
 		clim = clim, 
 		origin='lower', 
 		interpolation="nearest", 
@@ -70,7 +70,8 @@ def plotgradient(h, G, alphas, betas,
 	h.set_xticks([])
 	h.set_aspect('equal', adjustable='box')
 	h.axis([-0.5, G.shape[1]-0.5, -0.5, G.shape[0]-0.5])
-
+	return im
+	
 def gradientspace(coords, side):
 	"""
 	Converts a set of coordinates into integer locations within a

### 2017-01-22 Manuscripts/cogsci-2017/paper.tex
update to proceedings, figs

@@
 
-\section{PACKER: An Exemplar Model}
-
 The work presented by Ward \& Colleagues, as well as that of \citet{jern2013probabilistic}, undoubtedly captures a core principle of the creative use of conceptual knowledge: category generation is informed by the distributional characteristics of known concepts in the target domain.
@@
 In this paper, we introduce a novel exemplar-based approach to category generation, the PACKER model (\textit{Producing Alike and Contrasting Knowledge using Exemplar Representations}), which creates new categories by balancing two constraints: (1) new categories should be different from known categories (minimizing between-class similarity), and (2) new categories should be internally coherent (maximizing within-class similarity).
+We then report a behavioral experiment testing the predictions of the PACKER model, and report formal simulations comparing PACKER to the model proposed by \cite{jern2013probabilistic}.
 
-% Is this evident?: "we are not suggesting Jern \& Kemp have failed to capture a core aspect of generation, but that there are other properties that can be explained through one of the field's oldest traditions."
+\section{PACKER: An Exemplar Model}
+
+\begin{figure*}
+    \begin{center}
+    \inputpgf{figs/}{example-prob-spaces.pgf}
+    \caption{PACKER generation of a category `B' example, following exposure to one member of category `A' and one member of category `B'.
+Areas in which generation is not likely are shaded white; high probability areas are shaded blue.
+\textit{Left}: Predictions given $\{\phi = -1$, $\gamma = 0\}$ (contrast influence only).
+\textit{Center}: Predictions given $\{\phi = 0$, $\gamma = 1\}$ (target influence only).
+\textit{Right}: Predictions given $\{\phi = -1$, $\gamma = 1\}$ (both constraints considered).
+}
+    \label{fig:example-prob-spaces}
+    \end{center}
+\end{figure*}
 
@@
 The attention weights are of interest in explaining individual differences in generation strategy (discussed below), but for our formal simulations they are set uniformly.
+
+% Note something about euclidean distance for integral dims?  Or is the field over that?
 
@@
 
-\subsubsection{Summary.}
+\subsection{Summary}
 
@@
 
-\begin{figure*}
-    \begin{center}
-    \inputpgf{figs/}{example-prob-spaces.pgf}
-    \caption{Predicted generation of a category `B' example, following exposure to one member of category `A' and one member of category `B'.
-Areas in which generation is not likely are shaded white; high probability areas are shaded blue.
-\textit{Left}: Predictions given $\{\phi = -1$, $\gamma = 0\}$ (contrast influence only).
-\textit{Center}: Predictions given $\{\phi = 0$, $\gamma = 1\}$ (target influence only).
-\textit{Right}: Predictions given $\{\phi = -1$, $\gamma = 1\}$ (both constraints considered).
-}
-    \label{fig:example-prob-spaces}
-    \end{center}
-\end{figure*}
-
-Although PACKER appeals to the well-studied exemplar framework, it remains unique as an account of conceptual generation because, unlike the accounts put forward by \citet{jern2013probabilistic} and Ward \& colleagues, it specifies precisely how generated categories should \textit{differ} from known classes in the domain.
-In this paper, we explore whether an exemplar-similarity approach can offer insight into the creative use of conceptual knowledge.
-In particular, we focus on a key difference between PACKER and previous accounts: unlike the model proposed by \citeauthor{jern2013probabilistic}, the distributional properties of known categories affects performance in PACKER only insofar as it constrains inter-exemplar similarities.
+In the sections below, we report a behavioral experiment testing a key difference between PACKER and previous accounts: unlike the model proposed by \citeauthor{jern2013probabilistic}, the distributional properties of known categories affects performance in PACKER only insofar as it constrains inter-exemplar similarities.
 Instead, the \textit{location} of known categories in the domain is crucial because it constrains remaining possible locations for new categories.
+We then report formal simulations evaluating the fit of the PACKER and \citet{jern2013probabilistic} models to our data.
 
@@
 
-The behavioral experiment described below was designed to test the prediction made that the \textit{location} of contrast categories (as opposed to their structure), influences generation.
+The behavioral experiment described below was designed to test whether the \textit{location} of contrast categories (as opposed to their structure), influences generation.
 As such, our experiment manipulates only the location of the known categories within the domain (the structure of the categories is held constant).
-
-The experiment follows the paradigm set by \citet{jern2013probabilistic}: first, participants are exposed to members of a known category (`Alpha', or `A'), and are then asked to generate exemplars belonging to a new category (`Beta', or `B').
-In both conditions, members of the Alpha category are tightly clustered, with equal variance on both features and no correlation between features.
-See Figure \ref{fig:middle-bottom-conditions} for a depiction of the experimental conditions.
-Our manipulation is fairly `weak': the conditions have only a slight difference in the Y-Axis position of the Alpha category.
-In the `Middle' condition the Alpha category is placed in the center of the space, in the `Bottom' condition the Alpha category is placed in the bottom-center of the space.
 
@@
 
+The experiment follows the paradigm set by \citet{jern2013probabilistic}: first, participants are exposed to members of a known category (`Alpha', or `A'), and are then asked to generate exemplars belonging to a new category (`Beta', or `B').
+In both conditions, members of the Alpha category are tightly clustered, with equal variance on both features and no correlation between features.
+Our manipulation is fairly `weak' (see Figure \ref{fig:middle-bottom-conditions}): the conditions have only a slight difference in the Y-Axis position of the Alpha category.
+In the `Middle' condition the Alpha category is placed in the center of the space, in the `Bottom' condition the Alpha category is placed in the bottom-center of the space.
+
 Because the distributional characteristics of the domain do not differ between the Middle and Bottom conditions, the \citet{jern2013probabilistic} model predicts that there should be no difference in the categories generated within each domain.
@@
 Conversely, the bottom areas of the space should be less frequently used for generation in the Bottom condition compared to Middle, as stimuli in this area possess greater similarity to the Alpha category.
+% there has GOT to be a more concise way of expressing those final two sentences...
 
@@
 On each trial, a single exemplar from the Alpha category was presented, and participants were given as much time as they desired before moving on.
+Participants were shown a preview of the range of possible colors and sizes prior to training.
 
@@
 Neither the members of the Alpha category nor the previously generated Beta examples were visible during generation.
+Participants were not informed if they generated a member of the Alpha category, but could not complete the trial if they generated the same Beta two times.
 
@@
 Because the conditions differ only in their location along the vertical axis (see Figure \ref{fig:middle-bottom-conditions}), our main interest is in the generation of Beta examples above and below the contrast category.
-A highly detailed picture of these data is shown in Figure \ref{fig:middle-bottom-yranges}, wherein each participant's generated category is depicted in terms of its use of the vertical axis.
-As is evident in Figure \ref{fig:middle-bottom-yranges}, we observed a great deal of variance in generation strategy: whereas some participants generated all four Beta examples within a narrow range of the Y axis, others generated Beta examples along a fairly wide range.
+A highly detailed depiction of our results is shown in Figure \ref{fig:middle-bottom-yranges}, wherein each participant's generated category is depicted in terms of its use of the vertical axis.
+As is evident in Figure \ref{fig:middle-bottom-yranges}, we observed a great deal of individual differences in generation strategy: whereas some participants generated all four Beta examples within a narrow range of the Y axis, others generated Beta examples along a fairly wide range.
 
@@
 
-To obtain a more abstracted view of our results, we determined the number of participants in each condition who placed at least one Beta example within the top `rows' of the space (one of the top two possible Y axis values), as well as the number of participants who placed at least one example in one of the bottom rows.
+To more specifically evaluate the key predictions made by PACKER, we determined the number of participants in each condition who placed at least one Beta example within the top `rows' of the space (one of the top two possible Y axis values), as well as the number of participants who placed at least one example in one of the bottom rows.
 The resulting contingency tables data are shown in Table \ref{table:subset-table}.
@@
 
-Fisher's Exact Tests reveal that a greater number of participants assigned to the Middle condition generated at least one Beta example in the bottom rows of the space, $p < 0.001$.
-Participants in the Bottom condition were more marginally more likely to generate at least one Beta example in the top rows of the space, $p = 0.088$.
-The conditions did not differ in the number of participants who placed at least one Beta example in the top rows \textit{and} at least one Beta example in the bottom rows, $p = 0.14$.
+Fisher's Exact Tests reveal that a greater number of participants assigned to the Middle condition generated a Beta example in the bottom rows of the space, $p < 0.001$.
+Participants in the Bottom condition were more marginally more likely to generate a Beta example in the top rows of the space, $p = 0.088$.
+The conditions did not differ in the number of participants who placed Beta examples in the top rows \textit{and} the bottom rows, $p = 0.14$.
 Finally, a greater number of Middle participants generated all four Beta examples in the bottom rows of the space, $p = 0.003$, and a greater number of Bottom participants generated all four examples in the top rows of the space, $p = 0.017$.
+
+These results are broadly supportive of PACKER's predictions, and are not predicted by existing accounts \citep[i.e.,][]{jern2013probabilistic}.
+However, the results described above are somewhat commonsense: they simply demonstrate that Beta examples tend to be generated in areas not occupied by the Alphas.
+Beyond these effects, we observed a great deal of differences in generation strategy.
+Figure \ref{fig:range-diff-gradient} depicts the average difference in range between the two features (e.g., horizontal -- vertical) across the participant-generated categories, with respect to the location of each category's members.
+These data reveal highly systematic patterns of generation: whereas many participants generated categories with more vertical range (i.e., `column'-like categories), others generated categories with horizontal range (`row'-like categories).
+Furthermore, these two different types of categories appear in distinct locations of the domain space.
+Whereas vertically aligned categories more often appear to the left or right of the Alpha class, horizontally aligned categories appear above and below the Alpha class.
+Thus, even beyond generating Betas in locations not occupied by the Alphas, participants appear to modify the internal structure of their categories in order to maximize distance from the Alphas.
+
+\begin{figure}[ht!]
+    \begin{center}
+    \inputpgf{figs/}{range-diff-gradient.pgf}
+    \caption{Generated category structure as a function of location in the domain space.
+Orange areas in each gradient correspond to stimuli that were commonly generated into category possessing greater vertical (Y-Axis) range.
+Purple areas correspond to categories possessing greater horizontal range.
+White areas correspond to equal range along both features (or infrequent generation).}
+    \label{fig:range-diff-gradient}
+    \end{center}
+\end{figure}
 
@@
 
+Our behavioral experiment revealed strong effects of the role of category location on generation. Not only are participants more likely to generate Betas in locations
+
 \section{Simulations}
+
+We conducted formal simulations to our data using the PACKER model described above, as well as a conjugate implementation of the hierarchical sampling model proposed by \citet{jern2013probabilistic}.
+
+% do we need to describe the implementation?
+
+\subsection{Model-Fitting}
+In order to obtain an overall sense of each model's ability to explain our results, we fitted each model trial-wise, maximizing the log-likelihood of the its predictions against our results.
+For each trial, models were initialized with the participant's Alpha category (Middle or Bottom), and the configuration of the participant's Beta category during that trial.
+Four parameters were fitted for the PACKER model: $c$, $\phi$, $\gamma$, and $\theta$.
+Likewise, four parameters were fitted to the \cite{jern2013probabilistic} model: $\kappa$, $\gamma$, $\nu$, and $\theta$.
+Parameters were not allowed to vary between participants or conditions -- the goal was to obtain the best-fit to our entire dataset.
+
 \section{Discussion}
+
+The creative use of conceptual knowledge is a highly intriguing yet understudied topic in category learning research.
+In this paper, we presented a novel exemplar-based approach to explaining category generation.
+
 \section{Acknowledgments}
 Support for this research was provided by the Office of the Vice Chancellor for Research and Graduate Education at the University of Wisconsin - Madison with funding from the Wisconsin Alumni Research Foundation.
-We would like to thank Kenneth J.
-Kurtz for helpful comments on this work.
+We would like to thank Kenneth Kurtz for helpful comments on this work.
+We also thank Alan Jern \& Charles Kemp for providing code and data from their experiments.
 % any other thanks?

### 2017-01-22 Experiments/middle-bottom/analysis/plot.generation.by.distance.py

@@ -1,0 +1,66 @@
+import sqlite3, sys
+import pandas as pd
+import numpy as np
+import matplotlib.pyplot as plt
+import seaborn as sns
+sns.set_style("whitegrid")
+
+sys.path.insert(0, "../../../Modules/") # generate-categories/Modules
+import utils
+
+# import data
+con = sqlite3.connect('../data/experiment.db')
+info = pd.read_sql_query("SELECT participant, condition from participants", con)
+generation = pd.read_sql_query("SELECT * from generation", con)
+alphas = pd.read_sql_query("SELECT * from alphas", con)
+stimuli = pd.read_sql_query("SELECT * from stimuli", con).as_matrix()
+con.close()
+
+generation = pd.merge(generation, info, on = 'participant')
+
+ngenerations = pd.DataFrame(dict(
+	condition = [],
+	stimulus = [],
+	count = []
+))
+
+for c in pd.unique(info.condition):
+	for i in range(stimuli.shape[0]):
+		count = sum((generation.condition == c) & (generation.stimulus ==i))
+		row = dict(condition = c, stimulus = i, count = count)
+		ngenerations = ngenerations.append(row, ignore_index = True)
+
+
+
+fh = plt.figure(figsize = (3.5,3))
+condition_styles = dict(Middle = 'r-o', Bottom = 'b-o')
+for i, (c, rows) in enumerate(ngenerations.groupby('condition')):
+
+	As = stimuli[alphas[c],:]
+
+	D = utils.pdist(stimuli, As)
+	D = np.mean(D, axis = 1)
+	
+	x = np.unique(D)
+	y = []
+	e = []
+	for j in x:
+		nums = np.where(D == j)[0]
+		curr_rows = rows.loc[rows.stimulus.isin(nums)]
+		counts = curr_rows['count'].as_matrix()
+		y.append(np.mean(counts))
+		e.append(np.std(counts))
+	print y
+	plt.errorbar(x, y, yerr=e, fmt=condition_styles[c], label = c)
+
+plt.gca().xaxis.grid(False)
+# plt.axis([0.15, 1.35, -2, 34])
+
+plt.legend(loc = 'upper left', frameon = True)
+plt.xlabel('Distance From Alpha Category')
+plt.ylabel('Generation Frequency')
+
+fh.savefig('generation.by.distance.png', bbox_inches = 'tight')
+
+
+		

### 2017-01-22 Experiments/middle-bottom/analysis/plot.subsets.py

@@ -1,48 +1,0 @@
-import sqlite3, sys
-import numpy as np
-import pandas as pd
-
-import matplotlib.pyplot as plt
-
-sys.path.insert(0, "../../../Modules/")
-import utils
-
-np.set_printoptions(precision = 2)
-pd.set_option('precision', 2)
-
-# import data
-con = sqlite3.connect('../data/experiment.db')
-info = pd.read_sql_query("SELECT * from participants", con)
-stats = pd.read_sql_query("SELECT * from betastats", con)
-betas = pd.read_sql_query("SELECT participant, stimulus from generation", con)
-alphas = pd.read_sql_query("SELECT * from alphas", con)
-stimuli = pd.read_sql_query("SELECT * from stimuli", con).as_matrix()
-con.close()
-
-savedir = 'subsets'
-data = pd.merge(stats, info[['participant','condition']], on = 'participant')
-
-fields = ['top_and_bottom', 'bottom_only', 'top_only', 'bottom_used', 'top_used']
-
-
-
-for cstring, condition in data.groupby('condition'):
-	As = stimuli[alphas[cstring],:]
-	for F in fields:
-		for v, rows in condition.groupby(F):
-			if rows.shape[0] < 5: continue
-			pids = rows.participant
-			Bs = betas.loc[betas.participant.isin(pids), 'stimulus'].as_matrix()
-			
-			g = np.empty((stimuli.shape[0]))
-			for j in range(stimuli.shape[0]):
-				g[j] = sum(Bs==j)
-			g = g / float(sum(g))
-			g = utils.gradientroll(g,'roll')[:,:,0]
-			print np.max(g)
-
-			f = plt.figure(figsize = (3,3))
-			fname = savedir + '/' + cstring + '-' + F + '-' + str(v) + '.png'
-			utils.plotgradient(f.gca(), g, As, [], clim = (0, 0.2))
-			plt.title('n = ' + str(rows.shape[0]))
-			f.savefig(fname, bbox_inches='tight', transparent=False)

### 2017-01-22 Modules/utils/plot.py

@@ -1,20 +1,24 @@
 import numpy as np
 import utils
 
-def plotclasses(h, stimuli, alphas, betas):
-	textsettings = dict(
+def plotclasses(h, stimuli, alphas, betas,
+	textsettings = None):
+
+	final_textsettings = dict(
 		verticalalignment='center', 
 		horizontalalignment='center',
 		fontsize = 11.0)
 
+	final_textsettings.update(textsettings)
+
 	h.axis(np.array([-1, 1, -1, 1])*1.2)
 	for i in alphas:
 		x, y = stimuli[i,0], stimuli[i,1]
-		h.text(x,y,'A',color = 'r', **textsettings)
+		h.text(x,y,'A',color = 'r', **final_textsettings)
 
 	for i in betas:
 		x, y = stimuli[i,0], stimuli[i,1]
-		h.text(x,y,'B',color = 'b', **textsettings)
+		h.text(x,y,'B',color = 'b', **final_textsettings)
 
 	h.set_yticks([])
 	h.set_xticks([])

### 2017-01-23 Manuscripts/cogsci-2017/paper.tex
intro reorganize

@@
 
-In one of the earliest explorations on the creative use of conceptual knowledge, Ward \& colleagues \citep{marsh1999inadvertent,smith1993constraining,ward2002role,ward1994structured} provided clear evidence of the role of prior knowledge in concept generation.
+In one of the earliest explorations on the creative use of conceptual knowledge, Ward \& colleagues \citep{marsh1999inadvertent,smith1993constraining,ward2002role,ward1994structured,ward1995s} provided clear evidence of the role of prior knowledge in concept generation.
 In a canonical study, \citet{ward1994structured} asked participants to draw and describe novel species of plants and animals that might exist on other planets.
@@
 
-% early formal models, intro copy and tweak. [joe: did you ever get those papers?]
+An early account of these these results proposed that people complete the generation task by retrieving memories of known Earth plants and animals, and slightly changing them to create something new \citep{ward2002role,ward1995s}.
+This account, which we will refer to as the \textit{copy-and-tweak} account can be casted within a formal exemplar approach \citep[see][]{jern2013probabilistic}.
+However, due to the nature of the response formats used (e.g., drawing and describing examples), and due the prior knowledge assumed (e.g., knowledge of plants and animals on Earth), these experiments are not suitable for formal modeling.
 
-In a more recent report, \citet{jern2013probabilistic} developed a hierarchical model capable of explaining the influence of prior knowledge observed by \citet{ward1994structured}.
-In their model, categories are represented as distributions in multidimensional space, where known exemplars are assumed to have been generated from the underlying category distribution.
-Category distributions are in turn thought to be generated by an underlying domain distribution, representing common distributional characteristics of known categories (e.g., feature variance, feature correlations).
-When asked to generate exemplars from a novel category, people are thought to sample a new category from the domain distribution, and then sample exemplars from the category distribution.
-Thus, the generated exemplars tend to obey the same distributional properties of known categories.
-
-\citet{jern2013probabilistic} reported several experiments in support of their model's predictions.
-In their paradigm, participants are first exposed to members from a set of known categories within a $k$-dimensional feature space (in their experiments, $k=3$).
+In more recent work, \citet{jern2013probabilistic} reported several experiments on exemplar and category generation within an artificial category learning paradigm (and are thus suitable for formal modeling).
+In their paradigm, participants are first exposed to members from a set of known categories within a reduced feature space (e.g., size, hue, saturation).
 After observing the assigned examples, participants are asked to generate exemplars from a new category within the domain.
-Participants presented with a set of scales scales that can be used to adjust the feature values of the generated stimulus, and are given unlimited time to create each example.
+Participants provided with a set of scales scales that can be used to adjust the feature values of the generated stimulus (one scale per feature), and are given unlimited time to create each example.
 As in the classic \cite{ward1994structured} experiment, \citet{jern2013probabilistic} found that people generated categories possessing the same feature correlations as the known categories in the domain.
 
-The work presented by Ward \& Colleagues, as well as that of \citet{jern2013probabilistic}, undoubtedly captures a core principle of the creative use of conceptual knowledge: category generation is informed by the distributional characteristics of known concepts in the target domain.
-However, this work fails to describe the inverse: how do generated categories \textit{differ} from known categories?
-In this paper, we introduce a novel exemplar-based approach to category generation, the PACKER model (\textit{Producing Alike and Contrasting Knowledge using Exemplar Representations}), which creates new categories by balancing two constraints: (1) new categories should be different from known categories (minimizing between-class similarity), and (2) new categories should be internally coherent (maximizing within-class similarity).
-We then report a behavioral experiment testing the predictions of the PACKER model, and report formal simulations comparing PACKER to the model proposed by \cite{jern2013probabilistic}.
+\citet{jern2013probabilistic} developed a hierarchical model of generation in order to explain their results.
+In their model, categories are represented as probability distributions in a multidimensional space (with each stimulus feature corresponding to a dimension of the space).
+Observed exemplars (represented as points in this space) are viewed as samples from an underlying category distribution (e.g., a multivariate normal distribution).
+These category distributions are in turn viewed as samples from an underlying domain distribution, which represents the common distributional characteristics among known categories (e.g., feature variance, feature correlations).
+When asked to generate a novel class, people are thought to sample a category distribution from the domain distribution, and then sample exemplars from the category distribution.
+Thus, exemplars generated into novel classes tend to obey the same distributional properties of known categories.
 
-\section{PACKER: An Exemplar Model}
+The work presented by Ward \& Colleagues, as well as that of \citet{jern2013probabilistic}, undoubtedly captures a core principle of the creative use of conceptual knowledge: category generation is informed by prior knowledge of concepts in the target domain.
+The two models described above (copy-and-tweak, hierarchical sampling) are conceptually similar in that they assume learners \textit{copy} something about existing categories: individuals either directly copy-and-tweak stored observations, or copy the distributional structure of entire categories.
 
@@
     \caption{PACKER generation of a category `B' example, following exposure to one member of category `A' and one member of category `B'.
-Areas in which generation is not likely are shaded white; high probability areas are shaded blue.
 \textit{Left}: Predictions given $\{\phi = -1$, $\gamma = 0\}$ (contrast influence only).
@@
 \end{figure*}
+
+In this paper, we introduce a novel exemplar-based approach to category generation, the PACKER model (\textit{Producing Alike and Contrasting Knowledge using Exemplar Representations}), which creates new categories by balancing two constraints: (1) new categories should be different from known categories (minimizing between-class similarity), and (2) new categories should be internally coherent (maximizing within-class similarity).
+PACKER is a significant departure from the accounts of generation reviewed above in that it instantiates assumptions about how exemplars in novel classes should \textit{differ} from existing categories (rather than how they should align to them).
+Likewise, the PACKER model is of interest because it follows directly from the well-studied exemplar-similarity framework that has led the field for over three decades.
+
+In the sections below, we formally describe the PACKER model.
+We then report a behavioral experiment testing the model's predictions, and report formal simulations comparing PACKER to the copy-and-tweak and hierarchical sampling models.
+
+\section{PACKER: An Exemplar Model}
 
@@
 
-% Note something about euclidean distance for integral dims?  Or is the field over that?
+% Note something about minkowski metric? Or can we assume people know that.
 
@@
 
-The PACKER model described above explains exemplar and category generation under the traditional exemplar-similarity approach that has guided much of the field's theoretical and empirical work for over three decades.
+The PACKER model described above explains exemplar and category generation under a traditional exemplar-similarity approach.
 The model proposes that people generate categories by maximizing within-category similarity and minimizing between-category similarity.
-See Figure \ref{fig:example-prob-spaces} for a depiction of the model's behavior.
 
-In the sections below, we report a behavioral experiment testing a key difference between PACKER and previous accounts: unlike the model proposed by \citeauthor{jern2013probabilistic}, the distributional properties of known categories affects performance in PACKER only insofar as it constrains inter-exemplar similarities.
+PACKER is unique as a model of generation because it specifies precisely how generated classes should \textit{differ} from existing ones -- the copy-and-tweak and hierarchical sampling accounts explain why generated classes share many properties with known classes, but not how they should be distinct.
+Likewise, PACKER is unique in that the distributional properties of known categories affect generation only insofar as they constrain inter-exemplar similarities.
 Instead, the \textit{location} of known categories in the domain is crucial because it constrains remaining possible locations for new categories.
-We then report formal simulations evaluating the fit of the PACKER and \citet{jern2013probabilistic} models to our data.
+In the sections below, we report a behavioral experiment testing these predictions.
+We then report formal simulations with PACKER, a copy-and-tweak model, and the \citet{jern2013probabilistic} hierarchical sampling model.
 
@@
 The behavioral experiment described below was designed to test whether the \textit{location} of contrast categories (as opposed to their structure), influences generation.
-As such, our experiment manipulates only the location of the known categories within the domain (the structure of the categories is held constant).
+The experiment follows the paradigm developed by \citet{jern2013probabilistic}: first, participants are exposed to members of a known category (`Alpha', or `A'), and are then asked to generate exemplars belonging to a new category (`Beta', or `B').
+
+We developed two Alpha categories (see Figure \ref{fig:middle-bottom-conditions}), instantiating the two conditions of the experiment.
+In both conditions, members of the Alpha category are tightly clustered, with equal variance on both features and no correlation between features.
+Our manipulation is fairly `weak': the conditions have only a slight difference in the Y-Axis position of the Alpha category.
+In the `Middle' condition the Alpha category is placed in the center of the space, in the `Bottom' condition the Alpha category is placed in the bottom-center of the space.
 
@@
 
-The experiment follows the paradigm set by \citet{jern2013probabilistic}: first, participants are exposed to members of a known category (`Alpha', or `A'), and are then asked to generate exemplars belonging to a new category (`Beta', or `B').
-In both conditions, members of the Alpha category are tightly clustered, with equal variance on both features and no correlation between features.
-Our manipulation is fairly `weak' (see Figure \ref{fig:middle-bottom-conditions}): the conditions have only a slight difference in the Y-Axis position of the Alpha category.
-In the `Middle' condition the Alpha category is placed in the center of the space, in the `Bottom' condition the Alpha category is placed in the bottom-center of the space.
+Although our manipulation is minimal, the PACKER model is capable of predicting strong between-condition differences.
+The PACKER model proposes that the nature of the space not occupied by the Alpha category determines where members of the Beta category are likely to be generated.
+Thus, the lower areas of the stimulus space should be less frequently used for generation in the Bottom condition compared to the Middle (as these areas possess greater similarity to the Bottom Alpha category).
+Conversely, the upper areas of the stimulus space should be used for generation more frequently in the Bottom condition compared to Middle.
 
-Because the distributional characteristics of the domain do not differ between the Middle and Bottom conditions, the \citet{jern2013probabilistic} model predicts that there should be no difference in the categories generated within each domain.
-However, the PACKER model proposes that exemplars are generated such that each category has strong within-class similarity and weak between-class similarity.
-By consequence, the upper areas of the stimulus space should be used for generation more frequently in the Bottom condition compared to Middle, as stimuli in this area possess weaker similarity to members of the Alpha category.
-Conversely, the bottom areas of the space should be less frequently used for generation in the Bottom condition compared to Middle, as stimuli in this area possess greater similarity to the Alpha category.
-% there has GOT to be a more concise way of expressing those final two sentences...
+More generally, the PACKER model proposes that the probability a stimulus $y$ will be generated is a function of its similarity to contrast categories \textit{and} to members of the target category.
+Two more general predictions (not specific to either condition) follow from this proposal: (1) the location of beta examples should be positively related to distance from the Alpha category, (2) Beta examples should be more similar to one another than they are to members of the Alpha category.
+
+% THIS IS WHERE NOLAN STOPPED EDITING AFTER THE INTRO REFORMULATION
 

### 2017-01-23 Manuscripts/cogsci-2017/paper.tex
update

@@
 \usepackage{pslatex}
-\usepackage{amsmath}
+\usepackage{mathtools}
 \usepackage[natbibapa]{apacite}
@@
   s(x_i,x_j) = exp( -c \sum_{k}{|x_{ik} - x_{jk}|}w_k )
+  \label{eq:similarity}
 \end{equation}
@@
   ss(y, x) = \sum_j{\hat{f_j} s(y, x_j)}
+  \label{eq:packer-ss}
 \end{equation}
@@
     { \sum_i{ exp( { \theta \cdot ss(y_i, x) } ) } }
+    \label{eq:packer-choice}
 \end{equation}
@@
 
-We conducted formal simulations to our data using the PACKER model described above, as well as a conjugate implementation of the hierarchical sampling model proposed by \citet{jern2013probabilistic}.
+We conducted formal simulations to our data using the PACKER model described above, as well as an exemplar-based \textit{copy-and-tweak} model, and conjugate implementation of the hierarchical sampling model proposed by \citet{jern2013probabilistic}.
+Below, we briefly describe the formal basis of the copy-and-tweak and hierarchical sampling models.
+
+\subsection{Copy-and-Tweak}
+The copy-and-tweak model proposes that generation is a two-part process: first, learners retrieve a source exemplar from memory, and then they change it in some way so that the generated item is sufficiently dissimilar.
+In our model, the probability that a given exemplar $z$ is retrieved from memory is given by: $p(z) = exp(\hat{f}_z) / \sum_z{ exp(\hat{f}_z) }$.
+$\hat{f}_z$ is a function specifying each item's relative chance of being selected.
+As in Equation \ref{eq:packer-ss}, $\hat{f}_j$ may be set arbitrarily.
+However, in our simulations, $\hat{f}_z = -\gamma$ for members of contrast categories, and $\hat{f}_z = \gamma$ for members of the target category.
+The resulting free parameter $\gamma$ ($>0$) thus controls the probability that a member of the target category will be retrieved from memory.
+
+After a source example is retrieved, the similarity between generation candidates $y$ and the source is computed as per Equation \ref{eq:similarity}.
+The model's goal is to generate an item that is similar to $z$, but exceeds a dissimilarity threshold (such that the new item is similar-but-not-too-similar).
+Formally, the probability that $y$ will be generated based on a source $z$, is given by:
+
+\begin{equation}
+  \begin{cases} 
+    p(y|z) \propto exp(\theta \cdot s(y,z)), & \text{if } s(y,z) \leq \tau \\
+    p(y|z) = 0, & \text{otherwise} 
+  \end{cases} 
+\end{equation}
+where $\theta$ is a response determinism parameter, and $\tau$ is a similarity threshold ($0 < \tau \leq 1$) specifying the amount of similarity tolerated between $y$ and $z$.
+Values of $p(y|z)$ are then normalized such that $\sum_i{p(y_i|z)} = 1$.
+To obtain predictions not depending on a given source example, the model's predictions can be aggregated over all possible sources:
+
+\begin{equation}
+  p(y) = \sum_z{p(z)p(y|z) }
+\end{equation}
+
+Thus, the probability that an example $y$ will be selected is a function of its similarity to each of the possible sources, as well as the probability each source will be retrieved.
+
+\subsection{Hierarchical Sampling}
+
+In the \cite{jern2013probabilistic} model, examples are viewed as samples of an underlying category distribution, which is in turn viewed as a sample of an underlying domain distribution.
+As in the implementation reported by \cite{jern2013probabilistic}, we assume that exemplars are distributed according to a multivariate normal distribution with parameters ($\mu, \Sigma$).
+To obtain a conjugate model, we assume that the domain of categories is inverse-Wishart distributed.
+Space does not allow for a full derivation of this model, and so we will simply report how generated category parameters ($\mu_B, \Sigma_B$) are inferred.
 

### 2017-01-24 Manuscripts/cogsci-2017/paper.tex
another intro reorganization

@@
 \title{PACKER: An Exemplar Model of Category Generation}
- %JLA: I like an exemplar model of category generation better, the only issue is there already is an exemplar model. Need to figure out how to get PACKER or something like it as the acronym...
- % 		is that Jern & Kemp 2013 named their own exemplar model for category generation  ("copy-and-tweak"). We need to implement something like it...
 
@@
 One of the most intriguing capabilities of human cognition is the ability to creatively generate new ideas and concepts.
-% Some statement about how often people use concepts creatively, prevalence, etc
-The creative use of knowledge has, however, infrequently been the subject of scientific inquiry in the field of categorization.
-Most research on concepts and categorization has instead conformed to a traditional artificial classification learning paradigm \citep{kurtz2015human}, wherein the primary use of category knowledge is to discriminate between two or more experimenter-defined classes.
+The creative use of knowledge has, however, only infrequently been the subject of inquiry in the field of categorization.
+To date, most of what we understand about human concept generation comes from foundational studies in the field of creative cognition \citep[e.g.,][]{marsh1999inadvertent,smith1993constraining,ward2002role,ward1994structured,ward1995s}, though some work has recently been conducted using traditional classification learning approaches \citep{jern2013probabilistic}.
 
-In one of the earliest explorations on the creative use of conceptual knowledge, Ward \& colleagues \citep{marsh1999inadvertent,smith1993constraining,ward2002role,ward1994structured,ward1995s} provided clear evidence of the role of prior knowledge in concept generation.
-In a canonical study, \citet{ward1994structured} asked participants to draw and describe novel species of plants and animals that might exist on other planets.
-Generation was strongly constrained by prior knowledge of Earth plants and animals: people generated alien species using the same structural forms found on Earth (e.g., eyes, legs, wings), exemplars tended to obey the same feature correlations observed on Earth (e.g., feathers tended to co-occur with wings), and exemplars drawn from the same species were less variable than animals drawn from different species (as is the case on Earth).
+Existing work on the creative use of conceptual knowledge predominantly explores the role of prior knowledge in generating novel concepts.
+The core experimental phenomenon is that people generate categories possessing the same distributional properties as known categories.
+For example, in a canonical study \citet{ward1994structured} asked participants to draw and describe novel species of plants and animals that might exist on other planets.
+Generation was strongly constrained by prior knowledge of Earth plants and animals: people generated alien species using the same structural forms found on Earth (e.g., eyes, legs, wings) and possessing the same feature correlations observed on Earth (e.g., feathers co-occur with wings).
+Likewise, exemplars drawn from the same species were less variable than animals drawn from different species, and possessed features that were adaptive for their environment (as is the case on Earth).
 
-An early account of these these results proposed that people complete the generation task by retrieving memories of known Earth plants and animals, and slightly changing them to create something new \citep{ward2002role,ward1995s}.
-This account, which we will refer to as the \textit{copy-and-tweak} account can be casted within a formal exemplar approach \citep[see][]{jern2013probabilistic}.
-However, due to the nature of the response formats used (e.g., drawing and describing examples), and due the prior knowledge assumed (e.g., knowledge of plants and animals on Earth), these experiments are not suitable for formal modeling.
+Similar results were obtained more recently in a study conducted within a traditional categorization paradigm.
+\citet{jern2013probabilistic} trained participants on a set of experimenter-defined categories composed of exemplars within an artificial three-dimensional domain (e.g., 2D shapes varying in size, hue, and saturation).
+After a short training phase, participants were asked to generate exemplars from a new category.
+Participants were provided with a set of scales to adjust the feature values of each generated stimulus, and were given unlimited time to create each example.
+As in the classic \cite{ward1994structured} experiment, \citet{jern2013probabilistic} found that people generated artificial categories possessing the same feature variance and correlations as the experimenter-defined categories in the domain.
 
-In more recent work, \citet{jern2013probabilistic} reported several experiments on exemplar and category generation within an artificial category learning paradigm (and are thus suitable for formal modeling).
-In their paradigm, participants are first exposed to members from a set of known categories within a reduced feature space (e.g., size, hue, saturation).
-After observing the assigned examples, participants are asked to generate exemplars from a new category within the domain.
-Participants provided with a set of scales scales that can be used to adjust the feature values of the generated stimulus (one scale per feature), and are given unlimited time to create each example.
-As in the classic \cite{ward1994structured} experiment, \citet{jern2013probabilistic} found that people generated categories possessing the same feature correlations as the known categories in the domain.
+Results such as these have motivated formal accounts of generation that explicitly invoke the idea that people re-purpose existing knowledge to generate something new -- either by copying-and-tweaking an observation retrieved from memory \citep{ward2002role,ward1995s}, or by abstracting the distributional structure of categories in the domain \citep{jern2013probabilistic}.
 
-\citet{jern2013probabilistic} developed a hierarchical model of generation in order to explain their results.
+ % An early account, which \citet{jern2013probabilistic} termed the \textit{copy-and-tweak} account, proposed that people complete the generation task by retrieving observations from memory, and slightly changing them to create something new \citep{ward2002role,ward1995s}.
+This account can be casted within a formal exemplar approach -- first, participants retrieve an exemplar from memory, and then slightly change its features until it is sufficiently novel.
+
+% \citet{jern2013probabilistic} developed a hierarchical model of generation in order to explain their results.
 In their model, categories are represented as probability distributions in a multidimensional space (with each stimulus feature corresponding to a dimension of the space).
@@
 Thus, exemplars generated into novel classes tend to obey the same distributional properties of known categories.
-
-The work presented by Ward \& Colleagues, as well as that of \citet{jern2013probabilistic}, undoubtedly captures a core principle of the creative use of conceptual knowledge: category generation is informed by prior knowledge of concepts in the target domain.
-The two models described above (copy-and-tweak, hierarchical sampling) are conceptually similar in that they assume learners \textit{copy} something about existing categories: individuals either directly copy-and-tweak stored observations, or copy the distributional structure of entire categories.
 
@@
 In this paper, we introduce a novel exemplar-based approach to category generation, the PACKER model (\textit{Producing Alike and Contrasting Knowledge using Exemplar Representations}), which creates new categories by balancing two constraints: (1) new categories should be different from known categories (minimizing between-class similarity), and (2) new categories should be internally coherent (maximizing within-class similarity).
-PACKER is a significant departure from the accounts of generation reviewed above in that it instantiates assumptions about how exemplars in novel classes should \textit{differ} from existing categories (rather than how they should align to them).
-Likewise, the PACKER model is of interest because it follows directly from the well-studied exemplar-similarity framework that has led the field for over three decades.
+As such, PACKER is a significant departure from previous accounts of generation -- rather than proposing that people create new categories by re-purposing existing knowledge, PACKER explains generation exclusively in terms of the well-studied mechanics of exemplar representations and therefore possesses a rich connection to the wider body of research on category learning.
 
 In the sections below, we formally describe the PACKER model.
-We then report a behavioral experiment testing the model's predictions, and report formal simulations comparing PACKER to the copy-and-tweak and hierarchical sampling models.
+We then explore the model's predictions using a behavioral experiment and simulations.
+Rather than formally evaluate PACKER on the benchmarks reviewed above, we instead focus on testing predictions made by PACKER that distinguish the model from previous approaches.
 
@@
 \end{equation}
-where $w_k$ is a vector of attention parameters ($\sum_k{w_k} = 1$), weighting to the importance of each feature in the similarity computation.
-The $c$ parameter controls the specificity of examples: larger values of $c$ produce more \textit{specific} representations (exemplars that do not generalize broadly).
+where $w_k$ is a vector of attention parameters ($\sum_k{w_k} = 1$), weighting to the importance of each feature in the similarity computation, and $c$ ($<0$) is a specificity parameter controlling the spread of exemplar generalization.
 The attention weights are of interest in explaining individual differences in generation strategy (discussed below), but for our formal simulations they are set uniformly.
@@
 
-When prompted to generate a new example, the model considers both the summed similarity to examples from other categories as well as the summed similarity to examples in the target category.
-More formally, the summed similarity $ss$ between generation candidate $y$ and the model's stored exemplars $x$ can be computed as:
+When prompted to generate a new example, the model considers both the similarity to examples from other categories as well as the similarity to examples in the target category.
+More formally, the aggregated similarity $a$ between generation candidate $y$ and the model's stored exemplars $x$ can be computed as:
 \begin{equation}
-  ss(y, x) = \sum_j{\hat{f_j} s(y, x_j)}
+  a(y, x) = \sum_j{\hat{f_j} s(y, x_j)}
   \label{eq:packer-ss}
@@
 Because the model's main proposals are that new categories should be different from existing categories, and exemplars belonging to the same category should be similar to one another, PACKER is commonly simulated under the constraints that $\phi < 0$, $\gamma > 0$, thus, similarity to contrast categories is effectively subtracted from similarity to the target category.
-When $-\phi = \gamma$, negative summed similarities indicate that $y$ is more similar to members of contrast categories, and positive values indicate $y$ is more similar to members of the target category.
+When $-\phi = \gamma$, negative aggregate similarities indicate that $y$ is more similar to members of contrast categories, and positive values indicate $y$ is more similar to the target category.
 
 The probability that a given candidate $y$ will be generated is evaluated using the \citet{luce1977choice} choice axiom.
-Candidates with larger summed similarity are more likely to be generated compared to candidates with smaller summed similarity:
+Candidates with larger aggregate similarity are more likely to be generated compared to candidates with smaller aggregate similarity:
 \begin{equation}
 p(y) = \dfrac
-    { exp( { \theta \cdot ss(y, x) } ) }
-    { \sum_i{ exp( { \theta \cdot ss(y_i, x) } ) } }
+    { exp( { \theta \cdot a(y, x) } ) }
+    { \sum_i{ exp( { \theta \cdot a(y_i, x) } ) } }
     \label{eq:packer-choice}
@@
 
-The PACKER model described above explains exemplar and category generation under a traditional exemplar-similarity approach.
-The model proposes that people generate categories by maximizing within-category similarity and minimizing between-category similarity.
+The PACKER model described above explains exemplar and category generation under a traditional exemplar-similarity approach, proposing that people generate categories by maximizing within-category similarity and minimizing between-category similarity.
+By consequence, PACKER is unique as a model of generation in several respects.
+Not only does the model lack mechanisms for copying knowledge of existing categories, but it also specifies precisely how generated classes should \textit{differ} from existing ones.
+Thus, unlike previous accounts distributional properties of known categories affect generation only insofar as they constrain inter-exemplar similarities.
+Instead, the \textit{location} of known categories in the domain is crucial because it constrains remaining possible locations for new categories.
 
-PACKER is unique as a model of generation because it specifies precisely how generated classes should \textit{differ} from existing ones -- the copy-and-tweak and hierarchical sampling accounts explain why generated classes share many properties with known classes, but not how they should be distinct.
-Likewise, PACKER is unique in that the distributional properties of known categories affect generation only insofar as they constrain inter-exemplar similarities.
-Instead, the \textit{location} of known categories in the domain is crucial because it constrains remaining possible locations for new categories.
-In the sections below, we report a behavioral experiment testing these predictions.
-We then report formal simulations with PACKER, a copy-and-tweak model, and the \citet{jern2013probabilistic} hierarchical sampling model.
+\section{Alternative Models}
+
+\subsection{Copy-and-Tweak}
+The copy-and-tweak model proposes that generation is a two-part process: first, learners retrieve a source exemplar from memory, and then they change it in some way so that the generated item is sufficiently dissimilar.
+In our model, the probability that a given exemplar $z$ is retrieved from memory is given by:
+
+\begin{equation}
+p(z) = exp(\hat{f}_z) / \sum_z{ exp(\hat{f}_z) }
+\end{equation}
+
+$\hat{f}_z$ is a function specifying each item's relative chance of being selected.
+As in Equation \ref{eq:packer-ss}, $\hat{f}_j$ may be set arbitrarily.
+However, in our simulations, $\hat{f}_z = -\gamma$ for members of contrast categories, and $\hat{f}_z = \gamma$ for members of the target category.
+The resulting free parameter $\gamma$ ($>0$) thus controls the probability that a member of the target category will be retrieved from memory.
+
+After a source example is retrieved, the similarity between generation candidates $y$ and the source is computed as per Equation \ref{eq:similarity}.
+The model's goal is to generate an item that is similar to $z$, but exceeds a dissimilarity threshold (such that the new item is similar-but-not-too-similar).
+Formally, the probability that $y$ will be generated based on a source $z$, is given by:
+
+\begin{equation}
+    p(y|z)  = \dfrac
+    { exp(\theta \cdot s(y,z)) g(y,z,\tau)}
+    {\sum_i{exp(\theta \cdot s(y_i,z)) g(y_i,z,\tau)}} 
+\end{equation}
+where $\theta$ is a response determinism parameter.
+$g$ is a function specifying whether $y$ is sufficiently dissimilar to $z$.
+The dissimilarity threshold is specified by a free parameter, $\tau$ ($0<\tau\leq1$).
+$g = 1$ when $s(y,z) < \tau$ and $g = 0$ otherwise.
+
+To obtain predictions not depending on a given source example, the model's predictions can be aggregated over all possible sources:
+
+\begin{equation}
+  p(y) = \sum_z{p(z)p(y|z) }
+\end{equation}
+
+Thus, the probability that an example $y$ will be selected is a function of its similarity to each of the possible sources, as well as the probability each source will be retrieved.
+
+\subsection{Hierarchical Sampling}
+
+In the \cite{jern2013probabilistic} model, examples are viewed as samples of an underlying category distribution, which is in turn viewed as a sample of an underlying domain distribution.
+As in the implementation reported by \cite{jern2013probabilistic}, we assume that exemplars are distributed according to a multivariate normal distribution with parameters ($\mu, \Sigma$).
+To obtain a conjugate model, we assume that the domain of categories is inverse-Wishart distributed.
+Space does not allow for a full derivation of this model, and so we will simply report how generated category parameters ($\mu_B, \Sigma_B$) are inferred.
+
+% do we need to describe the implementation?
 
@@
 
+The PACKER model predicts
 The behavioral experiment described below was designed to test whether the \textit{location} of contrast categories (as opposed to their structure), influences generation.
@@
 61 Participants were assigned to the `Middle' condition, 61 were assigned to the `Bottom' condition.
-Stimuli were squares varying in color (RGB 25--230) and size (3.0--5.8cm), see Figure \ref{fig:stimuli-samples} for samples.
+Stimuli were squares varying in color (RGB 25--230) and size (3.0--5.8cm).
 The assignment of perceptual features (color, size) to axes of the domain space (x, y) was counterbalanced across participants.
-
-\begin{figure}
-    \begin{center}
-    \input{figs/stimuli-samples.pgf}
-    \caption{Sample stimuli (not drawn to scale).}
-    \label{fig:stimuli-samples}
-    \end{center}
-\end{figure}
 
@@
 
-\subsection{Copy-and-Tweak}
-The copy-and-tweak model proposes that generation is a two-part process: first, learners retrieve a source exemplar from memory, and then they change it in some way so that the generated item is sufficiently dissimilar.
-In our model, the probability that a given exemplar $z$ is retrieved from memory is given by: $p(z) = exp(\hat{f}_z) / \sum_z{ exp(\hat{f}_z) }$.
-$\hat{f}_z$ is a function specifying each item's relative chance of being selected.
-As in Equation \ref{eq:packer-ss}, $\hat{f}_j$ may be set arbitrarily.
-However, in our simulations, $\hat{f}_z = -\gamma$ for members of contrast categories, and $\hat{f}_z = \gamma$ for members of the target category.
-The resulting free parameter $\gamma$ ($>0$) thus controls the probability that a member of the target category will be retrieved from memory.
-
-After a source example is retrieved, the similarity between generation candidates $y$ and the source is computed as per Equation \ref{eq:similarity}.
-The model's goal is to generate an item that is similar to $z$, but exceeds a dissimilarity threshold (such that the new item is similar-but-not-too-similar).
-Formally, the probability that $y$ will be generated based on a source $z$, is given by:
-
-\begin{equation}
-  \begin{cases} 
-    p(y|z) \propto exp(\theta \cdot s(y,z)), & \text{if } s(y,z) \leq \tau \\
-    p(y|z) = 0, & \text{otherwise} 
-  \end{cases} 
-\end{equation}
-where $\theta$ is a response determinism parameter, and $\tau$ is a similarity threshold ($0 < \tau \leq 1$) specifying the amount of similarity tolerated between $y$ and $z$.
-Values of $p(y|z)$ are then normalized such that $\sum_i{p(y_i|z)} = 1$.
-To obtain predictions not depending on a given source example, the model's predictions can be aggregated over all possible sources:
-
-\begin{equation}
-  p(y) = \sum_z{p(z)p(y|z) }
-\end{equation}
-
-Thus, the probability that an example $y$ will be selected is a function of its similarity to each of the possible sources, as well as the probability each source will be retrieved.
-
-\subsection{Hierarchical Sampling}
-
-In the \cite{jern2013probabilistic} model, examples are viewed as samples of an underlying category distribution, which is in turn viewed as a sample of an underlying domain distribution.
-As in the implementation reported by \cite{jern2013probabilistic}, we assume that exemplars are distributed according to a multivariate normal distribution with parameters ($\mu, \Sigma$).
-To obtain a conjugate model, we assume that the domain of categories is inverse-Wishart distributed.
-Space does not allow for a full derivation of this model, and so we will simply report how generated category parameters ($\mu_B, \Sigma_B$) are inferred.
-
-% do we need to describe the implementation?
-
 \subsection{Model-Fitting}

### 2017-01-25 Manuscripts/cogsci-2017/paper.tex
drafted copytweak and hs

@@
 
- % An early account, which \citet{jern2013probabilistic} termed the \textit{copy-and-tweak} account, proposed that people complete the generation task by retrieving observations from memory, and slightly changing them to create something new \citep{ward2002role,ward1995s}.
-This account can be casted within a formal exemplar approach -- first, participants retrieve an exemplar from memory, and then slightly change its features until it is sufficiently novel.
-
-% \citet{jern2013probabilistic} developed a hierarchical model of generation in order to explain their results.
-In their model, categories are represented as probability distributions in a multidimensional space (with each stimulus feature corresponding to a dimension of the space).
-Observed exemplars (represented as points in this space) are viewed as samples from an underlying category distribution (e.g., a multivariate normal distribution).
-These category distributions are in turn viewed as samples from an underlying domain distribution, which represents the common distributional characteristics among known categories (e.g., feature variance, feature correlations).
-When asked to generate a novel class, people are thought to sample a category distribution from the domain distribution, and then sample exemplars from the category distribution.
-Thus, exemplars generated into novel classes tend to obey the same distributional properties of known categories.
-
 \begin{figure*}
@@
 
-In the sections below, we formally describe the PACKER model.
-We then explore the model's predictions using a behavioral experiment and simulations.
-Rather than formally evaluate PACKER on the benchmarks reviewed above, we instead focus on testing predictions made by PACKER that distinguish the model from previous approaches.
+In the sections below, we formally describe the PACKER model and explore its predictions in a behavioral experiment and simulations.
+Rather than formally evaluate PACKER on the benchmarks reviewed above, however, we instead focus on testing predictions made by PACKER that distinguish the model from previous approaches.
 
@@
 
-Because the model's main proposals are that new categories should be different from existing categories, and exemplars belonging to the same category should be similar to one another, PACKER is commonly simulated under the constraints that $\phi < 0$, $\gamma > 0$, thus, similarity to contrast categories is effectively subtracted from similarity to the target category.
+Because the model's main proposals are that new categories should be different from existing categories, and exemplars belonging to the same category should be similar to one another, PACKER is commonly simulated under the constraints that $\phi \leq 0$, $\gamma \geq 0$, thus, similarity to contrast categories is effectively subtracted from similarity to the target category.
 When $-\phi = \gamma$, negative aggregate similarities indicate that $y$ is more similar to members of contrast categories, and positive values indicate $y$ is more similar to the target category.
@@
 The PACKER model described above explains exemplar and category generation under a traditional exemplar-similarity approach, proposing that people generate categories by maximizing within-category similarity and minimizing between-category similarity.
-By consequence, PACKER is unique as a model of generation in several respects.
+These core design principles yield unique predictions about the generation.
 Not only does the model lack mechanisms for copying knowledge of existing categories, but it also specifies precisely how generated classes should \textit{differ} from existing ones.
-Thus, unlike previous accounts distributional properties of known categories affect generation only insofar as they constrain inter-exemplar similarities.
+Thus, unlike previous accounts, distributional properties of known categories affect generation only insofar as they constrain inter-exemplar similarities.
 Instead, the \textit{location} of known categories in the domain is crucial because it constrains remaining possible locations for new categories.
 
-\section{Alternative Models}
+\section{Previous Accounts of Category Generation}
+
+As noted above, existing accounts of category generation focus primarily on explaining the phenomenon that generated categories tend to possess distributional similarities with existing categories.
+\citet{jern2013probabilistic} were the first (as far as we know) to conduct formal simulations on category generation.
+Thus, the models they considered represent the strongest alternatives to PACKER.
+Below, we describe in detail two of these alternatives: a \textit{copy-and-tweak} account that was initially proposed within the creative cognition literature \citep{ward2002role,ward1995s}, and the \textit{hierarchical sampling} model developed by \citet{jern2013probabilistic}.
 
 \subsection{Copy-and-Tweak}
-The copy-and-tweak model proposes that generation is a two-part process: first, learners retrieve a source exemplar from memory, and then they change it in some way so that the generated item is sufficiently dissimilar.
-In our model, the probability that a given exemplar $z$ is retrieved from memory is given by:
+
+The \textit{copy-and-tweak} model implemented by \citet{jern2013probabilistic} instantiates one of the earliest accounts of category generation \citep[see][]{ward2002role,ward1995s}.
+The model proposes that generation is a two-part process: first, learners retrieve an observation from memory, and then they change it to create something new.
+\citet{jern2013probabilistic} interpreted this proposal in terms of a formal exemplar approach -- first, participants retrieve an exemplar from memory, and slightly change its features until it is sufficiently novel.
+In our implementation of the copy-and-tweak model, the probability that a given exemplar $z$ is retrieved from memory is given by:
 
@@
 \end{equation}
-
-$\hat{f}_z$ is a function specifying each item's relative chance of being selected.
+%
+where $\hat{f}_z$ is a function specifying each item's relative chance of being selected.
 As in Equation \ref{eq:packer-ss}, $\hat{f}_j$ may be set arbitrarily.
 However, in our simulations, $\hat{f}_z = -\gamma$ for members of contrast categories, and $\hat{f}_z = \gamma$ for members of the target category.
-The resulting free parameter $\gamma$ ($>0$) thus controls the probability that a member of the target category will be retrieved from memory.
+The resulting free parameter $\gamma$ ($>0$) thus controls the relative chances that a member of the target category will be retrieved as the source.
+\citet{jern2013probabilistic} did not implement any mechanisms to differentially weigh retrieval probabilities, and so our model encompasses theirs as a special case when $\gamma = 0$ (uniform probabilities).
 
-After a source example is retrieved, the similarity between generation candidates $y$ and the source is computed as per Equation \ref{eq:similarity}.
-The model's goal is to generate an item that is similar to $z$, but exceeds a dissimilarity threshold (such that the new item is similar-but-not-too-similar).
+After a source exemplar is retrieved, the similarity between generation candidates $y$ and the source is computed as per Equation \ref{eq:similarity}.
+The model's goal is to generate an item that is similar to $z$, but does not exceed a given similarity threshold: the generated item should be similar-but-not-too-similar to $z$.
 Formally, the probability that $y$ will be generated based on a source $z$, is given by:
@@
 \end{equation}
+% 
 where $\theta$ is a response determinism parameter.
-$g$ is a function specifying whether $y$ is sufficiently dissimilar to $z$.
-The dissimilarity threshold is specified by a free parameter, $\tau$ ($0<\tau\leq1$).
+$g$ is a function specifying whether $y$ exceeds the similarity threshold, $\tau$, which is treated as a free parameter ($0<\tau\leq1$).
 $g = 1$ when $s(y,z) < \tau$ and $g = 0$ otherwise.
+\citet{jern2013probabilistic} did not implement any similarity threshold mechanism, thus, their model is a special case when $\tau = 1$ (any exemplar may be selected).
 
@@
 \end{equation}
-
+% 
 Thus, the probability that an example $y$ will be selected is a function of its similarity to each of the possible sources, as well as the probability each source will be retrieved.
@@
 
-In the \cite{jern2013probabilistic} model, examples are viewed as samples of an underlying category distribution, which is in turn viewed as a sample of an underlying domain distribution.
-As in the implementation reported by \cite{jern2013probabilistic}, we assume that exemplars are distributed according to a multivariate normal distribution with parameters ($\mu, \Sigma$).
+\citet{jern2013probabilistic} reported several experiments revealing systematic flaws in the copy-and-tweak account.
+Instead, their results supported a hierarchical sampling model, in where categories are represented as probability distributions in a multidimensional space, with each stimulus feature corresponding to a dimension of the space.
+Observed exemplars (represented as points in this space) are viewed as samples from their underlying category distribution (e.g., a multivariate normal distribution).
+These category distributions are in turn viewed as samples from an underlying domain distribution, which represents the common distributional characteristics among known categories (e.g., feature variance, feature correlations).
+When asked to generate a novel class, people are thought to sample a category distribution from the domain distribution, and then sample exemplars from the category distribution.
+Thus, exemplars generated into novel classes tend to obey the same distributional properties of known categories.
+
+As in the implementation reported by \cite{jern2013probabilistic}, we assume that each category's exemplars are distributed according to a multivariate normal distribution with parameters ($\mu, \Sigma$).
 To obtain a conjugate model, we assume that the domain of categories is inverse-Wishart distributed.
-Space does not allow for a full derivation of this model, and so we will simply report how generated category parameters ($\mu_B, \Sigma_B$) are inferred.
+Space does not allow for a full derivation of this model, and so we will simply report how the target category parameters ($\mu_B, \Sigma_B$) are inferred.
 
-% do we need to describe the implementation?
+The \citet{jern2013probabilistic} implementation does not assume anything about the central tendency of categories, and in their model generated $\mu_B$ are sampled uniformly across all possible values.
+Thus, we infer $\mu_B$ with:
+
+\begin{equation}
+  \mu_B = \dfrac
+    {\kappa\mu_{0} + n_B \bar{x_B}}
+    {\kappa + n_B}
+    \label{eq:hs_mus}
+\end{equation}
+% 
+where $\mu_0$ is the prior on $\mu_B$.
+Because \citet{jern2013probabilistic} sample $\mu_B$ uniformly, we set $\mu_0$ to the center of the domain space.
+$\kappa$ ($>0$) is a free parameter, weighting the importance of $\mu_0$.
+$n_B$ and $\bar{x_B}$ are the target category sample size and mean, respectively.In the case where the target class is populated (i.e., $n_B > 0$), $\mu_{B}$ ends up lying somewhere between $\mu_{0}$ and $\bar{x_y}$, depending on $\kappa$ and $n_B$.
+When the target class has no members, Equation \ref{eq:hs_mus} reduces to $\mu_{B} = \mu_{0}$.
+Because we set $\mu_0$ to the center of the space, this outcome is the same as if we had integrated over all possible $\mu_0$.
+
+Unlike $\mu_B$, $\Sigma_B$ cannot be computed considering only the members of the target category.
+Instead, $\Sigma_B$ is influenced both by the distribution of the members of the target category as well as members of other categories through the domain distribution, $\Sigma_D$.
+In our implementation, $\Sigma_D$ is inferred based on the observed (empirical)  covariance matrices of the known categories in the domain, $C$.
+We assume these covariance matrices to be Wishart-distributed, and so $\Sigma_D$ can be computed as:
+
+\begin{equation}
+    \Sigma_D = \Sigma_0 + \sum_{i}{C_i}
+\end{equation}
+%
+where $\Sigma_{0}$ is a $k$-by-$k$ prior covariance matrix.
+We use a $k$-dimensional identity matrix $I_k$ multiplied element-wise against a free parameter, $\rho$, controlling the amount of variance assumed by the prior, $\Sigma_0 =  I_k\rho$.
+Thus, categories are assumed to have some degree of variance along each feature (specified by $\rho$), but are not assumed to possess feature-feature correlations.
+
+Finally, assuming $(\mu_B, \Sigma_B)$ are Normal-Inverse-Wishart distributed, $\Sigma_B$ can be computed as:
+
+\begin{equation}
+  \Sigma_B = [\Sigma_D \nu + C_B +
+    \dfrac
+    {\kappa n_B}
+    {\kappa + n_B}
+    (\bar{x_B}-\mu_B)(\bar{x_B}-\mu_B)^T
+  ] (\nu + n_B)^{-1}
+  \label{eq:Sigma_B}
+\end{equation}
+%
+where $\nu$ ($>k-1$) is an additional free parameter weighting the importance of $\Sigma_{D}$. When the target category has no members, Equation \ref{eq:Sigma_B} reduces to $\Sigma_B = \Sigma_D$.
+
+Generated exemplars are then assumed to be drawn from the multivariate normal distribution specified by $(\mu_{B}, \Sigma_{B})$. Thus, $p(y)$ can be computed as:
+
+\begin{equation}
+  p(y) = \dfrac
+    {exp( \theta \cdot Normal(y, \mu_{B}, \Sigma_{B}))}
+    {\sum_i exp( \theta \cdot Normal(y_i, \mu_{B}, \Sigma_{B}))} 
+\end{equation}
+\
+where $\theta$ is a response determinism parameter.
 
@@
 
-The PACKER model predicts
 The behavioral experiment described below was designed to test whether the \textit{location} of contrast categories (as opposed to their structure), influences generation.

### 2017-01-25 Manuscripts/cogsci-2017/paper.tex
update to figs

@@
 The core experimental phenomenon is that people generate categories possessing the same distributional properties as known categories.
-For example, in a canonical study \citet{ward1994structured} asked participants to draw and describe novel species of plants and animals that might exist on other planets.
+For example, in a canonical study \citet{ward1994structured} asked participants to draw and describe species of plants and animals that might exist on other planets.
 Generation was strongly constrained by prior knowledge of Earth plants and animals: people generated alien species using the same structural forms found on Earth (e.g., eyes, legs, wings) and possessing the same feature correlations observed on Earth (e.g., feathers co-occur with wings).
@@
 
-Similar results were obtained more recently in a study conducted within a traditional categorization paradigm.
+Similar results were obtained in a more recent study conducted within a traditional categorization paradigm.
 \citet{jern2013probabilistic} trained participants on a set of experimenter-defined categories composed of exemplars within an artificial three-dimensional domain (e.g., 2D shapes varying in size, hue, and saturation).
@@
 
-In the sections below, we formally describe the PACKER model and explore its predictions in a behavioral experiment and simulations.
-Rather than formally evaluate PACKER on the benchmarks reviewed above, however, we instead focus on testing predictions made by PACKER that distinguish the model from previous approaches.
+In the sections below, we formally describe the PACKER model and explore its predictions in a behavioral experiment.
+Rather than evaluate PACKER on the benchmarks reviewed above, however, we instead focus on testing predictions made by PACKER that distinguish it from previous approaches.
 
@@
 
-% Note something about minkowski metric? Or can we assume people know that.
-
 When prompted to generate a new example, the model considers both the similarity to examples from other categories as well as the similarity to examples in the target category.
-More formally, the aggregated similarity $a$ between generation candidate $y$ and the model's stored exemplars $x$ can be computed as:
+The aggregated similarity $a$ between generation candidate $y$ and the model's stored exemplars $x$ can be computed as:
 \begin{equation}
@@
 Although $\hat{f_j}$ may be set arbitrarily, in PACKER it is set according to class assignment.
-For known members of contrast categories, $\hat{f_j} = \phi$.
-For known members of the target category, $\hat{f_j} = \gamma$.
+$\hat{f_j} = \phi$ for known members of contrast categories, $\hat{f_j} = \gamma$ for members of the target category.
 $\phi$ and $\gamma$ are free parameters ($-\infty \leq \phi, \gamma \leq \infty$) controlling the contribution of target- and contrast-category similarity to generation.
@@
 The probability that a given candidate $y$ will be generated is evaluated using the \citet{luce1977choice} choice axiom.
-Candidates with larger aggregate similarity are more likely to be generated compared to candidates with smaller aggregate similarity:
+Candidates with greater values of $a$ are more likely to be generated than candidates with smaller values:
 \begin{equation}
@@
 The PACKER model described above explains exemplar and category generation under a traditional exemplar-similarity approach, proposing that people generate categories by maximizing within-category similarity and minimizing between-category similarity.
-These core design principles yield unique predictions about the generation.
-Not only does the model lack mechanisms for copying knowledge of existing categories, but it also specifies precisely how generated classes should \textit{differ} from existing ones.
-Thus, unlike previous accounts, distributional properties of known categories affect generation only insofar as they constrain inter-exemplar similarities.
-Instead, the \textit{location} of known categories in the domain is crucial because it constrains remaining possible locations for new categories.
+The underlying processes assumed by PACKER are highly similar to those in the GCM \citep{nosofsky1984choice}, with the only alteration being that the model's response is a function of the $a$ metric, which is the aggregate of potentially counteracting similarities (positive or negative values), rather than the relative summed similarity assumed in the original model.
+
+In a later section, we will explore the unique predictions yielded by these design principles. First, however, we contrast PACKER with existing approaches in the field.
 
@@
 
-As noted above, existing accounts of category generation focus primarily on explaining the phenomenon that generated categories tend to possess distributional similarities with existing categories.
+As noted in the introduction, existing accounts of category generation focus primarily on explaining the phenomenon that generated categories tend to possess distributional similarities with existing categories.
 \citet{jern2013probabilistic} were the first (as far as we know) to conduct formal simulations on category generation.
@@
 $g = 1$ when $s(y,z) < \tau$ and $g = 0$ otherwise.
-\citet{jern2013probabilistic} did not implement any similarity threshold mechanism, thus, their model is a special case when $\tau = 1$ (any exemplar may be selected).
+Thus, exemplars exceeding the similarity threshold will not be generated, but exemplars just under the threshold are most likely.
+\citet{jern2013probabilistic} did not implement any similarity threshold mechanism, and so their model is the special case where $\tau = 1$ (any exemplar may be selected).
 
@@
 % 
-Thus, the probability that an example $y$ will be selected is a function of its similarity to each of the possible sources, as well as the probability each source will be retrieved.
+Thus, the probability that an example $y$ will be selected is a function of its generation probability given each of the sources, and the probability each source will be retrieved.
 
@@
 \citet{jern2013probabilistic} reported several experiments revealing systematic flaws in the copy-and-tweak account.
-Instead, their results supported a hierarchical sampling model, in where categories are represented as probability distributions in a multidimensional space, with each stimulus feature corresponding to a dimension of the space.
+Instead, their results supported a hierarchical sampling model, in where categories are represented as probability distributions in a $k$-dimensional feature space.
 Observed exemplars (represented as points in this space) are viewed as samples from their underlying category distribution (e.g., a multivariate normal distribution).
@@
 $\kappa$ ($>0$) is a free parameter, weighting the importance of $\mu_0$.
-$n_B$ and $\bar{x_B}$ are the target category sample size and mean, respectively.In the case where the target class is populated (i.e., $n_B > 0$), $\mu_{B}$ ends up lying somewhere between $\mu_{0}$ and $\bar{x_y}$, depending on $\kappa$ and $n_B$.
+$n_B$ and $\bar{x_B}$ are the target category sample size and mean, respectively.
+In the case where the target class is populated (i.e., $n_B > 0$), $\mu_{B}$ ends up lying somewhere between $\mu_{0}$ and $\bar{x_y}$, depending on $\kappa$ and $n_B$.
 When the target class has no members, Equation \ref{eq:hs_mus} reduces to $\mu_{B} = \mu_{0}$.
@@
 
-The behavioral experiment described below was designed to test whether the \textit{location} of contrast categories (as opposed to their structure), influences generation.
+Whereas the copy-and-tweak and hierarchical sampling models both assume explicit mechanisms to re-purpose knowledge about known categories, the PACKER model forgoes these mechanisms entirely.
+Thus, whereas the distributional structure of observed categories will be the primary determinant of generation in the copy-and-tweak and hierarchical sampling models, these properties should only affect PACKER's generation insofar as they constrain inter-exemplar similarities.
+Instead, the \textit{location} of known categories in the domain is crucial because it constrains remaining possible locations for new categories.
+
+The behavioral experiment described below was designed to test this key prediction:  does the \textit{location} of contrast categories (as opposed to their structure), influence generation?
 The experiment follows the paradigm developed by \citet{jern2013probabilistic}: first, participants are exposed to members of a known category (`Alpha', or `A'), and are then asked to generate exemplars belonging to a new category (`Beta', or `B').
@@
 We developed two Alpha categories (see Figure \ref{fig:middle-bottom-conditions}), instantiating the two conditions of the experiment.
-In both conditions, members of the Alpha category are tightly clustered, with equal variance on both features and no correlation between features.
-Our manipulation is fairly `weak': the conditions have only a slight difference in the Y-Axis position of the Alpha category.
+In both conditions, members of the Alpha category are tightly clustered, with equal feature variance and no feature correlations.
+Our manipulation is fairly `weak': the conditions have only a slight difference in the y-axis position of the Alpha category.
 In the `Middle' condition the Alpha category is placed in the center of the space, in the `Bottom' condition the Alpha category is placed in the bottom-center of the space.
@@
 Although our manipulation is minimal, the PACKER model is capable of predicting strong between-condition differences.
-The PACKER model proposes that the nature of the space not occupied by the Alpha category determines where members of the Beta category are likely to be generated.
+PACKER proposes that the nature of the space not occupied by the Alpha category determines where members of the Beta category are likely to be generated.
 Thus, the lower areas of the stimulus space should be less frequently used for generation in the Bottom condition compared to the Middle (as these areas possess greater similarity to the Bottom Alpha category).
@@
 More generally, the PACKER model proposes that the probability a stimulus $y$ will be generated is a function of its similarity to contrast categories \textit{and} to members of the target category.
-Two more general predictions (not specific to either condition) follow from this proposal: (1) the location of beta examples should be positively related to distance from the Alpha category, (2) Beta examples should be more similar to one another than they are to members of the Alpha category.
-
-% THIS IS WHERE NOLAN STOPPED EDITING AFTER THE INTRO REFORMULATION
+Two more general predictions (not specific to either condition) follow from this proposal: (1) the location of Beta examples should be positively related to distance from the Alpha category, and (2) Beta examples should be more similar to one another than they are to members of the Alpha category.
 
@@
 
-Participants began the experiment with a short training phase (3 blocks of 4 trials), where they iteratively observed exemplars belonging to the `Alpha' category.
+Participants began the experiment with a short training phase (3 blocks of 4 trials), where they observed exemplars belonging to the `Alpha' category.
 Participants were instructed to learn as much as they can about the Alpha category, and that they would answer a series of test questions afterwards.
-On each trial, a single exemplar from the Alpha category was presented, and participants were given as much time as they desired before moving on.
+On each trial, a single Alpha category exemplar was presented, and participants were given as much time as they desired before moving on.
+Exemplars were randomly ordered within each block.
 Participants were shown a preview of the range of possible colors and sizes prior to training.
@@
 
-Because the conditions differ only in their location along the vertical axis (see Figure \ref{fig:middle-bottom-conditions}), our main interest is in the generation of Beta examples above and below the contrast category.
-A highly detailed depiction of our results is shown in Figure \ref{fig:middle-bottom-yranges}, wherein each participant's generated category is depicted in terms of its use of the vertical axis.
-As is evident in Figure \ref{fig:middle-bottom-yranges}, we observed a great deal of individual differences in generation strategy: whereas some participants generated all four Beta examples within a narrow range of the Y axis, others generated Beta examples along a fairly wide range.
+Because the conditions differ only in their location along the y-axis (see Figure \ref{fig:middle-bottom-conditions}), our main interest is in the generation of Beta examples above and below the contrast category.
+Several sample Beta categories are depicted in Figure \textbf{???????}.
+As is evident in Figure \textbf{???????}, we observed a great deal of individual differences in generation strategy: whereas some participants generated all four Beta examples within a narrow range, others generated Beta examples along a fairly wide range.
 
-\begin{figure*}
+\begin{figure}
     \begin{center}
-    \input{figs/middle-bottom-yranges.pgf}
+    \inputpgf{figs/}{distance.figs.pgf}
     \caption{Behavioral results.
-Each line shows the minimum and maximum value of a generated category along the Y (vertical) axis.
-Dots along each line represent the positions of individual exemplars in the category, and each participant's category is shown on a separate line.
-Participants are sorted by overall Y axis range, and then by condition.}
-    \label{fig:middle-bottom-yranges}
+\textit{Top}: Frequency of exemplar generation as a function of distance from the center of the Alpha category.
+\textit{Bottom}: Relation between average within- and Between-category distance for every participant.
+}
+    \label{fig:distance.figs}
     \end{center}
-\end{figure*}
+\end{figure}
 

### 2017-01-25 Experiments/middle-bottom/analysis/plot.distance.figs.py

@@ -1,0 +1,105 @@
+import sqlite3, sys
+import pandas as pd
+import numpy as np
+import matplotlib.pyplot as plt
+import seaborn as sns
+sns.set_style("whitegrid")
+
+sys.path.insert(0, "../../../Modules/") # generate-categories/Modules
+import utils
+
+# import data
+con = sqlite3.connect('../data/experiment.db')
+info = pd.read_sql_query("SELECT participant, condition from participants", con)
+generation = pd.read_sql_query("SELECT * from generation", con)
+alphas = pd.read_sql_query("SELECT * from alphas", con)
+stimuli = pd.read_sql_query("SELECT * from stimuli", con).as_matrix()
+stats = pd.read_sql_query("SELECT * from betastats", con)
+con.close()
+
+stats = pd.merge(stats, info, on = 'participant')
+generation = pd.merge(generation, info, on = 'participant')
+
+ngenerations = pd.DataFrame(dict(
+	condition = [],
+	stimulus = [],
+	count = []
+))
+
+for c in pd.unique(info.condition):
+	for i in range(stimuli.shape[0]):
+		count = sum((generation.condition == c) & (generation.stimulus ==i))
+		row = dict(condition = c, stimulus = i, count = count)
+		ngenerations = ngenerations.append(row, ignore_index = True)
+
+
+
+fh, ax = plt.subplots(2,1,figsize = (3.5,5.5))
+condition_styles = dict(Middle = 'r-o', Bottom = 'b-s')
+
+h = ax[0]
+for i, (c, rows) in enumerate(ngenerations.groupby('condition')):
+
+	As = stimuli[alphas[c],:]
+	D = utils.pdist(stimuli, As)
+	D = np.mean(D, axis = 1)
+	x = np.unique(D)
+	y = []
+	for j in x:
+		nums = np.where(D == j)[0]
+		curr_rows = rows.loc[rows.stimulus.isin(nums)]
+		counts = curr_rows['count'].as_matrix()
+		y.append(np.mean(counts))
+	print y
+
+	x = x - min(x)
+	x = x / max(x)
+	h.plot(x, y, condition_styles[c], alpha = 0.7, label = c)
+
+h.xaxis.grid(False)
+h.set_xticks([])
+h.legend(loc = 'upper left', frameon = True, framealpha = 1)
+
+
[…]

### 2017-01-25 Manuscripts/cogsci-2017/paper.tex
work on simulations section

@@
 
-As in the GCM, the similarity between two examples, $s(x_i, x_j)$, is defined as an inverse-exponential function of distance:
+As in the GCM, the similarity between two examples, $s(x_i, x_j)$, is an inverse-exponential function of distance:
 \begin{equation}
@@
 
-In a later section, we will explore the unique predictions yielded by these design principles. First, however, we contrast PACKER with existing approaches in the field.
+In later sections, we will explore the unique predictions yielded by these design principles. First, however, we contrast PACKER with existing approaches in the field.
 
@@
 As in Equation \ref{eq:packer-ss}, $\hat{f}_j$ may be set arbitrarily.
-However, in our simulations, $\hat{f}_z = -\gamma$ for members of contrast categories, and $\hat{f}_z = \gamma$ for members of the target category.
-The resulting free parameter $\gamma$ ($>0$) thus controls the relative chances that a member of the target category will be retrieved as the source.
-\citet{jern2013probabilistic} did not implement any mechanisms to differentially weigh retrieval probabilities, and so our model encompasses theirs as a special case when $\gamma = 0$ (uniform probabilities).
+However, in our simulations, $\hat{f}_z = -\lambda$ for members of contrast categories, and $\hat{f}_z = \lambda$ for members of the target category.
+The resulting free parameter $\lambda$ ($>0$) thus controls the relative chances that a member of the target category will be retrieved as the source.
+\citet{jern2013probabilistic} did not implement any mechanisms to differentially weigh retrieval probabilities, and so our model encompasses theirs as a special case when $\lambda = 0$ (uniform probabilities).
 
@@
 
+\begin{figure}[ht!]
+    \begin{center}
+    \inputpgf{figs/}{beta.samples.pgf}
+    \caption{Sample generated categories. }
+    \label{fig:beta.samples}
+    \end{center}
+\end{figure}
+
 \subsection{Procedure}
@@
 Because the conditions differ only in their location along the y-axis (see Figure \ref{fig:middle-bottom-conditions}), our main interest is in the generation of Beta examples above and below the contrast category.
-Several sample Beta categories are depicted in Figure \textbf{???????}.
-As is evident in Figure \textbf{???????}, we observed a great deal of individual differences in generation strategy: whereas some participants generated all four Beta examples within a narrow range, others generated Beta examples along a fairly wide range.
+Several sample Beta categories are depicted in Figure \ref{fig:beta.samples}.
+As is evident in Figure \ref{fig:beta.samples}, we observed a great deal of individual differences in generation strategy: whereas some participants generated all four Beta examples within a narrow y-axis range, others generated Beta examples along a wide range.
 
-\begin{figure}
-    \begin{center}
-    \inputpgf{figs/}{distance.figs.pgf}
-    \caption{Behavioral results.
-\textit{Top}: Frequency of exemplar generation as a function of distance from the center of the Alpha category.
-\textit{Bottom}: Relation between average within- and Between-category distance for every participant.
-}
-    \label{fig:distance.figs}
-    \end{center}
-\end{figure}
-
-To more specifically evaluate the key predictions made by PACKER, we determined the number of participants in each condition who placed at least one Beta example within the top `rows' of the space (one of the top two possible Y axis values), as well as the number of participants who placed at least one example in one of the bottom rows.
-The resulting contingency tables data are shown in Table \ref{table:subset-table}.
+To more specifically evaluate the key predictions made by PACKER, we determined the number of participants in each condition who placed at least one Beta example on the top `row' of the space (the maximum possible y-axis value), as well as the number of participants who placed at least one example on the bottom row.
+The resulting contingencies data are shown in Table \ref{table:subset-table}.
+Fisher's Exact Tests on these data reveal that a greater number of participants assigned to the Middle condition generated a Beta example in the bottom row of the space, $p < 0.001$, but the conditions did not differ in use of the top of the space, $p = 0.16$.
+A greater number of participants in the Middle condition placed Beta examples in the top rows \textit{and} the bottom rows, $p = 0.038$.
 
@@
 \begin{tabular}{ l r r}
-    \textbf{Middle}         & Used top rows & No top rows \\
+    \textbf{Middle}         & Used top row & No top row \\
     \hline
-    Used bottom rows        &  31 & 18  \\
-    No bottom rows          &  11 &  1  \\
+    Used bottom rows       &  28 & 18  \\
+    No bottom row          &  11 &  4  \\
     \\
-    \textbf{Bottom}         & Used top rows & No top rows \\
+    \textbf{Bottom}         & Used top row & No top row \\
     \hline
-    Used bottom rows        & 22 & 8 \\
-    No bottom rows          & 29 & 2 \\
+    Used bottom row        & 16 & 8 \\
+    No bottom row          & 31 & 6 \\
 \end{tabular}
@@
 
-Fisher's Exact Tests reveal that a greater number of participants assigned to the Middle condition generated a Beta example in the bottom rows of the space, $p < 0.001$.
-Participants in the Bottom condition were more marginally more likely to generate a Beta example in the top rows of the space, $p = 0.088$.
-The conditions did not differ in the number of participants who placed Beta examples in the top rows \textit{and} the bottom rows, $p = 0.14$.
-Finally, a greater number of Middle participants generated all four Beta examples in the bottom rows of the space, $p = 0.003$, and a greater number of Bottom participants generated all four examples in the top rows of the space, $p = 0.017$.
+The top panel of Figure \ref{fig:distance.figs} displays the stimulus generation frequency as a function of distance from the center of the Alpha category.
+There was a strong preference for stimuli that are dissimilar to the members of the Alpha category: maximally distant items were by far the most frequently generated.
+The bottom panel of Figure \ref{fig:distance.figs} depicts for each participant the distance between exemplars, within- and between-categories.
+As observed by \citet{ward1994structured}, most people generated Beta categories in which members are closer to one another than they are to members of the Alpha category.
+We did however, observe a notable subset of individuals with greater within-class distance.
+Inspection of these individuals reveals that they tended to adopted a `corners' approach, in which a Beta examples were placed almost exclusively in the furthest corners of the space, thereby producing more within-class distance.
 
-These results are broadly supportive of PACKER's predictions, and are not predicted by existing accounts \citep[i.e.,][]{jern2013probabilistic}.
-However, the results described above are somewhat commonsense: they simply demonstrate that Beta examples tend to be generated in areas not occupied by the Alphas.
-Beyond these effects, we observed a great deal of differences in generation strategy.
-Figure \ref{fig:range-diff-gradient} depicts the average difference in range between the two features (e.g., horizontal -- vertical) across the participant-generated categories, with respect to the location of each category's members.
-These data reveal highly systematic patterns of generation: whereas many participants generated categories with more vertical range (i.e., `column'-like categories), others generated categories with horizontal range (`row'-like categories).
-Furthermore, these two different types of categories appear in distinct locations of the domain space.
-Whereas vertically aligned categories more often appear to the left or right of the Alpha class, horizontally aligned categories appear above and below the Alpha class.
-Thus, even beyond generating Betas in locations not occupied by the Alphas, participants appear to modify the internal structure of their categories in order to maximize distance from the Alphas.
+\begin{figure}
+    \begin{center}
+    \inputpgf{figs/}{distance.figs.pgf}
+    \caption{Behavioral results.
+\textit{Top}: Frequency of exemplar generation as a function of distance from the center of the Alpha category.
+\textit{Bottom}: Relation between average within- and between-category distance for every participant.
+}
+    \label{fig:distance.figs}
+    \end{center}
+\end{figure}
 
-\begin{figure}[ht!]
+\subsection{Summary}
+
+Our results are broadly supportive of PACKER's predictions: people prefer to generate items that are maximally dissimilar from members of the Alpha category, and they tended to generate categories possessing greater within-class similarity than between-class similarity (with some notable exceptions, see Figure \ref{fig:distance.figs}).
+By consequence, we observed considerable differences in generation between the Middle and Bottom conditions: participants in the Bottom condition were less likely to use the bottom row of the stimulus space for generation, and participants in the Middle condition were more likely to create categories spanning the entire y-axis (utilizing the top and bottom row of the space).
+This latter result is especially interesting, as it shows that qualitatively different types of categories are generated, depending on only the location of the Alphas in the domain.
+
+The results described above are somewhat commonsense: they simply demonstrate that the location (rather than distributional structure) of existing categories imposes constraints on generation because people tend to generate examples in areas not occupied by existing categories.
+This principle, however, is not predicted by existing models of generation -- these models are instead designed to explain distributional correspondences between generated and existing categories.
+
+\section{Simulations}
+
+We conducted formal simulations to our data using the PACKER, copy-and-tweak, and hierarchical sampling models described above.
+In order to obtain an overall sense of each model's ability to explain our results, we fitted each model trial-wise, using a hill-climbing algorithm to maximize the log-likelihood of the model's predictions against our results.
+For each trial, models were initialized with the participant's Alpha category (Middle or Bottom), and the configuration of the participant's Beta category during that trial.
+
+Four free parameters were fitted in each model.
+The $c$, $\phi$, $\gamma$, and $\theta$ parameters were fitted for PACKER; $c$, $\lambda$, $\tau$, and $\theta$ were fitted for the copy-and-tweak model, and $\kappa$, $\rho$, $\nu$, and $\theta$ were fitted for the hierarchical sampling model.
+Note that each model possess as $\theta$ parameter fulfilling the same role (response determinism), and PACKER shares with copy-and-tweak a specificity parameter, $c$.
+Parameters were not allowed to vary between participants or conditions -- the goal was to obtain the best-fitting values to our entire dataset.
+
+\begin{table}
+\centering
+\caption{Model-fitting results.}
+\label{table:model-fits}
+\begin{tabular}{ l l l}
+\\
+ \textbf{PACKER} & \textbf{Copy \& Tweak} & \textbf{Hierarchical} \\
+ & & \textbf{Sampling} \\ \hline
+ $L = -1733$ & $L = -2040$ & $L = -2055$  \\ 
+ $c=0.56$ & $c=9.445$  & $\kappa<0.001$\\
+$\phi=-21.2$& $\lambda=15.115$ & $\rho=1.001$ \\ 
+$\gamma=18.707$  & $\tau=0.415$ & $\nu=0.451$  \\ 
+ $\theta=0.166$  & $\theta=7.068$ & $\theta=6.137$  \\ 
+\end{tabular}
+\end{table}
+
+Each model's best-fitting parameterization is shown in Table \ref{table:model-fits}.
+Overall, PACKER outperformed copy-and-tweak and the hierarchical sampling model by a considerable margin.
+The parameter settings associated with PACKER's best fit are exactly as expected: a strong preference for items that are similar to members of the target category but are dissimilar to members of the contrast category.
+A similar pattern of results was obtained when each participant's first trial was excluded (i.e., when the target category has no members).
+
+\subsection{Category structure \& location}
+
+\begin{figure}
     \begin{center}
@@
 
-\subsection{Summary}
-
-Our behavioral experiment revealed strong effects of the role of category location on generation. Not only are participants more likely to generate Betas in locations
-
-\section{Simulations}
-
-We conducted formal simulations to our data using the PACKER model described above, as well as an exemplar-based \textit{copy-and-tweak} model, and conjugate implementation of the hierarchical sampling model proposed by \citet{jern2013probabilistic}.
-Below, we briefly describe the formal basis of the copy-and-tweak and hierarchical sampling models.
-
-\subsection{Model-Fitting}
-In order to obtain an overall sense of each model's ability to explain our results, we fitted each model trial-wise, maximizing the log-likelihood of the its predictions against our results.
-For each trial, models were initialized with the participant's Alpha category (Middle or Bottom), and the configuration of the participant's Beta category during that trial.
-Four parameters were fitted for the PACKER model: $c$, $\phi$, $\gamma$, and $\theta$.
-Likewise, four parameters were fitted to the \cite{jern2013probabilistic} model: $\kappa$, $\gamma$, $\nu$, and $\theta$.
-Parameters were not allowed to vary between participants or conditions -- the goal was to obtain the best-fit to our entire dataset.
+% Beyond these effects, we observed a great deal of differences in generation strategy.
+Figure \ref{fig:range-diff-gradient} depicts the average difference in range between the two features (e.g., horizontal -- vertical) across the participant-generated categories, with respect to the location of each category's members.
+These data reveal highly systematic patterns of generation: whereas many participants generated categories with more vertical range (i.e., `column'-like categories), others generated categories with horizontal range (`row'-like categories).
+Furthermore, these two different types of categories appear in distinct locations of the domain space.
+Whereas vertically aligned categories more often appear to the left or right of the Alpha class, horizontally aligned categories appear above and below the Alpha class.
+Thus, even beyond generating Betas in locations not occupied by the Alphas, participants appear to modify the internal structure of their categories in order to maximize distance from the Alphas.
 

### 2017-01-25 Experiments/middle-bottom/analysis/plot.beta.samples.py

@@ -1,0 +1,52 @@
+import sqlite3
+import pandas as pd
+import numpy as np
+import matplotlib.pyplot as plt
+
+import sys
+sys.path.insert(0, "../../../Modules/") # generate-categories/Modules
+import utils
+
+con = sqlite3.connect('../data/experiment.db')
+participants = pd.read_sql_query("SELECT * from participants", con)
+generation = pd.read_sql_query("SELECT * from generation", con)
+alphas = pd.read_sql_query("SELECT * from alphas", con)
+stimuli = pd.read_sql_query("SELECT * from stimuli", con).as_matrix()
+con.close()
+
+samples = dict(
+	Middle = [
+		1, # 4 corner
+		7, # bottom row
+		68, # side column / corners
+		151, # bottom cluster
+	],
+	Bottom = [
+		11, # 4 corner
+		91, # top row / corners
+		106, # side column
+		117, # top cluster
+	]
+)
+
+fh, ax = plt.subplots(2,4, figsize = [3.2, 1.6])
+ax = ax.flat
+plotnum = 0
+for k, v in samples.items():
+	for pid in v:
+		h = ax[plotnum]
+
+		A = alphas[k]
+		B = generation.loc[generation.participant == pid, 'stimulus']
+		utils.plotclasses(h, stimuli, A, B, textsettings = dict(fontsize = 9))
+		plotnum += 1
+		
+
+plt.tight_layout(pad=-0.0, w_pad=-0.0)
+fh.savefig('beta.samples.png', bbox_inches = 'tight')
+
+import os, matplotlib
+os.environ["PATH"] += os.pathsep + '/Library/TeX/texbin/'
+opts = {'pgf.texsystem': 'pdflatex'}
+matplotlib.rcParams.update(opts)
+fh.savefig('../../../Manuscripts/cogsci-2017/figs/beta.samples.pgf', bbox_inches='tight', pad_inches=0.0)

### 2017-01-26 Manuscripts/cogsci-2017/paper.tex
draft finished [for now.....]

@@
 
-% % Should we bother reporting this? we never even looked at the data.
-% Following generation participants completed a generalization phase where they were asked to classify 81 stimuli sampled across the domain space (along a 9x9 grid).
-On each trial, a single stimulus was presented and participants were asked to classify the item into either the Alpha or Beta category.
-The members of the Alpha and Beta categories were presented during the generalization phase, but feedback was not provided on responses.
-
 \subsection{Results}
@@
 
-The top panel of Figure \ref{fig:distance.figs} displays the stimulus generation frequency as a function of distance from the center of the Alpha category.
-There was a strong preference for stimuli that are dissimilar to the members of the Alpha category: maximally distant items were by far the most frequently generated.
-The bottom panel of Figure \ref{fig:distance.figs} depicts for each participant the distance between exemplars, within- and between-categories.
-As observed by \citet{ward1994structured}, most people generated Beta categories in which members are closer to one another than they are to members of the Alpha category.
+To evaluate PACKER's other predictions, we computed the number of times each unique exemplar was generated, and plotted those data as a function of distance to the center of the Alpha category.
+These data, depicted in the top panel of Figure \ref{fig:distance.figs} reveal a strong preference for stimuli that are dissimilar to the members of the Alpha category: maximally distant items were by far the most frequently generated.
+
+Finally, we computed for each participant the average distance between exemplars belonging to the same and opposite categories.
+These data (depicted in the bottom panel of Figure \ref{fig:distance.figs}) show that, as observed by \citet{ward1994structured}, most people generated Beta categories in which members are closer to one another than they are to members of the Alpha category (i.e., more between- than within-category distance).
 We did however, observe a notable subset of individuals with greater within-class distance.
@@
     \caption{Behavioral results.
-\textit{Top}: Frequency of exemplar generation as a function of distance from the center of the Alpha category.
-\textit{Bottom}: Relation between average within- and between-category distance for every participant.
+\textit{Top}: Frequency of exemplar generation as a function of distance from the Alpha category.
+\textit{Bottom}: Within- vs.
+between-category distance for every participant.
 }
@@
 Note that each model possess as $\theta$ parameter fulfilling the same role (response determinism), and PACKER shares with copy-and-tweak a specificity parameter, $c$.
+Attention in PACKER and copy-and-tweak was set uniformly.
 Parameters were not allowed to vary between participants or conditions -- the goal was to obtain the best-fitting values to our entire dataset.
 
-\begin{table}
+Each model's best-fitting parameterization is shown in Table \ref{table:model-fits}.
+Overall, PACKER outperformed copy-and-tweak and the hierarchical sampling model by a considerable margin ($\sim300$ log-likelihood).
+The parameter settings associated with PACKER's best fit are exactly as expected: a strong preference for items that are similar to members of the target category but are dissimilar to members of the contrast category.
+A similar pattern of results was obtained when each participant's first trial was excluded (i.e., when the target category had no members).
+
+% JOE! I do not know how log-likes are supposed to be notated!
+
+\begin{table}[ht!]
 \centering
@@
  $L = -1733$ & $L = -2040$ & $L = -2055$  \\ 
- $c=0.56$ & $c=9.445$  & $\kappa<0.001$\\
-$\phi=-21.2$& $\lambda=15.115$ & $\rho=1.001$ \\ 
-$\gamma=18.707$  & $\tau=0.415$ & $\nu=0.451$  \\ 
- $\theta=0.166$  & $\theta=7.068$ & $\theta=6.137$  \\ 
+ $c=0.564$ & $c=9.445$  & $\kappa<0.001$\\
+$\phi=-13.608$& $\lambda=15.115$ & $\rho=1.001$ \\ 
+$\gamma=12.006$  & $\tau=0.415$ & $\nu=0.451$  \\ 
+ $\theta=0.259$  & $\theta=7.068$ & $\theta=6.137$  \\ 
 \end{tabular}
@@
 
-Each model's best-fitting parameterization is shown in Table \ref{table:model-fits}.
-Overall, PACKER outperformed copy-and-tweak and the hierarchical sampling model by a considerable margin.
-The parameter settings associated with PACKER's best fit are exactly as expected: a strong preference for items that are similar to members of the target category but are dissimilar to members of the contrast category.
-A similar pattern of results was obtained when each participant's first trial was excluded (i.e., when the target category has no members).
+The model-fitting results above follow directly from the design of our experiment.
+Because distributional structure is the primary determinant of generation in the copy-and-tweak and hierarchical sampling models, those models have no basis to predict any differences between Middle and Bottom conditions (which do not differ in distributional structure).
 
-\subsection{Category structure \& location}
+\subsection{Individual Differences}
 
-\begin{figure}
+Our behavioral results showed that, in general, generated categories consist of item that are (A) dissimilar to members of opposite categories, and (B) similar to members of the their own category.
+However, as shown in Figures \ref{fig:beta.samples} and \ref{fig:distance.figs}, we observed a great deal of individual differences in generation strategy.
+For example whereas many participants tightly clustered the Beta category, others distributed its members to the four corners of the domain.
+Likewise, whereas many participants created `column'-like categories (varying along the y-axis but not the x-axis), others created `row'-like categories.
+While it is important that any model be able to explain different types of performance under the same parameterization, these radical differences in generation are difficult to account for under a single set of parameters.
+
+In future work, we intend to explore more systematically the role of PACKER's parameters in different subtypes of performance.
+For example, whereas the tight cluster profile could be explained with a moderately negative $\phi$ and a strongly positive $\gamma$, the four corners profile could be simulated with a strongly negative $\phi$ and $\gamma\leq0$.
+Here, we quickly explore the role of the attention weights ($w$, in Equation \ref{eq:similarity}) in producing row-like or column-like categories.
+
+\begin{figure}[ht!]
     \begin{center}
     \inputpgf{figs/}{range-diff-gradient.pgf}
-    \caption{Generated category structure as a function of location in the domain space.
-Orange areas in each gradient correspond to stimuli that were commonly generated into category possessing greater vertical (Y-Axis) range.
-Purple areas correspond to categories possessing greater horizontal range.
+    \caption{Generated category structure as a function of location in the domain.
+Orange areas in each gradient correspond to stimuli that were commonly generated into category possessing greater y-axis range (columns).
+Purple areas correspond to categories possessing greater x-axis range.
 White areas correspond to equal range along both features (or infrequent generation).}
@@
 
-% Beyond these effects, we observed a great deal of differences in generation strategy.
-Figure \ref{fig:range-diff-gradient} depicts the average difference in range between the two features (e.g., horizontal -- vertical) across the participant-generated categories, with respect to the location of each category's members.
-These data reveal highly systematic patterns of generation: whereas many participants generated categories with more vertical range (i.e., `column'-like categories), others generated categories with horizontal range (`row'-like categories).
-Furthermore, these two different types of categories appear in distinct locations of the domain space.
-Whereas vertically aligned categories more often appear to the left or right of the Alpha class, horizontally aligned categories appear above and below the Alpha class.
-Thus, even beyond generating Betas in locations not occupied by the Alphas, participants appear to modify the internal structure of their categories in order to maximize distance from the Alphas.
+For each stimulus in the domain, we computed the difference in range between the features (e.g., $range(size) - range(color)$) across every generated category in which that stimulus was a member.
+Taking the average of these range differences yields a gradient describing how, on average, categories were distributed for each stimulus.
+These data, depicted in the top panels of Figure \ref{fig:range-diff-gradient} reveal a highly systematic relationship between category structure and category location.
+Whereas column-like categories more often include stimuli to the left or right of the Alpha class, row-like categories appear above and below the Alpha class.
+Thus, even beyond generating Betas in locations not occupied by the Alphas, participants appear to modify the distributional structure of their categories in order to maximize distance from the Alphas.
+
+In order to simulate this finding, we set the attention weight parameters in PACKER and copy-and-tweak separately for each participant.
+To simulate the hierarchical sampling model we set the domain covariance prior, $\Sigma_0$, separately for each participant.
+Each model's free parameters were set as in Table \ref{table:model-fits}.
+While there exist methods to find the optimal attention weights for a given classification \citep[see][]{vanpaemel2012using}, for ease of computation we simply approximated the weights based on the inverse of each feature's range: $w_k \propto range(k)^{-1}$.
+The ranges are not inverted for the hierarchical sampling model.
+Thus, the Alpha and Beta categories are assumed to be distinct along dimensions that the Betas do not vary on.
+We then simulated 50 Beta categories with each participant's weighting scheme to obtain a sense of how the relative importance of each dimension affects what types of categories are generated and where they are generated.
+
+The results, depicted in Figure \ref{fig:range-diff-gradient}, are highly supportive of PACKER's predictions.
+When the x-axis is more strongly weighted, PACKER creates column categories to the sides of the Alphas.
+Conversely, when the y-axis is weighted, PACKER creates row categories above and below the Betas.
+This behavior falls out from the nature of dimensional selective attention: dimensions with more weighting have a sharper similarity gradient.
+For example, when the x-axis is weighted, PACKER attempts to create Beta categories with more within-class similarity (less range), and less between-class similarity along the x-axis, resulting in column-like categories.
+
+The hierarchical sampling and copy-and-tweak models do not as often create categories with strong range differentials.
+However, when they do, the categories tend to show the opposite pattern: columns are placed above and below the Alphas, rows are to the side of the Alphas.
+In the hierarchical sampling model, the position of contrast categories is not considered when generating new items.
+Only the distributional structure of the contrast category and items already in the target category are considered.
+The model simulates row/column categories by assuming there is more x- or y-axis variance in the domain, essentially stretching the generated category distribution along that axis, and shrinking it along the opposite axis.
+However, because the Beta category distribution is centered in the middle of the space prior to the first trial, the model is most likely to generate items directly to the side of the Alphas if the x axis is weighted, or above/below the Alphas if the y-axis is weighted.
+Future generated items will follow from this distribution, and thus weighting the x-axis results row categories that possess the same y-axis position as the Alphas.
+Conversely, weighting the y-axis results in column categories that share the the same x-axis position as the Alphas.
+
+The copy-and-tweak model operates by a similar principle.
+When, for example, the x-axis is weighted, the range of examples exceeding the similarity threshold ($\tau$) becomes much sharper along the x-axis (items need to be much closer to be considered at all similar), and broader along the y-axis.
+On the first trial, one of the Alphas will be used as a source, and because the similarity gradient is so sharp along the x-axis, the model will generate a Beta that is above or below the source.
+Future examples will follow this same pattern, resulting in a column category that differs from the alphas along the y-axis, as opposed to the x-axis.
 
@@
 In this paper, we presented a novel exemplar-based approach to explaining category generation.
+The PACKER model, based on the Generalized Context Model \citep[GCM;][]{nosofsky1984choice}, proposes that categories are represented as a collection of exemplars stored in memory, and that members of generated categories should be similar to one another, yet dissimilar to members of opposing categories.
+
+Previous accounts of category generation \citep[see][]{jern2013probabilistic,ward1995s,ward1994structured} have focused primarily on explaining why generated categories tend to share distributional properties with known categories.
+Here we instead explored whether there are aspects of category generation that can be explained through the basic and well-understood principles of the exemplar framework.
+
+In a behavioral study and subsequent formal modeling, we found broad support for the PACKER model.
+Participants in our study more frequently generated items that are highly distant from members of contrast categories, and they tended to generate categories with more within-class than between-class similarity.
+Likewise, we found that the \textit{location} of contrast categories (rather than their distributional structure) shapes generation by imposing constraints on the areas of space that remain available for a new category.
+Our formal simulations reveal that PACKER provides a strong account of our observations.
+
+The PACKER model is, in general, highly expressive in its performance.
+Under different parameter settings it is capable of generating tightly clustered or highly distributed categories, and adjusting the distribution of categories along each feature (to create row-like or column-like categories; see Figure \ref{fig:beta.samples}).
+Future work will focus on exploring how individual differences in generation can be explained using PACKER, and whether the canonical effects observed by \citet{ward1994structured} and \citet{jern2013probabilistic} can be explained through exemplar-similarity alone.
 

### 2017-01-27 Manuscripts/cogsci-2017/paper.tex
intro, beginning of packer updated. section at 2:30, will start working again at 4pm

@@
 \usepackage[natbibapa]{apacite}
+
+%though some work has recently been conducted using traditional classification learning approaches . 
+%Similar results were obtained in a more recent study conducted within a traditional categorization paradigm.
+
+%%Results such as these have motivated formal accounts of generation that explicitly invoke the idea that people re-purpose existing knowledge to generate something new -- either by copying-and-tweaking an observation retrieved from memory \citep{ward2002role,ward1995s}, or by abstracting the distributional structure of categories in the domain \citep{jern2013probabilistic}.
+
+%Rather than evaluate PACKER on the benchmarks reviewed above, however, we instead focus on testing predictions made by PACKER that distinguish it from previous approaches.
+
+%The attention weights are of interest in explaining individual differences in generation strategy (discussed below), but for our formal simulations they are set uniformly.
 
@@
 One of the most intriguing capabilities of human cognition is the ability to creatively generate new ideas and concepts.
-The creative use of knowledge has, however, only infrequently been the subject of inquiry in the field of categorization.
-To date, most of what we understand about human concept generation comes from foundational studies in the field of creative cognition \citep[e.g.,][]{marsh1999inadvertent,smith1993constraining,ward2002role,ward1994structured,ward1995s}, though some work has recently been conducted using traditional classification learning approaches \citep{jern2013probabilistic}.
+The creative use of knowledge has only been infrequently the subject of formal inquiry.
+This paper focuses on one topic within creative cognition, category generation.
+Foundational work on category generation \citep[e.g.,][]{marsh1999inadvertent,smith1993constraining,ward2002role,ward1994structured,ward1995s} found that people tend to create new categories that are similar to related previously known categories: New exemplars from the new categories tend to be copied from exemplars for a similar categories, but tweaked on one or a few dimensions.
+Thus, people generate a new category containing similar distributional properties to the categories it is based on, while being different from those categories in some manner.
 
 Existing work on the creative use of conceptual knowledge predominantly explores the role of prior knowledge in generating novel concepts.
-The core experimental phenomenon is that people generate categories possessing the same distributional properties as known categories.
+A core experimental phenomenon is that people generate categories possessing the same distributional properties as known categories.
 For example, in a canonical study \citet{ward1994structured} asked participants to draw and describe species of plants and animals that might exist on other planets.
@@
 
-Similar results were obtained in a more recent study conducted within a traditional categorization paradigm.
-\citet{jern2013probabilistic} trained participants on a set of experimenter-defined categories composed of exemplars within an artificial three-dimensional domain (e.g., 2D shapes varying in size, hue, and saturation).
+Recent work has proposed and tested computational models of how new categories are generated based on prior knowledge \citep{jern2013probabilistic}.
+In their experiments, \citet{jern2013probabilistic} trained participants on a set of experimenter-defined categories composed of exemplars within an artificial three-dimensional domain (e.g., 2D shapes varying in size, hue, and saturation).
 After a short training phase, participants were asked to generate exemplars from a new category.
@@
 As in the classic \cite{ward1994structured} experiment, \citet{jern2013probabilistic} found that people generated artificial categories possessing the same feature variance and correlations as the experimenter-defined categories in the domain.
-
-Results such as these have motivated formal accounts of generation that explicitly invoke the idea that people re-purpose existing knowledge to generate something new -- either by copying-and-tweaking an observation retrieved from memory \citep{ward2002role,ward1995s}, or by abstracting the distributional structure of categories in the domain \citep{jern2013probabilistic}.
+They evaluated different possible computational models of how people generate categories (such as "copy-and-tweak" and exemplar accounts) and found that a hierarchical Bayesian model generating new categories by abstracting the distributional structure of the categories in the domain.
 
@@
 In this paper, we introduce a novel exemplar-based approach to category generation, the PACKER model (\textit{Producing Alike and Contrasting Knowledge using Exemplar Representations}), which creates new categories by balancing two constraints: (1) new categories should be different from known categories (minimizing between-class similarity), and (2) new categories should be internally coherent (maximizing within-class similarity).
-As such, PACKER is a significant departure from previous accounts of generation -- rather than proposing that people create new categories by re-purposing existing knowledge, PACKER explains generation exclusively in terms of the well-studied mechanics of exemplar representations and therefore possesses a rich connection to the wider body of research on category learning.
+As such, PACKER is a significant departure from previous accounts of generation -- rather than proposing that people create new categories that are mostly similar to related categories, PACKER creates new categories that are different from related categories.
+Further, it does so using the well-studied mechanics of exemplar representations and therefore possesses a rich connection to the wider body of research on category learning.
 
 In the sections below, we formally describe the PACKER model and explore its predictions in a behavioral experiment.
-Rather than evaluate PACKER on the benchmarks reviewed above, however, we instead focus on testing predictions made by PACKER that distinguish it from previous approaches.
+We compare its performance to a copy-and-tweak and hierarchical Bayesian model by examining their fits to aggregate results and individual differences.
 
@@
 
-The PACKER model is an extension of the highly influential Generalized Context Model of category learning (GCM; \citealp{nosofsky1984choice}).
+The PACKER model is an extension of the prominent Generalized Context Model of category learning (GCM; \citealp{nosofsky1984choice}).
 The model assumes that each category is represented by a collection of exemplars within in a $k$-dimensional psychological space, and that generation is constrained by both similarity to members of the target category (the category in which a stimulus is being generated) as well as similarity to members of other categories.
 
-As in the GCM, the similarity between two examples, $s(x_i, x_j)$, is an inverse-exponential function of distance:
+As in the GCM, the similarity between two examples, $s\left(x_i, x_j\right)$, is an inverse-exponential function of distance:
 \begin{equation}
-  s(x_i,x_j) = exp( -c \sum_{k}{|x_{ik} - x_{jk}|}w_k )
+  s\left(x_i,x_j\right) = \exp \left\{ -c \sum_{k}{ \left| x_{ik} - x_{jk} \right|}w_k \right\}
   \label{eq:similarity}
 \end{equation}
-where $w_k$ is a vector of attention parameters ($\sum_k{w_k} = 1$), weighting to the importance of each feature in the similarity computation, and $c$ ($<0$) is a specificity parameter controlling the spread of exemplar generalization.
-The attention weights are of interest in explaining individual differences in generation strategy (discussed below), but for our formal simulations they are set uniformly.
+where $w_k$ is the attention weighting of dimension $k$ ($\sum_k{w_k} = 1$), accounting for the relative importance of each dimension when calculating similarity, and $c$ ($c>0$) is a specificity parameter controlling the spread of exemplar generalization.
+For simplicity, most simulations will use uniform attention weights.
+We will use non-uniform attention weights in our discussion of individual differences.
+
+%%%%%%%% stop before section at 2:30. will resume after section at 4pm.
 

### 2017-01-28 Manuscripts/cogsci-2017/paper.tex
finished first pass through. headed home for dinner and will cut and work on it some more

@@
 
+%each example's degree of contribution toward generation.
+PACKER sets according $f_j$ .
+$\hat{f_j} = \phi$ for known members of contrast categories, $\hat{f_j} = \gamma$ for members of the target category.
+
 %%Results such as these have motivated formal accounts of generation that explicitly invoke the idea that people re-purpose existing knowledge to generate something new -- either by copying-and-tweaking an observation retrieved from memory \citep{ward2002role,ward1995s}, or by abstracting the distributional structure of categories in the domain \citep{jern2013probabilistic}.
@@
 %The attention weights are of interest in explaining individual differences in generation strategy (discussed below), but for our formal simulations they are set uniformly.
+
+% so that is 1 when specifying whether $y$ exceeds the similarity threshold, $\tau$, which is treated as a free parameter ($0<\tau\leq1$).
+$g = 1$ when $s(y,z) < \tau$ and $g = 0$ otherwise.
+Thus, exemplars exceeding the similarity threshold will not be generated, but exemplars just under the threshold are most likely.
+\citet{jern2013probabilistic} did not implement any similarity threshold mechanism, and so their model is the special case where $\tau = 1$ (any exemplar may be selected).
 
@@
 This paper focuses on one topic within creative cognition, category generation.
-Foundational work on category generation \citep[e.g.,][]{marsh1999inadvertent,smith1993constraining,ward2002role,ward1994structured,ward1995s} found that people tend to create new categories that are similar to related previously known categories: New exemplars from the new categories tend to be copied from exemplars for a similar categories, but tweaked on one or a few dimensions.
+Foundational work on category generation \citep[e.g.,][]{smith1993constraining,ward2002role,ward1994structured,ward1995s} found that people tend to create new categories that are similar to related previously known categories: New exemplars from the new categories tend to be copied from exemplars for a similar categories, but tweaked on one or a few dimensions.
 Thus, people generate a new category containing similar distributional properties to the categories it is based on, while being different from those categories in some manner.
@@
 
+%JLA: Add (a), (b) and (c) to figure and then use those to refer to each subpanel in the caption. It'll save space and be cleaner.
 \begin{figure*}
@@
 \end{equation}
-where $w_k$ is the attention weighting of dimension $k$ ($\sum_k{w_k} = 1$), accounting for the relative importance of each dimension when calculating similarity, and $c$ ($c>0$) is a specificity parameter controlling the spread of exemplar generalization.
+where $w_k$ is the attention weighting of dimension $k$ ($w_k \geq 0$ and $\sum_k{w_k} = 1$), accounting for the relative importance of each dimension when calculating similarity, and $c$ ($c>0$) is a specificity parameter controlling the spread of exemplar generalization.
 For simplicity, most simulations will use uniform attention weights.
@@
 
-%%%%%%%% stop before section at 2:30. will resume after section at 4pm.
+To generate a new example given examples from contrast and target categories, the model considers both the similarity to examples from the contrast category as well as the similarity to examples in the target category.
+The aggregated similarity $a$ between generation candidate $y$ and the model's stored exemplars $x$ can be computed as $a(y, x) = \sum_j{f(x_j) s(y, x_j)}$, where $f(x_j)$ is a function specifying the extent to which each exemplar contributes to the target category.
+For brevity, we will abbreviate $f(x_j)$ as $f_j$.
+PACKER sets $f_j$ depending on exemplar $j$'s category membership: $f_j = \phi$ if $x_j$ is a member of the contrast category, and $f_j = \gamma$ if $x_j$ is a member of the target category.
+$\phi$ and $\gamma$ are free parameters ($-\infty \leq \phi, \gamma \leq \infty$) controlling the contribution of contrast- and target-category similarity, respectively.
+Larger absolute values for either parameter produce greater consideration of that exemplar for generating the target category, with values of 0 producing no effect.
+A negative value for $f_j$ produces a `repelling' effect (target exemplars are less likely to be generated nearby $x_j$).
+Conversely, a positive value for $f_j$ produces a `pulling' effect (target exemplars are more likely to be generated nearby $x_j$).
 
-When prompted to generate a new example, the model considers both the similarity to examples from other categories as well as the similarity to examples in the target category.
-The aggregated similarity $a$ between generation candidate $y$ and the model's stored exemplars $x$ can be computed as:
-\begin{equation}
-  a(y, x) = \sum_j{\hat{f_j} s(y, x_j)}
-  \label{eq:packer-ss}
-\end{equation}
-where $\hat{f_j}$ is a function specifying each example's degree of contribution toward generation.
-Although $\hat{f_j}$ may be set arbitrarily, in PACKER it is set according to class assignment.
-$\hat{f_j} = \phi$ for known members of contrast categories, $\hat{f_j} = \gamma$ for members of the target category.
-$\phi$ and $\gamma$ are free parameters ($-\infty \leq \phi, \gamma \leq \infty$) controlling the contribution of target- and contrast-category similarity to generation.
-Larger absolute values for either parameter produce greater consideration of that type of similarity, with values of 0 producing no effect.
-Negative values produce a `repelling' effect (exemplars are less likely to be generated nearby).
-Conversely, positive values produce a `pulling' effect (exemplars are more likely to be generated nearby).
+Recall that we proposed that categories are generated by balancing two constraints: that new categories should be different from existing categories, and exemplars belonging to the same category should be similar to one another.
+This is encoded in PACKER when $\phi < 0$, and $\gamma > 0$.
+Negative values for $\phi$ encourages $y$ to be distant from the contrast category (as similarity to contrast category exemplars are subtracted during aggregation).
+Positive values for $\gamma$ encourage $y$ to be close to other exemplars of the target category.
+When $|\phi| = \gamma$, the "force" of repulsion from the contrast category is equal to the "force" of attraction to the target category.
 
-Because the model's main proposals are that new categories should be different from existing categories, and exemplars belonging to the same category should be similar to one another, PACKER is commonly simulated under the constraints that $\phi \leq 0$, $\gamma \geq 0$, thus, similarity to contrast categories is effectively subtracted from similarity to the target category.
-When $-\phi = \gamma$, negative aggregate similarities indicate that $y$ is more similar to members of contrast categories, and positive values indicate $y$ is more similar to the target category.
-
-The probability that a given candidate $y$ will be generated is evaluated using the \citet{luce1977choice} choice axiom.
+The probability that a given candidate $y$ will be generated is evaluated using an Exponentiated \citet{luce1977choice} choice rule.
 Candidates with greater values of $a$ are more likely to be generated than candidates with smaller values:
@@
 p(y) = \dfrac
-    { exp( { \theta \cdot a(y, x) } ) }
-    { \sum_i{ exp( { \theta \cdot a(y_i, x) } ) } }
+    { \exp( { \theta \cdot a(y, x) } ) }
+    { \sum_i{ \exp( { \theta \cdot a(y_i, x) } ) } }
     \label{eq:packer-choice}
 \end{equation}
-where $\theta$ ($\geq 0$) is a free parameter controlling overall response determinism.
+where $\theta$ ($\theta \geq 0$) is a free parameter controlling overall response determinism.
 
 \subsection{Summary}
-
-The PACKER model described above explains exemplar and category generation under a traditional exemplar-similarity approach, proposing that people generate categories by maximizing within-category similarity and minimizing between-category similarity.
+The proposed PACKER model formulates exemplar and category generation under a traditional exemplar-similarity approach, proposing that people generate categories by minimizing between-category similarity and maximizing within-category similarity.
 The underlying processes assumed by PACKER are highly similar to those in the GCM \citep{nosofsky1984choice}, with the only alteration being that the model's response is a function of the $a$ metric, which is the aggregate of potentially counteracting similarities (positive or negative values), rather than the relative summed similarity assumed in the original model.
 
-In later sections, we will explore the unique predictions yielded by these design principles. First, however, we contrast PACKER with existing approaches in the field.
+In later sections, we will explore the unique predictions yielded by these design principles. First, however, we contrast PACKER with other category generation models.
 
@@
 
-As noted in the introduction, existing accounts of category generation focus primarily on explaining the phenomenon that generated categories tend to possess distributional similarities with existing categories.
-\citet{jern2013probabilistic} were the first (as far as we know) to conduct formal simulations on category generation.
-Thus, the models they considered represent the strongest alternatives to PACKER.
-Below, we describe in detail two of these alternatives: a \textit{copy-and-tweak} account that was initially proposed within the creative cognition literature \citep{ward2002role,ward1995s}, and the \textit{hierarchical sampling} model developed by \citet{jern2013probabilistic}.
+Previous computational models of category generation have focused on capturing the tendency for people to produce new categories that have similar distributional properties to existing categories.
+To the best of our knowledge, \citet{jern2013probabilistic} were the first to analyze behavioral results quantiatively to computational models of category generation.
+Based on their work, we describe two alternative models: a formalization of the \textit{copy-and-tweak} hypothesis \citep{ward2002role,ward1995s}, and the \textit{hierarchical sampling} proposed by \citet{jern2013probabilistic}.
 
 \subsection{Copy-and-Tweak}
+%JLA: PLEASE CHECK WHAT I WROTE TO MAKE SURE IT"S TRUE OF THEIR MODEL
+The \textit{copy-and-tweak} model of \citet{jern2013probabilistic} formalizes one account of category generation \citep[see][]{ward2002role,ward1995s}.
+The model proposes that generation is a two-part process: First, learners retrieve an observation from memory, and then they tweak it to be different from the retrieved observation.
+\citet{jern2013probabilistic} interpreted this proposal in terms of an exemplar model, where participants score potential new exemplars as the summed exponentiated similarity of the potential new exemplar to every retrieved exemplar from the category, and sample from the normalized score distribution over potential exemplars.
+Given that this was intended for generating new exemplars within known category, we implemented a different formalization of the copy-and-tweak hypothesis.
 
-The \textit{copy-and-tweak} model implemented by \citet{jern2013probabilistic} instantiates one of the earliest accounts of category generation \citep[see][]{ward2002role,ward1995s}.
-The model proposes that generation is a two-part process: first, learners retrieve an observation from memory, and then they change it to create something new.
-\citet{jern2013probabilistic} interpreted this proposal in terms of a formal exemplar approach -- first, participants retrieve an exemplar from memory, and slightly change its features until it is sufficiently novel.
-In our implementation of the copy-and-tweak model, the probability that a given exemplar $z$ is retrieved from memory is given by:
+In our implementation of the copy-and-tweak model, the probability that a given exemplar $z$ is retrieved from memory is $p(z) = \exp(f_z) / \sum_z{ \exp(f_z) }$, where $f_z$ is a function specifying each item's relative chance of being selected.
+In our simulations, we set $f_z$ to be $-\lambda$ for members of contrast categories, and $f_z = \lambda$ for members of the target category.
+The resulting free parameter $\lambda$ ($>0$) thus controls the relative chances that a member of the target category will be retrieved as the source.
+%\citet{jern2013probabilistic} did not implement any mechanisms to differentially weigh retrieval probabilities, and so our model encompasses theirs as a special case when $\lambda = 0$ (uniform probabilities).
+JLA: (i reread their model and i don't think this is quite true)
 
-\begin{equation}
-p(z) = exp(\hat{f}_z) / \sum_z{ exp(\hat{f}_z) }
-\end{equation}
-%
-where $\hat{f}_z$ is a function specifying each item's relative chance of being selected.
-As in Equation \ref{eq:packer-ss}, $\hat{f}_j$ may be set arbitrarily.
-However, in our simulations, $\hat{f}_z = -\lambda$ for members of contrast categories, and $\hat{f}_z = \lambda$ for members of the target category.
-The resulting free parameter $\lambda$ ($>0$) thus controls the relative chances that a member of the target category will be retrieved as the source.
-\citet{jern2013probabilistic} did not implement any mechanisms to differentially weigh retrieval probabilities, and so our model encompasses theirs as a special case when $\lambda = 0$ (uniform probabilities).
-
-After a source exemplar is retrieved, the similarity between generation candidates $y$ and the source is computed as per Equation \ref{eq:similarity}.
-The model's goal is to generate an item that is similar to $z$, but does not exceed a given similarity threshold: the generated item should be similar-but-not-too-similar to $z$.
-Formally, the probability that $y$ will be generated based on a source $z$, is given by:
+After a source exemplar is retrieved, the similarity between generation candidates $y$ and the retrieved exemplar is computed as per Equation \ref{eq:similarity}.
+The model's goal is to generate an item that is similar to $z$, but does not exceed a given similarity threshold: The generated item should be similar, but not too similar to $z$.
+Formally, the probability that candidate exemplar $y$ will be generated based on a source exemplar $z$ is
 
@@
     p(y|z)  = \dfrac
-    { exp(\theta \cdot s(y,z)) g(y,z,\tau)}
-    {\sum_i{exp(\theta \cdot s(y_i,z)) g(y_i,z,\tau)}} 
+    { \exp \left\{\theta \cdot s(y,z) \right\} I\left(s(y,z) \leq \tau\right) }
+    {\sum_i{\exp \left\{ \theta \cdot s(y_i,z) \right\} I\left(s(y_i,z) \leq \tau\right)}} 
 \end{equation}
 % 
-where $\theta$ is a response determinism parameter.
-$g$ is a function specifying whether $y$ exceeds the similarity threshold, $\tau$, which is treated as a free parameter ($0<\tau\leq1$).
-$g = 1$ when $s(y,z) < \tau$ and $g = 0$ otherwise.
-Thus, exemplars exceeding the similarity threshold will not be generated, but exemplars just under the threshold are most likely.
-\citet{jern2013probabilistic} did not implement any similarity threshold mechanism, and so their model is the special case where $\tau = 1$ (any exemplar may be selected).
-
-To obtain predictions not depending on a given source example, the model's predictions can be aggregated over all possible sources:
-
-\begin{equation}
-  p(y) = \sum_z{p(z)p(y|z) }
-\end{equation}
-% 
-Thus, the probability that an example $y$ will be selected is a function of its generation probability given each of the sources, and the probability each source will be retrieved.
+where $\theta$ is a response determinism parameter, $\tau$ is the similarity threshold, and $I(\cdot)$ is the indicator function, which returns 1 when it is passed a true expression and 0 otherwise.
+So, $I\left(s(y,z) \leq \tau\right)$ is 1 when the candidate exemplar $y$ is far enough away from the source exemplar $z$.
+This ensures the source exemplar is tweaked "enough".
+When $\tau=1$, the threshold has no effect, which results in a model similar to \citet{jern2013probabilistic}.
+To obtain predictions not depending on a given source example, the model's predictions can be aggregated over all possible sources: $p(y) = \sum_z{p(z)p(y|z)}$.
+%Thus, the probability that $y$ will be selected is a function of its generation probability given each of the sources, and the probability each source will be retrieved.
 
 \subsection{Hierarchical Sampling}
-
-\citet{jern2013probabilistic} reported several experiments revealing systematic flaws in the copy-and-tweak account.
-Instead, their results supported a hierarchical sampling model, in where categories are represented as probability distributions in a $k$-dimensional feature space.
-Observed exemplars (represented as points in this space) are viewed as samples from their underlying category distribution (e.g., a multivariate normal distribution).
+% from other categories 
+%with independently generated means, but correlated covariance matrices.
+%Observed exemplars (represented as points in this space) are viewed as samples from their underlying category distribution (e.g., a multivariate normal distribution).
 These category distributions are in turn viewed as samples from an underlying domain distribution, which represents the common distributional characteristics among known categories (e.g., feature variance, feature correlations).
@@
 
-As in the implementation reported by \cite{jern2013probabilistic}, we assume that each category's exemplars are distributed according to a multivariate normal distribution with parameters ($\mu, \Sigma$).
-To obtain a conjugate model, we assume that the domain of categories is inverse-Wishart distributed.
-Space does not allow for a full derivation of this model, and so we will simply report how the target category parameters ($\mu_B, \Sigma_B$) are inferred.
+Based on several results inconsistent with the copy-and-tweak account, \citet{jern2013probabilistic} advocated a hierarchical Bayesian model.
+Exemplars of each category were generated from a multivariate Gaussian distribution over the dimensions of stimulus space.
+The mean of each category was independent, but the covariance matrices (encoding feature variances and correlations) were generated from a common prior distribution.
+New categories are generated then by generating a new mean (uniform over stimulus space) and covariance matrix from the common prior distribution.
+Because the shared prior distribution's parameters were unobserved, a hierarchical Bayesian model uses information from the previous categories (their feature variances and correlations) to generate the covariance matrix of the new category.
 
-The \citet{jern2013probabilistic} implementation does not assume anything about the central tendency of categories, and in their model generated $\mu_B$ are sampled uniformly across all possible values.
-Thus, we infer $\mu_B$ with:
+Following \citet{jern2013probabilistic}, we assume that each category's exemplars are distributed according to a multivariate normal distribution with parameters ($\mu, \Sigma$).\footnote{Following \citet{jern2013probabilistic} stimuli were put through a logistic function to constrain the domain to be between 0 and 1.} We assume that each category's covariance matrix is inverse-Wishart distributed with parameters ($v$, $\kappa,$ and $\Sigma_D$).\footnote{Note that \citet{jern2013probabilistic}'s model is slightly different, as they used a non-conjugate model.
+Their model acts very similar to ours and the differences in model fits to our results are minor.} $\Sigma_D$ is the covariance matrix shared between categories.
+We assumed the shared covariance matrix $\Sigma_D$ was generated from a Wishart distribution (again, due keep the model conjugate) with parameters $v_0$, $\kappa_0$, and $\Sigma_0$.
+We set $\nu_0 = \kappa_0 = 1$, and $\Sigma_0 = \rho {\bf I}$, where $\rho$ was a free parameter controlling the expected variance of dimensions (dimensions of the shared covariance are expected to be uncorrelated) and ${\bf I}$ is the identity matrix.
 
+To simplify the model predictions, we used {\em maximum a posteriori} (MAP) estimates for the hidden parameters and then generated new categories based on those estimates.
+Due to conjugacy, the MAP estimate for the shared covariance matrix $\Sigma_D = \Sigma_0 + \sum_c{C_c}$, where $C_c$ is the empirical covariance matrix of category $c$.
+The MAP estimate of the covariance matrix for category $B$ is
 \begin{equation}
-  \mu_B = \dfrac
-    {\kappa\mu_{0} + n_B \bar{x_B}}
-    {\kappa + n_B}
-    \label{eq:hs_mus}
-\end{equation}
-% 
-where $\mu_0$ is the prior on $\mu_B$.
-Because \citet{jern2013probabilistic} sample $\mu_B$ uniformly, we set $\mu_0$ to the center of the domain space.
-$\kappa$ ($>0$) is a free parameter, weighting the importance of $\mu_0$.
-$n_B$ and $\bar{x_B}$ are the target category sample size and mean, respectively.
-In the case where the target class is populated (i.e., $n_B > 0$), $\mu_{B}$ ends up lying somewhere between $\mu_{0}$ and $\bar{x_y}$, depending on $\kappa$ and $n_B$.
-When the target class has no members, Equation \ref{eq:hs_mus} reduces to $\mu_{B} = \mu_{0}$.
-Because we set $\mu_0$ to the center of the space, this outcome is the same as if we had integrated over all possible $\mu_0$.
-
-Unlike $\mu_B$, $\Sigma_B$ cannot be computed considering only the members of the target category.
-Instead, $\Sigma_B$ is influenced both by the distribution of the members of the target category as well as members of other categories through the domain distribution, $\Sigma_D$.
-In our implementation, $\Sigma_D$ is inferred based on the observed (empirical)  covariance matrices of the known categories in the domain, $C$.
-We assume these covariance matrices to be Wishart-distributed, and so $\Sigma_D$ can be computed as:
-
-\begin{equation}
-    \Sigma_D = \Sigma_0 + \sum_{i}{C_i}
-\end{equation}
-%
-where $\Sigma_{0}$ is a $k$-by-$k$ prior covariance matrix.
-We use a $k$-dimensional identity matrix $I_k$ multiplied element-wise against a free parameter, $\rho$, controlling the amount of variance assumed by the prior, $\Sigma_0 =  I_k\rho$.
-Thus, categories are assumed to have some degree of variance along each feature (specified by $\rho$), but are not assumed to possess feature-feature correlations.
-
-Finally, assuming $(\mu_B, \Sigma_B)$ are Normal-Inverse-Wishart distributed, $\Sigma_B$ can be computed as:
-
-\begin{equation}
-  \Sigma_B = [\Sigma_D \nu + C_B +
+  \Sigma_B = \left[ \Sigma_D \nu + C_B +
     \dfrac
@@
     (\bar{x_B}-\mu_B)(\bar{x_B}-\mu_B)^T
-  ] (\nu + n_B)^{-1}
+  \right] (\nu + n_B)^{-1}
   \label{eq:Sigma_B}
@@
 %
-where $\nu$ ($>k-1$) is an additional free parameter weighting the importance of $\Sigma_{D}$. When the target category has no members, Equation \ref{eq:Sigma_B} reduces to $\Sigma_B = \Sigma_D$.
+where $\nu$ ($>k-1$) is an additional free parameter (from the Inverse-Wishart prior on $\Sigma_B$) weighting the importance of $\Sigma_{D}$.
+When the target category has no members, $\Sigma_B = \Sigma_D$.
 
-Generated exemplars are then assumed to be drawn from the multivariate normal distribution specified by $(\mu_{B}, \Sigma_{B})$. Thus, $p(y)$ can be computed as:
-
+Generated exemplars are then assumed to be drawn from the multivariate normal distribution specified by $(\mu_{B}, \Sigma_{B})$. Thus, $p(y)$ is
 \begin{equation}
   p(y) = \dfrac
-    {exp( \theta \cdot Normal(y, \mu_{B}, \Sigma_{B}))}
-    {\sum_i exp( \theta \cdot Normal(y_i, \mu_{B}, \Sigma_{B}))} 
+    {\exp( \theta \cdot {\rm Normal}(y; \mu_{B}, \Sigma_{B}))}
+    {\sum_i \exp( \theta \cdot {\rm Normal}(y_i; \mu_{B}, \Sigma_{B}))} 
 \end{equation}
-\
-where $\theta$ is a response determinism parameter.
+where $\theta$ is a response determinism parameter and ${\rm Normal}(y; \mu, \Sigma)$ denotes a multivariate Normal density evaluated at $y$.
+%JLA: There's something funky going on with the notation. I got confused by your noation, so hopefully you can update what i did to be consistent with the rest.
+\section{Behavioral Experiment}
+A novel prediction of PACKER is that the contrast category's location within stimulus space should influence the generated category -- the generated category should be repulsed from the contrast category.
+The other models predict the location of the contrast category within stimulus space should not influence the generated category.
+This experiment tests this prediction and quantitatively evaluates the three models.
 
-\section{Behavioral Experiment}
-
-Whereas the copy-and-tweak and hierarchical sampling models both assume explicit mechanisms to re-purpose knowledge about known categories, the PACKER model forgoes these mechanisms entirely.
+%Whereas the copy-and-tweak and hierarchical sampling models both assume explicit mechanisms to re-purpose knowledge about known categories, the PACKER model forgoes these mechanisms entirely.
 Thus, whereas the distributional structure of observed categories will be the primary determinant of generation in the copy-and-tweak and hierarchical sampling models, these properties should only affect PACKER's generation insofar as they constrain inter-exemplar similarities.
@@
 
-The behavioral experiment described below was designed to test this key prediction:  does the \textit{location} of contrast categories (as opposed to their structure), influence generation?
-The experiment follows the paradigm developed by \citet{jern2013probabilistic}: first, participants are exposed to members of a known category (`Alpha', or `A'), and are then asked to generate exemplars belonging to a new category (`Beta', or `B').
+The behavioral experiment described below was designed to test this key prediction:  Does the \textit{location} of contrast categories, influence generation?
+The experiment follows the paradigm developed by \citet{jern2013probabilistic}: first, participants learn members of a known category (`Alpha', or `A'), and are then asked to generate exemplars belonging to a new category (`Beta', or `B').
+We developed two Alpha categories (see Figure \ref{fig:middle-bottom-conditions}).
+We did so by creating one of the Bottom Alpha category as a tight cluster and formed the Middle Alpha category by translating it to the center of stimulus space.
+As the two categories only differ in their location, the two categories should not differ in their distributional information.
 
-We developed two Alpha categories (see Figure \ref{fig:middle-bottom-conditions}), instantiating the two conditions of the experiment.
+%, instantiating the two conditions of the experiment.
 In both conditions, members of the Alpha category are tightly clustered, with equal feature variance and no feature correlations.
@@
 
-More generally, the PACKER model proposes that the probability a stimulus $y$ will be generated is a function of its similarity to contrast categories \textit{and} to members of the target category.
+More generally, PACKER proposes that the probability a stimulus $y$ will be generated is a function of its similarity to contrast categories \textit{and} to members of the target category.
 Two more general predictions (not specific to either condition) follow from this proposal: (1) the location of Beta examples should be positively related to distance from the Alpha category, and (2) Beta examples should be more similar to one another than they are to members of the Alpha category.
 
-\subsection{Participants \& Materials}
-
-We recruited 122 participants from Amazon Mechanical Turk.
-All participants were located in the United States.
-61 Participants were assigned to the `Middle' condition, 61 were assigned to the `Bottom' condition.
+\subsubsection{Participants \& Materials}
+We recruited 122 participants from Amazon Mechanical Turk from the US equally assigned to each condition.
 Stimuli were squares varying in color (RGB 25--230) and size (3.0--5.8cm).
@@
 Exemplars were randomly ordered within each block.
-Participants were shown a preview of the range of possible colors and sizes prior to training.
+Participants were shown the range of possible colors and sizes prior to training.
 
@@
 Participants were instructed that members of the Beta category could be quite similar or different depending on what they think makes the most sense for the category, but that they were not allowed to make the same example twice.
-
-As in the \citet{jern2013probabilistic} experiments, generation was completed using a sliding-scale interface.
-Participants were presented with two scales that could be used to change the color and size of the generated example.
+As in \citet{jern2013probabilistic}, generation was completed using a sliding-scale interface.
+Participants were presented with two scales that controlled the dimensions of the generated example.
 An on-screen preview of the example updated whenever one of the features was changed.
-Participants could generate any example along an evenly-spaced 9x9 grid, and they could change the example as much as they wanted before moving on.
+Participants could generate any example along an evenly-spaced 9x9 grid, except for any previously generated Beta exemplars.
 Neither the members of the Alpha category nor the previously generated Beta examples were visible during generation.
-Participants were not informed if they generated a member of the Alpha category, but could not complete the trial if they generated the same Beta two times.
 
@@
 
-Because the conditions differ only in their location along the y-axis (see Figure \ref{fig:middle-bottom-conditions}), our main interest is in the generation of Beta examples above and below the contrast category.
+Because the conditions differ only in their location along the y-axis (see Figure \ref{fig:middle-bottom-conditions}), we focus on how Beta exemplars are generated above and below the contrast category.
 Several sample Beta categories are depicted in Figure \ref{fig:beta.samples}.
-As is evident in Figure \ref{fig:beta.samples}, we observed a great deal of individual differences in generation strategy: whereas some participants generated all four Beta examples within a narrow y-axis range, others generated Beta examples along a wide range.
+As is evident in Figure \ref{fig:beta.samples}, we observed a great deal of individual differences in generation strategy: Whereas some participants generated all four Beta examples within a narrow y-axis range, others generated Beta examples along a wide range.
 
-To more specifically evaluate the key predictions made by PACKER, we determined the number of participants in each condition who placed at least one Beta example on the top `row' of the space (the maximum possible y-axis value), as well as the number of participants who placed at least one example on the bottom row.
+To evaluate the key predictions of PACKER, we determined the number of participants in each condition who placed at least one Beta examplar on the top and bottom `rows' of the space (the maximum and minimum possible y-axis value, respectively).
 The resulting contingencies data are shown in Table \ref{table:subset-table}.
-Fisher's Exact Tests on these data reveal that a greater number of participants assigned to the Middle condition generated a Beta example in the bottom row of the space, $p < 0.001$, but the conditions did not differ in use of the top of the space, $p = 0.16$.
-A greater number of participants in the Middle condition placed Beta examples in the top rows \textit{and} the bottom rows, $p = 0.038$.
+Fisher's Exact Tests reveal that more Middle participants generated a Beta exemplar in the bottom row , $p < 0.001$, but the conditions did not differ in use of the top of the space, $p = 0.16$.
+More Middle participants placed Beta exemplars in the top \textit{and} bottom rows, $p = 0.038$.
 
@@
 \begin{center} 
-\caption{Behavioral results.} 
 \label{table:subset-table} 
@@
 \end{tabular}
+\caption{Behavioral results.} 
 \end{center} 
 \end{table}
-
-To evaluate PACKER's other predictions, we computed the number of times each unique exemplar was generated, and plotted those data as a function of distance to the center of the Alpha category.
+To evaluate PACKER's other predictions, we computed the number of exemplars produced at different distances to the center of the Alpha category.
 These data, depicted in the top panel of Figure \ref{fig:distance.figs} reveal a strong preference for stimuli that are dissimilar to the members of the Alpha category: maximally distant items were by far the most frequently generated.
@@
 We did however, observe a notable subset of individuals with greater within-class distance.
-Inspection of these individuals reveals that they tended to adopted a `corners' approach, in which a Beta examples were placed almost exclusively in the furthest corners of the space, thereby producing more within-class distance.
+These individuals tended to adopt a `corners' approach, in which a Beta examples were placed almost exclusively in the corners of the space.
 
@@
     \caption{Behavioral results.
-\textit{Top}: Frequency of exemplar generation as a function of distance from the Alpha category.
+\textit{Top}: Frequency of exemplar generation as a function of distance from the Alpha category normalized by the maximum possible distance.
 \textit{Bottom}: Within- vs.
@@
 \subsection{Summary}
-
-Our results are broadly supportive of PACKER's predictions: people prefer to generate items that are maximally dissimilar from members of the Alpha category, and they tended to generate categories possessing greater within-class similarity than between-class similarity (with some notable exceptions, see Figure \ref{fig:distance.figs}).
+Our results support PACKER's predictions: People tend to generate items that are dissimilar from the contrast category Alpha and similar to the target category Beta (with some exceptions, see Figure \ref{fig:distance.figs}).
 By consequence, we observed considerable differences in generation between the Middle and Bottom conditions: participants in the Bottom condition were less likely to use the bottom row of the stimulus space for generation, and participants in the Middle condition were more likely to create categories spanning the entire y-axis (utilizing the top and bottom row of the space).
-This latter result is especially interesting, as it shows that qualitatively different types of categories are generated, depending on only the location of the Alphas in the domain.
+This latter result is especially interesting as it conflicts with previous results: Qualitatively different types of categories are generated (with very different distributional information), depending on only the location of the Alphas in the domain.
 
@@
 
-\section{Simulations}
-
-We conducted formal simulations to our data using the PACKER, copy-and-tweak, and hierarchical sampling models described above.
-In order to obtain an overall sense of each model's ability to explain our results, we fitted each model trial-wise, using a hill-climbing algorithm to maximize the log-likelihood of the model's predictions against our results.
-For each trial, models were initialized with the participant's Alpha category (Middle or Bottom), and the configuration of the participant's Beta category during that trial.
-
+\section{Model Evaluation}
+In this section, we evaluate the model fits to our results.
+To obtain an overall sense of each model's ability to explain our results, we fit each model by maximizing the log-likelihood of the model's predictions of the human results.
 Four free parameters were fitted in each model.
 The $c$, $\phi$, $\gamma$, and $\theta$ parameters were fitted for PACKER; $c$, $\lambda$, $\tau$, and $\theta$ were fitted for the copy-and-tweak model, and $\kappa$, $\rho$, $\nu$, and $\theta$ were fitted for the hierarchical sampling model.
-Note that each model possess as $\theta$ parameter fulfilling the same role (response determinism), and PACKER shares with copy-and-tweak a specificity parameter, $c$.
+Note that each model possess a $\theta$ parameter fulfilling the same role (response determinism), and PACKER shares with copy-and-tweak a specificity parameter, $c$.
 Attention in PACKER and copy-and-tweak was set uniformly.
@@
 Each model's best-fitting parameterization is shown in Table \ref{table:model-fits}.
-Overall, PACKER outperformed copy-and-tweak and the hierarchical sampling model by a considerable margin ($\sim300$ log-likelihood).
+Overall, PACKER outperformed copy-and-tweak and the hierarchical sampling model by a considerable margin ($\sim17\%$ improvement in log-likelihood).
 The parameter settings associated with PACKER's best fit are exactly as expected: a strong preference for items that are similar to members of the target category but are dissimilar to members of the contrast category.
-A similar pattern of results was obtained when each participant's first trial was excluded (i.e., when the target category had no members).
-
-% JOE! I do not know how log-likes are supposed to be notated!
-
+A similar pattern of results was obtained when we only considered the second to fourth exemplars generated by each participant.
+%JLA this is fine
 \begin{table}[ht!]
@@
 However, as shown in Figures \ref{fig:beta.samples} and \ref{fig:distance.figs}, we observed a great deal of individual differences in generation strategy.
-For example whereas many participants tightly clustered the Beta category, others distributed its members to the four corners of the domain.
-Likewise, whereas many participants created `column'-like categories (varying along the y-axis but not the x-axis), others created `row'-like categories.
-While it is important that any model be able to explain different types of performance under the same parameterization, these radical differences in generation are difficult to account for under a single set of parameters.
+Four representative patterns of the generated Beta category were observed: tightly clustered, four corners, `column'-like categories (varying along the y-axis but not the x-axis), others created `row'-like categories.
+While it is important that any model be able to explain different types of performance with the same parameter values, these radical differences in generation are difficult to account for under a single set of parameters.
 
@@
 For example, whereas the tight cluster profile could be explained with a moderately negative $\phi$ and a strongly positive $\gamma$, the four corners profile could be simulated with a strongly negative $\phi$ and $\gamma\leq0$.
-Here, we quickly explore the role of the attention weights ($w$, in Equation \ref{eq:similarity}) in producing row-like or column-like categories.
+Here, we briefly explore the role of the attention weights ($w_1$ and $w_2$, in Equation \ref{eq:similarity}) in producing row-like or column-like categories.
 
@@
 
-In order to simulate this finding, we set the attention weight parameters in PACKER and copy-and-tweak separately for each participant.
+To simulate this finding, we set the attention weight parameters in PACKER and copy-and-tweak separately for each participant.
 To simulate the hierarchical sampling model we set the domain covariance prior, $\Sigma_0$, separately for each participant.
-Each model's free parameters were set as in Table \ref{table:model-fits}.
+The other free parameters were set as in Table \ref{table:model-fits}.
 While there exist methods to find the optimal attention weights for a given classification \citep[see][]{vanpaemel2012using}, for ease of computation we simply approximated the weights based on the inverse of each feature's range: $w_k \propto range(k)^{-1}$.
-The ranges are not inverted for the hierarchical sampling model.
 Thus, the Alpha and Beta categories are assumed to be distinct along dimensions that the Betas do not vary on.
@@
 
-The results, depicted in Figure \ref{fig:range-diff-gradient}, are highly supportive of PACKER's predictions.
+The results, depicted in Figure \ref{fig:range-diff-gradient}, support PACKER's predictions.
 When the x-axis is more strongly weighted, PACKER creates column categories to the sides of the Alphas.
 Conversely, when the y-axis is weighted, PACKER creates row categories above and below the Betas.
-This behavior falls out from the nature of dimensional selective attention: dimensions with more weighting have a sharper similarity gradient.
-For example, when the x-axis is weighted, PACKER attempts to create Beta categories with more within-class similarity (less range), and less between-class similarity along the x-axis, resulting in column-like categories.
+This behavior falls out from the nature of selective attention: Dimensions with more weighting have a sharper similarity gradient.
+For example, when the x-axis is weighted more, PACKER favors Beta categories with more within-class similarity (less range), and less between-class similarity along the x-axis, resulting in column-like categories.
 
-The hierarchical sampling and copy-and-tweak models do not as often create categories with strong range differentials.
-However, when they do, the categories tend to show the opposite pattern: columns are placed above and below the Alphas, rows are to the side of the Alphas.
-In the hierarchical sampling model, the position of contrast categories is not considered when generating new items.
+The hierarchical sampling and copy-and-tweak models do not often create categories with strong range differentials.
+However, when they do, the categories tend to show the opposite pattern: Columns are placed above and below the Alphas, rows are to the side of the Alphas.
+In the hierarchical sampling model, the position of contrast category is not considered when generating new items.
 Only the distributional structure of the contrast category and items already in the target category are considered.
@@
 \section{Discussion}
-
-The creative use of conceptual knowledge is a highly intriguing yet understudied topic in category learning research.
+The creative use of conceptual knowledge is a fascinating yet understudied topic in categorization.
 In this paper, we presented a novel exemplar-based approach to explaining category generation.
-The PACKER model, based on the Generalized Context Model \citep[GCM;][]{nosofsky1984choice}, proposes that categories are represented as a collection of exemplars stored in memory, and that members of generated categories should be similar to one another, yet dissimilar to members of opposing categories.
-
-Previous accounts of category generation \citep[see][]{jern2013probabilistic,ward1995s,ward1994structured} have focused primarily on explaining why generated categories tend to share distributional properties with known categories.
-Here we instead explored whether there are aspects of category generation that can be explained through the basic and well-understood principles of the exemplar framework.
+The PACKER model proposes that categories are represented as a collection of exemplars stored in memory, and that members of generated categories should be similar to one another, yet dissimilar to members of opposing categories.
+Given that exemplar models can be viewed as Importance-Sampling approximations of Bayesian models \citep{shi10}, PACKER can be viewed as a rational process model and approximating the expected density of a new category based on a contrast category.
 
 In a behavioral study and subsequent formal modeling, we found broad support for the PACKER model.
-Participants in our study more frequently generated items that are highly distant from members of contrast categories, and they tended to generate categories with more within-class than between-class similarity.
-Likewise, we found that the \textit{location} of contrast categories (rather than their distributional structure) shapes generation by imposing constraints on the areas of space that remain available for a new category.
+Participants in our study more frequently generated items that are distant from members of contrast categories, and they tended to generate categories with more within-class than between-class similarity.
+Likewise, we found that the \textit{location} of contrast categories shapes generation by imposing constraints on the areas of space that remain available for a new category.
 Our formal simulations reveal that PACKER provides a strong account of our observations.
@@
 The PACKER model is, in general, highly expressive in its performance.
-Under different parameter settings it is capable of generating tightly clustered or highly distributed categories, and adjusting the distribution of categories along each feature (to create row-like or column-like categories; see Figure \ref{fig:beta.samples}).
+Under different parameter settings it is capable of generating tightly clustered or highly distributed categories, and adjusting the distribution of categories along each feature.
 Future work will focus on exploring how individual differences in generation can be explained using PACKER, and whether the canonical effects observed by \citet{ward1994structured} and \citet{jern2013probabilistic} can be explained through exemplar-similarity alone.
@@
 \section{Acknowledgments}
-Support for this research was provided by the Office of the Vice Chancellor for Research and Graduate Education at the University of Wisconsin - Madison with funding from the Wisconsin Alumni Research Foundation.
-We would like to thank Kenneth Kurtz for helpful comments on this work.
-We also thank Alan Jern \& Charles Kemp for providing code and data from their experiments.
+Support for this research was provided by the Office of the VCRGE at the UW - Madison with funding from the WARF.
+We thank Kenneth Kurtz for helpful comments, and Alan Jern and Charles Kemp for providing code and data.
 % any other thanks?

### 2017-01-28 Manuscripts/cogsci-2017/paper.tex
7 pages bishes! still working

@@
 
+%JLA: I took this one out because it's not really something we're dealing with in our paper.
+%Further, exemplars drawn from the same species were less variable than animals drawn from different species, and possessed features that were adaptive for their environment (as is the case on Earth).
+
 % set up PGF
@@
 \begin{abstract}
-FILLER
+FILLER (make very short)
 
@@
 
-One of the most intriguing capabilities of human cognition is the ability to creatively generate new ideas and concepts.
-The creative use of knowledge has only been infrequently the subject of formal inquiry.
+One of the most intriguing capabilities of human cognition is the ability to generate new ideas and concepts.
+The creative creation of knowledge is an infrequent subject of formal inquiry.
 This paper focuses on one topic within creative cognition, category generation.
-Foundational work on category generation \citep[e.g.,][]{smith1993constraining,ward2002role,ward1994structured,ward1995s} found that people tend to create new categories that are similar to related previously known categories: New exemplars from the new categories tend to be copied from exemplars for a similar categories, but tweaked on one or a few dimensions.
+Foundational work \citep[e.g.,][]{smith1993constraining,ward2002role,ward1994structured,ward1995s} found that people tend to create new categories that are similar to related previously known categories: New exemplars from the new categories tend to be copied from exemplars for a similar categories, but tweaked on one or a few dimensions.
 Thus, people generate a new category containing similar distributional properties to the categories it is based on, while being different from those categories in some manner.
 
-Existing work on the creative use of conceptual knowledge predominantly explores the role of prior knowledge in generating novel concepts.
-A core experimental phenomenon is that people generate categories possessing the same distributional properties as known categories.
-For example, in a canonical study \citet{ward1994structured} asked participants to draw and describe species of plants and animals that might exist on other planets.
-Generation was strongly constrained by prior knowledge of Earth plants and animals: people generated alien species using the same structural forms found on Earth (e.g., eyes, legs, wings) and possessing the same feature correlations observed on Earth (e.g., feathers co-occur with wings).
-Likewise, exemplars drawn from the same species were less variable than animals drawn from different species, and possessed features that were adaptive for their environment (as is the case on Earth).
+Existing work on the creative use of conceptual knowledge  explores the role of prior knowledge in generating novel concepts.
+A core phenomenon is that people generate new categories with similar distributional properties as existing categories.
+For example, \citet{ward1994structured} asked participants to draw and describe plants and animals that might exist on other planets.
+Generation was strongly constrained by prior knowledge of Earth plants and animals: People generated alien species with the same features as those found on Earth (e.g., eyes, legs, wings) and possessing the same feature correlations observed on Earth (e.g., feathers co-occur with wings).
 
 Recent work has proposed and tested computational models of how new categories are generated based on prior knowledge \citep{jern2013probabilistic}.
-In their experiments, \citet{jern2013probabilistic} trained participants on a set of experimenter-defined categories composed of exemplars within an artificial three-dimensional domain (e.g., 2D shapes varying in size, hue, and saturation).
+In their experiments, they trained participants on a set of experimenter-defined categories composed of exemplars within an artificial three-dimensional domain (e.g., 2D shapes varying in size, hue, and saturation).
 After a short training phase, participants were asked to generate exemplars from a new category.
@@
 \end{equation}
-where $\theta$ ($\theta \geq 0$) is a free parameter controlling overall response determinism.
+where $\theta$ ($\theta \geq 0$) controls response determinism.
 
 \subsection{Summary}
-The proposed PACKER model formulates exemplar and category generation under a traditional exemplar-similarity approach, proposing that people generate categories by minimizing between-category similarity and maximizing within-category similarity.
-The underlying processes assumed by PACKER are highly similar to those in the GCM \citep{nosofsky1984choice}, with the only alteration being that the model's response is a function of the $a$ metric, which is the aggregate of potentially counteracting similarities (positive or negative values), rather than the relative summed similarity assumed in the original model.
+The proposed PACKER model suggests people generate categories by minimizing between-category similarity and maximizing within-category similarity.
+The underlying processes assumed by PACKER are highly similar to those in the GCM \citep{nosofsky1984choice}, with the only alteration being that PACKER aggregates positive- and negative-valued similarities, rather than only aggregating positive-valued similarities.
 
@@
 
-Previous computational models of category generation have focused on capturing the tendency for people to produce new categories that have similar distributional properties to existing categories.
-To the best of our knowledge, \citet{jern2013probabilistic} were the first to analyze behavioral results quantiatively to computational models of category generation.
+Previous models of category generation focus on capturing the tendency for people to produce new categories that have similar distributional properties to existing categories.
+To the best of our knowledge, \citet{jern2013probabilistic} were the first to analyze behavioral results quantitatively to computational models.
 Based on their work, we describe two alternative models: a formalization of the \textit{copy-and-tweak} hypothesis \citep{ward2002role,ward1995s}, and the \textit{hierarchical sampling} proposed by \citet{jern2013probabilistic}.
@@
 
-Following \citet{jern2013probabilistic}, we assume that each category's exemplars are distributed according to a multivariate normal distribution with parameters ($\mu, \Sigma$).\footnote{Following \citet{jern2013probabilistic} stimuli were put through a logistic function to constrain the domain to be between 0 and 1.} We assume that each category's covariance matrix is inverse-Wishart distributed with parameters ($v$, $\kappa,$ and $\Sigma_D$).\footnote{Note that \citet{jern2013probabilistic}'s model is slightly different, as they used a non-conjugate model.
-Their model acts very similar to ours and the differences in model fits to our results are minor.} $\Sigma_D$ is the covariance matrix shared between categories.
-We assumed the shared covariance matrix $\Sigma_D$ was generated from a Wishart distribution (again, due keep the model conjugate) with parameters $v_0$, $\kappa_0$, and $\Sigma_0$.
-We set $\nu_0 = \kappa_0 = 1$, and $\Sigma_0 = \rho {\bf I}$, where $\rho$ was a free parameter controlling the expected variance of dimensions (dimensions of the shared covariance are expected to be uncorrelated) and ${\bf I}$ is the identity matrix.
+We assume that each category's exemplars are distributed according to a multivariate normal distribution with parameters ($\mu, \Sigma$).\footnote{Following \citet{jern2013probabilistic} stimuli were put through a logistic function to constrain the domain to be between 0 and 1.} We assume that each category's covariance matrix is inverse-Wishart distributed with parameters ($v$, $\kappa,$ and $\Sigma_D$).\footnote{Note that \citet{jern2013probabilistic}'s model is slightly different, as they used a non-conjugate model.
+Their model acts very similar to our version of it and receives comparable fits.} $\Sigma_D$ is the covariance matrix shared between categories.
+We assumed the shared covariance matrix $\Sigma_D$ was generated from a Wishart distribution (for conjugacy) with parameters $v_0$, $\kappa_0$, and $\Sigma_0$.
+We set $\nu_0 = \kappa_0 = 1$, and $\Sigma_0 = \rho {\bf I}$, where $\rho$ was a free parameter controlling the expected variance of dimensions (dimensions of the shared covariance matrix are expected to be uncorrelated) and ${\bf I}$ is the identity matrix.
 
@@
 %
-where $\nu$ ($>k-1$) is an additional free parameter (from the Inverse-Wishart prior on $\Sigma_B$) weighting the importance of $\Sigma_{D}$.
+where $\nu$ ($\nu>k-1$) is an additional free parameter (from the Inverse-Wishart prior on $\Sigma_B$) weighting the importance of $\Sigma_{D}$.
 When the target category has no members, $\Sigma_B = \Sigma_D$.

### 2017-01-28 Manuscripts/cogsci-2017/paper.tex
more cleaning. passing back. see email soon

@@
 
-The PACKER model is an extension of the prominent Generalized Context Model of category learning (GCM; \citealp{nosofsky1984choice}).
+The PACKER model is an extension of the prominent Generalized Context Model of category learning (GCM; \citealp{nosofsky1984abbr}).
 The model assumes that each category is represented by a collection of exemplars within in a $k$-dimensional psychological space, and that generation is constrained by both similarity to members of the target category (the category in which a stimulus is being generated) as well as similarity to members of other categories.
@@
 The proposed PACKER model suggests people generate categories by minimizing between-category similarity and maximizing within-category similarity.
-The underlying processes assumed by PACKER are highly similar to those in the GCM \citep{nosofsky1984choice}, with the only alteration being that PACKER aggregates positive- and negative-valued similarities, rather than only aggregating positive-valued similarities.
+The underlying processes assumed by PACKER are highly similar to those in the GCM, with the only alteration being that PACKER aggregates positive- and negative-valued similarities, rather than only aggregating positive-valued similarities.
 
@@
 %JLA: PLEASE CHECK WHAT I WROTE TO MAKE SURE IT"S TRUE OF THEIR MODEL 
-The \textit{copy-and-tweak} model of \citet{jern2013probabilistic} formalizes one account of category generation \citep[see][]{ward2002role,ward1995s}.
-The model proposes that generation is a two-part process: First, learners retrieve an observation from memory, and then they tweak it to be different from the retrieved observation.
-\citet{jern2013probabilistic} interpreted this proposal in terms of an exemplar model, where participants score potential new exemplars as the summed exponentiated similarity of the potential new exemplar to every retrieved exemplar from the category, and sample from the normalized score distribution over potential exemplars.
+The \textit{copy-and-tweak} model proposes that generation is a two-part process: First, learners retrieve an observation from memory, and then they tweak it to be different from the retrieved observation.
+\citet{jern2013probabilistic} interpreted this proposal in terms of an exemplar model, where participants score potential new exemplars as the summed exponentiated similarity of the new exemplar to every exemplar from the category, and sample from the normalized score distribution.
 Given that this was intended for generating new exemplars within known category, we implemented a different formalization of the copy-and-tweak hypothesis.
@@
 
-\subsection{Procedure}
+\subsubsection{Procedure}
 
@@
 
-\subsection{Results}
+\subsubsection{Results}
 
@@
 Several sample Beta categories are depicted in Figure \ref{fig:beta.samples}.
-As is evident in Figure \ref{fig:beta.samples}, we observed a great deal of individual differences in generation strategy: Whereas some participants generated all four Beta examples within a narrow y-axis range, others generated Beta examples along a wide range.
+As is evident in Figure \ref{fig:beta.samples}, we observed large individual differences in generation strategy: Whereas some participants generated all four Beta examples within a narrow y-axis range, others generated Beta examples along a wide range.
 
@@
 \subsection{Individual Differences}
+%In future work, we intend to explore more systematically the role of PACKER's parameters in different subtypes of performance.
+For example, whereas the tight cluster profile could be explained with a moderately negative $\phi$ and a strongly positive $\gamma$, the four corners profile could be simulated with a strongly negative $\phi$ and $\gamma\leq0$.
+%While it is important that any model be able to explain different types of performance with the same parameter values, these radical differences in generation are difficult to account for under a single set of parameters.
+Our behavioral results showed that, in general, generated categories consist of item that are dissimilar to members of opposite categories, and similar to members of the their own category.
+There were four typical patterns of Beta categories: tightly clustered, four corners, `column'-like categories (varying along the y-axis but not the x-axis), others created `row'-like categories (see Figures \ref{fig:beta.samples} and \ref{fig:distance.figs}).
+Here, we explore the role of PACKER's different attention weights ($w_1$ and $w_2$) in producing row-like or column-like categories.
 
-Our behavioral results showed that, in general, generated categories consist of item that are (A) dissimilar to members of opposite categories, and (B) similar to members of the their own category.
-However, as shown in Figures \ref{fig:beta.samples} and \ref{fig:distance.figs}, we observed a great deal of individual differences in generation strategy.
-Four representative patterns of the generated Beta category were observed: tightly clustered, four corners, `column'-like categories (varying along the y-axis but not the x-axis), others created `row'-like categories.
-While it is important that any model be able to explain different types of performance with the same parameter values, these radical differences in generation are difficult to account for under a single set of parameters.
-
-In future work, we intend to explore more systematically the role of PACKER's parameters in different subtypes of performance.
-For example, whereas the tight cluster profile could be explained with a moderately negative $\phi$ and a strongly positive $\gamma$, the four corners profile could be simulated with a strongly negative $\phi$ and $\gamma\leq0$.
-Here, we briefly explore the role of the attention weights ($w_1$ and $w_2$, in Equation \ref{eq:similarity}) in producing row-like or column-like categories.
-
+%JLA: Move PACKER to be between Behavioral and Copy and Tweak. Also, I think this will take up less space as a row figure.
 \begin{figure}[ht!]
@@
     \inputpgf{figs/}{range-diff-gradient.pgf}
-    \caption{Generated category structure as a function of location in the domain.
+    \caption{Generated category structure as a function of location.
 Orange areas in each gradient correspond to stimuli that were commonly generated into category possessing greater y-axis range (columns).
@@
 Whereas column-like categories more often include stimuli to the left or right of the Alpha class, row-like categories appear above and below the Alpha class.
-Thus, even beyond generating Betas in locations not occupied by the Alphas, participants appear to modify the distributional structure of their categories in order to maximize distance from the Alphas.
+Thus, participants modify the distributional structure of new categories to the maximize distance from the contrast category.
 
-To simulate this finding, we set the attention weight parameters in PACKER and copy-and-tweak separately for each participant.
+To simulate this finding, we set the attention weight parameters in PACKER and copy-and-tweak separately per participant.
 To simulate the hierarchical sampling model we set the domain covariance prior, $\Sigma_0$, separately for each participant.
@@
 The results, depicted in Figure \ref{fig:range-diff-gradient}, support PACKER's predictions.
-When the x-axis is more strongly weighted, PACKER creates column categories to the sides of the Alphas.
-Conversely, when the y-axis is weighted, PACKER creates row categories above and below the Betas.
-This behavior falls out from the nature of selective attention: Dimensions with more weighting have a sharper similarity gradient.
+When the x-axis is weighted more, PACKER creates column categories to the sides of the Alphas.
+Conversely, when the y-axis is weighted more, PACKER creates row categories above and below the Alphas.
+This behavior falls out from the nature of selective attention: Dimensions weighted more have a sharper similarity gradient.
 For example, when the x-axis is weighted more, PACKER favors Beta categories with more within-class similarity (less range), and less between-class similarity along the x-axis, resulting in column-like categories.

### 2017-01-29 Manuscripts/cogsci-2017/paper.tex
nbc pass back to joe

@@
 \begin{abstract}
-FILLER (make very short)
+
+The creative generation of new concepts is an intriguing yet understudied topic in cognitive science.
+In this paper, we present a novel exemplar-based approach to category generation: PACKER (\textit{Producing Alike and Contrasting Knowledge using Exemplar Representations}).
+PACKER's core design assumptions are that (\textit{a}) categories are represented as exemplars in a multidimensional psychological space, (\textit{b}) generated items should be similar to exemplars of the same category, and (\textit{c}) generated categories should be dissimilar to existing categories.
+A behavioral study and formal simulations reveal broad support for PACKER's predictions.
 
 \textbf{Keywords:} 
-Categorization; etc
+Categorization, exemplar models, category generation, creative cognition.
 \end{abstract}
@@
 
-One of the most intriguing capabilities of human cognition is the ability to generate new ideas and concepts.
-The creative creation of knowledge is an infrequent subject of formal inquiry.
-This paper focuses on one topic within creative cognition, category generation.
-Foundational work \citep[e.g.,][]{smith1993constraining,ward2002role,ward1994structured,ward1995s} found that people tend to create new categories that are similar to related previously known categories: New exemplars from the new categories tend to be copied from exemplars for a similar categories, but tweaked on one or a few dimensions.
-Thus, people generate a new category containing similar distributional properties to the categories it is based on, while being different from those categories in some manner.
+The creation of new concepts and ideas is among the most interesting -- yet infrequently studied -- capabilities of human cognition.
+This paper focuses on one topic within the broader field of creative cognition: category generation.
+Foundational work on this topic \citep[e.g.,][]{smith1993constraining,ward2002role,ward1994structured,ward1995s} has focused on the role of prior knowledge in generating novel concepts.
+A core phenomenon is that people generate categories with similar distributional properties as existing categories.
+For example, \citet{ward1994structured} asked participants to generate species of plants and animals that might exist on other planets.
+Generation was strongly constrained by prior knowledge of Earth species: People generated species with the same features as those found on Earth (e.g., eyes, legs, wings) and possessing the same feature correlations observed on Earth (e.g., feathers co-occur with wings).
 
-Existing work on the creative use of conceptual knowledge  explores the role of prior knowledge in generating novel concepts.
-A core phenomenon is that people generate new categories with similar distributional properties as existing categories.
-For example, \citet{ward1994structured} asked participants to draw and describe plants and animals that might exist on other planets.
-Generation was strongly constrained by prior knowledge of Earth plants and animals: People generated alien species with the same features as those found on Earth (e.g., eyes, legs, wings) and possessing the same feature correlations observed on Earth (e.g., feathers co-occur with wings).
-
-Recent work has proposed and tested computational models of how new categories are generated based on prior knowledge \citep{jern2013probabilistic}.
-In their experiments, they trained participants on a set of experimenter-defined categories composed of exemplars within an artificial three-dimensional domain (e.g., 2D shapes varying in size, hue, and saturation).
+Recent work has proposed and tested computational models of the processes underlying category generation.
+\citet{jern2013probabilistic} trained participants on a set of experimenter-defined categories composed of exemplars within an artificial three-dimensional domain (e.g., 2D shapes varying in size, hue, and saturation).
 After a short training phase, participants were asked to generate exemplars from a new category.
 Participants were provided with a set of scales to adjust the feature values of each generated stimulus, and were given unlimited time to create each example.
-As in the classic \cite{ward1994structured} experiment, \citet{jern2013probabilistic} found that people generated artificial categories possessing the same feature variance and correlations as the experimenter-defined categories in the domain.
-They evaluated different possible computational models of how people generate categories (such as "copy-and-tweak" and exemplar accounts) and found that a hierarchical Bayesian model generating new categories by abstracting the distributional structure of the categories in the domain.
+As in the classic \cite{ward1994structured} experiment, \citet{jern2013probabilistic} found that generated categories possessed the same feature variance and correlations as the experimenter-defined categories in the domain.
+They evaluated different possible computational models of how people generate categories (such as an exemplar-based "copy-and-tweak" model) and found that a hierarchical Bayesian model (which generates categories by abstracting the distributional structure of the domain) provided the strongest account of the behavioral results.
 
-%JLA: Add (a), (b) and (c) to figure and then use those to refer to each subpanel in the caption. It'll save space and be cleaner.
 \begin{figure*}
     \begin{center}
-    \inputpgf{figs/}{example-prob-spaces.pgf}
+    \inputpgf{figs/}{example.spaces.pgf}
     \caption{PACKER generation of a category `B' example, following exposure to one member of category `A' and one member of category `B'.
-\textit{Left}: Predictions given $\{\phi = -1$, $\gamma = 0\}$ (contrast influence only).
-\textit{Center}: Predictions given $\{\phi = 0$, $\gamma = 1\}$ (target influence only).
-\textit{Right}: Predictions given $\{\phi = -1$, $\gamma = 1\}$ (both constraints considered).
+\textit{(a)} Predictions based on contrast similarity only $\{\phi = -1$, $\gamma = 0\}$.
+\textit{(b)} Predictions based on target similarity only $\{\phi = 0$, $\gamma = 1\}$.
+\textit{(c)} Predictions with both constraints considered $\{\phi = -1$, $\gamma = 1\}$.
 }
@@
 
-In this paper, we introduce a novel exemplar-based approach to category generation, the PACKER model (\textit{Producing Alike and Contrasting Knowledge using Exemplar Representations}), which creates new categories by balancing two constraints: (1) new categories should be different from known categories (minimizing between-class similarity), and (2) new categories should be internally coherent (maximizing within-class similarity).
-As such, PACKER is a significant departure from previous accounts of generation -- rather than proposing that people create new categories that are mostly similar to related categories, PACKER creates new categories that are different from related categories.
+In this paper, we introduce a novel exemplar-based approach to category generation, PACKER (\textit{Producing Alike and Contrasting Knowledge using Exemplar Representations}), which creates  categories by balancing two constraints: (1) new categories should be different from known categories (minimizing between-class similarity), and (2) new categories should be internally coherent (maximizing within-class similarity).
+As such, PACKER is a significant departure from previous accounts of generation -- rather than proposing that people create categories by abstracting and re-using knowledge of related categories, PACKER first considers how the generated category should \textit{differ} from related categories.
 Further, it does so using the well-studied mechanics of exemplar representations and therefore possesses a rich connection to the wider body of research on category learning.
@@
 
-The PACKER model is an extension of the prominent Generalized Context Model of category learning (GCM; \citealp{nosofsky1984abbr}).
+The PACKER model is an extension of the prominent Generalized Context Model of category learning (GCM; \citealp{nosofsky1984choice}).
 The model assumes that each category is represented by a collection of exemplars within in a $k$-dimensional psychological space, and that generation is constrained by both similarity to members of the target category (the category in which a stimulus is being generated) as well as similarity to members of other categories.
@@
 
-To generate a new example given examples from contrast and target categories, the model considers both the similarity to examples from the contrast category as well as the similarity to examples in the target category.
+To generate a new example, the model considers both the similarity to examples from contrast categories as well as similarity to examples in the target category.
 The aggregated similarity $a$ between generation candidate $y$ and the model's stored exemplars $x$ can be computed as $a(y, x) = \sum_j{f(x_j) s(y, x_j)}$, where $f(x_j)$ is a function specifying the extent to which each exemplar contributes to the target category.
@@
 $\phi$ and $\gamma$ are free parameters ($-\infty \leq \phi, \gamma \leq \infty$) controlling the contribution of contrast- and target-category similarity, respectively.
-Larger absolute values for either parameter produce greater consideration of that exemplar for generating the target category, with values of 0 producing no effect.
-A negative value for $f_j$ produces a `repelling' effect (target exemplars are less likely to be generated nearby $x_j$).
-Conversely, a positive value for $f_j$ produces a `pulling' effect (target exemplars are more likely to be generated nearby $x_j$).
+Larger absolute values for either parameter produce greater consideration of those exemplars, with values of 0 producing no effect.
+A negative value for $f_j$ produces a `repelling' effect (exemplars are less likely to be generated nearby $x_j$).
+Conversely, a positive value for $f_j$ produces a `pulling' effect (exemplars are more likely to be generated nearby $x_j$).
 
 Recall that we proposed that categories are generated by balancing two constraints: that new categories should be different from existing categories, and exemplars belonging to the same category should be similar to one another.
-This is encoded in PACKER when $\phi < 0$, and $\gamma > 0$.
-Negative values for $\phi$ encourages $y$ to be distant from the contrast category (as similarity to contrast category exemplars are subtracted during aggregation).
+This is realized in PACKER when $\phi < 0$, and $\gamma > 0$.
+Negative values for $\phi$ encourage $y$ to be distant from the contrast category (as similarity to contrast category exemplars are subtracted during aggregation).
 Positive values for $\gamma$ encourage $y$ to be close to other exemplars of the target category.
@@
 Previous models of category generation focus on capturing the tendency for people to produce new categories that have similar distributional properties to existing categories.
-To the best of our knowledge, \citet{jern2013probabilistic} were the first to analyze behavioral results quantitatively to computational models.
+To the best of our knowledge, \citet{jern2013probabilistic} were the first to quantitatively evaluate computational models of generation.
 Based on their work, we describe two alternative models: a formalization of the \textit{copy-and-tweak} hypothesis \citep{ward2002role,ward1995s}, and the \textit{hierarchical sampling} proposed by \citet{jern2013probabilistic}.
@@
 \subsection{Copy-and-Tweak}
-%JLA: PLEASE CHECK WHAT I WROTE TO MAKE SURE IT"S TRUE OF THEIR MODEL 
-The \textit{copy-and-tweak} model proposes that generation is a two-part process: First, learners retrieve an observation from memory, and then they tweak it to be different from the retrieved observation.
-\citet{jern2013probabilistic} interpreted this proposal in terms of an exemplar model, where participants score potential new exemplars as the summed exponentiated similarity of the new exemplar to every exemplar from the category, and sample from the normalized score distribution.
-Given that this was intended for generating new exemplars within known category, we implemented a different formalization of the copy-and-tweak hypothesis.
+The \textit{copy-and-tweak} model proposes that generation is a two-part process: First, learners retrieve an observation of the target class from memory, and then they tweak it make something new.
+\citet{jern2013probabilistic} interpreted this proposal in terms of an exemplar model using the GCM \citep{nosofsky1984choice}.
+Formally, their model is equivalent to PACKER with $\{\phi = 0$, $\gamma = 1\}$ (see Figure \ref{fig:example-prob-spaces}).
+To allow for more flexible performance, and to explicitly formalize the `tweak' procedure, we slightly generalized their implementation.
 
-In our implementation of the copy-and-tweak model, the probability that a given exemplar $z$ is retrieved from memory is $p(z) = \exp(f_z) / \sum_z{ \exp(f_z) }$, where $f_z$ is a function specifying each item's relative chance of being selected.
-In our simulations, we set $f_z$ to be $-\lambda$ for members of contrast categories, and $f_z = \lambda$ for members of the target category.
-The resulting free parameter $\lambda$ ($>0$) thus controls the relative chances that a member of the target category will be retrieved as the source.
-%\citet{jern2013probabilistic} did not implement any mechanisms to differentially weigh retrieval probabilities, and so our model encompasses theirs as a special case when $\lambda = 0$ (uniform probabilities).
-JLA: (i reread their model and i don't think this is quite true)
-
+In our copy-and-tweak model, all observations $z$ from the target class are equally likely to be retrieved as the source: $p(z) = 1 / n$, where $n$ is the number of stored examples in the target class.
+If $n=0$, learners are assumed to generate an item at random.
 After a source exemplar is retrieved, the similarity between generation candidates $y$ and the retrieved exemplar is computed as per Equation \ref{eq:similarity}.
@@
 This ensures the source exemplar is tweaked "enough".
-When $\tau=1$, the threshold has no effect, which results in a model similar to \citet{jern2013probabilistic}.
+When $\tau=1$, the threshold has no effect, and our model is equivalent to that of \citet{jern2013probabilistic}.
 To obtain predictions not depending on a given source example, the model's predictions can be aggregated over all possible sources: $p(y) = \sum_z{p(z)p(y|z)}$.
-%Thus, the probability that $y$ will be selected is a function of its generation probability given each of the sources, and the probability each source will be retrieved.
 
@@
 
-We assume that each category's exemplars are distributed according to a multivariate normal distribution with parameters ($\mu, \Sigma$).\footnote{Following \citet{jern2013probabilistic} stimuli were put through a logistic function to constrain the domain to be between 0 and 1.} We assume that each category's covariance matrix is inverse-Wishart distributed with parameters ($v$, $\kappa,$ and $\Sigma_D$).\footnote{Note that \citet{jern2013probabilistic}'s model is slightly different, as they used a non-conjugate model.
+We assume that each category's exemplars are distributed according to a multivariate normal distribution with parameters ($\mu, \Sigma$).
+We assume that each category's covariance matrix is inverse-Wishart distributed with parameters ($v$, $\kappa,$ and $\Sigma_D$).\footnote{Note that \citet{jern2013probabilistic}'s model is slightly different, as they used a non-conjugate model.
 Their model acts very similar to our version of it and receives comparable fits.} $\Sigma_D$ is the covariance matrix shared between categories.
@@
 where $\nu$ ($\nu>k-1$) is an additional free parameter (from the Inverse-Wishart prior on $\Sigma_B$) weighting the importance of $\Sigma_{D}$.
-When the target category has no members, $\Sigma_B = \Sigma_D$.
+When the target category has no members (i.e., $n_B = 0$), items are generated at random.
 
@@
 %JLA: There's something funky going on with the notation. I got confused by your noation, so hopefully you can update what i did to be consistent with the rest.
+
 \section{Behavioral Experiment}
-A novel prediction of PACKER is that the contrast category's location within stimulus space should influence the generated category -- the generated category should be repulsed from the contrast category.
-The other models predict the location of the contrast category within stimulus space should not influence the generated category.
-This experiment tests this prediction and quantitatively evaluates the three models.
-
-%Whereas the copy-and-tweak and hierarchical sampling models both assume explicit mechanisms to re-purpose knowledge about known categories, the PACKER model forgoes these mechanisms entirely.
-Thus, whereas the distributional structure of observed categories will be the primary determinant of generation in the copy-and-tweak and hierarchical sampling models, these properties should only affect PACKER's generation insofar as they constrain inter-exemplar similarities.
-Instead, the \textit{location} of known categories in the domain is crucial because it constrains remaining possible locations for new categories.
-
-The behavioral experiment described below was designed to test this key prediction:  Does the \textit{location} of contrast categories, influence generation?
-The experiment follows the paradigm developed by \citet{jern2013probabilistic}: first, participants learn members of a known category (`Alpha', or `A'), and are then asked to generate exemplars belonging to a new category (`Beta', or `B').
-We developed two Alpha categories (see Figure \ref{fig:middle-bottom-conditions}).
-We did so by creating one of the Bottom Alpha category as a tight cluster and formed the Middle Alpha category by translating it to the center of stimulus space.
-As the two categories only differ in their location, the two categories should not differ in their distributional information.
-
-%, instantiating the two conditions of the experiment.
-In both conditions, members of the Alpha category are tightly clustered, with equal feature variance and no feature correlations.
-Our manipulation is fairly `weak': the conditions have only a slight difference in the y-axis position of the Alpha category.
-In the `Middle' condition the Alpha category is placed in the center of the space, in the `Bottom' condition the Alpha category is placed in the bottom-center of the space.
 
@@
 
+The copy-and-tweak and hierarchical sampling models were designed to explain effects of prior knowledge on the structure of categories, and so these models assume little about where in stimulus space categories are likely to be generated.
+Indeed, when there are no known examples of the target category, both models assume that generation is random.
+PACKER is thus unique in its prediction that contrast categories should influence both the structure and location of generated categories.
+The behavioral experiment described below was designed to test this key prediction.
+
+The experiment follows the paradigm developed by \citet{jern2013probabilistic}: first, participants learn members of a known category (`Alpha', or `A'), and are then asked to generate exemplars belonging to a new category (`Beta', or `B').
+We developed two Alpha categories (see Figure \ref{fig:middle-bottom-conditions}): the `Bottom' Alpha category is a tight cluster in the bottom-center of the space, and the `Middle' Alpha category is identical except that it lies in the center of stimulus space.
+The two categories only differ in their location, not in  distributional structure.
+
+%Whereas the copy-and-tweak and hierarchical sampling models both assume explicit mechanisms to re-purpose knowledge about known categories, the PACKER model forgoes these mechanisms entirely.
+Thus, whereas the distributional structure of observed categories will be the primary determinant of generation in the copy-and-tweak and hierarchical sampling models, these properties should only affect PACKER's generation insofar as they constrain inter-exemplar similarities.
+Instead, the \textit{location} of known categories in the domain is crucial because it constrains remaining possible locations for new categories.
+
+%, instantiating the two conditions of the experiment.
+In both conditions, members of the Alpha category are tightly clustered, with equal feature variance and no feature correlations.
+Our manipulation is fairly `weak': the conditions have only a slight difference in the y-axis position of the Alpha category.
+In the `Middle' condition the Alpha category is placed in the center of the space, in the `Bottom' condition the Alpha category is placed in the bottom-center of the space.
+
 Although our manipulation is minimal, the PACKER model is capable of predicting strong between-condition differences.
@@
 
-\subsubsection{Participants \& Materials}
-We recruited 122 participants from Amazon Mechanical Turk from the US equally assigned to each condition.
-Stimuli were squares varying in color (RGB 25--230) and size (3.0--5.8cm).
-The assignment of perceptual features (color, size) to axes of the domain space (x, y) was counterbalanced across participants.
-
-\begin{figure}[ht!]
+\begin{figure}
     \begin{center}
@@
 \end{figure}
+
+\subsubsection{Participants \& Materials}
+We recruited 122 participants from Amazon Mechanical Turk from the US equally assigned to each condition.
+Stimuli were squares varying in color (RGB 25--230) and size (3.0--5.8cm).
+The assignment of perceptual features (color, size) to axes of the domain space (x, y) was counterbalanced across participants.
 
@@
 Several sample Beta categories are depicted in Figure \ref{fig:beta.samples}.
-As is evident in Figure \ref{fig:beta.samples}, we observed large individual differences in generation strategy: Whereas some participants generated all four Beta examples within a narrow y-axis range, others generated Beta examples along a wide range.
+As is evident in Figure \ref{fig:beta.samples}, we observed broad individual differences in generation strategy: Whereas some participants generated all four Beta examples within a narrow y-axis range, others generated Beta examples along a wide range.
 
@@
 \begin{center} 
+\caption{Behavioral results.} 
 \label{table:subset-table} 
@@
 \end{tabular}
-\caption{Behavioral results.} 
 \end{center} 
 \end{table}
+
 To evaluate PACKER's other predictions, we computed the number of exemplars produced at different distances to the center of the Alpha category.
-These data, depicted in the top panel of Figure \ref{fig:distance.figs} reveal a strong preference for stimuli that are dissimilar to the members of the Alpha category: maximally distant items were by far the most frequently generated.
+These data, depicted in the left panel of Figure \ref{fig:distance.figs} reveal a strong preference for stimuli that are dissimilar to the members of the Alpha category: maximally distant items were by far the most frequently generated.
 
 Finally, we computed for each participant the average distance between exemplars belonging to the same and opposite categories.
-These data (depicted in the bottom panel of Figure \ref{fig:distance.figs}) show that, as observed by \citet{ward1994structured}, most people generated Beta categories in which members are closer to one another than they are to members of the Alpha category (i.e., more between- than within-category distance).
+These data (depicted in the right panel of Figure \ref{fig:distance.figs}) show that, as observed by \citet{ward1994structured}, most people generated Beta categories in which members are closer to one another than they are to members of the Alpha category (i.e., more between- than within-category distance).
 We did however, observe a notable subset of individuals with greater within-class distance.
 These individuals tended to adopt a `corners' approach, in which a Beta examples were placed almost exclusively in the corners of the space.
+
+\subsection{Summary}
+
+Our results support PACKER's predictions: People tend to generate items that are dissimilar from the contrast category and similar to the target category (with some exceptions, see Figure \ref{fig:distance.figs}).
+By consequence, we observed considerable differences in generation between the Middle and Bottom conditions: participants in the Bottom condition were less likely to use the bottom row of the stimulus space for generation, and participants in the Middle condition were more likely to create categories spanning the entire y-axis (utilizing the top and bottom row of the space).
+This latter result is especially interesting as it conflicts with previous results: Qualitatively different types of categories are generated (with very different distributional information), depending on only the location of the Alphas in the domain.
+
+The results described above are somewhat commonsense: they simply demonstrate that the location (rather than distributional structure) of existing categories imposes constraints on generation because people tend to generate examples in areas not occupied by existing categories.
+This principle, however, is not predicted by existing models of generation -- these models are instead designed to explain distributional correspondences between generated and existing categories.
 
@@
     \caption{Behavioral results.
-\textit{Top}: Frequency of exemplar generation as a function of distance from the Alpha category normalized by the maximum possible distance.
-\textit{Bottom}: Within- vs.
+\textit{Left}: Frequency of exemplar generation as a function of distance from the Alpha category normalized by the maximum possible distance.
+\textit{Right}: Within- vs.
 between-category distance for every participant.
@@
 
-\subsection{Summary}
-Our results support PACKER's predictions: People tend to generate items that are dissimilar from the contrast category Alpha and similar to the target category Beta (with some exceptions, see Figure \ref{fig:distance.figs}).
-By consequence, we observed considerable differences in generation between the Middle and Bottom conditions: participants in the Bottom condition were less likely to use the bottom row of the stimulus space for generation, and participants in the Middle condition were more likely to create categories spanning the entire y-axis (utilizing the top and bottom row of the space).
-This latter result is especially interesting as it conflicts with previous results: Qualitatively different types of categories are generated (with very different distributional information), depending on only the location of the Alphas in the domain.
-
-The results described above are somewhat commonsense: they simply demonstrate that the location (rather than distributional structure) of existing categories imposes constraints on generation because people tend to generate examples in areas not occupied by existing categories.
-This principle, however, is not predicted by existing models of generation -- these models are instead designed to explain distributional correspondences between generated and existing categories.
-
 \section{Model Evaluation}
-In this section, we evaluate the model fits to our results.
 To obtain an overall sense of each model's ability to explain our results, we fit each model by maximizing the log-likelihood of the model's predictions of the human results.
-Four free parameters were fitted in each model.
-The $c$, $\phi$, $\gamma$, and $\theta$ parameters were fitted for PACKER; $c$, $\lambda$, $\tau$, and $\theta$ were fitted for the copy-and-tweak model, and $\kappa$, $\rho$, $\nu$, and $\theta$ were fitted for the hierarchical sampling model.
+The $c$, $\phi$, $\gamma$, and $\theta$ parameters were fitted for PACKER; $c$, $\tau$, and $\theta$ were fitted for the copy-and-tweak model, and $\kappa$, $\rho$, $\nu$, and $\theta$ were fitted for the hierarchical sampling model.
 Note that each model possess a $\theta$ parameter fulfilling the same role (response determinism), and PACKER shares with copy-and-tweak a specificity parameter, $c$.
@@
 A similar pattern of results was obtained when we only considered the second to fourth exemplars generated by each participant.
-%JLA this is fine
-\begin{table}[ht!]
+
+Our model-fitting results follow directly from the design of our experiment.
+Because the copy-and-tweak and hierarchical sampling models are not influenced by the location of contrast categories within the space, they do not capture the broad tendency for generated categories to be dissimilar to existing classes.
+
+\begin{table}
 \centering
@@
  & & \textbf{Sampling} \\ \hline
- $L = -1733$ & $L = -2040$ & $L = -2055$  \\ 
- $c=0.564$ & $c=9.445$  & $\kappa<0.001$\\
-$\phi=-13.608$& $\lambda=15.115$ & $\rho=1.001$ \\ 
-$\gamma=12.006$  & $\tau=0.415$ & $\nu=0.451$  \\ 
- $\theta=0.259$  & $\theta=7.068$ & $\theta=6.137$  \\ 
+ $AIC = 3476$ & $AIC = 3937$ & $AIC = 4061$  \\ 
+ $c=0.565$ & $c=4.675$  & $\kappa<0.001$\\
+$\phi=-4.814$ & $\tau=0.895$ & $\nu=1.001$ \\ 
+$\gamma=4.250$ & $\theta=5.601$ &  $\rho=1.220$  \\ 
+ $\theta=0.731$ &  & $\theta = 7.147$  \\ 
 \end{tabular}
 \end{table}
-
-The model-fitting results above follow directly from the design of our experiment.
-Because distributional structure is the primary determinant of generation in the copy-and-tweak and hierarchical sampling models, those models have no basis to predict any differences between Middle and Bottom conditions (which do not differ in distributional structure).
 
@@
 %While it is important that any model be able to explain different types of performance with the same parameter values, these radical differences in generation are difficult to account for under a single set of parameters.
-Our behavioral results showed that, in general, generated categories consist of item that are dissimilar to members of opposite categories, and similar to members of the their own category.
-There were four typical patterns of Beta categories: tightly clustered, four corners, `column'-like categories (varying along the y-axis but not the x-axis), others created `row'-like categories (see Figures \ref{fig:beta.samples} and \ref{fig:distance.figs}).
-Here, we explore the role of PACKER's different attention weights ($w_1$ and $w_2$) in producing row-like or column-like categories.
+Our behavioral results showed that, in general, generated categories consist of items that are dissimilar to members of opposite categories, and similar to members of the their own category.
+However, beyond this broad description of performance, we observed a great deal of individual differences in generation style.
+Manually inspecting the data reveals four typical patterns (see Figure \ref{fig:beta.samples}): `corners' categories with one Beta example in each corner of the space, tight clusters, `column'-like categories varying along the y-axis but not the x-axis, and `row'-like categories varying along the x-axis but not the y-axis.
+Here, we explore the role of PACKER's attention weights ($w_1$ and $w_2$) in producing row-like or column-like categories.
 
-%JLA: Move PACKER to be between Behavioral and Copy and Tweak. Also, I think this will take up less space as a row figure.
-\begin{figure}[ht!]
+\begin{figure*}
     \begin{center}
@@
     \end{center}
-\end{figure}
+\end{figure*}
 
@@
 Taking the average of these range differences yields a gradient describing how, on average, categories were distributed for each stimulus.
-These data, depicted in the top panels of Figure \ref{fig:range-diff-gradient} reveal a highly systematic relationship between category structure and category location.
+These data, depicted in Figure \ref{fig:range-diff-gradient}, reveal a highly systematic relationship between category structure and category location.
 Whereas column-like categories more often include stimuli to the left or right of the Alpha class, row-like categories appear above and below the Alpha class.
@@
 To simulate this finding, we set the attention weight parameters in PACKER and copy-and-tweak separately per participant.
-To simulate the hierarchical sampling model we set the domain covariance prior, $\Sigma_0$, separately for each participant.
 The other free parameters were set as in Table \ref{table:model-fits}.
@@
 Thus, the Alpha and Beta categories are assumed to be distinct along dimensions that the Betas do not vary on.
+To simulate the hierarchical sampling model we set the domain covariance prior, $\Sigma_0$, proportional to the range of each feature (not using the inverse).
 We then simulated 50 Beta categories with each participant's weighting scheme to obtain a sense of how the relative importance of each dimension affects what types of categories are generated and where they are generated.
@@
 
-The hierarchical sampling and copy-and-tweak models do not often create categories with strong range differentials.
-However, when they do, the categories tend to show the opposite pattern: Columns are placed above and below the Alphas, rows are to the side of the Alphas.
-In the hierarchical sampling model, the position of contrast category is not considered when generating new items.
-Only the distributional structure of the contrast category and items already in the target category are considered.
-The model simulates row/column categories by assuming there is more x- or y-axis variance in the domain, essentially stretching the generated category distribution along that axis, and shrinking it along the opposite axis.
-However, because the Beta category distribution is centered in the middle of the space prior to the first trial, the model is most likely to generate items directly to the side of the Alphas if the x axis is weighted, or above/below the Alphas if the y-axis is weighted.
-Future generated items will follow from this distribution, and thus weighting the x-axis results row categories that possess the same y-axis position as the Alphas.
-Conversely, weighting the y-axis results in column categories that share the the same x-axis position as the Alphas.
-
-The copy-and-tweak model operates by a similar principle.
-When, for example, the x-axis is weighted, the range of examples exceeding the similarity threshold ($\tau$) becomes much sharper along the x-axis (items need to be much closer to be considered at all similar), and broader along the y-axis.
-On the first trial, one of the Alphas will be used as a source, and because the similarity gradient is so sharp along the x-axis, the model will generate a Beta that is above or below the source.
-Future examples will follow this same pattern, resulting in a column category that differs from the alphas along the y-axis, as opposed to the x-axis.
+Although differentially weighting the features results in different types of categories from the hierarchical sampling and copy-and-tweak models, the location of the Alpha category does not affect where items will be generated by these models.
+Thus, row- and column-like categories are not systematically generated in different areas of the stimulus space, resulting in the uniform predictions shown in Figure \ref{fig:range-diff-gradient}.
 
@@
 Likewise, we found that the \textit{location} of contrast categories shapes generation by imposing constraints on the areas of space that remain available for a new category.
-Our formal simulations reveal that PACKER provides a strong account of our observations.
+These effects are somewhat commonsense in that they simply reveal that people tend to generate items that are different from what already exists.
+However, existing models \citep[see][]{jern2013probabilistic} assume little about the role of contrast in category generation (in these models, the location of generated categories is randomly chosen).
+Our formal simulations reveal that PACKER provides a strong account of the phenomena observed behaviorally.
 

### 2017-01-29 Experiments/middle-bottom/analysis/plot.conditions.py

@@ -12,7 +12,7 @@ stimuli = pd.read_sql_query("SELECT * from stimuli", con).as_matrix()
 alphas = pd.read_sql_query("SELECT * from alphas", con)
 con.close()
 
-f, ax = plt.subplots(1, 2, figsize=(3.5, 1.75))
+f, ax = plt.subplots(1, 2, figsize=(3.5, 1.25))
 for i, k  in enumerate(list(alphas)):
 	h = ax[i]
 	funcs.plotclasses(h, stimuli, alphas[k], [])		
@@ -23,4 +23,4 @@ for i, k  in enumerate(list(alphas)):
 f.savefig('conditions.png', bbox_inches='tight', transparent=False)
 
 path = '../../../Manuscripts/cogsci-2017/figs/middle-bottom-conditions.pgf'
-# funcs.save_as_pgf(f, path)
+funcs.save_as_pgf(f, path)

### 2017-01-29 Experiments/middle-bottom/modeling/plot.example.spaces.py

@@ -6,6 +6,7 @@ execfile('Imports.py')
 from Modules.Classes import Packer
 import Modules.Funcs as funcs
 
+
 vals = np.linspace(-1, 1, 100).tolist()
 space = np.fliplr(funcs.cartesian([vals, vals]))
 
@@ -26,31 +27,31 @@ prob_spaces = {
     'Combination': Packer(cats,pos_neg)
 }
 
-f, ax = plt.subplots(1,3, figsize = (7.5, 2.5))
-counter = 0
+f, ax = plt.subplots(1,3, figsize = (7.5, 1.5))
 
-for k in ['Contrast Influence', 'Target Influence', 'Combination']:
+prefix = ['(a)','(b)','(c)']
+for i, k in enumerate(['Contrast Influence', 'Target Influence', 'Combination']):
     m = prob_spaces[k] 
-    h = ax[counter]
+    h = ax[i]
 
     ps = m.get_generation_ps(space,1)
     print max(ps)
 
     g = funcs.gradientroll(ps, 'roll')[:,:,0]
     im = funcs.plotgradient(h, g, A, B, cmap = 'Blues', beta_col = 'w')
-    h.set_title(k, fontsize = 11)
 
-    counter += 1
+    title = prefix[i] + ' ' + k
+    h.set_title(title, fontsize = 11)
 
 
 # add colorbar
 f.subplots_adjust(right=0.8)
-cbar = f.add_axes([0.83, 0.225, 0.03, 0.54])
+cbar = f.add_axes([0.83, 0.16, 0.03, 0.66])
 f.colorbar(im, cax=cbar, ticks = [0,0.000276585446215])
 cbar.set_yticklabels(['Lowest\nProbability', 'Greatest\nProbability'])
 cbar.tick_params(length = 0)
 
-f.savefig('example-prob-spaces.png', bbox_inches='tight', transparent=False)
+f.savefig('example.spaces.png', bbox_inches='tight', transparent=False)
 
-path = '../../../Manuscripts/cogsci-2017/figs/example-prob-spaces.pgf'
-# funcs.save_as_pgf(f, path)
+path = '../../../Manuscripts/cogsci-2017/figs/example.spaces.pgf'
+funcs.save_as_pgf(f, path)

### 2017-01-29 Modules/Classes/Optimize.py

@@ -83,8 +83,13 @@ def hillclimber(model_obj, init_params, dataset, stimuli, options):
 	print '\tIterations = ' + str(res.nit)
 	print '\tMessage = ' + str(res.message)
 
-	for k, v in zip(model_obj.parameter_names, res.x):
+	final_obj = model_obj(stimuli, res.x)
+	for k in final_obj.parameter_names:
+		v = getattr(final_obj, k)
 		print '\t' + k + ' = ' + str(v) + ','
 	print '\tLogLike = ' + str(res.fun)
+
+	AIC = 2.0*len(init_params) - 2.0* (-1.0 * res.fun)
+	print '\tAIC = ' + str(AIC)
 
 	return res

### 2017-01-30 Manuscripts/cogsci-2017/paper.tex
6 pagesgit pull

@@
 
-The creative generation of new concepts is an intriguing yet understudied topic in cognitive science.
-In this paper, we present a novel exemplar-based approach to category generation: PACKER (\textit{Producing Alike and Contrasting Knowledge using Exemplar Representations}).
-PACKER's core design assumptions are that (\textit{a}) categories are represented as exemplars in a multidimensional psychological space, (\textit{b}) generated items should be similar to exemplars of the same category, and (\textit{c}) generated categories should be dissimilar to existing categories.
-A behavioral study and formal simulations reveal broad support for PACKER's predictions.
+Generating new concepts is an intriguing yet understudied topic in cognitive science.
+In this paper, we present a novel exemplar model of category generation: PACKER (\textit{Producing Alike and Contrasting Knowledge using Exemplar Representations}).
+PACKER's core design assumptions are (1) categories are represented as exemplars in a multidimensional psychological space, (2) generated items should be similar to exemplars of the same category, and (3) generated categories should be dissimilar to existing categories.
+A behavioral study and formal simulations broadly support PACKER's predictions.
 
 \textbf{Keywords:} 
-Categorization, exemplar models, category generation, creative cognition.
+Categorization, exemplar models, category generation, creative cognition, computational modeling.
 \end{abstract}
@@
 
-Recent work has proposed and tested computational models of the processes underlying category generation.
-\citet{jern2013probabilistic} trained participants on a set of experimenter-defined categories composed of exemplars within an artificial three-dimensional domain (e.g., 2D shapes varying in size, hue, and saturation).
+Recent work has proposed and tested formal models of category generation.
+\citet{jern2013probabilistic} trained participants on experimenter-defined categories composed of exemplars within an artificial three-dimensional domain.
 After a short training phase, participants were asked to generate exemplars from a new category.
@@
 In the sections below, we formally describe the PACKER model and explore its predictions in a behavioral experiment.
-We compare its performance to a copy-and-tweak and hierarchical Bayesian model by examining their fits to aggregate results and individual differences.
+We compare its performance to copy-and-tweak and hierarchical Bayesian models by examining their fits to aggregate results and individual differences.
 
@@
 
-The PACKER model is an extension of the prominent Generalized Context Model of category learning (GCM; \citealp{nosofsky1984choice}).
-The model assumes that each category is represented by a collection of exemplars within in a $k$-dimensional psychological space, and that generation is constrained by both similarity to members of the target category (the category in which a stimulus is being generated) as well as similarity to members of other categories.
+The PACKER model is an extension of the Generalized Context Model of category learning (GCM; \citealp{nosofsky1984choice}).
+It assumes that each category is encoded by a set of exemplars within a $k$-dimensional psychological space, and that generation is constrained by both similarity to members of the target category (the category in which a stimulus is being generated) as well as similarity to members of other categories.
 
@@
 \end{equation}
-where $w_k$ is the attention weighting of dimension $k$ ($w_k \geq 0$ and $\sum_k{w_k} = 1$), accounting for the relative importance of each dimension when calculating similarity, and $c$ ($c>0$) is a specificity parameter controlling the spread of exemplar generalization.
-For simplicity, most simulations will use uniform attention weights.
-We will use non-uniform attention weights in our discussion of individual differences.
+where $w_k$ is the attention weighting of dimension $k$ ($w_k \geq 0$ and $\sum_k{w_k} = 1$), accounting for the relative importance of each dimension in similarity calculations, and $c$ ($c>0$) is a specificity parameter controlling the spread of exemplar generalization.
+For simplicity, simulations will use uniform attention weights, except in our discussion of individual differences.
 
-To generate a new example, the model considers both the similarity to examples from contrast categories as well as similarity to examples in the target category.
-The aggregated similarity $a$ between generation candidate $y$ and the model's stored exemplars $x$ can be computed as $a(y, x) = \sum_j{f(x_j) s(y, x_j)}$, where $f(x_j)$ is a function specifying the extent to which each exemplar contributes to the target category.
-For brevity, we will abbreviate $f(x_j)$ as $f_j$.
+To generate a new example, the model considers both the similarity to examples from contrast categories as well as the similarity to examples (if any exist) in the target category.
+The aggregated similarity $a$ between generation candidate $y$ and stored exemplars $x$ is $a(y, x) = \sum_j{f(x_j) s(y, x_j)}$, where $f(x_j)$ is a function specifying the extent to which each exemplar contributes to the target category.
+For brevity, we abbreviate $f(x_j)$ as $f_j$.
 PACKER sets $f_j$ depending on exemplar $j$'s category membership: $f_j = \phi$ if $x_j$ is a member of the contrast category, and $f_j = \gamma$ if $x_j$ is a member of the target category.
-$\phi$ and $\gamma$ are free parameters ($-\infty \leq \phi, \gamma \leq \infty$) controlling the contribution of contrast- and target-category similarity, respectively.
-Larger absolute values for either parameter produce greater consideration of those exemplars, with values of 0 producing no effect.
+$\phi$ and $\gamma$ are free parameters ($-\infty \leq \phi, \gamma \leq \infty$) controlling the contribution of contrast- and target-category similarity, respectively (see Fig \ref{fig:example-prob-spaces}).
+Larger absolute values result in greater consideration of those exemplars, with values of 0 eliminating their effect.
 A negative value for $f_j$ produces a `repelling' effect (exemplars are less likely to be generated nearby $x_j$).
-Conversely, a positive value for $f_j$ produces a `pulling' effect (exemplars are more likely to be generated nearby $x_j$).
+Conversely, a positive value for $f_j$ produces an `attracting' effect (exemplars are more likely to be generated nearby $x_j$).
 
-Recall that we proposed that categories are generated by balancing two constraints: that new categories should be different from existing categories, and exemplars belonging to the same category should be similar to one another.
+Recall that we proposed that categories are generated by balancing two constraints: New categories should be different from existing categories, and exemplars belonging to the same category should be similar to one another.
 This is realized in PACKER when $\phi < 0$, and $\gamma > 0$.
@@
 Positive values for $\gamma$ encourage $y$ to be close to other exemplars of the target category.
-When $|\phi| = \gamma$, the "force" of repulsion from the contrast category is equal to the "force" of attraction to the target category.
+When $|\phi| = \gamma$, the repulsion effect from the contrast category is equal to the attraction effect to the target category.
 
@@
 The proposed PACKER model suggests people generate categories by minimizing between-category similarity and maximizing within-category similarity.
-The underlying processes assumed by PACKER are highly similar to those in the GCM, with the only alteration being that PACKER aggregates positive- and negative-valued similarities, rather than only aggregating positive-valued similarities.
-
-In later sections, we will explore the unique predictions yielded by these design principles. First, however, we contrast PACKER with other category generation models.
+The underlying processes assumed by PACKER are highly similar to those in the GCM.
+The main difference is PACKER aggregates positive- and negative-valued similarities, rather than only aggregating positive-valued similarities.
+In later sections, we will explore the unique predictions yielded by these design principles.
+First, however, we contrast PACKER with other category generation models.
 
@@
 Previous models of category generation focus on capturing the tendency for people to produce new categories that have similar distributional properties to existing categories.
-To the best of our knowledge, \citet{jern2013probabilistic} were the first to quantitatively evaluate computational models of generation.
-Based on their work, we describe two alternative models: a formalization of the \textit{copy-and-tweak} hypothesis \citep{ward2002role,ward1995s}, and the \textit{hierarchical sampling} proposed by \citet{jern2013probabilistic}.
+To the best of our knowledge, \citet{jern2013probabilistic} were the first to evaluate computational models of generation.
+Based on their work, we describe two alternative models: a formalization of the \textit{copy-and-tweak} hypothesis \citep{ward1995s}, and the \textit{hierarchical sampling} hypothesis \citep{jern2013probabilistic}.
 
 \subsection{Copy-and-Tweak}
-The \textit{copy-and-tweak} model proposes that generation is a two-part process: First, learners retrieve an observation of the target class from memory, and then they tweak it make something new.
+The \textit{copy-and-tweak} model proposes that generation is a two-part process: First, learners retrieve an observation of the target class from memory, and then they tweak it to make something new.
 \citet{jern2013probabilistic} interpreted this proposal in terms of an exemplar model using the GCM \citep{nosofsky1984choice}.
@@
 
-In our copy-and-tweak model, all observations $z$ from the target class are equally likely to be retrieved as the source: $p(z) = 1 / n$, where $n$ is the number of stored examples in the target class.
-If $n=0$, learners are assumed to generate an item at random.
+In our copy-and-tweak model, all observations $z$ from the target class are equally likely to be retrieved as the source.
+The first exemplar is generated uniformly at random over stimulus space.
 After a source exemplar is retrieved, the similarity between generation candidates $y$ and the retrieved exemplar is computed as per Equation \ref{eq:similarity}.
-The model's goal is to generate an item that is similar to $z$, but does not exceed a given similarity threshold: The generated item should be similar, but not too similar to $z$.
+The model's goal is to generate an item that is similar to $z$, but does not exceed a given similarity threshold: The generated item should be copied, but tweaked.
 Formally, the probability that candidate exemplar $y$ will be generated based on a source exemplar $z$ is
@@
 where $\theta$ is a response determinism parameter, $\tau$ is the similarity threshold, and $I(\cdot)$ is the indicator function, which returns 1 when it is passed a true expression and 0 otherwise.
-So, $I\left(s(y,z) \leq \tau\right)$ is 1 when the candidate exemplar $y$ is far enough away from the source exemplar $z$.
-This ensures the source exemplar is tweaked "enough".
+So, $I\left(s(y,z) \leq \tau\right)$ is 1 when the candidate exemplar $y$ is tweaked "enough".
 When $\tau=1$, the threshold has no effect, and our model is equivalent to that of \citet{jern2013probabilistic}.
@@
 Based on several results inconsistent with the copy-and-tweak account, \citet{jern2013probabilistic} advocated a hierarchical Bayesian model.
-Exemplars of each category were generated from a multivariate Gaussian distribution over the dimensions of stimulus space.
-The mean of each category was independent, but the covariance matrices (encoding feature variances and correlations) were generated from a common prior distribution.
-New categories are generated then by generating a new mean (uniform over stimulus space) and covariance matrix from the common prior distribution.
+Exemplars of each category were generated from a multivariate Normal distribution over the dimensions of stimulus space.
+The mean of each category was independently generated, but the covariance matrix (encoding feature variances and correlations) was generated from a common prior distribution.
+New categories are produced by generating a new mean (uniform over stimulus space) and covariance matrix from the common prior distribution.
 Because the shared prior distribution's parameters were unobserved, a hierarchical Bayesian model uses information from the previous categories (their feature variances and correlations) to generate the covariance matrix of the new category.
 
-We assume that each category's exemplars are distributed according to a multivariate normal distribution with parameters ($\mu, \Sigma$).
-We assume that each category's covariance matrix is inverse-Wishart distributed with parameters ($v$, $\kappa,$ and $\Sigma_D$).\footnote{Note that \citet{jern2013probabilistic}'s model is slightly different, as they used a non-conjugate model.
+Each category's exemplars are assumed to be a multivariate Normal distribution with parameters ($\mu, \Sigma$).
+Each category's covariance matrix is assumed to be inverse-Wishart distributed with parameters ($v$, $\kappa,$ and $\Sigma_D$).\footnote{Note that \citet{jern2013probabilistic}'s model is slightly different, as they used a non-conjugate model.
 Their model acts very similar to our version of it and receives comparable fits.} $\Sigma_D$ is the covariance matrix shared between categories.
-We assumed the shared covariance matrix $\Sigma_D$ was generated from a Wishart distribution (for conjugacy) with parameters $v_0$, $\kappa_0$, and $\Sigma_0$.
-We set $\nu_0 = \kappa_0 = 1$, and $\Sigma_0 = \rho {\bf I}$, where $\rho$ was a free parameter controlling the expected variance of dimensions (dimensions of the shared covariance matrix are expected to be uncorrelated) and ${\bf I}$ is the identity matrix.
+We assume the shared covariance matrix $\Sigma_D$ is generated from a Wishart distribution (for conjugacy) with parameters $v_0$, $\kappa_0$, and $\Sigma_0$.
+We set $\nu_0 = \kappa_0 = 1$, and $\Sigma_0 = \rho {\bf I}$, where $\rho$ is a free parameter controlling the expected variance of dimensions (dimensions of the shared covariance matrix are expected to be uncorrelated) and ${\bf I}$ is the identity matrix.
 
@@
 
-Generated exemplars are then assumed to be drawn from the multivariate normal distribution specified by $(\mu_{B}, \Sigma_{B})$. Thus, $p(y)$ is
+Generated exemplars are drawn from a multivariate Normal distribution specified by $(\mu_{B}, \Sigma_{B})$. Thus, $p(y)$ is
 \begin{equation}
@@
 
-Although our manipulation is minimal, the PACKER model is capable of predicting strong between-condition differences.
+Although our manipulation is minimal, the PACKER model predicts strong between-condition differences.
 PACKER proposes that the nature of the space not occupied by the Alpha category determines where members of the Beta category are likely to be generated.
@@
 Participants were instructed that members of the Beta category could be quite similar or different depending on what they think makes the most sense for the category, but that they were not allowed to make the same example twice.
-As in \citet{jern2013probabilistic}, generation was completed using a sliding-scale interface.
-Participants were presented with two scales that controlled the dimensions of the generated example.
+As in \citet{jern2013probabilistic}, generation was completed using a sliding-scale interfaceTwo scales controlled the dimensions of the generated example.
 An on-screen preview of the example updated whenever one of the features was changed.
@@
 
-Because the conditions differ only in their location along the y-axis (see Figure \ref{fig:middle-bottom-conditions}), we focus on how Beta exemplars are generated above and below the contrast category.
+Because the conditions differ only in their location along the y-axis, we focus on how Beta exemplars are generated above and below the contrast category.
 Several sample Beta categories are depicted in Figure \ref{fig:beta.samples}.
@@
 To evaluate PACKER's other predictions, we computed the number of exemplars produced at different distances to the center of the Alpha category.
-These data, depicted in the left panel of Figure \ref{fig:distance.figs} reveal a strong preference for stimuli that are dissimilar to the members of the Alpha category: maximally distant items were by far the most frequently generated.
+These data (Figure \ref{fig:distance.figs} left) reveal a strong preference for stimuli that are dissimilar to the Alpha category members: maximally distant items were by far the most frequently generated.
 
 Finally, we computed for each participant the average distance between exemplars belonging to the same and opposite categories.
-These data (depicted in the right panel of Figure \ref{fig:distance.figs}) show that, as observed by \citet{ward1994structured}, most people generated Beta categories in which members are closer to one another than they are to members of the Alpha category (i.e., more between- than within-category distance).
+These data (Figure \ref{fig:distance.figs} right) show that, as observed by \citet{ward1994structured}, most people generated Beta categories in which members are closer to one another than they are to members of the Alpha category (i.e., more between- than within-category distance).
 We did however, observe a notable subset of individuals with greater within-class distance.
@@
 
-Our results support PACKER's predictions: People tend to generate items that are dissimilar from the contrast category and similar to the target category (with some exceptions, see Figure \ref{fig:distance.figs}).
-By consequence, we observed considerable differences in generation between the Middle and Bottom conditions: participants in the Bottom condition were less likely to use the bottom row of the stimulus space for generation, and participants in the Middle condition were more likely to create categories spanning the entire y-axis (utilizing the top and bottom row of the space).
-This latter result is especially interesting as it conflicts with previous results: Qualitatively different types of categories are generated (with very different distributional information), depending on only the location of the Alphas in the domain.
+Our results support PACKER's predictions: People tend to generate items that are dissimilar from the contrast category and similar to the target category.
+We observed considerable differences in generation between the Middle and Bottom conditions: Participants in the Bottom condition were less likely to use the bottom row of the stimulus space for generation, and participants in the Middle condition were more likely to create categories spanning the entire y-axis (utilizing the top and bottom row of the space).
+This latter result is especially interesting as it conflicts with previous results: Qualitatively different types of categories were generated, depending only on the location of the Alphas.
 
-The results described above are somewhat commonsense: they simply demonstrate that the location (rather than distributional structure) of existing categories imposes constraints on generation because people tend to generate examples in areas not occupied by existing categories.
-This principle, however, is not predicted by existing models of generation -- these models are instead designed to explain distributional correspondences between generated and existing categories.
+Some aspects of the results described above are somewhat commonsense: They demonstrate that the location of existing categories also imposes constraints on generation because people tend to generate examples in areas not occupied by existing categories.
+This principle, however, is not predicted by existing models of generation -- these models were designed to explain distributional correspondences between generated and existing categories.
 
@@
 
-Our model-fitting results follow directly from the design of our experiment.
-Because the copy-and-tweak and hierarchical sampling models are not influenced by the location of contrast categories within the space, they do not capture the broad tendency for generated categories to be dissimilar to existing classes.
+Our model-fitting results make sense given the design of the experiment.
+As the copy-and-tweak and hierarchical sampling models are not influenced by the location of contrast categories within the space, they do not capture the broad tendency for generating categories dissimilar to existing classes.
 
@@
 %While it is important that any model be able to explain different types of performance with the same parameter values, these radical differences in generation are difficult to account for under a single set of parameters.
-Our behavioral results showed that, in general, generated categories consist of items that are dissimilar to members of opposite categories, and similar to members of the their own category.
-However, beyond this broad description of performance, we observed a great deal of individual differences in generation style.
-Manually inspecting the data reveals four typical patterns (see Figure \ref{fig:beta.samples}): `corners' categories with one Beta example in each corner of the space, tight clusters, `column'-like categories varying along the y-axis but not the x-axis, and `row'-like categories varying along the x-axis but not the y-axis.
+Generally, our behavioral results showed that generated categories consist of items that are dissimilar to members of opposite categories, and similar to members of their own category.
+We also observed a great deal of individual differences in generation style.
+Manually inspecting the data reveals four typical patterns (see Figure \ref{fig:beta.samples}): `corners' categories with one Beta example in each corner of the space, tight clusters, `column'-like categories, and `row'-like categories.
 Here, we explore the role of PACKER's attention weights ($w_1$ and $w_2$) in producing row-like or column-like categories.
@@
 
-For each stimulus in the domain, we computed the difference in range between the features (e.g., $range(size) - range(color)$) across every generated category in which that stimulus was a member.
+For each stimulus in the domain, we computed the difference in range between the features ($range(size) - range(color)$) across every generated category that had the stimulus as member.
 Taking the average of these range differences yields a gradient describing how, on average, categories were distributed for each stimulus.
-These data, depicted in Figure \ref{fig:range-diff-gradient}, reveal a highly systematic relationship between category structure and category location.
+These data (Fig \ref{fig:range-diff-gradient}) reveal a systematic relationship between category structure and category location.
 Whereas column-like categories more often include stimuli to the left or right of the Alpha class, row-like categories appear above and below the Alpha class.
@@
 
-To simulate this finding, we set the attention weight parameters in PACKER and copy-and-tweak separately per participant.
+To simulate this finding, we set the attention weight parameters in PACKER and copy-and-tweak per participant.
 The other free parameters were set as in Table \ref{table:model-fits}.
-While there exist methods to find the optimal attention weights for a given classification \citep[see][]{vanpaemel2012using}, for ease of computation we simply approximated the weights based on the inverse of each feature's range: $w_k \propto range(k)^{-1}$.
+While there exist methods to find the optimal attention weights for a given classification \citep[see][]{vanpaemel2012using}, for simplicity we approximated the weights using the inverse of each feature's range: $w_k \propto range(k)^{-1}$.
 Thus, the Alpha and Beta categories are assumed to be distinct along dimensions that the Betas do not vary on.
-To simulate the hierarchical sampling model we set the domain covariance prior, $\Sigma_0$, proportional to the range of each feature (not using the inverse).
+To simulate the hierarchical sampling model we set the domain covariance prior, $\Sigma_0$, proportional to the range of each feature.
 We then simulated 50 Beta categories with each participant's weighting scheme to obtain a sense of how the relative importance of each dimension affects what types of categories are generated and where they are generated.
@@
 The PACKER model proposes that categories are represented as a collection of exemplars stored in memory, and that members of generated categories should be similar to one another, yet dissimilar to members of opposing categories.
-Given that exemplar models can be viewed as Importance-Sampling approximations of Bayesian models \citep{shi10}, PACKER can be viewed as a rational process model and approximating the expected density of a new category based on a contrast category.
+Exemplar models can be viewed as Importance-Sampling approximations of Bayesian models \citep{shi10}.
+So, PACKER can be viewed as a rational process model, approximating the expected density of a new category based on a contrast category.
 
@@
 Likewise, we found that the \textit{location} of contrast categories shapes generation by imposing constraints on the areas of space that remain available for a new category.
-These effects are somewhat commonsense in that they simply reveal that people tend to generate items that are different from what already exists.
-However, existing models \citep[see][]{jern2013probabilistic} assume little about the role of contrast in category generation (in these models, the location of generated categories is randomly chosen).
+These effects reveal that people tend to generate items that are different from what already exists.
+However, existing models \citep[see][]{jern2013probabilistic} do not account for these effects.
 Our formal simulations reveal that PACKER provides a strong account of the phenomena observed behaviorally.
@@
 Under different parameter settings it is capable of generating tightly clustered or highly distributed categories, and adjusting the distribution of categories along each feature.
-Future work will focus on exploring how individual differences in generation can be explained using PACKER, and whether the canonical effects observed by \citet{ward1994structured} and \citet{jern2013probabilistic} can be explained through exemplar-similarity alone.
+Future work will focus on exploring how individual differences in generation, and whether PACKER can explain previous results \citep{ward1994structured,jern2013probabilistic}.
 

### 2017-01-30 Manuscripts/cogsci-2017/paper.tex
update to manuscript

@@
 \usepackage[natbibapa]{apacite}
-
-%though some work has recently been conducted using traditional classification learning approaches . 
-%Similar results were obtained in a more recent study conducted within a traditional categorization paradigm.
-
-%each example's degree of contribution toward generation.
-PACKER sets according $f_j$ .
-$\hat{f_j} = \phi$ for known members of contrast categories, $\hat{f_j} = \gamma$ for members of the target category.
-
-%%Results such as these have motivated formal accounts of generation that explicitly invoke the idea that people re-purpose existing knowledge to generate something new -- either by copying-and-tweaking an observation retrieved from memory \citep{ward2002role,ward1995s}, or by abstracting the distributional structure of categories in the domain \citep{jern2013probabilistic}.
-
-%Rather than evaluate PACKER on the benchmarks reviewed above, however, we instead focus on testing predictions made by PACKER that distinguish it from previous approaches.
-
-%The attention weights are of interest in explaining individual differences in generation strategy (discussed below), but for our formal simulations they are set uniformly.
-
-% so that is 1 when specifying whether $y$ exceeds the similarity threshold, $\tau$, which is treated as a free parameter ($0<\tau\leq1$).
-$g = 1$ when $s(y,z) < \tau$ and $g = 0$ otherwise.
-Thus, exemplars exceeding the similarity threshold will not be generated, but exemplars just under the threshold are most likely.
-\citet{jern2013probabilistic} did not implement any similarity threshold mechanism, and so their model is the special case where $\tau = 1$ (any exemplar may be selected).
-
-%JLA: I took this one out because it's not really something we're dealing with in our paper.
-%Further, exemplars drawn from the same species were less variable than animals drawn from different species, and possessed features that were adaptive for their environment (as is the case on Earth).
 
@@
 
-Recent work has proposed and tested formal models of category generation.
+Recent work has proposed and tested formal models to explain these observations.
 \citet{jern2013probabilistic} trained participants on experimenter-defined categories composed of exemplars within an artificial three-dimensional domain.
@@
 As in the classic \cite{ward1994structured} experiment, \citet{jern2013probabilistic} found that generated categories possessed the same feature variance and correlations as the experimenter-defined categories in the domain.
-They evaluated different possible computational models of how people generate categories (such as an exemplar-based "copy-and-tweak" model) and found that a hierarchical Bayesian model (which generates categories by abstracting the distributional structure of the domain) provided the strongest account of the behavioral results.
+
+\citet{jern2013probabilistic} tested several different computational models on their data.
+Most relevant to the present investigation, they tested a `copy-and-tweak' model that generates items by copying and changing previous observations, and a hierarchical Bayesian model that uses the structure of known categories to infer the structure of new ones.
+The authors found that the hierarchical Bayesian model provided the strongest account of the behavioral results.
 
@@
 For brevity, we abbreviate $f(x_j)$ as $f_j$.
-PACKER sets $f_j$ depending on exemplar $j$'s category membership: $f_j = \phi$ if $x_j$ is a member of the contrast category, and $f_j = \gamma$ if $x_j$ is a member of the target category.
-$\phi$ and $\gamma$ are free parameters ($-\infty \leq \phi, \gamma \leq \infty$) controlling the contribution of contrast- and target-category similarity, respectively (see Fig \ref{fig:example-prob-spaces}).
+PACKER sets $f_j$ depending on exemplar $j$'s category membership: $f_j = \phi$ if $x_j$ is a member of a contrast category, and $f_j = \gamma$ if $x_j$ is a member of the target category.
+$\phi$ and $\gamma$ are free parameters ($-\infty \leq \phi, \gamma \leq \infty$) controlling the contribution of contrast- and target-category similarity, respectively (see Figure \ref{fig:example-prob-spaces}).
 Larger absolute values result in greater consideration of those exemplars, with values of 0 eliminating their effect.
@@
 
-Recall that we proposed that categories are generated by balancing two constraints: New categories should be different from existing categories, and exemplars belonging to the same category should be similar to one another.
-This is realized in PACKER when $\phi < 0$, and $\gamma > 0$.
-Negative values for $\phi$ encourage $y$ to be distant from the contrast category (as similarity to contrast category exemplars are subtracted during aggregation).
-Positive values for $\gamma$ encourage $y$ to be close to other exemplars of the target category.
-When $|\phi| = \gamma$, the repulsion effect from the contrast category is equal to the attraction effect to the target category.
+PACKER's core proposal is that new categories should be different from existing categories, and same-category exemplars should be similar to one another.
+This is realized when $\phi \leq 0$, and $\gamma \geq 0$.
+Negative $\phi$ values encourage $y$ to be distant from the contrast category (as similarity to contrast category exemplars are subtracted during aggregation).
+Positive $\gamma$ values encourage $y$ to be close to other exemplars of the target category.
+When $|\phi| = \gamma$, the repulsion effect from contrast categories is equal to the attraction effect to the target category.
 
@@
 The underlying processes assumed by PACKER are highly similar to those in the GCM.
-The main difference is PACKER aggregates positive- and negative-valued similarities, rather than only aggregating positive-valued similarities.
+The main difference is that PACKER aggregates positive- and negative-valued similarities, rather than only aggregating positive-valued similarities.
 In later sections, we will explore the unique predictions yielded by these design principles.
@@
 In our copy-and-tweak model, all observations $z$ from the target class are equally likely to be retrieved as the source.
-The first exemplar is generated uniformly at random over stimulus space.
+When there are no observations from the target class, the first exemplar is generated uniformly at random over stimulus space.
 After a source exemplar is retrieved, the similarity between generation candidates $y$ and the retrieved exemplar is computed as per Equation \ref{eq:similarity}.
@@
 Due to conjugacy, the MAP estimate for the shared covariance matrix $\Sigma_D = \Sigma_0 + \sum_c{C_c}$, where $C_c$ is the empirical covariance matrix of category $c$.
-The MAP estimate of the covariance matrix for category $B$ is
+The MAP estimate of the covariance matrix for the target category $B$ is
 \begin{equation}
@@
 where $\theta$ is a response determinism parameter and ${\rm Normal}(y; \mu, \Sigma)$ denotes a multivariate Normal density evaluated at $y$.
-%JLA: There's something funky going on with the notation. I got confused by your noation, so hopefully you can update what i did to be consistent with the rest.
 
@@
 We developed two Alpha categories (see Figure \ref{fig:middle-bottom-conditions}): the `Bottom' Alpha category is a tight cluster in the bottom-center of the space, and the `Middle' Alpha category is identical except that it lies in the center of stimulus space.
-The two categories only differ in their location, not in  distributional structure.
 
@@
 Participants were instructed that members of the Beta category could be quite similar or different depending on what they think makes the most sense for the category, but that they were not allowed to make the same example twice.
-As in \citet{jern2013probabilistic}, generation was completed using a sliding-scale interfaceTwo scales controlled the dimensions of the generated example.
+As in \citet{jern2013probabilistic}, generation was completed using a sliding-scale interface.
+Two scales controlled the features of the generated example.
 An on-screen preview of the example updated whenever one of the features was changed.
@@
 
-Because the conditions differ only in their location along the y-axis, we focus on how Beta exemplars are generated above and below the contrast category.
 Several sample Beta categories are depicted in Figure \ref{fig:beta.samples}.
+Because the conditions differ only in their location along the y-axis, we first focus on how Beta exemplars are generated above and below the contrast category.
 As is evident in Figure \ref{fig:beta.samples}, we observed broad individual differences in generation strategy: Whereas some participants generated all four Beta examples within a narrow y-axis range, others generated Beta examples along a wide range.
@@
     \hline
-    Used bottom rows       &  28 & 18  \\
+    Used bottom row       &  28 & 18  \\
     No bottom row          &  11 &  4  \\
@@
 We did however, observe a notable subset of individuals with greater within-class distance.
-These individuals tended to adopt a `corners' approach, in which a Beta examples were placed almost exclusively in the corners of the space.
+These individuals tended to adopt a `corners' approach, in which Beta examples were placed almost exclusively in the corners of the space.
 
@@
 
-Some aspects of the results described above are somewhat commonsense: They demonstrate that the location of existing categories also imposes constraints on generation because people tend to generate examples in areas not occupied by existing categories.
-This principle, however, is not predicted by existing models of generation -- these models were designed to explain distributional correspondences between generated and existing categories.
+Some aspects of the results described above are somewhat commonsense: They demonstrate that the location of existing categories imposes constraints on generation because people tend to generate examples in areas not occupied by existing categories.
+This principle, however, is not predicted by existing models of generation -- these models were designed to explain distributional correspondences between generated and existing categories, not effects of contrast.
 
@@
 
-Our model-fitting results make sense given the design of the experiment.
-As the copy-and-tweak and hierarchical sampling models are not influenced by the location of contrast categories within the space, they do not capture the broad tendency for generating categories dissimilar to existing classes.
+Our model-fitting results make sense given the assumptions made by each model.
+As the copy-and-tweak and hierarchical sampling models are not influenced by the location of contrast categories within the space, they do not capture the broad tendency for generated items to be dissimilar to existing classes.
+
+\subsection{Individual Differences}
+
+Generally, our behavioral results showed that members of generated categories are dissimilar to opposite categories, and similar to members of their own category.
+However, we also observed a great deal of individual differences in generation style.
+Manually inspecting the data reveals four typical patterns (see Figure \ref{fig:beta.samples}): `corners' categories with one Beta example in each corner of the space, tight clusters, `column'-like categories, and `row'-like categories.
+Here, we explore the role of PACKER's attention weights ($w_1$ and $w_2$) in producing row-like or column-like categories.
 
@@
 
-\subsection{Individual Differences}
-%In future work, we intend to explore more systematically the role of PACKER's parameters in different subtypes of performance.
-For example, whereas the tight cluster profile could be explained with a moderately negative $\phi$ and a strongly positive $\gamma$, the four corners profile could be simulated with a strongly negative $\phi$ and $\gamma\leq0$.
-%While it is important that any model be able to explain different types of performance with the same parameter values, these radical differences in generation are difficult to account for under a single set of parameters.
-Generally, our behavioral results showed that generated categories consist of items that are dissimilar to members of opposite categories, and similar to members of their own category.
-We also observed a great deal of individual differences in generation style.
-Manually inspecting the data reveals four typical patterns (see Figure \ref{fig:beta.samples}): `corners' categories with one Beta example in each corner of the space, tight clusters, `column'-like categories, and `row'-like categories.
-Here, we explore the role of PACKER's attention weights ($w_1$ and $w_2$) in producing row-like or column-like categories.
-
 \begin{figure*}
@@
 Taking the average of these range differences yields a gradient describing how, on average, categories were distributed for each stimulus.
-These data (Fig \ref{fig:range-diff-gradient}) reveal a systematic relationship between category structure and category location.
+These data (Figure \ref{fig:range-diff-gradient}) reveal a systematic relationship between category structure and location.
 Whereas column-like categories more often include stimuli to the left or right of the Alpha class, row-like categories appear above and below the Alpha class.
@@
 We then simulated 50 Beta categories with each participant's weighting scheme to obtain a sense of how the relative importance of each dimension affects what types of categories are generated and where they are generated.
+The results of these simulations are depicted in Figure \ref{fig:range-diff-gradient}.
 
-The results, depicted in Figure \ref{fig:range-diff-gradient}, support PACKER's predictions.
 When the x-axis is weighted more, PACKER creates column categories to the sides of the Alphas.
@@
 This behavior falls out from the nature of selective attention: Dimensions weighted more have a sharper similarity gradient.
-For example, when the x-axis is weighted more, PACKER favors Beta categories with more within-class similarity (less range), and less between-class similarity along the x-axis, resulting in column-like categories.
+For example, when the x-axis is weighted more, PACKER favors Beta categories with more within-class similarity (less range), and less between-class similarity along the x-axis, resulting in column-like categories that differ from the Alphas along the x-axis.
 
@@
 Likewise, we found that the \textit{location} of contrast categories shapes generation by imposing constraints on the areas of space that remain available for a new category.
-These effects reveal that people tend to generate items that are different from what already exists.
-However, existing models \citep[see][]{jern2013probabilistic} do not account for these effects.
+Existing models \citep[see][]{jern2013probabilistic}, possessing no mechanisms for category-contrast, do not account for these effects.
 Our formal simulations reveal that PACKER provides a strong account of the phenomena observed behaviorally.
@@
 Under different parameter settings it is capable of generating tightly clustered or highly distributed categories, and adjusting the distribution of categories along each feature.
-Future work will focus on exploring how individual differences in generation, and whether PACKER can explain previous results \citep{ward1994structured,jern2013probabilistic}.
+Future work will focus on exploring individual differences in generation, and whether PACKER can explain previous results in the field \citep{ward1994structured,jern2013probabilistic}.
 
@@
 We thank Kenneth Kurtz for helpful comments, and Alan Jern and Charles Kemp for providing code and data.
-% any other thanks?
 

### 2017-01-30 Manuscripts/cogsci-2017/paper.tex
slight edits

@@
 PACKER's core design assumptions are (1) categories are represented as exemplars in a multidimensional psychological space, (2) generated items should be similar to exemplars of the same category, and (3) generated categories should be dissimilar to existing categories.
-A behavioral study and formal simulations broadly support PACKER's predictions.
+A behavioral study reveals strong effects of contrast- and target-class similarity, which are directly predicted by the PACKER model but are not explained by existing formal approaches.
 
@@
 Participants in our study more frequently generated items that are distant from members of contrast categories, and they tended to generate categories with more within-class than between-class similarity.
-Likewise, we found that the \textit{location} of contrast categories shapes generation by imposing constraints on the areas of space that remain available for a new category.
-Existing models \citep[see][]{jern2013probabilistic}, possessing no mechanisms for category-contrast, do not account for these effects.
-Our formal simulations reveal that PACKER provides a strong account of the phenomena observed behaviorally.
+Likewise, we found that the \textit{location} of contrast categories (as opposed to their structure) shapes generation by imposing constraints on the areas of space that remain available for a new category.
+Formal simulations reveal that existing models \citep[see][]{jern2013probabilistic}, making no assumptions about category-contrast, do not account for these effects.
 
@@
 Under different parameter settings it is capable of generating tightly clustered or highly distributed categories, and adjusting the distribution of categories along each feature.
-Future work will focus on exploring individual differences in generation, and whether PACKER can explain previous results in the field \citep{ward1994structured,jern2013probabilistic}.
+Future work will focus on exploring the broad degree of individual differences we observed in generation, and whether PACKER can explain previous results in the field \citep{ward1994structured,jern2013probabilistic}.
 

### 2017-01-30 Manuscripts/model-descriptions/jern-kemp-conjugate/jern-kemp-conjugate.tex
distance analyses, documentation update

@@
 
-Working backwards, what is needed for simulation of exemplar generation is the probability of generating a stimulus $x$ given exposure to members of the target category $x_b$:
+Working backwards, what is needed for simulation of exemplar generation is the probability of generating a stimulus $y$ given exposure to members of the target category $x_B$:
 \
 \begin{equation}
-p(x | x_b) = ?
+p(y | x_B) = ?
 \end{equation}
 \
-where $x_b$ may be empty.
+where $x_B$ may be empty.
 Jern \& Kemp's model achieves this using a generative process.
-Category members ($x_a$ or $x_b$; more generally written as $x_y$) are assumed to have been generated using an underlying category distribution (specifically, multivariate normal):
+Category members ($x_A$ or $x_B$; more generally written as $x_C$) are assumed to have been generated using an underlying category distribution (specifically, multivariate normal):
 \
 \begin{equation}
-  x_y \sim Normal(\mu_{y}, \Sigma_{y})
+  x_C \sim Normal(\mu_{C}, \Sigma_{C})
 \end{equation}
 \
-$p(x | x_y)$ is proportional to the candidate's density under the target category's distribution.
-Thus, obtaining the category distribution parameters $(\mu_{y}, \Sigma_{y})$ is key for generation.
+$p(y | x_C)$ is proportional to the candidate's density under the target category's distribution.
+Thus, obtaining the category distribution parameters $(\mu_{C}, \Sigma_{C})$ is key for generation.
 This document describes how we compute these variables in a conjugate model.
 
-\subsection*{Computing $\mu_y$}
-Assuming $(\mu_y, \Sigma_y)$ are Normal-Inverse-Wishart distributed (unknown mean, unknown variance):
+\subsection*{Computing $\mu_C$}
+Assuming $(\mu_C, \Sigma_C)$ are Normal-Inverse-Wishart distributed (unknown mean, unknown variance):
 \
 \begin{equation}
-  \mu_y = \dfrac
-    {\kappa\mu_{0} + n_y \bar{x_y}}
-    {\kappa + n_y}
+  \mu_C = \dfrac
+    {\kappa\mu_{0} + n_C \bar{x_C}}
+    {\kappa + n_C}
     \label{eq:category_mus}
@@
     \item $\kappa$ is a scalar hyper-parameter, roughly weighting the importance of $\mu_{0}$. $\kappa$ must be greater than zero.
-    \item $n_y$ is the number of observations in $x_y$
-    \item $\bar{x_y}$ is the sample mean along $p$ dimensions
+    \item $n_C$ is the number of observations in $x_C$
+    \item $\bar{x_C}$ is the sample mean along $p$ dimensions
 \end{itemize}
 
-In the case of a populated class, $\mu_{y}$ ends up lying somewhere between $\mu_{0}$ and $\bar{x_y}$, depending on $\kappa_{0}$ and $n_y$.
-In the case of an empty class, $n_y = 0$, thus Equation \ref{eq:category_mus} reduces to $\mu_{y} = \mu_{0}$.
-Because we set $\mu_0$ to the center of the space, this outcome is the same as if we had integrated over all possible $\mu_y$.
+In the case of a populated class, $\mu_{C}$ ends up lying somewhere between $\mu_{0}$ and $\bar{x_C}$, depending on $\kappa_{0}$ and $n_C$.
+In the case of an empty class, $n_C = 0$, Equation \ref{eq:category_mus} reduces to $\mu_{C} = \mu_{0}$.
+Because we set $\mu_0$ to the center of the space, this outcome is the same as if we had integrated over all possible $\mu_C$.
+In practice, if $n_C = 0$, the model picks a stimulus at random from all candidates (uniform probabilities).
 
@@
 
-Unlike $\mu_y$, $\Sigma_y$ cannot be computed considering only the members of category $y$.
-Instead, $\Sigma_y$ is influenced both by the distribution of $x_y$ and by members of other categories through $\Sigma_D$.
+Unlike $\mu_C$, $\Sigma_C$ cannot be computed considering only the members of category $y$.
+Instead, $\Sigma_C$ is influenced both by the distribution of $x_C$ and by members of other categories through $\Sigma_D$.
 
@@
 \begin{equation}
-    \Sigma_D = \Sigma_0 + \sum_{y}{C_y}
+    \Sigma_D = \Sigma_0 + \sum_{C}{C_C}
 \end{equation}
@@
 $\Sigma_{0}$ is a $p$-by-$p$ prior covariance matrix.
-We use a $p$-dimensional identity matrix $I_p$ multiplied element-wise against a free parameter, $\gamma$, controlling the amount of variance assumed by the prior:
+We use a $p$-dimensional identity matrix $I_p$ multiplied element-wise against a free parameter, $\rho$, controlling the amount of variance assumed by the prior:
 \
 \begin{equation}
-    \Sigma_0 =  I_p\gamma
+    \Sigma_0 =  I_p\rho
 \end{equation}
 
-Thus, categories are assumed to have some degree of variance along each feature (specified by $\gamma$), but not are assumed to possess feature-feature correlations.
+Thus, categories are assumed to have some degree of variance along each feature (specified by $\rho$), but not are assumed to possess feature-feature correlations.
 
-\subsection*{Computing $\Sigma_y$}
+\subsection*{Computing $\Sigma_C$}
 
-Assuming $(\mu_y, \Sigma_y)$ are Normal-Inverse-Wishart distributed, $\Sigma_y$ can be computed as:
+Assuming $(\mu_C, \Sigma_C)$ are Normal-Inverse-Wishart distributed, $\Sigma_C$ can be computed as:
 
 \begin{equation}
-  \Sigma_y = [\Sigma_D\nu + C_y +
+  \Sigma_C = [\Sigma_D\nu + C_C +
     \dfrac
-    {\kappa n_y}
-    {\kappa + n_y}
-    (\bar{x_y}-\mu_y)(\bar{x_y}-\mu_y)^T
-  ] (\nu + n_y)^{-1}
+    {\kappa n_C}
+    {\kappa + n_C}
+    (\bar{x_C}-\mu_C)(\bar{x_C}-\mu_C)^T
+  ] (\nu + n_C)^{-1}
   \label{eq:Sigma_y}
@@
 \
-$\kappa$, $\bar{x_y}$, $C_y$, $n_y$, $\mu_0$,  are the same values as described above.
+$\kappa$, $\bar{x_C}$, $C_C$, $n_C$, $\mu_0$,  are the same values as described above.
 $\nu$ is an additional free parameter, weighting the importance of $\Sigma_{D}$.
 $\nu$ must be greater than $p-1$.
-When $x_b$ is empty, Equation \ref{eq:Sigma_y} reduces to $\Sigma_y = \Sigma_D$.
+When $x_b$ is empty, Equation \ref{eq:Sigma_y} reduces to $\Sigma_C = \Sigma_D$.
 
-\subsection*{Computing response probabilities $p(x | x_b)$}
+\subsection*{Computing response probabilities $p(y | x_C)$}
 
-$x$ are assumed to be drawn from a the distribution given by $Normal(\mu_{y}, \Sigma_{y})$. Thus,
+$x$ are assumed to be drawn from a the distribution given by $Normal(\mu_{C}, \Sigma_{C})$. Thus,
 
 \begin{equation}
-  p(x) \propto Normal(x | \mu_{b}, \Sigma_{b})
+  p(y) \propto Normal(y; \mu_{C}, \Sigma_{C})
 \end{equation}
 
-In practice, $p(x)$ is computed by first obtaining the relative density of every possible generation candidate $x_i$ under the category distribution.
+In practice, $p(y)$ is computed by first obtaining the relative density of every possible generation candidate $y_i$ under the category distribution.
 The end probability is a normalization of these values:
@@
   p(x) = \dfrac
-    {exp( \theta Normal(x | \mu_{b}, \Sigma_{b}))}
-    {\sum_i exp( \theta Normal(x_i |\mu_{b}, \Sigma_{b}))} 
+    {exp( \theta Normal(y ; \mu_{C}, \Sigma_{C}))}
+    {\sum_i exp( \theta Normal(y_i ; \mu_{C}, \Sigma_{C}))} 
 \end{equation}
@@
     \setlength\itemsep{-0.5em}
-    \item $\kappa$. Scalar, $\kappa>0$. Weights the importance of $\mu_{0}$ in inferring category $\mu_{y}$.
-    \item $\gamma$. Scalar, $\gamma>0$. Weights the importance of $\Sigma_{0}$ in inferring the domain $\Sigma_{D}$.
-    \item $\nu$. Scalar, $\nu > p-1$. Weights the importance of $\Sigma_{D}$ in inferring the domain $\Sigma_{y}$.
+    \item $\kappa$. Scalar, $\kappa>0$. Weights the importance of $\mu_{0}$ in inferring category $\mu_{C}$.
+    \item $\rho$. Scalar, $\rho>0$. Sets the assumed variance in the domain prior, $\Sigma_{0}$.
+    \item $\nu$. Scalar, $\nu > p-1$. Weights the importance of $\Sigma_{D}$ in inferring the domain $\Sigma_{C}$.
     \item $\theta$. Scalar, $\theta > 0$. Response determinism parameter.

### 2017-01-30 Manuscripts/cogsci-2017/paper.tex
minor last changes from joe for submission

@@
 PACKER's core design assumptions are (1) categories are represented as exemplars in a multidimensional psychological space, (2) generated items should be similar to exemplars of the same category, and (3) generated categories should be dissimilar to existing categories.
-A behavioral study reveals strong effects of contrast- and target-class similarity, which are directly predicted by the PACKER model but are not explained by existing formal approaches.
+A behavioral study reveals strong effects of contrast- and target-class similarity.
+These effects are novel empirical phenomena, which are directly predicted by the PACKER model but are not explained by existing formal approaches.
 
@@
 Most relevant to the present investigation, they tested a `copy-and-tweak' model that generates items by copying and changing previous observations, and a hierarchical Bayesian model that uses the structure of known categories to infer the structure of new ones.
-The authors found that the hierarchical Bayesian model provided the strongest account of the behavioral results.
+They found that the hierarchical Bayesian model provided the strongest account of the behavioral results.
 
@@
 where $w_k$ is the attention weighting of dimension $k$ ($w_k \geq 0$ and $\sum_k{w_k} = 1$), accounting for the relative importance of each dimension in similarity calculations, and $c$ ($c>0$) is a specificity parameter controlling the spread of exemplar generalization.
-For simplicity, simulations will use uniform attention weights, except in our discussion of individual differences.
+For simplicity, our simulations will use uniform attention weights, except in our discussion of individual differences.
 
@@
 We assume the shared covariance matrix $\Sigma_D$ is generated from a Wishart distribution (for conjugacy) with parameters $v_0$, $\kappa_0$, and $\Sigma_0$.
-We set $\nu_0 = \kappa_0 = 1$, and $\Sigma_0 = \rho {\bf I}$, where $\rho$ is a free parameter controlling the expected variance of dimensions (dimensions of the shared covariance matrix are expected to be uncorrelated) and ${\bf I}$ is the identity matrix.
+We set $\nu_0 = 4$, and $\Sigma_0 = \rho {\bf I}$, where $\rho$ is a free parameter controlling the expected variance of dimensions (dimensions of the shared covariance matrix are expected to be uncorrelated) and ${\bf I}$ is the identity matrix.
 
@@
 Although our manipulation is minimal, the PACKER model predicts strong between-condition differences.
-PACKER proposes that the nature of the space not occupied by the Alpha category determines where members of the Beta category are likely to be generated.
+According to PACKER, the nature of the space not occupied by the Alpha category should determine where members of the Beta category are likely to be generated.
 Thus, the lower areas of the stimulus space should be less frequently used for generation in the Bottom condition compared to the Middle (as these areas possess greater similarity to the Bottom Alpha category).
@@
 
-To evaluate the key predictions of PACKER, we determined the number of participants in each condition who placed at least one Beta examplar on the top and bottom `rows' of the space (the maximum and minimum possible y-axis value, respectively).
+To evaluate the key predictions of PACKER, we determined the number of participants in each condition who placed at least one Beta exemplar on the top and bottom `rows' of the space (the maximum and minimum possible y-axis value, respectively).
 The resulting contingencies data are shown in Table \ref{table:subset-table}.
@@
 Some aspects of the results described above are somewhat commonsense: They demonstrate that the location of existing categories imposes constraints on generation because people tend to generate examples in areas not occupied by existing categories.
-This principle, however, is not predicted by existing models of generation -- these models were designed to explain distributional correspondences between generated and existing categories, not effects of contrast.
+This principle, however, is novel and not predicted by existing models of generation -- these models were designed to explain distributional correspondences between generated and existing categories, not effects of contrast.
 
@@
 Each model's best-fitting parameterization is shown in Table \ref{table:model-fits}.
-Overall, PACKER outperformed copy-and-tweak and the hierarchical sampling model by a considerable margin ($\sim17\%$ improvement in log-likelihood).
+Overall, PACKER outperformed copy-and-tweak and the hierarchical sampling model by a considerable margin ($\sim12\%$ improvement in log-likelihood).
 The parameter settings associated with PACKER's best fit are exactly as expected: a strong preference for items that are similar to members of the target category but are dissimilar to members of the contrast category.

### 2017-03-03 Manuscripts/cog-psych/cogpsych.tex
made cogpsych paper skeleton

%!TEX output_directory = latex_out/

\documentclass[12pt]{article}
\usepackage[letterpaper, margin=1in, headheight=15pt]{geometry}
\usepackage{amsmath}
\usepackage{setspace}
\usepackage{pgfplots}
\usepackage{fancyhdr}
\usepackage{fancyvrb}
\usepackage[natbibapa]{apacite}

% set up PGF
\pgfplotsset{compat=1.6}

% set up header % 
\pagestyle{fancy}
\fancyhf{} % sets both header and footer to nothing
\renewcommand{\headrulewidth}{0pt}
\lhead{RUNNING HEAD: Contrast in Concept Generation}
\fancyhead[R]{\thepage}

% author info:
% https://www.elsevier.com/journals/cognitive-psychology/0010-0285/guide-for-authors

\begin{document}

% ------- TITLE PAGE ------- %
\begin{center}
\hfill
\\[1in]

The Role of Contrast in Concept Generation.
% ---- other ideas:
% - The Role of Contrast in Concept Generation.
% - Creating Something Different: The Role of Contrast in Concept Generation.
% - Something New: The Role of Contrast in Concept Generation.
% - What's New In Concept Generation? The Role of Category Contrast.

\vfill

Nolan Conaway\textsuperscript{1}, 
Kenneth J. Kurtz\textsuperscript{2}, 
\& Joseph L. Austerweil\textsuperscript{1}
\\[\baselineskip]
\textsuperscript{1}University of Wisconsin-Madison, Department of Psychology, Madison, WI, USA
\textsuperscript{2}Department of Psychology, Binghamton University, Binghamton, NY, USA
\\[1in]

\vfill

Author Note

Correspondence concerning this article should be addressed to: 
Joseph Austerweil, 1202 West Johnson Street, Madison, WI 53706.
E-mail: austerweil@wisc.edu

\end{center}
\clearpage

% ------- ABSTRACT PAGE ------- %
\doublespacing
\section{Abstract}

Filler!

\setlength\parindent{0.5in}
\textit{Keywords}: categorization, concepts, creativity, generation
\clearpage

% ------- BEGIN! ------- %
\begin{flushleft}

\section{Introduction}
\setlength\parindent{0.5in}

\section{Experiment 1}

\end{flushleft}

% references section
\clearpage
\setlength{\bibsep}{10pt}
\setstretch{1}
\bibliography{references.bib}
\clearpage

\end{document}

### 2017-03-07 Manuscripts/cog-psych/cogpsych.tex
misc snippets for cogpsych manuscript

@@
 \usepackage{fancyvrb}
+\usepackage{csquotes}
 \usepackage[natbibapa]{apacite}
@@
 \pgfplotsset{compat=1.6}
+\newcommand\inputpgf[2]{{
+\let\pgfimageWithoutPath\pgfimage
+\renewcommand{\pgfimage}[2][]{\pgfimageWithoutPath[##1]{#1/##2}}
+\input{#1/#2}
+}}
 
@@
 \renewcommand{\headrulewidth}{0pt}
-\lhead{RUNNING HEAD: Contrast in Concept Generation}
+\lhead{RUNNING HEAD: Similarity and Contrast in Concept Generation}
 \fancyhead[R]{\thepage}
@@
 
-The Role of Contrast in Concept Generation.
+Creating Something Different: Similarity and Contrast Effects in Concept Generation.
 % ---- other ideas:
@@
 
+The creation of novel concepts and ideas is a highly intriguing -- yet infrequently studied -- topic of research in human cognition.
+The creative use of conceptual knowledge is a core element of scientific investigation, wherein the generation of new ideas is critical to designing experiments and explaining observations.
+However, due to its complexity, little is presently known about the cognitive processes underlying the creative generation of new concepts.
+
+Much of what we know about conceptual generation comes from the foundational literature on creative cognition.
+In a classic series of reports, Ward \& colleagues \citep{ward1995s,ward1994structured,ward2002role,marsh1999inadvertent,ward2002role} established that category generation is highly constrained by prior knowledge.
+For example, \cite{ward1994structured} observed that participants asked to generate new species of alien animals tended to produce features possessed by earth species.
+Likewise...
+
+\section{Snippets}
+
+\cite{jern2013probabilistic} showed that difficult topic of creative generation could be studied through the well-developed methodology from the field of categorization.
+Specifically, they found that learners trained on  artificial categories typically generated new artificial categories with similar distributional properties.
+
+In this paper, we provide a systematic examination of conceptual generation using the well-developed methods and theoretical frameworks from the literature on human category learning.
+We also introduce a novel exemplar model of category generation, PACKER (\textit{Producing Alike and Contrasting Knowledge using Exemplar Representations}), which explains much of what we observe in the creative use of conceptual knowledge.
+One of the core principles of PACKER (as well as its name) are inspired by earlier work in category learning \citep[see][]{hidaka2011packing}.
+
 \section{Experiment 1}
+
+\subsection{Participants \& Materials}
+
+183 participants were recruited from Amazon Mechanical Turk.
+Participants were randomly assigned to one condition: 64 participants were assigned to the Cluster condition, 61 were assigned to the Row condition, and 58 were assigned to the XOR condition.
+Stimuli were squares varying in color (RGB 25--230) and side length (3.0--5.8cm).
+See Figure \ref{fig:sample-stimuli} for samples.
+The assignment of perceptual features (color, size) to axes of the domain space (x, y) was counterbalanced across participants.
+
+\begin{figure}
+    \begin{center}
+    \inputpgf{figs/}{stimuli-samples.pgf}
+    \caption{Sample stimuli used in Experiments 1 and 2.}
+    \label{fig:sample-stimuli}
+    \end{center}
+\end{figure}
+
+\subsection{Procedure}
+
+Participants began the experiment with a short training phase (3 blocks of 4 trials), where they observed exemplars belonging to the `Alpha' category.
+Participants were instructed to learn as much as they can about the Alpha category, and that they would answer a series of test questions afterwards.
+On each trial, a single Alpha category exemplar was presented, and participants were given as much time as they desired before moving on.
+Each block consisted of a single presentation of each of the members of the Alpha category, in a random order.
+Participants were shown the range of possible colors and sizes prior to training.
+
+Following the training phase, participants were asked to generate four examples belonging to another category called `Beta'.
+As in \citet{jern2013probabilistic}, generation was completed using a sliding-scale interface.
+Two scales controlled the features (color, size) of the generated example.
+An on-screen preview of the example updated whenever one of the features was changed.
+Participants could generate any example along an evenly-spaced 9x9 grid, except for any previously generated Beta exemplars.
+Neither the members of the Alpha category nor the previously generated Beta examples were visible during generation.
+Prior to beginning the generation phase, participants read the following instructions:
+
+\begin{displayquote}
+As it turns out, there is another category of geometric figures called "Beta".
+Instead of showing you examples of the Beta category, we would like to know what you think is likely to be in the Beta category.
+
+You will now be given the chance to create examples of any size or color in order to show what you expect about the Beta category.
+You will be asked to produce 4 Beta examples - they can be quite similar or quite different to each other, depending on what you think makes the most sense for the category.
+
+Each example needs to be unique, but the computer will let you know if you accidentally create a repeat.
+\end{displayquote}
+
+Following generation, participants completed a generalization phase wherein they classified novel examples into the Alpha and Beta categories without feedback.
+On each trial, a single example was presented, and participants were asked to classify it by clicking buttons labeled ``Alpha'' or ``Beta''.
+Participants classified a total of 81 items sampled along a 9x9 grid, including the members of the Alpha and Beta categories (randomly intermixed).
+These data were, however, collected to address a separate set of questions, and we do not discuss them in this paper.
+
+\section{Acknowledgments}
+A previous version of this work appeared in the Proceedings of the Thirty-Ninth Annual Conference of the Cognitive Science Society.
+Support for this research was provided by the Office of the VCRGE at the UW - Madison with funding from the WARF.
+We thank Alan Jern and Charles Kemp for providing code and data.
 
@@
 \clearpage
-\setlength{\bibsep}{10pt}
-\setstretch{1}
-\bibliography{references.bib}
+\bibliographystyle{apacite}
+\bibliography{citations.bib}
 \clearpage

### 2017-03-08 Manuscripts/cog-psych/cogpsych.tex
manuscript updates

@@
 Much of what we know about conceptual generation comes from the foundational literature on creative cognition.
-In a classic series of reports, Ward \& colleagues \citep{ward1995s,ward1994structured,ward2002role,marsh1999inadvertent,ward2002role} established that category generation is highly constrained by prior knowledge.
-For example, \cite{ward1994structured} observed that participants asked to generate new species of alien animals tended to produce features possessed by earth species.
-Likewise...
+In a classic series of reports, Ward \& colleagues \citep{ward1995s,ward1994structured,marsh1999inadvertent,ward2002role,smith1993constraining} established that category generation is highly constrained by prior knowledge: generated categories tend to consist of features observed in known categories, and they tend to exhibit the distributional properties as found in known categories.
+In a classic study, \cite{ward1994structured} asked participants to generate new species of alien animals by drawing and describing members of the species.
+People tended to generate species with the same features as on Earth (e.g., eyes, legs, wings), and possessing the same feature correlations as on Earth (e.g., feathers co-occur with wings).
+Likewise, aliens drawn from the same species tended to share more features with one another compared to members of opposite species.
+
+% [[NBC]] Explain copy and tweak here?
+
+Much of the work from this area \citep[e.g.,][]{smith1993constraining,marsh1999inadvertent} focuses on how sample cues (such as an example of a species generated by other participants) can drastically diminish creativity.
+However, the broader set of observations made by Ward \& colleagues provide a great deal of insight into the nature of creative generation.
+They indicate that people rely strongly on prior knowledge in the creative process, and people generate concepts in accord with what they already know.
+Problematically, however, the experimental paradigms employed in these studies were relatively uncontrolled.
+Specifically, in order to allow for maximum creative freedom, participants were not constrained as to what they could generate.
+While these procedures do ensure creative freedom, they render difficult the development of formal theories to explain conceptual generation.
+
+\cite{jern2013probabilistic} recently showed that creative generation could be studied in a more controlled manner through the well-developed methods of an artificial categorization paradigm \citep[see][]{kurtz2015human}.
+In their experiments 3 and 4, participants were exposed to members of experimenter-defined categories of "crystals" varying in size, hue, and saturation.
+Following a training phase during which the experimenter-defined categories were learned, participants were asked to generate novel categories of crystals.
+In a finding mirroring that of the \cite{ward1994structured} studies, \citeauthor{jern2013probabilistic} found that participants generated categories with the same distributional properties as the experimenter-defined categories: for example, after training on categories with a positive correlation between the size and saturation features (larger sized crystals were more saturated), participants generated novel categories with a positive correlation.
+
+The \cite{jern2013probabilistic} report thus represents an important step towards understanding the cognitive processes underlying creative generation: not only did they find that generation in an artificial domain resembles the more naturalistic cases observed by Ward \& colleagues, but they also found that theories of conceptual generation could be evaluated more formally through comparison to the participant-generated categories.
+
+% [[NBC]] explain their models here...
+
+It is worth noting, however, that the literature reviewed above provides only a limited view of category generation: the main focus of the published experiments has been on the distributional correspondences between learned and generated categories.
+In this paper, we seek to identify other important constraints on the processes involved in creative generation.
+We also introduce a novel exemplar-based model of category generation, PACKER (\textit{Producing Alike and Contrasting Knowledge using Exemplar Representations}), which explains much of what we observe in the creative use of conceptual knowledge.
+In the sections below, we introduce the PACKER model and its core theoretical claims.
+
+\subsection{Something Different: A Role For Contrast}
+
+In order to generate a novel concept, individuals must produce something that is in some capacity \textit{different} from what they already know.
+Thus, in a trivial sense, contrast can be viewed as a fundamental constraint on creative generation: new concepts must firstly be different from existing ones.
+However the role of contrast in category generation has not been systematically evaluated: how do people generate concepts such that the product is (A) different from what is already known, and (B) internally coherent?
+
+\clearpage
+
+\section{The PACKER Model}
+
+The PACKER model is an extension of the influential Generalized Context Model of category learning \citep[GCM;][]{nosofsky1984choice}.
+It assumes that each category is encoded by a set of exemplars within a $k$-dimensional psychological space, and that generation is constrained by both similarity to members of the target category (the category in which a stimulus is being generated) as well as similarity to members of other categories.
+
+As in the GCM, the similarity between two examples, $s\left(x_i, x_j\right)$, is an inverse-exponential function of distance:
+% 
+\begin{equation}
+  s\left(x_i,x_j\right) = \exp \left\{ -c \sum_{k}{ \left| x_{ik} - x_{jk} \right|}w_k \right\}
+  \label{eq:similarity}
+\end{equation}
+% 
+where $w_k$ is the attention weighting of dimension $k$ ($w_k \geq 0$ and $\sum_k{w_k} = 1$), accounting for the relative importance of each dimension in similarity calculations, and $c$ ($c>0$) is a specificity parameter controlling the spread of exemplar generalization.
+For simplicity, our simulations will use uniform attention weights, except in our discussion of individual differences.
+
+To generate a new example, the model considers both the similarity to examples from contrast categories as well as the similarity to examples (if any exist) in the target category.
+The aggregated similarity $a$ between generation candidate $y$ and stored exemplars $x$ is given by:
+% 
+\begin{equation}
+    a(y, x) = \sum_j{f(x_j) s(y, x_j)}
+\end{equation}
+% 
+where $f(x_j)$ is a function specifying the extent to the exemplar contributes to generation.
+PACKER sets $f(x_j)$ depending on exemplar $x_j$'s category membership: $f(x_j) = \phi$ if $x_j$ is a member of a contrast category, and $f(x_j) = \gamma$ if $x_j$ is a member of the target category.
+$\phi$ and $\gamma$ are free parameters ($-\infty \leq \phi, \gamma \leq \infty$) controlling the contribution of contrast- and target-category similarity, respectively.
+Larger absolute values result in greater consideration of those exemplars, with values of 0 eliminating their effect.
+A negative value for $f(x_j)$ produces a `repelling' effect (exemplars are less likely to be generated nearby $x_j$).
+Conversely, a positive value for $f(x_j)$ produces an `attracting' effect (exemplars are more likely to be generated nearby $x_j$).
+
+PACKER's core proposal is that new categories should be different from existing categories, and same-category exemplars should be similar to one another.
+This is realized when $\phi \leq 0$, and $\gamma \geq 0$.
+Negative $\phi$ values encourage $y$ to be distant from the contrast category (as similarity to contrast category exemplars are subtracted during aggregation).
+Positive $\gamma$ values encourage $y$ to be close to other exemplars of the target category.
+When $|\phi| = \gamma$, the repulsion effect from contrast categories is equal to the attraction effect to the target category.
+
+The probability that a given candidate $y$ will be generated is evaluated using an Exponentiated \citet{luce1977choice} choice rule.
+Candidates with greater values of $a$ are more likely to be generated than candidates with smaller values:
+% 
+\begin{equation}
+p(y) = \dfrac
+    { \exp \{ \theta \cdot a(y, x) \} } 
+    { \sum_i{ \exp \{ \theta \cdot a(y_i, x) \}  } }
+    \label{eq:packer-choice}
+\end{equation}
+% 
+where $\theta$ ($\theta \geq 0$) controls response determinism.
 
@@
 
-\cite{jern2013probabilistic} showed that difficult topic of creative generation could be studied through the well-developed methodology from the field of categorization.
-Specifically, they found that learners trained on  artificial categories typically generated new artificial categories with similar distributional properties.
+One of the core principles of PACKER (as well as its name) are inspired by earlier work in category learning \citep[see][]{hidaka2011packing}.
 
-In this paper, we provide a systematic examination of conceptual generation using the well-developed methods and theoretical frameworks from the literature on human category learning.
-We also introduce a novel exemplar model of category generation, PACKER (\textit{Producing Alike and Contrasting Knowledge using Exemplar Representations}), which explains much of what we observe in the creative use of conceptual knowledge.
-One of the core principles of PACKER (as well as its name) are inspired by earlier work in category learning \citep[see][]{hidaka2011packing}.
+\clearpage
 
@@
 Stimuli were squares varying in color (RGB 25--230) and side length (3.0--5.8cm).
-See Figure \ref{fig:sample-stimuli} for samples.
+These stimuli are slight variants of those used by \cite{conaway2016similar}, see Figure \ref{fig:sample-stimuli} for samples.
 The assignment of perceptual features (color, size) to axes of the domain space (x, y) was counterbalanced across participants.
@@
 
+\clearpage
+
+\section{Experiment 2}
+
+\subsection{Participants \& Materials}
+
+122 participants were recruited from Amazon Mechanical Turk.
+61 participants were randomly assigned to the Middle and Bottom conditions each.
+The stimuli were exactly as in Experiment 1.
+Again, the assignment of perceptual features (color, size) to axes of the domain space (x, y) was counterbalanced across participants.
+
+\subsection{Procedure}
+
+The procedure was exactly as in Experiment 1: participants first completed a short training phase, followed by the generation phase, followed by the generalization phase (data from this phase is not discussed in this report).
+
 \section{Acknowledgments}

### 2017-03-08 Manuscripts/cog-psych/cogpsych.tex
more manuscript updates!

@@
 \usepackage{csquotes}
+\usepackage{todonotes}
+\usepackage{verbatim}
 \usepackage[natbibapa]{apacite}
@@
 }}
+
+% set up notes-- different backgrounds for Nolan and Joe!
+\newcommand\nbcnote[1]{\todo[inline, backgroundcolor = yellow]{\textbf{NBC}: #1}}
+\newcommand\jlanote[1]{\todo[inline, backgroundcolor = lime]{\textbf{JLA}: #1}}
 
@@
 % ---- other ideas:
-% - The Role of Contrast in Concept Generation.
-% - Creating Something Different: The Role of Contrast in Concept Generation.
-% - Something New: The Role of Contrast in Concept Generation.
-% - What's New In Concept Generation? The Role of Category Contrast.
+% - The Constraining Influences of Contrast and Similartity in Concept Generation.
 
@@
 
-Filler!
+\jlanote{Joe! Did you know i made a command for you to enter in comments like this?? }
+
+\begin{verbatim}
+    \jlanote{Some note}
+\end{verbatim}
 
@@
 
-% [[NBC]] Explain copy and tweak here?
+\nbcnote{Explain Copy + Tweak here?}
 
@@
 
-% [[NBC]] explain their models here...
-
-It is worth noting, however, that the literature reviewed above provides only a limited view of category generation: the main focus of the published experiments has been on the distributional correspondences between learned and generated categories.
-In this paper, we seek to identify other important constraints on the processes involved in creative generation.
-We also introduce a novel exemplar-based model of category generation, PACKER (\textit{Producing Alike and Contrasting Knowledge using Exemplar Representations}), which explains much of what we observe in the creative use of conceptual knowledge.
-In the sections below, we introduce the PACKER model and its core theoretical claims.
+\nbcnote{explain hierarchical sampling here, maybe also their implementation of copy+tweak?
+Or maybe we should step away from copy+tweak as a relevant option.
+Or maybe we focus on systematically comparing their copy and tweak to PACKER with specific params?
+This copy and tweak issue makes for a mess...
+}
 
@@
 
-In order to generate a novel concept, individuals must produce something that is in some capacity \textit{different} from what they already know.
+It is worth noting that the literature reviewed above provides a somewhat limited view of category generation: the main concern of the published experiments has been on the distributional correspondences between learned and generated categories.
+In this paper, we seek to identify other important constraints on the processes involved in creative generation.
+Specifically, we focus on the constraints imposed by category contrast.
+That is, in order to generate a novel concept, individuals must produce something that is in some capacity \textit{different} from what they already know.
 Thus, in a trivial sense, contrast can be viewed as a fundamental constraint on creative generation: new concepts must firstly be different from existing ones.
-However the role of contrast in category generation has not been systematically evaluated: how do people generate concepts such that the product is (A) different from what is already known, and (B) internally coherent?
+
+Although it is evident that people are \textit{capable} of creating new concepts and categories, it is not altogether clear how new concepts are systematically made different from what is already known.
+The hierarchical sampling model developed by \cite{jern2013probabilistic}, for example, does not provide any strong claims about how generated categories should contrast with learned ones: it is simply assumed that categories are generated in a random location in the domain.
+
+The Copy+Tweak account evaluated by Ward \& colleagues \citep{ward1995s,ward2002role} makes a slightly stronger claim: new categories should be .....
+
+%
 
@@
 \section{Snippets}
+
+\nbcnote{Just some random snippets that i don't want to delete yet...}
+
+How do people generate concepts such that the product is (A) different from what is already known, and (B) internally coherent?
+
+We also introduce a novel exemplar-based model of category generation, PACKER (\textit{Producing Alike and Contrasting Knowledge using Exemplar Representations}), which explains much of what we observe in the creative use of conceptual knowledge.
+In the sections below, we introduce the PACKER model and its core theoretical claims.
 

### 2017-03-09 Manuscripts/model-descriptions/copy-and-tweak/copy-and-tweak.tex
started a latex describing approaches to copy and tweak

@@
 \usepackage[letterpaper, margin=1in, headheight=15pt]{geometry}
-\usepackage{setspace, amsmath}
+\usepackage{setspace}
+\usepackage{amsmath}
+\usepackage{csquotes}
+\usepackage{fancyvrb}
+\usepackage[natbibapa]{apacite}
 
@@
 
-\section{The Exemplar Copy-And-Tweak Model}
+\section{The Copy-And-Tweak Model}
 
-This model assumed that people represent categories as a collection of stored observations.
-When prompted to generate new examples, they copy and randomly tweak one of the stored examples.
-Thus, the model's generation is a two-part process:
+This model was originally referred to as the ``Path-of-Least-Resistance'' model \citep{ward1994structured,ward1995s,ward2002role}, and later re-interpreted as the ``Copy-and-Tweak'' model \citep{jern2013probabilistic}.
+Under each name it was interpreted differently, and indeed the core psychological theory can be reasonably instantiated in a variety of ways.
 
-\begin{enumerate}
-  \item Select a source example from memory.
-  \item Generate an example that is similar-but-not-too-similar to the source.
-\end{enumerate}
+In the broadest sense, the theory claims that people generate a new item by retrieving an observation from memory and changing its features.
+However, the manner in which observations can be retrieved and changed, was never formally specified.
+Below, i describe how the model has been interpreted.
 
-More formally, let $x$ be a $j$-exemplar by $k$-feature matrix, corresponding to every stored exemplar. The probability that a source example $z$ is selected is given by:
+\section{Ward's ``Path-of-Least-Resistance'' model}
 
-\begin{equation}
-  p(z) = exp(f_j) / \sum_j{ exp(f_j) }
-\end{equation}
+Ward never provided a formal specification of his model, though it is described briefly in \cite{ward1994structured,ward1995s} and tested empirically in \cite{ward2002role}.
+As far as I know, these are the main places where the model is described.
+Each of these papers describes a similar concept, with some slight differences.
+Below are key passages from each paper.
 
-$f_j$ is a vector of values specifying each item's relative chance of being selected.
-In theory, $f_j$ may be set arbitrarily, but in practice it is sensible to differentiate between members of the target and contrast categories.
-For example, $f_j = -\gamma$ for members of contrast categories, and $f_j = \gamma$ for members of the target category.
-The resulting free parameter $\gamma$ ($>0$) thus controls the degree favoritism for members of the target category, with larger values producing larger differentials.
+\subsection{\cite{ward1994structured}}
+\begin{displayquote}
+``...One simple way to do so would be to suggest that the stored representation that is used for making category decisions is the same information used to generate a novel member of the category.
+When subjects are given the task of generating a novel animal, for example, the label "animal" might lead them to retrieve typical exemplars of their animal category.
+They would then use those activated representations as a starting point for the new creation.
+Because all of the models contain information that would at least translate into characteristic features of known category members, they all would predict that newly generated exemplars will possess those characteristic features.
+That is, the characteristic properties that are so influential in category decision making should also be influential in the development of novel instances.
+Presumably, a person using imagination would create a novel entity that was similar to the stored representation by projecting characteristic properties of that representation onto the entity.''
+\end{displayquote}
 
-After a source exemplar $z$ is selected, similarity between candidate generation options $y$ is computed:
+\subsection{\cite{ward1995s}}
+\begin{displayquote}
+``The path-of-least-resistance model assumes that category representations include specific exemplars that are embedded within broader knowledge frameworks, which give the exemplars their categorical coherence.
+It also states that truly useful, creative innovations that are novel and appropriate to a task are more likely (though not guaranteed) to occur when individuals access broader knowledge structures than when they simply retrieve exemplars.
+The exemplars are seen as largely static, uninterpreted entities, in contrast to the dynamic and flexible information in broader knowledge structures.
+Although one might randomly vary attributes of retrieved exemplars, without the guidance provided by access to broader structures, there is little chance of those novel variations' being appropriate for a given task.
+Thus, to the extent that individuals have access to well-developed explanatory frameworks, they are in a position to modify existing designs or develop new ones.''
+\end{displayquote}
 
-\begin{equation}
-  s(y,z) = exp( -c \sum_k{|y_k - z_k|w_k})
-\end{equation}
+\subsection{\cite{ward2002role}}
+\begin{displayquote}
+``The focus on representativeness stems from the path-of-least-resistance model (Ward, 1994, 1995), which attributes the resemblance between newly generated entities and known ones to the approach people take in generating those novel entities.
+The model proposes that, although people can adopt a variety of strategies for developing new ideas, a predominant approach is to retrieve specific known instances of the relevant concept and to project the properties of those instances onto the novel idea.
+Furthermore, the selection of instances is assumed to be guided by their representativeness.
+Items that are the most representative of the concept are the ones most likely to be retrieved and used as starting points for new ideas.
+In generating an imaginary animal, for example, a person would tend to move toward a level of abstraction more specific than animal and to gravitate toward representative instances, such as mammal and dog, rather than less representative ones, such as fish or bat.''
+\end{displayquote}
 
-Where $c$ and $w_k$ are the standard specificity and attention weights discussed elsewhere. The probability a candidate stimulus will be generated is given by:
+\subsection{What can we conclude about Ward's interpretation?}
 
-\begin{equation}
-  p(y|z) \propto  
-  \begin{cases}
-     exp(\theta s(y,z)), & \text{if } s(y,z) \leq \phi \\
-    0, & \text{otherwise}
-  \end{cases}
-  \label{eq:generation-choice}
-\end{equation}
+In the 1994 paper, Ward argues that the generated item should share the characteristic properties of the retrieved item, and differ along non-characteristic properties.
+More formally, this can be construed as a claim about how much the generated item will differ from the source, according to how much the known category varies along each feature.
+If a category does not vary much along a feature, then that feature should be viewed as characteristic to the category, and should not be changed along that feature in generation.
 
-where $\theta$ is a response determinism parameter.
-$\phi$ is a similarity tolerance parameter ($0 \leq \phi \leq 1$) specifying the amount of dissimilarity tolerated between $y$ and $z$.
-Larger values allow for more similar examples to be generated.
-The final choice is made using a normalization of all possible values of $p(y|z)$.
+As an example, imagine asking someone to generate a new breed of dog.
+Dog breeds do not vary much in \textit{number of legs}: having four legs is characteristic of the dog category, and so generated breeds will probably not differ.
+However, dog breeds vary a lot in \textit{fur color pattern}, and so this feature is likely to be changed.
+Key to this concept is hierarchical organization: people are assumed to be generating examples of categories sharing a higher level domain (i.e., dog breeds are examples of dogs).
+So when asked to generate a breed, they retrieve a representation of a breed, not a specific observation.
+This claim makes the theory a lot more like that of \cite{jern2013probabilistic}, and is also mirrored in the passage from the 1995 chapter.
 
-\subsection*{The Practical Implementation}
+Mapping this example to the artificial domain, the new breed of dogs is of course the Beta category: each Beta example corresponds to a single member of the new breed.
+When asked to generate the Beta category, the theory assumes that people retrieve their knowledge of the Alpha category (the whole thing, not it's specific examples) and change it along non-characteristic features.
+So the claim is that members of the Beta category should actually be \textit{similar} to members of the Alpha category along features that the Alphas do not vary along, just as a generated breed of dog will also have 4 legs.
 
-In practice, however, it is useful to be able to generate model predictions more deterministically (i.e., without respect to any given source exemplar $z$).
-Generalizing things a little, the overall probability that an item $y$ will be generated is the sum of its generation probabilities given every possible source example $z$, $p(y|z)$, weighted by the probability each source will be selected, $p(z)$:
+\begin{figure}
+    \begin{center}
+    \begin{BVerbatim}
+|
+| AAAA   BBBB
+|_____________
+    \end{BVerbatim}
 
-\begin{equation}
-  p(y) = \sum_z{p(z)p(y|z) }
-\end{equation}
+    \caption{Generated category differs along the variable axis, but do not along homogeneous axis.}
+    \end{center}
+\end{figure}
 
-In this way, source-independent probabilities can be obtained without monte-carlo simulation.
-First, the probability of each source $p(z)$ is computed (which is also required for simulation).
-Then, the pairwise similarity $s(y,z)$ between each source and each generation candidate is computed.
-Then the generation probability $p(y|z)$ is computed as per Equation \ref{eq:generation-choice}.
-Then $p(y)$ is computed as above.
+The 2002 paper is more concerned with the process of selecting the source.
+The claim is that the most representative examples are the most likely to be spontaneously retrieved, and thus used as the source.
+This claim makes more sense in a natural category domain where, for example, there are multiple breeds of dogs that are graded in terms of typicality.
+In an artificial domain, there aren't many categories and we assume that some are not more representative than others.
+
+However, we do have strong theories about the typicality of examples \textit{within} a category.
+When generating new members of a known category, the probability of retrieval can be straightforwardly estimated using similarity measures.
+
+\section{\citeauthor{jern2013probabilistic}'s ``Copy-and-Tweak'' model}
+
+\section{Our Interpretation}
+
+The core concept of Ward's model is based on hierarchically organized domains of natural categories.
+The artificial domain differs in that it is not clear Alphas and Betas share some higher level category, and so its not clear they ought to have something in common.
+
+% references section
+\clearpage
+\bibliographystyle{apacite}
+\bibliography{citations.bib}
+\clearpage
 

### 2017-03-10 Manuscripts/model-descriptions/copy-and-tweak/copy-and-tweak.tex
drafted ward copytweak latex

--- 2017-01-17 Manuscripts/model-descriptions/copy-and-tweak/copy-and-tweak.tex
@@
 
-This model assumed that people represent categories as a collection of stored observations.
+This model assumes that people represent categories as a collection of stored observations.
 When prompted to generate new examples, they copy and randomly tweak one of the stored examples.
@@
 
-More formally, let $x$ be a $j$-exemplar by $k$-feature matrix, corresponding to every stored exemplar. The probability that a source example $z$ is selected is given by:
-
-\begin{equation}
-  p(z) = exp(f_j) / \sum_j{ exp(f_j) }
-\end{equation}
-
-$f_j$ is a vector of values specifying each item's relative chance of being selected.
-In theory, $f_j$ may be set arbitrarily, but in practice it is sensible to differentiate between members of the target and contrast categories.
-For example, $f_j = -\gamma$ for members of contrast categories, and $f_j = \gamma$ for members of the target category.
-The resulting free parameter $\gamma$ ($>0$) thus controls the degree favoritism for members of the target category, with larger values producing larger differentials.
-
+More formally, let $x$ be a $j$-exemplar by $k$-feature matrix, corresponding to the stored exemplars associated with the target category.
+The probability that a source example $z$ is selected is uniform across all exemplars.
 After a source exemplar $z$ is selected, similarity between candidate generation options $y$ is computed:
@@
 
-Where $c$ and $w_k$ are the standard specificity and attention weights discussed elsewhere. The probability a candidate stimulus will be generated is given by:
+Where $c$ and $w_k$ are the standard specificity and attention weights discussed elsewhere.
+The goal of the model is to generate an item that is similar but sufficiently tweaked from the source.
+Thus, probability a candidate stimulus will be generated is given by:
 
 \begin{equation}
-  p(y|z) \propto  
-  \begin{cases}
-     exp(\theta s(y,z)), & \text{if } s(y,z) \leq \phi \\
-    0, & \text{otherwise}
-  \end{cases}
+    p(y|z)  = \dfrac
+    { \exp \left\{\theta \cdot s(y,z) \right\} I\left(s(y,z) \leq \tau\right) }
+    {\sum_i{\exp \left\{ \theta \cdot s(y_i,z) \right\} I\left(s(y_i,z) \leq \tau\right)}} 
     \label{eq:generation-choice}
 \end{equation}
+% 
+where $\theta$ is a response determinism parameter, $\tau$ is the similarity threshold, and $I(\cdot)$ is the indicator function, which returns 1 when it is passed a true expression and 0 otherwise.
+So, $I\left(s(y,z) \leq \tau\right)$ is 1 when the candidate exemplar $y$ is tweaked enough.
+When $\tau=1$, the threshold has no effect.
 
-where $\theta$ is a response determinism parameter.
-$\phi$ is a similarity tolerance parameter ($0 \leq \phi \leq 1$) specifying the amount of dissimilarity tolerated between $y$ and $z$.
-Larger values allow for more similar examples to be generated.
-The final choice is made using a normalization of all possible values of $p(y|z)$.
-
-\subsection*{The Practical Implementation}
-
-In practice, however, it is useful to be able to generate model predictions more deterministically (i.e., without respect to any given source exemplar $z$).
-Generalizing things a little, the overall probability that an item $y$ will be generated is the sum of its generation probabilities given every possible source example $z$, $p(y|z)$, weighted by the probability each source will be selected, $p(z)$:
+To obtain predictions not depending on a given source example, the model's predictions can be aggregated over all possible sources:
 
@@
 
-In this way, source-independent probabilities can be obtained without monte-carlo simulation.
-First, the probability of each source $p(z)$ is computed (which is also required for simulation).
-Then, the pairwise similarity $s(y,z)$ between each source and each generation candidate is computed.
-Then the generation probability $p(y|z)$ is computed as per Equation \ref{eq:generation-choice}.
-Then $p(y)$ is computed as above.
-
 \end{document}

### 2017-03-10 Manuscripts/model-descriptions/copy-and-tweak/path-least-resistance.tex
drafted ward copytweak latex

%!TEX output_directory = latex_out/

\documentclass[12pt]{article}
\usepackage[letterpaper, margin=1in, headheight=15pt]{geometry}
\usepackage{setspace}
\usepackage{amsmath}
\usepackage{csquotes}
\usepackage{fancyvrb}
\usepackage[natbibapa]{apacite}

\begin{document}

\section{Ward's ``Path-of-Least-Resistance'' model}

This model was originally referred to as the ``Path-of-Least-Resistance'' model \citep{ward1994structured,ward1995s,ward2002role}, and later re-interpreted as the ``Copy-and-Tweak'' model \citep{jern2013probabilistic}. The core psychological theory can be reasonably instantiated in a variety of ways.

In the broadest sense, the theory claims that people generate a new item by retrieving \textit{something} from memory and changing it \textit{somehow}. But Ward never provided a formal specification of his model. It was only described briefly in \cite{ward1994structured,ward1995s} and tested empirically in \cite{ward2002role}. I have pasted passages from these papers in Appendix \ref{ap:passages}. Below I describe what we can interpret from what he has written.

\subsection{An interpretation}

The path of least resistance theory described by Ward is quite a bit different from what \cite{jern2013probabilistic} implemented. The theory is definitely formulated for a hierarchical natural-categories domain, and according to Ward, the theory is of a process, so it ostensibly describes the processes that operate on representations.

The first distinction is that when people select a source, they are selecting one of many \textit{subcategories} (not examples), all sharing a higher-level class. For example, people are picking among categories of \textit{animals} (e.g., \textit{cat}, \textit{dog}, \textit{bear}), each of which has exemplars. They are thought to pick a subclass according to its representativeness of the higher-level category, and then change the non-characteristic features among members of the subclass to make something new.

Problematically, in an artificial domain, there aren't many categories and we assume that they are all equal in representativeness. Furthermore, unlike in a natural categories domain, it is not clear Alphas and Betas \textit{should} share some higher level category (it wouldn't be unreasonable for participants to view them as opposites), and so its not clear they ought to be related. So it is worth noting that this model is not well-suited for the artificial domain.

Here's what it would look like if we tried to implement Ward's ideas as closely as possible. Since the Alphas are the only other class, they would of course be retrieved. The tweaking is based on how characteristic each feature is of the source: the location of the new category is the same as the source but moved in some (unspecified, maybe random) direction along one or more features that are viewed as uncharacteristic of the source. We can use range as a proxy for characteristic-ness (less range = more representative), so the new category will differ along axes that the sources varies on.

\begin{figure}[ht]
    \begin{center}
    \begin{BVerbatim}
|
| AAAA   BBBB
|_____________
    \end{BVerbatim}
    \caption{Generated category differs only along the variable axis.}
    \label{fig:AB-Sample}
    \end{center}
\end{figure}

That means that if the Alphas are a row (taking a single Y-axis value), the Betas will share the same Y-axis value and differ along the X-axis, which is the opposite of what we've observed. See Figure \ref{fig:AB-Sample} for an example.

\subsubsection{Open questions}

Critically, Ward makes no claims about how categories are represented. For example: \textit{What is the source}? Is it a prototype? Or a (presumably representative) example? Or the parameters to a distribution?

Likewise, there are no claims about how examples are generated once the source subclass has been tweaked. Maybe the tweaking specifies the prototype of the category, but how are examples of that category distributed? Is it about maximizing similarity like in an exemplar approach? Or is it a sampling approach like \cite{jern2013probabilistic}? This question is dependent on the nature of category representation, but involves additional assumptions about the generation process.

\section{Formalizing the Model}

I can think of several reasonable formalizations, all following from the core claims made by Ward. Later I will write that math in these.

\subsection{A Prototype Model}

\subsection{An Exemplar Model}

\subsection{A Hierarchical Prototype Model}

% references section
\clearpage
\bibliographystyle{apacite}
\bibliography{citations.bib}

\clearpage
\appendix
\section{Passages from Ward's papers}
\label{ap:passages}

\subsection{\cite{ward1994structured}}
\begin{displayquote}
``...One simple way to do so would be to suggest that the stored representation that is used for making category decisions is the same information used to generate a novel member of the category. When subjects are given the task of generating a novel animal, for example, the label "animal" might lead them to retrieve typical exemplars of their animal category. They would then use those activated representations as a starting point for the new creation. Because all of the models contain information that would at least translate into characteristic features of known category members, they all would predict that newly generated exemplars will possess those characteristic features. That is, the characteristic properties that are so influential in category decision making should also be influential in the development of novel instances. Presumably, a person using imagination would create a novel entity that was similar to the stored representation by projecting characteristic properties of that representation onto the entity.''
\end{displayquote}

\subsection{\cite{ward1995s}}
\begin{displayquote}
``The path-of-least-resistance model assumes that category representations include specific exemplars that are embedded within broader knowledge frameworks, which give the exemplars their categorical coherence. It also states that truly useful, creative innovations that are novel and appropriate to a task are more likely (though not guaranteed) to occur when individuals access broader knowledge structures than when they simply retrieve exemplars. The exemplars are seen as largely static, uninterpreted entities, in contrast to the dynamic and flexible information in broader knowledge structures. Although one might randomly vary attributes of retrieved exemplars, without the guidance provided by access to broader structures, there is little chance of those novel variations' being appropriate for a given task. Thus, to the extent that individuals have access to well-developed explanatory frameworks, they are in a position to modify existing designs or develop new ones.''
\end{displayquote}

\subsection{\cite{ward2002role}}
\begin{displayquote}
``The focus on representativeness stems from the path-of-least-resistance model (Ward, 1994, 1995), which attributes the resemblance between newly generated entities and known ones to the approach people take in generating those novel entities. The model proposes that, although people can adopt a variety of strategies for developing new ideas, a predominant approach is to retrieve specific known instances of the relevant concept and to project the properties of those instances onto the novel idea. Furthermore, the selection of instances is assumed to be guided by their representativeness. Items that are the most representative of the concept are the ones most likely to be retrieved and used as starting points for new ideas. In generating an imaginary animal, for example, a person would tend to move toward a level of abstraction more specific than animal and to gravitate toward representative instances, such as mammal and dog, rather than less representative ones, such as fish or bat.''
\end{displayquote}

\end{document}

### 2017-03-13 Manuscripts/model-descriptions/copy-and-tweak/polr-packer.tex
wrote packer polr combo latex

%!TEX output_directory = latex_out/

\documentclass[12pt]{article}
\usepackage[letterpaper, margin=1in, headheight=15pt]{geometry}
\usepackage{setspace}
\usepackage{amsmath}
\usepackage{csquotes}
\usepackage[natbibapa]{apacite}

\begin{document}

\section{Combining PACKER and POLR}

\subsection{Selecting a Source}

The Path of Least Resistance (POLR) model assumes that people select as a source a representation (for Ward, this is a subcategory, for us it will be an exemplar) according to it's representativeness: more representative exemplars are more easily retrieved, and thus more likely to be used as the source.

Within the exemplar view, the typicality of a stimulus within a category (a good proxy for representativeness) has been measured as the total similarity between the stimulus and the exemplars assigned to the category \citep[see][]{nosofsky1988exemplar}. The typicality $t$ of a stored example $x$ belonging to category $h$ is computed as the summed similarity between $x$ and the members of category $h$, $x_j \in h$.
\begin{equation}
  t(x, h) = \sum_{x_j \in h}{s(x,x_j)}
\end{equation}

If only one category is known, the probability that $x$ will be retrieved is a function of its relative typicality among all members of the category:

\begin{equation}
    p(x | h)  = \dfrac
    { \exp \left\{\theta_{source} \cdot t(x, h) \right\} }
    {\sum_i{\exp \left\{\theta_{source} \cdot t(x_i, h) \right\}}} 
    \label{eq:source_ps}
\end{equation}

In Ward's view, the probability a given category of exemplars will be retrieved is also based on representativeness. Thus, when there are multiple categories in memory, $p(x)  = p(x | h)p(h)$. We will assume that all categories are equally representative in the artificial domain, however, so $p(h)$ is uniform across all categories.

\subsection{Generating the Next Item}

Once an item is generated, PACKER's rules kick in. The probability a stimulus $y$ will be generated is a function of its relative similarity to the source $x$:

\begin{equation}
    p(y | x)  = \dfrac
    { \exp \{\theta_{generate} \cdot s(y, x) f(x) \} }
    {\sum_i{\exp \{ \theta_{generate} \cdot s(y_i, x) f(x) \}}}
\end{equation}
% 
where $f(x)$ is parameterized as in PACKER. To obtain predictions not depending on any given source, the new item can be drawn from a combined distribution:

\begin{equation}
    p(y) = \sum_j { p(y|x_j)p(x_j) }
\end{equation}

In doing so, this model differs from PACKER only in that the contribution of exemplars is weighted according to representativeness. When $\theta_{source} = 0$, the two models are identical.

\subsection{Feature Weighting}

An additional claim made by Ward is that people do not tweak along features entirely at random. Instead, they attempt to maintain commonalities along features that are viewed as characteristic of the category, only changing features that are viewed as unimportant to the source.

This can be implemented as a change to the feature weights on similarity judgments:

\begin{equation}
  s(x_i,x_j) = exp( {-c \sum_{k}{|x_{ik} - x_{jk}|}w_k} )
\end{equation}

Specifically, the feature weights $w_k$ can be set according to each feature's characteristic-ness in the source category, measured as the relative amount of variability of that feature within the category: characteristic features should be less variable.

\begin{equation}
    w_k = \dfrac
    { \exp \{-\theta_{wts} \cdot \text{range}(k) \} }
    {\sum_i{\exp \{ -\theta_{wts}  \cdot  \text{range}(k_i)) \}}}
\end{equation}

\textbf{figure out if that produces the ward-predicted results later...}

% references section
\clearpage
\bibliographystyle{apacite}
\bibliography{citations}

\end{document}

### 2017-03-15 Manuscripts/model-descriptions/copy-and-tweak/copy-and-tweak.tex
copytweak back to jk13 spec, added softmax function

@@
 
-\section{The Exemplar Copy-And-Tweak Model}
+\section{The Copy-And-Tweak Model}
 
-This model assumes that people represent categories as a collection of stored observations.
-When prompted to generate new examples, they copy and randomly tweak one of the stored examples.
-Thus, the model's generation is a two-part process:
+This model, initially implemented by \cite{jern2013probabilistic} to explain exemplar generation in a multinomial space, is broadly based on Ward's path-of-least-resistance model \citep{ward1994structured,ward1995s,ward2002role}, but is quite different in implementation.
+This document describes how to implement their copy-and-tweak approach within our 2D continuous domain.
+See Appendix \ref{ap:jk13ap} for description of the model provided by \cite{jern2013probabilistic}.
 
-\begin{enumerate}
-  \item Select a source example from memory.
-  \item Generate an example that is similar-but-not-too-similar to the source.
-\end{enumerate}
+The model proposes that, when prompted to generate new examples, people copy a known example from the target class and change it a small amount along each dimension.
+The core difference between this account and the hierarchical sampling account is that correlations between features are not maintained by a copy-and-tweak system.
+As a result, this issue was the focus of \cite{jern2013probabilistic}.
 
-More formally, let $x$ be a $j$-exemplar by $k$-feature matrix, corresponding to the stored exemplars associated with the target category.
-The probability that a source example $z$ is selected is uniform across all exemplars.
-After a source exemplar $z$ is selected, similarity between candidate generation options $y$ is computed:
+In order to generate a new exemplar, the model computes the similarity between the members of the target category $x$and possible generation option $y$.
+Similarity $s$ between two examples, $x_i$ and $y_i$, is an inverse exponential function of distance:
 
 \begin{equation}
-  s(y,z) = exp( -c \sum_k{|y_k - z_k|w_k})
+  s(x_i,y_i) = exp( -c \sum_k{|x_{ik} - y_{ik}|w_k})
 \end{equation}
 
-Where $c$ and $w_k$ are the standard specificity and attention weights discussed elsewhere.
-The goal of the model is to generate an item that is similar but sufficiently tweaked from the source.
-Thus, probability a candidate stimulus will be generated is given by:
+Where $c$ and $w_k$ are the standard specificity and attention weights discussed elsewhere \citep[see][]{nosofsky1984choice}.
 
 \begin{equation}
-    p(y|z)  = \dfrac
-    { \exp \left\{\theta \cdot s(y,z) \right\} I\left(s(y,z) \leq \tau\right) }
-    {\sum_i{\exp \left\{ \theta \cdot s(y_i,z) \right\} I\left(s(y_i,z) \leq \tau\right)}} 
+  s(x_i,y_i) = \sum_i{s(y, x_i)}
+\end{equation}
+
+The probability that a given candidate $y$ will be generated is based on its summed similarity to all members of the target category $x$: $\sum_i{s(y, x_i)}$.
+\begin{equation}
+    p(y)  = \dfrac
+    { \exp \left\{\theta \cdot \sum_i{s(y, x_i)} \right\} }
+    {\sum_j{\exp \left\{ \theta \cdot \sum_i{s(y_j, x_i)} \right\} }} 
     \label{eq:generation-choice}
@@
 % 
-where $\theta$ is a response determinism parameter, $\tau$ is the similarity threshold, and $I(\cdot)$ is the indicator function, which returns 1 when it is passed a true expression and 0 otherwise.
-So, $I\left(s(y,z) \leq \tau\right)$ is 1 when the candidate exemplar $y$ is tweaked enough.
-When $\tau=1$, the threshold has no effect.
+where $\theta$ is a response determinism parameter.
+Known members of the target category cannot be generated, and so their probability is manually set to 0.
+If there are no known members of the target category, generation is random (uniform probabilities over $y$).
 
-To obtain predictions not depending on a given source example, the model's predictions can be aggregated over all possible sources:
+\subsection{Parameter info}
+\begin{itemize}
+    \setlength\itemsep{-0.5em}
+    \item $c$. Scalar, $c>0$. Specificity parameter.
+    \item $\theta$. Scalar, $\theta > 0$. Response determinism parameter.
+\end{itemize}
+
+% references section
+\clearpage
+\bibliographystyle{apacite}
+\bibliography{citations.bib}
+
+\clearpage
+\appendix
+\section{Description from \cite{jern2013probabilistic}}
+\label{ap:jk13ap}
+
+\subsection{Passages from the Introduction}
+
+\subsubsection{Page 93}
+\begin{displayquote}
+Exemplar models like the GCM (Nosofsky, 1986) and ALCOVE (Kruschke, 1992) classify novel exemplars by computing their similarity to previously observed exemplars stored in memory.
+These models can be extended to account for generation by assuming that stored exemplars can be copied and changed in some way in order to produce novel exemplars.
+We will refer to the resulting approach as the copy-and-tweak account, and the same basic approach has been proposed by several previous authors (Barsalou \& Prinz, 1997; Barsalou, 1999; Ward, 1995; Ward, Patterson, Sifonis, Dodds, \& Saunders, 2002; Ward, Patterson, \& Sifonis, 2004).
+
+Although the copy-and-tweak account may seem conceptually different from the sampling account, in some cases the two approaches are equivalent.
+Consider, for example, a problem in which the exemplars are points in n-dimensional space.
+A copy-and-tweak account would retrieve one of these exemplars and change it slightly along each dimension.
+This procedure is conceptually identical to sampling from a sum of normal distributions centered on each observed exemplar, which is formally equivalent to the GCM (Ashby \& Alfonso-Reese, 1995; Jäkel, Schölkopf, \& Wichmann, 2008).
+Thus, in many cases, applying copy-and-tweak to an exemplar model will be equivalent to sampling from the generation distribution learned by a generative model.
+The relationship between copy-andtweak and the sampling account suggests that both accounts will make similar predictions for problems in which people generate from categories that correspond to clusters in a similarity space.
+The two accounts, however, will diverge in cases where category membership cannot naturally be captured by similarity.
+Our first set of experiments uses structured categories that are specifically designed to distinguish between the two accounts.
+\end{displayquote}
+
+\subsubsection{Pages 103-104}
+\begin{displayquote}
+As described earlier, the copy-and-tweak approach converts an exemplar model of classification into a model of generation.
+The approach proposes that new exemplars are generated by copying one of the exemplars stored in memory and then modifying some features of the copy.
+When we introduced the category used for Experiments 1 and 2, we noted that our stimuli were carefully chosen to distinguish between the structured sampling account and the copy-and-tweak account.
+The structured sampling model is capable of learning that pairs of letters are grouped into cohesive parts and therefore predicts that people will tend to generate novel combinations of parts.
+In contrast, a copy-and-tweak model will generate many exemplars that are not valid exemplars.
+For example, after observing the four exemplars in Fig.
+4b, a copy-and-tweak model might produce a new exemplar by copying the first exemplar and changing the top letter from D to X.
+
+We implemented a copy-and-tweak model based largely on the GCM (Nosofsky, 1986).
+The GCM makes classification judgments by computing the similarity between a novel exemplar and all of the stored exemplars, and similarity is computed as a function of pairwise distance between exemplars.
+Because our stimuli used discrete non-ordinal features, we used the Hamming distance, which measures the number of features that are different.
+As explained earlier, the GCM is equivalent to a generative model that maintains a probability distribution with peaks at the observed exemplars (Ashby \& Alfonso-Reese, 1995; Jakel et al., 2008).
+The generative model that corresponds to the GCM maintains a probability distribution over a continuous similarity space, and we implemented an analogous approach that maintains a probability distribution over our space of discrete flu genomes.
+\end{displayquote}
+
+\subsection{Their appendix}
+
+\begin{displayquote}
+
+The copy-and-tweak model uses similarity to observed exemplars as the basis for both classification and generation.
+We used a similarity rule derived from the GCM (Nosofsky, 1986) that defines the similarity $s$ between exemplars $x_1$ and $x_2$ as $s(x_1,x_2) = \exp (-c \cdot d(x_1,x_2))$, where $c$ is a scaling parameter and $d(x_1,x_2)$ is the distance between the two exemplars.
+We set $c$ to 1 and used the Hamming distance, which counts the number of features for which the two exemplars have different values.
+
+To derive classification predictions from the copy-and-tweak model for the rating task, we summed the similarities of each rating exemplar $x_r$ to each of the training exemplars $x_i$: $\text{rating}(x_r) =\sum_i{s(x_r,x_i)}$.
+We then applied a z-score transformation to produce the predictions in Fig.
+10.
+
+To derive generation predictions from the copy-and-tweak model, we computed similarity ratings for all possible exemplars and drew 2,000,000 samples from a Multinomial distribution based on the ratings:
 
 \begin{equation}
-  p(y) = \sum_z{p(z)p(y|z) }
+  \widetilde{x} \sim \text{Multinomial}([R \cdot \text{rating}(\widetilde{x}_1) ... \text{rating}(\widetilde{x}_n)])
 \end{equation}
+% 
+where $R = 1/\sum_i{\text{rating}(\widetilde{x}_i)}$ is a normalization constant.
+The reported predictions are based on a model which considers all possible exemplars that can be constructed from the 16 observed letters ($n = 16^4$ ).
+We also implemented an alternative model that considers only exemplars in which each letter can only appear in positions where it has been previously observed (n = 256).
+The two models produced similar predictions, but the former version performed slightly better.
+\end{displayquote}
 

### 2017-03-15 Modules/Classes/CopyTweak.py

@@ -1,36 +1,43 @@
 import numpy as np
 
 # imports from module
-import Modules.Funcs as funcs
-from Model import Model
+import Modules.Funcs as Funcs
+from Model import Exemplar
 
-class CopyTweak(Model):
+class CopyTweak(Exemplar):
 	"""
-		Exemplar similarity implementation of the copy-and-tweak model.
+		Continuous implementation of the copy-and-tweak model.
 	"""
 
 	model = 'Copy and Tweak'
 	parameter_names = [
 		'specificity', # > 0
-		'tolerance', # 0 < tolerance <= 1
 		'determinism' # > 0
 	]
 
+
 	@staticmethod
-	def rvs():
-		params = [
-			np.random.uniform(0.1, 6.0), # specificity
-			np.random.uniform(0.5, 1.0), # tolerance. biased high.
-			np.random.uniform(0.1, 6.0) # determinism
-		]
-		return params
+	def rvs(fmt = dict):
+		"""
+		Return random parameters in dict or list format.
+		"""
+		params = [np.random.uniform(0.1, 6.0), # specificity
+							np.random.uniform(0.1, 6.0)] # determinism
+
+		if fmt not in [dict, list]:
+			raise Exception('Format must be dict or list.')
+
+		if fmt == list:
+			return params
+		else:
+			return dict(zip(CopyTweak.parameter_names, params))
 
 
 	def _param_handler_(self):
 		super(CopyTweak, self)._param_handler_()
 		if self.specificity <= 0: self.specificity = 1e-10
-		if self.tolerance > 1.0: self.tolerance = 1.0
-		if self.determinism <= 0: self.determinism = 1e-10
+		if self.determinism < 0: self.determinism = 0.0
+
 
 	def get_generation_ps(self, stimuli, category):
 
@@ -40,36 +47,15 @@ class CopyTweak(Model):
 			ncandidates = stimuli.shape[0]
 			return np.ones(ncandidates) / float(ncandidates)
 
-		# get source probabilities
-		nsources = self.nexemplars[category]
-		sources = self.categories[category]
-		source_ps = np.ones(nsources) / float(nsources)
-
 		# get pairwise similarities
-		distances  = funcs.pdist(stimuli, sources, w = self.wts)
-		similarity = np.exp(-1.0 * self.specificity * distances)
+		similarity = self._sum_similarity(stimuli, self.categories[category])
+
+		# NaN out known members
+		known_members = Funcs.intersect2d(stimuli, self.categories[category])
[…]

### 2017-03-15 Modules/Classes/Model.py

@@ -2,7 +2,7 @@ import abc
 import numpy as np
 
 # imports from module
-import Modules.Funcs as utils
+import Modules.Funcs as Funcs
 
 
 class Model(object):
@@ -137,7 +137,7 @@ class Model(object):
 	def _update_(self):
 		"""
 		Generate descriptives about the known categories. 
-		The fields defined below change when items are generated or fotgotten,
+		The fields defined below change when items are generated or forgotten,
 		so this function is called whenever there is a change to model memory.
 		"""
 		self.ncategories = len(self.categories)
@@ -157,9 +157,9 @@ class Model(object):
 
 	def simulate_generation(self, stimuli, category, nexemplars = None):
 		"""
-		Simulate the generation of nexemplars, sourced from stimuli, into a category.
+		Simulate the generation of n-exemplars, sourced from stimuli, into a category.
 		The resulting category will be added to the model's memory, and the identity 
-		of the generated iutems wilol be returned.
+		of the generated items will be returned.
 		"""
 
 		if nexemplars is None:
@@ -175,7 +175,7 @@ class Model(object):
 
 			# compute probabilities, then pick an item
 			ps = self.get_generation_ps(stimuli, category)
-			num = utils.wpick(ps)
+			num = Funcs.wpick(ps)
 			values = np.atleast_2d(stimuli[num,:])
 			generated_examples.append(num)
 
@@ -188,3 +188,34 @@ class Model(object):
 			self._update_()
 
 		return generated_examples
+
+
+
+class Exemplar(Model):
+	"""		
+	Abstract base class for Exemplar models. Basically this is used to add 
+	a function computing summed similarity.
+	"""
+	__metaclass__ = abc.ABCMeta
+
+	def _sum_similarity(self, X, Y, 
+		param = 1.0,	
+		wts  = None, 
+		c = None ):
+		""" 
+		function to compute summed similarity along rows of 
+		X across all items in Y. Resulting array will have one element 
+		per row of X.
+
+		the "param" argument acts as a multiplier for similarities prior to summation
+		"c" and "wts" can be arbitrary if supplied, or based on the model attribute if not
+		"""
+
+		# set weights and c
+		if wts is None: wts = self.wts
+		if c is None: c = self.specificity
+
+		distance   = Funcs.pdist(np.atleast_2d(X), np.atleast_2d(Y), w = wts)
+		similarity = np.exp(-float(c) * distance)
+		similarity = similarity * float(param)
+		return np.sum(similarity, axis = 1)

### 2017-03-15 Modules/Classes/__init__.py

@@ -2,4 +2,3 @@ import Optimize
 from CopyTweak import CopyTweak
 from Packer import Packer
 from ConjugateJK13 import ConjugateJK13
-from PPOLR import PPOLR

### 2017-03-16 Manuscripts/cog-psych/cogpsych.tex
update to intro

@@
 
-Creating Something Different: Similarity and Contrast Effects in Concept Generation.
+Creating Something Different: Similarity and Contrast in Concept Generation.
 % ---- other ideas:
@@
 The creation of novel concepts and ideas is a highly intriguing -- yet infrequently studied -- topic of research in human cognition.
-The creative use of conceptual knowledge is a core element of scientific investigation, wherein the generation of new ideas is critical to designing experiments and explaining observations.
-However, due to its complexity, little is presently known about the cognitive processes underlying the creative generation of new concepts.
+The creative use of conceptual knowledge is, for example, a core element of scientific investigation, wherein the generation of new ideas is critical to designing experiments and explaining observations.
+However, due to its complexity, little is presently known about the cognitive processes underlying the generation of new concepts.
 
-Much of what we know about conceptual generation comes from the foundational literature on creative cognition.
+Much of what we know about concept generation comes from the foundational literature on creative cognition.
 In a classic series of reports, Ward \& colleagues \citep{ward1995s,ward1994structured,marsh1999inadvertent,ward2002role,smith1993constraining} established that category generation is highly constrained by prior knowledge: generated categories tend to consist of features observed in known categories, and they tend to exhibit the distributional properties as found in known categories.
@@
 
-\nbcnote{Explain Copy + Tweak here?}
-
 Much of the work from this area \citep[e.g.,][]{smith1993constraining,marsh1999inadvertent} focuses on how sample cues (such as an example of a species generated by other participants) can drastically diminish creativity.
@@
 They indicate that people rely strongly on prior knowledge in the creative process, and people generate concepts in accord with what they already know.
-Problematically, however, the experimental paradigms employed in these studies were relatively uncontrolled.
-Specifically, in order to allow for maximum creative freedom, participants were not constrained as to what they could generate.
-While these procedures do ensure creative freedom, they render difficult the development of formal theories to explain conceptual generation.
+
+Theoretical accounts of these effects were primarily grounded within the categorization literature.
+For example, the predominant ``Path of Least Resistance'' account \citep[see][]{ward1994structured,ward1995s,ward2002role} proposes that, when generating a new species of animal, people retrieve from memory a known subcategory of animals (e.g., \textit{bird}, \textit{dog}, \textit{horse}), and simply change some of the features to make something new.
+People were thought to change only features that are not characteristic of the retrieved category (e.g., if \textit{bird} was retrieved, the presence of \textit{wings} would not change, but \textit{color} might).
+This theory incorporates elements of the highly influential basic-level categories framework \citep{rosch1975cognitive,rosch1976basic}, as well as the exemplar view \citep{medin1978context,nosofsky1984choice,nosofsky1986attention}.
+Problematically, however, the experimental paradigms employed in these studies were relatively uncontrolled, precluding the development of formal approaches.
+Specifically, participants in these studies are typically asked to generate objects that they have a rich degree of prior knowledge about (e.g., plants, animals, tools, toys), and so it is difficult to apply formal models to the data.
 
@@
 Following a training phase during which the experimenter-defined categories were learned, participants were asked to generate novel categories of crystals.
-In a finding mirroring that of the \cite{ward1994structured} studies, \citeauthor{jern2013probabilistic} found that participants generated categories with the same distributional properties as the experimenter-defined categories: for example, after training on categories with a positive correlation between the size and saturation features (larger sized crystals were more saturated), participants generated novel categories with a positive correlation.
+In a finding mirroring that of the \cite{ward1994structured} studies, \citeauthor{jern2013probabilistic} found that participants generated categories with the same distributional properties as the experimenter-defined categories: for example, after training on categories with a positive correlation between the size and saturation features (larger sized crystals were more saturated), participants generated novel categories with the same positive correlation.
+This finding is notable, as it demonstrates that category generation can be studied in a well-known and highly controlled experimental paradigm.
 
-The \cite{jern2013probabilistic} report thus represents an important step towards understanding the cognitive processes underlying creative generation: not only did they find that generation in an artificial domain resembles the more naturalistic cases observed by Ward \& colleagues, but they also found that theories of conceptual generation could be evaluated more formally through comparison to the participant-generated categories.
+The authors evaluated the predictions of several formal models on their data.
+Most notably, they showed that a Bayesian hierarchical sampling model provided the strongest account.
+Their model represents categories as multivariate normal distributions within the space, and from exposure to members of observed categories the model induces a domain-wide representation of the distributional commonalities among the observed categories.
+When generating a new category, the model samples a new distribution from the domain, and thus the distributional properties of observed categories will be preserved within the generates category.
 
-\nbcnote{explain hierarchical sampling here, maybe also their implementation of copy+tweak?
-Or maybe we should step away from copy+tweak as a relevant option.
-Or maybe we focus on systematically comparing their copy and tweak to PACKER with specific params?
-This copy and tweak issue makes for a mess...
-}
+\nbcnote{Maybe we need a more technical view of the model here?}
+
+\cite{jern2013probabilistic} additionally tested a "copy-and-tweak" model that broadly resembles the earlier "Path of Least Resistance" account: the core proposal is that participants generate new items by copying stored examples from memory and tweaking them to generate something new.
+The copy-and-tweak model differs from the path of least resistance account in that it notably omits the hierarchical organization of categories, as well as selectivity in which features are changed.
+Instead, their copy-and-tweak model corresponds to a direct exemplar-similarity approach \citep[e.g.,][]{nosofsky1984choice}: participants are thought to generate new items according to their similarity to known members of the target category.
+This model provided a poor account of the observed data: indeed, the experiments devised by \citeauthor{jern2013probabilistic} were specifically designed to challenge this model.
+However, it's application is notable as a first step toward explaining category generation using well-known formal approaches from the categorization literature.
 
@@
 
-It is worth noting that the literature reviewed above provides a somewhat limited view of category generation: the main concern of the published experiments has been on the distributional correspondences between learned and generated categories.
+It is worth noting that the literature reviewed above provides a somewhat limited view of category generation: the main concern of the published experiments has been on the distributional correspondences between learned and generated categories, and as a result most of the modeling efforts have been towards explaining those effects.
 In this paper, we seek to identify other important constraints on the processes involved in creative generation.
@@
 The hierarchical sampling model developed by \cite{jern2013probabilistic}, for example, does not provide any strong claims about how generated categories should contrast with learned ones: it is simply assumed that categories are generated in a random location in the domain.
-
-The Copy+Tweak account evaluated by Ward \& colleagues \citep{ward1995s,ward2002role} makes a slightly stronger claim: new categories should be .....
-
-%
 

### 2017-03-16 Manuscripts/cog-psych/cogpsych.tex
intro drafted

@@
 
-Theoretical accounts of these effects were primarily grounded within the categorization literature.
+Theoretical accounts of these effects have primarily been grounded within the categorization literature.
 For example, the predominant ``Path of Least Resistance'' account \citep[see][]{ward1994structured,ward1995s,ward2002role} proposes that, when generating a new species of animal, people retrieve from memory a known subcategory of animals (e.g., \textit{bird}, \textit{dog}, \textit{horse}), and simply change some of the features to make something new.
@@
 Most notably, they showed that a Bayesian hierarchical sampling model provided the strongest account.
-Their model represents categories as multivariate normal distributions within the space, and from exposure to members of observed categories the model induces a domain-wide representation of the distributional commonalities among the observed categories.
+Their model represents categories as multivariate normal distributions within the space, and from exposure to category members the model induces a domain-wide representation of the distributional commonalities among the observed categories.
 When generating a new category, the model samples a new distribution from the domain, and thus the distributional properties of observed categories will be preserved within the generates category.
@@
 It is worth noting that the literature reviewed above provides a somewhat limited view of category generation: the main concern of the published experiments has been on the distributional correspondences between learned and generated categories, and as a result most of the modeling efforts have been towards explaining those effects.
-In this paper, we seek to identify other important constraints on the processes involved in creative generation.
-Specifically, we focus on the constraints imposed by category contrast.
-That is, in order to generate a novel concept, individuals must produce something that is in some capacity \textit{different} from what they already know.
+In this paper, we investigate another important constraint on the creative use of conceptual knowledge: category contrast.
+In order to generate a novel concept, individuals must produce something that is in some capacity \textit{different} from what they already know.
 Thus, in a trivial sense, contrast can be viewed as a fundamental constraint on creative generation: new concepts must firstly be different from existing ones.
 
-Although it is evident that people are \textit{capable} of creating new concepts and categories, it is not altogether clear how new concepts are systematically made different from what is already known.
-The hierarchical sampling model developed by \cite{jern2013probabilistic}, for example, does not provide any strong claims about how generated categories should contrast with learned ones: it is simply assumed that categories are generated in a random location in the domain.
+Although it is evident that people are \textit{capable} of creating new concepts and categories, it is not entirely clear how new concepts are systematically made different from what is already known.
+The hierarchical sampling model developed by \cite{jern2013probabilistic}, for example, does not provide any strong claims about how generated categories should contrast with learned ones: the model only proposes that generated categories are sampled from the same underlying domain distribution as observed categories, and will thus share a common distributional structure.
+The authors do not make predictions about the \textit{location} of the category within the domain (the perceptual instantiation of category members).
+Indeed, under a strict interpretation of the hierarchical Bayesian approach, the generated category should actually overlap substantially with the known categories (as it's location is drawn based on those categories).
+Alternatively, the model's placement of the new category may simply be random (this was the case in the simulations conducted by \citeauthor{jern2013probabilistic}).
+
+\nbcnote{Maybe tune that language to be more in line with the probabilistic world?}
+
+The copy-and-tweak model tested by \cite{jern2013probabilistic} also claims little about how generated categories should contrast with what is already known.
+In their simulations, the model was only tested on generation after the learner had been exposed to members of the target category, and so the model's ability to generate a new category was not evaluated.
+However, the model's generation is based exclusively on similarity to known members of the target category; when there are no members of the target category, generation is presumably random.
+
+\nbcnote{Maybe mention something about path of least resistance prediction that contrast is on non-characteristic dimensions? }
+
+\subsection{Overview}
+
+As noted in the sections above, research on the creative generation of novel concepts has thusfar focused on the finding that generated categories tend to possess distributional commonalities with known categories.
+However, a fundamental goal of concept generation is to create something \textit{new} (i.e., different from what is already known).
+The manner in which generated categories differ from known ones is, nonetheless, poorly understood: existing formal approaches do not make strong predictions about how creatively generated concepts should systematically differ from existing ones.
+
+In the sections below, we present two experiments demonstrating systematic effects of category contrast on creative generation.
+We also introduce a novel, exemplar-based model to explain these effects: PACKER (\textit{Producing Alike and Contrasting Knowledge using Exemplar Representations}).
+Finally, we report formal simulations comparing PACKER to the hierarchical sampling and copy-and-tweak models developed by \cite{jern2013probabilistic}.
 

### 2017-03-16 Manuscripts/cog-psych/cogpsych.tex
updates to e1, plot module

@@
 \usepackage{fancyhdr}
-\usepackage{fancyvrb}
 \usepackage{csquotes}
@@
 In the sections below, we present two experiments demonstrating systematic effects of category contrast on creative generation.
+Our experiments conform strictly to the artificial category learning paradigm from \cite{jern2013probabilistic}: participants are first exposed to a single, experimenter-defined category, and are then asked to generate members of a new category.
 We also introduce a novel, exemplar-based model to explain these effects: PACKER (\textit{Producing Alike and Contrasting Knowledge using Exemplar Representations}).
@@
 
+\section{Experiment 1}
+
+To begin our investigation, we developed a artificial, two dimensional domain of squares, varying in color and size (see Figure \ref{fig:sample-stimuli} for samples).
+We created three unique category types within this domain, shown in Figure \ref{fig:e1-conditions}.
+The category types were designed to possess strong differences in the distributional structure of the experimenter-defined (`A') category, allowing us to observe the effect of category contrast under a variety of types of prior knowledge.
+These conditions also provide an opportunity to replicate classic effects observed in previous reports \citep[e.g.,][]{jern2013probabilistic,ward1994structured,ward2002role,marsh1999inadvertent,smith1993constraining}.
+
+\nbcnote{This feels a little weak in motivation: how do these conditions specifically allow us to show is these effects?
+Did we come up with these out of thin air?
+So maybe foreshadow that you hardly need to look for these effects, they're \textit{everywhere}.
+Maybe there is also a problem that at this point it is not exactly clear what contrast means within this scenario.}
+
+\begin{figure}
+    \begin{center}
+    \inputpgf{figs/}{e1-conditions.pgf}
+    \caption{Category types tested in Experiment 1.}
+    \label{fig:e1-conditions}
+    \end{center}
+\end{figure}
+
+In the `Cluster' type, the experimenter-defined category is a tight cluster of examples in the space (e.g., all members of category `A' are small and dark in color).
+In the `Row' type, the category has a row pattern across the space: varying along one feature but not the other (e.g., members of category `A' are small and vary in color).
+In the `XOR' type, the experiment-defined category consists of two clusters separated in opposite corners of the space category, conforming to the exclusive-or logical structure (e.g., members of category `A' can be small and dark or large and light).
+
+\subsection{Participants \& Materials}
+
+183 participants were recruited from Amazon Mechanical Turk.
+Participants were randomly assigned to one condition: 64 participants were assigned to the Cluster condition, 61 were assigned to the Row condition, and 58 were assigned to the XOR condition.
+Stimuli were squares varying in color (RGB 25--230) and side length (3.0--5.8cm).
+These stimuli are slight variants of those used by \cite{conaway2016similar}, see Figure \ref{fig:sample-stimuli} for samples.
+The assignment of perceptual features (color, size) to axes of the domain space (x, y) was counterbalanced across participants.
+
+\begin{figure}
+    \begin{center}
+    \inputpgf{figs/}{stimuli-samples.pgf}
+    \caption{Sample stimuli used in Experiments 1 and 2.}
+    \label{fig:sample-stimuli}
+    \end{center}
+\end{figure}
+
+\subsection{Procedure}
+
+Participants began the experiment with a short training phase (3 blocks of 4 trials), where they observed exemplars belonging to the `Alpha' category.
+Participants were instructed to learn as much as they can about the Alpha category, and that they would answer a series of test questions afterwards.
+On each trial, a single Alpha category exemplar was presented, and participants were given as much time as they desired before moving on.
+Each block consisted of a single presentation of each of the members of the Alpha category, in a random order.
+Participants were shown the range of possible colors and sizes prior to training.
+
+Following the training phase, participants were asked to generate four examples belonging to another category called `Beta'.
+As in \citet{jern2013probabilistic}, generation was completed using a sliding-scale interface.
+Two scales controlled the features (color, size) of the generated example.
+An on-screen preview of the example updated whenever one of the features was changed.
+Participants could generate any example along an evenly-spaced 9x9 grid, except for any previously generated Beta exemplars.
+Neither the members of the Alpha category nor the previously generated Beta examples were visible during generation.
+Prior to beginning the generation phase, participants read the following instructions:
+
+\begin{displayquote}
+As it turns out, there is another category of geometric figures called ``Beta''.
+Instead of showing you examples of the Beta category, we would like to know what you think is likely to be in the Beta category.
+
+You will now be given the chance to create examples of any size or color in order to show what you expect about the Beta category.
+You will be asked to produce 4 Beta examples - they can be quite similar or quite different to each other, depending on what you think makes the most sense for the category.
+
+Each example needs to be unique, but the computer will let you know if you accidentally create a repeat.
+\end{displayquote}
+
+Following generation, participants completed a generalization phase wherein they classified novel examples into the Alpha and Beta categories without feedback.
+On each trial, a single example was presented, and participants were asked to classify it by clicking buttons labeled ``Alpha'' or ``Beta''.
+Participants classified a total of 81 items sampled along a 9x9 grid, including the members of the Alpha and Beta categories (randomly intermixed).
+These data were, however, collected to address a separate set of questions, and we do not discuss them in this paper.
+
 \clearpage
+
+\section{Experiment 2}
+
+\subsection{Participants \& Materials}
+
+122 participants were recruited from Amazon Mechanical Turk.
+61 participants were randomly assigned to the Middle and Bottom conditions each.
+The stimuli were exactly as in Experiment 1.
+Again, the assignment of perceptual features (color, size) to axes of the domain space (x, y) was counterbalanced across participants.
+
+\subsection{Procedure}
+
+The procedure was exactly as in Experiment 1: participants first completed a short training phase, followed by the generation phase, followed by the generalization phase (data from this phase is not discussed in this report).
 
@@
 
-\clearpage
-
-\section{Experiment 1}
-
-\subsection{Participants \& Materials}
-
-183 participants were recruited from Amazon Mechanical Turk.
-Participants were randomly assigned to one condition: 64 participants were assigned to the Cluster condition, 61 were assigned to the Row condition, and 58 were assigned to the XOR condition.
-Stimuli were squares varying in color (RGB 25--230) and side length (3.0--5.8cm).
-These stimuli are slight variants of those used by \cite{conaway2016similar}, see Figure \ref{fig:sample-stimuli} for samples.
-The assignment of perceptual features (color, size) to axes of the domain space (x, y) was counterbalanced across participants.
-
-\begin{figure}
-    \begin{center}
-    \inputpgf{figs/}{stimuli-samples.pgf}
-    \caption{Sample stimuli used in Experiments 1 and 2.}
-    \label{fig:sample-stimuli}
-    \end{center}
-\end{figure}
-
-\subsection{Procedure}
-
-Participants began the experiment with a short training phase (3 blocks of 4 trials), where they observed exemplars belonging to the `Alpha' category.
-Participants were instructed to learn as much as they can about the Alpha category, and that they would answer a series of test questions afterwards.
-On each trial, a single Alpha category exemplar was presented, and participants were given as much time as they desired before moving on.
-Each block consisted of a single presentation of each of the members of the Alpha category, in a random order.
-Participants were shown the range of possible colors and sizes prior to training.
-
-Following the training phase, participants were asked to generate four examples belonging to another category called `Beta'.
-As in \citet{jern2013probabilistic}, generation was completed using a sliding-scale interface.
-Two scales controlled the features (color, size) of the generated example.
-An on-screen preview of the example updated whenever one of the features was changed.
-Participants could generate any example along an evenly-spaced 9x9 grid, except for any previously generated Beta exemplars.
-Neither the members of the Alpha category nor the previously generated Beta examples were visible during generation.
-Prior to beginning the generation phase, participants read the following instructions:
-
-\begin{displayquote}
-As it turns out, there is another category of geometric figures called "Beta".
-Instead of showing you examples of the Beta category, we would like to know what you think is likely to be in the Beta category.
-
-You will now be given the chance to create examples of any size or color in order to show what you expect about the Beta category.
-You will be asked to produce 4 Beta examples - they can be quite similar or quite different to each other, depending on what you think makes the most sense for the category.
-
-Each example needs to be unique, but the computer will let you know if you accidentally create a repeat.
-\end{displayquote}
-
-Following generation, participants completed a generalization phase wherein they classified novel examples into the Alpha and Beta categories without feedback.
-On each trial, a single example was presented, and participants were asked to classify it by clicking buttons labeled ``Alpha'' or ``Beta''.
-Participants classified a total of 81 items sampled along a 9x9 grid, including the members of the Alpha and Beta categories (randomly intermixed).
-These data were, however, collected to address a separate set of questions, and we do not discuss them in this paper.
-
-\clearpage
-
-\section{Experiment 2}
-
-\subsection{Participants \& Materials}
-
-122 participants were recruited from Amazon Mechanical Turk.
-61 participants were randomly assigned to the Middle and Bottom conditions each.
-The stimuli were exactly as in Experiment 1.
-Again, the assignment of perceptual features (color, size) to axes of the domain space (x, y) was counterbalanced across participants.
-
-\subsection{Procedure}
-
-The procedure was exactly as in Experiment 1: participants first completed a short training phase, followed by the generation phase, followed by the generalization phase (data from this phase is not discussed in this report).
-
 \section{Acknowledgments}

### 2017-03-17 Manuscripts/cog-psych/cogpsych.tex
edited PACKER math, added copy-tweak relation section

@@
 
+\clearpage
+
 \section{The PACKER Model}
 
-The PACKER model is an extension of the influential Generalized Context Model of category learning \citep[GCM;][]{nosofsky1984choice}.
-It assumes that each category is encoded by a set of exemplars within a $k$-dimensional psychological space, and that generation is constrained by both similarity to members of the target category (the category in which a stimulus is being generated) as well as similarity to members of other categories.
+The PACKER model is an extension of the influential Generalized Context Model of category learning \citep[GCM;][]{nosofsky1984choice,nosofsky1986attention}.
+The GCM simulates categorization under the assumption that learners represent categories as a collection of exemplars, corresponding to the labeled stimuli they have observed.
+The exemplars are encoded within a $k$-dimensional psychological space, and categorization is based on the amount of similarity between the item to be categorized and the stored exemplars.
+Similarity between two examples, $s\left(x_i, x_j\right)$, is computed as an inverse exponential function of distance \citep[following][]{shepard1957stimulus,shepard1987toward}:
 
-As in the GCM, the similarity between two examples, $s\left(x_i, x_j\right)$, is an inverse-exponential function of distance:
-% 
 \begin{equation}
-  s\left(x_i,x_j\right) = \exp \left\{ -c \sum_{k}{ \left| x_{ik} - x_{jk} \right|}w_k \right\}
+  s\left(x_i,x_j\right) = \exp \left\{ -c \left[\sum_{k}{ w_k \left| x_{ik} - x_{jk} \right|^r }\right]^{1/r} \right\}
   \label{eq:similarity}
@@
 where $w_k$ is the attention weighting of dimension $k$ ($w_k \geq 0$ and $\sum_k{w_k} = 1$), accounting for the relative importance of each dimension in similarity calculations, and $c$ ($c>0$) is a specificity parameter controlling the spread of exemplar generalization.
-For simplicity, our simulations will use uniform attention weights, except in our discussion of individual differences.
+For simplicity, attention will be distributed uniformly in our simulations (unless otherwise noted).
+The value of $r$ depends on the nature of the experimental conditions being simulated: $r=1$ is appropriate for separable dimensions, whereas $r=2$ is appropriate for integral dimensions \citep[see][]{shepard1964attention,garner1974processing}.
 
-To generate a new example, the model considers both the similarity to examples from contrast categories as well as the similarity to examples (if any exist) in the target category.
+PACKER's core proposal is that generation is constrained by both similarity to members of the target category (the category in which a stimulus is being generated) as well as similarity to members of other categories: the most desirable generation candidates are similar to members of the target category and not similar to members of contrast categories.
+This is achieved by aggregating similarity across known exemplars differently according to class membership.
 The aggregated similarity $a$ between generation candidate $y$ and stored exemplars $x$ is given by:
@@
 % 
-where $f(x_j)$ is a function specifying the extent to the exemplar contributes to generation.
+where $f(x_j)$ is a function specifying each exemplar's contribution to generation.
 PACKER sets $f(x_j)$ depending on exemplar $x_j$'s category membership: $f(x_j) = \phi$ if $x_j$ is a member of a contrast category, and $f(x_j) = \gamma$ if $x_j$ is a member of the target category.
@@
 
-PACKER's core proposal is that new categories should be different from existing categories, and same-category exemplars should be similar to one another.
+\begin{figure}
+    \begin{center}
+    \inputpgf{figs/}{packer-examples.pgf}
+    \caption{PACKER generation of a category `B' example, following exposure to one member of category `A' and one member of category `B'.
+Predictions are shown for three different parameterizations: \textit{(a)} Predictions based on contrast similarity only.
+\textit{(b)} Predictions based on target similarity only.
+\textit{(c)} Predictions with both constraints considered.}
+    \label{fig:packer-examples}
+    \end{center}
+\end{figure}
+
+As noted above, PACKER proposes that new categories should be different from existing categories, and same-category exemplars should be similar to one another.
 This is realized when $\phi \leq 0$, and $\gamma \geq 0$.
-Negative $\phi$ values encourage $y$ to be distant from the contrast category (as similarity to contrast category exemplars are subtracted during aggregation).
+Negative $\phi$ values encourage $y$ to be distant from contrast categories (as similarity to contrast category exemplars are subtracted during aggregation).
 Positive $\gamma$ values encourage $y$ to be close to other exemplars of the target category.
 When $|\phi| = \gamma$, the repulsion effect from contrast categories is equal to the attraction effect to the target category.
+See Figure \ref{fig:packer-examples} for an illustration of how these parameters are used.
 
@@
 p(y) = \dfrac
-    { \exp \{ \theta \cdot a(y, x) \} } 
-    { \sum_i{ \exp \{ \theta \cdot a(y_i, x) \}  } }
+    { \exp  \{ \theta \cdot a \left( y, x \right) \} } 
+    { \sum_i{ \exp  \{ \theta \cdot a \left( y_i, x \right) \}  } }
     \label{eq:packer-choice}
@@
 % 
-where $\theta$ ($\theta \geq 0$) controls response determinism.
+where $\theta$ ($\theta \geq 0$) is a free parameter controlling response determinism.
+
+\subsection{Relation Between PACKER and Copy-And-Tweak}
+
+The PACKER model is in many senses similar to the copy-and-tweak model reported by \cite{jern2013probabilistic}: both models are exemplar-based, and both models generate new items according to their similarity to known members of the target class.
+PACKER diverges from from the copy-and-tweak model only in the inclusion of a contrast mechanism, enabling generation according to dissimilarity to members of opposing categories.
+By consequence, copy-and-tweak can be realized as a parameterization of the PACKER model that is insensitive to category contrast.
+Specifically, when $\phi = 0$ and $\gamma = 1$ (see Figure \ref{fig:packer-examples}, panel B), PACKER is not influenced by similarity to members of opposite categories, and is mathematically equivalent to a copy-and-tweak approach.
+
+In the sections below, we report simulations using this copy-and-tweak model.
+Due to their mathematical equivalence, the comparison between PACKER and copy-and-tweak provides a test of the explanatory value of the contrast mechanism: the account provided by copy-and-tweak will only equal that of PACKER if the contrast mechanism does not offer an advantage.
 

### 2017-03-17 Manuscripts/cog-psych/cogpsych.tex
added jk13 math section

@@
 
-PACKER's core proposal is that generation is constrained by both similarity to members of the target category (the category in which a stimulus is being generated) as well as similarity to members of other categories: the most desirable generation candidates are similar to members of the target category and not similar to members of contrast categories.
+PACKER's core proposal (as well as its name) was in part inspired by earlier work from category learning literature \citep[see][]{hidaka2011packing}.
+PACKER proposes that generation is constrained by both similarity to members of the target category (the category in which a stimulus is being generated) as well as similarity to members of other categories: the most desirable generation candidates are similar to members of the target category and not similar to members of contrast categories.
 This is achieved by aggregating similarity across known exemplars differently according to class membership.
 The aggregated similarity $a$ between generation candidate $y$ and stored exemplars $x$ is given by:
-% 
+
 \begin{equation}
@@
 p(y) = \dfrac
-    { \exp  \{ \theta \cdot a \left( y, x \right) \} } 
-    { \sum_i{ \exp  \{ \theta \cdot a \left( y_i, x \right) \}  } }
+    { \exp  \left \{ \theta \cdot a \left( y, x \right) \right \} } 
+    { \sum_i{ \exp  \left \{ \theta \cdot a \left( y_i, x \right) \right\}  } }
     \label{eq:packer-choice}
@@
 where $\theta$ ($\theta \geq 0$) is a free parameter controlling response determinism.
+
+\nbcnote{It is worth noting that PACKER is not the only \textit{possible} model that could incorporate a contrast mechanism.
+However, contrast falls out naturally from the exemplar view and ...
+Contrast could, for example, be implemented within...
+}
 
@@
 
-\section{Snippets}
+\subsection{The Hierarchical Sampling Model}
 
-\nbcnote{Just some random snippets that i don't want to delete yet...}
+\cite{jern2013probabilistic} demonstrated how a hierarchical Bayesian model could explain the distributional correspondences between observed and generated categories.
+In their model, exemplars of generated category were viewed as samples from a multivariate Normal distribution over the dimensions of stimulus space.
+The mean of the generated category was independent of the observed categories, but the covariance matrix (encoding feature variances and correlations) was based on a common prior distribution.
+Generating a new category was thus completed by sampling a new category mean (uniform over stimulus space) and covariance matrix from the common prior distribution.
+Because the shared prior distribution's parameters were unobserved, the hierarchical Bayesian approach was used to infer its parameters from the previous categories (their feature variances and correlations), and then to generate the covariance matrix of the new category.
 
-How do people generate concepts such that the product is (A) different from what is already known, and (B) internally coherent?
+In our implementation of their model, each category's exemplars are assumed to be sampled from a multivariate Normal distribution with parameters ($\mu, \Sigma$).
+Each category's covariance matrix is assumed to be inverse-Wishart distributed with parameters ($v$, $\kappa,$ and $\Sigma_D$).\footnote{Note that \citet{jern2013probabilistic}'s model is slightly different, as they used a non-conjugate model.
+Their model acts very similar to our version of it and receives comparable fits.} $\Sigma_D$ is the covariance matrix shared between categories.
+We assume the shared covariance matrix $\Sigma_D$ is generated from a Wishart distribution (for conjugacy) with parameters $v_0$, $\kappa_0$, and $\Sigma_0$.
+We set $\nu_0 = 4$, and $\Sigma_0 = \rho {\bf I}$, where $\rho$ is a free parameter controlling the expected variance of dimensions (dimensions of the shared covariance matrix are expected to be uncorrelated) and ${\bf I}$ is the identity matrix.
 
-We also introduce a novel exemplar-based model of category generation, PACKER (\textit{Producing Alike and Contrasting Knowledge using Exemplar Representations}), which explains much of what we observe in the creative use of conceptual knowledge.
-In the sections below, we introduce the PACKER model and its core theoretical claims.
+\nbcnote{check on that $\nu_0 = 4$ business}
 
-One of the core principles of PACKER (as well as its name) are inspired by earlier work in category learning \citep[see][]{hidaka2011packing}.
+To simplify the model predictions, we used {\em maximum a posteriori} (MAP) estimates for the hidden parameters and then generated new categories based on those estimates.
+Due to conjugacy, the MAP estimate for the shared covariance matrix $\Sigma_D = \Sigma_0 + \sum_c{C_c}$, where $C_c$ is the empirical covariance matrix of category $c$.
+The MAP estimate of the covariance matrix for the target category $B$ is
+\begin{equation}
+  \Sigma_B = \left[ \Sigma_D \nu + C_B +
+    \dfrac
+    {\kappa n_B}
+    {\kappa + n_B}
+    (\bar{x_B}-\mu_B)(\bar{x_B}-\mu_B)^T
+  \right] (\nu + n_B)^{-1}
+  \label{eq:Sigma_B}
+\end{equation}
+%
+where $\nu$ ($\nu>k-1$) is an additional free parameter (from the Inverse-Wishart prior on $\Sigma_B$) weighting the importance of $\Sigma_{D}$.
+When the target category has no members (i.e., $n_B = 0$), items are generated at random.
+
+Generated exemplars are drawn from a multivariate Normal distribution specified by $(\mu_{B}, \Sigma_{B})$. Thus, $p(y)$ is
+\begin{equation}
+  p(y) = \dfrac
+    {\exp \left \{ \theta \cdot {\rm Normal}(y; \mu_{B}, \Sigma_{B}) \right \} }
+    {\sum_i \exp \left \{ \theta \cdot {\rm Normal}(y_i; \mu_{B}, \Sigma_{B}) \right \} } 
+\end{equation}
+where $\theta$ is a response determinism parameter and ${\rm Normal}(y; \mu, \Sigma)$ denotes a multivariate Normal density evaluated at $y$.
 

### 2017-03-20 Manuscripts/cog-psych/cogpsych.tex
update prior to joes lookover

@@
 
+\begin{figure}
+    \begin{center}
+    \input{figs/e2-conditions.pgf}
+    \caption{Category types tested in Experiment 2.}
+    \label{fig:e2-conditions}
+    \end{center}
+\end{figure}
+
+We developed two Alpha categories (see Figure \ref{fig:e2-conditions}): the `Bottom' category is a tight cluster in the bottom-center of the space, and the `Middle' category is identical except that it lies in the center of stimulus space.
+Although this manipulation is minimal (the conditions differ only slightly in the Y-axis position of the experimenter-defined category), the PACKER model predicts strong between-condition differences.
+According to PACKER, the nature of the space not occupied by the existing categories should determine where members of the Beta category are likely to be generated.
+Thus, the lower areas of the stimulus space should be less frequently used for generation in the Bottom condition compared to the Middle (as these areas possess greater similarity to the Bottom category).
+Conversely, the upper areas of the stimulus space should be used for generation more frequently in the Bottom condition compared to Middle.
+
 \subsection{Participants \& Materials}
@@
 
-In the sections below, we report simulations using this copy-and-tweak model.
+In this paper, we report simulations using this copy-and-tweak model.
+The model we test is identical to PACKER, under the constraints that $\phi = 0$ and $\gamma = 1$, and is a continuous-dimension adaptation of the model tested by \cite{jern2013probabilistic}.
 Due to their mathematical equivalence, the comparison between PACKER and copy-and-tweak provides a test of the explanatory value of the contrast mechanism: the account provided by copy-and-tweak will only equal that of PACKER if the contrast mechanism does not offer an advantage.
@@
     {\kappa + n_B}
-    (\bar{x_B}-\mu_B)(\bar{x_B}-\mu_B)^T
+    (\bar{x}_B-\mu_B)(\bar{x}_B-\mu_B)^T
   \right] (\nu + n_B)^{-1}
@@
 When the target category has no members (i.e., $n_B = 0$), items are generated at random.
+
+\nbcnote{should explain Equation \ref{eq:Sigma_B} more fully now that we are not limited for space.}
 

### 2017-03-20 Manuscripts/cog-psych/cogpsych.tex
expanded/some reformulation of intro and some of background. restructuring

@@
 
+\jlanote{This is awesome!!!}
+
 \begin{verbatim}
@@
 
-The creation of novel concepts and ideas is a highly intriguing -- yet infrequently studied -- topic of research in human cognition.
+Creativity, innovation, imagination, and the creation of new ideas are among the most fascinating, important and difficult human capabilities to study.
+They are fascinating and important capabilities because they have produced some of the greatest scientific breakthroughs that revolutionized the world.
+Unfortunately, scientific breakthroughs are rare and the result of a great deal of cognitive effort, which it difficult to investigate the underlying cognitive representations and processes responsible for them using standard experimental methodology or formalize in computational models.
+Although we tend to focus on the most salient products of these processes (e.g., scientific breakthroughs), every person is likely to have generated a novel sentence, thought, and/or drawing.\footnote{\jlanote{We need citations/examples for this.
+In a compositional system, it's pretty clear this has to be the case, but I'd like to use someone else as a reference rather than arguing this myself.}} This observation provides a fantastic opportunity: By examining how people create new concepts in a domain amenable to formalization in computational models, we can investigate (some aspects of) creativity, innovation, and imagination scientifically.
+
+%The creation of novel concepts and ideas is a highly intriguing -- yet infrequently studied -- topic of research in human cognition.
 The creative use of conceptual knowledge is, for example, a core element of scientific investigation, wherein the generation of new ideas is critical to designing experiments and explaining observations.
@@
 
+Towards the aim of providing a formal account of the cognitive mechanisms involved in creativity and innovation, we explore a simplified form of creative cognition that can be studied using standard behavioral methodology and is amenable to formal modeling: category generation.
+In a category generation task, participants are given some background knowledge about a domain (a cover story and/or learning one or more categories in the domain) and then are asked to generate another category in the domain \citep{jern2013probabilistic,ward1994structured}.
+Unlike creativity and innovation, categorization is well studied from a breadth of perspectives, including behavioral, neural, comparative, and computational perspectives \citep{kurtz2015human,mack2013,margolis2015,pothoswills2011}.
+By analyzing a creative task using formal tools from the categorization literature,
+
+Previous work has focused on one key aspect of category generation: People tend to create new categories that have similar {\em statistical regularities} to previously known categories \citep{jern2013probabilistic,ward1994structured}.
+Although this is an important characteristic of generating new categories, it cannot be the only one.
+Taken to the extreme, the best "new" category in terms of having the same statistical regularities to other categories would be one that is identical to a prototypical previously observed category (and thus, not new at all).
+
+In this article, we propose a new constraint that guides category generation: "being different" or contrasting from other categories in the relevant domain.
+To some extent, new categories being different in some sense from previously observed ones has been implicitly assumed in much previous work.
+But "being different" is vague as there are many ways to be different from a previously observed category.
+Building on the largely successful exemplar modeling framework \citep{medin1978context,nosofsky1984choice}, we propose a novel exemplar model of category generation, {\em Producing Alike and Contrasting Knowledge using Exemplar Representations} (PACKER), that formalizes the constraint of "being different" from previous categories while similar to other exemplars generated so far in the new category.
+
+The outline of the article is as follows.
+First we describe the results of previous behavioral experiments investigating category generation.
+Next we review previous computational models of category generation.
+Then we propose our computational model, which produces new categories that are different from previous categories in the domain.
+Next we present two experiments examining how people generate new categories, one of which demonstrates that participants produce different categories when given one of two categories that do not differ in distributional information.
+We qualitatively and quantitatively analyze the performance of each model in capturing human category generation.
+We conclude with a discussion of the implications of our results for categorization and creative cognition, and directions for future work.
+
+\section{Prior work}
+
 Much of what we know about concept generation comes from the foundational literature on creative cognition.
-In a classic series of reports, Ward \& colleagues \citep{ward1995s,ward1994structured,marsh1999inadvertent,ward2002role,smith1993constraining} established that category generation is highly constrained by prior knowledge: generated categories tend to consist of features observed in known categories, and they tend to exhibit the distributional properties as found in known categories.
+In a classic series of reports, Ward \& colleagues \citep{ward1995s,ward1994structured,marsh1999inadvertent,ward2002role,smith1993constraining} established that category generation is highly constrained by prior knowledge: Generated categories tend to consist of features observed in known categories, and they tend to exhibit the distributional properties as found in known categories.
 In a classic study, \cite{ward1994structured} asked participants to generate new species of alien animals by drawing and describing members of the species.
@@
 
+\jlanote{Yeah, Here's how I would this section should be organized.
+Top is general behavioral findings/phenomena.
+Then subsections for each theory, which first is described in broad theoretical terms and then how that is formalized into a model}
+
 \cite{jern2013probabilistic} additionally tested a "copy-and-tweak" model that broadly resembles the earlier "Path of Least Resistance" account: the core proposal is that participants generate new items by copying stored examples from memory and tweaking them to generate something new.
@@
 This model provided a poor account of the observed data: indeed, the experiments devised by \citeauthor{jern2013probabilistic} were specifically designed to challenge this model.
-However, it's application is notable as a first step toward explaining category generation using well-known formal approaches from the categorization literature.
+However, its application is notable as a first step toward explaining category generation using well-known formal approaches from the categorization literature.
 
@@
 
-It is worth noting that the literature reviewed above provides a somewhat limited view of category generation: the main concern of the published experiments has been on the distributional correspondences between learned and generated categories, and as a result most of the modeling efforts have been towards explaining those effects.
+It is worth noting that the literature reviewed above provides a somewhat limited view of category generation: The main concern of the published experiments has been on the distributional correspondences between learned and generated categories, and as a result most of the modeling efforts have been towards explaining those effects.
 In this paper, we investigate another important constraint on the creative use of conceptual knowledge: category contrast.
@@
 The authors do not make predictions about the \textit{location} of the category within the domain (the perceptual instantiation of category members).
-Indeed, under a strict interpretation of the hierarchical Bayesian approach, the generated category should actually overlap substantially with the known categories (as it's location is drawn based on those categories).
-Alternatively, the model's placement of the new category may simply be random (this was the case in the simulations conducted by \citeauthor{jern2013probabilistic}).
+Indeed, under a strict interpretation of the hierarchical Bayesian approach presented by \citeauthor{jern2013probabilistic}, the most probable new category to be generated is located in {\em exactly} the same location as the given category.
+However, this is not simply an issue with their model -- Hierarchical Bayesian models assume the same underlying distribution generates all of the categories and thus, any differences between categories is due to {\em noise} and should not be {\em systematic}.
+The best a a hierarchical Bayesian model can do at capturing contrast is to assume that the new category is placed uniformly at random over stimulus space (this was the case in the simulations conducted by \citeauthor{jern2013probabilistic}).
 
 \nbcnote{Maybe tune that language to be more in line with the probabilistic world?}
+
+\jlanote{Don't worry about that for now. I'll do a pass on that sort of language later. Lets get the main stuff in here in the right places first.}
 
@@
 \nbcnote{Maybe mention something about path of least resistance prediction that contrast is on non-characteristic dimensions? }
+\jlanote{If we can make a strong case via verbal argument, than yeah. I think we decided not to model it, in which case we shouldn't go into detail on it.}
 
-\subsection{Overview}
-
-As noted in the sections above, research on the creative generation of novel concepts has thusfar focused on the finding that generated categories tend to possess distributional commonalities with known categories.
-However, a fundamental goal of concept generation is to create something \textit{new} (i.e., different from what is already known).
-The manner in which generated categories differ from known ones is, nonetheless, poorly understood: existing formal approaches do not make strong predictions about how creatively generated concepts should systematically differ from existing ones.
-
-In the sections below, we present two experiments demonstrating systematic effects of category contrast on creative generation.
-Our experiments conform strictly to the artificial category learning paradigm from \cite{jern2013probabilistic}: participants are first exposed to a single, experimenter-defined category, and are then asked to generate members of a new category.
-We also introduce a novel, exemplar-based model to explain these effects: PACKER (\textit{Producing Alike and Contrasting Knowledge using Exemplar Representations}).
-Finally, we report formal simulations comparing PACKER to the hierarchical sampling and copy-and-tweak models developed by \cite{jern2013probabilistic}.
-
-\section{Experiment 1}
-
-To begin our investigation, we developed a artificial, two dimensional domain of squares, varying in color and size (see Figure \ref{fig:sample-stimuli} for samples).
-We created three unique category types within this domain, shown in Figure \ref{fig:e1-conditions}.
-The category types were designed to possess strong differences in the distributional structure of the experimenter-defined (`A') category, allowing us to observe the effect of category contrast under a variety of types of prior knowledge.
-These conditions also provide an opportunity to replicate classic effects observed in previous reports \citep[e.g.,][]{jern2013probabilistic,ward1994structured,ward2002role,marsh1999inadvertent,smith1993constraining}.
-
-\nbcnote{This feels a little weak in motivation: how do these conditions specifically allow us to show is these effects?
-Did we come up with these out of thin air?
-So maybe foreshadow that you hardly need to look for these effects, they're \textit{everywhere}.
-Maybe there is also a problem that at this point it is not exactly clear what contrast means within this scenario.}
-
-\begin{figure}
-    \begin{center}
-    \inputpgf{figs/}{e1-conditions.pgf}
-    \caption{Category types tested in Experiment 1.}
-    \label{fig:e1-conditions}
-    \end{center}
-\end{figure}
-
-In the `Cluster' type, the experimenter-defined category is a tight cluster of examples in the space (e.g., all members of category `A' are small and dark in color).
-In the `Row' type, the category has a row pattern across the space: varying along one feature but not the other (e.g., members of category `A' are small and vary in color).
-In the `XOR' type, the experiment-defined category consists of two clusters separated in opposite corners of the space category, conforming to the exclusive-or logical structure (e.g., members of category `A' can be small and dark or large and light).
-
-\subsection{Participants \& Materials}
-
-183 participants were recruited from Amazon Mechanical Turk.
-Participants were randomly assigned to one condition: 64 participants were assigned to the Cluster condition, 61 were assigned to the Row condition, and 58 were assigned to the XOR condition.
-Stimuli were squares varying in color (RGB 25--230) and side length (3.0--5.8cm).
-These stimuli are slight variants of those used by \cite{conaway2016similar}, see Figure \ref{fig:sample-stimuli} for samples.
-The assignment of perceptual features (color, size) to axes of the domain space (x, y) was counterbalanced across participants.
-
-\begin{figure}
-    \begin{center}
-    \inputpgf{figs/}{stimuli-samples.pgf}
-    \caption{Sample stimuli used in Experiments 1 and 2.}
-    \label{fig:sample-stimuli}
-    \end{center}
-\end{figure}
-
-\subsection{Procedure}
-
-Participants began the experiment with a short training phase (3 blocks of 4 trials), where they observed exemplars belonging to the `Alpha' category.
-Participants were instructed to learn as much as they can about the Alpha category, and that they would answer a series of test questions afterwards.
-On each trial, a single Alpha category exemplar was presented, and participants were given as much time as they desired before moving on.
-Each block consisted of a single presentation of each of the members of the Alpha category, in a random order.
-Participants were shown the range of possible colors and sizes prior to training.
-
-Following the training phase, participants were asked to generate four examples belonging to another category called `Beta'.
-As in \citet{jern2013probabilistic}, generation was completed using a sliding-scale interface.
-Two scales controlled the features (color, size) of the generated example.
-An on-screen preview of the example updated whenever one of the features was changed.
-Participants could generate any example along an evenly-spaced 9x9 grid, except for any previously generated Beta exemplars.
-Neither the members of the Alpha category nor the previously generated Beta examples were visible during generation.
-Prior to beginning the generation phase, participants read the following instructions:
-
-\begin{displayquote}
-As it turns out, there is another category of geometric figures called ``Beta''.
-Instead of showing you examples of the Beta category, we would like to know what you think is likely to be in the Beta category.
-
-You will now be given the chance to create examples of any size or color in order to show what you expect about the Beta category.
-You will be asked to produce 4 Beta examples - they can be quite similar or quite different to each other, depending on what you think makes the most sense for the category.
-
-Each example needs to be unique, but the computer will let you know if you accidentally create a repeat.
-\end{displayquote}
-
-Following generation, participants completed a generalization phase wherein they classified novel examples into the Alpha and Beta categories without feedback.
-On each trial, a single example was presented, and participants were asked to classify it by clicking buttons labeled ``Alpha'' or ``Beta''.
-Participants classified a total of 81 items sampled along a 9x9 grid, including the members of the Alpha and Beta categories (randomly intermixed).
-These data were, however, collected to address a separate set of questions, and we do not discuss them in this paper.
-
-\clearpage
-
-\section{Experiment 2}
-
-\begin{figure}
-    \begin{center}
-    \input{figs/e2-conditions.pgf}
-    \caption{Category types tested in Experiment 2.}
-    \label{fig:e2-conditions}
-    \end{center}
-\end{figure}
-
-We developed two Alpha categories (see Figure \ref{fig:e2-conditions}): the `Bottom' category is a tight cluster in the bottom-center of the space, and the `Middle' category is identical except that it lies in the center of stimulus space.
-Although this manipulation is minimal (the conditions differ only slightly in the Y-axis position of the experimenter-defined category), the PACKER model predicts strong between-condition differences.
-According to PACKER, the nature of the space not occupied by the existing categories should determine where members of the Beta category are likely to be generated.
-Thus, the lower areas of the stimulus space should be less frequently used for generation in the Bottom condition compared to the Middle (as these areas possess greater similarity to the Bottom category).
-Conversely, the upper areas of the stimulus space should be used for generation more frequently in the Bottom condition compared to Middle.
-
-\subsection{Participants \& Materials}
-
-122 participants were recruited from Amazon Mechanical Turk.
-61 participants were randomly assigned to the Middle and Bottom conditions each.
-The stimuli were exactly as in Experiment 1.
-Again, the assignment of perceptual features (color, size) to axes of the domain space (x, y) was counterbalanced across participants.
-
-\subsection{Procedure}
-
-The procedure was exactly as in Experiment 1: participants first completed a short training phase, followed by the generation phase, followed by the generalization phase (data from this phase is not discussed in this report).
-
-\clearpage
-
-\section{The PACKER Model}
-
+\section{Formalizing category generation}
+\jlanote{this should come before the studies as formalizations of the different theories (some details may be delegated to Appendices).
+I like trying to integrate them with the verbal description of the theories rather than having a separate computational modeling section.
+But the latter works for sure and I'm not sure if the former will flow well...}
+\subsection{PACKER}
 The PACKER model is an extension of the influential Generalized Context Model of category learning \citep[GCM;][]{nosofsky1984choice,nosofsky1986attention}.
@@
 
+\jlanote{K, though if I remember, it's just conjugacy, in which case we can just cite a paper that has IW-W conjugacy.}
+
 Generated exemplars are drawn from a multivariate Normal distribution specified by $(\mu_{B}, \Sigma_{B})$. Thus, $p(y)$ is
@@
 where $\theta$ is a response determinism parameter and ${\rm Normal}(y; \mu, \Sigma)$ denotes a multivariate Normal density evaluated at $y$.
+
+\subsection{Synopsis and Prognosis}
+
+Research on the creative generation of novel concepts has focused on the finding that generated categories tend to possess distributional commonalities with known categories.
+However, a fundamental goal of concept generation is to create something \textit{new} (i.e., different from what is already known).
+The manner in which generated categories differ from known ones is, nonetheless, poorly understood: existing formal approaches do not make strong predictions about how creatively generated concepts should systematically differ from existing ones.
+
+In the sections below, we present two experiments demonstrating systematic effects of category contrast on creative generation.
+Our experiments conform strictly to the artificial category learning paradigm from \cite{jern2013probabilistic}: participants are first exposed to a single, experimenter-defined category, and are then asked to generate members of a new category.
+We also introduce a novel, exemplar-based model to explain these effects: PACKER (\textit{Producing Alike and Contrasting Knowledge using Exemplar Representations}).
+Finally, we report formal simulations comparing PACKER to the hierarchical sampling and copy-and-tweak models developed by \cite{jern2013probabilistic}.
+
+\section{Experiment 1}
+
+To begin our investigation, we developed a artificial, two dimensional domain of squares, varying in color and size (see Figure \ref{fig:sample-stimuli} for samples).
+We created three unique category types within this domain, shown in Figure \ref{fig:e1-conditions}.
+The category types were designed to possess strong differences in the distributional structure of the experimenter-defined (`A') category, allowing us to observe the effect of category contrast under a variety of types of prior knowledge.
+These conditions also provide an opportunity to replicate classic effects observed in previous reports \citep[e.g.,][]{jern2013probabilistic,ward1994structured,ward2002role,marsh1999inadvertent,smith1993constraining}.
+
+\nbcnote{This feels a little weak in motivation: how do these conditions specifically allow us to show is these effects?
+Did we come up with these out of thin air?
+So maybe foreshadow that you hardly need to look for these effects, they're \textit{everywhere}.
+Maybe there is also a problem that at this point it is not exactly clear what contrast means within this scenario.}
+
+\jlanote{Is it possible to get the main individual differences we find in this study from different parameterizations of PACKER?
+If so, we can motivate it from there.
+Alternatively, we may want to switch the order of the two studies and motivate the experiment from the next one as a .
+I like switching the order of the experiments less because I feel like Experiment 2 nails contrast down and Experiment 1 is more showing how it plays a role in more complex category types}
+
+\begin{figure}
+    \begin{center}
+    \inputpgf{figs/}{e1-conditions.pgf}
+    \caption{Category types tested in Experiment 1.}
+    \label{fig:e1-conditions}
+    \end{center}
+\end{figure}
+
+In the `Cluster' type, the experimenter-defined category is a tight cluster of examples in the space (e.g., all members of category `A' are small and dark in color).
+In the `Row' type, the category has a row pattern across the space: varying along one feature but not the other (e.g., members of category `A' are small and vary in color).
+In the `XOR' type, the experiment-defined category consists of two clusters separated in opposite corners of the space category, conforming to the exclusive-or logical structure (e.g., members of category `A' can be small and dark or large and light).
+
+\subsection{Participants \& Materials}
+
+183 participants were recruited from Amazon Mechanical Turk.
+Participants were randomly assigned to one condition: 64 participants were assigned to the Cluster condition, 61 were assigned to the Row condition, and 58 were assigned to the XOR condition.
+Stimuli were squares varying in color (RGB 25--230) and side length (3.0--5.8cm).
+These stimuli are slight variants of those used by \cite{conaway2016similar}, see Figure \ref{fig:sample-stimuli} for samples.
+The assignment of perceptual features (color, size) to axes of the domain space (x, y) was counterbalanced across participants.
+
+\begin{figure}
+    \begin{center}
+    \inputpgf{figs/}{stimuli-samples.pgf}
+    \caption{Sample stimuli used in Experiments 1 and 2.}
+    \label{fig:sample-stimuli}
+    \end{center}
+\end{figure}
+
+\subsection{Procedure}
+
+Participants began the experiment with a short training phase (3 blocks of 4 trials), where they observed exemplars belonging to the `Alpha' category.
+Participants were instructed to learn as much as they can about the Alpha category, and that they would answer a series of test questions afterwards.
+On each trial, a single Alpha category exemplar was presented, and participants were given as much time as they desired before moving on.
+Each block consisted of a single presentation of each of the members of the Alpha category, in a random order.
+Participants were shown the range of possible colors and sizes prior to training.
+
+Following the training phase, participants were asked to generate four examples belonging to another category called `Beta'.
+As in \citet{jern2013probabilistic}, generation was completed using a sliding-scale interface.
+Two scales controlled the features (color, size) of the generated example.
+An on-screen preview of the example updated whenever one of the features was changed.
+Participants could generate any example along an evenly-spaced 9x9 grid, except for any previously generated Beta exemplars.
+Neither the members of the Alpha category nor the previously generated Beta examples were visible during generation.
+Prior to beginning the generation phase, participants read the following instructions:
+
+\begin{displayquote}
+As it turns out, there is another category of geometric figures called ``Beta''.
+Instead of showing you examples of the Beta category, we would like to know what you think is likely to be in the Beta category.
+
+You will now be given the chance to create examples of any size or color in order to show what you expect about the Beta category.
+You will be asked to produce 4 Beta examples - they can be quite similar or quite different to each other, depending on what you think makes the most sense for the category.
+
+Each example needs to be unique, but the computer will let you know if you accidentally create a repeat.
+\end{displayquote}
+
+Following generation, participants completed a generalization phase wherein they classified novel examples into the Alpha and Beta categories without feedback.
+On each trial, a single example was presented, and participants were asked to classify it by clicking buttons labeled ``Alpha'' or ``Beta''.
+Participants classified a total of 81 items sampled along a 9x9 grid, including the members of the Alpha and Beta categories (randomly intermixed).
+These data were, however, collected to address a separate set of questions, and we do not discuss them in this paper.
+
+\clearpage
+
+\section{Experiment 2}
+
+\begin{figure}
+    \begin{center}
+    \input{figs/e2-conditions.pgf}
+    \caption{Category types tested in Experiment 2.}
+    \label{fig:e2-conditions}
+    \end{center}
+\end{figure}
+
+We developed two Alpha categories (see Figure \ref{fig:e2-conditions}): the `Bottom' category is a tight cluster in the bottom-center of the space, and the `Middle' category is identical except that it lies in the center of stimulus space.
+Although this manipulation is minimal (the conditions differ only slightly in the Y-axis position of the experimenter-defined category), the PACKER model predicts strong between-condition differences.
+According to PACKER, the nature of the space not occupied by the existing categories should determine where members of the Beta category are likely to be generated.
+Thus, the lower areas of the stimulus space should be less frequently used for generation in the Bottom condition compared to the Middle (as these areas possess greater similarity to the Bottom category).
+Conversely, the upper areas of the stimulus space should be used for generation more frequently in the Bottom condition compared to Middle.
+
+\subsection{Participants \& Materials}
+
+122 participants were recruited from Amazon Mechanical Turk.
+61 participants were randomly assigned to the Middle and Bottom conditions each.
+The stimuli were exactly as in Experiment 1.
+Again, the assignment of perceptual features (color, size) to axes of the domain space (x, y) was counterbalanced across participants.
+
+\subsection{Procedure}
+
+The procedure was exactly as in Experiment 1: participants first completed a short training phase, followed by the generation phase, followed by the generalization phase (data from this phase is not discussed in this report).
+
+\clearpage
 

### 2017-03-21 Manuscripts/cog-psych/cogpsych.tex
edits to manuscript

@@
 Although this is an important characteristic of generating new categories, it cannot be the only one.
-Taken to the extreme, the best "new" category in terms of having the same statistical regularities to other categories would be one that is identical to a prototypical previously observed category (and thus, not new at all).
+Taken to the extreme, the best ``new'' category in terms of having the same statistical regularities to other categories would be one that is identical to a prototypical previously observed category (and thus, not new at all).
 
-In this article, we propose a new constraint that guides category generation: "being different" or contrasting from other categories in the relevant domain.
-To some extent, new categories being different in some sense from previously observed ones has been implicitly assumed in much previous work.
-But "being different" is vague as there are many ways to be different from a previously observed category.
-Building on the largely successful exemplar modeling framework \citep{medin1978context,nosofsky1984choice}, we propose a novel exemplar model of category generation, {\em Producing Alike and Contrasting Knowledge using Exemplar Representations} (PACKER), that formalizes the constraint of "being different" from previous categories while similar to other exemplars generated so far in the new category.
+In this article, we propose a new constraint that guides category generation: ``being different'' or contrasting from other categories in the relevant domain.
+To some extent, this constraint has been implicitly assumed in much previous work: the idea that a new category should be ``different'' is vague, as there are many ways it could be different from a previously observed category.
+Building on the largely successful exemplar modeling framework \citep{medin1978context,nosofsky1984choice,nosofsky1986attention}, we propose a novel exemplar model of category generation, {\em Producing Alike and Contrasting Knowledge using Exemplar Representations} (PACKER), formalizing how new categories should differ from previous categories.
 
 The outline of the article is as follows.
-First we describe the results of previous behavioral experiments investigating category generation.
-Next we review previous computational models of category generation.
-Then we propose our computational model, which produces new categories that are different from previous categories in the domain.
-Next we present two experiments examining how people generate new categories, one of which demonstrates that participants produce different categories when given one of two categories that do not differ in distributional information.
-We qualitatively and quantitatively analyze the performance of each model in capturing human category generation.
+First we describe previous empirical work on the topic of category generation, as well as the computational approaches studied in those reports.
+Then we describe our novel computational model, which is designed to generate categories that systematically differ from existing categories in the domain.
+We present two experiments demonstrating strong and systematic effects of category contrast on creative generation, and we qualitatively and quantitatively analyze the performance of each model in capturing human category generation.
 We conclude with a discussion of the implications of our results for categorization and creative cognition, and directions for future work.
@@
 They indicate that people rely strongly on prior knowledge in the creative process, and people generate concepts in accord with what they already know.
-
+,
 Theoretical accounts of these effects have primarily been grounded within the categorization literature.
@@
 People were thought to change only features that are not characteristic of the retrieved category (e.g., if \textit{bird} was retrieved, the presence of \textit{wings} would not change, but \textit{color} might).
-This theory incorporates elements of the highly influential basic-level categories framework \citep{rosch1975cognitive,rosch1976basic}, as well as the exemplar view \citep{medin1978context,nosofsky1984choice,nosofsky1986attention}.
+This theory incorporates elements of the highly influential basic-level categories framework \citep{rosch1975cognitive,rosch1976basic}, as well as the exemplar view \citep{medin1978context,brooks1978nonanalytic}.
 Problematically, however, the experimental paradigms employed in these studies were relatively uncontrolled, precluding the development of formal approaches.
@@
 Most notably, they showed that a Bayesian hierarchical sampling model provided the strongest account.
-Their model represents categories as multivariate normal distributions within the space, and from exposure to category members the model induces a domain-wide representation of the distributional commonalities among the observed categories.
+Their model views observed examples as samples from an underlying category distribution, describing the location of the category in the space, as well as how it varies along each feature.
+In turn, each category is viewed as a sample from an underlying \textit{domain} distribution, specifying distributional commonalities among the observed categories.
+Generated categories are thought to stem from the same domain distribution as observed categories, thus the distributional properties of observed categories will be preserved within the generated category.
+
+% Here's what it used to be:
+% Their model represents categories as multivariate normal distributions within the space, and from exposure to category members the model induces a domain-wide representation of the distributional commonalities among the observed categories.
 When generating a new category, the model samples a new distribution from the domain, and thus the distributional properties of observed categories will be preserved within the generates category.
 
-\nbcnote{Maybe we need a more technical view of the model here?}
-
-\jlanote{Yeah, Here's how I would this section should be organized.
+% \jlanote{Yeah, Here's how I would this section should be organized.
 Top is general behavioral findings/phenomena.
@@
 The copy-and-tweak model differs from the path of least resistance account in that it notably omits the hierarchical organization of categories, as well as selectivity in which features are changed.
-Instead, their copy-and-tweak model corresponds to a direct exemplar-similarity approach \citep[e.g.,][]{nosofsky1984choice}: participants are thought to generate new items according to their similarity to known members of the target category.
+Instead, their copy-and-tweak model corresponds to a direct exemplar-similarity approach \citep[e.g.,][]{nosofsky1984choice,nosofsky1986attention}: participants are thought to generate new items according to their similarity to known members of the target category.
 This model provided a poor account of the observed data: indeed, the experiments devised by \citeauthor{jern2013probabilistic} were specifically designed to challenge this model.
@@
 
+% Joe says he'll do a pass on this language later:
 Although it is evident that people are \textit{capable} of creating new concepts and categories, it is not entirely clear how new concepts are systematically made different from what is already known.
@@
 
-\nbcnote{Maybe tune that language to be more in line with the probabilistic world?}
-
-\jlanote{Don't worry about that for now. I'll do a pass on that sort of language later. Lets get the main stuff in here in the right places first.}
-
 The copy-and-tweak model tested by \cite{jern2013probabilistic} also claims little about how generated categories should contrast with what is already known.
@@
 
-\nbcnote{Maybe mention something about path of least resistance prediction that contrast is on non-characteristic dimensions? }
-\jlanote{If we can make a strong case via verbal argument, than yeah. I think we decided not to model it, in which case we shouldn't go into detail on it.}
-
-\section{Formalizing category generation}
-\jlanote{this should come before the studies as formalizations of the different theories (some details may be delegated to Appendices).
+% \section{Formalizing category generation}
+% \jlanote{this should come before the studies as formalizations of the different theories (some details may be delegated to Appendices).
 I like trying to integrate them with the verbal description of the theories rather than having a separate computational modeling section.
 But the latter works for sure and I'm not sure if the former will flow well...}
-\subsection{PACKER}
-The PACKER model is an extension of the influential Generalized Context Model of category learning \citep[GCM;][]{nosofsky1984choice,nosofsky1986attention}.
-The GCM simulates categorization under the assumption that learners represent categories as a collection of exemplars, corresponding to the labeled stimuli they have observed.
-The exemplars are encoded within a $k$-dimensional psychological space, and categorization is based on the amount of similarity between the item to be categorized and the stored exemplars.
+
+\subsection{The PACKER Model}
+
+As noted above, the constraint that new concepts should differ from what is already known has been largely been overlooked in previous work.
+This is no doubt in part due to the vague definition of what it means for a concept to be ``different'': a generated category may be different from what is already known in any number of respects.
+Towards formalizing the role of contrast in creative generation, we developed the PACKER  ({\em Producing Alike and Contrasting Knowledge using Exemplar Representations}) model which explains category generation as a balance between two fundamental constraints: generated categories should not be similar to known categories, and exemplars within each category should be similar to one another.
+These ideas are implemented within the well-studied exemplar framework: the PACKER model is an extension of the influential Generalized Context Model of category learning \citep[GCM;][]{nosofsky1984choice,nosofsky1986attention}.
+
+\nbcnote{Maybe note something about how PACKER's basis in the exemplar view means that we're explaining category generation using well known ideas, and that exemplar models are viewed as highly principled.}
+
+Both PACKER and the GCM simulate categorization under the assumption that learners represent categories as a collection of exemplars, corresponding to the labeled stimuli they have observed.
+The exemplars are encoded within a $k$-dimensional psychological space, and model performance is based on the amount of similarity between the item to be categorized and the stored exemplars.
 Similarity between two examples, $s\left(x_i, x_j\right)$, is computed as an inverse exponential function of distance \citep[following][]{shepard1957stimulus,shepard1987toward}:
@@
 The value of $r$ depends on the nature of the experimental conditions being simulated: $r=1$ is appropriate for separable dimensions, whereas $r=2$ is appropriate for integral dimensions \citep[see][]{shepard1964attention,garner1974processing}.
+In our simulations, we set $r=1$ due to the separable nature of the stimulus dimensions used in our experiments (see Figure \ref{fig:sample-stimuli}).
 
@@
 
-\nbcnote{It is worth noting that PACKER is not the only \textit{possible} model that could incorporate a contrast mechanism.
-However, contrast falls out naturally from the exemplar view and ...
-Contrast could, for example, be implemented within...
-}
+It is worth noting that PACKER represents just one of many possible models that incorporate contrast in category generation.
+For example, although it may be possible to implement a contrast mechanism within \cite{jern2013probabilistic}'s hierarchical sampling model, as noted above this mechanism would be at odds with a strict interpretation of the hierarchical Bayesian framework.
+In contrast, these ideas emerge naturally from the exemplar view: we have not modified any of the core elements of the GCM in defining the PACKER model, we simply aggregated similarity slightly differently.
+\nbcnote{That line about importance sampling can go here.}
 
-\subsection{Relation Between PACKER and Copy-And-Tweak}
+\subsubsection{Relation Between PACKER and Copy-And-Tweak}
 
@@
 
-\subsection{The Hierarchical Sampling Model}
-
-\cite{jern2013probabilistic} demonstrated how a hierarchical Bayesian model could explain the distributional correspondences between observed and generated categories.
-In their model, exemplars of generated category were viewed as samples from a multivariate Normal distribution over the dimensions of stimulus space.
-The mean of the generated category was independent of the observed categories, but the covariance matrix (encoding feature variances and correlations) was based on a common prior distribution.
-Generating a new category was thus completed by sampling a new category mean (uniform over stimulus space) and covariance matrix from the common prior distribution.
-Because the shared prior distribution's parameters were unobserved, the hierarchical Bayesian approach was used to infer its parameters from the previous categories (their feature variances and correlations), and then to generate the covariance matrix of the new category.
-
-In our implementation of their model, each category's exemplars are assumed to be sampled from a multivariate Normal distribution with parameters ($\mu, \Sigma$).
-Each category's covariance matrix is assumed to be inverse-Wishart distributed with parameters ($v$, $\kappa,$ and $\Sigma_D$).\footnote{Note that \citet{jern2013probabilistic}'s model is slightly different, as they used a non-conjugate model.
-Their model acts very similar to our version of it and receives comparable fits.} $\Sigma_D$ is the covariance matrix shared between categories.
-We assume the shared covariance matrix $\Sigma_D$ is generated from a Wishart distribution (for conjugacy) with parameters $v_0$, $\kappa_0$, and $\Sigma_0$.
-We set $\nu_0 = 4$, and $\Sigma_0 = \rho {\bf I}$, where $\rho$ is a free parameter controlling the expected variance of dimensions (dimensions of the shared covariance matrix are expected to be uncorrelated) and ${\bf I}$ is the identity matrix.
-
-\nbcnote{check on that $\nu_0 = 4$ business}
-
-To simplify the model predictions, we used {\em maximum a posteriori} (MAP) estimates for the hidden parameters and then generated new categories based on those estimates.
-Due to conjugacy, the MAP estimate for the shared covariance matrix $\Sigma_D = \Sigma_0 + \sum_c{C_c}$, where $C_c$ is the empirical covariance matrix of category $c$.
-The MAP estimate of the covariance matrix for the target category $B$ is
-\begin{equation}
-\Sigma_B = \left[ \Sigma_D \nu + C_B +
-\dfrac
-{\kappa n_B}
-{\kappa + n_B}
-(\bar{x}_B-\mu_B)(\bar{x}_B-\mu_B)^T
-\right] (\nu + n_B)^{-1}
-\label{eq:Sigma_B}
-\end{equation}
-%
-where $\nu$ ($\nu>k-1$) is an additional free parameter (from the Inverse-Wishart prior on $\Sigma_B$) weighting the importance of $\Sigma_{D}$.
-When the target category has no members (i.e., $n_B = 0$), items are generated at random.
-
-\nbcnote{should explain Equation \ref{eq:Sigma_B} more fully now that we are not limited for space.}
-
-\jlanote{K, though if I remember, it's just conjugacy, in which case we can just cite a paper that has IW-W conjugacy.}
-
-Generated exemplars are drawn from a multivariate Normal distribution specified by $(\mu_{B}, \Sigma_{B})$. Thus, $p(y)$ is
-\begin{equation}
-p(y) = \dfrac
-{\exp \left \{ \theta \cdot {\rm Normal}(y; \mu_{B}, \Sigma_{B}) \right \} }
-{\sum_i \exp \left \{ \theta \cdot {\rm Normal}(y_i; \mu_{B}, \Sigma_{B}) \right \} } 
-\end{equation}
-where $\theta$ is a response determinism parameter and ${\rm Normal}(y; \mu, \Sigma)$ denotes a multivariate Normal density evaluated at $y$.
-
 \subsection{Synopsis and Prognosis}
@@
 The manner in which generated categories differ from known ones is, nonetheless, poorly understood: existing formal approaches do not make strong predictions about how creatively generated concepts should systematically differ from existing ones.
+Above, we introduced a novel, exemplar-based model formalizing the role of contrast in creative generation.
 
@@
 Our experiments conform strictly to the artificial category learning paradigm from \cite{jern2013probabilistic}: participants are first exposed to a single, experimenter-defined category, and are then asked to generate members of a new category.
-We also introduce a novel, exemplar-based model to explain these effects: PACKER (\textit{Producing Alike and Contrasting Knowledge using Exemplar Representations}).
-Finally, we report formal simulations comparing PACKER to the hierarchical sampling and copy-and-tweak models developed by \cite{jern2013probabilistic}.
+We then we report formal simulations comparing PACKER's account of our results to that of the hierarchical sampling and copy-and-tweak models developed by \cite{jern2013probabilistic}.
 
@@
 To begin our investigation, we developed a artificial, two dimensional domain of squares, varying in color and size (see Figure \ref{fig:sample-stimuli} for samples).
-We created three unique category types within this domain, shown in Figure \ref{fig:e1-conditions}.
-The category types were designed to possess strong differences in the distributional structure of the experimenter-defined (`A') category, allowing us to observe the effect of category contrast under a variety of types of prior knowledge.
-These conditions also provide an opportunity to replicate classic effects observed in previous reports \citep[e.g.,][]{jern2013probabilistic,ward1994structured,ward2002role,marsh1999inadvertent,smith1993constraining}.
+To foreshadow slightly, effects of contrast are strong and can be widely observed in category generation data.
+Accordingly, we began by testing the strength of these effects under a variety of learning conditions.
+This was achieved by training participants on categories possessing qualitatively different distributional structures: We created three unique category types within the domain, shown in Figure \ref{fig:e1-conditions}.
 
-\nbcnote{This feels a little weak in motivation: how do these conditions specifically allow us to show is these effects?
+Each panel of Figure \ref{fig:e1-conditions} shows the locations of exemplar belonging to a single, experimenter-defined category (`A', or `Alpha').
+In the `Cluster' type, category A is a tight cluster of examples in the space (e.g., all members are small and dark in color).
+In the `Row' type, the category has a row pattern across the space: varying along one feature but not the other (e.g., members are small and vary in color).
+In the `XOR' type, the experiment-defined category consists of two clusters separated in opposite corners of the space category, conforming to the exclusive-or logical structure (e.g., members can be small and dark or large and light).
+
+\nbcnote{what would contrast look like?}
+
+% \nbcnote{This feels a little weak in motivation: how do these conditions specifically allow us to show is these effects?
 Did we come up with these out of thin air?
@@
 Maybe there is also a problem that at this point it is not exactly clear what contrast means within this scenario.}
-
-\jlanote{Is it possible to get the main individual differences we find in this study from different parameterizations of PACKER?
+% \jlanote{Is it possible to get the main individual differences we find in this study from different parameterizations of PACKER?
 If so, we can motivate it from there.
@@
 \end{figure}
-
-In the `Cluster' type, the experimenter-defined category is a tight cluster of examples in the space (e.g., all members of category `A' are small and dark in color).
-In the `Row' type, the category has a row pattern across the space: varying along one feature but not the other (e.g., members of category `A' are small and vary in color).
-In the `XOR' type, the experiment-defined category consists of two clusters separated in opposite corners of the space category, conforming to the exclusive-or logical structure (e.g., members of category `A' can be small and dark or large and light).
 
@@
 
+\appendix
+
+\section{The Hierarchical Sampling Model}
+\label{ap:hsampling-definition}
+
+\cite{jern2013probabilistic} demonstrated how a hierarchical Bayesian model could explain the distributional correspondences between observed and generated categories.
+In their model, exemplars of generated category were viewed as samples from a multivariate Normal distribution over the dimensions of stimulus space.
+The mean of the generated category was independent of the observed categories, but the covariance matrix (encoding feature variances and correlations) was based on a common prior distribution.
+Generating a new category was thus completed by sampling a new category mean (uniform over stimulus space) and covariance matrix from the common prior distribution.
+Because the shared prior distribution's parameters were unobserved, the hierarchical Bayesian approach was used to infer its parameters from the previous categories (their feature variances and correlations), and then to generate the covariance matrix of the new category.
+
+In our implementation of their model, each category's exemplars are assumed to be sampled from a multivariate Normal distribution with parameters ($\mu, \Sigma$).
+Each category's covariance matrix is assumed to be inverse-Wishart distributed with parameters ($v$, $\kappa,$ and $\Sigma_D$).\footnote{Note that \citet{jern2013probabilistic}'s model is slightly different, as they used a non-conjugate model.
+Their model acts very similar to our version of it and receives comparable fits.} $\Sigma_D$ is the covariance matrix shared between categories.
+We assume the shared covariance matrix $\Sigma_D$ is generated from a Wishart distribution (for conjugacy) with parameters $v_0$, $\kappa_0$, and $\Sigma_0$.
+We set $\nu_0 = 4$, and $\Sigma_0 = \rho {\bf I}$, where $\rho$ is a free parameter controlling the expected variance of dimensions (dimensions of the shared covariance matrix are expected to be uncorrelated) and ${\bf I}$ is the identity matrix.
+
+\nbcnote{check on that $\nu_0 = 4$ business}
+
+To simplify the model predictions, we used {\em maximum a posteriori} (MAP) estimates for the hidden parameters and then generated new categories based on those estimates.
+Due to conjugacy, the MAP estimate for the shared covariance matrix $\Sigma_D = \Sigma_0 + \sum_c{C_c}$, where $C_c$ is the empirical covariance matrix of category $c$.
+The MAP estimate of the covariance matrix for the target category $B$ is
+\begin{equation}
+\Sigma_B = \left[ \Sigma_D \nu + C_B +
+\dfrac
+{\kappa n_B}
+{\kappa + n_B}
+(\bar{x}_B-\mu_B)(\bar{x}_B-\mu_B)^T
+\right] (\nu + n_B)^{-1}
+\label{eq:Sigma_B}
+\end{equation}
+%
+where $\nu$ ($\nu>k-1$) is an additional free parameter (from the Inverse-Wishart prior on $\Sigma_B$) weighting the importance of $\Sigma_{D}$.
+When the target category has no members (i.e., $n_B = 0$), items are generated at random.
+
+\nbcnote{should explain Equation \ref{eq:Sigma_B} more fully now that we are not limited for space.}
+
+\jlanote{K, though if I remember, it's just conjugacy, in which case we can just cite a paper that has IW-W conjugacy.}
+
+\nbcnote{It \textbf{is} just conjugacy, so we can just cite a paper (do you know of a paper or should I start digging?).
+We'll probably want to detail the what we did for priors and which free parameters we fitted though.}
+
+Generated exemplars are drawn from a multivariate Normal distribution specified by $(\mu_{B}, \Sigma_{B})$. Thus, $p(y)$ is
+\begin{equation}
+p(y) = \dfrac
+{\exp \left \{ \theta \cdot {\rm Normal}(y; \mu_{B}, \Sigma_{B}) \right \} }
+{\sum_i \exp \left \{ \theta \cdot {\rm Normal}(y_i; \mu_{B}, \Sigma_{B}) \right \} } 
+\end{equation}
+where $\theta$ is a response determinism parameter and ${\rm Normal}(y; \mu, \Sigma)$ denotes a multivariate Normal density evaluated at $y$.
+
 \end{document}

### 2017-03-22 Manuscripts/cog-psych/cogpsych.tex
edited to e1 results

@@
 The value of $r$ depends on the nature of the experimental conditions being simulated: $r=1$ is appropriate for separable dimensions, whereas $r=2$ is appropriate for integral dimensions \citep[see][]{shepard1964attention,garner1974processing}.
-In our simulations, we set $r=1$ due to the separable nature of the stimulus dimensions used in our experiments (see Figure \ref{fig:sample-stimuli}).
+In our simulations, we set $r=1$ due to the separable nature of the stimulus dimensions used in our experiments (see Figure \ref{fig:e1-conditions}).
 
@@
 As noted above, PACKER proposes that new categories should be different from existing categories, and same-category exemplars should be similar to one another.
-This is realized when $\phi \leq 0$, and $\gamma \geq 0$.
+This is realized when $\phi < 0$, and $\gamma > 0$.
 Negative $\phi$ values encourage $y$ to be distant from contrast categories (as similarity to contrast category exemplars are subtracted during aggregation).
@@
 
-To begin our investigation, we developed a artificial, two dimensional domain of squares, varying in color and size (see Figure \ref{fig:sample-stimuli} for samples).
+To begin our investigation, we developed an artificial, two dimensional domain of squares, varying in color and size (see Figure \ref{fig:e1-conditions}, panel A).
 To foreshadow slightly, effects of contrast are strong and can be widely observed in category generation data.
 Accordingly, we began by testing the strength of these effects under a variety of learning conditions.
-This was achieved by training participants on categories possessing qualitatively different distributional structures: We created three unique category types within the domain, shown in Figure \ref{fig:e1-conditions}.
+This was achieved by training participants on categories possessing qualitatively different distributional structures, as shown in Figure \ref{fig:e1-conditions}.
 
-Each panel of Figure \ref{fig:e1-conditions} shows the locations of exemplar belonging to a single, experimenter-defined category (`A', or `Alpha').
-In the `Cluster' type, category A is a tight cluster of examples in the space (e.g., all members are small and dark in color).
-In the `Row' type, the category has a row pattern across the space: varying along one feature but not the other (e.g., members are small and vary in color).
-In the `XOR' type, the experiment-defined category consists of two clusters separated in opposite corners of the space category, conforming to the exclusive-or logical structure (e.g., members can be small and dark or large and light).
+Panels B-D of Figure \ref{fig:e1-conditions} show the locations of exemplars belonging to the experimenter-defined categories (`A', or `Alpha') that participants were assigned to learn about prior to generating a new category.
+In the `Cluster' type, category A is a tight cluster of examples in the space.
+Perceptually instantiated, the members of category A might, for example, be large and dark in color.
+In the `Row' type, category A has a row pattern across the space, varying along one feature but not the other.
+Thus, its members might all be dark in color but would vary in size.
+Finally, in the `XOR' type, the experimenter-defined category consists of two clusters separated in opposite corners of the space category, conforming to the exclusive-or logical structure (e.g., members are small and dark or large and light).
 
-\nbcnote{what would contrast look like?}
+\begin{figure}
+    \begin{center}
+    \inputpgf{figs/}{e1-conditions.pgf}
+    \caption{Stimulus domain and category types tested in Experiment 1. }
+    \label{fig:e1-conditions}
+    \end{center}
+\end{figure}
 
-% \nbcnote{This feels a little weak in motivation: how do these conditions specifically allow us to show is these effects?
-Did we come up with these out of thin air?
-So maybe foreshadow that you hardly need to look for these effects, they're \textit{everywhere}.
-Maybe there is also a problem that at this point it is not exactly clear what contrast means within this scenario.}
+After learning about an experimenter-defined category, participants are asked to generate examples of a new category.
+Within this paradigm, an effect of category contrast would be realized if participants prefer to generate items in locations that are distant (i.e., perceptually dissimilar) from members of category A.
+However, generation is left unconstrained: For example, participants assigned to the Cluster condition may generate a tightly clustered category in the corner opposite of the experimenter-defined category.
+Alternatively, they may generate a tightly clustered category directly overlapping with the experimenter-defined category.
+Further, they may even generate an entirely different type of category (e.g., a row category).
+
+In this way, our first experiment also serves to replicate the classic finding that generated categories tend to share distributional properties with known categories in the domain \citep[see][]{jern2013probabilistic,ward1994structured}.
+From these results, we can predict that, in each condition, participants should generate categories that are distributionally similar to the experimenter-defined category: In the Cluster condition, generated categories should be tightly clustered.
+In the Row condition, generated categories should very more along the X-axis than the Y-axis.
+In XOR condition, generated categories should be widely distributed across both dimensions, and the two dimensions should be positively correlated.
+
+\nbcnote{I think that works as a decent motivation, do you think so?}
+
 % \jlanote{Is it possible to get the main individual differences we find in this study from different parameterizations of PACKER?
@@
 I like switching the order of the experiments less because I feel like Experiment 2 nails contrast down and Experiment 1 is more showing how it plays a role in more complex category types}
-
-\begin{figure}
-    \begin{center}
-    \inputpgf{figs/}{e1-conditions.pgf}
-    \caption{Category types tested in Experiment 1.}
-    \label{fig:e1-conditions}
-    \end{center}
-\end{figure}
 
@@
 Stimuli were squares varying in color (RGB 25--230) and side length (3.0--5.8cm).
-These stimuli are slight variants of those used by \cite{conaway2016similar}, see Figure \ref{fig:sample-stimuli} for samples.
+These stimuli are slight variants of those used by \cite{conaway2016similar}, see Figure \ref{fig:e1-conditions}.
 The assignment of perceptual features (color, size) to axes of the domain space (x, y) was counterbalanced across participants.
-
-\begin{figure}
-    \begin{center}
-    \inputpgf{figs/}{stimuli-samples.pgf}
-    \caption{Sample stimuli used in Experiments 1 and 2.}
-    \label{fig:sample-stimuli}
-    \end{center}
-\end{figure}
 
@@
 
+\subsection{Results}
+
+\subsection{Discussion}
+
 \clearpage
@@
 The procedure was exactly as in Experiment 1: participants first completed a short training phase, followed by the generation phase, followed by the generalization phase (data from this phase is not discussed in this report).
+
+\subsection{Results}
+
+\subsection{Discussion}
 

### 2017-03-22 Manuscripts/cog-psych/cogpsych.tex
edits to e1 results; working on compact e1 samples figure

@@
 
+To evaluate the role of contrast in category generation, we computed the number of times each stimulus was generated, as a function of its average city-block distance from members of the experimenter-defined ``Alpha'' category.
+These data, shown in Figure \ref{fig:e1-distanceplots}, reveal a clear pattern: examples that are highly distant from members of the experimenter-defined categories are more likely to be generated into a new category.
+
+\begin{figure}
+    \begin{center}
+    \inputpgf{figs/}{e1-distanceplots.pgf}
+    \caption{Experiment 1 results.
+\textit{Left}: Frequency of generation as a function of distance from members of the experimenter-defined category.
+\textit{Right}: Scatter plot of within-class versus between-class distance in each of the participant-generated categories.}
+
+    \nbcnote{Need to check on how nice the journal is about color figs!}
+
+    \label{fig:e1-distanceplots}
+    \end{center}
+\end{figure}
+
+Figure \ref{fig:e1-distanceplots} also depicts, for each participant, the average amount of distance between members of the generated category (\textit{within-class} distance) against the average amount of distance between members of the generated and experimenter-defined category (\textit{between-class} distance).
+These data also reveal a systematic pattern: the majority of participants generated categories with more between-class distance than within-class distance.
+That is, members of the generated category tended to be more similar to one another than to members of the experimenter-defined category.
+T-Tests comparing the amount of within- and between- class distance among the categories support this claim in each condition: Cluster, $t(63) = 11.43$, $p < 0.001$; Row, $t(60) = 13.16$, $p < 0.001$; and XOR, $t(57) = 3.64$, $p < 0.001$.
+These results provide clear evidence of an effect of category contrast: participants prefer to generate categories that are dissimilar to the learned category but maintain some level of internal cohesion.
+
+Clearly, participants tend to generate categories that differ from what is already known.
+However, the secondary goal of this experiment was to replicate the classic finding that generated and previously-known categories share distributional properties.
+To evaluate this, we computed, for each participant, the range of the generated category along each axis (X, Y), as well as the correlation between features within each category.
+These data, shown in Figure \ref{fig:e1-statsboxes}, reveal a broad-yet-incomplete replication.
+
+With respect to ranges along each axis (X, Y), the generated categories from each condition tend to reflect the ranges of the experimenter-defined categories.
+The categories generated in the Cluster condition had the less X-axis range compared to Row, $t(123) = 5.61$, $p < 0.001$, and XOR $t(120) = 2.68$, $p = 0.008$.
+Likewise, categories generated in the Row condition had less Y-axis range compared to Cluster, $t(123) = 4.57$, $p < 0.001$ and XOR $t(117) = 9.26$, $p < 0.001$, and categories from the Cluster condition had less Y-axis range compared to XOR $t(120) = 3.95$, $p < 0.001$.
+
+\begin{figure}
+    \begin{center}
+    \inputpgf{figs/}{e1-statsboxes.pgf}
+    \caption{Box-plots of the distributional statistics from the categories generated in Experiment 1. }
+    \nbcnote{Row has significantly \textit{less} X-range compared to XOR, which I do not know how to explain...}
+    \label{fig:e1-statsboxes}
+    \end{center}
+\end{figure}
+
+On average, the axis ranges of generated categories tends to reflect those of the experimenter-defined categories.
+However, this cannot be said for the correlation between features in the generated categories.
+
+\begin{figure}
+    \begin{center}
+    \inputpgf{figs/}{e1-samples.pgf}
+    \caption{Sample categories generated in Experiment 1. }
+    \label{fig:e1-samples}
+    \end{center}
+\end{figure}
+
+% ----  yrange: Cluster, Row
+% T = 4.5714, p = 1.16282e-05 Means: 0.9023 (0.7495), 0.3566 (0.5553),
+
+% ----  yrange: Cluster, XOR
+% T = -3.952, p = 0.0001313583    Means: 0.9023 (0.7495), 1.4224 (0.6858),
+
+% ----  yrange: Row, XOR
+% T = -9.2603, p = 0.0    Means: 0.3566 (0.5553), 1.4224 (0.6858),
+
+% ---- Row X vs. Y:
+% T = 11.1735, p = 0.0    Means: 1.5738 (0.5678), 0.3566 (0.5553),
+
+% ---- Cluster X vs. Y:
+% T = -0.1258, p = 0.9002773681   Means: 0.8828 (0.7768), 0.9023 (0.7495),
+
+% ---- XOR X vs. Y:
+% T = -1.331, p = 0.1885039643    Means: 1.2586 (0.7542), 1.4224 (0.6858),
+
+% ---- Cluster positive correlation?
+% 0.145647334125
+% 0.137594247941
+
+% ---- XOR negative correlation?
+% 0.0456817132064
+% 0.0213027864542
+
+% ---- XOR has more total range than Cluster?
+% T = -4.8918, p = 3.1371e-06 Means: 1.7852 (0.9012), 2.681 (1.1023),
+
+% ---- XOR has more total range than Row?
+% T = -4.3418, p = 3.0207e-05 Means: 1.9303 (0.7412), 2.681 (1.1023),
+
+% ----  yrange: Cluster, Row
+% T = 4.5714, p = 1.16282e-05 Means: 0.9023 (0.7495), 0.3566 (0.5553),
+
+% ----  yrange: Cluster, XOR
+% T = -3.952, p = 0.0001313583    Means: 0.9023 (0.7495), 1.4224 (0.6858),
+
+% ----  yrange: Row, XOR
+% T = -9.2603, p = 0.0    Means: 0.3566 (0.5553), 1.4224 (0.6858),
+
+% ----  correlation: Cluster, Row
+% T = 2.029, p = 0.044618689  Means: 0.0753 (0.4058), -0.0587 (0.3196),
+
+% ----  correlation: Cluster, XOR
+% T = 2.5588, p = 0.0117466248    Means: 0.0753 (0.4058), -0.1538 (0.5682),
+
+% ----  correlation: Row, XOR
+% T = 1.1219, p = 0.2641961314    Means: -0.0587 (0.3196), -0.1538 (0.5682),
+
 \subsection{Discussion}

### 2017-03-23 Manuscripts/cog-psych/cogpsych.tex
xor-cluster-row sample plot works

@@
 
+It is firstly worth noting that we observed a substantial degree of individual differences in our data.
+In Figure \ref{fig:e1-samples} we have plotted sample data from several participants, from which it evident that different participants assumed qualitatively distinct approaches to category generation.
+In this section we will focus on what can be learned from analyzing the data in aggregate, but in later sections we will explore how individual differences can be explained.
+
+\begin{figure}
+    \begin{center}
+    \inputpgf{figs/}{e1-samples.pgf}
+    \caption{Sample categories generated in Experiment 1. }
+    \label{fig:e1-samples}
+    \end{center}
+\end{figure}
+
 To evaluate the role of contrast in category generation, we computed the number of times each stimulus was generated, as a function of its average city-block distance from members of the experimenter-defined ``Alpha'' category.
@@
 However, this cannot be said for the correlation between features in the generated categories.
-
-\begin{figure}
-    \begin{center}
-    \inputpgf{figs/}{e1-samples.pgf}
-    \caption{Sample categories generated in Experiment 1. }
-    \label{fig:e1-samples}
-    \end{center}
-\end{figure}
-
-% ----  yrange: Cluster, Row
-% T = 4.5714, p = 1.16282e-05 Means: 0.9023 (0.7495), 0.3566 (0.5553),
-
-% ----  yrange: Cluster, XOR
-% T = -3.952, p = 0.0001313583    Means: 0.9023 (0.7495), 1.4224 (0.6858),
-
-% ----  yrange: Row, XOR
-% T = -9.2603, p = 0.0    Means: 0.3566 (0.5553), 1.4224 (0.6858),
-
-% ---- Row X vs. Y:
-% T = 11.1735, p = 0.0    Means: 1.5738 (0.5678), 0.3566 (0.5553),
-
-% ---- Cluster X vs. Y:
-% T = -0.1258, p = 0.9002773681   Means: 0.8828 (0.7768), 0.9023 (0.7495),
-
-% ---- XOR X vs. Y:
-% T = -1.331, p = 0.1885039643    Means: 1.2586 (0.7542), 1.4224 (0.6858),
+As in the experimenter-defined categories,
 
@@
 % 0.0213027864542
-
-% ---- XOR has more total range than Cluster?
-% T = -4.8918, p = 3.1371e-06 Means: 1.7852 (0.9012), 2.681 (1.1023),
-
-% ---- XOR has more total range than Row?
-% T = -4.3418, p = 3.0207e-05 Means: 1.9303 (0.7412), 2.681 (1.1023),
-
-% ----  yrange: Cluster, Row
-% T = 4.5714, p = 1.16282e-05 Means: 0.9023 (0.7495), 0.3566 (0.5553),
-
-% ----  yrange: Cluster, XOR
-% T = -3.952, p = 0.0001313583    Means: 0.9023 (0.7495), 1.4224 (0.6858),
-
-% ----  yrange: Row, XOR
-% T = -9.2603, p = 0.0    Means: 0.3566 (0.5553), 1.4224 (0.6858),
 

### 2017-03-23 Manuscripts/cog-psych/cogpsych.tex
drafted all of e1, started e2

@@
 
-Clearly, participants tend to generate categories that differ from what is already known.
-However, the secondary goal of this experiment was to replicate the classic finding that generated and previously-known categories share distributional properties.
-To evaluate this, we computed, for each participant, the range of the generated category along each axis (X, Y), as well as the correlation between features within each category.
-These data, shown in Figure \ref{fig:e1-statsboxes}, reveal a broad-yet-incomplete replication.
+The secondary goal of this experiment was to replicate the classic finding that generated and previously-known categories share distributional properties.
+To evaluate this, we computed, for each participant, the range of their generated category along each axis (X, Y), as well as the correlation between features.
+These data, shown in Figure \ref{fig:e1-statsboxes}, reveal broad individual differences: within each condition, participants generated categories spanning the entire X- and Y- axis as well as categories that spanned very little along each dimension.
+Likewise, in each condition participants generated categories possessing strongly positive, neutral, and strongly negative correlations between the dimensions.
+Comparing the distributional statistics between conditions yields a broad-yet-incomplete replication of the classic effect.
 
 With respect to ranges along each axis (X, Y), the generated categories from each condition tend to reflect the ranges of the experimenter-defined categories.
-The categories generated in the Cluster condition had the less X-axis range compared to Row, $t(123) = 5.61$, $p < 0.001$, and XOR $t(120) = 2.68$, $p = 0.008$.
-Likewise, categories generated in the Row condition had less Y-axis range compared to Cluster, $t(123) = 4.57$, $p < 0.001$ and XOR $t(117) = 9.26$, $p < 0.001$, and categories from the Cluster condition had less Y-axis range compared to XOR $t(120) = 3.95$, $p < 0.001$.
+The categories generated in the Cluster condition were less widely distributed compared to Row, $t(123) = 5.61$, $p < 0.001$, and XOR, $t(120) = 2.68$, $p = 0.008$.
+Likewise, categories generated in the Row condition had less Y-axis range compared to Cluster, $t(123) = 4.57$, $p < 0.001$ and XOR $t(117) = 9.26$, $p < 0.001$, and categories from the Cluster condition had less Y-axis range compared to XOR, $t(120) = 3.95$, $p < 0.001$.
+However, whereas the correlations in the Cluster and Row conditions were not systematically positive or negative ($ps > 0.1$), the generated categories in the XOR condition tended to possess a \textit{negatively} correlated dimensions, $t(57) = 2.04$, $p = 0.046$.
+This is the opposite of what can be expected, assuming learners are emulating the distributional structure of the experimenter-defined class (which possesses perfectly positively correlated features).
 
@@
 
-On average, the axis ranges of generated categories tends to reflect those of the experimenter-defined categories.
-However, this cannot be said for the correlation between features in the generated categories.
-As in the experimenter-defined categories,
-
-% ---- Cluster positive correlation?
-% 0.145647334125
-% 0.137594247941
-
-% ---- XOR negative correlation?
-% 0.0456817132064
-% 0.0213027864542
-
-% ----  correlation: Cluster, Row
-% T = 2.029, p = 0.044618689  Means: 0.0753 (0.4058), -0.0587 (0.3196),
-
-% ----  correlation: Cluster, XOR
-% T = 2.5588, p = 0.0117466248    Means: 0.0753 (0.4058), -0.1538 (0.5682),
-
-% ----  correlation: Row, XOR
-% T = 1.1219, p = 0.2641961314    Means: -0.0587 (0.3196), -0.1538 (0.5682),
-
 \subsection{Discussion}
 
-\clearpage
+In Experiment 1 we sought to test whether category contrast systematically influences categories that are creatively generated by participants.
+We found strong evidence for effects of category contrast: participants were more likely to generate stimuli that are more distant from (i.e., less similar to) members of a previously-learned category, and members of participant-generated categories tended to be more similar to one another than to members of previously-learned categories.
+We also partially replicated the classic finding that the distributional structure of generated categories reflects that of previously learned categories \citep{jern2013probabilistic,ward1994structured}: members of generated categories were more widely distributed along dimensions which were widely distributed in the experimenter-defined category.
+
+A notable divergence, however, was the finding that the feature correlations within the categories generated in the XOR condition tended oppose that of the experimenter-defined category.
+While this is difficult to account for within existing theoretical approaches \citep[i.e.,][]{jern2013probabilistic}, it can be concisely explained from a category contrast perspective.
+Within the XOR condition, individuals who seek to generate a category that is perceptually distinct from what is already known are left with only the upper-left and bottom-right quadrants, as members of the experimenter-defined XOR category lie in the bottom-right and top-left quadrants.
+If examples are generated into both of the available quadrants, the generated category will possess a strongly negative correlation, opposing that of the experimenter-defined class.
+
+The negative correlations observed in the XOR condition may therefore indicate something further about the role of category contrast: constraints on creative generation imposed by category contrast may not simply influence the \textit{location} of generated categories, but also their distributional structure.
+In Experiment 2, we test this claim more systematically.
 
 \section{Experiment 2}
+
+In Experiment 1, we observed that participants are less likely to generate stimuli that are highly similar to members of existing categories.
+Effectively, this demonstrates that members of known categories influence where new categories will be generated within the domain.
+A related question, however, concerns whether category contrast influences the distributional structure of generated categories.
+Specifically, does the shape of the unoccupied area within the domain influence the types of categories people generate?
 
@@
 
-We developed two Alpha categories (see Figure \ref{fig:e2-conditions}): the `Bottom' category is a tight cluster in the bottom-center of the space, and the `Middle' category is identical except that it lies in the center of stimulus space.
-Although this manipulation is minimal (the conditions differ only slightly in the Y-axis position of the experimenter-defined category), the PACKER model predicts strong between-condition differences.
+To address this question, we developed two new categories within the same domain as Experiment 1 (see Figure \ref{fig:e2-conditions}).
+The categories possess an identical distributional (both are tight clusters of examples with a neutral feature correlation), and only differ slightly in their Y-axis position: the `Bottom' category is a tight cluster in the bottom-center of the space, and the `Middle' category lies in the center of stimulus space.
+
+By consequence, approaches focusing on the emulation of distributional structure \citep[i.e.,][]{jern2013probabilistic} would fail to predict any differences in the types of categories generated after learning the
+
+Importantly, however, the slight change to the position along the Y-axis produces considerable changes to the shape of the space that remains unoccupied after learning.
+
+the PACKER model predicts strong between-condition differences.
 According to PACKER, the nature of the space not occupied by the existing categories should determine where members of the Beta category are likely to be generated.

### 2017-03-27 Manuscripts/cog-psych/cogpsych.tex
whoops quick fix
updates to e2 writeup; drafted transition from e1

--- 2017-03-23 Manuscripts/cog-psych/cogpsych.tex
@@
 
-It is firstly worth noting that we observed a substantial degree of individual differences in our data.
+To begin, it is worth noting that we observed a substantial degree of individual differences in our data.
 In Figure \ref{fig:e1-samples} we have plotted sample data from several participants, from which it evident that different participants assumed qualitatively distinct approaches to category generation.
@@
     \label{fig:e1-samples}
+    \nbcnote{I selected these at random, we can pick more representative samples later.}
     \end{center}
@@
 To evaluate the role of contrast in category generation, we computed the number of times each stimulus was generated, as a function of its average city-block distance from members of the experimenter-defined ``Alpha'' category.
-These data, shown in Figure \ref{fig:e1-distanceplots}, reveal a clear pattern: examples that are highly distant from members of the experimenter-defined categories are more likely to be generated into a new category.
+These data, shown in Figure \ref{fig:e1-distanceplots}, reveal a clear pattern: examples that are more distant from members of the experimenter-defined categories are more likely to be generated into a new category.
 
@@
 That is, members of the generated category tended to be more similar to one another than to members of the experimenter-defined category.
-T-Tests comparing the amount of within- and between- class distance among the categories support this claim in each condition: Cluster, $t(63) = 11.43$, $p < 0.001$; Row, $t(60) = 13.16$, $p < 0.001$; and XOR, $t(57) = 3.64$, $p < 0.001$.
+To formally evaluate this claim, we conducted T-Tests comparing the amount of within- and between- class distance in each condition.
+All conditions possessed greater between-class distance: Cluster, $t(63) = 11.43$, $p < 0.001$; Row, $t(60) = 13.16$, $p < 0.001$; and XOR, $t(57) = 3.64$, $p < 0.001$.
 These results provide clear evidence of an effect of category contrast: participants prefer to generate categories that are dissimilar to the learned category but maintain some level of internal cohesion.
 
-Clearly, participants tend to generate categories that differ from what is already known.
-However, the secondary goal of this experiment was to replicate the classic finding that generated and previously-known categories share distributional properties.
-To evaluate this, we computed, for each participant, the range of the generated category along each axis (X, Y), as well as the correlation between features within each category.
-These data, shown in Figure \ref{fig:e1-statsboxes}, reveal a broad-yet-incomplete replication.
+The secondary goal of this experiment was to replicate the classic finding that generated categories often possess the same distributional properties as previously-known categories.
+For each generated category, we computed the category range along each axis (X, Y), as well as the correlation between features.
+These data, shown in Figure \ref{fig:e1-statsboxes}, reveal broad individual differences: within each condition, participants generated categories spanning the entire X- and Y- axis as well as categories that spanned very little along each.
+Likewise, in each condition participants generated categories possessing strongly positive, neutral, and strongly negative correlations between the dimensions.
+Comparing the distributional statistics between conditions yields a broad-yet-incomplete replication of the classic effect.
 
 With respect to ranges along each axis (X, Y), the generated categories from each condition tend to reflect the ranges of the experimenter-defined categories.
-The categories generated in the Cluster condition had the less X-axis range compared to Row, $t(123) = 5.61$, $p < 0.001$, and XOR $t(120) = 2.68$, $p = 0.008$.
-Likewise, categories generated in the Row condition had less Y-axis range compared to Cluster, $t(123) = 4.57$, $p < 0.001$ and XOR $t(117) = 9.26$, $p < 0.001$, and categories from the Cluster condition had less Y-axis range compared to XOR $t(120) = 3.95$, $p < 0.001$.
+The categories generated in the Cluster condition were less widely distributed along the X-axis compared to Row, $t(123) = 5.61$, $p < 0.001$, and XOR, $t(120) = 2.68$, $p = 0.008$.
+Likewise, categories generated in the Row condition had less Y-axis range compared to Cluster, $t(123) = 4.57$, $p < 0.001$ and XOR $t(117) = 9.26$, $p < 0.001$, and categories from the Cluster condition had less Y-axis range compared to XOR, $t(120) = 3.95$, $p < 0.001$.
+However, whereas the correlations in the Cluster and Row conditions were not systematically positive or negative ($ps > 0.1$), the generated categories in the XOR condition tended to possess \textit{negatively} correlated dimensions, $t(57) = 2.04$, $p = 0.046$.
+This finding is notable, as it is the opposite of what would be expected, assuming learners are emulating the distributional structure of the experimenter-defined class (which possesses perfectly positively correlated features).
 
@@
 
-On average, the axis ranges of generated categories tends to reflect those of the experimenter-defined categories.
-However, this cannot be said for the correlation between features in the generated categories.
-As in the experimenter-defined categories,
-
-% ---- Cluster positive correlation?
-% 0.145647334125
-% 0.137594247941
-
-% ---- XOR negative correlation?
-% 0.0456817132064
-% 0.0213027864542
-
-% ----  correlation: Cluster, Row
-% T = 2.029, p = 0.044618689  Means: 0.0753 (0.4058), -0.0587 (0.3196),
-
-% ----  correlation: Cluster, XOR
-% T = 2.5588, p = 0.0117466248    Means: 0.0753 (0.4058), -0.1538 (0.5682),
-
-% ----  correlation: Row, XOR
-% T = 1.1219, p = 0.2641961314    Means: -0.0587 (0.3196), -0.1538 (0.5682),
-
 \subsection{Discussion}
 
-\clearpage
+In Experiment 1 we sought to test whether category contrast imposes constraints on creative generation.
+We found strong evidence for effects of category contrast: participants were more likely to generate stimuli that are more distant from (i.e., less similar to) members of a previously-learned category, and members of participant-generated categories tended to be more similar to one another than to members of previously-learned categories.
+We also partially replicated the classic finding that the distributional structure of generated categories reflects that of previously learned categories \citep{jern2013probabilistic,ward1994structured}: members of generated categories were more widely distributed along dimensions which were widely distributed in the experimenter-defined category.
+
+Notably, however, we also found that participants who learned an XOR category (composed of exemplars following a positive diagonal, see Figure \ref{fig:e1-conditions}) tended to generate items according to a \textit{negative} feature correlation -- the opposite of what was present in the previously learned category.
+While this may be difficult to account for under existing theoretical approaches (which assume generated categories follow the same distributional structure as known categories), it can be concisely explained from a category contrast perspective.
+Specifically, within the XOR condition, individuals who seek to generate a category that is perceptually distinct from what is already known are left with only the upper-left and bottom-right quadrants of the space, as members of the previously-learned XOR category lie in the bottom-left and top-right.
+If examples are generated into both of the available quadrants, the generated category will possess a strongly negative correlation, opposing that of the experimenter-defined class.
+
+Thus, while the core results of Experiment 1 indicate that generated categories systematically contrast with what is already known, the negative correlations observed in the XOR condition may signal something further.
+That is, the constraints on creative generation imposed by category contrast may not simply influence the \textit{location} of generated categories, but also their distributional structure.
+In Experiment 2, we test this claim more systematically.
 
 \section{Experiment 2}
+
+To test whether category contrast influences the distributional structure of generated categories, we repeated Experiment 1 using two new category types, depicted in Figure \ref{fig:e2-conditions}.
+The category types possess an identical distributional structure (both are tight clusters of examples with a neutral feature correlation), and only differ slightly in their Y-axis position: the `Bottom' category lies in the bottom-center of the space, and the `Middle' category lies in the center.
+Although this manipulation may be viewed as somewhat `weak', the distributional equality of these conditions is key to the design of the experiment: if the structure of previously learned categories were the only influence on the structure of generated categories, we should observe no difference between these two conditions with respect to distributional structure.
 
@@
 
-We developed two Alpha categories (see Figure \ref{fig:e2-conditions}): the `Bottom' category is a tight cluster in the bottom-center of the space, and the `Middle' category is identical except that it lies in the center of stimulus space.
-Although this manipulation is minimal (the conditions differ only slightly in the Y-axis position of the experimenter-defined category), the PACKER model predicts strong between-condition differences.
-According to PACKER, the nature of the space not occupied by the existing categories should determine where members of the Beta category are likely to be generated.
-Thus, the lower areas of the stimulus space should be less frequently used for generation in the Bottom condition compared to the Middle (as these areas possess greater similarity to the Bottom category).
-Conversely, the upper areas of the stimulus space should be used for generation more frequently in the Bottom condition compared to Middle.
+Alternatively, participants seeking to make a perceptually distinct category would be more likely to distribute members of the generated category into areas that are distant from members of known categories.
+Thus, if category contrast influences the distributional structure of the categories people generate, then we should observe different types of categories according to the shape of the space that is unoccupied by members of previously learned categories.
+Specifically, participants assigned to learn the Bottom category would be less likely to generate exemplars into the lower regions of the stimulus space (as these areas possess greater similarity to members of the Bottom category), preferring instead to distribute exemplars across the upper region of the space.
+This constraint is not imposed in the Middle condition, as the Middle category exemplars are equidistant to the upper and lower regions of the space -- accordingly, participants should be more likely to utilize both of these areas.
 
@@
 \subsection{Results}
+
+Because the conditions differ only in the Y-axis position of the experimenter-defined category, we began by comparing the conditions in terms of the frequency with which participants generated examples above and below the categories: we counted the number of participants in each condition who placed at least one `Beta' exemplar on the top and bottom `rows' of the space (the maximum and minimum possible y-axis value, respectively).
+The resulting contingencies data are shown in Table \ref{table:e2-subset-table}.
+Fisher's Exact Tests reveal that more Middle participants generated an exemplar in the bottom row , $p < 0.001$, but the conditions did not differ in use of the top of the space, $p = 0.16$.
+More Middle participants placed exemplars in the top \textit{and} bottom rows, $p = 0.038$.
+
+\begin{table}
+\begin{center} 
+\caption{Experiment 2 results.} 
+\label{table:e2-subset-table} 
+\vskip 0.12in
+\begin{tabular}{ l r r}
+    \textbf{Middle}         & Used top row & No top row \\
+    \hline
+    Used bottom row       &  28 & 18  \\
+    No bottom row          &  11 &  4  \\
+    \\
+    \textbf{Bottom}         & Used top row & No top row \\
+    \hline
+    Used bottom row        & 16 & 8 \\
+    No bottom row          & 31 & 6 \\
+\end{tabular}
+\end{center} 
+\end{table}
 

### 2017-03-28 Manuscripts/cog-psych/cogpsych.tex
edits to e2 results, added figures

@@
 
+This constraint, however, only serves to signify one possible account of category generation: the nature of PACKER's account can be drastically altered through changes to the $\phi$ and $\gamma$ parameters.
+Furthermore, by fitting these parameters to a dataset, one can describe the relative roles of between-class contrast and within-class similarity in generation.
+For example, as will be discussed in more detail below, PACKER is formally equivalent to the copy-and-tweak model proposed by \cite{jern2013probabilistic} when $\phi = 0$ and $\gamma = 1$.
+Likewise, when $\phi < 0$ and $\gamma = 0$, PACKER represents a contrast-only generation mode, relying exclusively on contrast when generating new categories.
+Finally, when $\phi = \gamma < 0$, PACKER represents a ``pure-packing'' approach, placing exemplars in any unoccupied area of the domain.
+Thus, beyond specifying PACKER's core proposal, the $\phi$ and $\gamma$ allow the us to address a wide variety of qualitatively distinct generation strategies.
+
 The probability that a given candidate $y$ will be generated is evaluated using an Exponentiated \citet{luce1977choice} choice rule.
@@
 
-\subsection{Participants \& Materials}
+\subsection{Participants, Materials, \& Procedure}
 
@@
 Again, the assignment of perceptual features (color, size) to axes of the domain space (x, y) was counterbalanced across participants.
-
-\subsection{Procedure}
-
 The procedure was exactly as in Experiment 1: participants first completed a short training phase, followed by the generation phase, followed by the generalization phase (data from this phase is not discussed in this report).
@@
 
-Because the conditions differ only in the Y-axis position of the experimenter-defined category, we began by comparing the conditions in terms of the frequency with which participants generated examples above and below the categories: we counted the number of participants in each condition who placed at least one `Beta' exemplar on the top and bottom `rows' of the space (the maximum and minimum possible y-axis value, respectively).
+\begin{figure}
+    \begin{center}
+    \inputpgf{figs/}{e2-samples.pgf}
+    \caption{Sample categories generated in Experiment 2. }
+    \label{fig:e2-samples}
+    \end{center}
+\end{figure}
+
+As in Experiment 1, we observed broad differences in the generation approach taken by different participants.
+To characterize the nature of these differences, Figure \ref{fig:e2-samples} depicts sample categories generated by participants.
+The data from each condition in Figure \ref{fig:e2-samples} is organized into four columns based on commonly observed patterns of generation: a `Cluster' type of tightly-clustered examples,   `Row' and `Column' types of exemplars widely distributed along the one axis but narrowly along the other, and a `Corners' type, wherein participants placed exemplars in disparate corners of the space.
+As before, in this section we focus no what can be concluded from analyzing the data in aggregate, but in a later section we will focus more specifically on explaining the individual differences.
+
+We began our analysis by again testing for the broad influence of category contrast on generation.
+As in Experiment 1, we computed the frequency each stimulus was generated as a function of its average distance from members of the experimenter defined category, as well as each participant's average within- and between- category distance.
+These data, shown in Figure \ref{fig:e2-distanceplots}, yield highly similar results.
+Stimuli that are more distant from members of the experimenter=-defined category were more frequently generated, and the categories in each condition tended to possess more between-class than within-class distance: Bottom, $t(60) = 5.5$, $p < 0.001$; Middle, $t(60) = 2.71$, $p = 0.009$.
+
+We did, however, observe a notable subgroup of participants in each condition who generated categories with more within-class than between-class distance.
+Upon manual inspection, many of these individuals appear to have assumed a `Corners' strategy, placing exemplars in disparate corners of the space, thus producing much more within-class distance, see Figure \ref{fig:e2-samples} for examples.
+
+\begin{figure}
+    \begin{center}
+    \inputpgf{figs/}{e2-distanceplots.pgf}
+    \caption{Experiment 2 results.
+\textit{Left}: Frequency of generation as a function of distance from members of the experimenter-defined category.
+\textit{Right}: Scatter plot of within-class versus between-class distance in each of the participant-generated categories.}
+    \label{fig:e2-distanceplots}
+    \end{center}
+\end{figure}
+
+Because the conditions differ only in the Y-axis position of the experimenter-defined category, we then compared the conditions in terms of the frequency with which participants generated examples above and below the categories.
+Specifically, we counted the number of participants in each condition who placed at least one `Beta' exemplar on the top and bottom `rows' of the space (the maximum and minimum possible y-axis value, respectively).
 The resulting contingencies data are shown in Table \ref{table:e2-subset-table}.
@@
 \subsection{Discussion}
+
+\section{Simulations}
+
+\section{General Discussion}
 

### 2017-03-28 Experiments/middle-bottom/analysis/plot.distance.figs.py

@@ -33,8 +33,8 @@ for c in pd.unique(info.condition):
 		ngenerations = ngenerations.append(row, ignore_index = True)
 
 
+fh, ax = plt.subplots(1,2,figsize = (6,2.7))
 
-fh, ax = plt.subplots(1,2,figsize = (3.45,1.85))
 styles = dict(Middle = '-o', Bottom = '-s')
 colors = dict(Middle = 'orange', Bottom = 'purple')
 
@@ -59,16 +59,16 @@ for i, (c, rows) in enumerate(ngenerations.groupby('condition')):
 
 h.xaxis.grid(False)
 h.set_xticks([])
-h.legend(loc = 'upper left', frameon = True, framealpha = 1, fontsize = 9)
+h.legend(loc = 'upper left', frameon = True, framealpha = 1, fontsize = 10)
 
 
 xax = h.axis()
-h.text(xax[0],xax[2] -1, 'Min', fontsize = 9, va = 'top')
-h.text(xax[1],xax[2] -1, 'Max', fontsize = 9, va = 'top', ha = 'right')
-h.set_xlabel('Distance')
+h.text(xax[0],xax[2] -1, 'Min', fontsize = 10, va = 'top')
+h.text(xax[1],xax[2] -1, 'Max', fontsize = 10, va = 'top', ha = 'right')
+h.set_xlabel('Distance',fontsize = 12)
 h.set_yticks(np.arange(0,35, 5))
-h.set_yticklabels(np.arange(0,35, 5),fontsize = 9)
-h.set_ylabel('Generations Per Stimulus', fontsize = 10)
+h.set_yticklabels(np.arange(0,35, 5),fontsize = 10)
+h.set_ylabel('Generations Per Stimulus', fontsize = 12)
 
 
 h = ax[1]
@@ -88,17 +88,18 @@ h.set_yticks([])
 
 h.axis([0, 1.5, 0, 1.5])
 # h.legend(loc = 'upper right', frameon = True, framealpha = 1, 
-# 	ncol = 2, columnspacing = 0.1, labelspacing = 0.1, fontsize = 10)
-h.set_xlabel('Within-Class Distance',fontsize = 9)
-h.set_ylabel('Between-Class Distance',fontsize = 9)
+# 	ncol = 2, columnspacing = 0.1, labelspacing = 0.1, fontsize = 12)
+h.set_xlabel('Within-Class Distance',fontsize = 12)
[…]

### 2017-03-28 Manuscripts/cog-psych/cogpsych.tex
e2 drafted

@@
 
+\nbcnote{That last paragraph was added since your last read, joe!}
+
 The probability that a given candidate $y$ will be generated is evaluated using an Exponentiated \citet{luce1977choice} choice rule.
@@
     \caption{Box-plots of the distributional statistics from the categories generated in Experiment 1. }
-    \nbcnote{Row has significantly \textit{less} X-range compared to XOR, which I do not know how to explain...}
+    \nbcnote{Weak spot: row has significantly \textit{less} X-range compared to XOR, which I do not know how to explain. For now I've ignored it.}
     \label{fig:e1-statsboxes}
@@
 Thus, if category contrast influences the distributional structure of the categories people generate, then we should observe different types of categories according to the shape of the space that is unoccupied by members of previously learned categories.
-Specifically, participants assigned to learn the Bottom category would be less likely to generate exemplars into the lower regions of the stimulus space (as these areas possess greater similarity to members of the Bottom category), preferring instead to distribute exemplars across the upper region of the space.
-This constraint is not imposed in the Middle condition, as the Middle category exemplars are equidistant to the upper and lower regions of the space -- accordingly, participants should be more likely to utilize both of these areas.
+The difference in the Y-axis position between the Bottom and Middle types produces a considerable change to the shape of the unoccupied space: participants assigned to learn the Bottom category would be less likely to generate exemplars into the lower regions of the stimulus space (as these areas possess greater similarity to members of the Bottom category), preferring instead to distribute exemplars across the upper region of the space.
+This constraint is lifted in the Middle condition, as the Middle category exemplars are equidistant to the upper and lower regions of the space.
+Accordingly, participants should be more likely to utilize both of these areas.
+
+\nbcnote{Weak point: this prediction is squishy-- why couldn't middle participants generate items above \textit{or} below?
+When we initially developed this, we were working with the pure-packing variant of PACKER, which is what predicted above-and-below for middle but not bottom.
+}
 
@@
 The data from each condition in Figure \ref{fig:e2-samples} is organized into four columns based on commonly observed patterns of generation: a `Cluster' type of tightly-clustered examples,   `Row' and `Column' types of exemplars widely distributed along the one axis but narrowly along the other, and a `Corners' type, wherein participants placed exemplars in disparate corners of the space.
-As before, in this section we focus no what can be concluded from analyzing the data in aggregate, but in a later section we will focus more specifically on explaining the individual differences.
+As before, in this section we focus on what can be concluded from analyzing the data in aggregate, but in later sections we will focus more specifically on explaining the individual differences.
+
+\begin{figure}
+    \begin{center}
+    \inputpgf{figs/}{e2-distanceplots.pgf}
+    \caption{Experiment 2 results.
+\textit{Left}: Frequency of generation as a function of distance from members of the experimenter-defined category.
+\textit{Right}: Scatter plot of within-class versus between-class distance in each of the participant-generated categories.}
+    \nbcnote{May want to change condition colors if purple-orange is used as the colormap for the range heat-maps in the simulations section.}
+    \label{fig:e2-distanceplots}
+    \end{center}
+\end{figure}
 
@@
 
-\begin{figure}
-    \begin{center}
-    \inputpgf{figs/}{e2-distanceplots.pgf}
-    \caption{Experiment 2 results.
-\textit{Left}: Frequency of generation as a function of distance from members of the experimenter-defined category.
-\textit{Right}: Scatter plot of within-class versus between-class distance in each of the participant-generated categories.}
-    \label{fig:e2-distanceplots}
-    \end{center}
-\end{figure}
+\nbcnote{ Weak spot: the obvious next analysis here would involve the same box-and-whisker plot as in Experiment 1.
+We ran this analysis but the conditions did not differ in Y-range due to the data being weirdly distributed.
+But it feels odd to ignore this point, so I've included the box-and-whisker plot, as well as the graph-of-many-lines figure, on a separate page at the end.}
 
@@
 The resulting contingencies data are shown in Table \ref{table:e2-subset-table}.
-Fisher's Exact Tests reveal that more Middle participants generated an exemplar in the bottom row , $p < 0.001$, but the conditions did not differ in use of the top of the space, $p = 0.16$.
-More Middle participants placed exemplars in the top \textit{and} bottom rows, $p = 0.038$.
 
@@
 
+Firstly, it should be noted that nearly every participant utilized the top and/or bottom rows: only $10 / 122$ participants generated their category entirely within the interior region.
+Fisher's Exact Tests comparing the conditions reveal that more Middle participants generated an exemplar in the bottom row, $p < 0.001$, again demonstrating the role of contrast guiding where exemplars are generated.
+The conditions did not differ in use of the top of the space, $p = 0.16$, however, more Middle participants placed exemplars in the top \textit{and} bottom rows, $p = 0.038$.
+The latter effect is of interest here, as it indicates that the shape of the unoccupied space exerts some influence on the distributional structure of generated categories: participants in the Middle condition were more likely to generate a category spanning the entire Y-axis.
+
 \subsection{Discussion}
+
+In Experiment 2, we replicated the core findings from Experiment 1: stimuli are more likely to be generated if they are distant from exemplars in other categories, and most participants generate categories with more between-class than within-class distance.
+However, we additionally found that the \textit{position} of a previously learned category (rather than its distributional structure) influences the types of categories people generate: participants who learned the `Middle' type were more likely to generate categories spanning the entire Y-axis of the space.
+Participants who learned the `Bottom' type were less likely to do so, presumably because of the presence of opposite category exemplars in the lower regions of the space.
+
+This finding is cannot be explained under the theoretical perspective that the distributional structure of previously learned categories is the sole determinant of the distributional structure of generated categories.
+However, the observed performance is sensible under a category contrast perspective: participants seeking to generate a perceptually distinct category will be more likely to use areas of space that are unoccupied by exemplars belonging to previously learned categories.
+In the Middle condition, the upper and lower regions of space are equidistant from members of the experimenter-defined category, whereas in the Bottom condition, the lower region of the space is closer to members of the experimenter-defined category.
+Thus, while Middle participants may form categories around the use of the equally unoccupied areas, the same is not true for the Bottom condition.
 
@@
 \clearpage
-
 \section{Acknowledgments}
-A previous version of this work appeared in the Proceedings of the Thirty-Ninth Annual Conference of the Cognitive Science Society.
+Previous versions of this work were presented at the Thirty-Ninth Annual Conference of the Cognitive Science Society and Forty-Ninth Annual Meeting of the Society for Mathematical Psychology.
 Support for this research was provided by the Office of the VCRGE at the UW - Madison with funding from the WARF.
 We thank Alan Jern and Charles Kemp for providing code and data.
-
 \end{flushleft}
@@
 \bibliography{citations.bib}
+\clearpage
+
+\begin{figure}
+    \begin{center}
+    \inputpgf{figs/}{e2-statsboxes.pgf}
+    \caption{Box-plots of the distributional statistics from the categories generated in Experiment 2. }
+    \nbcnote{Note the huge amount of variance in the range data, as well as the big Y-range difference in the medians.}
+
+    \label{fig:e2-statsboxes}
+    \end{center}
+\end{figure}
+
+\begin{figure}
+    \begin{center}
+    \inputpgf{figs/}{e2-yranges.pgf}
+    \caption{Y-Axis range and position of the categories generated in Experiment 2.
+Each line corresponds to a single category, with notches corresponding to the Y-axis position of exemplars within the category.
+Data from each participant is sorted by overall range, and then by condition.
+}
+    \label{fig:e2-statsboxes}
+    \end{center}
+\end{figure}
+
 \clearpage

### 2017-03-28 Experiments/middle-bottom/analysis/plot.yrange.data.py

@@ -6,6 +6,9 @@ import matplotlib.pyplot as plt
 np.set_printoptions(precision = 2)
 pd.set_option('precision', 2)
 
+execfile('Imports.py')
+import Modules.Funcs as funcs
+
 # import data
 con = sqlite3.connect('../data/experiment.db')
 info = pd.read_sql_query("SELECT * from participants", con)
@@ -22,7 +25,7 @@ df['x'] = stimuli.loc[df.stimulus, 'F1'].as_matrix()
 
 def plotlines(h, data, stats):
 	sort_df = stats.sort_values(by = ['yrange','condition'])
-	condition_colors = dict(Middle = 'r', Bottom = 'b')
+	condition_colors = dict(Middle = 'orange', Bottom = 'purple')
 	n = 0
 
 	for num, info in sort_df.iterrows():
@@ -53,7 +56,7 @@ def plotlines(h, data, stats):
 	h.set_yticks([])
 	h.set_xticks([])
 	h.set_ylabel('Y Axis Value', fontsize = 11)
-	h.set_xlabel('Y-Axis Range', fontsize = 11)
+	h.set_xlabel('Participant (sorted by overall range)', fontsize = 11)
 
 	for k, v in condition_colors.items():
 		plt.plot(np.NaN, np.NaN, '-', color = v, label = k)
@@ -67,5 +70,5 @@ plotlines(fh.gca(), df, stats)
 
 fh.savefig('yranges.pdf', bbox_inches = 'tight')
 
-# path = '../../../Manuscripts/cogsci-2017/figs/middle-bottom-yranges.pgf'
-# funcs.save_as_pgf(f, path)
+path = '../../../Manuscripts/cog-psych/figs/e2-yranges.pgf'
+funcs.save_as_pgf(fh, path)

### 2017-03-28 Manuscripts/cog-psych/cogpsych.tex
added simulation results, packer loglike fig

@@
 \usepackage{verbatim}
+\usepackage[toc]{appendix}
 \usepackage[natbibapa]{apacite}
@@
 Creating Something Different: Similarity and Contrast in Concept Generation.
-% ---- other ideas:
-% - The Constraining Influences of Contrast and Similartity in Concept Generation.
 
@@
 \doublespacing
-\section{Abstract}
+\section*{Abstract}
 
@@
 
+\label{section:PACKER-definition}
+
 As noted above, the constraint that new concepts should differ from what is already known has been largely been overlooked in previous work.
@@
 \subsubsection{Relation Between PACKER and Copy-And-Tweak}
-
+\label{section:copytweak-packer}
 The PACKER model is in many senses similar to the copy-and-tweak model reported by \cite{jern2013probabilistic}: both models are exemplar-based, and both models generate new items according to their similarity to known members of the target class.
@@
 
+Experiments 1 and 2 revealed systematic and strong effects of category contrast on creative generation.
+In this section, we report the results of simulations with formal models aimed at explaining our observations.
+Specifically, we present simulations from the PACKER model, as well as a `copy-and-tweak' model (discussed in section \ref{section:copytweak-packer}), defined as a variant of PACKER with the $\phi$ and $\gamma$ parameters held constant.
+The comparison of these two models serves to highlight the explanatory role of contrast within PACKER's framework: if contrast affords little explanatory advantage, then the two accounts should produce an equally strong account.
+We also present simulations from an implementation of the hierarchical sampling model proposed by \cite{jern2013probabilistic}, described in-depth in Appendix \ref{ap:hsampling-definition}.
+The comparison between the hierarchical sampling model and PACKER is meant to emphasize the role of emulation of distributional structure: whereas PACKER is insensitive to the distributional structure of learned categories (relying only on within- and between-class similarity), the hierarchical sampling model generates categories exclusively on the basis of knowledge of how existing classes are distributed.
+Our approach in this section is to first broadly evaluate and compare the quality of each model's account to our entire dataset (Experiments 1 and 2 combined), then we more specifically describe the strengths and weakness of each model's account.
+
+\subsection{Parameter-Fitting}
+
+To obtain a global measure of the quality of each model's account, we fitted the parameters of each model to our entire dataset (Experiments 1 and 2 combined), using a hill-climbing algorithm which maximized the log-likelihood of the model's predictions of the observed responses (1220 responses from 305 total participants).
+We fitted four parameters in the PACKER model ($c$, $\phi$, $\gamma$, and $\theta$; see Section \ref{section:PACKER-definition}), as well as four in the hierarchical sampling model ($\kappa$, $\rho$, $\nu$, and $\theta$; see Appendix \ref{ap:hsampling-definition}).
+We fitted only two parameters for the copy-and-tweak model ($c$, and $\theta$), as $\phi$ and $\gamma$ are held constant.
+Note that each model possesses a $\theta$ parameter fulfilling the same role (response determinism).
+Attention ($w$, see Equation \ref{eq:similarity}) in PACKER and copy-and-tweak was set uniformly.
+Parameters were not allowed to vary between participants or conditions -- the goal was to obtain the best-fitting values to our entire dataset.
+
+\begin{table}
+\centering
+\caption{Results of model-fitting to the combined datasets from Experiments 1 and 2.}
+\label{table:global-model-fits}
+\begin{tabular}{ l l l}
+\\
+ \textbf{PACKER}    & \textbf{Copy \& Tweak} & \textbf{Hierarchical Sampling} \\ 
+ \hline
+ $AIC = 9097$       & $AIC = 9842$          & $AIC = 10080$  \\ 
+ $L = -4545$        & $L = -4919$           & $L = -5036$  \\ 
+ $c=0.481$          & $c=3.189$             & $\kappa<0.001$\\
+ $\phi=-0.945$      & $\phi=0$ (fixed)      & $\nu=1.001$ \\ 
+ $\gamma=1.044$     & $\gamma=1$ (fixed)    & $\rho=5.313$  \\ 
+ $\theta=3.354$     & $\theta=2.968$        & $\theta = 7.949$  \\ 
+\end{tabular}
+\end{table}
+
+Table \ref{table:global-model-fits} contains the results of this fitting procedure.
+Due to the uneven number of fitted parameters among the models, we compare the model fits using the Akaike Information Criterion \citep[AIC;][]{akaike1974new}.
+Table \ref{table:global-model-fits} contains the AIC values of each model's best fit, as well as the corresponding log-likelihood ($L$) and the best-fitting parameter values.
+These results reveal strong model differentiation: the PACKER model achieved far better fits compared to the copy-and-tweak and hierarchical sampling models, and copy-and-tweak performed somewhat better than the hierarchical sampling model.
+While PACKER's advantage may tentatively be attributed to the model's sensitivity to category contrast (this will be explored in detail below), the advantage shown by copy-and-tweak over the hierarchical sampling model may be attributed to the assumption of an exemplar-based representation, as opposed to a prototype-based representation (via representing categories as multivariate normal distributions).
+
+More specifically, PACKER's best fitting parameter-setting reveals the...
+
+\begin{figure}
+    \begin{center}
+    \inputpgf{figs/}{packer-loglike.pgf}
+    \caption{PACKER's fit as a function of the between-class contrast parameter ($\phi$).
+To facilitate comparison, PACKER's other parameters ($c$, $\gamma$, $\theta$) were set to the best fitting values obtained for copy-and-tweak in Table \ref{table:global-model-fits}.
+Thus, in this figure PACKER is equivalent to copy-and-tweak when $\phi = 0$.}
+    \label{fig:packer-loglike}
+    \end{center}
+\end{figure}
+
+\nbcnote{Should include info on fitting to post-first trial.}
+
+\subsection{Individual Differences}
+
 \section{General Discussion}
@@
 }
-    \label{fig:e2-statsboxes}
+    \label{fig:e2-yranges}
     \end{center}
@@
 \appendix
+\numberwithin{equation}{section}
 

### 2017-03-28 Manuscripts/cog-psych/cogpsych.tex
global simulations drafted

@@
 These results reveal strong model differentiation: the PACKER model achieved far better fits compared to the copy-and-tweak and hierarchical sampling models, and copy-and-tweak performed somewhat better than the hierarchical sampling model.
-While PACKER's advantage may tentatively be attributed to the model's sensitivity to category contrast (this will be explored in detail below), the advantage shown by copy-and-tweak over the hierarchical sampling model may be attributed to the assumption of an exemplar-based representation, as opposed to a prototype-based representation (via representing categories as multivariate normal distributions).
+While PACKER's advantage may tentatively be attributed to the model's sensitivity to category contrast (this will be explored in detail below), the advantage shown by copy-and-tweak over the hierarchical sampling model may be attributed to its exemplar-based representation, as opposed to the prototype-based representation assumed in the hierarchical sampling model.
+\nbcnote{That last point is a little squishy.
+We don't have `proof' that we can point to for skeptical readers, but that's clearly the issue: many categories were not normally distributed, and the copying of structure really only works out in 2/5 conditions, and even then there's so much variability...}
+\nbcnote{Should include info on fits for only trials 2-4.}
 
-More specifically, PACKER's best fitting parameter-setting reveals the...
+Through comparison with the copy-and-tweak model, Figure \ref{fig:packer-loglike} more clearly demonstrates the explanatory gains yielded by PACKER's sensitivity to between-class contrast.
+Plotted is PACKER's log-likelihood as a function of the model's between-category parameter, $\phi$.
+The models other parameters ($c$, $\gamma$, $\theta$) were set according to copy-and-tweak's best fits from Table \ref{table:global-model-fits}, and thus when $\phi=0$, the models are equivalent.
+The figure clearly shows a ``sweet spot'':  a convex region in which PACKER achieves superior fits as a result of changes to the $\phi$ parameter.
+The best fitting values lie between -0.6 and -0.8, well beyond the 0 value assumed by the copy-and-tweak model (though note PACKER achieves even better fits when its other parameters are fitted, as in Table \ref{table:global-model-fits}).
 
@@
 \end{figure}
-
-\nbcnote{Should include info on fitting to post-first trial.}
 

### 2017-03-29 Manuscripts/cog-psych/cogpsych.tex
Looking great! i made some edits and a few comments.  things are progressing nicely

@@
 Kenneth J. Kurtz\textsuperscript{2}, 
-\& Joseph L. Austerweil\textsuperscript{1}
+and Joseph L. Austerweil\textsuperscript{1}
 \\[\baselineskip]
@@
 In a category generation task, participants are given some background knowledge about a domain (a cover story and/or learning one or more categories in the domain) and then are asked to generate another category in the domain \citep{jern2013probabilistic,ward1994structured}.
-Unlike creativity and innovation, categorization is well studied from a breadth of perspectives, including behavioral, neural, comparative, and computational perspectives \citep{kurtz2015human,mack2013,margolis2015,pothoswills2011}.
-By analyzing a creative task using formal tools from the categorization literature,
+Unlike creativity and innovation, categorization is well studied from a breadth of perspectives, including behavioral, neural, comparative, and computational  \citep{kurtz2015human,mack2013,margolis2015,pothoswills2011}.
+By analyzing a creative task using formal tools from the categorization literature, we can formalize some theories of creativity and test them empirically in a more rigorous manner.
 
@@
 Building on the largely successful exemplar modeling framework \citep{medin1978context,nosofsky1984choice,nosofsky1986attention}, we propose a novel exemplar model of category generation, {\em Producing Alike and Contrasting Knowledge using Exemplar Representations} (PACKER), formalizing how new categories should differ from previous categories.
+This model makes novel predictions about how contrast affects category generation, which we test using behavioral experiments.
 
@@
 Much of what we know about concept generation comes from the foundational literature on creative cognition.
-In a classic series of reports, Ward \& colleagues \citep{ward1995s,ward1994structured,marsh1999inadvertent,ward2002role,smith1993constraining} established that category generation is highly constrained by prior knowledge: Generated categories tend to consist of features observed in known categories, and they tend to exhibit the distributional properties as found in known categories.
+In a classic series of reports, Ward and colleagues \citep{ward1995s,ward1994structured,marsh1999inadvertent,ward2002role,smith1993constraining} established that category generation is highly constrained by prior knowledge: Generated categories tend to consist of features observed in known categories, and they tend to exhibit the distributional properties as found in known categories.
 In a classic study, \cite{ward1994structured} asked participants to generate new species of alien animals by drawing and describing members of the species.
@@
 Much of the work from this area \citep[e.g.,][]{smith1993constraining,marsh1999inadvertent} focuses on how sample cues (such as an example of a species generated by other participants) can drastically diminish creativity.
-However, the broader set of observations made by Ward \& colleagues provide a great deal of insight into the nature of creative generation.
+However, the broader set of observations made by Ward and colleagues provide a great deal of insight into the nature of creative generation.
 They indicate that people rely strongly on prior knowledge in the creative process, and people generate concepts in accord with what they already know.
-,
 Theoretical accounts of these effects have primarily been grounded within the categorization literature.
@@
 Problematically, however, the experimental paradigms employed in these studies were relatively uncontrolled, precluding the development of formal approaches.
-Specifically, participants in these studies are typically asked to generate objects that they have a rich degree of prior knowledge about (e.g., plants, animals, tools, toys), and so it is difficult to apply formal models to the data.
+Specifically, participants in these studies are typically asked to draw exemplars of novel categories by hand, and so it is difficult to apply formal models to the data.
 
-\cite{jern2013probabilistic} recently showed that creative generation could be studied in a more controlled manner through the well-developed methods of an artificial categorization paradigm \citep[see][]{kurtz2015human}.
-In their experiments 3 and 4, participants were exposed to members of experimenter-defined categories of "crystals" varying in size, hue, and saturation.
+\jlanote{The issue isn't prior knowledge actually -- Bayesian models can handle that gracefully -- it's the rich format of participant responses being 2-d sketches}
+
+\cite{jern2013probabilistic} recently showed that creative generation could be studied in a more controlled manner through the well-developed methods of an artificial categorization paradigm \citep[see][for a review]{kurtz2015human}.
+In Experiments 3 and 4 of their article, participants were exposed to members of experimenter-defined categories of "crystals" varying in size, hue, and saturation.
 Following a training phase during which the experimenter-defined categories were learned, participants were asked to generate novel categories of crystals.
@@
 This finding is notable, as it demonstrates that category generation can be studied in a well-known and highly controlled experimental paradigm.
+Note that generated exemplars for the novel category automatically contrasted from previous categories because the participants selected the size and saturation, but the exemplar's hue was set (and fixed) to a novel value.
+As such, hue was ignored in their analyses and may have satisfied any constraints contrast may have on category generation.
+
+\jlanote{It is hierarchical Bayesian, not Bayesian hierarchical.}
 
 The authors evaluated the predictions of several formal models on their data.
-Most notably, they showed that a Bayesian hierarchical sampling model provided the strongest account.
+Most notably, they showed that a hierarchical Bayesian sampling model provided the strongest account.
 Their model views observed examples as samples from an underlying category distribution, describing the location of the category in the space, as well as how it varies along each feature.
@@
 
-\cite{jern2013probabilistic} additionally tested a "copy-and-tweak" model that broadly resembles the earlier "Path of Least Resistance" account: the core proposal is that participants generate new items by copying stored examples from memory and tweaking them to generate something new.
+\cite{jern2013probabilistic} additionally tested a "copy-and-tweak" model that broadly resembles the earlier "Path of Least Resistance" account.
+The core proposal is that participants generate new items by copying stored examples from memory and tweaking them to generate something new.
 The copy-and-tweak model differs from the path of least resistance account in that it notably omits the hierarchical organization of categories, as well as selectivity in which features are changed.
-Instead, their copy-and-tweak model corresponds to a direct exemplar-similarity approach \citep[e.g.,][]{nosofsky1984choice,nosofsky1986attention}: participants are thought to generate new items according to their similarity to known members of the target category.
-This model provided a poor account of the observed data: indeed, the experiments devised by \citeauthor{jern2013probabilistic} were specifically designed to challenge this model.
+Instead, their copy-and-tweak model corresponds to a direct exemplar-similarity approach \citep[e.g.,][]{nosofsky1984choice,nosofsky1986attention}: The model generates new items according to their similarity to known members of the target category.
+This model provided a poor account of the observed data.
+Indeed, the experiments devised by \citeauthor{jern2013probabilistic} were specifically designed to challenge this model.
 However, its application is notable as a first step toward explaining category generation using well-known formal approaches from the categorization literature.
@@
 
-It is worth noting that the literature reviewed above provides a somewhat limited view of category generation: The main concern of the published experiments has been on the distributional correspondences between learned and generated categories, and as a result most of the modeling efforts have been towards explaining those effects.
+It is worth noting that the literature reviewed above provides a somewhat limited view of category generation.
+The main concern of the published experiments has been on the distributional correspondences between learned and generated categories, and as a result most of the modeling efforts have been towards explaining those effects.
 In this paper, we investigate another important constraint on the creative use of conceptual knowledge: category contrast.
-In order to generate a novel concept, individuals must produce something that is in some capacity \textit{different} from what they already know.
-Thus, in a trivial sense, contrast can be viewed as a fundamental constraint on creative generation: new concepts must firstly be different from existing ones.
+To generate a novel concept, individuals must produce something that is in some capacity \textit{different} from what they already know.
+Thus, contrast is a fundamental constraint on creative generation: new concepts must firstly be different from existing ones.
 
@@
 Although it is evident that people are \textit{capable} of creating new concepts and categories, it is not entirely clear how new concepts are systematically made different from what is already known.
-The hierarchical sampling model developed by \cite{jern2013probabilistic}, for example, does not provide any strong claims about how generated categories should contrast with learned ones: the model only proposes that generated categories are sampled from the same underlying domain distribution as observed categories, and will thus share a common distributional structure.
+The hierarchical sampling model developed by \cite{jern2013probabilistic} assumes that differences between generated category are only due to random variation.
+The model only proposes that generated categories are sampled from the same underlying domain distribution as observed categories, and will thus share a common distributional structure.
 The authors do not make predictions about the \textit{location} of the category within the domain (the perceptual instantiation of category members).
-Indeed, under a strict interpretation of the hierarchical Bayesian approach presented by \citeauthor{jern2013probabilistic}, the most probable new category to be generated is located in {\em exactly} the same location as the given category.
+Indeed, under a strict interpretation of their model, the most probable new category to be generated is located in {\em exactly} the same location and distributional information as the given category.
 However, this is not simply an issue with their model -- Hierarchical Bayesian models assume the same underlying distribution generates all of the categories and thus, any differences between categories is due to {\em noise} and should not be {\em systematic}.
-The best a a hierarchical Bayesian model can do at capturing contrast is to assume that the new category is placed uniformly at random over stimulus space (this was the case in the simulations conducted by \citeauthor{jern2013probabilistic}).
+The best a hierarchical Bayesian model can do at capturing contrast is to assume that the new category is placed uniformly at random over stimulus space.
+This defeats the purpose of a hierarchy as new category locations do not depend on the locations of previously observed categories.
 
@@
 As noted above, the constraint that new concepts should differ from what is already known has been largely been overlooked in previous work.
-This is no doubt in part due to the vague definition of what it means for a concept to be ``different'': a generated category may be different from what is already known in any number of respects.
+This is no doubt in part due to the vague definition of what it means for a concept to be ``different'': A generated category may be different from what is already known in any number of respects.
 Towards formalizing the role of contrast in creative generation, we developed the PACKER  ({\em Producing Alike and Contrasting Knowledge using Exemplar Representations}) model which explains category generation as a balance between two fundamental constraints: generated categories should not be similar to known categories, and exemplars within each category should be similar to one another.
@@
 
+\jlanote{Yeah, I think this would make a nice subsection in the general discussion.}
+
 Both PACKER and the GCM simulate categorization under the assumption that learners represent categories as a collection of exemplars, corresponding to the labeled stimuli they have observed.
 The exemplars are encoded within a $k$-dimensional psychological space, and model performance is based on the amount of similarity between the item to be categorized and the stored exemplars.
-Similarity between two examples, $s\left(x_i, x_j\right)$, is computed as an inverse exponential function of distance \citep[following][]{shepard1957stimulus,shepard1987toward}:
-
+Similarity between two examples, $s\left(x_i, x_j\right)$, is computed as an inverse exponential function of distance \citep[following][]{attneave1950,shepard1957stimulus,shepard1987toward}:
+\jlanote{attneave was first (though he was inspired by hypotheses made by householder and others, so it may date back even earlier).
+he framed it in terms of sum of log distances, but it's equivalent.
+funnily he said the biggest issue is violations of the triangle inequality via examples attributed to James.
+i can send you the article if you want....}
 \begin{equation}
@@
 
-PACKER's core proposal (as well as its name) was in part inspired by earlier work from category learning literature \citep[see][]{hidaka2011packing}.
-PACKER proposes that generation is constrained by both similarity to members of the target category (the category in which a stimulus is being generated) as well as similarity to members of other categories: the most desirable generation candidates are similar to members of the target category and not similar to members of contrast categories.
+PACKER (as well as its name) was in part inspired by earlier work from the categorization literature \citep{hidaka2011packing}.
+They argued that natural categories "pack" the values of features such that they fill the space with as much distance apart from one another, while maintaining items within the same category close together.
+Inspired by this idea, PACKER proposes that generation is constrained by both similarity to members of the target category (the category in which a stimulus is being generated) as well as similarity to members of other categories: the most desirable generation candidates are similar to members of the target category and not similar to members of contrast categories.
 This is achieved by aggregating similarity across known exemplars differently according to class membership.
@@
 
+\jlanote{Is our parameterization overdetermined?
+In other words, are there multiple parameter values that yield equivalent models?
+It seems like the sum of $\gamma$ and $\phi$ are related to $\theta$.
+Is that right?
+If this is the case, we need to either remove a parameter (if it is completely redundant) or place a constraint (e.g., $\gamma + \phi $ = 1)...}
+
 As noted above, PACKER proposes that new categories should be different from existing categories, and same-category exemplars should be similar to one another.
@@
 
-This constraint, however, only serves to signify one possible account of category generation: the nature of PACKER's account can be drastically altered through changes to the $\phi$ and $\gamma$ parameters.
+However, PAKCKER is only one possible exemplar-based account of category generation within our proposed framework.
+PACKER places specific constraints on the possible parameter values (i.e., $\phi <0$ and $\gamma > 0$).
+Other exemplar-based category generation models with drastically different behavior can be formalized in this framework by changing the constrains on the $\phi$ and $\gamma$ parameters.
 Furthermore, by fitting these parameters to a dataset, one can describe the relative roles of between-class contrast and within-class similarity in generation.
@@
 It is worth noting that PACKER represents just one of many possible models that incorporate contrast in category generation.
-For example, although it may be possible to implement a contrast mechanism within \cite{jern2013probabilistic}'s hierarchical sampling model, as noted above this mechanism would be at odds with a strict interpretation of the hierarchical Bayesian framework.
+For example, although it may be possible to extend \cite{jern2013probabilistic}'s hierarchical sampling model to include contrast, but this mechanism would be at odds with a conjugate hierarchical Bayesian framework.
+In fact, we will discuss integrating our approaches in the General Discussion.
 In contrast, these ideas emerge naturally from the exemplar view: we have not modified any of the core elements of the GCM in defining the PACKER model, we simply aggregated similarity slightly differently.
-\nbcnote{That line about importance sampling can go here.}
+\nbcnote{That line about importance sampling can go here.} \jlanote{I think I'll write a subsection about it/include it in the subsection on future directions.
+I may do a bit of math to show how it would be possible too.}
 
@@
 \label{section:copytweak-packer}
-The PACKER model is in many senses similar to the copy-and-tweak model reported by \cite{jern2013probabilistic}: both models are exemplar-based, and both models generate new items according to their similarity to known members of the target class.
-PACKER diverges from from the copy-and-tweak model only in the inclusion of a contrast mechanism, enabling generation according to dissimilarity to members of opposing categories.
+The PACKER model is some sense similar to the copy-and-tweak model reported by \cite{jern2013probabilistic}: both models are exemplar-based, and both models generate new items according to their similarity to known members of the target class.
+PACKER diverges from the copy-and-tweak model by including a contrast mechanism, enabling generation according to dissimilarity to members of opposing categories.
 By consequence, copy-and-tweak can be realized as a parameterization of the PACKER model that is insensitive to category contrast.
@@
 In this paper, we report simulations using this copy-and-tweak model.
-The model we test is identical to PACKER, under the constraints that $\phi = 0$ and $\gamma = 1$, and is a continuous-dimension adaptation of the model tested by \cite{jern2013probabilistic}.
-Due to their mathematical equivalence, the comparison between PACKER and copy-and-tweak provides a test of the explanatory value of the contrast mechanism: the account provided by copy-and-tweak will only equal that of PACKER if the contrast mechanism does not offer an advantage.
+This model fits within the exemplar-based category generation framework defined above, under the constraints that $\phi = 0$ and $\gamma = 1$, and is a continuous-dimension adaptation of the model tested by \cite{jern2013probabilistic}.
+By formalizing a model family where PACKER and copy-and-tweak are different parameterizations of models within the same framework, the comparison between PACKER and copy-and-tweak provides a test of the explanatory value of the contrast mechanism: the account provided by copy-and-tweak will only equal that of PACKER if the contrast mechanism does not offer an advantage (i.e., if $\phi \neq 0$ significantly improves model fits).
 
@@
 However, a fundamental goal of concept generation is to create something \textit{new} (i.e., different from what is already known).
-The manner in which generated categories differ from known ones is, nonetheless, poorly understood: existing formal approaches do not make strong predictions about how creatively generated concepts should systematically differ from existing ones.
-Above, we introduced a novel, exemplar-based model formalizing the role of contrast in creative generation.
+The manner in which generated categories differ from known ones is, nonetheless, poorly understood: Existing theories do not make strong predictions about how creatively generated concepts should systematically differ from existing ones.
+Above, we introduced a novel, exemplar-based model formalizing the roles of similarity and contrast in creative generation.
 
 In the sections below, we present two experiments demonstrating systematic effects of category contrast on creative generation.
-Our experiments conform strictly to the artificial category learning paradigm from \cite{jern2013probabilistic}: participants are first exposed to a single, experimenter-defined category, and are then asked to generate members of a new category.
+Our experiments are based on \cite{jern2013probabilistic}'s paradigm: participants are first exposed to a single, experimenter-defined category, and are then asked to generate members of a new category.
 We then we report formal simulations comparing PACKER's account of our results to that of the hierarchical sampling and copy-and-tweak models developed by \cite{jern2013probabilistic}.
@@
 This was achieved by training participants on categories possessing qualitatively different distributional structures, as shown in Figure \ref{fig:e1-conditions}.
+
+\jlanote{Can you make a note here about the counterbalancing we did and what that means about the structures?
+E.g., Does it mean cluster really is tested in all four quadrants of the currently depicted axes and both rows and columns are tested?
+I thought that was the case, but I wanted to double check and have you write it in a manner that is definitely what we did.}
 
@@
 Within this paradigm, an effect of category contrast would be realized if participants prefer to generate items in locations that are distant (i.e., perceptually dissimilar) from members of category A.
-However, generation is left unconstrained: For example, participants assigned to the Cluster condition may generate a tightly clustered category in the corner opposite of the experimenter-defined category.
+However, generation is left unconstrained.
+Critically, participants were not asked to generate something different in the prompt.
+For example, participants assigned to the Cluster condition may generate a tightly clustered category in the corner opposite of the experimenter-defined category.
 Alternatively, they may generate a tightly clustered category directly overlapping with the experimenter-defined category.
@@
 
-In this way, our first experiment also serves to replicate the classic finding that generated categories tend to share distributional properties with known categories in the domain \citep[see][]{jern2013probabilistic,ward1994structured}.
+In this way, we can analyze the results of the first experiment as a conceptual replication of the classic finding that generated categories tend to share distributional properties with known categories in the domain \citep[see][]{jern2013probabilistic,ward1994structured}.
 From these results, we can predict that, in each condition, participants should generate categories that are distributionally similar to the experimenter-defined category: In the Cluster condition, generated categories should be tightly clustered.
@@
 In XOR condition, generated categories should be widely distributed across both dimensions, and the two dimensions should be positively correlated.
+However, if contrast plays a role, exemplars in the generated categories of participants in the XOR condition may not be positively correlated (they may even be negatively correlated instead!).
 
 \nbcnote{I think that works as a decent motivation, do you think so?}
+
+\jlanote{Very good}
 
@@
 The assignment of perceptual features (color, size) to axes of the domain space (x, y) was counterbalanced across participants.
+
+\jlanote{was the direction of each dimension also counterbalanced (as in Expt 2)?}
 
@@
 \end{displayquote}
-
-Following generation, participants completed a generalization phase wherein they classified novel examples into the Alpha and Beta categories without feedback.
+\jlanote{were they allowed to create another Alpha or did those need to be unique too? we should be clear...}
+%Following generation, participants completed a generalization phase wherein they classified novel examples into the Alpha and Beta categories without feedback.
 On each trial, a single example was presented, and participants were asked to classify it by clicking buttons labeled ``Alpha'' or ``Beta''.
@@
     \nbcnote{I selected these at random, we can pick more representative samples later.}
+    \jlanote{ok!}
     \end{center}
@@
 That is, members of the generated category tended to be more similar to one another than to members of the experimenter-defined category.
-To formally evaluate this claim, we conducted T-Tests comparing the amount of within- and between- class distance in each condition.
+To formally evaluate this claim, we conducted t-tests comparing the amount of within- and between- class distance in each condition.
 All conditions possessed greater between-class distance: Cluster, $t(63) = 11.43$, $p < 0.001$; Row, $t(60) = 13.16$, $p < 0.001$; and XOR, $t(57) = 3.64$, $p < 0.001$.
@@
 
-The secondary goal of this experiment was to replicate the classic finding that generated categories often possess the same distributional properties as previously-known categories.
+The secondary goal of this experiment was to examine whether we replicate the classic result that generated categories often possess the same distributional properties as previously-known categories.
 For each generated category, we computed the category range along each axis (X, Y), as well as the correlation between features.
@@
 This finding is notable, as it is the opposite of what would be expected, assuming learners are emulating the distributional structure of the experimenter-defined class (which possesses perfectly positively correlated features).
+Further it is expected by our proposal that contrast is a fundamental principle for category generation.
 
@@
     \caption{Box-plots of the distributional statistics from the categories generated in Experiment 1. }
+    \jlanote{can you add a bit of an explanation of what boxplots are and how to interpret them? (short clearly)}
     \nbcnote{Weak spot: row has significantly \textit{less} X-range compared to XOR, which I do not know how to explain. For now I've ignored it.}
+  	\jlanote{It's not really a weak spot for us rather that their results don't fully replicate in my opinion.}
     \label{fig:e1-statsboxes}
@@
 In Experiment 1 we sought to test whether category contrast imposes constraints on creative generation.
-We found strong evidence for effects of category contrast: participants were more likely to generate stimuli that are more distant from (i.e., less similar to) members of a previously-learned category, and members of participant-generated categories tended to be more similar to one another than to members of previously-learned categories.
+We found strong evidence for effects of category contrast: Participants were more likely to generate stimuli that are more distant from (i.e., less similar to) members of a previously-learned category, and members of participant-generated categories tended to be more similar to one another than to members of previously-learned categories.
 We also partially replicated the classic finding that the distributional structure of generated categories reflects that of previously learned categories \citep{jern2013probabilistic,ward1994structured}: members of generated categories were more widely distributed along dimensions which were widely distributed in the experimenter-defined category.
@@
 The category types possess an identical distributional structure (both are tight clusters of examples with a neutral feature correlation), and only differ slightly in their Y-axis position: the `Bottom' category lies in the bottom-center of the space, and the `Middle' category lies in the center.
-Although this manipulation may be viewed as somewhat `weak', the distributional equality of these conditions is key to the design of the experiment: if the structure of previously learned categories were the only influence on the structure of generated categories, we should observe no difference between these two conditions with respect to distributional structure.
+This distributional equality of these conditions is key to the design of the experiment: if the structure of previously learned categories were the only influence on the structure of generated categories, we should observe no difference between these two conditions with respect to distributional structure.
 
@@
 }
+
+\jlanote{they could? wouldn't it depend on the extent that they care about within-category similarity?}
 
@@
     \nbcnote{May want to change condition colors if purple-orange is used as the colormap for the range heat-maps in the simulations section.}
+    \jlanote{sounds good to me.}
     \label{fig:e2-distanceplots}
@@
 These data, shown in Figure \ref{fig:e2-distanceplots}, yield highly similar results.
-Stimuli that are more distant from members of the experimenter=-defined category were more frequently generated, and the categories in each condition tended to possess more between-class than within-class distance: Bottom, $t(60) = 5.5$, $p < 0.001$; Middle, $t(60) = 2.71$, $p = 0.009$.
+Stimuli that are more distant from members of the experimenter-defined category were more frequently generated, and the categories in each condition tended to possess more between-class than within-class distance: Bottom, $t(60) = 5.5$, $p < 0.001$; Middle, $t(60) = 2.71$, $p = 0.009$.
 
@@
 
-\section{Simulations}
+\section{Model-based Analyses}
 
@@
 
+\jlanote{should do a nested model comparison with respect to PACKER v. copy \& tweak. they're nested, so should be doable...}
+
 Table \ref{table:global-model-fits} contains the results of this fitting procedure.
@@
 We don't have `proof' that we can point to for skeptical readers, but that's clearly the issue: many categories were not normally distributed, and the copying of structure really only works out in 2/5 conditions, and even then there's so much variability...}
+\jlanote{yeah, it's a small effect (in terms of $L$). }
 \nbcnote{Should include info on fits for only trials 2-4.}
@@
 The best fitting values lie between -0.6 and -0.8, well beyond the 0 value assumed by the copy-and-tweak model (though note PACKER achieves even better fits when its other parameters are fitted, as in Table \ref{table:global-model-fits}).
+Further, it demonstrates the robustness of the contrast effect.
 

### 2017-03-30 Manuscripts/cog-psych/cogpsych.tex
minor edits to manuscript

@@
 
-\jlanote{The issue isn't prior knowledge actually -- Bayesian models can handle that gracefully -- it's the rich format of participant responses being 2-d sketches}
-
 \cite{jern2013probabilistic} recently showed that creative generation could be studied in a more controlled manner through the well-developed methods of an artificial categorization paradigm \citep[see][for a review]{kurtz2015human}.
@@
 This finding is notable, as it demonstrates that category generation can be studied in a well-known and highly controlled experimental paradigm.
-Note that generated exemplars for the novel category automatically contrasted from previous categories because the participants selected the size and saturation, but the exemplar's hue was set (and fixed) to a novel value.
+
+% % ----- This point is not true -----
+% Note that generated exemplars for the novel category automatically contrasted from previous categories because the participants selected the size and saturation, but the exemplar's hue was set (and fixed) to a novel value.
 As such, hue was ignored in their analyses and may have satisfied any constraints contrast may have on category generation.
-
-\jlanote{It is hierarchical Bayesian, not Bayesian hierarchical.}
 
@@
 
-% Here's what it used to be:
-% Their model represents categories as multivariate normal distributions within the space, and from exposure to category members the model induces a domain-wide representation of the distributional commonalities among the observed categories.
-When generating a new category, the model samples a new distribution from the domain, and thus the distributional properties of observed categories will be preserved within the generates category.
-
-% \jlanote{Yeah, Here's how I would this section should be organized.
-Top is general behavioral findings/phenomena.
-Then subsections for each theory, which first is described in broad theoretical terms and then how that is formalized into a model}
-
-\cite{jern2013probabilistic} additionally tested a "copy-and-tweak" model that broadly resembles the earlier "Path of Least Resistance" account.
+\cite{jern2013probabilistic} additionally tested a ``copy-and-tweak'' model that broadly resembles the earlier ``Path of Least Resistance'' account.
 The core proposal is that participants generate new items by copying stored examples from memory and tweaking them to generate something new.
@@
 Instead, their copy-and-tweak model corresponds to a direct exemplar-similarity approach \citep[e.g.,][]{nosofsky1984choice,nosofsky1986attention}: The model generates new items according to their similarity to known members of the target category.
-This model provided a poor account of the observed data.
-Indeed, the experiments devised by \citeauthor{jern2013probabilistic} were specifically designed to challenge this model.
+This model provided a poor account of the observed data, as the experiments devised by \citeauthor{jern2013probabilistic} were specifically designed to challenge this model.
 However, its application is notable as a first step toward explaining category generation using well-known formal approaches from the categorization literature.
@@
 The copy-and-tweak model tested by \cite{jern2013probabilistic} also claims little about how generated categories should contrast with what is already known.
-In their simulations, the model was only tested on generation after the learner had been exposed to members of the target category, and so the model's ability to generate a new category was not evaluated.
+In their simulations, the model was only tested on generation after the learner had been exposed to members of the target category, and so the model's ability to generate a new category from scratch was not evaluated.
 However, the model's generation is based exclusively on similarity to known members of the target category; when there are no members of the target category, generation is presumably random.
@@
 
-\nbcnote{Maybe note something about how PACKER's basis in the exemplar view means that we're explaining category generation using well known ideas, and that exemplar models are viewed as highly principled.}
-
-\jlanote{Yeah, I think this would make a nice subsection in the general discussion.}
-
 Both PACKER and the GCM simulate categorization under the assumption that learners represent categories as a collection of exemplars, corresponding to the labeled stimuli they have observed.
@@
 Similarity between two examples, $s\left(x_i, x_j\right)$, is computed as an inverse exponential function of distance \citep[following][]{attneave1950,shepard1957stimulus,shepard1987toward}:
-\jlanote{attneave was first (though he was inspired by hypotheses made by householder and others, so it may date back even earlier).
-he framed it in terms of sum of log distances, but it's equivalent.
-funnily he said the biggest issue is violations of the triangle inequality via examples attributed to James.
-i can send you the article if you want....}
+
 \begin{equation}
@@
 PACKER (as well as its name) was in part inspired by earlier work from the categorization literature \citep{hidaka2011packing}.
-They argued that natural categories "pack" the values of features such that they fill the space with as much distance apart from one another, while maintaining items within the same category close together.
+They argued that natural categories "pack" the values of features such that different categories fill the space with distance between one another, while maintaining items within the same category close together.
 Inspired by this idea, PACKER proposes that generation is constrained by both similarity to members of the target category (the category in which a stimulus is being generated) as well as similarity to members of other categories: the most desirable generation candidates are similar to members of the target category and not similar to members of contrast categories.
@@
 
-However, PAKCKER is only one possible exemplar-based account of category generation within our proposed framework.
+However, PACKER is only one possible exemplar-based account of category generation within our proposed framework.
 PACKER places specific constraints on the possible parameter values (i.e., $\phi <0$ and $\gamma > 0$).
-Other exemplar-based category generation models with drastically different behavior can be formalized in this framework by changing the constrains on the $\phi$ and $\gamma$ parameters.
+Other exemplar-based category generation models with drastically different behavior can be formalized in this framework by changing the constraints on the $\phi$ and $\gamma$ parameters.
 Furthermore, by fitting these parameters to a dataset, one can describe the relative roles of between-class contrast and within-class similarity in generation.
@@
 Finally, when $\phi = \gamma < 0$, PACKER represents a ``pure-packing'' approach, placing exemplars in any unoccupied area of the domain.
-Thus, beyond specifying PACKER's core proposal, the $\phi$ and $\gamma$ allow the us to address a wide variety of qualitatively distinct generation strategies.
-
-\nbcnote{That last paragraph was added since your last read, joe!}
+Thus, beyond specifying PACKER's core proposal, the $\phi$ and $\gamma$ parameters allow the us to address a wide variety of qualitatively distinct generation strategies.
 
@@
 In contrast, these ideas emerge naturally from the exemplar view: we have not modified any of the core elements of the GCM in defining the PACKER model, we simply aggregated similarity slightly differently.
-\nbcnote{That line about importance sampling can go here.} \jlanote{I think I'll write a subsection about it/include it in the subsection on future directions.
-I may do a bit of math to show how it would be possible too.}
 
@@
 This model fits within the exemplar-based category generation framework defined above, under the constraints that $\phi = 0$ and $\gamma = 1$, and is a continuous-dimension adaptation of the model tested by \cite{jern2013probabilistic}.
-By formalizing a model family where PACKER and copy-and-tweak are different parameterizations of models within the same framework, the comparison between PACKER and copy-and-tweak provides a test of the explanatory value of the contrast mechanism: the account provided by copy-and-tweak will only equal that of PACKER if the contrast mechanism does not offer an advantage (i.e., if $\phi \neq 0$ significantly improves model fits).
+By formalizing a model family where PACKER and copy-and-tweak are different parameterizations of models within the same framework, the comparison between PACKER and copy-and-tweak provides a test of the explanatory value of the contrast mechanism: the account provided by copy-and-tweak will only equal that of PACKER if the contrast mechanism does not offer an advantage (i.e., if $\phi < 0$ significantly improves model fits).
 
@@
 
+\nbcnote{Make the point that we’re explaining category generation using well known ideas, and that exemplar models are viewed as highly principled.}
+
+\nbcnote{Discuss combining hierarchical sampling and contrast approaches.}
+
+\nbcnote{Joe says he'll put importance sampling in future directions.}
+
 \clearpage

### 2017-04-04 Manuscripts/model-descriptions/packer/packer.tex
rewrote packer description as a three-parameter model

@@
 
-First, let's assume exemplars $x$ are represented as points within an $k$-dimensional space.
-When prompted to generate the first member of a new class, participants seek out stimuli that have low similarity to all assigned examples.
-
-Distance between exemplars $X_i$ and $X_j$ is computed using the city-block metric:
+The similarity between exemplars $x_i$ and $x_j$ is computed using Shepard's law:
 
 \begin{equation}
-  d(x_i,x_j) = \sum_{k}{|x_{ik} - x_{jk}|}w_k
+s\left(x_i,x_j\right) = \exp \left\{ -c \left[\sum_{k}{ w_k \left| x_{ik} - x_{jk} \right|^r }\right]^{1/r} \right\}
+\label{eq:similarity}
 \end{equation}
 
-Where $w_k$ is a vector of feature weights, corresponding to the importance of each feature in the similarity computation.
-The attention weights are of interest in explaining individual differences in generation strategy, and they may have some relation to the relative dimensional variance of existing categories, but for the general case they can be uniform.
-
-The similarity between exemplars $x_i$ and $x_j$ is then computed using Shepard's law:
+When prompted to make a generation decision, participants are thought to consider both similarity to examples from other categories as well as similarity to examples in the target category.
+More formally, the aggregated similarity $a$ between candidate $y$ and the model's stored exemplars $x$ can be computed as:
 
 \begin{equation}
-  s(x_i,x_j) = exp( {-c d(x_i,x_j)} )
+  a(y, x) = \sum_j{f(x_j) s(y, x_j)}
 \end{equation}
 
-Where $c$ is a free parameter setting the specificity of examples.
-Larger values of $c$ produce more \textit{specific} exemplars -- only stimuli very close to the exemplar will be viewed as similar.
-Due to the nature of the exponential function, similarity values will lie in range $0-1$, with a value of $1$ indicating an exact match between $x_i$ and $x_j$.
+Where $f(x_j)$ is a function specifying each stored example's degree of contribution toward generation.
+Although $f(x_j)$ may be set arbitrarily, in PACKER it is set according to class assignment.
+For known members of the target category, $f(x_j) = \gamma$.
+For members of contrast categories, $f(x_j) = -1 + \gamma$.
+$\gamma$ is thus a free parameter ($0 \leq \gamma \leq 1$) controlling the trade-off between within-class similarity and between-class dissimilarity: $\gamma = 1$ produces exclusive consideration of same-category members, and $\gamma = 0$ produces exclusive consideration of opposite-category members.
+When $\gamma = 0.5$, the similarity to contrast categories is effectively subtracted from the similarity to the target category.
 
-When prompted to make a generation decision, participants are thought to consider both similarity to examples from other categories as well as similarity to examples in the target category.
-More formally, the summed similarity $ss$ between item $x_i$ and the model's stored exemplars $x_j$ can be computed as:
+The probability that a given item $y$ will be generated given the model's memory $x$ is computed using relative summed similarity values across all generation candidates $y_i$:
 
 \begin{equation}
-  ss(x_i, x_j) = \sum_j{f_j s(x_i, x_j)}
-\end{equation}
-
-Where $f_j$ is a function specifying each stored example's degree of contribution toward generation.
-Although $f_j$ may be set arbitrarily, for our purposes it is relevant to use the class assignment.
-For known members of contrast categories, $f_j = -\phi$.
-For known members of the target category, $f_j = \gamma$.
-$\phi$ and $\gamma$ are free parameters ($\geq 0$) that control the contribution of same- and opposite- category members to generation.
-Larger values for either parameter produce greater consideration of either constraint.
-Values of 0 for either parameter result in no effect for that type of similarity.
-
-This arrangement has a number of desirable properties.
-When $\phi = \gamma$, the similarity to contrast categories is effectively subtracted from the similarity to the target category -- negative values for $ss(x, x_j)$ thus indicate more contrast category similarity, and positive values indicate more target category similarity.
-Because $\phi$ and $\gamma$ are parameterized, this trade off can be managed explicitly.
-
-The probability that a given item $x$ will be generated given the model's memory $x_j$ is computed using relative summed similarity values across all generation candidates $x_i$:
-
-\begin{equation}
-p(x|x_j) = \dfrac
-    { exp( { \theta ss(x, x_j) } ) }
-    { \sum_i{ exp( { \theta ss(x_i, x_j) } ) } }
+p(y) = \dfrac
+{ \exp  \left \{ \theta \cdot a \left( y, x \right) \right \} } 
+{ \sum_i{ \exp  \left \{ \theta \cdot a \left( y_i, x \right) \right\}  } }
+\label{eq:packer-choice}
 \end{equation}
@@
 Where $\theta$ ($\geq 0$) is a free parameter controlling overall response determinism.
-Larger values of $\theta$ produce greater adherence to the relative probabilities of the choices (items with larger probabilities become more likely), and the special case of $\theta = 0$ produces random responding.
-Participants are not allowed to generate the same example twice, thus $p_x$ for these items is manually set to 0.
 

### 2017-04-04 Modules/tests/test_script.py

@@ -2,6 +2,12 @@ import matplotlib.pyplot as plt
 import numpy as np
 np.set_printoptions(precision = 2, linewidth = 120)
 
+
+
+
+
+
+
 execfile('Imports.py') 
 from Modules.Classes import CopyTweak, Packer, ConjugateJK13
 import Modules.Funcs as funcs
@@ -22,9 +28,4 @@ categories = [
 	stimuli[[8,65],:],
 	]
 
-M = ConjugateJK13(categories,ConjugateJK13.rvs())
-ps = M.get_generation_ps(stimuli,0)
-
-
-
-print ConjugateJK13.rvs()
+M = Packer(categories,Packer.rvs())

### 2017-04-04 Manuscripts/cog-psych/cogpsych.tex
update manuscript to reflect new packer params

@@
 
-%The creation of novel concepts and ideas is a highly intriguing -- yet infrequently studied -- topic of research in human cognition.
-The creative use of conceptual knowledge is, for example, a core element of scientific investigation, wherein the generation of new ideas is critical to designing experiments and explaining observations.
-However, due to its complexity, little is presently known about the cognitive processes underlying the generation of new concepts.
-
 Towards the aim of providing a formal account of the cognitive mechanisms involved in creativity and innovation, we explore a simplified form of creative cognition that can be studied using standard behavioral methodology and is amenable to formal modeling: category generation.
@@
 
-% Joe says he'll do a pass on this language later:
 Although it is evident that people are \textit{capable} of creating new concepts and categories, it is not entirely clear how new concepts are systematically made different from what is already known.
@@
 
-% \section{Formalizing category generation}
-% \jlanote{this should come before the studies as formalizations of the different theories (some details may be delegated to Appendices).
-I like trying to integrate them with the verbal description of the theories rather than having a separate computational modeling section.
-But the latter works for sure and I'm not sure if the former will flow well...}
+\subsubsection{Contrast Effects in Prior Work}
+
+Although existing accounts of creative generation broadly overlook the role of category contrast in determining what is novel versus familiar, effects of category contrast are undoubtedly ubiquitous within the published data.
+In each study, participants were \textit{successful} in creating something new: the question is about \textit{how} they were able to do so.
+To provide some concrete evidence of these effects within existing datasets, we conducted an analysis of the data from Experiment 3 of the report by \cite{jern2013probabilistic}.
+
+\nbcnote{Report analyses on the \cite{jern2013probabilistic} data here...}
 
@@
 PACKER (as well as its name) was in part inspired by earlier work from the categorization literature \citep{hidaka2011packing}.
-They argued that natural categories "pack" the values of features such that different categories fill the space with distance between one another, while maintaining items within the same category close together.
+They argued that natural categories ``pack'' the values of features such that different categories fill the space with distance between one another, while maintaining items within the same category close together.
 Inspired by this idea, PACKER proposes that generation is constrained by both similarity to members of the target category (the category in which a stimulus is being generated) as well as similarity to members of other categories: the most desirable generation candidates are similar to members of the target category and not similar to members of contrast categories.
@@
 where $f(x_j)$ is a function specifying each exemplar's contribution to generation.
-PACKER sets $f(x_j)$ depending on exemplar $x_j$'s category membership: $f(x_j) = \phi$ if $x_j$ is a member of a contrast category, and $f(x_j) = \gamma$ if $x_j$ is a member of the target category.
-$\phi$ and $\gamma$ are free parameters ($-\infty \leq \phi, \gamma \leq \infty$) controlling the contribution of contrast- and target-category similarity, respectively.
-Larger absolute values result in greater consideration of those exemplars, with values of 0 eliminating their effect.
-A negative value for $f(x_j)$ produces a `repelling' effect (exemplars are less likely to be generated nearby $x_j$).
-Conversely, a positive value for $f(x_j)$ produces an `attracting' effect (exemplars are more likely to be generated nearby $x_j$).
+A negative value for $f(x_j)$ produces a `repelling' effect (items are less likely to be generated nearby $x_j$), and a positive value produces an `attracting' effect (items are more likely to be generated nearby $x_j$).
+When $f(x_j)=0$, the exemplar does not contribute to generation.
+
+PACKER sets $f(x_j)$ depending on exemplar $x_j$'s category membership: $f(x_j) = \gamma$ if $x_j$ is a member of the target category, and $f(x_j) = \gamma - 1$ if $x_j$ is a member of a contrast category.
+$\gamma$ is thus a free parameter ($0 \leq \gamma \leq 1$) controlling the trade-off between within- and between-class similarity.
+For example, when $\gamma = 0.5$, $f(x_j) = 0.5$ for members of the target category and $f(x_j) = -0.5$ for members of other categories; thus, the model is likely to generate items that are similar to members of the target category but are not similar to members of other categories.
+In this way, $\gamma = 1$ produces exclusive consideration of target-category members, and $\gamma = 0$ produces exclusive consideration of opposite-category members.
+The $\gamma$ parameter thus specifies a wide breadth of possible approaches; by fitting it to a dataset, one can describe the relative roles of between-class contrast and within-class similarity in generation.
+See Figure \ref{fig:packer-examples} for an illustration of how the $\gamma$ parameter can be used.
 
@@
 		\caption{PACKER generation of a category `B' example, following exposure to one member of category `A' and one member of category `B'.
-Predictions are shown for three different parameterizations: \textit{(a)} Predictions based on contrast similarity only.
+Predictions are shown for three different parameterizations (differing only in $\gamma$): \textit{(a)} Predictions based on contrast similarity only.
 \textit{(b)} Predictions based on target similarity only.
@@
 
-\jlanote{Is our parameterization overdetermined?
-In other words, are there multiple parameter values that yield equivalent models?
-It seems like the sum of $\gamma$ and $\phi$ are related to $\theta$.
-Is that right?
-If this is the case, we need to either remove a parameter (if it is completely redundant) or place a constraint (e.g., $\gamma + \phi $ = 1)...}
-
-As noted above, PACKER proposes that new categories should be different from existing categories, and same-category exemplars should be similar to one another.
-This is realized when $\phi < 0$, and $\gamma > 0$.
-Negative $\phi$ values encourage $y$ to be distant from contrast categories (as similarity to contrast category exemplars are subtracted during aggregation).
-Positive $\gamma$ values encourage $y$ to be close to other exemplars of the target category.
-When $|\phi| = \gamma$, the repulsion effect from contrast categories is equal to the attraction effect to the target category.
-See Figure \ref{fig:packer-examples} for an illustration of how these parameters are used.
-
 However, PACKER is only one possible exemplar-based account of category generation within our proposed framework.
-PACKER places specific constraints on the possible parameter values (i.e., $\phi <0$ and $\gamma > 0$).
-Other exemplar-based category generation models with drastically different behavior can be formalized in this framework by changing the constraints on the $\phi$ and $\gamma$ parameters.
-Furthermore, by fitting these parameters to a dataset, one can describe the relative roles of between-class contrast and within-class similarity in generation.
-For example, as will be discussed in more detail below, PACKER is formally equivalent to the copy-and-tweak model proposed by \cite{jern2013probabilistic} when $\phi = 0$ and $\gamma = 1$.
-Likewise, when $\phi < 0$ and $\gamma = 0$, PACKER represents a contrast-only generation mode, relying exclusively on contrast when generating new categories.
-Finally, when $\phi = \gamma < 0$, PACKER represents a ``pure-packing'' approach, placing exemplars in any unoccupied area of the domain.
-Thus, beyond specifying PACKER's core proposal, the $\phi$ and $\gamma$ parameters allow the us to address a wide variety of qualitatively distinct generation strategies.
+PACKER places specific constraints on the possible values of $f(x_j)$, but other exemplar-based category generation models with drastically different behavior can be formalized in this framework by imposing alternative constraints.
+For example, as will be discussed in more detail below, PACKER is formally equivalent to the copy-and-tweak model proposed by \cite{jern2013probabilistic} when $\gamma = 1$.
+Likewise, when $\gamma = 0$, PACKER represents a contrast-only generation mode, relying exclusively on contrast when generating new categories.
+Finally, when $f(x_j) = -1$ for all $x_j$ (regardless of class membership), a ``pure-packing'' approach is yielded, generating items in any unoccupied area of the domain.
+Thus, the proposed framework may be used to describe a wide variety of qualitatively distinct generation strategies.
 
@@
 By consequence, copy-and-tweak can be realized as a parameterization of the PACKER model that is insensitive to category contrast.
-Specifically, when $\phi = 0$ and $\gamma = 1$ (see Figure \ref{fig:packer-examples}, panel B), PACKER is not influenced by similarity to members of opposite categories, and is mathematically equivalent to a copy-and-tweak approach.
+Specifically, when $\gamma = 1$ (see Figure \ref{fig:packer-examples}, panel B), $f(x_j) = 0$ for $x_j$ belonging to contrast categories; thus, PACKER is not influenced by these items, and is mathematically equivalent to a copy-and-tweak approach.
 
 In this paper, we report simulations using this copy-and-tweak model.
-This model fits within the exemplar-based category generation framework defined above, under the constraints that $\phi = 0$ and $\gamma = 1$, and is a continuous-dimension adaptation of the model tested by \cite{jern2013probabilistic}.
-By formalizing a model family where PACKER and copy-and-tweak are different parameterizations of models within the same framework, the comparison between PACKER and copy-and-tweak provides a test of the explanatory value of the contrast mechanism: the account provided by copy-and-tweak will only equal that of PACKER if the contrast mechanism does not offer an advantage (i.e., if $\phi < 0$ significantly improves model fits).
+This model fits within the exemplar-based category generation framework defined above, under the constraint that $\gamma = 1$, and is a continuous-dimension adaptation of the model tested by \cite{jern2013probabilistic}.
+By formalizing a model family where PACKER and copy-and-tweak are different parameterizations of models within the same framework, the comparison between PACKER and copy-and-tweak provides a test of the explanatory value of the contrast mechanism: the account provided by copy-and-tweak will only equal that of PACKER if the contrast mechanism does not offer an advantage (i.e., if $\gamma < 1$ significantly improves model fits).
 
@@
 Thus, its members might all be dark in color but would vary in size.
-Finally, in the `XOR' type, the experimenter-defined category consists of two clusters separated in opposite corners of the space category, conforming to the exclusive-or logical structure (e.g., members are small and dark or large and light).
+Finally, in the `XOR' type, the experimenter-defined category consists of two clusters separated in opposite corners of the space, conforming to the exclusive-or logical structure (e.g., members are small and dark or large and light).
 
@@
 In XOR condition, generated categories should be widely distributed across both dimensions, and the two dimensions should be positively correlated.
-However, if contrast plays a role, exemplars in the generated categories of participants in the XOR condition may not be positively correlated (they may even be negatively correlated instead!).
 
-\nbcnote{I think that works as a decent motivation, do you think so?}
-
-\jlanote{Very good}
-
-% \jlanote{Is it possible to get the main individual differences we find in this study from different parameterizations of PACKER?
-If so, we can motivate it from there.
-Alternatively, we may want to switch the order of the two studies and motivate the experiment from the next one as a .
-I like switching the order of the experiments less because I feel like Experiment 2 nails contrast down and Experiment 1 is more showing how it plays a role in more complex category types}
+Interestingly, the XOR condition also offers a dissociation between the roles of category contrast and the emulation of distributional structure: widely-distributed, positively-correlated categories would need to lie along the positive diagonal of the space (that is the only place they ``fit''), which is already occupied by the experimenter-defined category.
+Thus, if contrast plays a role, exemplars in the generated categories of participants in the XOR condition may not be positively correlated--they may be negatively correlated instead.
 
@@
 In this section, we report the results of simulations with formal models aimed at explaining our observations.
-Specifically, we present simulations from the PACKER model, as well as a `copy-and-tweak' model (discussed in section \ref{section:copytweak-packer}), defined as a variant of PACKER with the $\phi$ and $\gamma$ parameters held constant.
+Specifically, we present simulations from the PACKER model, as well as a `copy-and-tweak' model (discussed in section \ref{section:copytweak-packer}), defined as a variant of PACKER with the $\gamma$ parameter held constant.
 The comparison of these two models serves to highlight the explanatory role of contrast within PACKER's framework: if contrast affords little explanatory advantage, then the two accounts should produce an equally strong account.
@@
 To obtain a global measure of the quality of each model's account, we fitted the parameters of each model to our entire dataset (Experiments 1 and 2 combined), using a hill-climbing algorithm which maximized the log-likelihood of the model's predictions of the observed responses (1220 responses from 305 total participants).
-We fitted four parameters in the PACKER model ($c$, $\phi$, $\gamma$, and $\theta$; see Section \ref{section:PACKER-definition}), as well as four in the hierarchical sampling model ($\kappa$, $\rho$, $\nu$, and $\theta$; see Appendix \ref{ap:hsampling-definition}).
-We fitted only two parameters for the copy-and-tweak model ($c$, and $\theta$), as $\phi$ and $\gamma$ are held constant.
+We fitted three parameters in the PACKER model ($c$, $\gamma$, and $\theta$; see Section \ref{section:PACKER-definition}), as well as four in the hierarchical sampling model ($\kappa$, $\rho$, $\nu$, and $\theta$; see Appendix \ref{ap:hsampling-definition}).
+We fitted only two parameters for the copy-and-tweak model ($c$, and $\theta$), as $\gamma$ is held constant ($\gamma = 1$).
 Note that each model possesses a $\theta$ parameter fulfilling the same role (response determinism).
@@
  \hline
- $AIC = 9097$       & $AIC = 9842$          & $AIC = 10080$  \\ 
+ $AIC = 9095$       & $AIC = 9842$          & $AIC = 10080$     \\ 
  $L = -4545$        & $L = -4919$           & $L = -5036$       \\ 
- $c=0.481$          & $c=3.189$             & $\kappa<0.001$\\
- $\phi=-0.945$      & $\phi=0$ (fixed)      & $\nu=1.001$ \\ 
- $\gamma=1.044$     & $\gamma=1$ (fixed)    & $\rho=5.313$  \\ 
- $\theta=3.354$     & $\theta=2.968$        & $\theta = 7.949$  \\ 
+ $c = 0.482$        & $c = 3.187$           & $\kappa < 0.001$  \\
+ $\gamma = 0.525$   & $\gamma = 1$ (fixed)  & $\nu = 1.001$     \\ 
+ $\theta = 6.664$   & $\theta = 2.969$      & $\rho = 0.351$    \\ 
+                    &                       & $\theta = 7.951$  \\ 
 \end{tabular}
@@
 While PACKER's advantage may tentatively be attributed to the model's sensitivity to category contrast (this will be explored in detail below), the advantage shown by copy-and-tweak over the hierarchical sampling model may be attributed to its exemplar-based representation, as opposed to the prototype-based representation assumed in the hierarchical sampling model.
-\nbcnote{That last point is a little squishy.
-We don't have `proof' that we can point to for skeptical readers, but that's clearly the issue: many categories were not normally distributed, and the copying of structure really only works out in 2/5 conditions, and even then there's so much variability...}
-\jlanote{yeah, it's a small effect (in terms of $L$). }
+As observed in Figures \ref{fig:e1-samples} and \ref{fig:e2-samples}, the generated categories we observed were often widely distributed, with no items near the category prototype.
+This aspect of the data is inconsistent with the multivariate normal distributions used to represent categories in the \cite{jern2013probabilistic} model, but can be handled easily using an exemplar-based approach.
+
 \nbcnote{Should include info on fits for only trials 2-4.}
 
-Through comparison with the copy-and-tweak model, Figure \ref{fig:packer-loglike} more clearly demonstrates the explanatory gains yielded by PACKER's sensitivity to between-class contrast.
-Plotted is PACKER's log-likelihood as a function of the model's between-category parameter, $\phi$.
-The models other parameters ($c$, $\gamma$, $\theta$) were set according to copy-and-tweak's best fits from Table \ref{table:global-model-fits}, and thus when $\phi=0$, the models are equivalent.
-The figure clearly shows a ``sweet spot'':  a convex region in which PACKER achieves superior fits as a result of changes to the $\phi$ parameter.
-The best fitting values lie between -0.6 and -0.8, well beyond the 0 value assumed by the copy-and-tweak model (though note PACKER achieves even better fits when its other parameters are fitted, as in Table \ref{table:global-model-fits}).
-Further, it demonstrates the robustness of the contrast effect.
+Through comparison with the copy-and-tweak model, Figure \ref{fig:packer-loglike} more clearly demonstrates the explanatory gains yielded by PACKER's sensitivity to category contrast.
+Plotted is PACKER's log-likelihood as a function of $\gamma$ parameter.
+The model's other parameters ($c$, $\theta$) were set according to copy-and-tweak's best fits from Table \ref{table:global-model-fits}, and thus when $\gamma=1$, the models are equivalent.
+The figure clearly shows a ``sweet spot'':  a convex region in which PACKER achieves superior fits as a result of changes to $\gamma$.
+The best fitting values lie well below the value of 1 assumed by the copy-and-tweak model, which demonstrates the robustness of the contrast effect (though note PACKER achieves even better fits when its other parameters are fitted, as in Table \ref{table:global-model-fits}).
+Notably, however, the copy-and-tweak parameterization ($\gamma=1$) performs better than the `contrast-only' parameterization of the model ($\gamma = 0$).
+Thus, it is evident that the data is better explained when both within-class similarity and category contrast is considered.
 
@@
     \inputpgf{figs/}{packer-loglike.pgf}
-    \caption{PACKER's fit as a function of the between-class contrast parameter ($\phi$).
-To facilitate comparison, PACKER's other parameters ($c$, $\gamma$, $\theta$) were set to the best fitting values obtained for copy-and-tweak in Table \ref{table:global-model-fits}.
-Thus, in this figure PACKER is equivalent to copy-and-tweak when $\phi = 0$.}
+    \caption{PACKER's fit as a function of sensitivity to between-class contrast (using the $\gamma$ parameter).
+To facilitate comparison, PACKER's other parameters ($c$, $\theta$) were set to the best fitting values obtained for copy-and-tweak in Table \ref{table:global-model-fits}.
+Thus, in this figure PACKER is equivalent to copy-and-tweak when $\gamma = 1$.
+}
     \label{fig:packer-loglike}

### 2017-04-04 Experiments/cogpsych-code/global-model-fits.py

@@ -13,8 +13,7 @@ from Modules.Classes import ConjugateJK13
 with open("pickles/all_data_e1_e2.p", "rb" ) as f:
 	trials = pickle.load( f )
 
-print trials
-llll
+
 # options for the optimization routine
 options = dict(
 	method = 'Nelder-Mead',
@@ -22,6 +21,16 @@ options = dict(
 	tol = 0.01,
 ) 
 
+results = dict()
 for model_obj in [CopyTweak, Packer, ConjugateJK13]:
 
 	res = Simulation.hillclimber(model_obj, trials, options)
+	X = model_obj.params2dict(model_obj.clipper(res.x))
+	results[model_obj.model] = X
+
+# save final result in pickle
+with open('pickles/best_params_e1_e2.p','wb') as f:
+	pickle.dump(results, f)
+
+for k,v in results.items():
+	print k, v

### 2017-04-04 Experiments/cogpsych-code/plot.packer.examples.py

@@ -15,11 +15,11 @@ B = np.array([[ 0.25,  0.25]])
 cats = [A,B]
 
 # params for PACKER
-c, phi = 1.0, 3.0
+c, theta = 1.0, 3.
 
-pos =     [c,  0.0, 1.0, phi]
-neg =     [c, -1.0, 0.0, phi]
-pos_neg = [c, -1.0, 1.0, phi]
+pos =     [c, 1.0, theta]
+neg =     [c, 0.0, theta]
+pos_neg = [c, 0.5, theta]
 
 prob_spaces = {
     'Target Influence': Packer(cats,pos),
@@ -43,8 +43,14 @@ for i, k in enumerate(['Contrast Influence', 'Target Influence', 'Combination'])
     title = prefix[i] + ' ' + k
     h.set_title(title, fontsize = 11)
 
-    xlab  = '$\{ \phi = ' + str(int(m.between)) + '$, '
-    xlab += '$\gamma = ' + str(int(m.within)) + '\}$'
+    tradeoffstr = str(m.tradeoff)
+    if tradeoffstr in ['0.0', '1.0']:
+        tradeoffstr = tradeoffstr[0]
+    xlab  = '$\{'
+    xlab += 'c = ' + str(int(c)) + ','
+    xlab += '\gamma = ' + tradeoffstr  + ','
+    xlab += '\\theta = ' + str(int(theta))
+    xlab += '\}$'
     h.set_xlabel(xlab)
 
 
@@ -57,7 +63,7 @@ cbar.tick_params(length = 0)
 
 
 fname = 'packer-examples'
-# f.savefig('packer-examples.pdf', bbox_inches='tight', transparent=False)
+f.savefig('packer-examples.png', bbox_inches='tight', transparent=False)
 
 path = '../../Manuscripts/cog-psych/figs/packer-examples.pgf'
 funcs.save_as_pgf(f, path)

### 2017-04-04 Modules/Classes/Packer.py

@@ -4,9 +4,10 @@ import numpy as np
 import Modules.Funcs as Funcs
 from Model import Exemplar
 
+
 class Packer(Exemplar):
 	"""
-	Exemplar similarity implementation of the packing model.
+	The three-parameter PACKER Model
 	"""
 
 	model = 'PACKER'
@@ -45,5 +46,44 @@ class Packer(Exemplar):
 		aggregate[known_members] = np.nan
 
 		ps = Funcs.softmax(aggregate, theta = self.determinism)
+		return ps
+
+
+
+class CopyTweak(Exemplar):
+	"""
+		Continuous implementation of the copy-and-tweak model.
+	"""
+
+	model = 'Copy and Tweak'
+	parameter_names = ['specificity', 'determinism']
+	parameter_rules = dict(
+			specificity = dict(min = 1e-10),
+			determinism = dict(min = 0.0),
+		)
+
+	@staticmethod
+	def _make_rvs(fmt = dict):
+		""" Return random parameters """
+		return [np.random.uniform(0.1, 6.0), # specificity
+						np.random.uniform(0.1, 6.0)] # determinism
+
+	def get_generation_ps(self, stimuli, category):
+
+		# return uniform probabilities if there are no exemplars
+		target_is_populated = any(self.assignments == category)
+		if not target_is_populated:
+			ncandidates = stimuli.shape[0]
+			return np.ones(ncandidates) / float(ncandidates)
+
+		# get pairwise similarities
+		similarity = self._sum_similarity(stimuli, self.categories[category])
+
+		# NaN out known members
+		known_members = Funcs.intersect2d(stimuli, self.categories[category])
+		similarity[known_members] = np.nan
+
+		# get generation probabilities given each source
+		ps = Funcs.softmax(similarity, theta = self.determinism)
 		return ps
 

### 2017-04-04 Modules/Classes/Simulation.py

@@ -156,8 +156,8 @@ def hillclimber(model_obj, trials_obj, options, inits = None):
 	print '\tIterations = ' + str(res.nit)
 	print '\tMessage = ' + str(res.message)
 
-	X = model_obj.clipper(res.x)
-	for k, v in zip(model_obj.parameter_names, X):
+	X = model_obj.params2dict(model_obj.clipper(res.x))
+	for k, v in X.items():
 		print '\t' + k + ' = ' + str(v) + ','
 	print '\tLogLike = ' + str(res.fun)
 

### 2017-04-04 Modules/Classes/__init__.py

@@ -1,4 +1,3 @@
 import Simulation
-from CopyTweak import CopyTweak
-from Packer import Packer
+from Packer import Packer, CopyTweak
 from ConjugateJK13 import ConjugateJK13

### 2017-04-07 Manuscripts/cog-psych/cogpsych.tex
jk13 distance fig drafted

@@
 
-\jlanote{Can you make a note here about the counterbalancing we did and what that means about the structures?
-E.g., Does it mean cluster really is tested in all four quadrants of the currently depicted axes and both rows and columns are tested?
-I thought that was the case, but I wanted to double check and have you write it in a manner that is definitely what we did.}
-
 Panels B-D of Figure \ref{fig:e1-conditions} show the locations of exemplars belonging to the experimenter-defined categories (`A', or `Alpha') that participants were assigned to learn about prior to generating a new category.
@@
 Finally, in the `XOR' type, the experimenter-defined category consists of two clusters separated in opposite corners of the space, conforming to the exclusive-or logical structure (e.g., members are small and dark or large and light).
+
+It should be noted that in our experiments the assignment between the perceptual and conceptual dimensions (e.g., $X \rightarrow Size$, $Y \rightarrow Color$), as well as the direction of variation along each dimension (e.g., $dark \rightarrow light$ or $light \rightarrow dark$) was counterbalanced across participants.
+The category types in Figure \ref{fig:e1-conditions} are plotted in a conceptual space, rather than a perceptual space: each category type may have a different physical instantiation according to the counterbalance assignment.
+For example, the Cluster type  may be large and dark in color, or it may be small and light in color, depending on the assignment and direction of the dimensions.
+For this reason, below we will discuss generation within a conceptual space, rather than a physically instantiated one.
+Importantly, however, the conceptual organization of the category types remains constant across the counterbalanced assignments.
 
@@
 These stimuli are slight variants of those used by \cite{conaway2016similar}, see Figure \ref{fig:e1-conditions}.
-The assignment of perceptual features (color, size) to axes of the domain space (x, y) was counterbalanced across participants.
-
-\jlanote{was the direction of each dimension also counterbalanced (as in Expt 2)?}
+The assignment of perceptual features (color, size) to axes of the domain space (x, y), as well as the direction of variation along each axis (e.g., $dark \rightarrow light$ or $light \rightarrow dark$) was counterbalanced across participants.
 
@@
 An on-screen preview of the example updated whenever one of the features was changed.
-Participants could generate any example along an evenly-spaced 9x9 grid, except for any previously generated Beta exemplars.
+Participants could generate any example along an evenly-spaced 9x9 grid (including members of the Alpha category), except for any previously generated Beta exemplars.
 Neither the members of the Alpha category nor the previously generated Beta examples were visible during generation.
@@
 \end{displayquote}
-\jlanote{were they allowed to create another Alpha or did those need to be unique too? we should be clear...}
-%Following generation, participants completed a generalization phase wherein they classified novel examples into the Alpha and Beta categories without feedback.
-On each trial, a single example was presented, and participants were asked to classify it by clicking buttons labeled ``Alpha'' or ``Beta''.
-Participants classified a total of 81 items sampled along a 9x9 grid, including the members of the Alpha and Beta categories (randomly intermixed).
-These data were, however, collected to address a separate set of questions, and we do not discuss them in this paper.
 
@@
 The categories generated in the Cluster condition were less widely distributed along the X-axis compared to Row, $t(123) = 5.61$, $p < 0.001$, and XOR, $t(120) = 2.68$, $p = 0.008$.
-Likewise, categories generated in the Row condition had less Y-axis range compared to Cluster, $t(123) = 4.57$, $p < 0.001$ and XOR $t(117) = 9.26$, $p < 0.001$, and categories from the Cluster condition had less Y-axis range compared to XOR, $t(120) = 3.95$, $p < 0.001$.
+Categories generated in the XOR condition were also less widely distributed along the X-axis compared to Row, $t(117) = 2.56$, $p = 0.012$.
+This latter effect was not expected; however, the key finding is that categories from the Cluster condition tended to be more tightly distributed along the X-axis.
+
+Likewise, categories generated in the Row condition had less Y-axis range compared to Cluster, $t(123) = 4.57$, $p < 0.001$ and XOR, $t(117) = 9.26$, $p < 0.001$, and categories from the Cluster condition had less Y-axis range compared to XOR, $t(120) = 3.95$, $p < 0.001$.
 However, whereas the correlations in the Cluster and Row conditions were not systematically positive or negative ($ps > 0.1$), the generated categories in the XOR condition tended to possess \textit{negatively} correlated dimensions, $t(57) = 2.04$, $p = 0.046$.
 This finding is notable, as it is the opposite of what would be expected, assuming learners are emulating the distributional structure of the experimenter-defined class (which possesses perfectly positively correlated features).
-Further it is expected by our proposal that contrast is a fundamental principle for category generation.
+Further, it is expected by our proposal that contrast is a fundamental principle for category generation.
 
@@
     \inputpgf{figs/}{e1-statsboxes.pgf}
-    \caption{Box-plots of the distributional statistics from the categories generated in Experiment 1. }
-    \jlanote{can you add a bit of an explanation of what boxplots are and how to interpret them? (short clearly)}
-    \nbcnote{Weak spot: row has significantly \textit{less} X-range compared to XOR, which I do not know how to explain. For now I've ignored it.}
-  	\jlanote{It's not really a weak spot for us rather that their results don't fully replicate in my opinion.}
+    \caption{Box-plots of the distributional statistics from the categories generated in Experiment 1.
+Boxes depict the median and quartiles of each condition, with whiskers placed at 1.5 IQR.
+All points outside this region are marked individually.}
     \label{fig:e1-statsboxes}
@@
 The category types possess an identical distributional structure (both are tight clusters of examples with a neutral feature correlation), and only differ slightly in their Y-axis position: the `Bottom' category lies in the bottom-center of the space, and the `Middle' category lies in the center.
-This distributional equality of these conditions is key to the design of the experiment: if the structure of previously learned categories were the only influence on the structure of generated categories, we should observe no difference between these two conditions with respect to distributional structure.
+The distributional equality of these conditions is key to the design of the experiment: if the structure of previously learned categories were the only influence on the structure of generated categories, we should observe no difference between these two conditions with respect to distributional structure.
 
@@
 Accordingly, participants should be more likely to utilize both of these areas.
-
-\nbcnote{Weak point: this prediction is squishy-- why couldn't middle participants generate items above \textit{or} below?
-When we initially developed this, we were working with the pure-packing variant of PACKER, which is what predicted above-and-below for middle but not bottom.
-}
-
-\jlanote{they could? wouldn't it depend on the extent that they care about within-category similarity?}
+Thus, to the extent that category contrast influences distributional structure, we should observe more participants in the Middle condition that generate examples above \textit{and} below the experimenter-defined category.
 
@@
 61 participants were randomly assigned to the Middle and Bottom conditions each.
-The stimuli were exactly as in Experiment 1.
-Again, the assignment of perceptual features (color, size) to axes of the domain space (x, y) was counterbalanced across participants.
-The procedure was exactly as in Experiment 1: participants first completed a short training phase, followed by the generation phase, followed by the generalization phase (data from this phase is not discussed in this report).
+The stimuli and procedure were exactly as in Experiment 1: participants first completed a short training phase, followed by the generation phase.
 
@@
 
+\nbcnote{MADE IT HERE}
+
 \begin{table}
@@
 
-This finding is cannot be explained under the theoretical perspective that the distributional structure of previously learned categories is the sole determinant of the distributional structure of generated categories.
+This finding cannot be explained under the theoretical perspective that the distributional structure of previously learned categories is the sole determinant of the distributional structure of generated categories.
 However, the observed performance is sensible under a category contrast perspective: participants seeking to generate a perceptually distinct category will be more likely to use areas of space that are unoccupied by exemplars belonging to previously learned categories.

### 2017-04-08 Manuscripts/cog-psych/cogpsych.tex
drafted jk13 analysis section

@@
 
-\nbcnote{Report analyses on the \cite{jern2013probabilistic} data here...}
+Participants in their experiment were exposed to members of two experimenter-defined categories of `crystals' varying in hue, saturation, and size.
+Each category possessed a unique hue, but varied in saturation and size: In the `Positive' condition, there was a positive correlation between these features (i.e., larger sized crystals were more saturated), and in the `Negative' condition this relation was reversed.
+In the `Neutral' condition, there was no correlation between saturation and size.
+After learning about the categories from each condition, participants were asked to generate six exemplars belonging to a novel class.
+As noted above, \cite{jern2013probabilistic} found that the generated categories tended to follow the distributional properties of the experimenter-defined categories: generated categories were tightly distributed along the hue feature, and possessed the same saturation-size correlations as in the learned categories.
+\cite{jern2013probabilistic}, however, did not provide an analysis of how the generated categories \textit{differed} from the experimenter-defined categories.
+Here, we provide such an analysis.
+
+Because each experimenter-defined category in the \cite{jern2013probabilistic} experiment possessed a distinct hue, it is sensible to propose that participants would make their generated category distinct along this feature as well.
+Problematically, however, stimulus hue was measured on an circular (HSV) scale with limits $[0,1]$: both 0 and 1 corresponded to a `red' hue, and all other hues were represented between these values.
+By consequence, proximity on this dimension does not directly correspond to perceptual similarity.
+To overcome this issue, we categorized the hues of each participant's generated categories into one of six uniformly-spaced color groups: \{\textit{red}: 0 \& 1, \textit{yellow}: 0.167, \textit{green}: 0.333, \textit{teal}: 0.5, \textit{blue}: 0.667, \textit{pink}: 0.833\}.
+After each example's hue was categorized, we computed the amount of overlap between the hues of the six examples generated by each participant and the two hues possessed by the experimenter-defined categories.
+If contrast influences the hues of the generated categories, we should observe minimal overlap between the generated and experimenter-defined categories.
+
+\nbcnote{Joe should probably do a pass through that language. Maybe add a cite for proximity-similarity distinction?}
+
+\begin{figure}
+    \begin{center}
+        \inputpgf{figs/}{jk13-huecontrast.pgf}
+        \caption{Analysis of data from \cite{jern2013probabilistic}, Experiment 3.
+Plotted is the number of generated items that share a color group with one of the experimenter-defined classes.
+The ``Expected'' data follows a Binomial distribution with $p=2/6$, given there were two experimenter-defined classes, and 6 color groups.}
+        \label{fig:jk13-huecontrast}
+    \end{center}
+\end{figure}
+
+These data, shown in Figure \ref{fig:jk13-huecontrast}, reveal a clear pattern: The majority of participants generated categories possessing entirely distinct hues--with 0/6 exemplars sharing a hue with the experimenter-defined categories.
+These results can be compared to the predictions of a Binomial model, which proposes that participants generate hues at random.
+According to the model, if participants had generated hues at random, the probability that any given example will lie in the same color group as an experimenter-defined category is $p = 2/6$, as there were two experimenter-defined categories and six possible color groups.
+Thus, the number of shared hues is Binomially distributed with $p=2/6$ and $n=6$ (as that participants generated six examples in each condition).
+Comparison of \cite{jern2013probabilistic}'s results to the model's predictions reveals participants were far more likely to select distinct hues: participants tended to generate items that were perceptually distinct from the categories they had learned.
+
+\nbcnote{Some wrap-up ...}
 

### 2017-04-08 Modules/Funcs/Misc.py

@@ -64,6 +64,12 @@ def print2dspace(n, op = 'return'):
 		return vals
 
 
+def diffs1D(X, Y):
+	"""
+		Get pairwise differences in 1D
+	"""
+	return np.atleast_2d(X).transpose() - np.atleast_2d(np.array(Y)) 
+
 def histvec(X, bins, density = False):
 	"""
 	MATLAB-like histogram function, with items in vector X being placed
@@ -72,8 +78,7 @@ def histvec(X, bins, density = False):
 	if density = True, the histogram is normalized.
 	"""
 
-	D = np.atleast_2d(bins).transpose() - np.atleast_2d(np.array(X)) 
-	D = np.abs(D)
+	D = np.abs(diffs1D(bins, X))
 	assignment = np.argmin(D, axis=0)
 	counts = [np.sum(assignment==i) for i in range(len(bins))]
 	counts = np.array(counts)

### 2017-04-08 Manuscripts/cog-psych/cogpsych.tex
quick edits to the jk13 analysis section

--- 2017-04-07 Manuscripts/cog-psych/cogpsych.tex
@@
 
-Although existing accounts of creative generation broadly overlook the role of category contrast in determining what is novel versus familiar, effects of category contrast are undoubtedly ubiquitous within the published data.
+Although existing accounts of creative generation broadly overlook the role of category contrast in determining what is novel versus familiar, these effects are undoubtedly ubiquitous within the published data.
 In each study, participants were \textit{successful} in creating something new: the question is about \textit{how} they were able to do so.
-To provide some concrete evidence of these effects within existing datasets, we conducted an analysis of the data from Experiment 3 of the report by \cite{jern2013probabilistic}.
+To provide concrete evidence of  effects of category contrast within existing datasets, we conducted an analysis of Experiment 3 from \cite{jern2013probabilistic}.
 
-\nbcnote{Report analyses on the \cite{jern2013probabilistic} data here...}
+Participants in their experiment were exposed to members of two experimenter-defined categories of `crystals' varying in hue, saturation, and size.
+Each category possessed a unique hue, but varied in saturation and size: In the `Positive' condition, there was a positive correlation between these features (i.e., larger sized crystals were more saturated), and in the `Negative' condition this relation was reversed.
+In the `Neutral' condition, there was no correlation between saturation and size.
+After learning about the categories from each condition, participants were asked to generate six exemplars belonging to a novel class.
+As noted above, \cite{jern2013probabilistic} found that the generated categories tended to follow the distributional properties of the experimenter-defined categories: Generated categories were tightly distributed along the hue feature, and possessed the same saturation-size correlations as in the learned categories.
+\cite{jern2013probabilistic}, however, did not provide an analysis of how the generated categories \textit{differed} from the experimenter-defined categories.
+
+Because each experimenter-defined category in the \cite{jern2013probabilistic} experiment possessed a distinct hue, it is sensible to propose that participants would make their generated category distinct by using hues that differ from those present in the experimenter-defined category: If creative generation is influenced by category contrast, the hues of generated categories should be systematically different from those of the experimenter-defined categories.
+Problematically, however, stimulus hue was measured on an circular (HSV) scale with limits $[0,1]$: both 0 and 1 corresponded to a `red' hue, and all other possible hues were represented between these values.
+By consequence, proximity on the hue dimension does not directly correspond to perceptual similarity, and we cannot simply compare the hue values to determine the extent of the differences.
+
+To overcome this issue, we categorized the hues of each participant's generated categories into one of six uniformly-spaced color groups: \{\textit{red}: 0 \& 1, \textit{yellow}: 0.167, \textit{green}: 0.333, \textit{teal}: 0.5, \textit{blue}: 0.667, \textit{pink}: 0.833\}.
+Each generated crystal was placed into the color group closest to its hue value.
+After each example's hue was categorized, we computed the amount of overlap between the hues of the six hues generated by each participant and the two hues present in the experimenter-defined categories: If contrast influences the hues of the generated categories, we should observe minimal overlap between the generated and experimenter-defined categories.
+
+\nbcnote{Joe should probably do a pass through that language. Maybe add a cite for proximity-similarity distinction?}
+
+\begin{figure}
+    \begin{center}
+        \inputpgf{figs/}{jk13-huecontrast.pgf}
+        \caption{Analysis of data from \cite{jern2013probabilistic}, Experiment 3.
+Plotted is the number of generated items that share a color group with one of the experimenter-defined classes.
+The ``Expected'' data follows a Binomial distribution with $p=2/6$, given there were two experimenter-defined classes, and 6 color groups.}
+        \label{fig:jk13-huecontrast}
+    \end{center}
+\end{figure}
+
+These data, shown in Figure \ref{fig:jk13-huecontrast}, reveal a clear pattern: The majority of participants generated categories possessing entirely distinct hues; with 0/6 exemplars sharing a hue with the experimenter-defined categories.
+These results can be compared to the predictions of a Binomial model, which proposes that participants generate hues at random.
+According to the model, if hue selection is not systematic, the probability that any given example will lie in the same color group as an experimenter-defined category is $p = 2/6$, as there were two experimenter-defined categories and six possible color groups.
+Thus, the number of shared hues is Binomially distributed with $p=2/6$ and $n=6$ (as participants generated six examples in each condition).
+Comparison of the behavioral results to the model predictions reveals that participants were far more likely to select distinct hues: participants tended to generate items that were perceptually distinct from the categories they had learned.
+
+These results are notable in that they provide some early, concrete evidence that creative generation is influenced by category contrast.
+Taken alongside the analyses reported by \cite{jern2013probabilistic}, our results suggest that generated categories tend to be distinct from \textit{and} distributionally similar to what is already known.
+However, while these results indicate that contrast exerts \textit{some} influence, they do not precisely describe the nature of that influence.
+In the sections below, we propose a quantitative framework specifying the role of category contrast in creative generation.
 

### 2017-04-08 Manuscripts/cog-psych/cogpsych.tex
edits through simulations-individual differences

@@
 
-However, PACKER is only one possible exemplar-based account of category generation within our proposed framework.
-PACKER places specific constraints on the possible values of $f(x_j)$, but other exemplar-based category generation models with drastically different behavior can be formalized in this framework by imposing alternative constraints.
-For example, as will be discussed in more detail below, PACKER is formally equivalent to the copy-and-tweak model proposed by \cite{jern2013probabilistic} when $\gamma = 1$.
-Likewise, when $\gamma = 0$, PACKER represents a contrast-only generation mode, relying exclusively on contrast when generating new categories.
-Finally, when $f(x_j) = -1$ for all $x_j$ (regardless of class membership), a ``pure-packing'' approach is yielded, generating items in any unoccupied area of the domain.
-Thus, the proposed framework may be used to describe a wide variety of qualitatively distinct generation strategies.
-
 The probability that a given candidate $y$ will be generated is evaluated using an Exponentiated \citet{luce1977choice} choice rule.
@@
 
-It is worth noting that PACKER represents just one of many possible models that incorporate contrast in category generation.
-For example, although it may be possible to extend \cite{jern2013probabilistic}'s hierarchical sampling model to include contrast, but this mechanism would be at odds with a conjugate hierarchical Bayesian framework.
-In fact, we will discuss integrating our approaches in the General Discussion.
+It is worth noting that PACKER is only one possible exemplar-based account of category generation within our proposed framework.
+That is, PACKER places specific constraints on the possible values of $f(x_j)$, but other exemplar-based category generation models with drastically different behavior can be formalized in this framework by imposing alternative constraints.
+For example, as will be discussed in more detail below, PACKER is formally equivalent to the copy-and-tweak model proposed by \cite{jern2013probabilistic} when $\gamma = 1$.
+Likewise, when $\gamma = 0$, PACKER represents a contrast-only generation mode, relying exclusively on contrast when generating new categories.
+Finally, when $f(x_j) = -1$ for all $x_j$ (regardless of class membership), a ``pure-packing'' approach is yielded, generating items in any unoccupied area of the domain.
+Thus, the proposed framework may be used to describe a wide variety of qualitatively distinct generation strategies.
+
+It is also worth noting that PACKER represents just one of many possible models that incorporate contrast in category generation.
+For example, although it may be possible to extend \cite{jern2013probabilistic}'s hierarchical sampling model to include contrast, this mechanism would be at odds with a conjugate hierarchical Bayesian framework.
 In contrast, these ideas emerge naturally from the exemplar view: we have not modified any of the core elements of the GCM in defining the PACKER model, we simply aggregated similarity slightly differently.
+Thus, beyond formally specifying the role of contrast in creative generation, the PACKER account allows us to evaluate the well-understood principles of the exemplar view within the comparatively understudied field of conceptual generation.
+Nonetheless, we will discuss integrating our approaches in the General Discussion.
 
@@
 The manner in which generated categories differ from known ones is, nonetheless, poorly understood: Existing theories do not make strong predictions about how creatively generated concepts should systematically differ from existing ones.
-Above, we introduced a novel, exemplar-based model formalizing the roles of similarity and contrast in creative generation.
+Above, we provided some early evidence of the influence of category contrast in published data \citep[][Experiment 3]{jern2013probabilistic}, and we introduced a novel, exemplar-based model formalizing the roles of similarity and contrast in creative generation.
 
@@
 
-To begin our investigation, we developed an artificial, two dimensional domain of squares, varying in color and size (see Figure \ref{fig:e1-conditions}, panel A).
-To foreshadow slightly, effects of contrast are strong and can be widely observed in category generation data.
-Accordingly, we began by testing the strength of these effects under a variety of learning conditions.
+To begin our investigation, we sought to extend the early evidence from the \cite{jern2013probabilistic} data, under a variety of learning conditions and using more standard stimulus materials.
+We developed an artificial, two dimensional domain of squares, varying in color and size (see Figure \ref{fig:e1-conditions}, panel A).
+These features have been used in numerous classification learning studies \cite[e.g.,][]{conaway2016similar,conaway2016generalization,shepard1961learning,nosofsky1994comparing}.
+Unlike those used in the \cite{jern2013probabilistic} experiments, proximity along these dimensions aligns more directly with perceptual similarity, allowing us to more directly evaluate the role of category contrast in creative generation.
+In order to extend the evidence provided by the \cite{jern2013probabilistic} data, we began by testing effects of category contrast under a variety of learning conditions.
 This was achieved by training participants on categories possessing qualitatively different distributional structures, as shown in Figure \ref{fig:e1-conditions}.
+
+\nbcnote{can you think of other studies with size and color of squares as features?}
 
@@
 It should be noted that in our experiments the assignment between the perceptual and conceptual dimensions (e.g., $X \rightarrow Size$, $Y \rightarrow Color$), as well as the direction of variation along each dimension (e.g., $dark \rightarrow light$ or $light \rightarrow dark$) was counterbalanced across participants.
-The category types in Figure \ref{fig:e1-conditions} are plotted in a conceptual space, rather than a perceptual space: each category type may have a different physical instantiation according to the counterbalance assignment.
+The category types in Figure \ref{fig:e1-conditions} are plotted in a conceptual space, rather than a perceptual space: thus, while the conceptual organization of the category types remains constant, each category type may have a different physical instantiation according to the counterbalance assignment.
 For example, the Cluster type  may be large and dark in color, or it may be small and light in color, depending on the assignment and direction of the dimensions.
 For this reason, below we will discuss generation within a conceptual space, rather than a physically instantiated one.
-Importantly, however, the conceptual organization of the category types remains constant across the counterbalanced assignments.
 
@@
     \inputpgf{figs/}{e1-conditions.pgf}
-    \caption{Stimulus domain and category types tested in Experiment 1. }
+    \caption{Stimulus domain and category types tested in Experiment 1. Stimuli are not drawn to scale.}
     \label{fig:e1-conditions}
@@
 Participants were randomly assigned to one condition: 64 participants were assigned to the Cluster condition, 61 were assigned to the Row condition, and 58 were assigned to the XOR condition.
-Stimuli were squares varying in color (RGB 25--230) and side length (3.0--5.8cm).
-These stimuli are slight variants of those used by \cite{conaway2016similar}, see Figure \ref{fig:e1-conditions}.
+Stimuli were squares varying in color (RGB 25--230) and side length (3.0--5.8cm), see Figure \ref{fig:e1-conditions}.
 The assignment of perceptual features (color, size) to axes of the domain space (x, y), as well as the direction of variation along each axis (e.g., $dark \rightarrow light$ or $light \rightarrow dark$) was counterbalanced across participants.
@@
 
-In Experiment 1 we sought to test whether category contrast imposes constraints on creative generation.
-We found strong evidence for effects of category contrast: Participants were more likely to generate stimuli that are more distant from (i.e., less similar to) members of a previously-learned category, and members of participant-generated categories tended to be more similar to one another than to members of previously-learned categories.
+In Experiment 1 we sought to extend our analysis of the \cite{jern2013probabilistic} data by evaluating the influence of category contrast on creative generation, given qualitatively different types of prior knowledge.
+We found strong evidence for effects of category contrast in each condition: Participants were more likely to generate stimuli that are more distant from (i.e., less similar to) members of a previously-learned category, and members of participant-generated categories tended to be more similar to one another than to members of previously-learned categories.
 We also partially replicated the classic finding that the distributional structure of generated categories reflects that of previously learned categories \citep{jern2013probabilistic,ward1994structured}: members of generated categories were more widely distributed along dimensions which were widely distributed in the experimenter-defined category.
@@
 
-\nbcnote{MADE IT HERE}
-
 \begin{table}
@@
 Firstly, it should be noted that nearly every participant utilized the top and/or bottom rows: only $10 / 122$ participants generated their category entirely within the interior region.
-Fisher's Exact Tests comparing the conditions reveal that more Middle participants generated an exemplar in the bottom row, $p < 0.001$, again demonstrating the role of contrast guiding where exemplars are generated.
+Fisher's Exact Tests comparing the conditions reveal that more Middle participants generated an exemplar in the bottom row, $p < 0.001$, again demonstrating the role of contrast in guiding where exemplars are generated.
 The conditions did not differ in use of the top of the space, $p = 0.16$, however, more Middle participants placed exemplars in the top \textit{and} bottom rows, $p = 0.038$.
@@
 However, we additionally found that the \textit{position} of a previously learned category (rather than its distributional structure) influences the types of categories people generate: participants who learned the `Middle' type were more likely to generate categories spanning the entire Y-axis of the space.
-Participants who learned the `Bottom' type were less likely to do so, presumably because of the presence of opposite category exemplars in the lower regions of the space.
+Participants who learned the `Bottom' type were less likely to do so as a result of the presence of opposite category exemplars in the lower regions of the space.
 
-This finding cannot be explained under the theoretical perspective that the distributional structure of previously learned categories is the sole determinant of the distributional structure of generated categories.
-However, the observed performance is sensible under a category contrast perspective: participants seeking to generate a perceptually distinct category will be more likely to use areas of space that are unoccupied by exemplars belonging to previously learned categories.
+This finding cannot be explained under the viewpoint that the distributional structure of previously learned categories is the sole determinant of the distributional structure of generated categories.
+However, the observed performance is sensible given a category contrast perspective: participants seeking to generate a perceptually distinct category will be more likely to use areas of space that are unoccupied by exemplars belonging to previously learned categories.
 In the Middle condition, the upper and lower regions of space are equidistant from members of the experimenter-defined category, whereas in the Bottom condition, the lower region of the space is closer to members of the experimenter-defined category.
@@
 To facilitate comparison, PACKER's other parameters ($c$, $\theta$) were set to the best fitting values obtained for copy-and-tweak in Table \ref{table:global-model-fits}.
-Thus, in this figure PACKER is equivalent to copy-and-tweak when $\gamma = 1$.
 }

### 2017-04-08 Experiments/middle-bottom/analysis/plot.stats.figs.py
changed e2 colormap

@@ -4,6 +4,9 @@ import numpy as np
 import matplotlib.pyplot as plt
 import seaborn as sns
 sns.set_style("whitegrid")
+colors = ["#34495e", "#e74c3c"]
+sns.set_palette(colors)
+
 
 pd.set_option('display.width', 1000, 'precision', 2)
 
@@ -21,12 +24,11 @@ stats = pd.merge(stats, info, on = 'participant')
 
 
 fh, axes = plt.subplots(1,3,figsize = (7.5,2.5))
-colors = dict(Middle = 'orange', Bottom = 'purple')
 
 for i, col in enumerate(['xrange','yrange','correlation']):
 	ax = axes[i]
 	hs = sns.factorplot(x = 'condition', y = col, data= stats, ax = ax, kind = 'box', 
-		order = ['Middle', 'Bottom'], palette = 'PuOr')
+		order = ['Bottom', 'Middle'])
 
 	ax.set_title(col, fontsize = 12)
 	ax.set_ylabel('')

### 2017-04-08 Manuscripts/cog-psych/cogpsych.tex
changed e2 colormap
added range diff figure to manuscript

@@
 
-To evaluate the role of contrast in category generation, we computed the number of times each stimulus was generated, as a function of its average city-block distance from members of the experimenter-defined ``Alpha'' category.
+To evaluate the role of contrast, we computed the number of times each stimulus was generated, as a function of its average city-block distance from members of the experimenter-defined ``Alpha'' category.
 These data, shown in Figure \ref{fig:e1-distanceplots}, reveal a clear pattern: examples that are more distant from members of the experimenter-defined categories are more likely to be generated into a new category.
@@
 \textit{Right}: Scatter plot of within-class versus between-class distance in each of the participant-generated categories.}
-    \nbcnote{May want to change condition colors if purple-orange is used as the colormap for the range heat-maps in the simulations section.}
-    \jlanote{sounds good to me.}
     \label{fig:e2-distanceplots}
@@
 
+\begin{figure}[p]
+    \begin{center}
+    \inputpgf{figs/}{range-diff-gradients.pgf}
+    \caption{Behavioral and simulated relative range gradients.
+Each panel shows, for each stimulus, the dimensional orientation of the categories it was generated into: vertically aligned `columns' versus horizontally aligned `rows'.}
+    \label{fig:range-diff-gradients}
+    \end{center}
+\end{figure}
+
 \section{General Discussion}

### 2017-04-10 Manuscripts/cog-psych/cogpsych.tex
added stats to jk13 analysis, adjusted binning

@@
 By consequence, proximity on the hue dimension does not directly correspond to perceptual similarity, and we cannot simply compare the hue values to determine the extent of the differences.
+\nbcnote{Maybe add a cite for proximity-similarity distinction?}
 
-To overcome this issue, we categorized the hues of each participant's generated categories into one of six uniformly-spaced color groups: \{\textit{red}: 0 \& 1, \textit{yellow}: 0.167, \textit{green}: 0.333, \textit{teal}: 0.5, \textit{blue}: 0.667, \textit{pink}: 0.833\}.
-Each generated crystal was placed into the color group closest to its hue value.
-After each example's hue was categorized, we computed the amount of overlap between the hues of the six hues generated by each participant and the two hues present in the experimenter-defined categories: If contrast influences the hues of the generated categories, we should observe minimal overlap between the generated and experimenter-defined categories.
-
-\nbcnote{Joe should probably do a pass through that language. Maybe add a cite for proximity-similarity distinction?}
+To overcome this issue, we categorized all possible hues into one of eight uniformly-spaced color groups: \{\textit{Red}: $0-0.063$, $0.938-1$, \textit{Yellow}: $0.063-0.188$, \textit{Yellow-Green}: $0.188-0.313$, \textit{Green-Teal}: $0.313-0.438$, \textit{Teal}: $0.438-0.563$, \textit{Teal-Blue}: $0.563-0.688$, \textit{Purple}: $0.688-0.813$, \textit{Pink}: $0.813-0.938$\}.
+In the \cite{jern2013probabilistic} experiment, the hue of each experimenter-defined category was selected from one of six possible values, each of which falls into one of the color groups above.
+By categorizing the participant-generated crystals likewise, we can obtain a broad measure of category contrast by determining the proportion of participant-generated crystals that fall into the same groups as the experimenter-defined categories: If contrast influences the hues of the generated categories, we should observe minimal overlap between in the color groupings.
 
@@
 Plotted is the number of generated items that share a color group with one of the experimenter-defined classes.
-The ``Expected'' data follows a Binomial distribution with $p=2/6$, given there were two experimenter-defined classes, and 6 color groups.}
+The ``Expected'' data follows a Binomial distribution with $p=2/8$, given there were two experimenter-defined classes, and eight color groups.}
         \label{fig:jk13-huecontrast}
@@
 
-These data, shown in Figure \ref{fig:jk13-huecontrast}, reveal a clear pattern: The majority of participants generated categories possessing entirely distinct hues; with 0/6 exemplars sharing a hue with the experimenter-defined categories.
+These data, shown in Figure \ref{fig:jk13-huecontrast}, reveal a clear pattern: The majority of participants in each condition ($n = 22$) generated categories possessing entirely distinct hues; with 0/6 exemplars sharing a hue with the experimenter-defined categories.
 These results can be compared to the predictions of a Binomial model, which proposes that participants generate hues at random.
-According to the model, if hue selection is not systematic, the probability that any given example will lie in the same color group as an experimenter-defined category is $p = 2/6$, as there were two experimenter-defined categories and six possible color groups.
-Thus, the number of shared hues is Binomially distributed with $p=2/6$ and $n=6$ (as participants generated six examples in each condition).
-Comparison of the behavioral results to the model predictions reveals that participants were far more likely to select distinct hues: participants tended to generate items that were perceptually distinct from the categories they had learned.
+That is, if hue selection is not systematic, the probability that any given example will lie in the same color group as an experimenter-defined category is given by a Binomial distribution with $p = 2/8$, as there were two experimenter-defined categories and eight possible color groups.
+Chi-square goodness-of-fit tests reveal that the observed distribution in each condition is highly inconsistent with the random-hue model (all $\tilde{\chi}^2>200$, $p<0.001$): Participants tended to generate items that were perceptually distinct from the categories they had learned, and were less likely to generate hues possessed by members of the experimenter-defined categories.
 
@@
 Taken alongside the analyses reported by \cite{jern2013probabilistic}, our results suggest that generated categories tend to be distinct from \textit{and} distributionally similar to what is already known.
-However, while these results indicate that contrast exerts \textit{some} influence, they do not precisely describe the nature of that influence.
+However, it is worth noting that our analysis is still limited: the color groups defined above are somewhat arbitrary, and it is not clear that our color grouping is consistent with psychological color boundaries.
+While we did obtain similar results using a variety of alternative groupings, the hue dimension used in the \cite{jern2013probabilistic} study does not lend itself straightforwardly to the the computation of similarities, and thus we cannot be certain of whether our coding accurately approximates the psychological space of the stimuli.
+By consequence, although these results likely indicate that contrast exerts \textit{some} influence, they do not precisely describe the nature of that influence.
 In the sections below, we propose a quantitative framework specifying the role of category contrast in creative generation.
+
+\nbcnote{could use a cite on why hue dimension proximity does not match up with similarity. maybe also a note about how we can't be sure what hues were actually perceived due to norming, etc.}
 
@@
 
-To begin our investigation, we sought to extend the early evidence from the \cite{jern2013probabilistic} data, under a variety of learning conditions and using more standard stimulus materials.
+To begin our investigation, we sought to extend the early evidence obtained from our analysis of the \cite{jern2013probabilistic} data, under a variety of learning conditions and using more standard stimulus materials.
 We developed an artificial, two dimensional domain of squares, varying in color and size (see Figure \ref{fig:e1-conditions}, panel A).
@@
 To characterize the nature of these differences, Figure \ref{fig:e2-samples} depicts sample categories generated by participants.
-The data from each condition in Figure \ref{fig:e2-samples} is organized into four columns based on commonly observed patterns of generation: a `Cluster' type of tightly-clustered examples,   `Row' and `Column' types of exemplars widely distributed along the one axis but narrowly along the other, and a `Corners' type, wherein participants placed exemplars in disparate corners of the space.
+The data from each condition is organized into four columns based on commonly observed patterns of generation: a `Cluster' type of tightly-clustered examples,   `Row' and `Column' types of exemplars widely distributed along the one axis but narrowly along the other, and a `Corners' type, wherein participants placed exemplars in disparate corners of the space.
 As before, in this section we focus on what can be concluded from analyzing the data in aggregate, but in later sections we will focus more specifically on explaining the individual differences.
@@
 We ran this analysis but the conditions did not differ in Y-range due to the data being weirdly distributed.
-But it feels odd to ignore this point, so I've included the box-and-whisker plot, as well as the graph-of-many-lines figure, on a separate page at the end.}
+I've included the box-and-whisker plot, as well as the graph-of-many-lines figure, on a separate page at the end.}
 
@@
 Each panel shows, for each stimulus, the dimensional orientation of the categories it was generated into: vertically aligned `columns' versus horizontally aligned `rows'.}
+    \nbcnote{Need to run this on lando with many more samples}
     \label{fig:range-diff-gradients}

### 2017-04-10 Experiments/jernkemp2013-e3/JK13/CustomFuncs.py

@@ -1,19 +1,25 @@
 import numpy as np
 
 
-def dummycode_colors(X):
+def dummycode_colors(X, hues = None):
 	"""
 		function to convert hue value into a color category
 	"""
-
 	# naturalish categorization via https://en.wikipedia.org/wiki/File:HueScale.svg
-	hue_categories = np.array([[0,60,120,180,240,300,360]]).T / 360.0
+	if hues is None:
 
-	# 0.15s used by jk13, with added 1.0 at end
-	# hue_categories = np.array([[ 0,54,108,162,216,270,324, 360]]).T / 360.0
+		# # naturalish categorization via https://en.wikipedia.org/wiki/File:HueScale.svg
+		# hues = np.array([[0,60,120,180,240,300,360]]).T / 360.0
 
-	k = len(hue_categories)
-	D = np.abs(hue_categories - np.atleast_2d(np.array(X)))
+		# # Intermediates
+		# hues = np.array([[0,30,60,90,120,150,180,210,240,270,300,330,360]]).T / 360.0
+		hues = np.array([[0, 45, 90, 135, 180, 225, 270, 315, 360]]).T / 360.0
+
+		# # 0.15s used by jk13, with added 1.0 at end
+		# hues = np.array([[ 0,54,108,162,216,270,324, 360]]).T / 360.0
+
+	k = len(hues)
+	D = np.abs(hues - np.atleast_2d(np.array(X)))
 	assignment = np.argmin(D, axis=0)
 	assignment[assignment==(k-1)] = 0
 	return assignment

### 2017-04-10 Manuscripts/cog-psych/cogpsych.tex
hotfix-- added domain param to conjugatejk13 model

@@
 
+A key distinction between PACKER and copy-and-tweak, as well as the the hierarchical sampling model, is that, of the three models, only PACKER is capable of making strong predictions about the location of new category members when the target class is entirely novel (i.e., no member of the category has been observed).
+Under these circumstances, there are no examples to copy, and thus the copy-and-tweak model predicts that items are generated at random.
+Likewise, with no observations on which to condition the mean of the category distribution, the hierarchical sampling model also picks an item at random.
+Thus, it is possible that the failure of these models is simply due to their inability to explain each participant's first trial (generating the first item in the `Beta' category).
+We conducted an identical set of simulations as above, excluding this trial
+
 \nbcnote{Should include info on fits for only trials 2-4.}
@@
 
-\subsection{Individual Differences}
+\subsection{Category Location vs. Distributional Structure}
+
+, in relation to what is already known.
+That is, when the target category is entirely novel (i.e., no member of the category has been observed), both copy-and-tweak and the hierarchical sampling model essentially generate items at random.
+
+both the copy-and-tweak and hierarchical sampling models
+The copy-and-tweak and hierarchical sampling models predictions about where new categories should be generated in relation to existing category members
 
@@
 Each panel shows, for each stimulus, the dimensional orientation of the categories it was generated into: vertically aligned `columns' versus horizontally aligned `rows'.}
-    \nbcnote{Need to run this on lando with many more samples}
+    \nbcnote{Need to run this on lando with many more samples. current = 30}
     \label{fig:range-diff-gradients}
@@
 \end{figure}
+
+\subsection{Individual Differences}
+
+As noted in Experiments 1 and 2, we observed a great deal of individual differences in the types of categories generated.
+Within each condition, we observed tightly clustered and widely distributed categories, as well as row and column categories, and so forth.
+The simulations reported above serve to evaluate the models while considering the entire dataset, but a secondary goal of any formal account should be to provide some explanation of how different profiles of performance emerge.
+Many of the individual generation profiles we observed can be specifically targeted within the PACKER framework, simply by tuning the similarity trade-off parameter ($\gamma$) and the model's attention weights ($w$).
+
+\nbcnote{manual description of how to get rows, columns, 4 corners, etc}
+
+To systematically evaluate each model's ability to explain
 

### 2017-04-10 Experiments/cogpsych-code/construct-trial-set.py

@@ -33,3 +33,10 @@ trials = trials.add_frame(generation)
 
 with open('pickles/all_data_e1_e2.p','wb') as f:
 	pickle.dump(trials, f)
+
+# weed out first trials
+trials.Set = [i for i in trials.Set if len(i['categories'][1])>0]
+trials._update()
+
+with open('pickles/trials_2-4_e1_e2.p','wb') as f:
+	pickle.dump(trials, f)

### 2017-04-11 Manuscripts/cog-psych/cogpsych.tex
most of structure-location section written

@@
  \hline
- $AIC = 9095$       & $AIC = 9842$          & $AIC = 10080$     \\ 
- $L = -4545$        & $L = -4919$           & $L = -5036$       \\ 
+ $AIC = 9095$       & $AIC = 9842$          & $AIC = 9912$     \\ 
+ $L = -4545$        & $L = -4919$           & $L = -4952$       \\ 
  $c = 0.482$        & $c = 3.187$           & $\kappa < 0.001$  \\
- $\gamma = 0.525$   & $\gamma = 1$ (fixed)  & $\nu = 1.001$     \\ 
- $\theta = 6.664$   & $\theta = 2.969$      & $\rho = 0.351$    \\ 
-                    &                       & $\theta = 7.951$  \\ 
+ $\gamma = 0.525$   & $\gamma = 1$ (fixed)  & $\nu = 5.596$     \\ 
+ $\theta = 6.664$   & $\theta = 2.969$      & $\rho = 0.055$    \\ 
+                    &                       & $\theta = 3.174$  \\ 
 \end{tabular}
@@
 Thus, it is possible that the failure of these models is simply due to their inability to explain each participant's first trial (generating the first item in the `Beta' category).
-We conducted an identical set of simulations as above, excluding this trial
-
-\nbcnote{Should include info on fits for only trials 2-4.}
+We conducted an identical set of simulations as above, excluding this trial (leaving 915 responses in the dataset): Again, PACKER ($L = -3390$, $AIC = 6786$) achieved better fits than the copy-and-tweak ($L = -3579$, $AIC = 7162$) and hierarchical sampling ($L=-3612$, $AIC = 7232$) models.
 
@@
 
-, in relation to what is already known.
-That is, when the target category is entirely novel (i.e., no member of the category has been observed), both copy-and-tweak and the hierarchical sampling model essentially generate items at random.
+As noted above, while all three models make clear claims about the internal structure of generated categories, the copy-and-tweak and hierarchical sampling models do not make any claims about how generated categories should differ from what is already known.
+However, as we observed in Experiment 2, the distributional structure of a category is not always independent of its location within the domain.
+To demonstrate this point in a broader manner, we computed the X- and Y- axis ranges of every participant-generated category.
+Taking the difference between these values ($X-Y$) produces a measure of each category's orientation in the space: positive difference scores correspond to categories with more X-axis range (horizontally aligned, `Row' categories), whereas negative difference scores indicate the opposite (vertically aligned, `Column' categories).
+Neutral differences scores indicate there was an equal amount of X- and Y-axis range, which can be produced by a number of different category types (`Clusters', `Corners', etc; see Figures \ref{fig:e1-samples} and \ref{fig:e2-samples}).
+By plotting, for each possible stimulus, the difference scores of categories it was generated within, we can relate the distributional structure of generated categories to their location within the domain.
 
-both the copy-and-tweak and hierarchical sampling models
-The copy-and-tweak and hierarchical sampling models predictions about where new categories should be generated in relation to existing category members
+For each possible stimulus, we compiled the range differences across all the categories it was generated into.
+However, because many stimuli were infrequently generated (such items near members of the `Alpha' category), we cannot simply compute the empirical average, as infrequently generated stimuli would be likely to show artificially strong differences.
+Instead, we aggregated the scores using an empirical Bayesian analysis \citep{robbins1964empirical}: the aggregated score for each stimulus was viewed as a normal distribution with a $\mu_0 = 0$ and $\sigma$ set based on the observed scores $x$ for that stimulus.
+The aggregated difference $\mu_x$ was then computed as a Bayesian update:
+
+\begin{equation}
+\mu_x = \dfrac
+{ \mu_0/\sigma + \sigma^{-1}\sum{x}} 
+{ 1/\sigma + n_x/\sigma }
+\label{eq:rangediff-bayes}
+\end{equation}
+\nbcnote{I'm pretty sure some distributive-property trickery can be used to clean that up. In either case, Joe should take a look at that language.}
+% 
+Within this approach, the resulting aggregation is a trade-off between the number of generations and the strength of the range difference within each generated category.
+Infrequently generated stimuli, as well as those with both strongly positive and negative scores, are given neutral difference scores.
+The results of our analysis are shown in Figure \ref{fig:range-diff-gradients}.
+\nbcnote{Do you think we need to report the Gaussian smoothing?}
 
@@
     \inputpgf{figs/}{range-diff-gradients.pgf}
-    \caption{Behavioral and simulated relative range gradients.
+    \caption{Behavioral and simulated range difference gradients.
 Each panel shows, for each stimulus, the dimensional orientation of the categories it was generated into: vertically aligned `columns' versus horizontally aligned `rows'.}
@@
 \end{figure}
+
+These data reveal strong and consistent patterns across all the conditions we tested in Experiments 1 and 2: generated categories are more tightly distributed along the axis in which they are distinct.
+For example, in the `Cluster' condition, exemplars in the bottom-left of the space are more often generated into vertically aligned categories, and exemplars in the top-right are more often generated into horizontally aligned categories.
+Similarly, in the `Bottom' and `Middle' conditions, horizontally aligned categories are generated above and below the experimenter-defined categories, while vertically-aligned categories are generated to the sides.
+In the `Row' condition, most categories are horizontally aligned, and lie along the upper areas of the space.
+There are no strong range difference patterns in the XOR condition.
+
+These patterns of performance clearly depict the interdependence between the distributional structure and location of generated concepts.
+Our results can, in some sense, be interpreted in terms of local minimization of between-class similarity: by distributing the generated category away from members of the experimenter-defined category, participants may increase the degree of between-class distance without drastically altering the degree of within-class similarity.
+
+To attempt to explain our findings within the PACKER, copy-and-tweak- and hierarchical sampling models, we conducted simulations using an individual-differences approach.
+Each model was used to simulate each participant's generation separately, with the importance of each dimension set according to the relative range of the participant's generated category along each dimension.
+The other free parameters were set as in Table \ref{table:global-model-fits}.
+In the PACKER and copy-and-tweak models, the attention weights, $w$, specify the importance of each dimension in the computation of similarity.
+Traditionally, the attention weights are thought to reflect the diagnostic value of each dimension towards classifying members of the categories \citep{nosofsky1984choice,nosofsky1986attention,kruschke1992alcove}.
+Thus, while there exist methods to find the optimal attention weighting scheme given a classification \citep[see][]{vanpaemel2012using}, for simplicity we may assume that the Alpha and Beta categories are distinct along dimensions that the Betas do not vary on.
+Thus, the weighting for a given participant can be computed as:
+
+\begin{equation}
+w_k = \dfrac
+{\exp{ \{ -\theta_w \cdot\text{range}(k)}  \} } 
+{ \sum_k {\exp{ \{ -\theta_w \cdot\text{range}(k)}  \} } }
+\label{eq:range-weight}
+\end{equation}
+%
+
+In the hierarchical sampling model, the covariance matrix specifying the prior domain distribution can be used to manipulate the assumed category variance along each dimension, fulfilling a similar role.
+
+ % the weights using the inverse of each feature's range: $w_k \propto range(k)^{-1}$.
+
+% , we set the attention weight parameters in PACKER and copy-and-tweak per participant.
+The other free parameters were set as in Table \ref{table:model-fits}.
+While there exist methods to find the optimal attention weights for a given classification \citep[see][]{vanpaemel2012using}, for simplicity we approximated the weights using the inverse of each feature's range: $w_k \propto range(k)^{-1}$.
+Thus, the Alpha and Beta categories are assumed to be distinct along dimensions that the Betas do not vary on.
+To simulate the hierarchical sampling model we set the domain covariance prior, $\Sigma_0$, proportional to the range of each feature.
+We then simulated 50 Beta categories with each participant's weighting scheme to obtain a sense of how the relative importance of each dimension affects what types of categories are generated and where they are generated.
+The results of these simulations are depicted in Figure \ref{fig:range-diff-gradient}.
+
+\nbcnote{Weak spot: attention weighting feels unmotivated.}
 

### 2017-04-12 Manuscripts/cog-psych/cogpsych.tex
drafted location-structure section; fixed jk13 weighting

@@
 
+\subsection{Individual Differences}
+
+As noted in Experiments 1 and 2, we observed a great deal of individual differences in the types of categories generated.
+Within each condition, we observed tightly clustered and widely distributed categories, as well as row and column categories, and so forth.
+The simulations reported above serve to evaluate the models while considering the entire dataset, but a secondary goal of any formal account should be to provide some explanation of how different profiles of performance emerge.
+Many of the individual generation profiles we observed can be targeted within the PACKER framework, simply by tuning the model's parameters.
+In this section, we describe more specifically how the most frequently observed profile can be realized.
+
+\nbcnote{manual description of how to get rows, columns, 4 corners, etc}
+
+\nbcnote{If this is done correctly, we can motivate the use of attention weights below. }
+
 \subsection{Category Location vs. Distributional Structure}
@@
 Each model was used to simulate each participant's generation separately, with the importance of each dimension set according to the relative range of the participant's generated category along each dimension.
-The other free parameters were set as in Table \ref{table:global-model-fits}.
 In the PACKER and copy-and-tweak models, the attention weights, $w$, specify the importance of each dimension in the computation of similarity.
 Traditionally, the attention weights are thought to reflect the diagnostic value of each dimension towards classifying members of the categories \citep{nosofsky1984choice,nosofsky1986attention,kruschke1992alcove}.
-Thus, while there exist methods to find the optimal attention weighting scheme given a classification \citep[see][]{vanpaemel2012using}, for simplicity we may assume that the Alpha and Beta categories are distinct along dimensions that the Betas do not vary on.
+While there exist methods to find the optimal attention weighting scheme given a classification \citep[see][]{vanpaemel2012using}, for simplicity we may assume that the Alpha and Beta categories are distinct along dimensions that the Betas do not vary on.
 Thus, the weighting for a given participant can be computed as:
+
+\nbcnote{Weak spot: attention weighting feels unmotivated.}
 
@@
 % 
+where $\theta_w$ is a free parameter controlling how differences in range correspond to differences in weights (functioning similarity to the $\theta$ parameter in each of the models).
+In our simulations $\theta_w = 1.5$, though similar results can be obtained using a wide range of values.
+The resulting $w$ values are thus inversely proportional to the range of generated categories along each dimension, with less range corresponding to greater weighting.
 
-In the hierarchical sampling model, the covariance matrix specifying the prior domain distribution can be used to manipulate the assumed category variance along each dimension, fulfilling a similar role.
+In the hierarchical sampling model, the covariance matrix specifying the prior domain distribution, $\Sigma_0$, can be used to similar effect.
+This covariance matrix specifies the amount of variance assumed along each dimension (as well as the dimensional correlations) across the domain of categories.
+The covariance matrix for a newly generated category, $\Sigma_B$, is based on the assumed $\Sigma_0$ as well as the distributions of previously learned categories (see Appendix \ref{ap:hsampling-definition}).
+Thus, the importance of each feature can be coded into $\Sigma_0$ to alter the dimensional variance of generated categories.
+We incorporated the weights computed in Equation \ref{eq:range-weight} by setting the variance of each dimension according to: $\rho (1-w_k)n$, where $\rho$ is a free parameter specifying the overall assumed variance of the domain, and $n$ is the number of dimensions.
+Under this approach, evenly distributed weights correspond to an assumed variance of $\rho$.
+Likewise, larger values of $w$, which are produced when the generated category is tightly distributed along one dimension, correspond to smaller assumed variances.
 
- % the weights using the inverse of each feature's range: $w_k \propto range(k)^{-1}$.
+\nbcnote{This works for 2D, but not 3D and beyond (as $1-w_k$ would not produce a valid weighting).}
 
-% , we set the attention weight parameters in PACKER and copy-and-tweak per participant.
-The other free parameters were set as in Table \ref{table:model-fits}.
-While there exist methods to find the optimal attention weights for a given classification \citep[see][]{vanpaemel2012using}, for simplicity we approximated the weights using the inverse of each feature's range: $w_k \propto range(k)^{-1}$.
-Thus, the Alpha and Beta categories are assumed to be distinct along dimensions that the Betas do not vary on.
-To simulate the hierarchical sampling model we set the domain covariance prior, $\Sigma_0$, proportional to the range of each feature.
-We then simulated 50 Beta categories with each participant's weighting scheme to obtain a sense of how the relative importance of each dimension affects what types of categories are generated and where they are generated.
-The results of these simulations are depicted in Figure \ref{fig:range-diff-gradient}.
+Each model was used to simulate each participant's generation independently, with the importance of each dimension set according to the participant's generated category.
+The other free parameters within each model were set as in Table \ref{table:global-model-fits}.
+Every participant's generation was simulated \textbf{NUMBER} of times, resulting in \textbf{NUMBER * PARTICIPANTS} categories generated by each model.
+For comparison with our behavioral results, we then computed the range difference gradient identically as with the behavioral data.
+The results are shown in Figure \ref{fig:range-diff-gradients}.
 
-\nbcnote{Weak spot: attention weighting feels unmotivated.}
-
-\subsection{Individual Differences}
-
-As noted in Experiments 1 and 2, we observed a great deal of individual differences in the types of categories generated.
-Within each condition, we observed tightly clustered and widely distributed categories, as well as row and column categories, and so forth.
-The simulations reported above serve to evaluate the models while considering the entire dataset, but a secondary goal of any formal account should be to provide some explanation of how different profiles of performance emerge.
-Many of the individual generation profiles we observed can be specifically targeted within the PACKER framework, simply by tuning the similarity trade-off parameter ($\gamma$) and the model's attention weights ($w$).
-
-\nbcnote{manual description of how to get rows, columns, 4 corners, etc}
-
-To systematically evaluate each model's ability to explain
+As in the more traditional model evaluation procedure described above, PACKER provided a much closer match to our behavioral results than the copy-and-tweak and hierarchical sampling models.
+In all conditions, PACKER distributes categories similarly to the behavioral data: horizontally-aligned categories tend to be placed above and below members of the experimenter-defined category, and vertically-aligned categories tend to be placed to the sides.
+Conversely, because the copy-and-tweak and hierarchical sampling models are insensitive to category contrast, these models do not produce any systematic patterns of association between category location and distributional structure.
+The sole exception is within the `Row' condition, in which the majority of participants generated a `Row'-like category, widely distributed along the X-axis but not the Y-axis.
+In these case, both models are initialized with weights that produce Row categories, but because category contrast is not considered, categories are uniformly generated across the entire domain, rather than concentrated within the upper-regions as observed behaviorally.
 

### 2017-04-12 Manuscripts/cog-psych/cogpsych.tex
added representative samples for e1
added packer attention examples

@@
     \inputpgf{figs/}{e1-samples.pgf}
-    \caption{Sample categories generated in Experiment 1. }
+    \caption{Sample categories generated in Experiment 1. Representative samples from common generation profiles are shown.}
     \label{fig:e1-samples}
-    \nbcnote{I selected these at random, we can pick more representative samples later.}
-    \jlanote{ok!}
     \end{center}
@@
 As noted in Experiments 1 and 2, we observed a great deal of individual differences in the types of categories generated.
-Within each condition, we observed tightly clustered and widely distributed categories, as well as row and column categories, and so forth.
+Within each condition, we observed tightly clustered and widely distributed categories, as well as row and column categories, and so forth (see Figures \ref{fig:e1-samples} and \ref{fig:e2-samples}).
 The simulations reported above serve to evaluate the models while considering the entire dataset, but a secondary goal of any formal account should be to provide some explanation of how different profiles of performance emerge.
@@
 
-\nbcnote{manual description of how to get rows, columns, 4 corners, etc}
+To begin, the common profiles among the `Cluster' category types (including the `Cluster' condition of Experiment 1 and both conditions from Experiment 2) corresponded to: (A) a tight cluster of examples, (B) `row'- and `column'-like categories (varying widely along one dimension but not the other), and (C) a `corners' arrangement with examples placed into disparate corners of the space.
+While the tight cluster configuration can be considered PACKER's default generation mode (as it maximizes within-class similarity), the `row' and `column' types can be produced by altering PACKER's attention weights, $w$ (see Equation \ref{eq:similarity}).
+Traditionally, these weights are thought to reflect the diagnostic value of each dimension towards classifying the known category members \citep{nosofsky1984choice,nosofsky1986attention,kruschke1992alcove}, but within a generation context the weights specify the importance of within- and between-class similarity along each dimension.
+For example, if 100\% of attention is allocated along the X-axis, similarity along the Y-axis no longer influences performance: Thus, PACKER will create categories that are more widely distributed along the \textit{Y}-axis, as within-class similarity is not counted along that dimension.
+Further, PACKER's generated categories will be more likely to overlap with known categories along the Y-axis, as, again, between-class similarity is not counted along that dimension.
+Thus, differentially weighting one dimension will result in the generation of row- or column-like categories that are more widely distributed along, and overlap with existing classes along, the ignored axis.
+See Figure \ref{fig:packer-attention} for a depiction of how attention influences PACKER's performance.
+
+\begin{figure}
+    \begin{center}
+    \inputpgf{figs/}{packer-attention-examples.pgf}
+    \caption{PACKER generation of a category `B' example, following exposure to one member of category `A' and one member of category `B'.
+Predictions are shown for different attention settings: \textit{(a)} Increased weighting of the X-axis.
+\textit{(b)} Increased weighting of the Y-axis.
+\textit{(c)} Uniform weighting (identical to Figure \ref{fig:packer-examples}).}
+    \label{fig:packer-attention}
+    \end{center}
+\end{figure}
 
@@
 In the PACKER and copy-and-tweak models, the attention weights, $w$, specify the importance of each dimension in the computation of similarity.
-Traditionally, the attention weights are thought to reflect the diagnostic value of each dimension towards classifying members of the categories \citep{nosofsky1984choice,nosofsky1986attention,kruschke1992alcove}.
+As noted above, attention is usually thought to reflect the diagnostic value of each dimension in classification.
 While there exist methods to find the optimal attention weighting scheme given a classification \citep[see][]{vanpaemel2012using}, for simplicity we may assume that the Alpha and Beta categories are distinct along dimensions that the Betas do not vary on.

### 2017-04-12 Manuscripts/cog-psych/cogpsych.tex
wrote individual differences section. all simulations drafted

--- 2017-04-12 Manuscripts/cog-psych/cogpsych.tex
@@
 \subsection{Individual Differences}
+\label{section:individual-diffs}
 
 As noted in Experiments 1 and 2, we observed a great deal of individual differences in the types of categories generated.
-Within each condition, we observed tightly clustered and widely distributed categories, as well as row and column categories, and so forth.
+Within each condition, we observed tightly clustered and widely distributed categories, as well as row and column categories, and so forth (see Figures \ref{fig:e1-samples} and \ref{fig:e2-samples}).
 The simulations reported above serve to evaluate the models while considering the entire dataset, but a secondary goal of any formal account should be to provide some explanation of how different profiles of performance emerge.
 Many of the individual generation profiles we observed can be targeted within the PACKER framework, simply by tuning the model's parameters.
-In this section, we describe more specifically how the most frequently observed profile can be realized.
+In this section, we describe more specifically how the most frequently observed profiles can be realized.
 
-\nbcnote{manual description of how to get rows, columns, 4 corners, etc}
+By manual inspection, it is evident that the most common profiles of generation consist of: (A) a tightly-distributed `cluster' of examples, (B) `row'- and `column'-like arrangements (varying widely along one dimension but not the other), and (C) a `corners' arrangement with examples placed into disparate corners of the space.
+These four profiles are distinct in terms of the distribution of the generated category along each dimension: whereas the cluster profile is tightly distributed along both dimensions, the row and column profiles are tightly distributed along just one dimension.
+Finally, the corners profile is widely distributed along both dimensions.
 
-\nbcnote{If this is done correctly, we can motivate the use of attention weights below. }
+In the framework proposed by PACKER, the cluster and corners profiles arise based on different prioritization of within-class similarity versus between-class contrast, and the row and column profiles arise based on the prioritization of each dimension in the computation of similarity.
+For example, in the cluster profile, there is a high degree of within-class similarity along both dimensions, whereas in the corners profile there is minimal within-class similarity.
+Thus, PACKER's proposal is that these individual differences arise as a result of different priorities: While the tight cluster configuration can be considered PACKER's `default' mode (as it maximizes within-class similarity), the corners profile can be produced when between-class contrast is put at a higher priority (i.e., $\gamma = 0$).
+
+Likewise, in the row and column profiles, there is a high degree of within-class similarity along one dimension but not the other.
+These differences likely arise due to a differential focus on one dimension over another, and thus they can be produced by changes to PACKER's attention weights, $w$ (see Equation \ref{eq:similarity}).
+Traditionally, the attention weights in exemplar models are thought to reflect the diagnostic value of each dimension towards classifying the known category members \citep{nosofsky1984choice,nosofsky1986attention,kruschke1992alcove}, but within a generation context the weights specify the importance of within- and between-class similarity along each dimension.
+For example, if 100\% of attention is allocated along the X-axis, similarity along the Y-axis no longer influences performance: as a result, PACKER will create categories that are more widely distributed along the Y-axis, as similarity is not counted along that dimension.
+As a general principle, differentially weighting one dimension will result in the generation of categories that are more widely distributed along the ignored dimension, conforming to a row- or column-like arrangement.
+See Figure \ref{fig:packer-attention} for a depiction of how attention influences PACKER's performance.
+
+\begin{figure}
+    \begin{center}
+    \inputpgf{figs/}{packer-attention-examples.pgf}
+    \caption{PACKER generation of a category `B' example, following exposure to one member of category `A' and one member of category `B'.
+Predictions are shown for different attention settings: \textit{(a)} Increased weighting of the X-axis.
+\textit{(b)} Increased weighting of the Y-axis.
+\textit{(c)} Uniform weighting (identical to Figure \ref{fig:packer-examples}).}
+    \label{fig:packer-attention}
+    \end{center}
+\end{figure}
+
+Like in PACKER, changes to the weighting of the dimensions can also be used to produce row- and column-like categories in the copy-and-tweak and hierarchical sampling models.
+Indeed, as copy-and-tweak is simply a special case of the PACKER model, the attention weights operate exactly as described above.
+Within the hierarchical sampling model, the covariance matrix specifying the prior domain distribution, $\Sigma_0$, can be used to similar effect.
+This covariance matrix specifies the amount of variance assumed along each dimension (as well as the dimensional correlations) across the domain of categories.
+The covariance matrix for a newly generated category, $\Sigma_B$, is based on the assumed $\Sigma_0$ as well as the distributions of previously learned categories (see Appendix \ref{ap:hsampling-definition}).
+Thus, the importance of each feature can be coded into $\Sigma_0$ to alter the dimensional variance of generated categories.
+
+However, a key limitation within both in both of these models is the lack of interaction between the distributional structure of a category (in this case, a row-versus-column orientation) and the location of that category within the domain.
+While these models are \textit{capable} of generating row and column categories, there is no mechanism in place to ensure generated categories will be distinct from what is already known.
+In the section below, we explore the interdependence between distributional structure and location in creative generation.
 
@@
 To attempt to explain our findings within the PACKER, copy-and-tweak- and hierarchical sampling models, we conducted simulations using an individual-differences approach.
-Each model was used to simulate each participant's generation separately, with the importance of each dimension set according to the relative range of the participant's generated category along each dimension.
+As noted in Section \ref{section:individual-diffs}, row- and column-like categories can be produced by within model through changes to the weighting of each dimension.
+Given this information, we may use the models to simulate each participant's generation separately, with the importance of each dimension set according to the relative range of the participant's generated category along each dimension.
+
 In the PACKER and copy-and-tweak models, the attention weights, $w$, specify the importance of each dimension in the computation of similarity.
-Traditionally, the attention weights are thought to reflect the diagnostic value of each dimension towards classifying members of the categories \citep{nosofsky1984choice,nosofsky1986attention,kruschke1992alcove}.
 While there exist methods to find the optimal attention weighting scheme given a classification \citep[see][]{vanpaemel2012using}, for simplicity we may assume that the Alpha and Beta categories are distinct along dimensions that the Betas do not vary on.
 Thus, the weighting for a given participant can be computed as:
-
-\nbcnote{Weak spot: attention weighting feels unmotivated.}
 
@@
 % 
-where $\theta_w$ is a free parameter controlling how differences in range correspond to differences in weights (functioning similarity to the $\theta$ parameter in each of the models).
+where $\theta_w$ is a free parameter controlling how differences in range correspond to differences in weights (functioning similarity to the $\theta$ parameter in each of the models), and $\text{range}(k)$ is the range of examples generated by the participant along dimension $k$.
 In our simulations $\theta_w = 1.5$, though similar results can be obtained using a wide range of values.
@@
 Thus, the importance of each feature can be coded into $\Sigma_0$ to alter the dimensional variance of generated categories.
-We incorporated the weights computed in Equation \ref{eq:range-weight} by setting the variance of each dimension according to: $\rho (1-w_k)n$, where $\rho$ is a free parameter specifying the overall assumed variance of the domain, and $n$ is the number of dimensions.
+
+Unlike the PACKER and copy-and-tweak models, the hierarchical sampling model's dimensional variances correspond to the assumed variance of generated categories along each dimension (rather than the inverse of the variance).
+Thus, a slight transformation is needed to incorporate the weights computed in Equation \ref{eq:range-weight}.
+For the hierarchical sampling model, we computed the dimensional variances according to: $\rho (1-w_k)n$, where $\rho$ is a free parameter specifying the overall assumed variance of the domain, and $n$ is the number of dimensions.
 Under this approach, evenly distributed weights correspond to an assumed variance of $\rho$.

### 2017-04-13 Manuscripts/cog-psych/cogpsych.tex
discussion drafted

@@
 
+\nbcnote{I do not know exactly what you mean}
+
 Table \ref{table:global-model-fits} contains the results of this fitting procedure.
@@
     \inputpgf{figs/}{packer-loglike.pgf}
-    \caption{PACKER's fit as a function of sensitivity to between-class contrast (using the $\gamma$ parameter).
+    \caption{PACKER's fit as a function of its prioritization of within-class and between-class similarity (using the $\gamma$ parameter).
 To facilitate comparison, PACKER's other parameters ($c$, $\theta$) were set to the best fitting values obtained for copy-and-tweak in Table \ref{table:global-model-fits}.
@@
 
-\nbcnote{Make the point that we’re explaining category generation using well known ideas, and that exemplar models are viewed as highly principled.}
+The creative generation of novel concepts is an intriguing yet understudied topic in cognitive science.
+While the bulk of prior research on the topic has focused on the classic finding that generated concepts tend to be distributionally similar to known concepts, there has been little work addressing the role of contrast in creative generation: how is it that people are able to create something \textit{different} from what is already known?
+This issue is of fundamental importance: in order to successfully generate something new, it must be different from what is known.
+We developed a novel, exemplar-based model, PACKER, which formally specifies the role of contrast in generation.
+The model proposes that categories are represented as exemplars in a multidimensional psychological space, and generation is constrained both by within-class and between-class similarity: exemplars belonging to the same category should be similar to one another, and exemplars belonging to different categories should not be similar to one another.
 
-\nbcnote{Discuss combining hierarchical sampling and contrast approaches.}
+In addition to an analysis of published data \citep[][Experiment 3]{jern2013probabilistic}, we reported two experiments demonstrating systematic effects of category contrast in creative generation: members of participant-generated categories tended to be highly dissimilar from members of previously-learned categories, and were usually more similar to one another than to members of other categories.
+Finally, we conducted simulations comparing PACKER's account of our results to that of a ``copy-and-tweak'' model (realized as a variant of PACKER with no sensitivity to category contrast), and a hierarchical sampling model designed to explain the classic distributional similarity effect.
+In all simulations, we found that PACKER's sensitivity to contrast considerably enhanced the models account of generation.
 
-\nbcnote{Joe says he'll put importance sampling in future directions.}
+The PACKER model bridges two gaps in the literature.
+First, it provides a quantitative theory of the role of contrast in creative generation, a topic more or less overlooked by previous work.
+The lack of attention to this issue is likely due to the underconstrained notion of what it means for two concepts to be ``different''.
+In the PACKER model, this notion is interpreted using the theory of similarity employed by the exemplar view of categorization \citep{medin1978context,brooks1978nonanalytic,nosofsky1984choice,nosofsky1986attention}.
+This theory of similarity, tracing its foundations to classic research in stimulus generalization \citep{attneave1950,shepard1957stimulus,shepard1987toward}, has had broad influence in many fields within cognitive science.
+Thus, beyond providing a formal specification of category contrast, the PACKER model also aligns the relatively understudied topic of creative generation more closely with the well-developed literature on human categorization.
+The success of the PACKER model indicates that core theories from the categorization literature may offer a strong account of how people generate new concepts.
+
+The success of the PACKER model yields reveals keys insights about the nature of creative generation.
+First, by measuring PACKER's fit as a function of its prioritization of within- and between-class similarity, we observed that considering either constraint exclusively results in a relatively low-quality account.
+Instead, PACKER's best results were obtained when both constraints are considered, indicating that human learners do not generate novel concepts exclusively on the basis of within-class similarity or between class-contrast.
+This finding mirrors our behavioral results: it is clear that both constraints influence creative generation.
+
+This finding is extended by our observation of the broad interdependence between the distributional structure (feature variance, correlation) and physical instantiation (location within the domain space) of generated categories.
+In Experiment 2, we found that the unoccupied regions of the domain influenced the distributional structure of categories, and in Experiments 1 and 2 we observed that generated categories assume different distributional structures depending on how they contrast with known categories.
+This interdependence is straightforwardly explained under PACKER's dual-constraint proposal: after first establishing a location for their generated category, participants distribute exemplars in order to minimize between-class similarity while not drastically decreasing within-class similarity.
+
+However, PACKER does not provide a full account of what is known about category generation.
+Most notably, in this paper we have not evaluated the model's ability to explain the classic finding that generated categories tend to share distributional commonalities with previously learned categories \citep[see][]{jern2013probabilistic,ward1994structured,ward1995s,ward2002role,smith1993constraining,marsh1999inadvertent}.
+While we successfully replicated this effect in Experiment 1, we also found that its influence was limited in comparison to the fundamental constraints imposed by category contrast.
+Even within Experiment 1, we found systematic inconsistencies: by generating exemplars into unoccupied regions of the space, participants who learned an `XOR' category, composed of members that are widely distributed along both features and are positively correlated in space, tended to generate categories with an opposite (negative) correlation.
+More generally, PACKER's success over the hierarchical sampling model indicates that the emulation of distributional structure exerts only a limited influence in comparison to category contrast.
+
+Nonetheless, these classic effects are a core element of the phenomenology of creative generation, and it is worth noting that PACKER does not include any mechanisms explain them.
+Instead, through the development and evaluation of the PACKER model, we have sought to add new elements into such a phenomenology: specifically, the broad and strong influence of category contrast, and the interdependence between category location and distributional structure.
+It may be possible to combine the hierarchical sampling approach proposed by \cite{jern2013probabilistic} with PACKER's underlying claims to obtain a ``best of both worlds'' model, capable of explaining the role of contrast in creative generation, as well as the emulation of distributional structure.
+However, as noted in the introduction, the incorporation of category contrast is antithetical to the core principles of the hierarchical Bayesian approach.
+Thus, while such a model may provide an excellent fit to the data, its theoretical composition would be more or less of an amalgam, and thus it may not succeed in other areas of model evaluation (such as predicting and explaining novel phenomena).
+
+\nbcnote{I'm just spitballing there...}
+
+As it stands, however, PACKER's mathematical basis is subject to a wide variety of interpretations, leading to distinct views on the nature of creative generation. 
+\nbcnote{That sounds like a good spot to include importance sampling}
+
+\section{Conclusions}
 
@@
     \caption{Box-plots of the distributional statistics from the categories generated in Experiment 2. }
-    \nbcnote{Note the huge amount of variance in the range data, as well as the big Y-range difference in the medians.}
+    \nbcnote{Note the huge amount of variance in the range data, as well as the big median difference in Y-range.}
 

### 2017-04-16 Manuscripts/cog-psych/cogpsych.tex
pass back to Nolan

@@
 
-Creating Something Different: Similarity and Contrast in Concept Generation.
+Creating Something Different: Similarity and Contrast in Concept Generation
 
@@
 \\[\baselineskip]
-\textsuperscript{1}University of Wisconsin-Madison, Department of Psychology, Madison, WI, USA
+\textsuperscript{1}Department of Psychology, University of Wisconsin-Madison, Madison, WI, USA
 \textsuperscript{2}Department of Psychology, Binghamton University, Binghamton, NY, USA
@@
 They are fascinating and important capabilities because they have produced some of the greatest scientific breakthroughs that revolutionized the world.
-Unfortunately, scientific breakthroughs are rare and the result of a great deal of cognitive effort, which it difficult to investigate the underlying cognitive representations and processes responsible for them using standard experimental methodology or formalize in computational models.
-Although we tend to focus on the most salient products of these processes (e.g., scientific breakthroughs), every person is likely to have generated a novel sentence, thought, and/or drawing.\footnote{\jlanote{We need citations/examples for this.
-In a compositional system, it's pretty clear this has to be the case, but I'd like to use someone else as a reference rather than arguing this myself.}} This observation provides a fantastic opportunity: By examining how people create new concepts in a domain amenable to formalization in computational models, we can investigate (some aspects of) creativity, innovation, and imagination scientifically.
+Unfortunately, technological and scientific breakthroughs are rare and the result of a great deal of cognitive effort, which makes it difficult to investigate the underlying cognitive representations and processes responsible for them using standard experimental methodology or to formalize in computational models.
+Although we tend to focus on the most salient products of these processes (e.g., scientific breakthroughs), every person is likely to have generated many novel sentences, thoughts, and/or drawings given the combinatorial explosion of possible statements given a reasonably-sized set of primitive elements and combination rules in language, thought, and perception \citep{goldstone2003}.
+This observation provides a fantastic opportunity: By examining how people create new concepts in a domain amenable to formalization in computational models, we can investigate (some aspects of) creativity, innovation, and imagination scientifically.
 
@@
 
-Previous work has focused on one key aspect of category generation: People tend to create new categories that have similar {\em statistical regularities} to previously known categories \citep{jern2013probabilistic,ward1994structured}.
+Previous work in categorization has identified a number of basic principles underlying human category performance that can be explained by appealing to the structural properties of natural categories, such as  correlations between the features of category members and the relation of the differences of items within the same category to those in different categories\cite{roschmervis1975,regier2006}.
+Inspired by this work in categorization, previous work explored one key aspect of category generation: People tend to create new categories that have similar {\em statistical regularities} to previously known categories \citep{jern2013probabilistic,ward1994structured}.
 Although this is an important characteristic of generating new categories, it cannot be the only one.
@@
 
-In this article, we propose a new constraint that guides category generation: ``being different'' or contrasting from other categories in the relevant domain.
+Inspired by the importance of contrast in specific categorization domains \citep[e.g., color][]{regier2007}, we propose a new constraint that guides category generation: ``being different'' or contrasting from other categories in the relevant domain.
 To some extent, this constraint has been implicitly assumed in much previous work: the idea that a new category should be ``different'' is vague, as there are many ways it could be different from a previously observed category.
@@
 Problematically, however, the experimental paradigms employed in these studies were relatively uncontrolled, precluding the development of formal approaches.
-Specifically, participants in these studies are typically asked to draw exemplars of novel categories by hand, and so it is difficult to apply formal models to the data.
+Although this previous work has been incredibly useful in providing a conceptual sketch of generation theories, the fine and hand-drawn responses precludes formal analyses.
 
@@
 This finding is notable, as it demonstrates that category generation can be studied in a well-known and highly controlled experimental paradigm.
-
-% % ----- This point is not true -----
-% Note that generated exemplars for the novel category automatically contrasted from previous categories because the participants selected the size and saturation, but the exemplar's hue was set (and fixed) to a novel value.
-As such, hue was ignored in their analyses and may have satisfied any constraints contrast may have on category generation.
 
@@
 
-It is worth noting that the literature reviewed above provides a somewhat limited view of category generation.
-The main concern of the published experiments has been on the distributional correspondences between learned and generated categories, and as a result most of the modeling efforts have been towards explaining those effects.
+From the perspective of the categorization literature, prior work on category generation has only explored one of the possible constraints that guide category generation.
+The main concern of the published experiments has been on the distributional correspondences between learned and generated categories, and as a result most of the computational, theoretical, and empirical efforts have been towards explaining those effects.
 In this paper, we investigate another important constraint on the creative use of conceptual knowledge: category contrast.
 To generate a novel concept, individuals must produce something that is in some capacity \textit{different} from what they already know.
-Thus, contrast is a fundamental constraint on creative generation: new concepts must firstly be different from existing ones.
+Thus, contrast is a fundamental constraint on creative generation: new concepts must be different from existing ones.
 
@@
 The hierarchical sampling model developed by \cite{jern2013probabilistic} assumes that differences between generated category are only due to random variation.
-The model only proposes that generated categories are sampled from the same underlying domain distribution as observed categories, and will thus share a common distributional structure.
+The model assumes that generated categories are sampled from the same underlying domain distribution as observed categories, and will thus share a common distributional structure.
 The authors do not make predictions about the \textit{location} of the category within the domain (the perceptual instantiation of category members).
 Indeed, under a strict interpretation of their model, the most probable new category to be generated is located in {\em exactly} the same location and distributional information as the given category.
-However, this is not simply an issue with their model -- Hierarchical Bayesian models assume the same underlying distribution generates all of the categories and thus, any differences between categories is due to {\em noise} and should not be {\em systematic}.
+However, this is not simply an issue with their model but with the broader class of standard hierarchical Bayesian models.
+These models assume that at some point of the latent generative process the same underlying distribution generates all of the categories and thus, any differences between categories is due to {\em noise} and should not be {\em systematic}.
 The best a hierarchical Bayesian model can do at capturing contrast is to assume that the new category is placed uniformly at random over stimulus space.
@@
 
-\subsubsection{Contrast Effects in Prior Work}
+\subsubsection{Novel Analyses Demonstrating Contrast Effects in Prior Work}
 
-Although existing accounts of creative generation broadly overlook the role of category contrast in determining what is novel versus familiar, these effects are undoubtedly ubiquitous within the published data.
-In each study, participants were \textit{successful} in creating something new: the question is about \textit{how} they were able to do so.
-To provide concrete evidence of  effects of category contrast within existing datasets, we conducted an analysis of Experiment 3 from \cite{jern2013probabilistic}.
+Although existing accounts of creative generation broadly overlook the role of category contrast in determining what is novel versus familiar, if contrast affects category generation, we should find these effects within the experimental results of previous work.
+To test whether there is evidence of category contrast affecting category generation, we conducted a novel analysis of Experiment 3 from \cite{jern2013probabilistic}.
 
@@
 As noted above, \cite{jern2013probabilistic} found that the generated categories tended to follow the distributional properties of the experimenter-defined categories: Generated categories were tightly distributed along the hue feature, and possessed the same saturation-size correlations as in the learned categories.
-\cite{jern2013probabilistic}, however, did not provide an analysis of how the generated categories \textit{differed} from the experimenter-defined categories.
+\cite{jern2013probabilistic}, however, did not analyze or discuss how the generated categories \textit{differed} from the experimenter-defined categories.
 
-Because each experimenter-defined category in the \cite{jern2013probabilistic} experiment possessed a distinct hue, it is sensible to propose that participants would make their generated category distinct by using hues that differ from those present in the experimenter-defined category: If creative generation is influenced by category contrast, the hues of generated categories should be systematically different from those of the experimenter-defined categories.
-Problematically, however, stimulus hue was measured on an circular (HSV) scale with limits $[0,1]$: both 0 and 1 corresponded to a `red' hue, and all other possible hues were represented between these values.
-By consequence, proximity on the hue dimension does not directly correspond to perceptual similarity, and we cannot simply compare the hue values to determine the extent of the differences.
-\nbcnote{Maybe add a cite for proximity-similarity distinction?}
+Because each experimenter-defined category in the \cite{jern2013probabilistic} experiment possessed a distinct hue, it is sensible to propose that participants would generate a category with a hue distinct from the experimenter-provided categories, which all had a distinct hue shared by all members in the category.
+f creative generation is influenced by category contrast, the hues of generated categories should be systematically different from those of the experimenter-defined categories.
+Problematically, however, stimulus hue was presented using the Hue-Saturation-Value (HSV) color space, which is device-dependent and not perceptually normed such that perceived color similarity corresponds to proximity in the color space (as opposed to a color space such as CIELAB that is device-independent and equidistant sets of points correspond to pairs of colors that have the same perceptual similarity; \citet{wyszecki1967}).
+Further, they did not calibrate their monitor.
+Thus, we cannot know the precise values of colors presented to participants.
+As \cite{jern2013probabilistic} were interested in relations between the saturation and length of generated stimuli, this is not an issue for their analyses and results.
+But it means we cannot know precisely which colors were displayed to participants.
 
-To overcome this issue, we categorized all possible hues into one of eight uniformly-spaced color groups: \{\textit{Red}: $0-0.063$, $0.938-1$, \textit{Yellow}: $0.063-0.188$, \textit{Yellow-Green}: $0.188-0.313$, \textit{Green-Teal}: $0.313-0.438$, \textit{Teal}: $0.438-0.563$, \textit{Teal-Blue}: $0.563-0.688$, \textit{Purple}: $0.688-0.813$, \textit{Pink}: $0.813-0.938$\}.
+Although we cannot know precisely which colors were displayed or perceived, we can still analyze their results from a coarse perspective to see whether there is preliminary support for contrast in them.
+To do so, we binned all possible hues into one of eight uniformly-spaced color groups: \{\textit{Red}: $0-0.063$, $0.938-1$, \textit{Yellow}: $0.063-0.188$, \textit{Yellow-Green}: $0.188-0.313$, \textit{Green-Teal}: $0.313-0.438$, \textit{Teal}: $0.438-0.563$, \textit{Teal-Blue}: $0.563-0.688$, \textit{Purple}: $0.688-0.813$, \textit{Pink}: $0.813-0.938$\}.
 In the \cite{jern2013probabilistic} experiment, the hue of each experimenter-defined category was selected from one of six possible values, each of which falls into one of the color groups above.
@@
 These results can be compared to the predictions of a Binomial model, which proposes that participants generate hues at random.
-That is, if hue selection is not systematic, the probability that any given example will lie in the same color group as an experimenter-defined category is given by a Binomial distribution with $p = 2/8$, as there were two experimenter-defined categories and eight possible color groups.
-Chi-square goodness-of-fit tests reveal that the observed distribution in each condition is highly inconsistent with the random-hue model (all $\tilde{\chi}^2>200$, $p<0.001$): Participants tended to generate items that were perceptually distinct from the categories they had learned, and were less likely to generate hues possessed by members of the experimenter-defined categories.
+That is, if hue selection is not systematic, the probability that any given example will lie in the same color group as an experimenter-defined category is given by a Binomial distribution with $p = 2/8=1/4$, as there were two experimenter-defined categories and eight possible color groups.
+Chi-square goodness-of-fit tests reveal that the observed distribution in each condition is highly inconsistent with the hues being chosen at random (all $\tilde{\chi}^2>200$, $p<0.001$): Participants tended to generate items that were perceptually distinct from the categories they had learned, and were less likely to generate hues possessed by members of the experimenter-defined categories.
 
-These results are notable in that they provide some early, concrete evidence that creative generation is influenced by category contrast.
-Taken alongside the analyses reported by \cite{jern2013probabilistic}, our results suggest that generated categories tend to be distinct from \textit{and} distributionally similar to what is already known.
+Re-analyzing the results of \cite{jern2013probabilistic} provides preliminary support that contrast plays a role in category generation.
+Taken alongside the analyses reported by \cite{jern2013probabilistic}, our analysis suggests that generated categories tend to be distinct from \textit{and} distributionally similar to what is already known.
 However, it is worth noting that our analysis is still limited: the color groups defined above are somewhat arbitrary, and it is not clear that our color grouping is consistent with psychological color boundaries.
@@
 
-As noted above, the constraint that new concepts should differ from what is already known has been largely been overlooked in previous work.
+As noted above, the constraint that new concepts should differ from what is already known has been largely overlooked in previous work.
 This is no doubt in part due to the vague definition of what it means for a concept to be ``different'': A generated category may be different from what is already known in any number of respects.
-Towards formalizing the role of contrast in creative generation, we developed the PACKER  ({\em Producing Alike and Contrasting Knowledge using Exemplar Representations}) model which explains category generation as a balance between two fundamental constraints: generated categories should not be similar to known categories, and exemplars within each category should be similar to one another.
+Towards providing a more precise definition of the role of contrast in creative generation, we formalized contrast in a novel exemplar model, PACKER  ({\em Producing Alike and Contrasting Knowledge using Exemplar Representations}).
+PACKER explains category generation as a balance between two fundamental constraints: generated categories should not be similar to known categories, and exemplars within each category should be similar to one another.
 These ideas are implemented within the well-studied exemplar framework: the PACKER model is an extension of the influential Generalized Context Model of category learning \citep[GCM;][]{nosofsky1984choice,nosofsky1986attention}.
@@
 This is achieved by aggregating similarity across known exemplars differently according to class membership.
-The aggregated similarity $a$ between generation candidate $y$ and stored exemplars $x$ is given by:
+The aggregated similarity $a(y,x)$ between generation candidate $y$ and stored exemplars $x$ is given by:
 
@@
 The $\gamma$ parameter thus specifies a wide breadth of possible approaches; by fitting it to a dataset, one can describe the relative roles of between-class contrast and within-class similarity in generation.
-See Figure \ref{fig:packer-examples} for an illustration of how the $\gamma$ parameter can be used.
+See Figure \ref{fig:packer-examples} for an illustration of how $\gamma$ controls the relative influence of within category similarity and contrast to other categories.
 
@@
 The probability that a given candidate $y$ will be generated is evaluated using an Exponentiated \citet{luce1977choice} choice rule.
-Candidates with greater values of $a$ are more likely to be generated than candidates with smaller values:
+Candidates with greater values of $a(y,x)$ are more likely to be generated than candidates with smaller values:
 % 
 \begin{equation}
-p(y) = \dfrac
+p(y \mid x) = \dfrac
 { \exp  \left \{ \theta \cdot a \left( y, x \right) \right \} } 
@@
 Likewise, when $\gamma = 0$, PACKER represents a contrast-only generation mode, relying exclusively on contrast when generating new categories.
-Finally, when $f(x_j) = -1$ for all $x_j$ (regardless of class membership), a ``pure-packing'' approach is yielded, generating items in any unoccupied area of the domain.
+Finally, when $f(x_j) = -1$ for all $x_j$ (regardless of class membership), a ``pure-packing'' approach is yielded, generating items in unoccupied areas of the domain.
 Thus, the proposed framework may be used to describe a wide variety of qualitatively distinct generation strategies.
@@
 \label{section:copytweak-packer}
-The PACKER model is some sense similar to the copy-and-tweak model reported by \cite{jern2013probabilistic}: both models are exemplar-based, and both models generate new items according to their similarity to known members of the target class.
-PACKER diverges from the copy-and-tweak model by including a contrast mechanism, enabling generation according to dissimilarity to members of opposing categories.
+The PACKER model is similar to the copy-and-tweak model reported by \cite{jern2013probabilistic}: Both models are exemplar-based, and both models generate new items according to their similarity to known members of the target class.
+However, PACKER diverges from the copy-and-tweak model by including a contrast mechanism, enabling generation according to dissimilarity to members of opposing categories.
 By consequence, copy-and-tweak can be realized as a parameterization of the PACKER model that is insensitive to category contrast.
@@
 This model fits within the exemplar-based category generation framework defined above, under the constraint that $\gamma = 1$, and is a continuous-dimension adaptation of the model tested by \cite{jern2013probabilistic}.
-By formalizing a model family where PACKER and copy-and-tweak are different parameterizations of models within the same framework, the comparison between PACKER and copy-and-tweak provides a test of the explanatory value of the contrast mechanism: the account provided by copy-and-tweak will only equal that of PACKER if the contrast mechanism does not offer an advantage (i.e., if $\gamma < 1$ significantly improves model fits).
+By formalizing a model family where PACKER and copy-and-tweak are different parameterizations of models within the same framework, the comparison between PACKER and copy-and-tweak provides a test of the explanatory value of the contrast mechanism: The account provided by copy-and-tweak will only equal that of PACKER if the contrast mechanism does not offer an advantage (i.e., if $\gamma < 1$ significantly improves model fits).
+Note that the purpose of the article is to explore and formally analyze the role of contrast in category generation and thus, we leave extending PACKER to incorporate distributional factors \citep[as explored by][]{jern2013probabilistic} for future work.
 
@@
 The manner in which generated categories differ from known ones is, nonetheless, poorly understood: Existing theories do not make strong predictions about how creatively generated concepts should systematically differ from existing ones.
-Above, we provided some early evidence of the influence of category contrast in published data \citep[][Experiment 3]{jern2013probabilistic}, and we introduced a novel, exemplar-based model formalizing the roles of similarity and contrast in creative generation.
+Above, we provided encouraging initial support that contrast influences category generation \citep[][Experiment 3]{jern2013probabilistic}, and we introduced a novel, exemplar-based model formalizing the roles of similarity and contrast in creative generation.
 
-In the sections below, we present two experiments demonstrating systematic effects of category contrast on creative generation.
+In the sections below, we present two experiments demonstrating systematic effects of category contrast on creative generation inspired by factors influencing how PACKER generates new categories.
 Our experiments are based on \cite{jern2013probabilistic}'s paradigm: participants are first exposed to a single, experimenter-defined category, and are then asked to generate members of a new category.
-We then we report formal simulations comparing PACKER's account of our results to that of the hierarchical sampling and copy-and-tweak models developed by \cite{jern2013probabilistic}.
+We then report formal analyses comparing PACKER's account of our results to that of the hierarchical sampling and copy-and-tweak models developed by \cite{jern2013probabilistic}.
 
+\jlanote{I forget, but did JK13 explicitly have in their instructions that the new category be different? If so, we should mention the issue there as another problem with their study for us.}
 \section{Experiment 1}
@@
 To begin our investigation, we sought to extend the early evidence obtained from our analysis of the \cite{jern2013probabilistic} data, under a variety of learning conditions and using more standard stimulus materials.
-We developed an artificial, two dimensional domain of squares, varying in color and size (see Figure \ref{fig:e1-conditions}, panel A).
-These features have been used in numerous classification learning studies \cite[e.g.,][]{conaway2016similar,conaway2016generalization,shepard1961learning,nosofsky1994comparing}.
-Unlike those used in the \cite{jern2013probabilistic} experiments, proximity along these dimensions aligns more directly with perceptual similarity, allowing us to more directly evaluate the role of category contrast in creative generation.
-In order to extend the evidence provided by the \cite{jern2013probabilistic} data, we began by testing effects of category contrast under a variety of learning conditions.
-This was achieved by training participants on categories possessing qualitatively different distributional structures, as shown in Figure \ref{fig:e1-conditions}.
+We used an artificial stimulus design, two dimensional domain of squares, varying in color and size (see Figure \ref{fig:e1-conditions}, panel A).
+These dimensions have been used in numerous classification learning studies \cite[e.g.,][]{conaway2016similar,conaway2016generalization,shepard1961learning,nosofsky1994comparing}.
+Unlike those used in the \cite{jern2013probabilistic} experiments, proximity along these dimensions aligns more directly with perceptual similarity, allowing us to evaluate the role of category contrast in creative generation more precisely.
+To extend the evidence provided by the \cite{jern2013probabilistic} data, we tested the effects of category contrast after learning one category from a set of category structures.
+We did so by training participants on categories possessing qualitatively different distributional structures, as shown in Figure \ref{fig:e1-conditions}.
 
@@
 The assignment of perceptual features (color, size) to axes of the domain space (x, y), as well as the direction of variation along each axis (e.g., $dark \rightarrow light$ or $light \rightarrow dark$) was counterbalanced across participants.
+\jlanote{we should probably mention the discrepancy between the number of participants per condition and explain why it isn't anything to worry about.}
 
@@
 Participants were instructed to learn as much as they can about the Alpha category, and that they would answer a series of test questions afterwards.
-On each trial, a single Alpha category exemplar was presented, and participants were given as much time as they desired before moving on.
+On each trial, a single Alpha category exemplar was presented, and participants were given as much time as they desired to observe it before moving on to the next trial.
 Each block consisted of a single presentation of each of the members of the Alpha category, in a random order.
@@
 To begin, it is worth noting that we observed a substantial degree of individual differences in our data.
-In Figure \ref{fig:e1-samples} we have plotted sample data from several participants, from which it evident that different participants assumed qualitatively distinct approaches to category generation.
+In Figure \ref{fig:e1-samples} we have plotted sample data from several participants, from which it evident that different participants generated qualitatively different category structures.
 In this section we will focus on what can be learned from analyzing the data in aggregate, but in later sections we will explore how individual differences can be explained.
@@
     \inputpgf{figs/}{e1-samples.pgf}
-    \caption{Sample categories generated in Experiment 1. Representative samples from common generation profiles are shown.}
+    \caption{Sample categories generated by participants in Experiment 1. Representative samples from common generation profiles are shown.}
     \label{fig:e1-samples}
@@
 To evaluate the role of contrast, we computed the number of times each stimulus was generated, as a function of its average city-block distance from members of the experimenter-defined ``Alpha'' category.
-These data, shown in Figure \ref{fig:e1-distanceplots}, reveal a clear pattern: examples that are more distant from members of the experimenter-defined categories are more likely to be generated into a new category.
+These data, shown in Figure \ref{fig:e1-distanceplots}, reveal a clear pattern: Examples that are more distant from members of the experimenter-defined categories are more likely to be generated into a new category.
 
@@
     \nbcnote{Need to check on how nice the journal is about color figs!}
-
+	\jlanote{Is there something systematic about XOR for how the generated category varies for smaller/larger within-category distances. Also, should switch "class" to category here.}
     \label{fig:e1-distanceplots}
@@
 
-Figure \ref{fig:e1-distanceplots} also depicts, for each participant, the average amount of distance between members of the generated category (\textit{within-class} distance) against the average amount of distance between members of the generated and experimenter-defined category (\textit{between-class} distance).
+Figure \ref{fig:e1-distanceplots} also depicts, for each participant, the average distance of members within the generated category (\textit{within-class} distance) against the average distance between members of the generated and experimenter-defined category (\textit{between-class} distance).
 These data also reveal a systematic pattern: the majority of participants generated categories with more between-class distance than within-class distance.
@@
 
-The secondary goal of this experiment was to examine whether we replicate the classic result that generated categories often possess the same distributional properties as previously-known categories.
+A secondary goal of this experiment was to examine whether we replicate the classic result that generated categories often possess the same distributional properties as previously-known categories.
 For each generated category, we computed the category range along each axis (X, Y), as well as the correlation between features.
@@
 Likewise, in each condition participants generated categories possessing strongly positive, neutral, and strongly negative correlations between the dimensions.
-Comparing the distributional statistics between conditions yields a broad-yet-incomplete replication of the classic effect.
+Comparing the distributional statistics between conditions yields a broad-yet-misleading replication of the classic effect.
 
@@
 This latter effect was not expected; however, the key finding is that categories from the Cluster condition tended to be more tightly distributed along the X-axis.
-
+\jlanote{Was XOR slightly longer than row?}
 Likewise, categories generated in the Row condition had less Y-axis range compared to Cluster, $t(123) = 4.57$, $p < 0.001$ and XOR, $t(117) = 9.26$, $p < 0.001$, and categories from the Cluster condition had less Y-axis range compared to XOR, $t(120) = 3.95$, $p < 0.001$.
-However, whereas the correlations in the Cluster and Row conditions were not systematically positive or negative ($ps > 0.1$), the generated categories in the XOR condition tended to possess \textit{negatively} correlated dimensions, $t(57) = 2.04$, $p = 0.046$.
+As expected, the correlations in the Cluster and Row conditions were not systematically positive or negative ($ps > 0.1$).
+However, the generated categories in the XOR condition tended to possess \textit{negatively} correlated dimensions, $t(57) = 2.04$, $p = 0.046$.
 This finding is notable, as it is the opposite of what would be expected, assuming learners are emulating the distributional structure of the experimenter-defined class (which possesses perfectly positively correlated features).
-Further, it is expected by our proposal that contrast is a fundamental principle for category generation.
+Thus, there is more to category generation than the distributional structure of other categories in the domain!
+Further, as we will discuss in more detail in the model-based analysis section, it is expected by our proposal that contrast is a fundamental principle for category generation.
 
@@
 \section{Experiment 2}
-
-To test whether category contrast influences the distributional structure of generated categories, we repeated Experiment 1 using two new category types, depicted in Figure \ref{fig:e2-conditions}.
-The category types possess an identical distributional structure (both are tight clusters of examples with a neutral feature correlation), and only differ slightly in their Y-axis position: the `Bottom' category lies in the bottom-center of the space, and the `Middle' category lies in the center.
+To test whether category contrast influences the distributional structure of generated categories, we created two new category types (depicted in Figure ~\ref{fig:e2-conditions}) that do not differ in distributional structure, but we expect participants would generate different categories given each of them.
+The observed categories possess an identical distributional structure (both are tight clusters of examples with a neutral feature correlation), and only differ slightly in their Y-axis position: the `Bottom' category lies in the bottom-center of the space, and the `Middle' category lies in the center.
 The distributional equality of these conditions is key to the design of the experiment: if the structure of previously learned categories were the only influence on the structure of generated categories, we should observe no difference between these two conditions with respect to distributional structure.
@@
 
+\jlanote{I think we should show and explain that corner behavior mess it up (or whatever the reason is why it happens).
+I think it is more suspicious to not show it and from what I remember, it doesn't really affect our hypotheses.}
+
 Because the conditions differ only in the Y-axis position of the experimenter-defined category, we then compared the conditions in terms of the frequency with which participants generated examples above and below the categories.
@@
 In this section, we report the results of simulations with formal models aimed at explaining our observations.
-Specifically, we present simulations from the PACKER model, as well as a `copy-and-tweak' model (discussed in section \ref{section:copytweak-packer}), defined as a variant of PACKER with the $\gamma$ parameter held constant.
+Specifically, we present simulations from the PACKER model, as well as a `copy-and-tweak' model (discussed in Section \ref{section:copytweak-packer}), defined as a variant of PACKER with the $\gamma$ parameter constrained to be one.
 The comparison of these two models serves to highlight the explanatory role of contrast within PACKER's framework: if contrast affords little explanatory advantage, then the two accounts should produce an equally strong account.
 We also present simulations from an implementation of the hierarchical sampling model proposed by \cite{jern2013probabilistic}, described in-depth in Appendix \ref{ap:hsampling-definition}.
-The comparison between the hierarchical sampling model and PACKER is meant to emphasize the role of emulation of distributional structure: whereas PACKER is insensitive to the distributional structure of learned categories (relying only on within- and between-class similarity), the hierarchical sampling model generates categories exclusively on the basis of knowledge of how existing classes are distributed.
+The comparison between the hierarchical sampling model and PACKER is meant to emphasize the necessity of contrast and demonstrate it cannot be explained by emulating of distributional structure: whereas PACKER is insensitive to the distributional structure of learned categories (relying only on within- and between-class similarity), the hierarchical sampling model generates categories exclusively on the basis of knowledge of how existing classes are distributed.
 Our approach in this section is to first broadly evaluate and compare the quality of each model's account to our entire dataset (Experiments 1 and 2 combined), then we more specifically describe the strengths and weakness of each model's account.
@@
 \centering
-\caption{Results of model-fitting to the combined datasets from Experiments 1 and 2.}
+\caption{Results of model-fitting to the combined datasets from Experiments 1 and 2. Note that lower AIC values correspond to better model fits (adjusted for number of parameters)}
 \label{table:global-model-fits}
@@
 
+\jlanote{I did them.
+(it's just 2 * diff in log likelihood is $\chi^2$ distributed with df equal to difference in the number of parameters) You should check BIC and $AIC_c$ (easy to find formula for them on Wikipedia).
+I don't think we should list them all, but it'd be good to mention our results are robust}
+
 Table \ref{table:global-model-fits} contains the results of this fitting procedure.
-Due to the uneven number of fitted parameters among the models, we compare the model fits using the Akaike Information Criterion \citep[AIC;][]{akaike1974new}.
+Due to the uneven number of fitted parameters among the models, we compare the model fits using the Akaike Information Criterion \citep[AIC;][]{akaike1974new}, where smaller values correspond to better fits (discounted by the number of parameters).
 Table \ref{table:global-model-fits} contains the AIC values of each model's best fit, as well as the corresponding log-likelihood ($L$) and the best-fitting parameter values.
@@
 
-A key distinction between PACKER and copy-and-tweak, as well as the the hierarchical sampling model, is that, of the three models, only PACKER is capable of making strong predictions about the location of new category members when the target class is entirely novel (i.e., no member of the category has been observed).
+A key distinction between PACKER and copy-and-tweak, as well as the hierarchical sampling model, is that, of the three models, only PACKER is capable of making strong predictions about the location of new category members when the target class is entirely novel (i.e., no member of the category has been observed).
 Under these circumstances, there are no examples to copy, and thus the copy-and-tweak model predicts that items are generated at random.
@@
 We conducted an identical set of simulations as above, excluding this trial (leaving 915 responses in the dataset): Again, PACKER ($L = -3390$, $AIC = 6786$) achieved better fits than the copy-and-tweak ($L = -3579$, $AIC = 7162$) and hierarchical sampling ($L=-3612$, $AIC = 7232$) models.
+Because copy-and-tweak is nested within PACKER, we can use a likelihood ratio test to compare the two models.
+PACKER explains the aggregate data significantly better than copy-and-tweak ($\chi^2(1) = 748, p < 0.001$ for all data and $\chi^2(1) = 378, p < 0.001$ excluding the first example).
+Thus, participant results support that category generation includes a contrast mechanism.
 
-Through comparison with the copy-and-tweak model, Figure \ref{fig:packer-loglike} more clearly demonstrates the explanatory gains yielded by PACKER's sensitivity to category contrast.
+Through comparison with the copy-and-tweak model, Figure \ref{fig:packer-loglike} more clearly demonstrates the robustness of the explanatory gains yielded by PACKER's category contrast mechanism.
 Plotted is PACKER's log-likelihood as a function of $\gamma$ parameter.
@@
 In the section below, we explore the interdependence between distributional structure and location in creative generation.
-
+\jlanote{It's a combination of previous category location and where the target category is being generated, right?}
 \subsection{Category Location vs. Distributional Structure}
@@
 However, because many stimuli were infrequently generated (such items near members of the `Alpha' category), we cannot simply compute the empirical average, as infrequently generated stimuli would be likely to show artificially strong differences.
-Instead, we aggregated the scores using an empirical Bayesian analysis \citep{robbins1964empirical}: the aggregated score for each stimulus was viewed as a normal distribution with a $\mu_0 = 0$ and $\sigma$ set based on the observed scores $x$ for that stimulus.
+Instead, we aggregated the scores in an empirical Bayesian manner \citep{robbins1964empirical}: the aggregated score for each stimulus was viewed as a normal distribution with a $\mu_0 = 0$ and $\sigma$ set based on the observed scores $x$ for that stimulus.
 The aggregated difference $\mu_x$ was then computed as a Bayesian update:
@@
 \nbcnote{Do you think we need to report the Gaussian smoothing?}
-
+\jlanote{should mention it in the caption or a footnote}
 \begin{figure}[p]
@@
     \label{fig:range-diff-gradients}
+  	\jlanote{This is beautiful. I'd like to add one thing to it. On the left margin have something that says "Experiment 1" and groups the top 3 rows, and then "Experiment 2" for the bottom two.}
     \end{center}
@@
 where $\theta_w$ is a free parameter controlling how differences in range correspond to differences in weights (functioning similarity to the $\theta$ parameter in each of the models), and $\text{range}(k)$ is the range of examples generated by the participant along dimension $k$.
-In our simulations $\theta_w = 1.5$, though similar results can be obtained using a wide range of values.
+In our simulations $\theta_w = 1.5$, though the results are robust and similar for other $\theta_w$ values.
 The resulting $w$ values are thus inversely proportional to the range of generated categories along each dimension, with less range corresponding to greater weighting.
@@
 Unlike the PACKER and copy-and-tweak models, the hierarchical sampling model's dimensional variances correspond to the assumed variance of generated categories along each dimension (rather than the inverse of the variance).
-Thus, a slight transformation is needed to incorporate the weights computed in Equation \ref{eq:range-weight}.
-For the hierarchical sampling model, we computed the dimensional variances according to: $\rho (1-w_k)n$, where $\rho$ is a free parameter specifying the overall assumed variance of the domain, and $n$ is the number of dimensions.
-Under this approach, evenly distributed weights correspond to an assumed variance of $\rho$.
+Thus, a different transformation is appropriate for incorporating the weights computed in Equation \ref{eq:range-weight}.
+For the hierarchical sampling model, we computed the dimensional variances according to: $\lambda (1-w_k)d$, where $\lambda$ is a free parameter specifying the overall assumed variance of the domain, and $d$ is the number of dimensions.
+Under this approach, evenly distributed weights correspond to an assumed variance of $\lambda$.
 Likewise, larger values of $w$, which are produced when the generated category is tightly distributed along one dimension, correspond to smaller assumed variances.
@@
 \nbcnote{This works for 2D, but not 3D and beyond (as $1-w_k$ would not produce a valid weighting).}
+\jlanote{I renamed $\rho$ because $\rho$ is reserved for the correlation parameter of a covariance matrix (especially a $2 \times 2$).}
 
@@
 The lack of attention to this issue is likely due to the underconstrained notion of what it means for two concepts to be ``different''.
-In the PACKER model, this notion is interpreted using the theory of similarity employed by the exemplar view of categorization \citep{medin1978context,brooks1978nonanalytic,nosofsky1984choice,nosofsky1986attention}.
+In PACKER, this notion is interpreted using the theory of similarity employed by the exemplar view of categorization \citep{medin1978context,brooks1978nonanalytic,nosofsky1984choice,nosofsky1986attention}.
 This theory of similarity, tracing its foundations to classic research in stimulus generalization \citep{attneave1950,shepard1957stimulus,shepard1987toward}, has had broad influence in many fields within cognitive science.
-Thus, beyond providing a formal specification of category contrast, the PACKER model also aligns the relatively understudied topic of creative generation more closely with the well-developed literature on human categorization.
+Thus, beyond providing a formal specification of category contrast, PACKER also aligns the relatively understudied topic of creative generation more closely with the well-developed literature on human categorization.
 The success of the PACKER model indicates that core theories from the categorization literature may offer a strong account of how people generate new concepts.
 
-The success of the PACKER model yields reveals keys insights about the nature of creative generation.
+The success of PACKER yields reveals keys insights about the nature of creative generation.
 First, by measuring PACKER's fit as a function of its prioritization of within- and between-class similarity, we observed that considering either constraint exclusively results in a relatively low-quality account.
@@
 
-\nbcnote{I'm just spitballing there...}
-
 As it stands, however, PACKER's mathematical basis is subject to a wide variety of interpretations, leading to distinct views on the nature of creative generation. 
-\nbcnote{That sounds like a good spot to include importance sampling}
+%\nbcnote{That sounds like a good spot to include importance sampling}
+\jlanote{This is pretty disorganized at the moment. There's a lot of good in here though. Here's the structure I'd like to see. Please move and reorganize it into these subsections}
+Summary of results
+\subsection{Implications for creative cognition}
+more on the general theories/philosophy of creative cognition and what we bring to the table.
+\subsection{Contrast in categorization}
+More broad perspective of how this fits with categorization more broadly.
+Similar style model for explaining language/perceptual (phonemes and color) categories.
+Also, you should check on studies of how contrast affects categorization more broadly (i think there's some work on that).
+\subsection{Limitations and Future Directions}
+this is where the importance sampling connection will come in and will involve specifying the computational problem precisely.
+that hasn't been done yet for category generation and then perhaps to creativity more broadly.
+%discussion of color categories and how they evolve. how across cultures, they seem to maximize within-category similarity and minimize between-category simialrity
 

### 2017-04-19 Manuscripts/cog-psych/cogpsych.tex
edits to the intro to the intro

@@
 
-\jlanote{Joe! Did you know i made a command for you to enter in comments like this?? }
-
-\jlanote{This is awesome!!!}
-
-\begin{verbatim}
-    \jlanote{Some note}
-\end{verbatim}
+\nbcnote{Make sure to write this!}
 
 \setlength\parindent{0.5in}
-\textit{Keywords}: categorization, concepts, creativity, generation
+{\em Keywords}: categorization, concepts, creativity, generation
 \clearpage
@@
 Towards the aim of providing a formal account of the cognitive mechanisms involved in creativity and innovation, we explore a simplified form of creative cognition that can be studied using standard behavioral methodology and is amenable to formal modeling: category generation.
-In a category generation task, participants are given some background knowledge about a domain (a cover story and/or learning one or more categories in the domain) and then are asked to generate another category in the domain \citep{jern2013probabilistic,ward1994structured}.
+In a category generation task, participants are given some background knowledge about a domain (a cover story and/or learning one or more categories in the domain) and then are asked to generate another category in the domain.
 Unlike creativity and innovation, categorization is well studied from a breadth of perspectives, including behavioral, neural, comparative, and computational  \citep{kurtz2015human,mack2013,margolis2015,pothoswills2011}.
@@
 
-Previous work in categorization has identified a number of basic principles underlying human category performance that can be explained by appealing to the structural properties of natural categories, such as  correlations between the features of category members and the relation of the differences of items within the same category to those in different categories\cite{roschmervis1975,regier2006}.
-Inspired by this work in categorization, previous work explored one key aspect of category generation: People tend to create new categories that have similar {\em statistical regularities} to previously known categories \citep{jern2013probabilistic,ward1994structured}.
+Previous work in categorization has established that people are highly sensitive to the structural properties of categories, such as correlations between the features of category members and the relation between items within the same category and those in different categories \citep{roschmervis1975,regier2007,shepard1961learning}.
+Inspired by this work, previous research on the topic of category generation has explored a similar principle: People tend to create new categories that have similar {\em statistical regularities} as previously learned categories \citep{jern2013probabilistic,ward1994structured}.
 Although this is an important characteristic of generating new categories, it cannot be the only one.
-Taken to the extreme, the best ``new'' category in terms of having the same statistical regularities to other categories would be one that is identical to a prototypical previously observed category (and thus, not new at all).
+Taken to the extreme, the best ``new'' category in terms of having the same statistical regularities to other categories would identical to a known category that is representative of the domain (and thus, not new at all).
 
-Inspired by the importance of contrast in specific categorization domains \citep[e.g., color][]{regier2007}, we propose a new constraint that guides category generation: ``being different'' or contrasting from other categories in the relevant domain.
-To some extent, this constraint has been implicitly assumed in much previous work: the idea that a new category should be ``different'' is vague, as there are many ways it could be different from a previously observed category.
+To successfully generate something novel, what is generated must be different from what is already known.
+This fundamental constraint, ``being different'', or contrasting from other categories in the relevant domain, is the focus of our work.
+Beyond the assumption that participant-generated categories are indeed novel, this constraint has been overlooked in previous research: to our knowledge, there has not been any systematic investigation addressing how generated categories {\em differ} from what is already known.
+Although the idea of category contrast is ubiquitous throughout the categorization literature, and extends to a variety of other fields \citep[e.g., color;][]{regier2007}, the idea that a new category should be ``different'' is vague, as there are many ways it could be different from a previously observed category.
 Building on the largely successful exemplar modeling framework \citep{medin1978context,nosofsky1984choice,nosofsky1986attention}, we propose a novel exemplar model of category generation, {\em Producing Alike and Contrasting Knowledge using Exemplar Representations} (PACKER), formalizing how new categories should differ from previous categories.
@@
 Theoretical accounts of these effects have primarily been grounded within the categorization literature.
-For example, the predominant ``Path of Least Resistance'' account \citep[see][]{ward1994structured,ward1995s,ward2002role} proposes that, when generating a new species of animal, people retrieve from memory a known subcategory of animals (e.g., \textit{bird}, \textit{dog}, \textit{horse}), and simply change some of the features to make something new.
-People were thought to change only features that are not characteristic of the retrieved category (e.g., if \textit{bird} was retrieved, the presence of \textit{wings} would not change, but \textit{color} might).
+For example, the predominant ``Path of Least Resistance'' account \citep[see][]{ward1994structured,ward1995s,ward2002role} proposes that, when generating a new species of animal, people retrieve from memory a known subcategory of animals (e.g., {\em bird}, {\em dog}, {\em horse}), and simply change some of the features to make something new.
+People were thought to change only features that are not characteristic of the retrieved category (e.g., if {\em bird} was retrieved, the presence of {\em wings} would not change, but {\em color} might).
 This theory incorporates elements of the highly influential basic-level categories framework \citep{rosch1975cognitive,rosch1976basic}, as well as the exemplar view \citep{medin1978context,brooks1978nonanalytic}.
@@
 Their model views observed examples as samples from an underlying category distribution, describing the location of the category in the space, as well as how it varies along each feature.
-In turn, each category is viewed as a sample from an underlying \textit{domain} distribution, specifying distributional commonalities among the observed categories.
+In turn, each category is viewed as a sample from an underlying {\em domain} distribution, specifying distributional commonalities among the observed categories.
 Generated categories are thought to stem from the same domain distribution as observed categories, thus the distributional properties of observed categories will be preserved within the generated category.
@@
 In this paper, we investigate another important constraint on the creative use of conceptual knowledge: category contrast.
-To generate a novel concept, individuals must produce something that is in some capacity \textit{different} from what they already know.
+To generate a novel concept, individuals must produce something that is in some capacity {\em different} from what they already know.
 Thus, contrast is a fundamental constraint on creative generation: new concepts must be different from existing ones.
 
-Although it is evident that people are \textit{capable} of creating new concepts and categories, it is not entirely clear how new concepts are systematically made different from what is already known.
+Although it is evident that people are {\em capable} of creating new concepts and categories, it is not entirely clear how new concepts are systematically made different from what is already known.
 The hierarchical sampling model developed by \cite{jern2013probabilistic} assumes that differences between generated category are only due to random variation.
 The model assumes that generated categories are sampled from the same underlying domain distribution as observed categories, and will thus share a common distributional structure.
-The authors do not make predictions about the \textit{location} of the category within the domain (the perceptual instantiation of category members).
+The authors do not make predictions about the {\em location} of the category within the domain (the perceptual instantiation of category members).
 Indeed, under a strict interpretation of their model, the most probable new category to be generated is located in {\em exactly} the same location and distributional information as the given category.
@@
 As noted above, \cite{jern2013probabilistic} found that the generated categories tended to follow the distributional properties of the experimenter-defined categories: Generated categories were tightly distributed along the hue feature, and possessed the same saturation-size correlations as in the learned categories.
-\cite{jern2013probabilistic}, however, did not analyze or discuss how the generated categories \textit{differed} from the experimenter-defined categories.
+\cite{jern2013probabilistic}, however, did not analyze or discuss how the generated categories {\em differed} from the experimenter-defined categories.
 
@@
 Although we cannot know precisely which colors were displayed or perceived, we can still analyze their results from a coarse perspective to see whether there is preliminary support for contrast in them.
-To do so, we binned all possible hues into one of eight uniformly-spaced color groups: \{\textit{Red}: $0-0.063$, $0.938-1$, \textit{Yellow}: $0.063-0.188$, \textit{Yellow-Green}: $0.188-0.313$, \textit{Green-Teal}: $0.313-0.438$, \textit{Teal}: $0.438-0.563$, \textit{Teal-Blue}: $0.563-0.688$, \textit{Purple}: $0.688-0.813$, \textit{Pink}: $0.813-0.938$\}.
+To do so, we binned all possible hues into one of eight uniformly-spaced color groups: \{{\em Red}: $0-0.063$, $0.938-1$, {\em Yellow}: $0.063-0.188$, {\em Yellow-Green}: $0.188-0.313$, {\em Green-Teal}: $0.313-0.438$, {\em Teal}: $0.438-0.563$, {\em Teal-Blue}: $0.563-0.688$, {\em Purple}: $0.688-0.813$, {\em Pink}: $0.813-0.938$\}.
 In the \cite{jern2013probabilistic} experiment, the hue of each experimenter-defined category was selected from one of six possible values, each of which falls into one of the color groups above.
@@
 Re-analyzing the results of \cite{jern2013probabilistic} provides preliminary support that contrast plays a role in category generation.
-Taken alongside the analyses reported by \cite{jern2013probabilistic}, our analysis suggests that generated categories tend to be distinct from \textit{and} distributionally similar to what is already known.
+Taken alongside the analyses reported by \cite{jern2013probabilistic}, our analysis suggests that generated categories tend to be distinct from {\em and} distributionally similar to what is already known.
 However, it is worth noting that our analysis is still limited: the color groups defined above are somewhat arbitrary, and it is not clear that our color grouping is consistent with psychological color boundaries.
 While we did obtain similar results using a variety of alternative groupings, the hue dimension used in the \cite{jern2013probabilistic} study does not lend itself straightforwardly to the the computation of similarities, and thus we cannot be certain of whether our coding accurately approximates the psychological space of the stimuli.
-By consequence, although these results likely indicate that contrast exerts \textit{some} influence, they do not precisely describe the nature of that influence.
+By consequence, although these results likely indicate that contrast exerts {\em some} influence, they do not precisely describe the nature of that influence.
 In the sections below, we propose a quantitative framework specifying the role of category contrast in creative generation.
@@
 		\caption{PACKER generation of a category `B' example, following exposure to one member of category `A' and one member of category `B'.
-Predictions are shown for three different parameterizations (differing only in $\gamma$): \textit{(a)} Predictions based on contrast similarity only.
-\textit{(b)} Predictions based on target similarity only.
-\textit{(c)} Predictions with both constraints considered.}
+Predictions are shown for three different parameterizations (differing only in $\gamma$): {\em (a)} Predictions based on contrast similarity only.
+{\em (b)} Predictions based on target similarity only.
+{\em (c)} Predictions with both constraints considered.}
 		\label{fig:packer-examples}
@@
 Research on the creative generation of novel concepts has focused on the finding that generated categories tend to possess distributional commonalities with known categories.
-However, a fundamental goal of concept generation is to create something \textit{new} (i.e., different from what is already known).
+However, a fundamental goal of concept generation is to create something {\em new} (i.e., different from what is already known).
 The manner in which generated categories differ from known ones is, nonetheless, poorly understood: Existing theories do not make strong predictions about how creatively generated concepts should systematically differ from existing ones.
@@
     \caption{Experiment 1 results.
-\textit{Left}: Frequency of generation as a function of distance from members of the experimenter-defined category.
-\textit{Right}: Scatter plot of within-class versus between-class distance in each of the participant-generated categories.}
+{\em Left}: Frequency of generation as a function of distance from members of the experimenter-defined category.
+{\em Right}: Scatter plot of within-class versus between-class distance in each of the participant-generated categories.}
 
@@
 
-Figure \ref{fig:e1-distanceplots} also depicts, for each participant, the average distance of members within the generated category (\textit{within-class} distance) against the average distance between members of the generated and experimenter-defined category (\textit{between-class} distance).
+Figure \ref{fig:e1-distanceplots} also depicts, for each participant, the average distance of members within the generated category ({\em within-class} distance) against the average distance between members of the generated and experimenter-defined category ({\em between-class} distance).
 These data also reveal a systematic pattern: the majority of participants generated categories with more between-class distance than within-class distance.
@@
 As expected, the correlations in the Cluster and Row conditions were not systematically positive or negative ($ps > 0.1$).
-However, the generated categories in the XOR condition tended to possess \textit{negatively} correlated dimensions, $t(57) = 2.04$, $p = 0.046$.
+However, the generated categories in the XOR condition tended to possess {\em negatively} correlated dimensions, $t(57) = 2.04$, $p = 0.046$.
 This finding is notable, as it is the opposite of what would be expected, assuming learners are emulating the distributional structure of the experimenter-defined class (which possesses perfectly positively correlated features).
@@
 
-Notably, however, we also found that participants who learned an XOR category (composed of exemplars following a positive diagonal, see Figure \ref{fig:e1-conditions}) tended to generate items according to a \textit{negative} feature correlation -- the opposite of what was present in the previously learned category.
+Notably, however, we also found that participants who learned an XOR category (composed of exemplars following a positive diagonal, see Figure \ref{fig:e1-conditions}) tended to generate items according to a {\em negative} feature correlation -- the opposite of what was present in the previously learned category.
 While this may be difficult to account for under existing theoretical approaches (which assume generated categories follow the same distributional structure as known categories), it can be concisely explained from a category contrast perspective.
@@
 Thus, while the core results of Experiment 1 indicate that generated categories systematically contrast with what is already known, the negative correlations observed in the XOR condition may signal something further.
-That is, the constraints on creative generation imposed by category contrast may not simply influence the \textit{location} of generated categories, but also their distributional structure.
+That is, the constraints on creative generation imposed by category contrast may not simply influence the {\em location} of generated categories, but also their distributional structure.
 In Experiment 2, we test this claim more systematically.
@@
 Accordingly, participants should be more likely to utilize both of these areas.
-Thus, to the extent that category contrast influences distributional structure, we should observe more participants in the Middle condition that generate examples above \textit{and} below the experimenter-defined category.
+Thus, to the extent that category contrast influences distributional structure, we should observe more participants in the Middle condition that generate examples above {\em and} below the experimenter-defined category.
 
@@
     \caption{Experiment 2 results.
-\textit{Left}: Frequency of generation as a function of distance from members of the experimenter-defined category.
-\textit{Right}: Scatter plot of within-class versus between-class distance in each of the participant-generated categories.}
+{\em Left}: Frequency of generation as a function of distance from members of the experimenter-defined category.
+{\em Right}: Scatter plot of within-class versus between-class distance in each of the participant-generated categories.}
     \label{fig:e2-distanceplots}
@@
 Fisher's Exact Tests comparing the conditions reveal that more Middle participants generated an exemplar in the bottom row, $p < 0.001$, again demonstrating the role of contrast in guiding where exemplars are generated.
-The conditions did not differ in use of the top of the space, $p = 0.16$, however, more Middle participants placed exemplars in the top \textit{and} bottom rows, $p = 0.038$.
+The conditions did not differ in use of the top of the space, $p = 0.16$, however, more Middle participants placed exemplars in the top {\em and} bottom rows, $p = 0.038$.
 The latter effect is of interest here, as it indicates that the shape of the unoccupied space exerts some influence on the distributional structure of generated categories: participants in the Middle condition were more likely to generate a category spanning the entire Y-axis.
@@
 In Experiment 2, we replicated the core findings from Experiment 1: stimuli are more likely to be generated if they are distant from exemplars in other categories, and most participants generate categories with more between-class than within-class distance.
-However, we additionally found that the \textit{position} of a previously learned category (rather than its distributional structure) influences the types of categories people generate: participants who learned the `Middle' type were more likely to generate categories spanning the entire Y-axis of the space.
+However, we additionally found that the {\em position} of a previously learned category (rather than its distributional structure) influences the types of categories people generate: participants who learned the `Middle' type were more likely to generate categories spanning the entire Y-axis of the space.
 Participants who learned the `Bottom' type were less likely to do so as a result of the presence of opposite category exemplars in the lower regions of the space.
@@
     \caption{PACKER generation of a category `B' example, following exposure to one member of category `A' and one member of category `B'.
-Predictions are shown for different attention settings: \textit{(a)} Increased weighting of the X-axis.
-\textit{(b)} Increased weighting of the Y-axis.
-\textit{(c)} Uniform weighting (identical to Figure \ref{fig:packer-examples}).}
+Predictions are shown for different attention settings: {\em (a)} Increased weighting of the X-axis.
+{\em (b)} Increased weighting of the Y-axis.
+{\em (c)} Uniform weighting (identical to Figure \ref{fig:packer-examples}).}
     \label{fig:packer-attention}
@@
 However, a key limitation within both in both of these models is the lack of interaction between the distributional structure of a category (in this case, a row-versus-column orientation) and the location of that category within the domain.
-While these models are \textit{capable} of generating row and column categories, there is no mechanism in place to ensure generated categories will be distinct from what is already known.
+While these models are {\em capable} of generating row and column categories, there is no mechanism in place to ensure generated categories will be distinct from what is already known.
 In the section below, we explore the interdependence between distributional structure and location in creative generation.
@@
 The creative generation of novel concepts is an intriguing yet understudied topic in cognitive science.
-While the bulk of prior research on the topic has focused on the classic finding that generated concepts tend to be distributionally similar to known concepts, there has been little work addressing the role of contrast in creative generation: how is it that people are able to create something \textit{different} from what is already known?
+While the bulk of prior research on the topic has focused on the classic finding that generated concepts tend to be distributionally similar to known concepts, there has been little work addressing the role of contrast in creative generation: how is it that people are able to create something {\em different} from what is already known?
 This issue is of fundamental importance: in order to successfully generate something new, it must be different from what is known.

### 2017-04-19 Manuscripts/cog-psych/cogpsych.tex
edits up to packer intro

@@
 
+\nbcnote{More strongly link this to creativity: what do these findings indicate?}
+
 Much of the work from this area \citep[e.g.,][]{smith1993constraining,marsh1999inadvertent} focuses on how sample cues (such as an example of a species generated by other participants) can drastically diminish creativity.
@@
 For example, the predominant ``Path of Least Resistance'' account \citep[see][]{ward1994structured,ward1995s,ward2002role} proposes that, when generating a new species of animal, people retrieve from memory a known subcategory of animals (e.g., {\em bird}, {\em dog}, {\em horse}), and simply change some of the features to make something new.
-People were thought to change only features that are not characteristic of the retrieved category (e.g., if {\em bird} was retrieved, the presence of {\em wings} would not change, but {\em color} might).
+People are thought to change only features that are not characteristic of the retrieved category (e.g., if {\em bird} was retrieved, the presence of {\em wings} would not change, but {\em color} might).
 This theory incorporates elements of the highly influential basic-level categories framework \citep{rosch1975cognitive,rosch1976basic}, as well as the exemplar view \citep{medin1978context,brooks1978nonanalytic}.
-Problematically, however, the experimental paradigms employed in these studies were relatively uncontrolled, precluding the development of formal approaches.
-Although this previous work has been incredibly useful in providing a conceptual sketch of generation theories, the fine and hand-drawn responses precludes formal analyses.
+Problematically, however, while this work is been incredibly useful in providing a conceptual sketch of generation theories, the hand-drawn responses relied on in the experiments paradigms precludes the development of formal approaches.
 
@@
 
-From the perspective of the categorization literature, prior work on category generation has only explored one of the possible constraints that guide category generation.
+From the perspective of the categorization literature, prior work on category generation has explored only one of the possible constraints that guide category generation.
 The main concern of the published experiments has been on the distributional correspondences between learned and generated categories, and as a result most of the computational, theoretical, and empirical efforts have been towards explaining those effects.
-In this paper, we investigate another important constraint on the creative use of conceptual knowledge: category contrast.
+In this paper, we investigate another important constraint: category contrast.
 To generate a novel concept, individuals must produce something that is in some capacity {\em different} from what they already know.
-Thus, contrast is a fundamental constraint on creative generation: new concepts must be different from existing ones.
+By consequence, contrast should be viewed a primary constraint on creative generation: new concepts must be different from existing ones.
 
@@
 
-Although existing accounts of creative generation broadly overlook the role of category contrast in determining what is novel versus familiar, if contrast affects category generation, we should find these effects within the experimental results of previous work.
-To test whether there is evidence of category contrast affecting category generation, we conducted a novel analysis of Experiment 3 from \cite{jern2013probabilistic}.
+Although existing accounts of creative generation broadly overlook the role of category contrast in determining what is novel versus familiar, it is assumed that learners in previous experiments were {\em successful} in creating new categories.
+Thus, effects of category contrast should be observable within the experimental results of these studies.
+To provide a test of the influence of category contrast in creative generation, we conducted a novel analysis of Experiment 3 from \cite{jern2013probabilistic}.
+
+\nbcnote{Does it feel strange to focus on Experiment 3 specifically? Do you think wee need to explain why this is the `right' experiment for our question?}
 
@@
 
-Because each experimenter-defined category in the \cite{jern2013probabilistic} experiment possessed a distinct hue, it is sensible to propose that participants would generate a category with a hue distinct from the experimenter-provided categories, which all had a distinct hue shared by all members in the category.
-f creative generation is influenced by category contrast, the hues of generated categories should be systematically different from those of the experimenter-defined categories.
-Problematically, however, stimulus hue was presented using the Hue-Saturation-Value (HSV) color space, which is device-dependent and not perceptually normed such that perceived color similarity corresponds to proximity in the color space (as opposed to a color space such as CIELAB that is device-independent and equidistant sets of points correspond to pairs of colors that have the same perceptual similarity; \citet{wyszecki1967}).
-Further, they did not calibrate their monitor.
-Thus, we cannot know the precise values of colors presented to participants.
-As \cite{jern2013probabilistic} were interested in relations between the saturation and length of generated stimuli, this is not an issue for their analyses and results.
-But it means we cannot know precisely which colors were displayed to participants.
+Because each experimenter-defined category in the \cite{jern2013probabilistic} experiment possessed a distinct hue shared by all members of the category, it is sensible to propose that participants would generate a category with a hue distinct from the experimenter-provided categories.
+If creative generation were influenced by category contrast in this way, the hues of generated categories should be systematically different from those of the experimenter-defined categories.
+Problematically, however, stimulus hue was presented using the Hue-Saturation-Value (HSV) color space, which is device-dependent and not perceptually normed such that perceived color similarity corresponds to proximity in the color space (as opposed to a color space such as CIELAB that is device-independent and equidistant sets of points correspond to pairs of colors that have the same perceptual similarity; \citealp{wyszecki1967}).
+Further, they did not calibrate their monitor, and so we cannot know the precise values of colors presented to participants.
+As \cite{jern2013probabilistic} were interested in relations between the saturation and length of examples in generated into novel categories, this is not an issue for their analyses and results.
+However, these issues pose a significant challenge to evaluating contrast between the experimenter-defined and participant-generated categories along the hue dimension.
 
-Although we cannot know precisely which colors were displayed or perceived, we can still analyze their results from a coarse perspective to see whether there is preliminary support for contrast in them.
+Although we cannot know precisely which colors were displayed or perceived, we can still analyze their results from a coarse perspective to see whether there is preliminary support for contrast.
 To do so, we binned all possible hues into one of eight uniformly-spaced color groups: \{{\em Red}: $0-0.063$, $0.938-1$, {\em Yellow}: $0.063-0.188$, {\em Yellow-Green}: $0.188-0.313$, {\em Green-Teal}: $0.313-0.438$, {\em Teal}: $0.438-0.563$, {\em Teal-Blue}: $0.563-0.688$, {\em Purple}: $0.688-0.813$, {\em Pink}: $0.813-0.938$\}.
@@
 In the sections below, we propose a quantitative framework specifying the role of category contrast in creative generation.
-
-\nbcnote{could use a cite on why hue dimension proximity does not match up with similarity. maybe also a note about how we can't be sure what hues were actually perceived due to norming, etc.}
 

### 2017-04-20 Manuscripts/cog-psych/cogpsych.tex
edits through results

@@
 PACKER explains category generation as a balance between two fundamental constraints: generated categories should not be similar to known categories, and exemplars within each category should be similar to one another.
-These ideas are implemented within the well-studied exemplar framework: the PACKER model is an extension of the influential Generalized Context Model of category learning \citep[GCM;][]{nosofsky1984choice,nosofsky1986attention}.
+These ideas are implemented within the well-studied exemplar framework -- the PACKER model is an extension of the influential Generalized Context Model of category learning \citep[GCM;][]{nosofsky1984choice,nosofsky1986attention}.
 
@@
 PACKER sets $f(x_j)$ depending on exemplar $x_j$'s category membership: $f(x_j) = \gamma$ if $x_j$ is a member of the target category, and $f(x_j) = \gamma - 1$ if $x_j$ is a member of a contrast category.
-$\gamma$ is thus a free parameter ($0 \leq \gamma \leq 1$) controlling the trade-off between within- and between-class similarity.
+$\gamma$ is thus a free parameter ($0 \leq \gamma \leq 1$) controlling the trade-off between within- and between-category similarity.
 For example, when $\gamma = 0.5$, $f(x_j) = 0.5$ for members of the target category and $f(x_j) = -0.5$ for members of other categories; thus, the model is likely to generate items that are similar to members of the target category but are not similar to members of other categories.
 In this way, $\gamma = 1$ produces exclusive consideration of target-category members, and $\gamma = 0$ produces exclusive consideration of opposite-category members.
-The $\gamma$ parameter thus specifies a wide breadth of possible approaches; by fitting it to a dataset, one can describe the relative roles of between-class contrast and within-class similarity in generation.
+The $\gamma$ parameter thus specifies a wide breadth of possible approaches; by fitting it to a dataset, one can describe the relative roles of between-category contrast and within-category similarity in generation.
 See Figure \ref{fig:packer-examples} for an illustration of how $\gamma$ controls the relative influence of within category similarity and contrast to other categories.
@@
 It is also worth noting that PACKER represents just one of many possible models that incorporate contrast in category generation.
-For example, although it may be possible to extend \cite{jern2013probabilistic}'s hierarchical sampling model to include contrast, this mechanism would be at odds with a conjugate hierarchical Bayesian framework.
+For example, although it may be possible to extend \cite{jern2013probabilistic}'s hierarchical sampling model to include contrast, this mechanism would be at odds with a hierarchical Bayesian framework.
 In contrast, these ideas emerge naturally from the exemplar view: we have not modified any of the core elements of the GCM in defining the PACKER model, we simply aggregated similarity slightly differently.
@@
 
-\jlanote{I forget, but did JK13 explicitly have in their instructions that the new category be different? If so, we should mention the issue there as another problem with their study for us.}
 \section{Experiment 1}
@@
 Unlike those used in the \cite{jern2013probabilistic} experiments, proximity along these dimensions aligns more directly with perceptual similarity, allowing us to evaluate the role of category contrast in creative generation more precisely.
-To extend the evidence provided by the \cite{jern2013probabilistic} data, we tested the effects of category contrast after learning one category from a set of category structures.
-We did so by training participants on categories possessing qualitatively different distributional structures, as shown in Figure \ref{fig:e1-conditions}.
+To extend the evidence provided by the \cite{jern2013probabilistic} data, we tested the effects of category contrast after learning one category from a set of qualitatively distinct category structures, as shown in Figure \ref{fig:e1-conditions}.
 
@@
 From these results, we can predict that, in each condition, participants should generate categories that are distributionally similar to the experimenter-defined category: In the Cluster condition, generated categories should be tightly clustered.
-In the Row condition, generated categories should very more along the X-axis than the Y-axis.
+In the Row condition, generated categories should vary more along the X-axis than the Y-axis.
 In XOR condition, generated categories should be widely distributed across both dimensions, and the two dimensions should be positively correlated.
@@
 Interestingly, the XOR condition also offers a dissociation between the roles of category contrast and the emulation of distributional structure: widely-distributed, positively-correlated categories would need to lie along the positive diagonal of the space (that is the only place they ``fit''), which is already occupied by the experimenter-defined category.
-Thus, if contrast plays a role, exemplars in the generated categories of participants in the XOR condition may not be positively correlated--they may be negatively correlated instead.
+Thus, if contrast plays a role, exemplars in the generated categories of participants in the XOR condition may not be positively correlated -- they may be negatively correlated instead.
 
@@
 {\em Left}: Frequency of generation as a function of distance from members of the experimenter-defined category.
-{\em Right}: Scatter plot of within-class versus between-class distance in each of the participant-generated categories.}
-
-    \nbcnote{Need to check on how nice the journal is about color figs!}
-	\jlanote{Is there something systematic about XOR for how the generated category varies for smaller/larger within-category distances. Also, should switch "class" to category here.}
+{\em Right}: Scatter plot of within-category versus between-category distance in each of the participant-generated categories.}
     \label{fig:e1-distanceplots}
@@
 
-Figure \ref{fig:e1-distanceplots} also depicts, for each participant, the average distance of members within the generated category ({\em within-class} distance) against the average distance between members of the generated and experimenter-defined category ({\em between-class} distance).
-These data also reveal a systematic pattern: the majority of participants generated categories with more between-class distance than within-class distance.
+Figure \ref{fig:e1-distanceplots} also depicts, for each participant, the average distance of members within the generated category ({\em within-category} distance) against the average distance between members of the generated and experimenter-defined category ({\em between-category} distance).
+These data also reveal a systematic pattern: the majority of participants generated categories with more between-category distance than within-category distance.
 That is, members of the generated category tended to be more similar to one another than to members of the experimenter-defined category.
 To formally evaluate this claim, we conducted t-tests comparing the amount of within- and between- class distance in each condition.
-All conditions possessed greater between-class distance: Cluster, $t(63) = 11.43$, $p < 0.001$; Row, $t(60) = 13.16$, $p < 0.001$; and XOR, $t(57) = 3.64$, $p < 0.001$.
+All conditions possessed greater between-category distance: Cluster, $t(63) = 11.43$, $p < 0.001$; Row, $t(60) = 13.16$, $p < 0.001$; and XOR, $t(57) = 3.64$, $p < 0.001$.
+The narrow distribution of between-category distances in the XOR condition reflects the widely distributed nature of the experimenter-defined category: all possible generation candidates are distant from at least one cluster of the experimenter-defined category.
 These results provide clear evidence of an effect of category contrast: participants prefer to generate categories that are dissimilar to the learned category but maintain some level of internal cohesion.
@@
 {\em Left}: Frequency of generation as a function of distance from members of the experimenter-defined category.
-{\em Right}: Scatter plot of within-class versus between-class distance in each of the participant-generated categories.}
+{\em Right}: Scatter plot of within-category versus between-category distance in each of the participant-generated categories.}
     \label{fig:e2-distanceplots}
@@
 These data, shown in Figure \ref{fig:e2-distanceplots}, yield highly similar results.
-Stimuli that are more distant from members of the experimenter-defined category were more frequently generated, and the categories in each condition tended to possess more between-class than within-class distance: Bottom, $t(60) = 5.5$, $p < 0.001$; Middle, $t(60) = 2.71$, $p = 0.009$.
+Stimuli that are more distant from members of the experimenter-defined category were more frequently generated, and the categories in each condition tended to possess more between-category than within-category distance: Bottom, $t(60) = 5.5$, $p < 0.001$; Middle, $t(60) = 2.71$, $p = 0.009$.
 
-We did, however, observe a notable subgroup of participants in each condition who generated categories with more within-class than between-class distance.
-Upon manual inspection, many of these individuals appear to have assumed a `Corners' strategy, placing exemplars in disparate corners of the space, thus producing much more within-class distance, see Figure \ref{fig:e2-samples} for examples.
+We did, however, observe a notable subgroup of participants in each condition who generated categories with more within-category than between-category distance.
+Upon manual inspection, many of these individuals appear to have assumed a `Corners' strategy, placing exemplars in disparate corners of the space, thus producing much more within-category distance, see Figure \ref{fig:e2-samples} for examples.
 
@@
 
-In Experiment 2, we replicated the core findings from Experiment 1: stimuli are more likely to be generated if they are distant from exemplars in other categories, and most participants generate categories with more between-class than within-class distance.
+In Experiment 2, we replicated the core findings from Experiment 1: stimuli are more likely to be generated if they are distant from exemplars in other categories, and most participants generate categories with more between-category than within-category distance.
 However, we additionally found that the {\em position} of a previously learned category (rather than its distributional structure) influences the types of categories people generate: participants who learned the `Middle' type were more likely to generate categories spanning the entire Y-axis of the space.
@@
 We also present simulations from an implementation of the hierarchical sampling model proposed by \cite{jern2013probabilistic}, described in-depth in Appendix \ref{ap:hsampling-definition}.
-The comparison between the hierarchical sampling model and PACKER is meant to emphasize the necessity of contrast and demonstrate it cannot be explained by emulating of distributional structure: whereas PACKER is insensitive to the distributional structure of learned categories (relying only on within- and between-class similarity), the hierarchical sampling model generates categories exclusively on the basis of knowledge of how existing classes are distributed.
+The comparison between the hierarchical sampling model and PACKER is meant to emphasize the necessity of contrast and demonstrate it cannot be explained by emulating of distributional structure: whereas PACKER is insensitive to the distributional structure of learned categories (relying only on within- and between-category similarity), the hierarchical sampling model generates categories exclusively on the basis of knowledge of how existing classes are distributed.
 Our approach in this section is to first broadly evaluate and compare the quality of each model's account to our entire dataset (Experiments 1 and 2 combined), then we more specifically describe the strengths and weakness of each model's account.
@@
 Notably, however, the copy-and-tweak parameterization ($\gamma=1$) performs better than the `contrast-only' parameterization of the model ($\gamma = 0$).
-Thus, it is evident that the data is better explained when both within-class similarity and category contrast is considered.
+Thus, it is evident that the data is better explained when both within-category similarity and category contrast is considered.
 
@@
     \inputpgf{figs/}{packer-loglike.pgf}
-    \caption{PACKER's fit as a function of its prioritization of within-class and between-class similarity (using the $\gamma$ parameter).
+    \caption{PACKER's fit as a function of its prioritization of within-category and between-category similarity (using the $\gamma$ parameter).
 To facilitate comparison, PACKER's other parameters ($c$, $\theta$) were set to the best fitting values obtained for copy-and-tweak in Table \ref{table:global-model-fits}.
@@
 
-In the framework proposed by PACKER, the cluster and corners profiles arise based on different prioritization of within-class similarity versus between-class contrast, and the row and column profiles arise based on the prioritization of each dimension in the computation of similarity.
-For example, in the cluster profile, there is a high degree of within-class similarity along both dimensions, whereas in the corners profile there is minimal within-class similarity.
-Thus, PACKER's proposal is that these individual differences arise as a result of different priorities: While the tight cluster configuration can be considered PACKER's `default' mode (as it maximizes within-class similarity), the corners profile can be produced when between-class contrast is put at a higher priority (i.e., $\gamma = 0$).
+In the framework proposed by PACKER, the cluster and corners profiles arise based on different prioritization of within-category similarity versus between-category contrast, and the row and column profiles arise based on the prioritization of each dimension in the computation of similarity.
+For example, in the cluster profile, there is a high degree of within-category similarity along both dimensions, whereas in the corners profile there is minimal within-category similarity.
+Thus, PACKER's proposal is that these individual differences arise as a result of different priorities: While the tight cluster configuration can be considered PACKER's `default' mode (as it maximizes within-category similarity), the corners profile can be produced when between-category contrast is put at a higher priority (i.e., $\gamma = 0$).
 
-Likewise, in the row and column profiles, there is a high degree of within-class similarity along one dimension but not the other.
+Likewise, in the row and column profiles, there is a high degree of within-category similarity along one dimension but not the other.
 These differences likely arise due to a differential focus on one dimension over another, and thus they can be produced by changes to PACKER's attention weights, $w$ (see Equation \ref{eq:similarity}).
-Traditionally, the attention weights in exemplar models are thought to reflect the diagnostic value of each dimension towards classifying the known category members \citep{nosofsky1984choice,nosofsky1986attention,kruschke1992alcove}, but within a generation context the weights specify the importance of within- and between-class similarity along each dimension.
+Traditionally, the attention weights in exemplar models are thought to reflect the diagnostic value of each dimension towards classifying the known category members \citep{nosofsky1984choice,nosofsky1986attention,kruschke1992alcove}, but within a generation context the weights specify the importance of within- and between-category similarity along each dimension.
 For example, if 100\% of attention is allocated along the X-axis, similarity along the Y-axis no longer influences performance: as a result, PACKER will create categories that are more widely distributed along the Y-axis, as similarity is not counted along that dimension.
@@
 These patterns of performance clearly depict the interdependence between the distributional structure and location of generated concepts.
-Our results can, in some sense, be interpreted in terms of local minimization of between-class similarity: by distributing the generated category away from members of the experimenter-defined category, participants may increase the degree of between-class distance without drastically altering the degree of within-class similarity.
+Our results can, in some sense, be interpreted in terms of local minimization of between-category similarity: by distributing the generated category away from members of the experimenter-defined category, participants may increase the degree of between-category distance without drastically altering the degree of within-category similarity.
 
@@
 We developed a novel, exemplar-based model, PACKER, which formally specifies the role of contrast in generation.
-The model proposes that categories are represented as exemplars in a multidimensional psychological space, and generation is constrained both by within-class and between-class similarity: exemplars belonging to the same category should be similar to one another, and exemplars belonging to different categories should not be similar to one another.
+The model proposes that categories are represented as exemplars in a multidimensional psychological space, and generation is constrained both by within-category and between-category similarity: exemplars belonging to the same category should be similar to one another, and exemplars belonging to different categories should not be similar to one another.
 
@@
 The success of PACKER yields reveals keys insights about the nature of creative generation.
-First, by measuring PACKER's fit as a function of its prioritization of within- and between-class similarity, we observed that considering either constraint exclusively results in a relatively low-quality account.
-Instead, PACKER's best results were obtained when both constraints are considered, indicating that human learners do not generate novel concepts exclusively on the basis of within-class similarity or between class-contrast.
+First, by measuring PACKER's fit as a function of its prioritization of within- and between-category similarity, we observed that considering either constraint exclusively results in a relatively low-quality account.
+Instead, PACKER's best results were obtained when both constraints are considered, indicating that human learners do not generate novel concepts exclusively on the basis of within-category similarity or between class-contrast.
 This finding mirrors our behavioral results: it is clear that both constraints influence creative generation.
@@
 In Experiment 2, we found that the unoccupied regions of the domain influenced the distributional structure of categories, and in Experiments 1 and 2 we observed that generated categories assume different distributional structures depending on how they contrast with known categories.
-This interdependence is straightforwardly explained under PACKER's dual-constraint proposal: after first establishing a location for their generated category, participants distribute exemplars in order to minimize between-class similarity while not drastically decreasing within-class similarity.
+This interdependence is straightforwardly explained under PACKER's dual-constraint proposal: after first establishing a location for their generated category, participants distribute exemplars in order to minimize between-category similarity while not drastically decreasing within-category similarity.
 

### 2017-04-20 Manuscripts/cog-psych/cogpsych.tex
edits through e2

@@
 Figure \ref{fig:e1-distanceplots} also depicts, for each participant, the average distance of members within the generated category ({\em within-category} distance) against the average distance between members of the generated and experimenter-defined category ({\em between-category} distance).
-These data also reveal a systematic pattern: the majority of participants generated categories with more between-category distance than within-category distance.
+The narrow distribution of between-category distances in the XOR condition reflects the widely distributed nature of the experimenter-defined category: all possible generation candidates are distant from at least one cluster of the experimenter-defined category.
+These data reveal a systematic pattern: the majority of participants generated categories with more between-category distance than within-category distance.
 That is, members of the generated category tended to be more similar to one another than to members of the experimenter-defined category.
@@
 All conditions possessed greater between-category distance: Cluster, $t(63) = 11.43$, $p < 0.001$; Row, $t(60) = 13.16$, $p < 0.001$; and XOR, $t(57) = 3.64$, $p < 0.001$.
-The narrow distribution of between-category distances in the XOR condition reflects the widely distributed nature of the experimenter-defined category: all possible generation candidates are distant from at least one cluster of the experimenter-defined category.
 These results provide clear evidence of an effect of category contrast: participants prefer to generate categories that are dissimilar to the learned category but maintain some level of internal cohesion.
@@
 This latter effect was not expected; however, the key finding is that categories from the Cluster condition tended to be more tightly distributed along the X-axis.
-\jlanote{Was XOR slightly longer than row?}
+
 Likewise, categories generated in the Row condition had less Y-axis range compared to Cluster, $t(123) = 4.57$, $p < 0.001$ and XOR, $t(117) = 9.26$, $p < 0.001$, and categories from the Cluster condition had less Y-axis range compared to XOR, $t(120) = 3.95$, $p < 0.001$.
@@
 This finding is notable, as it is the opposite of what would be expected, assuming learners are emulating the distributional structure of the experimenter-defined class (which possesses perfectly positively correlated features).
-Thus, there is more to category generation than the distributional structure of other categories in the domain!
-Further, as we will discuss in more detail in the model-based analysis section, it is expected by our proposal that contrast is a fundamental principle for category generation.
+This indicates that there is more to category generation than the distributional structure of other categories in the domain.
+Further, as we will discuss in more detail in the model-based analysis section, it is expected by our proposal that contrast is a fundamental principle in category generation.
 
@@
 \section{Experiment 2}
-To test whether category contrast influences the distributional structure of generated categories, we created two new category types (depicted in Figure ~\ref{fig:e2-conditions}) that do not differ in distributional structure, but we expect participants would generate different categories given each of them.
-The observed categories possess an identical distributional structure (both are tight clusters of examples with a neutral feature correlation), and only differ slightly in their Y-axis position: the `Bottom' category lies in the bottom-center of the space, and the `Middle' category lies in the center.
+
+In order to test whether category contrast influences the distributional structure of generated categories, we sought to identify conditions in which differences in the distributional structure of generated categories cannot be explained by the distributional structure of the experimenter-defined category.
+We created two new category types (depicted in Figure ~\ref{fig:e2-conditions}) that possess an identical distributional structure (both are tight clusters of examples with a neutral feature correlation), and only differ slightly in their Y-axis position: the `Bottom' category lies in the bottom-center of the space, and the `Middle' category lies in the center.
 The distributional equality of these conditions is key to the design of the experiment: if the structure of previously learned categories were the only influence on the structure of generated categories, we should observe no difference between these two conditions with respect to distributional structure.
@@
 
-Alternatively, participants seeking to make a perceptually distinct category would be more likely to distribute members of the generated category into areas that are distant from members of known categories.
+Critically, however, participants seeking to make a perceptually distinct category would be more likely to distribute members of the generated category into areas that are distant from members of known categories.
 Thus, if category contrast influences the distributional structure of the categories people generate, then we should observe different types of categories according to the shape of the space that is unoccupied by members of previously learned categories.
@@
 
-\nbcnote{ Weak spot: the obvious next analysis here would involve the same box-and-whisker plot as in Experiment 1.
-We ran this analysis but the conditions did not differ in Y-range due to the data being weirdly distributed.
-I've included the box-and-whisker plot, as well as the graph-of-many-lines figure, on a separate page at the end.}
+To address out main question, we followed this analysis by comparing the conditions on measures of distributional structure.
+For each generated category, we computed the range of exemplars along each axis (X, Y), as well as the correlation between features.
+These data, shown in Figure \ref{fig:e2-statsboxes}, again demonstrate the degree of individual differences observed in our study: in each condition, we observed tightly clustered and widely distributed categories along each dimension, as well as positively, negatively, and uncorrelated categories.
 
-\jlanote{I think we should show and explain that corner behavior mess it up (or whatever the reason is why it happens).
-I think it is more suspicious to not show it and from what I remember, it doesn't really affect our hypotheses.}
+\begin{figure}
+    \begin{center}
+    \inputpgf{figs/}{e2-statsboxes.pgf}
+    \caption{Box-plots of the distributional statistics from the categories generated in Experiment 2. }
+    \label{fig:e2-statsboxes}
+    \end{center}
+\end{figure}
 
-Because the conditions differ only in the Y-axis position of the experimenter-defined category, we then compared the conditions in terms of the frequency with which participants generated examples above and below the categories.
+As noted above, if the distributional structure of generated categories is influenced by the shape of the space not occupied by members of known categories, then participants in the Middle condition would be more likely to place exemplars in the upper {\em and} lower regions of the space, as members of the experimenter-defined category are equidistant from these regions.
+Participants in the Bottom condition should in turn be less likely to generate category members in the bottom regions because members of the experimenter-defined category are located there.
+These predictions can be addressed within the Y-axis ranges of the generated categories: If Middle participants utilize the upper and lower regions of the space, their categories should vary more along the Y-axis.
+T-Tests comparing the conditions on the distributional statistics, however, reveal few between-group differences: the conditions do not differ with respect to X-axis range, Y-axis range, or feature correlations ($ps>0.17$).
+
+However, our ability to detect differences between the conditions is, in this case, diminished due to the non-normality of the data.
+Figure \ref{fig:e2-yranges} depicts the Y-axis position of the exemplars generated within each participant's category.
+The categories are sorted by overall range, then by condition assignment.
+These data reveal that there were nearly as many participants who generated categories spanning the entire Y-axis as those who generated categories spanning almost none of the Y-axis.
+The non-normality of the Y-axis range distributions thus requires that we use a different approach to addressing the experiment's main question.
+
+\begin{figure}
+    \begin{center}
+    \inputpgf{figs/}{e2-yranges.pgf}
+    \caption{Y-Axis range and position of the categories generated in Experiment 2.
+Each line corresponds to a single category, with notches corresponding to the Y-axis position of exemplars within the category (notches may overlap).
+Data from each participant is sorted by overall range, and then by condition.
+}
+    \label{fig:e2-yranges}
+    \end{center}
+\end{figure}
+
+Because our main prediction concerns the generation of exemplars within the upper and lower regions of the domain, we compared the conditions in terms of the frequency with which participants generated examples above and below the categories.
 Specifically, we counted the number of participants in each condition who placed at least one `Beta' exemplar on the top and bottom `rows' of the space (the maximum and minimum possible y-axis value, respectively).

### 2017-04-20 Manuscripts/cog-psych/cogpsych.tex
distributive-property'd the heck out of the emprical bayes formula

@@
 To obtain a global measure of the quality of each model's account, we fitted the parameters of each model to our entire dataset (Experiments 1 and 2 combined), using a hill-climbing algorithm which maximized the log-likelihood of the model's predictions of the observed responses (1220 responses from 305 total participants).
-We fitted three parameters in the PACKER model ($c$, $\gamma$, and $\theta$; see Section \ref{section:PACKER-definition}), as well as four in the hierarchical sampling model ($\kappa$, $\rho$, $\nu$, and $\theta$; see Appendix \ref{ap:hsampling-definition}).
+We fitted three parameters in the PACKER model ($c$, $\gamma$, and $\theta$; see Section \ref{section:PACKER-definition}), as well as four in the hierarchical sampling model ($\kappa$, $\lambda$, $\nu$, and $\theta$; see Appendix \ref{ap:hsampling-definition}).
 We fitted only two parameters for the copy-and-tweak model ($c$, and $\theta$), as $\gamma$ is held constant ($\gamma = 1$).
@@
  $\gamma = 0.525$   & $\gamma = 1$ (fixed)  & $\nu = 5.596$     \\ 
- $\theta = 6.664$   & $\theta = 2.969$      & $\rho = 0.055$    \\ 
+ $\theta = 6.664$   & $\theta = 2.969$      & $\lambda = 0.055$    \\ 
                     &                       & $\theta = 3.174$  \\ 
@@
 \end{table}
-
-\jlanote{should do a nested model comparison with respect to PACKER v. copy \& tweak. they're nested, so should be doable...}
-
-\nbcnote{I do not know exactly what you mean}
-
-\jlanote{I did them.
-(it's just 2 * diff in log likelihood is $\chi^2$ distributed with df equal to difference in the number of parameters) You should check BIC and $AIC_c$ (easy to find formula for them on Wikipedia).
-I don't think we should list them all, but it'd be good to mention our results are robust}
 
@@
 We conducted an identical set of simulations as above, excluding this trial (leaving 915 responses in the dataset): Again, PACKER ($L = -3390$, $AIC = 6786$) achieved better fits than the copy-and-tweak ($L = -3579$, $AIC = 7162$) and hierarchical sampling ($L=-3612$, $AIC = 7232$) models.
-Because copy-and-tweak is nested within PACKER, we can use a likelihood ratio test to compare the two models.
-PACKER explains the aggregate data significantly better than copy-and-tweak ($\chi^2(1) = 748, p < 0.001$ for all data and $\chi^2(1) = 378, p < 0.001$ excluding the first example).
-Thus, participant results support that category generation includes a contrast mechanism.
+Finally, because copy-and-tweak is nested within PACKER, we can use a likelihood ratio test to compare the two models.
+PACKER explains the aggregate data significantly better than copy-and-tweak ($\chi^2(1) = 748, p < 0.001$ for all data and $\chi^2(1) = 378, p < 0.001$ excluding the first example), providing more targeted evidence that category generation is better explained when contrast is considered.
+
+\nbcnote{report something about robustness to model comparison metric (BIC and $AIC_c$).}
 
@@
 In the section below, we explore the interdependence between distributional structure and location in creative generation.
-\jlanote{It's a combination of previous category location and where the target category is being generated, right?}
+
 \subsection{Category Location vs. Distributional Structure}
@@
 
+% ORIGINAL
+% \begin{equation}
+% \mu_x = \dfrac
+% { \mu_0/\sigma + \sigma^{-1}\sum{x}} 
+% { 1/\sigma + n_x/\sigma }
+% \label{eq:rangediff-bayes}
+% \end{equation}
+
+% CLEANER
+% \begin{equation}
+% \mu_x = \dfrac
+% { (\mu_0 + \sum{x}) / \sigma } 
+% { (1 + n_x) / \sigma }
+% \label{eq:rangediff-bayes}
+% \end{equation}
+
+% WHATTTT
 \begin{equation}
 \mu_x = \dfrac
-{ \mu_0/\sigma + \sigma^{-1}\sum{x}} 
-{ 1/\sigma + n_x/\sigma }
+{ \mu_0 + \sum{x} } 
+{ 1 + n_x }
 \label{eq:rangediff-bayes}
 \end{equation}
-\nbcnote{I'm pretty sure some distributive-property trickery can be used to clean that up. In either case, Joe should take a look at that language.}
 % 
@@
 Infrequently generated stimuli, as well as those with both strongly positive and negative scores, are given neutral difference scores.
-The results of our analysis are shown in Figure \ref{fig:range-diff-gradients}.
-\nbcnote{Do you think we need to report the Gaussian smoothing?}
-\jlanote{should mention it in the caption or a footnote}
+The results of our analysis\footnote{Prior to plotting, data was processed using a Gaussian filter with $\sigma = 0.8$} are shown in Figure \ref{fig:range-diff-gradients}.
+
 \begin{figure}[p]
@@
 \nbcnote{This works for 2D, but not 3D and beyond (as $1-w_k$ would not produce a valid weighting).}
-\jlanote{I renamed $\rho$ because $\rho$ is reserved for the correlation parameter of a covariance matrix (especially a $2 \times 2$).}
+\jlanote{I renamed $\lambda$ because $\lambda$ is reserved for the correlation parameter of a covariance matrix (especially a $2 \times 2$).}
 
@@
 
-\begin{figure}
-    \begin{center}
-    \inputpgf{figs/}{e2-statsboxes.pgf}
-    \caption{Box-plots of the distributional statistics from the categories generated in Experiment 2. }
-    \nbcnote{Note the huge amount of variance in the range data, as well as the big median difference in Y-range.}
-
-    \label{fig:e2-statsboxes}
-    \end{center}
-\end{figure}
-
-\begin{figure}
-    \begin{center}
-    \inputpgf{figs/}{e2-yranges.pgf}
-    \caption{Y-Axis range and position of the categories generated in Experiment 2.
-Each line corresponds to a single category, with notches corresponding to the Y-axis position of exemplars within the category.
-Data from each participant is sorted by overall range, and then by condition.
-}
-    \label{fig:e2-yranges}
-    \end{center}
-\end{figure}
-
-\clearpage
-
 \appendix
@@
 We assume the shared covariance matrix $\Sigma_D$ is generated from a Wishart distribution (for conjugacy) with parameters $v_0$, $\kappa_0$, and $\Sigma_0$.
-We set $\nu_0 = 4$, and $\Sigma_0 = \rho {\bf I}$, where $\rho$ is a free parameter controlling the expected variance of dimensions (dimensions of the shared covariance matrix are expected to be uncorrelated) and ${\bf I}$ is the identity matrix.
+We set $\nu_0 = 4$, and $\Sigma_0 = \lambda {\bf I}$, where $\lambda$ is a free parameter controlling the expected variance of dimensions (dimensions of the shared covariance matrix are expected to be uncorrelated) and ${\bf I}$ is the identity matrix.
 

### 2017-04-21 Manuscripts/cog-psych/cogpsych.tex
empirical bayes was normal inverse gamma THE WHOLE TIME

@@
 However, because many stimuli were infrequently generated (such items near members of the `Alpha' category), we cannot simply compute the empirical average, as infrequently generated stimuli would be likely to show artificially strong differences.
-Instead, we aggregated the scores in an empirical Bayesian manner \citep{robbins1964empirical}: the aggregated score for each stimulus was viewed as a normal distribution with a $\mu_0 = 0$ and $\sigma$ set based on the observed scores $x$ for that stimulus.
-The aggregated difference $\mu_x$ was then computed as a Bayesian update:
+Instead, we used a Bayesian analysis to estimate the mean $\mu_x$ on the assumption that the scores $x$ for each stimulus are normally distributed with an unknown mean and unknown standard deviation.
+The conjugate normal-inverse-gamma distribution provides a straightforward method for this estimation:
 
-% ORIGINAL
-% \begin{equation}
-% \mu_x = \dfrac
-% { \mu_0/\sigma + \sigma^{-1}\sum{x}} 
-% { 1/\sigma + n_x/\sigma }
-% \label{eq:rangediff-bayes}
-% \end{equation}
-
-% CLEANER
-% \begin{equation}
-% \mu_x = \dfrac
-% { (\mu_0 + \sum{x}) / \sigma } 
-% { (1 + n_x) / \sigma }
-% \label{eq:rangediff-bayes}
-% \end{equation}
-
-% WHATTTT
 \begin{equation}
 \mu_x = \dfrac
-{ \mu_0 + \sum{x} } 
-{ 1 + n_x }
+    { \nu_0 \mu_0 + \sum{x} } 
+    { \nu_0 + n }
 \label{eq:rangediff-bayes}
 \end{equation}
-% 
+\
+where $\mu_0$ is the prior mean, $\nu_0$ is a prior scale parameter (controlling the weighting of the $\mu_0$), and $n$ is the number of categories in which the stimulus was a member (i.e., the number of scores in $x$).
+The default assumption is that there is an equal amount of range along the X- and Y-axes, and so we set $\mu_0 = 0$.
+Likewise, to give a moderate amount of weighting to the prior mean we set $\nu_0 = 1$, though our results are robust to a range of values.
 Within this approach, the resulting aggregation is a trade-off between the number of generations and the strength of the range difference within each generated category.
@@
 The results of our analysis\footnote{Prior to plotting, data was processed using a Gaussian filter with $\sigma = 0.8$} are shown in Figure \ref{fig:range-diff-gradients}.
+
+\nbcnote{Those two paragraphs are, i think, either: (A) too specific and could be in an appendix, or (B) totally straightforward to anyone who knows about Bayesian stats, and so we just need a normal inverse gamma cite.}
 

### 2017-04-21 Manuscripts/cog-psych/cogpsych.tex
added experiment markers to range difference figure

@@
 Within this approach, the resulting aggregation is a trade-off between the number of generations and the strength of the range difference within each generated category.
-Infrequently generated stimuli, as well as those with both strongly positive and negative scores, are given neutral difference scores.
-The results of our analysis\footnote{Prior to plotting, data was processed using a Gaussian filter with $\sigma = 0.8$} are shown in Figure \ref{fig:range-diff-gradients}.
+Infrequently generated stimuli, as well as those with both strongly positive and negative scores, are given neutral difference scores\footnote{Prior to plotting, data was also processed using a Gaussian filter with $\sigma = 0.8$}.
 
@@
     \label{fig:range-diff-gradients}
-  	\jlanote{This is beautiful. I'd like to add one thing to it. On the left margin have something that says "Experiment 1" and groups the top 3 rows, and then "Experiment 2" for the bottom two.}
     \end{center}
@@
 
+The results of our analysis are shown in Figure \ref{fig:range-diff-gradients}.
 These data reveal strong and consistent patterns across all the conditions we tested in Experiments 1 and 2: generated categories are more tightly distributed along the axis in which they are distinct.
@@
 To attempt to explain our findings within the PACKER, copy-and-tweak- and hierarchical sampling models, we conducted simulations using an individual-differences approach.
-As noted in Section \ref{section:individual-diffs}, row- and column-like categories can be produced by within model through changes to the weighting of each dimension.
+As noted in Section \ref{section:individual-diffs}, row- and column-like categories can be produced by each model through changes to the weighting of each dimension.
 Given this information, we may use the models to simulate each participant's generation separately, with the importance of each dimension set according to the relative range of the participant's generated category along each dimension.
@@
 
-In the hierarchical sampling model, the covariance matrix specifying the prior domain distribution, $\Sigma_0$, can be used to similar effect.
-This covariance matrix specifies the amount of variance assumed along each dimension (as well as the dimensional correlations) across the domain of categories.
-The covariance matrix for a newly generated category, $\Sigma_B$, is based on the assumed $\Sigma_0$ as well as the distributions of previously learned categories (see Appendix \ref{ap:hsampling-definition}).
-Thus, the importance of each feature can be coded into $\Sigma_0$ to alter the dimensional variance of generated categories.
-
 Unlike the PACKER and copy-and-tweak models, the hierarchical sampling model's dimensional variances correspond to the assumed variance of generated categories along each dimension (rather than the inverse of the variance).
@@
 
-\nbcnote{This works for 2D, but not 3D and beyond (as $1-w_k$ would not produce a valid weighting).}
-\jlanote{I renamed $\lambda$ because $\lambda$ is reserved for the correlation parameter of a covariance matrix (especially a $2 \times 2$).}
+\nbcnote{This works for 2D, but not 3D and beyond (as $1-w_k$ would not produce a valid weighting). Maybe we should make a note of that, or design something more general.}
 
@@
 The sole exception is within the `Row' condition, in which the majority of participants generated a `Row'-like category, widely distributed along the X-axis but not the Y-axis.
-In these case, both models are initialized with weights that produce Row categories, but because category contrast is not considered, categories are uniformly generated across the entire domain, rather than concentrated within the upper-regions as observed behaviorally.
+In these cases, both models are initialized with weights that produce Row categories, but because category contrast is not considered, categories are uniformly generated across the entire domain, rather than concentrated within the upper-regions as observed behaviorally.
 
@@
 \jlanote{This is pretty disorganized at the moment. There's a lot of good in here though. Here's the structure I'd like to see. Please move and reorganize it into these subsections}
+
 Summary of results

### 2017-04-21 Experiments/cogpsych-code/plot-stats-gradients.py

@@ -15,11 +15,12 @@ STAT_LIMS =  (-2.0, 2.0)
 # STAT_OF_INTEREST = 'correlation'
 # STAT_LIMS =  (-1.0, 1.0)
 
-# prior mu for empirical bayes
+# prior mu for normal inverse gamma
 PRIOR_MU = 0.0
+PRIOR_NU = 1.0
 
 # simulation params
-N_SAMPLES = 1
+N_SAMPLES = 30
 WT_THETA = 1.5
 
 # plotting settings
@@ -39,12 +40,11 @@ stimuli = pd.read_sql_query("SELECT * from stimuli", con).as_matrix()
 con.close()
 
 # get 'observed' dataframe with columns:
-# condition, stimulus, mean, var, size
+# condition, stimulus, mean, size
 observed = pd.merge(generation, stats[['participant', STAT_OF_INTEREST]], on='participant')
 observed = pd.merge(observed, info[['participant', 'condition']], on='participant')
-observed = observed.groupby(['condition','stimulus'])[STAT_OF_INTEREST].agg(['mean', 'var', 'size'])
+observed = observed.groupby(['condition','stimulus'])[STAT_OF_INTEREST].agg(['mean', 'size'])
 observed = observed.reset_index()
-observed.loc[pd.isnull(observed['var']),'var'] = 1.0
 
 # store all data (models and behavioral alike) here
 all_data = dict(Behavioral = observed)
@@ -104,9 +104,8 @@ for model_name, model_obj in name_2_object.items():
 
     # aggregate over simulations, add to all data
     model_data = model_data.groupby(['condition','stimulus'])[STAT_OF_INTEREST]
-    model_data = model_data.agg(['mean', 'var', 'size'])
+    model_data = model_data.agg(['mean', 'size'])
     model_data = model_data.reset_index()
-    model_data.loc[pd.isnull(model_data['var']),'var'] = 1.0
     model_data['size'] /= float(N_SAMPLES)
     all_data[model_obj.model] = model_data
 
@@ -128,9 +127,7 @@ for rownum, c in enumerate(row_order):
         for i, row in df.groupby('stimulus'):
             n = row['size'].as_matrix()
             sumx = row['mean'].as_matrix() * n
-            sig = row['var'].as_matrix()
-            if sig == 0: sig = 0.001
-            vals[int(i)] = (PRIOR_MU/sig +  sumx / sig) / (1.0/sig + n/sig)
[…]

### 2017-04-23 Manuscripts/cog-psych/cogpsych.tex
drafted contrast discussion, creative cognition discussion

@@
 In addition to an analysis of published data \citep[][Experiment 3]{jern2013probabilistic}, we reported two experiments demonstrating systematic effects of category contrast in creative generation: members of participant-generated categories tended to be highly dissimilar from members of previously-learned categories, and were usually more similar to one another than to members of other categories.
+Likewise, we observed broad interdependence between the distributional structure (feature variance, correlation) and physical instantiation (location within the domain space) of generated categories: In Experiment 2, we found that the unoccupied regions of the domain influenced the distributional structure of categories, and in both Experiments we observed that participants generated categories that were oriented to increase contrast with what was already known.
+
 Finally, we conducted simulations comparing PACKER's account of our results to that of a ``copy-and-tweak'' model (realized as a variant of PACKER with no sensitivity to category contrast), and a hierarchical sampling model designed to explain the classic distributional similarity effect.
 In all simulations, we found that PACKER's sensitivity to contrast considerably enhanced the models account of generation.
-
-The PACKER model bridges two gaps in the literature.
-First, it provides a quantitative theory of the role of contrast in creative generation, a topic more or less overlooked by previous work.
-The lack of attention to this issue is likely due to the underconstrained notion of what it means for two concepts to be ``different''.
-In PACKER, this notion is interpreted using the theory of similarity employed by the exemplar view of categorization \citep{medin1978context,brooks1978nonanalytic,nosofsky1984choice,nosofsky1986attention}.
-This theory of similarity, tracing its foundations to classic research in stimulus generalization \citep{attneave1950,shepard1957stimulus,shepard1987toward}, has had broad influence in many fields within cognitive science.
-Thus, beyond providing a formal specification of category contrast, PACKER also aligns the relatively understudied topic of creative generation more closely with the well-developed literature on human categorization.
-The success of the PACKER model indicates that core theories from the categorization literature may offer a strong account of how people generate new concepts.
-
-The success of PACKER yields reveals keys insights about the nature of creative generation.
-First, by measuring PACKER's fit as a function of its prioritization of within- and between-category similarity, we observed that considering either constraint exclusively results in a relatively low-quality account.
+Further, by measuring PACKER's fit as a function of its prioritization of within- and between-category similarity, we observed that considering either constraint exclusively results in a relatively low-quality account.
 Instead, PACKER's best results were obtained when both constraints are considered, indicating that human learners do not generate novel concepts exclusively on the basis of within-category similarity or between class-contrast.
@@
 
-This finding is extended by our observation of the broad interdependence between the distributional structure (feature variance, correlation) and physical instantiation (location within the domain space) of generated categories.
-In Experiment 2, we found that the unoccupied regions of the domain influenced the distributional structure of categories, and in Experiments 1 and 2 we observed that generated categories assume different distributional structures depending on how they contrast with known categories.
-This interdependence is straightforwardly explained under PACKER's dual-constraint proposal: after first establishing a location for their generated category, participants distribute exemplars in order to minimize between-category similarity while not drastically decreasing within-category similarity.
+\subsection{Similarity and Contrast in Cognition}
+\jlanote{More broad perspective of how this fits with categorization more broadly.
+Similar style model for explaining language/perceptual (phonemes and color) categories.
+Also, you should check on studies of how contrast affects categorization more broadly (i think there's some work on that).}
 
-However, PACKER does not provide a full account of what is known about category generation.
-Most notably, in this paper we have not evaluated the model's ability to explain the classic finding that generated categories tend to share distributional commonalities with previously learned categories \citep[see][]{jern2013probabilistic,ward1994structured,ward1995s,ward2002role,smith1993constraining,marsh1999inadvertent}.
+The proposal that category contrast is a primary constraint in creative generation is, in some ways, entirely commonsense.
+In order to successful create something \textit{new}, it must be different from what is known.
+Beyond its role in creative generation, category contrast is also of fundamental importance in categorization more broadly: All other factors held constant, new categories are easier to learn if they are dissimilar to members of other categories, and knowledge of highly distinct categories is applied more accurately than that of ill-defined categories\footnote{\nbcnote{This is an uncontroversial claim but we should have some cites.
+I just can't think of a super appropriate one right now...}}.
+Likewise, the act of forming category representations affects similarity judgments about category members and nonmembers, with category members being viewed as more similar to one another than members of other categories \citep{goldstone1994influences,goldstone2001altering} .
+
+%discussion of color categories and how they evolve. how across cultures, they seem to maximize within-category similarity and minimize between-category simialrity
+
+Indeed, issues relating to the trade-off between within- and between-class similarity arise in a variety of domains.
+For example, \cite{regier2007} showed that the partitioning of color categories reflects such a trade-off in a psychological space -- colors are partitioned into groups with members that are viewed as highly similar to one another yet distinct from other colors.
+A similar trade-off can be observed in phoneme perception...
+\nbcnote {I got nothing there...}
+
+The above passages serve to demonstrate that the dual forces of within-class similarity and between-class contrast influence cognitive function in a wide variety of domains.
+The PACKER model is unique in that it extends this trade-off to the domain of creative generation, and allows us to begin to understand the processes involved in creative generation using our well-developed knowledge of human categorization.
+
+\subsection{Implications for Creative Cognition}
+
+A central focus of the creative cognition approach \citep{finke1992creative,smith1995creative} has been to explain acts of creativity in terms of the mental processes that are commonly studied in cognitive psychology and cognitive science.
+The PACKER model advances this goal by applying the well-understood representational and process assumptions of the exemplar view to the act of generating new categories.
+In doing so, the model offers an additional interpretation of many of the central findings within the literature on creative cognition.
+For example, perhaps the most foundational principle from this literature concerns the limiting influence of prior knowledge: individuals create new categories composed of features from existing classes, and what is created can be influenced drastically through the introduction of cues or examples \citep{marsh1999inadvertent,smith1993constraining}.
+In this paper, we have identified another important aspect of the constraining influence of prior knowledge: what is generated cannot be the same as what is already known.
+Our results demonstrate that this influence is concisely explained in terms of a trade-off between within-category similarity and between-category dissimilarity.
+
+Interestingly, the leading account within the creative cognition literature, the Path of Least Resistance account \citep{ward1994structured,ward1995s}, also explains generation in terms of an exemplar-based retrieval process.
+This account was designed to explain the creative generation of natural categories (e.g., new species of plants and animals) and as a result relies strongly on the hierarchical organization of these categories: individuals are thought to retrieve an example of the higher-level category being generated (e.g., \textit{bird} may be retrieved from the category \textit{animal}), and then systematically alter what was retrieved to make something new.
+As the PACKER model does not assume knowledge is hierarchically organized \citep[this is true of the exemplar view more boradly, see][]{murphy2016exemplar}, the model may be viewed as a formal instantiation of the Path of Least Resistance for application in a traditional artificial categorization domain (in which there is no established hierarchy of categories).
+PACKER's success in explaining generation within an artificial domain motivates future work exploring the nature of category contrast within a more naturalistic setting.
+
+\subsection{Limitations and Future Directions}
+\jlanote{this is where the importance sampling connection will come in and will involve specifying the computational problem precisely.
+that hasn't been done yet for category generation and then perhaps to creativity more broadly.}
+
+Although successful in explaining our results, PACKER does not provide a full account of what is known about category generation.
+Most notably, in this paper we have not evaluated the model's ability to explain the classic finding that generated categories tend to share distributional commonalities with previously learned categories \citep[see][]{jern2013probabilistic,ward1994structured}.
 While we successfully replicated this effect in Experiment 1, we also found that its influence was limited in comparison to the fundamental constraints imposed by category contrast.
@@
 
-Nonetheless, these classic effects are a core element of the phenomenology of creative generation, and it is worth noting that PACKER does not include any mechanisms explain them.
+Nonetheless, these classic effects are a core element of the phenomenology of creative generation, and PACKER does not include any mechanisms explain them.
 Instead, through the development and evaluation of the PACKER model, we have sought to add new elements into such a phenomenology: specifically, the broad and strong influence of category contrast, and the interdependence between category location and distributional structure.
@@
 
-As it stands, however, PACKER's mathematical basis is subject to a wide variety of interpretations, leading to distinct views on the nature of creative generation. 
-%\nbcnote{That sounds like a good spot to include importance sampling}
-\jlanote{This is pretty disorganized at the moment. There's a lot of good in here though. Here's the structure I'd like to see. Please move and reorganize it into these subsections}
-
-Summary of results
-\subsection{Implications for creative cognition}
-more on the general theories/philosophy of creative cognition and what we bring to the table.
-\subsection{Contrast in categorization}
-More broad perspective of how this fits with categorization more broadly.
-Similar style model for explaining language/perceptual (phonemes and color) categories.
-Also, you should check on studies of how contrast affects categorization more broadly (i think there's some work on that).
-\subsection{Limitations and Future Directions}
-this is where the importance sampling connection will come in and will involve specifying the computational problem precisely.
-that hasn't been done yet for category generation and then perhaps to creativity more broadly.
-%discussion of color categories and how they evolve. how across cultures, they seem to maximize within-category similarity and minimize between-category simialrity
+Although an overall goal of this work is to advance the study of creative thought by explaining generation through quantitative approaches, PACKER is not easily aligned with many of the myriad alternative theoretical viewpoints in the study of creativity \citep[see][]{kozbelt2010theories}.
+Instead, in this paper we have studied a highly complex behavior (creative generation) as it applies within a well-established domain (category learning).
 

### 2017-04-23 Manuscripts/cog-psych/cogpsych.tex
edits to creative cognition, future directions sections

@@
 \section*{Abstract}
-
-\nbcnote{Make sure to write this!}
 
@@
 
+\begin{figure}
+    \begin{center}
+    \inputpgf{figs/}{e2-samples.pgf}
+    \caption{Sample categories generated in Experiment 2. }
+    \label{fig:e2-samples}
+    \end{center}
+\end{figure}
+
 \subsection{Participants, Materials, \& Procedure}
@@
 \subsection{Results}
-
-\begin{figure}
-    \begin{center}
-    \inputpgf{figs/}{e2-samples.pgf}
-    \caption{Sample categories generated in Experiment 2. }
-    \label{fig:e2-samples}
-    \end{center}
-\end{figure}
 
@@
     \inputpgf{figs/}{e2-yranges.pgf}
-    \caption{Y-Axis range and position of the categories generated in Experiment 2.
-Each line corresponds to a single category, with notches corresponding to the Y-axis position of exemplars within the category (notches may overlap).
-Data from each participant is sorted by overall range, and then by condition.
+    \caption{Y-Axis range and position of the participant-generated categories from Experiment 2.
+Each line corresponds to a participant's category, with notches corresponding to the Y-axis position of exemplars within the category (notches may overlap).
+Participants are sorted by overall range, and then by condition.
 }
@@
 The resulting contingencies data are shown in Table \ref{table:e2-subset-table}.
+
+Firstly, it should be noted that nearly every participant utilized the top and/or bottom rows: only $10 / 122$ participants generated their category entirely within the interior region.
+Fisher's Exact Tests comparing the conditions reveal that more Middle participants generated an exemplar in the bottom row, $p < 0.001$, again demonstrating the role of contrast in guiding where exemplars are generated.
+The conditions did not differ in use of the top of the space, $p = 0.16$, however, more Middle participants placed exemplars in the top {\em and} bottom rows, $p = 0.038$.
+The latter effect is of interest here, as it indicates that the shape of the unoccupied space exerts some influence on the distributional structure of generated categories: participants in the Middle condition were more likely to generate a category spanning the entire Y-axis.
 
@@
 \end{table}
-
-Firstly, it should be noted that nearly every participant utilized the top and/or bottom rows: only $10 / 122$ participants generated their category entirely within the interior region.
-Fisher's Exact Tests comparing the conditions reveal that more Middle participants generated an exemplar in the bottom row, $p < 0.001$, again demonstrating the role of contrast in guiding where exemplars are generated.
-The conditions did not differ in use of the top of the space, $p = 0.16$, however, more Middle participants placed exemplars in the top {\em and} bottom rows, $p = 0.038$.
-The latter effect is of interest here, as it indicates that the shape of the unoccupied space exerts some influence on the distributional structure of generated categories: participants in the Middle condition were more likely to generate a category spanning the entire Y-axis.
 
@@
 While the bulk of prior research on the topic has focused on the classic finding that generated concepts tend to be distributionally similar to known concepts, there has been little work addressing the role of contrast in creative generation: how is it that people are able to create something {\em different} from what is already known?
-This issue is of fundamental importance: in order to successfully generate something new, it must be different from what is known.
+This issue is of fundamental importance: In order to successfully generate something new, it must be in some sense distinct.
 We developed a novel, exemplar-based model, PACKER, which formally specifies the role of contrast in generation.
@@
 
-In addition to an analysis of published data \citep[][Experiment 3]{jern2013probabilistic}, we reported two experiments demonstrating systematic effects of category contrast in creative generation: members of participant-generated categories tended to be highly dissimilar from members of previously-learned categories, and were usually more similar to one another than to members of other categories.
-Likewise, we observed broad interdependence between the distributional structure (feature variance, correlation) and physical instantiation (location within the domain space) of generated categories: In Experiment 2, we found that the unoccupied regions of the domain influenced the distributional structure of categories, and in both Experiments we observed that participants generated categories that were oriented to increase contrast with what was already known.
+In addition to an analysis of published data \citep[][Experiment 3]{jern2013probabilistic}, we reported two experiments demonstrating systematic effects of category contrast in creative generation.
+Members of participant-generated categories tended to be highly dissimilar from members of previously-learned categories, and were usually more similar to one another than to members of other categories.
+We also observed broad interdependence between the distributional structure (feature variance, correlation) and physical instantiation (location within the domain space) of generated categories: In Experiment 2, we found that the unoccupied regions of the domain influenced the distributional structure of categories, and in both Experiments we observed that participants distributed their generated categories to increase contrast with what was already known.
 
-Finally, we conducted simulations comparing PACKER's account of our results to that of a ``copy-and-tweak'' model (realized as a variant of PACKER with no sensitivity to category contrast), and a hierarchical sampling model designed to explain the classic distributional similarity effect.
+We conducted simulations comparing PACKER's account of our results to that of a ``copy-and-tweak'' model (realized as a variant of PACKER with no sensitivity to category contrast), and a hierarchical sampling model designed to explain the classic distributional similarity effect.
 In all simulations, we found that PACKER's sensitivity to contrast considerably enhanced the models account of generation.
@@
 Instead, PACKER's best results were obtained when both constraints are considered, indicating that human learners do not generate novel concepts exclusively on the basis of within-category similarity or between class-contrast.
-This finding mirrors our behavioral results: it is clear that both constraints influence creative generation.
+This finding mirrors our behavioral results and demonstrates that both constraints influence creative generation.
 
@@
 In order to successful create something \textit{new}, it must be different from what is known.
-Beyond its role in creative generation, category contrast is also of fundamental importance in categorization more broadly: All other factors held constant, new categories are easier to learn if they are dissimilar to members of other categories, and knowledge of highly distinct categories is applied more accurately than that of ill-defined categories\footnote{\nbcnote{This is an uncontroversial claim but we should have some cites.
+Beyond its role in creative generation, category contrast is also of fundamental importance in categorization more broadly.
+All other factors held constant, new categories are easier to learn if they are dissimilar to members of other categories, and knowledge of highly distinct categories is applied more accurately than that of ill-defined categories\footnote{\nbcnote{This is an uncontroversial claim but we should have some cites.
 I just can't think of a super appropriate one right now...}}.
@@
 
-Indeed, issues relating to the trade-off between within- and between-class similarity arise in a variety of domains.
+Even beyond the field of categorization, one can find instances of the trade-off between within and between-class similarity scattered throughout the research in cognitive science.
 For example, \cite{regier2007} showed that the partitioning of color categories reflects such a trade-off in a psychological space -- colors are partitioned into groups with members that are viewed as highly similar to one another yet distinct from other colors.
@@
 The above passages serve to demonstrate that the dual forces of within-class similarity and between-class contrast influence cognitive function in a wide variety of domains.
-The PACKER model is unique in that it extends this trade-off to the domain of creative generation, and allows us to begin to understand the processes involved in creative generation using our well-developed knowledge of human categorization.
+The PACKER model is notable in that it successfully interprets this trade-off within the domain of creative generation, and allows us to begin to understand the relatively understudied processes involved in creative generation through our more well-developed knowledge of human categorization.
 
@@
 
-A central focus of the creative cognition approach \citep{finke1992creative,smith1995creative} has been to explain acts of creativity in terms of the mental processes that are commonly studied in cognitive psychology and cognitive science.
-The PACKER model advances this goal by applying the well-understood representational and process assumptions of the exemplar view to the act of generating new categories.
-In doing so, the model offers an additional interpretation of many of the central findings within the literature on creative cognition.
+Although the focus of this paper has been to address the processes involved in creative generation using the empirical and quantitative toolkit of traditional categorization research, our findings and approach have relevant implications for research in creative cognition.
+A central focus of the creative cognition approach has been to explain acts of creativity in terms of the mental representations and processes that are commonly studied in cognitive psychology and cognitive science \citep{finke1992creative,smith1995creative}.
+However, unlike other fields in the study of cognition, creative cognition research rarely employs quantitative models to evaluate the explanatory value of such representations and processes.
+Our modeling results provide a concrete example of how formal approaches may be used to gain insight into the nature of creative cognition.
+
+In addition to demonstrating the utility of formal modeling in the study creative cognition, the PACKER model more specifically offers an additional interpretation of some of the field's most central findings.
 For example, perhaps the most foundational principle from this literature concerns the limiting influence of prior knowledge: individuals create new categories composed of features from existing classes, and what is created can be influenced drastically through the introduction of cues or examples \citep{marsh1999inadvertent,smith1993constraining}.
-In this paper, we have identified another important aspect of the constraining influence of prior knowledge: what is generated cannot be the same as what is already known.
-Our results demonstrate that this influence is concisely explained in terms of a trade-off between within-category similarity and between-category dissimilarity.
+In this paper, we have identified another important aspect of the constraining influence of prior knowledge: What is generated cannot be the same as what is already known.
+The results of our simulations with PACKER demonstrate that this influence is concisely explained in terms of a trade-off between within-category similarity and between-category dissimilarity.
 
-Interestingly, the leading account within the creative cognition literature, the Path of Least Resistance account \citep{ward1994structured,ward1995s}, also explains generation in terms of an exemplar-based retrieval process.
+Similarly, PACKER may offer an additional interpretation of existing accounts of creative generation.
+Most notably, a leading account within the creative cognition literature, the Path of Least Resistance \citep{ward1994structured,ward1995s}, also explains generation in terms of an exemplar-based retrieval process.
 This account was designed to explain the creative generation of natural categories (e.g., new species of plants and animals) and as a result relies strongly on the hierarchical organization of these categories: individuals are thought to retrieve an example of the higher-level category being generated (e.g., \textit{bird} may be retrieved from the category \textit{animal}), and then systematically alter what was retrieved to make something new.
-As the PACKER model does not assume knowledge is hierarchically organized \citep[this is true of the exemplar view more boradly, see][]{murphy2016exemplar}, the model may be viewed as a formal instantiation of the Path of Least Resistance for application in a traditional artificial categorization domain (in which there is no established hierarchy of categories).
+As the PACKER model does not assume knowledge is hierarchically organized \citep[this is true of the exemplar view more broadly, see][]{murphy2016exemplar}, the model may be viewed as a formal instantiation of the Path of Least Resistance for application in a traditional artificial categorization domain (in which there is no established hierarchy of categories).
 PACKER's success in explaining generation within an artificial domain motivates future work exploring the nature of category contrast within a more naturalistic setting.
@@
 
-Although an overall goal of this work is to advance the study of creative thought by explaining generation through quantitative approaches, PACKER is not easily aligned with many of the myriad alternative theoretical viewpoints in the study of creativity \citep[see][]{kozbelt2010theories}.
-Instead, in this paper we have studied a highly complex behavior (creative generation) as it applies within a well-established domain (category learning).
+Finally, although an overall goal of this work is to advance the study of creative thought by explaining generation through the empirical and quantitative approaches from research in categorization, these approaches are not easily aligned with many of the myriad alternative theoretical viewpoints in the study of creativity \citep[for a review see][]{kozbelt2010theories}, such as those based on free association \citep{mednick1962associative} and conceptual combination \citep{estes2002emergence,murphy1988comprehending}.
+Instead, we reduced our focus by studying a highly complex behavior (creative generation) as it applies within a well-established domain (artificial category learning).
+Importantly, however, our findings do provide a clue about a more general, computational-level interpretation of creative generation.
+That is, in this paper we have made clear that contrast is a key constraint, as it is required in order to successful create something new.
+However, it cannot be the only constraint: It is not enough for a generated category to be distinct, as categories generated purely on the basis of contrast would not be distributionally coherent.
+Thus, a more general interpretation of creative generation would need to concern both the distinctness and coherence of what is created.
+
+\nbcnote{I think importance sampling should fit in there pretty well.}
 

### 2017-04-23 Manuscripts/cog-psych/cogpsych.tex
drafted abstract

@@
 \section*{Abstract}
+
+The ability to creatively generate new concepts and ideas is among the most fascinating aspects of human cognition, but we do not have a strong understanding of the cognitive processes and representations underlying creative generation.
+In this paper, we study the generation of new categories using the empirical and quantitative toolkit of traditional artificial category learning.
+We report two experiments demonstrating that contrast between what is known and what is created is of fundamental importance in the creative process.
+We propose a novel, exemplar-based approach explaining our results, and compare the model's performance to two key alternatives.
+Our experiments and simulations demonstrate specifically how category contrast influences creative generation (category contrast), and how it can interact with other constraints in order to produce different types of creations.
+Our work also serves as a concrete example of how well-established and highly controlled experimental frameworks (such as those in the field of category learning) can be used to make progress in an intriguing yet understudied domain.
 
@@
 
+The generation of new concepts and ideas is a highly interesting topic, but it is difficult to study in a controlled experimental environment.
+In this paper, we have provided such an examination of category generation as it applies within a traditional artificial categorization categorization experiment.
+Our experiments and simulations revealed a core constraint upon creative generation (category contrast), and showed how it can interact with other constraints in order to produce different types of creations.
+More generally, our work serves as a concrete example of how well-established and highly controlled experimental frameworks (such as those in the field of category learning) can be used to study a less well-defined topic, and how existing tools can be used to make progress in an intriguing yet understudied domain.
+
 \clearpage

### 2017-04-23 Manuscripts/cog-psych/cogpsych.tex
drafted conclusions

@@
 The generation of new concepts and ideas is a highly interesting topic, but it is difficult to study in a controlled experimental environment.
-In this paper, we have provided such an examination of category generation as it applies within a traditional artificial categorization categorization experiment.
-Our experiments and simulations revealed a core constraint upon creative generation (category contrast), and showed how it can interact with other constraints in order to produce different types of creations.
-More generally, our work serves as a concrete example of how well-established and highly controlled experimental frameworks (such as those in the field of category learning) can be used to study a less well-defined topic, and how existing tools can be used to make progress in an intriguing yet understudied domain.
+In this paper, we have provided such an examination of category generation as it applies within an artificial categorization categorization experiment.
+Extending the literature on creative cognition, our experiments provide a detailed picture of the role of category contrast in generation: People seek to create concepts that are distinct from what they already know, and the nature what is created can be influenced by what \textit{doesn't} yet exist.
+Our simulations with traditional exemplar models, as well as a hierarchical sampling model, reinforce the claim that category contrast is of fundamental importance, and, more generally, demonstrate that popular explanatory approaches from basic research in cognitive science can offer a precise, quantitative account of a behavior as complex as that of creative generation.
 
@@
 In our implementation of their model, each category's exemplars are assumed to be sampled from a multivariate Normal distribution with parameters ($\mu, \Sigma$).
-Each category's covariance matrix is assumed to be inverse-Wishart distributed with parameters ($v$, $\kappa,$ and $\Sigma_D$).\footnote{Note that \citet{jern2013probabilistic}'s model is slightly different, as they used a non-conjugate model.
-Their model acts very similar to our version of it and receives comparable fits.} $\Sigma_D$ is the covariance matrix shared between categories.
+Each category's covariance matrix is assumed to be inverse-Wishart distributed with parameters: $v$, $\kappa,$ and $\Sigma_D$\footnote{Note that \citet{jern2013probabilistic}'s model is slightly different, as they used a non-conjugate model.
+Their model acts very similar to our version of it and receives comparable fits.}.
+$\Sigma_D$ is the covariance matrix shared between categories.
 We assume the shared covariance matrix $\Sigma_D$ is generated from a Wishart distribution (for conjugacy) with parameters $v_0$, $\kappa_0$, and $\Sigma_0$.

### 2017-04-23 Manuscripts/cog-psych/cogpsych.tex
ready to pass to joe

@@
 In our implementation of their model, each category's exemplars are assumed to be sampled from a multivariate Normal distribution with parameters ($\mu, \Sigma$).
-Each category's covariance matrix is assumed to be inverse-Wishart distributed with parameters: $v$, $\kappa,$ and $\Sigma_D$\footnote{Note that \citet{jern2013probabilistic}'s model is slightly different, as they used a non-conjugate model.
-Their model acts very similar to our version of it and receives comparable fits.}.
+Each category's covariance matrix is assumed to be inverse-Wishart distributed with parameters: $v$, $\kappa,$ and $\Sigma_D$
+\footnote{Note that \citet{jern2013probabilistic}'s model is slightly different, as they used a non-conjugate model. Their model acts very similar to our version of it and receives comparable fits.}. 
 $\Sigma_D$ is the covariance matrix shared between categories.

### 2017-04-25 Manuscripts/cogsci-2017/paper.tex
cogsci revisions: changed packer to 3 param model, changed copy and tweak accordingly

@@
     \begin{center}
-    \inputpgf{figs/}{example.spaces.pgf}
-    \caption{PACKER generation of a category `B' example, following exposure to one member of category `A' and one member of category `B'.
-\textit{(a)} Predictions based on contrast similarity only $\{\phi = -1$, $\gamma = 0\}$.
-\textit{(b)} Predictions based on target similarity only $\{\phi = 0$, $\gamma = 1\}$.
-\textit{(c)} Predictions with both constraints considered $\{\phi = -1$, $\gamma = 1\}$.
-}
+    \inputpgf{figs/}{packer-examples.pgf}
+    \caption{PACKER generation of a category `B' example, following exposure to one member of category `A' and category `B'.
+The panels differ in how the trade-off between within- and between-category similarity is managed (via the $\gamma$ parameter).}
     \label{fig:example-prob-spaces}
@@
 where $w_k$ is the attention weighting of dimension $k$ ($w_k \geq 0$ and $\sum_k{w_k} = 1$), accounting for the relative importance of each dimension in similarity calculations, and $c$ ($c>0$) is a specificity parameter controlling the spread of exemplar generalization.
-For simplicity, our simulations will use uniform attention weights, except in our discussion of individual differences.
+For simplicity, our simulations will use uniform attention weights, except when otherwise noted.
 
 To generate a new example, the model considers both the similarity to examples from contrast categories as well as the similarity to examples (if any exist) in the target category.
-The aggregated similarity $a$ between generation candidate $y$ and stored exemplars $x$ is $a(y, x) = \sum_j{f(x_j) s(y, x_j)}$, where $f(x_j)$ is a function specifying the extent to which each exemplar contributes to the target category.
-For brevity, we abbreviate $f(x_j)$ as $f_j$.
-PACKER sets $f_j$ depending on exemplar $j$'s category membership: $f_j = \phi$ if $x_j$ is a member of a contrast category, and $f_j = \gamma$ if $x_j$ is a member of the target category.
-$\phi$ and $\gamma$ are free parameters ($-\infty \leq \phi, \gamma \leq \infty$) controlling the contribution of contrast- and target-category similarity, respectively (see Figure \ref{fig:example-prob-spaces}).
-Larger absolute values result in greater consideration of those exemplars, with values of 0 eliminating their effect.
-A negative value for $f_j$ produces a `repelling' effect (exemplars are less likely to be generated nearby $x_j$).
-Conversely, a positive value for $f_j$ produces an `attracting' effect (exemplars are more likely to be generated nearby $x_j$).
+The aggregated similarity $a$ between generation candidate $y$ and stored exemplars $x$ is:
 
+\begin{equation}
+  a(y, x) = \sum_j{f(x_j) s(y, x_j)}
+  \label{eq:aggregation}
+\end{equation}
+\
+where $f(x_j)$ is a function specifying the extent to which each exemplar contributes to the generation.
+A negative value for $f(x_j)$ produces a `repelling' effect (items are less likely to be generated nearby $x_j$), and a positive value produces an `attracting' effect (items are more likely to be generated nearby $x_j$).
+When $f(x_j)=0$, the exemplar does not contribute to generation.
+
+PACKER sets $f(x_j)$ depending on exemplar $j$'s category membership: $f(x_j) = \gamma$ if $x_j$ is a member of the target category, and $f(x_j) = 1 - \gamma$ if $x_j$ is a member of a contrast category.
+$\gamma$ is thus a free parameter ($0 \leq \gamma \leq 1$) controlling the trade-off between within- and between-category similarity.
 PACKER's core proposal is that new categories should be different from existing categories, and same-category exemplars should be similar to one another.
-This is realized when $\phi \leq 0$, and $\gamma \geq 0$.
-Negative $\phi$ values encourage $y$ to be distant from the contrast category (as similarity to contrast category exemplars are subtracted during aggregation).
-Positive $\gamma$ values encourage $y$ to be close to other exemplars of the target category.
-When $|\phi| = \gamma$, the repulsion effect from contrast categories is equal to the attraction effect to the target category.
+This is realized when $\gamma$ assumes an intermediate value: For example, when $\gamma = 0.5$, $f(x_j) = 0.5$ for members of the target category and $f(x_j) = -0.5$ for members of other categories; thus, the model is likely to generate items that are similar to members of the target category but are not similar to members of other categories.
+However, more extreme values can be used to produce different behavior, see Figure \ref{fig:example-prob-spaces}.
 
@@
 \citet{jern2013probabilistic} interpreted this proposal in terms of an exemplar model using the GCM \citep{nosofsky1984choice}.
-Formally, their model is equivalent to PACKER with $\{\phi = 0$, $\gamma = 1\}$ (see Figure \ref{fig:example-prob-spaces}).
-To allow for more flexible performance, and to explicitly formalize the `tweak' procedure, we slightly generalized their implementation.
+Formally, their model is equivalent to PACKER with $\gamma = 1$ (see Figure \ref{fig:example-prob-spaces}).
+In this case,  $f(x_j) = 1$ for members of the target category and $f(x_j) = 0$ for members of other categories; thus, the model considers only target-class similarity, and when no members of the target class are known, the model generates items at random.
+Thus, in our work we provide simulations from the copy-and-tweak account as a variant of PACKER with a fixed $\gamma$ parameter.
 
-In our copy-and-tweak model, all observations $z$ from the target class are equally likely to be retrieved as the source.
-When there are no observations from the target class, the first exemplar is generated uniformly at random over stimulus space.
-After a source exemplar is retrieved, the similarity between generation candidates $y$ and the retrieved exemplar is computed as per Equation \ref{eq:similarity}.
-The model's goal is to generate an item that is similar to $z$, but does not exceed a given similarity threshold: The generated item should be copied, but tweaked.
-Formally, the probability that candidate exemplar $y$ will be generated based on a source exemplar $z$ is
-
-\begin{equation}
-    p(y|z)  = \dfrac
-    { \exp \left\{\theta \cdot s(y,z) \right\} I\left(s(y,z) \leq \tau\right) }
-    {\sum_i{\exp \left\{ \theta \cdot s(y_i,z) \right\} I\left(s(y_i,z) \leq \tau\right)}} 
-\end{equation}
-% 
-where $\theta$ is a response determinism parameter, $\tau$ is the similarity threshold, and $I(\cdot)$ is the indicator function, which returns 1 when it is passed a true expression and 0 otherwise.
-So, $I\left(s(y,z) \leq \tau\right)$ is 1 when the candidate exemplar $y$ is tweaked "enough".
-When $\tau=1$, the threshold has no effect, and our model is equivalent to that of \citet{jern2013probabilistic}.
-To obtain predictions not depending on a given source example, the model's predictions can be aggregated over all possible sources: $p(y) = \sum_z{p(z)p(y|z)}$.
+Formalizing a model family where PACKER and copy-and-tweak are different parameterizations within the same framework is useful because comparison between the models thus provides a test of the explanatory value of the contrast mechanism: The account provided by copy-and-tweak will only equal that of PACKER if the contrast mechanism does not offer an advantage (i.e., if $\gamma < 1$ significantly improves model fits).
 
@@
 We recruited 122 participants from Amazon Mechanical Turk from the US equally assigned to each condition.
-Stimuli were squares varying in color (RGB 25--230) and size (3.0--5.8cm).
+Stimuli were squares varying in color (grayscale 9.8\%-90.2\%) and size (3.0--5.8cm).
 The assignment of perceptual features (color, size) to axes of the domain space (x, y) was counterbalanced across participants.

### 2017-04-25 Manuscripts/cogsci-2017/paper.tex
re-ran cogsci simulations, redid heatmaps

@@
 We developed two Alpha categories (see Figure \ref{fig:middle-bottom-conditions}): the `Bottom' Alpha category is a tight cluster in the bottom-center of the space, and the `Middle' Alpha category is identical except that it lies in the center of stimulus space.
-
-%Whereas the copy-and-tweak and hierarchical sampling models both assume explicit mechanisms to re-purpose knowledge about known categories, the PACKER model forgoes these mechanisms entirely.
-Thus, whereas the distributional structure of observed categories will be the primary determinant of generation in the copy-and-tweak and hierarchical sampling models, these properties should only affect PACKER's generation insofar as they constrain inter-exemplar similarities.
-Instead, the \textit{location} of known categories in the domain is crucial because it constrains remaining possible locations for new categories.
-
-%, instantiating the two conditions of the experiment.
-In both conditions, members of the Alpha category are tightly clustered, with equal feature variance and no feature correlations.
-Our manipulation is fairly `weak': the conditions have only a slight difference in the y-axis position of the Alpha category.
-In the `Middle' condition the Alpha category is placed in the center of the space, in the `Bottom' condition the Alpha category is placed in the bottom-center of the space.
 
@@
 To obtain an overall sense of each model's ability to explain our results, we fit each model by maximizing the log-likelihood of the model's predictions of the human results.
-The $c$, $\phi$, $\gamma$, and $\theta$ parameters were fitted for PACKER; $c$, $\tau$, and $\theta$ were fitted for the copy-and-tweak model, and $\kappa$, $\rho$, $\nu$, and $\theta$ were fitted for the hierarchical sampling model.
-Note that each model possess a $\theta$ parameter fulfilling the same role (response determinism), and PACKER shares with copy-and-tweak a specificity parameter, $c$.
+The $c$, $\gamma$, and $\theta$ parameters were fitted for PACKER; $c$, and $\theta$ were fitted for the copy-and-tweak model ($\gamma$ was fixed at 1), and $\kappa$, $\lambda$, $\nu$, and $\theta$ were fitted for the hierarchical sampling model.
+Note that each model possess a $\theta$ parameter fulfilling the same role (response determinism).
 Attention in PACKER and copy-and-tweak was set uniformly.
@@
 Each model's best-fitting parameterization is shown in Table \ref{table:model-fits}.
-Overall, PACKER outperformed copy-and-tweak and the hierarchical sampling model by a considerable margin ($\sim12\%$ improvement in log-likelihood).
+Overall, PACKER outperformed copy-and-tweak and the hierarchical sampling model by a considerable margin ($\sim11\%$ improvement in log-likelihood).
 The parameter settings associated with PACKER's best fit are exactly as expected: a strong preference for items that are similar to members of the target category but are dissimilar to members of the contrast category.
@@
 
-\subsection{Individual Differences}
+\subsection{Relation Between Category Structure \& Location}
 
@@
 Manually inspecting the data reveals four typical patterns (see Figure \ref{fig:beta.samples}): `corners' categories with one Beta example in each corner of the space, tight clusters, `column'-like categories, and `row'-like categories.
-Here, we explore the role of PACKER's attention weights ($w_1$ and $w_2$) in producing row-like or column-like categories.
+Additionally, a manual inspection reveals that each of these category types tended to be generated into distinct regions of the domain, suggesting a link between the distributional structure of a category and its location.
 
@@
  & & \textbf{Sampling} \\ \hline
- $AIC = 3476$ & $AIC = 3937$ & $AIC = 4061$  \\ 
- $c=0.565$ & $c=4.675$  & $\kappa<0.001$\\
-$\phi=-4.814$ & $\tau=0.895$ & $\nu=1.001$ \\ 
-$\gamma=4.250$ & $\theta=5.601$ &  $\rho=1.220$  \\ 
- $\theta=0.731$ &  & $\theta = 7.147$  \\ 
+ $AIC = 3474$ & $AIC = 3914$ & $AIC = 3972$  \\ 
+ $c=0.565$      & $c=4.894$  & $\kappa<0.001$\\
+ $\gamma=0.469$ & $\gamma=1$ (fixed) & $\nu=4.660$ \\ 
+ $\theta=6.632$ & $\theta=3.712$ &  $\lambda=0.423$  \\ 
+  &  & $\theta = 2.771$  \\ 
 \end{tabular}
@@
     \begin{center}
-    \inputpgf{figs/}{range-diff-gradient.pgf}
+    \inputpgf{figs/}{range-diff-gradients.pgf}
     \caption{Generated category structure as a function of location.

### 2017-04-25 Experiments/middle-bottom/modeling/format-data-for-models.py

@@ -1,45 +1,0 @@
-import sqlite3, pickle
-import pandas as pd
-
-# get behavioral data
-con = sqlite3.connect('../data/experiment.db')
-participants = pd.read_sql_query("SELECT participant, condition from participants", con)
-generation = pd.read_sql_query("SELECT * from generation", con)
-alphas = pd.read_sql_query("SELECT * from alphas", con)
-stimuli = pd.read_sql_query("SELECT * from stimuli", con).as_matrix()
-con.close()
-
-# set index to make life easier
-participants = participants.set_index('participant')
-
-# create list of all trials to simulate
-trials = []
-for pid, rows in generation.groupby('participant'):
-	
-	condition = participants.loc[pid, 'condition']
-	A =  stimuli[alphas[condition]]
-	
-	for t in range(4):
-		assigned = list( rows.loc[rows.trial <  t, 'stimulus'] )
-		response = int(  rows.loc[rows.trial == t, 'stimulus'] )
-		
-		if not any(assigned): 
-			categories = [A] 
-		else: 
-			B = stimuli[assigned,:]
-			categories = [A, B] 
-
-		trials.append(dict(
-			participant = pid,
-			trial = t,
-			categories = categories,
-			response = response
-		))
-
-print trials
-
-# save to pickle
-with open('data.pickle','wb') as f:
-	pickle.dump([trials, stimuli], f)
-
-

### 2017-04-25 Manuscripts/cogsci-2017/paper.tex
changed individual differences section to strucutre x location

@@
 Manually inspecting the data reveals four typical patterns (see Figure \ref{fig:beta.samples}): `corners' categories with one Beta example in each corner of the space, tight clusters, `column'-like categories, and `row'-like categories.
-Additionally, a manual inspection reveals that each of these category types tended to be generated into distinct regions of the domain, suggesting a link between the distributional structure of a category and its location.
+This informal inspection also reveals that each of these category types tended to be generated into distinct regions of the domain, suggesting a link between category location and distributional structure.
 
@@
 
-For each stimulus in the domain, we computed the difference in range between the features ($range(size) - range(color)$) across every generated category that had the stimulus as member.
-Taking the average of these range differences yields a gradient describing how, on average, categories were distributed for each stimulus.
+To more systematically evaluate this possibility, we computed, for each stimulus in the domain, the difference in range between the features ($range(size) - range(color)$) across every generated category that had the stimulus as member.
+Aggregating over these range differences yields a gradient describing how categories tended to distributed for each stimulus.
 These data (Figure \ref{fig:range-diff-gradient}) reveal a systematic relationship between category structure and location.
@@
 The other free parameters were set as in Table \ref{table:model-fits}.
-While there exist methods to find the optimal attention weights for a given classification \citep[see][]{vanpaemel2012using}, for simplicity we approximated the weights using the inverse of each feature's range: $w_k \propto range(k)^{-1}$.
-Thus, the Alpha and Beta categories are assumed to be distinct along dimensions that the Betas do not vary on.
-To simulate the hierarchical sampling model we set the domain covariance prior, $\Sigma_0$, proportional to the range of each feature.
+While there exist methods to find the optimal attention weights for a given classification \citep[see][]{vanpaemel2012using}, for simplicity we approximated the weights using proportionally to inverse of each feature's range: Thus, the Alpha and Beta categories are assumed to be distinct along dimensions that the Betas do not vary on.
+To simulate the hierarchical sampling model we set the domain covariance prior, $\Sigma_0$, proportional to the range (not inverted) of each feature: Thus new categories were distributed more widely along the features that each participant used more widely.
 We then simulated 50 Beta categories with each participant's weighting scheme to obtain a sense of how the relative importance of each dimension affects what types of categories are generated and where they are generated.

### 2017-04-25 Manuscripts/cogsci-2017/paper.tex
retooled experiment predictions

@@
 
-The copy-and-tweak and hierarchical sampling models were designed to explain effects of prior knowledge on the structure of categories, and so these models assume little about where in stimulus space categories are likely to be generated.
+The copy-and-tweak and hierarchical sampling models were designed to explain effects of prior knowledge on the structure of categories, but they do not make any assumptions about the role of between-category contrast.
 Indeed, when there are no known examples of the target category, both models assume that generation is random.

### 2017-05-01 Manuscripts/cog-psych/cogpsych.tex
Last pass for JLA before submitting it to Dr. Kurtz... There are a few minor things for you to look over in the doc before sending it, but not too much

@@
 The ability to creatively generate new concepts and ideas is among the most fascinating aspects of human cognition, but we do not have a strong understanding of the cognitive processes and representations underlying creative generation.
-In this paper, we study the generation of new categories using the empirical and quantitative toolkit of traditional artificial category learning.
+In this paper, we study the generation of new categories using the computational and behavioral toolkit of traditional artificial category learning.
+Previous work in this domain has focused on how the statistical structure of known categories generalizes to generated categories, indifferent to contrast between the known and generated categories.
 We report two experiments demonstrating that contrast between what is known and what is created is of fundamental importance in the creative process.
 We propose a novel, exemplar-based approach explaining our results, and compare the model's performance to two key alternatives.
-Our experiments and simulations demonstrate specifically how category contrast influences creative generation (category contrast), and how it can interact with other constraints in order to produce different types of creations.
+Our experiments and simulations demonstrate specifically how category contrast influences creative generation, and how it can interact with other constraints to produce different types of creations.
 Our work also serves as a concrete example of how well-established and highly controlled experimental frameworks (such as those in the field of category learning) can be used to make progress in an intriguing yet understudied domain.
@@
 \setlength\parindent{0.5in}
-{\em Keywords}: categorization, concepts, creativity, generation
+{\em Keywords}: categorization, concepts, creativity, generation, computational modeling, exemplar models
 \clearpage
@@
 They are fascinating and important capabilities because they have produced some of the greatest scientific breakthroughs that revolutionized the world.
-Unfortunately, technological and scientific breakthroughs are rare and the result of a great deal of cognitive effort, which makes it difficult to investigate the underlying cognitive representations and processes responsible for them using standard experimental methodology or to formalize in computational models.
+Unfortunately, technological and scientific breakthroughs are rare and the result of a great deal of cognitive effort, which makes it difficult to investigate the underlying cognitive representations and processes responsible for them using standard experimental methodology or to formalize them in computational models.
 Although we tend to focus on the most salient products of these processes (e.g., scientific breakthroughs), every person is likely to have generated many novel sentences, thoughts, and/or drawings given the combinatorial explosion of possible statements given a reasonably-sized set of primitive elements and combination rules in language, thought, and perception \citep{goldstone2003}.
-This observation provides a fantastic opportunity: By examining how people create new concepts in a domain amenable to formalization in computational models, we can investigate (some aspects of) creativity, innovation, and imagination scientifically.
+This observation provides a fantastic opportunity: By examining how people create new concepts in a domain amenable to formalizing in computational models, we can investigate (some aspects of) creativity, innovation, and imagination scientifically.
 
-Towards the aim of providing a formal account of the cognitive mechanisms involved in creativity and innovation, we explore a simplified form of creative cognition that can be studied using standard behavioral methodology and is amenable to formal modeling: category generation.
+Towards the aim of providing a formal account of the cognitive mechanisms involved in creativity and innovation, we explore a form of creative cognition that can be studied using standard behavioral methodology and is amenable to formal modeling: category generation.
 In a category generation task, participants are given some background knowledge about a domain (a cover story and/or learning one or more categories in the domain) and then are asked to generate another category in the domain.
@@
 Although this is an important characteristic of generating new categories, it cannot be the only one.
-Taken to the extreme, the best ``new'' category in terms of having the same statistical regularities to other categories would identical to a known category that is representative of the domain (and thus, not new at all).
+Taken to the extreme, the best ``new'' category in terms of having the same statistical regularities to other categories would be identical to a known category that is representative of the domain (and thus, not new at all).
 
@@
 This fundamental constraint, ``being different'', or contrasting from other categories in the relevant domain, is the focus of our work.
-Beyond the assumption that participant-generated categories are indeed novel, this constraint has been overlooked in previous research: to our knowledge, there has not been any systematic investigation addressing how generated categories {\em differ} from what is already known.
+Beyond the assumption that participant-generated categories are indeed novel, this constraint has been overlooked in previous research: To our knowledge, there has not been any systematic investigation addressing how generated categories {\em differ} from what is already known.
 Although the idea of category contrast is ubiquitous throughout the categorization literature, and extends to a variety of other fields \citep[e.g., color;][]{regier2007}, the idea that a new category should be ``different'' is vague, as there are many ways it could be different from a previously observed category.
@@
 The outline of the article is as follows.
-First we describe previous empirical work on the topic of category generation, as well as the computational approaches studied in those reports.
+First we describe previous empirical work on the topic of category generation, as well as the computational formalizations of the theories in those reports.
 Then we describe our novel computational model, which is designed to generate categories that systematically differ from existing categories in the domain.
@@
 Much of what we know about concept generation comes from the foundational literature on creative cognition.
-In a classic series of reports, Ward and colleagues \citep{ward1995s,ward1994structured,marsh1999inadvertent,ward2002role,smith1993constraining} established that category generation is highly constrained by prior knowledge: Generated categories tend to consist of features observed in known categories, and they tend to exhibit the distributional properties as found in known categories.
-In a classic study, \cite{ward1994structured} asked participants to generate new species of alien animals by drawing and describing members of the species.
+In a series of reports, Ward and colleagues \citep{ward1995s,ward1994structured,marsh1999inadvertent,ward2002role,smith1993constraining} established that category generation is highly constrained by prior knowledge: Generated categories tend to consist of features observed in known categories, and they tend to exhibit the distributional properties as found in known categories.
+In a seminal study, \cite{ward1994structured} asked participants to generate new species of alien animals by drawing and describing members of the species.
 People tended to generate species with the same features as on Earth (e.g., eyes, legs, wings), and possessing the same feature correlations as on Earth (e.g., feathers co-occur with wings).
@@
 \nbcnote{More strongly link this to creativity: what do these findings indicate?}
+\jlanote{ok with me now.}
 
-Much of the work from this area \citep[e.g.,][]{smith1993constraining,marsh1999inadvertent} focuses on how sample cues (such as an example of a species generated by other participants) can drastically diminish creativity.
-However, the broader set of observations made by Ward and colleagues provide a great deal of insight into the nature of creative generation.
+The broader set of observations made by Ward and colleagues provide a great deal of insight into the nature of creative generation.
 They indicate that people rely strongly on prior knowledge in the creative process, and people generate concepts in accord with what they already know.
+Much of the work from this area \citep[e.g.,][]{smith1993constraining,marsh1999inadvertent} focuses on how information provided to participants (such as an example of a species generated by other participants) can drastically diminish creativity.
 Theoretical accounts of these effects have primarily been grounded within the categorization literature.
@@
 This theory incorporates elements of the highly influential basic-level categories framework \citep{rosch1975cognitive,rosch1976basic}, as well as the exemplar view \citep{medin1978context,brooks1978nonanalytic}.
-Problematically, however, while this work is been incredibly useful in providing a conceptual sketch of generation theories, the hand-drawn responses relied on in the experiments paradigms precludes the development of formal approaches.
+While this work is been incredibly useful in providing a conceptual sketch of generation theories, the hand-drawn responses relied on in the experiments paradigms precludes the development of formal approaches.
 
@@
 Following a training phase during which the experimenter-defined categories were learned, participants were asked to generate novel categories of crystals.
-In a finding mirroring that of the \cite{ward1994structured} studies, \citeauthor{jern2013probabilistic} found that participants generated categories with the same distributional properties as the experimenter-defined categories: for example, after training on categories with a positive correlation between the size and saturation features (larger sized crystals were more saturated), participants generated novel categories with the same positive correlation.
+In a finding mirroring that of the \cite{ward1994structured} studies, \cite{jern2013probabilistic} found that participants generated categories with the same distributional properties as the experimenter-defined categories.
+For example, after training on categories with a positive correlation between the size and saturation features (larger sized crystals were more saturated), participants generated novel categories with the same positive correlation.
 This finding is notable, as it demonstrates that category generation can be studied in a well-known and highly controlled experimental paradigm.
@@
 The authors evaluated the predictions of several formal models on their data.
-Most notably, they showed that a hierarchical Bayesian sampling model provided the strongest account.
+Most notably, they showed that a hierarchical Bayesian sampling model provided the strongest account of their results.
 Their model views observed examples as samples from an underlying category distribution, describing the location of the category in the space, as well as how it varies along each feature.
@@
 Instead, their copy-and-tweak model corresponds to a direct exemplar-similarity approach \citep[e.g.,][]{nosofsky1984choice,nosofsky1986attention}: The model generates new items according to their similarity to known members of the target category.
-This model provided a poor account of the observed data, as the experiments devised by \citeauthor{jern2013probabilistic} were specifically designed to challenge this model.
-However, its application is notable as a first step toward explaining category generation using well-known formal approaches from the categorization literature.
+This model provided a poor account of their results, as the experiments devised by \citeauthor{jern2013probabilistic} were specifically designed to challenge this model.
+However, its application is notable as a first step toward understanding creative cognition using well-known formal approaches from the categorization literature.
 
@@
 
-From the perspective of the categorization literature, prior work on category generation has explored only one of the possible constraints that guide category generation.
+Prior work on category generation has explored only one of the possible constraints that guide category generation.
 The main concern of the published experiments has been on the distributional correspondences between learned and generated categories, and as a result most of the computational, theoretical, and empirical efforts have been towards explaining those effects.
@@
 The authors do not make predictions about the {\em location} of the category within the domain (the perceptual instantiation of category members).
-Indeed, under a strict interpretation of their model, the most probable new category to be generated is located in {\em exactly} the same location and distributional information as the given category.
+Indeed, under a strict interpretation of their model, the most probable new category to be generated is located in {\em exactly} the same location and distributional information given a single category.
 However, this is not simply an issue with their model but with the broader class of standard hierarchical Bayesian models.
@@
 The best a hierarchical Bayesian model can do at capturing contrast is to assume that the new category is placed uniformly at random over stimulus space.
-This defeats the purpose of a hierarchy as new category locations do not depend on the locations of previously observed categories.
+This defeats the purpose of a hierarchy as new category locations do not depend on the locations of previously observed categories.\footnote{It is plausible that some hierarchical Bayesian model could be created that generates categories different from each other.
+However, this model would not be a standard application or extension of any pre-existing hierarchical Bayesian model.}
 
@@
 
+\jlanote{I'd add a footnote explaining why we focus on this one only}
+
 Participants in their experiment were exposed to members of two experimenter-defined categories of `crystals' varying in hue, saturation, and size.
@@
 
-Because each experimenter-defined category in the \cite{jern2013probabilistic} experiment possessed a distinct hue shared by all members of the category, it is sensible to propose that participants would generate a category with a hue distinct from the experimenter-provided categories.
+Because each experimenter-defined category in the \cite{jern2013probabilistic} experiment possessed a distinct hue shared by all members of the category, it is sensible that participants might generate a category with a hue distinct from the experimenter-provided categories.
 If creative generation were influenced by category contrast in this way, the hues of generated categories should be systematically different from those of the experimenter-defined categories.
-Problematically, however, stimulus hue was presented using the Hue-Saturation-Value (HSV) color space, which is device-dependent and not perceptually normed such that perceived color similarity corresponds to proximity in the color space (as opposed to a color space such as CIELAB that is device-independent and equidistant sets of points correspond to pairs of colors that have the same perceptual similarity; \citealp{wyszecki1967}).
+Unfortunately, stimulus hue was encoded and presented in the Hue-Saturation-Value (HSV) color space, which is device-dependent and not perceptually normed such that perceived color similarity corresponds to proximity in the color space (as opposed to a color space such as CIELAB that is device-independent and equidistant sets of points correspond to pairs of colors that have the same perceptual similarity; \citealp{wyszecki1967}).
 Further, they did not calibrate their monitor, and so we cannot know the precise values of colors presented to participants.
@@
 However, these issues pose a significant challenge to evaluating contrast between the experimenter-defined and participant-generated categories along the hue dimension.
+It is plausible that two uncalibrated monitors could display the same HSV color and the colors be perceived in different color categories (especially for color boundaries that vary over lightness, such as the yellow-brown boundary).
 
@@
 In the \cite{jern2013probabilistic} experiment, the hue of each experimenter-defined category was selected from one of six possible values, each of which falls into one of the color groups above.
+\jlanote{Note the two color bins we made that don't contain any of their values?}
 By categorizing the participant-generated crystals likewise, we can obtain a broad measure of category contrast by determining the proportion of participant-generated crystals that fall into the same groups as the experimenter-defined categories: If contrast influences the hues of the generated categories, we should observe minimal overlap between in the color groupings.
@@
 Plotted is the number of generated items that share a color group with one of the experimenter-defined classes.
-The ``Expected'' data follows a Binomial distribution with $p=2/8$, given there were two experimenter-defined classes, and eight color groups.}
+The ``Expected'' data follows a Binomial distribution with $p=2/8=1/4$, given there were two experimenter-defined classes, and eight color groups.}
         \label{fig:jk13-huecontrast}
@@
 That is, if hue selection is not systematic, the probability that any given example will lie in the same color group as an experimenter-defined category is given by a Binomial distribution with $p = 2/8=1/4$, as there were two experimenter-defined categories and eight possible color groups.
-Chi-square goodness-of-fit tests reveal that the observed distribution in each condition is highly inconsistent with the hues being chosen at random (all $\tilde{\chi}^2>200$, $p<0.001$): Participants tended to generate items that were perceptually distinct from the categories they had learned, and were less likely to generate hues possessed by members of the experimenter-defined categories.
-
+Chi-square goodness-of-fit tests reveal that the observed distribution in each condition is highly inconsistent with the hues being chosen at random (all $\chi^2>200$, $p<0.001$): Participants tended to generate items that were perceptually distinct from the categories they had learned, and were less likely to generate hues possessed by members of the experimenter-defined categories.
+\jlanote{need degrees of freedom for $\chi^2$ tests}
 Re-analyzing the results of \cite{jern2013probabilistic} provides preliminary support that contrast plays a role in category generation.
@@
 While we did obtain similar results using a variety of alternative groupings, the hue dimension used in the \cite{jern2013probabilistic} study does not lend itself straightforwardly to the the computation of similarities, and thus we cannot be certain of whether our coding accurately approximates the psychological space of the stimuli.
+This precludes exemplar modeling of their data as it is necessary to encode exemplars in psychological space to determine the similarity between exemplars.
 By consequence, although these results likely indicate that contrast exerts {\em some} influence, they do not precisely describe the nature of that influence.
@@
 Towards providing a more precise definition of the role of contrast in creative generation, we formalized contrast in a novel exemplar model, PACKER  ({\em Producing Alike and Contrasting Knowledge using Exemplar Representations}).
-PACKER explains category generation as a balance between two fundamental constraints: generated categories should not be similar to known categories, and exemplars within each category should be similar to one another.
+PACKER explains category generation as a balance between two fundamental constraints: The category to be generated should not be similar to known categories, and exemplars within each category should be similar to one another.
 These ideas are implemented within the well-studied exemplar framework -- the PACKER model is an extension of the influential Generalized Context Model of category learning \citep[GCM;][]{nosofsky1984choice,nosofsky1986attention}.
@@
 For simplicity, attention will be distributed uniformly in our simulations (unless otherwise noted).
-The value of $r$ depends on the nature of the experimental conditions being simulated: $r=1$ is appropriate for separable dimensions, whereas $r=2$ is appropriate for integral dimensions \citep[see][]{shepard1964attention,garner1974processing}.
+The value of $r$ depends on the nature of the experimental conditions being simulated: $r=1$ is appropriate for separable dimensions, whereas $r=2$ is appropriate for integral dimensions \citep[e.g.,][]{shepard1964attention,garner1974processing}.
 In our simulations, we set $r=1$ due to the separable nature of the stimulus dimensions used in our experiments (see Figure \ref{fig:e1-conditions}).
@@
 For example, as will be discussed in more detail below, PACKER is formally equivalent to the copy-and-tweak model proposed by \cite{jern2013probabilistic} when $\gamma = 1$.
-Likewise, when $\gamma = 0$, PACKER represents a contrast-only generation mode, relying exclusively on contrast when generating new categories.
+Likewise, when $\gamma = 0$, PACKER can represent a contrast-only generation mode, relying exclusively on contrast when generating new categories.
 Finally, when $f(x_j) = -1$ for all $x_j$ (regardless of class membership), a ``pure-packing'' approach is yielded, generating items in unoccupied areas of the domain.
@@
 For example, although it may be possible to extend \cite{jern2013probabilistic}'s hierarchical sampling model to include contrast, this mechanism would be at odds with a hierarchical Bayesian framework.
-In contrast, these ideas emerge naturally from the exemplar view: we have not modified any of the core elements of the GCM in defining the PACKER model, we simply aggregated similarity slightly differently.
+In contrast, these ideas emerge naturally from the exemplar view: We have not modified any of the core elements of the GCM in defining the PACKER model, we simply weight exemplar similarity during aggregation.
 Thus, beyond formally specifying the role of contrast in creative generation, the PACKER account allows us to evaluate the well-understood principles of the exemplar view within the comparatively understudied field of conceptual generation.
@@
 
-Panels B-D of Figure \ref{fig:e1-conditions} show the locations of exemplars belonging to the experimenter-defined categories (`A', or `Alpha') that participants were assigned to learn about prior to generating a new category.
+Panels B-D of Figure \ref{fig:e1-conditions} show the values of exemplar dimensions belonging to the experimenter-defined categories (`A', or `Alpha') that participants were assigned to learn about prior to generating a new category.
 In the `Cluster' type, category A is a tight cluster of examples in the space.
@@
 It should be noted that in our experiments the assignment between the perceptual and conceptual dimensions (e.g., $X \rightarrow Size$, $Y \rightarrow Color$), as well as the direction of variation along each dimension (e.g., $dark \rightarrow light$ or $light \rightarrow dark$) was counterbalanced across participants.
-The category types in Figure \ref{fig:e1-conditions} are plotted in a conceptual space, rather than a perceptual space: thus, while the conceptual organization of the category types remains constant, each category type may have a different physical instantiation according to the counterbalance assignment.
+The category types in Figure \ref{fig:e1-conditions} are plotted in a conceptual space, rather than a perceptual space.
+Thus, while the conceptual organization of the category types remains constant, each category type may have a different physical instantiation according to the counterbalance assignment.
 For example, the Cluster type  may be large and dark in color, or it may be small and light in color, depending on the assignment and direction of the dimensions.
@@
 
-In this way, we can analyze the results of the first experiment as a conceptual replication of the classic finding that generated categories tend to share distributional properties with known categories in the domain \citep[see][]{jern2013probabilistic,ward1994structured}.
+In this way, one aspect of the analyses of the experimental results is a conceptual replication of the classic finding that generated categories tend to share distributional properties with known categories in the domain \citep[see][]{jern2013probabilistic,ward1994structured}.
 From these results, we can predict that, in each condition, participants should generate categories that are distributionally similar to the experimenter-defined category: In the Cluster condition, generated categories should be tightly clustered.
@@
 Thus, if contrast plays a role, exemplars in the generated categories of participants in the XOR condition may not be positively correlated -- they may be negatively correlated instead.
+In this case, contrast and statistical regularities would interact, which would be inconsistent with the hierarchical sampling approach of creative cognition \citep{jern2013probabilistic}
 
-\subsection{Participants \& Materials}
+\subsection{Participants and Materials}
 
@@
 Participants were randomly assigned to one condition: 64 participants were assigned to the Cluster condition, 61 were assigned to the Row condition, and 58 were assigned to the XOR condition.
-Stimuli were squares varying in color (RGB 25--230) and side length (3.0--5.8cm), see Figure \ref{fig:e1-conditions}.
+Stimuli were squares varying in color (grayscale 9.8\%--90.2\%) and side length (3.0--5.8cm), see Figure \ref{fig:e1-conditions}.
 The assignment of perceptual features (color, size) to axes of the domain space (x, y), as well as the direction of variation along each axis (e.g., $dark \rightarrow light$ or $light \rightarrow dark$) was counterbalanced across participants.
@@
 As in \citet{jern2013probabilistic}, generation was completed using a sliding-scale interface.
-Two scales controlled the features (color, size) of the generated example.
+Two scales controlled the values of the two dimensions (color, size)  for the generated example.
 An on-screen preview of the example updated whenever one of the features was changed.
@@
 
-To begin, it is worth noting that we observed a substantial degree of individual differences in our data.
+We observed a substantial degree of individual differences in our data.
 In Figure \ref{fig:e1-samples} we have plotted sample data from several participants, from which it evident that different participants generated qualitatively different category structures.
-In this section we will focus on what can be learned from analyzing the data in aggregate, but in later sections we will explore how individual differences can be explained.
+In this section we will focus on analyzing the data in aggregate, but in later sections we will explore how individual differences can be explained.
 
@@
 These data, shown in Figure \ref{fig:e1-distanceplots}, reveal a clear pattern: Examples that are more distant from members of the experimenter-defined categories are more likely to be generated into a new category.
+This supports contrast being a fundamental constraint on creative cognition and statistical regularity alone being insufficient.
 
@@
 Figure \ref{fig:e1-distanceplots} also depicts, for each participant, the average distance of members within the generated category ({\em within-category} distance) against the average distance between members of the generated and experimenter-defined category ({\em between-category} distance).
-The narrow distribution of between-category distances in the XOR condition reflects the widely distributed nature of the experimenter-defined category: all possible generation candidates are distant from at least one cluster of the experimenter-defined category.
-These data reveal a systematic pattern: the majority of participants generated categories with more between-category distance than within-category distance.
+The narrow distribution of between-category distances in the XOR condition reflects the widely distributed nature of the experimenter-defined category, reducing the possible distances to the experimenter-defined category.
+These data reveal a systematic pattern: The majority of participants generated categories with greater between-category distance than within-category distance.
 That is, members of the generated category tended to be more similar to one another than to members of the experimenter-defined category.
-To formally evaluate this claim, we conducted t-tests comparing the amount of within- and between- class distance in each condition.
+To evaluate this claim quantitatively, we conducted t-tests comparing the amount of within- and between- class distance in each condition.
 All conditions possessed greater between-category distance: Cluster, $t(63) = 11.43$, $p < 0.001$; Row, $t(60) = 13.16$, $p < 0.001$; and XOR, $t(57) = 3.64$, $p < 0.001$.
-These results provide clear evidence of an effect of category contrast: participants prefer to generate categories that are dissimilar to the learned category but maintain some level of internal cohesion.
+These results provide further evidence of an effect of category contrast: Participants prefer to generate categories that are dissimilar to the learned category but maintain some level of internal cohesion.
 
@@
 Likewise, in each condition participants generated categories possessing strongly positive, neutral, and strongly negative correlations between the dimensions.
-Comparing the distributional statistics between conditions yields a broad-yet-misleading replication of the classic effect.
+Comparing the distributional statistics between conditions yields a broad yet, as we will see, misleading replication of the classic effect.
 
@@
 Categories generated in the XOR condition were also less widely distributed along the X-axis compared to Row, $t(117) = 2.56$, $p = 0.012$.
-This latter effect was not expected; however, the key finding is that categories from the Cluster condition tended to be more tightly distributed along the X-axis.
+This latter effect was not expected because the experimenter-defined category for XOR and Row had similar X-ranges.
+However, the key finding is that categories from the Cluster condition tended to be more tightly distributed along the X-axis.
 
@@
 This finding is notable, as it is the opposite of what would be expected, assuming learners are emulating the distributional structure of the experimenter-defined class (which possesses perfectly positively correlated features).
+We believe that the failure to replicate this finding is because participants in \cite{jern2013probabilistic} could differentiate the generated category on a third dimension (hue) to maintain the statistical regularities on the other two dimensions.
+Although the correlation in the XOR condition is significantly negative, it is clear from the box-plot in Figure ~\ref{fig:e1-statsboxes} that it would be inappropriate to make a strong conclusion (e.g., the median is close to zero).
+However, we can conclude with confidence that there are situations where people do not emulate the distributional structure of the given category.
 This indicates that there is more to category generation than the distributional structure of other categories in the domain.
-Further, as we will discuss in more detail in the model-based analysis section, it is expected by our proposal that contrast is a fundamental principle in category generation.
+Further, as we will discuss in more detail in the model-based analysis section, this is expected by our proposal that contrast is a fundamental principle in category generation.
 
@@
 
-Thus, while the core results of Experiment 1 indicate that generated categories systematically contrast with what is already known, the negative correlations observed in the XOR condition may signal something further.
+Thus, the core results of Experiment 1 indicate that generated categories systematically different from what is expected based on prior work.
+The negative (or zero) correlations observed in the XOR condition suggests an interesting interaction between contrasting from a given category and emulating statistical properties.
 That is, the constraints on creative generation imposed by category contrast may not simply influence the {\em location} of generated categories, but also their distributional structure.
@@
 
-In order to test whether category contrast influences the distributional structure of generated categories, we sought to identify conditions in which differences in the distributional structure of generated categories cannot be explained by the distributional structure of the experimenter-defined category.
-We created two new category types (depicted in Figure ~\ref{fig:e2-conditions}) that possess an identical distributional structure (both are tight clusters of examples with a neutral feature correlation), and only differ slightly in their Y-axis position: the `Bottom' category lies in the bottom-center of the space, and the `Middle' category lies in the center.
-The distributional equality of these conditions is key to the design of the experiment: if the structure of previously learned categories were the only influence on the structure of generated categories, we should observe no difference between these two conditions with respect to distributional structure.
+To test whether category contrast influences the distributional structure of generated categories, we sought to identify conditions in which differences in the distributional structure of generated categories cannot be explained by the distributional structure of the experimenter-defined category.
+We created two new category types (depicted in Figure ~~\ref{fig:e2-conditions}) that possess an identical distributional structure (both are tight clusters of examples with no correlation between features), and only differ in their Y-axis position: the `Bottom' category lies in the bottom-center of the space, and the `Middle' category lies in the center.
+The distributional equality of these conditions is key to the design of the experiment: If the structure of previously learned categories were the only influence on the structure of generated categories, we should observe no difference between these two conditions with respect to distributional structure.
 
@@
 
-Critically, however, participants seeking to make a perceptually distinct category would be more likely to distribute members of the generated category into areas that are distant from members of known categories.
+Critically, participants seeking to generate a distinct category, contrasting from the experimenter-given category, would be more likely to distribute members of the generated category into areas that are distant from members of known categories.
 Thus, if category contrast influences the distributional structure of the categories people generate, then we should observe different types of categories according to the shape of the space that is unoccupied by members of previously learned categories.
-The difference in the Y-axis position between the Bottom and Middle types produces a considerable change to the shape of the unoccupied space: participants assigned to learn the Bottom category would be less likely to generate exemplars into the lower regions of the stimulus space (as these areas possess greater similarity to members of the Bottom category), preferring instead to distribute exemplars across the upper region of the space.
+The difference in the Y-axis position between the Bottom and Middle types produces a considerable change to the shape of the unoccupied space.
+Participants assigned to learn the Bottom category should be less likely to generate exemplars into the lower regions of the stimulus space (as these areas possess greater similarity to members of the Bottom category), preferring instead to distribute exemplars across the upper region of the space.
 This constraint is lifted in the Middle condition, as the Middle category exemplars are equidistant to the upper and lower regions of the space.
 Accordingly, participants should be more likely to utilize both of these areas.
-Thus, to the extent that category contrast influences distributional structure, we should observe more participants in the Middle condition that generate examples above {\em and} below the experimenter-defined category.
+Thus, if category contrast influences the distributional structure of generated categories, we should observe more participants in the Middle condition that generate examples above {\em and} below the experimenter-defined category.
 
@@
 
-\subsection{Participants, Materials, \& Procedure}
+\subsection{Participants, Materials, and Procedure}
 
@@
 61 participants were randomly assigned to the Middle and Bottom conditions each.
-The stimuli and procedure were exactly as in Experiment 1: participants first completed a short training phase, followed by the generation phase.
+The stimuli and procedure were exactly as in Experiment 1.
+Participants first completed a short training phase, followed by the generation phase.
 
@@
 The data from each condition is organized into four columns based on commonly observed patterns of generation: a `Cluster' type of tightly-clustered examples,   `Row' and `Column' types of exemplars widely distributed along the one axis but narrowly along the other, and a `Corners' type, wherein participants placed exemplars in disparate corners of the space.
-As before, in this section we focus on what can be concluded from analyzing the data in aggregate, but in later sections we will focus more specifically on explaining the individual differences.
+As before, in this section we focus on analyzing the data in aggregate, but in later sections we will focus more specifically on explaining the individual differences.
 
@@
 
-We began our analysis by again testing for the broad influence of category contrast on generation.
-As in Experiment 1, we computed the frequency each stimulus was generated as a function of its average distance from members of the experimenter defined category, as well as each participant's average within- and between- category distance.
-These data, shown in Figure \ref{fig:e2-distanceplots}, yield highly similar results.
+We began our analysis by testing for the broad influence of category contrast on generation.
+As in Experiment 1, we computed the frequency each stimulus was generated as a function of its average distance from members of the experimenter-defined category, as well as each participant's average within- and between- category distance.
+These data, shown in Figure \ref{fig:e2-distanceplots}, yield very similar results.
 Stimuli that are more distant from members of the experimenter-defined category were more frequently generated, and the categories in each condition tended to possess more between-category than within-category distance: Bottom, $t(60) = 5.5$, $p < 0.001$; Middle, $t(60) = 2.71$, $p = 0.009$.
-
 We did, however, observe a notable subgroup of participants in each condition who generated categories with more within-category than between-category distance.
@@
 
-To address out main question, we followed this analysis by comparing the conditions on measures of distributional structure.
-For each generated category, we computed the range of exemplars along each axis (X, Y), as well as the correlation between features.
-These data, shown in Figure \ref{fig:e2-statsboxes}, again demonstrate the degree of individual differences observed in our study: in each condition, we observed tightly clustered and widely distributed categories along each dimension, as well as positively, negatively, and uncorrelated categories.
+To explore the aggregate distributional structure of the results, we computed the range of exemplars along each axis (X, Y), as well as the correlation between features.
+These data, shown in Figure \ref{fig:e2-statsboxes}, again demonstrate the degree of individual differences observed in our study.
+In each condition, we observed tightly clustered and widely distributed categories along each dimension, as well as positively, negatively, and uncorrelated categories.
 
@@
 
-However, our ability to detect differences between the conditions is, in this case, diminished due to the non-normality of the data.
+However, our ability to detect differences using a standard $t$-test between the conditions is, in this case, diminished due to the non-normality of the data (Shapiro-Wilk normality test $W=0.77, p< 0.001$ and $W=0.85, p < 0.001$ for the Bottom condition).
 Figure \ref{fig:e2-yranges} depicts the Y-axis position of the exemplars generated within each participant's category.
@@
 The conditions did not differ in use of the top of the space, $p = 0.16$, however, more Middle participants placed exemplars in the top {\em and} bottom rows, $p = 0.038$.
-The latter effect is of interest here, as it indicates that the shape of the unoccupied space exerts some influence on the distributional structure of generated categories: participants in the Middle condition were more likely to generate a category spanning the entire Y-axis.
+The latter effect is of interest here, as it indicates that the shape of the unoccupied space exerts some influence on the distributional structure of generated categories: Participants in the Middle condition were more likely to generate a category spanning the entire Y-axis.
+Thus, distributional structure of the generated categories can be influenced by category contrast alone because the distributional structure of the given categories did not differ between groups.
 
@@
 
-In Experiment 2, we replicated the core findings from Experiment 1: stimuli are more likely to be generated if they are distant from exemplars in other categories, and most participants generate categories with more between-category than within-category distance.
-However, we additionally found that the {\em position} of a previously learned category (rather than its distributional structure) influences the types of categories people generate: participants who learned the `Middle' type were more likely to generate categories spanning the entire Y-axis of the space.
+In Experiment 2, we replicated the core findings from Experiment 1.
+Stimuli are more likely to be generated if they are distant from exemplars in other categories, and most participants generate categories with more between-category than within-category distance.
+However, we additionally found that the {\em position} of a previously learned category (rather than its distributional structure) influences the types of categories people generate: Participants who learned the `Middle' type were more likely to generate categories spanning the entire Y-axis of the space.
 Participants who learned the `Bottom' type were less likely to do so as a result of the presence of opposite category exemplars in the lower regions of the space.
 
-This finding cannot be explained under the viewpoint that the distributional structure of previously learned categories is the sole determinant of the distributional structure of generated categories.
-However, the observed performance is sensible given a category contrast perspective: participants seeking to generate a perceptually distinct category will be more likely to use areas of space that are unoccupied by exemplars belonging to previously learned categories.
+This finding cannot be explained from the perspective that the distributional structure of previously learned categories is the sole determinant of the distributional structure of generated categories.
+However, the observed behavior is expected from a category contrast perspective: Participants seeking to generate a perceptually distinct category will be more likely to use areas of space that are unoccupied by exemplars belonging to previously learned categories.
 In the Middle condition, the upper and lower regions of space are equidistant from members of the experimenter-defined category, whereas in the Bottom condition, the lower region of the space is closer to members of the experimenter-defined category.
@@
 The comparison between the hierarchical sampling model and PACKER is meant to emphasize the necessity of contrast and demonstrate it cannot be explained by emulating of distributional structure: whereas PACKER is insensitive to the distributional structure of learned categories (relying only on within- and between-category similarity), the hierarchical sampling model generates categories exclusively on the basis of knowledge of how existing classes are distributed.
-Our approach in this section is to first broadly evaluate and compare the quality of each model's account to our entire dataset (Experiments 1 and 2 combined), then we more specifically describe the strengths and weakness of each model's account.
+Our approach in this section is to first broadly evaluate and compare the quality of each model's account to our entire dataset (Experiments 1 and 2 combined), then analyze the ability for each model to explain individual differences in each experiment, and lastly we more describe the strengths and weakness of each model's account of category generation.
 
@@
 Table \ref{table:global-model-fits} contains the AIC values of each model's best fit, as well as the corresponding log-likelihood ($L$) and the best-fitting parameter values.
-These results reveal strong model differentiation: the PACKER model achieved far better fits compared to the copy-and-tweak and hierarchical sampling models, and copy-and-tweak performed somewhat better than the hierarchical sampling model.
+These results reveal strong model differentiation: The PACKER model achieved far better fits compared to the copy-and-tweak and hierarchical sampling models, and copy-and-tweak performed somewhat better than the hierarchical sampling model.
 While PACKER's advantage may tentatively be attributed to the model's sensitivity to category contrast (this will be explored in detail below), the advantage shown by copy-and-tweak over the hierarchical sampling model may be attributed to its exemplar-based representation, as opposed to the prototype-based representation assumed in the hierarchical sampling model.
@@
 Finally, because copy-and-tweak is nested within PACKER, we can use a likelihood ratio test to compare the two models.
-PACKER explains the aggregate data significantly better than copy-and-tweak ($\chi^2(1) = 748, p < 0.001$ for all data and $\chi^2(1) = 378, p < 0.001$ excluding the first example), providing more targeted evidence that category generation is better explained when contrast is considered.
+PACKER explains the aggregate data significantly better than copy-and-tweak ($\chi^2(1) = 748, p < 0.001$ for all data and $\chi^2(1) = 378, p < 0.001$ excluding the first example), providing further evidence that category generation is better explained when contrast is considered.
 
 \nbcnote{report something about robustness to model comparison metric (BIC and $AIC_c$).}
+\jlanote{please -- i believe that they are derivable from log-likelihood, number of parameters, and number of data points}
 
 Through comparison with the copy-and-tweak model, Figure \ref{fig:packer-loglike} more clearly demonstrates the robustness of the explanatory gains yielded by PACKER's category contrast mechanism.
-Plotted is PACKER's log-likelihood as a function of $\gamma$ parameter.
+It displays the log-likelihood of the participants' results under PACKER as a function of $\gamma$ parameter.
 The model's other parameters ($c$, $\theta$) were set according to copy-and-tweak's best fits from Table \ref{table:global-model-fits}, and thus when $\gamma=1$, the models are equivalent.
@@
 Notably, however, the copy-and-tweak parameterization ($\gamma=1$) performs better than the `contrast-only' parameterization of the model ($\gamma = 0$).
-Thus, it is evident that the data is better explained when both within-category similarity and category contrast is considered.
+In sum, the data are better explained when both within-category similarity and category contrast is considered.
 
@@
 
-As noted in Experiments 1 and 2, we observed a great deal of individual differences in the types of categories generated.
+As noted in Experiments 1 and 2, we observed a great deal of individual differences in the types of categories that participants generated.
 Within each condition, we observed tightly clustered and widely distributed categories, as well as row and column categories, and so forth (see Figures \ref{fig:e1-samples} and \ref{fig:e2-samples}).
 The simulations reported above serve to evaluate the models while considering the entire dataset, but a secondary goal of any formal account should be to provide some explanation of how different profiles of performance emerge.
-Many of the individual generation profiles we observed can be targeted within the PACKER framework, simply by tuning the model's parameters.
+Many of the individual generation profiles we observed can be described within the PACKER framework, simply by tuning the model's parameters.
 In this section, we describe more specifically how the most frequently observed profiles can be realized.
@@
 By manual inspection, it is evident that the most common profiles of generation consist of: (A) a tightly-distributed `cluster' of examples, (B) `row'- and `column'-like arrangements (varying widely along one dimension but not the other), and (C) a `corners' arrangement with examples placed into disparate corners of the space.
-These four profiles are distinct in terms of the distribution of the generated category along each dimension: whereas the cluster profile is tightly distributed along both dimensions, the row and column profiles are tightly distributed along just one dimension.
+These four profiles are distinct in terms of the distribution of the generated category along each dimension: Whereas the cluster profile is tightly distributed along both dimensions, the row and column profiles are tightly distributed along just one dimension.
 Finally, the corners profile is widely distributed along both dimensions.
@@
 For example, in the cluster profile, there is a high degree of within-category similarity along both dimensions, whereas in the corners profile there is minimal within-category similarity.
-Thus, PACKER's proposal is that these individual differences arise as a result of different priorities: While the tight cluster configuration can be considered PACKER's `default' mode (as it maximizes within-category similarity), the corners profile can be produced when between-category contrast is put at a higher priority (i.e., $\gamma = 0$).
+Thus, PACKER's proposal is that these individual differences arise as a result of different priorities: While the tight cluster configuration can be considered PACKER's `default' mode (as it maximizes within-category similarity), the corners profile can be produced when between-category contrast is put at a higher priority (i.e., $\gamma$ near $0$).
 
-Likewise, in the row and column profiles, there is a high degree of within-category similarity along one dimension but not the other.
-These differences likely arise due to a differential focus on one dimension over another, and thus they can be produced by changes to PACKER's attention weights, $w$ (see Equation \ref{eq:similarity}).
+Likewise, in the row and column profiles, there is a large degree of within-category similarity along one dimension but not the other.
+These differences likely arise due to a differential focus on one dimension over another, and thus they can be produced by changes to PACKER's attention weights, $w_1$ and $w_2$ (see Equation \ref{eq:similarity}).
 Traditionally, the attention weights in exemplar models are thought to reflect the diagnostic value of each dimension towards classifying the known category members \citep{nosofsky1984choice,nosofsky1986attention,kruschke1992alcove}, but within a generation context the weights specify the importance of within- and between-category similarity along each dimension.
-For example, if 100\% of attention is allocated along the X-axis, similarity along the Y-axis no longer influences performance: as a result, PACKER will create categories that are more widely distributed along the Y-axis, as similarity is not counted along that dimension.
+For example, if all of attention is allocated along the X-axis ($w_1=1$ and $w_2=0$), similarity along the Y-axis no longer influences performance.
+As a result, PACKER will create categories that are more widely distributed along the Y-axis in this case, as similarity is not taken into account along that dimension.
 As a general principle, differentially weighting one dimension will result in the generation of categories that are more widely distributed along the ignored dimension, conforming to a row- or column-like arrangement.
@@
 
-Like in PACKER, changes to the weighting of the dimensions can also be used to produce row- and column-like categories in the copy-and-tweak and hierarchical sampling models.
+As in PACKER, changes to the weighting of the dimensions can also be used to produce row- and column-like categories in the copy-and-tweak and hierarchical sampling models.
 Indeed, as copy-and-tweak is simply a special case of the PACKER model, the attention weights operate exactly as described above.
@@
 
-However, a key limitation within both in both of these models is the lack of interaction between the distributional structure of a category (in this case, a row-versus-column orientation) and the location of that category within the domain.
+However, a key limitation of both of the copy-and-tweak and hierarchical sampling models is the lack of interaction between the distributional structure of a category (in this case, a row-versus-column orientation) and the location of that category within the domain.
 While these models are {\em capable} of generating row and column categories, there is no mechanism in place to ensure generated categories will be distinct from what is already known.
-In the section below, we explore the interdependence between distributional structure and location in creative generation.
+In the next subsection, we explore the interdependence between distributional structure and location in creative generation.
 
@@
 Instead, we used a Bayesian analysis to estimate the mean $\mu_x$ on the assumption that the scores $x$ for each stimulus are normally distributed with an unknown mean and unknown standard deviation.
-The conjugate normal-inverse-gamma distribution provides a straightforward method for this estimation:
+The conjugate Normal-Inverse Gamma distribution provides a straightforward method for this estimation:
 
@@
 The default assumption is that there is an equal amount of range along the X- and Y-axes, and so we set $\mu_0 = 0$.
-Likewise, to give a moderate amount of weighting to the prior mean we set $\nu_0 = 1$, though our results are robust to a range of values.
+Likewise, to give a moderate amount of weighting to the prior mean we set $\nu_0 = 1$, though the results are robust to a range of values.
 Within this approach, the resulting aggregation is a trade-off between the number of generations and the strength of the range difference within each generated category.
 Infrequently generated stimuli, as well as those with both strongly positive and negative scores, are given neutral difference scores\footnote{Prior to plotting, data was also processed using a Gaussian filter with $\sigma = 0.8$}.
-
-\nbcnote{Those two paragraphs are, i think, either: (A) too specific and could be in an appendix, or (B) totally straightforward to anyone who knows about Bayesian stats, and so we just need a normal inverse gamma cite.}
 
@@
     \caption{Behavioral and simulated range difference gradients.
-Each panel shows, for each stimulus, the dimensional orientation of the categories it was generated into: vertically aligned `columns' versus horizontally aligned `rows'.}
+Each panel shows, for each stimulus, the dimensional orientation of the categories it was generated into: vertically aligned `columns' (orange) versus horizontally aligned `rows' (purple).}
     \nbcnote{Need to run this on lando with many more samples. current = 30}
@@
 
-The results of our analysis are shown in Figure \ref{fig:range-diff-gradients}.
+The results of our analysis are shown in Figure \ref{fig:range-diff-gradients} for the experiment and model results.
+The left-most column of Figure \ref{fig:range-diff-gradients} displays the effect of category location and contrast on the distributional structure of the category generated by participants.
 These data reveal strong and consistent patterns across all the conditions we tested in Experiments 1 and 2: generated categories are more tightly distributed along the axis in which they are distinct.
@@
 These patterns of performance clearly depict the interdependence between the distributional structure and location of generated concepts.
-Our results can, in some sense, be interpreted in terms of local minimization of between-category similarity: by distributing the generated category away from members of the experimenter-defined category, participants may increase the degree of between-category distance without drastically altering the degree of within-category similarity.
+Our results can be interpreted in terms of local minimization of between-category similarity: by distributing the generated category away from members of the experimenter-defined category, participants may increase the degree of between-category distance without drastically altering the degree of within-category similarity.
 
-To attempt to explain our findings within the PACKER, copy-and-tweak- and hierarchical sampling models, we conducted simulations using an individual-differences approach.
+To explore how well PACKER, copy-and-tweak- and hierarchical sampling models explain our findings we conducted simulations using an individual-differences approach.
 As noted in Section \ref{section:individual-diffs}, row- and column-like categories can be produced by each model through changes to the weighting of each dimension.
@@
 \nbcnote{This works for 2D, but not 3D and beyond (as $1-w_k$ would not produce a valid weighting). Maybe we should make a note of that, or design something more general.}
+\jlanote{Isn't it just $1-\sum_{j\neq k}{w_j}$?}
 
@@
 Conversely, because the copy-and-tweak and hierarchical sampling models are insensitive to category contrast, these models do not produce any systematic patterns of association between category location and distributional structure.
-The sole exception is within the `Row' condition, in which the majority of participants generated a `Row'-like category, widely distributed along the X-axis but not the Y-axis.
+The sole exception is within the `Row' condition of Experiment 1, in which the majority of participants generated a `Row'-like category, widely distributed along the X-axis but not the Y-axis.
 In these cases, both models are initialized with weights that produce Row categories, but because category contrast is not considered, categories are uniformly generated across the entire domain, rather than concentrated within the upper-regions as observed behaviorally.
@@
 The creative generation of novel concepts is an intriguing yet understudied topic in cognitive science.
-While the bulk of prior research on the topic has focused on the classic finding that generated concepts tend to be distributionally similar to known concepts, there has been little work addressing the role of contrast in creative generation: how is it that people are able to create something {\em different} from what is already known?
-This issue is of fundamental importance: In order to successfully generate something new, it must be in some sense distinct.
+While the bulk of prior research on the topic has focused on the classic finding that generated concepts tend to be distributionally similar to known concepts, there has been little work addressing the role of contrast in creative generation: How is it that people are able to create something {\em different} from what is already known?
 We developed a novel, exemplar-based model, PACKER, which formally specifies the role of contrast in generation.
@@
 We conducted simulations comparing PACKER's account of our results to that of a ``copy-and-tweak'' model (realized as a variant of PACKER with no sensitivity to category contrast), and a hierarchical sampling model designed to explain the classic distributional similarity effect.
-In all simulations, we found that PACKER's sensitivity to contrast considerably enhanced the models account of generation.
+In all simulations, we found that PACKER's sensitivity to contrast considerably enhanced the models account of human category generation.
 Further, by measuring PACKER's fit as a function of its prioritization of within- and between-category similarity, we observed that considering either constraint exclusively results in a relatively low-quality account.
@@
 \subsection{Similarity and Contrast in Cognition}
-\jlanote{More broad perspective of how this fits with categorization more broadly.
-Similar style model for explaining language/perceptual (phonemes and color) categories.
-Also, you should check on studies of how contrast affects categorization more broadly (i think there's some work on that).}
-
 The proposal that category contrast is a primary constraint in creative generation is, in some ways, entirely commonsense.
-In order to successful create something \textit{new}, it must be different from what is known.
+To successful create something \textit{new}, it must be different from what is known.
 Beyond its role in creative generation, category contrast is also of fundamental importance in categorization more broadly.
@@
 
-Even beyond the field of categorization, one can find instances of the trade-off between within and between-class similarity scattered throughout the research in cognitive science.
+Beyond categorization, one can find instances of the trade-off between within and between-class similarity in linguistic categories over perceptual dimensions.
 For example, \cite{regier2007} showed that the partitioning of color categories reflects such a trade-off in a psychological space -- colors are partitioned into groups with members that are viewed as highly similar to one another yet distinct from other colors.
-A similar trade-off can be observed in phoneme perception...
-\nbcnote {I got nothing there...}
+A similar trade-off can be observed in phoneme categories.
+Different exemplars of the same phoneme must be similar to one another, while contrasting from other phonemes, such that a listener can infer the appropriate phoneme.
+This pattern has been found and modeled in the natural acoustics of American English vowels \citep{feldman2013,hillenbrand1995}.
+As linguistic categories must have been created at some point in human history, it is revealing that the constraints of emulating distributional structure across categories and having categories contrast from one another still bias human category generation today.
 
@@
 
-Although the focus of this paper has been to address the processes involved in creative generation using the empirical and quantitative toolkit of traditional categorization research, our findings and approach have relevant implications for research in creative cognition.
+Although the focus of this article has been to address the processes involved in creative generation using the empirical and quantitative toolkit of traditional categorization research, our findings and approach have relevant implications for research in creative cognition.
 A central focus of the creative cognition approach has been to explain acts of creativity in terms of the mental representations and processes that are commonly studied in cognitive psychology and cognitive science \citep{finke1992creative,smith1995creative}.
@@
 In addition to demonstrating the utility of formal modeling in the study creative cognition, the PACKER model more specifically offers an additional interpretation of some of the field's most central findings.
-For example, perhaps the most foundational principle from this literature concerns the limiting influence of prior knowledge: individuals create new categories composed of features from existing classes, and what is created can be influenced drastically through the introduction of cues or examples \citep{marsh1999inadvertent,smith1993constraining}.
+For example, perhaps the most foundational principle from this literature concerns the limiting influence of prior knowledge: Individuals create new categories composed of features from existing classes, and what is created can be influenced drastically through the introduction of cues or examples \citep{marsh1999inadvertent,smith1993constraining}.
 In this paper, we have identified another important aspect of the constraining influence of prior knowledge: What is generated cannot be the same as what is already known.
@@
 Most notably, in this paper we have not evaluated the model's ability to explain the classic finding that generated categories tend to share distributional commonalities with previously learned categories \citep[see][]{jern2013probabilistic,ward1994structured}.
-While we successfully replicated this effect in Experiment 1, we also found that its influence was limited in comparison to the fundamental constraints imposed by category contrast.
-Even within Experiment 1, we found systematic inconsistencies: by generating exemplars into unoccupied regions of the space, participants who learned an `XOR' category, composed of members that are widely distributed along both features and are positively correlated in space, tended to generate categories with an opposite (negative) correlation.
-More generally, PACKER's success over the hierarchical sampling model indicates that the emulation of distributional structure exerts only a limited influence in comparison to category contrast.
+While we successfully replicated this effect in Experiment 1, we also found that its influence was limited in comparison to the fundamental constraints imposed by  Even within Experiment 1, we found systematic inconsistencies: by generating exemplars into unoccupied regions of the space, participants who learned an `XOR' category, composed of members that are widely distributed along both features and are positively correlated in space, tended to generate categories with an opposite (negative) correlation.
+More generally, PACKER's success over the hierarchical sampling model indicates that the emulation of distributional structure exerts only a limited influence in comparison to category contrast in some situations.
 
-Nonetheless, these classic effects are a core element of the phenomenology of creative generation, and PACKER does not include any mechanisms explain them.
-Instead, through the development and evaluation of the PACKER model, we have sought to add new elements into such a phenomenology: specifically, the broad and strong influence of category contrast, and the interdependence between category location and distributional structure.
+Nonetheless, these classic effects are a core element of the phenomenology of creative generation, and PACKER does not include any mechanisms that explain them.
+Instead, through the development and evaluation of the PACKER model, we have sought to add new elements into such a phenomenology: The broad and strong influence of category contrast, and the interdependence between category location and distributional structure.
 It may be possible to combine the hierarchical sampling approach proposed by \cite{jern2013probabilistic} with PACKER's underlying claims to obtain a ``best of both worlds'' model, capable of explaining the role of contrast in creative generation, as well as the emulation of distributional structure.
-However, as noted in the introduction, the incorporation of category contrast is antithetical to the core principles of the hierarchical Bayesian approach.
-Thus, while such a model may provide an excellent fit to the data, its theoretical composition would be more or less of an amalgam, and thus it may not succeed in other areas of model evaluation (such as predicting and explaining novel phenomena).
+However, as noted in the introduction, the incorporation of category contrast is antithetical to the core principles of a traditional, semi-conjugate Bayesian approach.
+This suggests that category generation is a fundamentally different computational-level problem.
+Characterizing that problem and conducting a rational analysis is an important direction for future research.
+To that aim, we plan to explore the connection between exemplar modeling as an Importance Sampling approximation \citep{shi10exemplar}, and see what sort of computational-level problem PACKER approximates.
+Once formalized in probabilistic terms, it should also be straightforward to incorporate distributional factors into the model.
 
@@
 
-\nbcnote{I think importance sampling should fit in there pretty well.}
-
 \section{Conclusions}
@@
 In this paper, we have provided such an examination of category generation as it applies within an artificial categorization categorization experiment.
-Extending the literature on creative cognition, our experiments provide a detailed picture of the role of category contrast in generation: People seek to create concepts that are distinct from what they already know, and the nature what is created can be influenced by what \textit{doesn't} yet exist.
-Our simulations with traditional exemplar models, as well as a hierarchical sampling model, reinforce the claim that category contrast is of fundamental importance, and, more generally, demonstrate that popular explanatory approaches from basic research in cognitive science can offer a precise, quantitative account of a behavior as complex as that of creative generation.
+Extending the literature on creative cognition, our experiments provide a detailed picture of the role of category contrast in generation: People seek to create concepts that are distinct from what they already know, and the nature what is created can be influenced by what \textit{does not} yet exist.
+Our simulations with traditional exemplar models, as well as a hierarchical sampling model, provide strong support for the claim that category contrast is of fundamental importance, and, more generally, demonstrate that popular explanatory approaches from basic research in cognitive science can offer a precise, quantitative account of a behavior as complex as that of creative generation.
 

### 2017-05-01 Manuscripts/cog-psych/cogpsych.tex
nolan readthrough, fixed a couple of errors

@@
 In this paper, we study the generation of new categories using the computational and behavioral toolkit of traditional artificial category learning.
-Previous work in this domain has focused on how the statistical structure of known categories generalizes to generated categories, indifferent to contrast between the known and generated categories.
+Previous work in this domain has focused on how the statistical structure of known categories generalizes to generated categories, overlooking contrast between the known and generated categories.
 We report two experiments demonstrating that contrast between what is known and what is created is of fundamental importance in the creative process.
@@
 
-In Experiment 1 we sought to extend our analysis of the \cite{jern2013probabilistic} data by evaluating the influence of category contrast on creative generation, given qualitatively different types of prior knowledge.
+% In Experiment 1 we sought to extend our analysis of the \cite{jern2013probabilistic} data by evaluating the influence of category contrast on creative generation, given qualitatively different types of prior knowledge.
 We found strong evidence for effects of category contrast in each condition: Participants were more likely to generate stimuli that are more distant from (i.e., less similar to) members of a previously-learned category, and members of participant-generated categories tended to be more similar to one another than to members of previously-learned categories.
@@
 
-Thus, the core results of Experiment 1 indicate that generated categories systematically different from what is expected based on prior work.
+% Thus, the core results of Experiment 1 indicate that generated categories systematically differ from what can be expected based on prior work.
 The negative (or zero) correlations observed in the XOR condition suggests an interesting interaction between contrasting from a given category and emulating statistical properties.
@@
 
-However, our ability to detect differences using a standard $t$-test between the conditions is, in this case, diminished due to the non-normality of the data (Shapiro-Wilk normality test $W=0.77, p< 0.001$ and $W=0.85, p < 0.001$ for the Bottom condition).
+However, our ability to detect differences using a standard $t$-test between the conditions is, in this case, diminished due to the non-normality of the data (Shapiro-Wilk normality test $W=0.77, p< 0.001$ for the Middle condition and $W=0.85, p < 0.001$ for the Bottom condition).
 Figure \ref{fig:e2-yranges} depicts the Y-axis position of the exemplars generated within each participant's category.
@@
 We also present simulations from an implementation of the hierarchical sampling model proposed by \cite{jern2013probabilistic}, described in-depth in Appendix \ref{ap:hsampling-definition}.
-The comparison between the hierarchical sampling model and PACKER is meant to emphasize the necessity of contrast and demonstrate it cannot be explained by emulating of distributional structure: whereas PACKER is insensitive to the distributional structure of learned categories (relying only on within- and between-category similarity), the hierarchical sampling model generates categories exclusively on the basis of knowledge of how existing classes are distributed.
+The comparison between the hierarchical sampling model and PACKER is meant to emphasize the necessity of contrast and demonstrate it cannot be explained through the emulation of distributional structure: whereas PACKER is insensitive to the distributional structure of learned categories (relying only on within- and between-category similarity), the hierarchical sampling model generates categories exclusively on the basis of knowledge of how existing classes are distributed.
 Our approach in this section is to first broadly evaluate and compare the quality of each model's account to our entire dataset (Experiments 1 and 2 combined), then analyze the ability for each model to explain individual differences in each experiment, and lastly we more describe the strengths and weakness of each model's account of category generation.
@@
 
-For each possible stimulus, we compiled the range differences across all the categories it was generated into.
 However, because many stimuli were infrequently generated (such items near members of the `Alpha' category), we cannot simply compute the empirical average, as infrequently generated stimuli would be likely to show artificially strong differences.
@@
 Within this approach, the resulting aggregation is a trade-off between the number of generations and the strength of the range difference within each generated category.
-Infrequently generated stimuli, as well as those with both strongly positive and negative scores, are given neutral difference scores\footnote{Prior to plotting, data was also processed using a Gaussian filter with $\sigma = 0.8$}.
+Infrequently generated stimuli, as well as those with mixed positive and negative scores, are given neutral difference scores\footnote{Prior to plotting, data was also processed using a Gaussian filter with $\sigma = 0.8$}.
 
@@
 Most notably, in this paper we have not evaluated the model's ability to explain the classic finding that generated categories tend to share distributional commonalities with previously learned categories \citep[see][]{jern2013probabilistic,ward1994structured}.
-While we successfully replicated this effect in Experiment 1, we also found that its influence was limited in comparison to the fundamental constraints imposed by  Even within Experiment 1, we found systematic inconsistencies: by generating exemplars into unoccupied regions of the space, participants who learned an `XOR' category, composed of members that are widely distributed along both features and are positively correlated in space, tended to generate categories with an opposite (negative) correlation.
+While we successfully replicated this effect in Experiment 1, we also found that its influence was limited in comparison to the fundamental constraints imposed by  category contrast.
+Even within Experiment 1, we found systematic inconsistencies: by generating exemplars into unoccupied regions of the space, participants who learned an `XOR' category, composed of members that are widely distributed along both features and are positively correlated in space, tended to generate categories with an opposite (negative) correlation.
 More generally, PACKER's success over the hierarchical sampling model indicates that the emulation of distributional structure exerts only a limited influence in comparison to category contrast in some situations.

### 2017-06-02 Manuscripts/cog-psych/cogpsych.tex
outstanding, magnificent changes to the intro. best you'll ever see

@@
 \setlength\parindent{0.5in}
+
+The creation of new ideas is one of the most fascinating, important, and difficult human capabilities to study.
+Although we tend to focus on the most salient products of the processes underlying creative generation (e.g., scientific breakthroughs), every person is likely to have generated many novel concepts in ordinary life.
+Questions concerning the generation of novel concepts are typically addressed within research on creativity, which utilizes a diverse collection of empirical and theoretical approaches \citep[see][]{kaufman2010handbook}.
+However, due to the difficult nature of studying creative generation, this work infrequently employs computational methods to evaluate proposed processes and representations.
+
+In this paper, we approach the problem of creativity in terms of
+
+In particular, generating a member of a novel category can be considered as a `special case' application of existing category knowledge \cite{kemp2014taxonomy}: whereas research in categorization typically addresses how knowledge is used to classify items into known categories \citep[e.g.,][]{shepard1961learning,nosofsky1994comparing} or infer unobserved features of known category members \citep[see][]{markman2003category}.
+The generation of members of new categorires can be considered as a special case of such an inference
+
+Classification consist of selecting a category label given observed features, and feature inference consists of inferring the value of a missing feature given a set of observed features and a category label, generation of new category members requires individuals to infer all feature values given a novel category label.
+
+Many aspects of the creative process, however, are not altogether different from cognitive processes typically studied in laboratories.
+In particular, creating members of a new category can be considered a form of inference using one's knowledge of categories.
+
+Computational methods, nonetheless, have proven fruitful in advancing research in a wide variety
+
+% Question: How do people create new concepts?
+%     + [Maybe an example of this here, maybe not.]
+
+% Typically this question is approached from the lit on creativity, which uses a wide variety of approaches such as [X], [Y], [etc]
+
+% While traditional approaches have made great advances, they make it difficult to formalize the creative processes within a computational modeling framework.
+%     + [Maybe a line about how computational approaches have been fruitful in other fields, or how we'd ideally like to create machines with the creative capacity of humans.]
+
+% Many aspects of the creative process, however, are not altogether different from cognitive processes typically studied in laboratories.
+In particular, creating members of a new category can be considered a form of inference using one's knowledge of categories.
+%     + [Show some depiction of different types of inferences using category knowledge (classification, inference, generation), maybe that table.]
+%     + [Cite that Kemp taxonomy paper, Ken's PLM chapter.]
+
+% Under this framing, the generation of categories can be studied using toolkit of categorization.
+
+% However, this application of category knowledge has been broadly under-explored within the literature on categorization.
+%     + [Progress in this area not only informs our understanding of creativity, but also categorization]
+
+% Foundational work on the topic [ward intro] ...
+%     + [From here, we can probably append the intro as is, starting from the creative cognition section.]
 

### 2017-06-02 Manuscripts/cog-psych/cogpsych.tex
refining, PERFECTING, changes to the intro

@@
 Although we tend to focus on the most salient products of the processes underlying creative generation (e.g., scientific breakthroughs), every person is likely to have generated many novel concepts in ordinary life.
+
 Questions concerning the generation of novel concepts are typically addressed within research on creativity, which utilizes a diverse collection of empirical and theoretical approaches \citep[see][]{kaufman2010handbook}.
+
 However, due to the difficult nature of studying creative generation, this work infrequently employs computational methods to evaluate proposed processes and representations.
 
-In this paper, we approach the problem of creativity in terms of
-
-In particular, generating a member of a novel category can be considered as a `special case' application of existing category knowledge \cite{kemp2014taxonomy}: whereas research in categorization typically addresses how knowledge is used to classify items into known categories \citep[e.g.,][]{shepard1961learning,nosofsky1994comparing} or infer unobserved features of known category members \citep[see][]{markman2003category}.
-The generation of members of new categorires can be considered as a special case of such an inference
+In this paper, we assume a different approach to studying the generation of novel concepts.
+
+In particular, generating a member of a novel class can be considered a `special case' application of existing category knowledge \citep{kemp2014taxonomy,kurtz2015human}.
+Research in categorization typically focuses on classification \citep[predicting the category given feature information, see][]{shepard1961learning,nosofsky1994comparing} and inference \citep[predicting unobserved features given observed features and a category label, see][]{markman2003category}.
+Under this framing, the generation of new category members consists of inferring the {\em all} features given a novel category label.
+Thus, we can make progress in developing formal theories of creative generation by extending theories of categorization to the task of category generation.
+
+ whereas research in categorization typically addresses how knowledge is used to classify items into known categories  or infer unobserved features of known category members \citep[see][]{markman2003category}.
+The generation of members of new categories can be considered as a special case of such an inference
 

### 2017-06-02 Modules/tests/test_script.py

@@ -12,6 +12,9 @@ execfile('Imports.py')
 from Modules.Classes import CopyTweak, Packer, ConjugateJK13
 import Modules.Funcs as funcs
 
+print dir(funcs)
+lll
+
 # [[72 73 74 75 76 77 78 79 80]
 #  [63 64 65 66 67 68 69 70 71]
 #  [54 55 56 57 58 59 60 61 62]

### 2017-06-05 Manuscripts/cog-psych/cogpsych.tex
more intro updates

@@
 The creation of new ideas is one of the most fascinating, important, and difficult human capabilities to study.
-Although we tend to focus on the most salient products of the processes underlying creative generation (e.g., scientific breakthroughs), every person is likely to have generated many novel concepts in ordinary life.
-
-Questions concerning the generation of novel concepts are typically addressed within research on creativity, which utilizes a diverse collection of empirical and theoretical approaches \citep[see][]{kaufman2010handbook}.
-
-However, due to the difficult nature of studying creative generation, this work infrequently employs computational methods to evaluate proposed processes and representations.
-
-In this paper, we assume a different approach to studying the generation of novel concepts.
-
+We tend to focus on the most salient products of the processes underlying conceptual generation (e.g., scientific breakthroughs), but every person is likely to have generated many novel concepts in ordinary life -- indeed, the generation of novel sentences, works of art, and so on may also be considered as products of our capacity to create new concepts.
+Although this capacity is of great interest, it is also of great complexity and can be difficult to study in a typical laboratory setting.
+By consequence, we do not have a strong understanding of the representations and processes involved in the generation of new concepts.
+
+Questions concerning the generation of novel concepts are typically addressed within research on creativity, a tremendously diverse field employing a multitude of experimental and theoretical approaches \citep[see][]{kaufman2010handbook}.
+The processes underlying conceptual generation are commonly addressed using a creative cognition approach \citep{finke1992creative, smith1995creative}, in which the products of creative acts (such as drawings) are analyzed to obtain insights into how they might have been created.
+While these approaches have made great advances in our understanding of creativity, the focus on creative {\em products} makes difficult the task of formalizing proposed processes and representations within a computational modeling framework, and so the field does not benefit from the application of computational models to evaluate theories.
+
+\nbcnote{Maybe a bit about why that would be good?}
+
+However, many of these proposed processes are not altogether different from cognitive processes typically studied in laboratories.
 In particular, generating a member of a novel class can be considered a `special case' application of existing category knowledge \citep{kemp2014taxonomy,kurtz2015human}.
@@
 Thus, we can make progress in developing formal theories of creative generation by extending theories of categorization to the task of category generation.
-
- whereas research in categorization typically addresses how knowledge is used to classify items into known categories  or infer unobserved features of known category members \citep[see][]{markman2003category}.
-The generation of members of new categories can be considered as a special case of such an inference
-
-Classification consist of selecting a category label given observed features, and feature inference consists of inferring the value of a missing feature given a set of observed features and a category label, generation of new category members requires individuals to infer all feature values given a novel category label.
-
-Many aspects of the creative process, however, are not altogether different from cognitive processes typically studied in laboratories.
-In particular, creating members of a new category can be considered a form of inference using one's knowledge of categories.
-
-Computational methods, nonetheless, have proven fruitful in advancing research in a wide variety
 

### 2017-06-05 Manuscripts/cog-psych/cogpsych.tex
drafted intro addition

@@
 The creation of new ideas is one of the most fascinating, important, and difficult human capabilities to study.
-We tend to focus on the most salient products of the processes underlying conceptual generation (e.g., scientific breakthroughs), but every person is likely to have generated many novel concepts in ordinary life -- indeed, the generation of novel sentences, works of art, and so on may also be considered as products of our capacity to create new concepts.
-Although this capacity is of great interest, it is also of great complexity and can be difficult to study in a typical laboratory setting.
+We tend to focus on the most salient products conceptual generation (e.g., scientific breakthroughs), but every person is likely to have generated many novel concepts in ordinary life.
+Indeed, the generation of novel sentences, works of art, and so on may also be considered products of our capacity to create new concepts.
+However, while this capacity is of great interest, it is also of great complexity and can be difficult to study in a typical laboratory setting.
 By consequence, we do not have a strong understanding of the representations and processes involved in the generation of new concepts.
@@
 Questions concerning the generation of novel concepts are typically addressed within research on creativity, a tremendously diverse field employing a multitude of experimental and theoretical approaches \citep[see][]{kaufman2010handbook}.
-The processes underlying conceptual generation are commonly addressed using a creative cognition approach \citep{finke1992creative, smith1995creative}, in which the products of creative acts (such as drawings) are analyzed to obtain insights into how they might have been created.
-While these approaches have made great advances in our understanding of creativity, the focus on creative {\em products} makes difficult the task of formalizing proposed processes and representations within a computational modeling framework, and so the field does not benefit from the application of computational models to evaluate theories.
+The processes underlying conceptual generation are commonly investigated using a creative cognition approach \citep{finke1992creative, smith1995creative}, in which the products of creative acts (such as drawings) are analyzed to obtain insights into how they might have been created.
+These approaches have made great advances in our understanding of creativity, but the focus on creative {\em products} makes difficult the task of formalizing proposed processes and representations within a computational modeling framework, and so the field does not benefit from the application of computational models to evaluate theories.
 
@@
 
-However, many of these proposed processes are not altogether different from cognitive processes typically studied in laboratories.
+The act of generating a new concept is, however, not altogether different from the types of behaviors typically studied in cognitive psychology laboratories.
 In particular, generating a member of a novel class can be considered a `special case' application of existing category knowledge \citep{kemp2014taxonomy,kurtz2015human}.
-Research in categorization typically focuses on classification \citep[predicting the category given feature information, see][]{shepard1961learning,nosofsky1994comparing} and inference \citep[predicting unobserved features given observed features and a category label, see][]{markman2003category}.
-Under this framing, the generation of new category members consists of inferring the {\em all} features given a novel category label.
-Thus, we can make progress in developing formal theories of creative generation by extending theories of categorization to the task of category generation.
-
-% Question: How do people create new concepts?
-%     + [Maybe an example of this here, maybe not.]
-
-% Typically this question is approached from the lit on creativity, which uses a wide variety of approaches such as [X], [Y], [etc]
-
-% While traditional approaches have made great advances, they make it difficult to formalize the creative processes within a computational modeling framework.
-%     + [Maybe a line about how computational approaches have been fruitful in other fields, or how we'd ideally like to create machines with the creative capacity of humans.]
-
-% Many aspects of the creative process, however, are not altogether different from cognitive processes typically studied in laboratories.
-In particular, creating members of a new category can be considered a form of inference using one's knowledge of categories.
-%     + [Show some depiction of different types of inferences using category knowledge (classification, inference, generation), maybe that table.]
-%     + [Cite that Kemp taxonomy paper, Ken's PLM chapter.]
-
-% Under this framing, the generation of categories can be studied using toolkit of categorization.
-
-% However, this application of category knowledge has been broadly under-explored within the literature on categorization.
-%     + [Progress in this area not only informs our understanding of creativity, but also categorization]
-
-% Foundational work on the topic [ward intro] ...
-%     + [From here, we can probably append the intro as is, starting from the creative cognition section.]
-
-Creativity, innovation, imagination, and the creation of new ideas are among the most fascinating, important and difficult human capabilities to study.
+Research in categorization typically focuses on classification \citep[predicting an object's category given its features;][]{shepard1961learning,kurtz2013human} and inference \citep[predicting an object's unobserved features given observed features and a category label, see][]{markman2003category}.
+The generation of members of a new category consists of inferring the {\em all} features given a {\em novel} category label.
+Thus, we can make progress formalizing the processes involved in creative generation by extending theories of categorization to the special case.
+
+\nbcnote{Note that this problem has been under-studied in the categorization lit?}
+
+% % OLD INTRO:
+% Creativity, innovation, imagination, and the creation of new ideas are among the most fascinating, important and difficult human capabilities to study.
 They are fascinating and important capabilities because they have produced some of the greatest scientific breakthroughs that revolutionized the world.
@@
 
-Towards the aim of providing a formal account of the cognitive mechanisms involved in creativity and innovation, we explore a form of creative cognition that can be studied using standard behavioral methodology and is amenable to formal modeling: category generation.
+% Towards the aim of providing a formal account of the cognitive mechanisms involved in creativity and innovation, we explore a form of creative cognition that can be studied using standard behavioral methodology and is amenable to formal modeling: category generation.
 In a category generation task, participants are given some background knowledge about a domain (a cover story and/or learning one or more categories in the domain) and then are asked to generate another category in the domain.
@@
 This fundamental constraint, ``being different'', or contrasting from other categories in the relevant domain, is the focus of our work.
-Beyond the assumption that participant-generated categories are indeed novel, this constraint has been overlooked in previous research: To our knowledge, there has not been any systematic investigation addressing how generated categories {\em differ} from what is already known.
+Beyond the assumption that newly generated concepts are indeed novel, this constraint has been overlooked in previous research: To our knowledge, there has not been any systematic investigation addressing how generated categories {\em differ} from what is already known.
 Although the idea of category contrast is ubiquitous throughout the categorization literature, and extends to a variety of other fields \citep[e.g., color;][]{regier2007}, the idea that a new category should be ``different'' is vague, as there are many ways it could be different from a previously observed category.

### 2017-06-08 Manuscripts/cog-psych/cogpsych.tex
framing applied to remainder of document

@@
 
-\nbcnote{More strongly link this to creativity: what do these findings indicate?}
-\jlanote{ok with me now.}
-
 The broader set of observations made by Ward and colleagues provide a great deal of insight into the nature of creative generation.
@@
 
-\subsection{Something Different: A Role For Contrast}
+\section{Something Different: A Role For Contrast}
 
@@
 Thus, effects of category contrast should be observable within the experimental results of these studies.
-To provide a test of the influence of category contrast in creative generation, we conducted a novel analysis of Experiment 3 from \cite{jern2013probabilistic}.
-
-\nbcnote{Does it feel strange to focus on Experiment 3 specifically? Do you think wee need to explain why this is the `right' experiment for our question?}
-
-\jlanote{I'd add a footnote explaining why we focus on this one only}
+To provide a test of the influence of category contrast in creative generation, we conducted a novel analysis of Experiment 3 from \cite{jern2013probabilistic}\footnote{Although \cite{jern2013probabilistic} reported four experiments, we focus on Experiment 3 because their first two experiments tested generation of items into known categories, and their fourth experiment was identical to Experiment 3 but with a restricted generation space.}.
 
@@
 To do so, we binned all possible hues into one of eight uniformly-spaced color groups: \{{\em Red}: $0-0.063$, $0.938-1$, {\em Yellow}: $0.063-0.188$, {\em Yellow-Green}: $0.188-0.313$, {\em Green-Teal}: $0.313-0.438$, {\em Teal}: $0.438-0.563$, {\em Teal-Blue}: $0.563-0.688$, {\em Purple}: $0.688-0.813$, {\em Pink}: $0.813-0.938$\}.
-In the \cite{jern2013probabilistic} experiment, the hue of each experimenter-defined category was selected from one of six possible values, each of which falls into one of the color groups above.
-\jlanote{Note the two color bins we made that don't contain any of their values?}
+In the \cite{jern2013probabilistic} experiment, the hue of each experimenter-defined category was selected from one of six possible values, each of which falls into one of the color groups above (two color groups do not contain a possible hue for the experimenter-defined categories).
 By categorizing the participant-generated crystals likewise, we can obtain a broad measure of category contrast by determining the proportion of participant-generated crystals that fall into the same groups as the experimenter-defined categories: If contrast influences the hues of the generated categories, we should observe minimal overlap between in the color groupings.
@@
 That is, if hue selection is not systematic, the probability that any given example will lie in the same color group as an experimenter-defined category is given by a Binomial distribution with $p = 2/8=1/4$, as there were two experimenter-defined categories and eight possible color groups.
-Chi-square goodness-of-fit tests reveal that the observed distribution in each condition is highly inconsistent with the hues being chosen at random (all $\chi^2>200$, $p<0.001$): Participants tended to generate items that were perceptually distinct from the categories they had learned, and were less likely to generate hues possessed by members of the experimenter-defined categories.
-\jlanote{need degrees of freedom for $\chi^2$ tests}
+Chi-square goodness-of-fit tests reveal that the observed distribution in each condition is highly inconsistent with the hues being chosen at random, all $\chi^2(6,N=22)>200$, $p<0.001$.
+Participants tended to generate items that were perceptually distinct from the categories they had learned, and were less likely to generate hues possessed by members of the experimenter-defined categories.
+
 Re-analyzing the results of \cite{jern2013probabilistic} provides preliminary support that contrast plays a role in category generation.
@@
 In the sections below, we present two experiments demonstrating systematic effects of category contrast on creative generation inspired by factors influencing how PACKER generates new categories.
-Our experiments are based on \cite{jern2013probabilistic}'s paradigm: participants are first exposed to a single, experimenter-defined category, and are then asked to generate members of a new category.
+Our experiments are based on \cite{jern2013probabilistic}'s paradigm, which is a straightforward translation of the traditional artificial classification paradigm to the task of category generation: participants are first exposed to a single, experimenter-defined category, and are then asked to generate members of a new category.
 We then report formal analyses comparing PACKER's account of our results to that of the hierarchical sampling and copy-and-tweak models developed by \cite{jern2013probabilistic}.
@@
 To extend the evidence provided by the \cite{jern2013probabilistic} data, we tested the effects of category contrast after learning one category from a set of qualitatively distinct category structures, as shown in Figure \ref{fig:e1-conditions}.
-
-\nbcnote{can you think of other studies with size and color of squares as features?}
 
@@
 
-In this way, one aspect of the analyses of the experimental results is a conceptual replication of the classic finding that generated categories tend to share distributional properties with known categories in the domain \citep[see][]{jern2013probabilistic,ward1994structured}.
+In this way, our experimental results provide a conceptual replication of the classic finding that generated categories tend to share distributional properties with known categories in the domain \citep[see][]{jern2013probabilistic,ward1994structured}.
 From these results, we can predict that, in each condition, participants should generate categories that are distributionally similar to the experimenter-defined category: In the Cluster condition, generated categories should be tightly clustered.
@@
 Thus, if contrast plays a role, exemplars in the generated categories of participants in the XOR condition may not be positively correlated -- they may be negatively correlated instead.
-In this case, contrast and statistical regularities would interact, which would be inconsistent with the hierarchical sampling approach of creative cognition \citep{jern2013probabilistic}
+In this case, contrast and statistical regularities would interact, which would be inconsistent with the leading approaches in conceptual generation \citep{jern2013probabilistic}
 
@@
 183 participants were recruited from Amazon Mechanical Turk.
-Participants were randomly assigned to one condition: 64 participants were assigned to the Cluster condition, 61 were assigned to the Row condition, and 58 were assigned to the XOR condition.
+Each participant was randomly assigned to one condition: 64 participants were assigned to the Cluster condition, 61 were assigned to the Row condition, and 58 were assigned to the XOR condition (sample sizes differ due to random assignment).
 Stimuli were squares varying in color (grayscale 9.8\%--90.2\%) and side length (3.0--5.8cm), see Figure \ref{fig:e1-conditions}.
 The assignment of perceptual features (color, size) to axes of the domain space (x, y), as well as the direction of variation along each axis (e.g., $dark \rightarrow light$ or $light \rightarrow dark$) was counterbalanced across participants.
-\jlanote{we should probably mention the discrepancy between the number of participants per condition and explain why it isn't anything to worry about.}
 
@@
 
-% In Experiment 1 we sought to extend our analysis of the \cite{jern2013probabilistic} data by evaluating the influence of category contrast on creative generation, given qualitatively different types of prior knowledge.
+In Experiment 1 we sought to extend our analysis of the \cite{jern2013probabilistic} data by evaluating the influence of category contrast on creative generation, given qualitatively different types of prior knowledge.
 We found strong evidence for effects of category contrast in each condition: Participants were more likely to generate stimuli that are more distant from (i.e., less similar to) members of a previously-learned category, and members of participant-generated categories tended to be more similar to one another than to members of previously-learned categories.
@@
 
-% Thus, the core results of Experiment 1 indicate that generated categories systematically differ from what can be expected based on prior work.
+Thus, the core results of Experiment 1 indicate that generated categories systematically differ from what can be expected based on prior work.
 The negative (or zero) correlations observed in the XOR condition suggests an interesting interaction between contrasting from a given category and emulating statistical properties.
@@
 To test whether category contrast influences the distributional structure of generated categories, we sought to identify conditions in which differences in the distributional structure of generated categories cannot be explained by the distributional structure of the experimenter-defined category.
-We created two new category types (depicted in Figure ~~\ref{fig:e2-conditions}) that possess an identical distributional structure (both are tight clusters of examples with no correlation between features), and only differ in their Y-axis position: the `Bottom' category lies in the bottom-center of the space, and the `Middle' category lies in the center.
+We created two new category types (depicted in Figure \ref{fig:e2-conditions}) that possess an identical distributional structure (both are tight clusters of examples with no correlation between features), and only differ in their Y-axis position: the `Bottom' category lies in the bottom-center of the space, and the `Middle' category lies in the center.
 The distributional equality of these conditions is key to the design of the experiment: If the structure of previously learned categories were the only influence on the structure of generated categories, we should observe no difference between these two conditions with respect to distributional structure.
@@
 
-To explore the aggregate distributional structure of the results, we computed the range of exemplars along each axis (X, Y), as well as the correlation between features.
+To explore the distributional structure of the generated categories, we computed the range of exemplars along each axis (X, Y), as well as the correlation between features.
 These data, shown in Figure \ref{fig:e2-statsboxes}, again demonstrate the degree of individual differences observed in our study.
@@
 We also present simulations from an implementation of the hierarchical sampling model proposed by \cite{jern2013probabilistic}, described in-depth in Appendix \ref{ap:hsampling-definition}.
-The comparison between the hierarchical sampling model and PACKER is meant to emphasize the necessity of contrast and demonstrate it cannot be explained through the emulation of distributional structure: whereas PACKER is insensitive to the distributional structure of learned categories (relying only on within- and between-category similarity), the hierarchical sampling model generates categories exclusively on the basis of knowledge of how existing classes are distributed.
-Our approach in this section is to first broadly evaluate and compare the quality of each model's account to our entire dataset (Experiments 1 and 2 combined), then analyze the ability for each model to explain individual differences in each experiment, and lastly we more describe the strengths and weakness of each model's account of category generation.
+The comparison between the hierarchical sampling model and PACKER is meant to emphasize the necessity of contrast and demonstrate that generation cannot be explained entirely through the emulation of distributional structure: whereas PACKER is insensitive to the distributional structure of learned categories (relying only on within- and between-category similarity), the hierarchical sampling model generates categories exclusively on the basis of knowledge of how existing classes are distributed.
+
+Our approach in this section is to first broadly evaluate and compare the quality of each model's account to our entire dataset (Experiments 1 and 2 combined), then analyze the ability for each model to explain individual differences in each experiment, and lastly we describe the strengths and weakness of each model's account of category generation.
 
@@
 Due to the uneven number of fitted parameters among the models, we compare the model fits using the Akaike Information Criterion \citep[AIC;][]{akaike1974new}, where smaller values correspond to better fits (discounted by the number of parameters).
+Similar results were obtained with alternative model comparison metrics (e.g., BIC, \citealp{schwarz1978estimating}; $AIC_C$, \citealp{hurvich1989regression})
+
 Table \ref{table:global-model-fits} contains the AIC values of each model's best fit, as well as the corresponding log-likelihood ($L$) and the best-fitting parameter values.
@@
 We conducted an identical set of simulations as above, excluding this trial (leaving 915 responses in the dataset): Again, PACKER ($L = -3390$, $AIC = 6786$) achieved better fits than the copy-and-tweak ($L = -3579$, $AIC = 7162$) and hierarchical sampling ($L=-3612$, $AIC = 7232$) models.
+
 Finally, because copy-and-tweak is nested within PACKER, we can use a likelihood ratio test to compare the two models.
-PACKER explains the aggregate data significantly better than copy-and-tweak ($\chi^2(1) = 748, p < 0.001$ for all data and $\chi^2(1) = 378, p < 0.001$ excluding the first example), providing further evidence that category generation is better explained when contrast is considered.
-
-\nbcnote{report something about robustness to model comparison metric (BIC and $AIC_c$).}
-\jlanote{please -- i believe that they are derivable from log-likelihood, number of parameters, and number of data points}
+PACKER explains the aggregate data significantly better than copy-and-tweak ($\chi^2(1) = 749, p < 0.001$ for all data and $\chi^2(1) = 377, p < 0.001$ excluding the first example), providing further evidence that category generation is better explained when contrast is considered.
 
@@
 
-As in PACKER, changes to the weighting of the dimensions can also be used to produce row- and column-like categories in the copy-and-tweak and hierarchical sampling models.
-Indeed, as copy-and-tweak is simply a special case of the PACKER model, the attention weights operate exactly as described above.
-Within the hierarchical sampling model, the covariance matrix specifying the prior domain distribution, $\Sigma_0$, can be used to similar effect.
+As in PACKER, changes in the parameter settings of the copy-and-tweak model can also be used to produce different patterns of generation.
+Indeed, as copy-and-tweak is simply a special case of the PACKER model, the attention weights operate exactly as described above to produce row- and column-like categories.
+However, because the model is not influenced by category contrast, it is biased toward generating tightly clustered categories, as new items are always most likely to be generated near known examples of the target category.
+Thus, the lack of a contrast mechanism prevents the model from explaining why some individuals widely distribute their categories to the corners of the space.
+
+Similar to the copy-and-tweak model, the hierarchical sampling model possesses no mechanism to account for category contrast, and the model is most likely to generate new items that are similar to known examples of the target category.
+Although this precludes an account of why individuals might assume a cluster versus corners profile, the covariance matrix specifying the model's prior domain distribution, $\Sigma_0$, can be used to explain the generation of row-like and column-like categories.
 This covariance matrix specifies the amount of variance assumed along each dimension (as well as the dimensional correlations) across the domain of categories.
@@
 
-However, a key limitation of both of the copy-and-tweak and hierarchical sampling models is the lack of interaction between the distributional structure of a category (in this case, a row-versus-column orientation) and the location of that category within the domain.
-While these models are {\em capable} of generating row and column categories, there is no mechanism in place to ensure generated categories will be distinct from what is already known.
-In the next subsection, we explore the interdependence between distributional structure and location in creative generation.
+As noted above, while the copy-and-tweak and hierarchical sampling models possess mechanisms to explain row- and column-like categories, they cannot easily explain why some individuals widely distribute their generated categories into disparate corners of the space.
+This, however, reveals a more general limitation: According to the copy-and-tweak and hierarchical sampling models, the distributional structure of generated categories is {\em independent} of their location within the domain.
+For example, although the copy-and-tweak or hierarchical sampling models can be parameterized to generate row- or column-like categories, there is no mechanism in place to ensure that what is generated will be distinct from what is already known.
+In the next subsection, we explore this prediction through an analysis of the interdependence between distributional structure and location in creative generation.
 
@@
 Thus, a different transformation is appropriate for incorporating the weights computed in Equation \ref{eq:range-weight}.
-For the hierarchical sampling model, we computed the dimensional variances according to: $\lambda (1-w_k)d$, where $\lambda$ is a free parameter specifying the overall assumed variance of the domain, and $d$ is the number of dimensions.
+For the hierarchical sampling model, we computed the dimensional variances according to: $\lambda (1-w_k)d$, where $\lambda$ is a free parameter specifying the overall assumed variance of the domain, and $d$ is the number of dimensions\footnote{This calculation applies only in two-dimensional domains, where $w_2 = 1-w_1$.}.
 Under this approach, evenly distributed weights correspond to an assumed variance of $\lambda$.
@@
 
-\nbcnote{This works for 2D, but not 3D and beyond (as $1-w_k$ would not produce a valid weighting). Maybe we should make a note of that, or design something more general.}
-\jlanote{Isn't it just $1-\sum_{j\neq k}{w_j}$?}
-
 Each model was used to simulate each participant's generation independently, with the importance of each dimension set according to the participant's generated category.
 The other free parameters within each model were set as in Table \ref{table:global-model-fits}.
-Every participant's generation was simulated \textbf{NUMBER} of times, resulting in \textbf{NUMBER * PARTICIPANTS} categories generated by each model.
+Every participant's generation was simulated 2,000 times; multiplied against 305 total participants, each model generated 610,000 categories in total.
 For comparison with our behavioral results, we then computed the range difference gradient identically as with the behavioral data.
@@
 
-The creative generation of novel concepts is an intriguing yet understudied topic in cognitive science.
+The creative generation of concepts is an greatly intriguing topic, but it is also of great complexity and can be difficult to study in a traditional laboratory setting.
+Creative generation is most typically studied using a creative cognition approach \citep{finke1992creative,smith1995creative}, where creative products \citep[such as drawings of alien plants and animals;][]{ward1994structured} are analyzed to obtain insight into the processes and representations involved.
+While this approach has yielded a great deal of information about the nature of creative generation, the commonly employed response modalities (e.g., hand-drawn pictures) made difficult the development of computational models to formally evaluate the proposed representations and processes.
+However, the generation of novel concepts is not altogether different from the types of behaviors studied in cognitive psychology: In particular, it can be viewed as an form of inference using one's knowledge about existing categories \citep{kemp2014taxonomy,kurtz2015human}.
+Taking advantage of this insight, in this paper we investigated the generation of new categories using the empirical and computational toolkit from the field of categorization.
+
 While the bulk of prior research on the topic has focused on the classic finding that generated concepts tend to be distributionally similar to known concepts, there has been little work addressing the role of contrast in creative generation: How is it that people are able to create something {\em different} from what is already known?
@@
 Beyond its role in creative generation, category contrast is also of fundamental importance in categorization more broadly.
-All other factors held constant, new categories are easier to learn if they are dissimilar to members of other categories, and knowledge of highly distinct categories is applied more accurately than that of ill-defined categories\footnote{\nbcnote{This is an uncontroversial claim but we should have some cites.
-I just can't think of a super appropriate one right now...}}.
-Likewise, the act of forming category representations affects similarity judgments about category members and nonmembers, with category members being viewed as more similar to one another than members of other categories \citep{goldstone1994influences,goldstone2001altering} .
-
-%discussion of color categories and how they evolve. how across cultures, they seem to maximize within-category similarity and minimize between-category simialrity
+All other factors held constant, new categories are easier to learn if they are dissimilar to members of other categories, and knowledge of highly distinct categories is applied more accurately than that of ill-defined categories \citep{ashby1994categorization,imai1965discriminability}.
+Likewise, basic-level categories \citep{rosch1976basic} are thought to be abstracted in order to maximize within-category similarity while minimizing between-category similarity.
+Finally, the act of forming category representations affects similarity judgments about category members and nonmembers, with category members being viewed as more similar to one another than members of other categories \citep{goldstone1994influences,goldstone2001altering} .
 
@@
 
+\nbcnote{Something about analogy (similar relationally but different feature-ally); generated categories are alike in distributional structure but differently located in domain}
+
 The above passages serve to demonstrate that the dual forces of within-class similarity and between-class contrast influence cognitive function in a wide variety of domains.
@@
 \subsection{Limitations and Future Directions}
-\jlanote{this is where the importance sampling connection will come in and will involve specifying the computational problem precisely.
-that hasn't been done yet for category generation and then perhaps to creativity more broadly.}
 
@@
 
-Finally, although an overall goal of this work is to advance the study of creative thought by explaining generation through the empirical and quantitative approaches from research in categorization, these approaches are not easily aligned with many of the myriad alternative theoretical viewpoints in the study of creativity \citep[for a review see][]{kozbelt2010theories}, such as those based on free association \citep{mednick1962associative} and conceptual combination \citep{estes2002emergence,murphy1988comprehending}.
+Finally, although an overall goal of this work is to advance the study of creative thought by explaining generation through the empirical and quantitative approaches from research in categorization, these approaches are not easily aligned with many of the alternative theoretical viewpoints in the broader study of creativity \citep[for a review see][]{kozbelt2010theories}, such as those based on free association \citep{mednick1962associative} and conceptual combination \citep{estes2002emergence,murphy1988comprehending}.
 Instead, we reduced our focus by studying a highly complex behavior (creative generation) as it applies within a well-established domain (artificial category learning).
-Importantly, however, our findings do provide a clue about a more general, computational-level interpretation of creative generation.
-That is, in this paper we have made clear that contrast is a key constraint, as it is required in order to successful create something new.
-However, it cannot be the only constraint: It is not enough for a generated category to be distinct, as categories generated purely on the basis of contrast would not be distributionally coherent.
-Thus, a more general interpretation of creative generation would need to concern both the distinctness and coherence of what is created.
 
@@
 In this paper, we have provided such an examination of category generation as it applies within an artificial categorization categorization experiment.
-Extending the literature on creative cognition, our experiments provide a detailed picture of the role of category contrast in generation: People seek to create concepts that are distinct from what they already know, and the nature what is created can be influenced by what \textit{does not} yet exist.
+Extending the literature on creative cognition, our experiments provide a detailed picture of the role of category contrast in generation: People seek to create concepts that are distinct from what they already know, and the nature of what is created can be influenced by what does not yet exist.
 Our simulations with traditional exemplar models, as well as a hierarchical sampling model, provide strong support for the claim that category contrast is of fundamental importance, and, more generally, demonstrate that popular explanatory approaches from basic research in cognitive science can offer a precise, quantitative account of a behavior as complex as that of creative generation.

### 2017-06-08 Experiments/cogpsych-code/model.comparison.py

@@ -1,67 +1,0 @@
-import matplotlib.pyplot as plt
-import numpy as np
-from itertools import product
-import pandas as pd
-
-
-execfile('Imports.py')
-from Modules.Classes import CopyTweak, Packer, ConjugateJK13
-import Modules.Funcs as funcs
-
-pd.set_option('precision', 3)
-np.set_printoptions(precision = 3)
-
- # 72 73 74 75 76 77 78 79 80
- # 63 64 65 66 67 68 69 70 71
- # 54 55 56 57 58 59 60 61 62
- # 45 46 47 48 49 50 51 52 53
- # 36 37 38 39 40 41 42 43 44
- # 27 28 29 30 31 32 33 34 35
- # 18 19 20 21 22 23 24 25 26
- #  9 10 11 12 13 14 15 16 17
- #  0  1  2  3  4  5  6  7  8
-vals = np.linspace(-1, 1, 9).tolist()
-stimuli = np.fliplr(np.array(list(product(vals, vals))))
-
-
-alphas =  [30, 32, 48, 50]
-alphas =  [12, 30, 14, 32]
-betas = [0, 45]
-
-models = [ # these values copied on Jan 29!
-    [CopyTweak, dict(
-    specificity = 4.67536899146,
-    tolerance = 0.895345369763,
-    determinism = 5.60105216678,
-        )],
-    [Packer, dict(
-    specificity = 0.565028848775,
-    between = -4.81445541313,
-    within = 4.2500267818,
-    determinism = 0.731417901569,
-        )],
-    [ConjugateJK13, dict(
-    category_mean_bias = 1e-10,
-    category_variance_bias = 1.00000753396,
-    domain_variance_bias = 1.21974609442,
-    determinism = 7.14705600068,
-        )],
-]
-
-
-f, ax = plt.subplots(1, 3, figsize = (6.5, 2))
-pltnum = 0
-for FUN, params in models:
-    A = stimuli[alphas,:]
-    B = stimuli[betas,:]
-    
-    M = FUN([A, B], params) 
-    ps = M.get_generation_ps(stimuli, 1)
-
-    g = funcs.gradientroll(ps,'roll')[:,:,0]
-    h = ax[pltnum]
-    funcs.plotgradient(h, g, A, B, clim = (0, 0.15))
-    h.set_title(FUN.model, fontsize = 12)
-    pltnum += 1
-
-f.savefig('model.comparison.png', bbox_inches='tight', transparent=False)

### 2017-06-08 Manuscripts/cog-psych/cogpsych.tex
added alternative models bit, comparison of task to ward

@@
 The processes underlying conceptual generation are commonly investigated using a creative cognition approach \citep{finke1992creative, smith1995creative}, in which the products of creative acts (such as drawings) are analyzed to obtain insights into how they might have been created.
-These approaches have made great advances in our understanding of creativity, but the focus on creative {\em products} makes difficult the task of formalizing proposed processes and representations within a computational modeling framework, and so the field does not benefit from the application of computational models to evaluate theories.
-
-\nbcnote{Maybe a bit about why that would be good?}
+These approaches have made great advances in our understanding of creativity, but the focus on creative  products such as drawings makes difficult the task of formalizing proposed processes and representations within a computational modeling framework, and so work following this approach does not often employ computational models to provide formal tests of competing theories.
 
@@
 Thus, we can make progress formalizing the processes involved in creative generation by extending theories of categorization to the special case.
-
-\nbcnote{Note that this problem has been under-studied in the categorization lit?}
 
@@
 These ideas are implemented within the well-studied exemplar framework -- the PACKER model is an extension of the influential Generalized Context Model of category learning \citep[GCM;][]{nosofsky1984choice,nosofsky1986attention}.
+
+Although, as an exemplar model, one of PACKER's proposals is people represent categories in terms of a collection of stored exemplars, it is worth emphasizing that the underlying nature of human category representations is not the primary concern of our work.
+The choice to develop PACKER within an exemplar framework reflects the facts that exemplar models are well-studied, strongly theoretically motivated, and dominate much of the theoretical and empirical work in categorization.
+The focus of our work with PACKER concerns the dual constraints of within- and between-class similarity; it is not difficult to imagine how such constraints may be instantiated using alternative frameworks (e.g., \citealp{kurtz2007divergent,love2004sustain,smith2000thirty}; for a review, see \citealp{pothoswills2011}).
 
@@
 \subsection{Procedure}
+
+As noted in the introduction of this paper, the task of generating members of a new category can be aligned with common tasks studied in the categorization literature: Whereas classification consists of predicting an object's category label on the basis of its features, inference consists of predicting an observed feature, given a set of observed features and a category label.
+Generation thus consists of predicting {\em all} features of an object, given a novel category label.
+In keeping with this insight, we designed our generation task as an extension of the traditional artificial classification learning paradigm.
+The task differs from traditional work in creative cognition primarily through the use of an artificial domain, which enables the application of computational models.
+Unlike previous studies \citep[e.g.,][]{ward1994structured}, participants in our studies have no experience with the domain before the start of the experiment, and the experimenter-defined categories are not hierarchically structured (as are plants and animals).
 
@@
 Likewise, basic-level categories \citep{rosch1976basic} are thought to be abstracted in order to maximize within-category similarity while minimizing between-category similarity.
-Finally, the act of forming category representations affects similarity judgments about category members and nonmembers, with category members being viewed as more similar to one another than members of other categories \citep{goldstone1994influences,goldstone2001altering} .
+Finally, the act of forming category representations affects similarity judgments about category members and nonmembers, with category members being viewed as more similar to one another than members of other categories \citep{goldstone1994influences,goldstone2001altering}.
 
@@
 As linguistic categories must have been created at some point in human history, it is revealing that the constraints of emulating distributional structure across categories and having categories contrast from one another still bias human category generation today.
-
-\nbcnote{Something about analogy (similar relationally but different feature-ally); generated categories are alike in distributional structure but differently located in domain}
 

### 2017-06-08 Manuscripts/cog-psych/cogpsych.tex
redid conjugate jk13 section

@@
 
-In our implementation of their model, each category's exemplars are assumed to be sampled from a multivariate Normal distribution with parameters ($\mu, \Sigma$).
-Each category's covariance matrix is assumed to be inverse-Wishart distributed with parameters: $v$, $\kappa,$ and $\Sigma_D$
-\footnote{Note that \citet{jern2013probabilistic}'s model is slightly different, as they used a non-conjugate model. Their model acts very similar to our version of it and receives comparable fits.}. 
-$\Sigma_D$ is the covariance matrix shared between categories.
-We assume the shared covariance matrix $\Sigma_D$ is generated from a Wishart distribution (for conjugacy) with parameters $v_0$, $\kappa_0$, and $\Sigma_0$.
-We set $\nu_0 = 4$, and $\Sigma_0 = \lambda {\bf I}$, where $\lambda$ is a free parameter controlling the expected variance of dimensions (dimensions of the shared covariance matrix are expected to be uncorrelated) and ${\bf I}$ is the identity matrix.
-
-\nbcnote{check on that $\nu_0 = 4$ business}
-
-To simplify the model predictions, we used {\em maximum a posteriori} (MAP) estimates for the hidden parameters and then generated new categories based on those estimates.
-Due to conjugacy, the MAP estimate for the shared covariance matrix $\Sigma_D = \Sigma_0 + \sum_c{C_c}$, where $C_c$ is the empirical covariance matrix of category $c$.
-The MAP estimate of the covariance matrix for the target category $B$ is
+In our implementation of their model\footnote{Note that \citet{jern2013probabilistic}'s model is slightly different, as they used a non-conjugate model.
+Their model acts very similar to our version of it and receives comparable fits.}, each category's exemplars are viewed as samples from a multivariate Normal distribution with parameters ($\mu, \Sigma$).
+Category covariance matrices (specifying variance and covariance along $k$-dimensions, are assumed to be Normal-Inverse-Wishart distributed with parameters: $\nu$ ($>k-1$), $\kappa$ ($>0$), and $\Sigma_D$.
+$\nu$ and $\kappa$ are treated as free parameters in our simulations, and $\Sigma_D$ is the domain-wide covariance matrix from which all categories are viewed as samples.
+Assuming a given $\Sigma_D$, a category covariance matrix $\Sigma$ can be computed on the basis of its examples:
+
 \begin{equation}
-\Sigma_B = \left[ \Sigma_D \nu + C_B +
+\Sigma = \left[ \Sigma_D \nu + C +
 \dfrac
-{\kappa n_B}
-{\kappa + n_B}
-(\bar{x}_B-\mu_B)(\bar{x}_B-\mu_B)^T
-\right] (\nu + n_B)^{-1}
+{\kappa n}
+{\kappa + n}
+(\bar{x}-\mu)(\bar{x}-\mu)^T
+\right] (\nu + n)^{-1}
 \label{eq:Sigma_B}
@@
 % 
-where $\nu$ ($\nu>k-1$) is an additional free parameter (from the Inverse-Wishart prior on $\Sigma_B$) weighting the importance of $\Sigma_{D}$.
-When the target category has no members (i.e., $n_B = 0$), items are generated at random.
-
-\nbcnote{should explain Equation \ref{eq:Sigma_B} more fully now that we are not limited for space.}
-
-\jlanote{K, though if I remember, it's just conjugacy, in which case we can just cite a paper that has IW-W conjugacy.}
-
-\nbcnote{It \textbf{is} just conjugacy, so we can just cite a paper (do you know of a paper or should I start digging?).
-We'll probably want to detail the what we did for priors and which free parameters we fitted though.}
-
-Generated exemplars are drawn from a multivariate Normal distribution specified by $(\mu_{B}, \Sigma_{B})$. Thus, $p(y)$ is
+where $\bar{x}$ and $C$ are the empirical mean and covariance of the category's known members, and $n$ is the number of observed members of the category.
+When there are fewer than two known members of the category (and thus no covariance to speak of), $\Sigma = \Sigma_D\nu$.
+
+The category mean, $\mu$, can be computed as:
+
+\begin{equation}
+  \mu = \dfrac
+    {\kappa\mu_{0} + n \bar{x}}
+    {\kappa + n}
+    \label{eq:category_mus}
+\end{equation}
+% 
+where $\mu_{0}$ is the prior mean.
+In our simulations, $\mu_{0}$ is set to the center of the domain.
+However, when no examples of the target category have been observed, generation is assumed to be random.
+In practice, the model's best fits are achieved when the  $\kappa$ parameter, which controls the influence of $\mu_0$ on $\mu$, is set very close to zero (hence, the influence of $\mu_0$ is minimal).
+
+Importantly, the domain-wide covariance matrix $\Sigma_D$ is unobserved and needs to be inferred from the observed categories.
+For conjugacy, if $\Sigma_D$ is viewed as a sample from an Inverse-Wishart distribution with scale $\Sigma_0$, $\Sigma_D$ can be computed as:
+
+ \begin{equation}
+    \Sigma_D = \Sigma_0 + \sum_y{C_y}
+    \label{eq:category_mus}
+\end{equation}
+% 
+where $\Sigma_0$ is the prior covariance in the domain.
+In our simulations, $\Sigma_0 = \lambda {\bf I}$, where $\lambda$ is a free parameter controlling the expected variance of dimensions (dimensions of the domain covariance matrix are expected to be uncorrelated) and ${\bf I}$ is a $k$-by-$k$ identity matrix.
+
+Generated exemplars are drawn from a multivariate Normal distribution specified by $(\mu, \Sigma)$. Thus, $p(y)$ is
 \begin{equation}
 p(y) = \dfrac
-{\exp \left \{ \theta \cdot {\rm Normal}(y; \mu_{B}, \Sigma_{B}) \right \} }
-{\sum_i \exp \left \{ \theta \cdot {\rm Normal}(y_i; \mu_{B}, \Sigma_{B}) \right \} } 
+{\exp \left \{ \theta \cdot {\rm Normal}(y; \mu, \Sigma) \right \} }
+{\sum_i \exp \left \{ \theta \cdot {\rm Normal}(y_i; \mu, \Sigma) \right \} } 
 \end{equation}
+% 
 where $\theta$ is a response determinism parameter and ${\rm Normal}(y; \mu, \Sigma)$ denotes a multivariate Normal density evaluated at $y$.

### 2017-06-09 Manuscripts/cog-psych/cogpsych.tex
I SAY WITH GREAT SURETY THAT THIS IS THE MOST TREMENDOUS, GREATEST COMMIT OF ALL TIME

@@
 Correspondence concerning this article should be addressed to: 
-Joseph Austerweil, 1202 West Johnson Street, Madison, WI 53706.
+Joseph L. Austerweil, 1202 West Johnson Street, Madison, WI 53706.
 E-mail: austerweil@wisc.edu
@@
 
-The ability to creatively generate new concepts and ideas is among the most fascinating aspects of human cognition, but we do not have a strong understanding of the cognitive processes and representations underlying creative generation.
+The ability to creatively generate new concepts and ideas is among the most fascinating aspects of human cognition, but we do not have a strong understanding of the cognitive processes and representations underlying concept generation.
 In this paper, we study the generation of new categories using the computational and behavioral toolkit of traditional artificial category learning.
-Previous work in this domain has focused on how the statistical structure of known categories generalizes to generated categories, overlooking contrast between the known and generated categories.
-We report two experiments demonstrating that contrast between what is known and what is created is of fundamental importance in the creative process.
+Previous work in this domain has focused on how the statistical structure of known categories generalizes to generated categories, overlooking contrast between the known and generated categories as a factor.
+We report two experiments demonstrating that contrast between what is known and what is created is of fundamental importance for categorization and the creative process.
 We propose a novel, exemplar-based approach explaining our results, and compare the model's performance to two key alternatives.
@@
 Indeed, the generation of novel sentences, works of art, and so on may also be considered products of our capacity to create new concepts.
-However, while this capacity is of great interest, it is also of great complexity and can be difficult to study in a typical laboratory setting.
+However, while this capacity is of great interest, it is also rare and complex, making it difficult to study in a typical laboratory setting.
 By consequence, we do not have a strong understanding of the representations and processes involved in the generation of new concepts.
@@
 The processes underlying conceptual generation are commonly investigated using a creative cognition approach \citep{finke1992creative, smith1995creative}, in which the products of creative acts (such as drawings) are analyzed to obtain insights into how they might have been created.
-These approaches have made great advances in our understanding of creativity, but the focus on creative  products such as drawings makes difficult the task of formalizing proposed processes and representations within a computational modeling framework, and so work following this approach does not often employ computational models to provide formal tests of competing theories.
+These approaches have made great advances in our understanding of creativity, but the focus on particular types of creative products such as drawings makes difficult the task of formalizing proposed processes and representations within a computational modeling framework.
+As a consequence, this approach does not often employ computational models to provide formal tests of competing theories.
 
@@
 In particular, generating a member of a novel class can be considered a `special case' application of existing category knowledge \citep{kemp2014taxonomy,kurtz2015human}.
-Research in categorization typically focuses on classification \citep[predicting an object's category given its features;][]{shepard1961learning,kurtz2013human} and inference \citep[predicting an object's unobserved features given observed features and a category label, see][]{markman2003category}.
-The generation of members of a new category consists of inferring the {\em all} features given a {\em novel} category label.
+Research in categorization typically focuses on classification \citep[predicting an object's category given its features;][]{shepard1961learning,kurtz2013human} and inference \citep[predicting an object's unobserved features given observed features and a category label, see e.g.,][]{markman2003category}.
+The generation of members of a new category consists of inferring {\em all} features for a {\em novel} category label.
 Thus, we can make progress formalizing the processes involved in creative generation by extending theories of categorization to the special case.
@@
 This fundamental constraint, ``being different'', or contrasting from other categories in the relevant domain, is the focus of our work.
-Beyond the assumption that newly generated concepts are indeed novel, this constraint has been overlooked in previous research: To our knowledge, there has not been any systematic investigation addressing how generated categories {\em differ} from what is already known.
+Although implicitly assumed in some work, this constraint has been overlooked in previous research: To our knowledge, there has not been any systematic investigation addressing how generated categories {\em differ} from what is already known.
 Although the idea of category contrast is ubiquitous throughout the categorization literature, and extends to a variety of other fields \citep[e.g., color;][]{regier2007}, the idea that a new category should be ``different'' is vague, as there are many ways it could be different from a previously observed category.
@@
 The authors evaluated the predictions of several formal models on their data.
-Most notably, they showed that a hierarchical Bayesian sampling model provided the strongest account of their results.
+Most notably, they showed that a hierarchical Bayesian model provided the strongest account of their results.
 Their model views observed examples as samples from an underlying category distribution, describing the location of the category in the space, as well as how it varies along each feature.
@@
 The core proposal is that participants generate new items by copying stored examples from memory and tweaking them to generate something new.
-The copy-and-tweak model differs from the path of least resistance account in that it notably omits the hierarchical organization of categories, as well as selectivity in which features are changed.
+The copy-and-tweak model differs from the path of least resistance account in that it notably omits the hierarchical organization of categories, as well as selectivity in which features are changed \citep[both of which are factors in the Path of Least Resistance account;][]{ward2002role}.
 Instead, their copy-and-tweak model corresponds to a direct exemplar-similarity approach \citep[e.g.,][]{nosofsky1984choice,nosofsky1986attention}: The model generates new items according to their similarity to known members of the target category.
@@
 
-\section{Something Different: A Role For Contrast}
+\section{And Now for Something Different: A Role For Contrast}
 
@@
 Indeed, under a strict interpretation of their model, the most probable new category to be generated is located in {\em exactly} the same location and distributional information given a single category.
-However, this is not simply an issue with their model but with the broader class of standard hierarchical Bayesian models.
-These models assume that at some point of the latent generative process the same underlying distribution generates all of the categories and thus, any differences between categories is due to {\em noise} and should not be {\em systematic}.
-The best a hierarchical Bayesian model can do at capturing contrast is to assume that the new category is placed uniformly at random over stimulus space.
-This defeats the purpose of a hierarchy as new category locations do not depend on the locations of previously observed categories.\footnote{It is plausible that some hierarchical Bayesian model could be created that generates categories different from each other.
-However, this model would not be a standard application or extension of any pre-existing hierarchical Bayesian model.}
+However, this is not simply an issue with their model but with a broader class of standard hierarchical Bayesian models \citep[][e.g., ]{griffithsscn07,kemp07}.
+Many of these models assume that at some point of the latent generative process the same underlying distribution generates all of the categories and thus, any differences between categories is due to {\em noise} and should not be {\em systematic}.
+The best a standard hierarchical Bayesian model can do at capturing contrast is to assume that the new category is placed uniformly at random over stimulus space.
+This defeats the purpose of a hierarchy as it is ignored when determining a new category location!\footnote{It is plausible that some hierarchical Bayesian model could be created that generates categories different from each other.
+However, this model would not be a standard application or extension of most pre-existing hierarchical Bayesian models.}
 
@@
 
-\subsubsection{Novel Analyses Demonstrating Contrast Effects in Prior Work}
-
-Although existing accounts of creative generation broadly overlook the role of category contrast in determining what is novel versus familiar, it is assumed that learners in previous experiments were {\em successful} in creating new categories.
+\subsection{Novel Analyses Demonstrating Contrast Effects in Prior Work}
+
+Although existing accounts of creative generation broadly overlook the role of category contrast in determining what is novel versus familiar, it was implicitly assumed that learners in previous experiments were {\em successful} in creating new categories.
 Thus, effects of category contrast should be observable within the experimental results of these studies.
@@
 Unfortunately, stimulus hue was encoded and presented in the Hue-Saturation-Value (HSV) color space, which is device-dependent and not perceptually normed such that perceived color similarity corresponds to proximity in the color space (as opposed to a color space such as CIELAB that is device-independent and equidistant sets of points correspond to pairs of colors that have the same perceptual similarity; \citealp{wyszecki1967}).
-Further, they did not calibrate their monitor, and so we cannot know the precise values of colors presented to participants.
+Further, they did not calibrate their monitor, and so we cannot know the precise colors presented to participants.
 As \cite{jern2013probabilistic} were interested in relations between the saturation and length of examples in generated into novel categories, this is not an issue for their analyses and results.
@@
 
-Although we cannot know precisely which colors were displayed or perceived, we can still analyze their results from a coarse perspective to see whether there is preliminary support for contrast.
+Although we cannot know the precise colors that were displayed or perceived, we can still analyze their results from a coarse perspective to see whether there is preliminary support for contrast.
 To do so, we binned all possible hues into one of eight uniformly-spaced color groups: \{{\em Red}: $0-0.063$, $0.938-1$, {\em Yellow}: $0.063-0.188$, {\em Yellow-Green}: $0.188-0.313$, {\em Green-Teal}: $0.313-0.438$, {\em Teal}: $0.438-0.563$, {\em Teal-Blue}: $0.563-0.688$, {\em Purple}: $0.688-0.813$, {\em Pink}: $0.813-0.938$\}.
-In the \cite{jern2013probabilistic} experiment, the hue of each experimenter-defined category was selected from one of six possible values, each of which falls into one of the color groups above (two color groups do not contain a possible hue for the experimenter-defined categories).
+In the \cite{jern2013probabilistic} experiment, the hue of each experimenter-defined category was selected from one of six possible values, each of which falls into one of the color groups above (two color groups was not used as a possible hue for the experimenter-defined categories).
 By categorizing the participant-generated crystals likewise, we can obtain a broad measure of category contrast by determining the proportion of participant-generated crystals that fall into the same groups as the experimenter-defined categories: If contrast influences the hues of the generated categories, we should observe minimal overlap between in the color groupings.
@@
 Taken alongside the analyses reported by \cite{jern2013probabilistic}, our analysis suggests that generated categories tend to be distinct from {\em and} distributionally similar to what is already known.
-However, it is worth noting that our analysis is still limited: the color groups defined above are somewhat arbitrary, and it is not clear that our color grouping is consistent with psychological color boundaries.
+However, it is worth noting that our analysis is still limited: the color groups defined above are somewhat arbitrary, and it is not clear that our color grouping is consistent with the psychological color boundaries perceived by participants.
 While we did obtain similar results using a variety of alternative groupings, the hue dimension used in the \cite{jern2013probabilistic} study does not lend itself straightforwardly to the the computation of similarities, and thus we cannot be certain of whether our coding accurately approximates the psychological space of the stimuli.
@@
 It is also worth noting that PACKER represents just one of many possible models that incorporate contrast in category generation.
-For example, although it may be possible to extend \cite{jern2013probabilistic}'s hierarchical sampling model to include contrast, this mechanism would be at odds with a hierarchical Bayesian framework.
+For example, it may be possible to extend \cite{jern2013probabilistic}'s hierarchical Bayesian model to include contrast, this mechanism would be at odds with the standard hierarchical Bayesian framework.
 In contrast, these ideas emerge naturally from the exemplar view: We have not modified any of the core elements of the GCM in defining the PACKER model, we simply weight exemplar similarity during aggregation.
@@
 The manner in which generated categories differ from known ones is, nonetheless, poorly understood: Existing theories do not make strong predictions about how creatively generated concepts should systematically differ from existing ones.
-Above, we provided encouraging initial support that contrast influences category generation \citep[][Experiment 3]{jern2013probabilistic}, and we introduced a novel, exemplar-based model formalizing the roles of similarity and contrast in creative generation.
+Above, we found encouraging initial support that contrast influences category generation \citep[][Experiment 3]{jern2013probabilistic}, and we introduced a novel, exemplar-based model formalizing the roles of similarity and contrast in creative generation.
 
@@
 These dimensions have been used in numerous classification learning studies \cite[e.g.,][]{conaway2016similar,conaway2016generalization,shepard1961learning,nosofsky1994comparing}.
-Unlike those used in the \cite{jern2013probabilistic} experiments, proximity along these dimensions aligns more directly with perceptual similarity, allowing us to evaluate the role of category contrast in creative generation more precisely.
+Unlike those used in the \cite{jern2013probabilistic} experiments, distance on the physical dimensions aligns more directly with perceptual similarity, allowing us to evaluate the role of category contrast in creative generation more precisely.
 To extend the evidence provided by the \cite{jern2013probabilistic} data, we tested the effects of category contrast after learning one category from a set of qualitatively distinct category structures, as shown in Figure \ref{fig:e1-conditions}.
 
-Panels B-D of Figure \ref{fig:e1-conditions} show the values of exemplar dimensions belonging to the experimenter-defined categories (`A', or `Alpha') that participants were assigned to learn about prior to generating a new category.
+Figure \ref{fig:e1-conditions}b-d show the values of exemplar dimensions belonging to the experimenter-defined categories (`A', or `Alpha') that participants were assigned to learn about prior to generating a new category.
+Each participant learned one of the category types during training.
 In the `Cluster' type, category A is a tight cluster of examples in the space.
@@
 
-In this way, our experimental results provide a conceptual replication of the classic finding that generated categories tend to share distributional properties with known categories in the domain \citep[see][]{jern2013probabilistic,ward1994structured}.
+Our experimental results also provide a converging test of the classic finding that generated categories tend to share distributional properties with known categories in the domain \citep[see][]{jern2013probabilistic,ward1994structured}.
 From these results, we can predict that, in each condition, participants should generate categories that are distributionally similar to the experimenter-defined category: In the Cluster condition, generated categories should be tightly clustered.
@@
 Thus, if contrast plays a role, exemplars in the generated categories of participants in the XOR condition may not be positively correlated -- they may be negatively correlated instead.
-In this case, contrast and statistical regularities would interact, which would be inconsistent with the leading approaches in conceptual generation \citep{jern2013probabilistic}
+In this case, contrast and statistical regularities would interact, which would be inconsistent with the leading approach in conceptual generation \citep{jern2013probabilistic}
 
@@
 The task differs from traditional work in creative cognition primarily through the use of an artificial domain, which enables the application of computational models.
-Unlike previous studies \citep[e.g.,][]{ward1994structured}, participants in our studies have no experience with the domain before the start of the experiment, and the experimenter-defined categories are not hierarchically structured (as are plants and animals).
+Unlike most previous studies \citep[e.g.,][]{ward1994structured}, participants in our studies have no experience with the domain before the start of the experiment, and the experimenter-defined categories are not hierarchically structured (as are plants and animals).
 
@@
 We observed a substantial degree of individual differences in our data.
-In Figure \ref{fig:e1-samples} we have plotted sample data from several participants, from which it evident that different participants generated qualitatively different category structures.
+In Figure \ref{fig:e1-samples} we have plotted sample data from several participants, from which it is evident that different participants generated qualitatively different category structures.
 In this section we will focus on analyzing the data in aggregate, but in later sections we will explore how individual differences can be explained.
@@
 
-Thus, the core results of Experiment 1 indicate that generated categories systematically differ from what can be expected based on prior work.
+Thus, the core results of Experiment 1 indicate that generated categories can systematically differ from what we would expect based on prior work.
 The negative (or zero) correlations observed in the XOR condition suggests an interesting interaction between contrasting from a given category and emulating statistical properties.
@@
 To test whether category contrast influences the distributional structure of generated categories, we sought to identify conditions in which differences in the distributional structure of generated categories cannot be explained by the distributional structure of the experimenter-defined category.
-We created two new category types (depicted in Figure \ref{fig:e2-conditions}) that possess an identical distributional structure (both are tight clusters of examples with no correlation between features), and only differ in their Y-axis position: the `Bottom' category lies in the bottom-center of the space, and the `Middle' category lies in the center.
-The distributional equality of these conditions is key to the design of the experiment: If the structure of previously learned categories were the only influence on the structure of generated categories, we should observe no difference between these two conditions with respect to distributional structure.
+We created two new category types (depicted in Figure \ref{fig:e2-conditions}) that possess identical distributional structure (both are tight clusters of examples with no correlation between features), and only differ in their Y-axis position: the `Bottom' category lies in the bottom-center of the space, and the `Middle' category lies in the center.
+The distributional equality of these conditions is key to the design of the experiment: If the structure of previously learned categories were the only influence on the structure of generated categories, we should observe no difference in the categories participants generate between these two conditions.
+Will participants distribute their generated category differently between conditions due only to the differences in the available stimulus space for generating a new category?
 
@@
 
-Critically, participants seeking to generate a distinct category, contrasting from the experimenter-given category, would be more likely to distribute members of the generated category into areas that are distant from members of known categories.
-Thus, if category contrast influences the distributional structure of the categories people generate, then we should observe different types of categories according to the shape of the space that is unoccupied by members of previously learned categories.
-The difference in the Y-axis position between the Bottom and Middle types produces a considerable change to the shape of the unoccupied space.
+If category contrast influences the distributional structure of the categories people generate, then we should observe different types of categories according to the shape of the space that is unoccupied by members of previously learned categories.
+The difference in the Y-axis position between the Bottom and Middle conditions produces a considerable change to the shape of the unoccupied space.
 Participants assigned to learn the Bottom category should be less likely to generate exemplars into the lower regions of the stimulus space (as these areas possess greater similarity to members of the Bottom category), preferring instead to distribute exemplars across the upper region of the space.
@@
 Participants first completed a short training phase, followed by the generation phase.
+The only difference from Experiment 1 was the category types given to participants.
 
@@
 To characterize the nature of these differences, Figure \ref{fig:e2-samples} depicts sample categories generated by participants.
-The data from each condition is organized into four columns based on commonly observed patterns of generation: a `Cluster' type of tightly-clustered examples,   `Row' and `Column' types of exemplars widely distributed along the one axis but narrowly along the other, and a `Corners' type, wherein participants placed exemplars in disparate corners of the space.
+The data from each condition are organized into four columns based on commonly observed patterns of generation: a `Cluster' type of tightly-clustered examples,   `Row' and `Column' types of exemplars widely distributed along the one axis but narrowly along the other, and a `Corners' type, wherein participants placed exemplars in disparate corners of the space.
 As before, in this section we focus on analyzing the data in aggregate, but in later sections we will focus more specifically on explaining the individual differences.
@@
 These data, shown in Figure \ref{fig:e2-distanceplots}, yield very similar results.
-Stimuli that are more distant from members of the experimenter-defined category were more frequently generated, and the categories in each condition tended to possess more between-category than within-category distance: Bottom, $t(60) = 5.5$, $p < 0.001$; Middle, $t(60) = 2.71$, $p = 0.009$.
+Participants generated stimuli that are distant from members of the experimenter-defined category, and the categories in each condition tended to possess more between-category than within-category distance: Bottom, $t(60) = 5.5$, $p < 0.001$; Middle, $t(60) = 2.71$, $p < 0.01$.
 We did, however, observe a notable subgroup of participants in each condition who generated categories with more within-category than between-category distance.
@@
 As noted above, if the distributional structure of generated categories is influenced by the shape of the space not occupied by members of known categories, then participants in the Middle condition would be more likely to place exemplars in the upper {\em and} lower regions of the space, as members of the experimenter-defined category are equidistant from these regions.
-Participants in the Bottom condition should in turn be less likely to generate category members in the bottom regions because members of the experimenter-defined category are located there.
-These predictions can be addressed within the Y-axis ranges of the generated categories: If Middle participants utilize the upper and lower regions of the space, their categories should vary more along the Y-axis.
+Participants in the Bottom condition should be less likely to generate category members in the bottom regions because members of the experimenter-defined category are located there.
+One way to test these predictions is to analyze the Y-axis ranges of the generated categories: If Middle participants utilize the upper and lower regions of the space, their categories should vary more along the Y-axis.
 T-Tests comparing the conditions on the distributional statistics, however, reveal few between-group differences: the conditions do not differ with respect to X-axis range, Y-axis range, or feature correlations ($ps>0.17$).
@@
 Specifically, we present simulations from the PACKER model, as well as a `copy-and-tweak' model (discussed in Section \ref{section:copytweak-packer}), defined as a variant of PACKER with the $\gamma$ parameter constrained to be one.
-The comparison of these two models serves to highlight the explanatory role of contrast within PACKER's framework: if contrast affords little explanatory advantage, then the two accounts should produce an equally strong account.
-We also present simulations from an implementation of the hierarchical sampling model proposed by \cite{jern2013probabilistic}, described in-depth in Appendix \ref{ap:hsampling-definition}.
-The comparison between the hierarchical sampling model and PACKER is meant to emphasize the necessity of contrast and demonstrate that generation cannot be explained entirely through the emulation of distributional structure: whereas PACKER is insensitive to the distributional structure of learned categories (relying only on within- and between-category similarity), the hierarchical sampling model generates categories exclusively on the basis of knowledge of how existing classes are distributed.
+The comparison of these two models serves to highlight the explanatory role of contrast within PACKER's framework: If contrast affords little explanatory advantage, then the two accounts should produce an equally strong account.
+We also present simulations from an implementation of the hierarchical Bayesian model proposed by \cite{jern2013probabilistic}, described in-depth in Appendix \ref{ap:hsampling-definition}.
+The comparison between the hierarchical Bayesian model and PACKER is meant to emphasize the necessity of contrast and demonstrate that generation cannot be explained entirely through the emulation of distributional structure.
+Whereas PACKER is insensitive to the distributional structure of learned categories (relying only on within- and between-category similarity), the hierarchical Bayesian model generates categories exclusively on the basis of knowledge of how existing classes are distributed.
 
@@
 To obtain a global measure of the quality of each model's account, we fitted the parameters of each model to our entire dataset (Experiments 1 and 2 combined), using a hill-climbing algorithm which maximized the log-likelihood of the model's predictions of the observed responses (1220 responses from 305 total participants).
-We fitted three parameters in the PACKER model ($c$, $\gamma$, and $\theta$; see Section \ref{section:PACKER-definition}), as well as four in the hierarchical sampling model ($\kappa$, $\lambda$, $\nu$, and $\theta$; see Appendix \ref{ap:hsampling-definition}).
+We fitted three parameters in the PACKER model ($c$, $\gamma$, and $\theta$; see Section \ref{section:PACKER-definition}), as well as four in the hierarchical Bayesian model ($\kappa$, $\lambda$, $\nu$, and $\theta$; see Appendix \ref{ap:hsampling-definition}).
 We fitted only two parameters for the copy-and-tweak model ($c$, and $\theta$), as $\gamma$ is held constant ($\gamma = 1$).
@@
 \centering
-\caption{Results of model-fitting to the combined datasets from Experiments 1 and 2. Note that lower AIC values correspond to better model fits (adjusted for number of parameters)}
+\caption{Results of model-fitting to the combined datasets from Experiments 1 and 2. Note that smaller AIC values correspond to better model fits (adjusted for number of parameters)}
 \label{table:global-model-fits}
@@
 \\
- \textbf{PACKER}    & \textbf{Copy \& Tweak} & \textbf{Hierarchical Sampling} \\ 
+ \textbf{PACKER}    & \textbf{Copy \& Tweak} & \textbf{Hierarchical Bayesian} \\ 
  \hline
@@
 Due to the uneven number of fitted parameters among the models, we compare the model fits using the Akaike Information Criterion \citep[AIC;][]{akaike1974new}, where smaller values correspond to better fits (discounted by the number of parameters).
-Similar results were obtained with alternative model comparison metrics (e.g., BIC, \citealp{schwarz1978estimating}; $AIC_C$, \citealp{hurvich1989regression})
+The same qualitative results were obtained with alternative model comparison metrics (e.g., BIC, \citealp{schwarz1978estimating}; $AIC_C$, \citealp{hurvich1989regression})
 
 Table \ref{table:global-model-fits} contains the AIC values of each model's best fit, as well as the corresponding log-likelihood ($L$) and the best-fitting parameter values.
-These results reveal strong model differentiation: The PACKER model achieved far better fits compared to the copy-and-tweak and hierarchical sampling models, and copy-and-tweak performed somewhat better than the hierarchical sampling model.
-While PACKER's advantage may tentatively be attributed to the model's sensitivity to category contrast (this will be explored in detail below), the advantage shown by copy-and-tweak over the hierarchical sampling model may be attributed to its exemplar-based representation, as opposed to the prototype-based representation assumed in the hierarchical sampling model.
+These results reveal strong model differentiation: The PACKER model achieved far better fits compared to the copy-and-tweak and hierarchical Bayesian models, and copy-and-tweak performed somewhat better than the hierarchical Bayesian model.
+While PACKER's advantage may tentatively be attributed to the model's sensitivity to category contrast (this will be explored in detail below), the advantage shown by copy-and-tweak over the hierarchical Bayesian model may be attributed to its exemplar-based representation, as opposed to the prototype-based representation assumed by the hierarchical Bayesian model.
 As observed in Figures \ref{fig:e1-samples} and \ref{fig:e2-samples}, the generated categories we observed were often widely distributed, with no items near the category prototype.
@@
 
-A key distinction between PACKER and copy-and-tweak, as well as the hierarchical sampling model, is that, of the three models, only PACKER is capable of making strong predictions about the location of new category members when the target class is entirely novel (i.e., no member of the category has been observed).
+A key distinction between PACKER and copy-and-tweak, as well as the hierarchical Bayesian model, is that, of the three models, only PACKER is capable of making strong predictions about the location of new category members when the target class is entirely novel (i.e., no member of the category has been observed).
 Under these circumstances, there are no examples to copy, and thus the copy-and-tweak model predicts that items are generated at random.
-Likewise, with no observations on which to condition the mean of the category distribution, the hierarchical sampling model also picks an item at random.
+Likewise, with no observations on which to condition the mean of the category distribution, the hierarchical Bayesian model also picks an item at random.
 Thus, it is possible that the failure of these models is simply due to their inability to explain each participant's first trial (generating the first item in the `Beta' category).
-We conducted an identical set of simulations as above, excluding this trial (leaving 915 responses in the dataset): Again, PACKER ($L = -3390$, $AIC = 6786$) achieved better fits than the copy-and-tweak ($L = -3579$, $AIC = 7162$) and hierarchical sampling ($L=-3612$, $AIC = 7232$) models.
+We conducted an identical set of simulations as above, excluding this trial (leaving 915 responses in the dataset): Again, PACKER ($L = -3390$, $AIC = 6786$) achieved better fits than the copy-and-tweak ($L = -3579$, $AIC = 7162$) and hierarchical Bayesian ($L=-3612$, $AIC = 7232$) models.
 
@@
 As noted in Experiments 1 and 2, we observed a great deal of individual differences in the types of categories that participants generated.
-Within each condition, we observed tightly clustered and widely distributed categories, as well as row and column categories, and so forth (see Figures \ref{fig:e1-samples} and \ref{fig:e2-samples}).
+Within each condition, there were a wide variety of category types, such as row and column categories (see Figures \ref{fig:e1-samples} and \ref{fig:e2-samples}).
 The simulations reported above serve to evaluate the models while considering the entire dataset, but a secondary goal of any formal account should be to provide some explanation of how different profiles of performance emerge.
-Many of the individual generation profiles we observed can be described within the PACKER framework, simply by tuning the model's parameters.
+Many of the individual generation profiles we observed can be described within the PACKER framework, simply by tuning the model's parameters in a principled manner.
 In this section, we describe more specifically how the most frequently observed profiles can be realized.
@@
 
-Similar to the copy-and-tweak model, the hierarchical sampling model possesses no mechanism to account for category contrast, and the model is most likely to generate new items that are similar to known examples of the target category.
+Similar to the copy-and-tweak model, the hierarchical Bayesian model possesses no mechanism to account for category contrast, and the model is most likely to generate new items that are similar to known examples of the target category.
 Although this precludes an account of why individuals might assume a cluster versus corners profile, the covariance matrix specifying the model's prior domain distribution, $\Sigma_0$, can be used to explain the generation of row-like and column-like categories.
@@
 
-As noted above, while the copy-and-tweak and hierarchical sampling models possess mechanisms to explain row- and column-like categories, they cannot easily explain why some individuals widely distribute their generated categories into disparate corners of the space.
-This, however, reveals a more general limitation: According to the copy-and-tweak and hierarchical sampling models, the distributional structure of generated categories is {\em independent} of their location within the domain.
-For example, although the copy-and-tweak or hierarchical sampling models can be parameterized to generate row- or column-like categories, there is no mechanism in place to ensure that what is generated will be distinct from what is already known.
+As noted above, while the copy-and-tweak and hierarchical Bayesian models possess mechanisms to explain row- and column-like categories, they cannot easily explain why some individuals widely distribute their generated categories into disparate corners of the space.
+This, however, reveals a more general limitation: According to the copy-and-tweak and hierarchical Bayesian models, the distributional structure of generated categories is {\em independent} of their location within the domain.
+For example, although the copy-and-tweak or hierarchical Bayesian models can be parameterized to generate row- or column-like categories, there is no mechanism in place to ensure that what is generated will be distinct from what is already known.
 In the next subsection, we explore this prediction through an analysis of the interdependence between distributional structure and location in creative generation.
@@
 
-As noted above, while all three models make clear claims about the internal structure of generated categories, the copy-and-tweak and hierarchical sampling models do not make any claims about how generated categories should differ from what is already known.
+As noted above, while all three models make clear claims about the internal structure of generated categories, the copy-and-tweak and hierarchical Bayesian models do not make any claims about how generated categories should differ from what is already known.
 However, as we observed in Experiment 2, the distributional structure of a category is not always independent of its location within the domain.
-To demonstrate this point in a broader manner, we computed the X- and Y- axis ranges of every participant-generated category.
+To demonstrate this point in more depth, we computed the X- and Y- axis ranges of every participant-generated category.
 Taking the difference between these values ($X-Y$) produces a measure of each category's orientation in the space: positive difference scores correspond to categories with more X-axis range (horizontally aligned, `Row' categories), whereas negative difference scores indicate the opposite (vertically aligned, `Column' categories).
@@
 
-To explore how well PACKER, copy-and-tweak- and hierarchical sampling models explain our findings we conducted simulations using an individual-differences approach.
+To explore how well PACKER, copy-and-tweak- and hierarchical Bayesian models explain our findings we conducted simulations using an individual-differences approach.
 As noted in Section \ref{section:individual-diffs}, row- and column-like categories can be produced by each model through changes to the weighting of each dimension.
@@
 where $\theta_w$ is a free parameter controlling how differences in range correspond to differences in weights (functioning similarity to the $\theta$ parameter in each of the models), and $\text{range}(k)$ is the range of examples generated by the participant along dimension $k$.
-In our simulations $\theta_w = 1.5$, though the results are robust and similar for other $\theta_w$ values.
+We used $\theta_w = 1.5$ in our simulations, though the results are robust and similar for other $\theta_w$ values.
 The resulting $w$ values are thus inversely proportional to the range of generated categories along each dimension, with less range corresponding to greater weighting.
 
-Unlike the PACKER and copy-and-tweak models, the hierarchical sampling model's dimensional variances correspond to the assumed variance of generated categories along each dimension (rather than the inverse of the variance).
+Unlike the PACKER and copy-and-tweak models, the hierarchical Bayesian model's dimensional variances correspond to the assumed variance of generated categories along each dimension (rather than the inverse of the variance).
 Thus, a different transformation is appropriate for incorporating the weights computed in Equation \ref{eq:range-weight}.
-For the hierarchical sampling model, we computed the dimensional variances according to: $\lambda (1-w_k)d$, where $\lambda$ is a free parameter specifying the overall assumed variance of the domain, and $d$ is the number of dimensions\footnote{This calculation applies only in two-dimensional domains, where $w_2 = 1-w_1$.}.
+For the hierarchical Bayesian model, we computed the dimensional variances according to: $\lambda (1-w_k)d$, where $\lambda$ is a free parameter specifying the overall assumed variance of the domain, and $d$ is the number of dimensions\footnote{This calculation applies only in two-dimensional domains, where $w_2 = 1-w_1$.}.
 Under this approach, evenly distributed weights correspond to an assumed variance of $\lambda$.
@@
 The other free parameters within each model were set as in Table \ref{table:global-model-fits}.
-Every participant's generation was simulated 2,000 times; multiplied against 305 total participants, each model generated 610,000 categories in total.
+Every participant's generation was simulated 2,000 times; given the 305 participants tested across the two experiments, each model generated 610,000 categories in total.
 For comparison with our behavioral results, we then computed the range difference gradient identically as with the behavioral data.
@@
 
-As in the more traditional model evaluation procedure described above, PACKER provided a much closer match to our behavioral results than the copy-and-tweak and hierarchical sampling models.
+As in the more traditional model evaluation procedure described above, PACKER provided a much closer match to our behavioral results than the copy-and-tweak and hierarchical Bayesian models.
 In all conditions, PACKER distributes categories similarly to the behavioral data: horizontally-aligned categories tend to be placed above and below members of the experimenter-defined category, and vertically-aligned categories tend to be placed to the sides.
-Conversely, because the copy-and-tweak and hierarchical sampling models are insensitive to category contrast, these models do not produce any systematic patterns of association between category location and distributional structure.
+Conversely, because the copy-and-tweak and hierarchical Bayesian models are insensitive to category contrast, these models do not produce any systematic patterns of association between category location and distributional structure.
 The sole exception is within the `Row' condition of Experiment 1, in which the majority of participants generated a `Row'-like category, widely distributed along the X-axis but not the Y-axis.
@@
 
-The creative generation of concepts is an greatly intriguing topic, but it is also of great complexity and can be difficult to study in a traditional laboratory setting.
+The creative generation of concepts is an immensely intriguing topic, but it is also complex and difficult to study in a traditional laboratory setting.
 Creative generation is most typically studied using a creative cognition approach \citep{finke1992creative,smith1995creative}, where creative products \citep[such as drawings of alien plants and animals;][]{ward1994structured} are analyzed to obtain insight into the processes and representations involved.
-While this approach has yielded a great deal of information about the nature of creative generation, the commonly employed response modalities (e.g., hand-drawn pictures) made difficult the development of computational models to formally evaluate the proposed representations and processes.
-However, the generation of novel concepts is not altogether different from the types of behaviors studied in cognitive psychology: In particular, it can be viewed as an form of inference using one's knowledge about existing categories \citep{kemp2014taxonomy,kurtz2015human}.
+While this approach has yielded a great deal of information about the nature of creative generation, the commonly employed response modalities (e.g., hand-drawn pictures) made it difficult to develop computational models capable of formally evaluating the proposed representations and processes.
+However, the generation of novel concepts is not altogether different from other behaviors studied in cognitive psychology: In particular, it can be viewed as an form of inference using one's knowledge about existing categories \citep{kemp2014taxonomy,kurtz2015human}.
 Taking advantage of this insight, in this paper we investigated the generation of new categories using the empirical and computational toolkit from the field of categorization.
@@
 Members of participant-generated categories tended to be highly dissimilar from members of previously-learned categories, and were usually more similar to one another than to members of other categories.
-We also observed broad interdependence between the distributional structure (feature variance, correlation) and physical instantiation (location within the domain space) of generated categories: In Experiment 2, we found that the unoccupied regions of the domain influenced the distributional structure of categories, and in both Experiments we observed that participants distributed their generated categories to increase contrast with what was already known.
-
-We conducted simulations comparing PACKER's account of our results to that of a ``copy-and-tweak'' model (realized as a variant of PACKER with no sensitivity to category contrast), and a hierarchical sampling model designed to explain the classic distributional similarity effect.
-In all simulations, we found that PACKER's sensitivity to contrast considerably enhanced the models account of human category generation.
+We also observed broad interdependence between the distributional structure (feature variance, correlation) and physical instantiation (location within the stimulus space) of generated categories: In Experiment 2, we found that the unoccupied regions of the domain influenced the distributional structure of categories, and in both experiments we observed that participants distributed their generated categories to increase contrast with what was already known.
+
+We conducted simulations comparing PACKER's account of our results to the other major proposals for category generation: a ``copy-and-tweak'' model (realized as a variant of PACKER with no sensitivity to category contrast), and a hierarchical Bayesian model designed to explain the classic distributional similarity effect.
+In all simulations, we found that PACKER's sensitivity to contrast captured a previously unexplained and unexplored aspect of human category generation.
 Further, by measuring PACKER's fit as a function of its prioritization of within- and between-category similarity, we observed that considering either constraint exclusively results in a relatively low-quality account.
@@
 
-The above passages serve to demonstrate that the dual forces of within-class similarity and between-class contrast influence cognitive function in a wide variety of domains.
+The dual forces of within-class similarity and between-class contrast influence cognitive functions in a wide variety of domains.
 The PACKER model is notable in that it successfully interprets this trade-off within the domain of creative generation, and allows us to begin to understand the relatively understudied processes involved in creative generation through our more well-developed knowledge of human categorization.
@@
 In this paper, we have identified another important aspect of the constraining influence of prior knowledge: What is generated cannot be the same as what is already known.
+Further, there is systematicity in how generated categories differ from prior knowledge.
 The results of our simulations with PACKER demonstrate that this influence is concisely explained in terms of a trade-off between within-category similarity and between-category dissimilarity.
@@
 Even within Experiment 1, we found systematic inconsistencies: by generating exemplars into unoccupied regions of the space, participants who learned an `XOR' category, composed of members that are widely distributed along both features and are positively correlated in space, tended to generate categories with an opposite (negative) correlation.
-More generally, PACKER's success over the hierarchical sampling model indicates that the emulation of distributional structure exerts only a limited influence in comparison to category contrast in some situations.
+More generally, PACKER's success over the hierarchical Bayesian model indicates that the emulation of distributional structure exerts only a limited influence in comparison to category contrast in some situations.
 
@@
 Instead, through the development and evaluation of the PACKER model, we have sought to add new elements into such a phenomenology: The broad and strong influence of category contrast, and the interdependence between category location and distributional structure.
-It may be possible to combine the hierarchical sampling approach proposed by \cite{jern2013probabilistic} with PACKER's underlying claims to obtain a ``best of both worlds'' model, capable of explaining the role of contrast in creative generation, as well as the emulation of distributional structure.
+It may be possible to combine the hierarchical Bayesian approach proposed by \cite{jern2013probabilistic} with PACKER's underlying claims to obtain a ``best of both worlds'' model, capable of explaining the role of contrast in creative generation, as well as the emulation of distributional structure.
 However, as noted in the introduction, the incorporation of category contrast is antithetical to the core principles of a traditional, semi-conjugate Bayesian approach.
-This suggests that category generation is a fundamentally different computational-level problem.
+This suggests that category generation is a fundamentally different computational-level problem \citep[different from those posed by][]{jern2013probabilistic,kemp2014taxonomy}.
 Characterizing that problem and conducting a rational analysis is an important direction for future research.
@@
 Once formalized in probabilistic terms, it should also be straightforward to incorporate distributional factors into the model.
+Alternatively, it may be possible to integrate the core principles of PACKER into other categorization models, such as SUSTAIN \citep{lovemg04} or DIVA \citep{}.
 
@@
 Instead, we reduced our focus by studying a highly complex behavior (creative generation) as it applies within a well-established domain (artificial category learning).
+We hope that future work incorporates and highlights the importance of contrast into theories of creativity.
 
@@
 Extending the literature on creative cognition, our experiments provide a detailed picture of the role of category contrast in generation: People seek to create concepts that are distinct from what they already know, and the nature of what is created can be influenced by what does not yet exist.
-Our simulations with traditional exemplar models, as well as a hierarchical sampling model, provide strong support for the claim that category contrast is of fundamental importance, and, more generally, demonstrate that popular explanatory approaches from basic research in cognitive science can offer a precise, quantitative account of a behavior as complex as that of creative generation.
+Our simulations with traditional exemplar models, as well as a hierarchical Bayesian model, provide strong support for the claim that category contrast is of fundamental importance.
+More generally, our results demonstrate that popular explanatory approaches from basic research in cognitive science can offer a precise, quantitative account of a behavior as complex as that of creative generation.
 
@@
 
-\section{The Hierarchical Sampling Model}
+\section{The Hierarchical Bayesian Model of Concept Generation}
 \label{ap:hsampling-definition}
@@
 
-In our implementation of their model\footnote{Note that \citet{jern2013probabilistic}'s model is slightly different, as they used a non-conjugate model.
+In our implementation of their model\footnote{Note that \citet{jern2013probabilistic}'s model is slightly different, as they used a semi-conjugate model.
 Their model acts very similar to our version of it and receives comparable fits.}, each category's exemplars are viewed as samples from a multivariate Normal distribution with parameters ($\mu, \Sigma$).

### 2017-06-09 Manuscripts/cog-psych/cogpsych.tex
nolan's lawyer reports that he feels vindicated after the changes to the code

@@
 Indeed, the generation of novel sentences, works of art, and so on may also be considered products of our capacity to create new concepts.
-However, while this capacity is of great interest, it is also rare and complex, making it difficult to study in a typical laboratory setting.
+However, while this capacity is of great interest, it is also highly complex, making it difficult to study in a typical laboratory setting.
 By consequence, we do not have a strong understanding of the representations and processes involved in the generation of new concepts.
@@
 The copy-and-tweak model differs from the path of least resistance account in that it notably omits the hierarchical organization of categories, as well as selectivity in which features are changed \citep[both of which are factors in the Path of Least Resistance account;][]{ward2002role}.
-Instead, their copy-and-tweak model corresponds to a direct exemplar-similarity approach \citep[e.g.,][]{nosofsky1984choice,nosofsky1986attention}: The model generates new items according to their similarity to known members of the target category.
-This model provided a poor account of their results, as the experiments devised by \citeauthor{jern2013probabilistic} were specifically designed to challenge this model.
+Instead, their copy-and-tweak model corresponds to a direct exemplar-similarity approach \citep[e.g.,][]{nosofsky1984choice,nosofsky1986attention}, generating new items according to their similarity to known members of the target category.
+The copy-and-tweak model provided a poor account of their results, as the experiments devised by \citeauthor{jern2013probabilistic} were specifically designed to challenge it.
 However, its application is notable as a first step toward understanding creative cognition using well-known formal approaches from the categorization literature.
@@
 The authors do not make predictions about the {\em location} of the category within the domain (the perceptual instantiation of category members).
-Indeed, under a strict interpretation of their model, the most probable new category to be generated is located in {\em exactly} the same location and distributional information given a single category.
+Under a strict interpretation of their model, given knowledge of a single category within the domain, the most probable new category to be generated is located in {\em exactly} the same location and possesses an identical distributional structure.
 However, this is not simply an issue with their model but with a broader class of standard hierarchical Bayesian models \citep[][e.g., ]{griffithsscn07,kemp07}.

### 2017-06-09 Manuscripts/cog-psych/cogpsych.tex
after these changes, we are going to start WINNING again

@@
 To do so, we binned all possible hues into one of eight uniformly-spaced color groups: \{{\em Red}: $0-0.063$, $0.938-1$, {\em Yellow}: $0.063-0.188$, {\em Yellow-Green}: $0.188-0.313$, {\em Green-Teal}: $0.313-0.438$, {\em Teal}: $0.438-0.563$, {\em Teal-Blue}: $0.563-0.688$, {\em Purple}: $0.688-0.813$, {\em Pink}: $0.813-0.938$\}.
-In the \cite{jern2013probabilistic} experiment, the hue of each experimenter-defined category was selected from one of six possible values, each of which falls into one of the color groups above (two color groups was not used as a possible hue for the experimenter-defined categories).
+In the \cite{jern2013probabilistic} experiment, the hue of each experimenter-defined category was selected from one of six possible values, each of which falls into one of the color groups above (two color groups were not used as a possible hue for the experimenter-defined categories).
 By categorizing the participant-generated crystals likewise, we can obtain a broad measure of category contrast by determining the proportion of participant-generated crystals that fall into the same groups as the experimenter-defined categories: If contrast influences the hues of the generated categories, we should observe minimal overlap between in the color groupings.
@@
 While we did obtain similar results using a variety of alternative groupings, the hue dimension used in the \cite{jern2013probabilistic} study does not lend itself straightforwardly to the the computation of similarities, and thus we cannot be certain of whether our coding accurately approximates the psychological space of the stimuli.
-This precludes exemplar modeling of their data as it is necessary to encode exemplars in psychological space to determine the similarity between exemplars.
+This precludes traditional applications of categorization models to their data as it is usually necessary to encode objects in psychological space in order to accurately determine the similarity between objects.
 By consequence, although these results likely indicate that contrast exerts {\em some} influence, they do not precisely describe the nature of that influence.
@@
 It is also worth noting that PACKER represents just one of many possible models that incorporate contrast in category generation.
-For example, it may be possible to extend \cite{jern2013probabilistic}'s hierarchical Bayesian model to include contrast, this mechanism would be at odds with the standard hierarchical Bayesian framework.
+For example, it may be possible to extend \cite{jern2013probabilistic}'s hierarchical Bayesian model to include contrast, but this mechanism would be at odds with the standard hierarchical Bayesian framework.
 In contrast, these ideas emerge naturally from the exemplar view: We have not modified any of the core elements of the GCM in defining the PACKER model, we simply weight exemplar similarity during aggregation.
@@
 To begin our investigation, we sought to extend the early evidence obtained from our analysis of the \cite{jern2013probabilistic} data, under a variety of learning conditions and using more standard stimulus materials.
-We used an artificial stimulus design, two dimensional domain of squares, varying in color and size (see Figure \ref{fig:e1-conditions}, panel A).
+We used an artificial stimulus design: A two dimensional domain of squares, varying in color and size (see Figure \ref{fig:e1-conditions}, panel A).
 These dimensions have been used in numerous classification learning studies \cite[e.g.,][]{conaway2016similar,conaway2016generalization,shepard1961learning,nosofsky1994comparing}.
@@
 The task differs from traditional work in creative cognition primarily through the use of an artificial domain, which enables the application of computational models.
-Unlike most previous studies \citep[e.g.,][]{ward1994structured}, participants in our studies have no experience with the domain before the start of the experiment, and the experimenter-defined categories are not hierarchically structured (as are plants and animals).
+The use of an artificial domain also requires the addition of a training phase, during which participants learn about the categories in the domain.
+As a result,, unlike most previous studies \citep[e.g.,][]{ward1994structured}, participants in our studies have no experience with the domain before the start of the experiment, and the experimenter-defined categories are not hierarchically structured (as are plants and animals).
 
@@
 Thus, the core results of Experiment 1 indicate that generated categories can systematically differ from what we would expect based on prior work.
-The negative (or zero) correlations observed in the XOR condition suggests an interesting interaction between contrasting from a given category and emulating statistical properties.
+The negative (or null) correlations observed in the XOR condition suggests an interesting interaction between contrasting from a given category and emulating statistical properties.
 That is, the constraints on creative generation imposed by category contrast may not simply influence the {\em location} of generated categories, but also their distributional structure.
@@
 Due to the uneven number of fitted parameters among the models, we compare the model fits using the Akaike Information Criterion \citep[AIC;][]{akaike1974new}, where smaller values correspond to better fits (discounted by the number of parameters).
-The same qualitative results were obtained with alternative model comparison metrics (e.g., BIC, \citealp{schwarz1978estimating}; $AIC_C$, \citealp{hurvich1989regression})
+The same qualitative results were obtained with alternative model comparison metrics (e.g., BIC, \citealp{schwarz1978estimating}; $AIC_C$, \citealp{hurvich1989regression}).
 
@@
 Each panel shows, for each stimulus, the dimensional orientation of the categories it was generated into: vertically aligned `columns' (orange) versus horizontally aligned `rows' (purple).}
-    \nbcnote{Need to run this on lando with many more samples. current = 30}
     \label{fig:range-diff-gradients}
@@
 Once formalized in probabilistic terms, it should also be straightforward to incorporate distributional factors into the model.
-Alternatively, it may be possible to integrate the core principles of PACKER into other categorization models, such as SUSTAIN \citep{lovemg04} or DIVA \citep{}.
+Alternatively, it may be possible to integrate the core principles of PACKER into other categorization models \citep[e.g.,][]{love2004sustain,kurtz2015human,smith2000thirty}.
 
@@
 In our implementation of their model\footnote{Note that \citet{jern2013probabilistic}'s model is slightly different, as they used a semi-conjugate model.
-Their model acts very similar to our version of it and receives comparable fits.}, each category's exemplars are viewed as samples from a multivariate Normal distribution with parameters ($\mu, \Sigma$).
-Category covariance matrices (specifying variance and covariance along $k$-dimensions, are assumed to be Normal-Inverse-Wishart distributed with parameters: $\nu$ ($>k-1$), $\kappa$ ($>0$), and $\Sigma_D$.
+Their model acts very similarly to our version.}, each category's exemplars are viewed as samples from a multivariate Normal distribution with parameters ($\mu, \Sigma$).
+Category covariance matrices (specifying variance and covariance along $k$-dimensions), are assumed to be Normal-Inverse-Wishart distributed with parameters: $\nu$ ($>k-1$), $\kappa$ ($>0$), and $\Sigma_D$.
 $\nu$ and $\kappa$ are treated as free parameters in our simulations, and $\Sigma_D$ is the domain-wide covariance matrix from which all categories are viewed as samples.

### 2017-06-09 Manuscripts/cog-psych/cover-letter.tex
cover letter drafted; really tremendous work, outstanding

%!TEX output_directory = latex_out/

\documentclass[12pt]{article}
\usepackage[letterpaper, margin=1in, headheight=15pt]{geometry}

\setlength{\parindent}{0em}
\setlength{\parskip}{1em}

\begin{document}

We are pleased to submit to {\em Cognitive Psychology} a manuscript entitled ``Creating Something Different: Similarity and Contrast in Concept Generation.''

The manuscript reports two experiments and formal modeling on the topic of category generation. Previous work on the topic has not considered the integral role of contrast in ensuring generated categories differ from what is already known. Our behavioral experiments demonstrate such an influence, and our modeling efforts show that a novel exemplar-based approach (based on the Generalized Context Model) can explain the influence of category contrast in the generation process.

We expect that a broad range of readers will find our results noteworthy and compelling.  The work reported in the manuscript has not been submitted for journal publication. A previous version of this work is to appear in the Proceedings of the Thirty-Ninth Annual Conference of the Cognitive Science Society (Conaway \& Austerweil, 2017).

Thank you for considering our work for publication. Correspondence regarding the manuscript may be addressed to:

Joseph L. Austerweil \\
1202 West Johnson Street,\\ 
Madison, WI 53706.\\
E-mail: austerweil@wisc.edu

\end{document}

### 2017-06-16 Manuscripts/cog-psych/cogpsych.tex
added kjk passages, ready for joe's lookover

@@
 \subsection{Similarity and Contrast in Cognition}
+
 The proposal that category contrast is a primary constraint in creative generation is, in some ways, entirely commonsense.
@@
 PACKER's success in explaining generation within an artificial domain motivates future work exploring the nature of category contrast within a more naturalistic setting.
+
+\subsection{Implications For Categorization}
+
+Although our work has mainly addressed the processes involved in the creative generation of  concepts, in this paper we have studied these processes as they apply within an artificial categorization domain.
+By consequence, our findings also provide an advance in our understanding categorization more broadly.
+
+Categorization research addresses the representations and processes that underlie the learning and use of categories.
+Category learning tasks are generally about figuring out which items belong to which category.
+Once learned, categories are generally used to classify new stimuli and to make inferences beyond the available information.
+Our work is fairly unique in that people learn a single category through positive examples and then figure out another category that would make sense in the domain.
+In other words, only one category is provided to be learned and the use that is required of that category is to generate a contrast category.
+Such propagating of categories (producing one from another) may not seem like a fundamental or ubiquitous cognitive activity, but consider navigating a highly unfamiliar environment —  it would be practical, perhaps even critical, to generate expectations about what other kinds of things you are likely to encounter based on those that you already have.
+In the present studies, we have learned something about the form that such expectations are likely to take.
+
+We can think of the category generation task in our studies as asking a person to formulate an idea about what set of items in the domain are most interestingly not members of the original category.
+To meet this condition, the items must take some form of coherence that aligns with that of the original category and some form of distinctiveness relative to the original category.
+Reflecting the basic level of organization in natural categories, it makes sense to generate a set of items that possess strong within-category coherence (by importing or systematically transforming the internal structure of the original category) as well as strong between-category differentiation (by creating maximum contrast with the original category).
+In this sense one can see the patterns of performance in the category generation task as recapitulating the order of semantic organization.
 

### 2017-06-17 Manuscripts/cog-psych/cogpsych.tex
minor updates to manuscript and bibliography

@@
 In this paper, we study the generation of new categories using the computational and behavioral toolkit of traditional artificial category learning.
-Previous work in this domain has focused on how the statistical structure of known categories generalizes to generated categories, overlooking contrast between the known and generated categories as a factor.
+Previous work in this domain has focused on how the statistical structure of known categories generalizes to generated categories, overlooking the role of contrast between the known and generated categories as a factor.
 We report two experiments demonstrating that contrast between what is known and what is created is of fundamental importance for categorization and the creative process.
@@
 
-The creation of new ideas is one of the most fascinating, important, and difficult human capabilities to study.
-We tend to focus on the most salient products conceptual generation (e.g., scientific breakthroughs), but every person is likely to have generated many novel concepts in ordinary life.
+The creation of new ideas is one of the most fascinating and important human capabilities.
+Yet, it is also one of the most difficult human capabilities to study.
+We tend to focus on the most salient products of conceptual generation (e.g., scientific breakthroughs), but every person is likely to have generated many novel concepts in ordinary life.
 Indeed, the generation of novel sentences, works of art, and so on may also be considered products of our capacity to create new concepts.
@@
 Questions concerning the generation of novel concepts are typically addressed within research on creativity, a tremendously diverse field employing a multitude of experimental and theoretical approaches \citep[see][]{kaufman2010handbook}.
-The processes underlying conceptual generation are commonly investigated using a creative cognition approach \citep{finke1992creative, smith1995creative}, in which the products of creative acts (such as drawings) are analyzed to obtain insights into how they might have been created.
-These approaches have made great advances in our understanding of creativity, but the focus on particular types of creative products such as drawings makes difficult the task of formalizing proposed processes and representations within a computational modeling framework.
+The processes underlying conceptual generation are commonly investigated using the creative cognition approach \citep{finke1992creative, smith1995creative}, in which the products of creative acts (such as drawings) are analyzed to obtain insights into how they might have been created.
+These approaches have made great advances in our understanding of creativity, but the focus on particular types of creative products such as drawings makes it difficult to formalize the proposed processes and representations within a computational modeling framework.
 As a consequence, this approach does not often employ computational models to provide formal tests of competing theories.
@@
 This theory incorporates elements of the highly influential basic-level categories framework \citep{rosch1975cognitive,rosch1976basic}, as well as the exemplar view \citep{medin1978context,brooks1978nonanalytic}.
-While this work is been incredibly useful in providing a conceptual sketch of generation theories, the hand-drawn responses relied on in the experiments paradigms precludes the development of formal approaches.
+While this work is been incredibly useful in providing a conceptual sketch of generation theories, the hand-drawn responses used in the experiments paradigms precludes the development of formal approaches that can be used to test the theories in a fine-grained manner.
 
@@
 The copy-and-tweak model provided a poor account of their results, as the experiments devised by \citeauthor{jern2013probabilistic} were specifically designed to challenge it.
-However, its application is notable as a first step toward understanding creative cognition using well-known formal approaches from the categorization literature.
+However, its application is notable as a first step toward understanding theories developed in the creative cognition literature using well-known formal approaches from the categorization literature.
 
@@
 Prior work on category generation has explored only one of the possible constraints that guide category generation.
-The main concern of the published experiments has been on the distributional correspondences between learned and generated categories, and as a result most of the computational, theoretical, and empirical efforts have been towards explaining those effects.
+The main concern of the published experiments has been on the distributional correspondences between learned and generated categories, and as a result most of the computational, theoretical, and empirical efforts have been focused on explaining those effects.
 In this paper, we investigate another important constraint: category contrast.
 To generate a novel concept, individuals must produce something that is in some capacity {\em different} from what they already know.
-By consequence, contrast should be viewed a primary constraint on creative generation: new concepts must be different from existing ones.
+By consequence, we propose that contrast should be a primary constraint on creative generation: New concepts should be different from existing ones.
 
 Although it is evident that people are {\em capable} of creating new concepts and categories, it is not entirely clear how new concepts are systematically made different from what is already known.
-The hierarchical sampling model developed by \cite{jern2013probabilistic} assumes that differences between generated category are only due to random variation.
+The hierarchical Bayesian model developed by \cite{jern2013probabilistic} assumes that differences between observed and generated categories are only due to random variation.
 The model assumes that generated categories are sampled from the same underlying domain distribution as observed categories, and will thus share a common distributional structure.
-The authors do not make predictions about the {\em location} of the category within the domain (the perceptual instantiation of category members).
+The model does not make predictions about the {\em location} of the category within the domain (the perceptual instantiation of category members).
 Under a strict interpretation of their model, given knowledge of a single category within the domain, the most probable new category to be generated is located in {\em exactly} the same location and possesses an identical distributional structure.
-However, this is not simply an issue with their model but with a broader class of standard hierarchical Bayesian models \citep[][e.g., ]{griffithsscn07,kemp07}.
-Many of these models assume that at some point of the latent generative process the same underlying distribution generates all of the categories and thus, any differences between categories is due to {\em noise} and should not be {\em systematic}.
+However, this is not simply an issue with their model but with a broader class of standard hierarchical Bayesian models \citep[e.g., ][]{griffithsscn07,kemp07}.
+Many of these models assume that at some point of the latent generative process the same underlying distribution generates all of the categories and thus, any differences between categories are due to {\em noise} and should not be {\em systematic}.
 The best a standard hierarchical Bayesian model can do at capturing contrast is to assume that the new category is placed uniformly at random over stimulus space.
-This defeats the purpose of a hierarchy as it is ignored when determining a new category location!\footnote{It is plausible that some hierarchical Bayesian model could be created that generates categories different from each other.
-However, this model would not be a standard application or extension of most pre-existing hierarchical Bayesian models.}
+But, this defeats the purpose of a hierarchy as it is ignored when determining a new category location!\footnote{It is plausible that some hierarchical Bayesian model could be created that generates categories different from each other.
+However, this model would not be a standard application or extension of most pre-existing hierarchical Bayesian models.
+The generative process would need to include a component that promotes contrast.}
 
@@
 Taken alongside the analyses reported by \cite{jern2013probabilistic}, our analysis suggests that generated categories tend to be distinct from {\em and} distributionally similar to what is already known.
-However, it is worth noting that our analysis is still limited: the color groups defined above are somewhat arbitrary, and it is not clear that our color grouping is consistent with the psychological color boundaries perceived by participants.
+However, it is worth noting that our analysis is still limited: The color groups defined above are imprecise, and it is not clear that our color grouping is consistent with the psychological color boundaries perceived by participants.
 While we did obtain similar results using a variety of alternative groupings, the hue dimension used in the \cite{jern2013probabilistic} study does not lend itself straightforwardly to the the computation of similarities, and thus we cannot be certain of whether our coding accurately approximates the psychological space of the stimuli.
@@
 PACKER explains category generation as a balance between two fundamental constraints: The category to be generated should not be similar to known categories, and exemplars within each category should be similar to one another.
-These ideas are implemented within the well-studied exemplar framework -- the PACKER model is an extension of the influential Generalized Context Model of category learning \citep[GCM;][]{nosofsky1984choice,nosofsky1986attention}.
+These ideas are implemented within the well-studied exemplar framework -- the PACKER model is an extension of the influential Generalized Context Model of categorization \citep[GCM;][]{nosofsky1984choice,nosofsky1986attention}.
 
@@
 The choice to develop PACKER within an exemplar framework reflects the facts that exemplar models are well-studied, strongly theoretically motivated, and dominate much of the theoretical and empirical work in categorization.
-The focus of our work with PACKER concerns the dual constraints of within- and between-class similarity; it is not difficult to imagine how such constraints may be instantiated using alternative frameworks (e.g., \citealp{kurtz2007divergent,love2004sustain,smith2000thirty}; for a review, see \citealp{pothoswills2011}).
+The focus of our work with PACKER concerns the dual constraints of within- and between-class similarity; it is not difficult to imagine how such constraints may be instantiated using alternative frameworks (e.g., \citealp{kurtz2007divergent,love2004sustain,smith2000thirty}; for a review of categorization models, see \citealp{pothoswills2011}).
 
@@
 The $\gamma$ parameter thus specifies a wide breadth of possible approaches; by fitting it to a dataset, one can describe the relative roles of between-category contrast and within-category similarity in generation.
-See Figure \ref{fig:packer-examples} for an illustration of how $\gamma$ controls the relative influence of within category similarity and contrast to other categories.
+See Figure \ref{fig:packer-examples} for an illustration of how $\gamma$ controls the relative influence of within category similarity and contrast to other categories when generating a new exemplar.
 
@@
 
-The probability that a given candidate $y$ will be generated is evaluated using an Exponentiated \citet{luce1977choice} choice rule.
+The probability that a candidate $y$ will be generated is evaluated using an Exponentiated \citet{luce1977choice} choice rule.
 Candidates with greater values of $a(y,x)$ are more likely to be generated than candidates with smaller values:
@@
 In the sections below, we present two experiments demonstrating systematic effects of category contrast on creative generation inspired by factors influencing how PACKER generates new categories.
-Our experiments are based on \cite{jern2013probabilistic}'s paradigm, which is a straightforward translation of the traditional artificial classification paradigm to the task of category generation: participants are first exposed to a single, experimenter-defined category, and are then asked to generate members of a new category.
-We then report formal analyses comparing PACKER's account of our results to that of the hierarchical sampling and copy-and-tweak models developed by \cite{jern2013probabilistic}.
+Our experiments are based on \cite{jern2013probabilistic}'s paradigm, which is a straightforward translation of the traditional artificial classification paradigm to the task of category generation: Participants are first exposed to a single, experimenter-defined category, and are then asked to generate members of a new category.
+We then report formal analyses comparing PACKER's account of our results to that of the hierarchical Bayesian and copy-and-tweak models developed by \cite{jern2013probabilistic}.
 
@@
 
-Our experimental results also provide a converging test of the classic finding that generated categories tend to share distributional properties with known categories in the domain \citep[see][]{jern2013probabilistic,ward1994structured}.
+Our experimental results also provide a converging test of the classic finding that generated categories tend to share distributional properties with known categories in the domain \citep{jern2013probabilistic,ward1994structured}.
 From these results, we can predict that, in each condition, participants should generate categories that are distributionally similar to the experimenter-defined category: In the Cluster condition, generated categories should be tightly clustered.
@@
 The use of an artificial domain also requires the addition of a training phase, during which participants learn about the categories in the domain.
-As a result,, unlike most previous studies \citep[e.g.,][]{ward1994structured}, participants in our studies have no experience with the domain before the start of the experiment, and the experimenter-defined categories are not hierarchically structured (as are plants and animals).
+As a result, unlike most previous studies \citep[e.g.,][]{ward1994structured}, participants in our studies have no experience with the domain before the start of the experiment, and the experimenter-defined categories are not hierarchically structured (as are plants and animals).
 
@@
 For each generated category, we computed the category range along each axis (X, Y), as well as the correlation between features.
-These data, shown in Figure \ref{fig:e1-statsboxes}, reveal broad individual differences: within each condition, participants generated categories spanning the entire X- and Y- axis as well as categories that spanned very little along each.
+These data, shown in Figure \ref{fig:e1-statsboxes}, reveal broad individual differences: Within each condition, participants generated categories spanning the entire X- and Y- axis as well as categories that spanned very little along each.
 Likewise, in each condition participants generated categories possessing strongly positive, neutral, and strongly negative correlations between the dimensions.
@@
 We believe that the failure to replicate this finding is because participants in \cite{jern2013probabilistic} could differentiate the generated category on a third dimension (hue) to maintain the statistical regularities on the other two dimensions.
-Although the correlation in the XOR condition is significantly negative, it is clear from the box-plot in Figure ~\ref{fig:e1-statsboxes} that it would be inappropriate to make a strong conclusion (e.g., the median is close to zero).
+Although the correlation in the XOR condition is significantly negative, it is clear from the box-plot in Figure~\ref{fig:e1-statsboxes} that it would be inappropriate to make a strong conclusion (e.g., the median is close to zero).
 However, we can conclude with confidence that there are situations where people do not emulate the distributional structure of the given category.
@@
 We found strong evidence for effects of category contrast in each condition: Participants were more likely to generate stimuli that are more distant from (i.e., less similar to) members of a previously-learned category, and members of participant-generated categories tended to be more similar to one another than to members of previously-learned categories.
-We also partially replicated the classic finding that the distributional structure of generated categories reflects that of previously learned categories \citep{jern2013probabilistic,ward1994structured}: members of generated categories were more widely distributed along dimensions which were widely distributed in the experimenter-defined category.
+We also partially replicated the classic finding that the distributional structure of generated categories reflects that of previously learned categories \citep{jern2013probabilistic,ward1994structured}: Members of generated categories were more widely distributed along dimensions which were widely distributed in the experimenter-defined category.
 
@@
 We created two new category types (depicted in Figure \ref{fig:e2-conditions}) that possess identical distributional structure (both are tight clusters of examples with no correlation between features), and only differ in their Y-axis position: the `Bottom' category lies in the bottom-center of the space, and the `Middle' category lies in the center.
-The distributional equality of these conditions is key to the design of the experiment: If the structure of previously learned categories were the only influence on the structure of generated categories, we should observe no difference in the categories participants generate between these two conditions.
+The distributional equality of these conditions is key to the design of the experiment: If the distributional structure of previously learned categories were the only influence on the generated categories, we should observe no difference in the categories participants generate between these two conditions.
 Will participants distribute their generated category differently between conditions due only to the differences in the available stimulus space for generating a new category?
@@
 61 participants were randomly assigned to the Middle and Bottom conditions each.
-The stimuli and procedure were exactly as in Experiment 1.
+The stimulus space and procedure were exactly as in Experiment 1.
 Participants first completed a short training phase, followed by the generation phase.
@@
 The comparison between the hierarchical Bayesian model and PACKER is meant to emphasize the necessity of contrast and demonstrate that generation cannot be explained entirely through the emulation of distributional structure.
+Each model has complementary strengths and weaknesses.
 Whereas PACKER is insensitive to the distributional structure of learned categories (relying only on within- and between-category similarity), the hierarchical Bayesian model generates categories exclusively on the basis of knowledge of how existing classes are distributed.
@@
 Table \ref{table:global-model-fits} contains the results of this fitting procedure.
-Due to the uneven number of fitted parameters among the models, we compare the model fits using the Akaike Information Criterion \citep[AIC;][]{akaike1974new}, where smaller values correspond to better fits (discounted by the number of parameters).
+Due to the uneven number of fitted parameters among the models, we compare the model fits using the Akaike Information Criterion \citep[AIC;][]{akaike1974new}, where smaller values correspond to better fits (discounted by model complexity as measured by the number of parameters).
 The same qualitative results were obtained with alternative model comparison metrics (e.g., BIC, \citealp{schwarz1978estimating}; $AIC_C$, \citealp{hurvich1989regression}).
@@
 As observed in Figures \ref{fig:e1-samples} and \ref{fig:e2-samples}, the generated categories we observed were often widely distributed, with no items near the category prototype.
-This aspect of the data is inconsistent with the multivariate normal distributions used to represent categories in the \cite{jern2013probabilistic} model, but can be handled easily using an exemplar-based approach.
+This aspect of the data is inconsistent using the multivariate normal distributions as prototypes to represent categories in the \cite{jern2013probabilistic} model, but can be easily accounted for using an exemplar-based approach.
 
@@
 Thus, it is possible that the failure of these models is simply due to their inability to explain each participant's first trial (generating the first item in the `Beta' category).
-We conducted an identical set of simulations as above, excluding this trial (leaving 915 responses in the dataset): Again, PACKER ($L = -3390$, $AIC = 6786$) achieved better fits than the copy-and-tweak ($L = -3579$, $AIC = 7162$) and hierarchical Bayesian ($L=-3612$, $AIC = 7232$) models.
+To ensure this is not driving our results, we conducted an identical set of simulations as above, excluding this trial (leaving 915 responses in the dataset): Again, PACKER ($L = -3390$, $AIC = 6786$) achieved better fits than the copy-and-tweak ($L = -3579$, $AIC = 7162$) and hierarchical Bayesian ($L=-3612$, $AIC = 7232$) models.
 
@@
 Through comparison with the copy-and-tweak model, Figure \ref{fig:packer-loglike} more clearly demonstrates the robustness of the explanatory gains yielded by PACKER's category contrast mechanism.
-It displays the log-likelihood of the participants' results under PACKER as a function of $\gamma$ parameter.
+It displays the log-likelihood of the participants' results under PACKER as a function of the $\gamma$ parameter.
 The model's other parameters ($c$, $\theta$) were set according to copy-and-tweak's best fits from Table \ref{table:global-model-fits}, and thus when $\gamma=1$, the models are equivalent.
 The figure clearly shows a ``sweet spot'':  a convex region in which PACKER achieves superior fits as a result of changes to $\gamma$.
-The best fitting values lie well below the value of 1 assumed by the copy-and-tweak model, which demonstrates the robustness of the contrast effect (though note PACKER achieves even better fits when its other parameters are fitted, as in Table \ref{table:global-model-fits}).
+The best fitting values lie well below the value of 1 assumed by the copy-and-tweak model, which demonstrates the robustness of the contrast effect (though note PACKER achieves even better fits when its parameters are fitted together, as in Table \ref{table:global-model-fits}).
 Notably, however, the copy-and-tweak parameterization ($\gamma=1$) performs better than the `contrast-only' parameterization of the model ($\gamma = 0$).
@@
 As noted above, while all three models make clear claims about the internal structure of generated categories, the copy-and-tweak and hierarchical Bayesian models do not make any claims about how generated categories should differ from what is already known.
-However, as we observed in Experiment 2, the distributional structure of a category is not always independent of its location within the domain.
+However, as we observed in the results of Experiment 2, the distributional structure of a category is not always independent of its location within the domain.
 To demonstrate this point in more depth, we computed the X- and Y- axis ranges of every participant-generated category.
@@
 Within this approach, the resulting aggregation is a trade-off between the number of generations and the strength of the range difference within each generated category.
-Infrequently generated stimuli, as well as those with mixed positive and negative scores, are given neutral difference scores\footnote{Prior to plotting, data was also processed using a Gaussian filter with $\sigma = 0.8$}.
+Infrequently generated stimuli, as well as those with mixed positive and negative scores, are given neutral difference scores.\footnote{Prior to plotting, data was also processed using a Gaussian filter with $\sigma = 0.8$.}
 
@@
 These patterns of performance clearly depict the interdependence between the distributional structure and location of generated concepts.
-Our results can be interpreted in terms of local minimization of between-category similarity: by distributing the generated category away from members of the experimenter-defined category, participants may increase the degree of between-category distance without drastically altering the degree of within-category similarity.
+Our results can be interpreted in terms of local minimization of between-category similarity: By distributing the generated category away from members of the experimenter-defined category, participants may increase the degree of between-category distance without drastically altering the degree of within-category similarity.
 
@@
 % 
-where $\theta_w$ is a free parameter controlling how differences in range correspond to differences in weights (functioning similarity to the $\theta$ parameter in each of the models), and $\text{range}(k)$ is the range of examples generated by the participant along dimension $k$.
+where $\theta_w$ is a free parameter controlling how differences in range correspond to differences in weights (functioning similarly to the $\theta$ parameter in each of the models), and $\text{range}(k)$ is the range of examples generated by the participant along dimension $k$.
 We used $\theta_w = 1.5$ in our simulations, though the results are robust and similar for other $\theta_w$ values.
@@
 
-As in the more traditional model evaluation procedure described above, PACKER provided a much closer match to our behavioral results than the copy-and-tweak and hierarchical Bayesian models.
-In all conditions, PACKER distributes categories similarly to the behavioral data: horizontally-aligned categories tend to be placed above and below members of the experimenter-defined category, and vertically-aligned categories tend to be placed to the sides.
+As in the more traditional model evaluation analysis described above, PACKER provided a much closer match to our behavioral results than the copy-and-tweak and hierarchical Bayesian models.
+In all conditions, PACKER distributes categories similarly to the behavioral data: Horizontally-aligned categories tend to be placed above and below members of the experimenter-defined category, and vertically-aligned categories tend to be placed to the sides.
 Conversely, because the copy-and-tweak and hierarchical Bayesian models are insensitive to category contrast, these models do not produce any systematic patterns of association between category location and distributional structure.
@@
 We developed a novel, exemplar-based model, PACKER, which formally specifies the role of contrast in generation.
-The model proposes that categories are represented as exemplars in a multidimensional psychological space, and generation is constrained both by within-category and between-category similarity: exemplars belonging to the same category should be similar to one another, and exemplars belonging to different categories should not be similar to one another.
+The model proposes that categories are represented as exemplars in a multidimensional psychological space, and generation is constrained both by within-category and between-category similarity: Exemplars belonging to the same category should be similar to one another, and exemplars belonging to different categories should not be similar to one another.
 
@@
 Most notably, a leading account within the creative cognition literature, the Path of Least Resistance \citep{ward1994structured,ward1995s}, also explains generation in terms of an exemplar-based retrieval process.
-This account was designed to explain the creative generation of natural categories (e.g., new species of plants and animals) and as a result relies strongly on the hierarchical organization of these categories: individuals are thought to retrieve an example of the higher-level category being generated (e.g., \textit{bird} may be retrieved from the category \textit{animal}), and then systematically alter what was retrieved to make something new.
-As the PACKER model does not assume knowledge is hierarchically organized \citep[this is true of the exemplar view more broadly, see][]{murphy2016exemplar}, the model may be viewed as a formal instantiation of the Path of Least Resistance for application in a traditional artificial categorization domain (in which there is no established hierarchy of categories).
+This account was designed to explain the creative generation of natural categories (e.g., new species of plants and animals) and as a result relies strongly on the hierarchical organization of these categories: Individuals are thought to retrieve an example of the higher-level category being generated (e.g., \textit{bird} may be retrieved from the category \textit{animal}), and then systematically alter what was retrieved to make something new.
+As the PACKER model does not assume knowledge is hierarchically organized \citep[this is true of the exemplar view more broadly, see][]{murphy2016exemplar}, the model may be viewed as a formal instantiation of the Path of Least Resistance for application in a traditional artificial categorization domain (when there is no established hierarchy of categories).
 PACKER's success in explaining generation within an artificial domain motivates future work exploring the nature of category contrast within a more naturalistic setting.
@@
 \begin{equation}
-p(y) = \dfrac
+p(y \mid x) = \dfrac
 {\exp \left \{ \theta \cdot {\rm Normal}(y; \mu, \Sigma) \right \} }

### 2017-06-17 Manuscripts/cog-psych/cover-letter.tex
cover letter updates -- ready for you to submit once you look it over. if your changes are only minor, no need for me to look at the article or cover letter again.

@@
-%!TEX output_directory = latex_out/
+\documentclass{letter}
 
-\documentclass[12pt]{article}
+\setlength{\topmargin}{-0.8in}
+\setlength{\textheight}{8in}
+\setlength{\oddsidemargin}{0.2in}
+\setlength{\textwidth}{6.1in}
+
 \usepackage[letterpaper, margin=1in, headheight=15pt]{geometry}
@@
 
+\address{Joseph L. Austerweil \\ Department of Psychology \\ University of Wisconsin - Madison \\ 1202 West Johnson Street,\\ Madison, WI 53706.\\ E-mail: austerweil@wisc.edu}
+
+\signature{Joseph L. Austerweil \\ Assistant Professor of Psychology and Computer Science (affiliate) \\ University of Wisconsin - Madison}
+
 \begin{document}
+\date{June 17 2017}
+\begin{letter}{Dr. Gordon Logan \\ Editor \\ {\em Cognitive Psychology}}
 
-We are pleased to submit to {\em Cognitive Psychology} a manuscript entitled ``Creating Something Different: Similarity and Contrast in Concept Generation.''
+\vspace{30mm}
 
-The manuscript reports two experiments and formal modeling on the topic of category generation.
-Previous work on the topic has not considered the integral role of contrast in ensuring generated categories differ from what is already known.
+\opening{Dear Dr. Logan,}
+
+Nolan Conaway, Kenneth Kurtz, and I are pleased to submit to {\em Cognitive Psychology} a manuscript entitled ``Creating Something Different: Similarity and Contrast in Concept Generation.''
+
+The manuscript proposes and investigates a novel principle underlying how people generate categories: contrast (being different from other relevant categories).
+To formulate the principle formally, we present a novel exemplar-based approach (based on the Generalized Context Model) and test its predictions of how contrast should influence category generation in two experiments.
+Previous work on the topic has not considered the fundamental role of contrast in ensuring generated categories differ from what is already known.
+This includes an article recently published in {\em Cognitive Psychology} by Alan Jern and Charles Kemp, entitled "A Probabilistic Account of exemplar and category generation".
 Our behavioral experiments demonstrate such an influence, and our modeling efforts show that a novel exemplar-based approach (based on the Generalized Context Model) can explain the influence of category contrast in the generation process.
+We demonstrate that previous models of category generation are not influenced by contrast and are unable to account for the results of our experiments.
 
-We expect that a broad range of readers will find our results noteworthy and compelling.
+Creating new categories is a fundamental task tackled by thinking.
+Yet, outside of the previously mentioned article, there are no formal investigations of the topic.
+Further, to the best of our knowledge, we are the first authors to propose and explore the importance of contrast in category generation.
+These are novel theoretical, computational, and empirical advances that have important implications for researchers studying human thinking.
+Thus, we expect that a broad range of {\em Cognitive Psychology} readers will find our manuscript relevant.
 The work reported in the manuscript has not been submitted for journal publication.
@@
 
-Thank you for considering our work for publication. Correspondence regarding the manuscript may be addressed to:
+Given his expertise in thinking and categorization, we recommend Associate Editor {\em Art Markman} as the action editor for the manuscript.
+Thank you for considering our work for publication.
+Correspondence regarding the manuscript may be addressed to Joseph L.
+Austerweil whose contact information is above.
 
-Joseph L. Austerweil \\
-1202 West Johnson Street,\\ 
-Madison, WI 53706.\\
-E-mail: austerweil@wisc.edu
+\closing{Yours sincerely,}
+
+\end{letter}
 

### 2017-06-17 Manuscripts/cog-psych/cogpsych.tex
nolan's final comb completed

@@
 In this paper, we study the generation of new categories using the computational and behavioral toolkit of traditional artificial category learning.
-Previous work in this domain has focused on how the statistical structure of known categories generalizes to generated categories, overlooking the role of contrast between the known and generated categories as a factor.
+Previous work in this domain has focused on how the statistical structure of known categories generalizes to generated categories, overlooking the degree of contrast between the known and generated categories as a factor.
 We report two experiments demonstrating that contrast between what is known and what is created is of fundamental importance for categorization and the creative process.
@@
 The core proposal is that participants generate new items by copying stored examples from memory and tweaking them to generate something new.
-The copy-and-tweak model differs from the path of least resistance account in that it notably omits the hierarchical organization of categories, as well as selectivity in which features are changed \citep[both of which are factors in the Path of Least Resistance account;][]{ward2002role}.
+The copy-and-tweak model differs from the Path of Least Resistance account in that it notably omits the hierarchical organization of categories, as well as selectivity in which features are changed \citep[both of which are factors in the Path of Least Resistance account;][]{ward2002role}.
 Instead, their copy-and-tweak model corresponds to a direct exemplar-similarity approach \citep[e.g.,][]{nosofsky1984choice,nosofsky1986attention}, generating new items according to their similarity to known members of the target category.
@@
 Under a strict interpretation of their model, given knowledge of a single category within the domain, the most probable new category to be generated is located in {\em exactly} the same location and possesses an identical distributional structure.
-However, this is not simply an issue with their model but with a broader class of standard hierarchical Bayesian models \citep[e.g., ][]{griffithsscn07,kemp07}.
+This is not an issue with their model specifically, but with a broader class of standard hierarchical Bayesian models \citep[e.g., ][]{griffithsscn07,kemp07}.
 Many of these models assume that at some point of the latent generative process the same underlying distribution generates all of the categories and thus, any differences between categories are due to {\em noise} and should not be {\em systematic}.
@@
 In their simulations, the model was only tested on generation after the learner had been exposed to members of the target category, and so the model's ability to generate a new category from scratch was not evaluated.
-However, the model's generation is based exclusively on similarity to known members of the target category; when there are no members of the target category, generation is presumably random.
+However, the model's generation is based exclusively on similarity to known members of the {\em target} category; when there are no members of the target category, generation is presumably random.
 
@@
 Thus, effects of category contrast should be observable within the experimental results of these studies.
-To provide a test of the influence of category contrast in creative generation, we conducted a novel analysis of Experiment 3 from \cite{jern2013probabilistic}\footnote{Although \cite{jern2013probabilistic} reported four experiments, we focus on Experiment 3 because their first two experiments tested generation of items into known categories, and their fourth experiment was identical to Experiment 3 but with a restricted generation space.}.
+To provide a test of the influence of category contrast within existing data, we conducted a novel analysis of Experiment 3 from \cite{jern2013probabilistic}\footnote{Although \cite{jern2013probabilistic} reported four experiments, we focus on Experiment 3 because their first two experiments tested generation of items into known categories, and their fourth experiment was identical to Experiment 3 but with a restricted generation space.}.
 
@@
 Further, they did not calibrate their monitor, and so we cannot know the precise colors presented to participants.
-As \cite{jern2013probabilistic} were interested in relations between the saturation and length of examples in generated into novel categories, this is not an issue for their analyses and results.
+As \cite{jern2013probabilistic} were interested in relations between the saturation and length of examples in generated into novel categories, these issues do not undercut their analyses and results.
 However, these issues pose a significant challenge to evaluating contrast between the experimenter-defined and participant-generated categories along the hue dimension.
@@
 
-Re-analyzing the results of \cite{jern2013probabilistic} provides preliminary support that contrast plays a role in category generation.
+Re-analyzing the results from \cite{jern2013probabilistic} provides preliminary support that contrast plays a role in category generation.
 Taken alongside the analyses reported by \cite{jern2013probabilistic}, our analysis suggests that generated categories tend to be distinct from {\em and} distributionally similar to what is already known.
@@
 Although, as an exemplar model, one of PACKER's proposals is people represent categories in terms of a collection of stored exemplars, it is worth emphasizing that the underlying nature of human category representations is not the primary concern of our work.
-The choice to develop PACKER within an exemplar framework reflects the facts that exemplar models are well-studied, strongly theoretically motivated, and dominate much of the theoretical and empirical work in categorization.
+The choice to develop PACKER within an exemplar framework reflects the facts that exemplar models have been throughly evaluated, are strongly theoretically motivated, and dominate much of the theoretical and empirical work in categorization.
 The focus of our work with PACKER concerns the dual constraints of within- and between-class similarity; it is not difficult to imagine how such constraints may be instantiated using alternative frameworks (e.g., \citealp{kurtz2007divergent,love2004sustain,smith2000thirty}; for a review of categorization models, see \citealp{pothoswills2011}).
@@
 
-PACKER (as well as its name) was in part inspired by earlier work from the categorization literature \citep{hidaka2011packing}.
-They argued that natural categories ``pack'' the values of features such that different categories fill the space with distance between one another, while maintaining items within the same category close together.
+PACKER (as well as its name) was in part inspired by earlier work from the categorization literature.
+Specifically, \cite{hidaka2011packing} argued that natural categories ``pack'' the values of features such that different categories fill the domain space with distance between one another, while keeping items within the same category close together.
 Inspired by this idea, PACKER proposes that generation is constrained by both similarity to members of the target category (the category in which a stimulus is being generated) as well as similarity to members of other categories: the most desirable generation candidates are similar to members of the target category and not similar to members of contrast categories.
@@
 These dimensions have been used in numerous classification learning studies \cite[e.g.,][]{conaway2016similar,conaway2016generalization,shepard1961learning,nosofsky1994comparing}.
-Unlike those used in the \cite{jern2013probabilistic} experiments, distance on the physical dimensions aligns more directly with perceptual similarity, allowing us to evaluate the role of category contrast in creative generation more precisely.
+Unlike those used in the \cite{jern2013probabilistic} experiments, distance on these physical dimensions aligns more directly with perceptual similarity, allowing us to evaluate the role of category contrast in creative generation more precisely.
 To extend the evidence provided by the \cite{jern2013probabilistic} data, we tested the effects of category contrast after learning one category from a set of qualitatively distinct category structures, as shown in Figure \ref{fig:e1-conditions}.
 
-Figure \ref{fig:e1-conditions}b-d show the values of exemplar dimensions belonging to the experimenter-defined categories (`A', or `Alpha') that participants were assigned to learn about prior to generating a new category.
+Figures \ref{fig:e1-conditions}b-d show the values of exemplar dimensions belonging to the experimenter-defined categories (`A', or `Alpha') that participants were assigned to learn about prior to generating a new category.
 Each participant learned one of the category types during training.
@@
 The use of an artificial domain also requires the addition of a training phase, during which participants learn about the categories in the domain.
-As a result, unlike most previous studies \citep[e.g.,][]{ward1994structured}, participants in our studies have no experience with the domain before the start of the experiment, and the experimenter-defined categories are not hierarchically structured (as are plants and animals).
+As a result, unlike most previous studies \citep[e.g.,][]{ward1994structured}, participants in our studies have no experience with the domain before the start of the experiment, and the experimenter-defined categories are not hierarchically structured (as are many natural categories).
 
@@
 These data, shown in Figure \ref{fig:e1-distanceplots}, reveal a clear pattern: Examples that are more distant from members of the experimenter-defined categories are more likely to be generated into a new category.
-This supports contrast being a fundamental constraint on creative cognition and statistical regularity alone being insufficient.
+This supports the notion that contrast is a fundamental constraint on creative cognition and that suggests that statistical regularity alone is insufficient.
+
+Figure \ref{fig:e1-distanceplots} also depicts, for each participant, the average distance of members within the generated category ({\em within-category} distance) against the average distance between members of the generated and experimenter-defined category ({\em between-category} distance).
+The narrow distribution of between-category distances in the XOR condition reflects the widely distributed nature of the experimenter-defined category, reducing the possible distances to members of the participant-generated category.
+These data reveal a systematic pattern: The majority of participants generated categories with greater between-category distance than within-category distance.
+That is, members of the generated category tended to be more similar to one another than to members of the experimenter-defined category.
+To evaluate this claim quantitatively, we conducted t-tests comparing the amount of within- and between- class distance in each condition.
+All conditions possessed greater between-category distance: Cluster, $t(63) = 11.43$, $p < 0.001$; Row, $t(60) = 13.16$, $p < 0.001$; and XOR, $t(57) = 3.64$, $p < 0.001$.
+These results provide further evidence of an effect of category contrast: Participants prefer to generate categories that are dissimilar to the learned category but maintain some level of internal cohesion.
 
@@
 
-Figure \ref{fig:e1-distanceplots} also depicts, for each participant, the average distance of members within the generated category ({\em within-category} distance) against the average distance between members of the generated and experimenter-defined category ({\em between-category} distance).
-The narrow distribution of between-category distances in the XOR condition reflects the widely distributed nature of the experimenter-defined category, reducing the possible distances to the experimenter-defined category.
-These data reveal a systematic pattern: The majority of participants generated categories with greater between-category distance than within-category distance.
-That is, members of the generated category tended to be more similar to one another than to members of the experimenter-defined category.
-To evaluate this claim quantitatively, we conducted t-tests comparing the amount of within- and between- class distance in each condition.
-All conditions possessed greater between-category distance: Cluster, $t(63) = 11.43$, $p < 0.001$; Row, $t(60) = 13.16$, $p < 0.001$; and XOR, $t(57) = 3.64$, $p < 0.001$.
-These results provide further evidence of an effect of category contrast: Participants prefer to generate categories that are dissimilar to the learned category but maintain some level of internal cohesion.
-
 A secondary goal of this experiment was to examine whether we replicate the classic result that generated categories often possess the same distributional properties as previously-known categories.
@@
 Categories generated in the XOR condition were also less widely distributed along the X-axis compared to Row, $t(117) = 2.56$, $p = 0.012$.
-This latter effect was not expected because the experimenter-defined category for XOR and Row had similar X-ranges.
+This latter effect was not expected because the experimenter-defined categories for XOR and Row had similar X-ranges.
 However, the key finding is that categories from the Cluster condition tended to be more tightly distributed along the X-axis.
-
-Likewise, categories generated in the Row condition had less Y-axis range compared to Cluster, $t(123) = 4.57$, $p < 0.001$ and XOR, $t(117) = 9.26$, $p < 0.001$, and categories from the Cluster condition had less Y-axis range compared to XOR, $t(120) = 3.95$, $p < 0.001$.
-As expected, the correlations in the Cluster and Row conditions were not systematically positive or negative ($ps > 0.1$).
-However, the generated categories in the XOR condition tended to possess {\em negatively} correlated dimensions, $t(57) = 2.04$, $p = 0.046$.
-This finding is notable, as it is the opposite of what would be expected, assuming learners are emulating the distributional structure of the experimenter-defined class (which possesses perfectly positively correlated features).
-We believe that the failure to replicate this finding is because participants in \cite{jern2013probabilistic} could differentiate the generated category on a third dimension (hue) to maintain the statistical regularities on the other two dimensions.
-Although the correlation in the XOR condition is significantly negative, it is clear from the box-plot in Figure~\ref{fig:e1-statsboxes} that it would be inappropriate to make a strong conclusion (e.g., the median is close to zero).
-However, we can conclude with confidence that there are situations where people do not emulate the distributional structure of the given category.
-This indicates that there is more to category generation than the distributional structure of other categories in the domain.
-Further, as we will discuss in more detail in the model-based analysis section, this is expected by our proposal that contrast is a fundamental principle in category generation.
 
@@
 
+Likewise, categories generated in the Row condition had less Y-axis range compared to Cluster, $t(123) = 4.57$, $p < 0.001$ and XOR, $t(117) = 9.26$, $p < 0.001$, and categories from the Cluster condition had less Y-axis range compared to XOR, $t(120) = 3.95$, $p < 0.001$.
+As expected, the correlations in the Cluster and Row conditions were not systematically positive or negative ($ps > 0.1$).
+However, the generated categories in the XOR condition tended to possess {\em negatively} correlated dimensions, $t(57) = 2.04$, $p = 0.046$.
+This finding is notable, as it is the opposite of what would be expected, assuming learners are emulating the distributional structure of the experimenter-defined class (which possesses perfectly positively correlated features).
+
+We believe that the failure to replicate this finding is because participants in \cite{jern2013probabilistic} could differentiate the generated category on a third dimension (hue) to maintain the statistical regularities on the other two dimensions.
+Although the correlation in the XOR condition is significantly negative, it is clear from the box-plot in Figure~\ref{fig:e1-statsboxes} that it would be inappropriate to make a strong conclusion (e.g., the median is close to zero).
+However, we can conclude with confidence that there are situations where people do not emulate the distributional structure of the given category.
+This indicates that there is more to category generation than the emulation of distributional structure of other categories in the domain.
+Further, as we will discuss in more detail in the model-based analysis section, this is expected by our proposal that contrast is a fundamental principle in category generation.
+
 \subsection{Discussion}
@@
 Thus, the core results of Experiment 1 indicate that generated categories can systematically differ from what we would expect based on prior work.
-The negative (or null) correlations observed in the XOR condition suggests an interesting interaction between contrasting from a given category and emulating statistical properties.
+The negative (or null) correlations observed in the XOR condition suggests an interesting interaction between contrast with a given category and emulation of statistical properties.
 That is, the constraints on creative generation imposed by category contrast may not simply influence the {\em location} of generated categories, but also their distributional structure.
@@
 
-\section{Experiment 2}
-
-To test whether category contrast influences the distributional structure of generated categories, we sought to identify conditions in which differences in the distributional structure of generated categories cannot be explained by the distributional structure of the experimenter-defined category.
-We created two new category types (depicted in Figure \ref{fig:e2-conditions}) that possess identical distributional structure (both are tight clusters of examples with no correlation between features), and only differ in their Y-axis position: the `Bottom' category lies in the bottom-center of the space, and the `Middle' category lies in the center.
-The distributional equality of these conditions is key to the design of the experiment: If the distributional structure of previously learned categories were the only influence on the generated categories, we should observe no difference in the categories participants generate between these two conditions.
-Will participants distribute their generated category differently between conditions due only to the differences in the available stimulus space for generating a new category?
+\section{Experiment 2.}
 
@@
 
-If category contrast influences the distributional structure of the categories people generate, then we should observe different types of categories according to the shape of the space that is unoccupied by members of previously learned categories.
+To test whether category contrast influences the distributional structure of generated categories, we sought to identify conditions in which differences in the distributional structure of generated categories cannot be explained by the distributional structure of the experimenter-defined category.
+We created two new category types (depicted in Figure \ref{fig:e2-conditions}) that possess identical distributional structures (both are tight clusters of examples with no correlation between features), as they only differ in their Y-axis position: the `Bottom' category lies in the bottom-center of the space, and the `Middle' category lies in the center.
+The distributional equality of these conditions is key to the design of the experiment: If the distributional structure of previously learned categories were the only influence on the generated categories, we should observe no difference in the categories participants generate between these two conditions.
+Will participants distribute their generated category differently between conditions due to the differences in the available stimulus space for generating a new category?
+
+If category contrast influences the distributional structure of the categories people generate, then we should observe different types of categories according to the shape of the space that is {\em unoccupied} by members of previously learned categories.
 The difference in the Y-axis position between the Bottom and Middle conditions produces a considerable change to the shape of the unoccupied space.
@@
 
-\begin{figure}
-    \begin{center}
-    \inputpgf{figs/}{e2-samples.pgf}
-    \caption{Sample categories generated in Experiment 2. }
-    \label{fig:e2-samples}
-    \end{center}
-\end{figure}
-
 \subsection{Participants, Materials, and Procedure}
@@
 \subsection{Results}
+
+\begin{figure}
+    \begin{center}
+    \inputpgf{figs/}{e2-samples.pgf}
+    \caption{Sample categories generated in Experiment 2. }
+    \label{fig:e2-samples}
+    \end{center}
+\end{figure}
 
@@
 
-However, our ability to detect differences using a standard $t$-test between the conditions is, in this case, diminished due to the non-normality of the data (Shapiro-Wilk normality test $W=0.77, p< 0.001$ for the Middle condition and $W=0.85, p < 0.001$ for the Bottom condition).
+However, our ability to detect differences in Y-Axis range using a standard $t$-test between the conditions is, in this case, diminished due to the non-normality of the data (Shapiro-Wilk normality test $W=0.77, p< 0.001$ for the Middle condition and $W=0.85, p < 0.001$ for the Bottom condition).
 Figure \ref{fig:e2-yranges} depicts the Y-axis position of the exemplars generated within each participant's category.
@@
 Because our main prediction concerns the generation of exemplars within the upper and lower regions of the domain, we compared the conditions in terms of the frequency with which participants generated examples above and below the categories.
-Specifically, we counted the number of participants in each condition who placed at least one `Beta' exemplar on the top and bottom `rows' of the space (the maximum and minimum possible y-axis value, respectively).
+Specifically, we counted the number of participants in each condition who placed at least one `Beta' exemplar on the top and bottom `rows' of the space (the maximum and minimum possible Y-axis value, respectively).
 The resulting contingencies data are shown in Table \ref{table:e2-subset-table}.
@@
 The latter effect is of interest here, as it indicates that the shape of the unoccupied space exerts some influence on the distributional structure of generated categories: Participants in the Middle condition were more likely to generate a category spanning the entire Y-axis.
-Thus, distributional structure of the generated categories can be influenced by category contrast alone because the distributional structure of the given categories did not differ between groups.
+Thus, distributional structure of the generated categories can be influenced by category contrast alone.
 
@@
 The comparison between the hierarchical Bayesian model and PACKER is meant to emphasize the necessity of contrast and demonstrate that generation cannot be explained entirely through the emulation of distributional structure.
-Each model has complementary strengths and weaknesses.
-Whereas PACKER is insensitive to the distributional structure of learned categories (relying only on within- and between-category similarity), the hierarchical Bayesian model generates categories exclusively on the basis of knowledge of how existing classes are distributed.
+Each model has complementary strengths and weaknesses: Whereas PACKER is insensitive to the distributional structure of learned categories (relying only on within- and between-category similarity), the hierarchical Bayesian model generates categories exclusively on the basis of knowledge of how existing classes are distributed.
 
@@
 As observed in Figures \ref{fig:e1-samples} and \ref{fig:e2-samples}, the generated categories we observed were often widely distributed, with no items near the category prototype.
-This aspect of the data is inconsistent using the multivariate normal distributions as prototypes to represent categories in the \cite{jern2013probabilistic} model, but can be easily accounted for using an exemplar-based approach.
+This aspect of the data is inconsistent with the multivariate normal distributions (similar to prototypes) used to represent categories in the \cite{jern2013probabilistic} model, but can be easily accounted for using an exemplar-based approach.
 
@@
 Under these circumstances, there are no examples to copy, and thus the copy-and-tweak model predicts that items are generated at random.
-Likewise, with no observations on which to condition the mean of the category distribution, the hierarchical Bayesian model also picks an item at random.
+Likewise, with no observations on which to condition the  category distribution, the hierarchical Bayesian model also picks an item at random.
 Thus, it is possible that the failure of these models is simply due to their inability to explain each participant's first trial (generating the first item in the `Beta' category).
@@
 For example, if all of attention is allocated along the X-axis ($w_1=1$ and $w_2=0$), similarity along the Y-axis no longer influences performance.
-As a result, PACKER will create categories that are more widely distributed along the Y-axis in this case, as similarity is not taken into account along that dimension.
+As a result, PACKER will create categories that are more widely distributed along the Y-axis, as similarity is not taken into account along that dimension.
 As a general principle, differentially weighting one dimension will result in the generation of categories that are more widely distributed along the ignored dimension, conforming to a row- or column-like arrangement.
@@
 
-However, because many stimuli were infrequently generated (such items near members of the `Alpha' category), we cannot simply compute the empirical average, as infrequently generated stimuli would be likely to show artificially strong differences.
+However, because many stimuli were infrequently generated (such items near members of the `Alpha' category), we cannot simply compute the empirical average of the difference scores, as infrequently generated stimuli would be likely to show artificially strong differences.
 Instead, we used a Bayesian analysis to estimate the mean $\mu_x$ on the assumption that the scores $x$ for each stimulus are normally distributed with an unknown mean and unknown standard deviation.
@@
 Within this approach, the resulting aggregation is a trade-off between the number of generations and the strength of the range difference within each generated category.
-Infrequently generated stimuli, as well as those with mixed positive and negative scores, are given neutral difference scores.\footnote{Prior to plotting, data was also processed using a Gaussian filter with $\sigma = 0.8$.}
+Infrequently generated stimuli, as well as those with mixed positive and negative scores, are given neutral difference scores.
 
@@
 
-The results of our analysis are shown in Figure \ref{fig:range-diff-gradients} for the experiment and model results.
+The results of our analysis are shown in Figure \ref{fig:range-diff-gradients} for the experiment and model results\footnote{Prior to plotting, data was also processed using a Gaussian filter with $\sigma = 0.8$.}.
 The left-most column of Figure \ref{fig:range-diff-gradients} displays the effect of category location and contrast on the distributional structure of the category generated by participants.
-These data reveal strong and consistent patterns across all the conditions we tested in Experiments 1 and 2: generated categories are more tightly distributed along the axis in which they are distinct.
+These data reveal strong and consistent patterns across all the conditions we tested in Experiments 1 and 2: Generated categories are more tightly distributed along the axis in which they are distinct.
 For example, in the `Cluster' condition, exemplars in the bottom-left of the space are more often generated into vertically aligned categories, and exemplars in the top-right are more often generated into horizontally aligned categories.
@@
 
-To explore how well PACKER, copy-and-tweak- and hierarchical Bayesian models explain our findings we conducted simulations using an individual-differences approach.
+To explore how well the PACKER, copy-and-tweak, and hierarchical Bayesian models explain our findings, we conducted simulations using an individual-differences approach.
 As noted in Section \ref{section:individual-diffs}, row- and column-like categories can be produced by each model through changes to the weighting of each dimension.
@@
 Thus, a different transformation is appropriate for incorporating the weights computed in Equation \ref{eq:range-weight}.
-For the hierarchical Bayesian model, we computed the dimensional variances according to: $\lambda (1-w_k)d$, where $\lambda$ is a free parameter specifying the overall assumed variance of the domain, and $d$ is the number of dimensions\footnote{This calculation applies only in two-dimensional domains, where $w_2 = 1-w_1$.}.
+For the hierarchical Bayesian model, we computed the dimensional variances according to: $\lambda (1-w_k) 2$, where $\lambda$ is a free parameter specifying the overall assumed variance of the domain, and $2$ corresponds to the number of dimensions in our experiments\footnote{This calculation applies only in two-dimensional domains, where $w_2 = 1-w_1$.}.
 Under this approach, evenly distributed weights correspond to an assumed variance of $\lambda$.
@@
 Creative generation is most typically studied using a creative cognition approach \citep{finke1992creative,smith1995creative}, where creative products \citep[such as drawings of alien plants and animals;][]{ward1994structured} are analyzed to obtain insight into the processes and representations involved.
-While this approach has yielded a great deal of information about the nature of creative generation, the commonly employed response modalities (e.g., hand-drawn pictures) made it difficult to develop computational models capable of formally evaluating the proposed representations and processes.
+While this approach has yielded a great deal of information about the nature of creative generation, the commonly employed response modalities (e.g., hand-drawn pictures) make it difficult to develop computational models capable of formally evaluating the proposed representations and processes.
 However, the generation of novel concepts is not altogether different from other behaviors studied in cognitive psychology: In particular, it can be viewed as an form of inference using one's knowledge about existing categories \citep{kemp2014taxonomy,kurtz2015human}.
@@
 In other words, only one category is provided to be learned and the use that is required of that category is to generate a contrast category.
-Such propagating of categories (producing one from another) may not seem like a fundamental or ubiquitous cognitive activity, but consider navigating a highly unfamiliar environment —  it would be practical, perhaps even critical, to generate expectations about what other kinds of things you are likely to encounter based on those that you already have.
+Such propagating of categories (producing one from another) may not seem like a fundamental or ubiquitous cognitive activity, but consider navigating a highly unfamiliar environment -- it would be practical, perhaps even critical, to generate expectations about what other kinds of things you are likely to encounter based on those that you already have.
 In the present studies, we have learned something about the form that such expectations are likely to take.
 
-We can think of the category generation task in our studies as asking a person to formulate an idea about what set of items in the domain are most interestingly not members of the original category.
+We can think of the category generation task in our studies as asking a person to formulate an idea about what set of items in the domain are most interestingly {\em not} members of the original category.
 To meet this condition, the items must take some form of coherence that aligns with that of the original category and some form of distinctiveness relative to the original category.
@@
 This suggests that category generation is a fundamentally different computational-level problem \citep[different from those posed by][]{jern2013probabilistic,kemp2014taxonomy}.
+
 Characterizing that problem and conducting a rational analysis is an important direction for future research.
@@
 Once formalized in probabilistic terms, it should also be straightforward to incorporate distributional factors into the model.
-Alternatively, it may be possible to integrate the core principles of PACKER into other categorization models \citep[e.g.,][]{love2004sustain,kurtz2015human,smith2000thirty}.
-
-Finally, although an overall goal of this work is to advance the study of creative thought by explaining generation through the empirical and quantitative approaches from research in categorization, these approaches are not easily aligned with many of the alternative theoretical viewpoints in the broader study of creativity \citep[for a review see][]{kozbelt2010theories}, such as those based on free association \citep{mednick1962associative} and conceptual combination \citep{estes2002emergence,murphy1988comprehending}.
+Alternatively, it may be possible to integrate the core principles of PACKER into other categorization models \citep[e.g.,][]{love2004sustain,kurtz2007divergent,smith2000thirty}.
+
+Finally, although an overall goal of this work is to advance the study of creative thought by explaining generation through empirical and quantitative approaches from research in categorization, these approaches are not easily aligned with many of the alternative theoretical viewpoints in the broader study of creativity \citep[for a review see][]{kozbelt2010theories}, such as those based on free association \citep{mednick1962associative} and conceptual combination \citep{estes2002emergence,murphy1988comprehending}.
 Instead, we reduced our focus by studying a highly complex behavior (creative generation) as it applies within a well-established domain (artificial category learning).
@@
 \cite{jern2013probabilistic} demonstrated how a hierarchical Bayesian model could explain the distributional correspondences between observed and generated categories.
-In their model, exemplars of generated category were viewed as samples from a multivariate Normal distribution over the dimensions of stimulus space.
+In their model, exemplars of generated categories were viewed as samples from a multivariate Normal distribution over the dimensions of stimulus space.
 The mean of the generated category was independent of the observed categories, but the covariance matrix (encoding feature variances and correlations) was based on a common prior distribution.

### 2017-06-18 Manuscripts/cog-psych/cover-letter.tex
checked over the cover letter (_very_ fancy)

@@
+%!TEX output_directory = latex_out/
+
 \documentclass{letter}
@@
 
-The manuscript proposes and investigates a novel principle underlying how people generate categories: contrast (being different from other relevant categories).
-To formulate the principle formally, we present a novel exemplar-based approach (based on the Generalized Context Model) and test its predictions of how contrast should influence category generation in two experiments.
+The manuscript proposes and investigates a novel principle underlying how people generate novel concepts: contrast (being different from other relevant categories).
+To formulate the principle mathematically, we present a novel exemplar-based approach (based on the Generalized Context Model) and test its predictions regarding how contrast should influence category generation in two experiments.
 Previous work on the topic has not considered the fundamental role of contrast in ensuring generated categories differ from what is already known.
-This includes an article recently published in {\em Cognitive Psychology} by Alan Jern and Charles Kemp, entitled "A Probabilistic Account of exemplar and category generation".
-Our behavioral experiments demonstrate such an influence, and our modeling efforts show that a novel exemplar-based approach (based on the Generalized Context Model) can explain the influence of category contrast in the generation process.
+This includes an article recently published in {\em Cognitive Psychology} by Alan Jern and Charles Kemp, entitled ``A Probabilistic Account of exemplar and category generation.''
+
+Our behavioral experiments demonstrate such an influence, and our modeling efforts show that our novel exemplar-based approach can explain the influence of category contrast in the generation process.
 We demonstrate that previous models of category generation are not influenced by contrast and are unable to account for the results of our experiments.

### 2018-04-06 Presentations/mathpsych_2018/abstract.tex
PACKER and CopyTweak now using classic Luce instead of Exponentiated Luce. Fixed it with a variable baseline similarity to avoid the issue of impossible probabilities when tradeoff goes towards .5 . However, fits do not look good - tradeoff values for all fits are more than .9 . In other news, updated some readmes, and also wrote up some useful stuff for running scripts on CHTC.

Research into categorization has primarily focused on how people learn
categories and extend category membership to novel exemplars. However, it is
unclear how people generate new categories. In previous work, we presented
PACKER (\emph{Producing Alike and Contrasting Knowledge using Exemplar
Representations}), an exemplar model that accounts for category generation by
measuring not only the similarity of an exemplar with a target category but also
its dissimilarity with contrast categories. As predicted by PACKER, we found
that people generated new categories that contrast from a given category. In
this work, we examine whether category contrast also plays a role in category
learning. Specifically, we conducted a novel experiment where participants
learned categories generated by participants in our previous work. We
demonstrate how category contrast allows PACKER to explain empirical data in
both category learning and category generation paradigms better than competing
theories, providing evidence for category contrast as a fundamental factor in
categorization.

### 2018-04-06 Presentations/mathpsych_2018/main.tex
PACKER and CopyTweak now using classic Luce instead of Exponentiated Luce. Fixed it with a variable baseline similarity to avoid the issue of impossible probabilities when tradeoff goes towards .5 . However, fits do not look good - tradeoff values for all fits are more than .9 . In other news, updated some readmes, and also wrote up some useful stuff for running scripts on CHTC.

%%
\documentclass[man,12pt]{apa6}
\title{Learning generated categories: Applying PACKER to category assignments}
\shorttitle{PACKER in category learning}
\fourauthors{Shi Xian Liew}{Nolan Conaway}{Kenneth Kurtz}{Joseph Austerweil}
\fouraffiliations{University of Wisconsin -- Madison}{Shutterstock}{Binghamton University}{University of Wisconsin -- Madison}

%Turn off ligatures for easy copypasta
\usepackage{fontspec}
\setmainfont[Ligatures={NoRequired,NoCommon,NoContextual}]{Times New Roman}

\leftheader{Beitzel}

\abstract{\input{abstract}}

%\keywords{APA style, demonstration}

\begin{document}
\maketitle
\end{document}

### 2018-04-06 Experiments/cat-assign/analysis/compile.py

@@ -27,6 +27,7 @@ c.close()
 # been matched - ignore subsequent matches.
 data = []
 matchList = [];
+stimList = [];
 for i, row in workerInfo.iterrows():
 	
 	if not row.Complete: continue
@@ -51,6 +52,12 @@ for i, row in workerInfo.iterrows():
         # If pptmatch has already occurred, don't add data
         if pdata['info']['pptmatch'] in matchList: continue
         matchList.append(pdata['info']['pptmatch'])
+        # if the data configuration is already present, also don't add
+        stim = pdata['info']['stimuli']
+        stim.sort()
+        if stim in stimList: continue
+        stimList.append(stim)
+
 	data.append(json.loads(S))
 
 # create participant table

### 2018-04-06 Experiments/cogpsych-code/chtc_code/run_python.sh

@@ -1,0 +1,25 @@
+#!/bin/bash
+pwd
+ls
+# untar your Python installation
+tar -xzf python.tar.gz
+tar -xzf working.tar.gz
+# make sure the script will use your Python installation, 
+# and the working directory as it's home location
+export PATH=$(pwd)/python/bin:$PATH
+mkdir home
+export HOME=$(pwd)/home
+# run your script
+cp -r chtc/ generating-categories
+cd generating-categories/Experiments/cogpsych-code/
+python global_model_gridsearch_CHTC.py $1
+
+
+#tar the pickles
+tar -czvf pickles$1.tar.gz pickles/
+
+#Move it to main directory
+cd ~
+cd ..
+mv generating-categories/Experiments/cogpsych-code/pickles$1.tar.gz pickles$1.tar.gz
+ls

### 2018-04-06 Experiments/cogpsych-code/matlabtests/SimplePacker.m

@@ -53,7 +53,7 @@ else
     categories = [reshape(categories,nStimTrain-1,1);1];
 end
 
-categories
+
 
 pset = nan(nStimTest,numel(tradeoff));
 distanceset = cell(1,numel(tradeoff));

### 2018-04-06 Experiments/cogpsych-code/matlabtests/getData.m

@@ -1,0 +1,71 @@
+function [data_k,data_total,data_p,nstim_axes,stimTrainIdxAll,stimTestIdx,task,p2El] = getData(data)
+
+switch data
+    case 'nosofsky1986'
+        %Nosofsky1986
+        assignCat1raw = load('n86cat1.txt'); %ordered as stimuli, then conditions, then participants
+        assignCat2raw = load('n86cat2.txt');
+        task = 'assign';
+        %p_observed = assignCat1/(assignCat1 + assignCat2);
+        
+        %pool data across participants
+        assignCat1pPpt = reshape(assignCat1raw,64,2);%cat 1 assignments per participant
+        assignCat1Pool = sum(assignCat1pPpt,2);
+        assignCat1 = reshape(assignCat1Pool,16,4)'; %note the transposition
+        
+        assignCat2pPpt = reshape(assignCat2raw,64,2);%cat 2 assignments per participant
+        assignCat2Pool = sum(assignCat2pPpt,2);
+        assignCat2 = reshape(assignCat2Pool,16,4)'; %note the transposition
+        %Define stimulus space
+        nstim_axes = [4,4]; %This tells us that the stimulis are organised as a 4 by 4 array
+        nstim = prod(nstim_axes);
+
+%         categoriesSet = repmat([ones(1,4),ones(1,4)*2],nconditions,1);
+        
+        stimTrainIdxAll = [
+            %|-------cat 1-------|   |-------cat 2-------|
+            0     3     5     6     9    10    12    15 %dimensional
+            3     6     9    12     0     5    10    15 %crisscross
+            5     6     9    10     2     4    11    13 %intext
+            2     5     8    12     3     7    10    13] ; %diagonal
+        stimTrainIdxAll = stimTrainIdxAll + 1; %Adjust indexing from 1 (instead of 0)
+        stimTestIdx = 1:nstim; %test entire stimulus space
+
+        data_k = assignCat1;
+        data_total = assignCat1+assignCat2;
+        data_p = data_k./data_total;
+        p2El = 16;
+%         stimCoords = ndspace(4,2);%entire stimulus space
+    case 'NGPMG1994'
+        nstim_axes = [2,2,2]; %This tells us that the stimuli are organised as a 4 by 4 array
+        nstim = prod(nstim_axes);
+        ntrials = 200;
+        task = 'error';
[…]

### 2018-04-06 Experiments/cogpsych-code/matlabtests/testModels (Awesomeness's conflicted copy 2018-03-27).m

@@ -1,0 +1,160 @@
+%Load data
+dataSet = {'nosofsky1986','NGPMG1994'};
+data = 1;
+[data_k,data_total,data_p,nstim_axes] = getData(dataSet{data});
+%Nosofsky1986
+% assignCat1raw = load('n86cat1.txt'); %ordered as stimuli, then conditions, then participants
+% assignCat2raw = load('n86cat2.txt');
+% task = 'assign';
+% %p_observed = assignCat1/(assignCat1 + assignCat2);
+% 
+% %pool data across participants
+% assignCat1pPpt = reshape(assignCat1raw,64,2);%cat 1 assignments per participant
+% assignCat1Pool = sum(assignCat1pPpt,2); 
+% assignCat1 = reshape(assignCat1Pool,16,4)'; %note the transposition
+% 
+% assignCat2pPpt = reshape(assignCat2raw,64,2);%cat 2 assignments per participant
+% assignCat2Pool = sum(assignCat2pPpt,2); 
+% assignCat2 = reshape(assignCat2Pool,16,4)'; %note the transposition
+% 
+% %Define stimulus space
+% nstim_axes = [4,4]; %This tells us that the stimulis are organised as a 4 by 4 array
+% nstim = prod(nstim_axes);
+% 
+% 
+% nconditions = 4;
+% stimIdx = 1:nstim;
+% categoriesSet = repmat([ones(1,4),ones(1,4)*2],nconditions,1);
+% 
+% stimTrainIdxAll = [     
+%    %|-------cat 1-------|   |-------cat 2-------|
+%      0     3     5     6     9    10    12    15 %dimensional
+%      3     6     9    12     0     5    10    15 %crisscross
+%      5     6     9    10     2     4    11    13 %intext
+%      2     5     8    12     3     7    10    13] ; %diagonal
+% stimTrainIdxAll = stimTrainIdxAll + 1; %Adjust indexing from 1 (instead of 0)
+% 
+% 
+% data_k = assignCat1;
+% data_total = assignCat1+assignCat2;
+% data_p = data_k./data_total;
+nstim = prod(nstim_axes);
+stimCoords = ndspace(nstim_axes(1),numel(nstim_axes));%entire stimulus space
+stimTestIdx = 1:nstim; %test entire stimulus space
+normsteep_toggle = true;
+
[…]

### 2018-04-06 Experiments/cogpsych-code/matlabtests/testModels.m

@@ -1,56 +1,36 @@
 %Load data
+dataSet = {'nosofsky1986','NGPMG1994'};
+data = 1;
+[data_k,data_total,data_p,nstim_axes,stimTrainIdxAll,stimTestIdx,task,p2El] = getData(dataSet{data});
 
-%Nosofsky1986
-assignCat1raw = load('n86cat1.txt'); %ordered as stimuli, then conditions, then participants
-assignCat2raw = load('n86cat2.txt');
-task = 'assign';
-%p_observed = assignCat1/(assignCat1 + assignCat2);
-
-%pool data across participants
-assignCat1pPpt = reshape(assignCat1raw,64,2);%cat 1 assignments per participant
-assignCat1Pool = sum(assignCat1pPpt,2); 
-assignCat1 = reshape(assignCat1Pool,16,4)'; %note the transposition
-
-assignCat2pPpt = reshape(assignCat2raw,64,2);%cat 2 assignments per participant
-assignCat2Pool = sum(assignCat2pPpt,2); 
-assignCat2 = reshape(assignCat2Pool,16,4)'; %note the transposition
-
-%Define stimulus space
-nstim_axes = [4,4]; %This tells us that the stimulis are organised as a 4 by 4 array
-nstim = prod(nstim_axes);
-
-
-nconditions = 4;
-stimIdx = 1:nstim;
+nconditions = size(stimTrainIdxAll,1);
 categoriesSet = repmat([ones(1,4),ones(1,4)*2],nconditions,1);
-
-stimTrainIdxAll = [     
-   %|-------cat 1-------|   |-------cat 2-------|
-     0     3     5     6     9    10    12    15 %dimensional
-     3     6     9    12     0     5    10    15 %crisscross
-     5     6     9    10     2     4    11    13 %intext
-     2     5     8    12     3     7    10    13] ; %diagonal
-stimTrainIdxAll = stimTrainIdxAll + 1; %Adjust indexing from 1 (instead of 0)
-
-
-data_k = assignCat1;
-data_total = assignCat1+assignCat2;
-data_p = data_k./data_total;
-stimCoords = ndspace(4,2);%entire stimulus space
-stimTestIdx = 1:nstim; %test entire stimulus space
+nstim = prod(nstim_axes);
+stimCoords = ndspace(nstim_axes(1),numel(nstim_axes));%entire stimulus space
+% task = 'assign';
[…]

### 2018-04-06 Experiments/cogpsych-code/plot-stats-gradients.py

@@ -83,6 +83,8 @@ for model_name, model_obj in name_2_object.items():
         else:
             params['wts'] = np.array([0.5, 0.5])
 
+        # assume no baselinesim
+        params['baselinesim'] = 0
         # simulate
         model = model_obj([As], params)
         for j in range(N_SAMPLES):   
@@ -93,8 +95,11 @@ for model_name, model_obj in name_2_object.items():
             all_stats = funcs.stats_battery(stimuli[nums,:], As)
 
             # convert to row for df
-            rows = dict(condition = [pcond] *4, stimulus = nums)
-            rows[STAT_OF_INTEREST] = [all_stats[STAT_OF_INTEREST]]*4
+            # handle variable number of generated exemplars
+            numgen = len(nums)
+            rows = dict(condition = [pcond] *numgen, stimulus = nums)
+            rows[STAT_OF_INTEREST] = [all_stats[STAT_OF_INTEREST]]*numgen
+            
             model_data = model_data.append(pd.DataFrame(rows), ignore_index = True)
 
         print '\t' + str(i)

### 2018-04-06 Experiments/cogpsych-code/plot_heat_temp.py

@@ -1,0 +1,131 @@
+#Temporary file to generate ps plots
+#Gives a good idea of the distribution of generation probabilities at each step og exemplar generation
+#given some parameter and alpha stimuli values
+import pickle, math
+import pandas as pd
+import sqlite3
+execfile('Imports.py')
+import Modules.Funcs as funcs
+from Modules.Classes import Simulation
+from Modules.Classes import CopyTweak
+from Modules.Classes import Packer
+from Modules.Classes import ConjugateJK13
+from scipy.stats.stats import pearsonr
+import numpy as np
+import matplotlib.pyplot as plt
+import seaborn as sns
+sns.set_style("whitegrid")
+#Close figures
+plt.close()
+
+#load data
+dataname_def = 'pooled'#'nosofsky1986'#'NGPMG1994'
+participant_def = 204 #'all'
+unique_trials_def = 'all'
+dataname = dataname_def
+execfile('validate_data.py')
+
+# get data from pickle
+with open(pickledir+src, "rb" ) as f:
+    trials = pickle.load( f )
+trials.task = task
+
+pptTrialObj = Simulation.extractPptData(trials,participant_def,unique_trials_def)
+#Get best parms
+with open(pickledir+'gs_'+dst, "rb" ) as f:
+    best_params_t = pickle.load( f )
+#Rebuild it into a smaller dict
+best_params = dict()
+for modelname in best_params_t.keys():    
+    best_params[modelname] = dict()
+    for i,parmname in enumerate(best_params_t[modelname]['parmnames']):
+        parmval = best_params_t[modelname]['bestparmsll']
+        best_params[modelname][parmname] = parmval[i]
+
+#paramsT = dict(params)
+models = [Packer,CopyTweak,ConjugateJK13]
+#paramsP = dict(determinism = 2,specificity=.5,tradeoff=.5,wts=[.5,.5])
+#paramsCT = dict(paramsP)
+#paramsCT.pop('tradeoff')
+
+paramsP = best_params[Packer.model]
+paramsP['baselinesim'] = 0
+#paramsP['wts'] = paramsT['wts']
+paramsCT = best_params[CopyTweak.model]
+paramsCT['baselinesim'] = 0
+#paramsCT['wts'] = paramsT['wts']
+paramsJK = best_params[ConjugateJK13.model]
+paramSet = [paramsP,paramsCT,paramsJK]
+STAT_LIMS =  (-1.0, 1.0)
+
[…]

### 2018-04-06 Experiments/cogpsych-code/plot_modelppt_ease.py

@@ -1,0 +1,254 @@
+#Get the loglikelihoods of a given data set as a measure of how easy the model
+#can generate that dataset
+import pickle, math
+import pandas as pd
+import sqlite3
+execfile('Imports.py')
+import Modules.Funcs as funcs
+from Modules.Classes import Simulation
+from Modules.Classes import CopyTweak
+from Modules.Classes import Packer
+from Modules.Classes import ConjugateJK13
+from scipy.stats.stats import pearsonr
+import numpy as np
+import matplotlib.pyplot as plt
+import seaborn as sns
+sns.set_style("whitegrid")
+
+#Specify simulation values
+N_SAMPLES = 10000
+WT_THETA = 1.5
+MIN_LL = 1e-10
+
+# Specify default dataname
+dataname_def = 'midbot'#'nosofsky1986'#'NGPMG1994'
+participant_def = 'all'
+unique_trials_def = 'all'
+dataname = dataname_def
+execfile('validate_data.py')
+# get data from pickle
+with open(pickledir+src, "rb" ) as f:
+	trials = pickle.load( f )
+
+# get best params pickle
+with open("pickles/gs_best_params_all_data_e1_e2.p", "rb" ) as f:
+    best_params_t = pickle.load( f )
+#Rebuild it into a smaller dict
+best_params = dict()
+for modelname in best_params_t.keys():    
+    best_params[modelname] = dict()
+    for i,parmname in enumerate(best_params_t[modelname]['parmnames']):
+        parmval = best_params_t[modelname]['bestparmsll']
+        best_params[modelname][parmname] = parmval[i]
+modelList = [ConjugateJK13,CopyTweak,Packer]                            
+
+#Prepare matched database    
+matchdb='../cat-assign/data_utilities/cmp_midbot.db'
+        
+unique_trials = 'all'
+trials.task = task
+
+#Get learning data
+data_assign_file = '../cat-assign/data/experiment.db'
+con = sqlite3.connect(data_assign_file)
+info = pd.read_sql_query("SELECT * from participants", con)
+assignment = pd.read_sql_query("SELECT * FROM assignment", con)
+stimuli = pd.read_sql_query("SELECT * from stimuli", con).as_matrix()
+con.close()
+#Get generation data
+data_generate_file = 'experiment-midbot.db'
+con = sqlite3.connect(data_generate_file)
+stats = pd.read_sql_query("SELECT * from betastats", con)
+con.close()
+
[…]

### 2018-04-06 Experiments/cogpsych-code/temp.py

@@ -49,7 +49,7 @@ unique_trials = 'all'
 trials.task = task
 
 #Get learning data
-data_assign_file = '../cat-assign/data/experiment.db'
+data_assign_file = '../cat-assign/data/experiment260318.db'
 con = sqlite3.connect(data_assign_file)
 info = pd.read_sql_query("SELECT * from participants", con)
 assignment = pd.read_sql_query("SELECT * FROM assignment", con)
@@ -62,6 +62,9 @@ stats = pd.read_sql_query("SELECT * from betastats", con)
 con.close()
 
 #Get unique ppts
+# pptlist = []
+# for i,row in info.iterrows():
+#         pptlist += [row['pptmatch']]
 pptlist = np.array([]);
 for trial in trials.Set:    
     pptlist = np.concatenate((pptlist,trial['participant']))
@@ -73,7 +76,7 @@ modeleaseDB = "pickles/modelease_all_data_e1_e2.p"
 try:
     with open(modeleaseDB, "rb" ) as f:
         ll_global = pickle.load( f )
-        ll_loadSuccess = True
+        ll_loadSuccess = False #True
 except:
     ll_global = dict()
     ll_loadSuccess = False
@@ -207,9 +210,8 @@ for model_obj in modelList:
             ppt  = row.participant
             pptAssign = assignment.loc[assignment['participant']==ppt].sort_values('trial')
             nTrials = len(pptAssign)
-            error = [];
             accuracyEl = float(sum(pptAssign.correctcat == pptAssign.response))/nTrials
-            error.append(1-accuracyEl)
+
 
             pptmatch = row.pptmatch
             #Prepare to plot configuration

### 2018-04-12 Experiments/cogpsych-code/global_model_gridsearch_CHTC.py
New PACKER and CopyTweak implemented with gamma (tradeoff) as a multiplier for the contrast-category members only. Lets see how thje fits go! Still need to optimise the chtc code - there should be at least one more commit before sending this off to fit.

@@ -1,4 +1,4 @@
-import pickle, sys
+import pickle, sys, os
 import pandas as pd
 import numpy as np
 import time
@@ -17,6 +17,7 @@ unique_trials_def = 'all'
 nchunks = 1000 #number of CHTC instances to run
 #Allow for input arguments at the shell
 narg = len(sys.argv)
+
 if __name__ == "__main__" and narg>1:
 
         # if len(sys.argv)<4:
@@ -40,10 +41,18 @@ else:
         participant = participant_def
         unique_trials = unique_trials_def
         runchunk = 93;
-datasets = ['pooled','pooled-no1st','nosofsky1986','nosofsky1989','NGPMG1994']
+datasets = ['pooled','pooled-no1st','nosofsky1986','nosofsky1989','NGPMG1994']        
+
+
+#Check that output directory exists, otherwise create it
+pickledir = 'pickles/'
+outputdir = pickledir + 'newpickles/'
+if not os.path.isdir(outputdir):
+        os.system('mkdir ' + outputdir)
 
 for dataname in datasets:
     execfile('validate_data.py')
+    
     #add chunk number to dst
     dst = dst[0:-2] + '_chunk' + str(runchunk) + '.p'
     dst_error = dst[0:-2] + '_error.p'
@@ -180,7 +189,7 @@ for dataname in datasets:
         
 
     # save final result in pickle
-    with open(pickledir+'chtc_gs_'+dst,'wb') as f:
+    with open(outputdir + 'chtc_gs_'+dst,'wb') as f:
         #pass 
         pickle.dump(results, f)
 

### 2018-04-12 Experiments/cogpsych-code/plot_packer_loglike_t.py

@@ -22,7 +22,7 @@ with open("pickles/best_params_all_data_e1_e2.p", "rb" ) as f:
 start_params = best_params[CopyTweak.model]
 
 # set up grid of param values
-gamma_grid = np.linspace(0, 1.0, 100)
+gamma_grid = np.linspace(0, 1.50, 100)
 loglikes = np.empty(gamma_grid.shape)
 
 # add task type to trials object
@@ -32,6 +32,7 @@ trials.task = 'generate'
 for i, val in enumerate(gamma_grid):
 	curr = start_params.copy()
 	curr['tradeoff'] = val
+        curr = Packer.parmxform(curr, direction = 1)
 	loglikes[i] = -trials.loglike(curr, Packer)
 
 
@@ -42,14 +43,14 @@ for i in range(len(gamma_grid)):
 fh = plt.figure(figsize=(5,3))
 
 # copytweak annotation
-copytweak_loglike = loglikes[gamma_grid == 1]
+copytweak_loglike = loglikes[gamma_grid == 0]
 plt.plot([min(gamma_grid), max(gamma_grid)], [copytweak_loglike, copytweak_loglike], '--', linewidth = 1, color='gray')
 plt.text(0.5, copytweak_loglike, 'Target Only (Copy \& Tweak)', ha = 'center', va = 'bottom')
 
 # contrast only annotation
-contrast_loglike = loglikes[gamma_grid == 0]
-plt.plot([min(gamma_grid), max(gamma_grid)], [contrast_loglike, contrast_loglike], '--', linewidth = 1, color='gray')
-plt.text(0.5, contrast_loglike, 'Contrast Only', ha = 'center', va = 'bottom')
+contrast_loglike = loglikes[gamma_grid == 2]
+#plt.plot([min(gamma_grid), max(gamma_grid)], [contrast_loglike, contrast_loglike], '--', linewidth = 1, color='gray')
+#plt.text(0.5, contrast_loglike, 'Contrast Only', ha = 'center', va = 'bottom')
 
 # PACKER annotation
 x = float(gamma_grid[loglikes == max(loglikes)])
@@ -64,7 +65,7 @@ xticks = np.arange(min(gamma_grid),max(gamma_grid)+0.2,0.2)
 xticklabels = [str(i) for i in xticks]
 xticklabels[0] = '0'
 xticklabels[-1] = '1'
-plt.xticks(xticks)
[…]

### 2018-04-12 Experiments/cogpsych-code/analyse_chtc.py
Temp.

@@ -14,11 +14,14 @@ from Modules.Classes import ConjugateJK13
 
 
 tardir = 'chtctar/'
-maintarname = 'allpickles050418.tar.gz'
+privdir = os.path.join(tardir,'private') #private working folder that git ignores
+if not os.path.isdir(privdir):
+    os.system('mkdir {}'.format(privdir))
+maintarname = 'allpickles110418.tar.gz'
 appendkey = ['finalparmsll','startparms','chunkstartparms']
 removekey = ['bestparmsll','chunkidx']
 #Go through each tarball and find the chtc file
-allfiles = os.listdir(tardir)
+#allfiles = os.listdir(tardir)
 data = dict()
 datasetsAll = []
 modelsAll = []
@@ -33,7 +36,9 @@ for maintarmember in maintar.getmembers(): #for checkfile in allfiles:
         #Print chunk currently processing
         print_ct = funcs.printProg(tarct,print_ct,steps = 20, breakline = 40)
         tarct += 1
-        tar = tarfile.open(tardir+filename, "r:gz")
+        #extract file
+        maintar.extract(maintarmember,privdir)
+        tar = tarfile.open(os.path.join(privdir,filename), "r:gz")
         for member in tar.getmembers():            
             checkmembername = re.search('chtc_gs_best_params_(.*)_chunk{}\.p'.format(chunk),member.name)            
             if checkmembername != None:
@@ -59,7 +64,9 @@ for maintarmember in maintar.getmembers(): #for checkfile in allfiles:
                         #Otherwise, just append the stuff as ndicated by appendkey
                         for key in appendkey:
                             data[dataset][model][key] = np.concatenate((data[dataset][model][key], datachunk[model][key]),0)
-
+        #Remove file from private folder after use
+        os.system('rm {}'.format(os.path.join(privdir,filename)))
+        
 #Find best parms for each model in each dataset
 pickledir = 'pickles/'
 for dataset in datasetsAll:
@@ -81,7 +88,7 @@ for dataset in datasetsAll:
     # save final result in pickle    
     dst = dataset
     results = data[dataset]
-    with open('{}chtc_gs_best_params_{}.p'.format(pickledir,dst),'wb') as f:
-        #pass 
-        pickle.dump(results, f)
+    # with open('{}chtc_gs_best_params_{}.p'.format(pickledir,dst),'wb') as f:
+    #     #pass 
+    #     pickle.dump(results, f)
 

### 2018-04-16 Manuscripts/model-descriptions/packer/gamma_invariance.tex
New PACKER and CopyTweak implemented with gamma (tradeoff) as a multiplier for the contrast-category members only. Lets see how thje fits go! Still need to optimise the chtc code - there should be at least one more commit before sending this off to fit.
New fits seem good. CHTC code cleaned up. Still working on Hierarchical Bayes with representation?

--- 2017-04-04 Manuscripts/model-descriptions/packer/packer.tex
@@
-%!TEX output_directory = latex_out/
-
-\documentclass[12pt]{article}
-\usepackage[letterpaper, margin=1in, headheight=15pt]{geometry}
+\documentclass[12pt]{article} \usepackage[letterpaper, margin=1in, headheight=15pt]{geometry}
 \usepackage{amsmath, listings}
@@
 
-\section{The PACKER Model: Producing Alike or Contrasting Knowledge with Exemplar Representations}
+\section{Description of the non-variant effect of gamma on generation and assignment probabilities
+in PACKER}
 
-This model assumes that people represent categories as exemplars in a multidimensional space, and that generation is constrained by both similarity to members of the category being generated \textit{and} dissimilarity to members of other categories.
-The model assumes people generate categories that are dissimilar to known categories and have strong within-class similarity.
+This document refers to the version of PACKER as described in the manuscript submitted to Cognitive
+Psychology (sometime in 2017; henceforth referred to as the manuscript). In this version, PACKER
+performs very well in predicting the generation of new category exemplars. In particular, the
+tradeoff parameter (henceforth $\gamma$) actually affects the predictions generated in an expected
+way. However, when trying to produce predictions of category assignments, varying $\gamma$ does not
+affect the predictions in any way. The following section of this document describes why this is
+occurring.
+
+Equation 2 in the manuscript specifies that the aggregated similarity $a$ between generation
+candidate $y_i$ and a vector of stored exemplars $x$ can be expressed as:
+
+\begin{equation} a(y_i, x) = \sum_j{f(x_j) s(y_i, x_j)}
+\label{eq:aggsim}
+\end{equation} where $f(x_j) = \gamma$ when $x_j$ is a member of a target category and $f(x_j) =
+\gamma - 1$ when $x_j$ is a member of a contrast category.
+
+We can express Equation \ref{eq:aggsim} here a little differently:
+
+\begin{equation} a(y_i,x) = \gamma \sum_k{s(y_i,x_k)} + (\gamma-1) \sum_l{s(y_i,x_l)}
+\label{eq:aggsimsplit}
+\end{equation} where $k$ iterates over members of the target category and $l$ iterates over members
+of the contrast category. To make the algebra a little less tedious we can simplify some of the
+representations here. Let $t_i = \sum_k{s(y_i,x_k)}$ and $c_i = \sum_l{s(y_i,x_l)}$, which allows us to
+express Equation \ref{eq:aggsimsplit} as:
+
+\begin{equation} a(y_i,x) = \gamma t_i + (\gamma-1) c_i .
+\label{eq:aggsimsimp}
+\end{equation}
+
+According to the implementation of PACKER as seen in the manuscript, an exponentiated Luce's choice
+rule is used to find the probability $p$ of generating candidate $y_i$:
+
+\begin{equation} p(y_i) = \dfrac{\exp \{\theta \cdot a(y_i, x)\}} {\sum_i{\exp \{\theta \cdot
+a(y_i,x)\}}} .
+\end{equation} 
+Thus far, these equations only apply to predictions of category generation. However, predictions of
+category assignments require some reconceptualization of these equations, although most of the
+equations retain the same structure. Specifically, for Equations \ref{eq:aggsim} to
+\ref{eq:aggsimsimp}, $y_i$ now represents the assignment of candidate $y$ to target category $i$.
+
+expanding $a(y_i,x)$ with Equation \ref{eq:aggsimsimp}:
+\begin{equation} p(y_i) = \dfrac{\exp \{\theta \cdot \gamma t_i + (\gamma-1) c_i\}} {\sum_i {\exp
+      \{\theta \cdot \gamma t_i + (\gamma-1) c_i\}}}
+\label{eq:expluceexp}
+\end{equation}
+
+Whereas the denominator in Equation \ref{eq:expluceexp} is the summation of aggregated similarities
+over all generation candidates in category generation, category assignment predictions require that
+the denominator specifies the summation over all category assignments. Specifically, $i$ in the
+summation term represents each category, as opposed to each generation candidate.
+
+This model assumes that people represent categories as exemplars in a multidimensional space, and
+that generation is constrained by both similarity to members of the category being generated
+\textit{and} dissimilarity to members of other categories. The model assumes people generate
+categories that are dissimilar to known categories and have strong within-class similarity.
 
@@
 
-\begin{equation}
-s\left(x_i,x_j\right) = \exp \left\{ -c \left[\sum_{k}{ w_k \left| x_{ik} - x_{jk} \right|^r }\right]^{1/r} \right\}
+\begin{equation} s\left(x_i,x_j\right) = \exp \left\{ -c \left[\sum_{k}{ w_k \left| x_{ik} - x_{jk}
+\right|^r }\right]^{1/r} \right\}
 \label{eq:similarity}
@@
 
-When prompted to make a generation decision, participants are thought to consider both similarity to examples from other categories as well as similarity to examples in the target category.
-More formally, the aggregated similarity $a$ between candidate $y$ and the model's stored exemplars $x$ can be computed as:
+When prompted to make a generation decision, participants are thought to consider both similarity to
+examples from other categories as well as similarity to examples in the target category. More
+formally, the aggregated similarity $a$ between candidate $y$ and the model's stored exemplars $x$
+can be computed as:
 
-\begin{equation}
-  a(y, x) = \sum_j{f(x_j) s(y, x_j)}
+\begin{equation} a(y, x) = \sum_j{f(x_j) s(y, x_j)}
 \end{equation}
 
-Where $f(x_j)$ is a function specifying each stored example's degree of contribution toward generation.
-Although $f(x_j)$ may be set arbitrarily, in PACKER it is set according to class assignment.
-For known members of the target category, $f(x_j) = \gamma$.
-For members of contrast categories, $f(x_j) = -1 + \gamma$.
-$\gamma$ is thus a free parameter ($0 \leq \gamma \leq 1$) controlling the trade-off between within-class similarity and between-class dissimilarity: $\gamma = 1$ produces exclusive consideration of same-category members, and $\gamma = 0$ produces exclusive consideration of opposite-category members.
-When $\gamma = 0.5$, the similarity to contrast categories is effectively subtracted from the similarity to the target category.
+Where $f(x_j)$ is a function specifying each stored example's degree of contribution toward
+generation. Although $f(x_j)$ may be set arbitrarily, in PACKER it is set according to class
+assignment. For known members of the target category, $f(x_j) = \gamma$. For members of contrast
+categories, $f(x_j) = \gamma - 1$. $\gamma$ is thus a free parameter ($0 \leq \gamma \leq 1$)
+controlling the trade-off between within-class similarity and between-class dissimilarity: $\gamma =
+1$ produces exclusive consideration of same-category members, and $\gamma = 0$ produces exclusive
+consideration of opposite-category members. When $\gamma = 0.5$, the similarity to contrast
+categories is effectively subtracted from the similarity to the target category.
 
-The probability that a given item $y$ will be generated given the model's memory $x$ is computed using relative summed similarity values across all generation candidates $y_i$:
+The probability that a given item $y$ will be generated given the model's memory $x$ is computed
+using relative summed similarity values across all generation candidates $y_i$:
 
-\begin{equation}
-p(y) = \dfrac
-{ \exp  \left \{ \theta \cdot a \left( y, x \right) \right \} } 
-{ \sum_i{ \exp  \left \{ \theta \cdot a \left( y_i, x \right) \right\}  } }
+\begin{equation} p(y) = \dfrac { \exp \left \{ \theta \cdot a \left( y, x \right) \right \} } {
+\sum_i{ \exp \left \{ \theta \cdot a \left( y_i, x \right) \right\} } }
 \label{eq:packer-choice}

### 2018-04-20 Manuscripts/model-descriptions/packer/gamma_invariance.tex
Trying out another version of PACKER, where this time tradeoff is implemented as weights for both contrast and target similarities.

@@
-\documentclass[12pt]{article} \usepackage[letterpaper, margin=1in, headheight=15pt]{geometry}
-\usepackage{amsmath, listings}
+\documentclass[12pt]{article} \usepackage[letterpaper, margin=1in,
+headheight=15pt]{geometry} \usepackage{amsmath, listings}
 
@@
 
-\section{Description of the non-variant effect of gamma on generation and assignment probabilities
-in PACKER}
+\section{Description of the non-variant effect of gamma on generation and
+assignment probabilities in PACKER}
 
-This document refers to the version of PACKER as described in the manuscript submitted to Cognitive
-Psychology (sometime in 2017; henceforth referred to as the manuscript). In this version, PACKER
-performs very well in predicting the generation of new category exemplars. In particular, the
-tradeoff parameter (henceforth $\gamma$) actually affects the predictions generated in an expected
-way. However, when trying to produce predictions of category assignments, varying $\gamma$ does not
-affect the predictions in any way. The following section of this document describes why this is
-occurring.
+This document refers to the version of PACKER as described in the manuscript
+submitted to Cognitive Psychology (sometime in 2017; henceforth referred to as
+the manuscript). In this version, PACKER performs very well in predicting the
+generation of new category exemplars. In particular, the tradeoff parameter
+(henceforth $\gamma$) actually affects the predictions generated in an expected
+way. However, when trying to produce predictions of category assignments,
+varying $\gamma$ does not affect the predictions in any way. The following
+section of this document describes why this is occurring.
 
-Equation 2 in the manuscript specifies that the aggregated similarity $a$ between generation
-candidate $y_i$ and a vector of stored exemplars $x$ can be expressed as:
+Equation 2 in the manuscript specifies that the aggregated similarity $a$
+between generation candidate $y_i$ and a vector of stored exemplars $x$ can be
+expressed as:
 
@@
 \label{eq:aggsim}
-\end{equation} where $f(x_j) = \gamma$ when $x_j$ is a member of a target category and $f(x_j) =
-\gamma - 1$ when $x_j$ is a member of a contrast category.
+\end{equation} where $f(x_j) = \gamma$ when $x_j$ is a member of a target
+category $T$ and $f(x_j) = \gamma - 1$ when $x_j$ is a member of a contrast
+category $C$.
 
@@
 
-\begin{equation} a(y_i,x) = \gamma \sum_k{s(y_i,x_k)} + (\gamma-1) \sum_l{s(y_i,x_l)}
+\begin{equation} a(y_i,x) = \gamma \sum_{k \in K_T}{s(y_i,x_k)} + (\gamma-1)
+\sum_{k \in K_C}{s(y_i,x_k)}
 \label{eq:aggsimsplit}
-\end{equation} where $k$ iterates over members of the target category and $l$ iterates over members
-of the contrast category. To make the algebra a little less tedious we can simplify some of the
-representations here. Let $t_i = \sum_k{s(y_i,x_k)}$ and $c_i = \sum_l{s(y_i,x_l)}$, which allows us to
-express Equation \ref{eq:aggsimsplit} as:
+\end{equation} where $K_T$ is the set of all target category members and $K_C$ is
+the set of all members of the contrast category. To make the algebra a little
+less tedious we can simplify some of the representations here. Let $t_i =
+\sum_{k \in K_T}{s(y_i,x_k)}$ and $c_i = \sum_{k \in K_C}{s(y_i,x_k)}$, which allows
+us to express Equation \ref{eq:aggsimsplit} as:
 
@@
 
-According to the implementation of PACKER as seen in the manuscript, an exponentiated Luce's choice
-rule is used to find the probability $p$ of generating candidate $y_i$:
+According to the implementation of PACKER in the manuscript, an exponentiated
+Luce's choice rule is used to find the probability $p$ of generating candidate
+$y_i$:
 
-\begin{equation} p(y_i) = \dfrac{\exp \{\theta \cdot a(y_i, x)\}} {\sum_i{\exp \{\theta \cdot
-a(y_i,x)\}}} .
-\end{equation} 
-Thus far, these equations only apply to predictions of category generation. However, predictions of
-category assignments require some reconceptualization of these equations, although most of the
-equations retain the same structure. Specifically, for Equations \ref{eq:aggsim} to
-\ref{eq:aggsimsimp}, $y_i$ now represents the assignment of candidate $y$ to target category $i$.
+\begin{equation} p(y_i) = \dfrac{\exp \{\theta \cdot a(y_i, x)\}} {\sum_i{\exp
+\{\theta \cdot a(y_i,x)\}}} .
+\label{expluce}
+\end{equation} Thus far, Equation \ref{expluce} only apply to predictions of
+category generation. In order to extend PACKER to category assignment
+predictions, we need to find not the probability of generating candidate $y_i$,
+but rather, the probability of assigning candidate $y_i$ to the target category. 
+ these equations only apply to predictions of category
+generation. However, predictions of category assignments require some
+reconceptualization of these equations, although most of the equations retain
+the same structure. Specifically, for Equations \ref{eq:aggsim} to
+\ref{eq:expluce}, $i$ now iterates over each category (as opposed to each
+generation candidate; i.e., $i \in K$ $ y_i$ now represents the assignment of candidate $y$ to
+target category $i$.
 
-expanding $a(y_i,x)$ with Equation \ref{eq:aggsimsimp}:
-\begin{equation} p(y_i) = \dfrac{\exp \{\theta \cdot \gamma t_i + (\gamma-1) c_i\}} {\sum_i {\exp
-      \{\theta \cdot \gamma t_i + (\gamma-1) c_i\}}}
+Substituting $a(y_i,x)$ with Equation \ref{eq:aggsimsimp}:
+\begin{equation} p(y_i) = \dfrac{\exp \{\theta \cdot \gamma t_i + (\gamma-1)
+c_i\}} {\sum_i {\exp \{\theta \cdot \gamma t_i + (\gamma-1) c_i\}}}
 \label{eq:expluceexp}
@@
 
-Whereas the denominator in Equation \ref{eq:expluceexp} is the summation of aggregated similarities
-over all generation candidates in category generation, category assignment predictions require that
-the denominator specifies the summation over all category assignments. Specifically, $i$ in the
-summation term represents each category, as opposed to each generation candidate.
+Whereas the denominator in Equation \ref{eq:expluceexp} is the summation of
+aggregated similarities over all generation candidates in category generation,
+category assignment predictions require that the denominator specifies the
+summation over all category assignments. Specifically, $i$ in the summation term
+represents each category, as opposed to each generation candidate.
 
-This model assumes that people represent categories as exemplars in a multidimensional space, and
-that generation is constrained by both similarity to members of the category being generated
-\textit{and} dissimilarity to members of other categories. The model assumes people generate
-categories that are dissimilar to known categories and have strong within-class similarity.
+This model assumes that people represent categories as exemplars in a
+multidimensional space, and that generation is constrained by both similarity to
+members of the category being generated \textit{and} dissimilarity to members of
+other categories. The model assumes people generate categories that are
+dissimilar to known categories and have strong within-class similarity.
 
-The similarity between exemplars $x_i$ and $x_j$ is computed using Shepard's law:
+The similarity between exemplars $x_i$ and $x_j$ is computed using Shepard's
+law:
 
-\begin{equation} s\left(x_i,x_j\right) = \exp \left\{ -c \left[\sum_{k}{ w_k \left| x_{ik} - x_{jk}
-\right|^r }\right]^{1/r} \right\}
+\begin{equation} s\left(x_i,x_j\right) = \exp \left\{ -c \left[\sum_{k}{ w_k
+\left| x_{ik} - x_{jk} \right|^r }\right]^{1/r} \right\}
 \label{eq:similarity}
@@
 
-When prompted to make a generation decision, participants are thought to consider both similarity to
-examples from other categories as well as similarity to examples in the target category. More
-formally, the aggregated similarity $a$ between candidate $y$ and the model's stored exemplars $x$
-can be computed as:
+When prompted to make a generation decision, participants are thought to
+consider both similarity to examples from other categories as well as similarity
+to examples in the target category. More formally, the aggregated similarity $a$
+between candidate $y$ and the model's stored exemplars $x$ can be computed as:
 
@@
 
-Where $f(x_j)$ is a function specifying each stored example's degree of contribution toward
-generation. Although $f(x_j)$ may be set arbitrarily, in PACKER it is set according to class
-assignment. For known members of the target category, $f(x_j) = \gamma$. For members of contrast
-categories, $f(x_j) = \gamma - 1$. $\gamma$ is thus a free parameter ($0 \leq \gamma \leq 1$)
-controlling the trade-off between within-class similarity and between-class dissimilarity: $\gamma =
-1$ produces exclusive consideration of same-category members, and $\gamma = 0$ produces exclusive
-consideration of opposite-category members. When $\gamma = 0.5$, the similarity to contrast
-categories is effectively subtracted from the similarity to the target category.
+Where $f(x_j)$ is a function specifying each stored example's degree of
+contribution toward generation. Although $f(x_j)$ may be set arbitrarily, in
+PACKER it is set according to class assignment. For known members of the target
+category, $f(x_j) = \gamma$. For members of contrast categories, $f(x_j) =
+\gamma - 1$. $\gamma$ is thus a free parameter ($0 \leq \gamma \leq 1$)
+controlling the trade-off between within-class similarity and between-class
+dissimilarity: $\gamma = 1$ produces exclusive consideration of same-category
+members, and $\gamma = 0$ produces exclusive consideration of opposite-category
+members. When $\gamma = 0.5$, the similarity to contrast categories is
+effectively subtracted from the similarity to the target category.
 
-The probability that a given item $y$ will be generated given the model's memory $x$ is computed
-using relative summed similarity values across all generation candidates $y_i$:
+The probability that a given item $y$ will be generated given the model's memory
+$x$ is computed using relative summed similarity values across all generation
+candidates $y_i$:
 
-\begin{equation} p(y) = \dfrac { \exp \left \{ \theta \cdot a \left( y, x \right) \right \} } {
-\sum_i{ \exp \left \{ \theta \cdot a \left( y_i, x \right) \right\} } }
+\begin{equation} p(y) = \dfrac { \exp \left \{ \theta \cdot a \left( y, x
+\right) \right \} } { \sum_i{ \exp \left \{ \theta \cdot a \left( y_i, x \right)
+\right\} } }
 \label{eq:packer-choice}
@@
 
-Where $\theta$ ($\geq 0$) is a free parameter controlling overall response determinism.
+Where $\theta$ ($\geq 0$) is a free parameter controlling overall response
+determinism.
 

### 2018-04-20 Modules/Classes/ConjugateJK13.py

@@ -100,6 +100,13 @@ class ConjugateJK13(Model):
                         #target_dist = np.ones(mu.shape) * np.nan
                         density = np.ones(len(stimuli)) * np.nan
                 else:
+                        # #170418 Implementing representational draws
+                        # # See equation 7 in Tenenbaum & Griffiths 2001 cogsci proceedings paper
+                        # S = categories[category]
+                        # mmu = m-mu
+                        # Vinv = np.linalg.inv(V)
+                        # rep = N * log(S) - N * np.dot(np.dot(mmu.transpose(),Vinv),mmu) - trace(np.dot(S,Vinv))
+                        
                         target_dist = multivariate_normal(mean = mu, cov = Sigma)
                         density = target_dist.pdf(stimuli)
                         

### 2018-05-24 Manuscripts/cog-psych/revision/methods_results_e3.tex

% \documentclass[12pt]{article} \usepackage[letterpaper, margin=1in,
% headheight=15pt]{geometry} \usepackage{amsmath, listings}

% \begin{document}

\section{Participants, Materials, and Procedure}

Experiment 2 recruited 122 participants who each generated one set of four Alpha
and four Beta category exemplars. Among these 122 generated category sets were
102 unique sets (i.e., they contained a unique collection of Alpha and Beta
exemplars). Consequently, for Experiment 3, we recruited 102 participants with
one participant presented with a different unique category set.

Participants observed four blocks of eight trials. Each trial began with the
presentation of a fixation cross for 500 ms. This was followed by the
presentation of one exemplar randomly sampled without replacement from the
unique category set. Participants were tasked with assigning the presented
exemplar to either the Alpha or Beta category with no time limit imposed.
Feedback was automatically displayed for 2500 ms after each response.

\section{Results}

Overall accuracy of the participants was high, with a mean error rate of $.19$
($SD = .19$).

[Place error curve plots here, not unlike what NGPMG1994 has done]

[Move on to modelling analyses. Include correlation plots, don't worry about
individual fits. Also include the varying of PACKER's performance (and maybe
representativeness?) on correlation with changing tradeoff. ]

[Note Joe's comments on figure formatting (font size, remove redundant y-axes, etc)]

### 2018-06-22 Manuscripts/cog-psych/revision/methods_results_e3.tex
Working on fitting models to maximise correlations, but is slow. Unfeasible with multiple models. Somewhat optimised code still takes ages - I think I will need to restructure the script completely.

@@
-% \documentclass[12pt]{article} \usepackage[letterpaper, margin=1in,
-% headheight=15pt]{geometry} \usepackage{amsmath, listings}
+% \documentclass[12pt]{article} \usepackage[letterpaper, margin=1in, %
+%headheight=15pt]{geometry} \usepackage{amsmath, listings}
 
 % \begin{document}
+\section{Experiment 3.}
 
-\section{Participants, Materials, and Procedure}
+[I'm assuming Joe will include some preamble for the experiment, but otherwise
+it should be fairly straightforward for me to churn out one anyway.]
+
+\subsection{Participants, Materials, and Procedure}
 
@@
 
-\section{Results}
+\subsection{Results}
 
 Overall accuracy of the participants was high, with a mean error rate of $.19$
-($SD = .19$).
+($SD = .19$). Error rates for each block are presented in Figure
+\ref{fig:learningcurve}.
 
-[Place error curve plots here, not unlike what NGPMG1994 has done]
+\begin{figure}
+    \begin{center}
+    \includegraphics[width=\textwidth/2]{figs/e3-learningcurve.pdf}
+    \caption{Average error rate for each successive block. Green discontinuous
+      line represents the overall mean error rate.}
+    \label{fig:learningcurve}
+    \end{center}
+\end{figure}
 
-[Move on to modelling analyses. Include correlation plots, don't worry about
-individual fits. Also include the varying of PACKER's performance (and maybe
-representativeness?) on correlation with changing tradeoff. ]
+In the previous section (or Section X.X), we demonstrated that the best
+performing models were the two contrast models. In this section, our goal is to
+investigate if contrast is also important in category learning tasks.
 
-[Note Joe's comments on figure formatting (font size, remove redundant y-axes, etc)]
+It is also the
+goal of this paper to demonstrate that there is an association between the
+model's [justifying the errorvscat plot is harder than I thought] To compare the
+performances of each model, we analysed the correlation between each model's fit
+to a participant's unique category set and the participant's error rate.
+Intuitively, a well-performing model should be able to closely fit (i.e., easily
+generate) A positive correlation would indicate that the model is We used the
+same set of optimised parameters found from Section X.X, with individual values
+for selective attention.
+
+\begin{figure}
+    \begin{center}
+    \includegraphics[width=\textwidth]{figs/perror_corr.pdf}
+    \caption{Correlation between observed participant error and model fit.}
+    \label{fig:perror_corr}
+    \end{center}
+\end{figure}
+
+To emphasize the strong influence of contrast in maximizing the association
+between category generation and category learning, we computed the correlations
+over wide range of $\theta_{contrast}$ values, with the other parameter values of
+PACKER held constant at their optimized levels. As presented in Figure
+\ref{fig:packer-corr}, the correlation quickly increases with increasing weight
+on contrast, reaching a plateau above values of around 4.0.
+
+\begin{figure}
+    \begin{center}
+    \includegraphics[width=0.8\textwidth]{figs/packer-corr.pdf}
+    \caption{Correlation between PACKER's fit and participant error as a
+      function of the $\theta_{contrast}$ parameter. To facilitate comparison,
+      PACKER's other parameters ($c$, $\theta_{target}$) were set to the best
+      fitting values obtained for copy-and-tweak in Table
+      \ref{table:global-model-fits}. Grey dashed line represents the correlation
+      between Copy \& Tweak's fit and participant error.}
+    \label{fig:packer-corr}
+    \end{center}
+\end{figure}
+
+The specific reason for why contrast as implemented in PACKER is superior to
+contrast as implemented in the representativeness model is still not clear.
+
+[Note Joe's comments on figure formatting (font size, remove redundant y-axes,
+etc)]

### 2018-06-22 Manuscripts/model-descriptions/jern-kemp-conjugate/representative.tex
Working on fitting models to maximise correlations, but is slow. Unfeasible with multiple models. Somewhat optimised code still takes ages - I think I will need to restructure the script completely.

--- 2017-01-30 Manuscripts/model-descriptions/jern-kemp-conjugate/jern-kemp-conjugate.tex
@@
 
-\section*{Updated Version of the Jern \& Kemp (2013) Model}
+\section*{Conjugate Implementation of the Jern \& Kemp (2013) Model with Representativeness}
 
-Working backwards, what is needed for simulation of exemplar generation is the probability of generating a stimulus $y$ given exposure to members of the target category $x_B$:
-\
+The representativeness model is very similar to the Jern \& Kemp (2013)
+hierarchical Bayesian model. The essential difference here is the ultimate
+formulation of response probabilities. Where in the original hierarchical
+Bayesian model the exemplars are drawn from Gaussian distributions, in the
+representativeness model each exemplar is drawn in proportion to the exemplar's
+representativeness. The representativeness of an exemplar is defined by
+Tenenbaum and Griffiths (2012) as the relative evidence that is provided by the
+exemplar $x$ for a given hypothesis $h$ compared to all other hypotheses $h'$:
+
 \begin{equation}
-p(y | x_B) = ?
+  R(x,h) = \log \dfrac{p(x|h)}{\sum_{h'\neq h}{p(x|h')p(h')}}
+\label{representativeness}
 \end{equation}
-\
-where $x_B$ may be empty.
-Jern \& Kemp's model achieves this using a generative process.
-Category members ($x_A$ or $x_B$; more generally written as $x_C$) are assumed to have been generated using an underlying category distribution (specifically, multivariate normal):
-\
-\begin{equation}
-  x_C \sim Normal(\mu_{C}, \Sigma_{C})
-\end{equation}
-\
-$p(y | x_C)$ is proportional to the candidate's density under the target category's distribution.
-Thus, obtaining the category distribution parameters $(\mu_{C}, \Sigma_{C})$ is key for generation.
-This document describes how we compute these variables in a conjugate model.
+
+For simplicity, and consistency with the original hierarchical Bayesian model,
+we define $h_{C}$ as a multivariate normal distribution for a particular
+category $C$ parameterised by $\mu_C$ and $\Sigma_C$. For completeness, this
+document will repeat some information from the hierarchical Bayesian model
+document and describee how we compute these variables.
 
@@
 Because we set $\mu_0$ to the center of the space, this outcome is the same as if we had integrated over all possible $\mu_C$.
+
 In practice, if $n_C = 0$, the model picks a stimulus at random from all candidates (uniform probabilities).
@@
 \
-$\Sigma_{0}$ is a $p$-by-$p$ prior covariance matrix.
-We use a $p$-dimensional identity matrix $I_p$ multiplied element-wise against a free parameter, $\rho$, controlling the amount of variance assumed by the prior:
+$\Sigma_{0}$ is a $d$-by-$d$ prior covariance matrix.
+We use a $d$-dimensional identity matrix $I_d$ multiplied element-wise against a free parameter, $\lambda$, controlling the amount of variance assumed by the prior:
 \
 \begin{equation}
-    \Sigma_0 =  I_p\rho
+    \Sigma_0 =  \lambda I_d
 \end{equation}
 
-Thus, categories are assumed to have some degree of variance along each feature (specified by $\rho$), but not are assumed to possess feature-feature correlations.
+Thus, categories are assumed to have some degree of variance along each feature (specified by $\lambda$), but not are assumed to possess feature-feature correlations.
+Differences in the assumed variance among the features, similarly to weighting in an exemplar model, can be implemented through a small change to the equation.
+Specifically, the variance assumed of dimension $k$ is given by:
+\
+\begin{equation}
+    \Sigma_{0k} =  \lambda w_k d
+\end{equation}
+\
+where $d$ is the number of dimensions, and $w_k$ ($0 \leq w_k \geq 1$, $\sum_k{w_k} = 1$) indicates the dimension's relative share of the total assumed variability.
+Under this system, evenly distributed weights result in uniformly assumed variances, equal to $\lambda$.
 
@@
 $\nu$ is an additional free parameter, weighting the importance of $\Sigma_{D}$.
-$\nu$ must be greater than $p-1$.
+$\nu$ must be greater than $d-1$.
 When $x_b$ is empty, Equation \ref{eq:Sigma_y} reduces to $\Sigma_C = \Sigma_D$.
@@
 
-$x$ are assumed to be drawn from a the distribution given by $Normal(\mu_{C}, \Sigma_{C})$. Thus,
+As mentioned earlier, this is the point where the representativeness model
+diverges from the hierarchical Bayesian model. Specifically, the probability of
+generating exemplar $x$ is proportional to its representativeness:
 
 \begin{equation}
-  p(y) \propto Normal(y; \mu_{C}, \Sigma_{C})
+  p(x) \propto R(x,h) ,
 \end{equation}
+where $h$ is a multivariate normal distribution parameterised by $\mu_{C}$ and
+$\Sigma_{C}$.
 
-In practice, $p(y)$ is computed by first obtaining the relative density of every possible generation candidate $y_i$ under the category distribution.
-The end probability is a normalization of these values:
+In practice, $p(x)$ is computed by first obtaining the representativeness of
+every possible generation candidate $x_i$. The end probability is a
+normalization of these values:
 
@@
   p(x) = \dfrac
-    {exp( \theta Normal(y ; \mu_{C}, \Sigma_{C}))}
-    {\sum_i exp( \theta Normal(y_i ; \mu_{C}, \Sigma_{C}))} 
+    {\exp( \theta \cdot R(x,h))}
+    {\sum_i \exp( \theta \cdot R(x_i,h_i))} 
 \end{equation}
@@
     \item $\kappa$. Scalar, $\kappa>0$. Weights the importance of $\mu_{0}$ in inferring category $\mu_{C}$.
-    \item $\rho$. Scalar, $\rho>0$. Sets the assumed variance in the domain prior, $\Sigma_{0}$.
-    \item $\nu$. Scalar, $\nu > p-1$. Weights the importance of $\Sigma_{D}$ in inferring the domain $\Sigma_{C}$.
+    \item $\lambda$. Scalar, $\lambda>0$. Sets the assumed variance in the domain prior, $\Sigma_{0}$.
+    \item $\nu$. Scalar, $\nu > d-1$. Weights the importance of $\Sigma_{D}$ in inferring the domain $\Sigma_{C}$.
     \item $\theta$. Scalar, $\theta > 0$. Response determinism parameter.

### 2018-06-22 Modules/Classes/Simulation.py

@@ -271,7 +271,9 @@ class Trialset(object):
                             self.stimrange).get_generation_ps(self.stimuli, 1,self.task)
                 idc0 = trial['response'][0] 
                 idc1 = trial['response'][1] 
-
+                ps_add = np.concatenate([ps0[idc0],ps1[idc1]])
+ 
+                #190618 after much thought, error should be exactly the same as assign though...
                 #Actual category exemplars
                 correctcat = trial['categories']
 
@@ -297,7 +299,7 @@ class Trialset(object):
                         wrongps   = np.concatenate([wrongps,ps[wrongresp]])
                         
 
-                ps_add = np.concatenate([correctps,wrongps])
+                #ps_add = np.concatenate([correctps,wrongps])
 
                                 
             # check for nans and zeros
@@ -320,6 +322,35 @@ class Trialset(object):
         else:
             return -1.0 * loglike
 
+    def correlate(self, params, model, alphas, fixedparams = None):
+        """Returns the correlation between model generation fit and observed
+        participant error. This functions extracts the data for each individual
+        participant in the trialset object, computes the observed error from that
+        and calculates the model fit (negative LL) to that participant's data.        
+
+        """
+        from scipy.stats import stats as ss        
+        ppts = self.participants
+        ppt_err_rate = []
+        ll = []
+        for ppt in ppts:
+            pptTrialObj = extractPptData(self.ppt)
+            nresp = 0.0
+            nresp_correct = 0.0
+            for trial in pptTrialObj.Set:
+                correctcats = trial['categories']
+                #Iterate over each category
+                for i,correctcat in enumerate(correctcats):
+                    #update number of responses                    
+                    nresp += len(trial['response'][i])
+                    for correct in correctcat:
+                        nresp_correct += trial['response'][i].count(correct)
+            ppt_err_rate += [1-(nresp_correct/nresp)]
+            #get the total ll of all permutations of the beta category
+            ll += [pptTrialObj.loglike(params, model)]
+        correlation = ss.pearsonr(ppt_err_rate,ll)
+
+        
 def hillclimber(model_obj, trials_obj, options, fixedparams = None, inits = None, results = True,callbackstyle='none'):
     """
     Run an optimization routine.
@@ -379,6 +410,67 @@ def hillclimber(model_obj, trials_obj, options, fixedparams = None, inits = None
         print '\tAIC = ' + str(AIC)
[…]

### 2018-06-28 Manuscripts/cog-psych/revision/methods_results_e3.tex
Likelihood ratio tests for correlation might not be appropriate. Until I find a better solution, run bootstraps first.

@@
 Experiment 2 recruited 122 participants who each generated one set of four Alpha
-and four Beta category exemplars. Among these 122 generated category sets were
-102 unique sets (i.e., they contained a unique collection of Alpha and Beta
-exemplars). Consequently, for Experiment 3, we recruited 102 participants with
-one participant presented with a different unique category set.
+and four Beta category exemplars. Experiment 3 recruited the same number of
+participants.
 
@@
 
+Among the 122 originally generated category sets were 102 unique sets (i.e.,
+they contained a unique collection of Alpha and Beta exemplars). Consequently,
+for our subsequent analyses for this experiment we only use data from the first
+102 participants that were presented with a unique category set.
+
 Overall accuracy of the participants was high, with a mean error rate of $.19$
-($SD = .19$). Error rates for each block are presented in Figure
+($SD = .19$; min max?). Error rates for each block are presented in Figure
 \ref{fig:learningcurve}.

### 2018-06-28 Experiments/cogpsych-code/globa-model-grid-search-parallel.py

@@ -1,179 +1,0 @@
-import pickle
-import pandas as pd
-import numpy as np
-import time
-execfile('Imports.py')
-import Modules.Funcs as funcs
-from Modules.Classes import Simulation
-from Modules.Classes import CopyTweak
-from Modules.Classes import Packer
-from Modules.Classes import ConjugateJK13
-
-# Specify default dataname
-dataname_def = 'nosofsky1989'
-participant_def = 'all'
-unique_trials_def = 'all'
-
-#Allow for input arguments at the shell
-if __name__ == "__main__":
-        import sys
-
-        if len(sys.argv)<4:
-                unique_trials = unique_trials_def
-        else:
-                unique_trials = int(sys.argv[3])
-        if len(sys.argv)<3:
-                participant = participant_def
-        else:
-                participant = int(sys.argv[2])
-        if len(sys.argv)<2:
-                dataname = dataname_def
-        else:
-                dataname = sys.argv[1]        
-else:
-        dataname = dataname_def
-        participant = participant_def
-        unique_trials = unique_trials_def
-
-
-s = 'Invalid data name supplied. Please select one of these options:'
-choices = ['pooled','pooled-no1st','nosofsky1986','nosofsky1989']
-
-dataname = funcs.valData(dataname,s,choices)
-
-# Data
-if dataname == 'pooled':
-        # all data
-        src = "pickles/all_data_e1_e2.p"
-        dst = "pickles/gs_best_params_all_data_e1_e2.p"
-        task = "generate"
-elif dataname == 'pooled-no1st':
-        # trials 2-4
-        src = "pickles/trials_2-4_e1_e2.p"
-        dst = "pickles/gs_best_params_trials_2-4_e1_e2.p"
-        task = "generate"
-elif dataname == 'nosofsky1986':
-        # nosofsky data
-        src = "pickles/nosofsky1986.p"
-        dst = "pickles/gs_best_params_nosofsky1986.p"
-        task = "assign"
-elif dataname == 'nosofsky1989':
-        # nosofsky data
-        src = "pickles/nosofsky1989.p"
-        dst = "pickles/gs_best_params_nosofsky1989.p"
-        task = "assign"
-else:        
-        raise Exception('Invalid data name specified.')
-
-dst_error = dst[0:-2] + '_error.p'
-
-print 'Grid Searching Data: ' + dataname
-
-# get data from pickle
-with open(src, "rb" ) as f:
-	trials = pickle.load( f )
-
-trials.task = task
-
[…]

### 2018-06-28 Experiments/cogpsych-code/plot_modelppt_ease.py

@@ -16,12 +16,21 @@ import matplotlib.pyplot as plt
 import seaborn as sns
 sns.set_style("whitegrid")
 
+pearson = True
+if pearson:
+    corrtype = 'p'
+else:
+    corrtype = 's'
+
+savefilename='modelvsppt{}.pdf'.format(corrtype)
+bestparmdb = "pickles/chtc_gs_best_params_corr{}.p".format(corrtype)
+modeleaseDB = "pickles/modelease_corr{}.p".format(corrtype)
 #Some plotting options
 font = {'family' : 'DejaVu Sans',
         'weight' : 'regular',
         'size'   : 20}
 show_p = True #show pearson r in plots
-show_s = True #show spearman tho in plots
+show_s = False #show spearman tho in plots
 
 #Specify simulation values
 N_SAMPLES = 10000
@@ -44,7 +53,7 @@ with open(pickledir+src, "rb" ) as f:
 
 # get best params pickle
 #bestparmdb = "pickles/chtc_gs_best_params_all_data_e1_e2.p"
-bestparmdb = "pickles/chtc_gs_best_params_corr.p"
+#bestparmdb = "pickles/chtc_gs_best_params_corrs.p"
 with open(bestparmdb, "rb" ) as f:
     best_params_t = pickle.load( f )
 
@@ -87,9 +96,7 @@ for i,row in info.iterrows():
 
 
 pptlist = np.unique(pptlist)
-
 #see if ll_global exists as a pickle, otherwise construct new ll
-modeleaseDB = "pickles/modelease_corr.p"
 try:
     with open(modeleaseDB, "rb" ) as f:
         ll_global = pickle.load( f )
@@ -177,7 +184,6 @@ for model_obj in modelList:
         #Save pickle for faster running next time
         with open(modeleaseDB, "wb" ) as f:
             pickle.dump(ll_global, f)
-
 #fh,axs = plt.subplots(2,int(np.ceil(len(modelList)/2)), figsize=(20,8))
 fh,axs = plt.subplots(2,int(np.ceil(len(modelList)/2)), figsize=(15,15))
 plt.tight_layout(w_pad=1,h_pad=5.0,rect=(.05,.05,.95,.95))
@@ -196,12 +202,35 @@ for m,model_obj in enumerate(modelList):
     model_loc = modelPlotOrder==model_obj
     ax = axs[model_loc][0]
     #Plot figure
[…]

### 2018-06-28 Experiments/cogpsych-code/validate_data.py

@@ -60,3 +60,4 @@ else:
 
 src = '{}.p'.format(filename)
 dst = 'best_params_{}.p'.format(filename)
+bestparmchtc = 'chtc_gs_best_params_{}.p'.format(filename)

### 2018-06-28 Modules/Classes/ConjugateJK13.py

@@ -94,7 +94,7 @@ class ConjugateJK13(Model):
 
         return (mu,Sigma)
 
-    def get_generation_ps(self, stimuli, category, task='generate'):
+    def get_generation_ps(self, stimuli, category, task='generate',seedrng=False):
 
         # random response if there are no target members.
         target_is_populated = any(self.assignments == category)
@@ -243,7 +243,10 @@ class RepresentJK13(Model):
 
         return (mu,Sigma)
 
-    def get_generation_ps(self, stimuli, category, task='generate'):
+    def get_generation_ps(self, stimuli, category, task='generate', seedrng = False):
+        if seedrng:
+            np.random.seed(234983) #some arbitrary seed value here
+
         target_is_populated = any(self.assignments == category)                
         mu,Sigma = self.get_musig(stimuli, category)
                 

### 2018-06-28 Modules/Classes/Simulation.py

@@ -189,7 +189,7 @@ class Trialset(object):
         return self
                 
 
-    def loglike(self, params, model, fixedparams = None, whole_array=False, parmxform = True):
+    def loglike(self, params, model, fixedparams = None, whole_array=False, parmxform = True,seedrng = False):
         """
         Evaluate a model object's log-likelihood on the
         trial set based on the provided parameters.
@@ -212,6 +212,8 @@ class Trialset(object):
         # iterate over trials
         loglike = 0
         ps_list = np.array([])
+        if self.task is '':
+            raise Exception('Task for trialset object not defined.')
         task = self.task
         for idx, trial in enumerate(self.Set):
             # format categories
@@ -220,15 +222,15 @@ class Trialset(object):
             # ps0 = np.zeros(ps1.shape)
             if task == 'generate':                                
                 # compute probabilities of generating exemplar in cat 1
-                ps = model(categories, params, self.stimrange).get_generation_ps(self.stimuli, 1,self.task)
+                ps = model(categories, params, self.stimrange).get_generation_ps(self.stimuli, 1,self.task,seedrng = seedrng)
                 ps_add = ps[trial['response']]
             elif task=='assign':
                 #Compute probabilities of assigning exemplar to cat 0
-                ps0 = model(categories, params, self.stimrange).get_generation_ps(self.stimuli, 0,self.task)
+                ps0 = model(categories, params, self.stimrange).get_generation_ps(self.stimuli, 0,self.task,seedrng = seedrng)
                 #Compute probabilities of assigning exemplar to cat 1
                 ps1 = model(categories, params,
                             self.stimrange).get_generation_ps(self.stimuli,
-                                                              1,self.task)
+                                                              1,self.task,seedrng = seedrng)
 
 
                 idc0 = trial['response'][0]
@@ -264,10 +266,10 @@ class Trialset(object):
                 #idc_err = trial['response'][0]
                 #Compute probabilities of assigning exemplar to cat 0
                 ps0 = model(categories, params,
-                            self.stimrange).get_generation_ps(self.stimuli, 0,self.task)
+                            self.stimrange).get_generation_ps(self.stimuli, 0,self.task,seedrng = seedrng)
                 #Compute probabilities of assigning exemplar to cat 1
                 ps1 = model(categories, params,
[…]

### 2018-07-02 Manuscripts/cog-psych/revision/cogpsych.tex
Working on fitting models to maximise correlations, but is slow. Unfeasible with multiple models. Somewhat optimised code still takes ages - I think I will need to restructure the script completely.

--- 2017-06-17 Manuscripts/cog-psych/cogpsych.tex
@@
 \renewcommand{\headrulewidth}{0pt}
-\lhead{RUNNING HEAD: Similarity and Contrast in Concept Generation}
+\lhead{RUNNING HEAD: Similarity, Contrast, and Representativeness in Categorization}
 \fancyhead[R]{\thepage}
-
-% author info:
-% https://www.elsevier.com/journals/cognitive-psychology/0010-0285/guide-for-authors
 
@@
 
-Creating Something Different: Similarity and Contrast in Concept Generation
+Creating and Learning Something Different: \\
+Similarity, Contrast, and Representativeness in Categorization
 
 \vfill
-
+Joseph L. Austerweil\textsuperscript{1}
 Nolan Conaway\textsuperscript{1}, 
-Kenneth J. Kurtz\textsuperscript{2}, 
-and Joseph L. Austerweil\textsuperscript{1}
+Shi X. Liew\textsuperscript{1},
+and Kenneth J. Kurtz\textsuperscript{2}
 \\[\baselineskip]
@@
 \setlength\parindent{0.5in}
-{\em Keywords}: categorization, concepts, creativity, generation, computational modeling, exemplar models
+{\em Keywords}: categorization; concepts; creativity; generation; computational modeling; exemplar models
 \clearpage
 
-% ------- BEGIN! ------- %
-\begin{flushleft}
-
-\section{Introduction}
-\setlength\parindent{0.5in}
-
-The creation of new ideas is one of the most fascinating and important human capabilities.
-Yet, it is also one of the most difficult human capabilities to study.
-We tend to focus on the most salient products of conceptual generation (e.g., scientific breakthroughs), but every person is likely to have generated many novel concepts in ordinary life.
-Indeed, the generation of novel sentences, works of art, and so on may also be considered products of our capacity to create new concepts.
-However, while this capacity is of great interest, it is also highly complex, making it difficult to study in a typical laboratory setting.
-By consequence, we do not have a strong understanding of the representations and processes involved in the generation of new concepts.
-
-Questions concerning the generation of novel concepts are typically addressed within research on creativity, a tremendously diverse field employing a multitude of experimental and theoretical approaches \citep[see][]{kaufman2010handbook}.
-The processes underlying conceptual generation are commonly investigated using the creative cognition approach \citep{finke1992creative, smith1995creative}, in which the products of creative acts (such as drawings) are analyzed to obtain insights into how they might have been created.
-These approaches have made great advances in our understanding of creativity, but the focus on particular types of creative products such as drawings makes it difficult to formalize the proposed processes and representations within a computational modeling framework.
-As a consequence, this approach does not often employ computational models to provide formal tests of competing theories.
-
-The act of generating a new concept is, however, not altogether different from the types of behaviors typically studied in cognitive psychology laboratories.
-In particular, generating a member of a novel class can be considered a `special case' application of existing category knowledge \citep{kemp2014taxonomy,kurtz2015human}.
-Research in categorization typically focuses on classification \citep[predicting an object's category given its features;][]{shepard1961learning,kurtz2013human} and inference \citep[predicting an object's unobserved features given observed features and a category label, see e.g.,][]{markman2003category}.
-The generation of members of a new category consists of inferring {\em all} features for a {\em novel} category label.
-Thus, we can make progress formalizing the processes involved in creative generation by extending theories of categorization to the special case.
-
-% % OLD INTRO:
+% % OLD OLD INTRO:
 % Creativity, innovation, imagination, and the creation of new ideas are among the most fascinating, important and difficult human capabilities to study.
@@
 
-Previous work in categorization has established that people are highly sensitive to the structural properties of categories, such as correlations between the features of category members and the relation between items within the same category and those in different categories \citep{roschmervis1975,regier2007,shepard1961learning}.
+%cut from submission intro:
+
+%Yet, it is also one of the most difficult human capabilities to study.
+We tend to focus on the most salient products of conceptual generation (e.g., scientific breakthroughs), but every person is likely to have generated many novel concepts in ordinary life.
+Indeed, the generation of novel sentences, works of art, and so on may also be considered products of our capacity to create new concepts.
+However, while this capacity is of great interest, it is also highly complex, making it difficult to study in a typical laboratory setting.
+By consequence, we do not have a strong understanding of the representations and processes involved in the generation of new concepts.
+
+%Questions concerning the generation of novel concepts are typically addressed within research on creativity, a tremendously diverse field employing a multitude of experimental and theoretical approaches \citep[see][]{kaufman2010handbook}.
+The processes underlying conceptual generation are commonly investigated using the creative cognition approach \citep{finke1992creative, smith1995creative}, in which the products of creative acts (such as drawings) are analyzed to obtain insights into how they might have been created.
+These approaches have made great advances in our understanding of creativity, but the focus on particular types of creative products such as drawings makes it difficult to formalize the proposed processes and representations within a computational modeling framework.
+As a consequence, this approach does not often employ computational models to provide formal tests of competing theories.
+
+% ------- BEGIN! ------- %
+\begin{flushleft}
+
+\section{Introduction}
+\setlength\parindent{0.5in}
+
+Creating and learning new ideas is one of the most fascinating and important human capabilities.
+For example, cell phones and smart phones are newly created categories of objects that young and old adults learned and have become reliant on in the last few decades.
+In fact, computational reearchers have identified generating novel objects and objectives as one of the defining, and most difficult to formalize, characteristics of intelligent life \citep{lake2017,lehman2011,taylor2016}.
+Yet, the cognitive mechanisms that enable people to innovate and learn new innovations are not well understood and are understudied by scientists.
+In part, this is due to the inherent difficulty of designing and conducting experiments that test these capabilities.
+How would a scientist devise an experiment that induces a participant to create and learn new categories that are as interesting as cell phones and smart phones?
+In this article, we investigate the prior expectations people use to learn and generate new categories using computational models and behavioral experiments.
+
+Generating and learning new concepts are, however, not altogether different from the types of behaviors typically studied in cognitive psychology laboratories.
+In particular, generating a member of a novel class can be considered a `special case' application of existing category knowledge \citep{kemp2014taxonomy,kurtz2015human}.
+Research in categorization typically focuses on what properties of a category affect the classification of an object's category given its features \citep{shepard1961learning,kurtz2013human}, and the prediction of an object's unobserved features given some of its other features and/or its category \citep{markman2003category}.
+The generation of members of a new category consists of inferring {\em all} features for a {\em novel} category label.
+Thus, we can make progress formalizing the processes involved in creative generation by extending theories of categorization to this special case.
+Then, we can examine what factors of the generated categories affect the ease at which people learn the category.
+
+Previous work in category learning and generation has established that people are highly sensitive to the structural properties of categories, such as correlations between the features of category members and the relation between items within the same category and those in different categories \citep{roschmervis1975,regier2007,shepard1961learning,smith1995}.
 Inspired by this work, previous research on the topic of category generation has explored a similar principle: People tend to create new categories that have similar {\em statistical regularities} as previously learned categories \citep{jern2013probabilistic,ward1994structured}.
@@
 
+\include{methods_results_e3}
+
 \section{General Discussion}

### 2018-07-04 Manuscripts/cog-psych/revision/cogpsych.tex
rough draft of new intro

@@
 For example, cell phones and smart phones are newly created categories of objects that young and old adults learned and have become reliant on in the last few decades.
-In fact, computational reearchers have identified generating novel objects and objectives as one of the defining, and most difficult to formalize, characteristics of intelligent life \citep{lake2017,lehman2011,taylor2016}.
+In fact, computational researchers have identified generating novel objects and objectives as one of the defining, and most difficult to formalize, characteristics of intelligent life \citep{lake2017,lehman2011,taylor2016}.
 Yet, the cognitive mechanisms that enable people to innovate and learn new innovations are not well understood and are understudied by scientists.
@@
 How would a scientist devise an experiment that induces a participant to create and learn new categories that are as interesting as cell phones and smart phones?
-In this article, we investigate the prior expectations people use to learn and generate new categories using computational models and behavioral experiments.
+In this article, we use computational models and behavioral experiments to investigate the prior expectations people have when they learn and generate new categories.
 
@@
 The generation of members of a new category consists of inferring {\em all} features for a {\em novel} category label.
-Thus, we can make progress formalizing the processes involved in creative generation by extending theories of categorization to this special case.
-Then, we can examine what factors of the generated categories affect the ease at which people learn the category.
+Thus, we can make progress formalizing the processes involved in category generation by extending theories of categorization to this case.
+Then, to ensure that these principles reflect basic principles of categorization and are not just specific to category generation, we can examine what factors of the generated categories affect the ease at which people learn the category.
 
@@
 Although the idea of category contrast is ubiquitous throughout the categorization literature, and extends to a variety of other fields \citep[e.g., color;][]{regier2007}, the idea that a new category should be ``different'' is vague, as there are many ways it could be different from a previously observed category.
-Building on the largely successful exemplar modeling framework \citep{medin1978context,nosofsky1984choice,nosofsky1986attention}, we propose a novel exemplar model of category generation, {\em Producing Alike and Contrasting Knowledge using Exemplar Representations} (PACKER), formalizing how new categories should differ from previous categories.
-This model makes novel predictions about how contrast affects category generation, which we test using behavioral experiments.
+
+We examine two different definitions of how categories should differ from one another: exemplar contrast or representative of the alternative category.
+To formalize the first technique, we build on the largely successful exemplar modeling framework that \citep{medin1978context,nosofsky1984choice,nosofsky1986attention}.
+In doing so, we propose a novel exemplar model of category generation, {\em Producing Alike and Contrasting Knowledge using Exemplar Representations} (PACKER), formalizing how new categories should differ from previous categories.
+It embodies the exemplar contrast principle by incorporating a factor that repulses members of opposite categories from one another.
+To foreshadow, we find that this notion of contrast predicts human category learning performance, above and beyond other category learning models in a novel behavioral experiment.
+This demonstrates that contrast plays a central role in other important categorization tasks, which is not specific to category generation
+
+To formalize the hypothesis that people expect a new category to be representative of the opposite of the current category, or contrast as {\em representativeness} \citep{kahneman73}, we use the Bayesian formulation of representativeness \citep{tenenbaum01}, which has been used successfully to capture human performance in color categorization \citep{abbott16}, image category learning \citep{abbott11}, and language grammar learning \citep{rafferty10}.
+We do so by extending the state-of-the-art category generation model \citep{jern2013probabilistic} and find that  human category generation is best capture as generating representative, rather than probable, samples.
 

### 2018-07-10 Manuscripts/cog-psych/revision/cogpsych.tex
major updates to intro framing to incorporate representativeness; minor updates elsewhere to mark where it should go and to our discussion of what hierarchical Bayes can do in prior work

@@
 \section*{Abstract}
-
+\jlanote{need to update}
 The ability to creatively generate new concepts and ideas is among the most fascinating aspects of human cognition, but we do not have a strong understanding of the cognitive processes and representations underlying concept generation.
@@
 To foreshadow, we find that this notion of contrast predicts human category learning performance, above and beyond other category learning models in a novel behavioral experiment.
-This demonstrates that contrast plays a central role in other important categorization tasks, which is not specific to category generation
-
-To formalize the hypothesis that people expect a new category to be representative of the opposite of the current category, or contrast as {\em representativeness} \citep{kahneman73}, we use the Bayesian formulation of representativeness \citep{tenenbaum01}, which has been used successfully to capture human performance in color categorization \citep{abbott16}, image category learning \citep{abbott11}, and language grammar learning \citep{rafferty10}.
+This demonstrates that contrast plays a central role in other important categorization tasks, which is not specific to category generation.
+
+The second hypothesis for how contrast affects categorization is contrast as {\em representativeness} \citep{kahneman73}.
+The representativeness heuristic states that people make judgments regarding the probability of an outcome based on how representative it is of the observed evidence.
+The heuristic is a powerful theoretical construct that has been used to capture a range of complex patterns of human judgments \citep{kahneman73,tversky74,tversky83}, especially those that deviate from normative theory (as given by Bayesian inference).
+For example, the coinflip sequence TTHTH is perceived to be more random than TTTTT (despite both having the same probability of having been generated from a fair coin).
+Although there is a healthy debate as to whether perceived randomness {\em really} deviates from normative theory \citep{griffiths18,hahn09}, the representativeness heuristic remains one of the main explanations of human judgments \citep{reimers18}.
+To formalize the hypothesis that people expect a new category to be representative of the opposite of the current category, or contrast as {\em representativeness}, we use the Bayesian formulation of representativeness \citep{tenenbaum01}, which has been used successfully to capture human performance in color categorization \citep{abbott16}, image category learning \citep{abbott11}, and language grammar learning \citep{rafferty10}.
 We do so by extending the state-of-the-art category generation model \citep{jern2013probabilistic} and find that  human category generation is best capture as generating representative, rather than probable, samples.
@@
 First we describe previous empirical work on the topic of category generation, as well as the computational formalizations of the theories in those reports.
-Then we describe our novel computational model, which is designed to generate categories that systematically differ from existing categories in the domain.
+We then present two hypotheses for how contrast might affect categorization and formalize them in computational models.
+The first is a novel computational model, which is designed to generate categories that systematically differ from existing categories in the domain.
+The second assumes the goal of category generation is to create representative samples of the opposite of the observed categories.
 We present two experiments demonstrating strong and systematic effects of category contrast on creative generation, and we qualitatively and quantitatively analyze the performance of each model in capturing human category generation.
@@
 
-The broader set of observations made by Ward and colleagues provide a great deal of insight into the nature of creative generation.
-They indicate that people rely strongly on prior knowledge in the creative process, and people generate concepts in accord with what they already know.
-Much of the work from this area \citep[e.g.,][]{smith1993constraining,marsh1999inadvertent} focuses on how information provided to participants (such as an example of a species generated by other participants) can drastically diminish creativity.
+The broader set of observations made by Ward and colleagues provide a great deal of insight into the role of prior knowledge in constraining category generation.
+Much of the work from this area \citep[e.g.,][]{smith1993constraining,marsh1999inadvertent} focuses on how information provided to participants (such as an example of a species generated by other participants) can drastically diminish difference of a new category from pre-existing categories.
 Theoretical accounts of these effects have primarily been grounded within the categorization literature.
@@
 
-\cite{jern2013probabilistic} recently showed that creative generation could be studied in a more controlled manner through the well-developed methods of an artificial categorization paradigm \citep[see][for a review]{kurtz2015human}.
+\cite{jern2013probabilistic} recently showed that concept generation could be studied in a more controlled manner through the well-developed methods of an artificial categorization paradigm \citep[see][for a review]{kurtz2015human}.
 In Experiments 3 and 4 of their article, participants were exposed to members of experimenter-defined categories of "crystals" varying in size, hue, and saturation.
 Following a training phase during which the experimenter-defined categories were learned, participants were asked to generate novel categories of crystals.
-In a finding mirroring that of the \cite{ward1994structured} studies, \cite{jern2013probabilistic} found that participants generated categories with the same distributional properties as the experimenter-defined categories.
+In a finding mirroring \cite{ward1994structured}, \cite{jern2013probabilistic} found that participants generated categories with the same distributional properties as the experimenter-defined categories.
 For example, after training on categories with a positive correlation between the size and saturation features (larger sized crystals were more saturated), participants generated novel categories with the same positive correlation.
@@
 
-Prior work on category generation has explored only one of the possible constraints that guide category generation.
-The main concern of the published experiments has been on the distributional correspondences between learned and generated categories, and as a result most of the computational, theoretical, and empirical efforts have been focused on explaining those effects.
-In this paper, we investigate another important constraint: category contrast.
+Prior work on category generation has explored only one factor: distributional correspondences between learned and generated categories.
+As a result most of the computational, theoretical, and empirical efforts have been focused on explaining those effects.
+In this paper, we investigate another important factor: category contrast.
 To generate a novel concept, individuals must produce something that is in some capacity {\em different} from what they already know.
-By consequence, we propose that contrast should be a primary constraint on creative generation: New concepts should be different from existing ones.
+By consequence, we propose that contrast should be a primary constraint that guides how people create new categories: New categories should be different from existing ones.
+
+\jlanote{check to make sure the next paragraph has been updated enough}
 
@@
 Many of these models assume that at some point of the latent generative process the same underlying distribution generates all of the categories and thus, any differences between categories are due to {\em noise} and should not be {\em systematic}.
-The best a standard hierarchical Bayesian model can do at capturing contrast is to assume that the new category is placed uniformly at random over stimulus space.
+If the goal of creating a new category given others in a domain is generating from the posterior predictive distribution then, the best a standard hierarchical Bayesian model without a notion of contrast built into the model can do at capturing contrast is to assume that the new category is placed uniformly at random over stimulus space.
 But, this defeats the purpose of a hierarchy as it is ignored when determining a new category location!\footnote{It is plausible that some hierarchical Bayesian model could be created that generates categories different from each other.
 However, this model would not be a standard application or extension of most pre-existing hierarchical Bayesian models.
-The generative process would need to include a component that promotes contrast.}
+The generative process would need to include a component that promotes contrast.} Note that this does not mean location information and contrast cannot be captured using hierarchical Bayesian models.
+In fact, our model of contrast as representativeness is a hierarchical Bayesian model, but with a different goal than predicting the next category.
 
@@
 
-Although existing accounts of creative generation broadly overlook the role of category contrast in determining what is novel versus familiar, it was implicitly assumed that learners in previous experiments were {\em successful} in creating new categories.
-Thus, effects of category contrast should be observable within the experimental results of these studies.
-To provide a test of the influence of category contrast within existing data, we conducted a novel analysis of Experiment 3 from \cite{jern2013probabilistic}\footnote{Although \cite{jern2013probabilistic} reported four experiments, we focus on Experiment 3 because their first two experiments tested generation of items into known categories, and their fourth experiment was identical to Experiment 3 but with a restricted generation space.}.
+Although existing accounts of category generation broadly overlook the role of category contrast in determining what is novel versus familiar, it was implicitly assumed that learners in previous experiments were {\em successful} in creating new categories.
+Thus, if contrast plays a robust role in category genration, its effects should be discernable within the experimental results of these studies.
+To provide a test of the influence of category contrast within existing data, we conducted a novel analysis of Experiment 3 from \cite{jern2013probabilistic}.\footnote{Although \cite{jern2013probabilistic} reported four experiments, we focus on Experiment 3 because their first two experiments tested generation of items into known categories, and their fourth experiment was identical to Experiment 3 but with a restricted generation space.}
 
@@
 Note that the purpose of the article is to explore and formally analyze the role of contrast in category generation and thus, we leave extending PACKER to incorporate distributional factors \citep[as explored by][]{jern2013probabilistic} for future work.
+
+\subsection{Contrast as representativeness: An extension to Jern and Kemp (2013)}
+
+TODO: fill in Bayesian definition of representativeness and its application in JK2013.
 
@@
 
+\subsection{Representativeness in Categorization}
+MAKE SURE WE CITE AND DESCRIBE \citet{thomas98}!
+
 \subsection{Limitations and Future Directions}

### 2018-07-10 Manuscripts/cog-psych/revision/cogpsych.tex
minor fixes to bib file and intro

@@
 
-Previous work in category learning and generation has established that people are highly sensitive to the structural properties of categories, such as correlations between the features of category members and the relation between items within the same category and those in different categories \citep{roschmervis1975,regier2007,shepard1961learning,smith1995}.
+Previous work in category learning and generation has established that people are highly sensitive to the structural properties of categories, such as correlations between the features of category members and the relation between items within the same category and those in different categories \citep{roschmervis1975,regier2007,shepard1961learning,smith1995creative}.
 Inspired by this work, previous research on the topic of category generation has explored a similar principle: People tend to create new categories that have similar {\em statistical regularities} as previously learned categories \citep{jern2013probabilistic,ward1994structured}.
@@
 
-The second hypothesis for how contrast affects categorization is contrast as {\em representativeness} \citep{kahneman73}.
-The representativeness heuristic states that people make judgments regarding the probability of an outcome based on how representative it is of the observed evidence.
-The heuristic is a powerful theoretical construct that has been used to capture a range of complex patterns of human judgments \citep{kahneman73,tversky74,tversky83}, especially those that deviate from normative theory (as given by Bayesian inference).
-For example, the coinflip sequence TTHTH is perceived to be more random than TTTTT (despite both having the same probability of having been generated from a fair coin).
+The second hypothesis for how contrast might affect categorization is contrast as {\em representativeness}.
+The representativeness heuristic \citep{kahneman73} states that people make judgments regarding an outcome based on how representative it is of the evidence given in the current context.
+The heuristic is a powerful theoretical construct that has been used to capture a range of complex patterns of human judgments \citep{kahneman73,tversky74,tversky83}, especially those that deviate from normative theory (as given by a straightforward application of Bayes' rule).
+For example, the coinflip sequence TTHTH is perceived to be more random (due to it being representative of a coinflip sequence) than TTTTT (despite both having the same probability of having been generated from a fair coin).
 Although there is a healthy debate as to whether perceived randomness {\em really} deviates from normative theory \citep{griffiths18,hahn09}, the representativeness heuristic remains one of the main explanations of human judgments \citep{reimers18}.

### 2018-07-10 Manuscripts/cog-psych/revision/cogpsych.tex
to xian

@@
 The second assumes the goal of category generation is to create representative samples of the opposite of the observed categories.
-We present two experiments demonstrating strong and systematic effects of category contrast on creative generation, and we qualitatively and quantitatively analyze the performance of each model in capturing human category generation.
+We present two experiments demonstrating strong and systematic effects of category contrast on concept generation, and we qualitatively and quantitatively analyze the performance of each model in capturing human category generation.
+Afterwards, we find support for category contrast predicting the ease at which people learn a pair of novel categories in an empirical experiment.
 We conclude with a discussion of the implications of our results for categorization and creative cognition, and directions for future work.
@@
 
-\subsection{The PACKER Model}
+\subsection{Contrast as exemplar dissimilarity: The PACKER Model}
 
@@
 
-In the sections below, we present two experiments demonstrating systematic effects of category contrast on creative generation inspired by factors influencing how PACKER generates new categories.
+In the sections below, we present two experiments demonstrating systematic effects of category contrast on category generation inspired by factors influencing how the contrast models generate new categories.
 Our experiments are based on \cite{jern2013probabilistic}'s paradigm, which is a straightforward translation of the traditional artificial classification paradigm to the task of category generation: Participants are first exposed to a single, experimenter-defined category, and are then asked to generate members of a new category.
-We then report formal analyses comparing PACKER's account of our results to that of the hierarchical Bayesian and copy-and-tweak models developed by \cite{jern2013probabilistic}.
+We then report formal analyses comparing the exemplar dissimilarity and the representativeness accounts of our results to that of the hierarchical Bayesian and copy-and-tweak models developed by \cite{jern2013probabilistic}.
+After analyzing the category generation studies, we present third experiment to demonstrate the effect of contrast in category learning.
+This ensures that our results are not idiosyncratic to category generation.
 
@@
 \subsection{Discussion}
-
 In Experiment 1 we sought to extend our analysis of the \cite{jern2013probabilistic} data by evaluating the influence of category contrast on creative generation, given qualitatively different types of prior knowledge.
@@
 
-\subsection{Representativeness in Categorization}
+\subsection{Contrast and Representativeness in Categorization}
+Xian: Please write about the two hypotheses and that followup work needs to look into why the Bayesian representativeness model doesn't do as well at explaining category learning ease, but does better at category generation.
+
 MAKE SURE WE CITE AND DESCRIBE \citet{thomas98}!
@@
 \subsection{Limitations and Future Directions}
-
-Although successful in explaining our results, PACKER does not provide a full account of what is known about category generation.
+Although we have demonstrated that contrast plays a role in categorization, our results were equivocal as to whether contrast is governed by exemplar dissimilarity or representativeness.
+
+Although successful in explaining our results in all three studies (albeit not as well as the representativeness model for Experiments 1 and 2), PACKER does not provide a full account of what is known about category generation.
 Most notably, in this paper we have not evaluated the model's ability to explain the classic finding that generated categories tend to share distributional commonalities with previously learned categories \citep[see][]{jern2013probabilistic,ward1994structured}.
@@
 Even within Experiment 1, we found systematic inconsistencies: by generating exemplars into unoccupied regions of the space, participants who learned an `XOR' category, composed of members that are widely distributed along both features and are positively correlated in space, tended to generate categories with an opposite (negative) correlation.
-More generally, PACKER's success over the hierarchical Bayesian model indicates that the emulation of distributional structure exerts only a limited influence in comparison to category contrast in some situations.
+More generally, PACKER inherits the strengths and weaknesses of exemplar models of categorization: It provides a simple and flexible model that explains many results, but deviates systematically from human performance in some cases.
 
@@
 Once formalized in probabilistic terms, it should also be straightforward to incorporate distributional factors into the model.
-Alternatively, it may be possible to integrate the core principles of PACKER into other categorization models \citep[e.g.,][]{love2004sustain,kurtz2007divergent,smith2000thirty}.
-
-Finally, although an overall goal of this work is to advance the study of creative thought by explaining generation through empirical and quantitative approaches from research in categorization, these approaches are not easily aligned with many of the alternative theoretical viewpoints in the broader study of creativity \citep[for a review see][]{kozbelt2010theories}, such as those based on free association \citep{mednick1962associative} and conceptual combination \citep{estes2002emergence,murphy1988comprehending}.
+This would unite PACKER with the Bayesian representativeness model, and they would differ in terms of their assumptions about how people represent category distributions (as exemplars or prototypes, respectively).
+Alternatively, it may be possible to integrate the core principles of contrast into other categorization models \citep[e.g.,][]{love2004sustain,kurtz2007divergent,smith2000thirty}.
+
+%Finally, although an overall goal of this work is to advance the study of creative thought by explaining generation through empirical and quantitative approaches from research in categorization, these approaches are not easily aligned with many of the alternative theoretical viewpoints in the broader study of creativity \citep[for a review see][]{kozbelt2010theories}, such as those based on free association \citep{mednick1962associative} and conceptual combination \citep{estes2002emergence,murphy1988comprehending}.
 Instead, we reduced our focus by studying a highly complex behavior (creative generation) as it applies within a well-established domain (artificial category learning).
@@
 The generation of new concepts and ideas is a highly interesting topic, but it is difficult to study in a controlled experimental environment.
-In this paper, we have provided such an examination of category generation as it applies within an artificial categorization categorization experiment.
+In this paper, we have provided such an examination of category generation as it applies within an artificial categorization experiment.
 Extending the literature on creative cognition, our experiments provide a detailed picture of the role of category contrast in generation: People seek to create concepts that are distinct from what they already know, and the nature of what is created can be influenced by what does not yet exist.

### 2018-07-10 Manuscripts/cog-psych/revision/methods_results_e3.tex
to xian

@@
 % \begin{document}
-\section{Experiment 3.}
+\section{Experiment 3}
+Experiments 1 and 2 clearly establish the importance of contrast in category generation.
+Follow-up model-based analyses illustrated that both contrast models account for participant performance better than models that do not take into account contrast.
+We also found more support for contrast as representativeness hypothesis over contrast as exemplar dissimilarity hypothesis.
 
-[I'm assuming Joe will include some preamble for the experiment, but otherwise
-it should be fairly straightforward for me to churn out one anyway.]
+Although category generation is an intriguing aspect of human cognition, it has played a relatively limited role in the broader categorization literature.
+To establish that contrast plays a broader role in categorization, we conducted a traditional category learning study.
+In it, participants learned two categories, where the categories learned by each participant in Experiment 3 was yoked to the Alpha shown and Beta generated by each participant in Experiment 2.
+If contrast plays a broader role in categorization, then Beta categories rated as "better" according to the contrast models should be easier for participants to learn.
 

### 2018-07-10 Manuscripts/cog-psych/revision/cogpsych.tex
cut paragraph

@@
 
-Nonetheless, these classic effects are a core element of the phenomenology of creative generation, and PACKER does not include any mechanisms that explain them.
+%Nonetheless, these classic effects are a core element of the phenomenology of creative generation, and PACKER does not include any mechanisms that explain them.
 Instead, through the development and evaluation of the PACKER model, we have sought to add new elements into such a phenomenology: The broad and strong influence of category contrast, and the interdependence between category location and distributional structure.

### 2018-07-17 Manuscripts/cog-psych/revision/cogpsych.tex
Added representativeness to manuscript. Updated figures and results. Discussion in various places and some reviewer-related responses still need work.

--- 2018-07-10 Manuscripts/cog-psych/revision/cogpsych.tex
@@
 %!TEX output_directory = latex_out/
-
 \documentclass[12pt]{article}
@@
 \newcommand\jlanote[1]{\todo[inline, backgroundcolor = lime]{\textbf{JLA}: #1}}
+\newcommand\sxlnote[1]{\todo[inline, backgroundcolor = cyan]{\textbf{SXL}: #1}}
 
@@
 Nolan Conaway\textsuperscript{1}, 
-Shi X. Liew\textsuperscript{1},
+Shi Xian Liew\textsuperscript{1},
 and Kenneth J. Kurtz\textsuperscript{2}
@@
 \jlanote{need to update} 
-The ability to creatively generate new concepts and ideas is among the most fascinating aspects of human cognition, but we do not have a strong understanding of the cognitive processes and representations underlying concept generation.
-In this paper, we study the generation of new categories using the computational and behavioral toolkit of traditional artificial category learning.
-Previous work in this domain has focused on how the statistical structure of known categories generalizes to generated categories, overlooking the degree of contrast between the known and generated categories as a factor.
-We report two experiments demonstrating that contrast between what is known and what is created is of fundamental importance for categorization and the creative process.
-We propose a novel, exemplar-based approach explaining our results, and compare the model's performance to two key alternatives.
-Our experiments and simulations demonstrate specifically how category contrast influences creative generation, and how it can interact with other constraints to produce different types of creations.
-Our work also serves as a concrete example of how well-established and highly controlled experimental frameworks (such as those in the field of category learning) can be used to make progress in an intriguing yet understudied domain.
+\sxlnote{updated} 
+%
+The ability to creatively generate new concepts and ideas is among the most
+fascinating aspects of human cognition, but we do not have a strong
+understanding of the cognitive processes and representations underlying concept
+generation. In this paper, we study the generation of new categories using the
+computational and behavioral toolkit of traditional artificial category
+learning. Previous work in this domain has focused on how the statistical
+structure of known categories generalizes to generated categories, overlooking
+the degree of contrast between the known and generated categories as a factor.
+We report three experiments demonstrating that contrast between what is known
+and what is created is of fundamental importance for categorization and the
+creative process. We propose two novel approaches to modeling category contrast:
+one focused on exemplar dissimilarity and another on the representativeness
+heuristic. These new models are compared to two key alternatives. Our
+experiments and simulations demonstrate specifically how category contrast
+influences creative generation, and how it can interact with other constraints
+to produce different types of creations. Our work also serves as a concrete
+example of how well-established and highly controlled experimental frameworks
+(such as those in the field of category learning) can be used to make progress
+in an intriguing yet understudied domain.
 
@@
 
-Creating and learning new ideas is one of the most fascinating and important human capabilities.
-For example, cell phones and smart phones are newly created categories of objects that young and old adults learned and have become reliant on in the last few decades.
-In fact, computational researchers have identified generating novel objects and objectives as one of the defining, and most difficult to formalize, characteristics of intelligent life \citep{lake2017,lehman2011,taylor2016}.
-Yet, the cognitive mechanisms that enable people to innovate and learn new innovations are not well understood and are understudied by scientists.
-In part, this is due to the inherent difficulty of designing and conducting experiments that test these capabilities.
-How would a scientist devise an experiment that induces a participant to create and learn new categories that are as interesting as cell phones and smart phones?
-In this article, we use computational models and behavioral experiments to investigate the prior expectations people have when they learn and generate new categories.
-
-Generating and learning new concepts are, however, not altogether different from the types of behaviors typically studied in cognitive psychology laboratories.
-In particular, generating a member of a novel class can be considered a `special case' application of existing category knowledge \citep{kemp2014taxonomy,kurtz2015human}.
-Research in categorization typically focuses on what properties of a category affect the classification of an object's category given its features \citep{shepard1961learning,kurtz2013human}, and the prediction of an object's unobserved features given some of its other features and/or its category \citep{markman2003category}.
-The generation of members of a new category consists of inferring {\em all} features for a {\em novel} category label.
-Thus, we can make progress formalizing the processes involved in category generation by extending theories of categorization to this case.
-Then, to ensure that these principles reflect basic principles of categorization and are not just specific to category generation, we can examine what factors of the generated categories affect the ease at which people learn the category.
-
-Previous work in category learning and generation has established that people are highly sensitive to the structural properties of categories, such as correlations between the features of category members and the relation between items within the same category and those in different categories \citep{roschmervis1975,regier2007,shepard1961learning,smith1995creative}.
-Inspired by this work, previous research on the topic of category generation has explored a similar principle: People tend to create new categories that have similar {\em statistical regularities} as previously learned categories \citep{jern2013probabilistic,ward1994structured}.
-Although this is an important characteristic of generating new categories, it cannot be the only one.
-Taken to the extreme, the best ``new'' category in terms of having the same statistical regularities to other categories would be identical to a known category that is representative of the domain (and thus, not new at all).
-
-To successfully generate something novel, what is generated must be different from what is already known.
-This fundamental constraint, ``being different'', or contrasting from other categories in the relevant domain, is the focus of our work.
-Although implicitly assumed in some work, this constraint has been overlooked in previous research: To our knowledge, there has not been any systematic investigation addressing how generated categories {\em differ} from what is already known.
-Although the idea of category contrast is ubiquitous throughout the categorization literature, and extends to a variety of other fields \citep[e.g., color;][]{regier2007}, the idea that a new category should be ``different'' is vague, as there are many ways it could be different from a previously observed category.
-
-We examine two different definitions of how categories should differ from one another: exemplar contrast or representative of the alternative category.
-To formalize the first technique, we build on the largely successful exemplar modeling framework that \citep{medin1978context,nosofsky1984choice,nosofsky1986attention}.
-In doing so, we propose a novel exemplar model of category generation, {\em Producing Alike and Contrasting Knowledge using Exemplar Representations} (PACKER), formalizing how new categories should differ from previous categories.
-It embodies the exemplar contrast principle by incorporating a factor that repulses members of opposite categories from one another.
-To foreshadow, we find that this notion of contrast predicts human category learning performance, above and beyond other category learning models in a novel behavioral experiment.
-This demonstrates that contrast plays a central role in other important categorization tasks, which is not specific to category generation.
-
-The second hypothesis for how contrast might affect categorization is contrast as {\em representativeness}.
-The representativeness heuristic \citep{kahneman73} states that people make judgments regarding an outcome based on how representative it is of the evidence given in the current context.
-The heuristic is a powerful theoretical construct that has been used to capture a range of complex patterns of human judgments \citep{kahneman73,tversky74,tversky83}, especially those that deviate from normative theory (as given by a straightforward application of Bayes' rule).
-For example, the coinflip sequence TTHTH is perceived to be more random (due to it being representative of a coinflip sequence) than TTTTT (despite both having the same probability of having been generated from a fair coin).
-Although there is a healthy debate as to whether perceived randomness {\em really} deviates from normative theory \citep{griffiths18,hahn09}, the representativeness heuristic remains one of the main explanations of human judgments \citep{reimers18}.
-To formalize the hypothesis that people expect a new category to be representative of the opposite of the current category, or contrast as {\em representativeness}, we use the Bayesian formulation of representativeness \citep{tenenbaum01}, which has been used successfully to capture human performance in color categorization \citep{abbott16}, image category learning \citep{abbott11}, and language grammar learning \citep{rafferty10}.
-We do so by extending the state-of-the-art category generation model \citep{jern2013probabilistic} and find that  human category generation is best capture as generating representative, rather than probable, samples.
-
-The outline of the article is as follows.
-First we describe previous empirical work on the topic of category generation, as well as the computational formalizations of the theories in those reports.
-We then present two hypotheses for how contrast might affect categorization and formalize them in computational models.
-The first is a novel computational model, which is designed to generate categories that systematically differ from existing categories in the domain.
-The second assumes the goal of category generation is to create representative samples of the opposite of the observed categories.
-We present two experiments demonstrating strong and systematic effects of category contrast on concept generation, and we qualitatively and quantitatively analyze the performance of each model in capturing human category generation.
-Afterwards, we find support for category contrast predicting the ease at which people learn a pair of novel categories in an empirical experiment.
-We conclude with a discussion of the implications of our results for categorization and creative cognition, and directions for future work.
+Creating and learning new ideas is one of the most fascinating and important
+human capabilities. For example, cell phones and smart phones are newly created
+categories of objects that young and old adults learned and have become reliant
+on in the last few decades. In fact, computational researchers have identified
+generating novel objects and objectives as one of the defining, and most
+difficult to formalize, characteristics of intelligent life
+\citep{lake2017,lehman2011,taylor2016}. Yet, the cognitive mechanisms that
+enable people to innovate and learn new innovations are not well understood and
+are understudied by scientists. In part, this is due to the inherent difficulty
+of designing and conducting experiments that test these capabilities. How would
+a scientist devise an experiment that induces a participant to create and learn
+new categories that are as interesting as cell phones and smart phones? In this
+article, we use computational models and behavioral experiments to investigate
+the prior expectations people have when they learn and generate new categories.
+
+Generating and learning new concepts are, however, not altogether different from
+the types of behaviors typically studied in cognitive psychology laboratories.
+In particular, generating a member of a novel class can be considered a `special
+case' application of existing category knowledge
+\citep{kemp2014taxonomy,kurtz2015human}. Research in categorization typically
+focuses on what properties of a category affect the classification of an
+object's category given its features \citep{shepard1961learning,kurtz2013human},
+and the prediction of an object's unobserved features given some of its other
+features and/or its category \citep{markman2003category}. The generation of
+members of a new category consists of inferring {\em all} features for a {\em
+novel} category label. Thus, we can make progress formalizing the processes
+involved in category generation by extending theories of categorization to this
+case. Then, to ensure that these principles reflect basic principles of
+categorization and are not just specific to category generation, we can examine
+what factors of the generated categories affect the ease at which people learn
+the category.
+
+Previous work in category learning and generation has established that people
+are highly sensitive to the structural properties of categories, such as
+correlations between the features of category members and the relation between
+items within the same category and those in different categories
+\citep{roschmervis1975,regier2007,shepard1961learning,smith1995creative}.
+Inspired by this work, previous research on the topic of category generation has
+explored a similar principle: People tend to create new categories that have
+similar {\em statistical regularities} as previously learned categories
+\citep{jern2013probabilistic,ward1994structured}. Although this is an important
+characteristic of generating new categories, it cannot be the only one. Taken to
+the extreme, the best ``new'' category in terms of having the same statistical
+regularities to other categories would be identical to a known category that is
+representative of the domain (and thus, not new at all).
+
+\sxlnote{I think we can address Charles' point on real world examples of how
+  contrast can be important around here.}
+To successfully generate something novel, what is generated must be different
+from what is already known. This fundamental constraint, ``being different'', or
+contrasting from other categories in the relevant domain, is the focus of our
+work. Although implicitly assumed in some work, this constraint has been
+overlooked in previous research: To our knowledge, there has not been any
+systematic investigation addressing how generated categories {\em differ} from
+what is already known. Although the idea of category contrast is ubiquitous
+throughout the categorization literature, and extends to a variety of other
+fields \citep[e.g., color;][]{regier2007}, the idea that a new category should
+be ``different'' is vague, as there are many ways it could be different from a
+previously observed category.
+
+We examine two different definitions of how categories should differ from one
+another: exemplar dissimilarity or representativeness of the alternative
+category. To formalize the first technique, we build on the largely successful
+exemplar modeling framework that
+\citep{medin1978context,nosofsky1984choice,nosofsky1986attention}. In doing so,
+we propose a novel exemplar model of category generation, {\em Producing Alike
+  and Contrasting Knowledge using Exemplar Representations} (PACKER),
+formalizing how new categories should differ from previous categories. It
+embodies the exemplar dissimilarity principle by incorporating a factor that
+repulses members of opposite categories from one another. To foreshadow, we find
+that this notion of contrast predicts human category learning performance, above
+and beyond other category learning models in a novel behavioral experiment. This
+demonstrates that contrast plays a central role in other important
+categorization tasks, which is not specific to category generation.
+
+The second hypothesis for how contrast might affect categorization is contrast
+as {\em representativeness}. The representativeness heuristic \citep{kahneman73}
+states that people make judgments regarding an outcome based on how
+representative it is of the evidence given in the current context. The heuristic
+is a powerful theoretical construct that has been used to capture a range of
+complex patterns of human judgments \citep{kahneman73,tversky74,tversky83},
+especially those that deviate from normative theory (as given by a
+straightforward application of Bayes' rule). For example, the coinflip sequence
+TTHTH is perceived to be more random (due to it being representative of a
+coinflip sequence) than TTTTT (despite both having the same probability of
+having been generated from a fair coin). Although there is a healthy debate as
+to whether perceived randomness {\em really} deviates from normative theory
+\citep{griffiths18,hahn09}, the representativeness heuristic remains one of the
+main explanations of human judgments \citep{reimers18}. To formalize the
+hypothesis that people expect a new category to be representative of the
+opposite of the current category, or contrast as {\em representativeness}, we
+use the Bayesian formulation of representativeness \citep{tenenbaum01}, which
+has been used successfully to capture human performance in color categorization
+\citep{abbott16}, image category learning \citep{abbott11}, and language grammar
+learning \citep{rafferty10}. We do so by extending the state-of-the-art category
+generation model \citep{jern2013probabilistic} and find that human category
+generation is best captured as generating representative, rather than probable,
+samples.
+
+The outline of the article is as follows. First we describe previous empirical
+work on the topic of category generation, as well as the computational
+formalizations of the theories in those reports. We then present two hypotheses
+for how contrast might affect categorization and formalize them in computational
+models. The first is a novel computational model, which is designed to generate
+categories that systematically differ from existing categories in the domain.
+The second assumes the goal of category generation is to create representative
+samples of the opposite of the observed categories. We present two experiments
+demonstrating strong and systematic effects of category contrast on concept
+generation, and we qualitatively and quantitatively analyze the performance of
+each model in capturing human category generation. Afterwards, we find support
+for category contrast predicting the ease at which people learn a pair of novel
+categories in an empirical experiment. We conclude with a discussion of the
+implications of our results for categorization and creative cognition, and
+directions for future work.
 
@@
 
-Much of what we know about concept generation comes from the foundational literature on creative cognition.
-In a series of reports, Ward and colleagues \citep{ward1995s,ward1994structured,marsh1999inadvertent,ward2002role,smith1993constraining} established that category generation is highly constrained by prior knowledge: Generated categories tend to consist of features observed in known categories, and they tend to exhibit the distributional properties as found in known categories.
-In a seminal study, \cite{ward1994structured} asked participants to generate new species of alien animals by drawing and describing members of the species.
-People tended to generate species with the same features as on Earth (e.g., eyes, legs, wings), and possessing the same feature correlations as on Earth (e.g., feathers co-occur with wings).
-Likewise, aliens drawn from the same species tended to share more features with one another compared to members of opposite species.
-
-The broader set of observations made by Ward and colleagues provide a great deal of insight into the role of prior knowledge in constraining category generation.
-Much of the work from this area \citep[e.g.,][]{smith1993constraining,marsh1999inadvertent} focuses on how information provided to participants (such as an example of a species generated by other participants) can drastically diminish difference of a new category from pre-existing categories.
-Theoretical accounts of these effects have primarily been grounded within the categorization literature.
-For example, the predominant ``Path of Least Resistance'' account \citep[see][]{ward1994structured,ward1995s,ward2002role} proposes that, when generating a new species of animal, people retrieve from memory a known subcategory of animals (e.g., {\em bird}, {\em dog}, {\em horse}), and simply change some of the features to make something new.
-People are thought to change only features that are not characteristic of the retrieved category (e.g., if {\em bird} was retrieved, the presence of {\em wings} would not change, but {\em color} might).
-This theory incorporates elements of the highly influential basic-level categories framework \citep{rosch1975cognitive,rosch1976basic}, as well as the exemplar view \citep{medin1978context,brooks1978nonanalytic}.
-While this work is been incredibly useful in providing a conceptual sketch of generation theories, the hand-drawn responses used in the experiments paradigms precludes the development of formal approaches that can be used to test the theories in a fine-grained manner.
-
-\cite{jern2013probabilistic} recently showed that concept generation could be studied in a more controlled manner through the well-developed methods of an artificial categorization paradigm \citep[see][for a review]{kurtz2015human}.
-In Experiments 3 and 4 of their article, participants were exposed to members of experimenter-defined categories of "crystals" varying in size, hue, and saturation.
-Following a training phase during which the experimenter-defined categories were learned, participants were asked to generate novel categories of crystals.
-In a finding mirroring \cite{ward1994structured}, \cite{jern2013probabilistic} found that participants generated categories with the same distributional properties as the experimenter-defined categories.
-For example, after training on categories with a positive correlation between the size and saturation features (larger sized crystals were more saturated), participants generated novel categories with the same positive correlation.
-This finding is notable, as it demonstrates that category generation can be studied in a well-known and highly controlled experimental paradigm.
+Much of what we know about concept generation comes from the foundational
+literature on creative cognition. In a series of reports, Ward and colleagues
+\citep{ward1995s,ward1994structured,marsh1999inadvertent,ward2002role,smith1993constraining}
+established that category generation is highly constrained by prior knowledge:
+Generated categories tend to consist of features observed in known categories,
+and they tend to exhibit the distributional properties as found in known
+categories. In a seminal study, \cite{ward1994structured} asked participants to
+generate new species of alien animals by drawing and describing members of the
+species. People tended to generate species with the same features as on Earth
+(e.g., eyes, legs, wings), and possessing the same feature correlations as on
+Earth (e.g., feathers co-occur with wings). Likewise, aliens drawn from the same
+species tended to share more features with one another compared to members of
+opposite species.
+
+\sxlnote{Added thomas98 description here} It is worth pointing out that the
+observation of regularities in distributional properties across categories is
+not confined to the work of Ward and colleagues. Most notably, \citet{thomas98}
+found that when people were asked to generate feature values for exemplars with
+a missing feature (specifically circles of different sizes, each with a radial
+line of varying orientation), the resulting set of exemplars tended to share the
+same feature correlations as exemplars from previously learned categories. [the
+transition to the next point isn't great, though]
+
+The broader set of observations made by Ward and colleagues provide a great deal
+of insight into the role of prior knowledge in constraining category generation.
+Much of the work from this area
+\citep[e.g.,][]{smith1993constraining,marsh1999inadvertent} focuses on how
+information provided to participants (such as an example of a species generated
+by other participants) can drastically diminish difference of a new category
+from pre-existing categories. Theoretical accounts of these effects have
+primarily been grounded within the categorization literature. For example, the
+predominant ``Path of Least Resistance'' account
+\citep[see][]{ward1994structured,ward1995s,ward2002role} proposes that, when
+generating a new species of animal, people retrieve from memory a known
+subcategory of animals (e.g., {\em bird}, {\em dog}, {\em horse}), and simply
+change some of the features to make something new. People are thought to change
+only features that are not characteristic of the retrieved category (e.g., if
+{\em bird} was retrieved, the presence of {\em wings} would not change, but {\em
+color} might). This theory incorporates elements of the highly influential
+basic-level categories framework \citep{rosch1975cognitive,rosch1976basic}, as
+well as the exemplar view \citep{medin1978context,brooks1978nonanalytic}. While
+this work is been incredibly useful in providing a conceptual sketch of
+generation theories, the hand-drawn responses used in the experiments paradigms
+precludes the development of formal approaches that can be used to test the
+theories in a fine-grained manner.
+
+\cite{jern2013probabilistic} recently showed that concept generation could be
+studied in a more controlled manner through the well-developed methods of an
+artificial categorization paradigm \citep[see][for a review]{kurtz2015human}. In
+Experiments 3 and 4 of their article, participants were exposed to members of
+experimenter-defined categories of "crystals" varying in size, hue, and
+saturation. Following a training phase during which the experimenter-defined
+categories were learned, participants were asked to generate novel categories of
+crystals. In a finding mirroring \cite{ward1994structured},
+\cite{jern2013probabilistic} found that participants generated categories with
+the same distributional properties as the experimenter-defined categories. For
+example, after training on categories with a positive correlation between the
+size and saturation features (larger sized crystals were more saturated),
+participants generated novel categories with the same positive correlation. This
+finding is notable, as it demonstrates that category generation can be studied
+in a well-known and highly controlled experimental paradigm.
 
 The authors evaluated the predictions of several formal models on their data.
-Most notably, they showed that a hierarchical Bayesian model provided the strongest account of their results.
-Their model views observed examples as samples from an underlying category distribution, describing the location of the category in the space, as well as how it varies along each feature.
-In turn, each category is viewed as a sample from an underlying {\em domain} distribution, specifying distributional commonalities among the observed categories.
-Generated categories are thought to stem from the same domain distribution as observed categories, thus the distributional properties of observed categories will be preserved within the generated category.
-
-\cite{jern2013probabilistic} additionally tested a ``copy-and-tweak'' model that broadly resembles the earlier ``Path of Least Resistance'' account.
-The core proposal is that participants generate new items by copying stored examples from memory and tweaking them to generate something new.
-The copy-and-tweak model differs from the Path of Least Resistance account in that it notably omits the hierarchical organization of categories, as well as selectivity in which features are changed \citep[both of which are factors in the Path of Least Resistance account;][]{ward2002role}.
-Instead, their copy-and-tweak model corresponds to a direct exemplar-similarity approach \citep[e.g.,][]{nosofsky1984choice,nosofsky1986attention}, generating new items according to their similarity to known members of the target category.
-The copy-and-tweak model provided a poor account of their results, as the experiments devised by \citeauthor{jern2013probabilistic} were specifically designed to challenge it.
-However, its application is notable as a first step toward understanding theories developed in the creative cognition literature using well-known formal approaches from the categorization literature.
+Most notably, they showed that a hierarchical Bayesian model provided the
+strongest account of their results. Their model views observed examples as
+samples from an underlying category distribution, describing the location of the
+category in the space, as well as how it varies along each feature. In turn,
+each category is viewed as a sample from an underlying {\em domain}
+distribution, specifying distributional commonalities among the observed
+categories. Generated categories are thought to stem from the same domain
+distribution as observed categories, thus the distributional properties of
+observed categories will be preserved within the generated category.
+
+\cite{jern2013probabilistic} additionally tested a ``copy-and-tweak'' model that
+broadly resembles the earlier ``Path of Least Resistance'' account. The core
+proposal is that participants generate new items by copying stored examples from
+memory and tweaking them to generate something new. The copy-and-tweak model
+differs from the Path of Least Resistance account in that it notably omits the
+hierarchical organization of categories, as well as selectivity in which
+features are changed \citep[both of which are factors in the Path of Least
+Resistance account;][]{ward2002role}. Instead, their copy-and-tweak model
+corresponds to a direct exemplar-similarity approach
+\citep[e.g.,][]{nosofsky1984choice,nosofsky1986attention}, generating new items
+according to their similarity to known members of the target category. The
+copy-and-tweak model provided a poor account of their results, as the
+experiments devised by \citeauthor{jern2013probabilistic} were specifically
+designed to challenge it. However, its application is notable as a first step
+toward understanding theories developed in the creative cognition literature
+using well-known formal approaches from the categorization literature.
 
@@
 
-Prior work on category generation has explored only one factor: distributional correspondences between learned and generated categories.
-As a result most of the computational, theoretical, and empirical efforts have been focused on explaining those effects.
-In this paper, we investigate another important factor: category contrast.
-To generate a novel concept, individuals must produce something that is in some capacity {\em different} from what they already know.
-By consequence, we propose that contrast should be a primary constraint that guides how people create new categories: New categories should be different from existing ones.
+Prior work on category generation has explored only one factor: distributional
+correspondences between learned and generated categories. As a result most of
+the computational, theoretical, and empirical efforts have been focused on
+explaining those effects. In this paper, we investigate another important
+factor: category contrast. To generate a novel concept, individuals must produce
+something that is in some capacity {\em different} from what they already know.
+By consequence, we propose that contrast should be a primary constraint that
+guides how people create new categories: New categories should be different from
+existing ones.
 
 \jlanote{check to make sure the next paragraph has been updated enough}
-
-Although it is evident that people are {\em capable} of creating new concepts and categories, it is not entirely clear how new concepts are systematically made different from what is already known.
-The hierarchical Bayesian model developed by \cite{jern2013probabilistic} assumes that differences between observed and generated categories are only due to random variation.
-The model assumes that generated categories are sampled from the same underlying domain distribution as observed categories, and will thus share a common distributional structure.
-The model does not make predictions about the {\em location} of the category within the domain (the perceptual instantiation of category members).
-Under a strict interpretation of their model, given knowledge of a single category within the domain, the most probable new category to be generated is located in {\em exactly} the same location and possesses an identical distributional structure.
-This is not an issue with their model specifically, but with a broader class of standard hierarchical Bayesian models \citep[e.g., ][]{griffithsscn07,kemp07}.
-Many of these models assume that at some point of the latent generative process the same underlying distribution generates all of the categories and thus, any differences between categories are due to {\em noise} and should not be {\em systematic}.
-If the goal of creating a new category given others in a domain is generating from the posterior predictive distribution then, the best a standard hierarchical Bayesian model without a notion of contrast built into the model can do at capturing contrast is to assume that the new category is placed uniformly at random over stimulus space.
-But, this defeats the purpose of a hierarchy as it is ignored when determining a new category location!\footnote{It is plausible that some hierarchical Bayesian model could be created that generates categories different from each other.
-However, this model would not be a standard application or extension of most pre-existing hierarchical Bayesian models.
-The generative process would need to include a component that promotes contrast.} Note that this does not mean location information and contrast cannot be captured using hierarchical Bayesian models.
-In fact, our model of contrast as representativeness is a hierarchical Bayesian model, but with a different goal than predicting the next category.
-
-The copy-and-tweak model tested by \cite{jern2013probabilistic} also claims little about how generated categories should contrast with what is already known.
-In their simulations, the model was only tested on generation after the learner had been exposed to members of the target category, and so the model's ability to generate a new category from scratch was not evaluated.
-However, the model's generation is based exclusively on similarity to known members of the {\em target} category; when there are no members of the target category, generation is presumably random.
-
+\sxlnote{I think it's updated enough! If given a chance I might edit this a
+  little to be less harsh on hierarchical bayesian models in general though I
+  think, since the representativeness model is one of them?}
+Although it is evident that people are {\em capable} of creating new concepts
+and categories, it is not entirely clear how new concepts are systematically
+made different from what is already known. The hierarchical Bayesian model
+developed by \cite{jern2013probabilistic} assumes that differences between
+observed and generated categories are only due to random variation. The model
+assumes that generated categories are sampled from the same underlying domain
+distribution as observed categories, and will thus share a common distributional
+structure. The model does not make predictions about the {\em location} of the
+category within the domain (the perceptual instantiation of category members).
+Under a strict interpretation of their model, given knowledge of a single
+category within the domain, the most probable new category to be generated is
+located in {\em exactly} the same location and possesses an identical
+distributional structure. This is not an issue with their model specifically,
+but with a broader class of standard hierarchical Bayesian models \citep[e.g.,
+][]{griffithsscn07,kemp07}. Many of these models assume that at some point of
+the latent generative process the same underlying distribution generates all of
+the categories and thus, any differences between categories are due to {\em
+noise} and should not be {\em systematic}. If the goal of creating a new
+category given others in a domain is generating from the posterior predictive
+distribution then the best a standard hierarchical Bayesian model without a
+notion of contrast built into the model can do at capturing contrast is to
+assume that the new category is placed uniformly at random over stimulus space.
+But, this defeats the purpose of a hierarchy as it is ignored when determining a
+new category location!\footnote{It is plausible that some hierarchical Bayesian
+model could be created that generates categories different from each other.
+However, this model would not be a standard application or extension of most
+pre-existing hierarchical Bayesian models. The generative process would need to
+include a component that promotes contrast.} Note that this does not mean
+location information and contrast cannot be captured using hierarchical Bayesian
+models. In fact, our model of contrast as representativeness is a hierarchical
+Bayesian model, but with a different goal than predicting the next category.
+
+The copy-and-tweak model tested by \cite{jern2013probabilistic} also claims
+little about how generated categories should contrast with what is already
+known. In their simulations, the model was only tested on generation after the
+learner had been exposed to members of the target category, and so the model's
+ability to generate a new category from scratch was not evaluated. However, the
+model's generation is based exclusively on similarity to known members of the
+{\em target} category; when there are no members of the target category,
+generation is presumably random.
+
+\sxlnote{Charles suggested we either remove the section below, or move it to the
+  discussion. I've tried finding a good place for it in the discussion but it
+  doesn't fit very well, and moving it takes away this nice motivation for
+  our further investigation into contrast. I think that keeping it here in the
+  introduction might work out, provided Charles finds my argument for why the
+  regularity assumption is too strong convincing enough. More details a little
+  further down.}
 \subsection{Novel Analyses Demonstrating Contrast Effects in Prior Work}
 
-Although existing accounts of category generation broadly overlook the role of category contrast in determining what is novel versus familiar, it was implicitly assumed that learners in previous experiments were {\em successful} in creating new categories.
-Thus, if contrast plays a robust role in category genration, its effects should be discernable within the experimental results of these studies.
-To provide a test of the influence of category contrast within existing data, we conducted a novel analysis of Experiment 3 from \cite{jern2013probabilistic}.\footnote{Although \cite{jern2013probabilistic} reported four experiments, we focus on Experiment 3 because their first two experiments tested generation of items into known categories, and their fourth experiment was identical to Experiment 3 but with a restricted generation space.}
-
-Participants in their experiment were exposed to members of two experimenter-defined categories of `crystals' varying in hue, saturation, and size.
-Each category possessed a unique hue, but varied in saturation and size: In the `Positive' condition, there was a positive correlation between these features (i.e., larger sized crystals were more saturated), and in the `Negative' condition this relation was reversed.
-In the `Neutral' condition, there was no correlation between saturation and size.
-After learning about the categories from each condition, participants were asked to generate six exemplars belonging to a novel class.
-As noted above, \cite{jern2013probabilistic} found that the generated categories tended to follow the distributional properties of the experimenter-defined categories: Generated categories were tightly distributed along the hue feature, and possessed the same saturation-size correlations as in the learned categories.
-\cite{jern2013probabilistic}, however, did not analyze or discuss how the generated categories {\em differed} from the experimenter-defined categories.
-
-Because each experimenter-defined category in the \cite{jern2013probabilistic} experiment possessed a distinct hue shared by all members of the category, it is sensible that participants might generate a category with a hue distinct from the experimenter-provided categories.
-If creative generation were influenced by category contrast in this way, the hues of generated categories should be systematically different from those of the experimenter-defined categories.
-Unfortunately, stimulus hue was encoded and presented in the Hue-Saturation-Value (HSV) color space, which is device-dependent and not perceptually normed such that perceived color similarity corresponds to proximity in the color space (as opposed to a color space such as CIELAB that is device-independent and equidistant sets of points correspond to pairs of colors that have the same perceptual similarity; \citealp{wyszecki1967}).
-Further, they did not calibrate their monitor, and so we cannot know the precise colors presented to participants.
-As \cite{jern2013probabilistic} were interested in relations between the saturation and length of examples in generated into novel categories, these issues do not undercut their analyses and results.
-However, these issues pose a significant challenge to evaluating contrast between the experimenter-defined and participant-generated categories along the hue dimension.
-It is plausible that two uncalibrated monitors could display the same HSV color and the colors be perceived in different color categories (especially for color boundaries that vary over lightness, such as the yellow-brown boundary).
-
-Although we cannot know the precise colors that were displayed or perceived, we can still analyze their results from a coarse perspective to see whether there is preliminary support for contrast.
-To do so, we binned all possible hues into one of eight uniformly-spaced color groups: \{{\em Red}: $0-0.063$, $0.938-1$, {\em Yellow}: $0.063-0.188$, {\em Yellow-Green}: $0.188-0.313$, {\em Green-Teal}: $0.313-0.438$, {\em Teal}: $0.438-0.563$, {\em Teal-Blue}: $0.563-0.688$, {\em Purple}: $0.688-0.813$, {\em Pink}: $0.813-0.938$\}.
-In the \cite{jern2013probabilistic} experiment, the hue of each experimenter-defined category was selected from one of six possible values, each of which falls into one of the color groups above (two color groups were not used as a possible hue for the experimenter-defined categories).
-By categorizing the participant-generated crystals likewise, we can obtain a broad measure of category contrast by determining the proportion of participant-generated crystals that fall into the same groups as the experimenter-defined categories: If contrast influences the hues of the generated categories, we should observe minimal overlap between in the color groupings.
+Although existing accounts of category generation broadly overlook the role of
+category contrast in determining what is novel versus familiar, it was
+implicitly assumed that learners in previous experiments were {\em successful}
+in creating new categories. Thus, if contrast plays a robust role in category
+genration, its effects should be discernable within the experimental results of
+these studies. To provide a test of the influence of category contrast within
+existing data, we conducted a novel analysis of Experiment 3 from
+\cite{jern2013probabilistic}.\footnote{Although \cite{jern2013probabilistic}
+reported four experiments, we focus on Experiment 3 because their first two
+experiments tested generation of items into known categories, and their fourth
+experiment was identical to Experiment 3 but with a restricted generation
+space.}
+
+Participants in their experiment were exposed to members of two
+experimenter-defined categories of `crystals' varying in hue, saturation, and
+size. Each category possessed a unique hue, but varied in saturation and size:
+In the `Positive' condition, there was a positive correlation between these
+features (i.e., larger sized crystals were more saturated), and in the
+`Negative' condition this relation was reversed. In the `Neutral' condition,
+there was no correlation between saturation and size. After learning about the
+categories from each condition, participants were asked to generate six
+exemplars belonging to a novel class. As noted above,
+\cite{jern2013probabilistic} found that the generated categories tended to
+follow the distributional properties of the experimenter-defined categories:
+Generated categories were tightly distributed along the hue feature, and
+possessed the same saturation-size correlations as in the learned categories.
+\cite{jern2013probabilistic}, however, did not analyze or discuss how the
+generated categories {\em differed} from the experimenter-defined categories.
+
+Because each experimenter-defined category in the \cite{jern2013probabilistic}
+experiment possessed a distinct hue shared by all members of the category, it is
+sensible that participants might generate a category with a hue distinct from
+the experimenter-provided categories. If creative generation were influenced by
+category contrast in this way, the hues of generated categories should be
+systematically different from those of the experimenter-defined categories.
+Unfortunately, stimulus hue was encoded and presented in the
+Hue-Saturation-Value (HSV) color space, which is device-dependent and not
+perceptually normed such that perceived color similarity corresponds to
+proximity in the color space (as opposed to a color space such as CIELAB that is
+device-independent and equidistant sets of points correspond to pairs of colors
+that have the same perceptual similarity; \citealp{wyszecki1967}). Further, they
+did not calibrate their monitor, and so we cannot know the precise colors
+presented to participants. As \cite{jern2013probabilistic} were interested in
+relations between the saturation and length of examples in generated into novel
+categories, these issues do not undercut their analyses and results. However,
+these issues pose a significant challenge to evaluating contrast between the
+experimenter-defined and participant-generated categories along the hue
+dimension. It is plausible that two uncalibrated monitors could display the same
+HSV color and the colors be perceived in different color categories (especially
+for color boundaries that vary over lightness, such as the yellow-brown
+boundary).
+
+Although we cannot know the precise colors that were displayed or perceived, we
+can still analyze their results from a coarse perspective to see whether there
+is preliminary support for contrast. To do so, we binned all possible hues into
+one of eight uniformly-spaced color groups: \{{\em Red}: $0-0.063$, $0.938-1$,
+{\em Yellow}: $0.063-0.188$, {\em Yellow-Green}: $0.188-0.313$, {\em
+Green-Teal}: $0.313-0.438$, {\em Teal}: $0.438-0.563$, {\em Teal-Blue}:
+$0.563-0.688$, {\em Purple}: $0.688-0.813$, {\em Pink}: $0.813-0.938$\}. In the
+\cite{jern2013probabilistic} experiment, the hue of each experimenter-defined
+category was selected from one of six possible values, each of which falls into
+one of the color groups above (two color groups were not used as a possible hue
+for the experimenter-defined categories). By categorizing the
+participant-generated crystals likewise, we can obtain a broad measure of
+category contrast by determining the proportion of participant-generated
+crystals that fall into the same groups as the experimenter-defined categories:
+If contrast influences the hues of the generated categories, we should observe
+minimal overlap between in the color groupings.
 
 \begin{figure}
-    \begin{center}
-        \inputpgf{figs/}{jk13-huecontrast.pgf}
-        \caption{Analysis of data from \cite{jern2013probabilistic}, Experiment 3.
-Plotted is the number of generated items that share a color group with one of the experimenter-defined classes.
-The ``Expected'' data follows a Binomial distribution with $p=2/8=1/4$, given there were two experimenter-defined classes, and eight color groups.}
+    \begin{center} \inputpgf{figs/}{jk13-huecontrast.pgf}
+        \caption{Analysis of data from \cite{jern2013probabilistic}, Experiment
+3. Plotted is the number of generated items that share a color group with one of
+the experimenter-defined classes. The ``Expected'' data follows a Binomial
+distribution with $p=2/8=1/4$, given there were two experimenter-defined
+classes, and eight color groups.}
         \label{fig:jk13-huecontrast}
@@
 
-These data, shown in Figure \ref{fig:jk13-huecontrast}, reveal a clear pattern: The majority of participants in each condition ($n = 22$) generated categories possessing entirely distinct hues; with 0/6 exemplars sharing a hue with the experimenter-defined categories.
-These results can be compared to the predictions of a Binomial model, which proposes that participants generate hues at random.
-That is, if hue selection is not systematic, the probability that any given example will lie in the same color group as an experimenter-defined category is given by a Binomial distribution with $p = 2/8=1/4$, as there were two experimenter-defined categories and eight possible color groups.
-Chi-square goodness-of-fit tests reveal that the observed distribution in each condition is highly inconsistent with the hues being chosen at random, all $\chi^2(6,N=22)>200$, $p<0.001$.
-Participants tended to generate items that were perceptually distinct from the categories they had learned, and were less likely to generate hues possessed by members of the experimenter-defined categories.
-
-Re-analyzing the results from \cite{jern2013probabilistic} provides preliminary support that contrast plays a role in category generation.
-Taken alongside the analyses reported by \cite{jern2013probabilistic}, our analysis suggests that generated categories tend to be distinct from {\em and} distributionally similar to what is already known.
-However, it is worth noting that our analysis is still limited: The color groups defined above are imprecise, and it is not clear that our color grouping is consistent with the psychological color boundaries perceived by participants.
-While we did obtain similar results using a variety of alternative groupings, the hue dimension used in the \cite{jern2013probabilistic} study does not lend itself straightforwardly to the the computation of similarities, and thus we cannot be certain of whether our coding accurately approximates the psychological space of the stimuli.
-This precludes traditional applications of categorization models to their data as it is usually necessary to encode objects in psychological space in order to accurately determine the similarity between objects.
-By consequence, although these results likely indicate that contrast exerts {\em some} influence, they do not precisely describe the nature of that influence.
-In the sections below, we propose a quantitative framework specifying the role of category contrast in creative generation.
+These data, shown in Figure \ref{fig:jk13-huecontrast}, reveal a clear pattern:
+The majority of participants in each condition ($n = 22$) generated categories
+possessing entirely distinct hues; with 0/6 exemplars sharing a hue with the
+experimenter-defined categories. These results can be compared to the
+predictions of a Binomial model, which proposes that participants generate hues
+at random. That is, if hue selection is not systematic, the probability that any
+given example will lie in the same color group as an experimenter-defined
+category is given by a Binomial distribution with $p = 2/8=1/4$, as there were
+two experimenter-defined categories and eight possible color groups. Chi-square
+goodness-of-fit tests reveal that the observed distribution in each condition is
+highly inconsistent with the hues being chosen at random, all
+$\chi^2(6,N=22)>200$, $p<0.001$. Participants tended to generate items that were
+perceptually distinct from the categories they had learned, and were less likely
+to generate hues possessed by members of the experimenter-defined categories.
+
+\sxlnote{Added this bit on regularity below - I'm a little undecided as to leave
+  it in the main text or place as a gigantic footnote? It is a bit of a
+  digression, and it's really just here to address Charles' point.... If we want
+  to keep this in the main text, I'll have to amend the following paragraph
+  ``Re-analyzing the results...'' to improve the transition.}
+%
+It is possible, however, that this data can be explained by a process that does
+not involve contrast. Specifically, since participants were trained on exemplars
+that share the same hue within a category, it is plausible that they generate
+exemplars that all share a common hue. If this is the case, given a $6/8 = .75$
+probability that a hue distinct from the experimenter-defined categories is
+selected, there is a corresponding $.75$ probability (or about 17 out of 22
+participants) that the generated category will have no exemplars that share a
+hue with any of the experimenter-defined categories. In addition, that there is
+a $2/8 = .25$ probability (about 6 out of 22 participants) that the generated
+category will have all six exemplars that share a hue with any of the
+experimenter-defined categories. This prediction of regularity across categories
+is much closer to the data than the prediction from a Binomial model and does
+not require any mechanism of category contrast.
+
+However, upon analysing the consistency of hues in the generated categories, it
+is clear that participants do not generally produce categories with entirely
+similar hues across all exemplars. Specifically, only .55 of all generated
+categories comprised exemplars with a common hue. Consequently, the assumption
+of regularity across categories may be inappropriate for this data set, leaving
+the contrast-based explanation as the most plausible for our purposes.
+
+Re-analyzing the results from \cite{jern2013probabilistic} provides preliminary
+support that contrast plays a role in category generation. Taken alongside the
+analyses reported by \cite{jern2013probabilistic}, our analysis suggests that
+generated categories tend to be distinct from {\em and} distributionally similar
+to what is already known. However, it is worth noting that our analysis is still
+limited: The color groups defined above are imprecise, and it is not clear that
+our color grouping is consistent with the psychological color boundaries
+perceived by participants. While we did obtain similar results using a variety
+of alternative groupings, the hue dimension used in the
+\cite{jern2013probabilistic} study does not lend itself straightforwardly to the
+the computation of similarities, and thus we cannot be certain of whether our
+coding accurately approximates the psychological space of the stimuli. This
+precludes traditional applications of categorization models to their data as it
+is usually necessary to encode objects in psychological space in order to
+accurately determine the similarity between objects. By consequence, although
+these results likely indicate that contrast exerts {\em some} influence, they do
+not precisely describe the nature of that influence. In the sections below, we
+propose a quantitative framework specifying the role of category contrast in
+creative generation.
 
@@
 
-As noted above, the constraint that new concepts should differ from what is already known has been largely overlooked in previous work.
-This is no doubt in part due to the vague definition of what it means for a concept to be ``different'': A generated category may be different from what is already known in any number of respects.
-Towards providing a more precise definition of the role of contrast in creative generation, we formalized contrast in a novel exemplar model, PACKER  ({\em Producing Alike and Contrasting Knowledge using Exemplar Representations}).
-PACKER explains category generation as a balance between two fundamental constraints: The category to be generated should not be similar to known categories, and exemplars within each category should be similar to one another.
-These ideas are implemented within the well-studied exemplar framework -- the PACKER model is an extension of the influential Generalized Context Model of categorization \citep[GCM;][]{nosofsky1984choice,nosofsky1986attention}.
-
-Although, as an exemplar model, one of PACKER's proposals is people represent categories in terms of a collection of stored exemplars, it is worth emphasizing that the underlying nature of human category representations is not the primary concern of our work.
-The choice to develop PACKER within an exemplar framework reflects the facts that exemplar models have been throughly evaluated, are strongly theoretically motivated, and dominate much of the theoretical and empirical work in categorization.
-The focus of our work with PACKER concerns the dual constraints of within- and between-class similarity; it is not difficult to imagine how such constraints may be instantiated using alternative frameworks (e.g., \citealp{kurtz2007divergent,love2004sustain,smith2000thirty}; for a review of categorization models, see \citealp{pothoswills2011}).
-
-Both PACKER and the GCM simulate categorization under the assumption that learners represent categories as a collection of exemplars, corresponding to the labeled stimuli they have observed.
-The exemplars are encoded within a $k$-dimensional psychological space, and model performance is based on the amount of similarity between the item to be categorized and the stored exemplars.
-Similarity between two examples, $s\left(x_i, x_j\right)$, is computed as an inverse exponential function of distance \citep[following][]{attneave1950,shepard1957stimulus,shepard1987toward}:
-
-\begin{equation}
-s\left(x_i,x_j\right) = \exp \left\{ -c \left[\sum_{k}{ w_k \left| x_{ik} - x_{jk} \right|^r }\right]^{1/r} \right\}
+As noted above, the constraint that new concepts should differ from what is
+already known has been largely overlooked in previous work. This is no doubt in
+part due to the vague definition of what it means for a concept to be
+``different'': A generated category may be different from what is already known
+in any number of respects. Towards providing a more precise definition of the
+role of contrast in creative generation, we formalized contrast in a novel
+exemplar model, PACKER ({\em Producing Alike and Contrasting Knowledge using
+Exemplar Representations}). PACKER explains category generation as a balance
+between two fundamental constraints: The category to be generated should not be
+similar to known categories, and exemplars within each category should be
+similar to one another. These ideas are implemented within the well-studied
+exemplar framework -- the PACKER model is an extension of the influential
+Generalized Context Model of categorization
+\citep[GCM;][]{nosofsky1984choice,nosofsky1986attention}.
+
+Although, as an exemplar model, one of PACKER's proposals is people represent
+categories in terms of a collection of stored exemplars, it is worth emphasizing
+that the underlying nature of human category representations is not the primary
+concern of our work. The choice to develop PACKER within an exemplar framework
+reflects the facts that exemplar models have been throughly evaluated, are
+strongly theoretically motivated, and dominate much of the theoretical and
+empirical work in categorization. The focus of our work with PACKER concerns the
+dual constraints of within- and between-class similarity; it is not difficult to
+imagine how such constraints may be instantiated using alternative frameworks
+(e.g., \citealp{kurtz2007divergent,love2004sustain,smith2000thirty}; for a
+review of categorization models, see \citealp{pothoswills2011}).
+
+Both PACKER and the GCM simulate categorization under the assumption that
+learners represent categories as a collection of exemplars, corresponding to the
+labeled stimuli they have observed. The exemplars are encoded within a
+$k$-dimensional psychological space, and model performance is based on the
+amount of similarity between the item to be categorized and the stored
+exemplars. Similarity between two examples, $s\left(x_i, x_j\right)$, is
+computed as an inverse exponential function of distance
+\citep[following][]{attneave1950,shepard1957stimulus,shepard1987toward}:
+
+\begin{equation} s\left(x_i,x_j\right) = \exp \left\{ -c \left[\sum_{k}{ w_k
+\left| x_{ik} - x_{jk} \right|^r }\right]^{1/r} \right\}
 \label{eq:similarity}
@@
 % 
-where $w_k$ is the attention weighting of dimension $k$ ($w_k \geq 0$ and $\sum_k{w_k} = 1$), accounting for the relative importance of each dimension in similarity calculations, and $c$ ($c>0$) is a specificity parameter controlling the spread of exemplar generalization.
-For simplicity, attention will be distributed uniformly in our simulations (unless otherwise noted).
-The value of $r$ depends on the nature of the experimental conditions being simulated: $r=1$ is appropriate for separable dimensions, whereas $r=2$ is appropriate for integral dimensions \citep[e.g.,][]{shepard1964attention,garner1974processing}.
-In our simulations, we set $r=1$ due to the separable nature of the stimulus dimensions used in our experiments (see Figure \ref{fig:e1-conditions}).
-
-PACKER (as well as its name) was in part inspired by earlier work from the categorization literature.
-Specifically, \cite{hidaka2011packing} argued that natural categories ``pack'' the values of features such that different categories fill the domain space with distance between one another, while keeping items within the same category close together.
-Inspired by this idea, PACKER proposes that generation is constrained by both similarity to members of the target category (the category in which a stimulus is being generated) as well as similarity to members of other categories: the most desirable generation candidates are similar to members of the target category and not similar to members of contrast categories.
-This is achieved by aggregating similarity across known exemplars differently according to class membership.
-The aggregated similarity $a(y,x)$ between generation candidate $y$ and stored exemplars $x$ is given by:
-
-\begin{equation}
-a(y, x) = \sum_j{f(x_j) s(y, x_j)}
+where $w_k$ is the attention weighting of dimension $k$ ($w_k \geq 0$ and
+$\sum_k{w_k} = 1$), accounting for the relative importance of each dimension in
+similarity calculations, and $c$ ($c>0$) is a specificity parameter controlling
+the spread of exemplar generalization. For simplicity, attention will be
+distributed uniformly in our simulations (unless otherwise noted). The value of
+$r$ depends on the nature of the experimental conditions being simulated: $r=1$
+is appropriate for separable dimensions, whereas $r=2$ is appropriate for
+integral dimensions \citep[e.g.,][]{shepard1964attention,garner1974processing}.
+In our simulations, we set $r=1$ due to the separable nature of the stimulus
+dimensions used in our experiments (see Figure \ref{fig:e1-conditions}).
+
+PACKER (as well as its name) was in part inspired by earlier work from the
+categorization literature. Specifically, \cite{hidaka2011packing} argued that
+natural categories ``pack'' the values of features such that different
+categories fill the domain space with distance between one another, while
+keeping items within the same category close together. Inspired by this idea,
+PACKER proposes that generation is constrained by both similarity to members of
+the target category (the category in which a stimulus is being generated) as
+well as similarity to members of other categories: the most desirable generation
+candidates are similar to members of the target category and not similar to
+members of contrast categories. This is achieved by aggregating similarity
+across known exemplars differently according to class membership. The aggregated
+similarity $a(y,x)$ between generation candidate $y$ and stored exemplars $x$ is
+given by:
+
+\begin{equation} a(y, x) = \sum_j{f(x_j) s(y, x_j)}
 \end{equation}
 % 
-where $f(x_j)$ is a function specifying each exemplar's contribution to generation.
-A negative value for $f(x_j)$ produces a `repelling' effect (items are less likely to be generated nearby $x_j$), and a positive value produces an `attracting' effect (items are more likely to be generated nearby $x_j$).
-When $f(x_j)=0$, the exemplar does not contribute to generation.
-
-PACKER sets $f(x_j)$ depending on exemplar $x_j$'s category membership: $f(x_j) = \gamma$ if $x_j$ is a member of the target category, and $f(x_j) = \gamma - 1$ if $x_j$ is a member of a contrast category.
-$\gamma$ is thus a free parameter ($0 \leq \gamma \leq 1$) controlling the trade-off between within- and between-category similarity.
-For example, when $\gamma = 0.5$, $f(x_j) = 0.5$ for members of the target category and $f(x_j) = -0.5$ for members of other categories; thus, the model is likely to generate items that are similar to members of the target category but are not similar to members of other categories.
-In this way, $\gamma = 1$ produces exclusive consideration of target-category members, and $\gamma = 0$ produces exclusive consideration of opposite-category members.
-The $\gamma$ parameter thus specifies a wide breadth of possible approaches; by fitting it to a dataset, one can describe the relative roles of between-category contrast and within-category similarity in generation.
-See Figure \ref{fig:packer-examples} for an illustration of how $\gamma$ controls the relative influence of within category similarity and contrast to other categories when generating a new exemplar.
+where $f(x_j)$ is a function specifying each exemplar's contribution to
+generation. A negative value for $f(x_j)$ produces a `repelling' effect (items
+are less likely to be generated nearby $x_j$), and a positive value produces an
+`attracting' effect (items are more likely to be generated nearby $x_j$). When
+$f(x_j)=0$, the exemplar does not contribute to generation.
+
+PACKER sets $f(x_j)$ depending on exemplar $x_j$'s category membership:
+$f(x_j) = \theta_t$ if $x_j$ is a member of the target category, and
+$f(x_j) = -\theta_c$ if $x_j$ is a member of a contrast category.
+$\theta_t$ and $\theta_c$ are free parameters
+($0 \leq (\theta_t,\theta_c) \leq 1$) controlling the trade-off
+between within- and between-category similarity. For example, when
+$\theta_t = \theta_c = 0.5$, $f(x_j) = 0.5$ for members of the
+target category and $f(x_j) = -0.5$ for members of other categories; thus, the
+model is likely to generate items that are similar to members of the target
+category but are not similar to members of other categories. In this way,
+$\theta_t > 0$ with $\theta_c = 0$ produces exclusive consideration
+of target-category members, and $\theta_c > 0$ with $\theta_t = 0$
+produces exclusive consideration of contrast-category members. The combination
+of $\theta_t$ and $\theta_c$ parameters thus specifies a wide
+breadth of possible approaches; by fitting it to a dataset, one can describe the
+relative roles of between-category contrast and within-category similarity in
+generation. See Figure \ref{fig:packer-examples} for an illustration of how
+these parameters control the relative influence of within-category similarity
+and contrast to other categories when generating a new exemplar.
 
 \begin{figure}
-	\begin{center}
-		\inputpgf{figs/}{packer-examples.pgf}
-		\caption{PACKER generation of a category `B' example, following exposure to one member of category `A' and one member of category `B'.
-Predictions are shown for three different parameterizations (differing only in $\gamma$): {\em (a)} Predictions based on contrast similarity only.
-{\em (b)} Predictions based on target similarity only.
-{\em (c)} Predictions with both constraints considered.}
+	\begin{center} \inputpgf{figs/}{packer-examples.pgf}
+      \caption{PACKER generation of a category `B' example, following exposure
+        to one member of category `A' and one member of category `B'.
+        Predictions are shown for three different parameterizations (differing
+        in values of $\theta_t$ and $\theta_c$): {\em (a)}
+        Predictions based on contrast similarity only. {\em (b)}
+        Predictions based on target similarity only. {\em (c)} Predictions with
+        both constraints considered.}
 		\label{fig:packer-examples}
@@
 
-The probability that a candidate $y$ will be generated is evaluated using an Exponentiated \citet{luce1977choice} choice rule.
-Candidates with greater values of $a(y,x)$ are more likely to be generated than candidates with smaller values:
+The probability that a candidate $y$ will be generated is evaluated using an
+Exponentiated \citet{luce1977choice} choice rule. Candidates with greater values
+of $a(y,x)$ are more likely to be generated than candidates with smaller values:
 % 
-\begin{equation}
-p(y \mid x) = \dfrac
-{ \exp  \left \{ \theta \cdot a \left( y, x \right) \right \} } 
-{ \sum_i{ \exp  \left \{ \theta \cdot a \left( y_i, x \right) \right\}  } }
+\begin{equation} p(y \mid x) = \dfrac { \exp \left \{ a \left( y, x
+\right) \right \} } { \sum_i{ \exp \left \{ a \left( y_i, x \right)
+\right\} } }
 \label{eq:packer-choice}
@@
 % 
-where $\theta$ ($\theta \geq 0$) is a free parameter controlling response determinism.
-
-It is worth noting that PACKER is only one possible exemplar-based account of category generation within our proposed framework.
-That is, PACKER places specific constraints on the possible values of $f(x_j)$, but other exemplar-based category generation models with drastically different behavior can be formalized in this framework by imposing alternative constraints.
-For example, as will be discussed in more detail below, PACKER is formally equivalent to the copy-and-tweak model proposed by \cite{jern2013probabilistic} when $\gamma = 1$.
-Likewise, when $\gamma = 0$, PACKER can represent a contrast-only generation mode, relying exclusively on contrast when generating new categories.
-Finally, when $f(x_j) = -1$ for all $x_j$ (regardless of class membership), a ``pure-packing'' approach is yielded, generating items in unoccupied areas of the domain.
-Thus, the proposed framework may be used to describe a wide variety of qualitatively distinct generation strategies.
-
-It is also worth noting that PACKER represents just one of many possible models that incorporate contrast in category generation.
-For example, it may be possible to extend \cite{jern2013probabilistic}'s hierarchical Bayesian model to include contrast, but this mechanism would be at odds with the standard hierarchical Bayesian framework.
-In contrast, these ideas emerge naturally from the exemplar view: We have not modified any of the core elements of the GCM in defining the PACKER model, we simply weight exemplar similarity during aggregation.
-Thus, beyond formally specifying the role of contrast in creative generation, the PACKER account allows us to evaluate the well-understood principles of the exemplar view within the comparatively understudied field of conceptual generation.
-Nonetheless, we will discuss integrating our approaches in the General Discussion.
+%where $\theta$ ($\theta \geq 0$) is a free parameter controlling response
+%determinism.
+
+It is worth noting that PACKER is only one possible exemplar-based account of
+category generation within our proposed framework. That is, PACKER places
+specific constraints on the possible values of $f(x_j)$, but other
+exemplar-based category generation models with drastically different behavior
+can be formalized in this framework by imposing alternative constraints. For
+example, as will be discussed in more detail below, PACKER is formally
+equivalent to the copy-and-tweak model proposed by \cite{jern2013probabilistic}
+when $\theta_c = 0$ and $\theta_t > 0$. Likewise, when
+$\theta_t = 0$ and $\theta_c > 0$, PACKER can represent a
+contrast-only generation mode, relying exclusively on contrast when generating
+new categories. Finally, when $f(x_j) < 0$ for all $x_j$ (regardless of class
+membership), a ``pure-packing'' approach is yielded, generating items in
+unoccupied areas of the domain. Thus, the proposed framework may be used to
+describe a wide variety of qualitatively distinct generation strategies.
+
+It is also worth noting that PACKER represents just one of many possible models
+that incorporate contrast in category generation. For example, it may be
+possible to extend \cite{jern2013probabilistic}'s hierarchical Bayesian model to
+include contrast, but this mechanism would be at odds with the standard
+hierarchical Bayesian framework. In contrast, these ideas emerge naturally from
+the exemplar view: We have not modified any of the core elements of the GCM in
+defining the PACKER model, we simply weight exemplar similarity during
+aggregation. Thus, beyond formally specifying the role of contrast in creative
+generation, the PACKER account allows us to evaluate the well-understood
+principles of the exemplar view within the comparatively understudied field of
+conceptual generation. Nonetheless, we will discuss integrating our approaches
+in the General Discussion.
 
 \subsubsection{Relation Between PACKER and Copy-And-Tweak}
-\label{section:copytweak-packer}
-The PACKER model is similar to the copy-and-tweak model reported by \cite{jern2013probabilistic}: Both models are exemplar-based, and both models generate new items according to their similarity to known members of the target class.
-However, PACKER diverges from the copy-and-tweak model by including a contrast mechanism, enabling generation according to dissimilarity to members of opposing categories.
-By consequence, copy-and-tweak can be realized as a parameterization of the PACKER model that is insensitive to category contrast.
-Specifically, when $\gamma = 1$ (see Figure \ref{fig:packer-examples}, panel B), $f(x_j) = 0$ for $x_j$ belonging to contrast categories; thus, PACKER is not influenced by these items, and is mathematically equivalent to a copy-and-tweak approach.
-
-In this paper, we report simulations using this copy-and-tweak model.
-This model fits within the exemplar-based category generation framework defined above, under the constraint that $\gamma = 1$, and is a continuous-dimension adaptation of the model tested by \cite{jern2013probabilistic}.
-By formalizing a model family where PACKER and copy-and-tweak are different parameterizations of models within the same framework, the comparison between PACKER and copy-and-tweak provides a test of the explanatory value of the contrast mechanism: The account provided by copy-and-tweak will only equal that of PACKER if the contrast mechanism does not offer an advantage (i.e., if $\gamma < 1$ significantly improves model fits).
-Note that the purpose of the article is to explore and formally analyze the role of contrast in category generation and thus, we leave extending PACKER to incorporate distributional factors \citep[as explored by][]{jern2013probabilistic} for future work.
-
-\subsection{Contrast as representativeness: An extension to Jern and Kemp (2013)}
-
-TODO: fill in Bayesian definition of representativeness and its application in JK2013.
+\label{section:copytweak-packer} The PACKER model is similar to the
+copy-and-tweak model reported by \cite{jern2013probabilistic}: Both models are
+exemplar-based, and both models generate new items according to their similarity
+to known members of the target class. However, PACKER diverges from the
+copy-and-tweak model by including a contrast mechanism, enabling generation
+according to dissimilarity to members of opposing categories. By consequence,
+copy-and-tweak can be realized as a parameterization of the PACKER model that is
+insensitive to category contrast. Specifically, when $\theta_c = 0$ and
+$\theta_t > 0$ (see Figure \ref{fig:packer-examples}, panel B),
+$f(x_j) = 0$ for $x_j$ belonging to contrast categories; thus, PACKER is not
+influenced by these items, and is mathematically equivalent to a copy-and-tweak
+approach.
+
+In this paper, we report simulations using this copy-and-tweak model. This model
+fits within the exemplar-based category generation framework defined above, and
+is a continuous-dimension adaptation of the model tested by
+\cite{jern2013probabilistic}. By formalizing a model family where PACKER and
+copy-and-tweak are different parameterizations of models within the same
+framework, the comparison between PACKER and copy-and-tweak provides a test of
+the explanatory value of the contrast mechanism based on exemplar dissimilarity:
+The account provided by copy-and-tweak will only equal that of PACKER if the
+contrast mechanism does not offer an advantage (i.e., if $\theta_c > 0$
+significantly improves model fits). Note that the purpose of the article is to
+explore and formally analyze the role of contrast in category generation and
+thus, we leave extending PACKER to incorporate distributional factors \citep[as
+explored by][]{jern2013probabilistic} for future work.
+
+\subsection{Contrast as representativeness: An extension to Jern and Kemp
+(2013)}
+\label{section:representativeness-definition}
+
+An alternate conceptualization of category contrast is the idea of
+representativeness -- exemplars are generated such that they are more
+representative of the target category and thus more distinct from and less
+similar to other categories. We adopt the definition specified by
+\cite{tenenbaum01}, where the representativeness $R$ of an item (or in our case,
+an exemplar) $y$ is the relative amount of evidence that is provided by $y$ for
+a given hypothesis $h$ compared to all other hypotheses $H'$. More formally,
+\begin{equation}
+  R(x,h) = \log \dfrac{p(x|h)}{\sum_{h'\in H'}{p(x|h')p(h')}} , 
+\label{eq:representativeness}
+\end{equation}
+
+In Equation \ref{eq:representativeness}, $p(h')$ is the prior probability
+distribution on the hypothesis space that excludes $h$. Note that when there are
+only two hypotheses (e.g. two categories) involved in the domain, this prior
+takes the value of $1$ since there is ever only one alternative hypothesis.
+
+This Bayesian formalization of representativeness makes it straightforward to
+implement into the existing hieriarchical Bayesian model of category generation
+developed by \cite{jern2013probabilistic}. Specifically, both models assume that
+exemplars are sampled from a given category distribution ($h$ in
+Equation \ref{eq:representativeness}). Each category distribution is a
+multivariate normal parameterized by a location vector $\mu_k$ and covariance
+matrix $\Sigma_k$ (i.e., $h = \mathcal{N}(\mu_k,\Sigma_k)$). The parameters
+$\mu_k$ and $\Sigma_k$ are assumed to be samples from a prior
+normal-inverse-Wishart (NIW) distribution parameterized by
+$\mu_0,\Sigma_0,\kappa,\nu$. Mathematically,
+\begin{align}
+  \mu_k, \Sigma_k|\mu_0,\Sigma_0,\kappa,\nu &\sim
+                                              \textrm{NIW}(\mu_0,\Sigma_0,\kappa,\nu)
+                                              \label{eq:mvnsample}\\
+  y|C &\sim \mathcal{N}(\mu_k,\Sigma_k) \label{eq:exemplarsample}
+\end{align}
+
+In our simulations, we set the prior mean $\mu_0$ to the center of the stimulus
+space and $\Sigma_0 = \lambda \mathbf{I}$ where $\lambda$ is a free parameter
+and $\mathbf{I}$ is a $d$-by-$d$ identity matrix with $d$ representing the
+number of dimensions or features in the domain. The $\kappa$ and $\nu$
+parameters are freely estimated within the constraints $\kappa > d-1$ and
+$\nu > 0$.
+
+With the assumption of the NIW prior, the location vector $\mu_k$ can be computed
+with:
+\begin{equation}
+  \mu_k = \dfrac
+    {\kappa\mu_{0} + n_k \bar{x}_k}
+    {\kappa + n_k}
+    \label{eq:category_mus}
+\end{equation}
+where $n_k$ is the number of observed exemplars in category $k$ and $\bar{x}_k$
+is the observed category mean. Note that if there are no observed exemplars in
+the target category (i.e., if the model is generating a completely novel
+category), Equation \ref{eq:category_mus} simplifies to $\mu_k = \mu_0$.
+
+The NIW prior also allows us to infer the category covariance matrix $\Sigma_k$
+by computing the following:
+\begin{equation}
+  \Sigma_k = [\Sigma_D\nu + C_k +
+    \dfrac
+    {\kappa n_k}
+    {\kappa + n_k}
+    (\bar{x_k}-\mu_k)(\bar{x_k}-\mu_k)^T
+  ] (\nu + n_k)^{-1}
+  \label{eq:Sigma_y}
+\end{equation}
+where $C_k$ is the observed category covariance and $\Sigma_D$ is the domain
+covariance matrix from which $\Sigma_k$ samples are obtained. We can
+infer $\Sigma_D$ from the observed category covariances $C$ and the prior
+covariance $\Sigma_0$. Specifically,
+\begin{equation}
+    \Sigma_D = \Sigma_0 + \sum_{k}{C_k}
+\end{equation}
+
+At this point, both the representativeness model and the hierarchical Bayesian
+model from \cite{jern2013probabilistic} are largely identical. However, both
+models diverge in their computation of exemplar generation probabilities. While
+the original hierarchical Bayesian model produces novel exemplars with
+probabilities proportional to the multivariate normal likelihood $p(y|h)$, the
+representativeness model generates new exemplars with probabilities weighted by
+their representativeness $R(y,h)$. In practice, the probability of generating a
+particular candidate $y$ is obtained using an Exponentiated
+\citet{luce1977choice} choice rule:
+\begin{equation}
+  p(y) = \dfrac
+    {\exp( \theta \cdot R(y,h))}
+    {\sum_k \exp(\theta  \cdot R(y_k,h_k))} .
+\label{eq:representative-choice}
+\end{equation}
+where $\theta$ is a freely estimated response determinism parameter (constrained
+such that $\theta \geq 0$).
+
+Despite both the original hierarchical Bayesian model and our representativeness
+model sharing identical hierarchical structures, their distinct response
+processes can yield very different exemplars. Specifically, the hierarchical
+Bayesian model emphasizes the generation of exemplars that maintain
+distributional commonalities across categories. In contrast, the
+representativeness model focuses on the generation of exemplars that are
+representative of the underlying distribution for a target category. Generally,
+this representativeness mechanism results in the generation of exemplars that
+are less similar (i.e., more different) to contrast categories. We see this
+illustrated in Figure \ref{fig:representative-examples}, where the
+representativeness model displays a strong preference for generating exemplars
+from target category 'B' that are further away from the contrast category 'A'.
+
+\begin{figure}
+	\begin{center} \inputpgf{figs/}{hbm-examples.pgf}
+		\caption{Generation of a category `B' example, following exposure
+to one member of category `A' and one member of category `B'. Predictions are
+shown for {\em(a)} the original hierarchical Bayesian model and {\em(b)} the
+representativeness model.}
+		\label{fig:representative-examples}
+	\end{center}
+\end{figure}
 
@@
 
-Research on the creative generation of novel concepts has focused on the finding that generated categories tend to possess distributional commonalities with known categories.
-However, a fundamental goal of concept generation is to create something {\em new} (i.e., different from what is already known).
-The manner in which generated categories differ from known ones is, nonetheless, poorly understood: Existing theories do not make strong predictions about how creatively generated concepts should systematically differ from existing ones.
-Above, we found encouraging initial support that contrast influences category generation \citep[][Experiment 3]{jern2013probabilistic}, and we introduced a novel, exemplar-based model formalizing the roles of similarity and contrast in creative generation.
-
-In the sections below, we present two experiments demonstrating systematic effects of category contrast on category generation inspired by factors influencing how the contrast models generate new categories.
-Our experiments are based on \cite{jern2013probabilistic}'s paradigm, which is a straightforward translation of the traditional artificial classification paradigm to the task of category generation: Participants are first exposed to a single, experimenter-defined category, and are then asked to generate members of a new category.
-We then report formal analyses comparing the exemplar dissimilarity and the representativeness accounts of our results to that of the hierarchical Bayesian and copy-and-tweak models developed by \cite{jern2013probabilistic}.
-After analyzing the category generation studies, we present third experiment to demonstrate the effect of contrast in category learning.
-This ensures that our results are not idiosyncratic to category generation.
+Research on the creative generation of novel concepts has focused on the finding
+that generated categories tend to possess distributional commonalities with
+known categories. However, a fundamental goal of concept generation is to create
+something {\em new} (i.e., different from what is already known). The manner in
+which generated categories differ from known ones is, nonetheless, poorly
+understood: Existing theories do not make strong predictions about how
+creatively generated concepts should systematically differ from existing ones.
+Above, we found encouraging initial support that contrast influences category
+generation \citep[][Experiment 3]{jern2013probabilistic}, and we introduced two
+novel approaches to formalizing the roles of similarity and contrast in creative
+generation.
+
+In the sections below, we present two experiments demonstrating systematic
+effects of category contrast on category generation inspired by factors
+influencing how the contrast models generate new categories. Our experiments are
+based on \cite{jern2013probabilistic}'s paradigm, which is a straightforward
+translation of the traditional artificial classification paradigm to the task of
+category generation: Participants are first exposed to a single,
+experimenter-defined category, and are then asked to generate members of a new
+category. We then report formal analyses comparing the exemplar dissimilarity
+and the representativeness accounts of our results to that of the hierarchical
+Bayesian and copy-and-tweak models developed by \cite{jern2013probabilistic}.
+After analyzing the category generation studies, we present third experiment to
+demonstrate the effect of contrast in category learning. This ensures that our
+results are not idiosyncratic to category generation.
 
@@
 
-To begin our investigation, we sought to extend the early evidence obtained from our analysis of the \cite{jern2013probabilistic} data, under a variety of learning conditions and using more standard stimulus materials.
-We used an artificial stimulus design: A two dimensional domain of squares, varying in color and size (see Figure \ref{fig:e1-conditions}, panel A).
-These dimensions have been used in numerous classification learning studies \cite[e.g.,][]{conaway2016similar,conaway2016generalization,shepard1961learning,nosofsky1994comparing}.
-Unlike those used in the \cite{jern2013probabilistic} experiments, distance on these physical dimensions aligns more directly with perceptual similarity, allowing us to evaluate the role of category contrast in creative generation more precisely.
-To extend the evidence provided by the \cite{jern2013probabilistic} data, we tested the effects of category contrast after learning one category from a set of qualitatively distinct category structures, as shown in Figure \ref{fig:e1-conditions}.
-
-Figures \ref{fig:e1-conditions}b-d show the values of exemplar dimensions belonging to the experimenter-defined categories (`A', or `Alpha') that participants were assigned to learn about prior to generating a new category.
-Each participant learned one of the category types during training.
-In the `Cluster' type, category A is a tight cluster of examples in the space.
-Perceptually instantiated, the members of category A might, for example, be large and dark in color.
-In the `Row' type, category A has a row pattern across the space, varying along one feature but not the other.
-Thus, its members might all be dark in color but would vary in size.
-Finally, in the `XOR' type, the experimenter-defined category consists of two clusters separated in opposite corners of the space, conforming to the exclusive-or logical structure (e.g., members are small and dark or large and light).
-
-It should be noted that in our experiments the assignment between the perceptual and conceptual dimensions (e.g., $X \rightarrow Size$, $Y \rightarrow Color$), as well as the direction of variation along each dimension (e.g., $dark \rightarrow light$ or $light \rightarrow dark$) was counterbalanced across participants.
-The category types in Figure \ref{fig:e1-conditions} are plotted in a conceptual space, rather than a perceptual space.
-Thus, while the conceptual organization of the category types remains constant, each category type may have a different physical instantiation according to the counterbalance assignment.
-For example, the Cluster type  may be large and dark in color, or it may be small and light in color, depending on the assignment and direction of the dimensions.
-For this reason, below we will discuss generation within a conceptual space, rather than a physically instantiated one.
+To begin our investigation, we sought to extend the early evidence obtained from
+our analysis of the \cite{jern2013probabilistic} data, under a variety of
+learning conditions and using more standard stimulus materials. We used an
+artificial stimulus design: A two dimensional domain of squares, varying in
+color and size (see Figure \ref{fig:e1-conditions}, panel A). These dimensions
+have been used in numerous classification learning studies
+\cite[e.g.,][]{conaway2016similar,conaway2016generalization,shepard1961learning,nosofsky1994comparing}.
+Unlike those used in the \cite{jern2013probabilistic} experiments, distance on
+these physical dimensions aligns more directly with perceptual similarity,
+allowing us to evaluate the role of category contrast in creative generation
+more precisely. To extend the evidence provided by the
+\cite{jern2013probabilistic} data, we tested the effects of category contrast
+after learning one category from a set of qualitatively distinct category
+structures, as shown in Figure \ref{fig:e1-conditions}.
+
+Figures \ref{fig:e1-conditions}b-d show the values of exemplar dimensions
+belonging to the experimenter-defined categories (`A', or `Alpha') that
+participants were assigned to learn about prior to generating a new category.
+Each participant learned one of the category types during training. In the
+`Cluster' type, category A is a tight cluster of examples in the space.
+Perceptually instantiated, the members of category A might, for example, be
+large and dark in color. In the `Row' type, category A has a row pattern across
+the space, varying along one feature but not the other. Thus, its members might
+all be dark in color but would vary in size. Finally, in the `XOR' type, the
+experimenter-defined category consists of two clusters separated in opposite
+corners of the space, conforming to the exclusive-or logical structure (e.g.,
+members are small and dark or large and light).
+
+It should be noted that in our experiments the assignment between the perceptual
+and conceptual dimensions (e.g., $X \rightarrow Size$, $Y \rightarrow Color$),
+as well as the direction of variation along each dimension (e.g., $dark
+\rightarrow light$ or $light \rightarrow dark$) was counterbalanced across
+participants. The category types in Figure \ref{fig:e1-conditions} are plotted
+in a conceptual space, rather than a perceptual space. Thus, while the
+conceptual organization of the category types remains constant, each category
+type may have a different physical instantiation according to the counterbalance
+assignment. For example, the Cluster type may be large and dark in color, or it
+may be small and light in color, depending on the assignment and direction of
+the dimensions. For this reason, below we will discuss generation within a
+conceptual space, rather than a physically instantiated one.
 
 \begin{figure}
-    \begin{center}
-    \inputpgf{figs/}{e1-conditions.pgf}
-    \caption{Stimulus domain and category types tested in Experiment 1. Stimuli are not drawn to scale.}
+    \begin{center} \inputpgf{figs/}{e1-conditions.pgf}
+    \caption{Stimulus domain and category types tested in Experiment 1. Stimuli
+are not drawn to scale.}
     \label{fig:e1-conditions}
@@
 
-After learning about an experimenter-defined category, participants are asked to generate examples of a new category.
-Within this paradigm, an effect of category contrast would be realized if participants prefer to generate items in locations that are distant (i.e., perceptually dissimilar) from members of category A.
-However, generation is left unconstrained.
-Critically, participants were not asked to generate something different in the prompt.
-For example, participants assigned to the Cluster condition may generate a tightly clustered category in the corner opposite of the experimenter-defined category.
-Alternatively, they may generate a tightly clustered category directly overlapping with the experimenter-defined category.
-Further, they may even generate an entirely different type of category (e.g., a row category).
-
-Our experimental results also provide a converging test of the classic finding that generated categories tend to share distributional properties with known categories in the domain \citep{jern2013probabilistic,ward1994structured}.
-From these results, we can predict that, in each condition, participants should generate categories that are distributionally similar to the experimenter-defined category: In the Cluster condition, generated categories should be tightly clustered.
-In the Row condition, generated categories should vary more along the X-axis than the Y-axis.
-In XOR condition, generated categories should be widely distributed across both dimensions, and the two dimensions should be positively correlated.
-
-Interestingly, the XOR condition also offers a dissociation between the roles of category contrast and the emulation of distributional structure: widely-distributed, positively-correlated categories would need to lie along the positive diagonal of the space (that is the only place they ``fit''), which is already occupied by the experimenter-defined category.
-Thus, if contrast plays a role, exemplars in the generated categories of participants in the XOR condition may not be positively correlated -- they may be negatively correlated instead.
-In this case, contrast and statistical regularities would interact, which would be inconsistent with the leading approach in conceptual generation \citep{jern2013probabilistic}
+After learning about an experimenter-defined category, participants are asked to
+generate examples of a new category. Within this paradigm, an effect of category
+contrast would be realized if participants prefer to generate items in locations
+that are distant (i.e., perceptually dissimilar) from members of category A.
+However, generation is left unconstrained. Critically, participants were not
+asked to generate something different in the prompt. For example, participants
+assigned to the Cluster condition may generate a tightly clustered category in
+the corner opposite of the experimenter-defined category. Alternatively, they
+may generate a tightly clustered category directly overlapping with the
+experimenter-defined category. Further, they may even generate an entirely
+different type of category (e.g., a row category).
+
+Our experimental results also provide a converging test of the classic finding
+that generated categories tend to share distributional properties with known
+categories in the domain \citep{jern2013probabilistic,ward1994structured}. From
+these results, we can predict that, in each condition, participants should
+generate categories that are distributionally similar to the
+experimenter-defined category: In the Cluster condition, generated categories
+should be tightly clustered. In the Row condition, generated categories should
+vary more along the X-axis than the Y-axis. In XOR condition, generated
+categories should be widely distributed across both dimensions, and the two
+dimensions should be positively correlated.
+
+Interestingly, the XOR condition also offers a dissociation between the roles of
+category contrast and the emulation of distributional structure:
+widely-distributed, positively-correlated categories would need to lie along the
+positive diagonal of the space (that is the only place they ``fit''), which is
+already occupied by the experimenter-defined category. Thus, if contrast plays a
+role, exemplars in the generated categories of participants in the XOR condition
+may not be positively correlated -- they may be negatively correlated instead.
+In this case, contrast and statistical regularities would interact, which would
+be inconsistent with the leading approach in conceptual generation
+\citep{jern2013probabilistic}
 
@@
 
-183 participants were recruited from Amazon Mechanical Turk.
-Each participant was randomly assigned to one condition: 64 participants were assigned to the Cluster condition, 61 were assigned to the Row condition, and 58 were assigned to the XOR condition (sample sizes differ due to random assignment).
-Stimuli were squares varying in color (grayscale 9.8\%--90.2\%) and side length (3.0--5.8cm), see Figure \ref{fig:e1-conditions}.
-The assignment of perceptual features (color, size) to axes of the domain space (x, y), as well as the direction of variation along each axis (e.g., $dark \rightarrow light$ or $light \rightarrow dark$) was counterbalanced across participants.
+183 participants were recruited from Amazon Mechanical Turk. Each participant
+was randomly assigned to one condition: 64 participants were assigned to the
+Cluster condition, 61 were assigned to the Row condition, and 58 were assigned
+to the XOR condition (sample sizes differ due to random assignment). Stimuli
+were squares varying in color (grayscale 9.8\%--90.2\%) and side length
+(3.0--5.8cm), see Figure \ref{fig:e1-conditions}. The assignment of perceptual
+features (color, size) to axes of the domain space (x, y), as well as the
+direction of variation along each axis (e.g., $dark \rightarrow light$ or $light
+\rightarrow dark$) was counterbalanced across participants.
 
@@
 
-As noted in the introduction of this paper, the task of generating members of a new category can be aligned with common tasks studied in the categorization literature: Whereas classification consists of predicting an object's category label on the basis of its features, inference consists of predicting an observed feature, given a set of observed features and a category label.
-Generation thus consists of predicting {\em all} features of an object, given a novel category label.
-In keeping with this insight, we designed our generation task as an extension of the traditional artificial classification learning paradigm.
-The task differs from traditional work in creative cognition primarily through the use of an artificial domain, which enables the application of computational models.
-The use of an artificial domain also requires the addition of a training phase, during which participants learn about the categories in the domain.
-As a result, unlike most previous studies \citep[e.g.,][]{ward1994structured}, participants in our studies have no experience with the domain before the start of the experiment, and the experimenter-defined categories are not hierarchically structured (as are many natural categories).
-
-Participants began the experiment with a short training phase (3 blocks of 4 trials), where they observed exemplars belonging to the `Alpha' category.
-Participants were instructed to learn as much as they can about the Alpha category, and that they would answer a series of test questions afterwards.
-On each trial, a single Alpha category exemplar was presented, and participants were given as much time as they desired to observe it before moving on to the next trial.
-Each block consisted of a single presentation of each of the members of the Alpha category, in a random order.
-Participants were shown the range of possible colors and sizes prior to training.
-
-Following the training phase, participants were asked to generate four examples belonging to another category called `Beta'.
-As in \citet{jern2013probabilistic}, generation was completed using a sliding-scale interface.
-Two scales controlled the values of the two dimensions (color, size)  for the generated example.
-An on-screen preview of the example updated whenever one of the features was changed.
-Participants could generate any example along an evenly-spaced 9x9 grid (including members of the Alpha category), except for any previously generated Beta exemplars.
-Neither the members of the Alpha category nor the previously generated Beta examples were visible during generation.
-Prior to beginning the generation phase, participants read the following instructions:
-
-\begin{displayquote}
-As it turns out, there is another category of geometric figures called ``Beta''.
-Instead of showing you examples of the Beta category, we would like to know what you think is likely to be in the Beta category.
-
-You will now be given the chance to create examples of any size or color in order to show what you expect about the Beta category.
-You will be asked to produce 4 Beta examples - they can be quite similar or quite different to each other, depending on what you think makes the most sense for the category.
-
-Each example needs to be unique, but the computer will let you know if you accidentally create a repeat.
+As noted in the introduction of this paper, the task of generating members of a
+new category can be aligned with common tasks studied in the categorization
+literature: Whereas classification consists of predicting an object's category
+label on the basis of its features, inference consists of predicting an observed
+feature, given a set of observed features and a category label. Generation thus
+consists of predicting {\em all} features of an object, given a novel category
+label. In keeping with this insight, we designed our generation task as an
+extension of the traditional artificial classification learning paradigm. The
+task differs from traditional work in creative cognition primarily through the
+use of an artificial domain, which enables the application of computational
+models. The use of an artificial domain also requires the addition of a training
+phase, during which participants learn about the categories in the domain. As a
+result, unlike most previous studies \citep[e.g.,][]{ward1994structured},
+participants in our studies have no experience with the domain before the start
+of the experiment, and the experimenter-defined categories are not
+hierarchically structured (as are many natural categories).
+
+Participants began the experiment with a short training phase (3 blocks of 4
+trials), where they observed exemplars belonging to the `Alpha' category.
+Participants were instructed to learn as much as they can about the Alpha
+category, and that they would answer a series of test questions afterwards. On
+each trial, a single Alpha category exemplar was presented, and participants
+were given as much time as they desired to observe it before moving on to the
+next trial. Each block consisted of a single presentation of each of the members
+of the Alpha category, in a random order. Participants were shown the range of
+possible colors and sizes prior to training.
+
+Following the training phase, participants were asked to generate four examples
+belonging to another category called `Beta'. As in
+\citet{jern2013probabilistic}, generation was completed using a sliding-scale
+interface. Two scales controlled the values of the two dimensions (color, size)
+for the generated example. An on-screen preview of the example updated whenever
+one of the features was changed. Participants could generate any example along
+an evenly-spaced 9x9 grid (including members of the Alpha category), except for
+any previously generated Beta exemplars. Neither the members of the Alpha
+category nor the previously generated Beta examples were visible during
+generation. Prior to beginning the generation phase, participants read the
+following instructions:
+
+\begin{displayquote} As it turns out, there is another category of geometric
+figures called ``Beta''. Instead of showing you examples of the Beta category,
+we would like to know what you think is likely to be in the Beta category.
+
+You will now be given the chance to create examples of any size or color in
+order to show what you expect about the Beta category. You will be asked to
+produce 4 Beta examples - they can be quite similar or quite different to each
+other, depending on what you think makes the most sense for the category.
+
+Each example needs to be unique, but the computer will let you know if you
+accidentally create a repeat.
 \end{displayquote}
@@
 
-We observed a substantial degree of individual differences in our data.
-In Figure \ref{fig:e1-samples} we have plotted sample data from several participants, from which it is evident that different participants generated qualitatively different category structures.
-In this section we will focus on analyzing the data in aggregate, but in later sections we will explore how individual differences can be explained.
+We observed a substantial degree of individual differences in our data. In
+Figure \ref{fig:e1-samples} we have plotted sample data from several
+participants, from which it is evident that different participants generated
+qualitatively different category structures. In this section we will focus on
+analyzing the data in aggregate, but in later sections we will explore how
+individual differences can be explained.
 
 \begin{figure}
-    \begin{center}
-    \inputpgf{figs/}{e1-samples.pgf}
-    \caption{Sample categories generated by participants in Experiment 1. Representative samples from common generation profiles are shown.}
+    \begin{center} \inputpgf{figs/}{e1-samples.pgf}
+    \caption{Sample categories generated by participants in Experiment 1.
+Representative samples from common generation profiles are shown.}
     \label{fig:e1-samples}
@@
 
-To evaluate the role of contrast, we computed the number of times each stimulus was generated, as a function of its average city-block distance from members of the experimenter-defined ``Alpha'' category.
-These data, shown in Figure \ref{fig:e1-distanceplots}, reveal a clear pattern: Examples that are more distant from members of the experimenter-defined categories are more likely to be generated into a new category.
-This supports the notion that contrast is a fundamental constraint on creative cognition and that suggests that statistical regularity alone is insufficient.
-
-Figure \ref{fig:e1-distanceplots} also depicts, for each participant, the average distance of members within the generated category ({\em within-category} distance) against the average distance between members of the generated and experimenter-defined category ({\em between-category} distance).
-The narrow distribution of between-category distances in the XOR condition reflects the widely distributed nature of the experimenter-defined category, reducing the possible distances to members of the participant-generated category.
-These data reveal a systematic pattern: The majority of participants generated categories with greater between-category distance than within-category distance.
-That is, members of the generated category tended to be more similar to one another than to members of the experimenter-defined category.
-To evaluate this claim quantitatively, we conducted t-tests comparing the amount of within- and between- class distance in each condition.
-All conditions possessed greater between-category distance: Cluster, $t(63) = 11.43$, $p < 0.001$; Row, $t(60) = 13.16$, $p < 0.001$; and XOR, $t(57) = 3.64$, $p < 0.001$.
-These results provide further evidence of an effect of category contrast: Participants prefer to generate categories that are dissimilar to the learned category but maintain some level of internal cohesion.
+To evaluate the role of contrast, we computed the number of times each stimulus
+was generated, as a function of its average city-block distance from members of
+the experimenter-defined ``Alpha'' category. These data, shown in Figure
+\ref{fig:e1-distanceplots}, reveal a clear pattern: Examples that are more
+distant from members of the experimenter-defined categories are more likely to
+be generated into a new category. This supports the notion that contrast is a
+fundamental constraint on creative cognition and that suggests that statistical
+regularity alone is insufficient.
+
+Figure \ref{fig:e1-distanceplots} also depicts, for each participant, the
+average distance of members within the generated category ({\em within-category}
+distance) against the average distance between members of the generated and
+experimenter-defined category ({\em between-category} distance). The narrow
+distribution of between-category distances in the XOR condition reflects the
+widely distributed nature of the experimenter-defined category, reducing the
+possible distances to members of the participant-generated category. These data
+reveal a systematic pattern: The majority of participants generated categories
+with greater between-category distance than within-category distance. That is,
+members of the generated category tended to be more similar to one another than
+to members of the experimenter-defined category. To evaluate this claim
+quantitatively, we conducted t-tests comparing the amount of within- and
+between- class distance in each condition. All conditions possessed greater
+between-category distance: Cluster, $t(63) = 11.43$, $p < 0.001$; Row, $t(60) =
+13.16$, $p < 0.001$; and XOR, $t(57) = 3.64$, $p < 0.001$. These results provide
+further evidence of an effect of category contrast: Participants prefer to
+generate categories that are dissimilar to the learned category but maintain
+some level of internal cohesion.
 
 \begin{figure}
-    \begin{center}
-    \inputpgf{figs/}{e1-distanceplots.pgf}
-    \caption{Experiment 1 results.
-{\em Left}: Frequency of generation as a function of distance from members of the experimenter-defined category.
-{\em Right}: Scatter plot of within-category versus between-category distance in each of the participant-generated categories.}
+    \begin{center} \inputpgf{figs/}{e1-distanceplots.pgf}
+    \caption{Experiment 1 results. {\em Left}: Frequency of generation as a
+function of distance from members of the experimenter-defined category. {\em
+Right}: Scatter plot of within-category versus between-category distance in each
+of the participant-generated categories.}
     \label{fig:e1-distanceplots}
@@
 
-A secondary goal of this experiment was to examine whether we replicate the classic result that generated categories often possess the same distributional properties as previously-known categories.
-For each generated category, we computed the category range along each axis (X, Y), as well as the correlation between features.
-These data, shown in Figure \ref{fig:e1-statsboxes}, reveal broad individual differences: Within each condition, participants generated categories spanning the entire X- and Y- axis as well as categories that spanned very little along each.
-Likewise, in each condition participants generated categories possessing strongly positive, neutral, and strongly negative correlations between the dimensions.
-Comparing the distributional statistics between conditions yields a broad yet, as we will see, misleading replication of the classic effect.
-
-With respect to ranges along each axis (X, Y), the generated categories from each condition tend to reflect the ranges of the experimenter-defined categories.
-The categories generated in the Cluster condition were less widely distributed along the X-axis compared to Row, $t(123) = 5.61$, $p < 0.001$, and XOR, $t(120) = 2.68$, $p = 0.008$.
-Categories generated in the XOR condition were also less widely distributed along the X-axis compared to Row, $t(117) = 2.56$, $p = 0.012$.
-This latter effect was not expected because the experimenter-defined categories for XOR and Row had similar X-ranges.
-However, the key finding is that categories from the Cluster condition tended to be more tightly distributed along the X-axis.
+A secondary goal of this experiment was to examine whether we replicate the
+classic result that generated categories often possess the same distributional
+properties as previously-known categories. For each generated category, we
+computed the category range along each axis (X, Y), as well as the correlation
+between features. These data, shown in Figure \ref{fig:e1-statsboxes}, reveal
+broad individual differences: Within each condition, participants generated
+categories spanning the entire X- and Y- axis as well as categories that spanned
+very little along each. Likewise, in each condition participants generated
+categories possessing strongly positive, neutral, and strongly negative
+correlations between the dimensions. Comparing the distributional statistics
+between conditions yields a broad yet, as we will see, misleading replication of
+the classic effect.
+
+With respect to ranges along each axis (X, Y), the generated categories from
+each condition tend to reflect the ranges of the experimenter-defined
+categories. The categories generated in the Cluster condition were less widely
+distributed along the X-axis compared to Row, $t(123) = 5.61$, $p < 0.001$, and
+XOR, $t(120) = 2.68$, $p = 0.008$. Categories generated in the XOR condition
+were also less widely distributed along the X-axis compared to Row, $t(117) =
+2.56$, $p = 0.012$. This latter effect was not expected because the
+experimenter-defined categories for XOR and Row had similar X-ranges. However,
+the key finding is that categories from the Cluster condition tended to be more
+tightly distributed along the X-axis.
 
 \begin{figure}
-    \begin{center}
-    \inputpgf{figs/}{e1-statsboxes.pgf}
-    \caption{Box-plots of the distributional statistics from the categories generated in Experiment 1.
-Boxes depict the median and quartiles of each condition, with whiskers placed at 1.5 IQR.
-All points outside this region are marked individually.}
+    \begin{center} \inputpgf{figs/}{e1-statsboxes.pgf}
+    \caption{Box-plots of the distributional statistics from the categories
+generated in Experiment 1. Boxes depict the median and quartiles of each
+condition, with whiskers placed at 1.5 IQR. All points outside this region are
+marked individually.}
     \label{fig:e1-statsboxes}
@@
 
-Likewise, categories generated in the Row condition had less Y-axis range compared to Cluster, $t(123) = 4.57$, $p < 0.001$ and XOR, $t(117) = 9.26$, $p < 0.001$, and categories from the Cluster condition had less Y-axis range compared to XOR, $t(120) = 3.95$, $p < 0.001$.
-As expected, the correlations in the Cluster and Row conditions were not systematically positive or negative ($ps > 0.1$).
-However, the generated categories in the XOR condition tended to possess {\em negatively} correlated dimensions, $t(57) = 2.04$, $p = 0.046$.
-This finding is notable, as it is the opposite of what would be expected, assuming learners are emulating the distributional structure of the experimenter-defined class (which possesses perfectly positively correlated features).
-
-We believe that the failure to replicate this finding is because participants in \cite{jern2013probabilistic} could differentiate the generated category on a third dimension (hue) to maintain the statistical regularities on the other two dimensions.
-Although the correlation in the XOR condition is significantly negative, it is clear from the box-plot in Figure~\ref{fig:e1-statsboxes} that it would be inappropriate to make a strong conclusion (e.g., the median is close to zero).
-However, we can conclude with confidence that there are situations where people do not emulate the distributional structure of the given category.
-This indicates that there is more to category generation than the emulation of distributional structure of other categories in the domain.
-Further, as we will discuss in more detail in the model-based analysis section, this is expected by our proposal that contrast is a fundamental principle in category generation.
-
-\subsection{Discussion}
-In Experiment 1 we sought to extend our analysis of the \cite{jern2013probabilistic} data by evaluating the influence of category contrast on creative generation, given qualitatively different types of prior knowledge.
-We found strong evidence for effects of category contrast in each condition: Participants were more likely to generate stimuli that are more distant from (i.e., less similar to) members of a previously-learned category, and members of participant-generated categories tended to be more similar to one another than to members of previously-learned categories.
-We also partially replicated the classic finding that the distributional structure of generated categories reflects that of previously learned categories \citep{jern2013probabilistic,ward1994structured}: Members of generated categories were more widely distributed along dimensions which were widely distributed in the experimenter-defined category.
-
-Notably, however, we also found that participants who learned an XOR category (composed of exemplars following a positive diagonal, see Figure \ref{fig:e1-conditions}) tended to generate items according to a {\em negative} feature correlation -- the opposite of what was present in the previously learned category.
-While this may be difficult to account for under existing theoretical approaches (which assume generated categories follow the same distributional structure as known categories), it can be concisely explained from a category contrast perspective.
-Specifically, within the XOR condition, individuals who seek to generate a category that is perceptually distinct from what is already known are left with only the upper-left and bottom-right quadrants of the space, as members of the previously-learned XOR category lie in the bottom-left and top-right.
-If examples are generated into both of the available quadrants, the generated category will possess a strongly negative correlation, opposing that of the experimenter-defined class.
-
-Thus, the core results of Experiment 1 indicate that generated categories can systematically differ from what we would expect based on prior work.
-The negative (or null) correlations observed in the XOR condition suggests an interesting interaction between contrast with a given category and emulation of statistical properties.
-That is, the constraints on creative generation imposed by category contrast may not simply influence the {\em location} of generated categories, but also their distributional structure.
-In Experiment 2, we test this claim more systematically.
+Likewise, categories generated in the Row condition had less Y-axis range
+compared to Cluster, $t(123) = 4.57$, $p < 0.001$ and XOR, $t(117) = 9.26$, $p <
+0.001$, and categories from the Cluster condition had less Y-axis range compared
+to XOR, $t(120) = 3.95$, $p < 0.001$. As expected, the correlations in the
+Cluster and Row conditions were not systematically positive or negative ($ps >
+0.1$). However, the generated categories in the XOR condition tended to possess
+{\em negatively} correlated dimensions, $t(57) = 2.04$, $p = 0.046$. This
+finding is notable, as it is the opposite of what would be expected, assuming
+learners are emulating the distributional structure of the experimenter-defined
+class (which possesses perfectly positively correlated features).
+
+We believe that the failure to replicate this finding is because participants in
+\cite{jern2013probabilistic} could differentiate the generated category on a
+third dimension (hue) to maintain the statistical regularities on the other two
+dimensions. Although the correlation in the XOR condition is significantly
+negative, it is clear from the box-plot in Figure~\ref{fig:e1-statsboxes} that
+it would be inappropriate to make a strong conclusion (e.g., the median is close
+to zero). However, we can conclude with confidence that there are situations
+where people do not emulate the distributional structure of the given category.
+This indicates that there is more to category generation than the emulation of
+distributional structure of other categories in the domain. Further, as we will
+discuss in more detail in the model-based analysis section, this is expected by
+our proposal that contrast is a fundamental principle in category generation.
+
+\subsection{Discussion} In Experiment 1 we sought to extend our analysis of the
+\cite{jern2013probabilistic} data by evaluating the influence of category
+contrast on creative generation, given qualitatively different types of prior
+knowledge. We found strong evidence for effects of category contrast in each
+condition: Participants were more likely to generate stimuli that are more
+distant from (i.e., less similar to) members of a previously-learned category,
+and members of participant-generated categories tended to be more similar to one
+another than to members of previously-learned categories. We also partially
+replicated the classic finding that the distributional structure of generated
+categories reflects that of previously learned categories
+\citep{jern2013probabilistic,ward1994structured}: Members of generated
+categories were more widely distributed along dimensions which were widely
+distributed in the experimenter-defined category.
+
+Notably, however, we also found that participants who learned an XOR category
+(composed of exemplars following a positive diagonal, see Figure
+\ref{fig:e1-conditions}) tended to generate items according to a {\em negative}
+feature correlation -- the opposite of what was present in the previously
+learned category. While this may be difficult to account for under existing
+theoretical approaches (which assume generated categories follow the same
+distributional structure as known categories), it can be concisely explained
+from a category contrast perspective. Specifically, within the XOR condition,
+individuals who seek to generate a category that is perceptually distinct from
+what is already known are left with only the upper-left and bottom-right
+quadrants of the space, as members of the previously-learned XOR category lie in
+the bottom-left and top-right. If examples are generated into both of the
+available quadrants, the generated category will possess a strongly negative
+correlation, opposing that of the experimenter-defined class.
+
+Thus, the core results of Experiment 1 indicate that generated categories can
+systematically differ from what we would expect based on prior work. The
+negative (or null) correlations observed in the XOR condition suggests an
+interesting interaction between contrast with a given category and emulation of
+statistical properties. That is, the constraints on creative generation imposed
+by category contrast may not simply influence the {\em location} of generated
+categories, but also their distributional structure. In Experiment 2, we test
+this claim more systematically.
 
@@
 \begin{figure}
-    \begin{center}
-    \input{figs/e2-conditions.pgf}
+    \begin{center} \input{figs/e2-conditions.pgf}
     \caption{Category types tested in Experiment 2.}
@@
 
-To test whether category contrast influences the distributional structure of generated categories, we sought to identify conditions in which differences in the distributional structure of generated categories cannot be explained by the distributional structure of the experimenter-defined category.
-We created two new category types (depicted in Figure \ref{fig:e2-conditions}) that possess identical distributional structures (both are tight clusters of examples with no correlation between features), as they only differ in their Y-axis position: the `Bottom' category lies in the bottom-center of the space, and the `Middle' category lies in the center.
-The distributional equality of these conditions is key to the design of the experiment: If the distributional structure of previously learned categories were the only influence on the generated categories, we should observe no difference in the categories participants generate between these two conditions.
-Will participants distribute their generated category differently between conditions due to the differences in the available stimulus space for generating a new category?
-
-If category contrast influences the distributional structure of the categories people generate, then we should observe different types of categories according to the shape of the space that is {\em unoccupied} by members of previously learned categories.
-The difference in the Y-axis position between the Bottom and Middle conditions produces a considerable change to the shape of the unoccupied space.
-Participants assigned to learn the Bottom category should be less likely to generate exemplars into the lower regions of the stimulus space (as these areas possess greater similarity to members of the Bottom category), preferring instead to distribute exemplars across the upper region of the space.
-This constraint is lifted in the Middle condition, as the Middle category exemplars are equidistant to the upper and lower regions of the space.
-Accordingly, participants should be more likely to utilize both of these areas.
-Thus, if category contrast influences the distributional structure of generated categories, we should observe more participants in the Middle condition that generate examples above {\em and} below the experimenter-defined category.
+To test whether category contrast influences the distributional structure of
+generated categories, we sought to identify conditions in which differences in
+the distributional structure of generated categories cannot be explained by the
+distributional structure of the experimenter-defined category. We created two
+new category types (depicted in Figure \ref{fig:e2-conditions}) that possess
+identical distributional structures (both are tight clusters of examples with no
+correlation between features), as they only differ in their Y-axis position: the
+`Bottom' category lies in the bottom-center of the space, and the `Middle'
+category lies in the center. The distributional equality of these conditions is
+key to the design of the experiment: If the distributional structure of
+previously learned categories were the only influence on the generated
+categories, we should observe no difference in the categories participants
+generate between these two conditions. Will participants distribute their
+generated category differently between conditions due to the differences in the
+available stimulus space for generating a new category?
+
+If category contrast influences the distributional structure of the categories
+people generate, then we should observe different types of categories according
+to the shape of the space that is {\em unoccupied} by members of previously
+learned categories. The difference in the Y-axis position between the Bottom and
+Middle conditions produces a considerable change to the shape of the unoccupied
+space. Participants assigned to learn the Bottom category should be less likely
+to generate exemplars into the lower regions of the stimulus space (as these
+areas possess greater similarity to members of the Bottom category), preferring
+instead to distribute exemplars across the upper region of the space. This
+constraint is lifted in the Middle condition, as the Middle category exemplars
+are equidistant to the upper and lower regions of the space. Accordingly,
+participants should be more likely to utilize both of these areas. Thus, if
+category contrast influences the distributional structure of generated
+categories, we should observe more participants in the Middle condition that
+generate examples above {\em and} below the experimenter-defined category.
 
@@
 
-122 participants were recruited from Amazon Mechanical Turk.
-61 participants were randomly assigned to the Middle and Bottom conditions each.
-The stimulus space and procedure were exactly as in Experiment 1.
-Participants first completed a short training phase, followed by the generation phase.
-The only difference from Experiment 1 was the category types given to participants.
+122 participants were recruited from Amazon Mechanical Turk. 61 participants
+were randomly assigned to the Middle and Bottom conditions each. The stimulus
+space and procedure were exactly as in Experiment 1. Participants first
+completed a short training phase, followed by the generation phase. The only
+difference from Experiment 1 was the category types given to participants.
 
@@
 \begin{figure}
-    \begin{center}
-    \inputpgf{figs/}{e2-samples.pgf}
+    \begin{center} \inputpgf{figs/}{e2-samples.pgf}
     \caption{Sample categories generated in Experiment 2. }
@@
 
-As in Experiment 1, we observed broad differences in the generation approach taken by different participants.
-To characterize the nature of these differences, Figure \ref{fig:e2-samples} depicts sample categories generated by participants.
-The data from each condition are organized into four columns based on commonly observed patterns of generation: a `Cluster' type of tightly-clustered examples,   `Row' and `Column' types of exemplars widely distributed along the one axis but narrowly along the other, and a `Corners' type, wherein participants placed exemplars in disparate corners of the space.
-As before, in this section we focus on analyzing the data in aggregate, but in later sections we will focus more specifically on explaining the individual differences.
+As in Experiment 1, we observed broad differences in the generation approach
+taken by different participants. To characterize the nature of these
+differences, Figure \ref{fig:e2-samples} depicts sample categories generated by
+participants. The data from each condition are organized into four columns based
+on commonly observed patterns of generation: a `Cluster' type of
+tightly-clustered examples, `Row' and `Column' types of exemplars widely
+distributed along the one axis but narrowly along the other, and a `Corners'
+type, wherein participants placed exemplars in disparate corners of the space.
+As before, in this section we focus on analyzing the data in aggregate, but in
+later sections we will focus more specifically on explaining the individual
+differences.
 
 \begin{figure}
-    \begin{center}
-    \inputpgf{figs/}{e2-distanceplots.pgf}
-    \caption{Experiment 2 results.
-{\em Left}: Frequency of generation as a function of distance from members of the experimenter-defined category.
-{\em Right}: Scatter plot of within-category versus between-category distance in each of the participant-generated categories.}
+    \begin{center} \inputpgf{figs/}{e2-distanceplots.pgf}
+    \caption{Experiment 2 results. {\em Left}: Frequency of generation as a
+function of distance from members of the experimenter-defined category. {\em
+Right}: Scatter plot of within-category versus between-category distance in each
+of the participant-generated categories.}
     \label{fig:e2-distanceplots}
@@
 
-We began our analysis by testing for the broad influence of category contrast on generation.
-As in Experiment 1, we computed the frequency each stimulus was generated as a function of its average distance from members of the experimenter-defined category, as well as each participant's average within- and between- category distance.
-These data, shown in Figure \ref{fig:e2-distanceplots}, yield very similar results.
-Participants generated stimuli that are distant from members of the experimenter-defined category, and the categories in each condition tended to possess more between-category than within-category distance: Bottom, $t(60) = 5.5$, $p < 0.001$; Middle, $t(60) = 2.71$, $p < 0.01$.
-We did, however, observe a notable subgroup of participants in each condition who generated categories with more within-category than between-category distance.
-Upon manual inspection, many of these individuals appear to have assumed a `Corners' strategy, placing exemplars in disparate corners of the space, thus producing much more within-category distance, see Figure \ref{fig:e2-samples} for examples.
-
-To explore the distributional structure of the generated categories, we computed the range of exemplars along each axis (X, Y), as well as the correlation between features.
-These data, shown in Figure \ref{fig:e2-statsboxes}, again demonstrate the degree of individual differences observed in our study.
-In each condition, we observed tightly clustered and widely distributed categories along each dimension, as well as positively, negatively, and uncorrelated categories.
+We began our analysis by testing for the broad influence of category contrast on
+generation. As in Experiment 1, we computed the frequency each stimulus was
+generated as a function of its average distance from members of the
+experimenter-defined category, as well as each participant's average within- and
+between- category distance. These data, shown in Figure
+\ref{fig:e2-distanceplots}, yield very similar results. Participants generated
+stimuli that are distant from members of the experimenter-defined category, and
+the categories in each condition tended to possess more between-category than
+within-category distance: Bottom, $t(60) = 5.5$, $p < 0.001$; Middle, $t(60) =
+2.71$, $p < 0.01$. We did, however, observe a notable subgroup of participants
+in each condition who generated categories with more within-category than
+between-category distance. Upon manual inspection, many of these individuals
+appear to have assumed a `Corners' strategy, placing exemplars in disparate
+corners of the space, thus producing much more within-category distance, see
+Figure \ref{fig:e2-samples} for examples.
+
+To explore the distributional structure of the generated categories, we computed
+the range of exemplars along each axis (X, Y), as well as the correlation
+between features. These data, shown in Figure \ref{fig:e2-statsboxes}, again
+demonstrate the degree of individual differences observed in our study. In each
+condition, we observed tightly clustered and widely distributed categories along
+each dimension, as well as positively, negatively, and uncorrelated categories.
 
 \begin{figure}
-    \begin{center}
-    \inputpgf{figs/}{e2-statsboxes.pgf}
-    \caption{Box-plots of the distributional statistics from the categories generated in Experiment 2. }
+    \begin{center} \inputpgf{figs/}{e2-statsboxes.pgf}
+    \caption{Box-plots of the distributional statistics from the categories
+generated in Experiment 2. }
     \label{fig:e2-statsboxes}
@@
 
-As noted above, if the distributional structure of generated categories is influenced by the shape of the space not occupied by members of known categories, then participants in the Middle condition would be more likely to place exemplars in the upper {\em and} lower regions of the space, as members of the experimenter-defined category are equidistant from these regions.
-Participants in the Bottom condition should be less likely to generate category members in the bottom regions because members of the experimenter-defined category are located there.
-One way to test these predictions is to analyze the Y-axis ranges of the generated categories: If Middle participants utilize the upper and lower regions of the space, their categories should vary more along the Y-axis.
-T-Tests comparing the conditions on the distributional statistics, however, reveal few between-group differences: the conditions do not differ with respect to X-axis range, Y-axis range, or feature correlations ($ps>0.17$).
-
-However, our ability to detect differences in Y-Axis range using a standard $t$-test between the conditions is, in this case, diminished due to the non-normality of the data (Shapiro-Wilk normality test $W=0.77, p< 0.001$ for the Middle condition and $W=0.85, p < 0.001$ for the Bottom condition).
-Figure \ref{fig:e2-yranges} depicts the Y-axis position of the exemplars generated within each participant's category.
-The categories are sorted by overall range, then by condition assignment.
-These data reveal that there were nearly as many participants who generated categories spanning the entire Y-axis as those who generated categories spanning almost none of the Y-axis.
-The non-normality of the Y-axis range distributions thus requires that we use a different approach to addressing the experiment's main question.
+As noted above, if the distributional structure of generated categories is
+influenced by the shape of the space not occupied by members of known
+categories, then participants in the Middle condition would be more likely to
+place exemplars in the upper {\em and} lower regions of the space, as members of
+the experimenter-defined category are equidistant from these regions.
+Participants in the Bottom condition should be less likely to generate category
+members in the bottom regions because members of the experimenter-defined
+category are located there. One way to test these predictions is to analyze the
+Y-axis ranges of the generated categories: If Middle participants utilize the
+upper and lower regions of the space, their categories should vary more along
+the Y-axis. T-Tests comparing the conditions on the distributional statistics,
+however, reveal few between-group differences: the conditions do not differ with
+respect to X-axis range, Y-axis range, or feature correlations ($ps>0.17$).
+
+However, our ability to detect differences in Y-Axis range using a standard
+$t$-test between the conditions is, in this case, diminished due to the
+non-normality of the data (Shapiro-Wilk normality test $W=0.77, p< 0.001$ for
+the Middle condition and $W=0.85, p < 0.001$ for the Bottom condition). Figure
+\ref{fig:e2-yranges} depicts the Y-axis position of the exemplars generated
+within each participant's category. The categories are sorted by overall range,
+then by condition assignment. These data reveal that there were nearly as many
+participants who generated categories spanning the entire Y-axis as those who
+generated categories spanning almost none of the Y-axis. The non-normality of
+the Y-axis range distributions thus requires that we use a different approach to
+addressing the experiment's main question.
 
 \begin{figure}
-    \begin{center}
-    \inputpgf{figs/}{e2-yranges.pgf}
-    \caption{Y-Axis range and position of the participant-generated categories from Experiment 2.
-Each line corresponds to a participant's category, with notches corresponding to the Y-axis position of exemplars within the category (notches may overlap).
-Participants are sorted by overall range, and then by condition.
-}
+    \begin{center} \inputpgf{figs/}{e2-yranges.pgf}
+    \caption{Y-Axis range and position of the participant-generated categories
+from Experiment 2. Each line corresponds to a participant's category, with
+notches corresponding to the Y-axis position of exemplars within the category
+(notches may overlap). Participants are sorted by overall range, and then by
+condition. }
     \label{fig:e2-yranges}
@@
 
-Because our main prediction concerns the generation of exemplars within the upper and lower regions of the domain, we compared the conditions in terms of the frequency with which participants generated examples above and below the categories.
-Specifically, we counted the number of participants in each condition who placed at least one `Beta' exemplar on the top and bottom `rows' of the space (the maximum and minimum possible Y-axis value, respectively).
-The resulting contingencies data are shown in Table \ref{table:e2-subset-table}.
-
-Firstly, it should be noted that nearly every participant utilized the top and/or bottom rows: only $10 / 122$ participants generated their category entirely within the interior region.
-Fisher's Exact Tests comparing the conditions reveal that more Middle participants generated an exemplar in the bottom row, $p < 0.001$, again demonstrating the role of contrast in guiding where exemplars are generated.
-The conditions did not differ in use of the top of the space, $p = 0.16$, however, more Middle participants placed exemplars in the top {\em and} bottom rows, $p = 0.038$.
-The latter effect is of interest here, as it indicates that the shape of the unoccupied space exerts some influence on the distributional structure of generated categories: Participants in the Middle condition were more likely to generate a category spanning the entire Y-axis.
-Thus, distributional structure of the generated categories can be influenced by category contrast alone.
+Because our main prediction concerns the generation of exemplars within the
+upper and lower regions of the domain, we compared the conditions in terms of
+the frequency with which participants generated examples above and below the
+categories. Specifically, we counted the number of participants in each
+condition who placed at least one `Beta' exemplar on the top and bottom `rows'
+of the space (the maximum and minimum possible Y-axis value, respectively). The
+resulting contingencies data are shown in Table \ref{table:e2-subset-table}.
+
+Firstly, it should be noted that nearly every participant utilized the top
+and/or bottom rows: only $10 / 122$ participants generated their category
+entirely within the interior region. Fisher's Exact Tests comparing the
+conditions reveal that more Middle participants generated an exemplar in the
+bottom row, $p < 0.001$, again demonstrating the role of contrast in guiding
+where exemplars are generated. The conditions did not differ in use of the top
+of the space, $p = 0.16$, however, more Middle participants placed exemplars in
+the top {\em and} bottom rows, $p = 0.038$. The latter effect is of interest
+here, as it indicates that the shape of the unoccupied space exerts some
+influence on the distributional structure of generated categories: Participants
+in the Middle condition were more likely to generate a category spanning the
+entire Y-axis. Thus, distributional structure of the generated categories can be
+influenced by category contrast alone.
 
@@
 \caption{Experiment 2 results.}
-\label{table:e2-subset-table} 
-\vskip 0.12in
-\begin{tabular}{ l r r}
-    \textbf{Middle}         & Used top row & No top row \\
-    \hline
-    Used bottom row       &  28 & 18  \\
-    No bottom row          &  11 &  4  \\
-    \\
-    \textbf{Bottom}         & Used top row & No top row \\
-    \hline
-    Used bottom row        & 16 & 8 \\
-    No bottom row          & 31 & 6 \\
+\label{table:e2-subset-table} \vskip 0.12in
+\begin{tabular}{ l r r} \textbf{Middle} & Used top row & No top row \\ \hline
+Used bottom row & 28 & 18 \\ No bottom row & 11 & 4 \\ \\ \textbf{Bottom} & Used
+top row & No top row \\ \hline Used bottom row & 16 & 8 \\ No bottom row & 31 &
+6 \\
 \end{tabular}
@@
 
-In Experiment 2, we replicated the core findings from Experiment 1.
-Stimuli are more likely to be generated if they are distant from exemplars in other categories, and most participants generate categories with more between-category than within-category distance.
-However, we additionally found that the {\em position} of a previously learned category (rather than its distributional structure) influences the types of categories people generate: Participants who learned the `Middle' type were more likely to generate categories spanning the entire Y-axis of the space.
-Participants who learned the `Bottom' type were less likely to do so as a result of the presence of opposite category exemplars in the lower regions of the space.
-
-This finding cannot be explained from the perspective that the distributional structure of previously learned categories is the sole determinant of the distributional structure of generated categories.
-However, the observed behavior is expected from a category contrast perspective: Participants seeking to generate a perceptually distinct category will be more likely to use areas of space that are unoccupied by exemplars belonging to previously learned categories.
-In the Middle condition, the upper and lower regions of space are equidistant from members of the experimenter-defined category, whereas in the Bottom condition, the lower region of the space is closer to members of the experimenter-defined category.
-Thus, while Middle participants may form categories around the use of the equally unoccupied areas, the same is not true for the Bottom condition.
+In Experiment 2, we replicated the core findings from Experiment 1. Stimuli are
+more likely to be generated if they are distant from exemplars in other
+categories, and most participants generate categories with more between-category
+than within-category distance. However, we additionally found that the {\em
+position} of a previously learned category (rather than its distributional
+structure) influences the types of categories people generate: Participants who
+learned the `Middle' type were more likely to generate categories spanning the
+entire Y-axis of the space. Participants who learned the `Bottom' type were less
+likely to do so as a result of the presence of opposite category exemplars in
+the lower regions of the space.
+
+This finding cannot be explained from the perspective that the distributional
+structure of previously learned categories is the sole determinant of the
+distributional structure of generated categories. However, the observed behavior
+is expected from a category contrast perspective: Participants seeking to
+generate a perceptually distinct category will be more likely to use areas of
+space that are unoccupied by exemplars belonging to previously learned
+categories. In the Middle condition, the upper and lower regions of space are
+equidistant from members of the experimenter-defined category, whereas in the
+Bottom condition, the lower region of the space is closer to members of the
+experimenter-defined category. Thus, while Middle participants may form
+categories around the use of the equally unoccupied areas, the same is not true
+for the Bottom condition.
 
 \section{Model-based Analyses}
-
-Experiments 1 and 2 revealed systematic and strong effects of category contrast on creative generation.
-In this section, we report the results of simulations with formal models aimed at explaining our observations.
-Specifically, we present simulations from the PACKER model, as well as a `copy-and-tweak' model (discussed in Section \ref{section:copytweak-packer}), defined as a variant of PACKER with the $\gamma$ parameter constrained to be one.
-The comparison of these two models serves to highlight the explanatory role of contrast within PACKER's framework: If contrast affords little explanatory advantage, then the two accounts should produce an equally strong account.
-We also present simulations from an implementation of the hierarchical Bayesian model proposed by \cite{jern2013probabilistic}, described in-depth in Appendix \ref{ap:hsampling-definition}.
-The comparison between the hierarchical Bayesian model and PACKER is meant to emphasize the necessity of contrast and demonstrate that generation cannot be explained entirely through the emulation of distributional structure.
-Each model has complementary strengths and weaknesses: Whereas PACKER is insensitive to the distributional structure of learned categories (relying only on within- and between-category similarity), the hierarchical Bayesian model generates categories exclusively on the basis of knowledge of how existing classes are distributed.
-
-Our approach in this section is to first broadly evaluate and compare the quality of each model's account to our entire dataset (Experiments 1 and 2 combined), then analyze the ability for each model to explain individual differences in each experiment, and lastly we describe the strengths and weakness of each model's account of category generation.
+\sxlnote{For this section I've added the existence of the representativeness
+  model but I think discussion of its performance is still lacking compared to PACKER.}
+Experiments 1 and 2 revealed systematic and strong effects of category contrast
+on creative generation. In this section, we report the results of simulations
+with formal models aimed at explaining our observations. Specifically, we
+present simulations from the two novel contrast models: PACKER and the
+representativeness model, and compare them to two models that do not incorporate
+contrast: the copy-and-tweak model (discussed in Section
+\ref{section:copytweak-packer}) and an implementation of the hierarchical
+Bayesian model proposed by \cite{jern2013probabilistic}, described in-depth in
+Appendix \ref{ap:hsampling-definition}. For our simulations here, the
+copy-and-tweak model is defined as a variant of PACKER with the $\theta_c$
+parameter constrained to be zero. The comparison of this set of models serves to
+highlight the explanatory role of contrast in categorization: If contrast
+affords little explanatory advantage, then there should be few differences in
+performance between PACKER and copy-and-tweak, or between the representativeness
+and hierarchical Bayesian model. The comparison between these models can also
+emphasize the necessity of contrast and demonstrate that generation cannot be
+explained entirely through the emulation of distributional structure. Each model
+has complementary strengths and weaknesses: Whereas PACKER and copy-and-tweak
+are insensitive to the distributional structure of learned categories (relying
+only on exemplar similarities), the representativeness and hierarchical Bayesian
+model generates categories exclusively on the basis of knowledge of how existing
+classes are distributed.
+
+Our approach in this section is to first broadly evaluate and compare the
+quality of each model's account to our entire dataset (Experiments 1 and 2
+combined), then analyze the ability for each model to explain individual
+differences in each experiment, and lastly we describe the strengths and
+weakness of each model's account of category generation.
 
@@
 
-To obtain a global measure of the quality of each model's account, we fitted the parameters of each model to our entire dataset (Experiments 1 and 2 combined), using a hill-climbing algorithm which maximized the log-likelihood of the model's predictions of the observed responses (1220 responses from 305 total participants).
-We fitted three parameters in the PACKER model ($c$, $\gamma$, and $\theta$; see Section \ref{section:PACKER-definition}), as well as four in the hierarchical Bayesian model ($\kappa$, $\lambda$, $\nu$, and $\theta$; see Appendix \ref{ap:hsampling-definition}).
-We fitted only two parameters for the copy-and-tweak model ($c$, and $\theta$), as $\gamma$ is held constant ($\gamma = 1$).
-Note that each model possesses a $\theta$ parameter fulfilling the same role (response determinism).
-Attention ($w$, see Equation \ref{eq:similarity}) in PACKER and copy-and-tweak was set uniformly.
-Parameters were not allowed to vary between participants or conditions -- the goal was to obtain the best-fitting values to our entire dataset.
-
-\begin{table}
-\centering
-\caption{Results of model-fitting to the combined datasets from Experiments 1 and 2. Note that smaller AIC values correspond to better model fits (adjusted for number of parameters)}
+To obtain a global measure of the quality of each model's account, we fitted the
+parameters of each model to our entire dataset (Experiments 1 and 2 combined),
+using a hill-climbing algorithm which maximized the log-likelihood of the
+model's predictions of the observed responses (1220 responses from 305 total
+participants). We fitted three parameters in the PACKER model ($c$, $\theta_t$,
+and $\theta_c$; see Section \ref{section:PACKER-definition}), as well as four in
+the representativeness model and the hierarchical Bayesian model ($\kappa$,
+$\lambda$, $\nu$, and $\theta$; see Section
+\ref{section:representativeness-definition} and Appendix
+\ref{ap:hsampling-definition} respectively). We fitted only two parameters for
+the copy-and-tweak model ($c$, and $\theta_t$), as $\theta_c$ is held constant
+($\theta_c = 0$). Attention ($w$, see Equation \ref{eq:similarity}) in PACKER
+and copy-and-tweak was set uniformly. Parameters were not allowed to vary
+between participants or conditions -- the goal was to obtain the best-fitting
+values to our entire dataset.
+
+\begin{table} \centering
+\caption{Results of model-fitting to the combined datasets from Experiments 1
+and 2. Note that smaller AIC values correspond to better model fits (adjusted
+for number of parameters)}
 \label{table:global-model-fits}
-\begin{tabular}{ l l l}
+\begin{tabular}{ l l l l} \\ \textbf{PACKER} & \textbf{Copy \& Tweak} &
+\textbf{Representativeness} & \textbf{Hierarchical Bayesian} \\ \hline $AIC =
+  9095$ & $AIC = 9842$ & $AIC = 8878$ &  $AIC =
+9912$ \\ $L = -4545$ & $L = -4919$ & $L = -4435$ & $L = -4952$ \\ $c = 0.483$ &
+                                                                                $c = 3.197$ & $\kappa = 11.232$ &
+$\kappa < 0.001$ \\ $\theta_c = 3.16$ & $\theta_c = 0$ (fixed) & $\nu = 2.844$ & $\nu = 5.597$ \\
+$\theta_t = 3.494$ & $\theta_t = 2.972$ & $\lambda = 1.917$& $\lambda
+                                                                       = 0.055$
+  \\ & &$\theta = 5.885 $& $\theta = 3.175$
 \\
- \textbf{PACKER}    & \textbf{Copy \& Tweak} & \textbf{Hierarchical Bayesian} \\ 
- \hline
- $AIC = 9095$       & $AIC = 9842$          & $AIC = 9912$     \\ 
- $L = -4545$        & $L = -4919$           & $L = -4952$       \\ 
- $c = 0.482$        & $c = 3.187$           & $\kappa < 0.001$  \\
- $\gamma = 0.525$   & $\gamma = 1$ (fixed)  & $\nu = 5.596$     \\ 
- $\theta = 6.664$   & $\theta = 2.969$      & $\lambda = 0.055$    \\ 
-                    &                       & $\theta = 3.174$  \\ 
 \end{tabular}
@@
 
-Table \ref{table:global-model-fits} contains the results of this fitting procedure.
-Due to the uneven number of fitted parameters among the models, we compare the model fits using the Akaike Information Criterion \citep[AIC;][]{akaike1974new}, where smaller values correspond to better fits (discounted by model complexity as measured by the number of parameters).
-The same qualitative results were obtained with alternative model comparison metrics (e.g., BIC, \citealp{schwarz1978estimating}; $AIC_C$, \citealp{hurvich1989regression}).
-
-Table \ref{table:global-model-fits} contains the AIC values of each model's best fit, as well as the corresponding log-likelihood ($L$) and the best-fitting parameter values.
-These results reveal strong model differentiation: The PACKER model achieved far better fits compared to the copy-and-tweak and hierarchical Bayesian models, and copy-and-tweak performed somewhat better than the hierarchical Bayesian model.
-While PACKER's advantage may tentatively be attributed to the model's sensitivity to category contrast (this will be explored in detail below), the advantage shown by copy-and-tweak over the hierarchical Bayesian model may be attributed to its exemplar-based representation, as opposed to the prototype-based representation assumed by the hierarchical Bayesian model.
-As observed in Figures \ref{fig:e1-samples} and \ref{fig:e2-samples}, the generated categories we observed were often widely distributed, with no items near the category prototype.
-This aspect of the data is inconsistent with the multivariate normal distributions (similar to prototypes) used to represent categories in the \cite{jern2013probabilistic} model, but can be easily accounted for using an exemplar-based approach.
-
-A key distinction between PACKER and copy-and-tweak, as well as the hierarchical Bayesian model, is that, of the three models, only PACKER is capable of making strong predictions about the location of new category members when the target class is entirely novel (i.e., no member of the category has been observed).
-Under these circumstances, there are no examples to copy, and thus the copy-and-tweak model predicts that items are generated at random.
-Likewise, with no observations on which to condition the  category distribution, the hierarchical Bayesian model also picks an item at random.
-Thus, it is possible that the failure of these models is simply due to their inability to explain each participant's first trial (generating the first item in the `Beta' category).
-To ensure this is not driving our results, we conducted an identical set of simulations as above, excluding this trial (leaving 915 responses in the dataset): Again, PACKER ($L = -3390$, $AIC = 6786$) achieved better fits than the copy-and-tweak ($L = -3579$, $AIC = 7162$) and hierarchical Bayesian ($L=-3612$, $AIC = 7232$) models.
-
-Finally, because copy-and-tweak is nested within PACKER, we can use a likelihood ratio test to compare the two models.
-PACKER explains the aggregate data significantly better than copy-and-tweak ($\chi^2(1) = 749, p < 0.001$ for all data and $\chi^2(1) = 377, p < 0.001$ excluding the first example), providing further evidence that category generation is better explained when contrast is considered.
-
-Through comparison with the copy-and-tweak model, Figure \ref{fig:packer-loglike} more clearly demonstrates the robustness of the explanatory gains yielded by PACKER's category contrast mechanism.
-It displays the log-likelihood of the participants' results under PACKER as a function of the $\gamma$ parameter.
-The model's other parameters ($c$, $\theta$) were set according to copy-and-tweak's best fits from Table \ref{table:global-model-fits}, and thus when $\gamma=1$, the models are equivalent.
-The figure clearly shows a ``sweet spot'':  a convex region in which PACKER achieves superior fits as a result of changes to $\gamma$.
-The best fitting values lie well below the value of 1 assumed by the copy-and-tweak model, which demonstrates the robustness of the contrast effect (though note PACKER achieves even better fits when its parameters are fitted together, as in Table \ref{table:global-model-fits}).
-Notably, however, the copy-and-tweak parameterization ($\gamma=1$) performs better than the `contrast-only' parameterization of the model ($\gamma = 0$).
-In sum, the data are better explained when both within-category similarity and category contrast is considered.
+Table \ref{table:global-model-fits} contains the results of this fitting
+procedure. Due to the uneven number of fitted parameters among the models, we
+compare the model fits using the Akaike Information Criterion
+\citep[AIC;][]{akaike1974new}, where smaller values correspond to better fits
+(discounted by model complexity as measured by the number of parameters). The
+same qualitative results were obtained with alternative model comparison metrics
+(e.g., BIC, \citealp{schwarz1978estimating}; $AIC_C$,
+\citealp{hurvich1989regression}).
+
+Table \ref{table:global-model-fits} contains the AIC values of each model's best
+fit, as well as the corresponding log-likelihood ($L$) and the best-fitting
+parameter values. These results reveal strong model differentiation: both
+contrast models (PACKER and the representativeness model) achieved far better
+fits compared to their non-contrast counterparts: copy-and-tweak and the
+hierarchical Bayesian model respectively. Interestingly between the contrast
+models, the hierarchical model (representativeness) outperformed the
+exemplar-based theory (PACKER), whereas between the non-contrast models, the
+reverse is observed. Specifically, here the exemplar-based model
+(copy-and-tweak) performed somewhat better than the hierarchical Bayesian model.
+
+\sxlnote{The argument in the following two paragraphs appear to be challenged by
+  the success of the representativeness model. I might suggest removing them altogether?}
+While PACKER's advantage may tentatively be attributed to the model's
+sensitivity to category contrast (this will be explored in detail below), the
+advantage shown by copy-and-tweak over the hierarchical Bayesian model may be
+attributed to its exemplar-based representation, as opposed to the
+prototype-based representation assumed by the hierarchical Bayesian model. As
+observed in Figures \ref{fig:e1-samples} and \ref{fig:e2-samples}, the generated
+categories we observed were often widely distributed, with no items near the
+category prototype. This aspect of the data is inconsistent with the
+multivariate normal distributions (similar to prototypes) used to represent
+categories in the \cite{jern2013probabilistic} model, but can be easily
+accounted for using an exemplar-based approach.
+
+A key distinction between PACKER and the other three models is that only PACKER
+is capable of making strong predictions about the location of new category
+members when the target class is entirely novel (i.e., no member of the category
+has been observed). Under these circumstances, there are no examples to copy,
+and thus the copy-and-tweak model predicts that items are generated at random.
+Likewise, with no observations on which to condition the category distribution,
+the hierarchical Bayesian and representativeness models also picks an item at
+random. Crucially, however, the next generated exemplar from the
+representativeness model is strongly influenced by contrast as
+representativeness. Thus, it is possible that the failure of these models is
+simply due to their inability to explain each participant's first trial
+(generating the first item in the `Beta' category). To ensure this is not
+driving our results, we conducted an identical set of simulations as above,
+excluding this trial (leaving 915 responses in the dataset): Again, PACKER
+($L = -3390$, $AIC = 6786$) achieved better fits than the copy-and-tweak
+($L = -3579$, $AIC = 7162$) and hierarchical Bayesian ($L=-3612$, $AIC = 7232$)
+models.
+
+Finally, because copy-and-tweak is nested within PACKER, we can use a likelihood
+ratio test to compare the two models. PACKER explains the aggregate data
+significantly better than copy-and-tweak ($\chi^2(1) = 749, p < 0.001$ for all
+data and $\chi^2(1) = 377, p < 0.001$ excluding the first example), providing
+further evidence that category generation is better explained when contrast is
+considered.
+
+Through comparison with the copy-and-tweak model, Figure
+\ref{fig:packer-loglike} more clearly demonstrates the robustness of the
+explanatory gains yielded by PACKER's category contrast mechanism. It displays
+the log-likelihood of the participants' results under PACKER as a function of
+the $\theta_c$ parameter. The model's other parameters ($c$,
+$\theta_t$) were set according to copy-and-tweak's best fits from Table
+\ref{table:global-model-fits}, and thus when $\theta_c=0$, the models are
+equivalent. The figure clearly shows a ``sweet spot'': a convex region in which
+PACKER achieves superior fits as a result of changes to $\theta_c$. The
+best fitting values lie well above the value of 0 assumed by the copy-and-tweak
+model, which demonstrates the robustness of the contrast effect (though note
+PACKER achieves even better fits when its parameters are fitted together, as in
+Table \ref{table:global-model-fits}). In sum, the data are better explained when
+both within-category similarity and category contrast is considered.
 
 \begin{figure}
-    \begin{center}
-    \inputpgf{figs/}{packer-loglike.pgf}
-    \caption{PACKER's fit as a function of its prioritization of within-category and between-category similarity (using the $\gamma$ parameter).
-To facilitate comparison, PACKER's other parameters ($c$, $\theta$) were set to the best fitting values obtained for copy-and-tweak in Table \ref{table:global-model-fits}.
-}
+    \begin{center} \inputpgf{figs/}{packer-loglike.pgf}
+      \caption{PACKER's fit as a function of its prioritization of
+        within-category and between-category similarity (using the
+        $\theta_c$ parameter) . To facilitate comparison, PACKER's other
+        parameters ($c$, $\theta_t$) were set to the best fitting values
+        obtained for copy-and-tweak in Table \ref{table:global-model-fits}. }
     \label{fig:packer-loglike}
@@
 
-As noted in Experiments 1 and 2, we observed a great deal of individual differences in the types of categories that participants generated.
-Within each condition, there were a wide variety of category types, such as row and column categories (see Figures \ref{fig:e1-samples} and \ref{fig:e2-samples}).
-The simulations reported above serve to evaluate the models while considering the entire dataset, but a secondary goal of any formal account should be to provide some explanation of how different profiles of performance emerge.
-Many of the individual generation profiles we observed can be described within the PACKER framework, simply by tuning the model's parameters in a principled manner.
-In this section, we describe more specifically how the most frequently observed profiles can be realized.
-
-By manual inspection, it is evident that the most common profiles of generation consist of: (A) a tightly-distributed `cluster' of examples, (B) `row'- and `column'-like arrangements (varying widely along one dimension but not the other), and (C) a `corners' arrangement with examples placed into disparate corners of the space.
-These four profiles are distinct in terms of the distribution of the generated category along each dimension: Whereas the cluster profile is tightly distributed along both dimensions, the row and column profiles are tightly distributed along just one dimension.
-Finally, the corners profile is widely distributed along both dimensions.
-
-In the framework proposed by PACKER, the cluster and corners profiles arise based on different prioritization of within-category similarity versus between-category contrast, and the row and column profiles arise based on the prioritization of each dimension in the computation of similarity.
-For example, in the cluster profile, there is a high degree of within-category similarity along both dimensions, whereas in the corners profile there is minimal within-category similarity.
-Thus, PACKER's proposal is that these individual differences arise as a result of different priorities: While the tight cluster configuration can be considered PACKER's `default' mode (as it maximizes within-category similarity), the corners profile can be produced when between-category contrast is put at a higher priority (i.e., $\gamma$ near $0$).
-
-Likewise, in the row and column profiles, there is a large degree of within-category similarity along one dimension but not the other.
-These differences likely arise due to a differential focus on one dimension over another, and thus they can be produced by changes to PACKER's attention weights, $w_1$ and $w_2$ (see Equation \ref{eq:similarity}).
-Traditionally, the attention weights in exemplar models are thought to reflect the diagnostic value of each dimension towards classifying the known category members \citep{nosofsky1984choice,nosofsky1986attention,kruschke1992alcove}, but within a generation context the weights specify the importance of within- and between-category similarity along each dimension.
-For example, if all of attention is allocated along the X-axis ($w_1=1$ and $w_2=0$), similarity along the Y-axis no longer influences performance.
-As a result, PACKER will create categories that are more widely distributed along the Y-axis, as similarity is not taken into account along that dimension.
-As a general principle, differentially weighting one dimension will result in the generation of categories that are more widely distributed along the ignored dimension, conforming to a row- or column-like arrangement.
-See Figure \ref{fig:packer-attention} for a depiction of how attention influences PACKER's performance.
+As noted in Experiments 1 and 2, we observed a great deal of individual
+differences in the types of categories that participants generated. Within each
+condition, there were a wide variety of category types, such as row and column
+categories (see Figures \ref{fig:e1-samples} and \ref{fig:e2-samples}). The
+simulations reported above serve to evaluate the models while considering the
+entire dataset, but a secondary goal of any formal account should be to provide
+some explanation of how different profiles of performance emerge. Many of the
+individual generation profiles we observed can be described within the PACKER
+framework, simply by tuning the model's parameters in a principled manner. In
+this section, we describe more specifically how the most frequently observed
+profiles can be realized.
+
+By manual inspection, it is evident that the most common profiles of generation
+consist of: (A) a tightly-distributed `cluster' of examples, (B) `row'- and
+`column'-like arrangements (varying widely along one dimension but not the
+other), and (C) a `corners' arrangement with examples placed into disparate
+corners of the space. These four profiles are distinct in terms of the
+distribution of the generated category along each dimension: Whereas the cluster
+profile is tightly distributed along both dimensions, the row and column
+profiles are tightly distributed along just one dimension. Finally, the corners
+profile is widely distributed along both dimensions.
+
+In the framework proposed by PACKER, the cluster and corners profiles arise
+based on different prioritization of within-category similarity versus
+between-category contrast, and the row and column profiles arise based on the
+prioritization of each dimension in the computation of similarity. For example,
+in the cluster profile, there is a high degree of within-category similarity
+along both dimensions, whereas in the corners profile there is minimal
+within-category similarity. Thus, PACKER's proposal is that these individual
+differences arise as a result of different priorities: While the tight cluster
+configuration can be considered PACKER's `default' mode (as it maximizes
+within-category similarity), the corners profile can be produced when
+between-category contrast is put at a higher priority (i.e., $\theta_c > \theta_t$).
+
+Likewise, in the row and column profiles, there is a large degree of
+within-category similarity along one dimension but not the other. These
+differences likely arise due to a differential focus on one dimension over
+another, and thus they can be produced by changes to PACKER's attention weights,
+$w_1$ and $w_2$ (see Equation \ref{eq:similarity}). Traditionally, the attention
+weights in exemplar models are thought to reflect the diagnostic value of each
+dimension towards classifying the known category members
+\citep{nosofsky1984choice,nosofsky1986attention,kruschke1992alcove}, but within
+a generation context the weights specify the importance of within- and
+between-category similarity along each dimension. For example, if all of
+attention is allocated along the X-axis ($w_1=1$ and $w_2=0$), similarity along
+the Y-axis no longer influences performance. As a result, PACKER will create
+categories that are more widely distributed along the Y-axis, as similarity is
+not taken into account along that dimension. As a general principle,
+differentially weighting one dimension will result in the generation of
+categories that are more widely distributed along the ignored dimension,
+conforming to a row- or column-like arrangement. See Figure
+\ref{fig:packer-attention} for a depiction of how attention influences PACKER's
+performance.
 
 \begin{figure}
-    \begin{center}
-    \inputpgf{figs/}{packer-attention-examples.pgf}
-    \caption{PACKER generation of a category `B' example, following exposure to one member of category `A' and one member of category `B'.
-Predictions are shown for different attention settings: {\em (a)} Increased weighting of the X-axis.
-{\em (b)} Increased weighting of the Y-axis.
-{\em (c)} Uniform weighting (identical to Figure \ref{fig:packer-examples}).}
+    \begin{center} \inputpgf{figs/}{packer-attention-examples.pgf}
+    \caption{PACKER generation of a category `B' example, following exposure to
+one member of category `A' and one member of category `B'. Predictions are shown
+for different attention settings: {\em (a)} Increased weighting of the X-axis.
+{\em (b)} Increased weighting of the Y-axis. {\em (c)} Uniform weighting
+(identical to Figure \ref{fig:packer-examples}).}
     \label{fig:packer-attention}
@@
 
-As in PACKER, changes in the parameter settings of the copy-and-tweak model can also be used to produce different patterns of generation.
-Indeed, as copy-and-tweak is simply a special case of the PACKER model, the attention weights operate exactly as described above to produce row- and column-like categories.
-However, because the model is not influenced by category contrast, it is biased toward generating tightly clustered categories, as new items are always most likely to be generated near known examples of the target category.
-Thus, the lack of a contrast mechanism prevents the model from explaining why some individuals widely distribute their categories to the corners of the space.
-
-Similar to the copy-and-tweak model, the hierarchical Bayesian model possesses no mechanism to account for category contrast, and the model is most likely to generate new items that are similar to known examples of the target category.
-Although this precludes an account of why individuals might assume a cluster versus corners profile, the covariance matrix specifying the model's prior domain distribution, $\Sigma_0$, can be used to explain the generation of row-like and column-like categories.
-This covariance matrix specifies the amount of variance assumed along each dimension (as well as the dimensional correlations) across the domain of categories.
-The covariance matrix for a newly generated category, $\Sigma_B$, is based on the assumed $\Sigma_0$ as well as the distributions of previously learned categories (see Appendix \ref{ap:hsampling-definition}).
-Thus, the importance of each feature can be coded into $\Sigma_0$ to alter the dimensional variance of generated categories.
-
-As noted above, while the copy-and-tweak and hierarchical Bayesian models possess mechanisms to explain row- and column-like categories, they cannot easily explain why some individuals widely distribute their generated categories into disparate corners of the space.
-This, however, reveals a more general limitation: According to the copy-and-tweak and hierarchical Bayesian models, the distributional structure of generated categories is {\em independent} of their location within the domain.
-For example, although the copy-and-tweak or hierarchical Bayesian models can be parameterized to generate row- or column-like categories, there is no mechanism in place to ensure that what is generated will be distinct from what is already known.
-In the next subsection, we explore this prediction through an analysis of the interdependence between distributional structure and location in creative generation.
+As in PACKER, changes in the parameter settings of the copy-and-tweak model can
+also be used to produce different patterns of generation. Indeed, as
+copy-and-tweak is simply a special case of the PACKER model, the attention
+weights operate exactly as described above to produce row- and column-like
+categories. However, because the model is not influenced by category contrast,
+it is biased toward generating tightly clustered categories, as new items are
+always most likely to be generated near known examples of the target category.
+Thus, the lack of a contrast mechanism prevents the model from explaining why
+some individuals widely distribute their categories to the corners of the space.
+
+Similar to the copy-and-tweak model, the hierarchical Bayesian model possesses
+no mechanism to account for category contrast, and the model is most likely to
+generate new items that are similar to known examples of the target category.
+Although this precludes an account of why individuals might assume a cluster
+versus corners profile, the covariance matrix specifying the model's prior
+domain distribution, $\Sigma_0$, can be used to explain the generation of
+row-like and column-like categories. This covariance matrix specifies the amount
+of variance assumed along each dimension (as well as the dimensional
+correlations) across the domain of categories. The covariance matrix for a newly
+generated category, $\Sigma_B$, is based on the assumed $\Sigma_0$ as well as
+the distributions of previously learned categories (see Appendix
+\ref{ap:hsampling-definition}). Thus, the importance of each feature can be
+coded into $\Sigma_0$ to alter the dimensional variance of generated categories.
+
+As noted above, while the copy-and-tweak and hierarchical Bayesian models
+possess mechanisms to explain row- and column-like categories, they cannot
+easily explain why some individuals widely distribute their generated categories
+into disparate corners of the space. This, however, reveals a more general
+limitation: According to the copy-and-tweak and hierarchical Bayesian models,
+the distributional structure of generated categories is {\em independent} of
+their location within the domain. For example, although the copy-and-tweak or
+hierarchical Bayesian models can be parameterized to generate row- or
+column-like categories, there is no mechanism in place to ensure that what is
+generated will be distinct from what is already known. In the next subsection,
+we explore this prediction through an analysis of the interdependence between
+distributional structure and location in creative generation.
 
 \subsection{Category Location vs. Distributional Structure}
-
-As noted above, while all three models make clear claims about the internal structure of generated categories, the copy-and-tweak and hierarchical Bayesian models do not make any claims about how generated categories should differ from what is already known.
-However, as we observed in the results of Experiment 2, the distributional structure of a category is not always independent of its location within the domain.
-To demonstrate this point in more depth, we computed the X- and Y- axis ranges of every participant-generated category.
-Taking the difference between these values ($X-Y$) produces a measure of each category's orientation in the space: positive difference scores correspond to categories with more X-axis range (horizontally aligned, `Row' categories), whereas negative difference scores indicate the opposite (vertically aligned, `Column' categories).
-Neutral differences scores indicate there was an equal amount of X- and Y-axis range, which can be produced by a number of different category types (`Clusters', `Corners', etc; see Figures \ref{fig:e1-samples} and \ref{fig:e2-samples}).
-By plotting, for each possible stimulus, the difference scores of categories it was generated within, we can relate the distributional structure of generated categories to their location within the domain.
-
-However, because many stimuli were infrequently generated (such items near members of the `Alpha' category), we cannot simply compute the empirical average of the difference scores, as infrequently generated stimuli would be likely to show artificially strong differences.
-Instead, we used a Bayesian analysis to estimate the mean $\mu_x$ on the assumption that the scores $x$ for each stimulus are normally distributed with an unknown mean and unknown standard deviation.
-The conjugate Normal-Inverse Gamma distribution provides a straightforward method for this estimation:
-
-\begin{equation}
-\mu_x = \dfrac
-    { \nu_0 \mu_0 + \sum{x} } 
-    { \nu_0 + n }
+\label{section:individual-diff}
+As noted above, while all three models make clear claims about the internal
+structure of generated categories, the copy-and-tweak and hierarchical Bayesian
+models do not make any claims about how generated categories should differ from
+what is already known. However, as we observed in the results of Experiment 2,
+the distributional structure of a category is not always independent of its
+location within the domain. To demonstrate this point in more depth, we computed
+the X- and Y- axis ranges of every participant-generated category. Taking the
+difference between these values ($X-Y$) produces a measure of each category's
+orientation in the space: positive difference scores correspond to categories
+with more X-axis range (horizontally aligned, `Row' categories), whereas
+negative difference scores indicate the opposite (vertically aligned, `Column'
+categories). Neutral differences scores indicate there was an equal amount of X-
+and Y-axis range, which can be produced by a number of different category types
+(`Clusters', `Corners', etc; see Figures \ref{fig:e1-samples} and
+\ref{fig:e2-samples}). By plotting, for each possible stimulus, the difference
+scores of categories it was generated within, we can relate the distributional
+structure of generated categories to their location within the domain.
+
+However, because many stimuli were infrequently generated (such items near
+members of the `Alpha' category), we cannot simply compute the empirical average
+of the difference scores, as infrequently generated stimuli would be likely to
+show artificially strong differences. Instead, we used a Bayesian analysis to
+estimate the mean $\mu_x$ on the assumption that the scores $x$ for each
+stimulus are normally distributed with an unknown mean and unknown standard
+deviation. The conjugate Normal-Inverse Gamma distribution provides a
+straightforward method for this estimation:
+
+\begin{equation} \mu_x = \dfrac { \nu_0 \mu_0 + \sum{x} } { \nu_0 + n }
 \label{eq:rangediff-bayes}
-\end{equation}
-\
-where $\mu_0$ is the prior mean, $\nu_0$ is a prior scale parameter (controlling the weighting of the $\mu_0$), and $n$ is the number of categories in which the stimulus was a member (i.e., the number of scores in $x$).
-The default assumption is that there is an equal amount of range along the X- and Y-axes, and so we set $\mu_0 = 0$.
-Likewise, to give a moderate amount of weighting to the prior mean we set $\nu_0 = 1$, though the results are robust to a range of values.
-Within this approach, the resulting aggregation is a trade-off between the number of generations and the strength of the range difference within each generated category.
-Infrequently generated stimuli, as well as those with mixed positive and negative scores, are given neutral difference scores.
+\end{equation} \ where $\mu_0$ is the prior mean, $\nu_0$ is a prior scale
+parameter (controlling the weighting of the $\mu_0$), and $n$ is the number of
+categories in which the stimulus was a member (i.e., the number of scores in
+$x$). The default assumption is that there is an equal amount of range along the
+X- and Y-axes, and so we set $\mu_0 = 0$. Likewise, to give a moderate amount of
+weighting to the prior mean we set $\nu_0 = 1$, though the results are robust to
+a range of values. Within this approach, the resulting aggregation is a
+trade-off between the number of generations and the strength of the range
+difference within each generated category. Infrequently generated stimuli, as
+well as those with mixed positive and negative scores, are given neutral
+difference scores.
 
@@
     \begin{center} 
-    \inputpgf{figs/}{range-diff-gradients.pgf}
-    \caption{Behavioral and simulated range difference gradients.
-Each panel shows, for each stimulus, the dimensional orientation of the categories it was generated into: vertically aligned `columns' (orange) versus horizontally aligned `rows' (purple).}
+      \includegraphics[width=\textwidth]{figs/range-diff-gradients.pdf}
+      \caption{Behavioral and simulated range difference gradients. Each panel
+        shows, for each stimulus, the dimensional orientation of the categories
+        it was generated into: vertically aligned `columns' (orange) versus
+        horizontally aligned `rows' (purple).}
       \label{fig:range-diff-gradients}
@@
 
-The results of our analysis are shown in Figure \ref{fig:range-diff-gradients} for the experiment and model results\footnote{Prior to plotting, data was also processed using a Gaussian filter with $\sigma = 0.8$.}.
-The left-most column of Figure \ref{fig:range-diff-gradients} displays the effect of category location and contrast on the distributional structure of the category generated by participants.
-These data reveal strong and consistent patterns across all the conditions we tested in Experiments 1 and 2: Generated categories are more tightly distributed along the axis in which they are distinct.
-For example, in the `Cluster' condition, exemplars in the bottom-left of the space are more often generated into vertically aligned categories, and exemplars in the top-right are more often generated into horizontally aligned categories.
-Similarly, in the `Bottom' and `Middle' conditions, horizontally aligned categories are generated above and below the experimenter-defined categories, while vertically-aligned categories are generated to the sides.
-In the `Row' condition, most categories are horizontally aligned, and lie along the upper areas of the space.
-There are no strong range difference patterns in the XOR condition.
-
-These patterns of performance clearly depict the interdependence between the distributional structure and location of generated concepts.
-Our results can be interpreted in terms of local minimization of between-category similarity: By distributing the generated category away from members of the experimenter-defined category, participants may increase the degree of between-category distance without drastically altering the degree of within-category similarity.
-
-To explore how well the PACKER, copy-and-tweak, and hierarchical Bayesian models explain our findings, we conducted simulations using an individual-differences approach.
-As noted in Section \ref{section:individual-diffs}, row- and column-like categories can be produced by each model through changes to the weighting of each dimension.
-Given this information, we may use the models to simulate each participant's generation separately, with the importance of each dimension set according to the relative range of the participant's generated category along each dimension.
-
-In the PACKER and copy-and-tweak models, the attention weights, $w$, specify the importance of each dimension in the computation of similarity.
-While there exist methods to find the optimal attention weighting scheme given a classification \citep[see][]{vanpaemel2012using}, for simplicity we may assume that the Alpha and Beta categories are distinct along dimensions that the Betas do not vary on.
+The results of our analysis are shown in Figure \ref{fig:range-diff-gradients}
+for the experiment and model results\footnote{Prior to plotting, data was also
+processed using a Gaussian filter with $\sigma = 0.8$.}. The left-most column of
+Figure \ref{fig:range-diff-gradients} displays the effect of category location
+and contrast on the distributional structure of the category generated by
+participants. These data reveal strong and consistent patterns across all the
+conditions we tested in Experiments 1 and 2: Generated categories are more
+tightly distributed along the axis in which they are distinct. For example, in
+the `Cluster' condition, exemplars in the bottom-left of the space are more
+often generated into vertically aligned categories, and exemplars in the
+top-right are more often generated into horizontally aligned categories.
+Similarly, in the `Bottom' and `Middle' conditions, horizontally aligned
+categories are generated above and below the experimenter-defined categories,
+while vertically-aligned categories are generated to the sides. In the `Row'
+condition, most categories are horizontally aligned, and lie along the upper
+areas of the space. There are no strong range difference patterns in the XOR
+condition.
+
+These patterns of performance clearly depict the interdependence between the
+distributional structure and location of generated concepts. Our results can be
+interpreted in terms of local minimization of between-category similarity: By
+distributing the generated category away from members of the
+experimenter-defined category, participants may increase the degree of
+between-category distance without drastically altering the degree of
+within-category similarity.
+
+To explore how well the PACKER, copy-and-tweak, representativeness, and
+hierarchical Bayesian models explain our findings, we conducted simulations
+using an individual-differences approach. As noted in Section
+\ref{section:individual-diffs}, row- and column-like categories can be produced
+by each model through changes to the weighting of each dimension. Given this
+information, we may use the models to simulate each participant's generation
+separately, with the importance of each dimension set according to the relative
+range of the participant's generated category along each dimension.
+
+In the PACKER and copy-and-tweak models, the attention weights, $w$, specify the
+importance of each dimension in the computation of similarity. While there exist
+methods to find the optimal attention weighting scheme given a classification
+\citep[see][]{vanpaemel2012using}, for simplicity we may assume that the Alpha
+and Beta categories are distinct along dimensions that the Betas do not vary on.
 Thus, the weighting for a given participant can be computed as:
 
-\begin{equation}
-w_k = \dfrac
-{\exp{ \{ -\theta_w \cdot\text{range}(k)}  \} } 
-{ \sum_k {\exp{ \{ -\theta_w \cdot\text{range}(k)}  \} } }
+\begin{equation} w_k = \dfrac {\exp{ \{ -\theta_w \cdot\text{range}(k)} \} } {
+\sum_k {\exp{ \{ -\theta_w \cdot\text{range}(k)} \} } }
 \label{eq:range-weight}
@@
 % 
-where $\theta_w$ is a free parameter controlling how differences in range correspond to differences in weights (functioning similarly to the $\theta$ parameter in each of the models), and $\text{range}(k)$ is the range of examples generated by the participant along dimension $k$.
-We used $\theta_w = 1.5$ in our simulations, though the results are robust and similar for other $\theta_w$ values.
-The resulting $w$ values are thus inversely proportional to the range of generated categories along each dimension, with less range corresponding to greater weighting.
-
-Unlike the PACKER and copy-and-tweak models, the hierarchical Bayesian model's dimensional variances correspond to the assumed variance of generated categories along each dimension (rather than the inverse of the variance).
-Thus, a different transformation is appropriate for incorporating the weights computed in Equation \ref{eq:range-weight}.
-For the hierarchical Bayesian model, we computed the dimensional variances according to: $\lambda (1-w_k) 2$, where $\lambda$ is a free parameter specifying the overall assumed variance of the domain, and $2$ corresponds to the number of dimensions in our experiments\footnote{This calculation applies only in two-dimensional domains, where $w_2 = 1-w_1$.}.
-Under this approach, evenly distributed weights correspond to an assumed variance of $\lambda$.
-Likewise, larger values of $w$, which are produced when the generated category is tightly distributed along one dimension, correspond to smaller assumed variances.
-
-Each model was used to simulate each participant's generation independently, with the importance of each dimension set according to the participant's generated category.
-The other free parameters within each model were set as in Table \ref{table:global-model-fits}.
-Every participant's generation was simulated 2,000 times; given the 305 participants tested across the two experiments, each model generated 610,000 categories in total.
-For comparison with our behavioral results, we then computed the range difference gradient identically as with the behavioral data.
-The results are shown in Figure \ref{fig:range-diff-gradients}.
-
-As in the more traditional model evaluation analysis described above, PACKER provided a much closer match to our behavioral results than the copy-and-tweak and hierarchical Bayesian models.
-In all conditions, PACKER distributes categories similarly to the behavioral data: Horizontally-aligned categories tend to be placed above and below members of the experimenter-defined category, and vertically-aligned categories tend to be placed to the sides.
-Conversely, because the copy-and-tweak and hierarchical Bayesian models are insensitive to category contrast, these models do not produce any systematic patterns of association between category location and distributional structure.
-The sole exception is within the `Row' condition of Experiment 1, in which the majority of participants generated a `Row'-like category, widely distributed along the X-axis but not the Y-axis.
-In these cases, both models are initialized with weights that produce Row categories, but because category contrast is not considered, categories are uniformly generated across the entire domain, rather than concentrated within the upper-regions as observed behaviorally.
+where $\theta_w$ is a free parameter controlling how differences in range
+correspond to differences in weights (functioning similarly to the $\theta$
+parameter in each of the models), and $\text{range}(k)$ is the range of examples
+generated by the participant along dimension $k$. We used $\theta_w = 1.5$ in
+our simulations, though the results are robust and similar for other $\theta_w$
+values. The resulting $w$ values are thus inversely proportional to the range of
+generated categories along each dimension, with less range corresponding to
+greater weighting.
+
+Unlike the PACKER and copy-and-tweak models, the representativeness and
+hierarchical Bayesian model's dimensional variances correspond to the assumed
+variance of generated categories along each dimension (rather than the inverse
+of the variance). Thus, a different transformation is appropriate for
+incorporating the weights computed in Equation \ref{eq:range-weight}. For the
+hierarchical Bayesian model, we computed the dimensional variances according to:
+$\lambda (1-w_k) 2$, where $\lambda$ is a free parameter specifying the overall
+assumed variance of the domain, and $2$ corresponds to the number of dimensions
+in our experiments\footnote{This calculation applies only in two-dimensional
+  domains, where $w_2 = 1-w_1$.}. Under this approach, evenly distributed
+weights correspond to an assumed variance of $\lambda$. Likewise, larger values
+of $w$, which are produced when the generated category is tightly distributed
+along one dimension, correspond to smaller assumed variances.
+
+Each model was used to simulate each participant's generation independently,
+with the importance of each dimension set according to the participant's
+generated category. The other free parameters within each model were set as in
+Table \ref{table:global-model-fits}. Every participant's generation was
+simulated 2,000 times; given the 305 participants tested across the two
+experiments, each model generated 610,000 categories in total. For comparison
+with our behavioral results, we then computed the range difference gradient
+identically as with the behavioral data. The results are shown in Figure
+\ref{fig:range-diff-gradients}.
+
+As in the more traditional model evaluation analysis described above, the
+contrast models (i.e., PACKER and the representativeness model) provided a much
+closer match to our behavioral results than the copy-and-tweak and hierarchical
+Bayesian models. In all conditions, the contrast models distribute categories
+similarly to the behavioral data: Horizontally-aligned categories tend to be
+placed above and below members of the experimenter-defined category, and
+vertically-aligned categories tend to be placed to the sides. Conversely,
+because the copy-and-tweak and hierarchical Bayesian models are insensitive to
+category contrast, these models do not produce any systematic patterns of
+association between category location and distributional structure. The sole
+exception is within the `Row' condition of Experiment 1, in which the majority
+of participants generated a `Row'-like category, widely distributed along the
+X-axis but not the Y-axis. In these cases, both models are initialized with
+weights that produce Row categories, but because category contrast is not
+considered, categories are uniformly generated across the entire domain, rather
+than concentrated within the upper-regions as observed behaviorally.
 
@@
 
-The creative generation of concepts is an immensely intriguing topic, but it is also complex and difficult to study in a traditional laboratory setting.
-Creative generation is most typically studied using a creative cognition approach \citep{finke1992creative,smith1995creative}, where creative products \citep[such as drawings of alien plants and animals;][]{ward1994structured} are analyzed to obtain insight into the processes and representations involved.
-While this approach has yielded a great deal of information about the nature of creative generation, the commonly employed response modalities (e.g., hand-drawn pictures) make it difficult to develop computational models capable of formally evaluating the proposed representations and processes.
-However, the generation of novel concepts is not altogether different from other behaviors studied in cognitive psychology: In particular, it can be viewed as an form of inference using one's knowledge about existing categories \citep{kemp2014taxonomy,kurtz2015human}.
-Taking advantage of this insight, in this paper we investigated the generation of new categories using the empirical and computational toolkit from the field of categorization.
-
-While the bulk of prior research on the topic has focused on the classic finding that generated concepts tend to be distributionally similar to known concepts, there has been little work addressing the role of contrast in creative generation: How is it that people are able to create something {\em different} from what is already known?
-We developed a novel, exemplar-based model, PACKER, which formally specifies the role of contrast in generation.
-The model proposes that categories are represented as exemplars in a multidimensional psychological space, and generation is constrained both by within-category and between-category similarity: Exemplars belonging to the same category should be similar to one another, and exemplars belonging to different categories should not be similar to one another.
-
-In addition to an analysis of published data \citep[][Experiment 3]{jern2013probabilistic}, we reported two experiments demonstrating systematic effects of category contrast in creative generation.
-Members of participant-generated categories tended to be highly dissimilar from members of previously-learned categories, and were usually more similar to one another than to members of other categories.
-We also observed broad interdependence between the distributional structure (feature variance, correlation) and physical instantiation (location within the stimulus space) of generated categories: In Experiment 2, we found that the unoccupied regions of the domain influenced the distributional structure of categories, and in both experiments we observed that participants distributed their generated categories to increase contrast with what was already known.
-
-We conducted simulations comparing PACKER's account of our results to the other major proposals for category generation: a ``copy-and-tweak'' model (realized as a variant of PACKER with no sensitivity to category contrast), and a hierarchical Bayesian model designed to explain the classic distributional similarity effect.
-In all simulations, we found that PACKER's sensitivity to contrast captured a previously unexplained and unexplored aspect of human category generation.
-Further, by measuring PACKER's fit as a function of its prioritization of within- and between-category similarity, we observed that considering either constraint exclusively results in a relatively low-quality account.
-Instead, PACKER's best results were obtained when both constraints are considered, indicating that human learners do not generate novel concepts exclusively on the basis of within-category similarity or between class-contrast.
-This finding mirrors our behavioral results and demonstrates that both constraints influence creative generation.
+The creative generation of concepts is an immensely intriguing topic, but it is
+also complex and difficult to study in a traditional laboratory setting.
+Creative generation is most typically studied using a creative cognition
+approach \citep{finke1992creative,smith1995creative}, where creative products
+\citep[such as drawings of alien plants and animals;][]{ward1994structured} are
+analyzed to obtain insight into the processes and representations involved.
+While this approach has yielded a great deal of information about the nature of
+creative generation, the commonly employed response modalities (e.g., hand-drawn
+pictures) make it difficult to develop computational models capable of formally
+evaluating the proposed representations and processes. However, the generation
+of novel concepts is not altogether different from other behaviors studied in
+cognitive psychology: In particular, it can be viewed as an form of inference
+using one's knowledge about existing categories
+\citep{kemp2014taxonomy,kurtz2015human}. Taking advantage of this insight, in
+this paper we investigated the generation of new categories using the empirical
+and computational toolkit from the field of categorization.
+
+While the bulk of prior research on the topic has focused on the classic finding
+that generated concepts tend to be distributionally similar to known concepts,
+there has been little work addressing the role of contrast in creative
+generation: How is it that people are able to create something {\em different}
+from what is already known? We developed a novel, exemplar-based model, PACKER,
+which formally specifies the role of contrast in generation. The model proposes
+that categories are represented as exemplars in a multidimensional psychological
+space, and generation is constrained both by within-category and
+between-category similarity: Exemplars belonging to the same category should be
+similar to one another, and exemplars belonging to different categories should
+not be similar to one another.
+
+In addition to an analysis of published data \citep[][Experiment
+3]{jern2013probabilistic}, we reported two experiments demonstrating systematic
+effects of category contrast in creative generation. Members of
+participant-generated categories tended to be highly dissimilar from members of
+previously-learned categories, and were usually more similar to one another than
+to members of other categories. We also observed broad interdependence between
+the distributional structure (feature variance, correlation) and physical
+instantiation (location within the stimulus space) of generated categories: In
+Experiment 2, we found that the unoccupied regions of the domain influenced the
+distributional structure of categories, and in both experiments we observed that
+participants distributed their generated categories to increase contrast with
+what was already known.
+
+We conducted simulations comparing PACKER's account of our results to the other
+major proposals for category generation: a ``copy-and-tweak'' model (realized as
+a variant of PACKER with no sensitivity to category contrast), and a
+hierarchical Bayesian model designed to explain the classic distributional
+similarity effect. In all simulations, we found that PACKER's sensitivity to
+contrast captured a previously unexplained and unexplored aspect of human
+category generation. Further, by measuring PACKER's fit as a function of its
+prioritization of within- and between-category similarity, we observed that
+considering either constraint exclusively results in a relatively low-quality
+account. Instead, PACKER's best results were obtained when both constraints are
+considered, indicating that human learners do not generate novel concepts
+exclusively on the basis of within-category similarity or between
+class-contrast. This finding mirrors our behavioral results and demonstrates
+that both constraints influence creative generation.
 
@@
 
-The proposal that category contrast is a primary constraint in creative generation is, in some ways, entirely commonsense.
-To successful create something \textit{new}, it must be different from what is known.
-Beyond its role in creative generation, category contrast is also of fundamental importance in categorization more broadly.
-All other factors held constant, new categories are easier to learn if they are dissimilar to members of other categories, and knowledge of highly distinct categories is applied more accurately than that of ill-defined categories \citep{ashby1994categorization,imai1965discriminability}.
-Likewise, basic-level categories \citep{rosch1976basic} are thought to be abstracted in order to maximize within-category similarity while minimizing between-category similarity.
-Finally, the act of forming category representations affects similarity judgments about category members and nonmembers, with category members being viewed as more similar to one another than members of other categories \citep{goldstone1994influences,goldstone2001altering}.
-
-Beyond categorization, one can find instances of the trade-off between within and between-class similarity in linguistic categories over perceptual dimensions.
-For example, \cite{regier2007} showed that the partitioning of color categories reflects such a trade-off in a psychological space -- colors are partitioned into groups with members that are viewed as highly similar to one another yet distinct from other colors.
-A similar trade-off can be observed in phoneme categories.
-Different exemplars of the same phoneme must be similar to one another, while contrasting from other phonemes, such that a listener can infer the appropriate phoneme.
-This pattern has been found and modeled in the natural acoustics of American English vowels \citep{feldman2013,hillenbrand1995}.
-As linguistic categories must have been created at some point in human history, it is revealing that the constraints of emulating distributional structure across categories and having categories contrast from one another still bias human category generation today.
-
-The dual forces of within-class similarity and between-class contrast influence cognitive functions in a wide variety of domains.
-The PACKER model is notable in that it successfully interprets this trade-off within the domain of creative generation, and allows us to begin to understand the relatively understudied processes involved in creative generation through our more well-developed knowledge of human categorization.
+The proposal that category contrast is a primary constraint in creative
+generation is, in some ways, entirely commonsense. To successful create
+something \textit{new}, it must be different from what is known. Beyond its role
+in creative generation, category contrast is also of fundamental importance in
+categorization more broadly. All other factors held constant, new categories are
+easier to learn if they are dissimilar to members of other categories, and
+knowledge of highly distinct categories is applied more accurately than that of
+ill-defined categories \citep{ashby1994categorization,imai1965discriminability}.
+Likewise, basic-level categories \citep{rosch1976basic} are thought to be
+abstracted in order to maximize within-category similarity while minimizing
+between-category similarity. Finally, the act of forming category
+representations affects similarity judgments about category members and
+nonmembers, with category members being viewed as more similar to one another
+than members of other categories
+\citep{goldstone1994influences,goldstone2001altering,goldstone1996isolated}.
+
+Beyond categorization, one can find instances of the trade-off between within
+and between-class similarity in linguistic categories over perceptual
+dimensions. For example, \cite{regier2007} showed that the partitioning of color
+categories reflects such a trade-off in a psychological space -- colors are
+partitioned into groups with members that are viewed as highly similar to one
+another yet distinct from other colors. A similar trade-off can be observed in
+phoneme categories. Different exemplars of the same phoneme must be similar to
+one another, while contrasting from other phonemes, such that a listener can
+infer the appropriate phoneme. This pattern has been found and modeled in the
+natural acoustics of American English vowels
+\citep{feldman2013,hillenbrand1995}. As linguistic categories must have been
+created at some point in human history, it is revealing that the constraints of
+emulating distributional structure across categories and having categories
+contrast from one another still bias human category generation today.
+
+The dual forces of within-class similarity and between-class contrast influence
+cognitive functions in a wide variety of domains. The PACKER model is notable in
+that it successfully interprets this trade-off within the domain of creative
+generation, and allows us to begin to understand the relatively understudied
+processes involved in creative generation through our more well-developed
+knowledge of human categorization.
 
@@
 
-Although the focus of this article has been to address the processes involved in creative generation using the empirical and quantitative toolkit of traditional categorization research, our findings and approach have relevant implications for research in creative cognition.
-A central focus of the creative cognition approach has been to explain acts of creativity in terms of the mental representations and processes that are commonly studied in cognitive psychology and cognitive science \citep{finke1992creative,smith1995creative}.
-However, unlike other fields in the study of cognition, creative cognition research rarely employs quantitative models to evaluate the explanatory value of such representations and processes.
-Our modeling results provide a concrete example of how formal approaches may be used to gain insight into the nature of creative cognition.
-
-In addition to demonstrating the utility of formal modeling in the study creative cognition, the PACKER model more specifically offers an additional interpretation of some of the field's most central findings.
-For example, perhaps the most foundational principle from this literature concerns the limiting influence of prior knowledge: Individuals create new categories composed of features from existing classes, and what is created can be influenced drastically through the introduction of cues or examples \citep{marsh1999inadvertent,smith1993constraining}.
-In this paper, we have identified another important aspect of the constraining influence of prior knowledge: What is generated cannot be the same as what is already known.
-Further, there is systematicity in how generated categories differ from prior knowledge.
-The results of our simulations with PACKER demonstrate that this influence is concisely explained in terms of a trade-off between within-category similarity and between-category dissimilarity.
-
-Similarly, PACKER may offer an additional interpretation of existing accounts of creative generation.
-Most notably, a leading account within the creative cognition literature, the Path of Least Resistance \citep{ward1994structured,ward1995s}, also explains generation in terms of an exemplar-based retrieval process.
-This account was designed to explain the creative generation of natural categories (e.g., new species of plants and animals) and as a result relies strongly on the hierarchical organization of these categories: Individuals are thought to retrieve an example of the higher-level category being generated (e.g., \textit{bird} may be retrieved from the category \textit{animal}), and then systematically alter what was retrieved to make something new.
-As the PACKER model does not assume knowledge is hierarchically organized \citep[this is true of the exemplar view more broadly, see][]{murphy2016exemplar}, the model may be viewed as a formal instantiation of the Path of Least Resistance for application in a traditional artificial categorization domain (when there is no established hierarchy of categories).
-PACKER's success in explaining generation within an artificial domain motivates future work exploring the nature of category contrast within a more naturalistic setting.
+Although the focus of this article has been to address the processes involved in
+creative generation using the empirical and quantitative toolkit of traditional
+categorization research, our findings and approach have relevant implications
+for research in creative cognition. A central focus of the creative cognition
+approach has been to explain acts of creativity in terms of the mental
+representations and processes that are commonly studied in cognitive psychology
+and cognitive science \citep{finke1992creative,smith1995creative}. However,
+unlike other fields in the study of cognition, creative cognition research
+rarely employs quantitative models to evaluate the explanatory value of such
+representations and processes. Our modeling results provide a concrete example
+of how formal approaches may be used to gain insight into the nature of creative
+cognition.
+
+In addition to demonstrating the utility of formal modeling in the study
+creative cognition, the PACKER model more specifically offers an additional
+interpretation of some of the field's most central findings. For example,
+perhaps the most foundational principle from this literature concerns the
+limiting influence of prior knowledge: Individuals create new categories
+composed of features from existing classes, and what is created can be
+influenced drastically through the introduction of cues or examples
+\citep{marsh1999inadvertent,smith1993constraining}. In this paper, we have
+identified another important aspect of the constraining influence of prior
+knowledge: What is generated cannot be the same as what is already known.
+Further, there is systematicity in how generated categories differ from prior
+knowledge. The results of our simulations with PACKER demonstrate that this
+influence is concisely explained in terms of a trade-off between within-category
+similarity and between-category dissimilarity.
+
+Similarly, PACKER may offer an additional interpretation of existing accounts of
+creative generation. Most notably, a leading account within the creative
+cognition literature, the Path of Least Resistance
+\citep{ward1994structured,ward1995s}, also explains generation in terms of an
+exemplar-based retrieval process. This account was designed to explain the
+creative generation of natural categories (e.g., new species of plants and
+animals) and as a result relies strongly on the hierarchical organization of
+these categories: Individuals are thought to retrieve an example of the
+higher-level category being generated (e.g., \textit{bird} may be retrieved from
+the category \textit{animal}), and then systematically alter what was retrieved
+to make something new. As the PACKER model does not assume knowledge is
+hierarchically organized \citep[this is true of the exemplar view more broadly,
+see][]{murphy2016exemplar}, the model may be viewed as a formal instantiation of
+the Path of Least Resistance for application in a traditional artificial
+categorization domain (when there is no established hierarchy of categories).
+PACKER's success in explaining generation within an artificial domain motivates
+future work exploring the nature of category contrast within a more naturalistic
+setting.
 
@@
 
-Although our work has mainly addressed the processes involved in the creative generation of  concepts, in this paper we have studied these processes as they apply within an artificial categorization domain.
-By consequence, our findings also provide an advance in our understanding categorization more broadly.
-
-Categorization research addresses the representations and processes that underlie the learning and use of categories.
-Category learning tasks are generally about figuring out which items belong to which category.
-Once learned, categories are generally used to classify new stimuli and to make inferences beyond the available information.
-Our work is fairly unique in that people learn a single category through positive examples and then figure out another category that would make sense in the domain.
-In other words, only one category is provided to be learned and the use that is required of that category is to generate a contrast category.
-Such propagating of categories (producing one from another) may not seem like a fundamental or ubiquitous cognitive activity, but consider navigating a highly unfamiliar environment -- it would be practical, perhaps even critical, to generate expectations about what other kinds of things you are likely to encounter based on those that you already have.
-In the present studies, we have learned something about the form that such expectations are likely to take.
-
-We can think of the category generation task in our studies as asking a person to formulate an idea about what set of items in the domain are most interestingly {\em not} members of the original category.
-To meet this condition, the items must take some form of coherence that aligns with that of the original category and some form of distinctiveness relative to the original category.
-Reflecting the basic level of organization in natural categories, it makes sense to generate a set of items that possess strong within-category coherence (by importing or systematically transforming the internal structure of the original category) as well as strong between-category differentiation (by creating maximum contrast with the original category).
-In this sense one can see the patterns of performance in the category generation task as recapitulating the order of semantic organization.
-
-\subsection{Contrast and Representativeness in Categorization}
-Xian: Please write about the two hypotheses and that followup work needs to look into why the Bayesian representativeness model doesn't do as well at explaining category learning ease, but does better at category generation.
-
-MAKE SURE WE CITE AND DESCRIBE \citet{thomas98}!
-
-\subsection{Limitations and Future Directions}
-Although we have demonstrated that contrast plays a role in categorization, our results were equivocal as to whether contrast is governed by exemplar dissimilarity or representativeness.
-
-Although successful in explaining our results in all three studies (albeit not as well as the representativeness model for Experiments 1 and 2), PACKER does not provide a full account of what is known about category generation.
-Most notably, in this paper we have not evaluated the model's ability to explain the classic finding that generated categories tend to share distributional commonalities with previously learned categories \citep[see][]{jern2013probabilistic,ward1994structured}.
-While we successfully replicated this effect in Experiment 1, we also found that its influence was limited in comparison to the fundamental constraints imposed by category contrast.
-Even within Experiment 1, we found systematic inconsistencies: by generating exemplars into unoccupied regions of the space, participants who learned an `XOR' category, composed of members that are widely distributed along both features and are positively correlated in space, tended to generate categories with an opposite (negative) correlation.
-More generally, PACKER inherits the strengths and weaknesses of exemplar models of categorization: It provides a simple and flexible model that explains many results, but deviates systematically from human performance in some cases.
-
-Nonetheless, these classic effects are a core element of the phenomenology of creative generation, and PACKER does not include any mechanisms that explain them.
-Instead, through the development and evaluation of the PACKER model, we have sought to add new elements into such a phenomenology: The broad and strong influence of category contrast, and the interdependence between category location and distributional structure.
-It may be possible to combine the hierarchical Bayesian approach proposed by \cite{jern2013probabilistic} with PACKER's underlying claims to obtain a ``best of both worlds'' model, capable of explaining the role of contrast in creative generation, as well as the emulation of distributional structure.
-However, as noted in the introduction, the incorporation of category contrast is antithetical to the core principles of a traditional, semi-conjugate Bayesian approach.
-This suggests that category generation is a fundamentally different computational-level problem \citep[different from those posed by][]{jern2013probabilistic,kemp2014taxonomy}.
-
-Characterizing that problem and conducting a rational analysis is an important direction for future research.
-To that aim, we plan to explore the connection between exemplar modeling as an Importance Sampling approximation \citep{shi10exemplar}, and see what sort of computational-level problem PACKER approximates.
-Once formalized in probabilistic terms, it should also be straightforward to incorporate distributional factors into the model.
-This would unite PACKER with the Bayesian representativeness model, and they would differ in terms of their assumptions about how people represent category distributions (as exemplars or prototypes, respectively).
-Alternatively, it may be possible to integrate the core principles of contrast into other categorization models \citep[e.g.,][]{love2004sustain,kurtz2007divergent,smith2000thirty}.
-
-%Finally, although an overall goal of this work is to advance the study of creative thought by explaining generation through empirical and quantitative approaches from research in categorization, these approaches are not easily aligned with many of the alternative theoretical viewpoints in the broader study of creativity \citep[for a review see][]{kozbelt2010theories}, such as those based on free association \citep{mednick1962associative} and conceptual combination \citep{estes2002emergence,murphy1988comprehending}.
-Instead, we reduced our focus by studying a highly complex behavior (creative generation) as it applies within a well-established domain (artificial category learning).
-We hope that future work incorporates and highlights the importance of contrast into theories of creativity.
+Although our work has mainly addressed the processes involved in the creative
+generation of concepts, in this paper we have studied these processes as they
+apply within an artificial categorization domain. By consequence, our findings
+also provide an advance in our understanding categorization more broadly.
+
+Categorization research addresses the representations and processes that
+underlie the learning and use of categories. Category learning tasks are
+generally about figuring out which items belong to which category. Once learned,
+categories are generally used to classify new stimuli and to make inferences
+beyond the available information. Our work is fairly unique in that people learn
+a single category through positive examples and then figure out another category
+that would make sense in the domain. In other words, only one category is
+provided to be learned and the use that is required of that category is to
+generate a contrast category. Such propagating of categories (producing one from
+another) may not seem like a fundamental or ubiquitous cognitive activity, but
+consider navigating a highly unfamiliar environment -- it would be practical,
+perhaps even critical, to generate expectations about what other kinds of things
+you are likely to encounter based on those that you already have. In the present
+studies, we have learned something about the form that such expectations are
+likely to take.
+
+We can think of the category generation task in our studies as asking a person
+to formulate an idea about what set of items in the domain are most
+interestingly {\em not} members of the original category. To meet this
+condition, the items must take some form of coherence that aligns with that of
+the original category and some form of distinctiveness relative to the original
+category. Reflecting the basic level of organization in natural categories, it
+makes sense to generate a set of items that possess strong within-category
+coherence (by importing or systematically transforming the internal structure of
+the original category) as well as strong between-category differentiation (by
+creating maximum contrast with the original category, be it through
+exemplar-based dissimilarity or the maximization of exemplar-category
+representativeness). In this sense one can see the patterns of performance in
+the category generation task as recapitulating the order of semantic
+organization.
+
+\subsection{Exemplar dissimilarity and Representativeness in Categorization}
+
+\sxlnote{I'm struggling a little with this section! That said, I think this may
+  also be a good point to address Charles' concern re: the possible conditions
+  under which contrast are important.} 
+Although both PACKER and the
+representativeness model perform better than the other models, they appear to be
+capturing fundamentally different aspects of contrast in categorization.
+Specifically, while the representativeness model excelled in predicting our
+category generation data, PACKER was the best-fitting model in accounting for
+participant error in category learning. The source of this discrepancy is still
+not yet clear -- it could be an indication of a fundamental difference between
+the processes involved in generation versus learning where exemplar
+dissimilarity is a larger factor in category generation while representativeness
+exerts more influence in category learning.
+
+Xian: Please
+write about the two hypotheses and that followup work needs to look into why the
+Bayesian representativeness model doesn't do as well at explaining category
+learning ease, but does better at category generation.
+
+MAKE SURE WE CITE AND DESCRIBE \citet{thomas98}! \sxlnote{I moved the
+  description and citation of thomas98 to the introduction for justification of
+  why we think categories should share similar distributions.}
+\subsection{Limitations and Future Directions} Although we have demonstrated
+that contrast plays a role in categorization, our results were equivocal as to
+whether contrast is governed by exemplar dissimilarity or representativeness.
+
+Although successful in explaining our results in all three studies (albeit not
+as well as the representativeness model for Experiments 1 and 2), PACKER does
+not provide a full account of what is known about category generation. Most
+notably, in this paper we have not evaluated the model's ability to explain the
+classic finding that generated categories tend to share distributional
+commonalities with previously learned categories
+\citep[see][]{jern2013probabilistic,ward1994structured}. While we successfully
+replicated this effect in Experiment 1, we also found that its influence was
+limited in comparison to the fundamental constraints imposed by category
+contrast. Even within Experiment 1, we found systematic inconsistencies: by
+generating exemplars into unoccupied regions of the space, participants who
+learned an `XOR' category, composed of members that are widely distributed along
+both features and are positively correlated in space, tended to generate
+categories with an opposite (negative) correlation. More generally, PACKER
+inherits the strengths and weaknesses of exemplar models of categorization: It
+provides a simple and flexible model that explains many results, but deviates
+systematically from human performance in some cases.
+
+Nonetheless, these classic effects are a core element of the phenomenology of
+creative generation, and PACKER does not include any mechanisms that explain
+them. Instead, through the development and evaluation of the PACKER model, we
+have sought to add new elements into such a phenomenology: The broad and strong
+influence of category contrast, and the interdependence between category
+location and distributional structure. It may be possible to combine the
+hierarchical Bayesian approach proposed by \cite{jern2013probabilistic} with
+PACKER's underlying claims to obtain a ``best of both worlds'' model, capable of
+explaining the role of contrast in creative generation, as well as the emulation
+of distributional structure. However, as noted in the introduction, the
+incorporation of category contrast is antithetical to the core principles of a
+traditional, semi-conjugate Bayesian approach. This suggests that category
+generation is a fundamentally different computational-level problem
+\citep[different from those posed by][]{jern2013probabilistic,kemp2014taxonomy}.
+
+Characterizing that problem and conducting a rational analysis is an important
+direction for future research. To that aim, we plan to explore the connection
+between exemplar modeling as an Importance Sampling approximation
+\citep{shi10exemplar}, and see what sort of computational-level problem PACKER
+approximates. Once formalized in probabilistic terms, it should also be
+straightforward to incorporate distributional factors into the model. This would
+unite PACKER with the Bayesian representativeness model, and they would differ
+in terms of their assumptions about how people represent category distributions
+(as exemplars or prototypes, respectively). Alternatively, it may be possible to
+integrate the core principles of contrast into other categorization models
+\citep[e.g.,][]{love2004sustain,kurtz2007divergent,smith2000thirty}.
+
+Finally, although an overall goal of this work is to advance the study of
+creative thought by explaining generation through empirical and quantitative
+approaches from research in categorization, these approaches are not easily
+aligned with many of the alternative theoretical viewpoints in the broader study
+of creativity \citep[for a review see][]{kozbelt2010theories}, such as those
+based on free association \citep{mednick1962associative} and conceptual
+combination \citep{estes2002emergence,murphy1988comprehending}. Instead, we
+reduced our focus by studying a highly complex behavior (creative generation) as
+it applies within a well-established domain (artificial category learning). We
+hope that future work incorporates and highlights the importance of contrast
+into theories of creativity.
 
@@
 
-The generation of new concepts and ideas is a highly interesting topic, but it is difficult to study in a controlled experimental environment.
-In this paper, we have provided such an examination of category generation as it applies within an artificial categorization experiment.
-Extending the literature on creative cognition, our experiments provide a detailed picture of the role of category contrast in generation: People seek to create concepts that are distinct from what they already know, and the nature of what is created can be influenced by what does not yet exist.
-Our simulations with traditional exemplar models, as well as a hierarchical Bayesian model, provide strong support for the claim that category contrast is of fundamental importance.
-More generally, our results demonstrate that popular explanatory approaches from basic research in cognitive science can offer a precise, quantitative account of a behavior as complex as that of creative generation.
+The generation of new concepts and ideas is a highly interesting topic, but it
+is difficult to study in a controlled experimental environment. In this paper,
+we have provided such an examination of category generation as it applies within
+an artificial categorization experiment. Extending the literature on creative
+cognition, our experiments provide a detailed picture of the role of category
+contrast in generation: People seek to create concepts that are distinct from
+what they already know, and the nature of what is created can be influenced by
+what does not yet exist. Our simulations with traditional exemplar models, as
+well as a hierarchical Bayesian model, provide strong support for the claim that
+category contrast is of fundamental importance. More generally, our results
+demonstrate that popular explanatory approaches from basic research in cognitive
+science can offer a precise, quantitative account of a behavior as complex as
+that of creative generation.
 
 \clearpage
-\section{Acknowledgments}
-Previous versions of this work were presented at the Thirty-Ninth Annual Conference of the Cognitive Science Society and Forty-Ninth Annual Meeting of the Society for Mathematical Psychology.
-Support for this research was provided by the Office of the VCRGE at the UW - Madison with funding from the WARF.
-We thank Alan Jern and Charles Kemp for providing code and data.
+\section{Acknowledgments} Previous versions of this work were presented at the
+Thirty-Ninth Annual Conference of the Cognitive Science Society and Forty-Ninth
+Annual Meeting of the Society for Mathematical Psychology. Support for this
+research was provided by the Office of the VCRGE at the UW - Madison with
+funding from the WARF. We thank Alan Jern and Charles Kemp for providing code
+and data.
 \end{flushleft}
@@
 % references section
+\clearpage \bibliographystyle{apacite} \bibliography{citations.bib}
 \clearpage
-\bibliographystyle{apacite}
-\bibliography{citations.bib}
-\clearpage
-
-\appendix
-\numberwithin{equation}{section}
+
+\appendix \numberwithin{equation}{section}
 
@@
 
-\cite{jern2013probabilistic} demonstrated how a hierarchical Bayesian model could explain the distributional correspondences between observed and generated categories.
-In their model, exemplars of generated categories were viewed as samples from a multivariate Normal distribution over the dimensions of stimulus space.
-The mean of the generated category was independent of the observed categories, but the covariance matrix (encoding feature variances and correlations) was based on a common prior distribution.
-Generating a new category was thus completed by sampling a new category mean (uniform over stimulus space) and covariance matrix from the common prior distribution.
-Because the shared prior distribution's parameters were unobserved, the hierarchical Bayesian approach was used to infer its parameters from the previous categories (their feature variances and correlations), and then to generate the covariance matrix of the new category.
-
-In our implementation of their model\footnote{Note that \citet{jern2013probabilistic}'s model is slightly different, as they used a semi-conjugate model.
-Their model acts very similarly to our version.}, each category's exemplars are viewed as samples from a multivariate Normal distribution with parameters ($\mu, \Sigma$).
-Category covariance matrices (specifying variance and covariance along $k$-dimensions), are assumed to be Normal-Inverse-Wishart distributed with parameters: $\nu$ ($>k-1$), $\kappa$ ($>0$), and $\Sigma_D$.
-$\nu$ and $\kappa$ are treated as free parameters in our simulations, and $\Sigma_D$ is the domain-wide covariance matrix from which all categories are viewed as samples.
-Assuming a given $\Sigma_D$, a category covariance matrix $\Sigma$ can be computed on the basis of its examples:
-
-\begin{equation}
-\Sigma = \left[ \Sigma_D \nu + C +
-\dfrac
-{\kappa n}
-{\kappa + n}
-(\bar{x}-\mu)(\bar{x}-\mu)^T
-\right] (\nu + n)^{-1}
+\cite{jern2013probabilistic} demonstrated how a hierarchical Bayesian model
+could explain the distributional correspondences between observed and generated
+categories. In their model, exemplars of generated categories were viewed as
+samples from a multivariate Normal distribution over the dimensions of stimulus
+space. The mean of the generated category was independent of the observed
+categories, but the covariance matrix (encoding feature variances and
+correlations) was based on a common prior distribution. Generating a new
+category was thus completed by sampling a new category mean (uniform over
+stimulus space) and covariance matrix from the common prior distribution.
+Because the shared prior distribution's parameters were unobserved, the
+hierarchical Bayesian approach was used to infer its parameters from the
+previous categories (their feature variances and correlations), and then to
+generate the covariance matrix of the new category.
+
+In our implementation of their model\footnote{Note that
+\citet{jern2013probabilistic}'s model is slightly different, as they used a
+semi-conjugate model. Their model acts very similarly to our version.}, each
+category's exemplars are viewed as samples from a multivariate Normal
+distribution with parameters ($\mu, \Sigma$). Category covariance matrices
+(specifying variance and covariance along $k$-dimensions), are assumed to be
+Normal-Inverse-Wishart distributed with parameters: $\nu$ ($>k-1$), $\kappa$
+($>0$), and $\Sigma_D$. $\nu$ and $\kappa$ are treated as free parameters in our
+simulations, and $\Sigma_D$ is the domain-wide covariance matrix from which all
+categories are viewed as samples. Assuming a given $\Sigma_D$, a category
+covariance matrix $\Sigma$ can be computed on the basis of its examples:
+
+\begin{equation} \Sigma = \left[ \Sigma_D \nu + C + \dfrac {\kappa n} {\kappa +
+n} (\bar{x}-\mu)(\bar{x}-\mu)^T \right] (\nu + n)^{-1}
 \label{eq:Sigma_B}
@@
 % 
-where $\bar{x}$ and $C$ are the empirical mean and covariance of the category's known members, and $n$ is the number of observed members of the category.
-When there are fewer than two known members of the category (and thus no covariance to speak of), $\Sigma = \Sigma_D\nu$.
+where $\bar{x}$ and $C$ are the empirical mean and covariance of the category's
+known members, and $n$ is the number of observed members of the category. When
+there are fewer than two known members of the category (and thus no covariance
+to speak of), $\Sigma = \Sigma_D\nu$.
 
@@
 
-\begin{equation}
-  \mu = \dfrac
-    {\kappa\mu_{0} + n \bar{x}}
-    {\kappa + n}
-    \label{eq:category_mus}
+\begin{equation} \mu = \dfrac {\kappa\mu_{0} + n \bar{x}} {\kappa + n}
+    \label{eq:category_mus_jk13}
 \end{equation}
 % 
-where $\mu_{0}$ is the prior mean.
-In our simulations, $\mu_{0}$ is set to the center of the domain.
-However, when no examples of the target category have been observed, generation is assumed to be random.
-In practice, the model's best fits are achieved when the  $\kappa$ parameter, which controls the influence of $\mu_0$ on $\mu$, is set very close to zero (hence, the influence of $\mu_0$ is minimal).
-
-Importantly, the domain-wide covariance matrix $\Sigma_D$ is unobserved and needs to be inferred from the observed categories.
-For conjugacy, if $\Sigma_D$ is viewed as a sample from an Inverse-Wishart distribution with scale $\Sigma_0$, $\Sigma_D$ can be computed as:
-
- \begin{equation}
-    \Sigma_D = \Sigma_0 + \sum_y{C_y}
-    \label{eq:category_mus}
+where $\mu_{0}$ is the prior mean. In our simulations, $\mu_{0}$ is set to the
+center of the domain. However, when no examples of the target category have been
+observed, generation is assumed to be random. In practice, the model's best fits
+are achieved when the $\kappa$ parameter, which controls the influence of
+$\mu_0$ on $\mu$, is set very close to zero (hence, the influence of $\mu_0$ is
+minimal).
+
+Importantly, the domain-wide covariance matrix $\Sigma_D$ is unobserved and
+needs to be inferred from the observed categories. For conjugacy, if $\Sigma_D$
+is viewed as a sample from an Inverse-Wishart distribution with scale
+$\Sigma_0$, $\Sigma_D$ can be computed as:
+
+ \begin{equation} \Sigma_D = \Sigma_0 + \sum_y{C_y}
+    \label{eq:category_sigmas}
 \end{equation}
 % 
-where $\Sigma_0$ is the prior covariance in the domain.
-In our simulations, $\Sigma_0 = \lambda {\bf I}$, where $\lambda$ is a free parameter controlling the expected variance of dimensions (dimensions of the domain covariance matrix are expected to be uncorrelated) and ${\bf I}$ is a $k$-by-$k$ identity matrix.
-
-Generated exemplars are drawn from a multivariate Normal distribution specified by $(\mu, \Sigma)$. Thus, $p(y)$ is
-\begin{equation}
-p(y \mid x) = \dfrac
-{\exp \left \{ \theta \cdot {\rm Normal}(y; \mu, \Sigma) \right \} }
-{\sum_i \exp \left \{ \theta \cdot {\rm Normal}(y_i; \mu, \Sigma) \right \} } 
+where $\Sigma_0$ is the prior covariance in the domain. In our simulations,
+$\Sigma_0 = \lambda {\bf I}$, where $\lambda$ is a free parameter controlling
+the expected variance of dimensions (dimensions of the domain covariance matrix
+are expected to be uncorrelated) and ${\bf I}$ is a $k$-by-$k$ identity matrix.
+
+Generated exemplars are drawn from a multivariate Normal distribution specified
+by $(\mu, \Sigma)$. Thus, $p(y)$ is
+\begin{equation} p(y \mid x) = \dfrac {\exp \left \{ \theta \cdot {\rm
+Normal}(y; \mu, \Sigma) \right \} } {\sum_i \exp \left \{ \theta \cdot {\rm
+Normal}(y_i; \mu, \Sigma) \right \} }
 \end{equation}
 % 
-where $\theta$ is a response determinism parameter and ${\rm Normal}(y; \mu, \Sigma)$ denotes a multivariate Normal density evaluated at $y$.
+where $\theta$ is a response determinism parameter and ${\rm Normal}(y; \mu,
+\Sigma)$ denotes a multivariate Normal density evaluated at $y$.
 

### 2018-07-17 Manuscripts/cog-psych/revision/methods_results_e3.tex
Added representativeness to manuscript. Updated figures and results. Discussion in various places and some reviewer-related responses still need work.

@@
 \section{Experiment 3}
-Experiments 1 and 2 clearly establish the importance of contrast in category generation.
-Follow-up model-based analyses illustrated that both contrast models account for participant performance better than models that do not take into account contrast.
-We also found more support for contrast as representativeness hypothesis over contrast as exemplar dissimilarity hypothesis.
 
-Although category generation is an intriguing aspect of human cognition, it has played a relatively limited role in the broader categorization literature.
-To establish that contrast plays a broader role in categorization, we conducted a traditional category learning study.
-In it, participants learned two categories, where the categories learned by each participant in Experiment 3 was yoked to the Alpha shown and Beta generated by each participant in Experiment 2.
-If contrast plays a broader role in categorization, then Beta categories rated as "better" according to the contrast models should be easier for participants to learn.
+Experiments 1 and 2 clearly establish the importance of contrast in category
+generation. Follow-up model-based analyses illustrated that both contrast models
+account for participant performance better than models that do not take into
+account contrast. We also found more support for contrast as representativeness
+hypothesis over contrast as exemplar dissimilarity hypothesis.
+
+\sxlnote{I felt like there could be a little more explanation in terms of
+  justifying this experiment so I added an extra sentence (the second last in
+  the following paragraph) - I don't have too strong feelings about it so please
+  feel free to remove if you think it's too much. Also see the footnote
+  addressing the paper's lack of category learning fits. It's a bit of a cop
+  out, but entirely true!}
+
+Although category generation is an intriguing aspect of human cognition, it has
+played a relatively limited role in the broader categorization literature. To
+establish that contrast plays a broader role in categorization, we conducted a
+traditional category learning study. In it, participants learned two categories,
+where the categories learned by each participant in Experiment 3 was yoked to
+the Alpha shown and Beta generated by each participant in Experiment 2. Our goal
+is to investigate whether model performance (in generating categories) is
+consistent with human performance (in learning those categories).\footnote{It is
+  possible to adjust the present models of category generation such that they
+  can make category learning predictions. However, such an adjustment makes the
+  strong assumption that the very same processes governing category generation
+  also apply to category learning. Taking this approach would fall outside the
+  scope of the current investigation.} Specifically, if contrast plays a
+broader role in categorization, then Beta categories rated as "better" according
+to the contrast models should be easier for participants to learn.
 
@@
 
-\subsection{Results}
+\subsection{Results and Analysis}
 
@@
 Overall accuracy of the participants was high, with a mean error rate of $.19$
-($SD = .19$; min max?). Error rates for each block are presented in Figure
+($SD = .19$). Error rates for each block are presented in Figure
 \ref{fig:learningcurve}.
@@
 
-In the previous section (or Section X.X), we demonstrated that the best
-performing models were the two contrast models. In this section, our goal is to
-investigate if contrast is also important in category learning tasks.
+In order to analyze the consistency between model and human performance, we
+optimized each model such that the Pearson correlation between the model's
+negative log-likelihood of generating each unique category set and the
+participant's error in learning that category set is maximized. As with the
+analysis in Section \ref{section:individual-diff}, we simulated the weighting of
+each dimension independently depending on the generated categories.
 
-It is also the
-goal of this paper to demonstrate that there is an association between the
-model's [justifying the errorvscat plot is harder than I thought] To compare the
-performances of each model, we analysed the correlation between each model's fit
-to a participant's unique category set and the participant's error rate.
-Intuitively, a well-performing model should be able to closely fit (i.e., easily
-generate) A positive correlation would indicate that the model is We used the
-same set of optimised parameters found from Section X.X, with individual values
-for selective attention.
+The greatest correlations were observed when fitting with the contrast models.
+Specifically, PACKER correlated with human performance with
+$r = .65, 95\% CI$\footnote{Confidence intervals were obtained using a bootstrap
+  method with 1000 bootstrap samples.}$ [.47,.78]$, and the representativeness
+model correlated with human performance with $r = .56, 95\% CI [.38, .72]$.
+Copy-and-tweak yielded a correlation of $r = .45, 95\% CI [.29, .61]$ and the
+hierarchical Bayesian model yielded a similar correlation of
+$r = .46, 95\% CI [.30, .61]$. These results are graphically presented in Figure
+\ref{fig:perror_corr}.
 
@@
     \begin{center}
-    \includegraphics[width=\textwidth]{figs/perror_corr.pdf}
-    \caption{Correlation between observed participant error and model fit.}
+    \includegraphics[width=\textwidth]{figs/modelvspptp.pdf}
+    \caption{Scatterplots indicating the correlations between observed
+      participant error and model fit. Blue discontinuous line are best-fit
+      lines. Green discontinuous lines indicate boundaries generated by the 95\%
+      CI around the fitted correlations. }
     \label{fig:perror_corr}
@@
 between category generation and category learning, we computed the correlations
-over wide range of $\theta_{contrast}$ values, with the other parameter values of
-PACKER held constant at their optimized levels. As presented in Figure
+over wide range of $\theta_{contrast}$ values, with the other parameter values
+of PACKER held constant at their optimized levels. As presented in Figure
 \ref{fig:packer-corr}, the correlation quickly increases with increasing weight
@@
 function of the $\theta_{contrast}$ parameter. To facilitate comparison,
-      PACKER's other parameters ($c$, $\theta_{target}$) were set to the best
-      fitting values obtained for copy-and-tweak in Table
-      \ref{table:global-model-fits}. Grey dashed line represents the correlation
-      between Copy \& Tweak's fit and participant error.}
+PACKER's other parameters ($c$, $\theta_{target}$) were set to the best fitting
+values obtained for copy-and-tweak in Table \ref{table:global-model-fits}. Grey
+dashed line represents the correlation between Copy \& Tweak's fit and
+participant error.}
     \label{fig:packer-corr}
@@
 \end{figure}
-
-The specific reason for why contrast as implemented in PACKER is superior to
-contrast as implemented in the representativeness model is still not clear.
-
-[Note Joe's comments on figure formatting (font size, remove redundant y-axes,
-etc)]

### 2018-07-17 Experiments/cogpsych-code/plot-packer-loglike.py

@@ -1,80 +1,0 @@
-import pickle
-import pandas as pd
-import numpy as np
-import matplotlib.pyplot as plt
-
-execfile('Imports.py')
-import Modules.Funcs as funcs
-from Modules.Classes import Simulation
-from Modules.Classes import Packer, CopyTweak
-
-# get trials pickle
-# pickle can be "pickles/all_data_e1_e2.p", "pickles/nosofsky1986.p"
-with open("pickles/all_data_e1_e2.p", "rb" ) as f:
-	trials = pickle.load( f )
-
-# get best params pickle
-# pickle can be "pickles/best_params_all_data_e1_e2.p", "pickles/best_params_nosofsky1986.p"
-with open("pickles/best_params_all_data_e1_e2.p", "rb" ) as f:
-	best_params = pickle.load( f )
-
-# compute copytweak loglike
-start_params = best_params[CopyTweak.model]
-
-# set up grid of param values
-gamma_grid = np.linspace(0, 1.0, 100)
-loglikes = np.empty(gamma_grid.shape)
-
-# add task type to trials object
-trials.task = 'generate'
-
-# evaluate loglike at each grid point
-for i, val in enumerate(gamma_grid):
-	curr = start_params.copy()
-	curr['tradeoff'] = val
-	loglikes[i] = -trials.loglike(curr, Packer)
-
-
-for i in range(len(gamma_grid)):
-	print i, gamma_grid[i], loglikes[i]
-
-# plot it
-fh = plt.figure(figsize=(5,3))
-
-# copytweak annotation
-copytweak_loglike = loglikes[gamma_grid == 1]
-plt.plot([min(gamma_grid), max(gamma_grid)], [copytweak_loglike, copytweak_loglike], '--', linewidth = 1, color='gray')
-plt.text(0.5, copytweak_loglike, 'Target Only (Copy \& Tweak)', ha = 'center', va = 'bottom')
-
-# contrast only annotation
-contrast_loglike = loglikes[gamma_grid == 0]
-plt.plot([min(gamma_grid), max(gamma_grid)], [contrast_loglike, contrast_loglike], '--', linewidth = 1, color='gray')
-plt.text(0.5, contrast_loglike, 'Contrast Only', ha = 'center', va = 'bottom')
-
-# PACKER annotation
[…]

### 2018-07-17 Experiments/cogpsych-code/plot-stats-gradients.py

@@ -25,9 +25,9 @@ WT_THETA = 1.5
 
 # plotting settings
 fontsettings = dict(fontsize = 12.0)
-col_order = ['Behavioral', 'PACKER', 'Copy and Tweak', 'Hierarchical Sampling','Hierarchical Sampling With Representativeness']
+col_order = ['Behavioral', 'PACKER', 'Copy and Tweak','Hierarchical Sampling With Representativeness', 'Hierarchical Sampling']
 #col_order = ['Hierarchical Sampling With Representativeness']
-col_names_short = ['Behavioral', 'PACKER', 'Copy & Tweak', 'Hier. Sampling','Hier. Sampling w/ Rep']
+col_names_short = ['Behavioral', 'PACKER', 'Copy & Tweak','Representativeness', 'Hier. Bayes']
 row_order = ['Cluster','Row', 'XOR', 'Bottom', 'Middle']
 SMOOTHING_PARAM = 0.8
 
@@ -123,7 +123,7 @@ for model_name, model_obj in name_2_object.items():
     all_data[model_obj.model] = model_data
 
 # plotting
-f, ax = plt.subplots(5,len(name_2_object)+1,figsize = (10, 8))
+f, ax = plt.subplots(5,len(name_2_object)+1,figsize = (10, 10))
 for rownum, c in enumerate(row_order):
     A = stimuli[alphas[c],:]
     
@@ -179,11 +179,11 @@ cbar.set_xticklabels([
 ],**fontsettings)
 cbar.tick_params(length = 0)
 
-plt.tight_layout(w_pad=-8.0, h_pad= 1)
+plt.tight_layout(w_pad=-2.0, h_pad= .5)
 
-fname = 'gradients-t-' + STAT_OF_INTEREST
+fname = 'gradients-' + STAT_OF_INTEREST
 f.savefig(fname + '.pdf', bbox_inches='tight', transparent=False)
 #f.savefig(fname + '.png', bbox_inches='tight', transparent=False)
 
-#path = '../../Manuscripts/cog-psych/figs/range-diff-gradients.pgf'
-#funcs.save_as_pgf(f, path)
+path = '../../Manuscripts/cog-psych/revision/figs/range-diff-gradients.pgf'
+funcs.save_as_pgf(f, path)

### 2018-07-17 Experiments/cogpsych-code/temp.py

@@ -1,5 +1,6 @@
+
 # plotting
-f, ax = plt.subplots(5,len(name_2_object)+1,figsize = (10, 8))
+f, ax = plt.subplots(5,len(name_2_object)+1,figsize = (10, 10))
 for rownum, c in enumerate(row_order):
     A = stimuli[alphas[c],:]
     
@@ -55,11 +56,11 @@ cbar.set_xticklabels([
 ],**fontsettings)
 cbar.tick_params(length = 0)
 
-plt.tight_layout(w_pad=-8.0, h_pad= 1)
+plt.tight_layout(w_pad=-2.0, h_pad= .5)
 
-fname = 'gradients-t-' + STAT_OF_INTEREST
-#f.savefig(fname + '.pdf', bbox_inches='tight', transparent=False)
+fname = 'gradients-' + STAT_OF_INTEREST
+f.savefig(fname + '.pdf', bbox_inches='tight', transparent=False)
 #f.savefig(fname + '.png', bbox_inches='tight', transparent=False)
 
-#path = '../../Manuscripts/cog-psych/figs/range-diff-gradients.pgf'
-#funcs.save_as_pgf(f, path)
+path = '../../Manuscripts/cog-psych/revision/figs/range-diff-gradients.pgf'
+funcs.save_as_pgf(f, path)

### 2018-07-17 Modules/Classes/Model.py

@@ -128,7 +128,7 @@ class Model(object):
     @abc.abstractmethod
     def get_generation_ps(self, stimuli, category, task='generate',seedrng=False): pass
 
-    def __init__(self, categories, params, stimrange=[{'min':-1,'max':1}]):
+    def __init__(self, categories, params, stimrange=[{'min':-1,'max':1},{'min':-1,'max':1}]):
         """
         Initialize the model. "categories" should be a list of numpy
         arrays with the same number of columns (features). Items in 

### 2018-07-20 Manuscripts/cog-psych/revision/cogpsych.tex
major update. many things. to ken/xian

@@
 \section*{Abstract}
-\jlanote{need to update} 
-\sxlnote{updated} 
-%
-The ability to creatively generate new concepts and ideas is among the most
+The ability to generate new concepts and ideas is among the most
 fascinating aspects of human cognition, but we do not have a strong
@@
 We report three experiments demonstrating that contrast between what is known
-and what is created is of fundamental importance for categorization and the
-creative process. We propose two novel approaches to modeling category contrast:
+and what is created is of fundamental importance for categorization and innovation. We propose two novel approaches to modeling category contrast:
 one focused on exemplar dissimilarity and another on the representativeness
@@
 representative of the domain (and thus, not new at all).
-
-\sxlnote{I think we can address Charles' point on real world examples of how
-  contrast can be important around here.}
+Further, contrast has played a critical role studying the spread of new ideas in diverse fields, such as marketing \citep{berger16} and sociology \citep{rogers03}.
+
 To successfully generate something novel, what is generated must be different
@@
 demonstrates that contrast plays a central role in other important
-categorization tasks, which is not specific to category generation.
+categorization tasks, and it is not specific to category generation.
 
@@
 work on the topic of category generation, as well as the computational
-formalizations of the theories in those reports. We then present two hypotheses
+formalizations of their theoretical explanations. We then present two hypotheses
 for how contrast might affect categorization and formalize them in computational
-models. The first is a novel computational model, which is designed to generate
+models. The first is a novel exemplar model, which is designed to generate
 categories that systematically differ from existing categories in the domain.
@@
 each model in capturing human category generation. Afterwards, we find support
-for category contrast predicting the ease at which people learn a pair of novel
-categories in an empirical experiment. We conclude with a discussion of the
+for category contrast predicting the ease at which people learn a pair of 
+categories in a novel empirical experiment. We conclude with a discussion of the
 implications of our results for categorization and creative cognition, and
@@
 
-Much of what we know about concept generation comes from the foundational
+Much of what we know about concept generation and contrast comes from the foundational
 literature on creative cognition. In a series of reports, Ward and colleagues
@@
 Generated categories tend to consist of features observed in known categories,
-and they tend to exhibit the distributional properties as found in known
+and they tend to exhibit the distributional properties found in known
 categories. In a seminal study, \cite{ward1994structured} asked participants to
@@
 
-\sxlnote{Added thomas98 description here} It is worth pointing out that the
+\sxlnote{Added thomas98 description here} \jlanote{still need the contrast result from her work.} It is worth pointing out that the
 observation of regularities in distributional properties across categories is
-not confined to the work of Ward and colleagues. Most notably, \citet{thomas98}
+not confined to generating whole new categories, but it also occures when inferring an unobserved feature within traditional categorization paradigms. Most notably, \citet{thomas98}
 found that when people were asked to generate feature values for exemplars with
@@
 the same distributional properties as the experimenter-defined categories. For
-example, after training on categories with a positive correlation between the
+example, after learning categories with a positive correlation between the
 size and saturation features (larger sized crystals were more saturated),
@@
 differs from the Path of Least Resistance account in that it notably omits the
-hierarchical organization of categories, as well as selectivity in which
+hierarchical organization of categories, as well as selectivity to which
 features are changed \citep[both of which are factors in the Path of Least
@@
 experiments devised by \citeauthor{jern2013probabilistic} were specifically
-designed to challenge it. However, its application is notable as a first step
+designed to challenge it. Its application is notable as a first step
 toward understanding theories developed in the creative cognition literature
@@
 something that is in some capacity {\em different} from what they already know.
-By consequence, we propose that contrast should be a primary constraint that
-guides how people create new categories: New categories should be different from
-existing ones.
-
-\jlanote{check to make sure the next paragraph has been updated enough}
-\sxlnote{I think it's updated enough! If given a chance I might edit this a
-  little to be less harsh on hierarchical bayesian models in general though I
-  think, since the representativeness model is one of them?}
+As a consequence, we propose that contrast should be a primary constraint that
+underlies people's expectations about how a set of categories are organized: Categories should be different from each other.
+
 Although it is evident that people are {\em capable} of creating new concepts
@@
 distributional structure. This is not an issue with their model specifically,
-but with a broader class of standard hierarchical Bayesian models \citep[e.g.,
+but using a broader class of standard hierarchical Bayesian models without any additional features \citep[e.g.,
 ][]{griffithsscn07,kemp07}. Many of these models assume that at some point of
@@
 location information and contrast cannot be captured using hierarchical Bayesian
-models. In fact, our model of contrast as representativeness is a hierarchical
-Bayesian model, but with a different goal than predicting the next category.
+models. In fact, a model instantiating contrast as representativeness is a hierarchical
+Bayesian model. But, it requires an additional factor, which in our case is having a different goal than predicting the next category.
 
@@
 
-\sxlnote{Charles suggested we either remove the section below, or move it to the
-  discussion. I've tried finding a good place for it in the discussion but it
-  doesn't fit very well, and moving it takes away this nice motivation for
-  our further investigation into contrast. I think that keeping it here in the
-  introduction might work out, provided Charles finds my argument for why the
-  regularity assumption is too strong convincing enough. More details a little
-  further down.}
-\subsection{Novel Analyses Demonstrating Contrast Effects in Prior Work}
-
+\subsection{Contrast as exemplar dissimilarity: The PACKER Model}
+
+\label{section:PACKER-definition}
+
+As noted above, the constraint that new concepts should differ from what is
+already known has been largely overlooked in previous work. This is no doubt in
+part due to the vague definition of what it means for a concept to be
+``different'': A generated category may be different from what is already known
+in any number of respects. Towards providing a more precise definition of the
+role of contrast in creative generation, we formalized contrast in a novel
+exemplar model, PACKER ({\em Producing Alike and Contrasting Knowledge using
+Exemplar Representations}). PACKER explains category generation as a balance
+between two fundamental constraints: The category to be generated should not be
+similar to known categories, and exemplars within each category should be
+similar to one another. These ideas are implemented within the well-studied
+exemplar framework -- the PACKER model is an extension of the influential
+Generalized Context Model of categorization
+\citep[GCM;][]{nosofsky1984choice,nosofsky1986attention}.
+
+Although, as an exemplar model, one of PACKER's proposals is people represent
+categories in terms of a collection of stored exemplars, it is worth emphasizing
+that the underlying nature of human category representations is not the primary
+concern of our work. The choice to develop PACKER within an exemplar framework
+reflects the facts that exemplar models have been throughly evaluated, are
+strongly theoretically motivated, and dominate much of the theoretical and
+empirical work in categorization. The focus of our work with PACKER concerns the
+dual constraints of within- and between-class similarity; it is not difficult to
+imagine how such constraints may be instantiated using alternative frameworks
+(e.g., \citealp{kurtz2007divergent,love2004sustain,smith2000thirty}; for a
+review of categorization models, see \citealp{pothoswills2011}).
+
+Both PACKER and the GCM simulate categorization under the assumption that
+learners represent categories as a collection of exemplars, corresponding to the
+labeled stimuli they have observed. The exemplars are encoded within a
+$k$-dimensional psychological space, and model performance is based on the
+amount of similarity between the item to be categorized and the stored
+exemplars. Similarity between two examples, $s\left(x_i, x_j\right)$, is
+computed as an inverse exponential function of distance
+\citep[following][]{attneave1950,shepard1957stimulus,shepard1987toward}:
+
+\begin{equation} s\left(x_i,x_j\right) = \exp \left\{ -c \left[\sum_{k}{ w_k
+\left| x_{ik} - x_{jk} \right|^r }\right]^{1/r} \right\}
+\label{eq:similarity}
+\end{equation}
+% 
+where $w_k$ is the attention weighting of dimension $k$ ($w_k \geq 0$ and
+$\sum_k{w_k} = 1$), accounting for the relative importance of each dimension in
+similarity calculations, and $c$ ($c>0$) is a specificity parameter controlling
+the spread of exemplar generalization. For simplicity, attention will be
+distributed uniformly in our simulations (unless otherwise noted). The value of
+$r$ depends on the nature of the experimental conditions being simulated: $r=1$
+is appropriate for separable dimensions, whereas $r=2$ is appropriate for
+integral dimensions \citep[e.g.,][]{shepard1964attention,garner1974processing}.
+In our simulations, we set $r=1$ due to the separable nature of the stimulus
+dimensions used in our experiments (see Figure \ref{fig:e1-conditions}).
+
+PACKER (as well as its name) was in part inspired by earlier work from the
+categorization literature \citep{hikada2011packing,stewart2005}. \citet{hidaka2011packing} argued that
+natural categories ``pack'' the values of features such that different
+categories fill the domain space with distance between one another, while
+keeping items within the same category close together. Inspired by this idea,
+PACKER proposes that generation is constrained by both similarity to members of
+the target category (the category in which a stimulus is being generated) as
+well as similarity to members of other categories: the most desirable generation
+candidates are similar to members of the target category and not similar to
+members of contrast categories. This is achieved by aggregating similarity
+across known exemplars differently according to class membership. The aggregated
+similarity $a(y,x)$ between generation candidate $y$ and stored exemplars $x$ is
+given by:
+
+\begin{equation} a(y, x) = \sum_j{f(x_j) s(y, x_j)}
+\end{equation}
+% 
+where $f(x_j)$ is a function specifying each exemplar's contribution to
+generation. A negative value for $f(x_j)$ produces a `repelling' effect (items
+are less likely to be generated nearby $x_j$), and a positive value produces an
+`attracting' effect (items are more likely to be generated nearby $x_j$). When
+$f(x_j)=0$, the exemplar does not contribute to generation.
+
+\jlanote{I'm a bit confused on the theta constraints.
+it's clear they need to be non-negative, but you don't usually see both needing to be between 0 and 1 but not sum to some value.
+is this really what we ended up with?
+Something feels fishy to me with the math.}
+PACKER sets $f(x_j)$ depending on exemplar $x_j$'s category membership:
+$f(x_j) = \theta_t$ if $x_j$ is a member of the target category, and
+$f(x_j) = -\theta_c$ if $x_j$ is a member of a contrast category.
+$\theta_t$ and $\theta_c$ are free parameters
+($0 \leq (\theta_t,\theta_c) \leq 1$) controlling the trade-off
+between within- and between-category similarity. For example, when
+$\theta_t = \theta_c = 0.5$, $f(x_j) = 0.5$ for members of the
+target category and $f(x_j) = -0.5$ for members of other categories; thus, the
+model is likely to generate items that are similar to members of the target
+category but are not similar to members of other categories. In this way,
+$\theta_t > 0$ with $\theta_c = 0$ produces exclusive consideration
+of target-category members, and $\theta_c > 0$ with $\theta_t = 0$
+produces exclusive consideration of contrast-category members. The combination
+of $\theta_t$ and $\theta_c$ parameters thus specifies a wide
+breadth of possible approaches; by fitting it to a dataset, one can describe the
+relative roles of between-category contrast and within-category similarity in
+generation. See Figure \ref{fig:packer-examples} for an illustration of how
+these parameters control the relative influence of within-category similarity
+and contrast to other categories when generating a new exemplar.
+
+\begin{figure}
+	\begin{center} \inputpgf{figs/}{packer-examples.pgf}
+      \caption{PACKER generation of a category `B' example, following exposure
+        to one member of category `A' and one member of category `B'.
+        Predictions are shown for three different parameterizations (differing
+        in values of $\theta_t$ and $\theta_c$): {\em (a)}
+        Predictions based on contrast similarity only. {\em (b)}
+        Predictions based on target similarity only. {\em (c)} Predictions with
+        both constraints considered.}
+		\label{fig:packer-examples}
+	\end{center}
+\end{figure}
+
+The probability that a candidate $y$ will be generated is evaluated using an
+Exponentiated \citet{luce1977choice} choice rule. Candidates with greater values
+of $a(y,x)$ are more likely to be generated than candidates with smaller values:
+% 
+\begin{equation} p(y \mid x) = \dfrac { \exp \left \{ a \left( y, x
+\right) \right \} } { \sum_i{ \exp \left \{ a \left( y_i, x \right)
+\right\} } }
+\label{eq:packer-choice}
+\end{equation}
+% 
+%where $\theta$ ($\theta \geq 0$) is a free parameter controlling response
+%determinism.
+
+It is worth noting that PACKER is only one possible exemplar-based account of
+category generation within our proposed framework. That is, PACKER places
+specific constraints on the possible values of $f(x_j)$, but other
+exemplar-based category generation models with drastically different behavior
+can be formalized in this framework by imposing alternative constraints. For
+example, as will be discussed in more detail below, PACKER is formally
+equivalent to the copy-and-tweak model proposed by \cite{jern2013probabilistic}
+when $\theta_c = 0$ and $\theta_t > 0$. Likewise, when
+$\theta_t = 0$ and $\theta_c > 0$, PACKER can represent a
+contrast-only generation mode, relying exclusively on contrast when generating
+new categories. When $f(x_j) < 0$ for all $x_j$ (regardless of class
+membership), a ``pure-packing'' approach is yielded, generating items in
+unoccupied areas of the domain. Thus, the proposed framework may be used to
+describe a wide variety of qualitatively distinct generation strategies.
+
+\subsubsection{Relation Between PACKER and Copy-And-Tweak}
+\label{section:copytweak-packer} The PACKER model is similar to the
+copy-and-tweak model reported by \cite{jern2013probabilistic}: Both models are
+exemplar-based, and both models generate new items according to their similarity
+to known members of the target class. However, PACKER diverges from the
+copy-and-tweak model by including a contrast mechanism, enabling generation
+according to dissimilarity to members of opposing categories. By consequence,
+copy-and-tweak can be realized as a parameterization of the PACKER model that is
+insensitive to category contrast. Specifically, when $\theta_c = 0$ and
+$\theta_t > 0$ (see Figure \ref{fig:packer-examples}, panel B),
+$f(x_j) = 0$ for $x_j$ belonging to contrast categories; thus, PACKER is not
+influenced by these items, and is mathematically equivalent to a copy-and-tweak
+approach.
+
+In this paper, we report simulations using this copy-and-tweak model. This model
+fits within the exemplar-based category generation framework defined above, and
+is a continuous-dimension adaptation of the model tested by
+\cite{jern2013probabilistic}. By formalizing a model family where PACKER and
+copy-and-tweak are different parameterizations of models within the same
+framework, the comparison between PACKER and copy-and-tweak provides a test of
+the explanatory value of the contrast mechanism based on exemplar dissimilarity:
+The account provided by copy-and-tweak will only equal that of PACKER if the
+contrast mechanism does not offer an advantage (i.e., if $\theta_c > 0$
+significantly improves model fits). Note that the purpose of the article is to
+explore and formally analyze the role of contrast in categorization and
+thus, we leave extending PACKER to incorporate distributional factors \citep[as
+explored by][]{jern2013probabilistic} for future work.
+
+\subsection{Contrast as representativeness: An extension to Jern and Kemp
+(2013)}
+\label{section:representativeness-definition}
+
+An alternate conceptualization of category contrast is the idea of
+representativeness -- exemplars are generated such that they resemble the target category and thus more distinct from and less
+similar to other categories. We adopt the formalization given by
+\cite{tenenbaum01}, where the representativeness $R$ of an item (or in our case,
+an exemplar) $y$ is the relative amount of evidence that is provided by $y$ for
+a given hypothesis $h$ in a space of hypotheses $\mathcal{H}$ in contrast to all other hypotheses $\mathcal{H}^c=\mathcal{H} \setminus \{h\}$:
+\begin{equation}
+  R(x,h) = \log \dfrac{p(x|h)}{1-p(x|\mathcal{H}^c)} = \log \dfrac{p(x|h)}{\sum_{h'\in H'}{p(x|h')p(h')}} . 
+\label{eq:representativeness}
+\end{equation}
+
+In Equation \ref{eq:representativeness}, $p(h')$ is the prior probability
+distribution on the hypothesis space that excludes $h$. Note that when there are
+only two hypotheses (e.g. two categories) involved in the domain, this prior
+takes the value of $1$ since there is ever only one alternative hypothesis.
+
+The Bayesian formalization of representativeness makes it straightforward to
+extend the existing hierarchical Bayesian model of category generation
+developed by \cite{jern2013probabilistic}.\footnote{We thank Charles Kemp for the suggestion.} Specifically, both models assume that
+exemplars are sampled from a given category distribution ($h$ in
+Equation \ref{eq:representativeness}). Each category distribution is a
+multivariate normal parameterized by a location vector $\mu_k$ and covariance
+matrix $\Sigma_k$ (i.e., $h = \mathcal{N}(\mu_k,\Sigma_k)$). The parameters
+$\mu_k$ and $\Sigma_k$ are assumed to be samples from a prior
+normal-inverse-Wishart (NIW) distribution parameterized by
+$\mu_0,\Sigma_0,\kappa,\nu$. Mathematically,
+\begin{align}
+  \mu_k, \Sigma_k|\mu_0,\Sigma_0,\kappa,\nu &\sim
+                                              \textrm{NIW}(\mu_0,\Sigma_0,\kappa,\nu)
+                                              \label{eq:mvnsample}\\
+  y|C &\sim \mathcal{N}(\mu_k,\Sigma_k) \label{eq:exemplarsample}
+\end{align}
+
+In our simulations, we set the prior mean $\mu_0$ to the center of the stimulus
+space and $\Sigma_0 = \lambda \mathbf{I}$ where $\lambda$ is a free parameter
+and $\mathbf{I}$ is a $d$-by-$d$ identity matrix with $d$ representing the
+number of dimensions or features in the domain. The $\kappa$ and $\nu$
+parameters are freely estimated within the constraints $\kappa > d-1$ and
+$\nu > 0$.
+
+With the assumption of the NIW prior, the expected location vector $\mu_k$ given exemplars observed from category $k$ is
+\begin{equation}
+  \mu_k = \dfrac
+    {\kappa\mu_{0} + n_k \bar{x}_k}
+    {\kappa + n_k}
+    \label{eq:category_mus}
+\end{equation}
+where $n_k$ is the number of observed exemplars in category $k$ and $\bar{x}_k$
+is the observed category mean. Note that if there are no observed exemplars in
+the target category (i.e., if the model is generating a completely novel
+category), Equation \ref{eq:category_mus} simplifies to $\mu_k = \mu_0$.
+
+The NIW prior also allows us to infer the category covariance matrix $\Sigma_k$
+by computing the following:
+\begin{equation}
+  \Sigma_k = [\nu \Sigma_D+ C_k +
+    \dfrac
+    {\kappa n_k}
+    {\kappa + n_k}
+    (\bar{x_k}-\mu_k)(\bar{x_k}-\mu_k)^T
+  ] (\nu + n_k)^{-1}
+  \label{eq:Sigma_y}
+\end{equation}
+where $C_k$ is the observed category covariance and $\Sigma_D$ is the domain
+covariance matrix from which $\Sigma_k$ samples are obtained. We can
+infer $\Sigma_D$ from the observed category covariances $C$ and the prior
+covariance $\Sigma_0$. Specifically,
+\begin{equation}
+    \Sigma_D = \Sigma_0 + \sum_{k}{C_k}
+\end{equation}
+
+At this point, both the representativeness model and the hierarchical Bayesian
+model from \cite{jern2013probabilistic} are largely identical. However, both
+models diverge in their computation of exemplar generation probabilities. While
+the original hierarchical Bayesian model produces novel exemplars with
+probabilities proportional to the multivariate normal likelihood $p(y|h)$, the
+representativeness model generates new exemplars with probabilities weighted by
+their representativeness $R(y,h)$. In practice, the probability of generating a
+particular candidate $y$ is obtained using an Exponentiated
+\citet{luce1977choice} choice rule:
+\begin{equation}
+  p(y) = \dfrac
+    {\exp( \theta \cdot R(y,h))}
+    {\sum_k \exp(\theta  \cdot R(y_k,h_k))} .
+\label{eq:representative-choice}
+\end{equation}
+where $\theta$ is a freely estimated response determinism parameter (constrained
+such that $\theta \geq 0$).
+
+Despite both the original hierarchical Bayesian model and our representativeness
+model sharing identical hierarchical structures, their distinct response
+processes can yield very different exemplars. Specifically, the hierarchical
+Bayesian model emphasizes the generation of exemplars that maintain
+distributional commonalities across categories. In contrast, the
+representativeness model focuses on the generation of exemplars that are
+representative of the underlying distribution for a target category. \jlanote{The next sentence is weird sounding and should be cleaned up} Generally,
+this representativeness mechanism results in the generation of exemplars that
+are less similar (i.e., more different) to contrast categories. We see this
+illustrated in Figure \ref{fig:representative-examples}, where the
+representativeness model displays a strong preference for generating exemplars
+from target category 'B' that are further away from the contrast category 'A'.
+
+\begin{figure}
+	\begin{center} \inputpgf{figs/}{hbm-examples.pgf}
+		\caption{Generation of a category `B' example, following exposure
+to one member of category `A' and one member of category `B'. Predictions are
+shown for {\em(a)} the original hierarchical Bayesian model and {\em(b)} the
+representativeness model.}
+		\label{fig:representative-examples}
+	\end{center}
+\end{figure}
+
+\subsection{Synopsis and Prognosis}
+
+Research on categorization and category generation has focused on the finding
+that generated categories tend to possess distributional commonalities with
+known categories. However, a fundamental task of categorization is to create
+something {\em new} (i.e., different from what is already known). The manner in
+which generated categories differ from known ones is, nonetheless, poorly
+understood: Existing theories do not make strong predictions about how
+creatively generated concepts should systematically differ from existing ones.
+Above, we introduced two
+novel approaches to formalizing the roles of similarity and contrast in creative
+generation.
+
+In the sections below, we present two experiments demonstrating systematic
+effects of category contrast on category generation inspired by factors
+influencing how the contrast models generate new categories. Our experiments are
+based on \cite{jern2013probabilistic}'s paradigm, which is a straightforward
+translation of the traditional artificial classification paradigm to the task of
+category generation: Participants are first exposed to a single,
+experimenter-defined category, and are then asked to generate members of a new
+category. We then report quantitative analyses of our results comparing the exemplar dissimilarity
+and the representativeness accounts of our results to that of the hierarchical
+Bayesian and copy-and-tweak models developed by \cite{jern2013probabilistic}.
+After analyzing the category generation studies, we present third experiment to
+demonstrate the effect of contrast in category learning. This ensures that our
+results are not idiosyncratic to category generation.
+
+\section{Experiment 1}
+To begin our investigation, we examined category generation in a well-understood domain using a few disparate category types. We used an
+artificial stimulus design: A two dimensional domain of squares, varying in
+color and size (see Figure \ref{fig:e1-conditions}, panel A). These dimensions
+have been used in numerous classification learning studies
+\cite[e.g.,][]{conaway2016similar,conaway2016generalization,shepard1961learning,nosofsky1994comparing}.
+Unlike those used in the \cite{jern2013probabilistic} experiments, distance on
+these physical dimensions aligns more directly with perceptual similarity,
+allowing us to evaluate the role of  contrast in categorization
+more precisely.
+It also enables more straightforward comparisons to prior work.
+We tested the effects of category contrast after learning one category from a set of qualitatively distinct category structures, as shown in Figure \ref{fig:e1-conditions}.
+
+Figures \ref{fig:e1-conditions}b-d show the values of exemplar dimensions
+belonging to the experimenter-defined categories (`A', or `Alpha') that
+participants were assigned to learn about prior to generating a new category.
+Each participant learned one of the category types during training. In the
+`Cluster' type, category A is a tight cluster of examples in the space.
+Perceptually instantiated, the members of category A might, for example, be
+large and dark in color. In the `Row' type, category A has a row pattern across
+the space, varying along one feature but not the other. Thus, its members might
+all be dark in color but would vary in size. Finally, in the `XOR' type, the
+experimenter-defined category consists of two clusters separated in opposite
+corners of the space, conforming to the exclusive-or logical structure (e.g.,
+members are small and dark or large and light).
+
+It should be noted that in our experiments the assignment of the perceptual
+to conceptual dimensions (e.g., $X \rightarrow Size$, $Y \rightarrow Color$) and the direction of variation along each dimension (e.g., $dark
+\rightarrow light$ or $light \rightarrow dark$) were counterbalanced across
+participants. The category types in Figure \ref{fig:e1-conditions} are plotted
+in conceptual space, rather than perceptual space. Thus, while the
+conceptual organization of the category types remains constant, each category
+type may have a different physical instantiation according to the counterbalance
+assignment. For example, the Cluster type may be large and dark in color, or it
+may be small and light in color, depending on the assignment and direction of
+the dimensions. For this reason, below we will discuss generation within a
+conceptual space, rather than a physically instantiated one.
+
+\begin{figure}
+    \begin{center} \inputpgf{figs/}{e1-conditions.pgf}
+    \caption{Stimulus domain and category types tested in Experiment 1. Stimuli
+are not drawn to scale. Dimension and direction assignment (e.g., large to small or small to large) for color and size were counterbalanced over participants.}
+    \label{fig:e1-conditions}
+    \end{center}
+\end{figure}
+
+After learning about an experimenter-defined category, participants are asked to
+generate examples of a new category. Within this paradigm, an effect of category
+contrast would be realized if participants prefer to generate items in locations
+that are distant (i.e., perceptually dissimilar) from members of category A.
+However, generation is left unconstrained. Critically, participants were not
+asked to generate something different in the prompt. For example, participants
+assigned to the Cluster condition may generate a tightly clustered category in
+the corner opposite of the experimenter-defined category. Alternatively, they
+may generate a tightly clustered category directly overlapping with the
+experimenter-defined category. Further, they may even generate an entirely
+different type of category (e.g., a row category).
+
+Our experimental results also provide a converging test
+that generated categories tend to share distributional properties with known
+categories in the domain \citep{jern2013probabilistic,ward1994structured}. From
+these results, we can predict that, in each condition, participants should
+generate categories that are distributionally similar to the
+experimenter-defined category: In the Cluster condition, generated categories
+should be tightly clustered. In the Row condition, generated categories should
+vary more along the X-axis than the Y-axis. In XOR condition, generated
+categories should be widely distributed across both dimensions, and the two
+dimensions should be positively correlated.
+
+Interestingly, the XOR condition also offers a dissociation between the roles of
+category contrast and the emulation of distributional structure:
+widely-distributed, positively-correlated categories would need to lie along the
+positive diagonal of the space (that is the only place they ``fit''), which is
+already occupied by the experimenter-defined category. Thus, if contrast plays a
+role, exemplars in the generated categories of participants in the XOR condition
+may not be positively correlated -- they may be negatively correlated instead.
+In this case, contrast and statistical regularities would interact, which would
+be inconsistent with prevailing theories of category generation
+\citep{jern2013probabilistic}
+
+\subsection{Participants and Materials}
+
+183 participants were recruited from Amazon Mechanical Turk. Each participant
+was randomly assigned to one condition: 64 participants were assigned to the
+Cluster condition, 61 were assigned to the Row condition, and 58 were assigned
+to the XOR condition (sample sizes differ due to random assignment). Stimuli
+were squares varying in color (grayscale 9.8\%--90.2\%) and side length
+(3.0--5.8cm), see Figure \ref{fig:e1-conditions}. The assignment of perceptual
+features (color, size) to axes of the conceptual space (x, y)  and the
+direction of variation along each axis (e.g., $dark \rightarrow light$ or $light
+\rightarrow dark$) were counterbalanced across participants.
+
+\subsection{Procedure}
+
+As noted in the introduction of this paper, the task of generating members of a
+new category is well situated as a task in the categorization
+literature: Whereas classification consists of predicting an object's category
+label on the basis of its features, inference consists of predicting an observed
+feature, given a set of observed features and a category label. Generation thus
+consists of predicting {\em all} features of an object, given a novel category
+label. We designed our generation task as an
+extension of the traditional artificial classification learning paradigm. The
+task differs from traditional work in creative cognition primarily through the
+use of an artificial domain, which enables the application of computational
+models. The use of an artificial domain also requires the addition of a training
+phase, during which participants learn about the categories in the domain. As a
+result, unlike most previous studies \citep[e.g.,][]{ward1994structured},
+participants in our studies have no experience with the domain before the start
+of the experiment, and the experimenter-defined categories are not
+hierarchically structured (as are many natural categories).
+
+Participants began the experiment with a short training phase (3 blocks of 4
+trials), where they observed exemplars belonging to the `Alpha' category.
+Participants were instructed to learn as much as they can about the Alpha
+category, and that they would answer a series of test questions afterwards. On
+each trial, a single Alpha category exemplar was presented, and participants
+were given as much time as they desired to observe it before moving on to the
+next trial. Each block consisted of a single presentation of each of the members
+of the Alpha category, in a random order. Participants were shown the range of
+possible colors and sizes prior to training.
+
+Following the training phase, participants were asked to generate four examples
+belonging to another category called `Beta'. As in
+\citet{jern2013probabilistic}, generation was completed using a sliding-scale
+interface. Two scales controlled the values of the two dimensions (color, size)
+for the generated example. An on-screen preview of the example updated whenever
+one of the features was changed. Participants could generate any example along
+an evenly-spaced 9x9 grid (including members of the Alpha category), except for
+any previously generated Beta exemplars. Neither the members of the Alpha
+category nor the previously generated Beta examples were visible during
+generation. Prior to beginning the generation phase, participants read the
+following instructions:
+
+\begin{displayquote} As it turns out, there is another category of geometric
+figures called ``Beta''. Instead of showing you examples of the Beta category,
+we would like to know what you think is likely to be in the Beta category.
+
+You will now be given the chance to create examples of any size or color in
+order to show what you expect about the Beta category. You will be asked to
+produce 4 Beta examples - they can be quite similar or quite different to each
+other, depending on what you think makes the most sense for the category.
+
+Each example needs to be unique, but the computer will let you know if you
+accidentally create a repeat.
+\end{displayquote}
+
+\subsection{Results}
+
+We observed a substantial degree of individual differences in our data. In
+Figure \ref{fig:e1-samples} we have plotted sample data from several
+participants, from which it is evident that different participants generated
+qualitatively different category structures. In this section we will focus on
+analyzing the data in aggregate, but in later sections we will explore how these
+individual differences can be explained.
+
+\begin{figure}
+    \begin{center} \inputpgf{figs/}{e1-samples.pgf}
+    \caption{Sample categories generated by participants in Experiment 1.
+Representative samples from common generation profiles are shown.}
+    \label{fig:e1-samples}
+    \end{center}
+\end{figure}
+
+To evaluate the role of contrast, we computed the number of times each stimulus
+was generated, as a function of its average city-block distance from members of
+the experimenter-defined ``Alpha'' category. These data, shown in Figure
+\ref{fig:e1-distanceplots}, reveal a clear pattern: Examples that are more
+distant from members of the experimenter-defined categories are more likely to
+be generated into a new category. This supports the notion that contrast is a
+fundamental constraint on how categories are related to one another and that statistical
+regularity alone is insufficient.
+
+Figure \ref{fig:e1-distanceplots} also depicts, for each participant, the
+average distance of members within the generated category ({\em within-category}
+distance) against the average distance between members of the generated and
+experimenter-defined category ({\em between-category} distance). The narrow
+distribution of between-category distances in the XOR condition reflects the
+widely distributed nature of the experimenter-defined category, reducing the
+possible distances to members of the participant-generated category. These data
+reveal a systematic pattern: The majority of participants generated categories
+with greater between-category distance than within-category distance. That is,
+members of the generated category tended to be more similar to one another than
+to members of the experimenter-defined category. To evaluate this claim
+quantitatively, we conducted t-tests comparing the amount of within- and
+between- class distance in each condition. All conditions possessed greater
+between-category distance: Cluster, $t(63) = 11.43$, $p < 0.001$; Row, $t(60) =
+13.16$, $p < 0.001$; and XOR, $t(57) = 3.64$, $p < 0.001$. These results provide
+further evidence of an effect of category contrast: Participants prefer to
+generate categories that are dissimilar to the learned category but maintain
+some level of internal cohesion.
+
+\begin{figure}
+    \begin{center} \inputpgf{figs/}{e1-distanceplots.pgf}
+    \caption{Experiment 1 results. {\em Left}: Frequency of generation as a
+function of distance from members of the experimenter-defined category. {\em
+Right}: Scatter plot of within-category versus between-category distance in each
+of the participant-generated categories.}
+    \label{fig:e1-distanceplots}
+    \end{center}
+\end{figure}
+
+A secondary goal of this experiment was to examine whether we replicate the
+classic result that generated categories often possess the same distributional
+properties as previously-known categories.
+Given the increased emphasis of replication within psychology \citep{zwaan2017}, it is important as it serves as a conceptual replication of \citet{jern2013probabilistic}.
+For each generated category, we
+computed the category range along each axis (X, Y), as well as the correlation
+between features. These data, shown in Figure \ref{fig:e1-statsboxes}, reveal
+broad individual differences: Within each condition, participants generated
+categories spanning the entire X- and Y- axis as well as categories that spanned
+very little along each. Likewise, in each condition participants generated
+categories possessing strongly positive, neutral, and strongly negative
+correlations between the dimensions. Comparing the distributional statistics
+between conditions yields a broad yet, as we will see, misleading replication of
+the classic effect.
+
+With respect to ranges along each axis (X, Y), the generated categories from
+each condition tend to reflect the ranges of the experimenter-defined
+categories. The categories generated in the Cluster condition were less widely
+distributed along the X-axis compared to Row, $t(123) = 5.61$, $p < 0.001$, and
+XOR, $t(120) = 2.68$, $p < 0.01$. Categories generated in the XOR condition
+were also less widely distributed along the X-axis compared to Row, $t(117) =
+2.56$, $p < 0.05$. This latter effect was not expected because the
+experimenter-defined categories for XOR and Row had similar X-ranges. However,
+the key finding is that categories from the Cluster condition tended to be more
+tightly distributed along the X-axis.
+
+\begin{figure}
+    \begin{center} \inputpgf{figs/}{e1-statsboxes.pgf}
+    \caption{Box-plots of the distributional statistics from the categories
+generated in Experiment 1. Boxes depict the median and quartiles of each
+condition, with whiskers placed at 1.5 IQR. All points outside this region are
+marked individually.}
+    \label{fig:e1-statsboxes}
+    \end{center}
+\end{figure}
+
+Likewise, categories generated in the Row condition had less Y-axis range
+compared to Cluster, $t(123) = 4.57$, $p < 0.001$ and XOR, $t(117) = 9.26$, $p <
+0.001$, and categories from the Cluster condition had less Y-axis range compared
+to XOR, $t(120) = 3.95$, $p < 0.001$. As expected, the correlations in the
+Cluster and Row conditions were not systematically positive or negative ($ps >
+0.1$). However, the generated categories in the XOR condition tended to possess
+{\em negatively} correlated dimensions, $t(57) = 2.04$, $p = 0.046$. This
+finding is notable, as it is the opposite of what would be expected based on previous literature \citep{jern2013probabilistic}, assuming
+learners are emulating the distributional structure of the experimenter-defined
+class (which possesses perfectly positively correlated features).
+
+We believe that the failure to replicate dimension correlation emulation is because participants in
+\cite{jern2013probabilistic} could differentiate the generated category on a
+third dimension (hue) to maintain the statistical regularities on the other two
+dimensions. Although the correlation in the XOR condition is significantly
+negative, it is clear from the box-plot in Figure~\ref{fig:e1-statsboxes} that
+it would be inappropriate to make a strong conclusion (e.g., the median is close
+to zero). However, we can conclude with confidence that there are situations
+where people do not emulate the distributional structure of the given category.
+This indicates that there is more to category generation than the emulation of
+distributional structure of other categories in the domain. Further, as we will
+discuss in more detail in the model-based analysis section, this is expected by
+our proposal that contrast is a fundamental principle in categorization.
+
+\subsection{Discussion} In Experiment 1 we sought to extend our analysis of the
+\cite{jern2013probabilistic} data by evaluating the influence of category
+contrast on category generation, given qualitatively different types of categories. We found strong evidence for effects of category contrast in each
+condition: Participants were more likely to generate stimuli that are more
+distant from (i.e., less similar to) members of a previously-learned category,
+and members of participant-generated categories tended to be more similar to one
+another than to members of previously-learned categories. We also partially
+replicated the classic finding that the distributional structure of generated
+categories reflects that of previously learned categories
+\citep{jern2013probabilistic,ward1994structured}: Members of generated
+categories were more widely distributed along dimensions which were widely
+distributed in the experimenter-defined category.
+
+Notably, however, we also found that participants who learned an XOR category
+(composed of exemplars following a positive diagonal, see Figure
+\ref{fig:e1-conditions}) tended to generate items according to a {\em negative}
+feature correlation -- the opposite of what was present in the previously
+learned category. While this may be difficult to account for under existing
+theoretical approaches (which assume generated categories follow the same
+distributional structure as known categories), it can be concisely explained
+from a category contrast perspective. Specifically, within the XOR condition,
+individuals who seek to generate a category that is perceptually distinct from
+what is already known are left with only the upper-left and bottom-right
+quadrants of the space, as members of the previously-learned XOR category lie in
+the bottom-left and top-right. If examples are generated into both of the
+available quadrants, the generated category will possess a strongly negative
+correlation, opposing that of the experimenter-defined class.
+
+Thus, the core results of Experiment 1 indicate that generated categories can
+systematically differ from what we would expect based on prior work. The
+negative (or null) correlations observed in the XOR condition suggests an
+interesting interaction between contrast with a given category and emulation of
+statistical properties. That is, the constraints on creative generation imposed
+by category contrast may not simply influence the {\em location} of generated
+categories, but also their distributional structure. In Experiment 2, we test
+this claim more systematically.
+
+\section{Experiment 2.}
+
+\begin{figure}
+    \begin{center} \input{figs/e2-conditions.pgf}
+    \caption{Category types tested in Experiment 2.}
+    \label{fig:e2-conditions}
+    \end{center}
+\end{figure}
+
+To test whether category contrast influences the distributional structure of
+generated categories, we sought to identify conditions in which differences in
+the distributional structure of generated categories cannot be explained by the
+distributional structure of the experimenter-defined category. We created two
+new category types (depicted in Figure \ref{fig:e2-conditions}) that possess
+identical distributional structures (both are tight clusters of examples with no
+correlation between features), as they only differ in their Y-axis position: the
+`Bottom' category lies near the bottom of the space, and the `Middle'
+category lies in the center. The distributional equality of these conditions is
+key to the design of the experiment: If the distributional structure of
+previously learned categories were the only influence on the generated
+categories, we should observe no difference in the categories participants
+generate between these two conditions. Will participants distribute their
+generated category differently between conditions due to the differences in the
+available stimulus space for generating a new category?
+
+If category contrast influences the distributional structure of the categories
+people generate, then we should observe different types of categories according
+to the shape of the space that is {\em unoccupied} by members of previously
+learned categories. The difference in the Y-axis position between the Bottom and
+Middle conditions produces a considerable change to the shape of the unoccupied
+space. Participants assigned to learn the Bottom category should be less likely
+to generate exemplars into the lower regions of the stimulus space (as these
+areas possess greater similarity to members of the Bottom category), preferring
+instead to distribute exemplars across the upper region of the space. This
+constraint is lifted in the Middle condition, as the Middle category exemplars
+are equidistant to the upper and lower regions of the space. Accordingly,
+participants should be more likely to utilize both of these areas. Thus, if
+category contrast influences the distributional structure of generated
+categories, we should observe more participants in the Middle condition that
+generate examples above {\em and} below the experimenter-defined category.
+
+\subsection{Participants, Materials, and Procedure}
+
+122 participants were recruited from Amazon Mechanical Turk. 61 participants
+were randomly assigned to the Middle and Bottom conditions each. The stimulus
+space and procedure were exactly as in Experiment 1. Participants first
+completed a short training phase, followed by the generation phase. The only
+difference from Experiment 1 was the category types given to participants.
+
+\subsection{Results}
+
+\begin{figure}
+    \begin{center} \inputpgf{figs/}{e2-samples.pgf}
+    \caption{Sample categories generated in Experiment 2. }
+    \label{fig:e2-samples}
+    \end{center}
+\end{figure}
+
+As in Experiment 1, we observed broad differences in the generation approach
+taken by different participants. To characterize the nature of these
+differences, Figure \ref{fig:e2-samples} depicts sample categories generated by
+participants. The data from each condition are organized into four columns based
+on commonly observed patterns of generation: a `Cluster' type of
+tightly-clustered examples, `Row' and `Column' types of exemplars widely
+distributed along the one axis but narrowly along the other, and a `Corners'
+type, wherein participants placed exemplars in disparate corners of the space.
+As before, in this section we focus on analyzing the data in aggregate, but in
+later sections we will focus more specifically on explaining the individual
+differences.
+
+\begin{figure}
+    \begin{center} \inputpgf{figs/}{e2-distanceplots.pgf}
+    \caption{Experiment 2 results. {\em Left}: Frequency of generation as a
+function of distance from members of the experimenter-defined category. {\em
+Right}: Scatter plot of within-category versus between-category distance in each
+of the participant-generated categories.}
+    \label{fig:e2-distanceplots}
+    \end{center}
+\end{figure}
+
+We began our analysis by testing for the broad influence of category contrast on
+generation. As in Experiment 1, we computed the frequency each stimulus was
+generated as a function of its average distance from members of the
+experimenter-defined category, as well as each participant's average within- and
+between- category distance. These data, shown in Figure
+\ref{fig:e2-distanceplots}, yield very similar results. Participants generated
+stimuli that are distant from members of the experimenter-defined category, and
+the categories in each condition tended to possess more between-category than
+within-category distance: Bottom, $t(60) = 5.5$, $p < 0.001$; Middle, $t(60) =
+2.71$, $p < 0.01$. We did, however, observe a notable subgroup of participants
+in each condition who generated categories with more within-category than
+between-category distance. Upon manual inspection, many of these individuals
+appear to have assumed a `Corners' strategy, placing exemplars in disparate
+corners of the space, thus producing much more within-category distance, see
+Figure \ref{fig:e2-samples} for examples.
+
+To explore the distributional structure of the generated categories, we computed
+the range of exemplars along each axis (X, Y), as well as the correlation
+between features. These data, shown in Figure \ref{fig:e2-statsboxes}, again
+demonstrate the degree of individual differences observed in our study. In each
+condition, we observed tightly clustered and widely distributed categories along
+each dimension, as well as positively, negatively, and uncorrelated categories.
+
+\begin{figure}
+    \begin{center} \inputpgf{figs/}{e2-statsboxes.pgf}
+    \caption{Box-plots of the distributional statistics from the categories
+generated in Experiment 2. }
+    \label{fig:e2-statsboxes}
+    \end{center}
+\end{figure}
+
+As noted above, if the distributional structure of generated categories is
+influenced by the shape of the space not occupied by members of known
+categories, then participants in the Middle condition would be more likely to
+place exemplars in the upper {\em and} lower regions of the space, as members of
+the experimenter-defined category are equidistant from these regions.
+Participants in the Bottom condition should be less likely to generate category
+members in the bottom regions because members of the experimenter-defined
+category are located there. One way to test these predictions is to analyze the
+Y-axis ranges of the generated categories: If Middle participants utilize the
+upper and lower regions of the space, their categories should vary more along
+the Y-axis. T-Tests comparing the conditions on the distributional statistics,
+however, reveal few between-group differences: the conditions do not differ with
+respect to X-axis range, Y-axis range, or feature correlations ($ps>0.17$).
+
+However, our ability to detect differences in Y-Axis range using a standard
+$t$-test between the conditions is, in this case, diminished due to the
+non-normality of the data (Shapiro-Wilk normality test $W=0.77, p< 0.001$ for
+the Middle condition and $W=0.85, p < 0.001$ for the Bottom condition). Figure
+\ref{fig:e2-yranges} depicts the Y-axis position of the exemplars generated
+within each participant's category.
+Further, the median produced Y-Axis range is much smaller in the Bottom than Middle condition, whereas they are essentially identical in X-Axis range.
+The categories are sorted by overall range,
+then by condition assignment. These data reveal that there were nearly as many
+participants who generated categories spanning the entire Y-axis as those who
+generated categories spanning almost none of the Y-axis. The non-normality of
+the Y-axis range distributions thus requires that we use a different approach to
+addressing the experiment's main question.
+
+\begin{figure}
+    \begin{center} \inputpgf{figs/}{e2-yranges.pgf}
+    \caption{Y-Axis range and position of the participant-generated categories
+from Experiment 2. Each line corresponds to a participant's category, with
+notches corresponding to the Y-axis position of exemplars within the category
+(notches may overlap). Participants are sorted by overall range, and then by
+condition. }
+    \label{fig:e2-yranges}
+    \end{center}
+\end{figure}
+
+Because our main prediction concerns the generation of exemplars within the
+upper and lower regions of the domain, we compared the conditions in terms of
+the frequency with which participants generated examples above and below the
+categories. Specifically, we counted the number of participants in each
+condition who placed at least one `Beta' exemplar on the top and bottom `rows'
+of the space (the maximum and minimum possible Y-axis value, respectively). The
+resulting contingencies data are shown in Table \ref{table:e2-subset-table}.
+
+Firstly, it should be noted that nearly every participant utilized the top
+and/or bottom rows: only $10 / 122$ participants generated their category
+entirely within the interior region. Fisher's Exact Tests comparing the
+conditions reveal that more Middle participants generated an exemplar in the
+bottom row, $p < 0.001$, again demonstrating the role of contrast in guiding
+where exemplars are generated. The conditions did not differ in use of the top
+of the space, $p = 0.16$, however, more Middle participants placed exemplars in
+the top {\em and} bottom rows, $p < 0.05$. The latter effect is of particular interest
+here, as it indicates that the shape of the unoccupied space exerts some
+influence on the distributional structure of generated categories: Participants
+in the Middle condition were more likely to generate a category spanning the
+entire Y-axis. Thus, distributional structure of the generated categories can be
+influenced by category contrast alone.
+
+\begin{table}
+\begin{center}
+\caption{Experiment 2 results.}
+\label{table:e2-subset-table} \vskip 0.12in
+\begin{tabular}{ l r r} \textbf{Middle} & Used top row & No top row \\ \hline
+Used bottom row & 28 & 18 \\ No bottom row & 11 & 4 \\ \\ \textbf{Bottom} & Used
+top row & No top row \\ \hline Used bottom row & 16 & 8 \\ No bottom row & 31 &
+6 \\
+\end{tabular}
+\end{center}
+\end{table}
+
+\subsection{Discussion}
+
+In Experiment 2, we replicated the core findings from Experiment 1. Stimuli are
+more likely to be generated if they are distant from exemplars in other
+categories, and most participants generate categories with more between-category
+than within-category distance. However, we additionally found that the {\em
+position} of a previously learned category (rather than its distributional
+structure) influences the types of categories people generate: Participants who
+learned the `Middle' type were more likely to generate categories spanning the
+entire Y-axis of the space. Participants who learned the `Bottom' type were less
+likely to do so as a result of the presence of opposite category exemplars in
+the lower regions of the space.
+
+This finding cannot be explained from the perspective that the distributional
+structure of previously learned categories is the sole determinant of the
+distributional structure of generated categories. However, the observed behavior
+is expected from a category contrast perspective: Participants seeking to
+generate a perceptually distinct category will be more likely to use areas of
+space that are unoccupied by exemplars belonging to previously learned
+categories. In the Middle condition, the upper and lower regions of space are
+equidistant from members of the experimenter-defined category, whereas in the
+Bottom condition, the lower region of the space is closer to members of the
+experimenter-defined category. Thus, while Middle participants may form
+categories around the use of the equally unoccupied areas, the same is not true
+for the Bottom condition.
+
+\section{Model-based Analyses of Experiments 1 and 2}
+\sxlnote{For this section I've added the existence of the representativeness
+  model but I think discussion of its performance is still lacking compared to PACKER.}
+Experiments 1 and 2 revealed systematic and strong effects of category contrast
+on creative generation. In this section, we analyze the performance of the different formal models at explaining the experimental results. Specifically, we
+present simulations from the two novel contrast models: PACKER and the
+representativeness model, and compare them to two models that do not incorporate
+contrast: the copy-and-tweak model (discussed in Section
+\ref{section:copytweak-packer}) and an implementation of the hierarchical
+Bayesian model proposed by \cite{jern2013probabilistic}, described in-depth in
+Appendix \ref{ap:hsampling-definition}. For our simulations here, the
+copy-and-tweak model is defined as a variant of PACKER with the $\theta_c$
+parameter constrained to be zero. The comparison of this set of models serves to
+highlight the explanatory role of contrast in categorization: If contrast
+affords little explanatory advantage, then there should be few differences in
+performance between PACKER and copy-and-tweak, or between the representativeness
+and hierarchical Bayesian model. The comparison between these models can also
+emphasize the necessity of contrast and demonstrate that generation cannot be
+explained entirely through the emulation of distributional structure. Each model
+has complementary strengths and weaknesses: Whereas PACKER and copy-and-tweak
+are insensitive to the distributional structure of learned categories (relying
+only on exemplar similarities), the representativeness and hierarchical Bayesian
+model generates categories exclusively on the basis of knowledge of how existing
+classes are distributed.
+
+Our approach in this section is to first broadly evaluate and compare the
+quality of each model's account to our entire dataset (Experiments 1 and 2
+combined), then analyze the ability for each model to explain individual
+differences in each experiment, and lastly we describe the strengths and
+weakness of each model's account of category generation.
+
+\subsection{Parameter-Fitting}
+
+To obtain a global measure of the quality of each model's account, we fitted the
+parameters of each model to our entire dataset (Experiments 1 and 2 combined),
+using a hill-climbing algorithm which maximized the log-likelihood of the
+model's predictions of the observed responses (1220 responses from 305 total
+participants). We fitted three parameters in the PACKER model ($c$, $\theta_t$,
+and $\theta_c$; see Section \ref{section:PACKER-definition}), as well as four in
+the representativeness model and the hierarchical Bayesian model ($\kappa$,
+$\lambda$, $\nu$, and $\theta$; see Section
+\ref{section:representativeness-definition} and Appendix
+\ref{ap:hsampling-definition} respectively). We fitted only two parameters for
+the copy-and-tweak model ($c$, and $\theta_t$), as $\theta_c$ is held constant
+($\theta_c = 0$). Attention ($w$, see Equation \ref{eq:similarity}) in PACKER
+and copy-and-tweak was set uniformly. Parameters were not allowed to vary
+between participants or conditions -- the goal was to obtain the best-fitting
+values to our entire dataset.
+
+\begin{table} \centering
+\caption{Results of model-fitting to the combined datasets from Experiments 1
+and 2. Note that smaller AIC values correspond to better model fits (adjusted
+for number of parameters)}
+\label{table:global-model-fits}
+\begin{tabular}{ l l l l} \\ \textbf{PACKER} & \textbf{Copy \& Tweak} &
+\textbf{Representativeness} & \textbf{Hierarchical Bayesian} \\ \hline $AIC =
+  9095$ & $AIC = 9842$ & $AIC = 8878$ &  $AIC =
+9912$ \\ $L = -4545$ & $L = -4919$ & $L = -4435$ & $L = -4952$ \\ $c = 0.483$ &
+                                                                                $c = 3.197$ & $\kappa = 11.232$ &
+$\kappa < 0.001$ \\ $\theta_c = 3.16$ & $\theta_c = 0$ (fixed) & $\nu = 2.844$ & $\nu = 5.597$ \\
+$\theta_t = 3.494$ & $\theta_t = 2.972$ & $\lambda = 1.917$& $\lambda
+                                                                       = 0.055$
+  \\ & &$\theta = 5.885 $& $\theta = 3.175$
+\\
+\end{tabular}
+\end{table}
+
+Table \ref{table:global-model-fits} contains the model fits. Due to the uneven number of fitted parameters among the models, we
+compare the model fits using the Akaike Information Criterion
+\citep[AIC;][]{akaike1974new}, where smaller values correspond to better fits
+(discounted by model complexity as measured by the number of parameters). The
+same qualitative results were obtained with alternative model comparison metrics
+(e.g., BIC, \citealp{schwarz1978estimating}; $AIC_C$,
+\citealp{hurvich1989regression}). In addition to AIC, table \ref{table:global-model-fits} contains the corresponding log-likelihood ($L$) and the best-fitting
+parameter values. These results reveal strong model differentiation: both
+contrast models (PACKER and the representativeness model) achieved far better
+fits compared to their non-contrast counterparts: copy-and-tweak and the
+hierarchical Bayesian model respectively. Interestingly between the contrast
+models, the hierarchical model (representativeness) outperformed the
+exemplar-based theory (PACKER), whereas between the non-contrast models, the
+reverse is observed. Specifically, here the exemplar-based model
+(copy-and-tweak) performed somewhat better than the hierarchical Bayesian model.
+
+\sxlnote{The argument in the following two paragraphs appear to be challenged by
+  the success of the representativeness model. I might suggest removing them altogether?}
+While PACKER's advantage may tentatively be attributed to the model's
+sensitivity to category contrast (this will be explored in detail below), the
+advantage shown by copy-and-tweak over the hierarchical Bayesian model may be
+attributed to its exemplar-based representation of category B, as opposed to forcing a
+prototype-based representation as assumed by the hierarchical Bayesian model. As
+observed in Figures \ref{fig:e1-samples} and \ref{fig:e2-samples}, the generated
+categories we observed were often widely distributed, with no items near the
+category prototype. This aspect of the data is inconsistent with the
+multivariate normal distributions (similar to prototypes) used to represent
+categories in the \cite{jern2013probabilistic} model, but can be easily
+accounted for using an exemplar-based approach.
+Interestingly, representativeness using a prototype approach fits better than an exemplar-based approach.
+Understanding these differences more thoroughly is left for future research.
+
+A key distinction between PACKER and the other three models is that only PACKER
+is capable of making strong predictions about the location of new category
+members when the target class is entirely novel (i.e., no member of the category
+has been observed). Under these circumstances, there are no examples to copy,
+and thus the copy-and-tweak model predicts that items are generated at random.
+Likewise, with no observations on which to condition the category distribution,
+
+\jlanote{Do they really pick the first item at random? That doesn't seem right for the representativeness -- shouldn't it be constrained?}
+the hierarchical Bayesian and representativeness models also picks an item at
+random.
+
+Crucially, however, the next generated exemplar from the
+representativeness model is strongly influenced by contrast as
+representativeness. Thus, it is possible that the failure of these models is
+simply due to their inability to explain each participant's first trial
+(generating the first item in the `Beta' category). To ensure this is not
+driving our results, we conducted an identical set of simulations as above,
+excluding this trial (leaving 915 responses in the dataset): Again, PACKER
+($L = -3390$, $AIC = 6786$) achieved better fits than the copy-and-tweak
+($L = -3579$, $AIC = 7162$) and hierarchical Bayesian ($L=-3612$, $AIC = 7232$)
+models.
+\jlanote{last sentence missing representativeness fit}
+
+Finally, because copy-and-tweak is nested within PACKER, we can use a likelihood
+ratio test to compare the two models. PACKER explains the aggregate data
+significantly better than copy-and-tweak ($\chi^2(1) = 749, p < 0.001$ for all
+data and $\chi^2(1) = 377, p < 0.001$ excluding the first example), providing
+further evidence that category generation is better explained when contrast is
+considered.
+
+Through comparison with the copy-and-tweak model, Figure
+\ref{fig:packer-loglike} more clearly demonstrates the robustness of the
+explanatory gains yielded by PACKER's category contrast mechanism. It displays
+the log-likelihood of the participants' results under PACKER as a function of
+the $\theta_c$ parameter. The model's other parameters ($c$,
+$\theta_t$) were set according to copy-and-tweak's best fits from Table
+\ref{table:global-model-fits}, and thus when $\theta_c=0$, the models are
+equivalent. The figure clearly shows a ``sweet spot'': a convex region in which
+PACKER achieves superior fits as a result of changes to $\theta_c$. The
+best fitting values lie well above the value of 0 assumed by the copy-and-tweak
+model, which demonstrates the robustness of the contrast effect (though note
+PACKER achieves even better fits when its parameters are fitted together, as in
+Table \ref{table:global-model-fits}). In sum, the data are better explained when
+both within-category similarity and category contrast is considered.
+
+\begin{figure}
+    \begin{center} \inputpgf{figs/}{packer-loglike.pgf}
+    	\jlanote{fix x-axis label}
+      \caption{PACKER's fit as a function of its prioritization of
+        within-category and between-category similarity (using the
+        $\theta_c$ parameter) . To facilitate comparison, PACKER's other
+        parameters ($c$, $\theta_t$) were set to the best fitting values
+        obtained for copy-and-tweak in Table \ref{table:global-model-fits}. }
+    \label{fig:packer-loglike}
+    \end{center}
+\end{figure}
+
+\subsection{Individual Differences}
+\label{section:individual-diffs}
+
+As noted in Experiments 1 and 2, we observed a great deal of individual
+differences in the types of categories that participants generated. Within each
+condition, there were a wide variety of category types, such as row and column
+categories (see Figures \ref{fig:e1-samples} and \ref{fig:e2-samples}). The
+simulations reported above serve to evaluate the models while considering the
+entire dataset, but a secondary goal of any formal account should be to provide
+some explanation of how different profiles of performance emerge. Many of the
+individual generation profiles we observed can be described within the PACKER
+framework, simply by tuning the model's parameters in a principled manner. In
+this section, we describe more specifically how the most frequently observed
+profiles can be realized.
+
+By manual inspection, it is evident that the most common profiles of generation
+consist of: (A) a tightly-distributed `cluster' of examples, (B) `row'- and
+`column'-like arrangements (varying widely along one dimension but not the
+other), and (C) a `corners' arrangement with examples placed into disparate
+corners of the space. These four profiles are distinct in terms of the
+distribution of the generated category along each dimension: Whereas the cluster
+profile is tightly distributed along both dimensions, the row and column
+profiles are tightly distributed along just one dimension. Finally, the corners
+profile is widely distributed along both dimensions.
+
+In the framework proposed by PACKER, the cluster and corners profiles arise
+based on different prioritization of within-category similarity versus
+between-category contrast, and the row and column profiles arise based on the
+prioritization of each dimension in the computation of similarity. For example,
+in the cluster profile, there is a high degree of within-category similarity
+along both dimensions, whereas in the corners profile there is minimal
+within-category similarity. Thus, PACKER's proposal is that these individual
+differences arise as a result of different priorities: While the tight cluster
+configuration can be considered PACKER's `default' mode (as it maximizes
+within-category similarity), the corners profile can be produced when
+between-category contrast is put at a higher priority (i.e., $\theta_c > \theta_t$).
+
+Likewise, in the row and column profiles, there is a large degree of
+within-category similarity along one dimension but not the other. These
+differences likely arise due to a differential focus on one dimension over
+another, and thus they can be produced by changes to PACKER's attention weights,
+$w_1$ and $w_2$ (see Equation \ref{eq:similarity}). Traditionally, the attention
+weights in exemplar models are thought to reflect the diagnostic value of each
+dimension towards classifying the known category members
+\citep{nosofsky1984choice,nosofsky1986attention,kruschke1992alcove}, but within
+a generation context the weights specify the importance of within- and
+between-category similarity along each dimension. For example, if all of
+attention is allocated along the X-axis ($w_1=1$ and $w_2=0$), similarity along
+the Y-axis no longer influences performance. As a result, PACKER will create
+categories that are more widely distributed along the Y-axis, as similarity is
+not taken into account along that dimension. As a general principle,
+differentially weighting one dimension will result in the generation of
+categories that are more widely distributed along the ignored dimension,
+conforming to a row- or column-like arrangement. See Figure
+\ref{fig:packer-attention} for a depiction of how attention influences PACKER's
+performance.
+
+\begin{figure}
+    \begin{center} \inputpgf{figs/}{packer-attention-examples.pgf}
+    \caption{PACKER generation of a category `B' example, following exposure to
+one member of category `A' and one member of category `B'. Predictions are shown
+for different attention settings: {\em (a)} Increased weighting of the X-axis.
+{\em (b)} Increased weighting of the Y-axis. {\em (c)} Uniform weighting
+(identical to Figure \ref{fig:packer-examples}).}
+    \label{fig:packer-attention}
+    \end{center}
+\end{figure}
+
+As in PACKER, changes in the parameter settings of the copy-and-tweak model can
+also be used to produce different patterns of generation. Indeed, as
+copy-and-tweak is simply a special case of the PACKER model, the attention
+weights operate exactly as described above to produce row- and column-like
+categories. However, because the model is not influenced by category contrast,
+it is biased toward generating tightly clustered categories, as new items are
+always most likely to be generated near known examples of the target category.
+Thus, the lack of a contrast mechanism prevents the model from explaining why
+some individuals widely distribute their categories to the corners of the space.
+
+\jlanote{Xian: Please add a paragraph or two on the representativeness model here}
+
+Similar to the copy-and-tweak model, the hierarchical Bayesian model possesses
+no mechanism to account for category contrast, and the model is most likely to
+generate new items that are similar to known examples of the target category.
+Although this precludes an account of why individuals might assume a cluster
+versus corners profile, the covariance matrix specifying the model's prior
+domain distribution, $\Sigma_0$, can be used to explain the generation of
+row-like and column-like categories. This covariance matrix specifies the amount
+of variance assumed along each dimension (as well as the dimensional
+correlations) across the domain of categories. The covariance matrix for a newly
+generated category, $\Sigma_B$, is based on the assumed $\Sigma_0$ as well as
+the distributions of previously learned categories (see Appendix
+\ref{ap:hsampling-definition}). Thus, the importance of each feature can be
+coded into $\Sigma_0$ to alter the dimensional variance of generated categories.
+
+As noted above, while the copy-and-tweak and hierarchical Bayesian models
+possess mechanisms to explain row- and column-like categories, they cannot
+easily explain why some individuals widely distribute their generated categories
+into disparate corners of the space. This, however, reveals a more general
+limitation: According to the copy-and-tweak and hierarchical Bayesian models,
+the distributional structure of generated categories is {\em independent} of
+their location within the domain. For example, although the copy-and-tweak or
+hierarchical Bayesian models can be parameterized to generate row- or
+column-like categories, there is no mechanism in place to ensure that what is
+generated will be distinct from what is already known. In the next subsection,
+we explore this prediction through an analysis of the interdependence between
+distributional structure and location in creative generation.
+
+\subsection{Category Location vs. Distributional Structure}
+\label{section:individual-diff}
+As noted above, while all three models make clear claims about the internal
+structure of generated categories, the copy-and-tweak and hierarchical Bayesian
+models do not make any claims about how generated categories should differ from
+what is already known. However, as we observed in the results of Experiment 2,
+the distributional structure of a category is not always independent of its
+location within the domain. To demonstrate this point in more depth, we computed
+the X- and Y- axis ranges of every participant-generated category. Taking the
+difference between these values ($X-Y$) produces a measure of each category's
+orientation in the space: positive difference scores correspond to categories
+with more X-axis range (horizontally aligned, `Row' categories), whereas
+negative difference scores indicate the opposite (vertically aligned, `Column'
+categories). Neutral differences scores indicate there was an equal amount of X-
+and Y-axis range, which can be produced by a number of different category types
+(`Clusters', `Corners', etc; see Figures \ref{fig:e1-samples} and
+\ref{fig:e2-samples}). By plotting, for each possible stimulus, the difference
+scores of categories it was generated within, we can relate the distributional
+structure of generated categories to their location within the domain.
+
+However, because many stimuli were infrequently generated (such items near
+members of the `Alpha' category), we cannot simply compute the empirical average
+of the difference scores, as infrequently generated stimuli would be likely to
+show artificially strong differences. Instead, we used a Bayesian analysis to
+estimate the mean $\mu_x$ on the assumption that the scores $x$ for each
+stimulus are normally distributed with an unknown mean and unknown standard
+deviation. The conjugate Normal-Inverse Gamma distribution provides a
+straightforward method for this estimation:
+
+\begin{equation} \mu_x = \dfrac { \nu_0 \mu_0 + \sum{x} } { \nu_0 + n }
+\label{eq:rangediff-bayes}
+\end{equation} \ where $\mu_0$ is the prior mean, $\nu_0$ is a prior scale
+parameter (controlling the weighting of the $\mu_0$), and $n$ is the number of
+categories in which the stimulus was a member (i.e., the number of scores in
+$x$). The default assumption is that there is an equal amount of range along the
+X- and Y-axes, and so we set $\mu_0 = 0$. Likewise, to give a moderate amount of
+weighting to the prior mean we set $\nu_0 = 1$, though the results are robust to
+a range of values. Within this approach, the resulting aggregation is a
+trade-off between the number of generations and the strength of the range
+difference within each generated category. Infrequently generated stimuli, as
+well as those with mixed positive and negative scores, are given neutral
+difference scores.
+
+\begin{figure}[p]
+    \begin{center} 
+      \includegraphics[width=\textwidth]{figs/range-diff-gradients.pdf}
+      \caption{Behavioral and simulated range difference gradients. Each panel
+        shows, for each stimulus, the dimensional orientation of the categories
+        it was generated into: vertically aligned `columns' (orange) versus
+        horizontally aligned `rows' (purple).}
+      \label{fig:range-diff-gradients}
+    \end{center}
+\end{figure}
+
+The results of our analysis are shown in Figure \ref{fig:range-diff-gradients}
+for the experiment and model results\footnote{Prior to plotting, data were also
+processed using a Gaussian filter with $\sigma = 0.8$.}. The left-most column of
+Figure \ref{fig:range-diff-gradients} displays the effect of category location
+and contrast on the distributional structure of the category generated by
+participants. These data reveal strong and consistent patterns across all the
+conditions we tested in Experiments 1 and 2: Generated categories are more
+tightly distributed along the axis in which they are distinct. For example, in
+the `Cluster' condition, exemplars in the bottom-left of the space are more
+often generated into vertically aligned categories, and exemplars in the
+top-right are more often generated into horizontally aligned categories.
+Similarly, in the `Bottom' and `Middle' conditions, horizontally aligned
+categories are generated above and below the experimenter-defined categories,
+while vertically-aligned categories are generated to the sides. In the `Row'
+condition, most categories are horizontally aligned, and lie along the upper
+areas of the space. There are no strong range difference patterns in the XOR
+condition.
+
+These patterns of performance clearly depict the interdependence between the
+distributional structure and location of generated concepts. Our results can be
+interpreted in terms of local minimization of between-category similarity: By
+distributing the generated category away from members of the
+experimenter-defined category, participants may increase the degree of
+between-category distance without drastically altering the degree of
+within-category similarity.
+
+To explore how well the PACKER, copy-and-tweak, representativeness, and
+hierarchical Bayesian models explain our findings, we conducted simulations
+using an individual-differences approach. As noted in Section
+\ref{section:individual-diffs}, row- and column-like categories can be produced
+by each model through changes to the weighting of each dimension. Given this
+information, we may use the models to simulate each participant's generation
+separately, with the importance of each dimension set according to the relative
+range of the participant's generated category along each dimension.
+
+In the PACKER and copy-and-tweak models, the attention weights, $w$, specify the
+importance of each dimension in the computation of similarity. While there exist
+methods to find the optimal attention weighting scheme given a classification
+\citep[see][]{vanpaemel2012using}, for simplicity we assume that the Alpha
+and Beta categories are distinct along dimensions that the Betas do not vary on.
+In this case, the weighting for a given participant can be computed as:
+
+\begin{equation} w_k = \dfrac {\exp{ \{ -\theta_w \cdot\text{range}(k)} \} } {
+\sum_k {\exp{ \{ -\theta_w \cdot\text{range}(k)} \} } }
+\label{eq:range-weight}
+\end{equation}
+% 
+where $\theta_w$ is a free parameter controlling how differences in range
+correspond to differences in weights (functioning similarly to the $\theta$
+parameter in each of the models), and $\text{range}(k)$ is the range of examples
+generated by the participant along dimension $k$. We used $\theta_w = 1.5$ in
+our simulations, though the results are robust and similar for other $\theta_w$
+values. The resulting $w$ values are thus inversely proportional to the range of
+generated categories along each dimension, with less range corresponding to
+greater weighting.
+
+Unlike the PACKER and copy-and-tweak models, the representativeness and
+hierarchical Bayesian model's dimensional variances correspond to the assumed
+variance of generated categories along each dimension (rather than the inverse
+of the variance). Thus, a different transformation is appropriate for
+incorporating the weights computed in Equation \ref{eq:range-weight}. For the
+hierarchical Bayesian model, we computed the dimensional variances according to:
+$\lambda (1-w_k) 2$, where $\lambda$ is a free parameter specifying the overall
+assumed variance of the domain, and $2$ corresponds to the number of dimensions
+in our experiments\footnote{This calculation applies only in two-dimensional
+  domains, where $w_2 = 1-w_1$.}. Under this approach, evenly distributed
+weights correspond to an assumed variance of $\lambda$. Likewise, larger values
+of $w$, which are produced when the generated category is tightly distributed
+along one dimension, correspond to smaller assumed variances.
+
+Each model was used to simulate each participant's generation independently,
+with the importance of each dimension set according to the participant's
+generated category. The other free parameters within each model were set as in
+Table \ref{table:global-model-fits}. Every participant's generation was
+simulated 2,000 times; given the 305 participants tested across the two
+experiments, each model generated 610,000 categories in total. For comparison
+with our behavioral results, we then computed the range difference gradient
+identically as with the behavioral data. The results are shown in Figure
+\ref{fig:range-diff-gradients}.
+
+As in the more traditional model evaluation analysis described above, the
+contrast models (i.e., PACKER and the representativeness model) provided a much
+closer match to our behavioral results than the copy-and-tweak and hierarchical
+Bayesian models. In all conditions, the contrast models distribute categories
+similarly to the behavioral data: Horizontally-aligned categories tend to be
+placed above and below members of the experimenter-defined category, and
+vertically-aligned categories tend to be placed to the sides. Conversely,
+because the copy-and-tweak and hierarchical Bayesian models are insensitive to
+category contrast, these models do not produce any systematic patterns of
+association between category location and distributional structure. The sole
+exception is within the `Row' condition of Experiment 1, in which the majority
+of participants generated a `Row'-like category, widely distributed along the
+X-axis but not the Y-axis. In these cases, both models are initialized with
+weights that produce Row categories, but because category contrast is not
+considered, categories are uniformly generated across the entire domain, rather
+than concentrated within the upper-regions as observed behaviorally.
+
+\include{methods_results_e3}
+
+\section{General Discussion}
+The sensory impression of every stimulus and event is unique.
+Grouping distinct patterns of sensory information into categories is a fundamental task solved by the mind.
+Most work has focused on how people learn new categories that are provided to them through unlabeled, partially labeled, or fully labeled examples.
+How were these categories first determined?
+Some natural categories are likely to be the result of regularities in the dynamics of our environment.
+But, these are only a subset of the categories that people learn.
+Other categories, such as tools and ideas, were generated by people over time.
+What basic principles underlie how people generate categories?
+
+While the bulk of prior research on categorization has focused on the classic finding
+that generated concepts tend to be distributionally similar to known concepts,
+there has been little work addressing the role of contrast in creative
+generation: How is it that people are able to create something {\em different}
+from what is already known?
+\jlanote{Update this next section:}
+We developed a novel, exemplar-based model, PACKER,
+which formally specifies the role of contrast in generation. The model proposes
+that categories are represented as exemplars in a multidimensional psychological
+space, and generation is constrained both by within-category and
+between-category similarity: Exemplars belonging to the same category should be
+similar to one another, and exemplars belonging to different categories should
+not be similar to one another.
+
+We reported two experiments demonstrating systematic
+effects of contrast in category generation. Members of
+participant-generated categories tended to be highly dissimilar from members of
+previously-learned categories, and were usually more similar to one another than
+to members of other categories. We also observed broad interdependence between
+the distributional structure (feature variance, correlation) and physical
+instantiation (location within the stimulus space) of generated categories: In
+Experiment 2, we found that the unoccupied regions of the domain influenced the
+distributional structure of categories, and in both experiments we observed that
+participants distributed their generated categories to increase contrast with
+what was already known.
+
+We conducted simulations comparing contrast accounts of our results to the other
+major proposals for category generation from prior work: a ``copy-and-tweak'' model (realized as
+a variant of PACKER with no sensitivity to category contrast), and a
+hierarchical Bayesian model designed to explain the classic distributional
+similarity effect. In all simulations, we found that 
+contrast captured a previously unexplained and unexplored aspect of human
+category generation. In the case of contrast as exemplar dissimilarity, we operationalized contrast by measuring PACKER's fit as a function of its
+prioritization of within- and between-category similarity, we observed that
+considering either constraint exclusively results in a relatively low-quality
+account. Instead, PACKER's best results were obtained when both constraints are
+considered, indicating that human learners do not generate novel concepts
+exclusively on the basis of within-category similarity or between
+class-contrast. This finding mirrors our behavioral results and demonstrates
+that both constraints influence creative generation.
+\jlanote{Above needs some updating still. Also, do we want an outline of general discussion paragraph?}
+
+\subsection{Contrast Effects in Prior Work}
 Although existing accounts of category generation broadly overlook the role of
@@
 
-\sxlnote{Added this bit on regularity below - I'm a little undecided as to leave
-  it in the main text or place as a gigantic footnote? It is a bit of a
-  digression, and it's really just here to address Charles' point.... If we want
-  to keep this in the main text, I'll have to amend the following paragraph
-  ``Re-analyzing the results...'' to improve the transition.}
-%
 It is possible, however, that this data can be explained by a process that does
@@
 probability that a hue distinct from the experimenter-defined categories is
-selected, there is a corresponding $.75$ probability (or about 17 out of 22
-participants) that the generated category will have no exemplars that share a
-hue with any of the experimenter-defined categories. In addition, that there is
-a $2/8 = .25$ probability (about 6 out of 22 participants) that the generated
+selected, there is a corresponding 75\%  chance that the generated category will have no exemplars that share a
+hue with any of the experimenter-defined categories. Conversely, that there is
+a 25\% chance that the generated
 category will have all six exemplars that share a hue with any of the
@@
 
-However, upon analysing the consistency of hues in the generated categories, it
+\jlanote{Xian: Please do a Chi-Squared test between what is expected based on the logic above and the observed frequencies mentioned below.}
+However, upon analyzing the consistency of hues in the generated categories, it
 is clear that participants do not generally produce categories with entirely
-similar hues across all exemplars. Specifically, only .55 of all generated
+similar hues across all exemplars. Specifically, only 55\% of all generated
 categories comprised exemplars with a common hue. Consequently, the assumption
@@
 
-\subsection{Contrast as exemplar dissimilarity: The PACKER Model}
-
-\label{section:PACKER-definition}
-
-As noted above, the constraint that new concepts should differ from what is
-already known has been largely overlooked in previous work. This is no doubt in
-part due to the vague definition of what it means for a concept to be
-``different'': A generated category may be different from what is already known
-in any number of respects. Towards providing a more precise definition of the
-role of contrast in creative generation, we formalized contrast in a novel
-exemplar model, PACKER ({\em Producing Alike and Contrasting Knowledge using
-Exemplar Representations}). PACKER explains category generation as a balance
-between two fundamental constraints: The category to be generated should not be
-similar to known categories, and exemplars within each category should be
-similar to one another. These ideas are implemented within the well-studied
-exemplar framework -- the PACKER model is an extension of the influential
-Generalized Context Model of categorization
-\citep[GCM;][]{nosofsky1984choice,nosofsky1986attention}.
-
-Although, as an exemplar model, one of PACKER's proposals is people represent
-categories in terms of a collection of stored exemplars, it is worth emphasizing
-that the underlying nature of human category representations is not the primary
-concern of our work. The choice to develop PACKER within an exemplar framework
-reflects the facts that exemplar models have been throughly evaluated, are
-strongly theoretically motivated, and dominate much of the theoretical and
-empirical work in categorization. The focus of our work with PACKER concerns the
-dual constraints of within- and between-class similarity; it is not difficult to
-imagine how such constraints may be instantiated using alternative frameworks
-(e.g., \citealp{kurtz2007divergent,love2004sustain,smith2000thirty}; for a
-review of categorization models, see \citealp{pothoswills2011}).
-
-Both PACKER and the GCM simulate categorization under the assumption that
-learners represent categories as a collection of exemplars, corresponding to the
-labeled stimuli they have observed. The exemplars are encoded within a
-$k$-dimensional psychological space, and model performance is based on the
-amount of similarity between the item to be categorized and the stored
-exemplars. Similarity between two examples, $s\left(x_i, x_j\right)$, is
-computed as an inverse exponential function of distance
-\citep[following][]{attneave1950,shepard1957stimulus,shepard1987toward}:
-
-\begin{equation} s\left(x_i,x_j\right) = \exp \left\{ -c \left[\sum_{k}{ w_k
-\left| x_{ik} - x_{jk} \right|^r }\right]^{1/r} \right\}
-\label{eq:similarity}
-\end{equation}
-% 
-where $w_k$ is the attention weighting of dimension $k$ ($w_k \geq 0$ and
-$\sum_k{w_k} = 1$), accounting for the relative importance of each dimension in
-similarity calculations, and $c$ ($c>0$) is a specificity parameter controlling
-the spread of exemplar generalization. For simplicity, attention will be
-distributed uniformly in our simulations (unless otherwise noted). The value of
-$r$ depends on the nature of the experimental conditions being simulated: $r=1$
-is appropriate for separable dimensions, whereas $r=2$ is appropriate for
-integral dimensions \citep[e.g.,][]{shepard1964attention,garner1974processing}.
-In our simulations, we set $r=1$ due to the separable nature of the stimulus
-dimensions used in our experiments (see Figure \ref{fig:e1-conditions}).
-
-PACKER (as well as its name) was in part inspired by earlier work from the
-categorization literature. Specifically, \cite{hidaka2011packing} argued that
-natural categories ``pack'' the values of features such that different
-categories fill the domain space with distance between one another, while
-keeping items within the same category close together. Inspired by this idea,
-PACKER proposes that generation is constrained by both similarity to members of
-the target category (the category in which a stimulus is being generated) as
-well as similarity to members of other categories: the most desirable generation
-candidates are similar to members of the target category and not similar to
-members of contrast categories. This is achieved by aggregating similarity
-across known exemplars differently according to class membership. The aggregated
-similarity $a(y,x)$ between generation candidate $y$ and stored exemplars $x$ is
-given by:
-
-\begin{equation} a(y, x) = \sum_j{f(x_j) s(y, x_j)}
-\end{equation}
-% 
-where $f(x_j)$ is a function specifying each exemplar's contribution to
-generation. A negative value for $f(x_j)$ produces a `repelling' effect (items
-are less likely to be generated nearby $x_j$), and a positive value produces an
-`attracting' effect (items are more likely to be generated nearby $x_j$). When
-$f(x_j)=0$, the exemplar does not contribute to generation.
-
-PACKER sets $f(x_j)$ depending on exemplar $x_j$'s category membership:
-$f(x_j) = \theta_t$ if $x_j$ is a member of the target category, and
-$f(x_j) = -\theta_c$ if $x_j$ is a member of a contrast category.
-$\theta_t$ and $\theta_c$ are free parameters
-($0 \leq (\theta_t,\theta_c) \leq 1$) controlling the trade-off
-between within- and between-category similarity. For example, when
-$\theta_t = \theta_c = 0.5$, $f(x_j) = 0.5$ for members of the
-target category and $f(x_j) = -0.5$ for members of other categories; thus, the
-model is likely to generate items that are similar to members of the target
-category but are not similar to members of other categories. In this way,
-$\theta_t > 0$ with $\theta_c = 0$ produces exclusive consideration
-of target-category members, and $\theta_c > 0$ with $\theta_t = 0$
-produces exclusive consideration of contrast-category members. The combination
-of $\theta_t$ and $\theta_c$ parameters thus specifies a wide
-breadth of possible approaches; by fitting it to a dataset, one can describe the
-relative roles of between-category contrast and within-category similarity in
-generation. See Figure \ref{fig:packer-examples} for an illustration of how
-these parameters control the relative influence of within-category similarity
-and contrast to other categories when generating a new exemplar.
-
-\begin{figure}
-	\begin{center} \inputpgf{figs/}{packer-examples.pgf}
-      \caption{PACKER generation of a category `B' example, following exposure
-        to one member of category `A' and one member of category `B'.
-        Predictions are shown for three different parameterizations (differing
-        in values of $\theta_t$ and $\theta_c$): {\em (a)}
-        Predictions based on contrast similarity only. {\em (b)}
-        Predictions based on target similarity only. {\em (c)} Predictions with
-        both constraints considered.}
-		\label{fig:packer-examples}
-	\end{center}
-\end{figure}
-
-The probability that a candidate $y$ will be generated is evaluated using an
-Exponentiated \citet{luce1977choice} choice rule. Candidates with greater values
-of $a(y,x)$ are more likely to be generated than candidates with smaller values:
-% 
-\begin{equation} p(y \mid x) = \dfrac { \exp \left \{ a \left( y, x
-\right) \right \} } { \sum_i{ \exp \left \{ a \left( y_i, x \right)
-\right\} } }
-\label{eq:packer-choice}
-\end{equation}
-% 
-%where $\theta$ ($\theta \geq 0$) is a free parameter controlling response
-%determinism.
-
-It is worth noting that PACKER is only one possible exemplar-based account of
-category generation within our proposed framework. That is, PACKER places
-specific constraints on the possible values of $f(x_j)$, but other
-exemplar-based category generation models with drastically different behavior
-can be formalized in this framework by imposing alternative constraints. For
-example, as will be discussed in more detail below, PACKER is formally
-equivalent to the copy-and-tweak model proposed by \cite{jern2013probabilistic}
-when $\theta_c = 0$ and $\theta_t > 0$. Likewise, when
-$\theta_t = 0$ and $\theta_c > 0$, PACKER can represent a
-contrast-only generation mode, relying exclusively on contrast when generating
-new categories. Finally, when $f(x_j) < 0$ for all $x_j$ (regardless of class
-membership), a ``pure-packing'' approach is yielded, generating items in
-unoccupied areas of the domain. Thus, the proposed framework may be used to
-describe a wide variety of qualitatively distinct generation strategies.
-
-It is also worth noting that PACKER represents just one of many possible models
-that incorporate contrast in category generation. For example, it may be
-possible to extend \cite{jern2013probabilistic}'s hierarchical Bayesian model to
-include contrast, but this mechanism would be at odds with the standard
-hierarchical Bayesian framework. In contrast, these ideas emerge naturally from
-the exemplar view: We have not modified any of the core elements of the GCM in
-defining the PACKER model, we simply weight exemplar similarity during
-aggregation. Thus, beyond formally specifying the role of contrast in creative
-generation, the PACKER account allows us to evaluate the well-understood
-principles of the exemplar view within the comparatively understudied field of
-conceptual generation. Nonetheless, we will discuss integrating our approaches
-in the General Discussion.
-
-\subsubsection{Relation Between PACKER and Copy-And-Tweak}
-\label{section:copytweak-packer} The PACKER model is similar to the
-copy-and-tweak model reported by \cite{jern2013probabilistic}: Both models are
-exemplar-based, and both models generate new items according to their similarity
-to known members of the target class. However, PACKER diverges from the
-copy-and-tweak model by including a contrast mechanism, enabling generation
-according to dissimilarity to members of opposing categories. By consequence,
-copy-and-tweak can be realized as a parameterization of the PACKER model that is
-insensitive to category contrast. Specifically, when $\theta_c = 0$ and
-$\theta_t > 0$ (see Figure \ref{fig:packer-examples}, panel B),
-$f(x_j) = 0$ for $x_j$ belonging to contrast categories; thus, PACKER is not
-influenced by these items, and is mathematically equivalent to a copy-and-tweak
-approach.
-
-In this paper, we report simulations using this copy-and-tweak model. This model
-fits within the exemplar-based category generation framework defined above, and
-is a continuous-dimension adaptation of the model tested by
-\cite{jern2013probabilistic}. By formalizing a model family where PACKER and
-copy-and-tweak are different parameterizations of models within the same
-framework, the comparison between PACKER and copy-and-tweak provides a test of
-the explanatory value of the contrast mechanism based on exemplar dissimilarity:
-The account provided by copy-and-tweak will only equal that of PACKER if the
-contrast mechanism does not offer an advantage (i.e., if $\theta_c > 0$
-significantly improves model fits). Note that the purpose of the article is to
-explore and formally analyze the role of contrast in category generation and
-thus, we leave extending PACKER to incorporate distributional factors \citep[as
-explored by][]{jern2013probabilistic} for future work.
-
-\subsection{Contrast as representativeness: An extension to Jern and Kemp
-(2013)}
-\label{section:representativeness-definition}
-
-An alternate conceptualization of category contrast is the idea of
-representativeness -- exemplars are generated such that they are more
-representative of the target category and thus more distinct from and less
-similar to other categories. We adopt the definition specified by
-\cite{tenenbaum01}, where the representativeness $R$ of an item (or in our case,
-an exemplar) $y$ is the relative amount of evidence that is provided by $y$ for
-a given hypothesis $h$ compared to all other hypotheses $H'$. More formally,
-\begin{equation}
-  R(x,h) = \log \dfrac{p(x|h)}{\sum_{h'\in H'}{p(x|h')p(h')}} , 
-\label{eq:representativeness}
-\end{equation}
-
-In Equation \ref{eq:representativeness}, $p(h')$ is the prior probability
-distribution on the hypothesis space that excludes $h$. Note that when there are
-only two hypotheses (e.g. two categories) involved in the domain, this prior
-takes the value of $1$ since there is ever only one alternative hypothesis.
-
-This Bayesian formalization of representativeness makes it straightforward to
-implement into the existing hieriarchical Bayesian model of category generation
-developed by \cite{jern2013probabilistic}. Specifically, both models assume that
-exemplars are sampled from a given category distribution ($h$ in
-Equation \ref{eq:representativeness}). Each category distribution is a
-multivariate normal parameterized by a location vector $\mu_k$ and covariance
-matrix $\Sigma_k$ (i.e., $h = \mathcal{N}(\mu_k,\Sigma_k)$). The parameters
-$\mu_k$ and $\Sigma_k$ are assumed to be samples from a prior
-normal-inverse-Wishart (NIW) distribution parameterized by
-$\mu_0,\Sigma_0,\kappa,\nu$. Mathematically,
-\begin{align}
-  \mu_k, \Sigma_k|\mu_0,\Sigma_0,\kappa,\nu &\sim
-                                              \textrm{NIW}(\mu_0,\Sigma_0,\kappa,\nu)
-                                              \label{eq:mvnsample}\\
-  y|C &\sim \mathcal{N}(\mu_k,\Sigma_k) \label{eq:exemplarsample}
-\end{align}
-
-In our simulations, we set the prior mean $\mu_0$ to the center of the stimulus
-space and $\Sigma_0 = \lambda \mathbf{I}$ where $\lambda$ is a free parameter
-and $\mathbf{I}$ is a $d$-by-$d$ identity matrix with $d$ representing the
-number of dimensions or features in the domain. The $\kappa$ and $\nu$
-parameters are freely estimated within the constraints $\kappa > d-1$ and
-$\nu > 0$.
-
-With the assumption of the NIW prior, the location vector $\mu_k$ can be computed
-with:
-\begin{equation}
-  \mu_k = \dfrac
-    {\kappa\mu_{0} + n_k \bar{x}_k}
-    {\kappa + n_k}
-    \label{eq:category_mus}
-\end{equation}
-where $n_k$ is the number of observed exemplars in category $k$ and $\bar{x}_k$
-is the observed category mean. Note that if there are no observed exemplars in
-the target category (i.e., if the model is generating a completely novel
-category), Equation \ref{eq:category_mus} simplifies to $\mu_k = \mu_0$.
-
-The NIW prior also allows us to infer the category covariance matrix $\Sigma_k$
-by computing the following:
-\begin{equation}
-  \Sigma_k = [\Sigma_D\nu + C_k +
-    \dfrac
-    {\kappa n_k}
-    {\kappa + n_k}
-    (\bar{x_k}-\mu_k)(\bar{x_k}-\mu_k)^T
-  ] (\nu + n_k)^{-1}
-  \label{eq:Sigma_y}
-\end{equation}
-where $C_k$ is the observed category covariance and $\Sigma_D$ is the domain
-covariance matrix from which $\Sigma_k$ samples are obtained. We can
-infer $\Sigma_D$ from the observed category covariances $C$ and the prior
-covariance $\Sigma_0$. Specifically,
-\begin{equation}
-    \Sigma_D = \Sigma_0 + \sum_{k}{C_k}
-\end{equation}
-
-At this point, both the representativeness model and the hierarchical Bayesian
-model from \cite{jern2013probabilistic} are largely identical. However, both
-models diverge in their computation of exemplar generation probabilities. While
-the original hierarchical Bayesian model produces novel exemplars with
-probabilities proportional to the multivariate normal likelihood $p(y|h)$, the
-representativeness model generates new exemplars with probabilities weighted by
-their representativeness $R(y,h)$. In practice, the probability of generating a
-particular candidate $y$ is obtained using an Exponentiated
-\citet{luce1977choice} choice rule:
-\begin{equation}
-  p(y) = \dfrac
-    {\exp( \theta \cdot R(y,h))}
-    {\sum_k \exp(\theta  \cdot R(y_k,h_k))} .
-\label{eq:representative-choice}
-\end{equation}
-where $\theta$ is a freely estimated response determinism parameter (constrained
-such that $\theta \geq 0$).
-
-Despite both the original hierarchical Bayesian model and our representativeness
-model sharing identical hierarchical structures, their distinct response
-processes can yield very different exemplars. Specifically, the hierarchical
-Bayesian model emphasizes the generation of exemplars that maintain
-distributional commonalities across categories. In contrast, the
-representativeness model focuses on the generation of exemplars that are
-representative of the underlying distribution for a target category. Generally,
-this representativeness mechanism results in the generation of exemplars that
-are less similar (i.e., more different) to contrast categories. We see this
-illustrated in Figure \ref{fig:representative-examples}, where the
-representativeness model displays a strong preference for generating exemplars
-from target category 'B' that are further away from the contrast category 'A'.
-
-\begin{figure}
-	\begin{center} \inputpgf{figs/}{hbm-examples.pgf}
-		\caption{Generation of a category `B' example, following exposure
-to one member of category `A' and one member of category `B'. Predictions are
-shown for {\em(a)} the original hierarchical Bayesian model and {\em(b)} the
-representativeness model.}
-		\label{fig:representative-examples}
-	\end{center}
-\end{figure}
-
-\subsection{Synopsis and Prognosis}
-
-Research on the creative generation of novel concepts has focused on the finding
-that generated categories tend to possess distributional commonalities with
-known categories. However, a fundamental goal of concept generation is to create
-something {\em new} (i.e., different from what is already known). The manner in
-which generated categories differ from known ones is, nonetheless, poorly
-understood: Existing theories do not make strong predictions about how
-creatively generated concepts should systematically differ from existing ones.
-Above, we found encouraging initial support that contrast influences category
-generation \citep[][Experiment 3]{jern2013probabilistic}, and we introduced two
-novel approaches to formalizing the roles of similarity and contrast in creative
-generation.
-
-In the sections below, we present two experiments demonstrating systematic
-effects of category contrast on category generation inspired by factors
-influencing how the contrast models generate new categories. Our experiments are
-based on \cite{jern2013probabilistic}'s paradigm, which is a straightforward
-translation of the traditional artificial classification paradigm to the task of
-category generation: Participants are first exposed to a single,
-experimenter-defined category, and are then asked to generate members of a new
-category. We then report formal analyses comparing the exemplar dissimilarity
-and the representativeness accounts of our results to that of the hierarchical
-Bayesian and copy-and-tweak models developed by \cite{jern2013probabilistic}.
-After analyzing the category generation studies, we present third experiment to
-demonstrate the effect of contrast in category learning. This ensures that our
-results are not idiosyncratic to category generation.
-
-\section{Experiment 1}
-
-To begin our investigation, we sought to extend the early evidence obtained from
-our analysis of the \cite{jern2013probabilistic} data, under a variety of
-learning conditions and using more standard stimulus materials. We used an
-artificial stimulus design: A two dimensional domain of squares, varying in
-color and size (see Figure \ref{fig:e1-conditions}, panel A). These dimensions
-have been used in numerous classification learning studies
-\cite[e.g.,][]{conaway2016similar,conaway2016generalization,shepard1961learning,nosofsky1994comparing}.
-Unlike those used in the \cite{jern2013probabilistic} experiments, distance on
-these physical dimensions aligns more directly with perceptual similarity,
-allowing us to evaluate the role of category contrast in creative generation
-more precisely. To extend the evidence provided by the
-\cite{jern2013probabilistic} data, we tested the effects of category contrast
-after learning one category from a set of qualitatively distinct category
-structures, as shown in Figure \ref{fig:e1-conditions}.
-
-Figures \ref{fig:e1-conditions}b-d show the values of exemplar dimensions
-belonging to the experimenter-defined categories (`A', or `Alpha') that
-participants were assigned to learn about prior to generating a new category.
-Each participant learned one of the category types during training. In the
-`Cluster' type, category A is a tight cluster of examples in the space.
-Perceptually instantiated, the members of category A might, for example, be
-large and dark in color. In the `Row' type, category A has a row pattern across
-the space, varying along one feature but not the other. Thus, its members might
-all be dark in color but would vary in size. Finally, in the `XOR' type, the
-experimenter-defined category consists of two clusters separated in opposite
-corners of the space, conforming to the exclusive-or logical structure (e.g.,
-members are small and dark or large and light).
-
-It should be noted that in our experiments the assignment between the perceptual
-and conceptual dimensions (e.g., $X \rightarrow Size$, $Y \rightarrow Color$),
-as well as the direction of variation along each dimension (e.g., $dark
-\rightarrow light$ or $light \rightarrow dark$) was counterbalanced across
-participants. The category types in Figure \ref{fig:e1-conditions} are plotted
-in a conceptual space, rather than a perceptual space. Thus, while the
-conceptual organization of the category types remains constant, each category
-type may have a different physical instantiation according to the counterbalance
-assignment. For example, the Cluster type may be large and dark in color, or it
-may be small and light in color, depending on the assignment and direction of
-the dimensions. For this reason, below we will discuss generation within a
-conceptual space, rather than a physically instantiated one.
-
-\begin{figure}
-    \begin{center} \inputpgf{figs/}{e1-conditions.pgf}
-    \caption{Stimulus domain and category types tested in Experiment 1. Stimuli
-are not drawn to scale.}
-    \label{fig:e1-conditions}
-    \end{center}
-\end{figure}
-
-After learning about an experimenter-defined category, participants are asked to
-generate examples of a new category. Within this paradigm, an effect of category
-contrast would be realized if participants prefer to generate items in locations
-that are distant (i.e., perceptually dissimilar) from members of category A.
-However, generation is left unconstrained. Critically, participants were not
-asked to generate something different in the prompt. For example, participants
-assigned to the Cluster condition may generate a tightly clustered category in
-the corner opposite of the experimenter-defined category. Alternatively, they
-may generate a tightly clustered category directly overlapping with the
-experimenter-defined category. Further, they may even generate an entirely
-different type of category (e.g., a row category).
-
-Our experimental results also provide a converging test of the classic finding
-that generated categories tend to share distributional properties with known
-categories in the domain \citep{jern2013probabilistic,ward1994structured}. From
-these results, we can predict that, in each condition, participants should
-generate categories that are distributionally similar to the
-experimenter-defined category: In the Cluster condition, generated categories
-should be tightly clustered. In the Row condition, generated categories should
-vary more along the X-axis than the Y-axis. In XOR condition, generated
-categories should be widely distributed across both dimensions, and the two
-dimensions should be positively correlated.
-
-Interestingly, the XOR condition also offers a dissociation between the roles of
-category contrast and the emulation of distributional structure:
-widely-distributed, positively-correlated categories would need to lie along the
-positive diagonal of the space (that is the only place they ``fit''), which is
-already occupied by the experimenter-defined category. Thus, if contrast plays a
-role, exemplars in the generated categories of participants in the XOR condition
-may not be positively correlated -- they may be negatively correlated instead.
-In this case, contrast and statistical regularities would interact, which would
-be inconsistent with the leading approach in conceptual generation
-\citep{jern2013probabilistic}
-
-\subsection{Participants and Materials}
-
-183 participants were recruited from Amazon Mechanical Turk. Each participant
-was randomly assigned to one condition: 64 participants were assigned to the
-Cluster condition, 61 were assigned to the Row condition, and 58 were assigned
-to the XOR condition (sample sizes differ due to random assignment). Stimuli
-were squares varying in color (grayscale 9.8\%--90.2\%) and side length
-(3.0--5.8cm), see Figure \ref{fig:e1-conditions}. The assignment of perceptual
-features (color, size) to axes of the domain space (x, y), as well as the
-direction of variation along each axis (e.g., $dark \rightarrow light$ or $light
-\rightarrow dark$) was counterbalanced across participants.
-
-\subsection{Procedure}
-
-As noted in the introduction of this paper, the task of generating members of a
-new category can be aligned with common tasks studied in the categorization
-literature: Whereas classification consists of predicting an object's category
-label on the basis of its features, inference consists of predicting an observed
-feature, given a set of observed features and a category label. Generation thus
-consists of predicting {\em all} features of an object, given a novel category
-label. In keeping with this insight, we designed our generation task as an
-extension of the traditional artificial classification learning paradigm. The
-task differs from traditional work in creative cognition primarily through the
-use of an artificial domain, which enables the application of computational
-models. The use of an artificial domain also requires the addition of a training
-phase, during which participants learn about the categories in the domain. As a
-result, unlike most previous studies \citep[e.g.,][]{ward1994structured},
-participants in our studies have no experience with the domain before the start
-of the experiment, and the experimenter-defined categories are not
-hierarchically structured (as are many natural categories).
-
-Participants began the experiment with a short training phase (3 blocks of 4
-trials), where they observed exemplars belonging to the `Alpha' category.
-Participants were instructed to learn as much as they can about the Alpha
-category, and that they would answer a series of test questions afterwards. On
-each trial, a single Alpha category exemplar was presented, and participants
-were given as much time as they desired to observe it before moving on to the
-next trial. Each block consisted of a single presentation of each of the members
-of the Alpha category, in a random order. Participants were shown the range of
-possible colors and sizes prior to training.
-
-Following the training phase, participants were asked to generate four examples
-belonging to another category called `Beta'. As in
-\citet{jern2013probabilistic}, generation was completed using a sliding-scale
-interface. Two scales controlled the values of the two dimensions (color, size)
-for the generated example. An on-screen preview of the example updated whenever
-one of the features was changed. Participants could generate any example along
-an evenly-spaced 9x9 grid (including members of the Alpha category), except for
-any previously generated Beta exemplars. Neither the members of the Alpha
-category nor the previously generated Beta examples were visible during
-generation. Prior to beginning the generation phase, participants read the
-following instructions:
-
-\begin{displayquote} As it turns out, there is another category of geometric
-figures called ``Beta''. Instead of showing you examples of the Beta category,
-we would like to know what you think is likely to be in the Beta category.
-
-You will now be given the chance to create examples of any size or color in
-order to show what you expect about the Beta category. You will be asked to
-produce 4 Beta examples - they can be quite similar or quite different to each
-other, depending on what you think makes the most sense for the category.
-
-Each example needs to be unique, but the computer will let you know if you
-accidentally create a repeat.
-\end{displayquote}
-
-\subsection{Results}
-
-We observed a substantial degree of individual differences in our data. In
-Figure \ref{fig:e1-samples} we have plotted sample data from several
-participants, from which it is evident that different participants generated
-qualitatively different category structures. In this section we will focus on
-analyzing the data in aggregate, but in later sections we will explore how
-individual differences can be explained.
-
-\begin{figure}
-    \begin{center} \inputpgf{figs/}{e1-samples.pgf}
-    \caption{Sample categories generated by participants in Experiment 1.
-Representative samples from common generation profiles are shown.}
-    \label{fig:e1-samples}
-    \end{center}
-\end{figure}
-
-To evaluate the role of contrast, we computed the number of times each stimulus
-was generated, as a function of its average city-block distance from members of
-the experimenter-defined ``Alpha'' category. These data, shown in Figure
-\ref{fig:e1-distanceplots}, reveal a clear pattern: Examples that are more
-distant from members of the experimenter-defined categories are more likely to
-be generated into a new category. This supports the notion that contrast is a
-fundamental constraint on creative cognition and that suggests that statistical
-regularity alone is insufficient.
-
-Figure \ref{fig:e1-distanceplots} also depicts, for each participant, the
-average distance of members within the generated category ({\em within-category}
-distance) against the average distance between members of the generated and
-experimenter-defined category ({\em between-category} distance). The narrow
-distribution of between-category distances in the XOR condition reflects the
-widely distributed nature of the experimenter-defined category, reducing the
-possible distances to members of the participant-generated category. These data
-reveal a systematic pattern: The majority of participants generated categories
-with greater between-category distance than within-category distance. That is,
-members of the generated category tended to be more similar to one another than
-to members of the experimenter-defined category. To evaluate this claim
-quantitatively, we conducted t-tests comparing the amount of within- and
-between- class distance in each condition. All conditions possessed greater
-between-category distance: Cluster, $t(63) = 11.43$, $p < 0.001$; Row, $t(60) =
-13.16$, $p < 0.001$; and XOR, $t(57) = 3.64$, $p < 0.001$. These results provide
-further evidence of an effect of category contrast: Participants prefer to
-generate categories that are dissimilar to the learned category but maintain
-some level of internal cohesion.
-
-\begin{figure}
-    \begin{center} \inputpgf{figs/}{e1-distanceplots.pgf}
-    \caption{Experiment 1 results. {\em Left}: Frequency of generation as a
-function of distance from members of the experimenter-defined category. {\em
-Right}: Scatter plot of within-category versus between-category distance in each
-of the participant-generated categories.}
-    \label{fig:e1-distanceplots}
-    \end{center}
-\end{figure}
-
-A secondary goal of this experiment was to examine whether we replicate the
-classic result that generated categories often possess the same distributional
-properties as previously-known categories. For each generated category, we
-computed the category range along each axis (X, Y), as well as the correlation
-between features. These data, shown in Figure \ref{fig:e1-statsboxes}, reveal
-broad individual differences: Within each condition, participants generated
-categories spanning the entire X- and Y- axis as well as categories that spanned
-very little along each. Likewise, in each condition participants generated
-categories possessing strongly positive, neutral, and strongly negative
-correlations between the dimensions. Comparing the distributional statistics
-between conditions yields a broad yet, as we will see, misleading replication of
-the classic effect.
-
-With respect to ranges along each axis (X, Y), the generated categories from
-each condition tend to reflect the ranges of the experimenter-defined
-categories. The categories generated in the Cluster condition were less widely
-distributed along the X-axis compared to Row, $t(123) = 5.61$, $p < 0.001$, and
-XOR, $t(120) = 2.68$, $p = 0.008$. Categories generated in the XOR condition
-were also less widely distributed along the X-axis compared to Row, $t(117) =
-2.56$, $p = 0.012$. This latter effect was not expected because the
-experimenter-defined categories for XOR and Row had similar X-ranges. However,
-the key finding is that categories from the Cluster condition tended to be more
-tightly distributed along the X-axis.
-
-\begin{figure}
-    \begin{center} \inputpgf{figs/}{e1-statsboxes.pgf}
-    \caption{Box-plots of the distributional statistics from the categories
-generated in Experiment 1. Boxes depict the median and quartiles of each
-condition, with whiskers placed at 1.5 IQR. All points outside this region are
-marked individually.}
-    \label{fig:e1-statsboxes}
-    \end{center}
-\end{figure}
-
-Likewise, categories generated in the Row condition had less Y-axis range
-compared to Cluster, $t(123) = 4.57$, $p < 0.001$ and XOR, $t(117) = 9.26$, $p <
-0.001$, and categories from the Cluster condition had less Y-axis range compared
-to XOR, $t(120) = 3.95$, $p < 0.001$. As expected, the correlations in the
-Cluster and Row conditions were not systematically positive or negative ($ps >
-0.1$). However, the generated categories in the XOR condition tended to possess
-{\em negatively} correlated dimensions, $t(57) = 2.04$, $p = 0.046$. This
-finding is notable, as it is the opposite of what would be expected, assuming
-learners are emulating the distributional structure of the experimenter-defined
-class (which possesses perfectly positively correlated features).
-
-We believe that the failure to replicate this finding is because participants in
-\cite{jern2013probabilistic} could differentiate the generated category on a
-third dimension (hue) to maintain the statistical regularities on the other two
-dimensions. Although the correlation in the XOR condition is significantly
-negative, it is clear from the box-plot in Figure~\ref{fig:e1-statsboxes} that
-it would be inappropriate to make a strong conclusion (e.g., the median is close
-to zero). However, we can conclude with confidence that there are situations
-where people do not emulate the distributional structure of the given category.
-This indicates that there is more to category generation than the emulation of
-distributional structure of other categories in the domain. Further, as we will
-discuss in more detail in the model-based analysis section, this is expected by
-our proposal that contrast is a fundamental principle in category generation.
-
-\subsection{Discussion} In Experiment 1 we sought to extend our analysis of the
-\cite{jern2013probabilistic} data by evaluating the influence of category
-contrast on creative generation, given qualitatively different types of prior
-knowledge. We found strong evidence for effects of category contrast in each
-condition: Participants were more likely to generate stimuli that are more
-distant from (i.e., less similar to) members of a previously-learned category,
-and members of participant-generated categories tended to be more similar to one
-another than to members of previously-learned categories. We also partially
-replicated the classic finding that the distributional structure of generated
-categories reflects that of previously learned categories
-\citep{jern2013probabilistic,ward1994structured}: Members of generated
-categories were more widely distributed along dimensions which were widely
-distributed in the experimenter-defined category.
-
-Notably, however, we also found that participants who learned an XOR category
-(composed of exemplars following a positive diagonal, see Figure
-\ref{fig:e1-conditions}) tended to generate items according to a {\em negative}
-feature correlation -- the opposite of what was present in the previously
-learned category. While this may be difficult to account for under existing
-theoretical approaches (which assume generated categories follow the same
-distributional structure as known categories), it can be concisely explained
-from a category contrast perspective. Specifically, within the XOR condition,
-individuals who seek to generate a category that is perceptually distinct from
-what is already known are left with only the upper-left and bottom-right
-quadrants of the space, as members of the previously-learned XOR category lie in
-the bottom-left and top-right. If examples are generated into both of the
-available quadrants, the generated category will possess a strongly negative
-correlation, opposing that of the experimenter-defined class.
-
-Thus, the core results of Experiment 1 indicate that generated categories can
-systematically differ from what we would expect based on prior work. The
-negative (or null) correlations observed in the XOR condition suggests an
-interesting interaction between contrast with a given category and emulation of
-statistical properties. That is, the constraints on creative generation imposed
-by category contrast may not simply influence the {\em location} of generated
-categories, but also their distributional structure. In Experiment 2, we test
-this claim more systematically.
-
-\section{Experiment 2.}
-
-\begin{figure}
-    \begin{center} \input{figs/e2-conditions.pgf}
-    \caption{Category types tested in Experiment 2.}
-    \label{fig:e2-conditions}
-    \end{center}
-\end{figure}
-
-To test whether category contrast influences the distributional structure of
-generated categories, we sought to identify conditions in which differences in
-the distributional structure of generated categories cannot be explained by the
-distributional structure of the experimenter-defined category. We created two
-new category types (depicted in Figure \ref{fig:e2-conditions}) that possess
-identical distributional structures (both are tight clusters of examples with no
-correlation between features), as they only differ in their Y-axis position: the
-`Bottom' category lies in the bottom-center of the space, and the `Middle'
-category lies in the center. The distributional equality of these conditions is
-key to the design of the experiment: If the distributional structure of
-previously learned categories were the only influence on the generated
-categories, we should observe no difference in the categories participants
-generate between these two conditions. Will participants distribute their
-generated category differently between conditions due to the differences in the
-available stimulus space for generating a new category?
-
-If category contrast influences the distributional structure of the categories
-people generate, then we should observe different types of categories according
-to the shape of the space that is {\em unoccupied} by members of previously
-learned categories. The difference in the Y-axis position between the Bottom and
-Middle conditions produces a considerable change to the shape of the unoccupied
-space. Participants assigned to learn the Bottom category should be less likely
-to generate exemplars into the lower regions of the stimulus space (as these
-areas possess greater similarity to members of the Bottom category), preferring
-instead to distribute exemplars across the upper region of the space. This
-constraint is lifted in the Middle condition, as the Middle category exemplars
-are equidistant to the upper and lower regions of the space. Accordingly,
-participants should be more likely to utilize both of these areas. Thus, if
-category contrast influences the distributional structure of generated
-categories, we should observe more participants in the Middle condition that
-generate examples above {\em and} below the experimenter-defined category.
-
-\subsection{Participants, Materials, and Procedure}
-
-122 participants were recruited from Amazon Mechanical Turk. 61 participants
-were randomly assigned to the Middle and Bottom conditions each. The stimulus
-space and procedure were exactly as in Experiment 1. Participants first
-completed a short training phase, followed by the generation phase. The only
-difference from Experiment 1 was the category types given to participants.
-
-\subsection{Results}
-
-\begin{figure}
-    \begin{center} \inputpgf{figs/}{e2-samples.pgf}
-    \caption{Sample categories generated in Experiment 2. }
-    \label{fig:e2-samples}
-    \end{center}
-\end{figure}
-
-As in Experiment 1, we observed broad differences in the generation approach
-taken by different participants. To characterize the nature of these
-differences, Figure \ref{fig:e2-samples} depicts sample categories generated by
-participants. The data from each condition are organized into four columns based
-on commonly observed patterns of generation: a `Cluster' type of
-tightly-clustered examples, `Row' and `Column' types of exemplars widely
-distributed along the one axis but narrowly along the other, and a `Corners'
-type, wherein participants placed exemplars in disparate corners of the space.
-As before, in this section we focus on analyzing the data in aggregate, but in
-later sections we will focus more specifically on explaining the individual
-differences.
-
-\begin{figure}
-    \begin{center} \inputpgf{figs/}{e2-distanceplots.pgf}
-    \caption{Experiment 2 results. {\em Left}: Frequency of generation as a
-function of distance from members of the experimenter-defined category. {\em
-Right}: Scatter plot of within-category versus between-category distance in each
-of the participant-generated categories.}
-    \label{fig:e2-distanceplots}
-    \end{center}
-\end{figure}
-
-We began our analysis by testing for the broad influence of category contrast on
-generation. As in Experiment 1, we computed the frequency each stimulus was
-generated as a function of its average distance from members of the
-experimenter-defined category, as well as each participant's average within- and
-between- category distance. These data, shown in Figure
-\ref{fig:e2-distanceplots}, yield very similar results. Participants generated
-stimuli that are distant from members of the experimenter-defined category, and
-the categories in each condition tended to possess more between-category than
-within-category distance: Bottom, $t(60) = 5.5$, $p < 0.001$; Middle, $t(60) =
-2.71$, $p < 0.01$. We did, however, observe a notable subgroup of participants
-in each condition who generated categories with more within-category than
-between-category distance. Upon manual inspection, many of these individuals
-appear to have assumed a `Corners' strategy, placing exemplars in disparate
-corners of the space, thus producing much more within-category distance, see
-Figure \ref{fig:e2-samples} for examples.
-
-To explore the distributional structure of the generated categories, we computed
-the range of exemplars along each axis (X, Y), as well as the correlation
-between features. These data, shown in Figure \ref{fig:e2-statsboxes}, again
-demonstrate the degree of individual differences observed in our study. In each
-condition, we observed tightly clustered and widely distributed categories along
-each dimension, as well as positively, negatively, and uncorrelated categories.
-
-\begin{figure}
-    \begin{center} \inputpgf{figs/}{e2-statsboxes.pgf}
-    \caption{Box-plots of the distributional statistics from the categories
-generated in Experiment 2. }
-    \label{fig:e2-statsboxes}
-    \end{center}
-\end{figure}
-
-As noted above, if the distributional structure of generated categories is
-influenced by the shape of the space not occupied by members of known
-categories, then participants in the Middle condition would be more likely to
-place exemplars in the upper {\em and} lower regions of the space, as members of
-the experimenter-defined category are equidistant from these regions.
-Participants in the Bottom condition should be less likely to generate category
-members in the bottom regions because members of the experimenter-defined
-category are located there. One way to test these predictions is to analyze the
-Y-axis ranges of the generated categories: If Middle participants utilize the
-upper and lower regions of the space, their categories should vary more along
-the Y-axis. T-Tests comparing the conditions on the distributional statistics,
-however, reveal few between-group differences: the conditions do not differ with
-respect to X-axis range, Y-axis range, or feature correlations ($ps>0.17$).
-
-However, our ability to detect differences in Y-Axis range using a standard
-$t$-test between the conditions is, in this case, diminished due to the
-non-normality of the data (Shapiro-Wilk normality test $W=0.77, p< 0.001$ for
-the Middle condition and $W=0.85, p < 0.001$ for the Bottom condition). Figure
-\ref{fig:e2-yranges} depicts the Y-axis position of the exemplars generated
-within each participant's category. The categories are sorted by overall range,
-then by condition assignment. These data reveal that there were nearly as many
-participants who generated categories spanning the entire Y-axis as those who
-generated categories spanning almost none of the Y-axis. The non-normality of
-the Y-axis range distributions thus requires that we use a different approach to
-addressing the experiment's main question.
-
-\begin{figure}
-    \begin{center} \inputpgf{figs/}{e2-yranges.pgf}
-    \caption{Y-Axis range and position of the participant-generated categories
-from Experiment 2. Each line corresponds to a participant's category, with
-notches corresponding to the Y-axis position of exemplars within the category
-(notches may overlap). Participants are sorted by overall range, and then by
-condition. }
-    \label{fig:e2-yranges}
-    \end{center}
-\end{figure}
-
-Because our main prediction concerns the generation of exemplars within the
-upper and lower regions of the domain, we compared the conditions in terms of
-the frequency with which participants generated examples above and below the
-categories. Specifically, we counted the number of participants in each
-condition who placed at least one `Beta' exemplar on the top and bottom `rows'
-of the space (the maximum and minimum possible Y-axis value, respectively). The
-resulting contingencies data are shown in Table \ref{table:e2-subset-table}.
-
-Firstly, it should be noted that nearly every participant utilized the top
-and/or bottom rows: only $10 / 122$ participants generated their category
-entirely within the interior region. Fisher's Exact Tests comparing the
-conditions reveal that more Middle participants generated an exemplar in the
-bottom row, $p < 0.001$, again demonstrating the role of contrast in guiding
-where exemplars are generated. The conditions did not differ in use of the top
-of the space, $p = 0.16$, however, more Middle participants placed exemplars in
-the top {\em and} bottom rows, $p = 0.038$. The latter effect is of interest
-here, as it indicates that the shape of the unoccupied space exerts some
-influence on the distributional structure of generated categories: Participants
-in the Middle condition were more likely to generate a category spanning the
-entire Y-axis. Thus, distributional structure of the generated categories can be
-influenced by category contrast alone.
-
-\begin{table}
-\begin{center}
-\caption{Experiment 2 results.}
-\label{table:e2-subset-table} \vskip 0.12in
-\begin{tabular}{ l r r} \textbf{Middle} & Used top row & No top row \\ \hline
-Used bottom row & 28 & 18 \\ No bottom row & 11 & 4 \\ \\ \textbf{Bottom} & Used
-top row & No top row \\ \hline Used bottom row & 16 & 8 \\ No bottom row & 31 &
-6 \\
-\end{tabular}
-\end{center}
-\end{table}
-
-\subsection{Discussion}
-
-In Experiment 2, we replicated the core findings from Experiment 1. Stimuli are
-more likely to be generated if they are distant from exemplars in other
-categories, and most participants generate categories with more between-category
-than within-category distance. However, we additionally found that the {\em
-position} of a previously learned category (rather than its distributional
-structure) influences the types of categories people generate: Participants who
-learned the `Middle' type were more likely to generate categories spanning the
-entire Y-axis of the space. Participants who learned the `Bottom' type were less
-likely to do so as a result of the presence of opposite category exemplars in
-the lower regions of the space.
-
-This finding cannot be explained from the perspective that the distributional
-structure of previously learned categories is the sole determinant of the
-distributional structure of generated categories. However, the observed behavior
-is expected from a category contrast perspective: Participants seeking to
-generate a perceptually distinct category will be more likely to use areas of
-space that are unoccupied by exemplars belonging to previously learned
-categories. In the Middle condition, the upper and lower regions of space are
-equidistant from members of the experimenter-defined category, whereas in the
-Bottom condition, the lower region of the space is closer to members of the
-experimenter-defined category. Thus, while Middle participants may form
-categories around the use of the equally unoccupied areas, the same is not true
-for the Bottom condition.
-
-\section{Model-based Analyses}
-\sxlnote{For this section I've added the existence of the representativeness
-  model but I think discussion of its performance is still lacking compared to PACKER.}
-Experiments 1 and 2 revealed systematic and strong effects of category contrast
-on creative generation. In this section, we report the results of simulations
-with formal models aimed at explaining our observations. Specifically, we
-present simulations from the two novel contrast models: PACKER and the
-representativeness model, and compare them to two models that do not incorporate
-contrast: the copy-and-tweak model (discussed in Section
-\ref{section:copytweak-packer}) and an implementation of the hierarchical
-Bayesian model proposed by \cite{jern2013probabilistic}, described in-depth in
-Appendix \ref{ap:hsampling-definition}. For our simulations here, the
-copy-and-tweak model is defined as a variant of PACKER with the $\theta_c$
-parameter constrained to be zero. The comparison of this set of models serves to
-highlight the explanatory role of contrast in categorization: If contrast
-affords little explanatory advantage, then there should be few differences in
-performance between PACKER and copy-and-tweak, or between the representativeness
-and hierarchical Bayesian model. The comparison between these models can also
-emphasize the necessity of contrast and demonstrate that generation cannot be
-explained entirely through the emulation of distributional structure. Each model
-has complementary strengths and weaknesses: Whereas PACKER and copy-and-tweak
-are insensitive to the distributional structure of learned categories (relying
-only on exemplar similarities), the representativeness and hierarchical Bayesian
-model generates categories exclusively on the basis of knowledge of how existing
-classes are distributed.
-
-Our approach in this section is to first broadly evaluate and compare the
-quality of each model's account to our entire dataset (Experiments 1 and 2
-combined), then analyze the ability for each model to explain individual
-differences in each experiment, and lastly we describe the strengths and
-weakness of each model's account of category generation.
-
-\subsection{Parameter-Fitting}
-
-To obtain a global measure of the quality of each model's account, we fitted the
-parameters of each model to our entire dataset (Experiments 1 and 2 combined),
-using a hill-climbing algorithm which maximized the log-likelihood of the
-model's predictions of the observed responses (1220 responses from 305 total
-participants). We fitted three parameters in the PACKER model ($c$, $\theta_t$,
-and $\theta_c$; see Section \ref{section:PACKER-definition}), as well as four in
-the representativeness model and the hierarchical Bayesian model ($\kappa$,
-$\lambda$, $\nu$, and $\theta$; see Section
-\ref{section:representativeness-definition} and Appendix
-\ref{ap:hsampling-definition} respectively). We fitted only two parameters for
-the copy-and-tweak model ($c$, and $\theta_t$), as $\theta_c$ is held constant
-($\theta_c = 0$). Attention ($w$, see Equation \ref{eq:similarity}) in PACKER
-and copy-and-tweak was set uniformly. Parameters were not allowed to vary
-between participants or conditions -- the goal was to obtain the best-fitting
-values to our entire dataset.
-
-\begin{table} \centering
-\caption{Results of model-fitting to the combined datasets from Experiments 1
-and 2. Note that smaller AIC values correspond to better model fits (adjusted
-for number of parameters)}
-\label{table:global-model-fits}
-\begin{tabular}{ l l l l} \\ \textbf{PACKER} & \textbf{Copy \& Tweak} &
-\textbf{Representativeness} & \textbf{Hierarchical Bayesian} \\ \hline $AIC =
-  9095$ & $AIC = 9842$ & $AIC = 8878$ &  $AIC =
-9912$ \\ $L = -4545$ & $L = -4919$ & $L = -4435$ & $L = -4952$ \\ $c = 0.483$ &
-                                                                                $c = 3.197$ & $\kappa = 11.232$ &
-$\kappa < 0.001$ \\ $\theta_c = 3.16$ & $\theta_c = 0$ (fixed) & $\nu = 2.844$ & $\nu = 5.597$ \\
-$\theta_t = 3.494$ & $\theta_t = 2.972$ & $\lambda = 1.917$& $\lambda
-                                                                       = 0.055$
-  \\ & &$\theta = 5.885 $& $\theta = 3.175$
-\\
-\end{tabular}
-\end{table}
-
-Table \ref{table:global-model-fits} contains the results of this fitting
-procedure. Due to the uneven number of fitted parameters among the models, we
-compare the model fits using the Akaike Information Criterion
-\citep[AIC;][]{akaike1974new}, where smaller values correspond to better fits
-(discounted by model complexity as measured by the number of parameters). The
-same qualitative results were obtained with alternative model comparison metrics
-(e.g., BIC, \citealp{schwarz1978estimating}; $AIC_C$,
-\citealp{hurvich1989regression}).
-
-Table \ref{table:global-model-fits} contains the AIC values of each model's best
-fit, as well as the corresponding log-likelihood ($L$) and the best-fitting
-parameter values. These results reveal strong model differentiation: both
-contrast models (PACKER and the representativeness model) achieved far better
-fits compared to their non-contrast counterparts: copy-and-tweak and the
-hierarchical Bayesian model respectively. Interestingly between the contrast
-models, the hierarchical model (representativeness) outperformed the
-exemplar-based theory (PACKER), whereas between the non-contrast models, the
-reverse is observed. Specifically, here the exemplar-based model
-(copy-and-tweak) performed somewhat better than the hierarchical Bayesian model.
-
-\sxlnote{The argument in the following two paragraphs appear to be challenged by
-  the success of the representativeness model. I might suggest removing them altogether?}
-While PACKER's advantage may tentatively be attributed to the model's
-sensitivity to category contrast (this will be explored in detail below), the
-advantage shown by copy-and-tweak over the hierarchical Bayesian model may be
-attributed to its exemplar-based representation, as opposed to the
-prototype-based representation assumed by the hierarchical Bayesian model. As
-observed in Figures \ref{fig:e1-samples} and \ref{fig:e2-samples}, the generated
-categories we observed were often widely distributed, with no items near the
-category prototype. This aspect of the data is inconsistent with the
-multivariate normal distributions (similar to prototypes) used to represent
-categories in the \cite{jern2013probabilistic} model, but can be easily
-accounted for using an exemplar-based approach.
-
-A key distinction between PACKER and the other three models is that only PACKER
-is capable of making strong predictions about the location of new category
-members when the target class is entirely novel (i.e., no member of the category
-has been observed). Under these circumstances, there are no examples to copy,
-and thus the copy-and-tweak model predicts that items are generated at random.
-Likewise, with no observations on which to condition the category distribution,
-the hierarchical Bayesian and representativeness models also picks an item at
-random. Crucially, however, the next generated exemplar from the
-representativeness model is strongly influenced by contrast as
-representativeness. Thus, it is possible that the failure of these models is
-simply due to their inability to explain each participant's first trial
-(generating the first item in the `Beta' category). To ensure this is not
-driving our results, we conducted an identical set of simulations as above,
-excluding this trial (leaving 915 responses in the dataset): Again, PACKER
-($L = -3390$, $AIC = 6786$) achieved better fits than the copy-and-tweak
-($L = -3579$, $AIC = 7162$) and hierarchical Bayesian ($L=-3612$, $AIC = 7232$)
-models.
-
-Finally, because copy-and-tweak is nested within PACKER, we can use a likelihood
-ratio test to compare the two models. PACKER explains the aggregate data
-significantly better than copy-and-tweak ($\chi^2(1) = 749, p < 0.001$ for all
-data and $\chi^2(1) = 377, p < 0.001$ excluding the first example), providing
-further evidence that category generation is better explained when contrast is
-considered.
-
-Through comparison with the copy-and-tweak model, Figure
-\ref{fig:packer-loglike} more clearly demonstrates the robustness of the
-explanatory gains yielded by PACKER's category contrast mechanism. It displays
-the log-likelihood of the participants' results under PACKER as a function of
-the $\theta_c$ parameter. The model's other parameters ($c$,
-$\theta_t$) were set according to copy-and-tweak's best fits from Table
-\ref{table:global-model-fits}, and thus when $\theta_c=0$, the models are
-equivalent. The figure clearly shows a ``sweet spot'': a convex region in which
-PACKER achieves superior fits as a result of changes to $\theta_c$. The
-best fitting values lie well above the value of 0 assumed by the copy-and-tweak
-model, which demonstrates the robustness of the contrast effect (though note
-PACKER achieves even better fits when its parameters are fitted together, as in
-Table \ref{table:global-model-fits}). In sum, the data are better explained when
-both within-category similarity and category contrast is considered.
-
-\begin{figure}
-    \begin{center} \inputpgf{figs/}{packer-loglike.pgf}
-      \caption{PACKER's fit as a function of its prioritization of
-        within-category and between-category similarity (using the
-        $\theta_c$ parameter) . To facilitate comparison, PACKER's other
-        parameters ($c$, $\theta_t$) were set to the best fitting values
-        obtained for copy-and-tweak in Table \ref{table:global-model-fits}. }
-    \label{fig:packer-loglike}
-    \end{center}
-\end{figure}
-
-\subsection{Individual Differences}
-\label{section:individual-diffs}
-
-As noted in Experiments 1 and 2, we observed a great deal of individual
-differences in the types of categories that participants generated. Within each
-condition, there were a wide variety of category types, such as row and column
-categories (see Figures \ref{fig:e1-samples} and \ref{fig:e2-samples}). The
-simulations reported above serve to evaluate the models while considering the
-entire dataset, but a secondary goal of any formal account should be to provide
-some explanation of how different profiles of performance emerge. Many of the
-individual generation profiles we observed can be described within the PACKER
-framework, simply by tuning the model's parameters in a principled manner. In
-this section, we describe more specifically how the most frequently observed
-profiles can be realized.
-
-By manual inspection, it is evident that the most common profiles of generation
-consist of: (A) a tightly-distributed `cluster' of examples, (B) `row'- and
-`column'-like arrangements (varying widely along one dimension but not the
-other), and (C) a `corners' arrangement with examples placed into disparate
-corners of the space. These four profiles are distinct in terms of the
-distribution of the generated category along each dimension: Whereas the cluster
-profile is tightly distributed along both dimensions, the row and column
-profiles are tightly distributed along just one dimension. Finally, the corners
-profile is widely distributed along both dimensions.
-
-In the framework proposed by PACKER, the cluster and corners profiles arise
-based on different prioritization of within-category similarity versus
-between-category contrast, and the row and column profiles arise based on the
-prioritization of each dimension in the computation of similarity. For example,
-in the cluster profile, there is a high degree of within-category similarity
-along both dimensions, whereas in the corners profile there is minimal
-within-category similarity. Thus, PACKER's proposal is that these individual
-differences arise as a result of different priorities: While the tight cluster
-configuration can be considered PACKER's `default' mode (as it maximizes
-within-category similarity), the corners profile can be produced when
-between-category contrast is put at a higher priority (i.e., $\theta_c > \theta_t$).
-
-Likewise, in the row and column profiles, there is a large degree of
-within-category similarity along one dimension but not the other. These
-differences likely arise due to a differential focus on one dimension over
-another, and thus they can be produced by changes to PACKER's attention weights,
-$w_1$ and $w_2$ (see Equation \ref{eq:similarity}). Traditionally, the attention
-weights in exemplar models are thought to reflect the diagnostic value of each
-dimension towards classifying the known category members
-\citep{nosofsky1984choice,nosofsky1986attention,kruschke1992alcove}, but within
-a generation context the weights specify the importance of within- and
-between-category similarity along each dimension. For example, if all of
-attention is allocated along the X-axis ($w_1=1$ and $w_2=0$), similarity along
-the Y-axis no longer influences performance. As a result, PACKER will create
-categories that are more widely distributed along the Y-axis, as similarity is
-not taken into account along that dimension. As a general principle,
-differentially weighting one dimension will result in the generation of
-categories that are more widely distributed along the ignored dimension,
-conforming to a row- or column-like arrangement. See Figure
-\ref{fig:packer-attention} for a depiction of how attention influences PACKER's
-performance.
-
-\begin{figure}
-    \begin{center} \inputpgf{figs/}{packer-attention-examples.pgf}
-    \caption{PACKER generation of a category `B' example, following exposure to
-one member of category `A' and one member of category `B'. Predictions are shown
-for different attention settings: {\em (a)} Increased weighting of the X-axis.
-{\em (b)} Increased weighting of the Y-axis. {\em (c)} Uniform weighting
-(identical to Figure \ref{fig:packer-examples}).}
-    \label{fig:packer-attention}
-    \end{center}
-\end{figure}
-
-As in PACKER, changes in the parameter settings of the copy-and-tweak model can
-also be used to produce different patterns of generation. Indeed, as
-copy-and-tweak is simply a special case of the PACKER model, the attention
-weights operate exactly as described above to produce row- and column-like
-categories. However, because the model is not influenced by category contrast,
-it is biased toward generating tightly clustered categories, as new items are
-always most likely to be generated near known examples of the target category.
-Thus, the lack of a contrast mechanism prevents the model from explaining why
-some individuals widely distribute their categories to the corners of the space.
-
-Similar to the copy-and-tweak model, the hierarchical Bayesian model possesses
-no mechanism to account for category contrast, and the model is most likely to
-generate new items that are similar to known examples of the target category.
-Although this precludes an account of why individuals might assume a cluster
-versus corners profile, the covariance matrix specifying the model's prior
-domain distribution, $\Sigma_0$, can be used to explain the generation of
-row-like and column-like categories. This covariance matrix specifies the amount
-of variance assumed along each dimension (as well as the dimensional
-correlations) across the domain of categories. The covariance matrix for a newly
-generated category, $\Sigma_B$, is based on the assumed $\Sigma_0$ as well as
-the distributions of previously learned categories (see Appendix
-\ref{ap:hsampling-definition}). Thus, the importance of each feature can be
-coded into $\Sigma_0$ to alter the dimensional variance of generated categories.
-
-As noted above, while the copy-and-tweak and hierarchical Bayesian models
-possess mechanisms to explain row- and column-like categories, they cannot
-easily explain why some individuals widely distribute their generated categories
-into disparate corners of the space. This, however, reveals a more general
-limitation: According to the copy-and-tweak and hierarchical Bayesian models,
-the distributional structure of generated categories is {\em independent} of
-their location within the domain. For example, although the copy-and-tweak or
-hierarchical Bayesian models can be parameterized to generate row- or
-column-like categories, there is no mechanism in place to ensure that what is
-generated will be distinct from what is already known. In the next subsection,
-we explore this prediction through an analysis of the interdependence between
-distributional structure and location in creative generation.
-
-\subsection{Category Location vs. Distributional Structure}
-\label{section:individual-diff}
-As noted above, while all three models make clear claims about the internal
-structure of generated categories, the copy-and-tweak and hierarchical Bayesian
-models do not make any claims about how generated categories should differ from
-what is already known. However, as we observed in the results of Experiment 2,
-the distributional structure of a category is not always independent of its
-location within the domain. To demonstrate this point in more depth, we computed
-the X- and Y- axis ranges of every participant-generated category. Taking the
-difference between these values ($X-Y$) produces a measure of each category's
-orientation in the space: positive difference scores correspond to categories
-with more X-axis range (horizontally aligned, `Row' categories), whereas
-negative difference scores indicate the opposite (vertically aligned, `Column'
-categories). Neutral differences scores indicate there was an equal amount of X-
-and Y-axis range, which can be produced by a number of different category types
-(`Clusters', `Corners', etc; see Figures \ref{fig:e1-samples} and
-\ref{fig:e2-samples}). By plotting, for each possible stimulus, the difference
-scores of categories it was generated within, we can relate the distributional
-structure of generated categories to their location within the domain.
-
-However, because many stimuli were infrequently generated (such items near
-members of the `Alpha' category), we cannot simply compute the empirical average
-of the difference scores, as infrequently generated stimuli would be likely to
-show artificially strong differences. Instead, we used a Bayesian analysis to
-estimate the mean $\mu_x$ on the assumption that the scores $x$ for each
-stimulus are normally distributed with an unknown mean and unknown standard
-deviation. The conjugate Normal-Inverse Gamma distribution provides a
-straightforward method for this estimation:
-
-\begin{equation} \mu_x = \dfrac { \nu_0 \mu_0 + \sum{x} } { \nu_0 + n }
-\label{eq:rangediff-bayes}
-\end{equation} \ where $\mu_0$ is the prior mean, $\nu_0$ is a prior scale
-parameter (controlling the weighting of the $\mu_0$), and $n$ is the number of
-categories in which the stimulus was a member (i.e., the number of scores in
-$x$). The default assumption is that there is an equal amount of range along the
-X- and Y-axes, and so we set $\mu_0 = 0$. Likewise, to give a moderate amount of
-weighting to the prior mean we set $\nu_0 = 1$, though the results are robust to
-a range of values. Within this approach, the resulting aggregation is a
-trade-off between the number of generations and the strength of the range
-difference within each generated category. Infrequently generated stimuli, as
-well as those with mixed positive and negative scores, are given neutral
-difference scores.
-
-\begin{figure}[p]
-    \begin{center} 
-      \includegraphics[width=\textwidth]{figs/range-diff-gradients.pdf}
-      \caption{Behavioral and simulated range difference gradients. Each panel
-        shows, for each stimulus, the dimensional orientation of the categories
-        it was generated into: vertically aligned `columns' (orange) versus
-        horizontally aligned `rows' (purple).}
-      \label{fig:range-diff-gradients}
-    \end{center}
-\end{figure}
-
-The results of our analysis are shown in Figure \ref{fig:range-diff-gradients}
-for the experiment and model results\footnote{Prior to plotting, data was also
-processed using a Gaussian filter with $\sigma = 0.8$.}. The left-most column of
-Figure \ref{fig:range-diff-gradients} displays the effect of category location
-and contrast on the distributional structure of the category generated by
-participants. These data reveal strong and consistent patterns across all the
-conditions we tested in Experiments 1 and 2: Generated categories are more
-tightly distributed along the axis in which they are distinct. For example, in
-the `Cluster' condition, exemplars in the bottom-left of the space are more
-often generated into vertically aligned categories, and exemplars in the
-top-right are more often generated into horizontally aligned categories.
-Similarly, in the `Bottom' and `Middle' conditions, horizontally aligned
-categories are generated above and below the experimenter-defined categories,
-while vertically-aligned categories are generated to the sides. In the `Row'
-condition, most categories are horizontally aligned, and lie along the upper
-areas of the space. There are no strong range difference patterns in the XOR
-condition.
-
-These patterns of performance clearly depict the interdependence between the
-distributional structure and location of generated concepts. Our results can be
-interpreted in terms of local minimization of between-category similarity: By
-distributing the generated category away from members of the
-experimenter-defined category, participants may increase the degree of
-between-category distance without drastically altering the degree of
-within-category similarity.
-
-To explore how well the PACKER, copy-and-tweak, representativeness, and
-hierarchical Bayesian models explain our findings, we conducted simulations
-using an individual-differences approach. As noted in Section
-\ref{section:individual-diffs}, row- and column-like categories can be produced
-by each model through changes to the weighting of each dimension. Given this
-information, we may use the models to simulate each participant's generation
-separately, with the importance of each dimension set according to the relative
-range of the participant's generated category along each dimension.
-
-In the PACKER and copy-and-tweak models, the attention weights, $w$, specify the
-importance of each dimension in the computation of similarity. While there exist
-methods to find the optimal attention weighting scheme given a classification
-\citep[see][]{vanpaemel2012using}, for simplicity we may assume that the Alpha
-and Beta categories are distinct along dimensions that the Betas do not vary on.
-Thus, the weighting for a given participant can be computed as:
-
-\begin{equation} w_k = \dfrac {\exp{ \{ -\theta_w \cdot\text{range}(k)} \} } {
-\sum_k {\exp{ \{ -\theta_w \cdot\text{range}(k)} \} } }
-\label{eq:range-weight}
-\end{equation}
-% 
-where $\theta_w$ is a free parameter controlling how differences in range
-correspond to differences in weights (functioning similarly to the $\theta$
-parameter in each of the models), and $\text{range}(k)$ is the range of examples
-generated by the participant along dimension $k$. We used $\theta_w = 1.5$ in
-our simulations, though the results are robust and similar for other $\theta_w$
-values. The resulting $w$ values are thus inversely proportional to the range of
-generated categories along each dimension, with less range corresponding to
-greater weighting.
-
-Unlike the PACKER and copy-and-tweak models, the representativeness and
-hierarchical Bayesian model's dimensional variances correspond to the assumed
-variance of generated categories along each dimension (rather than the inverse
-of the variance). Thus, a different transformation is appropriate for
-incorporating the weights computed in Equation \ref{eq:range-weight}. For the
-hierarchical Bayesian model, we computed the dimensional variances according to:
-$\lambda (1-w_k) 2$, where $\lambda$ is a free parameter specifying the overall
-assumed variance of the domain, and $2$ corresponds to the number of dimensions
-in our experiments\footnote{This calculation applies only in two-dimensional
-  domains, where $w_2 = 1-w_1$.}. Under this approach, evenly distributed
-weights correspond to an assumed variance of $\lambda$. Likewise, larger values
-of $w$, which are produced when the generated category is tightly distributed
-along one dimension, correspond to smaller assumed variances.
-
-Each model was used to simulate each participant's generation independently,
-with the importance of each dimension set according to the participant's
-generated category. The other free parameters within each model were set as in
-Table \ref{table:global-model-fits}. Every participant's generation was
-simulated 2,000 times; given the 305 participants tested across the two
-experiments, each model generated 610,000 categories in total. For comparison
-with our behavioral results, we then computed the range difference gradient
-identically as with the behavioral data. The results are shown in Figure
-\ref{fig:range-diff-gradients}.
-
-As in the more traditional model evaluation analysis described above, the
-contrast models (i.e., PACKER and the representativeness model) provided a much
-closer match to our behavioral results than the copy-and-tweak and hierarchical
-Bayesian models. In all conditions, the contrast models distribute categories
-similarly to the behavioral data: Horizontally-aligned categories tend to be
-placed above and below members of the experimenter-defined category, and
-vertically-aligned categories tend to be placed to the sides. Conversely,
-because the copy-and-tweak and hierarchical Bayesian models are insensitive to
-category contrast, these models do not produce any systematic patterns of
-association between category location and distributional structure. The sole
-exception is within the `Row' condition of Experiment 1, in which the majority
-of participants generated a `Row'-like category, widely distributed along the
-X-axis but not the Y-axis. In these cases, both models are initialized with
-weights that produce Row categories, but because category contrast is not
-considered, categories are uniformly generated across the entire domain, rather
-than concentrated within the upper-regions as observed behaviorally.
-
-\include{methods_results_e3}
-
-\section{General Discussion}
-
-The creative generation of concepts is an immensely intriguing topic, but it is
-also complex and difficult to study in a traditional laboratory setting.
-Creative generation is most typically studied using a creative cognition
-approach \citep{finke1992creative,smith1995creative}, where creative products
-\citep[such as drawings of alien plants and animals;][]{ward1994structured} are
-analyzed to obtain insight into the processes and representations involved.
-While this approach has yielded a great deal of information about the nature of
-creative generation, the commonly employed response modalities (e.g., hand-drawn
-pictures) make it difficult to develop computational models capable of formally
-evaluating the proposed representations and processes. However, the generation
-of novel concepts is not altogether different from other behaviors studied in
-cognitive psychology: In particular, it can be viewed as an form of inference
-using one's knowledge about existing categories
-\citep{kemp2014taxonomy,kurtz2015human}. Taking advantage of this insight, in
-this paper we investigated the generation of new categories using the empirical
-and computational toolkit from the field of categorization.
-
-While the bulk of prior research on the topic has focused on the classic finding
-that generated concepts tend to be distributionally similar to known concepts,
-there has been little work addressing the role of contrast in creative
-generation: How is it that people are able to create something {\em different}
-from what is already known? We developed a novel, exemplar-based model, PACKER,
-which formally specifies the role of contrast in generation. The model proposes
-that categories are represented as exemplars in a multidimensional psychological
-space, and generation is constrained both by within-category and
-between-category similarity: Exemplars belonging to the same category should be
-similar to one another, and exemplars belonging to different categories should
-not be similar to one another.
-
-In addition to an analysis of published data \citep[][Experiment
-3]{jern2013probabilistic}, we reported two experiments demonstrating systematic
-effects of category contrast in creative generation. Members of
-participant-generated categories tended to be highly dissimilar from members of
-previously-learned categories, and were usually more similar to one another than
-to members of other categories. We also observed broad interdependence between
-the distributional structure (feature variance, correlation) and physical
-instantiation (location within the stimulus space) of generated categories: In
-Experiment 2, we found that the unoccupied regions of the domain influenced the
-distributional structure of categories, and in both experiments we observed that
-participants distributed their generated categories to increase contrast with
-what was already known.
-
-We conducted simulations comparing PACKER's account of our results to the other
-major proposals for category generation: a ``copy-and-tweak'' model (realized as
-a variant of PACKER with no sensitivity to category contrast), and a
-hierarchical Bayesian model designed to explain the classic distributional
-similarity effect. In all simulations, we found that PACKER's sensitivity to
-contrast captured a previously unexplained and unexplored aspect of human
-category generation. Further, by measuring PACKER's fit as a function of its
-prioritization of within- and between-category similarity, we observed that
-considering either constraint exclusively results in a relatively low-quality
-account. Instead, PACKER's best results were obtained when both constraints are
-considered, indicating that human learners do not generate novel concepts
-exclusively on the basis of within-category similarity or between
-class-contrast. This finding mirrors our behavioral results and demonstrates
-that both constraints influence creative generation.
-
 \subsection{Similarity and Contrast in Cognition}
-
-The proposal that category contrast is a primary constraint in creative
-generation is, in some ways, entirely commonsense. To successful create
-something \textit{new}, it must be different from what is known. Beyond its role
-in creative generation, category contrast is also of fundamental importance in
+The proposal that category contrast is a primary constraint in categorization is, in some ways, entirely commonsense.
+For categories to be useful, they should not all be identical, or in other words, they should be different.
+Thus, a newly generated category should be different than pre-existing categories.
+Beyond its role
+in category generation specifically, category contrast is also of fundamental importance in
 categorization more broadly. All other factors held constant, new categories are
@@
 cognitive functions in a wide variety of domains. The PACKER model is notable in
-that it successfully interprets this trade-off within the domain of creative
-generation, and allows us to begin to understand the relatively understudied
-processes involved in creative generation through our more well-developed
+that it explicitly interprets this trade-off within the domain of categorization, and allows us to begin to understand the relatively understudied
+processes involved in category generation through our more well-developed
 knowledge of human categorization.
@@
 \subsection{Implications for Creative Cognition}
-
-Although the focus of this article has been to address the processes involved in
-creative generation using the empirical and quantitative toolkit of traditional
-categorization research, our findings and approach have relevant implications
+\jlanote{this needs updating for representativeness.}
+Although the focus of this article has been to address the role of contrast in category generation and learning, our findings and approach have relevant implications
 for research in creative cognition. A central focus of the creative cognition
@@
   under which contrast are important.} 
+\jlanote{Ken, could you take a stab at this section please?}
 Although both PACKER and the
@@
 
-Xian: Please
-write about the two hypotheses and that followup work needs to look into why the
-Bayesian representativeness model doesn't do as well at explaining category
-learning ease, but does better at category generation.
-
-MAKE SURE WE CITE AND DESCRIBE \citet{thomas98}! \sxlnote{I moved the
-  description and citation of thomas98 to the introduction for justification of
-  why we think categories should share similar distributions.}
+%Xian: Please
+%write about the two hypotheses and that followup work needs to look into why the
+%Bayesian representativeness model doesn't do as well at explaining category
+%learning ease, but does better at category generation.
+
 \subsection{Limitations and Future Directions} Although we have demonstrated

### 2018-07-20 Manuscripts/cog-psych/revision/methods_results_e3.tex
major update. many things. to ken/xian

@@
 
-\sxlnote{I felt like there could be a little more explanation in terms of
-  justifying this experiment so I added an extra sentence (the second last in
-  the following paragraph) - I don't have too strong feelings about it so please
-  feel free to remove if you think it's too much. Also see the footnote
-  addressing the paper's lack of category learning fits. It's a bit of a cop
-  out, but entirely true!}
-
 Although category generation is an intriguing aspect of human cognition, it has
@@
 is to investigate whether model performance (in generating categories) is
-consistent with human performance (in learning those categories).\footnote{It is
-  possible to adjust the present models of category generation such that they
-  can make category learning predictions. However, such an adjustment makes the
-  strong assumption that the very same processes governing category generation
-  also apply to category learning. Taking this approach would fall outside the
-  scope of the current investigation.} Specifically, if contrast plays a
-broader role in categorization, then Beta categories rated as "better" according
+consistent with human performance (in learning those categories). Specifically, if contrast plays a broader role in categorization, then Beta categories rated as "better" according
 to the contrast models should be easier for participants to learn.
-
+%\footnote{It is possible to adjust the present models of category generation such that they can make category learning predictions.
+However, such an adjustment makes the strong assumption that the very same processes governing category generation also apply to category learning.
+Taking this approach would fall outside the scope of the current investigation.}
 \subsection{Participants, Materials, and Procedure}
@@
 Overall accuracy of the participants was high, with a mean error rate of $.19$
-($SD = .19$). Error rates for each block are presented in Figure
+($SD = .19$). Aggregated error rates for each block are presented in Figure
 \ref{fig:learningcurve}.
@@
 Specifically, PACKER correlated with human performance with
-$r = .65, 95\% CI$\footnote{Confidence intervals were obtained using a bootstrap
-  method with 1000 bootstrap samples.}$ [.47,.78]$, and the representativeness
-model correlated with human performance with $r = .56, 95\% CI [.38, .72]$.
-Copy-and-tweak yielded a correlation of $r = .45, 95\% CI [.29, .61]$ and the
+$r = .65, \ 95\% CI$\footnote{Confidence intervals were obtained using a bootstrap
+  method with 1000 bootstrap samples.}$ \ [.47,.78]$, and the representativeness
+model correlated with human performance with $r = .56, \ 95\%  CI \ [.38, .72]$.
+Copy-and-tweak yielded a correlation of $r = .45, 95\% CI \ [.29, .61]$ and the
 hierarchical Bayesian model yielded a similar correlation of
-$r = .46, 95\% CI [.30, .61]$. These results are graphically presented in Figure
+$r = .46, 95\% CI \ [.30, .61]$. These results are graphically presented in Figure
 \ref{fig:perror_corr}.

### 2018-09-03 Manuscripts/cog-psych/revision/cogpsych.tex
Need to investigate source of weird corr fits.

--- 2018-07-17 Manuscripts/cog-psych/revision/cogpsych.tex
@@
 $f(x_j) = \theta_t$ if $x_j$ is a member of the target category, and
-$f(x_j) = -\theta_c$ if $x_j$ is a member of a contrast category.
-$\theta_t$ and $\theta_c$ are free parameters
-($0 \leq (\theta_t,\theta_c) \leq 1$) controlling the trade-off
-between within- and between-category similarity. For example, when
-$\theta_t = \theta_c = 0.5$, $f(x_j) = 0.5$ for members of the
-target category and $f(x_j) = -0.5$ for members of other categories; thus, the
-model is likely to generate items that are similar to members of the target
-category but are not similar to members of other categories. In this way,
-$\theta_t > 0$ with $\theta_c = 0$ produces exclusive consideration
-of target-category members, and $\theta_c > 0$ with $\theta_t = 0$
-produces exclusive consideration of contrast-category members. The combination
-of $\theta_t$ and $\theta_c$ parameters thus specifies a wide
-breadth of possible approaches; by fitting it to a dataset, one can describe the
-relative roles of between-category contrast and within-category similarity in
-generation. See Figure \ref{fig:packer-examples} for an illustration of how
-these parameters control the relative influence of within-category similarity
-and contrast to other categories when generating a new exemplar.
+$f(x_j) = -\theta_c$ if $x_j$ is a member of a contrast category. $\theta_t$ and
+$\theta_c$ are free parameters ($0 \leq \theta_t,\theta_c$) controlling the
+trade-off between within- and between-category similarity. For example, when
+$\theta_t = \theta_c = 0.5$, $f(x_j) = 0.5$ for members of the target category
+and $f(x_j) = -0.5$ for members of other categories; thus, the model is likely
+to generate items that are similar to members of the target category but are not
+similar to members of other categories. In this way, $\theta_t > 0$ with
+$\theta_c = 0$ produces exclusive consideration of target-category members, and
+$\theta_c > 0$ with $\theta_t = 0$ produces exclusive consideration of
+contrast-category members. The combination of $\theta_t$ and $\theta_c$
+parameters thus specifies a wide breadth of possible approaches; by fitting it
+to a dataset, one can describe the relative roles of between-category contrast
+and within-category similarity in generation. See Figure
+\ref{fig:packer-examples} for an illustration of how these parameters control
+the relative influence of within-category similarity and contrast to other
+categories when generating a new exemplar.
 
@@
 representative of the underlying distribution for a target category. Generally,
-this representativeness mechanism results in the generation of exemplars that
-are less similar (i.e., more different) to contrast categories. We see this
-illustrated in Figure \ref{fig:representative-examples}, where the
-representativeness model displays a strong preference for generating exemplars
-from target category 'B' that are further away from the contrast category 'A'.
+this mechanism of representativeness results in the generation of exemplars that
+are less similar to exemplars from other categories. We see this illustrated in
+Figure \ref{fig:representative-examples}, where the representativeness model
+displays a strong preference for generating exemplars from target category 'B'
+that are further away from the contrast category 'A'.
 
@@
 In addition to demonstrating the utility of formal modeling in the study
-creative cognition, the PACKER model more specifically offers an additional
+creative cognition, the contrast models more specifically offer an additional
 interpretation of some of the field's most central findings. For example,
@@
 identified another important aspect of the constraining influence of prior
-knowledge: What is generated cannot be the same as what is already known.
+knowledge: What is generated tends to be different to what is already known.
 Further, there is systematicity in how generated categories differ from prior
-knowledge. The results of our simulations with PACKER demonstrate that this
-influence is concisely explained in terms of a trade-off between within-category
-similarity and between-category dissimilarity.
-
-Similarly, PACKER may offer an additional interpretation of existing accounts of
-creative generation. Most notably, a leading account within the creative
-cognition literature, the Path of Least Resistance
-\citep{ward1994structured,ward1995s}, also explains generation in terms of an
-exemplar-based retrieval process. This account was designed to explain the
-creative generation of natural categories (e.g., new species of plants and
-animals) and as a result relies strongly on the hierarchical organization of
-these categories: Individuals are thought to retrieve an example of the
-higher-level category being generated (e.g., \textit{bird} may be retrieved from
-the category \textit{animal}), and then systematically alter what was retrieved
-to make something new. As the PACKER model does not assume knowledge is
-hierarchically organized \citep[this is true of the exemplar view more broadly,
-see][]{murphy2016exemplar}, the model may be viewed as a formal instantiation of
-the Path of Least Resistance for application in a traditional artificial
-categorization domain (when there is no established hierarchy of categories).
-PACKER's success in explaining generation within an artificial domain motivates
-future work exploring the nature of category contrast within a more naturalistic
-setting.
+knowledge. The results of our simulations suggest that this conceptualization of
+difference can be addressed in at least two different ways. Specifically,
+simulations with PACKER demonstrate that the constraining influence of
+difference is concisely explained in terms of a trade-off between
+within-category similarity and between-category dissimilarity. Conversely, the
+representativeness model shows that this influence can also be the result of
+enhancing the representativeness of generated exemplars to their category. [xian edit]
+
+PACKER may offer an additional interpretation of existing accounts of creative
+generation. Most notably, a leading account within the creative cognition
+literature, the Path of Least Resistance \citep{ward1994structured,ward1995s},
+also explains generation in terms of an exemplar-based retrieval process. This
+account was designed to explain the creative generation of natural categories
+(e.g., new species of plants and animals) and as a result relies strongly on the
+hierarchical organization of these categories: Individuals are thought to
+retrieve an example of the higher-level category being generated (e.g.,
+\textit{bird} may be retrieved from the category \textit{animal}), and then
+systematically alter what was retrieved to make something new. As the PACKER
+model does not assume knowledge is hierarchically organized \citep[this is true
+of the exemplar view more broadly, see][]{murphy2016exemplar}, the model may be
+viewed as a formal instantiation of the Path of Least Resistance for application
+in a traditional artificial categorization domain (when there is no established
+hierarchy of categories). PACKER's success in explaining generation within an
+artificial domain motivates future work exploring the nature of category
+contrast within a more naturalistic setting.
 
@@
 
-\subsection{Exemplar dissimilarity and Representativeness in Categorization}
+\subsection{Disentangling Exemplar dissimilarity and Representativeness in Categorization}
 
@@
 exerts more influence in category learning.
+
+In this paper, we have developed two novel models of category generation, each
+involving a different conceptualization of contrast.  However, there is no theoretical constraint that dictates the It is also possible to treat the similarity
+
+A thorough investigation to disentangle the two contrast mechanisms here is
+beyond the scope of the current article. However, as a quick proof of concept it
+is useful to check if the addition of a representativeness mechanism to an
+exemplar model would yield any benefit at all.
 

### 2018-09-03 Experiments/middle-bottomABC/webapp/cgi-bin/config.py

@@ -8,13 +8,13 @@ print                                                             # blank line, 
 # - - - - - - - - - - - - - - - - - - - - - - - -
 # USER SHOULD SET THESE PARAMETERS
 # This is where all the data is stored (X.json and workers.db) 
-destination = "../../datatemp"
+destination = "/var/services/homes/xian/CloudStation/data/middle-bottomABC"
 assignmentdb = destination + "/assignments.db"
 
 # what are the conditions?
 conditions = ['Middle', 'Bottom']
 counterbalances = range(8)
-sessionorders = ['b-bc','bc-b']
+gentypes = range(3)
 
 #  ----- DATA ORGANIZATION -----
 # Database: ~/CloudStation/data/assignments.db
@@ -40,7 +40,7 @@ if not os.path.isfile(assignmentdb):
     (Participant INTEGER,
     Condition TEXT, 
     Counterbalance INTEGER, 
-    Sessionorder TEXT,
+    Gentype INTEGER,
     Complete INTEGER)'''
     c.execute(cmd)
     conn.commit()

### 2018-09-03 Experiments/middle-bottomABC/webapp/experiment.js

@@ -12,10 +12,6 @@ function startup() {
 	load_template("html/templates/observe.html", observation);
 	load_template("html/templates/generate.html", generation);
 	load_template("html/templates/generalize.html", generalization);
-
-	//allocate first session given the assigned sessionorder condition
-	sessionorderIdx = session.sessionorders.indexOf(data.info.sessionorder)
-	session.condition = session.types[sessionorderIdx][session.count]
 	
 	// get start time
 	data.info.start = Date.now();
@@ -24,7 +20,7 @@ function startup() {
 	savedata(data)
 
 	// BEGIN EXPERIMENT
-	inserthtml(observation.instructions[session.count]);
+	inserthtml(observation.instructions);
 };
 
 // ------------------------------------------------------ //

### 2018-09-04 Manuscripts/cog-psych/revision/_region_.tex
Primary revisions to manuscript: copytweak-rep in section 8.5, unsup cat, contrast in thomas98, some mention of machine learning generative models. In other news, updated and more or less finalised fits with the category [0] bug fixed in Simulation.py. TODO: Update the goldstate.

--- 2018-09-03 Manuscripts/cog-psych/revision/cogpsych.tex
@@
-%!TEX output_directory = latex_out/
+\message{ !name(cogpsych.tex)}%!TEX output_directory = latex_out/
 \documentclass[12pt]{article}
@@
 
+\message{ !name(cogpsych.tex) !offset(-3) }
+
 % ------- TITLE PAGE ------- %
@@
 \section*{Abstract}
-\jlanote{need to update} 
-\sxlnote{updated} 
 %
@@
 regularities to other categories would be identical to a known category that is
-representative of the domain (and thus, not new at all).
-
-\sxlnote{I think we can address Charles' point on real world examples of how
-  contrast can be important around here.}
+representative of the domain (and thus, not new at all). Further, contrast has
+played a critical role studying the spread of new ideas in diverse fields, such
+as marketing \citep{berger16} and sociology \citep{rogers03}.
+
 To successfully generate something novel, what is generated must be different
@@
 The second hypothesis for how contrast might affect categorization is contrast
-as {\em representativeness}. The representativeness heuristic \citep{kahneman73}
+as {\em representativeness}. The representativeness heuristic \citep{kahneman1972subjective}
 states that people make judgments regarding an outcome based on how
@@
 opposite species.
-
-\sxlnote{Added thomas98 description here} It is worth pointing out that the
-observation of regularities in distributional properties across categories is
-not confined to the work of Ward and colleagues. Most notably, \citet{thomas98}
-found that when people were asked to generate feature values for exemplars with
-a missing feature (specifically circles of different sizes, each with a radial
-line of varying orientation), the resulting set of exemplars tended to share the
-same feature correlations as exemplars from previously learned categories. [the
-transition to the next point isn't great, though]
 
@@
 
+It is worth pointing out that the early observation of regularities in
+distributional properties across categories is not confined to the work of Ward
+and colleagues. Most notably, \citet{thomas98} trained participants to classify
+circles of different sizes, each with a radial line of varying orientation, into
+two categories. During a surprise prediction phase, participants were asked to
+generate a value for a missing feature for a certain category exemplar given a
+value on the other feature. Their results revealed that exemplars tended to
+share the same feature correlations as exemplars from previously learned
+categories. Interestingly, this effect was not consistently observed when the
+given values of a feature fell outside the range of learned values for that
+particular category. For instance, if the learned Category A exemplars were
+generally 8 cm to 10 cm in radius and negatively correlated with the angle of
+the radial line, participants generally produced radial line orientations that
+were also negatively correlated with the exemplar size when the given size was
+between 8 cm and 10 cm. However, when the given size fell outside that range for
+that category (e.g., 14 cm for a Category A exemplar), some participants
+consistently produced radial line orientations that were not negatively
+correlated with the size and that also placed the exemplars in a significantly
+different location in the feature space. These results suggest that when tasked
+to produce exemplars that fall outside what was previously learned, people show
+some tendency to generate exemplars that are different (both distributionally
+and spatially) to what was learned. We explore this idea of contrast
+more deeply in the following section.
+
 \section{And Now for Something Different: A Role For Contrast}
@@
 
-\jlanote{check to make sure the next paragraph has been updated enough}
-\sxlnote{I think it's updated enough! If given a chance I might edit this a
-  little to be less harsh on hierarchical bayesian models in general though I
-  think, since the representativeness model is one of them?}
 Although it is evident that people are {\em capable} of creating new concepts
@@
 location information and contrast cannot be captured using hierarchical Bayesian
-models. In fact, our model of contrast as representativeness is a hierarchical
-Bayesian model, but with a different goal than predicting the next category.
+models. In fact, a model instantiating contrast as representativeness is a
+hierarchical BAyesian model. But, it requires an additional factor, which in our
+case is having a different goal than predicting the next category.
 
@@
 
-\sxlnote{Charles suggested we either remove the section below, or move it to the
-  discussion. I've tried finding a good place for it in the discussion but it
-  doesn't fit very well, and moving it takes away this nice motivation for
-  our further investigation into contrast. I think that keeping it here in the
-  introduction might work out, provided Charles finds my argument for why the
-  regularity assumption is too strong convincing enough. More details a little
-  further down.}
+\subsection{Contrast as exemplar dissimilarity: The PACKER Model}
+
+\label{section:PACKER-definition}
+
+As noted above, the constraint that new concepts should differ from what is
+already known has been largely overlooked in previous work. This is no doubt in
+part due to the vague definition of what it means for a concept to be
+``different'': A generated category may be different from what is already known
+in any number of respects. Towards providing a more precise definition of the
+role of contrast in creative generation, we formalized contrast in a novel
+exemplar model, PACKER ({\em Producing Alike and Contrasting Knowledge using
+Exemplar Representations}). PACKER explains category generation as a balance
+between two fundamental constraints: The category to be generated should not be
+similar to known categories, and exemplars within each category should be
+similar to one another. These ideas are implemented within the well-studied
+exemplar framework -- the PACKER model is an extension of the influential
+Generalized Context Model of categorization
+\citep[GCM;][]{nosofsky1984choice,nosofsky1986attention}.
+
+Although, as an exemplar model, one of PACKER's proposals is people represent
+categories in terms of a collection of stored exemplars, it is worth emphasizing
+that the underlying nature of human category representations is not the primary
+concern of our work. The choice to develop PACKER within an exemplar framework
+reflects the facts that exemplar models have been throughly evaluated, are
+strongly theoretically motivated, and dominate much of the theoretical and
+empirical work in categorization. The focus of our work with PACKER concerns the
+dual constraints of within- and between-class similarity; it is not difficult to
+imagine how such constraints may be instantiated using alternative frameworks
+(e.g., \citealp{kurtz2007divergent,love2004sustain,smith2000thirty}; for a
+review of categorization models, see \citealp{pothoswills2011}).
+
+Both PACKER and the GCM simulate categorization under the assumption that
+learners represent categories as a collection of exemplars, corresponding to the
+labeled stimuli they have observed. The exemplars are encoded within a
+$k$-dimensional psychological space, and model performance is based on the
+amount of similarity between the item to be categorized and the stored
+exemplars. Similarity between two examples, $s\left(x_i, x_j\right)$, is
+computed as an inverse exponential function of distance
+\citep[following][]{attneave1950,shepard1957stimulus,shepard1987toward}:
+
+\begin{equation} s\left(x_i,x_j\right) = \exp \left\{ -c \left[\sum_{k}{ w_k
+\left| x_{ik} - x_{jk} \right|^r }\right]^{1/r} \right\}
+\label{eq:similarity}
+\end{equation}
+% 
+where $w_k$ is the attention weighting of dimension $k$ ($w_k \geq 0$ and
+$\sum_k{w_k} = 1$), accounting for the relative importance of each dimension in
+similarity calculations, and $c$ ($c>0$) is a specificity parameter controlling
+the spread of exemplar generalization. For simplicity, attention will be
+distributed uniformly in our simulations (unless otherwise noted). The value of
+$r$ depends on the nature of the experimental conditions being simulated: $r=1$
+is appropriate for separable dimensions, whereas $r=2$ is appropriate for
+integral dimensions \citep[e.g.,][]{shepard1964attention,garner1974processing}.
+In our simulations, we set $r=1$ due to the separable nature of the stimulus
+dimensions used in our experiments (see Figure \ref{fig:e1-conditions}).
+
+PACKER (as well as its name) was in part inspired by earlier work from the
+categorization literature. Specifically, \cite{hidaka2011packing} argued that
+natural categories ``pack'' the values of features such that different
+categories fill the domain space with distance between one another, while
+keeping items within the same category close together. Inspired by this idea,
+PACKER proposes that generation is constrained by both similarity to members of
+the target category (the category in which a stimulus is being generated) as
+well as similarity to members of other categories: the most desirable generation
+candidates are similar to members of the target category and not similar to
+members of contrast categories. This is achieved by aggregating similarity
+across known exemplars differently according to class membership. The aggregated
+similarity $a(y,x)$ between generation candidate $y$ and stored exemplars $x$ is
+given by:
+
+\begin{equation} a(y, x) = \sum_j{f(x_j) s(y, x_j)}
+\end{equation}
+% 
+where $f(x_j)$ is a function specifying each exemplar's contribution to
+generation. A negative value for $f(x_j)$ produces a `repelling' effect (items
+are less likely to be generated nearby $x_j$), and a positive value produces an
+`attracting' effect (items are more likely to be generated nearby $x_j$). When
+$f(x_j)=0$, the exemplar does not contribute to generation.
+
+PACKER sets $f(x_j)$ depending on exemplar $x_j$'s category membership:
+$f(x_j) = \theta_t$ if $x_j$ is a member of the target category, and
+$f(x_j) = -\theta_c$ if $x_j$ is a member of a contrast category. $\theta_t$ and
+$\theta_c$ are free parameters ($0 \leq \theta_t,\theta_c$) controlling the
+trade-off between within- and between-category similarity. For example, when
+$\theta_t = \theta_c = 0.5$, $f(x_j) = 0.5$ for members of the target category
+and $f(x_j) = -0.5$ for members of other categories; thus, the model is likely
+to generate items that are similar to members of the target category but are not
+similar to members of other categories. In this way, $\theta_t > 0$ with
+$\theta_c = 0$ produces exclusive consideration of target-category members, and
+$\theta_c > 0$ with $\theta_t = 0$ produces exclusive consideration of
+contrast-category members. The combination of $\theta_t$ and $\theta_c$
+parameters thus specifies a wide breadth of possible approaches; by fitting it
+to a dataset, one can describe the relative roles of between-category contrast
+and within-category similarity in generation. See Figure
+\ref{fig:packer-examples} for an illustration of how these parameters control
+the relative influence of within-category similarity and contrast to other
+categories when generating a new exemplar.
+
+\begin{figure}
+	\begin{center} \inputpgf{figs/}{packer-examples.pgf}
+      \caption{PACKER generation of a category `B' example, following exposure
+        to one member of category `A' and one member of category `B'.
+        Predictions are shown for three different parameterizations (differing
+        in values of $\theta_t$ and $\theta_c$): {\em (a)}
+        Predictions based on contrast similarity only. {\em (b)}
+        Predictions based on target similarity only. {\em (c)} Predictions with
+        both constraints considered.}
+		\label{fig:packer-examples}
+	\end{center}
+\end{figure}
+
+The probability that a candidate $y$ will be generated is evaluated using an
+Exponentiated \citet{luce1977choice} choice rule. Candidates with greater values
+of $a(y,x)$ are more likely to be generated than candidates with smaller values:
+% 
+\begin{equation} p(y \mid x) = \dfrac { \exp \left \{ a \left( y, x
+\right) \right \} } { \sum_i{ \exp \left \{ a \left( y_i, x \right)
+\right\} } }
+\label{eq:packer-choice}
+\end{equation}
+% 
+%where $\theta$ ($\theta \geq 0$) is a free parameter controlling response
+%determinism.
+
+It is worth noting that PACKER is only one possible exemplar-based account of
+category generation within our proposed framework. That is, PACKER places
+specific constraints on the possible values of $f(x_j)$, but other
+exemplar-based category generation models with drastically different behavior
+can be formalized in this framework by imposing alternative constraints. For
+example, as will be discussed in more detail below, PACKER is formally
+equivalent to the copy-and-tweak model proposed by \cite{jern2013probabilistic}
+when $\theta_c = 0$ and $\theta_t > 0$. Likewise, when
+$\theta_t = 0$ and $\theta_c > 0$, PACKER can represent a
+contrast-only generation mode, relying exclusively on contrast when generating
+new categories. Finally, when $f(x_j) < 0$ for all $x_j$ (regardless of class
+membership), a ``pure-packing'' approach is yielded, generating items in
+unoccupied areas of the domain. Thus, the proposed framework may be used to
+describe a wide variety of qualitatively distinct generation strategies.
+
+It is also worth noting that PACKER represents just one of many possible models
+that incorporate contrast in category generation. For example, it may be
+possible to extend \cite{jern2013probabilistic}'s hierarchical Bayesian model to
+include contrast, but this mechanism would be at odds with the standard
+hierarchical Bayesian framework. In contrast, these ideas emerge naturally from
+the exemplar view: We have not modified any of the core elements of the GCM in
+defining the PACKER model, we simply weight exemplar similarity during
+aggregation. Thus, beyond formally specifying the role of contrast in creative
+generation, the PACKER account allows us to evaluate the well-understood
+principles of the exemplar view within the comparatively understudied field of
+conceptual generation. Nonetheless, we will discuss integrating our approaches
+in the General Discussion.
+
+\subsubsection{Relation Between PACKER and Copy-And-Tweak}
+\label{section:copytweak-packer} The PACKER model is similar to the
+copy-and-tweak model reported by \cite{jern2013probabilistic}: Both models are
+exemplar-based, and both models generate new items according to their similarity
+to known members of the target class. However, PACKER diverges from the
+copy-and-tweak model by including a contrast mechanism, enabling generation
+according to dissimilarity to members of opposing categories. By consequence,
+copy-and-tweak can be realized as a parameterization of the PACKER model that is
+insensitive to category contrast. Specifically, when $\theta_c = 0$ and
+$\theta_t > 0$ (see Figure \ref{fig:packer-examples}, panel B),
+$f(x_j) = 0$ for $x_j$ belonging to contrast categories; thus, PACKER is not
+influenced by these items, and is mathematically equivalent to a copy-and-tweak
+approach.
+
+In this paper, we report simulations using this copy-and-tweak model. This model
+fits within the exemplar-based category generation framework defined above, and
+is a continuous-dimension adaptation of the model tested by
+\cite{jern2013probabilistic}. By formalizing a model family where PACKER and
+copy-and-tweak are different parameterizations of models within the same
+framework, the comparison between PACKER and copy-and-tweak provides a test of
+the explanatory value of the contrast mechanism based on exemplar dissimilarity:
+The account provided by copy-and-tweak will only equal that of PACKER if the
+contrast mechanism does not offer an advantage (i.e., if $\theta_c > 0$
+significantly improves model fits). Note that the purpose of the article is to
+explore and formally analyze the role of contrast in category generation and
+thus, we leave extending PACKER to incorporate distributional factors \citep[as
+explored by][]{jern2013probabilistic} for future work.
+
+\subsection{Contrast as representativeness: An extension to Jern and Kemp
+(2013)}
+\label{section:representativeness-definition}
+
+An alternate conceptualization of category contrast is the idea of
+representativeness -- exemplars are generated such that they are more
+representative of the target category and thus more distinct from and less
+similar to other categories. We adopt the definition specified by
+\cite{tenenbaum01}, where the representativeness $R$ of an item (or in our case,
+an exemplar) $y$ is the relative amount of evidence that is provided by $y$ for
+a given hypothesis $h$ compared to all other hypotheses $H'$. More formally,
+\begin{equation}
+  R(x,h) = \log \dfrac{p(x|h)}{\sum_{h'\in H'}{p(x|h')p(h')}} , 
+\label{eq:representativeness}
+\end{equation}
+
+In Equation \ref{eq:representativeness}, $p(h')$ is the prior probability
+distribution on the hypothesis space that excludes $h$. Note that when there are
+only two hypotheses (e.g. two categories) involved in the domain, this prior
+takes the value of $1$ since there is ever only one alternative hypothesis.
+
+This Bayesian formalization of representativeness makes it straightforward to
+implement into the existing hieriarchical Bayesian model of category generation
+developed by \cite{jern2013probabilistic}. Specifically, both models assume that
+exemplars are sampled from a given category distribution ($h$ in
+Equation \ref{eq:representativeness}). Each category distribution is a
+multivariate normal parameterized by a location vector $\mu_k$ and covariance
+matrix $\Sigma_k$ (i.e., $h = \mathcal{N}(\mu_k,\Sigma_k)$). The parameters
+$\mu_k$ and $\Sigma_k$ are assumed to be samples from a prior
+normal-inverse-Wishart (NIW) distribution parameterized by
+$\mu_0,\Sigma_0,\kappa,\nu$. Mathematically,
+\begin{align}
+  \mu_k, \Sigma_k|\mu_0,\Sigma_0,\kappa,\nu &\sim
+                                              \textrm{NIW}(\mu_0,\Sigma_0,\kappa,\nu)
+                                              \label{eq:mvnsample}\\
+  y|C &\sim \mathcal{N}(\mu_k,\Sigma_k) \label{eq:exemplarsample}
+\end{align}
+
+In our simulations, we set the prior mean $\mu_0$ to the center of the stimulus
+space and $\Sigma_0 = \lambda \mathbf{I}$ where $\lambda$ is a free parameter
+and $\mathbf{I}$ is a $d$-by-$d$ identity matrix with $d$ representing the
+number of dimensions or features in the domain. The $\kappa$ and $\nu$
+parameters are freely estimated within the constraints $\kappa > d-1$ and
+$\nu > 0$.
+
+With the assumption of the NIW prior, the location vector $\mu_k$ can be computed
+with:
+\begin{equation}
+  \mu_k = \dfrac
+    {\kappa\mu_{0} + n_k \bar{x}_k}
+    {\kappa + n_k}
+    \label{eq:category_mus}
+\end{equation}
+where $n_k$ is the number of observed exemplars in category $k$ and $\bar{x}_k$
+is the observed category mean. Note that if there are no observed exemplars in
+the target category (i.e., if the model is generating a completely novel
+category), Equation \ref{eq:category_mus} simplifies to $\mu_k = \mu_0$.
+
+The NIW prior also allows us to infer the category covariance matrix $\Sigma_k$
+by computing the following:
+\begin{equation}
+  \Sigma_k = [\Sigma_D\nu + C_k +
+    \dfrac
+    {\kappa n_k}
+    {\kappa + n_k}
+    (\bar{x_k}-\mu_k)(\bar{x_k}-\mu_k)^T
+  ] (\nu + n_k)^{-1}
+  \label{eq:Sigma_y}
+\end{equation}
+where $C_k$ is the observed category covariance and $\Sigma_D$ is the domain
+covariance matrix from which $\Sigma_k$ samples are obtained. We can
+infer $\Sigma_D$ from the observed category covariances $C$ and the prior
+covariance $\Sigma_0$. Specifically,
+\begin{equation}
+    \Sigma_D = \Sigma_0 + \sum_{k}{C_k}
+\end{equation}
+
+At this point, both the representativeness model and the hierarchical Bayesian
+model from \cite{jern2013probabilistic} are largely identical. However, both
+models diverge in their computation of exemplar generation probabilities. While
+the original hierarchical Bayesian model produces novel exemplars with
+probabilities proportional to the multivariate normal likelihood $p(y|h)$, the
+representativeness model generates new exemplars with probabilities weighted by
+their representativeness $R(y,h)$. In practice, the probability of generating a
+particular candidate $y$ is obtained using an Exponentiated
+\citet{luce1977choice} choice rule:
+\begin{equation}
+  p(y) = \dfrac
+    {\exp( \theta \cdot R(y,h))}
+    {\sum_k \exp(\theta  \cdot R(y_k,h_k))} .
+\label{eq:representative-choice}
+\end{equation}
+where $\theta$ is a freely estimated response determinism parameter (constrained
+such that $\theta \geq 0$).
+
+Despite both the original hierarchical Bayesian model and our representativeness
+model sharing identical hierarchical structures, their distinct response
+processes can yield very different exemplars. Specifically, the hierarchical
+Bayesian model emphasizes the generation of exemplars that maintain
+distributional commonalities across categories. In contrast, the
+representativeness model focuses on the generation of exemplars that are
+representative of the underlying distribution for a target category.
+\jlanote{The next sentence is weird sounding and should be cleaned up.}
+\sxlnote{Ok! I hope it reads better now.} Generally, this mechanism of
+representativeness results in the generation of exemplars that are less similar
+to exemplars from other categories. We see this illustrated in Figure
+\ref{fig:representative-examples}, where the representativeness model displays a
+strong preference for generating exemplars from target category 'B' that are
+further away from the contrast category 'A'.
+
+\begin{figure}
+	\begin{center} \inputpgf{figs/}{hbm-examples.pgf}
+		\caption{Generation of a category `B' example, following exposure
+to one member of category `A' and one member of category `B'. Predictions are
+shown for {\em(a)} the original hierarchical Bayesian model and {\em(b)} the
+representativeness model.}
+		\label{fig:representative-examples}
+	\end{center}
+\end{figure}
+
+\subsection{Synopsis and Prognosis}
+
+Research on the creative generation of novel concepts has focused on the finding
+that generated categories tend to possess distributional commonalities with
+known categories. However, a fundamental goal of concept generation is to create
+something {\em new} (i.e., different from what is already known). The manner in
+which generated categories differ from known ones is, nonetheless, poorly
+understood: Existing theories do not make strong predictions about how
+creatively generated concepts should systematically differ from existing ones.
+Above, we found encouraging initial support that contrast influences category
+generation \citep[][Experiment 3]{jern2013probabilistic}, and we introduced two
+novel approaches to formalizing the roles of similarity and contrast in creative
+generation.
+
+In the sections below, we present two experiments demonstrating systematic
+effects of category contrast on category generation inspired by factors
+influencing how the contrast models generate new categories. Our experiments are
+based on \cite{jern2013probabilistic}'s paradigm, which is a straightforward
+translation of the traditional artificial classification paradigm to the task of
+category generation: Participants are first exposed to a single,
+experimenter-defined category, and are then asked to generate members of a new
+category. We then report formal analyses comparing the exemplar dissimilarity
+and the representativeness accounts of our results to that of the hierarchical
+Bayesian and copy-and-tweak models developed by \cite{jern2013probabilistic}.
+After analyzing the category generation studies, we present third experiment to
+demonstrate the effect of contrast in category learning. This ensures that our
+results are not idiosyncratic to category generation.
+
+\section{Experiment 1}
+
+To begin our investigation, we sought to extend the early evidence obtained from
+our analysis of the \cite{jern2013probabilistic} data, under a variety of
+learning conditions and using more standard stimulus materials. We used an
+artificial stimulus design: A two dimensional domain of squares, varying in
+color and size (see Figure \ref{fig:e1-conditions}, panel A). These dimensions
+have been used in numerous classification learning studies
+\cite[e.g.,][]{conaway2016similar,conaway2016generalization,shepard1961learning,nosofsky1994comparing}.
+Unlike those used in the \cite{jern2013probabilistic} experiments, distance on
+these physical dimensions aligns more directly with perceptual similarity,
+allowing us to evaluate the role of category contrast in creative generation
+more precisely. To extend the evidence provided by the
+\cite{jern2013probabilistic} data, we tested the effects of category contrast
+after learning one category from a set of qualitatively distinct category
+structures, as shown in Figure \ref{fig:e1-conditions}.
+
+Figures \ref{fig:e1-conditions}b-d show the values of exemplar dimensions
+belonging to the experimenter-defined categories (`A', or `Alpha') that
+participants were assigned to learn about prior to generating a new category.
+Each participant learned one of the category types during training. In the
+`Cluster' type, category A is a tight cluster of examples in the space.
+Perceptually instantiated, the members of category A might, for example, be
+large and dark in color. In the `Row' type, category A has a row pattern across
+the space, varying along one feature but not the other. Thus, its members might
+all be dark in color but would vary in size. Finally, in the `XOR' type, the
+experimenter-defined category consists of two clusters separated in opposite
+corners of the space, conforming to the exclusive-or logical structure (e.g.,
+members are small and dark or large and light).
+
+It should be noted that in our experiments the assignment between the perceptual
+and conceptual dimensions (e.g., $X \rightarrow Size$, $Y \rightarrow Color$),
+as well as the direction of variation along each dimension (e.g., $dark
+\rightarrow light$ or $light \rightarrow dark$) was counterbalanced across
+participants. The category types in Figure \ref{fig:e1-conditions} are plotted
+in a conceptual space, rather than a perceptual space. Thus, while the
+conceptual organization of the category types remains constant, each category
+type may have a different physical instantiation according to the counterbalance
+assignment. For example, the Cluster type may be large and dark in color, or it
+may be small and light in color, depending on the assignment and direction of
+the dimensions. For this reason, below we will discuss generation within a
+conceptual space, rather than a physically instantiated one.
+
+\begin{figure}
+    \begin{center} \inputpgf{figs/}{e1-conditions.pgf}
+    \caption{Stimulus domain and category types tested in Experiment 1. Stimuli
+are not drawn to scale.}
+    \label{fig:e1-conditions}
+    \end{center}
+\end{figure}
+
+After learning about an experimenter-defined category, participants are asked to
+generate examples of a new category. Within this paradigm, an effect of category
+contrast would be realized if participants prefer to generate items in locations
+that are distant (i.e., perceptually dissimilar) from members of category A.
+However, generation is left unconstrained. Critically, participants were not
+asked to generate something different in the prompt. For example, participants
+assigned to the Cluster condition may generate a tightly clustered category in
+the corner opposite of the experimenter-defined category. Alternatively, they
+may generate a tightly clustered category directly overlapping with the
+experimenter-defined category. Further, they may even generate an entirely
+different type of category (e.g., a row category).
+
+Our experimental results also provide a converging test of the classic finding
+that generated categories tend to share distributional properties with known
+categories in the domain \citep{jern2013probabilistic,ward1994structured}. From
+these results, we can predict that, in each condition, participants should
+generate categories that are distributionally similar to the
+experimenter-defined category: In the Cluster condition, generated categories
+should be tightly clustered. In the Row condition, generated categories should
+vary more along the X-axis than the Y-axis. In XOR condition, generated
+categories should be widely distributed across both dimensions, and the two
+dimensions should be positively correlated.
+
+Interestingly, the XOR condition also offers a dissociation between the roles of
+category contrast and the emulation of distributional structure:
+widely-distributed, positively-correlated categories would need to lie along the
+positive diagonal of the space (that is the only place they ``fit''), which is
+already occupied by the experimenter-defined category. Thus, if contrast plays a
+role, exemplars in the generated categories of participants in the XOR condition
+may not be positively correlated -- they may be negatively correlated instead.
+In this case, contrast and statistical regularities would interact, which would
+be inconsistent with the leading approach in conceptual generation
+\citep{jern2013probabilistic}
+
+\subsection{Participants and Materials}
+
+183 participants were recruited from Amazon Mechanical Turk. Each participant
+was randomly assigned to one condition: 64 participants were assigned to the
+Cluster condition, 61 were assigned to the Row condition, and 58 were assigned
+to the XOR condition (sample sizes differ due to random assignment). Stimuli
+were squares varying in color (grayscale 9.8\%--90.2\%) and side length
+(3.0--5.8cm), see Figure \ref{fig:e1-conditions}. The assignment of perceptual
+features (color, size) to axes of the domain space (x, y), as well as the
+direction of variation along each axis (e.g., $dark \rightarrow light$ or $light
+\rightarrow dark$) was counterbalanced across participants.
+
+\subsection{Procedure}
+
+As noted in the introduction of this paper, the task of generating members of a
+new category can be aligned with common tasks studied in the categorization
+literature: Whereas classification consists of predicting an object's category
+label on the basis of its features, inference consists of predicting an observed
+feature, given a set of observed features and a category label. Generation thus
+consists of predicting {\em all} features of an object, given a novel category
+label. In keeping with this insight, we designed our generation task as an
+extension of the traditional artificial classification learning paradigm. The
+task differs from traditional work in creative cognition primarily through the
+use of an artificial domain, which enables the application of computational
+models. The use of an artificial domain also requires the addition of a training
+phase, during which participants learn about the categories in the domain. As a
+result, unlike most previous studies \citep[e.g.,][]{ward1994structured},
+participants in our studies have no experience with the domain before the start
+of the experiment, and the experimenter-defined categories are not
+hierarchically structured (as are many natural categories).
+
+Participants began the experiment with a short training phase (3 blocks of 4
+trials), where they observed exemplars belonging to the `Alpha' category.
+Participants were instructed to learn as much as they can about the Alpha
+category, and that they would answer a series of test questions afterwards. On
+each trial, a single Alpha category exemplar was presented, and participants
+were given as much time as they desired to observe it before moving on to the
+next trial. Each block consisted of a single presentation of each of the members
+of the Alpha category, in a random order. Participants were shown the range of
+possible colors and sizes prior to training.
+
+Following the training phase, participants were asked to generate four examples
+belonging to another category called `Beta'. As in
+\citet{jern2013probabilistic}, generation was completed using a sliding-scale
+interface. Two scales controlled the values of the two dimensions (color, size)
+for the generated example. An on-screen preview of the example updated whenever
+one of the features was changed. Participants could generate any example along
+an evenly-spaced 9x9 grid (including members of the Alpha category), except for
+any previously generated Beta exemplars. Neither the members of the Alpha
+category nor the previously generated Beta examples were visible during
+generation. Prior to beginning the generation phase, participants read the
+following instructions:
+
+\begin{displayquote} As it turns out, there is another category of geometric
+figures called ``Beta''. Instead of showing you examples of the Beta category,
+we would like to know what you think is likely to be in the Beta category.
+
+You will now be given the chance to create examples of any size or color in
+order to show what you expect about the Beta category. You will be asked to
+produce 4 Beta examples - they can be quite similar or quite different to each
+other, depending on what you think makes the most sense for the category.
+
+Each example needs to be unique, but the computer will let you know if you
+accidentally create a repeat.
+\end{displayquote}
+
+\subsection{Results}
+
+We observed a substantial degree of individual differences in our data. In
+Figure \ref{fig:e1-samples} we have plotted sample data from several
+participants, from which it is evident that different participants generated
+qualitatively different category structures. In this section we will focus on
+analyzing the data in aggregate, but in later sections we will explore how
+individual differences can be explained.
+
+\begin{figure}
+    \begin{center} \inputpgf{figs/}{e1-samples.pgf}
+    \caption{Sample categories generated by participants in Experiment 1.
+Representative samples from common generation profiles are shown.}
+    \label{fig:e1-samples}
+    \end{center}
+\end{figure}
+
+To evaluate the role of contrast, we computed the number of times each stimulus
+was generated, as a function of its average city-block distance from members of
+the experimenter-defined ``Alpha'' category. These data, shown in Figure
+\ref{fig:e1-distanceplots}, reveal a clear pattern: Examples that are more
+distant from members of the experimenter-defined categories are more likely to
+be generated into a new category. This supports the notion that contrast is a
+fundamental constraint on creative cognition and that suggests that statistical
+regularity alone is insufficient.
+
+Figure \ref{fig:e1-distanceplots} also depicts, for each participant, the
+average distance of members within the generated category ({\em within-category}
+distance) against the average distance between members of the generated and
+experimenter-defined category ({\em between-category} distance). The narrow
+distribution of between-category distances in the XOR condition reflects the
+widely distributed nature of the experimenter-defined category, reducing the
+possible distances to members of the participant-generated category. These data
+reveal a systematic pattern: The majority of participants generated categories
+with greater between-category distance than within-category distance. That is,
+members of the generated category tended to be more similar to one another than
+to members of the experimenter-defined category. To evaluate this claim
+quantitatively, we conducted t-tests comparing the amount of within- and
+between- class distance in each condition. All conditions possessed greater
+between-category distance: Cluster, $t(63) = 11.43$, $p < 0.001$; Row, $t(60) =
+13.16$, $p < 0.001$; and XOR, $t(57) = 3.64$, $p < 0.001$. These results provide
+further evidence of an effect of category contrast: Participants prefer to
+generate categories that are dissimilar to the learned category but maintain
+some level of internal cohesion.
+
+\begin{figure}
+    \begin{center} \inputpgf{figs/}{e1-distanceplots.pgf}
+    \caption{Experiment 1 results. {\em Left}: Frequency of generation as a
+function of distance from members of the experimenter-defined category. {\em
+Right}: Scatter plot of within-category versus between-category distance in each
+of the participant-generated categories.}
+    \label{fig:e1-distanceplots}
+    \end{center}
+\end{figure}
+
+A secondary goal of this experiment was to examine whether we replicate the
+classic result that generated categories often possess the same distributional
+properties as previously-known categories. For each generated category, we
+computed the category range along each axis (X, Y), as well as the correlation
+between features. These data, shown in Figure \ref{fig:e1-statsboxes}, reveal
+broad individual differences: Within each condition, participants generated
+categories spanning the entire X- and Y- axis as well as categories that spanned
+very little along each. Likewise, in each condition participants generated
+categories possessing strongly positive, neutral, and strongly negative
+correlations between the dimensions. Comparing the distributional statistics
+between conditions yields a broad yet, as we will see, misleading replication of
+the classic effect.
+
+With respect to ranges along each axis (X, Y), the generated categories from
+each condition tend to reflect the ranges of the experimenter-defined
+categories. The categories generated in the Cluster condition were less widely
+distributed along the X-axis compared to Row, $t(123) = 5.61$, $p < 0.001$, and
+XOR, $t(120) = 2.68$, $p = 0.008$. Categories generated in the XOR condition
+were also less widely distributed along the X-axis compared to Row, $t(117) =
+2.56$, $p = 0.012$. This latter effect was not expected because the
+experimenter-defined categories for XOR and Row had similar X-ranges. However,
+the key finding is that categories from the Cluster condition tended to be more
+tightly distributed along the X-axis.
+
+\begin{figure}
+    \begin{center} \inputpgf{figs/}{e1-statsboxes.pgf}
+    \caption{Box-plots of the distributional statistics from the categories
+generated in Experiment 1. Boxes depict the median and quartiles of each
+condition, with whiskers placed at 1.5 IQR. All points outside this region are
+marked individually.}
+    \label{fig:e1-statsboxes}
+    \end{center}
+\end{figure}
+
+Likewise, categories generated in the Row condition had less Y-axis range
+compared to Cluster, $t(123) = 4.57$, $p < 0.001$ and XOR, $t(117) = 9.26$, $p <
+0.001$, and categories from the Cluster condition had less Y-axis range compared
+to XOR, $t(120) = 3.95$, $p < 0.001$. As expected, the correlations in the
+Cluster and Row conditions were not systematically positive or negative ($ps >
+0.1$). However, the generated categories in the XOR condition tended to possess
+{\em negatively} correlated dimensions, $t(57) = 2.04$, $p = 0.046$. This
+finding is notable, as it is the opposite of what would be expected, assuming
+learners are emulating the distributional structure of the experimenter-defined
+class (which possesses perfectly positively correlated features).
+
+We believe that the failure to replicate this finding is because participants in
+\cite{jern2013probabilistic} could differentiate the generated category on a
+third dimension (hue) to maintain the statistical regularities on the other two
+dimensions. Although the correlation in the XOR condition is significantly
+negative, it is clear from the box-plot in Figure~\ref{fig:e1-statsboxes} that
+it would be inappropriate to make a strong conclusion (e.g., the median is close
+to zero). However, we can conclude with confidence that there are situations
+where people do not emulate the distributional structure of the given category.
+This indicates that there is more to category generation than the emulation of
+distributional structure of other categories in the domain. Further, as we will
+discuss in more detail in the model-based analysis section, this is expected by
+our proposal that contrast is a fundamental principle in category generation.
+
+\subsection{Discussion} In Experiment 1 we sought to extend our analysis of the
+\cite{jern2013probabilistic} data by evaluating the influence of category
+contrast on creative generation, given qualitatively different types of prior
+knowledge. We found strong evidence for effects of category contrast in each
+condition: Participants were more likely to generate stimuli that are more
+distant from (i.e., less similar to) members of a previously-learned category,
+and members of participant-generated categories tended to be more similar to one
+another than to members of previously-learned categories. We also partially
+replicated the classic finding that the distributional structure of generated
+categories reflects that of previously learned categories
+\citep{jern2013probabilistic,ward1994structured}: Members of generated
+categories were more widely distributed along dimensions which were widely
+distributed in the experimenter-defined category.
+
+Notably, however, we also found that participants who learned an XOR category
+(composed of exemplars following a positive diagonal, see Figure
+\ref{fig:e1-conditions}) tended to generate items according to a {\em negative}
+feature correlation -- the opposite of what was present in the previously
+learned category. While this may be difficult to account for under existing
+theoretical approaches (which assume generated categories follow the same
+distributional structure as known categories), it can be concisely explained
+from a category contrast perspective. Specifically, within the XOR condition,
+individuals who seek to generate a category that is perceptually distinct from
+what is already known are left with only the upper-left and bottom-right
+quadrants of the space, as members of the previously-learned XOR category lie in
+the bottom-left and top-right. If examples are generated into both of the
+available quadrants, the generated category will possess a strongly negative
+correlation, opposing that of the experimenter-defined class.
+
+Thus, the core results of Experiment 1 indicate that generated categories can
+systematically differ from what we would expect based on prior work. The
+negative (or null) correlations observed in the XOR condition suggests an
+interesting interaction between contrast with a given category and emulation of
+statistical properties. That is, the constraints on creative generation imposed
+by category contrast may not simply influence the {\em location} of generated
+categories, but also their distributional structure. In Experiment 2, we test
+this claim more systematically.
+
+\section{Experiment 2.}
+
+\begin{figure}
+    \begin{center} \input{figs/e2-conditions.pgf}
+    \caption{Category types tested in Experiment 2.}
+    \label{fig:e2-conditions}
+    \end{center}
+\end{figure}
+
+To test whether category contrast influences the distributional structure of
+generated categories, we sought to identify conditions in which differences in
+the distributional structure of generated categories cannot be explained by the
+distributional structure of the experimenter-defined category. We created two
+new category types (depicted in Figure \ref{fig:e2-conditions}) that possess
+identical distributional structures (both are tight clusters of examples with no
+correlation between features), as they only differ in their Y-axis position: the
+`Bottom' category lies in the bottom-center of the space, and the `Middle'
+category lies in the center. The distributional equality of these conditions is
+key to the design of the experiment: If the distributional structure of
+previously learned categories were the only influence on the generated
+categories, we should observe no difference in the categories participants
+generate between these two conditions. Will participants distribute their
+generated category differently between conditions due to the differences in the
+available stimulus space for generating a new category?
+
+If category contrast influences the distributional structure of the categories
+people generate, then we should observe different types of categories according
+to the shape of the space that is {\em unoccupied} by members of previously
+learned categories. The difference in the Y-axis position between the Bottom and
+Middle conditions produces a considerable change to the shape of the unoccupied
+space. Participants assigned to learn the Bottom category should be less likely
+to generate exemplars into the lower regions of the stimulus space (as these
+areas possess greater similarity to members of the Bottom category), preferring
+instead to distribute exemplars across the upper region of the space. This
+constraint is lifted in the Middle condition, as the Middle category exemplars
+are equidistant to the upper and lower regions of the space. Accordingly,
+participants should be more likely to utilize both of these areas. Thus, if
+category contrast influences the distributional structure of generated
+categories, we should observe more participants in the Middle condition that
+generate examples above {\em and} below the experimenter-defined category.
+
+\subsection{Participants, Materials, and Procedure}
+
+122 participants were recruited from Amazon Mechanical Turk. 61 participants
+were randomly assigned to the Middle and Bottom conditions each. The stimulus
+space and procedure were exactly as in Experiment 1. Participants first
+completed a short training phase, followed by the generation phase. The only
+difference from Experiment 1 was the category types given to participants.
+
+\subsection{Results}
+
+\begin{figure}
+    \begin{center} \inputpgf{figs/}{e2-samples.pgf}
+    \caption{Sample categories generated in Experiment 2. }
+    \label{fig:e2-samples}
+    \end{center}
+\end{figure}
+
+As in Experiment 1, we observed broad differences in the generation approach
+taken by different participants. To characterize the nature of these
+differences, Figure \ref{fig:e2-samples} depicts sample categories generated by
+participants. The data from each condition are organized into four columns based
+on commonly observed patterns of generation: a `Cluster' type of
+tightly-clustered examples, `Row' and `Column' types of exemplars widely
+distributed along the one axis but narrowly along the other, and a `Corners'
+type, wherein participants placed exemplars in disparate corners of the space.
+As before, in this section we focus on analyzing the data in aggregate, but in
+later sections we will focus more specifically on explaining the individual
+differences.
+
+\begin{figure}
+    \begin{center} \inputpgf{figs/}{e2-distanceplots.pgf}
+    \caption{Experiment 2 results. {\em Left}: Frequency of generation as a
+function of distance from members of the experimenter-defined category. {\em
+Right}: Scatter plot of within-category versus between-category distance in each
+of the participant-generated categories.}
+    \label{fig:e2-distanceplots}
+    \end{center}
+\end{figure}
+
+We began our analysis by testing for the broad influence of category contrast on
+generation. As in Experiment 1, we computed the frequency each stimulus was
+generated as a function of its average distance from members of the
+experimenter-defined category, as well as each participant's average within- and
+between- category distance. These data, shown in Figure
+\ref{fig:e2-distanceplots}, yield very similar results. Participants generated
+stimuli that are distant from members of the experimenter-defined category, and
+the categories in each condition tended to possess more between-category than
+within-category distance: Bottom, $t(60) = 5.5$, $p < 0.001$; Middle, $t(60) =
+2.71$, $p < 0.01$. We did, however, observe a notable subgroup of participants
+in each condition who generated categories with more within-category than
+between-category distance. Upon manual inspection, many of these individuals
+appear to have assumed a `Corners' strategy, placing exemplars in disparate
+corners of the space, thus producing much more within-category distance, see
+Figure \ref{fig:e2-samples} for examples.
+
+To explore the distributional structure of the generated categories, we computed
+the range of exemplars along each axis (X, Y), as well as the correlation
+between features. These data, shown in Figure \ref{fig:e2-statsboxes}, again
+demonstrate the degree of individual differences observed in our study. In each
+condition, we observed tightly clustered and widely distributed categories along
+each dimension, as well as positively, negatively, and uncorrelated categories.
+
+\begin{figure}
+    \begin{center} \inputpgf{figs/}{e2-statsboxes.pgf}
+    \caption{Box-plots of the distributional statistics from the categories
+generated in Experiment 2. }
+    \label{fig:e2-statsboxes}
+    \end{center}
+\end{figure}
+
+As noted above, if the distributional structure of generated categories is
+influenced by the shape of the space not occupied by members of known
+categories, then participants in the Middle condition would be more likely to
+place exemplars in the upper {\em and} lower regions of the space, as members of
+the experimenter-defined category are equidistant from these regions.
+Participants in the Bottom condition should be less likely to generate category
+members in the bottom regions because members of the experimenter-defined
+category are located there. One way to test these predictions is to analyze the
+Y-axis ranges of the generated categories: If Middle participants utilize the
+upper and lower regions of the space, their categories should vary more along
+the Y-axis. T-Tests comparing the conditions on the distributional statistics,
+however, reveal few between-group differences: the conditions do not differ with
+respect to X-axis range, Y-axis range, or feature correlations ($ps>0.17$).
+
+However, our ability to detect differences in Y-Axis range using a standard
+$t$-test between the conditions is, in this case, diminished due to the
+non-normality of the data (Shapiro-Wilk normality test $W=0.77, p< 0.001$ for
+the Middle condition and $W=0.85, p < 0.001$ for the Bottom condition). Figure
+\ref{fig:e2-yranges} depicts the Y-axis position of the exemplars generated
+within each participant's category. The categories are sorted by overall range,
+then by condition assignment. These data reveal that there were nearly as many
+participants who generated categories spanning the entire Y-axis as those who
+generated categories spanning almost none of the Y-axis. The non-normality of
+the Y-axis range distributions thus requires that we use a different approach to
+addressing the experiment's main question.
+
+\begin{figure}
+    \begin{center} \inputpgf{figs/}{e2-yranges.pgf}
+    \caption{Y-Axis range and position of the participant-generated categories
+from Experiment 2. Each line corresponds to a participant's category, with
+notches corresponding to the Y-axis position of exemplars within the category
+(notches may overlap). Participants are sorted by overall range, and then by
+condition. }
+    \label{fig:e2-yranges}
+    \end{center}
+\end{figure}
+
+Because our main prediction concerns the generation of exemplars within the
+upper and lower regions of the domain, we compared the conditions in terms of
+the frequency with which participants generated examples above and below the
+categories. Specifically, we counted the number of participants in each
+condition who placed at least one `Beta' exemplar on the top and bottom `rows'
+of the space (the maximum and minimum possible Y-axis value, respectively). The
+resulting contingencies data are shown in Table \ref{table:e2-subset-table}.
+
+Firstly, it should be noted that nearly every participant utilized the top
+and/or bottom rows: only $10 / 122$ participants generated their category
+entirely within the interior region. Fisher's Exact Tests comparing the
+conditions reveal that more Middle participants generated an exemplar in the
+bottom row, $p < 0.001$, again demonstrating the role of contrast in guiding
+where exemplars are generated. The conditions did not differ in use of the top
+of the space, $p = 0.16$, however, more Middle participants placed exemplars in
+the top {\em and} bottom rows, $p = 0.038$. The latter effect is of interest
+here, as it indicates that the shape of the unoccupied space exerts some
+influence on the distributional structure of generated categories: Participants
+in the Middle condition were more likely to generate a category spanning the
+entire Y-axis. Thus, distributional structure of the generated categories can be
+influenced by category contrast alone.
+
+\begin{table}
+\begin{center}
+\caption{Experiment 2 results.}
+\label{table:e2-subset-table} \vskip 0.12in
+\begin{tabular}{ l r r} \textbf{Middle} & Used top row & No top row \\ \hline
+Used bottom row & 28 & 18 \\ No bottom row & 11 & 4 \\ \\ \textbf{Bottom} & Used
+top row & No top row \\ \hline Used bottom row & 16 & 8 \\ No bottom row & 31 &
+6 \\
+\end{tabular}
+\end{center}
+\end{table}
+
+\subsection{Discussion}
+
+In Experiment 2, we replicated the core findings from Experiment 1. Stimuli are
+more likely to be generated if they are distant from exemplars in other
+categories, and most participants generate categories with more between-category
+than within-category distance. However, we additionally found that the {\em
+position} of a previously learned category (rather than its distributional
+structure) influences the types of categories people generate: Participants who
+learned the `Middle' type were more likely to generate categories spanning the
+entire Y-axis of the space. Participants who learned the `Bottom' type were less
+likely to do so as a result of the presence of opposite category exemplars in
+the lower regions of the space.
+
+This finding cannot be explained from the perspective that the distributional
+structure of previously learned categories is the sole determinant of the
+distributional structure of generated categories. However, the observed behavior
+is expected from a category contrast perspective: Participants seeking to
+generate a perceptually distinct category will be more likely to use areas of
+space that are unoccupied by exemplars belonging to previously learned
+categories. In the Middle condition, the upper and lower regions of space are
+equidistant from members of the experimenter-defined category, whereas in the
+Bottom condition, the lower region of the space is closer to members of the
+experimenter-defined category. Thus, while Middle participants may form
+categories around the use of the equally unoccupied areas, the same is not true
+for the Bottom condition.
+
+\section{Model-based Analyses}
+Experiments 1 and 2 revealed systematic and strong effects of category contrast
+on creative generation. In this section, we report the results of simulations
+with formal models aimed at explaining our observations. Specifically, we
+present simulations from the two novel contrast models: PACKER and the
+representativeness model, and compare them to two models that do not incorporate
+contrast: the copy-and-tweak model (discussed in Section
+\ref{section:copytweak-packer}) and an implementation of the hierarchical
+Bayesian model proposed by \cite{jern2013probabilistic}, described in-depth in
+Appendix \ref{ap:hsampling-definition}. For our simulations here, the
+copy-and-tweak model is defined as a variant of PACKER with the $\theta_c$
+parameter constrained to be zero. The comparison of this set of models serves to
+highlight the explanatory role of contrast in categorization: If contrast
+affords little explanatory advantage, then there should be few differences in
+performance between PACKER and copy-and-tweak, or between the representativeness
+and hierarchical Bayesian model. The comparison between these models can also
+emphasize the necessity of contrast and demonstrate that generation cannot be
+explained entirely through the emulation of distributional structure. Each model
+has complementary strengths and weaknesses: Whereas PACKER and copy-and-tweak
+are insensitive to the distributional structure of learned categories (relying
+only on exemplar similarities), the representativeness and hierarchical Bayesian
+model generates categories exclusively on the basis of knowledge of how existing
+classes are distributed.
+
+Our approach in this section is to first broadly evaluate and compare the
+quality of each model's account to our entire dataset (Experiments 1 and 2
+combined), then analyze the ability for each model to explain individual
+differences in each experiment, and lastly we describe the strengths and
+weakness of each model's account of category generation.
+
+\subsection{Parameter-Fitting}
+
+To obtain a global measure of the quality of each model's account, we fitted the
+parameters of each model to our entire dataset (Experiments 1 and 2 combined),
+using a hill-climbing algorithm which maximized the log-likelihood of the
+model's predictions of the observed responses (1220 responses from 305 total
+participants). We fitted three parameters in the PACKER model ($c$, $\theta_t$,
+and $\theta_c$; see Section \ref{section:PACKER-definition}), as well as four in
+the representativeness model and the hierarchical Bayesian model ($\kappa$,
+$\lambda$, $\nu$, and $\theta$; see Section
+\ref{section:representativeness-definition} and Appendix
+\ref{ap:hsampling-definition} respectively). We fitted only two parameters for
+the copy-and-tweak model ($c$, and $\theta_t$), as $\theta_c$ is held constant
+($\theta_c = 0$). Attention ($w$, see Equation \ref{eq:similarity}) in PACKER
+and copy-and-tweak was set uniformly. Parameters were not allowed to vary
+between participants or conditions -- the goal was to obtain the best-fitting
+values to our entire dataset.
+
+\begin{table} \centering
+\caption{Results of model-fitting to the combined datasets from Experiments 1
+and 2. Note that smaller AIC values correspond to better model fits (adjusted
+for number of parameters)}
+\label{table:global-model-fits}
+\begin{tabular}{ l l l l} \\ \textbf{PACKER} & \textbf{Copy \& Tweak} &
+\textbf{Representativeness} & \textbf{Hierarchical Bayesian} \\ \hline $AIC =
+  9095$ & $AIC = 9842$ & $AIC = 8878$ &  $AIC =
+9912$ \\ $L = -4545$ & $L = -4919$ & $L = -4435$ & $L = -4952$ \\ $c = 0.483$ &
+                                                                                $c = 3.197$ & $\kappa = 11.232$ &
+$\kappa < 0.001$ \\ $\theta_c = 3.16$ & $\theta_c = 0$ (fixed) & $\nu = 2.844$ & $\nu = 5.597$ \\
+$\theta_t = 3.494$ & $\theta_t = 2.972$ & $\lambda = 1.917$& $\lambda
+                                                                       = 0.055$
+  \\ & &$\theta = 5.885 $& $\theta = 3.175$
+\\
+\end{tabular}
+\end{table}
+
+Table \ref{table:global-model-fits} contains the results of this fitting
+procedure. Due to the uneven number of fitted parameters among the models, we
+compare the model fits using the Akaike Information Criterion
+\citep[AIC;][]{akaike1974new}, where smaller values correspond to better fits
+(discounted by model complexity as measured by the number of parameters). The
+same qualitative results were obtained with alternative model comparison metrics
+(e.g., BIC, \citealp{schwarz1978estimating}; $AIC_C$,
+\citealp{hurvich1989regression}).
+
+Table \ref{table:global-model-fits} contains the AIC values of each model's best
+fit, as well as the corresponding log-likelihood ($L$) and the best-fitting
+parameter values. These results reveal strong model differentiation: both
+contrast models (PACKER and the representativeness model) achieved far better
+fits compared to their non-contrast counterparts: copy-and-tweak and the
+hierarchical Bayesian model respectively. Interestingly between the contrast
+models, the hierarchical model (representativeness) outperformed the
+exemplar-based theory (PACKER), whereas between the non-contrast models, the
+reverse is observed. Specifically, here the exemplar-based model
+(copy-and-tweak) performed somewhat better than the hierarchical Bayesian model.
+
+While PACKER's advantage may tentatively be attributed to the model's
+sensitivity to category contrast (this will be explored in detail below), the
+advantage shown by copy-and-tweak over the hierarchical Bayesian model may be
+attributed to its exemplar-based representation, as opposed to the
+prototype-based representation assumed by the hierarchical Bayesian model. As
+observed in Figures \ref{fig:e1-samples} and \ref{fig:e2-samples}, the generated
+categories we observed were often widely distributed, with no items near the
+category prototype. This aspect of the data is inconsistent with the
+multivariate normal distributions (similar to prototypes) used to represent
+categories in the \cite{jern2013probabilistic} model, but can be easily
+accounted for using an exemplar-based approach.
+
+A key distinction between the contrast and classical models is that only the
+contrast models are capable of making strong predictions about the location of
+new category members when the target class is entirely novel (i.e., no member of
+the category has been observed). Under these circumstances, there are no
+examples to copy, and thus the copy-and-tweak model predicts that items are
+generated at random. Likewise, with no observations on which to condition the
+category distribution, the hierarchical Bayesian model also picks an item at
+random.
+
+Thus, it is possible that the failure of the classical models is simply due to
+their inability to explain each participant's first trial (generating the first
+item in the `Beta' category). To ensure this is not driving our results, we
+conducted an identical set of simulations as above, excluding this trial
+(leaving 915 responses in the dataset): Again, the representativeness model ($L
+= -3286$, $AIC = 6580$) and PACKER ($L = -3377$,
+$AIC = 6759$) achieved better fits than the copy-and-tweak ($L = -3564$,
+$AIC = 7132$) and hierarchical Bayesian ($L = -3597$, $AIC = 7201$) models.
+
+Finally, because copy-and-tweak is nested within PACKER, we can use a likelihood
+ratio test to compare the two models. PACKER explains the aggregate data
+significantly better than copy-and-tweak ($\chi^2(1) = 747, p < 0.001$ for all
+data and $\chi^2(1) = 375, p < 0.001$ excluding the first example), providing
+further evidence that category generation is better explained when contrast is
+considered.
+
+Through comparison with the copy-and-tweak model, Figure
+\ref{fig:packer-loglike} more clearly demonstrates the robustness of the
+explanatory gains yielded by PACKER's category contrast mechanism. It displays
+the log-likelihood of the participants' results under PACKER as a function of
+the $\theta_c$ parameter. The model's other parameters ($c$,
+$\theta_t$) were set according to copy-and-tweak's best fits from Table
+\ref{table:global-model-fits}, and thus when $\theta_c=0$, the models are
+equivalent. The figure clearly shows a ``sweet spot'': a convex region in which
+PACKER achieves superior fits as a result of changes to $\theta_c$. The
+best fitting values lie well above the value of 0 assumed by the copy-and-tweak
+model, which demonstrates the robustness of the contrast effect (though note
+PACKER achieves even better fits when its parameters are fitted together, as in
+Table \ref{table:global-model-fits}). In sum, the data are better explained when
+both within-category similarity and category contrast is considered.
+
+\begin{figure}
+    \begin{center} \inputpgf{figs/}{packer-loglike.pgf}
+      \caption{PACKER's fit as a function of its prioritization of
+        within-category and between-category similarity (using the
+        $\theta_c$ parameter) . To facilitate comparison, PACKER's other
+        parameters ($c$, $\theta_t$) were set to the best fitting values
+        obtained for copy-and-tweak in Table \ref{table:global-model-fits}. }
+    \label{fig:packer-loglike}
+    \end{center}
+\end{figure}
+
+\subsection{Individual Differences}
+\label{section:individual-diffs}
+
+As noted in Experiments 1 and 2, we observed a great deal of individual
+differences in the types of categories that participants generated. Within each
+condition, there were a wide variety of category types, such as row and column
+categories (see Figures \ref{fig:e1-samples} and \ref{fig:e2-samples}). The
+simulations reported above serve to evaluate the models while considering the
+entire dataset, but a secondary goal of any formal account should be to provide
+some explanation of how different profiles of performance emerge. Many of the
+individual generation profiles we observed can be described within the PACKER
+framework, simply by tuning the model's parameters in a principled manner. In
+this section, we describe more specifically how the most frequently observed
+profiles can be realized.
+
+By manual inspection, it is evident that the most common profiles of generation
+consist of: (A) a tightly-distributed `cluster' of examples, (B) `row'- and
+`column'-like arrangements (varying widely along one dimension but not the
+other), and (C) a `corners' arrangement with examples placed into disparate
+corners of the space. These four profiles are distinct in terms of the
+distribution of the generated category along each dimension: Whereas the cluster
+profile is tightly distributed along both dimensions, the row and column
+profiles are tightly distributed along just one dimension. Finally, the corners
+profile is widely distributed along both dimensions.
+
+In the framework proposed by PACKER, the cluster and corners profiles arise
+based on different prioritization of within-category similarity versus
+between-category contrast, and the row and column profiles arise based on the
+prioritization of each dimension in the computation of similarity. For example,
+in the cluster profile, there is a high degree of within-category similarity
+along both dimensions, whereas in the corners profile there is minimal
+within-category similarity. Thus, PACKER's proposal is that these individual
+differences arise as a result of different priorities: While the tight cluster
+configuration can be considered PACKER's `default' mode (as it maximizes
+within-category similarity), the corners profile can be produced when
+between-category contrast is put at a higher priority (i.e., $\theta_c > \theta_t$).
+
+Likewise, in the row and column profiles, there is a large degree of
+within-category similarity along one dimension but not the other. These
+differences likely arise due to a differential focus on one dimension over
+another, and thus they can be produced by changes to PACKER's attention weights,
+$w_1$ and $w_2$ (see Equation \ref{eq:similarity}). Traditionally, the attention
+weights in exemplar models are thought to reflect the diagnostic value of each
+dimension towards classifying the known category members
+\citep{nosofsky1984choice,nosofsky1986attention,kruschke1992alcove}, but within
+a generation context the weights specify the importance of within- and
+between-category similarity along each dimension. For example, if all of
+attention is allocated along the X-axis ($w_1=1$ and $w_2=0$), similarity along
+the Y-axis no longer influences performance. As a result, PACKER will create
+categories that are more widely distributed along the Y-axis, as similarity is
+not taken into account along that dimension. As a general principle,
+differentially weighting one dimension will result in the generation of
+categories that are more widely distributed along the ignored dimension,
+conforming to a row- or column-like arrangement. See Figure
+\ref{fig:packer-attention} for a depiction of how attention influences PACKER's
+performance.
+
+\begin{figure}
+    \begin{center} \inputpgf{figs/}{packer-attention-examples.pgf}
+    \caption{PACKER generation of a category `B' example, following exposure to
+one member of category `A' and one member of category `B'. Predictions are shown
+for different attention settings: {\em (a)} Increased weighting of the X-axis.
+{\em (b)} Increased weighting of the Y-axis. {\em (c)} Uniform weighting
+(identical to Figure \ref{fig:packer-examples}).}
+    \label{fig:packer-attention}
+    \end{center}
+\end{figure}
+
+As in PACKER, changes in the parameter settings of the copy-and-tweak model can
+also be used to produce different patterns of generation. Indeed, as
+copy-and-tweak is simply a special case of the PACKER model, the attention
+weights operate exactly as described above to produce row- and column-like
+categories. However, because the model is not influenced by category contrast,
+it is biased toward generating tightly clustered categories, as new items are
+always most likely to be generated near known examples of the target category.
+Thus, the lack of a contrast mechanism prevents the model from explaining why
+some individuals widely distribute their categories to the corners of the space.
+\jlanote{Xian: Please add a paragraph or two on the representativeness model
+  here.}
+\sxlnote{Ok! I've kind of integrated it into one bigger paragraph with
+  hierarchical bayes.}
+
+Unlike PACKER, both the hierarchical Bayesian model and the representative model
+do not contain any explicit mechanism for controlling the weights of
+between-category contrast and within-category similarity. This means both models
+do not make strong predictions of cluster versus corner profiles. However, the
+covariance matrix specifying the model's prior domain distribution, $\Sigma_0$,
+can be used to explain the generation of row-like and column-like categories.
+This covariance matrix specifies the amount of variance assumed along each
+dimension (as well as the dimensional correlations) across the domain of
+categories. The covariance matrix for a newly generated category, $\Sigma_B$, is
+based on the assumed $\Sigma_0$ as well as the distributions of previously
+learned categories (see Appendix \ref{ap:hsampling-definition}). Thus, the
+importance of each feature can be coded into $\Sigma_0$ to alter the dimensional
+variance of generated categories. Because the hierarchical Bayesian model
+possesses no mechanism to account for category contrast, the model is most
+likely to generate new items that are similar to known examples of the target
+category with no regard for how different it is to the contrast category.
+However, the representativeness model predicts that new exemplars should be more
+representative (i.e., more similar to target and less similar to contrast
+categories), accounting for the tendency of row-like and column-like profiles
+occupying the edge of the feature space.
+
+As noted above, while the copy-and-tweak and hierarchical Bayesian models
+possess mechanisms to explain row- and column-like categories, they cannot
+easily explain why some individuals widely distribute their generated categories
+into disparate corners of the space. This, however, reveals a more general
+limitation: According to the copy-and-tweak and hierarchical Bayesian models,
+the distributional structure of generated categories is {\em independent} of
+their location within the domain. For example, although the copy-and-tweak or
+hierarchical Bayesian models can be parameterized to generate row- or
+column-like categories, there is no mechanism in place to ensure that what is
+generated will be distinct from what is already known. In the next subsection,
+we explore this prediction through an analysis of the interdependence between
+distributional structure and location in creative generation.
+
+\subsection{Category Location vs. Distributional Structure}
+\label{section:individual-diff}
+As noted above, while all three models make clear claims about the internal
+structure of generated categories, the copy-and-tweak and hierarchical Bayesian
+models do not make any claims about how generated categories should differ from
+what is already known. However, as we observed in the results of Experiment 2,
+the distributional structure of a category is not always independent of its
+location within the domain. To demonstrate this point in more depth, we computed
+the X- and Y- axis ranges of every participant-generated category. Taking the
+difference between these values ($X-Y$) produces a measure of each category's
+orientation in the space: positive difference scores correspond to categories
+with more X-axis range (horizontally aligned, `Row' categories), whereas
+negative difference scores indicate the opposite (vertically aligned, `Column'
+categories). Neutral differences scores indicate there was an equal amount of X-
+and Y-axis range, which can be produced by a number of different category types
+(`Clusters', `Corners', etc; see Figures \ref{fig:e1-samples} and
+\ref{fig:e2-samples}). By plotting, for each possible stimulus, the difference
+scores of categories it was generated within, we can relate the distributional
+structure of generated categories to their location within the domain.
+
+However, because many stimuli were infrequently generated (such items near
+members of the `Alpha' category), we cannot simply compute the empirical average
+of the difference scores, as infrequently generated stimuli would be likely to
+show artificially strong differences. Instead, we used a Bayesian analysis to
+estimate the mean $\mu_x$ on the assumption that the scores $x$ for each
+stimulus are normally distributed with an unknown mean and unknown standard
+deviation. The conjugate Normal-Inverse Gamma distribution provides a
+straightforward method for this estimation:
+
+\begin{equation} \mu_x = \dfrac { \nu_0 \mu_0 + \sum{x} } { \nu_0 + n }
+\label{eq:rangediff-bayes}
+\end{equation} \ where $\mu_0$ is the prior mean, $\nu_0$ is a prior scale
+parameter (controlling the weighting of the $\mu_0$), and $n$ is the number of
+categories in which the stimulus was a member (i.e., the number of scores in
+$x$). The default assumption is that there is an equal amount of range along the
+X- and Y-axes, and so we set $\mu_0 = 0$. Likewise, to give a moderate amount of
+weighting to the prior mean we set $\nu_0 = 1$, though the results are robust to
+a range of values. Within this approach, the resulting aggregation is a
+trade-off between the number of generations and the strength of the range
+difference within each generated category. Infrequently generated stimuli, as
+well as those with mixed positive and negative scores, are given neutral
+difference scores.
+
+\begin{figure}[p]
+    \begin{center} 
+      \includegraphics[width=\textwidth]{figs/range-diff-gradients.pdf}
+      \caption{Behavioral and simulated range difference gradients. Each panel
+        shows, for each stimulus, the dimensional orientation of the categories
+        it was generated into: vertically aligned `columns' (orange) versus
+        horizontally aligned `rows' (purple).}
+      \label{fig:range-diff-gradients}
+    \end{center}
+\end{figure}
+
+The results of our analysis are shown in Figure \ref{fig:range-diff-gradients}
+for the experiment and model results\footnote{Prior to plotting, data was also
+processed using a Gaussian filter with $\sigma = 0.8$.}. The left-most column of
+Figure \ref{fig:range-diff-gradients} displays the effect of category location
+and contrast on the distributional structure of the category generated by
+participants. These data reveal strong and consistent patterns across all the
+conditions we tested in Experiments 1 and 2: Generated categories are more
+tightly distributed along the axis in which they are distinct. For example, in
+the `Cluster' condition, exemplars in the bottom-left of the space are more
+often generated into vertically aligned categories, and exemplars in the
+top-right are more often generated into horizontally aligned categories.
+Similarly, in the `Bottom' and `Middle' conditions, horizontally aligned
+categories are generated above and below the experimenter-defined categories,
+while vertically-aligned categories are generated to the sides. In the `Row'
+condition, most categories are horizontally aligned, and lie along the upper
+areas of the space. There are no strong range difference patterns in the XOR
+condition.
+
+These patterns of performance clearly depict the interdependence between the
+distributional structure and location of generated concepts. Our results can be
+interpreted in terms of local minimization of between-category similarity: By
+distributing the generated category away from members of the
+experimenter-defined category, participants may increase the degree of
+between-category distance without drastically altering the degree of
+within-category similarity.
+
+To explore how well the PACKER, copy-and-tweak, representativeness, and
+hierarchical Bayesian models explain our findings, we conducted simulations
+using an individual-differences approach. As noted in Section
+\ref{section:individual-diffs}, row- and column-like categories can be produced
+by each model through changes to the weighting of each dimension. Given this
+information, we may use the models to simulate each participant's generation
+separately, with the importance of each dimension set according to the relative
+range of the participant's generated category along each dimension.
+
+In the PACKER and copy-and-tweak models, the attention weights, $w$, specify the
+importance of each dimension in the computation of similarity. While there exist
+methods to find the optimal attention weighting scheme given a classification
+\citep[see][]{vanpaemel2012using}, for simplicity we may assume that the Alpha
+and Beta categories are distinct along dimensions that the Betas do not vary on.
+Thus, the weighting for a given participant can be computed as:
+
+\begin{equation} w_k = \dfrac {\exp{ \{ -\theta_w \cdot\text{range}(k)} \} } {
+\sum_k {\exp{ \{ -\theta_w \cdot\text{range}(k)} \} } }
+\label{eq:range-weight}
+\end{equation}
+% 
+where $\theta_w$ is a free parameter controlling how differences in range
+correspond to differences in weights (functioning similarly to the $\theta$
+parameter in each of the models), and $\text{range}(k)$ is the range of examples
+generated by the participant along dimension $k$. We used $\theta_w = 1.5$ in
+our simulations, though the results are robust and similar for other $\theta_w$
+values. The resulting $w$ values are thus inversely proportional to the range of
+generated categories along each dimension, with less range corresponding to
+greater weighting.
+
+Unlike the PACKER and copy-and-tweak models, the representativeness and
+hierarchical Bayesian model's dimensional variances correspond to the assumed
+variance of generated categories along each dimension (rather than the inverse
+of the variance). Thus, a different transformation is appropriate for
+incorporating the weights computed in Equation \ref{eq:range-weight}. For the
+hierarchical Bayesian model, we computed the dimensional variances according to:
+$\lambda (1-w_k) 2$, where $\lambda$ is a free parameter specifying the overall
+assumed variance of the domain, and $2$ corresponds to the number of dimensions
+in our experiments\footnote{This calculation applies only in two-dimensional
+  domains, where $w_2 = 1-w_1$.}. Under this approach, evenly distributed
+weights correspond to an assumed variance of $\lambda$. Likewise, larger values
+of $w$, which are produced when the generated category is tightly distributed
+along one dimension, correspond to smaller assumed variances.
+
+Each model was used to simulate each participant's generation independently,
+with the importance of each dimension set according to the participant's
+generated category. The other free parameters within each model were set as in
+Table \ref{table:global-model-fits}. Every participant's generation was
+simulated 2,000 times; given the 305 participants tested across the two
+experiments, each model generated 610,000 categories in total. For comparison
+with our behavioral results, we then computed the range difference gradient
+identically as with the behavioral data. The results are shown in Figure
+\ref{fig:range-diff-gradients}.
+
+As in the more traditional model evaluation analysis described above, the
+contrast models (i.e., PACKER and the representativeness model) provided a much
+closer match to our behavioral results than the copy-and-tweak and hierarchical
+Bayesian models. In all conditions, the contrast models distribute categories
+similarly to the behavioral data: Horizontally-aligned categories tend to be
+placed above and below members of the experimenter-defined category, and
+vertically-aligned categories tend to be placed to the sides. Conversely,
+because the copy-and-tweak and hierarchical Bayesian models are insensitive to
+category contrast, these models do not produce any systematic patterns of
+association between category location and distributional structure. The sole
+exception is within the `Row' condition of Experiment 1, in which the majority
+of participants generated a `Row'-like category, widely distributed along the
+X-axis but not the Y-axis. In these cases, both models are initialized with
+weights that produce Row categories, but because category contrast is not
+considered, categories are uniformly generated across the entire domain, rather
+than concentrated within the upper-regions as observed behaviorally.
+
+\include{methods_results_e3}
+
+\section{General Discussion}
+The sensory impression of every stimulus and event is unique. Grouping distinct
+patterns of sensory information into categories is a fundamental task solved by
+the mind. Most work has focused on how people learn new categories that are
+provided to them through unlabeled, partially labeled, or fully-labeled
+examples. How were these categories first determined? Some natural categories
+are likely to be the result of regularities in the dynamics of our environment.
+But, these are only a subset of the categories that people learn. Other
+categories, such as tools and ideas, were generated by people over time. What
+basic principles underlie how people generate categories?
+
+% The creative generation of concepts is an immensely intriguing topic, but it is
+% also complex and difficult to study in a traditional laboratory setting.
+% Creative generation is most typically studied using a creative cognition
+% approach \citep{finke1992creative,smith1995creative}, where creative products
+% \citep[such as drawings of alien plants and animals;][]{ward1994structured} are
+% analyzed to obtain insight into the processes and representations involved.
+% While this approach has yielded a great deal of information about the nature of
+% creative generation, the commonly employed response modalities (e.g., hand-drawn
+% pictures) make it difficult to develop computational models capable of formally
+% evaluating the proposed representations and processes. However, the generation
+% of novel concepts is not altogether different from other behaviors studied in
+% cognitive psychology: In particular, it can be viewed as an form of inference
+% using one's knowledge about existing categories
+% \citep{kemp2014taxonomy,kurtz2015human}. Taking advantage of this insight, in
+% this paper we investigated the generation of new categories using the empirical
+% and computational toolkit from the field of categorization.
+
+While the bulk of prior research on the topic has focused on the classic finding
+that generated concepts tend to be distributionally similar to known concepts,
+there has been little work addressing the role of contrast in creative
+generation: How is it that people are able to create something {\em different}
+from what is already known?
+\jlanote{Update this next section:}
+\sxlnote{Updated}
+ We developed two novel models, each incorporating a
+different conceptualization of category contrast. One model is PACKER, which is
+an exemplar-based model, that formally specifies the role of contrast in
+generation as exemplar dissimilarity. Specifically, the model proposes that
+categories are represented as exemplars in a multidimensional psychological
+space, and generation is constrained both by within-category and
+between-category similarity: Exemplars belonging to the same category should be
+similar to one another, and exemplars belonging to different categories should
+not be similar to one another. The second model is a novel hierarchical
+Bayesian model involving a representativeness mechanism. This model generates
+exemplars that are more representative of (i.e., has more evidence from) the
+novel category compared to the learned category.
+
+We reported two experiments demonstrating systematic effects of category
+contrast in creative generation. Members of participant-generated categories
+tended to be highly dissimilar from members of previously-learned categories,
+and were usually more similar to one another than to members of other
+categories. We also observed broad interdependence between the distributional
+structure (feature variance, correlation) and physical instantiation (location
+within the stimulus space) of generated categories: In Experiment 2, we found
+that the unoccupied regions of the domain influenced the distributional
+structure of categories, and in both experiments we observed that participants
+distributed their generated categories to increase contrast with what was
+already known.
+
+We conducted simulations comparing the contrast models' account of our results
+to the classical proposals for category generation: a ``copy-and-tweak'' model
+(realized as a variant of PACKER with no sensitivity to category contrast), and
+a hierarchical Bayesian model designed to explain the classic distributional
+similarity effect. In all simulations, we found that the contrast models captured a previously unexplained and unexplored aspect of human
+category generation. In particular, by measuring PACKER's fit as a function of its
+prioritization of within- and between-category similarity, we observed that
+considering either constraint exclusively results in a relatively low-quality
+account. Instead, PACKER's best results were obtained when both constraints are
+considered, indicating that human learners do not generate novel concepts
+exclusively on the basis of within-category similarity or between
+class-contrast. This finding mirrors our behavioral results and demonstrates
+that both constraints influence creative generation when explained with an
+exemplar model. Ultimately, our category generation data was best explained by
+the representativeness model. The exact nature of its superiority over the other
+models is still not yet clear. However, as we explore in a later section, it is
+likely that the representativeness mechanism itself (i.e., independent of the
+hierarchical Bayesian foundation) captures a core aspect of both category
+generation and category learning.
+\jlanote{Above needs some updating still. Also, do we want an outline of general
+discussion paragraph?}
+\sxlnote{Updated. I tried writing an outline, but found that points got a little
+  repetitive after a bit, so I ditched that idea... Might have just been the way
+I wrote it though.}
+
 \subsection{Novel Analyses Demonstrating Contrast Effects in Prior Work}
@@
 not require any mechanism of category contrast.
+
+\jlanote{Xian: Please do a Chi-Squared test between what is expected based on
+  the logic above and the observed frequencies mentioned below}
+\sxlnote{I didn't do a chi-sq test because what appears below is not the
+  observed frequencies that correspond to the logic above, but is instead just
+  an indication of whether or not it is reasonable to assume the logic above
+  (and the argument is that it is not reasonable.)}
 
@@
 
-\subsection{Contrast as exemplar dissimilarity: The PACKER Model}
-
-\label{section:PACKER-definition}
-
-As noted above, the constraint that new concepts should differ from what is
-already known has been largely overlooked in previous work. This is no doubt in
-part due to the vague definition of what it means for a concept to be
-``different'': A generated category may be different from what is already known
-in any number of respects. Towards providing a more precise definition of the
-role of contrast in creative generation, we formalized contrast in a novel
-exemplar model, PACKER ({\em Producing Alike and Contrasting Knowledge using
-Exemplar Representations}). PACKER explains category generation as a balance
-between two fundamental constraints: The category to be generated should not be
-similar to known categories, and exemplars within each category should be
-similar to one another. These ideas are implemented within the well-studied
-exemplar framework -- the PACKER model is an extension of the influential
-Generalized Context Model of categorization
-\citep[GCM;][]{nosofsky1984choice,nosofsky1986attention}.
-
-Although, as an exemplar model, one of PACKER's proposals is people represent
-categories in terms of a collection of stored exemplars, it is worth emphasizing
-that the underlying nature of human category representations is not the primary
-concern of our work. The choice to develop PACKER within an exemplar framework
-reflects the facts that exemplar models have been throughly evaluated, are
-strongly theoretically motivated, and dominate much of the theoretical and
-empirical work in categorization. The focus of our work with PACKER concerns the
-dual constraints of within- and between-class similarity; it is not difficult to
-imagine how such constraints may be instantiated using alternative frameworks
-(e.g., \citealp{kurtz2007divergent,love2004sustain,smith2000thirty}; for a
-review of categorization models, see \citealp{pothoswills2011}).
-
-Both PACKER and the GCM simulate categorization under the assumption that
-learners represent categories as a collection of exemplars, corresponding to the
-labeled stimuli they have observed. The exemplars are encoded within a
-$k$-dimensional psychological space, and model performance is based on the
-amount of similarity between the item to be categorized and the stored
-exemplars. Similarity between two examples, $s\left(x_i, x_j\right)$, is
-computed as an inverse exponential function of distance
-\citep[following][]{attneave1950,shepard1957stimulus,shepard1987toward}:
-
-\begin{equation} s\left(x_i,x_j\right) = \exp \left\{ -c \left[\sum_{k}{ w_k
-\left| x_{ik} - x_{jk} \right|^r }\right]^{1/r} \right\}
-\label{eq:similarity}
-\end{equation}
-% 
-where $w_k$ is the attention weighting of dimension $k$ ($w_k \geq 0$ and
-$\sum_k{w_k} = 1$), accounting for the relative importance of each dimension in
-similarity calculations, and $c$ ($c>0$) is a specificity parameter controlling
-the spread of exemplar generalization. For simplicity, attention will be
-distributed uniformly in our simulations (unless otherwise noted). The value of
-$r$ depends on the nature of the experimental conditions being simulated: $r=1$
-is appropriate for separable dimensions, whereas $r=2$ is appropriate for
-integral dimensions \citep[e.g.,][]{shepard1964attention,garner1974processing}.
-In our simulations, we set $r=1$ due to the separable nature of the stimulus
-dimensions used in our experiments (see Figure \ref{fig:e1-conditions}).
-
-PACKER (as well as its name) was in part inspired by earlier work from the
-categorization literature. Specifically, \cite{hidaka2011packing} argued that
-natural categories ``pack'' the values of features such that different
-categories fill the domain space with distance between one another, while
-keeping items within the same category close together. Inspired by this idea,
-PACKER proposes that generation is constrained by both similarity to members of
-the target category (the category in which a stimulus is being generated) as
-well as similarity to members of other categories: the most desirable generation
-candidates are similar to members of the target category and not similar to
-members of contrast categories. This is achieved by aggregating similarity
-across known exemplars differently according to class membership. The aggregated
-similarity $a(y,x)$ between generation candidate $y$ and stored exemplars $x$ is
-given by:
-
-\begin{equation} a(y, x) = \sum_j{f(x_j) s(y, x_j)}
-\end{equation}
-% 
-where $f(x_j)$ is a function specifying each exemplar's contribution to
-generation. A negative value for $f(x_j)$ produces a `repelling' effect (items
-are less likely to be generated nearby $x_j$), and a positive value produces an
-`attracting' effect (items are more likely to be generated nearby $x_j$). When
-$f(x_j)=0$, the exemplar does not contribute to generation.
-
-PACKER sets $f(x_j)$ depending on exemplar $x_j$'s category membership:
-$f(x_j) = \theta_t$ if $x_j$ is a member of the target category, and
-$f(x_j) = -\theta_c$ if $x_j$ is a member of a contrast category. $\theta_t$ and
-$\theta_c$ are free parameters ($0 \leq \theta_t,\theta_c$) controlling the
-trade-off between within- and between-category similarity. For example, when
-$\theta_t = \theta_c = 0.5$, $f(x_j) = 0.5$ for members of the target category
-and $f(x_j) = -0.5$ for members of other categories; thus, the model is likely
-to generate items that are similar to members of the target category but are not
-similar to members of other categories. In this way, $\theta_t > 0$ with
-$\theta_c = 0$ produces exclusive consideration of target-category members, and
-$\theta_c > 0$ with $\theta_t = 0$ produces exclusive consideration of
-contrast-category members. The combination of $\theta_t$ and $\theta_c$
-parameters thus specifies a wide breadth of possible approaches; by fitting it
-to a dataset, one can describe the relative roles of between-category contrast
-and within-category similarity in generation. See Figure
-\ref{fig:packer-examples} for an illustration of how these parameters control
-the relative influence of within-category similarity and contrast to other
-categories when generating a new exemplar.
-
-\begin{figure}
-	\begin{center} \inputpgf{figs/}{packer-examples.pgf}
-      \caption{PACKER generation of a category `B' example, following exposure
-        to one member of category `A' and one member of category `B'.
-        Predictions are shown for three different parameterizations (differing
-        in values of $\theta_t$ and $\theta_c$): {\em (a)}
-        Predictions based on contrast similarity only. {\em (b)}
-        Predictions based on target similarity only. {\em (c)} Predictions with
-        both constraints considered.}
-		\label{fig:packer-examples}
-	\end{center}
-\end{figure}
-
-The probability that a candidate $y$ will be generated is evaluated using an
-Exponentiated \citet{luce1977choice} choice rule. Candidates with greater values
-of $a(y,x)$ are more likely to be generated than candidates with smaller values:
-% 
-\begin{equation} p(y \mid x) = \dfrac { \exp \left \{ a \left( y, x
-\right) \right \} } { \sum_i{ \exp \left \{ a \left( y_i, x \right)
-\right\} } }
-\label{eq:packer-choice}
-\end{equation}
-% 
-%where $\theta$ ($\theta \geq 0$) is a free parameter controlling response
-%determinism.
-
-It is worth noting that PACKER is only one possible exemplar-based account of
-category generation within our proposed framework. That is, PACKER places
-specific constraints on the possible values of $f(x_j)$, but other
-exemplar-based category generation models with drastically different behavior
-can be formalized in this framework by imposing alternative constraints. For
-example, as will be discussed in more detail below, PACKER is formally
-equivalent to the copy-and-tweak model proposed by \cite{jern2013probabilistic}
-when $\theta_c = 0$ and $\theta_t > 0$. Likewise, when
-$\theta_t = 0$ and $\theta_c > 0$, PACKER can represent a
-contrast-only generation mode, relying exclusively on contrast when generating
-new categories. Finally, when $f(x_j) < 0$ for all $x_j$ (regardless of class
-membership), a ``pure-packing'' approach is yielded, generating items in
-unoccupied areas of the domain. Thus, the proposed framework may be used to
-describe a wide variety of qualitatively distinct generation strategies.
-
-It is also worth noting that PACKER represents just one of many possible models
-that incorporate contrast in category generation. For example, it may be
-possible to extend \cite{jern2013probabilistic}'s hierarchical Bayesian model to
-include contrast, but this mechanism would be at odds with the standard
-hierarchical Bayesian framework. In contrast, these ideas emerge naturally from
-the exemplar view: We have not modified any of the core elements of the GCM in
-defining the PACKER model, we simply weight exemplar similarity during
-aggregation. Thus, beyond formally specifying the role of contrast in creative
-generation, the PACKER account allows us to evaluate the well-understood
-principles of the exemplar view within the comparatively understudied field of
-conceptual generation. Nonetheless, we will discuss integrating our approaches
-in the General Discussion.
-
-\subsubsection{Relation Between PACKER and Copy-And-Tweak}
-\label{section:copytweak-packer} The PACKER model is similar to the
-copy-and-tweak model reported by \cite{jern2013probabilistic}: Both models are
-exemplar-based, and both models generate new items according to their similarity
-to known members of the target class. However, PACKER diverges from the
-copy-and-tweak model by including a contrast mechanism, enabling generation
-according to dissimilarity to members of opposing categories. By consequence,
-copy-and-tweak can be realized as a parameterization of the PACKER model that is
-insensitive to category contrast. Specifically, when $\theta_c = 0$ and
-$\theta_t > 0$ (see Figure \ref{fig:packer-examples}, panel B),
-$f(x_j) = 0$ for $x_j$ belonging to contrast categories; thus, PACKER is not
-influenced by these items, and is mathematically equivalent to a copy-and-tweak
-approach.
-
-In this paper, we report simulations using this copy-and-tweak model. This model
-fits within the exemplar-based category generation framework defined above, and
-is a continuous-dimension adaptation of the model tested by
-\cite{jern2013probabilistic}. By formalizing a model family where PACKER and
-copy-and-tweak are different parameterizations of models within the same
-framework, the comparison between PACKER and copy-and-tweak provides a test of
-the explanatory value of the contrast mechanism based on exemplar dissimilarity:
-The account provided by copy-and-tweak will only equal that of PACKER if the
-contrast mechanism does not offer an advantage (i.e., if $\theta_c > 0$
-significantly improves model fits). Note that the purpose of the article is to
-explore and formally analyze the role of contrast in category generation and
-thus, we leave extending PACKER to incorporate distributional factors \citep[as
-explored by][]{jern2013probabilistic} for future work.
-
-\subsection{Contrast as representativeness: An extension to Jern and Kemp
-(2013)}
-\label{section:representativeness-definition}
-
-An alternate conceptualization of category contrast is the idea of
-representativeness -- exemplars are generated such that they are more
-representative of the target category and thus more distinct from and less
-similar to other categories. We adopt the definition specified by
-\cite{tenenbaum01}, where the representativeness $R$ of an item (or in our case,
-an exemplar) $y$ is the relative amount of evidence that is provided by $y$ for
-a given hypothesis $h$ compared to all other hypotheses $H'$. More formally,
-\begin{equation}
-  R(x,h) = \log \dfrac{p(x|h)}{\sum_{h'\in H'}{p(x|h')p(h')}} , 
-\label{eq:representativeness}
-\end{equation}
-
-In Equation \ref{eq:representativeness}, $p(h')$ is the prior probability
-distribution on the hypothesis space that excludes $h$. Note that when there are
-only two hypotheses (e.g. two categories) involved in the domain, this prior
-takes the value of $1$ since there is ever only one alternative hypothesis.
-
-This Bayesian formalization of representativeness makes it straightforward to
-implement into the existing hieriarchical Bayesian model of category generation
-developed by \cite{jern2013probabilistic}. Specifically, both models assume that
-exemplars are sampled from a given category distribution ($h$ in
-Equation \ref{eq:representativeness}). Each category distribution is a
-multivariate normal parameterized by a location vector $\mu_k$ and covariance
-matrix $\Sigma_k$ (i.e., $h = \mathcal{N}(\mu_k,\Sigma_k)$). The parameters
-$\mu_k$ and $\Sigma_k$ are assumed to be samples from a prior
-normal-inverse-Wishart (NIW) distribution parameterized by
-$\mu_0,\Sigma_0,\kappa,\nu$. Mathematically,
-\begin{align}
-  \mu_k, \Sigma_k|\mu_0,\Sigma_0,\kappa,\nu &\sim
-                                              \textrm{NIW}(\mu_0,\Sigma_0,\kappa,\nu)
-                                              \label{eq:mvnsample}\\
-  y|C &\sim \mathcal{N}(\mu_k,\Sigma_k) \label{eq:exemplarsample}
-\end{align}
-
-In our simulations, we set the prior mean $\mu_0$ to the center of the stimulus
-space and $\Sigma_0 = \lambda \mathbf{I}$ where $\lambda$ is a free parameter
-and $\mathbf{I}$ is a $d$-by-$d$ identity matrix with $d$ representing the
-number of dimensions or features in the domain. The $\kappa$ and $\nu$
-parameters are freely estimated within the constraints $\kappa > d-1$ and
-$\nu > 0$.
-
-With the assumption of the NIW prior, the location vector $\mu_k$ can be computed
-with:
-\begin{equation}
-  \mu_k = \dfrac
-    {\kappa\mu_{0} + n_k \bar{x}_k}
-    {\kappa + n_k}
-    \label{eq:category_mus}
-\end{equation}
-where $n_k$ is the number of observed exemplars in category $k$ and $\bar{x}_k$
-is the observed category mean. Note that if there are no observed exemplars in
-the target category (i.e., if the model is generating a completely novel
-category), Equation \ref{eq:category_mus} simplifies to $\mu_k = \mu_0$.
-
-The NIW prior also allows us to infer the category covariance matrix $\Sigma_k$
-by computing the following:
-\begin{equation}
-  \Sigma_k = [\Sigma_D\nu + C_k +
-    \dfrac
-    {\kappa n_k}
-    {\kappa + n_k}
-    (\bar{x_k}-\mu_k)(\bar{x_k}-\mu_k)^T
-  ] (\nu + n_k)^{-1}
-  \label{eq:Sigma_y}
-\end{equation}
-where $C_k$ is the observed category covariance and $\Sigma_D$ is the domain
-covariance matrix from which $\Sigma_k$ samples are obtained. We can
-infer $\Sigma_D$ from the observed category covariances $C$ and the prior
-covariance $\Sigma_0$. Specifically,
-\begin{equation}
-    \Sigma_D = \Sigma_0 + \sum_{k}{C_k}
-\end{equation}
-
-At this point, both the representativeness model and the hierarchical Bayesian
-model from \cite{jern2013probabilistic} are largely identical. However, both
-models diverge in their computation of exemplar generation probabilities. While
-the original hierarchical Bayesian model produces novel exemplars with
-probabilities proportional to the multivariate normal likelihood $p(y|h)$, the
-representativeness model generates new exemplars with probabilities weighted by
-their representativeness $R(y,h)$. In practice, the probability of generating a
-particular candidate $y$ is obtained using an Exponentiated
-\citet{luce1977choice} choice rule:
-\begin{equation}
-  p(y) = \dfrac
-    {\exp( \theta \cdot R(y,h))}
-    {\sum_k \exp(\theta  \cdot R(y_k,h_k))} .
-\label{eq:representative-choice}
-\end{equation}
-where $\theta$ is a freely estimated response determinism parameter (constrained
-such that $\theta \geq 0$).
-
-Despite both the original hierarchical Bayesian model and our representativeness
-model sharing identical hierarchical structures, their distinct response
-processes can yield very different exemplars. Specifically, the hierarchical
-Bayesian model emphasizes the generation of exemplars that maintain
-distributional commonalities across categories. In contrast, the
-representativeness model focuses on the generation of exemplars that are
-representative of the underlying distribution for a target category. Generally,
-this mechanism of representativeness results in the generation of exemplars that
-are less similar to exemplars from other categories. We see this illustrated in
-Figure \ref{fig:representative-examples}, where the representativeness model
-displays a strong preference for generating exemplars from target category 'B'
-that are further away from the contrast category 'A'.
-
-\begin{figure}
-	\begin{center} \inputpgf{figs/}{hbm-examples.pgf}
-		\caption{Generation of a category `B' example, following exposure
-to one member of category `A' and one member of category `B'. Predictions are
-shown for {\em(a)} the original hierarchical Bayesian model and {\em(b)} the
-representativeness model.}
-		\label{fig:representative-examples}
-	\end{center}
-\end{figure}
-
-\subsection{Synopsis and Prognosis}
-
-Research on the creative generation of novel concepts has focused on the finding
-that generated categories tend to possess distributional commonalities with
-known categories. However, a fundamental goal of concept generation is to create
-something {\em new} (i.e., different from what is already known). The manner in
-which generated categories differ from known ones is, nonetheless, poorly
-understood: Existing theories do not make strong predictions about how
-creatively generated concepts should systematically differ from existing ones.
-Above, we found encouraging initial support that contrast influences category
-generation \citep[][Experiment 3]{jern2013probabilistic}, and we introduced two
-novel approaches to formalizing the roles of similarity and contrast in creative
-generation.
-
-In the sections below, we present two experiments demonstrating systematic
-effects of category contrast on category generation inspired by factors
-influencing how the contrast models generate new categories. Our experiments are
-based on \cite{jern2013probabilistic}'s paradigm, which is a straightforward
-translation of the traditional artificial classification paradigm to the task of
-category generation: Participants are first exposed to a single,
-experimenter-defined category, and are then asked to generate members of a new
-category. We then report formal analyses comparing the exemplar dissimilarity
-and the representativeness accounts of our results to that of the hierarchical
-Bayesian and copy-and-tweak models developed by \cite{jern2013probabilistic}.
-After analyzing the category generation studies, we present third experiment to
-demonstrate the effect of contrast in category learning. This ensures that our
-results are not idiosyncratic to category generation.
-
-\section{Experiment 1}
-
-To begin our investigation, we sought to extend the early evidence obtained from
-our analysis of the \cite{jern2013probabilistic} data, under a variety of
-learning conditions and using more standard stimulus materials. We used an
-artificial stimulus design: A two dimensional domain of squares, varying in
-color and size (see Figure \ref{fig:e1-conditions}, panel A). These dimensions
-have been used in numerous classification learning studies
-\cite[e.g.,][]{conaway2016similar,conaway2016generalization,shepard1961learning,nosofsky1994comparing}.
-Unlike those used in the \cite{jern2013probabilistic} experiments, distance on
-these physical dimensions aligns more directly with perceptual similarity,
-allowing us to evaluate the role of category contrast in creative generation
-more precisely. To extend the evidence provided by the
-\cite{jern2013probabilistic} data, we tested the effects of category contrast
-after learning one category from a set of qualitatively distinct category
-structures, as shown in Figure \ref{fig:e1-conditions}.
-
-Figures \ref{fig:e1-conditions}b-d show the values of exemplar dimensions
-belonging to the experimenter-defined categories (`A', or `Alpha') that
-participants were assigned to learn about prior to generating a new category.
-Each participant learned one of the category types during training. In the
-`Cluster' type, category A is a tight cluster of examples in the space.
-Perceptually instantiated, the members of category A might, for example, be
-large and dark in color. In the `Row' type, category A has a row pattern across
-the space, varying along one feature but not the other. Thus, its members might
-all be dark in color but would vary in size. Finally, in the `XOR' type, the
-experimenter-defined category consists of two clusters separated in opposite
-corners of the space, conforming to the exclusive-or logical structure (e.g.,
-members are small and dark or large and light).
-
-It should be noted that in our experiments the assignment between the perceptual
-and conceptual dimensions (e.g., $X \rightarrow Size$, $Y \rightarrow Color$),
-as well as the direction of variation along each dimension (e.g., $dark
-\rightarrow light$ or $light \rightarrow dark$) was counterbalanced across
-participants. The category types in Figure \ref{fig:e1-conditions} are plotted
-in a conceptual space, rather than a perceptual space. Thus, while the
-conceptual organization of the category types remains constant, each category
-type may have a different physical instantiation according to the counterbalance
-assignment. For example, the Cluster type may be large and dark in color, or it
-may be small and light in color, depending on the assignment and direction of
-the dimensions. For this reason, below we will discuss generation within a
-conceptual space, rather than a physically instantiated one.
-
-\begin{figure}
-    \begin{center} \inputpgf{figs/}{e1-conditions.pgf}
-    \caption{Stimulus domain and category types tested in Experiment 1. Stimuli
-are not drawn to scale.}
-    \label{fig:e1-conditions}
-    \end{center}
-\end{figure}
-
-After learning about an experimenter-defined category, participants are asked to
-generate examples of a new category. Within this paradigm, an effect of category
-contrast would be realized if participants prefer to generate items in locations
-that are distant (i.e., perceptually dissimilar) from members of category A.
-However, generation is left unconstrained. Critically, participants were not
-asked to generate something different in the prompt. For example, participants
-assigned to the Cluster condition may generate a tightly clustered category in
-the corner opposite of the experimenter-defined category. Alternatively, they
-may generate a tightly clustered category directly overlapping with the
-experimenter-defined category. Further, they may even generate an entirely
-different type of category (e.g., a row category).
-
-Our experimental results also provide a converging test of the classic finding
-that generated categories tend to share distributional properties with known
-categories in the domain \citep{jern2013probabilistic,ward1994structured}. From
-these results, we can predict that, in each condition, participants should
-generate categories that are distributionally similar to the
-experimenter-defined category: In the Cluster condition, generated categories
-should be tightly clustered. In the Row condition, generated categories should
-vary more along the X-axis than the Y-axis. In XOR condition, generated
-categories should be widely distributed across both dimensions, and the two
-dimensions should be positively correlated.
-
-Interestingly, the XOR condition also offers a dissociation between the roles of
-category contrast and the emulation of distributional structure:
-widely-distributed, positively-correlated categories would need to lie along the
-positive diagonal of the space (that is the only place they ``fit''), which is
-already occupied by the experimenter-defined category. Thus, if contrast plays a
-role, exemplars in the generated categories of participants in the XOR condition
-may not be positively correlated -- they may be negatively correlated instead.
-In this case, contrast and statistical regularities would interact, which would
-be inconsistent with the leading approach in conceptual generation
-\citep{jern2013probabilistic}
-
-\subsection{Participants and Materials}
-
-183 participants were recruited from Amazon Mechanical Turk. Each participant
-was randomly assigned to one condition: 64 participants were assigned to the
-Cluster condition, 61 were assigned to the Row condition, and 58 were assigned
-to the XOR condition (sample sizes differ due to random assignment). Stimuli
-were squares varying in color (grayscale 9.8\%--90.2\%) and side length
-(3.0--5.8cm), see Figure \ref{fig:e1-conditions}. The assignment of perceptual
-features (color, size) to axes of the domain space (x, y), as well as the
-direction of variation along each axis (e.g., $dark \rightarrow light$ or $light
-\rightarrow dark$) was counterbalanced across participants.
-
-\subsection{Procedure}
-
-As noted in the introduction of this paper, the task of generating members of a
-new category can be aligned with common tasks studied in the categorization
-literature: Whereas classification consists of predicting an object's category
-label on the basis of its features, inference consists of predicting an observed
-feature, given a set of observed features and a category label. Generation thus
-consists of predicting {\em all} features of an object, given a novel category
-label. In keeping with this insight, we designed our generation task as an
-extension of the traditional artificial classification learning paradigm. The
-task differs from traditional work in creative cognition primarily through the
-use of an artificial domain, which enables the application of computational
-models. The use of an artificial domain also requires the addition of a training
-phase, during which participants learn about the categories in the domain. As a
-result, unlike most previous studies \citep[e.g.,][]{ward1994structured},
-participants in our studies have no experience with the domain before the start
-of the experiment, and the experimenter-defined categories are not
-hierarchically structured (as are many natural categories).
-
-Participants began the experiment with a short training phase (3 blocks of 4
-trials), where they observed exemplars belonging to the `Alpha' category.
-Participants were instructed to learn as much as they can about the Alpha
-category, and that they would answer a series of test questions afterwards. On
-each trial, a single Alpha category exemplar was presented, and participants
-were given as much time as they desired to observe it before moving on to the
-next trial. Each block consisted of a single presentation of each of the members
-of the Alpha category, in a random order. Participants were shown the range of
-possible colors and sizes prior to training.
-
-Following the training phase, participants were asked to generate four examples
-belonging to another category called `Beta'. As in
-\citet{jern2013probabilistic}, generation was completed using a sliding-scale
-interface. Two scales controlled the values of the two dimensions (color, size)
-for the generated example. An on-screen preview of the example updated whenever
-one of the features was changed. Participants could generate any example along
-an evenly-spaced 9x9 grid (including members of the Alpha category), except for
-any previously generated Beta exemplars. Neither the members of the Alpha
-category nor the previously generated Beta examples were visible during
-generation. Prior to beginning the generation phase, participants read the
-following instructions:
-
-\begin{displayquote} As it turns out, there is another category of geometric
-figures called ``Beta''. Instead of showing you examples of the Beta category,
-we would like to know what you think is likely to be in the Beta category.
-
-You will now be given the chance to create examples of any size or color in
-order to show what you expect about the Beta category. You will be asked to
-produce 4 Beta examples - they can be quite similar or quite different to each
-other, depending on what you think makes the most sense for the category.
-
-Each example needs to be unique, but the computer will let you know if you
-accidentally create a repeat.
-\end{displayquote}
-
-\subsection{Results}
-
-We observed a substantial degree of individual differences in our data. In
-Figure \ref{fig:e1-samples} we have plotted sample data from several
-participants, from which it is evident that different participants generated
-qualitatively different category structures. In this section we will focus on
-analyzing the data in aggregate, but in later sections we will explore how
-individual differences can be explained.
-
-\begin{figure}
-    \begin{center} \inputpgf{figs/}{e1-samples.pgf}
-    \caption{Sample categories generated by participants in Experiment 1.
-Representative samples from common generation profiles are shown.}
-    \label{fig:e1-samples}
-    \end{center}
-\end{figure}
-
-To evaluate the role of contrast, we computed the number of times each stimulus
-was generated, as a function of its average city-block distance from members of
-the experimenter-defined ``Alpha'' category. These data, shown in Figure
-\ref{fig:e1-distanceplots}, reveal a clear pattern: Examples that are more
-distant from members of the experimenter-defined categories are more likely to
-be generated into a new category. This supports the notion that contrast is a
-fundamental constraint on creative cognition and that suggests that statistical
-regularity alone is insufficient.
-
-Figure \ref{fig:e1-distanceplots} also depicts, for each participant, the
-average distance of members within the generated category ({\em within-category}
-distance) against the average distance between members of the generated and
-experimenter-defined category ({\em between-category} distance). The narrow
-distribution of between-category distances in the XOR condition reflects the
-widely distributed nature of the experimenter-defined category, reducing the
-possible distances to members of the participant-generated category. These data
-reveal a systematic pattern: The majority of participants generated categories
-with greater between-category distance than within-category distance. That is,
-members of the generated category tended to be more similar to one another than
-to members of the experimenter-defined category. To evaluate this claim
-quantitatively, we conducted t-tests comparing the amount of within- and
-between- class distance in each condition. All conditions possessed greater
-between-category distance: Cluster, $t(63) = 11.43$, $p < 0.001$; Row, $t(60) =
-13.16$, $p < 0.001$; and XOR, $t(57) = 3.64$, $p < 0.001$. These results provide
-further evidence of an effect of category contrast: Participants prefer to
-generate categories that are dissimilar to the learned category but maintain
-some level of internal cohesion.
-
-\begin{figure}
-    \begin{center} \inputpgf{figs/}{e1-distanceplots.pgf}
-    \caption{Experiment 1 results. {\em Left}: Frequency of generation as a
-function of distance from members of the experimenter-defined category. {\em
-Right}: Scatter plot of within-category versus between-category distance in each
-of the participant-generated categories.}
-    \label{fig:e1-distanceplots}
-    \end{center}
-\end{figure}
-
-A secondary goal of this experiment was to examine whether we replicate the
-classic result that generated categories often possess the same distributional
-properties as previously-known categories. For each generated category, we
-computed the category range along each axis (X, Y), as well as the correlation
-between features. These data, shown in Figure \ref{fig:e1-statsboxes}, reveal
-broad individual differences: Within each condition, participants generated
-categories spanning the entire X- and Y- axis as well as categories that spanned
-very little along each. Likewise, in each condition participants generated
-categories possessing strongly positive, neutral, and strongly negative
-correlations between the dimensions. Comparing the distributional statistics
-between conditions yields a broad yet, as we will see, misleading replication of
-the classic effect.
-
-With respect to ranges along each axis (X, Y), the generated categories from
-each condition tend to reflect the ranges of the experimenter-defined
-categories. The categories generated in the Cluster condition were less widely
-distributed along the X-axis compared to Row, $t(123) = 5.61$, $p < 0.001$, and
-XOR, $t(120) = 2.68$, $p = 0.008$. Categories generated in the XOR condition
-were also less widely distributed along the X-axis compared to Row, $t(117) =
-2.56$, $p = 0.012$. This latter effect was not expected because the
-experimenter-defined categories for XOR and Row had similar X-ranges. However,
-the key finding is that categories from the Cluster condition tended to be more
-tightly distributed along the X-axis.
-
-\begin{figure}
-    \begin{center} \inputpgf{figs/}{e1-statsboxes.pgf}
-    \caption{Box-plots of the distributional statistics from the categories
-generated in Experiment 1. Boxes depict the median and quartiles of each
-condition, with whiskers placed at 1.5 IQR. All points outside this region are
-marked individually.}
-    \label{fig:e1-statsboxes}
-    \end{center}
-\end{figure}
-
-Likewise, categories generated in the Row condition had less Y-axis range
-compared to Cluster, $t(123) = 4.57$, $p < 0.001$ and XOR, $t(117) = 9.26$, $p <
-0.001$, and categories from the Cluster condition had less Y-axis range compared
-to XOR, $t(120) = 3.95$, $p < 0.001$. As expected, the correlations in the
-Cluster and Row conditions were not systematically positive or negative ($ps >
-0.1$). However, the generated categories in the XOR condition tended to possess
-{\em negatively} correlated dimensions, $t(57) = 2.04$, $p = 0.046$. This
-finding is notable, as it is the opposite of what would be expected, assuming
-learners are emulating the distributional structure of the experimenter-defined
-class (which possesses perfectly positively correlated features).
-
-We believe that the failure to replicate this finding is because participants in
-\cite{jern2013probabilistic} could differentiate the generated category on a
-third dimension (hue) to maintain the statistical regularities on the other two
-dimensions. Although the correlation in the XOR condition is significantly
-negative, it is clear from the box-plot in Figure~\ref{fig:e1-statsboxes} that
-it would be inappropriate to make a strong conclusion (e.g., the median is close
-to zero). However, we can conclude with confidence that there are situations
-where people do not emulate the distributional structure of the given category.
-This indicates that there is more to category generation than the emulation of
-distributional structure of other categories in the domain. Further, as we will
-discuss in more detail in the model-based analysis section, this is expected by
-our proposal that contrast is a fundamental principle in category generation.
-
-\subsection{Discussion} In Experiment 1 we sought to extend our analysis of the
-\cite{jern2013probabilistic} data by evaluating the influence of category
-contrast on creative generation, given qualitatively different types of prior
-knowledge. We found strong evidence for effects of category contrast in each
-condition: Participants were more likely to generate stimuli that are more
-distant from (i.e., less similar to) members of a previously-learned category,
-and members of participant-generated categories tended to be more similar to one
-another than to members of previously-learned categories. We also partially
-replicated the classic finding that the distributional structure of generated
-categories reflects that of previously learned categories
-\citep{jern2013probabilistic,ward1994structured}: Members of generated
-categories were more widely distributed along dimensions which were widely
-distributed in the experimenter-defined category.
-
-Notably, however, we also found that participants who learned an XOR category
-(composed of exemplars following a positive diagonal, see Figure
-\ref{fig:e1-conditions}) tended to generate items according to a {\em negative}
-feature correlation -- the opposite of what was present in the previously
-learned category. While this may be difficult to account for under existing
-theoretical approaches (which assume generated categories follow the same
-distributional structure as known categories), it can be concisely explained
-from a category contrast perspective. Specifically, within the XOR condition,
-individuals who seek to generate a category that is perceptually distinct from
-what is already known are left with only the upper-left and bottom-right
-quadrants of the space, as members of the previously-learned XOR category lie in
-the bottom-left and top-right. If examples are generated into both of the
-available quadrants, the generated category will possess a strongly negative
-correlation, opposing that of the experimenter-defined class.
-
-Thus, the core results of Experiment 1 indicate that generated categories can
-systematically differ from what we would expect based on prior work. The
-negative (or null) correlations observed in the XOR condition suggests an
-interesting interaction between contrast with a given category and emulation of
-statistical properties. That is, the constraints on creative generation imposed
-by category contrast may not simply influence the {\em location} of generated
-categories, but also their distributional structure. In Experiment 2, we test
-this claim more systematically.
-
-\section{Experiment 2.}
-
-\begin{figure}
-    \begin{center} \input{figs/e2-conditions.pgf}
-    \caption{Category types tested in Experiment 2.}
-    \label{fig:e2-conditions}
-    \end{center}
-\end{figure}
-
-To test whether category contrast influences the distributional structure of
-generated categories, we sought to identify conditions in which differences in
-the distributional structure of generated categories cannot be explained by the
-distributional structure of the experimenter-defined category. We created two
-new category types (depicted in Figure \ref{fig:e2-conditions}) that possess
-identical distributional structures (both are tight clusters of examples with no
-correlation between features), as they only differ in their Y-axis position: the
-`Bottom' category lies in the bottom-center of the space, and the `Middle'
-category lies in the center. The distributional equality of these conditions is
-key to the design of the experiment: If the distributional structure of
-previously learned categories were the only influence on the generated
-categories, we should observe no difference in the categories participants
-generate between these two conditions. Will participants distribute their
-generated category differently between conditions due to the differences in the
-available stimulus space for generating a new category?
-
-If category contrast influences the distributional structure of the categories
-people generate, then we should observe different types of categories according
-to the shape of the space that is {\em unoccupied} by members of previously
-learned categories. The difference in the Y-axis position between the Bottom and
-Middle conditions produces a considerable change to the shape of the unoccupied
-space. Participants assigned to learn the Bottom category should be less likely
-to generate exemplars into the lower regions of the stimulus space (as these
-areas possess greater similarity to members of the Bottom category), preferring
-instead to distribute exemplars across the upper region of the space. This
-constraint is lifted in the Middle condition, as the Middle category exemplars
-are equidistant to the upper and lower regions of the space. Accordingly,
-participants should be more likely to utilize both of these areas. Thus, if
-category contrast influences the distributional structure of generated
-categories, we should observe more participants in the Middle condition that
-generate examples above {\em and} below the experimenter-defined category.
-
-\subsection{Participants, Materials, and Procedure}
-
-122 participants were recruited from Amazon Mechanical Turk. 61 participants
-were randomly assigned to the Middle and Bottom conditions each. The stimulus
-space and procedure were exactly as in Experiment 1. Participants first
-completed a short training phase, followed by the generation phase. The only
-difference from Experiment 1 was the category types given to participants.
-
-\subsection{Results}
-
-\begin{figure}
-    \begin{center} \inputpgf{figs/}{e2-samples.pgf}
-    \caption{Sample categories generated in Experiment 2. }
-    \label{fig:e2-samples}
-    \end{center}
-\end{figure}
-
-As in Experiment 1, we observed broad differences in the generation approach
-taken by different participants. To characterize the nature of these
-differences, Figure \ref{fig:e2-samples} depicts sample categories generated by
-participants. The data from each condition are organized into four columns based
-on commonly observed patterns of generation: a `Cluster' type of
-tightly-clustered examples, `Row' and `Column' types of exemplars widely
-distributed along the one axis but narrowly along the other, and a `Corners'
-type, wherein participants placed exemplars in disparate corners of the space.
-As before, in this section we focus on analyzing the data in aggregate, but in
-later sections we will focus more specifically on explaining the individual
-differences.
-
-\begin{figure}
-    \begin{center} \inputpgf{figs/}{e2-distanceplots.pgf}
-    \caption{Experiment 2 results. {\em Left}: Frequency of generation as a
-function of distance from members of the experimenter-defined category. {\em
-Right}: Scatter plot of within-category versus between-category distance in each
-of the participant-generated categories.}
-    \label{fig:e2-distanceplots}
-    \end{center}
-\end{figure}
-
-We began our analysis by testing for the broad influence of category contrast on
-generation. As in Experiment 1, we computed the frequency each stimulus was
-generated as a function of its average distance from members of the
-experimenter-defined category, as well as each participant's average within- and
-between- category distance. These data, shown in Figure
-\ref{fig:e2-distanceplots}, yield very similar results. Participants generated
-stimuli that are distant from members of the experimenter-defined category, and
-the categories in each condition tended to possess more between-category than
-within-category distance: Bottom, $t(60) = 5.5$, $p < 0.001$; Middle, $t(60) =
-2.71$, $p < 0.01$. We did, however, observe a notable subgroup of participants
-in each condition who generated categories with more within-category than
-between-category distance. Upon manual inspection, many of these individuals
-appear to have assumed a `Corners' strategy, placing exemplars in disparate
-corners of the space, thus producing much more within-category distance, see
-Figure \ref{fig:e2-samples} for examples.
-
-To explore the distributional structure of the generated categories, we computed
-the range of exemplars along each axis (X, Y), as well as the correlation
-between features. These data, shown in Figure \ref{fig:e2-statsboxes}, again
-demonstrate the degree of individual differences observed in our study. In each
-condition, we observed tightly clustered and widely distributed categories along
-each dimension, as well as positively, negatively, and uncorrelated categories.
-
-\begin{figure}
-    \begin{center} \inputpgf{figs/}{e2-statsboxes.pgf}
-    \caption{Box-plots of the distributional statistics from the categories
-generated in Experiment 2. }
-    \label{fig:e2-statsboxes}
-    \end{center}
-\end{figure}
-
-As noted above, if the distributional structure of generated categories is
-influenced by the shape of the space not occupied by members of known
-categories, then participants in the Middle condition would be more likely to
-place exemplars in the upper {\em and} lower regions of the space, as members of
-the experimenter-defined category are equidistant from these regions.
-Participants in the Bottom condition should be less likely to generate category
-members in the bottom regions because members of the experimenter-defined
-category are located there. One way to test these predictions is to analyze the
-Y-axis ranges of the generated categories: If Middle participants utilize the
-upper and lower regions of the space, their categories should vary more along
-the Y-axis. T-Tests comparing the conditions on the distributional statistics,
-however, reveal few between-group differences: the conditions do not differ with
-respect to X-axis range, Y-axis range, or feature correlations ($ps>0.17$).
-
-However, our ability to detect differences in Y-Axis range using a standard
-$t$-test between the conditions is, in this case, diminished due to the
-non-normality of the data (Shapiro-Wilk normality test $W=0.77, p< 0.001$ for
-the Middle condition and $W=0.85, p < 0.001$ for the Bottom condition). Figure
-\ref{fig:e2-yranges} depicts the Y-axis position of the exemplars generated
-within each participant's category. The categories are sorted by overall range,
-then by condition assignment. These data reveal that there were nearly as many
-participants who generated categories spanning the entire Y-axis as those who
-generated categories spanning almost none of the Y-axis. The non-normality of
-the Y-axis range distributions thus requires that we use a different approach to
-addressing the experiment's main question.
-
-\begin{figure}
-    \begin{center} \inputpgf{figs/}{e2-yranges.pgf}
-    \caption{Y-Axis range and position of the participant-generated categories
-from Experiment 2. Each line corresponds to a participant's category, with
-notches corresponding to the Y-axis position of exemplars within the category
-(notches may overlap). Participants are sorted by overall range, and then by
-condition. }
-    \label{fig:e2-yranges}
-    \end{center}
-\end{figure}
-
-Because our main prediction concerns the generation of exemplars within the
-upper and lower regions of the domain, we compared the conditions in terms of
-the frequency with which participants generated examples above and below the
-categories. Specifically, we counted the number of participants in each
-condition who placed at least one `Beta' exemplar on the top and bottom `rows'
-of the space (the maximum and minimum possible Y-axis value, respectively). The
-resulting contingencies data are shown in Table \ref{table:e2-subset-table}.
-
-Firstly, it should be noted that nearly every participant utilized the top
-and/or bottom rows: only $10 / 122$ participants generated their category
-entirely within the interior region. Fisher's Exact Tests comparing the
-conditions reveal that more Middle participants generated an exemplar in the
-bottom row, $p < 0.001$, again demonstrating the role of contrast in guiding
-where exemplars are generated. The conditions did not differ in use of the top
-of the space, $p = 0.16$, however, more Middle participants placed exemplars in
-the top {\em and} bottom rows, $p = 0.038$. The latter effect is of interest
-here, as it indicates that the shape of the unoccupied space exerts some
-influence on the distributional structure of generated categories: Participants
-in the Middle condition were more likely to generate a category spanning the
-entire Y-axis. Thus, distributional structure of the generated categories can be
-influenced by category contrast alone.
-
-\begin{table}
-\begin{center}
-\caption{Experiment 2 results.}
-\label{table:e2-subset-table} \vskip 0.12in
-\begin{tabular}{ l r r} \textbf{Middle} & Used top row & No top row \\ \hline
-Used bottom row & 28 & 18 \\ No bottom row & 11 & 4 \\ \\ \textbf{Bottom} & Used
-top row & No top row \\ \hline Used bottom row & 16 & 8 \\ No bottom row & 31 &
-6 \\
-\end{tabular}
-\end{center}
-\end{table}
-
-\subsection{Discussion}
-
-In Experiment 2, we replicated the core findings from Experiment 1. Stimuli are
-more likely to be generated if they are distant from exemplars in other
-categories, and most participants generate categories with more between-category
-than within-category distance. However, we additionally found that the {\em
-position} of a previously learned category (rather than its distributional
-structure) influences the types of categories people generate: Participants who
-learned the `Middle' type were more likely to generate categories spanning the
-entire Y-axis of the space. Participants who learned the `Bottom' type were less
-likely to do so as a result of the presence of opposite category exemplars in
-the lower regions of the space.
-
-This finding cannot be explained from the perspective that the distributional
-structure of previously learned categories is the sole determinant of the
-distributional structure of generated categories. However, the observed behavior
-is expected from a category contrast perspective: Participants seeking to
-generate a perceptually distinct category will be more likely to use areas of
-space that are unoccupied by exemplars belonging to previously learned
-categories. In the Middle condition, the upper and lower regions of space are
-equidistant from members of the experimenter-defined category, whereas in the
-Bottom condition, the lower region of the space is closer to members of the
-experimenter-defined category. Thus, while Middle participants may form
-categories around the use of the equally unoccupied areas, the same is not true
-for the Bottom condition.
-
-\section{Model-based Analyses}
-\sxlnote{For this section I've added the existence of the representativeness
-  model but I think discussion of its performance is still lacking compared to PACKER.}
-Experiments 1 and 2 revealed systematic and strong effects of category contrast
-on creative generation. In this section, we report the results of simulations
-with formal models aimed at explaining our observations. Specifically, we
-present simulations from the two novel contrast models: PACKER and the
-representativeness model, and compare them to two models that do not incorporate
-contrast: the copy-and-tweak model (discussed in Section
-\ref{section:copytweak-packer}) and an implementation of the hierarchical
-Bayesian model proposed by \cite{jern2013probabilistic}, described in-depth in
-Appendix \ref{ap:hsampling-definition}. For our simulations here, the
-copy-and-tweak model is defined as a variant of PACKER with the $\theta_c$
-parameter constrained to be zero. The comparison of this set of models serves to
-highlight the explanatory role of contrast in categorization: If contrast
-affords little explanatory advantage, then there should be few differences in
-performance between PACKER and copy-and-tweak, or between the representativeness
-and hierarchical Bayesian model. The comparison between these models can also
-emphasize the necessity of contrast and demonstrate that generation cannot be
-explained entirely through the emulation of distributional structure. Each model
-has complementary strengths and weaknesses: Whereas PACKER and copy-and-tweak
-are insensitive to the distributional structure of learned categories (relying
-only on exemplar similarities), the representativeness and hierarchical Bayesian
-model generates categories exclusively on the basis of knowledge of how existing
-classes are distributed.
-
-Our approach in this section is to first broadly evaluate and compare the
-quality of each model's account to our entire dataset (Experiments 1 and 2
-combined), then analyze the ability for each model to explain individual
-differences in each experiment, and lastly we describe the strengths and
-weakness of each model's account of category generation.
-
-\subsection{Parameter-Fitting}
-
-To obtain a global measure of the quality of each model's account, we fitted the
-parameters of each model to our entire dataset (Experiments 1 and 2 combined),
-using a hill-climbing algorithm which maximized the log-likelihood of the
-model's predictions of the observed responses (1220 responses from 305 total
-participants). We fitted three parameters in the PACKER model ($c$, $\theta_t$,
-and $\theta_c$; see Section \ref{section:PACKER-definition}), as well as four in
-the representativeness model and the hierarchical Bayesian model ($\kappa$,
-$\lambda$, $\nu$, and $\theta$; see Section
-\ref{section:representativeness-definition} and Appendix
-\ref{ap:hsampling-definition} respectively). We fitted only two parameters for
-the copy-and-tweak model ($c$, and $\theta_t$), as $\theta_c$ is held constant
-($\theta_c = 0$). Attention ($w$, see Equation \ref{eq:similarity}) in PACKER
-and copy-and-tweak was set uniformly. Parameters were not allowed to vary
-between participants or conditions -- the goal was to obtain the best-fitting
-values to our entire dataset.
-
-\begin{table} \centering
-\caption{Results of model-fitting to the combined datasets from Experiments 1
-and 2. Note that smaller AIC values correspond to better model fits (adjusted
-for number of parameters)}
-\label{table:global-model-fits}
-\begin{tabular}{ l l l l} \\ \textbf{PACKER} & \textbf{Copy \& Tweak} &
-\textbf{Representativeness} & \textbf{Hierarchical Bayesian} \\ \hline $AIC =
-  9095$ & $AIC = 9842$ & $AIC = 8878$ &  $AIC =
-9912$ \\ $L = -4545$ & $L = -4919$ & $L = -4435$ & $L = -4952$ \\ $c = 0.483$ &
-                                                                                $c = 3.197$ & $\kappa = 11.232$ &
-$\kappa < 0.001$ \\ $\theta_c = 3.16$ & $\theta_c = 0$ (fixed) & $\nu = 2.844$ & $\nu = 5.597$ \\
-$\theta_t = 3.494$ & $\theta_t = 2.972$ & $\lambda = 1.917$& $\lambda
-                                                                       = 0.055$
-  \\ & &$\theta = 5.885 $& $\theta = 3.175$
-\\
-\end{tabular}
-\end{table}
-
-Table \ref{table:global-model-fits} contains the results of this fitting
-procedure. Due to the uneven number of fitted parameters among the models, we
-compare the model fits using the Akaike Information Criterion
-\citep[AIC;][]{akaike1974new}, where smaller values correspond to better fits
-(discounted by model complexity as measured by the number of parameters). The
-same qualitative results were obtained with alternative model comparison metrics
-(e.g., BIC, \citealp{schwarz1978estimating}; $AIC_C$,
-\citealp{hurvich1989regression}).
-
-Table \ref{table:global-model-fits} contains the AIC values of each model's best
-fit, as well as the corresponding log-likelihood ($L$) and the best-fitting
-parameter values. These results reveal strong model differentiation: both
-contrast models (PACKER and the representativeness model) achieved far better
-fits compared to their non-contrast counterparts: copy-and-tweak and the
-hierarchical Bayesian model respectively. Interestingly between the contrast
-models, the hierarchical model (representativeness) outperformed the
-exemplar-based theory (PACKER), whereas between the non-contrast models, the
-reverse is observed. Specifically, here the exemplar-based model
-(copy-and-tweak) performed somewhat better than the hierarchical Bayesian model.
-
-\sxlnote{The argument in the following two paragraphs appear to be challenged by
-  the success of the representativeness model. I might suggest removing them altogether?}
-While PACKER's advantage may tentatively be attributed to the model's
-sensitivity to category contrast (this will be explored in detail below), the
-advantage shown by copy-and-tweak over the hierarchical Bayesian model may be
-attributed to its exemplar-based representation, as opposed to the
-prototype-based representation assumed by the hierarchical Bayesian model. As
-observed in Figures \ref{fig:e1-samples} and \ref{fig:e2-samples}, the generated
-categories we observed were often widely distributed, with no items near the
-category prototype. This aspect of the data is inconsistent with the
-multivariate normal distributions (similar to prototypes) used to represent
-categories in the \cite{jern2013probabilistic} model, but can be easily
-accounted for using an exemplar-based approach.
-
-A key distinction between PACKER and the other three models is that only PACKER
-is capable of making strong predictions about the location of new category
-members when the target class is entirely novel (i.e., no member of the category
-has been observed). Under these circumstances, there are no examples to copy,
-and thus the copy-and-tweak model predicts that items are generated at random.
-Likewise, with no observations on which to condition the category distribution,
-the hierarchical Bayesian and representativeness models also picks an item at
-random. Crucially, however, the next generated exemplar from the
-representativeness model is strongly influenced by contrast as
-representativeness. Thus, it is possible that the failure of these models is
-simply due to their inability to explain each participant's first trial
-(generating the first item in the `Beta' category). To ensure this is not
-driving our results, we conducted an identical set of simulations as above,
-excluding this trial (leaving 915 responses in the dataset): Again, PACKER
-($L = -3390$, $AIC = 6786$) achieved better fits than the copy-and-tweak
-($L = -3579$, $AIC = 7162$) and hierarchical Bayesian ($L=-3612$, $AIC = 7232$)
-models.
-
-Finally, because copy-and-tweak is nested within PACKER, we can use a likelihood
-ratio test to compare the two models. PACKER explains the aggregate data
-significantly better than copy-and-tweak ($\chi^2(1) = 749, p < 0.001$ for all
-data and $\chi^2(1) = 377, p < 0.001$ excluding the first example), providing
-further evidence that category generation is better explained when contrast is
-considered.
-
-Through comparison with the copy-and-tweak model, Figure
-\ref{fig:packer-loglike} more clearly demonstrates the robustness of the
-explanatory gains yielded by PACKER's category contrast mechanism. It displays
-the log-likelihood of the participants' results under PACKER as a function of
-the $\theta_c$ parameter. The model's other parameters ($c$,
-$\theta_t$) were set according to copy-and-tweak's best fits from Table
-\ref{table:global-model-fits}, and thus when $\theta_c=0$, the models are
-equivalent. The figure clearly shows a ``sweet spot'': a convex region in which
-PACKER achieves superior fits as a result of changes to $\theta_c$. The
-best fitting values lie well above the value of 0 assumed by the copy-and-tweak
-model, which demonstrates the robustness of the contrast effect (though note
-PACKER achieves even better fits when its parameters are fitted together, as in
-Table \ref{table:global-model-fits}). In sum, the data are better explained when
-both within-category similarity and category contrast is considered.
-
-\begin{figure}
-    \begin{center} \inputpgf{figs/}{packer-loglike.pgf}
-      \caption{PACKER's fit as a function of its prioritization of
-        within-category and between-category similarity (using the
-        $\theta_c$ parameter) . To facilitate comparison, PACKER's other
-        parameters ($c$, $\theta_t$) were set to the best fitting values
-        obtained for copy-and-tweak in Table \ref{table:global-model-fits}. }
-    \label{fig:packer-loglike}
-    \end{center}
-\end{figure}
-
-\subsection{Individual Differences}
-\label{section:individual-diffs}
-
-As noted in Experiments 1 and 2, we observed a great deal of individual
-differences in the types of categories that participants generated. Within each
-condition, there were a wide variety of category types, such as row and column
-categories (see Figures \ref{fig:e1-samples} and \ref{fig:e2-samples}). The
-simulations reported above serve to evaluate the models while considering the
-entire dataset, but a secondary goal of any formal account should be to provide
-some explanation of how different profiles of performance emerge. Many of the
-individual generation profiles we observed can be described within the PACKER
-framework, simply by tuning the model's parameters in a principled manner. In
-this section, we describe more specifically how the most frequently observed
-profiles can be realized.
-
-By manual inspection, it is evident that the most common profiles of generation
-consist of: (A) a tightly-distributed `cluster' of examples, (B) `row'- and
-`column'-like arrangements (varying widely along one dimension but not the
-other), and (C) a `corners' arrangement with examples placed into disparate
-corners of the space. These four profiles are distinct in terms of the
-distribution of the generated category along each dimension: Whereas the cluster
-profile is tightly distributed along both dimensions, the row and column
-profiles are tightly distributed along just one dimension. Finally, the corners
-profile is widely distributed along both dimensions.
-
-In the framework proposed by PACKER, the cluster and corners profiles arise
-based on different prioritization of within-category similarity versus
-between-category contrast, and the row and column profiles arise based on the
-prioritization of each dimension in the computation of similarity. For example,
-in the cluster profile, there is a high degree of within-category similarity
-along both dimensions, whereas in the corners profile there is minimal
-within-category similarity. Thus, PACKER's proposal is that these individual
-differences arise as a result of different priorities: While the tight cluster
-configuration can be considered PACKER's `default' mode (as it maximizes
-within-category similarity), the corners profile can be produced when
-between-category contrast is put at a higher priority (i.e., $\theta_c > \theta_t$).
-
-Likewise, in the row and column profiles, there is a large degree of
-within-category similarity along one dimension but not the other. These
-differences likely arise due to a differential focus on one dimension over
-another, and thus they can be produced by changes to PACKER's attention weights,
-$w_1$ and $w_2$ (see Equation \ref{eq:similarity}). Traditionally, the attention
-weights in exemplar models are thought to reflect the diagnostic value of each
-dimension towards classifying the known category members
-\citep{nosofsky1984choice,nosofsky1986attention,kruschke1992alcove}, but within
-a generation context the weights specify the importance of within- and
-between-category similarity along each dimension. For example, if all of
-attention is allocated along the X-axis ($w_1=1$ and $w_2=0$), similarity along
-the Y-axis no longer influences performance. As a result, PACKER will create
-categories that are more widely distributed along the Y-axis, as similarity is
-not taken into account along that dimension. As a general principle,
-differentially weighting one dimension will result in the generation of
-categories that are more widely distributed along the ignored dimension,
-conforming to a row- or column-like arrangement. See Figure
-\ref{fig:packer-attention} for a depiction of how attention influences PACKER's
-performance.
-
-\begin{figure}
-    \begin{center} \inputpgf{figs/}{packer-attention-examples.pgf}
-    \caption{PACKER generation of a category `B' example, following exposure to
-one member of category `A' and one member of category `B'. Predictions are shown
-for different attention settings: {\em (a)} Increased weighting of the X-axis.
-{\em (b)} Increased weighting of the Y-axis. {\em (c)} Uniform weighting
-(identical to Figure \ref{fig:packer-examples}).}
-    \label{fig:packer-attention}
-    \end{center}
-\end{figure}
-
-As in PACKER, changes in the parameter settings of the copy-and-tweak model can
-also be used to produce different patterns of generation. Indeed, as
-copy-and-tweak is simply a special case of the PACKER model, the attention
-weights operate exactly as described above to produce row- and column-like
-categories. However, because the model is not influenced by category contrast,
-it is biased toward generating tightly clustered categories, as new items are
-always most likely to be generated near known examples of the target category.
-Thus, the lack of a contrast mechanism prevents the model from explaining why
-some individuals widely distribute their categories to the corners of the space.
-
-Similar to the copy-and-tweak model, the hierarchical Bayesian model possesses
-no mechanism to account for category contrast, and the model is most likely to
-generate new items that are similar to known examples of the target category.
-Although this precludes an account of why individuals might assume a cluster
-versus corners profile, the covariance matrix specifying the model's prior
-domain distribution, $\Sigma_0$, can be used to explain the generation of
-row-like and column-like categories. This covariance matrix specifies the amount
-of variance assumed along each dimension (as well as the dimensional
-correlations) across the domain of categories. The covariance matrix for a newly
-generated category, $\Sigma_B$, is based on the assumed $\Sigma_0$ as well as
-the distributions of previously learned categories (see Appendix
-\ref{ap:hsampling-definition}). Thus, the importance of each feature can be
-coded into $\Sigma_0$ to alter the dimensional variance of generated categories.
-
-As noted above, while the copy-and-tweak and hierarchical Bayesian models
-possess mechanisms to explain row- and column-like categories, they cannot
-easily explain why some individuals widely distribute their generated categories
-into disparate corners of the space. This, however, reveals a more general
-limitation: According to the copy-and-tweak and hierarchical Bayesian models,
-the distributional structure of generated categories is {\em independent} of
-their location within the domain. For example, although the copy-and-tweak or
-hierarchical Bayesian models can be parameterized to generate row- or
-column-like categories, there is no mechanism in place to ensure that what is
-generated will be distinct from what is already known. In the next subsection,
-we explore this prediction through an analysis of the interdependence between
-distributional structure and location in creative generation.
-
-\subsection{Category Location vs. Distributional Structure}
-\label{section:individual-diff}
-As noted above, while all three models make clear claims about the internal
-structure of generated categories, the copy-and-tweak and hierarchical Bayesian
-models do not make any claims about how generated categories should differ from
-what is already known. However, as we observed in the results of Experiment 2,
-the distributional structure of a category is not always independent of its
-location within the domain. To demonstrate this point in more depth, we computed
-the X- and Y- axis ranges of every participant-generated category. Taking the
-difference between these values ($X-Y$) produces a measure of each category's
-orientation in the space: positive difference scores correspond to categories
-with more X-axis range (horizontally aligned, `Row' categories), whereas
-negative difference scores indicate the opposite (vertically aligned, `Column'
-categories). Neutral differences scores indicate there was an equal amount of X-
-and Y-axis range, which can be produced by a number of different category types
-(`Clusters', `Corners', etc; see Figures \ref{fig:e1-samples} and
-\ref{fig:e2-samples}). By plotting, for each possible stimulus, the difference
-scores of categories it was generated within, we can relate the distributional
-structure of generated categories to their location within the domain.
-
-However, because many stimuli were infrequently generated (such items near
-members of the `Alpha' category), we cannot simply compute the empirical average
-of the difference scores, as infrequently generated stimuli would be likely to
-show artificially strong differences. Instead, we used a Bayesian analysis to
-estimate the mean $\mu_x$ on the assumption that the scores $x$ for each
-stimulus are normally distributed with an unknown mean and unknown standard
-deviation. The conjugate Normal-Inverse Gamma distribution provides a
-straightforward method for this estimation:
-
-\begin{equation} \mu_x = \dfrac { \nu_0 \mu_0 + \sum{x} } { \nu_0 + n }
-\label{eq:rangediff-bayes}
-\end{equation} \ where $\mu_0$ is the prior mean, $\nu_0$ is a prior scale
-parameter (controlling the weighting of the $\mu_0$), and $n$ is the number of
-categories in which the stimulus was a member (i.e., the number of scores in
-$x$). The default assumption is that there is an equal amount of range along the
-X- and Y-axes, and so we set $\mu_0 = 0$. Likewise, to give a moderate amount of
-weighting to the prior mean we set $\nu_0 = 1$, though the results are robust to
-a range of values. Within this approach, the resulting aggregation is a
-trade-off between the number of generations and the strength of the range
-difference within each generated category. Infrequently generated stimuli, as
-well as those with mixed positive and negative scores, are given neutral
-difference scores.
-
-\begin{figure}[p]
-    \begin{center} 
-      \includegraphics[width=\textwidth]{figs/range-diff-gradients.pdf}
-      \caption{Behavioral and simulated range difference gradients. Each panel
-        shows, for each stimulus, the dimensional orientation of the categories
-        it was generated into: vertically aligned `columns' (orange) versus
-        horizontally aligned `rows' (purple).}
-      \label{fig:range-diff-gradients}
-    \end{center}
-\end{figure}
-
-The results of our analysis are shown in Figure \ref{fig:range-diff-gradients}
-for the experiment and model results\footnote{Prior to plotting, data was also
-processed using a Gaussian filter with $\sigma = 0.8$.}. The left-most column of
-Figure \ref{fig:range-diff-gradients} displays the effect of category location
-and contrast on the distributional structure of the category generated by
-participants. These data reveal strong and consistent patterns across all the
-conditions we tested in Experiments 1 and 2: Generated categories are more
-tightly distributed along the axis in which they are distinct. For example, in
-the `Cluster' condition, exemplars in the bottom-left of the space are more
-often generated into vertically aligned categories, and exemplars in the
-top-right are more often generated into horizontally aligned categories.
-Similarly, in the `Bottom' and `Middle' conditions, horizontally aligned
-categories are generated above and below the experimenter-defined categories,
-while vertically-aligned categories are generated to the sides. In the `Row'
-condition, most categories are horizontally aligned, and lie along the upper
-areas of the space. There are no strong range difference patterns in the XOR
-condition.
-
-These patterns of performance clearly depict the interdependence between the
-distributional structure and location of generated concepts. Our results can be
-interpreted in terms of local minimization of between-category similarity: By
-distributing the generated category away from members of the
-experimenter-defined category, participants may increase the degree of
-between-category distance without drastically altering the degree of
-within-category similarity.
-
-To explore how well the PACKER, copy-and-tweak, representativeness, and
-hierarchical Bayesian models explain our findings, we conducted simulations
-using an individual-differences approach. As noted in Section
-\ref{section:individual-diffs}, row- and column-like categories can be produced
-by each model through changes to the weighting of each dimension. Given this
-information, we may use the models to simulate each participant's generation
-separately, with the importance of each dimension set according to the relative
-range of the participant's generated category along each dimension.
-
-In the PACKER and copy-and-tweak models, the attention weights, $w$, specify the
-importance of each dimension in the computation of similarity. While there exist
-methods to find the optimal attention weighting scheme given a classification
-\citep[see][]{vanpaemel2012using}, for simplicity we may assume that the Alpha
-and Beta categories are distinct along dimensions that the Betas do not vary on.
-Thus, the weighting for a given participant can be computed as:
-
-\begin{equation} w_k = \dfrac {\exp{ \{ -\theta_w \cdot\text{range}(k)} \} } {
-\sum_k {\exp{ \{ -\theta_w \cdot\text{range}(k)} \} } }
-\label{eq:range-weight}
-\end{equation}
-% 
-where $\theta_w$ is a free parameter controlling how differences in range
-correspond to differences in weights (functioning similarly to the $\theta$
-parameter in each of the models), and $\text{range}(k)$ is the range of examples
-generated by the participant along dimension $k$. We used $\theta_w = 1.5$ in
-our simulations, though the results are robust and similar for other $\theta_w$
-values. The resulting $w$ values are thus inversely proportional to the range of
-generated categories along each dimension, with less range corresponding to
-greater weighting.
-
-Unlike the PACKER and copy-and-tweak models, the representativeness and
-hierarchical Bayesian model's dimensional variances correspond to the assumed
-variance of generated categories along each dimension (rather than the inverse
-of the variance). Thus, a different transformation is appropriate for
-incorporating the weights computed in Equation \ref{eq:range-weight}. For the
-hierarchical Bayesian model, we computed the dimensional variances according to:
-$\lambda (1-w_k) 2$, where $\lambda$ is a free parameter specifying the overall
-assumed variance of the domain, and $2$ corresponds to the number of dimensions
-in our experiments\footnote{This calculation applies only in two-dimensional
-  domains, where $w_2 = 1-w_1$.}. Under this approach, evenly distributed
-weights correspond to an assumed variance of $\lambda$. Likewise, larger values
-of $w$, which are produced when the generated category is tightly distributed
-along one dimension, correspond to smaller assumed variances.
-
-Each model was used to simulate each participant's generation independently,
-with the importance of each dimension set according to the participant's
-generated category. The other free parameters within each model were set as in
-Table \ref{table:global-model-fits}. Every participant's generation was
-simulated 2,000 times; given the 305 participants tested across the two
-experiments, each model generated 610,000 categories in total. For comparison
-with our behavioral results, we then computed the range difference gradient
-identically as with the behavioral data. The results are shown in Figure
-\ref{fig:range-diff-gradients}.
-
-As in the more traditional model evaluation analysis described above, the
-contrast models (i.e., PACKER and the representativeness model) provided a much
-closer match to our behavioral results than the copy-and-tweak and hierarchical
-Bayesian models. In all conditions, the contrast models distribute categories
-similarly to the behavioral data: Horizontally-aligned categories tend to be
-placed above and below members of the experimenter-defined category, and
-vertically-aligned categories tend to be placed to the sides. Conversely,
-because the copy-and-tweak and hierarchical Bayesian models are insensitive to
-category contrast, these models do not produce any systematic patterns of
-association between category location and distributional structure. The sole
-exception is within the `Row' condition of Experiment 1, in which the majority
-of participants generated a `Row'-like category, widely distributed along the
-X-axis but not the Y-axis. In these cases, both models are initialized with
-weights that produce Row categories, but because category contrast is not
-considered, categories are uniformly generated across the entire domain, rather
-than concentrated within the upper-regions as observed behaviorally.
-
-\include{methods_results_e3}
-
-\section{General Discussion}
-
-The creative generation of concepts is an immensely intriguing topic, but it is
-also complex and difficult to study in a traditional laboratory setting.
-Creative generation is most typically studied using a creative cognition
-approach \citep{finke1992creative,smith1995creative}, where creative products
-\citep[such as drawings of alien plants and animals;][]{ward1994structured} are
-analyzed to obtain insight into the processes and representations involved.
-While this approach has yielded a great deal of information about the nature of
-creative generation, the commonly employed response modalities (e.g., hand-drawn
-pictures) make it difficult to develop computational models capable of formally
-evaluating the proposed representations and processes. However, the generation
-of novel concepts is not altogether different from other behaviors studied in
-cognitive psychology: In particular, it can be viewed as an form of inference
-using one's knowledge about existing categories
-\citep{kemp2014taxonomy,kurtz2015human}. Taking advantage of this insight, in
-this paper we investigated the generation of new categories using the empirical
-and computational toolkit from the field of categorization.
-
-While the bulk of prior research on the topic has focused on the classic finding
-that generated concepts tend to be distributionally similar to known concepts,
-there has been little work addressing the role of contrast in creative
-generation: How is it that people are able to create something {\em different}
-from what is already known? We developed a novel, exemplar-based model, PACKER,
-which formally specifies the role of contrast in generation. The model proposes
-that categories are represented as exemplars in a multidimensional psychological
-space, and generation is constrained both by within-category and
-between-category similarity: Exemplars belonging to the same category should be
-similar to one another, and exemplars belonging to different categories should
-not be similar to one another.
-
-In addition to an analysis of published data \citep[][Experiment
-3]{jern2013probabilistic}, we reported two experiments demonstrating systematic
-effects of category contrast in creative generation. Members of
-participant-generated categories tended to be highly dissimilar from members of
-previously-learned categories, and were usually more similar to one another than
-to members of other categories. We also observed broad interdependence between
-the distributional structure (feature variance, correlation) and physical
-instantiation (location within the stimulus space) of generated categories: In
-Experiment 2, we found that the unoccupied regions of the domain influenced the
-distributional structure of categories, and in both experiments we observed that
-participants distributed their generated categories to increase contrast with
-what was already known.
-
-We conducted simulations comparing PACKER's account of our results to the other
-major proposals for category generation: a ``copy-and-tweak'' model (realized as
-a variant of PACKER with no sensitivity to category contrast), and a
-hierarchical Bayesian model designed to explain the classic distributional
-similarity effect. In all simulations, we found that PACKER's sensitivity to
-contrast captured a previously unexplained and unexplored aspect of human
-category generation. Further, by measuring PACKER's fit as a function of its
-prioritization of within- and between-category similarity, we observed that
-considering either constraint exclusively results in a relatively low-quality
-account. Instead, PACKER's best results were obtained when both constraints are
-considered, indicating that human learners do not generate novel concepts
-exclusively on the basis of within-category similarity or between
-class-contrast. This finding mirrors our behavioral results and demonstrates
-that both constraints influence creative generation.
-
 \subsection{Similarity and Contrast in Cognition}
@@
 \subsection{Implications for Creative Cognition}
-
+\jlanote{this needs updating for representativeness}
+\sxlnote{Updated.}
 Although the focus of this article has been to address the processes involved in
@@
 In addition to demonstrating the utility of formal modeling in the study
-creative cognition, the contrast models more specifically offer an additional
+creative cognition, the contrast models here specifically offer an additional
 interpretation of some of the field's most central findings. For example,
@@
 representativeness model shows that this influence can also be the result of
-enhancing the representativeness of generated exemplars to their category. [xian edit]
+enhancing the representativeness of generated exemplars to their category.
 
@@
 
-\subsection{Disentangling Exemplar dissimilarity and Representativeness in Categorization}
-
-\sxlnote{I'm struggling a little with this section! That said, I think this may
-  also be a good point to address Charles' concern re: the possible conditions
-  under which contrast are important.} 
-Although both PACKER and the
-representativeness model perform better than the other models, they appear to be
-capturing fundamentally different aspects of contrast in categorization.
-Specifically, while the representativeness model excelled in predicting our
-category generation data, PACKER was the best-fitting model in accounting for
-participant error in category learning. The source of this discrepancy is still
-not yet clear -- it could be an indication of a fundamental difference between
-the processes involved in generation versus learning where exemplar
-dissimilarity is a larger factor in category generation while representativeness
-exerts more influence in category learning.
-
-In this paper, we have developed two novel models of category generation, each
-involving a different conceptualization of contrast.  However, there is no theoretical constraint that dictates the It is also possible to treat the similarity
-
-A thorough investigation to disentangle the two contrast mechanisms here is
-beyond the scope of the current article. However, as a quick proof of concept it
-is useful to check if the addition of a representativeness mechanism to an
-exemplar model would yield any benefit at all.
-
-Xian: Please
-write about the two hypotheses and that followup work needs to look into why the
-Bayesian representativeness model doesn't do as well at explaining category
-learning ease, but does better at category generation.
-
-MAKE SURE WE CITE AND DESCRIBE \citet{thomas98}! \sxlnote{I moved the
-  description and citation of thomas98 to the introduction for justification of
-  why we think categories should share similar distributions.}
+The current work also suggests exciting directions for related investigations
+into unsupervised categorization. While the unsupervised categorization
+literature is primarily interested in the generation of categories within a set
+of observed exemplars where no prior categories are learned, our category
+generation studies are focused on the production of novel category exemplars
+themselves. Despite this distinction, both types of categorization research are
+ultimately interested in the question of how categories can be formed. The results
+of our current work indicate that the formation of categories is constrained by
+some measure of contrast. To our knowledge, this explicit investigation of
+category contrast has not yet been attempted in the unsupervised categorization
+literature.
+
+Interestingly, the unidimensional sort bias that is commonly observed in
+unsupervised categorization
+\citep{imai1965discriminability,milton2004influence,ahn1992two} was not
+consistently observed in our tasks. To clarify, the unidimensional sort bias is
+the observation that categories are typically constructed solely by focusing on
+a single dimension or feature. In our experiments we only observed this bias in
+the row condition of Experiment 1, whereas in other conditions participants
+generated categories that typically varied along both dimensions. This may be
+unsurprising at face value, since participants in conditions other than the row
+condition were trained on a prior category that varied on both dimensions.
+However, these results suggest that the unidimensional sort bias is not
+universal to all types of category construction -- or more specifically, the
+unidimensional bias may not apply when participants need to generate categories
+with entirely new exemplars.
+
+\subsection{Exemplar dissimilarity and Representativeness in Categorization} 
+\sxlnote{Hmmm.. ok after quite a few revisions I am still not quite satisfied
+  with this section. It doesn't feel quite ...tight, but I think it's time for
+  me to say stop and pass it on.)
+Although both PACKER and the representativeness model perform better than the
+classical models, they appear to be capturing fundamentally different aspects of
+contrast in categorization. Specifically, while the representativeness model
+excelled in predicting our category generation data, PACKER was the best-fitting
+model in accounting for participant error in category learning. Given the
+history of their development, this is perhaps not entirely surprising. PACKER
+is implemented as an extension of the GCM, which was originally developed for (and
+has been highly successful in) category learning \citep{nosofsky1986attention,
+  nosofsky1994comparing}. Conversely, the representativeness mechanism was
+conceived as a heuristic describing the relative extent to which an individual
+observation is generated from some unobserved population
+\citep{kahneman1972subjective,abbott16}. This generative nature of the
+representativeness mechanism intuitively lends itself to the prediction of
+category generation data.
+
+While the contrast models were born quite naturally out of the corresponding
+classical models, it is worth considering if similar benefits in prediction
+would be seen if the contrast mechanisms were applied to different classical
+models. Although there is no clear way to adapt the exemplar dissimilarity
+mechanism to the current hierarchical Bayesian model, it is a fairly
+straightforward process applying the representativeness mechanism to an exemplar
+model. Specifically, by treating the similarity measure (Equation
+\ref{eq:similarity}) in the current implementation of copy-and-tweak as a
+density estimate for the representativeness mechanism (Equation
+\ref{eq:representativeness}), we gain a new model of representativeness that is
+based on exemplar similarity instead of a multivariate Gaussian
+likelihood.\footnote{An analogous implementation of the representativeness
+  mechanism to PACKER results in both $\theta_c$ and $\theta_t$ adding to
+  constant value that is independent of the similarity to contrast and target
+  exemplars respectively. Consequently, it becomes equivalent to a
+  copy-and-tweak model with representativeness.}
+
+By subjecting this preliminary model to the same analysis as the other models,
+we find that it substantially outperforms its classical counterpart, but not to
+the extent of PACKER or the representativeness model. Specifically, fitting this
+model to the category generation data yielded $L = -4815$ and $AIC=9633$ for the
+entire set of data from Experiments 1 and 2, and $L = -3474$, $AIC=6953$ for
+data excluding the first trials. When used to predict the error rates of
+participants in the category learning task (Experiment 3), the performance of
+the copy-and-tweak with representativeness model correlated with participant
+error with $r=.56, p<.001$. These results suggest that the predictive advantages
+to including the representativeness mechanism is not limited to a hierarchical
+Bayesian model, but can also be found when applied to an exemplar model.
+Although its performance is not as strong as either of the fully-developed
+contrast models, there is a benefit to the representativeness mechanism that is
+independent from any interaction with a hierarchical Bayesian framework.
+
+Ultimately, the source of the discrepancy between the performance of the two
+contrast mechanisms in category generation versus category learning is still not
+entirely clear. Our results could be an indication of a fundamental difference
+between the processes involved where exemplar dissimilarity is a larger factor
+in category generation while representativeness exerts more influence in
+category learning. While a thorough investigation is beyond the scope of the
+current article, the preliminary results of this section indicate that further
+work to disentagle these two mechanisms would be promising. In the following
+section, we explore some other limitations of the current work and provide more
+recommendations for future research.
+
 \subsection{Limitations and Future Directions} Although we have demonstrated
@@
 based on free association \citep{mednick1962associative} and conceptual
-combination \citep{estes2002emergence,murphy1988comprehending}. Instead, we
-reduced our focus by studying a highly complex behavior (creative generation) as
-it applies within a well-established domain (artificial category learning). We
-hope that future work incorporates and highlights the importance of contrast
-into theories of creativity.
+combination \citep{estes2002emergence,murphy1988comprehending}. In addition, the
+machine learning literature has taken very interesting approaches using neural
+networks to address the overall problem of creative generation
+\citep[e.g.,][]{goodfellow2014gan,ho2016gail,kingma2016,chen2016infogan}. In
+contrast to these varied investigations, we reduced our focus by studying a
+highly complex behavior (creative generation) as it applies within a
+well-established domain (artificial category learning). We hope that future work
+incorporates and highlights the importance of contrast into theories of
+creativity.
 
@@
 \end{document}
+
+\message{ !name(cogpsych.tex) !offset(-2294) }

### 2018-09-04 Manuscripts/cog-psych/revision/cogpsych.tex
Primary revisions to manuscript: copytweak-rep in section 8.5, unsup cat, contrast in thomas98, some mention of machine learning generative models. In other news, updated and more or less finalised fits with the category [0] bug fixed in Simulation.py. TODO: Update the goldstate.

@@
 
-\sxlnote{Charles suggested we either remove the section below, or move it to the
-  discussion. I've tried finding a good place for it in the discussion but it
-  doesn't fit very well, and moving it takes away this nice motivation for
-  our further investigation into contrast. I think that keeping it here in the
-  introduction might work out, provided Charles finds my argument for why the
-  regularity assumption is too strong convincing enough. More details a little
-  further down.}
+\subsection{Contrast as exemplar dissimilarity: The PACKER Model}
+
+\label{section:PACKER-definition}
+
+As noted above, the constraint that new concepts should differ from what is
+already known has been largely overlooked in previous work. This is no doubt in
+part due to the vague definition of what it means for a concept to be
+``different'': A generated category may be different from what is already known
+in any number of respects. Towards providing a more precise definition of the
+role of contrast in creative generation, we formalized contrast in a novel
+exemplar model, PACKER ({\em Producing Alike and Contrasting Knowledge using
+Exemplar Representations}). PACKER explains category generation as a balance
+between two fundamental constraints: The category to be generated should not be
+similar to known categories, and exemplars within each category should be
+similar to one another. These ideas are implemented within the well-studied
+exemplar framework -- the PACKER model is an extension of the influential
+Generalized Context Model of categorization
+\citep[GCM;][]{nosofsky1984choice,nosofsky1986attention}.
+
+Although, as an exemplar model, one of PACKER's proposals is people represent
+categories in terms of a collection of stored exemplars, it is worth emphasizing
+that the underlying nature of human category representations is not the primary
+concern of our work. The choice to develop PACKER within an exemplar framework
+reflects the facts that exemplar models have been throughly evaluated, are
+strongly theoretically motivated, and dominate much of the theoretical and
+empirical work in categorization. The focus of our work with PACKER concerns the
+dual constraints of within- and between-class similarity; it is not difficult to
+imagine how such constraints may be instantiated using alternative frameworks
+(e.g., \citealp{kurtz2007divergent,love2004sustain,smith2000thirty}; for a
+review of categorization models, see \citealp{pothoswills2011}).
+
+Both PACKER and the GCM simulate categorization under the assumption that
+learners represent categories as a collection of exemplars, corresponding to the
+labeled stimuli they have observed. The exemplars are encoded within a
+$k$-dimensional psychological space, and model performance is based on the
+amount of similarity between the item to be categorized and the stored
+exemplars. Similarity between two examples, $s\left(x_i, x_j\right)$, is
+computed as an inverse exponential function of distance
+\citep[following][]{attneave1950,shepard1957stimulus,shepard1987toward}:
+
+\begin{equation} s\left(x_i,x_j\right) = \exp \left\{ -c \left[\sum_{k}{ w_k
+\left| x_{ik} - x_{jk} \right|^r }\right]^{1/r} \right\}
+\label{eq:similarity}
+\end{equation}
+% 
+where $w_k$ is the attention weighting of dimension $k$ ($w_k \geq 0$ and
+$\sum_k{w_k} = 1$), accounting for the relative importance of each dimension in
+similarity calculations, and $c$ ($c>0$) is a specificity parameter controlling
+the spread of exemplar generalization. For simplicity, attention will be
+distributed uniformly in our simulations (unless otherwise noted). The value of
+$r$ depends on the nature of the experimental conditions being simulated: $r=1$
+is appropriate for separable dimensions, whereas $r=2$ is appropriate for
+integral dimensions \citep[e.g.,][]{shepard1964attention,garner1974processing}.
+In our simulations, we set $r=1$ due to the separable nature of the stimulus
+dimensions used in our experiments (see Figure \ref{fig:e1-conditions}).
+
+PACKER (as well as its name) was in part inspired by earlier work from the
+categorization literature. Specifically, \cite{hidaka2011packing} argued that
+natural categories ``pack'' the values of features such that different
+categories fill the domain space with distance between one another, while
+keeping items within the same category close together. Inspired by this idea,
+PACKER proposes that generation is constrained by both similarity to members of
+the target category (the category in which a stimulus is being generated) as
+well as similarity to members of other categories: the most desirable generation
+candidates are similar to members of the target category and not similar to
+members of contrast categories. This is achieved by aggregating similarity
+across known exemplars differently according to class membership. The aggregated
+similarity $a(y,x)$ between generation candidate $y$ and stored exemplars $x$ is
+given by:
+
+\begin{equation} a(y, x) = \sum_j{f(x_j) s(y, x_j)}
+\end{equation}
+% 
+where $f(x_j)$ is a function specifying each exemplar's contribution to
+generation. A negative value for $f(x_j)$ produces a `repelling' effect (items
+are less likely to be generated nearby $x_j$), and a positive value produces an
+`attracting' effect (items are more likely to be generated nearby $x_j$). When
+$f(x_j)=0$, the exemplar does not contribute to generation.
+
+PACKER sets $f(x_j)$ depending on exemplar $x_j$'s category membership:
+$f(x_j) = \theta_t$ if $x_j$ is a member of the target category, and
+$f(x_j) = -\theta_c$ if $x_j$ is a member of a contrast category. $\theta_t$ and
+$\theta_c$ are free parameters ($0 \leq \theta_t,\theta_c$) controlling the
+trade-off between within- and between-category similarity. For example, when
+$\theta_t = \theta_c = 0.5$, $f(x_j) = 0.5$ for members of the target category
+and $f(x_j) = -0.5$ for members of other categories; thus, the model is likely
+to generate items that are similar to members of the target category but are not
+similar to members of other categories. In this way, $\theta_t > 0$ with
+$\theta_c = 0$ produces exclusive consideration of target-category members, and
+$\theta_c > 0$ with $\theta_t = 0$ produces exclusive consideration of
+contrast-category members. The combination of $\theta_t$ and $\theta_c$
+parameters thus specifies a wide breadth of possible approaches; by fitting it
+to a dataset, one can describe the relative roles of between-category contrast
+and within-category similarity in generation. See Figure
+\ref{fig:packer-examples} for an illustration of how these parameters control
+the relative influence of within-category similarity and contrast to other
+categories when generating a new exemplar.
+
+\begin{figure}
+	\begin{center} \inputpgf{figs/}{packer-examples.pgf}
+      \caption{PACKER generation of a category `B' example, following exposure
+        to one member of category `A' and one member of category `B'.
+        Predictions are shown for three different parameterizations (differing
+        in values of $\theta_t$ and $\theta_c$): {\em (a)}
+        Predictions based on contrast similarity only. {\em (b)}
+        Predictions based on target similarity only. {\em (c)} Predictions with
+        both constraints considered.}
+		\label{fig:packer-examples}
+	\end{center}
+\end{figure}
+
+The probability that a candidate $y$ will be generated is evaluated using an
+Exponentiated \citet{luce1977choice} choice rule. Candidates with greater values
+of $a(y,x)$ are more likely to be generated than candidates with smaller values:
+% 
+\begin{equation} p(y \mid x) = \dfrac { \exp \left \{ a \left( y, x
+\right) \right \} } { \sum_i{ \exp \left \{ a \left( y_i, x \right)
+\right\} } }
+\label{eq:packer-choice}
+\end{equation}
+% 
+%where $\theta$ ($\theta \geq 0$) is a free parameter controlling response
+%determinism.
+
+It is worth noting that PACKER is only one possible exemplar-based account of
+category generation within our proposed framework. That is, PACKER places
+specific constraints on the possible values of $f(x_j)$, but other
+exemplar-based category generation models with drastically different behavior
+can be formalized in this framework by imposing alternative constraints. For
+example, as will be discussed in more detail below, PACKER is formally
+equivalent to the copy-and-tweak model proposed by \cite{jern2013probabilistic}
+when $\theta_c = 0$ and $\theta_t > 0$. Likewise, when
+$\theta_t = 0$ and $\theta_c > 0$, PACKER can represent a
+contrast-only generation mode, relying exclusively on contrast when generating
+new categories. Finally, when $f(x_j) < 0$ for all $x_j$ (regardless of class
+membership), a ``pure-packing'' approach is yielded, generating items in
+unoccupied areas of the domain. Thus, the proposed framework may be used to
+describe a wide variety of qualitatively distinct generation strategies.
+
+It is also worth noting that PACKER represents just one of many possible models
+that incorporate contrast in category generation. For example, it may be
+possible to extend \cite{jern2013probabilistic}'s hierarchical Bayesian model to
+include contrast, but this mechanism would be at odds with the standard
+hierarchical Bayesian framework. In contrast, these ideas emerge naturally from
+the exemplar view: We have not modified any of the core elements of the GCM in
+defining the PACKER model, we simply weight exemplar similarity during
+aggregation. Thus, beyond formally specifying the role of contrast in creative
+generation, the PACKER account allows us to evaluate the well-understood
+principles of the exemplar view within the comparatively understudied field of
+conceptual generation. Nonetheless, we will discuss integrating our approaches
+in the General Discussion.
+
+\subsubsection{Relation Between PACKER and Copy-And-Tweak}
+\label{section:copytweak-packer} The PACKER model is similar to the
+copy-and-tweak model reported by \cite{jern2013probabilistic}: Both models are
+exemplar-based, and both models generate new items according to their similarity
+to known members of the target class. However, PACKER diverges from the
+copy-and-tweak model by including a contrast mechanism, enabling generation
+according to dissimilarity to members of opposing categories. By consequence,
+copy-and-tweak can be realized as a parameterization of the PACKER model that is
+insensitive to category contrast. Specifically, when $\theta_c = 0$ and
+$\theta_t > 0$ (see Figure \ref{fig:packer-examples}, panel B),
+$f(x_j) = 0$ for $x_j$ belonging to contrast categories; thus, PACKER is not
+influenced by these items, and is mathematically equivalent to a copy-and-tweak
+approach.
+
+In this paper, we report simulations using this copy-and-tweak model. This model
+fits within the exemplar-based category generation framework defined above, and
+is a continuous-dimension adaptation of the model tested by
+\cite{jern2013probabilistic}. By formalizing a model family where PACKER and
+copy-and-tweak are different parameterizations of models within the same
+framework, the comparison between PACKER and copy-and-tweak provides a test of
+the explanatory value of the contrast mechanism based on exemplar dissimilarity:
+The account provided by copy-and-tweak will only equal that of PACKER if the
+contrast mechanism does not offer an advantage (i.e., if $\theta_c > 0$
+significantly improves model fits). Note that the purpose of the article is to
+explore and formally analyze the role of contrast in category generation and
+thus, we leave extending PACKER to incorporate distributional factors \citep[as
+explored by][]{jern2013probabilistic} for future work.
+
+\subsection{Contrast as representativeness: An extension to Jern and Kemp
+(2013)}
+\label{section:representativeness-definition}
+
+An alternate conceptualization of category contrast is the idea of
+representativeness -- exemplars are generated such that they are more
+representative of the target category and thus more distinct from and less
+similar to other categories. We adopt the definition specified by
+\cite{tenenbaum01}, where the representativeness $R$ of an item (or in our case,
+an exemplar) $y$ is the relative amount of evidence that is provided by $y$ for
+a given hypothesis $h$ compared to all other hypotheses $H'$. More formally,
+\begin{equation}
+  R(x,h) = \log \dfrac{p(x|h)}{\sum_{h'\in H'}{p(x|h')p(h')}} , 
+\label{eq:representativeness}
+\end{equation}
+
+In Equation \ref{eq:representativeness}, $p(h')$ is the prior probability
+distribution on the hypothesis space that excludes $h$. Note that when there are
+only two hypotheses (e.g. two categories) involved in the domain, this prior
+takes the value of $1$ since there is ever only one alternative hypothesis.
+
+This Bayesian formalization of representativeness makes it straightforward to
+implement into the existing hieriarchical Bayesian model of category generation
+developed by \cite{jern2013probabilistic}. Specifically, both models assume that
+exemplars are sampled from a given category distribution ($h$ in
+Equation \ref{eq:representativeness}). Each category distribution is a
+multivariate normal parameterized by a location vector $\mu_k$ and covariance
+matrix $\Sigma_k$ (i.e., $h = \mathcal{N}(\mu_k,\Sigma_k)$). The parameters
+$\mu_k$ and $\Sigma_k$ are assumed to be samples from a prior
+normal-inverse-Wishart (NIW) distribution parameterized by
+$\mu_0,\Sigma_0,\kappa,\nu$. Mathematically,
+\begin{align}
+  \mu_k, \Sigma_k|\mu_0,\Sigma_0,\kappa,\nu &\sim
+                                              \textrm{NIW}(\mu_0,\Sigma_0,\kappa,\nu)
+                                              \label{eq:mvnsample}\\
+  y|C &\sim \mathcal{N}(\mu_k,\Sigma_k) \label{eq:exemplarsample}
+\end{align}
+
+In our simulations, we set the prior mean $\mu_0$ to the center of the stimulus
+space and $\Sigma_0 = \lambda \mathbf{I}$ where $\lambda$ is a free parameter
+and $\mathbf{I}$ is a $d$-by-$d$ identity matrix with $d$ representing the
+number of dimensions or features in the domain. The $\kappa$ and $\nu$
+parameters are freely estimated within the constraints $\kappa > d-1$ and
+$\nu > 0$.
+
+With the assumption of the NIW prior, the location vector $\mu_k$ can be computed
+with:
+\begin{equation}
+  \mu_k = \dfrac
+    {\kappa\mu_{0} + n_k \bar{x}_k}
+    {\kappa + n_k}
+    \label{eq:category_mus}
+\end{equation}
+where $n_k$ is the number of observed exemplars in category $k$ and $\bar{x}_k$
+is the observed category mean. Note that if there are no observed exemplars in
+the target category (i.e., if the model is generating a completely novel
+category), Equation \ref{eq:category_mus} simplifies to $\mu_k = \mu_0$.
+
+The NIW prior also allows us to infer the category covariance matrix $\Sigma_k$
+by computing the following:
+\begin{equation}
+  \Sigma_k = [\Sigma_D\nu + C_k +
+    \dfrac
+    {\kappa n_k}
+    {\kappa + n_k}
+    (\bar{x_k}-\mu_k)(\bar{x_k}-\mu_k)^T
+  ] (\nu + n_k)^{-1}
+  \label{eq:Sigma_y}
+\end{equation}
+where $C_k$ is the observed category covariance and $\Sigma_D$ is the domain
+covariance matrix from which $\Sigma_k$ samples are obtained. We can
+infer $\Sigma_D$ from the observed category covariances $C$ and the prior
+covariance $\Sigma_0$. Specifically,
+\begin{equation}
+    \Sigma_D = \Sigma_0 + \sum_{k}{C_k}
+\end{equation}
+
+At this point, both the representativeness model and the hierarchical Bayesian
+model from \cite{jern2013probabilistic} are largely identical. However, both
+models diverge in their computation of exemplar generation probabilities. While
+the original hierarchical Bayesian model produces novel exemplars with
+probabilities proportional to the multivariate normal likelihood $p(y|h)$, the
+representativeness model generates new exemplars with probabilities weighted by
+their representativeness $R(y,h)$. In practice, the probability of generating a
+particular candidate $y$ is obtained using an Exponentiated
+\citet{luce1977choice} choice rule:
+\begin{equation}
+  p(y) = \dfrac
+    {\exp( \theta \cdot R(y,h))}
+    {\sum_k \exp(\theta  \cdot R(y_k,h_k))} .
+\label{eq:representative-choice}
+\end{equation}
+where $\theta$ is a freely estimated response determinism parameter (constrained
+such that $\theta \geq 0$).
+
+Despite both the original hierarchical Bayesian model and our representativeness
+model sharing identical hierarchical structures, their distinct response
+processes can yield very different exemplars. Specifically, the hierarchical
+Bayesian model emphasizes the generation of exemplars that maintain
+distributional commonalities across categories. In contrast, the
+representativeness model focuses on the generation of exemplars that are
+representative of the underlying distribution for a target category.
+\jlanote{The next sentence is weird sounding and should be cleaned up.}
+\sxlnote{Ok! I hope it reads better now.} Generally, this mechanism of
+representativeness results in the generation of exemplars that are less similar
+to exemplars from other categories. We see this illustrated in Figure
+\ref{fig:representative-examples}, where the representativeness model displays a
+strong preference for generating exemplars from target category 'B' that are
+further away from the contrast category 'A'.
+
+\begin{figure}
+	\begin{center} \inputpgf{figs/}{hbm-examples.pgf}
+		\caption{Generation of a category `B' example, following exposure
+to one member of category `A' and one member of category `B'. Predictions are
+shown for {\em(a)} the original hierarchical Bayesian model and {\em(b)} the
+representativeness model.}
+		\label{fig:representative-examples}
+	\end{center}
+\end{figure}
+
+\subsection{Synopsis and Prognosis}
+
+Research on the creative generation of novel concepts has focused on the finding
+that generated categories tend to possess distributional commonalities with
+known categories. However, a fundamental goal of concept generation is to create
+something {\em new} (i.e., different from what is already known). The manner in
+which generated categories differ from known ones is, nonetheless, poorly
+understood: Existing theories do not make strong predictions about how
+creatively generated concepts should systematically differ from existing ones.
+Above, we found encouraging initial support that contrast influences category
+generation \citep[][Experiment 3]{jern2013probabilistic}, and we introduced two
+novel approaches to formalizing the roles of similarity and contrast in creative
+generation.
+
+In the sections below, we present two experiments demonstrating systematic
+effects of category contrast on category generation inspired by factors
+influencing how the contrast models generate new categories. Our experiments are
+based on \cite{jern2013probabilistic}'s paradigm, which is a straightforward
+translation of the traditional artificial classification paradigm to the task of
+category generation: Participants are first exposed to a single,
+experimenter-defined category, and are then asked to generate members of a new
+category. We then report formal analyses comparing the exemplar dissimilarity
+and the representativeness accounts of our results to that of the hierarchical
+Bayesian and copy-and-tweak models developed by \cite{jern2013probabilistic}.
+After analyzing the category generation studies, we present third experiment to
+demonstrate the effect of contrast in category learning. This ensures that our
+results are not idiosyncratic to category generation.
+
+\section{Experiment 1}
+
+To begin our investigation, we sought to extend the early evidence obtained from
+our analysis of the \cite{jern2013probabilistic} data, under a variety of
+learning conditions and using more standard stimulus materials. We used an
+artificial stimulus design: A two dimensional domain of squares, varying in
+color and size (see Figure \ref{fig:e1-conditions}, panel A). These dimensions
+have been used in numerous classification learning studies
+\cite[e.g.,][]{conaway2016similar,conaway2016generalization,shepard1961learning,nosofsky1994comparing}.
+Unlike those used in the \cite{jern2013probabilistic} experiments, distance on
+these physical dimensions aligns more directly with perceptual similarity,
+allowing us to evaluate the role of category contrast in creative generation
+more precisely. To extend the evidence provided by the
+\cite{jern2013probabilistic} data, we tested the effects of category contrast
+after learning one category from a set of qualitatively distinct category
+structures, as shown in Figure \ref{fig:e1-conditions}.
+
+Figures \ref{fig:e1-conditions}b-d show the values of exemplar dimensions
+belonging to the experimenter-defined categories (`A', or `Alpha') that
+participants were assigned to learn about prior to generating a new category.
+Each participant learned one of the category types during training. In the
+`Cluster' type, category A is a tight cluster of examples in the space.
+Perceptually instantiated, the members of category A might, for example, be
+large and dark in color. In the `Row' type, category A has a row pattern across
+the space, varying along one feature but not the other. Thus, its members might
+all be dark in color but would vary in size. Finally, in the `XOR' type, the
+experimenter-defined category consists of two clusters separated in opposite
+corners of the space, conforming to the exclusive-or logical structure (e.g.,
+members are small and dark or large and light).
+
+It should be noted that in our experiments the assignment between the perceptual
+and conceptual dimensions (e.g., $X \rightarrow Size$, $Y \rightarrow Color$),
+as well as the direction of variation along each dimension (e.g., $dark
+\rightarrow light$ or $light \rightarrow dark$) was counterbalanced across
+participants. The category types in Figure \ref{fig:e1-conditions} are plotted
+in a conceptual space, rather than a perceptual space. Thus, while the
+conceptual organization of the category types remains constant, each category
+type may have a different physical instantiation according to the counterbalance
+assignment. For example, the Cluster type may be large and dark in color, or it
+may be small and light in color, depending on the assignment and direction of
+the dimensions. For this reason, below we will discuss generation within a
+conceptual space, rather than a physically instantiated one.
+
+\begin{figure}
+    \begin{center} \inputpgf{figs/}{e1-conditions.pgf}
+    \caption{Stimulus domain and category types tested in Experiment 1. Stimuli
+are not drawn to scale.}
+    \label{fig:e1-conditions}
+    \end{center}
+\end{figure}
+
+After learning about an experimenter-defined category, participants are asked to
+generate examples of a new category. Within this paradigm, an effect of category
+contrast would be realized if participants prefer to generate items in locations
+that are distant (i.e., perceptually dissimilar) from members of category A.
+However, generation is left unconstrained. Critically, participants were not
+asked to generate something different in the prompt. For example, participants
+assigned to the Cluster condition may generate a tightly clustered category in
+the corner opposite of the experimenter-defined category. Alternatively, they
+may generate a tightly clustered category directly overlapping with the
+experimenter-defined category. Further, they may even generate an entirely
+different type of category (e.g., a row category).
+
+Our experimental results also provide a converging test of the classic finding
+that generated categories tend to share distributional properties with known
+categories in the domain \citep{jern2013probabilistic,ward1994structured}. From
+these results, we can predict that, in each condition, participants should
+generate categories that are distributionally similar to the
+experimenter-defined category: In the Cluster condition, generated categories
+should be tightly clustered. In the Row condition, generated categories should
+vary more along the X-axis than the Y-axis. In XOR condition, generated
+categories should be widely distributed across both dimensions, and the two
+dimensions should be positively correlated.
+
+Interestingly, the XOR condition also offers a dissociation between the roles of
+category contrast and the emulation of distributional structure:
+widely-distributed, positively-correlated categories would need to lie along the
+positive diagonal of the space (that is the only place they ``fit''), which is
+already occupied by the experimenter-defined category. Thus, if contrast plays a
+role, exemplars in the generated categories of participants in the XOR condition
+may not be positively correlated -- they may be negatively correlated instead.
+In this case, contrast and statistical regularities would interact, which would
+be inconsistent with the leading approach in conceptual generation
+\citep{jern2013probabilistic}
+
+\subsection{Participants and Materials}
+
+183 participants were recruited from Amazon Mechanical Turk. Each participant
+was randomly assigned to one condition: 64 participants were assigned to the
+Cluster condition, 61 were assigned to the Row condition, and 58 were assigned
+to the XOR condition (sample sizes differ due to random assignment). Stimuli
+were squares varying in color (grayscale 9.8\%--90.2\%) and side length
+(3.0--5.8cm), see Figure \ref{fig:e1-conditions}. The assignment of perceptual
+features (color, size) to axes of the domain space (x, y), as well as the
+direction of variation along each axis (e.g., $dark \rightarrow light$ or $light
+\rightarrow dark$) was counterbalanced across participants.
+
+\subsection{Procedure}
+
+As noted in the introduction of this paper, the task of generating members of a
+new category can be aligned with common tasks studied in the categorization
+literature: Whereas classification consists of predicting an object's category
+label on the basis of its features, inference consists of predicting an observed
+feature, given a set of observed features and a category label. Generation thus
+consists of predicting {\em all} features of an object, given a novel category
+label. In keeping with this insight, we designed our generation task as an
+extension of the traditional artificial classification learning paradigm. The
+task differs from traditional work in creative cognition primarily through the
+use of an artificial domain, which enables the application of computational
+models. The use of an artificial domain also requires the addition of a training
+phase, during which participants learn about the categories in the domain. As a
+result, unlike most previous studies \citep[e.g.,][]{ward1994structured},
+participants in our studies have no experience with the domain before the start
+of the experiment, and the experimenter-defined categories are not
+hierarchically structured (as are many natural categories).
+
+Participants began the experiment with a short training phase (3 blocks of 4
+trials), where they observed exemplars belonging to the `Alpha' category.
+Participants were instructed to learn as much as they can about the Alpha
+category, and that they would answer a series of test questions afterwards. On
+each trial, a single Alpha category exemplar was presented, and participants
+were given as much time as they desired to observe it before moving on to the
+next trial. Each block consisted of a single presentation of each of the members
+of the Alpha category, in a random order. Participants were shown the range of
+possible colors and sizes prior to training.
+
+Following the training phase, participants were asked to generate four examples
+belonging to another category called `Beta'. As in
+\citet{jern2013probabilistic}, generation was completed using a sliding-scale
+interface. Two scales controlled the values of the two dimensions (color, size)
+for the generated example. An on-screen preview of the example updated whenever
+one of the features was changed. Participants could generate any example along
+an evenly-spaced 9x9 grid (including members of the Alpha category), except for
+any previously generated Beta exemplars. Neither the members of the Alpha
+category nor the previously generated Beta examples were visible during
+generation. Prior to beginning the generation phase, participants read the
+following instructions:
+
+\begin{displayquote} As it turns out, there is another category of geometric
+figures called ``Beta''. Instead of showing you examples of the Beta category,
+we would like to know what you think is likely to be in the Beta category.
+
+You will now be given the chance to create examples of any size or color in
+order to show what you expect about the Beta category. You will be asked to
+produce 4 Beta examples - they can be quite similar or quite different to each
+other, depending on what you think makes the most sense for the category.
+
+Each example needs to be unique, but the computer will let you know if you
+accidentally create a repeat.
+\end{displayquote}
+
+\subsection{Results}
+
+We observed a substantial degree of individual differences in our data. In
+Figure \ref{fig:e1-samples} we have plotted sample data from several
+participants, from which it is evident that different participants generated
+qualitatively different category structures. In this section we will focus on
+analyzing the data in aggregate, but in later sections we will explore how
+individual differences can be explained.
+
+\begin{figure}
+    \begin{center} \inputpgf{figs/}{e1-samples.pgf}
+    \caption{Sample categories generated by participants in Experiment 1.
+Representative samples from common generation profiles are shown.}
+    \label{fig:e1-samples}
+    \end{center}
+\end{figure}
+
+To evaluate the role of contrast, we computed the number of times each stimulus
+was generated, as a function of its average city-block distance from members of
+the experimenter-defined ``Alpha'' category. These data, shown in Figure
+\ref{fig:e1-distanceplots}, reveal a clear pattern: Examples that are more
+distant from members of the experimenter-defined categories are more likely to
+be generated into a new category. This supports the notion that contrast is a
+fundamental constraint on creative cognition and that suggests that statistical
+regularity alone is insufficient.
+
+Figure \ref{fig:e1-distanceplots} also depicts, for each participant, the
+average distance of members within the generated category ({\em within-category}
+distance) against the average distance between members of the generated and
+experimenter-defined category ({\em between-category} distance). The narrow
+distribution of between-category distances in the XOR condition reflects the
+widely distributed nature of the experimenter-defined category, reducing the
+possible distances to members of the participant-generated category. These data
+reveal a systematic pattern: The majority of participants generated categories
+with greater between-category distance than within-category distance. That is,
+members of the generated category tended to be more similar to one another than
+to members of the experimenter-defined category. To evaluate this claim
+quantitatively, we conducted t-tests comparing the amount of within- and
+between- class distance in each condition. All conditions possessed greater
+between-category distance: Cluster, $t(63) = 11.43$, $p < 0.001$; Row, $t(60) =
+13.16$, $p < 0.001$; and XOR, $t(57) = 3.64$, $p < 0.001$. These results provide
+further evidence of an effect of category contrast: Participants prefer to
+generate categories that are dissimilar to the learned category but maintain
+some level of internal cohesion.
+
+\begin{figure}
+    \begin{center} \inputpgf{figs/}{e1-distanceplots.pgf}
+    \caption{Experiment 1 results. {\em Left}: Frequency of generation as a
+function of distance from members of the experimenter-defined category. {\em
+Right}: Scatter plot of within-category versus between-category distance in each
+of the participant-generated categories.}
+    \label{fig:e1-distanceplots}
+    \end{center}
+\end{figure}
+
+A secondary goal of this experiment was to examine whether we replicate the
+classic result that generated categories often possess the same distributional
+properties as previously-known categories. For each generated category, we
+computed the category range along each axis (X, Y), as well as the correlation
+between features. These data, shown in Figure \ref{fig:e1-statsboxes}, reveal
+broad individual differences: Within each condition, participants generated
+categories spanning the entire X- and Y- axis as well as categories that spanned
+very little along each. Likewise, in each condition participants generated
+categories possessing strongly positive, neutral, and strongly negative
+correlations between the dimensions. Comparing the distributional statistics
+between conditions yields a broad yet, as we will see, misleading replication of
+the classic effect.
+
+With respect to ranges along each axis (X, Y), the generated categories from
+each condition tend to reflect the ranges of the experimenter-defined
+categories. The categories generated in the Cluster condition were less widely
+distributed along the X-axis compared to Row, $t(123) = 5.61$, $p < 0.001$, and
+XOR, $t(120) = 2.68$, $p = 0.008$. Categories generated in the XOR condition
+were also less widely distributed along the X-axis compared to Row, $t(117) =
+2.56$, $p = 0.012$. This latter effect was not expected because the
+experimenter-defined categories for XOR and Row had similar X-ranges. However,
+the key finding is that categories from the Cluster condition tended to be more
+tightly distributed along the X-axis.
+
+\begin{figure}
+    \begin{center} \inputpgf{figs/}{e1-statsboxes.pgf}
+    \caption{Box-plots of the distributional statistics from the categories
+generated in Experiment 1. Boxes depict the median and quartiles of each
+condition, with whiskers placed at 1.5 IQR. All points outside this region are
+marked individually.}
+    \label{fig:e1-statsboxes}
+    \end{center}
+\end{figure}
+
+Likewise, categories generated in the Row condition had less Y-axis range
+compared to Cluster, $t(123) = 4.57$, $p < 0.001$ and XOR, $t(117) = 9.26$, $p <
+0.001$, and categories from the Cluster condition had less Y-axis range compared
+to XOR, $t(120) = 3.95$, $p < 0.001$. As expected, the correlations in the
+Cluster and Row conditions were not systematically positive or negative ($ps >
+0.1$). However, the generated categories in the XOR condition tended to possess
+{\em negatively} correlated dimensions, $t(57) = 2.04$, $p = 0.046$. This
+finding is notable, as it is the opposite of what would be expected, assuming
+learners are emulating the distributional structure of the experimenter-defined
+class (which possesses perfectly positively correlated features).
+
+We believe that the failure to replicate this finding is because participants in
+\cite{jern2013probabilistic} could differentiate the generated category on a
+third dimension (hue) to maintain the statistical regularities on the other two
+dimensions. Although the correlation in the XOR condition is significantly
+negative, it is clear from the box-plot in Figure~\ref{fig:e1-statsboxes} that
+it would be inappropriate to make a strong conclusion (e.g., the median is close
+to zero). However, we can conclude with confidence that there are situations
+where people do not emulate the distributional structure of the given category.
+This indicates that there is more to category generation than the emulation of
+distributional structure of other categories in the domain. Further, as we will
+discuss in more detail in the model-based analysis section, this is expected by
+our proposal that contrast is a fundamental principle in category generation.
+
+\subsection{Discussion} In Experiment 1 we sought to extend our analysis of the
+\cite{jern2013probabilistic} data by evaluating the influence of category
+contrast on creative generation, given qualitatively different types of prior
+knowledge. We found strong evidence for effects of category contrast in each
+condition: Participants were more likely to generate stimuli that are more
+distant from (i.e., less similar to) members of a previously-learned category,
+and members of participant-generated categories tended to be more similar to one
+another than to members of previously-learned categories. We also partially
+replicated the classic finding that the distributional structure of generated
+categories reflects that of previously learned categories
+\citep{jern2013probabilistic,ward1994structured}: Members of generated
+categories were more widely distributed along dimensions which were widely
+distributed in the experimenter-defined category.
+
+Notably, however, we also found that participants who learned an XOR category
+(composed of exemplars following a positive diagonal, see Figure
+\ref{fig:e1-conditions}) tended to generate items according to a {\em negative}
+feature correlation -- the opposite of what was present in the previously
+learned category. While this may be difficult to account for under existing
+theoretical approaches (which assume generated categories follow the same
+distributional structure as known categories), it can be concisely explained
+from a category contrast perspective. Specifically, within the XOR condition,
+individuals who seek to generate a category that is perceptually distinct from
+what is already known are left with only the upper-left and bottom-right
+quadrants of the space, as members of the previously-learned XOR category lie in
+the bottom-left and top-right. If examples are generated into both of the
+available quadrants, the generated category will possess a strongly negative
+correlation, opposing that of the experimenter-defined class.
+
+Thus, the core results of Experiment 1 indicate that generated categories can
+systematically differ from what we would expect based on prior work. The
+negative (or null) correlations observed in the XOR condition suggests an
+interesting interaction between contrast with a given category and emulation of
+statistical properties. That is, the constraints on creative generation imposed
+by category contrast may not simply influence the {\em location} of generated
+categories, but also their distributional structure. In Experiment 2, we test
+this claim more systematically.
+
+\section{Experiment 2.}
+
+\begin{figure}
+    \begin{center} \input{figs/e2-conditions.pgf}
+    \caption{Category types tested in Experiment 2.}
+    \label{fig:e2-conditions}
+    \end{center}
+\end{figure}
+
+To test whether category contrast influences the distributional structure of
+generated categories, we sought to identify conditions in which differences in
+the distributional structure of generated categories cannot be explained by the
+distributional structure of the experimenter-defined category. We created two
+new category types (depicted in Figure \ref{fig:e2-conditions}) that possess
+identical distributional structures (both are tight clusters of examples with no
+correlation between features), as they only differ in their Y-axis position: the
+`Bottom' category lies in the bottom-center of the space, and the `Middle'
+category lies in the center. The distributional equality of these conditions is
+key to the design of the experiment: If the distributional structure of
+previously learned categories were the only influence on the generated
+categories, we should observe no difference in the categories participants
+generate between these two conditions. Will participants distribute their
+generated category differently between conditions due to the differences in the
+available stimulus space for generating a new category?
+
+If category contrast influences the distributional structure of the categories
+people generate, then we should observe different types of categories according
+to the shape of the space that is {\em unoccupied} by members of previously
+learned categories. The difference in the Y-axis position between the Bottom and
+Middle conditions produces a considerable change to the shape of the unoccupied
+space. Participants assigned to learn the Bottom category should be less likely
+to generate exemplars into the lower regions of the stimulus space (as these
+areas possess greater similarity to members of the Bottom category), preferring
+instead to distribute exemplars across the upper region of the space. This
+constraint is lifted in the Middle condition, as the Middle category exemplars
+are equidistant to the upper and lower regions of the space. Accordingly,
+participants should be more likely to utilize both of these areas. Thus, if
+category contrast influences the distributional structure of generated
+categories, we should observe more participants in the Middle condition that
+generate examples above {\em and} below the experimenter-defined category.
+
+\subsection{Participants, Materials, and Procedure}
+
+122 participants were recruited from Amazon Mechanical Turk. 61 participants
+were randomly assigned to the Middle and Bottom conditions each. The stimulus
+space and procedure were exactly as in Experiment 1. Participants first
+completed a short training phase, followed by the generation phase. The only
+difference from Experiment 1 was the category types given to participants.
+
+\subsection{Results}
+
+\begin{figure}
+    \begin{center} \inputpgf{figs/}{e2-samples.pgf}
+    \caption{Sample categories generated in Experiment 2. }
+    \label{fig:e2-samples}
+    \end{center}
+\end{figure}
+
+As in Experiment 1, we observed broad differences in the generation approach
+taken by different participants. To characterize the nature of these
+differences, Figure \ref{fig:e2-samples} depicts sample categories generated by
+participants. The data from each condition are organized into four columns based
+on commonly observed patterns of generation: a `Cluster' type of
+tightly-clustered examples, `Row' and `Column' types of exemplars widely
+distributed along the one axis but narrowly along the other, and a `Corners'
+type, wherein participants placed exemplars in disparate corners of the space.
+As before, in this section we focus on analyzing the data in aggregate, but in
+later sections we will focus more specifically on explaining the individual
+differences.
+
+\begin{figure}
+    \begin{center} \inputpgf{figs/}{e2-distanceplots.pgf}
+    \caption{Experiment 2 results. {\em Left}: Frequency of generation as a
+function of distance from members of the experimenter-defined category. {\em
+Right}: Scatter plot of within-category versus between-category distance in each
+of the participant-generated categories.}
+    \label{fig:e2-distanceplots}
+    \end{center}
+\end{figure}
+
+We began our analysis by testing for the broad influence of category contrast on
+generation. As in Experiment 1, we computed the frequency each stimulus was
+generated as a function of its average distance from members of the
+experimenter-defined category, as well as each participant's average within- and
+between- category distance. These data, shown in Figure
+\ref{fig:e2-distanceplots}, yield very similar results. Participants generated
+stimuli that are distant from members of the experimenter-defined category, and
+the categories in each condition tended to possess more between-category than
+within-category distance: Bottom, $t(60) = 5.5$, $p < 0.001$; Middle, $t(60) =
+2.71$, $p < 0.01$. We did, however, observe a notable subgroup of participants
+in each condition who generated categories with more within-category than
+between-category distance. Upon manual inspection, many of these individuals
+appear to have assumed a `Corners' strategy, placing exemplars in disparate
+corners of the space, thus producing much more within-category distance, see
+Figure \ref{fig:e2-samples} for examples.
+
+To explore the distributional structure of the generated categories, we computed
+the range of exemplars along each axis (X, Y), as well as the correlation
+between features. These data, shown in Figure \ref{fig:e2-statsboxes}, again
+demonstrate the degree of individual differences observed in our study. In each
+condition, we observed tightly clustered and widely distributed categories along
+each dimension, as well as positively, negatively, and uncorrelated categories.
+
+\begin{figure}
+    \begin{center} \inputpgf{figs/}{e2-statsboxes.pgf}
+    \caption{Box-plots of the distributional statistics from the categories
+generated in Experiment 2. }
+    \label{fig:e2-statsboxes}
+    \end{center}
+\end{figure}
+
+As noted above, if the distributional structure of generated categories is
+influenced by the shape of the space not occupied by members of known
+categories, then participants in the Middle condition would be more likely to
+place exemplars in the upper {\em and} lower regions of the space, as members of
+the experimenter-defined category are equidistant from these regions.
+Participants in the Bottom condition should be less likely to generate category
+members in the bottom regions because members of the experimenter-defined
+category are located there. One way to test these predictions is to analyze the
+Y-axis ranges of the generated categories: If Middle participants utilize the
+upper and lower regions of the space, their categories should vary more along
+the Y-axis. T-Tests comparing the conditions on the distributional statistics,
+however, reveal few between-group differences: the conditions do not differ with
+respect to X-axis range, Y-axis range, or feature correlations ($ps>0.17$).
+
+However, our ability to detect differences in Y-Axis range using a standard
+$t$-test between the conditions is, in this case, diminished due to the
+non-normality of the data (Shapiro-Wilk normality test $W=0.77, p< 0.001$ for
+the Middle condition and $W=0.85, p < 0.001$ for the Bottom condition). Figure
+\ref{fig:e2-yranges} depicts the Y-axis position of the exemplars generated
+within each participant's category. The categories are sorted by overall range,
+then by condition assignment. These data reveal that there were nearly as many
+participants who generated categories spanning the entire Y-axis as those who
+generated categories spanning almost none of the Y-axis. The non-normality of
+the Y-axis range distributions thus requires that we use a different approach to
+addressing the experiment's main question.
+
+\begin{figure}
+    \begin{center} \inputpgf{figs/}{e2-yranges.pgf}
+    \caption{Y-Axis range and position of the participant-generated categories
+from Experiment 2. Each line corresponds to a participant's category, with
+notches corresponding to the Y-axis position of exemplars within the category
+(notches may overlap). Participants are sorted by overall range, and then by
+condition. }
+    \label{fig:e2-yranges}
+    \end{center}
+\end{figure}
+
+Because our main prediction concerns the generation of exemplars within the
+upper and lower regions of the domain, we compared the conditions in terms of
+the frequency with which participants generated examples above and below the
+categories. Specifically, we counted the number of participants in each
+condition who placed at least one `Beta' exemplar on the top and bottom `rows'
+of the space (the maximum and minimum possible Y-axis value, respectively). The
+resulting contingencies data are shown in Table \ref{table:e2-subset-table}.
+
+Firstly, it should be noted that nearly every participant utilized the top
+and/or bottom rows: only $10 / 122$ participants generated their category
+entirely within the interior region. Fisher's Exact Tests comparing the
+conditions reveal that more Middle participants generated an exemplar in the
+bottom row, $p < 0.001$, again demonstrating the role of contrast in guiding
+where exemplars are generated. The conditions did not differ in use of the top
+of the space, $p = 0.16$, however, more Middle participants placed exemplars in
+the top {\em and} bottom rows, $p = 0.038$. The latter effect is of interest
+here, as it indicates that the shape of the unoccupied space exerts some
+influence on the distributional structure of generated categories: Participants
+in the Middle condition were more likely to generate a category spanning the
+entire Y-axis. Thus, distributional structure of the generated categories can be
+influenced by category contrast alone.
+
+\begin{table}
+\begin{center}
+\caption{Experiment 2 results.}
+\label{table:e2-subset-table} \vskip 0.12in
+\begin{tabular}{ l r r} \textbf{Middle} & Used top row & No top row \\ \hline
+Used bottom row & 28 & 18 \\ No bottom row & 11 & 4 \\ \\ \textbf{Bottom} & Used
+top row & No top row \\ \hline Used bottom row & 16 & 8 \\ No bottom row & 31 &
+6 \\
+\end{tabular}
+\end{center}
+\end{table}
+
+\subsection{Discussion}
+
+In Experiment 2, we replicated the core findings from Experiment 1. Stimuli are
+more likely to be generated if they are distant from exemplars in other
+categories, and most participants generate categories with more between-category
+than within-category distance. However, we additionally found that the {\em
+position} of a previously learned category (rather than its distributional
+structure) influences the types of categories people generate: Participants who
+learned the `Middle' type were more likely to generate categories spanning the
+entire Y-axis of the space. Participants who learned the `Bottom' type were less
+likely to do so as a result of the presence of opposite category exemplars in
+the lower regions of the space.
+
+This finding cannot be explained from the perspective that the distributional
+structure of previously learned categories is the sole determinant of the
+distributional structure of generated categories. However, the observed behavior
+is expected from a category contrast perspective: Participants seeking to
+generate a perceptually distinct category will be more likely to use areas of
+space that are unoccupied by exemplars belonging to previously learned
+categories. In the Middle condition, the upper and lower regions of space are
+equidistant from members of the experimenter-defined category, whereas in the
+Bottom condition, the lower region of the space is closer to members of the
+experimenter-defined category. Thus, while Middle participants may form
+categories around the use of the equally unoccupied areas, the same is not true
+for the Bottom condition.
+
+\section{Model-based Analyses}
+Experiments 1 and 2 revealed systematic and strong effects of category contrast
+on creative generation. In this section, we report the results of simulations
+with formal models aimed at explaining our observations. Specifically, we
+present simulations from the two novel contrast models: PACKER and the
+representativeness model, and compare them to two models that do not incorporate
+contrast: the copy-and-tweak model (discussed in Section
+\ref{section:copytweak-packer}) and an implementation of the hierarchical
+Bayesian model proposed by \cite{jern2013probabilistic}, described in-depth in
+Appendix \ref{ap:hsampling-definition}. For our simulations here, the
+copy-and-tweak model is defined as a variant of PACKER with the $\theta_c$
+parameter constrained to be zero. The comparison of this set of models serves to
+highlight the explanatory role of contrast in categorization: If contrast
+affords little explanatory advantage, then there should be few differences in
+performance between PACKER and copy-and-tweak, or between the representativeness
+and hierarchical Bayesian model. The comparison between these models can also
+emphasize the necessity of contrast and demonstrate that generation cannot be
+explained entirely through the emulation of distributional structure. Each model
+has complementary strengths and weaknesses: Whereas PACKER and copy-and-tweak
+are insensitive to the distributional structure of learned categories (relying
+only on exemplar similarities), the representativeness and hierarchical Bayesian
+model generates categories exclusively on the basis of knowledge of how existing
+classes are distributed.
+
+Our approach in this section is to first broadly evaluate and compare the
+quality of each model's account to our entire dataset (Experiments 1 and 2
+combined), then analyze the ability for each model to explain individual
+differences in each experiment, and lastly we describe the strengths and
+weakness of each model's account of category generation.
+
+\subsection{Parameter-Fitting}
+
+To obtain a global measure of the quality of each model's account, we fitted the
+parameters of each model to our entire dataset (Experiments 1 and 2 combined),
+using a hill-climbing algorithm which maximized the log-likelihood of the
+model's predictions of the observed responses (1220 responses from 305 total
+participants). We fitted three parameters in the PACKER model ($c$, $\theta_t$,
+and $\theta_c$; see Section \ref{section:PACKER-definition}), as well as four in
+the representativeness model and the hierarchical Bayesian model ($\kappa$,
+$\lambda$, $\nu$, and $\theta$; see Section
+\ref{section:representativeness-definition} and Appendix
+\ref{ap:hsampling-definition} respectively). We fitted only two parameters for
+the copy-and-tweak model ($c$, and $\theta_t$), as $\theta_c$ is held constant
+($\theta_c = 0$). Attention ($w$, see Equation \ref{eq:similarity}) in PACKER
+and copy-and-tweak was set uniformly. Parameters were not allowed to vary
+between participants or conditions -- the goal was to obtain the best-fitting
+values to our entire dataset.
+
+\begin{table} \centering
+\caption{Results of model-fitting to the combined datasets from Experiments 1
+and 2. Note that smaller AIC values correspond to better model fits (adjusted
+for number of parameters)}
+\label{table:global-model-fits}
+\begin{tabular}{ l l l l} \\ \textbf{PACKER} & \textbf{Copy \& Tweak} &
+\textbf{Representativeness} & \textbf{Hierarchical Bayesian} \\ \hline $AIC =
+  9069$ & $AIC = 9813$ & $AIC = 8783$ &  $AIC =
+9881$ \\ $L = -4531$ & $L = -4905$ & $L = -4388$ & $L = -4937$ \\ $c = 0.51$ &
+                                                                                $c = 3.22$ & $\kappa = 12.23$ &
+$\kappa < 0.001$ \\ $\theta_c = 3.09$ & $\theta_c = 0$ (fixed) & $\nu = 1.00$ & $\nu = 5.44$ \\
+$\theta_t = 3.47$ & $\theta_t = 3.00$ & $\lambda = 7.04$& $\lambda
+                                                                       = 0.06$
+  \\ & &$\theta = 10.21 $& $\theta = 3.09$
+\\
+\end{tabular}
+\end{table}
+
+Table \ref{table:global-model-fits} contains the results of this fitting
+procedure. Due to the uneven number of fitted parameters among the models, we
+compare the model fits using the Akaike Information Criterion
+\citep[AIC;][]{akaike1974new}, where smaller values correspond to better fits
+(discounted by model complexity as measured by the number of parameters). The
+same qualitative results were obtained with alternative model comparison metrics
+(e.g., BIC, \citealp{schwarz1978estimating}; $AIC_C$,
+\citealp{hurvich1989regression}).
+
+Table \ref{table:global-model-fits} contains the AIC values of each model's best
+fit, as well as the corresponding log-likelihood ($L$) and the best-fitting
+parameter values. These results reveal strong model differentiation: both
+contrast models (PACKER and the representativeness model) achieved far better
+fits compared to their non-contrast counterparts: copy-and-tweak and the
+hierarchical Bayesian model respectively. Interestingly between the contrast
+models, the hierarchical model (representativeness) outperformed the
+exemplar-based theory (PACKER), whereas between the non-contrast models, the
+reverse is observed. Specifically, here the exemplar-based model
+(copy-and-tweak) performed somewhat better than the hierarchical Bayesian model.
+
+While PACKER's advantage may tentatively be attributed to the model's
+sensitivity to category contrast (this will be explored in detail below), the
+advantage shown by copy-and-tweak over the hierarchical Bayesian model may be
+attributed to its exemplar-based representation, as opposed to the
+prototype-based representation assumed by the hierarchical Bayesian model. As
+observed in Figures \ref{fig:e1-samples} and \ref{fig:e2-samples}, the generated
+categories we observed were often widely distributed, with no items near the
+category prototype. This aspect of the data is inconsistent with the
+multivariate normal distributions (similar to prototypes) used to represent
+categories in the \cite{jern2013probabilistic} model, but can be easily
+accounted for using an exemplar-based approach.
+
+A key distinction between the contrast and classical models is that only the
+contrast models are capable of making strong predictions about the location of
+new category members when the target class is entirely novel (i.e., no member of
+the category has been observed). Under these circumstances, there are no
+examples to copy, and thus the copy-and-tweak model predicts that items are
+generated at random. Likewise, with no observations on which to condition the
+category distribution, the hierarchical Bayesian model also picks an item at
+random.
+
+Thus, it is possible that the failure of the classical models is simply due to
+their inability to explain each participant's first trial (generating the first
+item in the `Beta' category). To ensure this is not driving our results, we
+conducted an identical set of simulations as above, excluding this trial
+(leaving 915 responses in the dataset): Again, the representativeness model ($L
+= -3286$, $AIC = 6580$) and PACKER ($L = -3377$,
+$AIC = 6759$) achieved better fits than the copy-and-tweak ($L = -3564$,
+$AIC = 7132$) and hierarchical Bayesian ($L = -3597$, $AIC = 7201$) models.
+
+Finally, because copy-and-tweak is nested within PACKER, we can use a likelihood
+ratio test to compare the two models. PACKER explains the aggregate data
+significantly better than copy-and-tweak ($\chi^2(1) = 747, p < 0.001$ for all
+data and $\chi^2(1) = 375, p < 0.001$ excluding the first example), providing
+further evidence that category generation is better explained when contrast is
+considered.
+
+Through comparison with the copy-and-tweak model, Figure
+\ref{fig:packer-loglike} more clearly demonstrates the robustness of the
+explanatory gains yielded by PACKER's category contrast mechanism. It displays
+the log-likelihood of the participants' results under PACKER as a function of
+the $\theta_c$ parameter. The model's other parameters ($c$,
+$\theta_t$) were set according to copy-and-tweak's best fits from Table
+\ref{table:global-model-fits}, and thus when $\theta_c=0$, the models are
+equivalent. The figure clearly shows a ``sweet spot'': a convex region in which
+PACKER achieves superior fits as a result of changes to $\theta_c$. The
+best fitting values lie well above the value of 0 assumed by the copy-and-tweak
+model, which demonstrates the robustness of the contrast effect (though note
+PACKER achieves even better fits when its parameters are fitted together, as in
+Table \ref{table:global-model-fits}). In sum, the data are better explained when
+both within-category similarity and category contrast is considered.
+
+\begin{figure}
+    \begin{center} 
+      \includegraphics[width=\textwidth]{figs/packer-loglike.pdf}
+      \caption{PACKER's fit as a function of its prioritization of
+        within-category and between-category similarity (using the $\theta_c$
+        parameter) . To facilitate comparison, PACKER's other parameters ($c$,
+        $\theta_t$) were set to the best fitting values obtained for
+        copy-and-tweak in Table \ref{table:global-model-fits}. The black diamond
+        marker indicates the log-likelihood for the point where
+        $\theta_c=\theta_t$. }
+    \label{fig:packer-loglike}
+    \end{center}
+\end{figure}
+
+\subsection{Individual Differences}
+\label{section:individual-diffs}
+
+As noted in Experiments 1 and 2, we observed a great deal of individual
+differences in the types of categories that participants generated. Within each
+condition, there were a wide variety of category types, such as row and column
+categories (see Figures \ref{fig:e1-samples} and \ref{fig:e2-samples}). The
+simulations reported above serve to evaluate the models while considering the
+entire dataset, but a secondary goal of any formal account should be to provide
+some explanation of how different profiles of performance emerge. Many of the
+individual generation profiles we observed can be described within the PACKER
+framework, simply by tuning the model's parameters in a principled manner. In
+this section, we describe more specifically how the most frequently observed
+profiles can be realized.
+
+By manual inspection, it is evident that the most common profiles of generation
+consist of: (A) a tightly-distributed `cluster' of examples, (B) `row'- and
+`column'-like arrangements (varying widely along one dimension but not the
+other), and (C) a `corners' arrangement with examples placed into disparate
+corners of the space. These four profiles are distinct in terms of the
+distribution of the generated category along each dimension: Whereas the cluster
+profile is tightly distributed along both dimensions, the row and column
+profiles are tightly distributed along just one dimension. Finally, the corners
+profile is widely distributed along both dimensions.
+
+In the framework proposed by PACKER, the cluster and corners profiles arise
+based on different prioritization of within-category similarity versus
+between-category contrast, and the row and column profiles arise based on the
+prioritization of each dimension in the computation of similarity. For example,
+in the cluster profile, there is a high degree of within-category similarity
+along both dimensions, whereas in the corners profile there is minimal
+within-category similarity. Thus, PACKER's proposal is that these individual
+differences arise as a result of different priorities: While the tight cluster
+configuration can be considered PACKER's `default' mode (as it maximizes
+within-category similarity), the corners profile can be produced when
+between-category contrast is put at a higher priority (i.e., $\theta_c > \theta_t$).
+
+Likewise, in the row and column profiles, there is a large degree of
+within-category similarity along one dimension but not the other. These
+differences likely arise due to a differential focus on one dimension over
+another, and thus they can be produced by changes to PACKER's attention weights,
+$w_1$ and $w_2$ (see Equation \ref{eq:similarity}). Traditionally, the attention
+weights in exemplar models are thought to reflect the diagnostic value of each
+dimension towards classifying the known category members
+\citep{nosofsky1984choice,nosofsky1986attention,kruschke1992alcove}, but within
+a generation context the weights specify the importance of within- and
+between-category similarity along each dimension. For example, if all of
+attention is allocated along the X-axis ($w_1=1$ and $w_2=0$), similarity along
+the Y-axis no longer influences performance. As a result, PACKER will create
+categories that are more widely distributed along the Y-axis, as similarity is
+not taken into account along that dimension. As a general principle,
+differentially weighting one dimension will result in the generation of
+categories that are more widely distributed along the ignored dimension,
+conforming to a row- or column-like arrangement. See Figure
+\ref{fig:packer-attention} for a depiction of how attention influences PACKER's
+performance.
+
+\begin{figure}
+    \begin{center} \inputpgf{figs/}{packer-attention-examples.pgf}
+    \caption{PACKER generation of a category `B' example, following exposure to
+one member of category `A' and one member of category `B'. Predictions are shown
+for different attention settings: {\em (a)} Increased weighting of the X-axis.
+{\em (b)} Increased weighting of the Y-axis. {\em (c)} Uniform weighting
+(identical to Figure \ref{fig:packer-examples}).}
+    \label{fig:packer-attention}
+    \end{center}
+\end{figure}
+
+As in PACKER, changes in the parameter settings of the copy-and-tweak model can
+also be used to produce different patterns of generation. Indeed, as
+copy-and-tweak is simply a special case of the PACKER model, the attention
+weights operate exactly as described above to produce row- and column-like
+categories. However, because the model is not influenced by category contrast,
+it is biased toward generating tightly clustered categories, as new items are
+always most likely to be generated near known examples of the target category.
+Thus, the lack of a contrast mechanism prevents the model from explaining why
+some individuals widely distribute their categories to the corners of the space.
+\jlanote{Xian: Please add a paragraph or two on the representativeness model
+  here.}
+\sxlnote{Ok! I've kind of integrated it into one bigger paragraph with
+  hierarchical bayes.}
+
+Unlike PACKER, both the hierarchical Bayesian model and the representative model
+do not contain any explicit mechanism for controlling the weights of
+between-category contrast and within-category similarity. This means both models
+do not make strong predictions of cluster versus corner profiles. However, the
+covariance matrix specifying the model's prior domain distribution, $\Sigma_0$,
+can be used to explain the generation of row-like and column-like categories.
+This covariance matrix specifies the amount of variance assumed along each
+dimension (as well as the dimensional correlations) across the domain of
+categories. The covariance matrix for a newly generated category, $\Sigma_B$, is
+based on the assumed $\Sigma_0$ as well as the distributions of previously
+learned categories (see Appendix \ref{ap:hsampling-definition}). Thus, the
+importance of each feature can be coded into $\Sigma_0$ to alter the dimensional
+variance of generated categories. Because the hierarchical Bayesian model
+possesses no mechanism to account for category contrast, the model is most
+likely to generate new items that are similar to known examples of the target
+category with no regard for how different it is to the contrast category.
+However, the representativeness model predicts that new exemplars should be more
+representative (i.e., more similar to target and less similar to contrast
+categories), accounting for the tendency of row-like and column-like profiles
+occupying the edge of the feature space.
+
+As noted above, while the copy-and-tweak and hierarchical Bayesian models
+possess mechanisms to explain row- and column-like categories, they cannot
+easily explain why some individuals widely distribute their generated categories
+into disparate corners of the space. This, however, reveals a more general
+limitation: According to the copy-and-tweak and hierarchical Bayesian models,
+the distributional structure of generated categories is {\em independent} of
+their location within the domain. For example, although the copy-and-tweak or
+hierarchical Bayesian models can be parameterized to generate row- or
+column-like categories, there is no mechanism in place to ensure that what is
+generated will be distinct from what is already known. In the next subsection,
+we explore this prediction through an analysis of the interdependence between
+distributional structure and location in creative generation.
+
+\subsection{Category Location vs. Distributional Structure}
+\label{section:individual-diff}
+As noted above, while all three models make clear claims about the internal
+structure of generated categories, the copy-and-tweak and hierarchical Bayesian
+models do not make any claims about how generated categories should differ from
+what is already known. However, as we observed in the results of Experiment 2,
+the distributional structure of a category is not always independent of its
+location within the domain. To demonstrate this point in more depth, we computed
+the X- and Y- axis ranges of every participant-generated category. Taking the
+difference between these values ($X-Y$) produces a measure of each category's
+orientation in the space: positive difference scores correspond to categories
+with more X-axis range (horizontally aligned, `Row' categories), whereas
+negative difference scores indicate the opposite (vertically aligned, `Column'
+categories). Neutral differences scores indicate there was an equal amount of X-
+and Y-axis range, which can be produced by a number of different category types
+(`Clusters', `Corners', etc; see Figures \ref{fig:e1-samples} and
+\ref{fig:e2-samples}). By plotting, for each possible stimulus, the difference
+scores of categories it was generated within, we can relate the distributional
+structure of generated categories to their location within the domain.
+
+However, because many stimuli were infrequently generated (such items near
+members of the `Alpha' category), we cannot simply compute the empirical average
+of the difference scores, as infrequently generated stimuli would be likely to
+show artificially strong differences. Instead, we used a Bayesian analysis to
+estimate the mean $\mu_x$ on the assumption that the scores $x$ for each
+stimulus are normally distributed with an unknown mean and unknown standard
+deviation. The conjugate Normal-Inverse Gamma distribution provides a
+straightforward method for this estimation:
+
+\begin{equation} \mu_x = \dfrac { \nu_0 \mu_0 + \sum{x} } { \nu_0 + n }
+\label{eq:rangediff-bayes}
+\end{equation} \ where $\mu_0$ is the prior mean, $\nu_0$ is a prior scale
+parameter (controlling the weighting of the $\mu_0$), and $n$ is the number of
+categories in which the stimulus was a member (i.e., the number of scores in
+$x$). The default assumption is that there is an equal amount of range along the
+X- and Y-axes, and so we set $\mu_0 = 0$. Likewise, to give a moderate amount of
+weighting to the prior mean we set $\nu_0 = 1$, though the results are robust to
+a range of values. Within this approach, the resulting aggregation is a
+trade-off between the number of generations and the strength of the range
+difference within each generated category. Infrequently generated stimuli, as
+well as those with mixed positive and negative scores, are given neutral
+difference scores.
+
+\begin{figure}[p]
+    \begin{center} 
+      \includegraphics[width=\textwidth]{figs/range-diff-gradients.pdf}
+      \caption{Behavioral and simulated range difference gradients. Each panel
+        shows, for each stimulus, the dimensional orientation of the categories
+        it was generated into: vertically aligned `columns' (orange) versus
+        horizontally aligned `rows' (purple).}
+      \label{fig:range-diff-gradients}
+    \end{center}
+\end{figure}
+
+The results of our analysis are shown in Figure \ref{fig:range-diff-gradients}
+for the experiment and model results\footnote{Prior to plotting, data was also
+processed using a Gaussian filter with $\sigma = 0.8$.}. The left-most column of
+Figure \ref{fig:range-diff-gradients} displays the effect of category location
+and contrast on the distributional structure of the category generated by
+participants. These data reveal strong and consistent patterns across all the
+conditions we tested in Experiments 1 and 2: Generated categories are more
+tightly distributed along the axis in which they are distinct. For example, in
+the `Cluster' condition, exemplars in the bottom-left of the space are more
+often generated into vertically aligned categories, and exemplars in the
+top-right are more often generated into horizontally aligned categories.
+Similarly, in the `Bottom' and `Middle' conditions, horizontally aligned
+categories are generated above and below the experimenter-defined categories,
+while vertically-aligned categories are generated to the sides. In the `Row'
+condition, most categories are horizontally aligned, and lie along the upper
+areas of the space. There are no strong range difference patterns in the XOR
+condition.
+
+These patterns of performance clearly depict the interdependence between the
+distributional structure and location of generated concepts. Our results can be
+interpreted in terms of local minimization of between-category similarity: By
+distributing the generated category away from members of the
+experimenter-defined category, participants may increase the degree of
+between-category distance without drastically altering the degree of
+within-category similarity.
+
+To explore how well the PACKER, copy-and-tweak, representativeness, and
+hierarchical Bayesian models explain our findings, we conducted simulations
+using an individual-differences approach. As noted in Section
+\ref{section:individual-diffs}, row- and column-like categories can be produced
+by each model through changes to the weighting of each dimension. Given this
+information, we may use the models to simulate each participant's generation
+separately, with the importance of each dimension set according to the relative
+range of the participant's generated category along each dimension.
+
+In the PACKER and copy-and-tweak models, the attention weights, $w$, specify the
+importance of each dimension in the computation of similarity. While there exist
+methods to find the optimal attention weighting scheme given a classification
+\citep[see][]{vanpaemel2012using}, for simplicity we may assume that the Alpha
+and Beta categories are distinct along dimensions that the Betas do not vary on.
+Thus, the weighting for a given participant can be computed as:
+
+\begin{equation} w_k = \dfrac {\exp{ \{ -\theta_w \cdot\text{range}(k)} \} } {
+\sum_k {\exp{ \{ -\theta_w \cdot\text{range}(k)} \} } }
+\label{eq:range-weight}
+\end{equation}
+% 
+where $\theta_w$ is a free parameter controlling how differences in range
+correspond to differences in weights (functioning similarly to the $\theta$
+parameter in each of the models), and $\text{range}(k)$ is the range of examples
+generated by the participant along dimension $k$. We used $\theta_w = 1.5$ in
+our simulations, though the results are robust and similar for other $\theta_w$
+values. The resulting $w$ values are thus inversely proportional to the range of
+generated categories along each dimension, with less range corresponding to
+greater weighting.
+
+Unlike the PACKER and copy-and-tweak models, the representativeness and
+hierarchical Bayesian model's dimensional variances correspond to the assumed
+variance of generated categories along each dimension (rather than the inverse
+of the variance). Thus, a different transformation is appropriate for
+incorporating the weights computed in Equation \ref{eq:range-weight}. For the
+hierarchical Bayesian model, we computed the dimensional variances according to:
+$\lambda (1-w_k) 2$, where $\lambda$ is a free parameter specifying the overall
+assumed variance of the domain, and $2$ corresponds to the number of dimensions
+in our experiments\footnote{This calculation applies only in two-dimensional
+  domains, where $w_2 = 1-w_1$.}. Under this approach, evenly distributed
+weights correspond to an assumed variance of $\lambda$. Likewise, larger values
+of $w$, which are produced when the generated category is tightly distributed
+along one dimension, correspond to smaller assumed variances.
+
+Each model was used to simulate each participant's generation independently,
+with the importance of each dimension set according to the participant's
+generated category. The other free parameters within each model were set as in
+Table \ref{table:global-model-fits}. Every participant's generation was
+simulated 2,000 times; given the 305 participants tested across the two
+experiments, each model generated 610,000 categories in total. For comparison
+with our behavioral results, we then computed the range difference gradient
+identically as with the behavioral data. The results are shown in Figure
+\ref{fig:range-diff-gradients}.
+
+As in the more traditional model evaluation analysis described above, the
+contrast models (i.e., PACKER and the representativeness model) provided a much
+closer match to our behavioral results than the copy-and-tweak and hierarchical
+Bayesian models. In all conditions, the contrast models distribute categories
+similarly to the behavioral data: Horizontally-aligned categories tend to be
+placed above and below members of the experimenter-defined category, and
+vertically-aligned categories tend to be placed to the sides. Conversely,
+because the copy-and-tweak and hierarchical Bayesian models are insensitive to
+category contrast, these models do not produce any systematic patterns of
+association between category location and distributional structure. The sole
+exception is within the `Row' condition of Experiment 1, in which the majority
+of participants generated a `Row'-like category, widely distributed along the
+X-axis but not the Y-axis. In these cases, both models are initialized with
+weights that produce Row categories, but because category contrast is not
+considered, categories are uniformly generated across the entire domain, rather
+than concentrated within the upper-regions as observed behaviorally.
+
+\include{methods_results_e3}
+
+\section{General Discussion}
+The sensory impression of every stimulus and event is unique. Grouping distinct
+patterns of sensory information into categories is a fundamental task solved by
+the mind. Most work has focused on how people learn new categories that are
+provided to them through unlabeled, partially labeled, or fully-labeled
+examples. How were these categories first determined? Some natural categories
+are likely to be the result of regularities in the dynamics of our environment.
+But, these are only a subset of the categories that people learn. Other
+categories, such as tools and ideas, were generated by people over time. What
+basic principles underlie how people generate categories?
+
+% The creative generation of concepts is an immensely intriguing topic, but it is
+% also complex and difficult to study in a traditional laboratory setting.
+% Creative generation is most typically studied using a creative cognition
+% approach \citep{finke1992creative,smith1995creative}, where creative products
+% \citep[such as drawings of alien plants and animals;][]{ward1994structured} are
+% analyzed to obtain insight into the processes and representations involved.
+% While this approach has yielded a great deal of information about the nature of
+% creative generation, the commonly employed response modalities (e.g., hand-drawn
+% pictures) make it difficult to develop computational models capable of formally
+% evaluating the proposed representations and processes. However, the generation
+% of novel concepts is not altogether different from other behaviors studied in
+% cognitive psychology: In particular, it can be viewed as an form of inference
+% using one's knowledge about existing categories
+% \citep{kemp2014taxonomy,kurtz2015human}. Taking advantage of this insight, in
+% this paper we investigated the generation of new categories using the empirical
+% and computational toolkit from the field of categorization.
+
+While the bulk of prior research on the topic has focused on the classic finding
+that generated concepts tend to be distributionally similar to known concepts,
+there has been little work addressing the role of contrast in creative
+generation: How is it that people are able to create something {\em different}
+from what is already known? \jlanote{Update this next section:}
+\sxlnote{Updated} We developed two novel models, each incorporating a different
+conceptualization of category contrast. One model is PACKER, which is an
+exemplar-based model that formally specifies the role of contrast in generation
+as exemplar dissimilarity. Specifically, the model proposes that categories are
+represented as exemplars in a multidimensional psychological space, and
+generation is constrained both by within-category and between-category
+similarity: Exemplars belonging to the same category should be similar to one
+another, and exemplars belonging to different categories should not be similar
+to one another. The second model is a novel hierarchical Bayesian model with a
+representativeness mechanism. This model generates exemplars that are more
+representative of (i.e., has greater relative evidence from) the novel category
+compared to the learned category.
+
+We reported two experiments demonstrating systematic effects of category
+contrast in creative generation. Members of participant-generated categories
+tended to be highly dissimilar from members of previously-learned categories,
+and were usually more similar to one another than to members of other
+categories. We also observed broad interdependence between the distributional
+structure (feature variance, correlation) and physical instantiation (location
+within the stimulus space) of generated categories: In Experiment 2, we found
+that the unoccupied regions of the domain influenced the distributional
+structure of categories, and in both experiments we observed that participants
+distributed their generated categories to increase contrast with what was
+already known.
+
+We conducted simulations comparing the contrast models' account of our results
+to the classical proposals for category generation: a ``copy-and-tweak'' model
+(realized as a variant of PACKER with no sensitivity to category contrast), and
+a hierarchical Bayesian model designed to explain the classic distributional
+similarity effect. In all simulations, we found that the contrast models captured a previously unexplained and unexplored aspect of human
+category generation. In particular, by measuring PACKER's fit as a function of its
+prioritization of within- and between-category similarity, we observed that
+considering either constraint exclusively results in a relatively low-quality
+account. Instead, PACKER's best results were obtained when both constraints are
+considered, indicating that human learners do not generate novel concepts
+exclusively on the basis of within-category similarity or between
+class-contrast. This finding mirrors our behavioral results and demonstrates
+that both constraints influence creative generation when explained with an
+exemplar model. Ultimately, our category generation data was best explained by
+the representativeness model. The exact nature of its superiority over the other
+models is still not yet clear. However, as we explore in a later section, it is
+likely that the representativeness mechanism itself (i.e., independent of the
+hierarchical Bayesian foundation) captures a core aspect of both category
+generation and category learning.
+\jlanote{Above needs some updating still. Also, do we want an outline of general
+discussion paragraph?}
+\sxlnote{Updated. I tried writing an outline, but found that points got a little
+  repetitive after a bit, so I ditched that idea...}
+
 \subsection{Novel Analyses Demonstrating Contrast Effects in Prior Work}
@@
 not require any mechanism of category contrast.
+
+\jlanote{Xian: Please do a Chi-Squared test between what is expected based on
+  the logic above and the observed frequencies mentioned below} \sxlnote{I
+  didn't do a chi-sq test because what appears below is not the observed
+  frequencies that correspond to the logic above, but is instead an indication
+  of whether or not it is reasonable to assume the logic above (and our argument
+  is that it is not reasonable.)}
 
@@
 
-\subsection{Disentangling Exemplar dissimilarity and Representativeness in Categorization}
-
-\sxlnote{I'm struggling a little with this section! That said, I think this may
-  also be a good point to address Charles' concern re: the possible conditions
-  under which contrast are important.} 
-Although both PACKER and the
-representativeness model perform better than the other models, they appear to be
-capturing fundamentally different aspects of contrast in categorization.
-Specifically, while the representativeness model excelled in predicting our
-category generation data, PACKER was the best-fitting model in accounting for
-participant error in category learning. The source of this discrepancy is still
-not yet clear -- it could be an indication of a fundamental difference between
-the processes involved in generation versus learning where exemplar
-dissimilarity is a larger factor in category generation while representativeness
-exerts more influence in category learning.
-
-In this paper, we have developed two novel models of category generation, each
-involving a different conceptualization of contrast.  However, there is no theoretical constraint that dictates the It is also possible to treat the similarity
-
-A thorough investigation to disentangle the two contrast mechanisms here is
-beyond the scope of the current article. However, as a quick proof of concept it
-is useful to check if the addition of a representativeness mechanism to an
-exemplar model would yield any benefit at all.
-
-Xian: Please
-write about the two hypotheses and that followup work needs to look into why the
-Bayesian representativeness model doesn't do as well at explaining category
-learning ease, but does better at category generation.
-
-MAKE SURE WE CITE AND DESCRIBE \citet{thomas98}! \sxlnote{I moved the
-  description and citation of thomas98 to the introduction for justification of
-  why we think categories should share similar distributions.}
+The current work also suggests exciting directions for related investigations
+into unsupervised categorization. While the unsupervised categorization
+literature is primarily interested in the generation of categories within a set
+of observed exemplars where no prior categories are learned, our category
+generation studies are focused on the production of novel category exemplars
+themselves. Despite this distinction, both types of categorization research are
+ultimately interested in the question of how categories can be formed. The results
+of our current work indicate that the formation of categories is constrained by
+some measure of contrast. To our knowledge, this explicit investigation of
+category contrast has not yet been attempted in the unsupervised categorization
+literature.
+
+Interestingly, the unidimensional sort bias that is commonly observed in
+unsupervised categorization
+\citep{imai1965discriminability,milton2004influence,ahn1992two} was not
+consistently observed in our tasks. To clarify, the unidimensional sort bias is
+the observation that categories are typically constructed solely by focusing on
+a single dimension or feature. In our experiments we only observed this bias in
+the row condition of Experiment 1, whereas in other conditions participants
+generated categories that typically varied along both dimensions. This may be
+unsurprising at face value, since participants in conditions other than the row
+condition were trained on a prior category that varied on both dimensions.
+However, these results suggest that the unidimensional sort bias is not
+universal to all types of category construction -- or more specifically, the
+unidimensional bias may not apply when participants need to generate categories
+with entirely new exemplars.
+
+\subsection{Exemplar dissimilarity and Representativeness in Categorization} 
+\sxlnote{Hmmm.. ok after quite a few revisions I am still not quite satisfied
+  with this section. It doesn't feel quite ...tight, but I think it's time for
+  me to say stop and pass it on.}
+Although both PACKER and the representativeness model perform better than the
+classical models, they appear to be capturing fundamentally different aspects of
+contrast in categorization. Specifically, while the representativeness model
+excelled in predicting our category generation data, PACKER was the best-fitting
+model in accounting for participant error in category learning. Given the
+history of their development, this is perhaps not entirely surprising. PACKER
+is implemented as an extension of the GCM, which was originally developed for (and
+has been highly successful in) category learning \citep{nosofsky1986attention,
+  nosofsky1994comparing}. Conversely, the representativeness mechanism was
+conceived as a heuristic describing the relative extent to which an individual
+observation is generated from some unobserved population
+\citep{kahneman1972subjective,abbott16}. This generative nature of the
+representativeness mechanism intuitively lends itself to the prediction of
+category generation data.
+
+While the contrast models were born quite naturally out of the corresponding
+classical models, it is worth considering if similar benefits in prediction
+would be seen if the contrast mechanisms were applied to different classical
+models. Although there is no clear way to adapt the exemplar dissimilarity
+mechanism to the current hierarchical Bayesian model, it is a fairly
+straightforward process applying the representativeness mechanism to an exemplar
+model. Specifically, by treating the similarity measure (Equation
+\ref{eq:similarity}) in the current implementation of copy-and-tweak as a
+density estimate for the representativeness mechanism (Equation
+\ref{eq:representativeness}), we gain a new model of representativeness that is
+based on exemplar similarity instead of a multivariate Gaussian
+likelihood.\footnote{An analogous implementation of the representativeness
+  mechanism to PACKER results in both $\theta_c$ and $\theta_t$ adding to
+  constant value that is independent of the similarity to contrast and target
+  exemplars respectively. Consequently, it becomes equivalent to a
+  copy-and-tweak model with representativeness.}
+
+By subjecting this preliminary model to the same analysis as the other models,
+we find that it substantially outperforms its classical counterpart, but not to
+the extent of PACKER or the representativeness model. Specifically, fitting this
+model to the category generation data yielded $L = -4815$ and $AIC=9633$ for the
+entire set of data from Experiments 1 and 2, and $L = -3474$, $AIC=6953$ for
+data excluding the first trials. When used to predict the error rates of
+participants in the category learning task (Experiment 3), the performance of
+the copy-and-tweak with representativeness model correlated with participant
+error with $r=.56, p<.001$. These results suggest that the predictive advantages
+to including the representativeness mechanism is not limited to a hierarchical
+Bayesian model, but can also be found when applied to an exemplar model.
+Although its performance is not as strong as either of the fully-developed
+contrast models, there is a benefit to the representativeness mechanism that is
+independent from any interaction with a hierarchical Bayesian framework.
+
+Ultimately, the source of the discrepancy between the performance of the two
+contrast mechanisms in category generation versus category learning is still not
+entirely clear. Our results could be an indication of a fundamental difference
+between the processes involved where exemplar dissimilarity is a larger factor
+in category generation while representativeness exerts more influence in
+category learning. While a thorough investigation is beyond the scope of the
+current article, the preliminary results of this section indicate that further
+work to disentagle these two mechanisms would be promising. In the following
+section, we explore some other limitations of the current work and provide more
+recommendations for future research.
+
 \subsection{Limitations and Future Directions} Although we have demonstrated

### 2018-09-04 Manuscripts/cog-psych/revision/methods_results_e3.tex
Primary revisions to manuscript: copytweak-rep in section 8.5, unsup cat, contrast in thomas98, some mention of machine learning generative models. In other news, updated and more or less finalised fits with the category [0] bug fixed in Simulation.py. TODO: Update the goldstate.

@@
 is to investigate whether model performance (in generating categories) is
-consistent with human performance (in learning those categories). Specifically, if contrast plays a broader role in categorization, then Beta categories rated as "better" according
+consistent with human performance (in learning those categories).
+%\footnote{It is
+  % possible to adjust the present models of category generation such that they
+  % can make category learning predictions. However, such an adjustment makes the
+  % strong assumption that the very same processes governing category generation
+  % also apply to category learning. Taking this approach would fall outside the
+  % scope of the current investigation.} 
+Specifically, if contrast plays a
+broader role in categorization, then Beta categories rated as "better" according
 to the contrast models should be easier for participants to learn.
-%\footnote{It is possible to adjust the present models of category generation such that they can make category learning predictions.
-However, such an adjustment makes the strong assumption that the very same processes governing category generation also apply to category learning.
-Taking this approach would fall outside the scope of the current investigation.}
+
 \subsection{Participants, Materials, and Procedure}
@@
 Overall accuracy of the participants was high, with a mean error rate of $.19$
-($SD = .19$). Aggregated error rates for each block are presented in Figure
+($SD = .19$). Error rates for each block are presented in Figure
 \ref{fig:learningcurve}.
@@
 Specifically, PACKER correlated with human performance with
-$r = .65, \ 95\% CI$\footnote{Confidence intervals were obtained using a bootstrap
-  method with 1000 bootstrap samples.}$ \ [.47,.78]$, and the representativeness
-model correlated with human performance with $r = .56, \ 95\%  CI \ [.38, .72]$.
-Copy-and-tweak yielded a correlation of $r = .45, 95\% CI \ [.29, .61]$ and the
+$r = .65, 95\% CI$\footnote{Confidence intervals were obtained using a bootstrap
+  method with 1000 bootstrap samples.}$ [.48,.78]$, and the representativeness
+model correlated with human performance with $r = .59, 95\% CI [.40, .73]$.
+Copy-and-tweak yielded a correlation of $r = .45, 95\% CI [.28, .62]$ and the
 hierarchical Bayesian model yielded a similar correlation of
-$r = .46, 95\% CI \ [.30, .61]$. These results are graphically presented in Figure
+$r = .46, 95\% CI [.30, .62]$. These results are graphically presented in Figure
 \ref{fig:perror_corr}.

### 2018-09-04 Experiments/cogpsych-code/build_send_chtc.py

@@ -15,9 +15,11 @@ rmsubmit = 'cd {};./rmold_submitnew.sh'.format(send2dir) #this script on server 
 execfile('build_chtc.py')
 os.chdir('../../..')
 #upload the new batch
+print 'Uploading latest working batch to {}.'.format(send2dir)
 os.system('scp -i ~/Dropbox/.ssh/chtc working.tar.gz liew2@submit-5.chtc.wisc.edu:{}'.format(send2dir))
 
 #remotely remove the old files and submit
+print 'Running rm and submit scripts on CHTC server...'
 os.system('ssh -i ~/Dropbox/.ssh/chtc -t liew2@submit-5.chtc.wisc.edu \'{}\' '.format(rmsubmit))
 #os.system('ssh -i ~/Dropbox/.ssh/chtc liew2@submit-5.chtc.wisc.edu')
 

### 2018-09-04 Experiments/cogpsych-code/plot_modelppt_ease.py

@@ -25,10 +25,10 @@ if pearson:
 else:
     corrtype = 's'
 #Bootstrap parameters
-nbootstraps = 10
+nbootstraps = 1000
 
 savefilename='modelvsppt{}_t.pdf'.format(corrtype)
-bestparmdb = "pickles/chtc_gs_best_params_corr{}_new.p".format(corrtype)
+bestparmdb = "pickles/chtc_gs_best_params_corr{}.p".format(corrtype)
 modeleaseDB = "pickles/modelease_corr{}.p".format(corrtype)
 #Some plotting options
 font = {'family' : 'DejaVu Sans',
@@ -69,12 +69,12 @@ for modelname in best_params_t.keys():
     for i,parmname in enumerate(best_params_t[modelname]['parmnames']):
         parmval = best_params_t[modelname]['bestparmsll']
         best_params[modelname][parmname] = parmval[i]
-#modelList = [Packer,CopyTweak,ConjugateJK13,RepresentJK13]
-modelList = [CopyTweak,CopyTweakRep,Packer, RepresentJK13,]                            
+modelList = [Packer,CopyTweak,ConjugateJK13,RepresentJK13]
+#modelList = [CopyTweak,CopyTweakRep,Packer, RepresentJK13,]                            
 
 #Specify plot order
-#modelPlotOrder = np.array([[Packer,RepresentJK13],[CopyTweak,ConjugateJK13]])
-modelPlotOrder = np.array([[CopyTweak,CopyTweakRep],[Packer,RepresentJK13]])
+modelPlotOrder = np.array([[Packer,RepresentJK13],[CopyTweak,ConjugateJK13]])
+#modelPlotOrder = np.array([[CopyTweak,CopyTweakRep],[Packer,RepresentJK13]])
 
 #Prepare matched database    
 matchdb='../cat-assign/data_utilities/cmp_midbot.db'
@@ -135,6 +135,7 @@ options = dict(
 if 'tso' not in ll_global.keys():
     pptdata,tso = funcs.prep_corrvar(info,assignment,stimuli,stats,WT_THETA)
     ll_global['tso'] = tso
+    ll_global['pptdata'] = pptdata
 else:
     tso = ll_global['tso']
 
@@ -175,7 +176,7 @@ for model_obj in modelList:
             #Get loglikelihoods
             raw_array_ll = Simulation.loglike_allperm(params, model_obj, categories, stimuli, permute_category = 1)
             ll_list += [raw_array_ll]
[…]

### 2018-09-04 Experiments/cogpsych-code/print_gs_results_csv.py

@@ -41,6 +41,7 @@ for dname in data_names:
 
 #Iterate though each data file
 data = dict()
+parmnames_global = {}
 for i,file in enumerate(gsfiles):
     with open(pickledir+file,'rb') as f:
         fulldata = pickle.load(f)        
@@ -50,6 +51,7 @@ for i,file in enumerate(gsfiles):
     for mn in model_names:
         #with open(os.path.join(writedir,file[0:len(file)-2]+'_'+modelname_print+'.csv'),'wb') as f:
         parmnames = fulldata[mn]['parmnames']
+        parmnames_global[mn] = parmnames
         nparms = len(parmnames)
         nllAIC = fulldata[mn]['bestparmsll'][[nparms,nparms+1]]
         parmvals = fulldata[mn]['bestparmsll']
@@ -69,7 +71,7 @@ with open(writefile+'.csv','wb') as f:
     model_names = data.keys()
     for mn in model_names:
         model_print = funcs.getModelName(mn,fetch='short') #fetch short version of model name
-        parmnames = fulldata[mn]['parmnames']
+        parmnames = parmnames_global[mn]
         for parm in parmnames:
             parmvals = []
             for dn in data_names:

### 2018-09-11 Manuscripts/cog-psych/revision/cogpsych.tex
Final revisions by Joe. Mostly cleaned up various loose bits. Now off to Ken.

--- 2018-07-20 Manuscripts/cog-psych/revision/cogpsych.tex
@@
 \section*{Abstract}
-The ability to generate new concepts and ideas is among the most
-fascinating aspects of human cognition, but we do not have a strong
-understanding of the cognitive processes and representations underlying concept
-generation. In this paper, we study the generation of new categories using the
-computational and behavioral toolkit of traditional artificial category
-learning. Previous work in this domain has focused on how the statistical
-structure of known categories generalizes to generated categories, overlooking
-the degree of contrast between the known and generated categories as a factor.
-We report three experiments demonstrating that contrast between what is known
-and what is created is of fundamental importance for categorization and innovation. We propose two novel approaches to modeling category contrast:
-one focused on exemplar dissimilarity and another on the representativeness
-heuristic. These new models are compared to two key alternatives. Our
-experiments and simulations demonstrate specifically how category contrast
-influences creative generation, and how it can interact with other constraints
-to produce different types of creations. Our work also serves as a concrete
-example of how well-established and highly controlled experimental frameworks
-(such as those in the field of category learning) can be used to make progress
-in an intriguing yet understudied domain.
+The ability to generate new concepts and ideas is among the most fascinating
+aspects of human cognition, but we do not have a strong understanding of the
+cognitive processes and representations underlying concept generation. In this
+paper, we study the generation of new categories using the computational and
+behavioral toolkit of traditional artificial category learning. Previous work in
+this domain has focused on how the statistical structure of known categories
+generalizes to generated categories, overlooking the extent that the contrast
+between the known and generated categories is a factor. We report three
+experiments demonstrating that contrast between what is known and what is
+created is of fundamental importance for categorization. We propose two novel
+approaches to modeling category contrast: one focused on exemplar dissimilarity
+and another on the representativeness heuristic. Our experiments and
+computational analyses demonstrate that both models capture different aspects of
+contrast's role in categorization. Our work also serves as a concrete example of
+how well-established categorization models can be applied beyond category learning.
 
 \setlength\parindent{0.5in}
-{\em Keywords}: categorization; concepts; creativity; generation; computational modeling; exemplar models
+{\em Keywords}: categorization; concepts; category learning; generation;
+computational modeling; exemplar models; representativeness
 \clearpage
@@
 a scientist devise an experiment that induces a participant to create and learn
-new categories that are as interesting as cell phones and smart phones? In this
-article, we use computational models and behavioral experiments to investigate
-the prior expectations people have when they learn and generate new categories.
+new categories that are as interesting as cell phones and smart phones? As a
+step towards these goals, in this article we investigate the expectations that
+people have when they learn and generate new categories using computational
+models and behavioral experiments.
 
@@
 \citep{kemp2014taxonomy,kurtz2015human}. Research in categorization typically
-focuses on what properties of a category affect the classification of an
-object's category given its features \citep{shepard1961learning,kurtz2013human},
-and the prediction of an object's unobserved features given some of its other
-features and/or its category \citep{markman2003category}. The generation of
-members of a new category consists of inferring {\em all} features for a {\em
-novel} category label. Thus, we can make progress formalizing the processes
-involved in category generation by extending theories of categorization to this
-case. Then, to ensure that these principles reflect basic principles of
-categorization and are not just specific to category generation, we can examine
-what factors of the generated categories affect the ease at which people learn
-the category.
+focuses on what properties of a category affect human learning of an object's
+category given its features \citep{shepard1961learning,kurtz2013human}, or the
+prediction of an object's unobserved features given some of its other features
+and/or its category \citep{markman2003category}. The generation of members of a
+new category consists of inferring {\em all} features for a {\em novel} category
+label. Thus, we can make progress formalizing the processes involved in category
+generation by extending theories of categorization to this case. Then, to ensure
+that these principles reflect basic principles of categorization and are not
+just specific to category generation, we can examine what factors of the
+generated categories affect the ease at which people learn the category.
 
@@
 regularities to other categories would be identical to a known category that is
-representative of the domain (and thus, not new at all).
-Further, contrast has played a critical role studying the spread of new ideas in diverse fields, such as marketing \citep{berger16} and sociology \citep{rogers03}.
+representative of the domain (and thus, not new at all). Contrast from other
+known categories should play a role. Further, scientists in other fields, such
+as marketing \citep{berger16} and sociology \citep{rogers03}, highlight the
+critical importance of contrast in the creation of new ideas.
 
@@
 We examine two different definitions of how categories should differ from one
-another: exemplar dissimilarity or representativeness of the alternative
+another: exemplar dissimilarity and representativeness of the alternative
 category. To formalize the first technique, we build on the largely successful
@@
 
-The outline of the article is as follows. First we describe previous empirical
-work on the topic of category generation, as well as the computational
-formalizations of their theoretical explanations. We then present two hypotheses
-for how contrast might affect categorization and formalize them in computational
-models. The first is a novel exemplar model, which is designed to generate
-categories that systematically differ from existing categories in the domain.
-The second assumes the goal of category generation is to create representative
-samples of the opposite of the observed categories. We present two experiments
-demonstrating strong and systematic effects of category contrast on concept
-generation, and we qualitatively and quantitatively analyze the performance of
-each model in capturing human category generation. Afterwards, we find support
-for category contrast predicting the ease at which people learn a pair of 
-categories in a novel empirical experiment. We conclude with a discussion of the
-implications of our results for categorization and creative cognition, and
-directions for future work.
+The outline of the article is as follows. First we describe previous
+computational formalizations of theories of category generation and empirical
+work investigating them. We then present two hypotheses for how contrast might
+affect categorization and formalize them in computational models. The first is a
+novel exemplar model, which is designed to generate categories that
+systematically differ from existing categories in the domain. The second assumes
+that the goal of category generation is to create representative samples of the
+opposite of the observed categories. We present two experiments demonstrating
+strong and systematic effects of category contrast on concept generation, and we
+qualitatively and quantitatively analyze the performance of each model in
+capturing human category generation. Afterwards, we find support for category
+contrast predicting the ease at which people learn a pair of categories in a
+novel empirical experiment. We conclude with a discussion of the implications of
+our results for categorization and directions for future
+work.
 
@@
 
-Much of what we know about concept generation and contrast comes from the foundational
-literature on creative cognition. In a series of reports, Ward and colleagues
+Much of what we know about concept generation and contrast comes from the
+foundational literature on creative cognition. In a series of reports, Ward and
+colleagues
 \citep{ward1995s,ward1994structured,marsh1999inadvertent,ward2002role,smith1993constraining}
@@
 
-\sxlnote{Added thomas98 description here} \jlanote{still need the contrast result from her work.} It is worth pointing out that the
-observation of regularities in distributional properties across categories is
-not confined to generating whole new categories, but it also occures when inferring an unobserved feature within traditional categorization paradigms. Most notably, \citet{thomas98}
-found that when people were asked to generate feature values for exemplars with
-a missing feature (specifically circles of different sizes, each with a radial
-line of varying orientation), the resulting set of exemplars tended to share the
-same feature correlations as exemplars from previously learned categories. [the
-transition to the next point isn't great, though]
-
 The broader set of observations made by Ward and colleagues provide a great deal
@@
 information provided to participants (such as an example of a species generated
-by other participants) can drastically diminish difference of a new category
+by other participants) can drastically diminish the difference of a new category
 from pre-existing categories. Theoretical accounts of these effects have
@@
 well as the exemplar view \citep{medin1978context,brooks1978nonanalytic}. While
-this work is been incredibly useful in providing a conceptual sketch of
-generation theories, the hand-drawn responses used in the experiments paradigms
-precludes the development of formal approaches that can be used to test the
-theories in a fine-grained manner.
+this work has been incredibly useful in providing a conceptual sketch of
+generation theories, its qualitative nature and the hand-drawn responses used in
+the experiments paradigms precludes the development of formal approaches that
+can be used to test them in a fine-grained manner.
 
@@
 the same distributional properties as the experimenter-defined categories. For
-example, after learning categories with a positive correlation between the
-size and saturation features (larger sized crystals were more saturated),
-participants generated novel categories with the same positive correlation. This
-finding is notable, as it demonstrates that category generation can be studied
-in a well-known and highly controlled experimental paradigm.
+example, after learning categories with a positive correlation between the size
+and saturation features (larger sized crystals were more saturated),
+participants generated novel categories with the same positive correlation. By
+replicating \cite{ward1994structured}, they demonstrated that category generation
+can be studied in a well-known and highly controlled experimental paradigm.
 
@@
 
-\section{And Now for Something Different: A Role For Contrast}
+It is worth pointing out that the early observation of regularities in
+distributional properties across categories is not confined to the work of Ward
+and colleagues. Most notably, \citet{thomas98} trained participants to classify
+circles of different sizes, each with a radial line of varying orientation, into
+two categories. During a surprise prediction phase, participants were asked to
+generate a value for a missing feature for a certain category exemplar given a
+value on the other feature. Their results revealed that exemplars tended to
+share the same feature correlations as exemplars from previously learned
+categories. Interestingly, this effect was not consistently observed when the
+given values of a feature fell outside the range of learned values for that
+particular category. For instance, if the learned Category A exemplars were
+generally 8 cm to 10 cm in radius and negatively correlated with the angle of
+the radial line, participants generally produced radial line orientations that
+were also negatively correlated with the exemplar size when the given size was
+between 8 cm and 10 cm. However, when the given size fell outside that range for
+that category (e.g., 14 cm for a Category A exemplar), some participants
+consistently produced radial line orientations that were not negatively
+correlated with the size and that also placed the exemplars in a significantly
+different location in the feature space. These results suggest that when tasked
+to produce exemplars that fall outside what was previously learned, people show
+some tendency to generate exemplars that are different (both distributionally
+and spatially) to what was learned. We explore this idea of contrast
+more deeply in the following section.
+
+\section{And Now for Something (Completely) Different: The Role of Contrast}
 
 Prior work on category generation has explored only one factor: distributional
-correspondences between learned and generated categories. As a result most of
+correspondences between learned and generated categories. As a result, most of
 the computational, theoretical, and empirical efforts have been focused on
@@
 As a consequence, we propose that contrast should be a primary constraint that
-underlies people's expectations about how a set of categories are organized: Categories should be different from each other.
+underlies people's expectations about how a set of categories are organized:
+Categories should be different from each other.
 
@@
 distributional structure. This is not an issue with their model specifically,
-but using a broader class of standard hierarchical Bayesian models without any additional features \citep[e.g.,
-][]{griffithsscn07,kemp07}. Many of these models assume that at some point of
-the latent generative process the same underlying distribution generates all of
-the categories and thus, any differences between categories are due to {\em
-noise} and should not be {\em systematic}. If the goal of creating a new
-category given others in a domain is generating from the posterior predictive
-distribution then the best a standard hierarchical Bayesian model without a
-notion of contrast built into the model can do at capturing contrast is to
-assume that the new category is placed uniformly at random over stimulus space.
-But, this defeats the purpose of a hierarchy as it is ignored when determining a
-new category location!\footnote{It is plausible that some hierarchical Bayesian
-model could be created that generates categories different from each other.
-However, this model would not be a standard application or extension of most
-pre-existing hierarchical Bayesian models. The generative process would need to
-include a component that promotes contrast.} Note that this does not mean
-location information and contrast cannot be captured using hierarchical Bayesian
-models. In fact, a model instantiating contrast as representativeness is a hierarchical
-Bayesian model. But, it requires an additional factor, which in our case is having a different goal than predicting the next category.
+but using a broader class of standard hierarchical Bayesian models without any
+additional features \citep[e.g., ][]{griffithsscn07,kemp07}. Many of these
+models assume that at some point of the latent generative process the same
+underlying distribution generates all of the categories and thus, any
+differences between categories are due to {\em noise} and should not be {\em
+  systematic}. If the goal of creating a new category given others in a domain
+is generating from the posterior predictive distribution then the best a
+standard hierarchical Bayesian model without a notion of contrast built into the
+model can do at capturing contrast is to assume that the new category is placed
+uniformly at random over stimulus space. But, this defeats the purpose of a
+hierarchy as it is ignored when determining a new category location!\footnote{It
+  is plausible that some hierarchical Bayesian models could be created that
+  generates categories different from each other. However, this model would not
+  be a standard application or extension of most pre-existing hierarchical
+  Bayesian models. The generative process would need to include a component that
+  presumes contrast.} Note that this does not mean location information and
+contrast cannot be captured using hierarchical Bayesian models. In fact, a model
+instantiating contrast as representativeness is a hierarchical Bayesian model.
+But, it requires an additional factor, which in our case is having a different
+goal than predicting the next category.
 
@@
 in any number of respects. Towards providing a more precise definition of the
-role of contrast in creative generation, we formalized contrast in a novel
-exemplar model, PACKER ({\em Producing Alike and Contrasting Knowledge using
-Exemplar Representations}). PACKER explains category generation as a balance
-between two fundamental constraints: The category to be generated should not be
-similar to known categories, and exemplars within each category should be
-similar to one another. These ideas are implemented within the well-studied
-exemplar framework -- the PACKER model is an extension of the influential
-Generalized Context Model of categorization
-\citep[GCM;][]{nosofsky1984choice,nosofsky1986attention}.
+role of contrast in generation, we formalized contrast in a novel exemplar
+model, PACKER ({\em Producing Alike and Contrasting Knowledge using Exemplar
+  Representations}). PACKER explains category generation as a balance between
+two fundamental constraints: The category to be generated should not be similar
+to known categories, and exemplars within each category should be similar to one
+another. These ideas are implemented within the well-studied exemplar framework
+-- the PACKER model is an extension of the influential Generalized Context Model
+of categorization \citep[GCM;][]{nosofsky1984choice,nosofsky1986attention}.
 
 Although, as an exemplar model, one of PACKER's proposals is people represent
-categories in terms of a collection of stored exemplars, it is worth emphasizing
-that the underlying nature of human category representations is not the primary
-concern of our work. The choice to develop PACKER within an exemplar framework
-reflects the facts that exemplar models have been throughly evaluated, are
-strongly theoretically motivated, and dominate much of the theoretical and
-empirical work in categorization. The focus of our work with PACKER concerns the
-dual constraints of within- and between-class similarity; it is not difficult to
-imagine how such constraints may be instantiated using alternative frameworks
-(e.g., \citealp{kurtz2007divergent,love2004sustain,smith2000thirty}; for a
-review of categorization models, see \citealp{pothoswills2011}).
+categories in terms of a collection of stored exemplars, we did not pick it
+assuming it is the correct model of human categorization. The choice to develop
+PACKER within an exemplar framework reflects the facts that exemplar models have
+been throughly evaluated, are strongly theoretically motivated, and dominate
+much of the theoretical and empirical work in categorization. The focus of our
+work with PACKER concerns the dual constraints of within- and between-class
+similarity; it is not difficult to imagine how such constraints may be
+instantiated using alternative frameworks (e.g.,
+\citealp{kurtz2007divergent,love2004sustain,smith2000thirty}; for a review of
+categorization models, see \citealp{pothoswills2011}).
 
@@
 PACKER (as well as its name) was in part inspired by earlier work from the
-categorization literature \citep{hikada2011packing,stewart2005}. \citet{hidaka2011packing} argued that
+categorization literature \citep{hidaka2011packing,stewart2005}. \citet{hidaka2011packing} argued that
 natural categories ``pack'' the values of features such that different
@@
 
-\jlanote{I'm a bit confused on the theta constraints.
-it's clear they need to be non-negative, but you don't usually see both needing to be between 0 and 1 but not sum to some value.
-is this really what we ended up with?
-Something feels fishy to me with the math.}
 PACKER sets $f(x_j)$ depending on exemplar $x_j$'s category membership:
 $f(x_j) = \theta_t$ if $x_j$ is a member of the target category, and
-$f(x_j) = -\theta_c$ if $x_j$ is a member of a contrast category.
-$\theta_t$ and $\theta_c$ are free parameters
-($0 \leq (\theta_t,\theta_c) \leq 1$) controlling the trade-off
-between within- and between-category similarity. For example, when
-$\theta_t = \theta_c = 0.5$, $f(x_j) = 0.5$ for members of the
-target category and $f(x_j) = -0.5$ for members of other categories; thus, the
-model is likely to generate items that are similar to members of the target
-category but are not similar to members of other categories. In this way,
-$\theta_t > 0$ with $\theta_c = 0$ produces exclusive consideration
-of target-category members, and $\theta_c > 0$ with $\theta_t = 0$
-produces exclusive consideration of contrast-category members. The combination
-of $\theta_t$ and $\theta_c$ parameters thus specifies a wide
-breadth of possible approaches; by fitting it to a dataset, one can describe the
-relative roles of between-category contrast and within-category similarity in
-generation. See Figure \ref{fig:packer-examples} for an illustration of how
-these parameters control the relative influence of within-category similarity
-and contrast to other categories when generating a new exemplar.
+$f(x_j) = -\theta_c$ if $x_j$ is a member of a contrast category. $\theta_t$ and
+$\theta_c$ are free parameters ($0 \leq \theta_t,\theta_c$) controlling the
+trade-off between within- and between-category similarity. For example, when
+$\theta_t = \theta_c = 0.5$, $f(x_j) = 0.5$ for members of the target category
+and $f(x_j) = -0.5$ for members of other categories; thus, the model is likely
+to generate items that are similar to members of the target category but are not
+similar to members of other categories. In this way, $\theta_t > 0$ with
+$\theta_c = 0$ produces exclusive consideration of target-category members, and
+$\theta_c > 0$ with $\theta_t = 0$ produces exclusive consideration of
+contrast-category members. The combination of $\theta_t$ and $\theta_c$
+parameters thus specifies a wide breadth of possible approaches; by fitting it
+to a dataset, one can describe the relative roles of between-category contrast
+and within-category similarity in generation. See Figure
+\ref{fig:packer-examples} for an illustration of how these parameters control
+the relative influence of within-category similarity and contrast to other
+categories when generating a new exemplar.
 
@@
 %determinism.
-
 It is worth noting that PACKER is only one possible exemplar-based account of
@@
 copy-and-tweak model by including a contrast mechanism, enabling generation
-according to dissimilarity to members of opposing categories. By consequence,
+according to dissimilarity to members of opposing categories. As a consequence,
 copy-and-tweak can be realized as a parameterization of the PACKER model that is
 insensitive to category contrast. Specifically, when $\theta_c = 0$ and
-$\theta_t > 0$ (see Figure \ref{fig:packer-examples}, panel B),
-$f(x_j) = 0$ for $x_j$ belonging to contrast categories; thus, PACKER is not
-influenced by these items, and is mathematically equivalent to a copy-and-tweak
-approach.
+$\theta_t > 0$ (see Figure \ref{fig:packer-examples}, panel B), $f(x_j) = 0$ for
+$x_j$ belonging to contrast categories; thus, PACKER is not influenced by these
+items, and is mathematically equivalent to a copy-and-tweak approach.
 
@@
 significantly improves model fits). Note that the purpose of the article is to
-explore and formally analyze the role of contrast in categorization and
-thus, we leave extending PACKER to incorporate distributional factors \citep[as
-explored by][]{jern2013probabilistic} for future work.
+explore and formally analyze the role of contrast in categorization and thus, we
+leave extending PACKER to incorporate distributional factors for future work.
 
@@
 An alternate conceptualization of category contrast is the idea of
-representativeness -- exemplars are generated such that they resemble the target category and thus more distinct from and less
-similar to other categories. We adopt the formalization given by
-\cite{tenenbaum01}, where the representativeness $R$ of an item (or in our case,
-an exemplar) $y$ is the relative amount of evidence that is provided by $y$ for
-a given hypothesis $h$ in a space of hypotheses $\mathcal{H}$ in contrast to all other hypotheses $\mathcal{H}^c=\mathcal{H} \setminus \{h\}$:
+representativeness -- exemplars are generated such that they resemble the target
+category and thus more distinct from and less similar to other categories. We
+adopt the formalization given by \cite{tenenbaum01}, where the
+representativeness $R(y,h)$ of an item $y$ (or in our case, an exemplar) is the
+relative amount of evidence that is provided by $y$ for a given hypothesis $h$
+in a space of hypotheses $\mathcal{H}$ in contrast to all other hypotheses
+$\mathcal{H}^c=\mathcal{H} \setminus \{h\}$:
 \begin{equation}
@@
 \end{equation}
-
 In Equation \ref{eq:representativeness}, $p(h')$ is the prior probability
@@
 In our simulations, we set the prior mean $\mu_0$ to the center of the stimulus
-space and $\Sigma_0 = \lambda \mathbf{I}$ where $\lambda$ is a free parameter
-and $\mathbf{I}$ is a $d$-by-$d$ identity matrix with $d$ representing the
-number of dimensions or features in the domain. The $\kappa$ and $\nu$
-parameters are freely estimated within the constraints $\kappa > d-1$ and
-$\nu > 0$.
-
-With the assumption of the NIW prior, the expected location vector $\mu_k$ given exemplars observed from category $k$ is
+space the prior variance to be isotropic ($\Sigma_0 = \lambda \mathbf{I}$ where
+$\lambda$ is a free parameter and $\mathbf{I}$ is a $d$-by-$d$ identity matrix
+with $d$ representing the number of dimensions or features in the domain). The
+$\kappa$ and $\nu$ parameters are freely estimated within the constraints
+$\kappa > d-1$ and $\nu > 0$.
+
+With the assumption of the NIW prior, the expected location vector $\mu_k$ given exemplars observed from category $k$ is:
 \begin{equation}
@@
 probabilities proportional to the multivariate normal likelihood $p(y|h)$, the
-representativeness model generates new exemplars with probabilities weighted by
-their representativeness $R(y,h)$. In practice, the probability of generating a
+representativeness model generates new exemplars according to their
+representativeness $R(y,h)$. In practice, the probability of generating a
 particular candidate $y$ is obtained using an Exponentiated
@@
 representativeness model focuses on the generation of exemplars that are
-representative of the underlying distribution for a target category. \jlanote{The next sentence is weird sounding and should be cleaned up} Generally,
-this representativeness mechanism results in the generation of exemplars that
-are less similar (i.e., more different) to contrast categories. We see this
-illustrated in Figure \ref{fig:representative-examples}, where the
-representativeness model displays a strong preference for generating exemplars
-from target category 'B' that are further away from the contrast category 'A'.
+representative of the underlying distribution for a target category. Generally,
+this mechanism of representativeness results in the generation of exemplars that
+are less similar to exemplars from other categories. We see this illustrated in
+Figure \ref{fig:representative-examples}, where the representativeness model
+displays a strong preference for generating exemplars from target category 'B'
+that are further away from the contrast category 'A'.
 
@@
 to one member of category `A' and one member of category `B'. Predictions are
-shown for {\em(a)} the original hierarchical Bayesian model and {\em(b)} the
+shown for (a) the original hierarchical Bayesian model and (b) the
 representativeness model.}
@@
 which generated categories differ from known ones is, nonetheless, poorly
-understood: Existing theories do not make strong predictions about how
-creatively generated concepts should systematically differ from existing ones.
-Above, we introduced two
-novel approaches to formalizing the roles of similarity and contrast in creative
-generation.
+understood: Existing theories do not make strong predictions about how generated
+concepts should systematically differ from existing ones. Above, we introduced
+two novel approaches to formalizing the roles of similarity and contrast in
+category generation.
 
@@
 artificial stimulus design: A two dimensional domain of squares, varying in
-color and size (see Figure \ref{fig:e1-conditions}, panel A). These dimensions
+color and size (see Figure \ref{fig:e1-conditions}a). These dimensions
 have been used in numerous classification learning studies
@@
 be inconsistent with prevailing theories of category generation
-\citep{jern2013probabilistic}
+\citep{jern2013probabilistic}.
 
@@
 trials), where they observed exemplars belonging to the `Alpha' category.
-Participants were instructed to learn as much as they can about the Alpha
+Participants were instructed to learn as much as they can about the `Alpha'
 category, and that they would answer a series of test questions afterwards. On
-each trial, a single Alpha category exemplar was presented, and participants
+each trial, a single `Alpha' category exemplar was presented, and participants
 were given as much time as they desired to observe it before moving on to the
 next trial. Each block consisted of a single presentation of each of the members
-of the Alpha category, in a random order. Participants were shown the range of
+of the `Alpha' category, in a random order. Participants were shown the range of
 possible colors and sizes prior to training.
@@
 one of the features was changed. Participants could generate any example along
-an evenly-spaced 9x9 grid (including members of the Alpha category), except for
-any previously generated Beta exemplars. Neither the members of the Alpha
-category nor the previously generated Beta examples were visible during
+an evenly-spaced 9x9 grid (including members of the `Alpha' category), except for
+any previously generated `Beta' exemplars. Neither the members of the `Alpha'
+category nor the previously generated `Beta' examples were visible during
 generation. Prior to beginning the generation phase, participants read the
@@
 between- class distance in each condition. All conditions possessed greater
-between-category distance: Cluster, $t(63) = 11.43$, $p < 0.001$; Row, $t(60) =
-13.16$, $p < 0.001$; and XOR, $t(57) = 3.64$, $p < 0.001$. These results provide
+between-category distance: Cluster, $t(63) = 11.43$, $p < .001$; Row, $t(60) =
+13.16$, $p < .001$; and XOR, $t(57) = 3.64$, $p < .001$. These results provide
 further evidence of an effect of category contrast: Participants prefer to
@@
 properties as previously-known categories.
-Given the increased emphasis of replication within psychology \citep{zwaan2017}, it is important as it serves as a conceptual replication of \citet{jern2013probabilistic}.
+Given the increased emphasis of replication within psychology \citep{zwaan2018making}, it is important as it serves as a conceptual replication of \citet{jern2013probabilistic}.
 For each generated category, we
@@
 categories. The categories generated in the Cluster condition were less widely
-distributed along the X-axis compared to Row, $t(123) = 5.61$, $p < 0.001$, and
-XOR, $t(120) = 2.68$, $p < 0.01$. Categories generated in the XOR condition
+distributed along the X-axis compared to Row, $t(123) = 5.61$, $p < .001$, and
+XOR, $t(120) = 2.68$, $p < .01$. Categories generated in the XOR condition
 were also less widely distributed along the X-axis compared to Row, $t(117) =
-2.56$, $p < 0.05$. This latter effect was not expected because the
+2.56$, $p = .046$. This latter effect was not expected because the
 experimenter-defined categories for XOR and Row had similar X-ranges. However,
@@
 Likewise, categories generated in the Row condition had less Y-axis range
-compared to Cluster, $t(123) = 4.57$, $p < 0.001$ and XOR, $t(117) = 9.26$, $p <
-0.001$, and categories from the Cluster condition had less Y-axis range compared
-to XOR, $t(120) = 3.95$, $p < 0.001$. As expected, the correlations in the
+compared to Cluster, $t(123) = 4.57$, $p < .001$ and XOR, $t(117) = 9.26$, $p <
+.001$, and categories from the Cluster condition had less Y-axis range compared
+to XOR, $t(120) = 3.95$, $p < .001$. As expected, the correlations in the
 Cluster and Row conditions were not systematically positive or negative ($ps >
-0.1$). However, the generated categories in the XOR condition tended to possess
-{\em negatively} correlated dimensions, $t(57) = 2.04$, $p = 0.046$. This
+.1$). However, the generated categories in the XOR condition tended to possess
+{\em negatively} correlated dimensions, $t(57) = 2.04$, $p = .046$. This
 finding is notable, as it is the opposite of what would be expected based on previous literature \citep{jern2013probabilistic}, assuming
@@
 
-We believe that the failure to replicate dimension correlation emulation is because participants in
-\cite{jern2013probabilistic} could differentiate the generated category on a
-third dimension (hue) to maintain the statistical regularities on the other two
-dimensions. Although the correlation in the XOR condition is significantly
+We believe that the failure to replicate dimension correlation emulation is
+because participants in \cite{jern2013probabilistic} could differentiate the
+generated category on a third dimension (hue) to maintain the statistical
+regularities on the other two dimensions. In addition, in our XOR condition the
+stimuli were constrained to the corners of the feature space, whereas in the
+positive diagonal condition of \cite{jern2013probabilistic} the stimuli had no
+such constraint. Although the correlation in the XOR condition is significantly
 negative, it is clear from the box-plot in Figure~\ref{fig:e1-statsboxes} that
@@
 interesting interaction between contrast with a given category and emulation of
-statistical properties. That is, the constraints on creative generation imposed
-by category contrast may not simply influence the {\em location} of generated
+statistical properties. That is, the constraints on generation imposed by
+category contrast may not simply influence the {\em location} of generated
 categories, but also their distributional structure. In Experiment 2, we test
@@
 
-\section{Experiment 2.}
+\section{Experiment 2}
 
@@
 correlation between features), as they only differ in their Y-axis position: the
-`Bottom' category lies near the bottom of the space, and the `Middle'
-category lies in the center. The distributional equality of these conditions is
-key to the design of the experiment: If the distributional structure of
-previously learned categories were the only influence on the generated
-categories, we should observe no difference in the categories participants
-generate between these two conditions. Will participants distribute their
-generated category differently between conditions due to the differences in the
-available stimulus space for generating a new category?
+`Bottom' category lies near the bottom of the space, and the `Middle' category
+lies in the center. The distributional equality of these conditions is key to
+the design of the experiment: If the distributional structure of previously
+learned categories were the only influence on the generated categories, we
+should observe no difference in the categories participants generate between
+these two conditions. Will participants distribute their generated category
+differently between conditions due to the differences in the available empty
+stimulus space for generating a new category?
 
@@
 the categories in each condition tended to possess more between-category than
-within-category distance: Bottom, $t(60) = 5.5$, $p < 0.001$; Middle, $t(60) =
-2.71$, $p < 0.01$. We did, however, observe a notable subgroup of participants
+within-category distance: Bottom, $t(60) = 5.5$, $p < .001$; Middle, $t(60) =
+2.71$, $p < .01$. We did, however, observe a notable subgroup of participants
 in each condition who generated categories with more within-category than
@@
 condition, we observed tightly clustered and widely distributed categories along
-each dimension, as well as positively, negatively, and uncorrelated categories.
+each dimension. Although most participants generated uncorrelated categories in
+both conditions, many still produced positively and negatively correlated
+categories.
 
@@
     \caption{Box-plots of the distributional statistics from the categories
-generated in Experiment 2. }
+generated in Experiment 2. Boxes depict the median and quartiles of each
+condition, with whiskers placed at 1.5 IQR. All points outside this region are
+marked individually.}
     \label{fig:e2-statsboxes}
@@
 $t$-test between the conditions is, in this case, diminished due to the
-non-normality of the data (Shapiro-Wilk normality test $W=0.77, p< 0.001$ for
-the Middle condition and $W=0.85, p < 0.001$ for the Bottom condition). Figure
+non-normality of the data (Shapiro-Wilk normality test $W=0.77, p < .001$ for
+the Middle condition and $W=0.85, p < .001$ for the Bottom condition). Figure
 \ref{fig:e2-yranges} depicts the Y-axis position of the exemplars generated
@@
 conditions reveal that more Middle participants generated an exemplar in the
-bottom row, $p < 0.001$, again demonstrating the role of contrast in guiding
+bottom row, $p < .001$, again demonstrating the role of contrast in guiding
 where exemplars are generated. The conditions did not differ in use of the top
-of the space, $p = 0.16$, however, more Middle participants placed exemplars in
-the top {\em and} bottom rows, $p < 0.05$. The latter effect is of particular interest
+of the space, $p = .16$, however, more Middle participants placed exemplars in
+the top {\em and} bottom rows, $p = .04$. The latter effect is of particular interest
 here, as it indicates that the shape of the unoccupied space exerts some
@@
 in the Middle condition were more likely to generate a category spanning the
-entire Y-axis. Thus, distributional structure of the generated categories can be
-influenced by category contrast alone.
+entire Y-axis. Thus, the distributional structure of the generated categories
+can be influenced without any change to the distributional structure of the
+given category. Rather, it can be affected by category contrast alone.
 
@@
 \section{Model-based Analyses of Experiments 1 and 2}
-\sxlnote{For this section I've added the existence of the representativeness
-  model but I think discussion of its performance is still lacking compared to PACKER.}
+
 Experiments 1 and 2 revealed systematic and strong effects of category contrast
-on creative generation. In this section, we analyze the performance of the different formal models at explaining the experimental results. Specifically, we
+on category generation. In this section, we analyze the performance of the
+different formal models at explaining the experimental results. Specifically, we
 present simulations from the two novel contrast models: PACKER and the
@@
 has complementary strengths and weaknesses: Whereas PACKER and copy-and-tweak
-are insensitive to the distributional structure of learned categories (relying
-only on exemplar similarities), the representativeness and hierarchical Bayesian
-model generates categories exclusively on the basis of knowledge of how existing
-classes are distributed.
+are relatively insensitive to the distributional structure of learned categories
+(relying only on exemplar similarities), the representativeness and hierarchical
+Bayesian model generates categories exclusively on the basis of knowledge of how
+existing classes are distributed.
 
@@
 \textbf{Representativeness} & \textbf{Hierarchical Bayesian} \\ \hline $AIC =
-  9095$ & $AIC = 9842$ & $AIC = 8878$ &  $AIC =
-9912$ \\ $L = -4545$ & $L = -4919$ & $L = -4435$ & $L = -4952$ \\ $c = 0.483$ &
-                                                                                $c = 3.197$ & $\kappa = 11.232$ &
-$\kappa < 0.001$ \\ $\theta_c = 3.16$ & $\theta_c = 0$ (fixed) & $\nu = 2.844$ & $\nu = 5.597$ \\
-$\theta_t = 3.494$ & $\theta_t = 2.972$ & $\lambda = 1.917$& $\lambda
-                                                                       = 0.055$
-  \\ & &$\theta = 5.885 $& $\theta = 3.175$
+  9069$ & $AIC = 9813$ & $AIC = 8783$ &  $AIC =
+9881$ \\ $L = -4531$ & $L = -4905$ & $L = -4388$ & $L = -4937$ \\ $c = 0.51$ &
+                                                                                $c = 3.22$ & $\kappa = 12.23$ &
+$\kappa < 0.001$ \\ $\theta_c = 3.09$ & $\theta_c = 0$ (fixed) & $\nu = 1.00$ & $\nu = 5.44$ \\
+$\theta_t = 3.47$ & $\theta_t = 3.00$ & $\lambda = 7.04$& $\lambda
+                                                                       = 0.06$
+  \\ & &$\theta = 10.21 $& $\theta = 3.09$
 \\
@@
 
-\sxlnote{The argument in the following two paragraphs appear to be challenged by
-  the success of the representativeness model. I might suggest removing them altogether?}
 While PACKER's advantage may tentatively be attributed to the model's
@@
 
-A key distinction between PACKER and the other three models is that only PACKER
-is capable of making strong predictions about the location of new category
-members when the target class is entirely novel (i.e., no member of the category
-has been observed). Under these circumstances, there are no examples to copy,
-and thus the copy-and-tweak model predicts that items are generated at random.
-Likewise, with no observations on which to condition the category distribution,
-
-\jlanote{Do they really pick the first item at random? That doesn't seem right for the representativeness -- shouldn't it be constrained?}
-the hierarchical Bayesian and representativeness models also picks an item at
+A key distinction between the contrast and classical models is that only the
+contrast models are capable of making strong predictions about the location of
+new category members when the target class is entirely novel (i.e., no member of
+the category has been observed). Under these circumstances, there are no
+examples to copy, and thus the copy-and-tweak model predicts that items are
+generated at random. Likewise, with no observations on which to condition the
+category distribution, the hierarchical Bayesian model also picks an item at
 random.
 
-Crucially, however, the next generated exemplar from the
-representativeness model is strongly influenced by contrast as
-representativeness. Thus, it is possible that the failure of these models is
-simply due to their inability to explain each participant's first trial
-(generating the first item in the `Beta' category). To ensure this is not
-driving our results, we conducted an identical set of simulations as above,
-excluding this trial (leaving 915 responses in the dataset): Again, PACKER
-($L = -3390$, $AIC = 6786$) achieved better fits than the copy-and-tweak
-($L = -3579$, $AIC = 7162$) and hierarchical Bayesian ($L=-3612$, $AIC = 7232$)
-models.
-\jlanote{last sentence missing representativeness fit}
+Thus, it is possible that the failure of the classical models is simply due to
+their inability to explain each participant's first trial (generating the first
+item in the `Beta' category). To ensure this is not driving our results, we
+conducted an identical set of simulations as above, excluding this trial
+(leaving 915 responses in the dataset): Again, the representativeness model ($L
+= -3286$, $AIC = 6580$) and PACKER ($L = -3377$,
+$AIC = 6759$) achieved better fits than the copy-and-tweak ($L = -3564$,
+$AIC = 7132$) and hierarchical Bayesian ($L = -3597$, $AIC = 7201$) models.
 
@@
 ratio test to compare the two models. PACKER explains the aggregate data
-significantly better than copy-and-tweak ($\chi^2(1) = 749, p < 0.001$ for all
-data and $\chi^2(1) = 377, p < 0.001$ excluding the first example), providing
+significantly better than copy-and-tweak ($\chi^2(1) = 747, p < .001$ for all
+data and $\chi^2(1) = 375, p < .001$ excluding the first example), providing
 further evidence that category generation is better explained when contrast is
@@
 \begin{figure}
-    \begin{center} \inputpgf{figs/}{packer-loglike.pgf}
-    	\jlanote{fix x-axis label}
+    \begin{center}
+      \includegraphics[width=\textwidth]{figs/packer-loglike.pdf}    
       \caption{PACKER's fit as a function of its prioritization of
-        within-category and between-category similarity (using the
-        $\theta_c$ parameter) . To facilitate comparison, PACKER's other
-        parameters ($c$, $\theta_t$) were set to the best fitting values
-        obtained for copy-and-tweak in Table \ref{table:global-model-fits}. }
+        within-category and between-category similarity (using the $\theta_c$
+        parameter) . To facilitate comparison, PACKER's other parameters ($c$,
+        $\theta_t$) were set to the best fitting values obtained for
+        copy-and-tweak in Table \ref{table:global-model-fits}. The black diamond
+        marker indicates the log-likelihood for the point where
+        $\theta_c=\theta_t$. }
     \label{fig:packer-loglike}
@@
 some explanation of how different profiles of performance emerge. Many of the
-individual generation profiles we observed can be described within the PACKER
-framework, simply by tuning the model's parameters in a principled manner. In
-this section, we describe more specifically how the most frequently observed
-profiles can be realized.
+individual generation profiles we observed can be described with the models
+simply by tuning the model's parameters in a principled manner. In this section,
+we describe more specifically how the most frequently observed profiles can be
+realized.
 
@@
 
-\jlanote{Xian: Please add a paragraph or two on the representativeness model here}
-
-Similar to the copy-and-tweak model, the hierarchical Bayesian model possesses
-no mechanism to account for category contrast, and the model is most likely to
-generate new items that are similar to known examples of the target category.
-Although this precludes an account of why individuals might assume a cluster
-versus corners profile, the covariance matrix specifying the model's prior
-domain distribution, $\Sigma_0$, can be used to explain the generation of
+For both the hierarchical Bayesian and the representativeness models, the prior
+domain covariance matrix $\Sigma_0$ can be used to explain the generation of
 row-like and column-like categories. This covariance matrix specifies the amount
@@
 coded into $\Sigma_0$ to alter the dimensional variance of generated categories.
-
-As noted above, while the copy-and-tweak and hierarchical Bayesian models
-possess mechanisms to explain row- and column-like categories, they cannot
-easily explain why some individuals widely distribute their generated categories
-into disparate corners of the space. This, however, reveals a more general
-limitation: According to the copy-and-tweak and hierarchical Bayesian models,
-the distributional structure of generated categories is {\em independent} of
-their location within the domain. For example, although the copy-and-tweak or
-hierarchical Bayesian models can be parameterized to generate row- or
-column-like categories, there is no mechanism in place to ensure that what is
-generated will be distinct from what is already known. In the next subsection,
-we explore this prediction through an analysis of the interdependence between
-distributional structure and location in creative generation.
+Because the hierarchical Bayesian model possesses no mechanism to account for
+category contrast, the model is most likely to generate new items that are
+similar to known examples of the target category with no regard for how
+different it is to the contrast category. However, the representativeness model
+predicts that new exemplars should provide more relative evidence to the `Beta'
+category, accounting for the tendency of row-like and column-like profiles
+occupying the edge of the feature space.
+
+While the copy-and-tweak and hierarchical Bayesian models possess mechanisms to
+explain row- and column-like categories, they cannot easily explain why some
+individuals widely distribute their generated categories into disparate corners
+of the space. This, however, reveals a more general limitation: According to the
+copy-and-tweak and hierarchical Bayesian models, the distributional structure of
+generated categories is {\em independent} of their location within the domain.
+For example, although the copy-and-tweak or hierarchical Bayesian models can be
+parameterized to generate row- or column-like categories, there is no mechanism
+in place to ensure that what is generated will be distinct from what is already
+known. In the next subsection, we explore this prediction through an analysis of
+the interdependence between distributional structure and location in category
+generation.
 
@@
 methods to find the optimal attention weighting scheme given a classification
-\citep[see][]{vanpaemel2012using}, for simplicity we assume that the Alpha
-and Beta categories are distinct along dimensions that the Betas do not vary on.
-In this case, the weighting for a given participant can be computed as:
+\citep[see][]{vanpaemel2012using}, for simplicity we assume that the `Alpha' and
+`Beta' categories are distinct along dimensions that the `Beta' exemplarss do
+not vary on. In this case, the weighting for a given participant can be computed
+as:
 
@@
 
-\include{methods_results_e3}
+%Experiment 3 starts here
+\section{Experiment 3}
+
+Experiments 1 and 2 clearly establish the importance of contrast in category
+generation. Follow-up model-based analyses illustrated that both contrast models
+account for participant performance better than models that do not take into
+account contrast. We also found more support for contrast as representativeness
+hypothesis over contrast as exemplar dissimilarity hypothesis.
+
+Although category generation is an intriguing aspect of human cognition, it has
+played a relatively limited role in the broader categorization literature. To
+establish that contrast plays a broader role in categorization, we conducted a
+traditional category learning study. In it, participants learned two categories,
+where the categories learned by each participant in Experiment 3 was yoked to
+the `Alpha' shown and `Beta' generated by each participant in Experiment 2. Our goal
+is to investigate whether model performance (in generating categories) is
+consistent with human performance (in learning those categories). Specifically, if contrast plays a broader role in categorization, then `Beta' categories rated as "better" according
+to the contrast models should be easier for participants to learn.
+%\footnote{It is possible to adjust the present models of category generation such that they can make category learning predictions.
+However, such an adjustment makes the strong assumption that the very same processes governing category generation also apply to category learning.
+Taking this approach would fall outside the scope of the current investigation.}
+\subsection{Participants, Materials, and Procedure}
+
+Experiment 2 recruited 122 participants who each generated one set of four `Alpha'
+and four `Beta' category exemplars. Experiment 3 recruited the same number of
+participants.
+
+Participants observed four blocks of eight trials. Each trial began with the
+presentation of a fixation cross for 500 ms. This was followed by the
+presentation of one exemplar randomly sampled without replacement from the
+unique category set. Participants were tasked with assigning the presented
+exemplar to either the `Alpha' or `Beta' category with no time limit imposed.
+Feedback was automatically displayed for 2500 ms after each response.
+
+\subsection{Results and Analysis}
+
+Among the 122 originally generated category sets were 102 unique sets (i.e.,
+they contained a unique collection of `Alpha' and `Beta' exemplars). Consequently,
+for our subsequent analyses for this experiment we only use data from the first
+102 participants that were presented with a unique category set.
+
+Overall accuracy of the participants was high, with a mean error rate of $.19$
+($SD = .19$). Aggregated error rates for each block are presented in Figure
+\ref{fig:learningcurve}.
+
+\begin{figure}
+    \begin{center}
+    \includegraphics[width=\textwidth/2]{figs/e3-learningcurve.pdf}
+    \caption{Average error rate for each successive block. Green discontinuous
+line represents the overall mean error rate.}
+    \label{fig:learningcurve}
+    \end{center}
+\end{figure}
+
+In order to analyze the consistency between model and human performance, we
+optimized each model such that the Pearson correlation between the model's
+negative log-likelihood of generating each unique category set and the
+participant's error in learning that category set is maximized. As with the
+analysis in Section \ref{section:individual-diff}, we simulated the weighting of
+each dimension independently depending on the generated categories.
+
+The greatest correlations were observed when fitting the contrast models.
+Specifically, PACKER correlated with human performance with
+$r = .65, 95\%\,CI\, [.48,.78]$, and the representativeness
+model correlated with human performance with $r = .59, 95\%\,CI\,[.40, .73]$\footnote{Confidence intervals were obtained using a bootstrap
+  method with 1000 bootstrap samples.}.
+Copy-and-tweak yielded a correlation of $r = .45, 95\%\,CI\,[.28, .62]$ and the
+hierarchical Bayesian model yielded a similar correlation of
+$r = .46, 95\%\,CI\, [.30, .62]$. These results are graphically presented in Figure
+\ref{fig:perror_corr}.
+
+\begin{figure}
+    \begin{center}
+    \includegraphics[width=\textwidth]{figs/modelvspptp.pdf}
+    \caption{Scatterplots indicating the correlations between observed
+      participant error and model fit. Blue discontinuous line are best-fit
+      lines. Green discontinuous lines indicate boundaries generated by the 95\%
+      CI around the fitted correlations. }
+    \label{fig:perror_corr}
+    \end{center}
+\end{figure}
+
+To emphasize the strong influence of contrast in maximizing the association
+between category generation and category learning, we computed the correlations
+over wide range of $\theta_{contrast}$ values, with the other parameter values
+of PACKER held constant at their optimized levels. As presented in Figure
+\ref{fig:packer-corr}, the correlation quickly increases with increasing weight
+on contrast, reaching a plateau above values of around 4.0.
+
+\begin{figure}
+    \begin{center}
+    \includegraphics[width=0.8\textwidth]{figs/packer-corr.pdf}
+    \caption{Correlation between PACKER's fit and participant error as a
+function of the $\theta_{contrast}$ parameter. To facilitate comparison,
+PACKER's other parameters ($c$, $\theta_{target}$) were set to the best fitting
+values obtained for copy-and-tweak in Table \ref{table:global-model-fits}. Grey
+dashed line represents the correlation between Copy \& Tweak's fit and
+participant error.}
+    \label{fig:packer-corr}
+    \end{center}
+\end{figure}
 
@@
 
-While the bulk of prior research on categorization has focused on the classic finding
-that generated concepts tend to be distributionally similar to known concepts,
-there has been little work addressing the role of contrast in creative
+While the bulk of prior research on categorization has focused on the classic
+finding that generated concepts tend to be distributionally similar to known
+concepts, there has been little work addressing the role of contrast in category
 generation: How is it that people are able to create something {\em different}
-from what is already known?
-\jlanote{Update this next section:}
-We developed a novel, exemplar-based model, PACKER,
-which formally specifies the role of contrast in generation. The model proposes
-that categories are represented as exemplars in a multidimensional psychological
+from what is already known? We developed two novel models, each incorporating a
+different conceptualization of category contrast. One model is PACKER, which is
+an exemplar-based model that formally specifies the role of contrast in
+generation as exemplar dissimilarity. Specifically, the model proposes that
+categories are represented as exemplars in a multidimensional psychological
 space, and generation is constrained both by within-category and
@@
 similar to one another, and exemplars belonging to different categories should
-not be similar to one another.
-
-We reported two experiments demonstrating systematic
-effects of contrast in category generation. Members of
-participant-generated categories tended to be highly dissimilar from members of
-previously-learned categories, and were usually more similar to one another than
-to members of other categories. We also observed broad interdependence between
-the distributional structure (feature variance, correlation) and physical
-instantiation (location within the stimulus space) of generated categories: In
-Experiment 2, we found that the unoccupied regions of the domain influenced the
-distributional structure of categories, and in both experiments we observed that
-participants distributed their generated categories to increase contrast with
-what was already known.
-
-We conducted simulations comparing contrast accounts of our results to the other
-major proposals for category generation from prior work: a ``copy-and-tweak'' model (realized as
-a variant of PACKER with no sensitivity to category contrast), and a
-hierarchical Bayesian model designed to explain the classic distributional
-similarity effect. In all simulations, we found that 
-contrast captured a previously unexplained and unexplored aspect of human
-category generation. In the case of contrast as exemplar dissimilarity, we operationalized contrast by measuring PACKER's fit as a function of its
+not be similar to one another. The second model is a novel hierarchical Bayesian
+model with a representativeness mechanism. This model generates exemplars that
+are more representative of (i.e., has greater relative evidence from) the novel
+category compared to the learned category.
+
+We reported two experiments demonstrating systematic effects of category
+contrast in category generation. Members of participant-generated categories
+tended to be highly dissimilar from members of previously-learned categories,
+and were usually more similar to one another than to members of other
+categories. We also observed broad interdependence between the distributional
+structure (feature variance, correlation) and physical instantiation (location
+within the stimulus space) of generated categories: In Experiment 2, we found
+that the unoccupied regions of the domain influenced the distributional
+structure of categories, and in both experiments we observed that participants
+distributed their generated categories to increase contrast with what was
+already known.
+
+We conducted simulations comparing the contrast models' account of our results
+to the classical proposals for category generation: a ``copy-and-tweak'' model
+(realized as a variant of PACKER with no sensitivity to category contrast), and
+a hierarchical Bayesian model designed to explain the classic distributional
+similarity effect. In all simulations, we found that the contrast models
+captured a previously unexplained and unexplored aspect of human category
+generation. In particular, by measuring PACKER's fit as a function of its
 prioritization of within- and between-category similarity, we observed that
@@
 class-contrast. This finding mirrors our behavioral results and demonstrates
-that both constraints influence creative generation.
-\jlanote{Above needs some updating still. Also, do we want an outline of general discussion paragraph?}
-
-\subsection{Contrast Effects in Prior Work}
+that both constraints influence generation when explained with an exemplar
+model. Ultimately, our category generation data were best explained by the
+representativeness model. The exact nature of its superiority over the other
+models is still not yet clear. However, as we explore in a later section, it is
+likely that the representativeness mechanism itself (i.e., independent of the
+hierarchical Bayesian foundation) captures a core aspect of both category
+generation and category learning.
+
+\subsection{Finding Contrast Effects in Prior Work}
+
 Although existing accounts of category generation broadly overlook the role of
@@
 sensible that participants might generate a category with a hue distinct from
-the experimenter-provided categories. If creative generation were influenced by
+the experimenter-provided categories. If category generation were influenced by
 category contrast in this way, the hues of generated categories should be
@@
 the experimenter-defined classes. The ``Expected'' data follows a Binomial
-			distribution with $p=2/8=1/4$, given there were two experimenter-defined
+distribution with $p = 2/8 = 1/4$, given there were two experimenter-defined
 classes, and eight color groups.}
@@
 highly inconsistent with the hues being chosen at random, all
-$\chi^2(6,N=22)>200$, $p<0.001$. Participants tended to generate items that were
+$\chi^2(6,N=22)>200$, $p < .001$. Participants tended to generate items that were
 perceptually distinct from the categories they had learned, and were less likely
@@
 that share the same hue within a category, it is plausible that they generate
-exemplars that all share a common hue. If this is the case, given a $6/8 = .75$
+exemplars that all share a common hue. If this were the case, given a $6/8 = .75$
 probability that a hue distinct from the experimenter-defined categories is
-selected, there is a corresponding 75\%  chance that the generated category will have no exemplars that share a
-hue with any of the experimenter-defined categories. Conversely, that there is
-a 25\% chance that the generated
+selected, there is a corresponding $.75$ probability (or about 17 out of 22
+participants) that the generated category will have no exemplars that share a
+hue with any  of the experimenter-defined categories. In addition, that there is
+a $2/8 = .25$ probability (about 6 out of 22 participants) that the generated
 category will have all six exemplars that share a hue with any of the
@@
 
-\jlanote{Xian: Please do a Chi-Squared test between what is expected based on the logic above and the observed frequencies mentioned below.}
 However, upon analyzing the consistency of hues in the generated categories, it
@@
 
-Re-analyzing the results from \cite{jern2013probabilistic} provides preliminary
-support that contrast plays a role in category generation. Taken alongside the
-analyses reported by \cite{jern2013probabilistic}, our analysis suggests that
-generated categories tend to be distinct from {\em and} distributionally similar
-to what is already known. However, it is worth noting that our analysis is still
-limited: The color groups defined above are imprecise, and it is not clear that
-our color grouping is consistent with the psychological color boundaries
-perceived by participants. While we did obtain similar results using a variety
-of alternative groupings, the hue dimension used in the
-\cite{jern2013probabilistic} study does not lend itself straightforwardly to the
-the computation of similarities, and thus we cannot be certain of whether our
+Re-analyzing the results from \cite{jern2013probabilistic} provides some
+corroborating support for contrast playing a role in category generation. Taken
+alongside the analyses reported by \cite{jern2013probabilistic}, our analysis
+suggests that generated categories tend to be distinct from {\em and}
+distributionally similar to what is already known. However, it is worth noting
+that our analysis is still limited: The color groups defined above are
+imprecise, and it is not clear that our color grouping is consistent with the
+psychological color boundaries perceived by participants. While we did obtain
+similar results using a variety of alternative groupings, the hue dimension used
+in the \cite{jern2013probabilistic} study does not lend itself straightforwardly
+to the computation of similarities, and thus we cannot be certain of whether our
 coding accurately approximates the psychological space of the stimuli. This
@@
 these results likely indicate that contrast exerts {\em some} influence, they do
-not precisely describe the nature of that influence. In the sections below, we
-propose a quantitative framework specifying the role of category contrast in
-creative generation.
+not precisely describe the nature of that influence.
 
@@
 cognitive functions in a wide variety of domains. The PACKER model is notable in
-that it explicitly interprets this trade-off within the domain of categorization, and allows us to begin to understand the relatively understudied
-processes involved in category generation through our more well-developed
-knowledge of human categorization.
+that it explicitly interprets this trade-off within the domain of
+categorization, and allows us to begin to understand the relatively understudied
+processes involved in category generation through our more well-developed tools
+for understanding human categorization.
 
 \subsection{Implications for Creative Cognition}
-\jlanote{this needs updating for representativeness.}
+
 Although the focus of this article has been to address the role of contrast in category generation and learning, our findings and approach have relevant implications
@@
 
-In addition to demonstrating the utility of formal modeling in the study
-creative cognition, the PACKER model more specifically offers an additional
+In addition to demonstrating the utility of formal modeling for studying
+creative cognition, the contrast models here specifically offer an additional
 interpretation of some of the field's most central findings. For example,
@@
 Further, there is systematicity in how generated categories differ from prior
-knowledge. The results of our simulations with PACKER demonstrate that this
-influence is concisely explained in terms of a trade-off between within-category
-similarity and between-category dissimilarity.
-
-Similarly, PACKER may offer an additional interpretation of existing accounts of
-creative generation. Most notably, a leading account within the creative
-cognition literature, the Path of Least Resistance
-\citep{ward1994structured,ward1995s}, also explains generation in terms of an
-exemplar-based retrieval process. This account was designed to explain the
-creative generation of natural categories (e.g., new species of plants and
-animals) and as a result relies strongly on the hierarchical organization of
-these categories: Individuals are thought to retrieve an example of the
-higher-level category being generated (e.g., \textit{bird} may be retrieved from
-the category \textit{animal}), and then systematically alter what was retrieved
-to make something new. As the PACKER model does not assume knowledge is
-hierarchically organized \citep[this is true of the exemplar view more broadly,
-see][]{murphy2016exemplar}, the model may be viewed as a formal instantiation of
-the Path of Least Resistance for application in a traditional artificial
-categorization domain (when there is no established hierarchy of categories).
-PACKER's success in explaining generation within an artificial domain motivates
-future work exploring the nature of category contrast within a more naturalistic
-setting.
+knowledge. The results of our simulations suggest that this conceptualization of
+difference can be addressed in at least two different ways. Specifically,
+simulations with PACKER demonstrate that the constraining influence of
+difference is concisely explained in terms of a trade-off between
+within-category similarity and between-category dissimilarity. Conversely, the
+representativeness model shows that this influence can also be the result of
+enhancing the representativeness of generated exemplars to their category.
+
+PACKER may offer an additional interpretation of existing accounts of creative
+generation. Most notably, a leading account within the creative cognition
+literature, the Path of Least Resistance \citep{ward1994structured,ward1995s},
+also explains generation in terms of an exemplar-based retrieval process. This
+account was designed to explain the creative generation of natural categories
+(e.g., new species of plants and animals) and as a result relies strongly on the
+hierarchical organization of these categories: Individuals are thought to
+retrieve an example of the higher-level category being generated (e.g.,
+\textit{bird} may be retrieved from the category \textit{animal}), and then
+systematically alter what was retrieved to make something new. As the PACKER
+model does not assume knowledge is hierarchically organized \citep[this is true
+of the exemplar view more broadly, see][]{murphy2016exemplar}, the model may be
+viewed as a formal instantiation of the Path of Least Resistance for application
+in a traditional artificial categorization domain (when there is no established
+hierarchy of categories). PACKER's success in explaining generation within an
+artificial domain motivates future work exploring the nature of category
+contrast within a more naturalistic setting.
+
+The broader study of creativity currently involves a wide breadth of different
+approaches,\citep[for a review see][]{kozbelt2010theories}, such as those based
+on free association \citep{mednick1962associative} and conceptual combination
+\citep{estes2002emergence,murphy1988comprehending}. In addition, recent work in
+the machine learning literature has explored using neural networks to address
+the overall problem of creative generation
+\citep[e.g.,][]{goodfellow2014gan,ho2016gail,kingma2016,chen2016infogan}. In
+contrast to these varied investigations, we reduced our focus by studying a
+highly complex behavior (category generation) as it applies within a
+well-established domain (artificial category learning). We hope that future work
+incorporates and highlights the importance of contrast into theories of
+creativity.
 
 \subsection{Implications For Categorization}
-
-Although our work has mainly addressed the processes involved in the creative
-generation of concepts, in this paper we have studied these processes as they
-apply within an artificial categorization domain. By consequence, our findings
-also provide an advance in our understanding categorization more broadly.
 
@@
 beyond the available information. Our work is fairly unique in that people learn
-a single category through positive examples and then figure out another category
+a single category through positive examples and then create another category
 that would make sense in the domain. In other words, only one category is
 provided to be learned and the use that is required of that category is to
-generate a contrast category. Such propagating of categories (producing one from
-another) may not seem like a fundamental or ubiquitous cognitive activity, but
-consider navigating a highly unfamiliar environment -- it would be practical,
-perhaps even critical, to generate expectations about what other kinds of things
-you are likely to encounter based on those that you already have. In the present
-studies, we have learned something about the form that such expectations are
-likely to take.
+generate a contrast category. Such scenarios have direct application to
+real-world situations. For instance, consider being stranded in an unfamiliar
+environment where you eat and get sick from a plant with certain features. In
+avoiding such plants and seeking alternative sources of nutrition, it would
+behoove you to imagine what features an edible, safe plant would possess (e.g.,
+if the poisonous plant was bright red, perhaps a non-red plant would be safe to
+consume). In the present studies, we have learned something about the form that
+such expectations are likely to take.
 
@@
 
-\subsection{Exemplar dissimilarity and Representativeness in Categorization}
-
-\sxlnote{I'm struggling a little with this section! That said, I think this may
-  also be a good point to address Charles' concern re: the possible conditions
-  under which contrast are important.} 
-\jlanote{Ken, could you take a stab at this section please?}
-Although both PACKER and the
-representativeness model perform better than the other models, they appear to be
-capturing fundamentally different aspects of contrast in categorization.
-Specifically, while the representativeness model excelled in predicting our
-category generation data, PACKER was the best-fitting model in accounting for
-participant error in category learning. The source of this discrepancy is still
-not yet clear -- it could be an indication of a fundamental difference between
-the processes involved in generation versus learning where exemplar
-dissimilarity is a larger factor in category generation while representativeness
-exerts more influence in category learning.
-
-%Xian: Please
-%write about the two hypotheses and that followup work needs to look into why the
-%Bayesian representativeness model doesn't do as well at explaining category
-%learning ease, but does better at category generation.
-
-\subsection{Limitations and Future Directions} Although we have demonstrated
-that contrast plays a role in categorization, our results were equivocal as to
-whether contrast is governed by exemplar dissimilarity or representativeness.
+The current work also suggests exciting directions for related investigations
+into unsupervised categorization. While the unsupervised categorization
+literature is primarily interested in the generation of categories within a set
+of observed exemplars where no prior categories are learned, our category
+generation studies are focused on the production of novel category exemplars
+themselves. Despite this distinction, both types of categorization research are
+ultimately interested in the question of how categories can be formed. The results
+of our current work indicate that the formation of categories is constrained by
+some measure of contrast. To our knowledge, this explicit investigation of
+category contrast has not yet been attempted in the unsupervised categorization
+literature.
+
+Interestingly, the unidimensional sort bias that is commonly observed in
+unsupervised cateogorization
+\citep{imai1965discriminability,milton2004influence,ahn1992two}, where
+participants sort unlabelled family resemblance category exemplars by focusing
+on a single feature, was not consistently observed in our tasks. In our
+experiments we only observed this bias in the row condition of Experiment 1,
+whereas in other conditions participants generated categories that typically
+varied along both dimensions. This may be unsurprising at face value, since
+participants in conditions other than the row condition were trained on a prior
+category that varied on both dimensions. However, these results suggest that the
+unidimensional sort bias is not universal to all types of category construction
+-- or more specifically, the unidimensional bias may not apply when participants
+need to generate categories with entirely new exemplars.
+
+\subsection{Exemplar Dissimilarity and Representativeness in Categorization} 
+Although both PACKER and the representativeness model perform better than the
+classical models, they appear to be capturing fundamentally different aspects of
+contrast in categorization. Specifically, while the representativeness model
+excelled in predicting our category generation data, PACKER was the best-fitting
+model in accounting for participant error in category learning. Given the
+history of their development, this is perhaps not entirely surprising. PACKER
+is implemented as an extension of the GCM, which was originally developed for (and
+has been highly successful in) category learning \citep{nosofsky1986attention,
+  nosofsky1994comparing}. Conversely, the representativeness mechanism was
+conceived as a heuristic describing the relative extent to which an individual
+observation is generated from some unobserved population
+\citep{kahneman1972subjective,abbott16}. This generative nature of the
+representativeness mechanism intuitively lends itself to the prediction of
+category generation data.
+
+While the contrast models were born quite naturally out of the corresponding
+classical models, it is worth considering if similar benefits in prediction
+would be seen if the contrast mechanisms were applied to different classical
+models. Although there is no clear way to adapt the exemplar dissimilarity
+mechanism to the current hierarchical Bayesian model, it is a fairly
+straightforward process applying the representativeness mechanism to an exemplar
+model. Specifically, by treating the similarity measure (Equation
+\ref{eq:similarity}) in the current implementation of copy-and-tweak as a
+density estimate for the representativeness mechanism (Equation
+\ref{eq:representativeness}), we gain a new model of representativeness that is
+based on exemplar similarity instead of a multivariate Gaussian
+likelihood.\footnote{An analogous implementation of the representativeness
+  mechanism to PACKER results in both $\theta_c$ and $\theta_t$ adding to
+  constant value that is independent of the similarity to contrast and target
+  exemplars respectively. Consequently, it becomes equivalent to a
+  copy-and-tweak model with representativeness.}
+
+How well does a representativeness model that uses exemplar similarity to
+represent the category’s density over features explain our experimental results?
+We find that this model substantially outperforms its classical counterpart, but
+not to the extent of PACKER or the representativeness model. Specifically,
+fitting this model to the category generation data yielded $L = -4815$ and
+$AIC=9633$ for the entire set of data from Experiments 1 and 2, and $L = -3474$,
+$AIC=6953$ for data excluding the first trials. When used to predict the error
+rates of participants in the category learning task (Experiment 3), the
+performance of the copy-and-tweak with representativeness model correlated with
+participant error with $r=.56, p < .001$. These results suggest that the
+predictive advantages to including the representativeness mechanism is not
+limited to a hierarchical Bayesian model, but can also be found when applied to
+an exemplar model. Although its performance is not as strong as either of the
+fully-developed contrast models, there is a benefit to the representativeness
+mechanism that is independent from any interaction with a hierarchical Bayesian
+framework.
+
+Ultimately, the source of the discrepancy between the performance of the two
+contrast mechanisms in category generation versus category learning is still not
+entirely clear. Our results could be an indication of a fundamental difference
+between the processes involved where exemplar dissimilarity is a larger factor
+in category generation while representativeness exerts more influence in
+category learning. Alternatively, it is possible they are both approximating
+some factor that we still have not fully identified. While a thorough
+investigation is beyond the scope of the current article, the preliminary
+results of this section indicate that further work to disentangle these two
+mechanisms would be promising. In the following section, we explore some other
+limitations of the current work and provide more recommendations for future
+research.
+
+\subsection{Limitations and Future Directions}
 
@@
 Nonetheless, these classic effects are a core element of the phenomenology of
-creative generation, and PACKER does not include any mechanisms that explain
+category generation, and PACKER does not include any mechanisms that explain
 them. Instead, through the development and evaluation of the PACKER model, we
@@
 PACKER's underlying claims to obtain a ``best of both worlds'' model, capable of
-explaining the role of contrast in creative generation, as well as the emulation
+explaining the role of contrast in category generation, as well as the emulation
 of distributional structure. However, as noted in the introduction, the
@@
 (as exemplars or prototypes, respectively). Alternatively, it may be possible to
-integrate the core principles of contrast into other categorization models
+integrate the core principles of either model of contrast into other
+categorization models
 \citep[e.g.,][]{love2004sustain,kurtz2007divergent,smith2000thirty}.
-
-Finally, although an overall goal of this work is to advance the study of
-creative thought by explaining generation through empirical and quantitative
-approaches from research in categorization, these approaches are not easily
-aligned with many of the alternative theoretical viewpoints in the broader study
-of creativity \citep[for a review see][]{kozbelt2010theories}, such as those
-based on free association \citep{mednick1962associative} and conceptual
-combination \citep{estes2002emergence,murphy1988comprehending}. Instead, we
-reduced our focus by studying a highly complex behavior (creative generation) as
-it applies within a well-established domain (artificial category learning). We
-hope that future work incorporates and highlights the importance of contrast
-into theories of creativity.
 
@@
 well as a hierarchical Bayesian model, provide strong support for the claim that
-category contrast is of fundamental importance. More generally, our results
-demonstrate that popular explanatory approaches from basic research in cognitive
-science can offer a precise, quantitative account of a behavior as complex as
-that of creative generation.
+category contrast is of fundamental importance to categorization. More broadly,
+our results demonstrate that popular explanatory approaches from basic research
+in cognitive science can offer a precise, quantitative account of a behavior as
+complex as that of category generation, and in exploring them, we can increase
+our understanding of categorization as well.
 
@@
 research was provided by the Office of the VCRGE at the UW - Madison with
-funding from the WARF. We thank Alan Jern and Charles Kemp for providing code
-and data.
+funding from the WARF. We thank Charles Kemp for feedback on a previous version
+of this manuscript. We also thank him and Alan Jern for providing code and data.
 \end{flushleft}

### 2018-09-17 Manuscripts/cog-psych/revision/cogpsych.tex
Re manuscript: some new paragraphs (not yet vetted by Joe) on real world example and discussion on possible influences of contrast. Re: new exp, individual plots and some scripts to make life a little easier.

@@
 contrasting from other categories in the relevant domain, is the focus of our
-work. Although implicitly assumed in some work, this constraint has been
+work. Intuitively, the influence of contrast may be expected in a number of
+real-life scenarios involving the generation of new categories. For instance, a
+college student who has been eating instant noodles for a week may be more
+inclined to seek out something healthier and less carbohydrate-heavy (e.g., a
+Caesar salad) compared to something similar (a different flavour of instant
+noodles). An actor who is dissatisfied with being typecast as a villain may seek
+out roles as heroes as opposed to playing other antagonistic characters. A more
+compelling example is seen in recent evidence by \cite{askin2017} who found that
+new songs that were more different to their peers were generally more popular
+than new songs that were similar to others. It appears that once there is the
+motivation to generate something new, there is an associated motivation to
+generate something \emph{different}.
+
+Although implicitly assumed in some work, this constraint has been
 overlooked in previous research: To our knowledge, there has not been any
@@
 In our simulations, we set the prior mean $\mu_0$ to the center of the stimulus
-space the prior variance to be isotropic ($\Sigma_0 = \lambda \mathbf{I}$ where
+space and the prior variance to be isotropic ($\Sigma_0 = \lambda \mathbf{I}$ where
 $\lambda$ is a free parameter and $\mathbf{I}$ is a $d$-by-$d$ identity matrix
@@
 The broader study of creativity currently involves a wide breadth of different
-approaches,\citep[for a review see][]{kozbelt2010theories}, such as those based
+approaches\citep[for a review see][]{kozbelt2010theories}, such as those based
 on free association \citep{mednick1962associative} and conceptual combination
@@
 Interestingly, the unidimensional sort bias that is commonly observed in
-unsupervised cateogorization
+unsupervised categorization
 \citep{imai1965discriminability,milton2004influence,ahn1992two}, where
@@
   constant value that is independent of the similarity to contrast and target
-  exemplars respectively. Consequently, it becomes equivalent to a
+  exemplars respectively. Consequently, it is formally equivalent to a
   copy-and-tweak model with representativeness.}
@@
 
+The present work focuses on the influence of contrast on categorization,
+primarily by exploring two different conceptualizations of contrast. However,
+throughout this paper we have only explored one type of category generation
+problem, that is the generation of new categories of squares that vary only in
+size and shade. It is possible that the strength of the influence of contrast on
+category generation decisions varies depending on the nature of the category
+generation problem, for instance, in situations where the boundaries of the
+domain are not clearly defined. To provide some intuition for this, consider the
+following example: an entomologist is asked to draw a new type of insect that
+they have never seen. Clearly, features of the new insect will be constrained by
+the definition of an insect: it will be an arthropod with six legs and a
+relatively hard exoskeleton. However, there are no strict limitations on other
+features such as how large the insect needs to be, or if the insect should have
+a particular type of pattern on its body. In this scenario, it is plausible that
+the entomologist will draw an insect that is fairly similar to insects that
+currently exist (e.g., small in size with a camouflage pattern that matches its
+environment) as opposed to something completely different (e.g., a horse-sized
+beetle with roman numerals on its back), indicating that contrast plays only a
+minor role in the decision making process. In this situation it appears that
+there is a combination of clearly defined constraints on the domain boundary
+that are imposed by the external environment (insects must be arthropods with
+six legs) but also ill-defined constraints that are imposed by the observer
+(insects are probably small and have patterns on their body that suit the
+environment). However, the relation of these contraints (and the interactions
+between them) to the role of contrast is not currently well understood.
+
+The definition of the domain boundary is only one possible factor that could
+influence the effect of contrast on categorization. It is also plausible that
+other conditions such as the nature of the instruction (``Generate a new
+category that is Not Alpha'' vs ``Generate a new category that is Beta'') and
+the boundedness of the features (using features such as hue and orientation that
+do have have strict minima or maxima) can affect the extent to which contrast
+influences categorization. Though our current contrast models can allow the
+influence of contrast to vary, they do not make any predictions regarding the
+specific conditions under which contrast can vary. Consequently,
+addressing this issue would be an immensely promising avenue for future
+research.
+
 \section{Conclusions}

### 2018-09-17 Experiments/middle-bottomABC/analysis/plot.heatmaps.py

@@ -17,29 +17,31 @@ alphas = pd.read_sql_query("SELECT * from alphas", con)
 generation = pd.read_sql_query("SELECT * from generation", con)
 con.close()
 
+#generation = generation.loc[generation.trial<4]
+
 from scipy.ndimage.filters import gaussian_filter
 
 clim = (0.0, 0.10)
 
-f, ax = plt.subplots(1,2, figsize = (4,2))
+f, ax = plt.subplots(2,3, figsize = (4,2))
 for i, (K, rows) in enumerate(infodf.groupby('condition')):
-
-	pids = rows.participant
-	betas = generation.loc[generation.participant.isin(pids), 'stimulus']
-	
-	counts = np.array([sum(betas==j) for j in range(stimuli.shape[0])])
-	ps = counts / float(sum(counts))
-	g = funcs.gradientroll(ps,'roll')[:,:,0]
-
-	# g = gaussian_filter(g, 0.8)
-
-	print(g)
-	print(K, np.max(g))
-
-	h = ax[i]
-	im = funcs.plotgradient(h, g, stimuli[alphas[K],:], [], clim = clim)
-	h.set_title(K, fontsize = 12)
-
+    for gentype,rowss in rows.groupby('gentype'):
+        pids = rowss.participant
+        betas = generation.loc[generation.participant.isin(pids), 'stimulus']
+        
+        counts = np.array([sum(betas==j) for j in range(stimuli.shape[0])])
+        ps = counts / float(sum(counts))
+        g = funcs.gradientroll(ps,'roll')[:,:,0]
+        
+        # g = gaussian_filter(g, 0.8)
+        
+        print(g)
+        print(K, np.max(g))
+        
+        h = ax[i,gentype]
+        im = funcs.plotgradient(h, g, stimuli[alphas[K],:], [], clim = clim)
+        h.set_title(K, fontsize = 12)
+        
 
 # add colorbar
 f.subplots_adjust(right=0.8)
@@ -47,6 +49,7 @@ cbar = f.add_axes([0.83, 0.225, 0.03, 0.54])
 f.colorbar(im, cax=cbar, ticks = clim)
 cbar.set_yticklabels(['Lowest\nProbability', 'Greatest\nProbability'])
 cbar.tick_params(length = 0)
+plt.tight_layout(w_pad=-8, h_pad= .5)
 
 f.savefig('heatmaps.pdf', bbox_inches='tight', transparent=False)
 

### 2018-09-17 Experiments/middle-bottomABC/fetchdata.py

@@ -13,3 +13,4 @@ os.system('ssh -i ~/Dropbox/.ssh/luke -p 1202 -t xian@alab.psych.wisc.edu \'{};{
 os.system('scp -i ~/Dropbox/.ssh/luke -P 1202 xian@alab.psych.wisc.edu:{} data/'.format(os.path.join(serverdir,zipfile)))
 #untar data
 os.system('tar -xvzf data/{} -C data/'.format(zipfile))        
+print 'Data extracted to the data folder. Feel free to compile by going into the analysis folder and running \'python compile.py\'.'

### 2018-09-17 Modules/Funcs/Plot.py

@@ -1,138 +1,142 @@
 import numpy as np
 
 def save_as_pgf(fh, path, 
-	texpath = '/Library/TeX/texbin/',
-	custom_opts = None,
-	pad_inches = 0.01):
-	""" 
-	Wrapper to save a pgf file. 
-	"""
-	
-	opts = {
-				'pgf.texsystem': 'pdflatex', 
-				'pgf.rcfonts': True,
-		}
-	if custom_opts is not None:
-		opts.update(textsettings)
+    texpath = '/Library/TeX/texbin/',
+    custom_opts = None,
+    pad_inches = 0.01):
+    """ 
+    Wrapper to save a pgf file. 
+    """
+    
+    opts = {
+        'pgf.texsystem': 'pdflatex', 
+        'pgf.rcfonts': True,
+        }
+    if custom_opts is not None:
+        opts.update(textsettings)
 
-	import os, matplotlib
-	os.environ["PATH"] += os.pathsep + texpath
-	matplotlib.rcParams.update(opts)
-	fh.savefig(path, bbox_inches='tight', pad_inches=pad_inches)
+    import os, matplotlib
+    os.environ["PATH"] += os.pathsep + texpath
+    matplotlib.rcParams.update(opts)
+    fh.savefig(path, bbox_inches='tight', pad_inches=pad_inches)
 
 
 def plotclasses(h, stimuli, alphas, betas,
-	textsettings = None, 
-	spinewidth = 0.5):
+    textsettings = None, 
+                spinewidth = 0.5, betastr = 'B', betacol = [0,0,.5]):
 
-	final_textsettings = dict(
-		verticalalignment='center', 
-		horizontalalignment='center',
-		fontsize = 12.0,
-		fontname = 'sans-serif')
+    final_textsettings = dict(
+        verticalalignment='center', 
+        horizontalalignment='center',
+        fontsize = 12.0,
+        fontname = 'sans-serif')
 
-	if textsettings is not None:
-		final_textsettings.update(textsettings)
+    if textsettings is not None:
+        final_textsettings.update(textsettings)
 
-	h.axis(np.array([-1, 1, -1, 1])*1.2)
-	for i in alphas:
-		x, y = stimuli[i,0], stimuli[i,1]
-		h.text(x, y, 'A', color = [0.5,0,0], **final_textsettings)
+    h.axis(np.array([-1, 1, -1, 1])*1.2)
+    for i in alphas:
[…]

### 2018-09-18 Manuscripts/cog-psych/revision/cogpsych.tex
Manuscript: fixed minor typo, cleaned a few sentences. This version sent as minor update to Ken and Nolan.

@@
 \citep[see][]{vanpaemel2012using}, for simplicity we assume that the `Alpha' and
-`Beta' categories are distinct along dimensions that the `Beta' exemplarss do
+`Beta' categories are distinct along dimensions that the `Beta' exemplars do
 not vary on. In this case, the weighting for a given participant can be computed
@@
 throughout this paper we have only explored one type of category generation
-problem, that is the generation of new categories of squares that vary only in
-size and shade. It is possible that the strength of the influence of contrast on
+problem, that is the generation of new categories of new categories in an
+artificial domain. It is possible that the strength of the influence of contrast on
 category generation decisions varies depending on the nature of the category

### 2018-09-19 Manuscripts/cog-psych/revision/cogpsych.tex
Incorporated minor updates suggested by Ken (in particular the unidimensional sort bias bit.)

@@
 noodles). An actor who is dissatisfied with being typecast as a villain may seek
-out roles as heroes as opposed to playing other antagonistic characters. A more
-compelling example is seen in recent evidence by \cite{askin2017} who found that
+out roles as heroes as opposed to playing other antagonistic characters. An
+especially compelling example can be seen in \cite{askin2017} who found that
 new songs that were more different to their peers were generally more popular
@@
 -- or more specifically, the unidimensional bias may not apply when participants
-need to generate categories with entirely new exemplars.
+need to generate categories with entirely new exemplars. Additionally, the
+unidimensional sort bias is observed in category construction tasks in which
+participants develop a category-level organization of a provided domain while
+category generation tasks involve generating a new category relative to one that
+was just acquired (one might say that participants are forging a domain from a
+category rather than categories from a domain). Since the more creative task of
+generating a category from scratch (with respect to an established category)
+does not invoke the unidimensional bias, this suggests that a more creatively
+demanding version of a sort task might also reduce the bias.
 

### 2018-09-19 Experiments/middle-bottomABC/analysis/descriptives.py

@@ -13,40 +13,47 @@ con.close()
 print participants.shape
 
 # counts per condition
-print(participants.groupby('condition').size())
-print 
+#print(participants.groupby('condition').size())
+print(participants.groupby(['condition','gentype']).size())
+
 
 participants = pd.merge(participants, counterbalance, on = 'counterbalance')
 print pd.pivot_table(
-	data = participants,
-	columns = 'xax',
-	index = 'condition',
-	aggfunc = 'size'
-	)
+    data = participants,
+    columns = 'xax',
+    index = 'condition',
+    aggfunc = 'size'
+    )
 
 
 
 stats = pd.merge(stats, participants, on = 'participant')
 from scipy.stats import ttest_ind, mannwhitneyu, ttest_rel
 cols = ['area','between','within',
-				'correlation', 
-				'drange', 'xrange', 'yrange', 'xstd', 'ystd']
-
+                'correlation', 
+                'drange', 'xrange', 'yrange', 'xstd', 'ystd']
+conds = ['B0','B1','B2','M0','M1','M2']
 for i in cols:
-	gs = list(stats.groupby('condition')[i])
-	d = dict(gs)
-	ms = dict([(k, np.mean(v)) for k,v in d.items()])
-	p = mannwhitneyu(d['Middle'], d['Bottom']).pvalue
+    gs = list(stats.groupby('condition')[i])
+    gs = list(stats.groupby(['condition','gentype'])[i])
+    d = dict(gs)
+    ms = dict([(k, np.mean(v)) for k,v in d.items()])
+    #p = mannwhitneyu(d['Middle'], d['Bottom']).pvalue
 
-	S = i
-	for k,v in d.items():
-		S += '\t' + k + ' = ' + str(round(np.mean(v),3))
-	S += '\t' + 'p = ' + str(round(p,3))
-	print S
+    S = i
+    tempdict = {}
+    for k,v in d.items():
+        kstr = str(k[0][0]) + str(k[1])
+        tempdict[kstr] = round(np.mean(v),3)
+        #S += '\t' + kstr + ' = ' + str(round(np.mean(v),3))
+    #S += '\t' + 'p = ' + str(round(p,3))
+    for cond in conds:
+        S += '\t' + cond + ' = ' + str(tempdict[cond])
+    print S
 
 
 for j, rows in stats.groupby('condition'):
-	g1 = rows['within'].as_matrix()
[…]

### 2018-09-19 Experiments/middle-bottomABC/analysis/plot.individuals.py

@@ -38,5 +38,7 @@ for i, row in info.iterrows():
     funcs.plotclasses(ax, stimuli, palphas, pbetas, betastr=betastr,betacol = betacol)
     
     fname = os.path.join(savedir,condition + '-' + gentypeStr_p + '-' + str(pid) + '.png')
+    ax.text(.9,1.25,str(pid))
     f.savefig(fname, bbox_inches='tight', transparent=False)
     plt.cla()
+

### 2019-03-06 Manuscripts/cogsci-2019/paper2019.tex
Heaps of stuff: Two new experiments looking at Not Alphas and Gammas, Cogsci 2019 draft and AMPC 2019 abstract + slides, some model vs model correlation plots (still a work in progress.

% 
% Annual Cognitive Science Conference
% Sample LaTeX Paper -- Proceedings Format
%

% Original : Ashwin Ram (ashwin@cc.gatech.edu)       04/01/1994
% Modified : Johanna Moore (jmoore@cs.pitt.edu)      03/17/1995
% Modified : David Noelle (noelle@ucsd.edu)          03/15/1996
% Modified : Pat Langley (langley@cs.stanford.edu)   01/26/1997
% Latex2e corrections by Ramin Charles Nakisa        01/28/1997 
% Modified : Tina Eliassi-Rad (eliassi@cs.wisc.edu)  01/31/1998
% Modified : Trisha Yannuzzi (trisha@ircs.upenn.edu) 12/28/1999 (in process)
% Modified : Mary Ellen Foster (M.E.Foster@ed.ac.uk) 12/11/2000
% Modified : Ken Forbus                              01/23/2004
% Modified : Eli M. Silk (esilk@pitt.edu)            05/24/2005
% Modified : Niels Taatgen (taatgen@cmu.edu)         10/24/2006
% Modified : David Noelle (dnoelle@ucmerced.edu)     11/19/2014
% Modified : Roger Levy (rplevy@mit.edu)     12/31/2018

%% Change "letterpaper" in the following line to "a4paper" if you must.

\documentclass[10pt,letterpaper]{article}

\usepackage{cogsci}

%\cogscifinalcopy % Uncomment this line for the final submission

\usepackage{graphicx} %for figures
\usepackage{pslatex}
\usepackage{apacite}
\usepackage{float} % Roger Levy added this and changed figure/table
                   % placement to [H] for conformity to Word template,
                   % though floating tables and figures to top is
                   % still generally recommended!

%\usepackage[none]{hyphenat} % Sometimes it can be useful to turn off
%hyphenation for purposes such as spell checking of the resulting
%PDF.  Uncomment this block to turn off hyphenation.

%\setlength\titlebox{4.5cm}
% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 4.5cm (the original size).
%%If you do, we reserve the right to require you to change it back in
%%the camera-ready version, which could interfere with the timely
%%appearance of your paper in the Proceedings.

%\title{[working title] Calling it what it is: Novel categories are distinct from Not-categories}

\title{Novel categories are distinct from "Not"-categories}

\author{{\large \bf Shi Xian Liew (liew2@wisc.edu)} \\
  Department of Psychology, 1202 W. Johnson Street \\
  Madison, WI 53706 USA
  \AND {\large \bf Joseph L. Austerweil (austerweil@wisc.edu)} \\
  Department of Psychology, 1202 W. Johnson Street \\
  Madison, WI 53706 USA}

\begin{document}

\maketitle

\begin{abstract}

Category generation is an exciting but still poorly understood topic in cognitive science. One issue that is consistently being overlooked is the conflation of generating new, independently identified categories with the generation of categories that are the negation of a known category. This study represents the first attempt at delineating these two separate types of generated categories. We specifically test for differences in the distributional structure of generated categories, demonstrating that categories identified as \emph{not} what was known are larger and wider-spread compared to categories that were identified with a specific label. We also observe consistency in distributional structure across multiple generated categories, replicating and extending previous findings. These results are discussed in the context of providing a foundation for future modeling work.

\textbf{Keywords:} 
categorization; category generation; contrast; category learning; 
\end{abstract}

\section{Introduction}
% Within the broader field of categorization, the study of category generation has
% been gaining traction in recent decades. This specific topic is primarily
% concerned with
People are remarkable in their capacity to innovate new and different ideas. Is creating a new idea the same as creating a different idea? Consider a jazz musician who wants to generate music in a new genre. Is that the same as wanting to generate music that is {\em not} jazz? Both new categories of music are distinct from jazz. The former is identified as its own category,   the latter is identified in relation to a known category.

While categorization researchers have primarily focused their effort on classification (associating an exemplar with a category given its features), and inference (predicting exemplar features given its category),
work on category generation -- predicting all exemplar features for a novel
category -- is relatively scarce. Recent work has established a few key
findings. Earlier studies have shown that generated
categories tend to share distributional statistics with learned categories \cite{jern2013probabilistic,thomas98,ward1994structured}. More recently, 
\citeA{austerweil2018catgen} have found that category contrast is an important
factor in category generation and learning -- computational models sensitive to the
differences between categories were a better fit to generated categories and differences in learning different generated categories than models
which did not take categories contrast into account.

Although previous work has established a few key findings, the processes involved in category generation are still not well understood. In particular, it is currently unclear whether generating a
new category is itself distinct from identifying the set of exemplar features that are simply not what was previously identified in a known category.

 In order to better understand the nature of generated categories, it may be necessary to distinguish between these two possible types of generated categories: one driven by a motivation to be not what is known -- a \emph{category-by-negation}, and another driven by its own identity -- an \emph{independently identified category}.

Current models of category generation do not explicitly make this distinction. One of the first computational models of category generation, \citeA{jern2013probabilistic}'s hierarchical sampling model, is a Bayesian model that reproduces distributionally similar categories by assuming that the covariance matrix of features of exemplars from generated categories is generated by the same prior that generated the covariance matrix of the known categories. An alternative proposed model, PACKER \cite{austerweil2018catgen,conaway2017packer}, is an extension of the classic Generalized Context Model \cite<GCM;>{nosofsky1984choice,nosofsky1986attention}. It explicitly incorporates contrast into the similarity function by including a penalty for being similar to exemplars from a given category. A contrast parameter allows more different candidate categories to be weighted more heavily than similar candidate categories. Both models are flexible enough to describe how generated categories can be distributionally similar (or different) from an experimenter-defined category. However, they do not distinguish between  generating an independently identified category and a category-by-negation because there is no mechanism to account for these different identities.

This issue also extends to the methodology applied in studying category generation. In \citeA{ward1994structured}, participants were told to generate aliens that belonged to a different species from a prior group of aliens, without applying any specific label to this new group of aliens. Similarly, \citeA{jern2013probabilistic} instructed participants to generate a new, different type of crystal after having observed two different types of crystals. In contrast, \citeA{conaway2017packer} prompted participants to generate exemplars from a novel "Beta" category, while avoiding an explicit instruction for participants to create something "different". In each of these studies there appears to be implicit assumption that generating an independently identified category is equivalent to generating a category-by-negation.

In this paper, we challenge this null assumption by positing that the explicit association of categories-by-negation to its known counterpart (i.e., the explicit identification of the to-be-generated category as \emph{not} a known category) should result in the observer taking advantage of the entire area of the feature space not occupied by the known category. In contrast, observers generating independently identified categories should be less focused on the unoccupied feature space and instead construct their categories based on the structure of known categories. From this we can predict that categories-by-negation should occupy larger areas of the feature space compared to independently identified categories. In addition, the similarity in distributional statistics should extend not only from the learned category to the first independently identified category to be generated (as previous studies have found), but also between subsequent independently identified categories. We test these predictions by adapting and extending a category generation experiment by \citeA{conaway2017packer} and explore the implications of its results.

% In this paper, we investigate how the identity of the generated category affects its structure. Specifically, we predict that categories-by-negation should occupy larger areas of the feature space compared to independently identified categories. Independently identified categories are predicted to share greater distributional similarities with other independently identified categories. We test these predictions by adapting and extending a category generation experiment by \citeA{austerweil2018catgen}.

% [Started reading more about Packing Theory and I don't think we really have a good case about it. For one thing, I think it can be argued to support both the Not-Alpha and Beta type groups? I've moved Packing theory to the discussion instead.] One hypothesis that can be useful in guiding our intuitions about how generated categories can differ is the packing hypothesis \cite{hidaka2011packing}. Their Packing Theory suggests that categories are would inform us that the variation in the range of
% feature values of 'Not-Alpha' exemplars should be larger than than the variation
% in the range of feature values of 'Beta' and 'Gamma' exemplars.

% While \citeA{austerweil2018catgen} found that new categories tend to be different from known categories, this does not indicate an equivalence between a newly generated category and a category that is not what was known.

% There might be no difference in the resulting new composition if the musician identifies the new work with a specific name (independently identified category) or that the musician identifies the new work as simply not-jazz (category-by-negation). Both are meant to be distinct categories from jazz, but the former is identified as its own category and the latter is identified as a category relative to the previous category.

% If the category is identified by some arbitrary name, that should not offer any more or less information to the observer than stating that the category is not a known category;  consequently, the generated category should not be different in either case. [convoluted; revise if there's time] However, this argument assumes that a category-by-negation is labeled as arbitrarily as an independently identified category.

% Labelling the to-be-generated category as "" [Gotta work on the names here.] For instance, suppose a musician has released an album of jazz music. For her next album, she does not know what particular genre she wants to release -- she simply knows that it is not jazz. The question we are asking is, is there a difference between this and a genre called "goose" (for example)?
% The current state of affairs is that studies and models don't distinguish between the two.

% The generation of a new category involves at least the concept that the category is novel in some sense, and that it is a category. Typically, in laboratory tasks the novelty of the category is communicated to the observer by indicating that this to-be-generated category is \emph{different to} the prior observed category.

% Suppose a band of musicians has released an album of jazz music. For their next composition, the band leader suggests that they do something different to jazz. The double-bassist agrees, but adds that while they should make a new type of music, they should also give this new genre a new name. The band leader is confused -- whether or not their new music is given a name is irrelevant so long as it is different to jazz. The double-bassist might argue that the name gives the band 'focus'. When creating a new category of objects or exemplars, is there a difference in the resulting category

% When motivated to produce something novel, do For
% instance, if a musician is creating a new album, is there a difference in the
% content of the new album when they are trying to compose something that is not
% one of their previous albums, compared to creating something that is a new genre?
% [needs work ugh]

\section{Experiment}
%JLA: If possible it would be good to cite CA2017 as much as we can rather than the manuscript. It's supposed to be double blind and citing a particular manuscript in preparation all the time may cue reviewers to the identity of the authors...
Our current experiment closely mirrors the the experimental design of \citeA{austerweil2018catgen} and \citeA{conaway2017packer}, where participants are first trained on a category named 'Alpha' before being tasked to generate exemplars from a new category. While the original authors were primarily interested in the measuring the location of generated categories relative to learned categories, our current investigation focuses on analyzing the differences in distributional statistics across different generated categories. Consequently, in addition to varying the shape and location of the Alpha category (i.e., across different Alpha conditions where the position of the Alpha exemplars are systematically varied), we include an additional independent variable comprising three different generation conditions: a Not-Alpha condition, where participants generate a new category that is not the learned Alpha category; a Beta-Only condition, where participants generate a new category named 'Beta'; and a Beta-and-Gamma condition, where participants generate a category 'Beta' as well as a category 'Gamma'. The resulting 3X3 design is applied in a between-subjects fashion -- participants can be in only one of the nine unique conditions.

The main advantage of adapting the experiments by \citeA{austerweil2018catgen} \citeA{conaway2017packer} is that the simplicity of their stimuli allow for a straightforward test of the distributional similarities across known and generated categories. In addition, the variety of Alpha categories used (i.e., the different shapes of the Alpha categories) also allows us to observe the effect of generated category identity across multiple scenarios.

In line with our earlier predictions, we expect that the Not-Alpha conditions on average generate categories that are larger in area and more widely dispersed than categories from the Beta-Only as well as Beta-and-Gamma conditions. In addition, we predict that within the Beta-and-Gamma conditions, the generated Beta and generated Gamma categories should be distributionally similar.

% No interaction effects between the factors (i.e., Alpha $\times$ Generation conditions) are expected.

\subsection{Method}

\subsubsection{Participants and materials}

We recruited 240 participants through Amazon Mechanical Turk and randomly
assigned them to one of the nine unique conditions. Sample sizes of each
condition are presented in Table \ref{table:samplesize}.

Stimuli were squares that varied along two dimensions: color (grayscale
9.8\% -- 90.2\%) and size (3.0 -- 5.8cm on each side). The assignment of perceptual
features (color, size) to axes of the domain space (x, y), as well as the
direction of variation along each axis (e.g., increasing or decreasing size) was
counterbalanced across participants. Feature values were evenly-spaced on a
50X50 grid, giving a near-continuous space from which exemplars can be generated. An example of the feature space is presented in Figure \ref{fig:alphas}a.

Similar to \citeA{austerweil2018catgen}, our three Alpha conditions can be described as Cluster, Row, and Diagonal. In the Cluster condition, Alpha exemplars occupy a small area towards one corner of the feature space. In the Row condition, the exemplars are nearly equal along one feature, while equally spread out along the other feature. The Diagonal condition is an adaptation of the XOR condition in \citeA{austerweil2018catgen}, except instead of exemplars situated at the corners, they are equally spaced along the diagonal of the feature space in a similar fashion to the diagonal conditions of \citeA{jern2013probabilistic}. In addition, exemplar feature values are slightly jittered from their original values in \citeA{austerweil2018catgen} in order to ensure that exemplars are not completely identical along a single feature. The exact same amount of jitter is applied to all Alpha exemplars within a given Alpha condition. The locations of these different Alpha categories in the feature space are presented in Figures \ref{fig:alphas}b to \ref{fig:alphas}d.

\begin{figure*}%[H]
\begin{center}
\includegraphics[width=\textwidth]{Figures/conditions.png}
\end{center}
\caption{(a) Example of stimuli located at the corners of the feature space. (b-d) Locations of Alpha category exemplars for each Alpha condition.} 
\label{fig:alphas}
\end{figure*}

\subsubsection{Procedure}
In the first phase of the experiment, participants learned the Alpha category
exemplars by observing a unique exemplar on each trial. This
was repeated over a total of three blocks (four trials per block -- one
corresponding to each unique exemplar), with the order of exemplar presentation
randomized within each block. Participants were allowed to spend as much time as
they wanted on each trial and were also shown the full range of possible feature
values prior to training.

The next phase comprised a series of generation trials. Depending on their generation condition, participants were asked to generate either eight exemplars from a category that was 'Not-Alpha' (Not-Alpha generation condition), eight exemplars from a category called 'Beta' (Beta-Only generation condition), or four exemplars from a category 'Beta' \emph{and}
four exemplars from a third category 'Gamma' (Beta-and-Gamma generation condition). Exemplars were generated once per trial using two on-screen sliding scales, with each scale controlling the individual features (color and size) of the generated exemplar. Feature values could take any one of 50 evenly-spaced values between the specified boundaries. Previously generated exemplars were not allowed to be generated a second time. Participants were shown an on-screen preview of their exemplar on each trial as they interacted with the sliders, but could not see previously generated exemplars or exemplars from the Alpha category.

\begin{table}[H]
\begin{center} 
\caption{Sample sizes for each condition.} 
\label{table:samplesize} 
\vskip 0.12in
\begin{tabular}{llll} 
\hline
Generation Condition    &  Cluster & Row & Diagonal \\
\hline
Not-Alpha      &   26 & 28 & 25 \\
Only Beta      &   30 & 27 & 25 \\
Beta-and-Gamma &   26 & 27 & 26 \\
\hline
\end{tabular} 
\end{center} 
\end{table}

\subsubsection{Analyses}

We analyze our data in two stages. First, to provide a coarse overview of the distribution of different patterns of generated categories, we classify the generated categories into
six different profiles: Positives, where the correlation between the dimensions
is more than $r$; Negatives, where the correlation between the dimensions is less
than $-r$; Rows, where the range of values across the x dimension is at least
five times more than the range across the y dimension; Columns, where the range
of y dimension values is at least five times more than the range of x dimension
values; Clusters, where the ranges across both dimensions are less than $a$; and
Dispersed, where the ranges across both dimensions are more than $a$. Next, we compare the generated categories on four key distributional measures: ranges of each feature, the feature correlations, and the area enclosed by the generated exemplars in the feature space (i.e., their convex hull). Differences along each of these statistics are performed using Bayesian $t$-tests \cite{rouder2009bayesian}, yielding Bayes factors ($BF_{01}$) which indicate evidence for the null hypothesis when $BF_{01} > 1$, with larger values indicating greater evidence for the null hypothesis. $BF_{01} < 1$ indicates evidence for the alternative hypothesis. Interpretations of the sizes of Bayes factors are guided by \citeA{jeffreys1961}.

\subsection{Results}

We found that setting $r = .7$ and $a = .25$ was useful in capturing the different profiles of generated categories. A representative sample of each profile is shown in Figure
\ref{fig:samples} and the frequency plot of the different profiles is presented
in Figure \ref{fig:profile}.

The most striking patterns to note here are the high frequencies of Row category profiles from participants in the Row conditions, and the high frequencies of Dispersed category profiles from participants in the Not-Alpha conditions. The former indicates that the distributional similarities between learned and generated categories are especially strong in the Row conditions, while the latter provides preliminary evidence that generated categories from the Not-Alpha conditions tends to be more widely dispersed. Also noteworthy are the low counts of both Positive and Negative category profiles across the whole data set -- in contrast to the Row conditions, this indicates low similarity in distributional structure between the learned and generated categories for the Diagonal condition.

%Note that the trained category exemplars (labeled as A in the figure)

\begin{figure}[H]
\begin{center}
    \includegraphics[width=.5\textwidth]{Figures/samples.pdf}
\end{center}
\caption{Representative samples of the six different category generation profiles.} 
\label{fig:samples}
\end{figure}

\begin{figure*}%[H]
\begin{center}
    \includegraphics[width=\textwidth]{Figures/profile_freq.png}
\end{center}
\caption{Frequencies of category generation profiles broken down by Alpha condition (left plot) and generation condition (right plot).}
\label{fig:profile}
\end{figure*}

Focusing on the distributional statistics, when broken down by the Alpha conditions (Figure \ref{fig:stats_cond}), we found that the Cluster conditions tended to have categories with a smaller range of both features, and with correspondingly small sizes. The Row conditions produced categories that are high on the x dimension but not the y dimension. These observations indicate, as was observed in \citeA{austerweil2018catgen}, that the distribution statistics were carried over from the known category to the generated category in these two Alpha conditions. However, the Diagonal conditions did not have similar distributional statistics observed -- instead of a positive correlation, the Diagonal conditions tended to produce large, dispersed categories. \citeA{austerweil2018catgen} also observed a similar effect (although they found more evidence for a negative correlation). Overall, in alignment with previous research, at least two out of the three Alpha conditions in our experiment generated categories that share distributional statistics as the known category.

More interestingly, when the data is broken down by the generation conditions (Figure \ref{fig:stats_instr}), we found that compared to the Not-Alpha conditions, there is moderate evidence showing Beta-Only conditions with a lower y dimension range ($t(162) = 3.00, BF_{01} = 0.16$) and moderate to strong evidence that their generated categories are smaller in area ($t(162) = 3.16, BF_{01} = 0.10$). There is moderate evidence that the Not-Alpha and Beta conditions share equal range of x dimension values ($t(162) = 1.33, BF_{01} = 4.82$). With Beta-and-Gamma conditions, we find greater evidence for smaller and tighter categories compared to the Not-Alpha conditions. Specifically, there is very strong evidence that both Beta and Gamma categories from the Beta-and-Gamma condition are smaller in both x ($t(156) = 4.83, BF_{01} = 2.55 \times 10^{-4}$;$t(156) = 4.21, BF_{01} = 2.94 \times 10^{-3}$, respectively) and y ($t(156) = 4.57, BF_{01} = 7.40 \times 10^{-4}$;$t(156) = 7.56, BF_{01} = 4.30 \times 10^{-10}$, respectively) ranges compared to the Not-Alpha conditions. Similarly, there is very strong evidence that both categories in the Beta-and-Gamma conditions are smaller in area than the Not-Alpha conditions ($t(156) = 5.70, BF_{01} = 5.41 \times 10^{-6}$;$t(156) = 7.69, BF_{01} = 2.07 \times 10^{-10}$, respectively). Overall, when comparing the Not-Alpha conditions to categories from other generation conditions, we consistently find moderate to very strong evidence that Not-Alpha categories are more widely dispersed (in their range values) and also larger (in their area), with the only exception being the comparison of x dimension ranges between the Not-Alpha and Beta-Only conditions.

When comparing the distributions of the Beta and Gamma categories generated within the Beta-and-Gamma conditions, we find an overall weak-to-moderate evidence for equal distributional statistics. Specifically, there was moderate evidence for equal x dimension ranges ($t(150) = 0.56, BF_{01} = 9.47$) and weak evidence for both feature correlations and area size ($t(150) = 1.85, BF_{01} = 2.10$;$t(150) = 1.77, BF_{01} = 2.40$, respectively). When measured on their y dimension ranges, we found weak evidence for lower values from Gamma categories compared to the Beta categories ($t(150) = 2.47, BF_{01} = 0.59$).

% \begin{table}[H]
% \begin{center} 
% \caption{Results of Bayesian $t$-test between the generation conditions: pairwise comparisons of Not-Alpha (A') and Beta-Only (B) conditions, as well as the Beta (BC-B) and Gamma (BC-C) categories separately from the Beta-and-Gamma condition.} 
% \label{table:samplesize} 
% \vskip 0.12in
% \begin{tabular}{llcc} 
% \hline
% Group 1 & Group 2  & $t$(df) & $BF_{01}$ \\
% \hline
% A'   & B    &  3.16(162) & $9.81\times10^{-2}$ \\
% A'   & BC-B &  5.70(156) & $5.41\times10^{-6}$ \\
% A'   & BC-C &  7.69(156) & $2.07\times10^{-10}$ \\
% B    & BC-B &  2.48(156) & $0.58$ \\
% B    & BC-C &  4.27(156) & $2.35\times10^{-3}$ \\
% BC-B & BC-C &  1.77(150) & $2.40$      \\
% \hline
% \end{tabular} 
% \end{center} 
% \end{table}

\begin{figure*}%[H]
\begin{center}
    \includegraphics[width=\textwidth]{Figures/stats_cond.pdf}
\end{center}
\caption{Box-plots of the distributional statistics from the generated categories. Boxes depict the median and quartiles of each Alpha condition, with whiskers placed at 1.5 inter-quartile range. All points outside this region are marked individually.} 
\label{fig:stats_cond}
\end{figure*}

\begin{figure*}%[H]
\begin{center}
    \includegraphics[width=\textwidth]{Figures/stats_instr.pdf}
\end{center}
\caption{Box-plots of the distributional statistics from the generated categories. Boxes depict the median and quartiles of each generation condition, with the Not-A condition denoted by A' and the Beta-Only condition denoted by B. Data from the Beta-and-Gamma conditions are separated into the Beta groups (denoted BC-B) and the Gamma groups (denoted BC-C) generated in that condition. Whiskers are placed at 1.5 inter-quartile range. All points outside this region are marked individually.} 
\label{fig:stats_instr}
\end{figure*}

\section{Discussion}
At first glance, it seems reasonable to assume that generating a new category X after learning category Y is the same as generating category Not-X. An independently identified category should already be a category-by-negation (in that an independently identified category is not what was known). Further, if the categories are already identified by arbitrary labels, then it may be easy to assume that identifying the negation of a known category cannot add any additional information in category generation.

Our results have indicated otherwise. Specifically, when tasked to produce categories-by-negation, participants tended to generate wider and larger new categories compared to when tasked with producing independently identified categories. To our knowledge, this paper represents the first piece of evidence distinguishing these separate types of categories.

Aside from demonstrating a new effect, this study has continued to show the robustness of the distributional similarities between learned and generated categories. In this sense, the results from the different Alpha conditions are similar to those observed in \citeA{austerweil2018catgen}. The generated categories from the Cluster condition tend to possess lower x and y ranges, with a correspondingly smaller area, and the generated categories from the Row condition tend to adopt Row-type profiles. There is also similar lack of categories with positively correlated features from the Diagonal condition.

However, one notable difference is that while \citeA{austerweil2018catgen} found evidence of negatively-correlated generated categories in their XOR condition, we found no evidence of negatively-correlated generated categories in our comparable Diagonal condition. \citeA{austerweil2018catgen} explained that the presence of negatively-correlated features in this condition was indicated the effect of category contrast -- that is, categories with negatively correlated features are generated because they are particularly different to categories with positively correlated features. It is possible that due to the reduced strength of the positive correlation in our study compared to \citeA{austerweil2018catgen} (because of the addition of noise to the feature values), participants were no longer as sensitive to the negative correlations in the experimenter-defined category and therefore started to produce uncorrelated but widely dispersed categories.

Beyond replicating previous studies, the current study has demonstrated that the consistency in distributional statistics can persist beyond the first generated category. However, evidence showing this was ultimately weak. One possible reason for this is the relatively small feature space employed in the tasks. The generation of a second novel category is necessarily more constrained in the feature space than the generation of the first novel category, possibly resulting in differences in distributional structure. By exploring stimuli with less defined boundaries (e.g., orientation), we may expect to see greater consistency in distributional structure over multiple generated categories.

Although we have observed participants generating multiple independently identified categories, we do not want to imply that categories-by-negation can only happen once. It would be worth investigating how participants might proceed to generate additional categories-by-negation (e.g., by asking observers to generate a category that is Not-Alpha and Not Beta). Packing Theory \cite{hidaka2011packing} -- a hypothesis that suggests categories can be neatly 'packed' into the feature space -- may indicate that successive categories-by-negation are generated in a fashion that preferentially occupies spaces between observed categories. Further, although none of the current models of category generation can directly account for the effects observed in this study, they may be useful components in a larger category generation framework. For instance, future work may consider implementing the hierarchical sampling model from \citeA{jern2013probabilistic} in a framework of overhypotheses \cite{kemp07}, where a prior can be placed over a category identity space, allowing models to behave differently under different regions of generated category identity.

Ultimately, the nature of newly generated categories appears to vary depending on the identity they were associated with. The extent to which they may differ, and the mechanisms driving these differences represent fascinating areas for future research.

% \section{Acknowledgments}

% In the \textbf{initial submission}, please \textbf{do not include
%   acknowledgements}, to preserve anonymity.  In the \textbf{final submission},
% place acknowledgments (including funding information) in a section \textbf{at
% the end of the paper}.

\bibliographystyle{apacite}

\setlength{\bibleftmargin}{.125in}
\setlength{\bibindent}{-\bibleftmargin}

\bibliography{citations.bib}

\end{document}

### 2019-03-06 Manuscripts/model-descriptions/copy-and-tweak/cp-rep.tex
Heaps of stuff: Two new experiments looking at Not Alphas and Gammas, Cogsci 2019 draft and AMPC 2019 abstract + slides, some model vs model correlation plots (still a work in progress.

--- 2017-03-15 Manuscripts/model-descriptions/copy-and-tweak/copy-and-tweak.tex
@@
 
-\documentclass[12pt]{article}
-\usepackage[letterpaper, margin=1in, headheight=15pt]{geometry}
-\usepackage{setspace}
-\usepackage{amsmath}
-\usepackage{csquotes}
-\usepackage{fancyvrb}
-\usepackage[natbibapa]{apacite}
+\documentclass[12pt]{article} \usepackage[letterpaper, margin=1in,
+headheight=15pt]{geometry} \usepackage{setspace} \usepackage{amsmath}
+\usepackage{csquotes} \usepackage{fancyvrb} \usepackage[natbibapa]{apacite}
 
@@
 
-\section{The Copy-And-Tweak Model}
+\section{The Copy-And-Tweak Model with Representativeness}
 
-This model, initially implemented by \cite{jern2013probabilistic} to explain exemplar generation in a multinomial space, is broadly based on Ward's path-of-least-resistance model \citep{ward1994structured,ward1995s,ward2002role}, but is quite different in implementation.
-This document describes how to implement their copy-and-tweak approach within our 2D continuous domain.
-See Appendix \ref{ap:jk13ap} for description of the model provided by \cite{jern2013probabilistic}.
+This model is in essence the copy-and-tweak model as discussed in
+\cite{jern2013probabilistic}, but with the categorization probabilities (for
+category generation and presumably extendable to category learning) generated by
+the representativeness heuristic as formally defined by \cite{tenenbaum01}.
+Specifically, the representativeness of an exemplar $x$ for a given hypothesis
+(e.g., a category) is the relative evidence that is provided by the exemplar for
+a given hypothesis $h$ compared to all other hypotheses $h'$:
 
-The model proposes that, when prompted to generate new examples, people copy a known example from the target class and change it a small amount along each dimension.
-The core difference between this account and the hierarchical sampling account is that correlations between features are not maintained by a copy-and-tweak system.
-As a result, this issue was the focus of \cite{jern2013probabilistic}.
-
-In order to generate a new exemplar, the model computes the similarity between the members of the target category $x$and possible generation option $y$.
-Similarity $s$ between two examples, $x_i$ and $y_i$, is an inverse exponential function of distance:
-
+[Note to Xian from Xian: This is still a work in progress - pls remember to come
+back to it.]
 \begin{equation}
-  s(x_i,y_i) = exp( -c \sum_k{|x_{ik} - y_{ik}|w_k})
+  R(x,h) = \log \dfrac{p(x|h)}{\sum_{h'\neq h}{p(x|h')p(h')}}
+\label{representativeness}
 \end{equation}
 
-Where $c$ and $w_k$ are the standard specificity and attention weights discussed elsewhere \citep[see][]{nosofsky1984choice}.
+This model,
+initially implemented by \cite{jern2013probabilistic} to explain exemplar
+generation in a multinomial space, is broadly based on Ward's
+path-of-least-resistance model
+\citep{ward1994structured,ward1995s,ward2002role}, but is quite different in
+implementation. This document describes how to implement their copy-and-tweak
+approach within our 2D continuous domain. See Appendix \ref{ap:jk13ap} for
+description of the model provided by \cite{jern2013probabilistic}.
 
-\begin{equation}
-  s(x_i,y_i) = \sum_i{s(y, x_i)}
+The model proposes that, when prompted to generate new examples, people copy a
+known example from the target class and change it a small amount along each
+dimension. The core difference between this account and the hierarchical
+sampling account is that correlations between features are not maintained by a
+copy-and-tweak system. As a result, this issue was the focus of
+\cite{jern2013probabilistic}.
+
+In order to generate a new exemplar, the model computes the similarity between
+the members of the target category $x$and possible generation option $y$.
+Similarity $s$ between two examples, $x_i$ and $y_i$, is an inverse exponential
+function of distance:
+
+\begin{equation} s(x_i,y_i) = exp( -c \sum_k{|x_{ik} - y_{ik}|w_k})
 \end{equation}
 
-The probability that a given candidate $y$ will be generated is based on its summed similarity to all members of the target category $x$: $\sum_i{s(y, x_i)}$.
-\begin{equation}
-    p(y)  = \dfrac
-    { \exp \left\{\theta \cdot \sum_i{s(y, x_i)} \right\} }
-    {\sum_j{\exp \left\{ \theta \cdot \sum_i{s(y_j, x_i)} \right\} }} 
+Where $c$ and $w_k$ are the standard specificity and attention weights discussed
+elsewhere \citep[see][]{nosofsky1984choice}.
+
+\begin{equation} s(x_i,y_i) = \sum_i{s(y, x_i)}
+\end{equation}
+
+The probability that a given candidate $y$ will be generated is based on its
+summed similarity to all members of the target category $x$: $\sum_i{s(y,
+x_i)}$.
+\begin{equation} p(y) = \dfrac { \exp \left\{\theta \cdot \sum_i{s(y, x_i)}
+\right\} } {\sum_j{\exp \left\{ \theta \cdot \sum_i{s(y_j, x_i)} \right\} }}
     \label{eq:generation-choice}
@@
 % 
-where $\theta$ is a response determinism parameter.
-Known members of the target category cannot be generated, and so their probability is manually set to 0.
-If there are no known members of the target category, generation is random (uniform probabilities over $y$).
+where $\theta$ is a response determinism parameter. Known members of the target
+category cannot be generated, and so their probability is manually set to 0. If
+there are no known members of the target category, generation is random (uniform
+probabilities over $y$).
 
 \subsection{Parameter info}
-\begin{itemize}
-    \setlength\itemsep{-0.5em}
+\begin{itemize} \setlength\itemsep{-0.5em}
     \item $c$. Scalar, $c>0$. Specificity parameter.
@@
 % references section
-\clearpage
-\bibliographystyle{apacite}
-\bibliography{citations.bib}
+\clearpage \bibliographystyle{apacite} \bibliography{citations.bib}
 
@@
 \subsubsection{Page 93}
-\begin{displayquote}
-Exemplar models like the GCM (Nosofsky, 1986) and ALCOVE (Kruschke, 1992) classify novel exemplars by computing their similarity to previously observed exemplars stored in memory.
-These models can be extended to account for generation by assuming that stored exemplars can be copied and changed in some way in order to produce novel exemplars.
-We will refer to the resulting approach as the copy-and-tweak account, and the same basic approach has been proposed by several previous authors (Barsalou \& Prinz, 1997; Barsalou, 1999; Ward, 1995; Ward, Patterson, Sifonis, Dodds, \& Saunders, 2002; Ward, Patterson, \& Sifonis, 2004).
+\begin{displayquote} Exemplar models like the GCM (Nosofsky, 1986) and ALCOVE
+(Kruschke, 1992) classify novel exemplars by computing their similarity to
+previously observed exemplars stored in memory. These models can be extended to
+account for generation by assuming that stored exemplars can be copied and
+changed in some way in order to produce novel exemplars. We will refer to the
+resulting approach as the copy-and-tweak account, and the same basic approach
+has been proposed by several previous authors (Barsalou \& Prinz, 1997;
+Barsalou, 1999; Ward, 1995; Ward, Patterson, Sifonis, Dodds, \& Saunders, 2002;
+Ward, Patterson, \& Sifonis, 2004).
 
-Although the copy-and-tweak account may seem conceptually different from the sampling account, in some cases the two approaches are equivalent.
-Consider, for example, a problem in which the exemplars are points in n-dimensional space.
-A copy-and-tweak account would retrieve one of these exemplars and change it slightly along each dimension.
-This procedure is conceptually identical to sampling from a sum of normal distributions centered on each observed exemplar, which is formally equivalent to the GCM (Ashby \& Alfonso-Reese, 1995; Jäkel, Schölkopf, \& Wichmann, 2008).
-Thus, in many cases, applying copy-and-tweak to an exemplar model will be equivalent to sampling from the generation distribution learned by a generative model.
-The relationship between copy-andtweak and the sampling account suggests that both accounts will make similar predictions for problems in which people generate from categories that correspond to clusters in a similarity space.
-The two accounts, however, will diverge in cases where category membership cannot naturally be captured by similarity.
-Our first set of experiments uses structured categories that are specifically designed to distinguish between the two accounts.
+Although the copy-and-tweak account may seem conceptually different from the
+sampling account, in some cases the two approaches are equivalent. Consider, for
+example, a problem in which the exemplars are points in n-dimensional space. A
+copy-and-tweak account would retrieve one of these exemplars and change it
+slightly along each dimension. This procedure is conceptually identical to
+sampling from a sum of normal distributions centered on each observed exemplar,
+which is formally equivalent to the GCM (Ashby \& Alfonso-Reese, 1995; Jäkel,
+Schölkopf, \& Wichmann, 2008). Thus, in many cases, applying copy-and-tweak to
+an exemplar model will be equivalent to sampling from the generation
+distribution learned by a generative model. The relationship between
+copy-andtweak and the sampling account suggests that both accounts will make
+similar predictions for problems in which people generate from categories that
+correspond to clusters in a similarity space. The two accounts, however, will
+diverge in cases where category membership cannot naturally be captured by
+similarity. Our first set of experiments uses structured categories that are
+specifically designed to distinguish between the two accounts.
 \end{displayquote}
@@
 \subsubsection{Pages 103-104}
-\begin{displayquote}
-As described earlier, the copy-and-tweak approach converts an exemplar model of classification into a model of generation.
-The approach proposes that new exemplars are generated by copying one of the exemplars stored in memory and then modifying some features of the copy.
-When we introduced the category used for Experiments 1 and 2, we noted that our stimuli were carefully chosen to distinguish between the structured sampling account and the copy-and-tweak account.
-The structured sampling model is capable of learning that pairs of letters are grouped into cohesive parts and therefore predicts that people will tend to generate novel combinations of parts.
-In contrast, a copy-and-tweak model will generate many exemplars that are not valid exemplars.
-For example, after observing the four exemplars in Fig.
-4b, a copy-and-tweak model might produce a new exemplar by copying the first exemplar and changing the top letter from D to X.
+\begin{displayquote} As described earlier, the copy-and-tweak approach converts
+an exemplar model of classification into a model of generation. The approach
+proposes that new exemplars are generated by copying one of the exemplars stored
+in memory and then modifying some features of the copy. When we introduced the
+category used for Experiments 1 and 2, we noted that our stimuli were carefully
+chosen to distinguish between the structured sampling account and the
+copy-and-tweak account. The structured sampling model is capable of learning
+that pairs of letters are grouped into cohesive parts and therefore predicts
+that people will tend to generate novel combinations of parts. In contrast, a
+copy-and-tweak model will generate many exemplars that are not valid exemplars.
+For example, after observing the four exemplars in Fig. 4b, a copy-and-tweak
+model might produce a new exemplar by copying the first exemplar and changing
+the top letter from D to X.
 
 We implemented a copy-and-tweak model based largely on the GCM (Nosofsky, 1986).
-The GCM makes classification judgments by computing the similarity between a novel exemplar and all of the stored exemplars, and similarity is computed as a function of pairwise distance between exemplars.
-Because our stimuli used discrete non-ordinal features, we used the Hamming distance, which measures the number of features that are different.
-As explained earlier, the GCM is equivalent to a generative model that maintains a probability distribution with peaks at the observed exemplars (Ashby \& Alfonso-Reese, 1995; Jakel et al., 2008).
-The generative model that corresponds to the GCM maintains a probability distribution over a continuous similarity space, and we implemented an analogous approach that maintains a probability distribution over our space of discrete flu genomes.
+The GCM makes classification judgments by computing the similarity between a
+novel exemplar and all of the stored exemplars, and similarity is computed as a
+function of pairwise distance between exemplars. Because our stimuli used
+discrete non-ordinal features, we used the Hamming distance, which measures the
+number of features that are different. As explained earlier, the GCM is
+equivalent to a generative model that maintains a probability distribution with
+peaks at the observed exemplars (Ashby \& Alfonso-Reese, 1995; Jakel et al.,
+2008). The generative model that corresponds to the GCM maintains a probability
+distribution over a continuous similarity space, and we implemented an analogous
+approach that maintains a probability distribution over our space of discrete
+flu genomes.
 \end{displayquote}
@@
 
-The copy-and-tweak model uses similarity to observed exemplars as the basis for both classification and generation.
-We used a similarity rule derived from the GCM (Nosofsky, 1986) that defines the similarity $s$ between exemplars $x_1$ and $x_2$ as $s(x_1,x_2) = \exp (-c \cdot d(x_1,x_2))$, where $c$ is a scaling parameter and $d(x_1,x_2)$ is the distance between the two exemplars.
-We set $c$ to 1 and used the Hamming distance, which counts the number of features for which the two exemplars have different values.
+The copy-and-tweak model uses similarity to observed exemplars as the basis for
+both classification and generation. We used a similarity rule derived from the
+GCM (Nosofsky, 1986) that defines the similarity $s$ between exemplars $x_1$ and
+$x_2$ as $s(x_1,x_2) = \exp (-c \cdot d(x_1,x_2))$, where $c$ is a scaling
+parameter and $d(x_1,x_2)$ is the distance between the two exemplars. We set $c$
+to 1 and used the Hamming distance, which counts the number of features for
+which the two exemplars have different values.
 
-To derive classification predictions from the copy-and-tweak model for the rating task, we summed the similarities of each rating exemplar $x_r$ to each of the training exemplars $x_i$: $\text{rating}(x_r) =\sum_i{s(x_r,x_i)}$.
-We then applied a z-score transformation to produce the predictions in Fig.
-10.
+To derive classification predictions from the copy-and-tweak model for the
+rating task, we summed the similarities of each rating exemplar $x_r$ to each of
+the training exemplars $x_i$: $\text{rating}(x_r) =\sum_i{s(x_r,x_i)}$. We then
+applied a z-score transformation to produce the predictions in Fig. 10.
 
-To derive generation predictions from the copy-and-tweak model, we computed similarity ratings for all possible exemplars and drew 2,000,000 samples from a Multinomial distribution based on the ratings:
+To derive generation predictions from the copy-and-tweak model, we computed
+similarity ratings for all possible exemplars and drew 2,000,000 samples from a
+Multinomial distribution based on the ratings:
 
-\begin{equation}
-  \widetilde{x} \sim \text{Multinomial}([R \cdot \text{rating}(\widetilde{x}_1) ... \text{rating}(\widetilde{x}_n)])
+\begin{equation} \widetilde{x} \sim \text{Multinomial}([R \cdot
+\text{rating}(\widetilde{x}_1) ... \text{rating}(\widetilde{x}_n)])
 \end{equation}
 % 
-where $R = 1/\sum_i{\text{rating}(\widetilde{x}_i)}$ is a normalization constant.
-The reported predictions are based on a model which considers all possible exemplars that can be constructed from the 16 observed letters ($n = 16^4$ ).
-We also implemented an alternative model that considers only exemplars in which each letter can only appear in positions where it has been previously observed (n = 256).
-The two models produced similar predictions, but the former version performed slightly better.
+where $R = 1/\sum_i{\text{rating}(\widetilde{x}_i)}$ is a normalization
+constant. The reported predictions are based on a model which considers all
+possible exemplars that can be constructed from the 16 observed letters ($n =
+16^4$ ). We also implemented an alternative model that considers only exemplars
+in which each letter can only appear in positions where it has been previously
+observed (n = 256). The two models produced similar predictions, but the former
+version performed slightly better.
 \end{displayquote}

### 2019-03-06 Presentations/ampc_2019/abstract.tex
Heaps of stuff: Two new experiments looking at Not Alphas and Gammas, Cogsci 2019 draft and AMPC 2019 abstract + slides, some model vs model correlation plots (still a work in progress.

--- 2019-03-06 Manuscripts/cogsci-2019/paper2019.tex
@@
 
-\title{Novel categories are distinct from "Not"-categories}
+\title{}
 
@@
 
-Category generation is an exciting but still poorly understood topic in cognitive science.
-One issue that is consistently being overlooked is the conflation of generating new, independently identified categories with the generation of categories that are the negation of a known category.
-This study represents the first attempt at delineating these two separate types of generated categories.
-We specifically test for differences in the distributional structure of generated categories, demonstrating that categories identified as \emph{not} what was known are larger and wider-spread compared to categories that were identified with a specific label.
-We also observe consistency in distributional structure across multiple generated categories, replicating and extending previous findings.
-These results are discussed in the context of providing a foundation for future modeling work.
+The topic of category generation is a growing
+However, there is the general assumption that generating a new category B is the same as generating a new category
+Not-A. generating a new category is equivalent to generating a category that is
+not
+
+Understanding how new categories are generated and the factors influencing their structure is a growing area of interest
+in cognitive science. One issue that has been overlooked, however, is the assumption that generating a distinct novel
+category is equivalent to generating a category that is \emph{not} a known category. In this study, we provide a series
+of preliminary work challenging this assumption by distinguishing two types of generated categories. Extending category
+generation tasks from past research, we find that generated categories identified as \emph{not} what was known
+consistently tend to be larger compared to generated categories that were identified with a specific label. We also
+replicate a known effect demonstrating that generated categories share similar distributional properties as known
+categories. Our findings provide the first steps in building a formal computational account of category generation.
 
@@
 \end{abstract}
-
-\section{Introduction}
-% Within the broader field of categorization, the study of category generation has
-% been gaining traction in recent decades. This specific topic is primarily
-% concerned with
-People are remarkable in their capacity to innovate new and different ideas.
-Is creating a new idea the same as creating a different idea?
-Consider a jazz musician who wants to generate music in a new genre.
-Is that the same as wanting to generate music that is {\em not} jazz?
-Both new categories of music are distinct from jazz.
-The former is identified as its own category,   the latter is identified in relation to a known category.
-
-While categorization researchers have primarily focused their effort on classification (associating an exemplar with a category given its features), and inference (predicting exemplar features given its category),
-work on category generation -- predicting all exemplar features for a novel
-category -- is relatively scarce. Recent work has established a few key
-findings. Earlier studies have shown that generated
-categories tend to share distributional statistics with learned categories \cite{jern2013probabilistic,thomas98,ward1994structured}. More recently, 
-\citeA{austerweil2018catgen} have found that category contrast is an important
-factor in category generation and learning -- computational models sensitive to the
-differences between categories were a better fit to generated categories and differences in learning different generated categories than models
-which did not take categories contrast into account.
-
-Although previous work has established a few key findings, the processes involved in category generation are still not well understood. In particular, it is currently unclear whether generating a
-new category is itself distinct from identifying the set of exemplar features that are simply not what was previously identified in a known category.
-
- In order to better understand the nature of generated categories, it may be necessary to distinguish between these two possible types of generated categories: one driven by a motivation to be not what is known -- a \emph{category-by-negation}, and another driven by its own identity -- an \emph{independently identified category}.
-
-Current models of category generation do not explicitly make this distinction.
-One of the first computational models of category generation, \citeA{jern2013probabilistic}'s hierarchical sampling model, is a Bayesian model that reproduces distributionally similar categories by assuming that the covariance matrix of features of exemplars from generated categories is generated by the same prior that generated the covariance matrix of the known categories.
-An alternative proposed model, PACKER \cite{austerweil2018catgen,conaway2017packer}, is an extension of the classic Generalized Context Model \cite<GCM;>{nosofsky1984choice,nosofsky1986attention}.
-It explicitly incorporates contrast into the similarity function by including a penalty for being similar to exemplars from a given category.
-A contrast parameter allows more different candidate categories to be weighted more heavily than similar candidate categories.
-Both models are flexible enough to describe how generated categories can be distributionally similar (or different) from an experimenter-defined category.
-However, they do not distinguish between  generating an independently identified category and a category-by-negation because there is no mechanism to account for these different identities.
-
-This issue also extends to the methodology applied in studying category generation.
-In \citeA{ward1994structured}, participants were told to generate aliens that belonged to a different species from a prior group of aliens, without applying any specific label to this new group of aliens.
-Similarly, \citeA{jern2013probabilistic} instructed participants to generate a new, different type of crystal after having observed two different types of crystals.
-In contrast, \citeA{conaway2017packer} prompted participants to generate exemplars from a novel "Beta" category, while avoiding an explicit instruction for participants to create something "different".
-In each of these studies there appears to be implicit assumption that generating an independently identified category is equivalent to generating a category-by-negation.
-
-In this paper, we challenge this null assumption by positing that the explicit association of categories-by-negation to its known counterpart (i.e., the explicit identification of the to-be-generated category as \emph{not} a known category) should result in the observer taking advantage of the entire area of the feature space not occupied by the known category.
-In contrast, observers generating independently identified categories should be less focused on the unoccupied feature space and instead construct their categories based on the structure of known categories.
-From this we can predict that categories-by-negation should occupy larger areas of the feature space compared to independently identified categories.
-In addition, the similarity in distributional statistics should extend not only from the learned category to the first independently identified category to be generated (as previous studies have found), but also between subsequent independently identified categories.
-We test these predictions by adapting and extending a category generation experiment by \citeA{conaway2017packer} and explore the implications of its results.
-
-% In this paper, we investigate how the identity of the generated category affects its structure.
-Specifically, we predict that categories-by-negation should occupy larger areas of the feature space compared to independently identified categories.
-Independently identified categories are predicted to share greater distributional similarities with other independently identified categories.
-We test these predictions by adapting and extending a category generation experiment by \citeA{austerweil2018catgen}.
-
-% [Started reading more about Packing Theory and I don't think we really have a good case about it.
-For one thing, I think it can be argued to support both the Not-Alpha and Beta type groups?
-I've moved Packing theory to the discussion instead.] One hypothesis that can be useful in guiding our intuitions about how generated categories can differ is the packing hypothesis \cite{hidaka2011packing}.
-Their Packing Theory suggests that categories are would inform us that the variation in the range of
-% feature values of 'Not-Alpha' exemplars should be larger than than the variation
-% in the range of feature values of 'Beta' and 'Gamma' exemplars.
-
-% While \citeA{austerweil2018catgen} found that new categories tend to be different from known categories, this does not indicate an equivalence between a newly generated category and a category that is not what was known.
-
-% There might be no difference in the resulting new composition if the musician identifies the new work with a specific name (independently identified category) or that the musician identifies the new work as simply not-jazz (category-by-negation).
-Both are meant to be distinct categories from jazz, but the former is identified as its own category and the latter is identified as a category relative to the previous category.
-
-% If the category is identified by some arbitrary name, that should not offer any more or less information to the observer than stating that the category is not a known category;  consequently, the generated category should not be different in either case.
-[convoluted; revise if there's time] However, this argument assumes that a category-by-negation is labeled as arbitrarily as an independently identified category.
-
-% Labelling the to-be-generated category as "" [Gotta work on the names here.] For instance, suppose a musician has released an album of jazz music.
-For her next album, she does not know what particular genre she wants to release -- she simply knows that it is not jazz.
-The question we are asking is, is there a difference between this and a genre called "goose" (for example)?
-% The current state of affairs is that studies and models don't distinguish between the two.
-
-% The generation of a new category involves at least the concept that the category is novel in some sense, and that it is a category.
-Typically, in laboratory tasks the novelty of the category is communicated to the observer by indicating that this to-be-generated category is \emph{different to} the prior observed category.
-
-% Suppose a band of musicians has released an album of jazz music.
-For their next composition, the band leader suggests that they do something different to jazz.
-The double-bassist agrees, but adds that while they should make a new type of music, they should also give this new genre a new name.
-The band leader is confused -- whether or not their new music is given a name is irrelevant so long as it is different to jazz.
-The double-bassist might argue that the name gives the band 'focus'.
-When creating a new category of objects or exemplars, is there a difference in the resulting category
-
-% When motivated to produce something novel, do For
-% instance, if a musician is creating a new album, is there a difference in the
-% content of the new album when they are trying to compose something that is not
-% one of their previous albums, compared to creating something that is a new genre?
-% [needs work ugh]
-
-\section{Experiment}
-%JLA: If possible it would be good to cite CA2017 as much as we can rather than the manuscript.
-It's supposed to be double blind and citing a particular manuscript in preparation all the time may cue reviewers to the identity of the authors...
-Our current experiment closely mirrors the the experimental design of \citeA{austerweil2018catgen} and \citeA{conaway2017packer}, where participants are first trained on a category named 'Alpha' before being tasked to generate exemplars from a new category.
-While the original authors were primarily interested in the measuring the location of generated categories relative to learned categories, our current investigation focuses on analyzing the differences in distributional statistics across different generated categories.
-Consequently, in addition to varying the shape and location of the Alpha category (i.e., across different Alpha conditions where the position of the Alpha exemplars are systematically varied), we include an additional independent variable comprising three different generation conditions: a Not-Alpha condition, where participants generate a new category that is not the learned Alpha category; a Beta-Only condition, where participants generate a new category named 'Beta'; and a Beta-and-Gamma condition, where participants generate a category 'Beta' as well as a category 'Gamma'.
-The resulting 3X3 design is applied in a between-subjects fashion -- participants can be in only one of the nine unique conditions.
-
-The main advantage of adapting the experiments by \citeA{austerweil2018catgen} \citeA{conaway2017packer} is that the simplicity of their stimuli allow for a straightforward test of the distributional similarities across known and generated categories.
-In addition, the variety of Alpha categories used (i.e., the different shapes of the Alpha categories) also allows us to observe the effect of generated category identity across multiple scenarios.
-
-In line with our earlier predictions, we expect that the Not-Alpha conditions on average generate categories that are larger in area and more widely dispersed than categories from the Beta-Only as well as Beta-and-Gamma conditions.
-In addition, we predict that within the Beta-and-Gamma conditions, the generated Beta and generated Gamma categories should be distributionally similar.
-
-% No interaction effects between the factors (i.e., Alpha $\times$ Generation conditions) are expected.
-
-\subsection{Method}
-
-\subsubsection{Participants and materials}
-
-We recruited 240 participants through Amazon Mechanical Turk and randomly
-assigned them to one of the nine unique conditions. Sample sizes of each
-condition are presented in Table \ref{table:samplesize}.
-
-Stimuli were squares that varied along two dimensions: color (grayscale
-9.8\% -- 90.2\%) and size (3.0 -- 5.8cm on each side). The assignment of perceptual
-features (color, size) to axes of the domain space (x, y), as well as the
-direction of variation along each axis (e.g., increasing or decreasing size) was
-counterbalanced across participants. Feature values were evenly-spaced on a
-50X50 grid, giving a near-continuous space from which exemplars can be generated. An example of the feature space is presented in Figure \ref{fig:alphas}a.
-
-Similar to \citeA{austerweil2018catgen}, our three Alpha conditions can be described as Cluster, Row, and Diagonal.
-In the Cluster condition, Alpha exemplars occupy a small area towards one corner of the feature space.
-In the Row condition, the exemplars are nearly equal along one feature, while equally spread out along the other feature.
-The Diagonal condition is an adaptation of the XOR condition in \citeA{austerweil2018catgen}, except instead of exemplars situated at the corners, they are equally spaced along the diagonal of the feature space in a similar fashion to the diagonal conditions of \citeA{jern2013probabilistic}.
-In addition, exemplar feature values are slightly jittered from their original values in \citeA{austerweil2018catgen} in order to ensure that exemplars are not completely identical along a single feature.
-The exact same amount of jitter is applied to all Alpha exemplars within a given Alpha condition.
-The locations of these different Alpha categories in the feature space are presented in Figures \ref{fig:alphas}b to \ref{fig:alphas}d.
-
-\begin{figure*}%[H]
-\begin{center}
-\includegraphics[width=\textwidth]{Figures/conditions.png}
-\end{center}
-\caption{(a) Example of stimuli located at the corners of the feature space. (b-d) Locations of Alpha category exemplars for each Alpha condition.} 
-\label{fig:alphas}
-\end{figure*}
-
-\subsubsection{Procedure}
-In the first phase of the experiment, participants learned the Alpha category
-exemplars by observing a unique exemplar on each trial. This
-was repeated over a total of three blocks (four trials per block -- one
-corresponding to each unique exemplar), with the order of exemplar presentation
-randomized within each block. Participants were allowed to spend as much time as
-they wanted on each trial and were also shown the full range of possible feature
-values prior to training.
-
-The next phase comprised a series of generation trials.
-Depending on their generation condition, participants were asked to generate either eight exemplars from a category that was 'Not-Alpha' (Not-Alpha generation condition), eight exemplars from a category called 'Beta' (Beta-Only generation condition), or four exemplars from a category 'Beta' \emph{and}
-four exemplars from a third category 'Gamma' (Beta-and-Gamma generation condition).
-Exemplars were generated once per trial using two on-screen sliding scales, with each scale controlling the individual features (color and size) of the generated exemplar.
-Feature values could take any one of 50 evenly-spaced values between the specified boundaries.
-Previously generated exemplars were not allowed to be generated a second time.
-Participants were shown an on-screen preview of their exemplar on each trial as they interacted with the sliders, but could not see previously generated exemplars or exemplars from the Alpha category.
-
-\begin{table}[H]
-\begin{center} 
-\caption{Sample sizes for each condition.} 
-\label{table:samplesize} 
-\vskip 0.12in
-\begin{tabular}{llll} 
-\hline
-Generation Condition    &  Cluster & Row & Diagonal \\
-\hline
-Not-Alpha      &   26 & 28 & 25 \\
-Only Beta      &   30 & 27 & 25 \\
-Beta-and-Gamma &   26 & 27 & 26 \\
-\hline
-\end{tabular} 
-\end{center} 
-\end{table}
-
-\subsubsection{Analyses}
-
-We analyze our data in two stages. First, to provide a coarse overview of the distribution of different patterns of generated categories, we classify the generated categories into
-six different profiles: Positives, where the correlation between the dimensions
-is more than $r$; Negatives, where the correlation between the dimensions is less
-than $-r$; Rows, where the range of values across the x dimension is at least
-five times more than the range across the y dimension; Columns, where the range
-of y dimension values is at least five times more than the range of x dimension
-values; Clusters, where the ranges across both dimensions are less than $a$; and
-Dispersed, where the ranges across both dimensions are more than $a$.
-Next, we compare the generated categories on four key distributional measures: ranges of each feature, the feature correlations, and the area enclosed by the generated exemplars in the feature space (i.e., their convex hull).
-Differences along each of these statistics are performed using Bayesian $t$-tests \cite{rouder2009bayesian}, yielding Bayes factors ($BF_{01}$) which indicate evidence for the null hypothesis when $BF_{01} > 1$, with larger values indicating greater evidence for the null hypothesis.
-$BF_{01} < 1$ indicates evidence for the alternative hypothesis.
-Interpretations of the sizes of Bayes factors are guided by \citeA{jeffreys1961}.
-
-\subsection{Results}
-
-We found that setting $r = .7$ and $a = .25$ was useful in capturing the different profiles of generated categories. A representative sample of each profile is shown in Figure
-\ref{fig:samples} and the frequency plot of the different profiles is presented
-in Figure \ref{fig:profile}.
-
-The most striking patterns to note here are the high frequencies of Row category profiles from participants in the Row conditions, and the high frequencies of Dispersed category profiles from participants in the Not-Alpha conditions.
-The former indicates that the distributional similarities between learned and generated categories are especially strong in the Row conditions, while the latter provides preliminary evidence that generated categories from the Not-Alpha conditions tends to be more widely dispersed.
-Also noteworthy are the low counts of both Positive and Negative category profiles across the whole data set -- in contrast to the Row conditions, this indicates low similarity in distributional structure between the learned and generated categories for the Diagonal condition.
-
-%Note that the trained category exemplars (labeled as A in the figure)
-
-\begin{figure}[H]
-\begin{center}
-    \includegraphics[width=.5\textwidth]{Figures/samples.pdf}
-\end{center}
-\caption{Representative samples of the six different category generation profiles.} 
-\label{fig:samples}
-\end{figure}
-
-\begin{figure*}%[H]
-\begin{center}
-    \includegraphics[width=\textwidth]{Figures/profile_freq.png}
-\end{center}
-\caption{Frequencies of category generation profiles broken down by Alpha condition (left plot) and generation condition (right plot).}
-\label{fig:profile}
-\end{figure*}
-
-Focusing on the distributional statistics, when broken down by the Alpha conditions (Figure \ref{fig:stats_cond}), we found that the Cluster conditions tended to have categories with a smaller range of both features, and with correspondingly small sizes.
-The Row conditions produced categories that are high on the x dimension but not the y dimension.
-These observations indicate, as was observed in \citeA{austerweil2018catgen}, that the distribution statistics were carried over from the known category to the generated category in these two Alpha conditions.
-However, the Diagonal conditions did not have similar distributional statistics observed -- instead of a positive correlation, the Diagonal conditions tended to produce large, dispersed categories.
-\citeA{austerweil2018catgen} also observed a similar effect (although they found more evidence for a negative correlation).
-Overall, in alignment with previous research, at least two out of the three Alpha conditions in our experiment generated categories that share distributional statistics as the known category.
-
-More interestingly, when the data is broken down by the generation conditions (Figure \ref{fig:stats_instr}), we found that compared to the Not-Alpha conditions, there is moderate evidence showing Beta-Only conditions with a lower y dimension range ($t(162) = 3.00, BF_{01} = 0.16$) and moderate to strong evidence that their generated categories are smaller in area ($t(162) = 3.16, BF_{01} = 0.10$).
-There is moderate evidence that the Not-Alpha and Beta conditions share equal range of x dimension values ($t(162) = 1.33, BF_{01} = 4.82$).
-With Beta-and-Gamma conditions, we find greater evidence for smaller and tighter categories compared to the Not-Alpha conditions.
-Specifically, there is very strong evidence that both Beta and Gamma categories from the Beta-and-Gamma condition are smaller in both x ($t(156) = 4.83, BF_{01} = 2.55 \times 10^{-4}$;$t(156) = 4.21, BF_{01} = 2.94 \times 10^{-3}$, respectively) and y ($t(156) = 4.57, BF_{01} = 7.40 \times 10^{-4}$;$t(156) = 7.56, BF_{01} = 4.30 \times 10^{-10}$, respectively) ranges compared to the Not-Alpha conditions.
-Similarly, there is very strong evidence that both categories in the Beta-and-Gamma conditions are smaller in area than the Not-Alpha conditions ($t(156) = 5.70, BF_{01} = 5.41 \times 10^{-6}$;$t(156) = 7.69, BF_{01} = 2.07 \times 10^{-10}$, respectively).
-Overall, when comparing the Not-Alpha conditions to categories from other generation conditions, we consistently find moderate to very strong evidence that Not-Alpha categories are more widely dispersed (in their range values) and also larger (in their area), with the only exception being the comparison of x dimension ranges between the Not-Alpha and Beta-Only conditions.
-
-When comparing the distributions of the Beta and Gamma categories generated within the Beta-and-Gamma conditions, we find an overall weak-to-moderate evidence for equal distributional statistics.
-Specifically, there was moderate evidence for equal x dimension ranges ($t(150) = 0.56, BF_{01} = 9.47$) and weak evidence for both feature correlations and area size ($t(150) = 1.85, BF_{01} = 2.10$;$t(150) = 1.77, BF_{01} = 2.40$, respectively).
-When measured on their y dimension ranges, we found weak evidence for lower values from Gamma categories compared to the Beta categories ($t(150) = 2.47, BF_{01} = 0.59$).
-
-% \begin{table}[H]
-% \begin{center} 
-% \caption{Results of Bayesian $t$-test between the generation conditions: pairwise comparisons of Not-Alpha (A') and Beta-Only (B) conditions, as well as the Beta (BC-B) and Gamma (BC-C) categories separately from the Beta-and-Gamma condition.}
-% \label{table:samplesize} 
-% \vskip 0.12in
-% \begin{tabular}{llcc} 
-% \hline
-% Group 1 & Group 2  & $t$(df) & $BF_{01}$ \\
-% \hline
-% A'   & B    &  3.16(162) & $9.81\times10^{-2}$ \\
-% A'   & BC-B &  5.70(156) & $5.41\times10^{-6}$ \\
-% A'   & BC-C &  7.69(156) & $2.07\times10^{-10}$ \\
-% B    & BC-B &  2.48(156) & $0.58$ \\
-% B    & BC-C &  4.27(156) & $2.35\times10^{-3}$ \\
-% BC-B & BC-C &  1.77(150) & $2.40$      \\
-% \hline
-% \end{tabular} 
-% \end{center} 
-% \end{table}
-
-\begin{figure*}%[H]
-\begin{center}
-    \includegraphics[width=\textwidth]{Figures/stats_cond.pdf}
-\end{center}
-\caption{Box-plots of the distributional statistics from the generated categories.
-Boxes depict the median and quartiles of each Alpha condition, with whiskers placed at 1.5 inter-quartile range.
-All points outside this region are marked individually.}
-\label{fig:stats_cond}
-\end{figure*}
-
-\begin{figure*}%[H]
-\begin{center}
-    \includegraphics[width=\textwidth]{Figures/stats_instr.pdf}
-\end{center}
-\caption{Box-plots of the distributional statistics from the generated categories.
-Boxes depict the median and quartiles of each generation condition, with the Not-A condition denoted by A' and the Beta-Only condition denoted by B.
-Data from the Beta-and-Gamma conditions are separated into the Beta groups (denoted BC-B) and the Gamma groups (denoted BC-C) generated in that condition.
-Whiskers are placed at 1.5 inter-quartile range.
-All points outside this region are marked individually.}
-\label{fig:stats_instr}
-\end{figure*}
-
-\section{Discussion}
-At first glance, it seems reasonable to assume that generating a new category X after learning category Y is the same as generating category Not-X.
-An independently identified category should already be a category-by-negation (in that an independently identified category is not what was known).
-Further, if the categories are already identified by arbitrary labels, then it may be easy to assume that identifying the negation of a known category cannot add any additional information in category generation.
-
-Our results have indicated otherwise.
-Specifically, when tasked to produce categories-by-negation, participants tended to generate wider and larger new categories compared to when tasked with producing independently identified categories.
-To our knowledge, this paper represents the first piece of evidence distinguishing these separate types of categories.
-
-Aside from demonstrating a new effect, this study has continued to show the robustness of the distributional similarities between learned and generated categories.
-In this sense, the results from the different Alpha conditions are similar to those observed in \citeA{austerweil2018catgen}.
-The generated categories from the Cluster condition tend to possess lower x and y ranges, with a correspondingly smaller area, and the generated categories from the Row condition tend to adopt Row-type profiles.
-There is also similar lack of categories with positively correlated features from the Diagonal condition.
-
-However, one notable difference is that while \citeA{austerweil2018catgen} found evidence of negatively-correlated generated categories in their XOR condition, we found no evidence of negatively-correlated generated categories in our comparable Diagonal condition.
-\citeA{austerweil2018catgen} explained that the presence of negatively-correlated features in this condition was indicated the effect of category contrast -- that is, categories with negatively correlated features are generated because they are particularly different to categories with positively correlated features.
-It is possible that due to the reduced strength of the positive correlation in our study compared to \citeA{austerweil2018catgen} (because of the addition of noise to the feature values), participants were no longer as sensitive to the negative correlations in the experimenter-defined category and therefore started to produce uncorrelated but widely dispersed categories.
-
-Beyond replicating previous studies, the current study has demonstrated that the consistency in distributional statistics can persist beyond the first generated category.
-However, evidence showing this was ultimately weak.
-One possible reason for this is the relatively small feature space employed in the tasks.
-The generation of a second novel category is necessarily more constrained in the feature space than the generation of the first novel category, possibly resulting in differences in distributional structure.
-By exploring stimuli with less defined boundaries (e.g., orientation), we may expect to see greater consistency in distributional structure over multiple generated categories.
-
-Although we have observed participants generating multiple independently identified categories, we do not want to imply that categories-by-negation can only happen once.
-It would be worth investigating how participants might proceed to generate additional categories-by-negation (e.g., by asking observers to generate a category that is Not-Alpha and Not Beta).
-Packing Theory \cite{hidaka2011packing} -- a hypothesis that suggests categories can be neatly 'packed' into the feature space -- may indicate that successive categories-by-negation are generated in a fashion that preferentially occupies spaces between observed categories.
-Further, although none of the current models of category generation can directly account for the effects observed in this study, they may be useful components in a larger category generation framework.
-For instance, future work may consider implementing the hierarchical sampling model from \citeA{jern2013probabilistic} in a framework of overhypotheses \cite{kemp07}, where a prior can be placed over a category identity space, allowing models to behave differently under different regions of generated category identity.
-
-Ultimately, the nature of newly generated categories appears to vary depending on the identity they were associated with.
-The extent to which they may differ, and the mechanisms driving these differences represent fascinating areas for future research.
 
@@
 
-\bibliography{citations.bib}
+%\bibliography{citations.bib}
 

### 2019-03-06 Experiments/cogpsych-code/print_gs_results_csv.py

@@ -12,7 +12,7 @@ import numpy as np
 pickledir = 'pickles/'
 writefile = 'globalfits'
 prefix = 'chtc_gs_best_params'
-printdata = ['e1_e2']#'all'
+printdata = ['e1_e2','inc150','corrp']#'all'
 if not printdata is 'all':
     gsfiles = []
     data_names = []

### 2019-03-06 Experiments/cogpsych-code/status_chtc.py

@@ -8,4 +8,4 @@ narg = len(sys.argv)
 if __name__ == "__main__" and narg>1:
     fetchdir = sys.argv[1]
 cmd = 'condor_q; cd {};cat output/gencat*.err;'.format(fetchdir)
-os.system('ssh -i ~/Dropbox/.ssh/chtc -t liew2@submit-5.chtc.wisc.edu \'{}\' '.format(cmd))
+os.system('ssh -i ~/Dropbox/.ssh/chtc -t liew2@submit-1.chtc.wisc.edu \'{}\' '.format(cmd))

### 2019-03-06 Experiments/middle-bottomABC/analysis/.#plot.distance.figs.py

@@ -1,1 +1,0 @@
-sourguyxian@Awesomeness2018.49537

### 2019-03-06 Experiments/middle-bottomABC/analysis/gmm.py

@@ -1,0 +1,67 @@
+import numpy as np, pandas as pd, matplotlib.pyplot as plt, seaborn as sns
+import pymc3 as pm, theano.tensor as tt
+sns.set_context('paper')
+sns.set_style('darkgrid')
+
+# simulate data from a known mixture distribution
+np.random.seed(12345) # set random seed for reproducibility
+
+k = 3
+ndata = 500
+spread = 5
+centersX = np.array([-spread, 0, spread])
+centersY = np.array([-spread+2,1,spread-1.4])
+#centers = np.array([[-5,2],[3,5],[1,1]])
+# simulate data from mixture distribution
+v = np.random.randint(0, k, ndata)
+dataX = centersX[v] + np.random.randn(ndata)
+dataY = centersY[v] + np.random.randn(ndata)
+#data = centers[v] + np.reshape(np.random.randn(ndata*2),(ndata,2))
+data = np.concatenate([np.atleast_2d(dataX),np.atleast_2d(dataY)],0)
+data = data.transpose()
+plt.hist2d(dataX,dataY,bins=20);
+
+# setup model
+model = pm.Model()
+with model:
+    # cluster sizes
+    p = pm.Dirichlet('p', a=pm.floatX(0.1 * np.ones(2)),shape = (2,))
+    #p = pm.Dirichlet('p', a=np.array([1., 1., 1.]), shape=k)
+    # ensure all clusters have some points
+    p_min_potential = pm.Potential('p_min_potential', tt.switch(tt.min(p) < .1, -np.inf, 0))
+    # cluster centers
+    means = [pm.MvNormal('mu_%d'%i,
+                         pm.floatX(np.zeros(2)),
+                         tau=pm.floatX(.1*np.eye(2)),
+                         shape=(2,))
+             for i in range(2)]
+                         #mu=[0, 0, 0], sd=15, shape=k)
+    # break symmetry
+    #order_means_potential = pm.Potential('order_means_potential',
+    #                        tt.switch(means[1]-means[0] < 0, -np.inf, 0)
+    #                        + tt.switch(means[2]-means[1] < 0, -np.inf, 0))
+    # measurement error
+    sigma = pm.LKJCholeskyCov('packed_L', n=2,
+                              eta=2., sd_dist=pm.HalfCauchy.dist(2.5))
+    L = pm.expand_packed_triangular(2, sigma)
+    #sd = pm.Uniform('sd', lower=0, upper=20)
+    # latent cluster of each observation
+    category = pm.Categorical('category',
[…]

### 2019-03-06 Experiments/middle-bottomABC/webapp/cgi-bin/config.py

@@ -8,7 +8,7 @@ print                                                             # blank line, 
 # - - - - - - - - - - - - - - - - - - - - - - - -
 # USER SHOULD SET THESE PARAMETERS
 # This is where all the data is stored (X.json and workers.db) 
-destination = "/var/services/homes/xian/CloudStation/data/middle-bottomABC"
+destination = "../../datatemp"
 assignmentdb = destination + "/assignments.db"
 
 # what are the conditions?

### 2019-03-06 Experiments/xor-cluster-rowABC/analysis/descriptives.py

@@ -1,0 +1,59 @@
+import sqlite3, sys
+import pandas as pd
+import numpy as np
+
+pd.set_option('display.width', 120, 'precision', 2)
+
+con = sqlite3.connect('../data/experiment.db')
+participants = pd.read_sql_query("SELECT * from participants", con)
+counterbalance = pd.read_sql_query("SELECT * from counterbalance", con)
+stats = pd.read_sql_query("SELECT * from betastats", con)
+con.close()
+
+print participants.shape
+
+# counts per condition
+#print(participants.groupby('condition').size())
+print(participants.groupby(['condition','gentype']).size())
+
+
+participants = pd.merge(participants, counterbalance, on = 'counterbalance')
+print pd.pivot_table(
+    data = participants,
+    columns = 'xax',
+    index = 'condition',
+    aggfunc = 'size'
+    )
+
+
+
+stats = pd.merge(stats, participants, on = 'participant')
+from scipy.stats import ttest_ind, mannwhitneyu, ttest_rel
+cols = ['area','between','within',
+                'correlation', 
+                'drange', 'xrange', 'yrange', 'xstd', 'ystd']
+conds = ['C0','C1','C2','R0','R1','R2','X0','X1','X2']
+for i in cols:
+    gs = list(stats.groupby('condition')[i])
+    gs = list(stats.groupby(['condition','gentype'])[i])
+    d = dict(gs)
+    ms = dict([(k, np.mean(v)) for k,v in d.items()])
+    #p = mannwhitneyu(d['Middle'], d['Bottom']).pvalue
+
+    S = i
+    tempdict = {}
+    for k,v in d.items():
+        kstr = str(k[0][0]) + str(k[1])
+        tempdict[kstr] = round(np.mean(v),3)
+        #S += '\t' + kstr + ' = ' + str(round(np.mean(v),3))
+    #S += '\t' + 'p = ' + str(round(p,3))
+    for cond in conds:
+        S += '\t' + cond + ' = ' + str(tempdict[cond])
+    print S
+
+
+for j, rows in stats.groupby('condition'):
+    g1 = rows['within'].as_matrix()
+    g2 = rows['between'].as_matrix()
+    print j, ttest_rel(g1, g2)
+

### 2019-03-06 Experiments/xor-cluster-rowABC/analysis/plot.heatmaps.py

@@ -1,0 +1,59 @@
+import sqlite3
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+
+np.set_printoptions(precision = 2)
+
+execfile('Imports.py')
+import Modules.Funcs as funcs
+
+
+
+con = sqlite3.connect('../data/experiment.db')
+infodf = pd.read_sql_query("SELECT * from participants", con)
+stimuli = pd.read_sql_query("SELECT * from stimuli", con).as_matrix()
+alphas = pd.read_sql_query("SELECT * from alphas", con)
+generation = pd.read_sql_query("SELECT * from generation", con)
+con.close()
+
+#generation = generation.loc[generation.trial<4]
+
+from scipy.ndimage.filters import gaussian_filter
+
+clim = (0.0, 0.10)
+
+f, ax = plt.subplots(3,3, figsize = (5,4))
+for i, (K, rows) in enumerate(infodf.groupby('condition')):
+    for gentype,rowss in rows.groupby('gentype'):
+        pids = rowss.participant
+        betas = generation.loc[generation.participant.isin(pids), 'stimulus']
+        
+        counts = np.array([sum(betas==j) for j in range(stimuli.shape[0])])
+        ps = counts / float(sum(counts))
+        g = funcs.gradientroll(ps,'roll')[:,:,0]
+        
+        # g = gaussian_filter(g, 0.8)
+        
+        print(g)
+        print(K, np.max(g))
+        
+        h = ax[i,gentype]
+        im = funcs.plotgradient(h, g, stimuli[alphas[K],:], [], clim = clim)
+        h.set_title(K, fontsize = 12)
+        
+
+# add colorbar
+f.subplots_adjust(right=0.8)
+cbar = f.add_axes([0.83, 0.225, 0.03, 0.54])
+f.colorbar(im, cax=cbar, ticks = clim)
+cbar.set_yticklabels(['Lowest\nProbability', 'Greatest\nProbability'])
+cbar.tick_params(length = 0)
+plt.tight_layout(w_pad=-8, h_pad= .5)
+
+f.savefig('heatmaps.pdf', bbox_inches='tight', transparent=False)
+
+path = '../../../Manuscripts/cogsci-2017/figs/behavioral-heatmaps.pgf'
+# funcs.save_as_pgf(f, path)
+
+

### 2019-03-06 Experiments/xor-cluster-rowABC/analysis/plot.individuals.py

@@ -1,0 +1,86 @@
+import sqlite3, os
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+import pickle
+
+execfile('Imports.py')
+import Modules.Funcs as funcs
+
+pd.set_option('precision', 2)
+
+# import data
+con = sqlite3.connect('../data/experiment.db')
+info = pd.read_sql_query("SELECT * from participants", con)
+df = pd.read_sql_query("SELECT * from generation", con)
+alphas = pd.read_sql_query("SELECT * from alphas", con)
+stimuli = pd.read_sql_query("SELECT * from stimuli", con).as_matrix()
+stats = pd.read_sql_query("SELECT * from betastats", con)
+con.close()
+
+savedir = 'individuals'
+gentypeStr = ['N','B','C'] #not alpha, only beta, beta-gamma
+gentypeStrDisp = ['A\'','B','C'] #not alpha, only beta, beta-gamma
+gentypeCols = [[.3,0,.5],[0,0,.5],[0,.5,0]]
+f, ax= plt.subplots(1,1, figsize=(1.6, 1.6))
+
+var_thresh = .25
+cor_thresh = .7
+groupcount = {'Positive':0,'Negative':0,'Cluster':0,'Dispersed':0,'Row':0,'Column':0}
+data_arr = []
+for i, row in info.iterrows():
+    pid, condition, gentype = int(row.participant), row.condition, row.gentype
+    gentypeStr_p = gentypeStr[gentype]
+
+    palphas = alphas[condition]
+    #pbetas_all = df.stimulus[df.participant == pid]
+    if gentype==2:
+        pdf = df.loc[df.participant==pid]
+        betastr = [gentypeStrDisp[1] if pdf_row.category=='Beta' else gentypeStrDisp[2] for ii,pdf_row in pdf.iterrows() ]
+        betacol = [gentypeCols[1] if pdf_row.category=='Beta' else gentypeCols[2] for ii,pdf_row in pdf.iterrows() ]
+        pbetas_list = [pdf.stimulus[pdf.category=='Beta']]
+        pbetas_list += [pdf.stimulus[pdf.category=='Gamma']]
+    else:
+        pbetas_list = [df.stimulus[df.participant == pid]]
+        betastr = gentypeStrDisp[gentype]
+        betacol = gentypeCols[gentype]
+    #Add thresholds for classifying category shapes
+    #get covariance matrix
+    printstr = ''
+    for ii,pbetas in enumerate(pbetas_list):
[…]

### 2019-03-06 Experiments/xor-cluster-rowABC/send2luke.py

@@ -1,0 +1,23 @@
+#Package up webapp dir and send to server (Luke)
+
+import os, sys
+
+localdir = 'webapp'
+serverdirbase = '/var/services/web/experiments/'
+serverdir = 'catgen_xcrABC'
+
+#tarzip local
+os.system('tar -cvzf {}.tar.gz {}'.format(localdir,localdir))
+#scp it over 
+os.system('scp -i ~/Dropbox/.ssh/luke -P 1202 {}.tar.gz xian@alab.psych.wisc.edu:{}'.format(localdir,serverdirbase))
+
+#mk new dir on host and untar the tar
+cdcmd = 'cd {}'.format(serverdirbase)
+untarcmd = 'tar -xvzf {}.tar.gz '.format(localdir)
+rmcmd = 'if [ -d {} ]; then rm -r {}; fi'.format(serverdir,serverdir)
+renamecmd = 'mv {} {}'.format(localdir,serverdir)
+#rm the tar
+removecmd = 'rm {}.tar.gz'.format(localdir)
+#run it all
+os.system('ssh -i ~/Dropbox/.ssh/luke -p 1202 -t xian@alab.psych.wisc.edu  \'{};{};{};{};{}\' '.format(cdcmd,untarcmd,rmcmd,renamecmd,removecmd))
+

### 2019-03-06 Experiments/xor-cluster-rowABC/webapp/js/generation.js

@@ -1,0 +1,216 @@
+function generate() {
+	stage.innerHTML = '';	
+	
+	// some function globals
+	var dupemessage;
+	var continuebutton;
+	var stimulusdiv;
+	var color_control
+	var size_control
+
+	//reset counter
+	generation.counter = 0
+
+	//get category type
+	if (data.info.gentype==2){ //beta-gamma condition
+		generation.category = generation.bcnames[generation.countbc]
+		generation.ntrials = generation.ntrialsbase
+	} else {
+		if (data.info.gentype==1) { //beta-only condition
+			generation.category = 'Beta'
+		} else { //not-alpha condition
+			generation.category = 'Not-Alpha'
+		}
+		generation.ntrials = generation.ntrialsbase*2
+	}
+
+	// function to start a new trial
+	function init() {		
+		// replace existing ui.
+		stage.innerHTML = generation.ui;
+		stage.style.visibility = 'hidden'; // hide everything during setup
+		if (generation.useknobs){
+			document.getElementById('size_holder').innerHTML =
+				'<input type = "text"'+
+				'class = "dial"'+
+				'id = "size_control">'
+			document.getElementById('color_holder').innerHTML =
+				'<input type = "text"'+
+				'class = "dial"'+
+				'id = "color_control">'
+			cc_old = 10;
+			sc_old = 10;
+			$(".dial").knob({				
+				'min':0,
+				'max':stimsteps-1,
+				'width':100,
+				'cursor':true,
+				'thickness':.3,
+				'fgColor': "#222222",
+				'displayInput':false,
+				'change': function() { generate_handler() }
+			});
+			$(".dial").val(cc_old)
+		} else {
+			document.getElementById('size_holder').innerHTML =
+				'<input type = "range"'+
+		  		'class = "feature_range"'+
+		  		'id = "size_control"'+
+				'min = "0"'+
+		  		'max = "10"'+
+		  		'step = "1">'
+		  					
+			document.getElementById('color_holder').innerHTML =
+				'<input type = "range"'+
+		  		'class = "feature_range"'+
+		  		'id = "color_control"'+
+		  		'min = "0"'+
+		  		'max = "10"'+
+		  		'step = "1">'
+		}
+		
+		// define UI elements
+		continuebutton = document.getElementById('continuebutton');
+		stimulusdiv = document.getElementById('stimulus');
+		dupemessage = document.getElementById('dupemessage');
+		color_control = document.getElementById('color_control');
[…]

### 2019-03-06 Experiments/xor-cluster-rowABC/webapp/js/server.js

@@ -1,0 +1,110 @@
+// Functions to handle communication with servers
+
+
+// see if the worker is eligible
+function checkworker() {
+	stage.innerHTML =  "Waiting for server...";
+
+	var args = {
+		workerId: worker.workerId,
+		Stimuli: data.experiment.Stimuli,
+		Experiment: data.experiment.Experiment,
+		Paradigm: data.experiment.Paradigm
+	}
+
+	$.ajax({
+			type: "POST",
+			url: "cgi-bin/check-worker.cgi",
+			data: JSON.stringify(args),
+		}).done( function(o) {
+			var res = JSON.parse(o);
+
+			// get assignments for experiment
+			if (res.status == 'go') {
+				getassignments();
+
+			// show a "return HIT" screen
+			} else if (res.status == 'exposed') {
+				inserthtml('html/return_message.html');
+			}
+			
+		});
+}
+
+
+// get ID, condition, and counterbalance assignments
+function getassignments() {
+	$.ajax({
+		type: "POST",
+		url: "cgi-bin/get-assignments.cgi",
+		error: function(x,e){console.log('Some ajax error in getassignments(). ' + x.responseText)}
+	}).done( function(o) {
+		console.log(o)
+		var res = JSON.parse(o);
+		data.info.participant = res.data.participant;
+		data.info.condition = res.data.condition;
+		data.info.counterbalance = res.data.counterbalance;
+		data.info.gentype = res.data.gentype
+	  	startup()			
+	});
+}
+
+
+// function to mark the worker as exposed 
+function markexposed() {
+
+	var args = {
+		workerId: worker.workerId,
+		Stimuli: data.experiment.Stimuli,
+		Experiment: data.experiment.Experiment,
+		Paradigm: data.experiment.Paradigm
+	}
+
+	$.ajax({
+			type: "POST",
+			url: "cgi-bin/mark-exposed.cgi",
+			data: JSON.stringify(args),
+		}).done( function(o) {
+			
+			var res = JSON.parse(o)
+
+			if (res.status == 'lab') {
+				console.log('Hi Nolan!')
+				data.info.lab = true;
+			} else {
+				data.info.lab = false;
+			}
+		})
+}
+
+
+// function to mark the participant as complete on the server
+function markcomplete() {
+	$.ajax({
+			type: "POST",
+		  url: "cgi-bin/mark-complete.cgi",
+		  data: JSON.stringify(data.info.participant),
[…]

### 2019-03-06 Modules/Classes/Simulation.py

@@ -669,8 +669,8 @@ def extractPptData(trial_obj, ppt = 'all', unique_trials = 'all'):
                     extractIdx = np.array(pptcats==np.array(round(i)))
                     responsecats = np.array(responsecats)
                     pptcats = np.array(pptcats)
-                    respList = responsecats[extractIdx]
-                    pptList = pptcats[extractIdx]
+                    respList = np.append(respList,responsecats[extractIdx])
+                    pptList = np.append(pptList,pptcats[extractIdx])
                         
         elif trial_obj.task is 'assign' or trial_obj.task is 'error':
             #iterate over categories of responses        
@@ -691,8 +691,8 @@ def extractPptData(trial_obj, ppt = 'all', unique_trials = 'all'):
         else:
             raise ValueError('trialset.task not specified. Please specify this as \'generate\' or \'assign\' in your script.')
                                 
-        output_obj.Set[ti]['response'] = respList
-        output_obj.Set[ti]['participant'] = pptList
+        output_obj.Set[ti]['response'] = respList.astype(int)
+        output_obj.Set[ti]['participant'] = pptList.astype(int)
     #Clean up
     #cleanIdx = np.ones(len(output_obj.Set),dtype=bool)
     output_objTemp = cp.deepcopy(output_obj.Set)

### 2019-05-09 Manuscripts/cogsci-2019/paper2019.tex
Mostly work on cogsci2019 submission. Added new paras for clarity, new sentences as suggested by reviewers, figures made grayscale, added screenshots, fixed typos, overall a little more awesome.

@@
 
-%\cogscifinalcopy % Uncomment this line for the final submission
+\cogscifinalcopy % Uncomment this line for the final submission
 
@@
 \usepackage{pslatex}
-\usepackage{apacite}
+\usepackage[natbibapa]{apacite}
 \usepackage{float} % Roger Levy added this and changed figure/table
@@
 
-\title{Novel categories are distinct from "Not"-categories}
+\title{Novel categories are distinct from ``Not''-categories}
 
@@
 
-Category generation is an exciting but still poorly understood topic in cognitive science.
-One issue that is consistently being overlooked is the conflation of generating new, independently identified categories with the generation of categories that are the negation of a known category.
-This study represents the first attempt at delineating these two separate types of generated categories.
-We specifically test for differences in the distributional structure of generated categories, demonstrating that categories identified as \emph{not} what was known are larger and wider-spread compared to categories that were identified with a specific label.
-We also observe consistency in distributional structure across multiple generated categories, replicating and extending previous findings.
-These results are discussed in the context of providing a foundation for future modeling work.
+  The categorization literature often considers two types of categories as equivalent: (a) standard categories and (b)
+  negation categories. For example, category learning studies typically conflate learning categories A and B with
+  learning categories A and NOT A. This study represents the first attempt at delineating these two separate types of
+  generated categories. We specifically test for differences in the distributional structure of generated categories,
+  demonstrating that categories identified as \emph{not} what was known are larger and wider-spread compared to
+  categories that were identified with a specific label. We also observe consistency in distributional structure across
+  multiple generated categories, replicating and extending previous findings. These results are discussed in the context
+  of providing a foundation for future modeling work.
 
@@
 % concerned with
-People are remarkable in their capacity to innovate new and different ideas.
+%People are remarkable in their capacity to innovate new and different ideas.
 Is creating a new idea the same as creating a different idea?
@@
 Both new categories of music are distinct from jazz.
-The former is identified as its own category,   the latter is identified in relation to a known category.
+The former is identified as its own category and the latter is identified in relation to a known category.
 
-While categorization researchers have primarily focused their effort on classification (associating an exemplar with a category given its features), and inference (predicting exemplar features given its category),
-work on category generation -- predicting all exemplar features for a novel
-category -- is relatively scarce. Recent work has established a few key
+People are remarkable in their capacity to innovate new and different ideas. Is creating a new idea the same as creating
+a different idea? Consider a restaurant that serves one meal per night. Their chef cooked red curry last night and wants
+to create and cook a new dish tonight. Is that the same as wanting to create and cook a new dish that is {\em not} red
+curry? While the former is identified as its own category, the latter is identified in relation to a known category.
+
+While categorization researchers have primarily focused their effort on classification (associating an exemplar with a
+category given its features), and inference (predicting exemplar features given its category), work on category
+generation -- predicting all exemplar features for a novel category -- is relatively scarce. This is surprising because
+category generation is not an uncommon phenomenon -- people are constantly challenged to generate novel categories, such
+as a new meal plan for the week, a new music playlist for an upcoming road trip, or a new exercise regimen to stay
+healthy.
+
+Recent category generation work has established a few key
 findings. Earlier studies have shown that generated
-categories tend to share distributional statistics with learned categories \cite{jern2013probabilistic,thomas98,ward1994structured}. More recently, 
-\citeA{austerweil2018catgen} have found that category contrast is an important
-factor in category generation and learning -- computational models sensitive to the
-differences between categories were a better fit to generated categories and differences in learning different generated categories than models
+categories tend to share distributional statistics with learned categories \citep{ward1994structured,jern2013probabilistic,thomas98}. More recently, 
+\cite{austerweil2018catgen} and \cite{conaway2017packer} have found that category contrast is an important factor in category generation and learning -- computational models sensitive to the
+differences between categories were a better fit to generated categories than models
 which did not take categories contrast into account.
 
-Although previous work has established a few key findings, the processes involved in category generation are still not well understood. In particular, it is currently unclear whether generating a
-new category is itself distinct from identifying the set of exemplar features that are simply not what was previously identified in a known category.
-
- In order to better understand the nature of generated categories, it may be necessary to distinguish between these two possible types of generated categories: one driven by a motivation to be not what is known -- a \emph{category-by-negation}, and another driven by its own identity -- an \emph{independently identified category}.
+Although previous work has established a few key findings, the basic phenomena and processes involved in category
+generation are still not well understood. In this article, we examine whether given a known category in a domain,
+generating a new category is different from generating a category that is not the learned category. If the only category
+is A, most formal accounts of categorization would consider generating a new category B as equivalent to generating not
+A. In order to better understand the nature of generated categories, it is necessary to distinguish between these two
+possible types of generated categories: one driven by its own identity -- an \emph{independently identified category},
+and another driven by  a motivation to be not what is known -- a
+\emph{category-by-negation}.
 
 Current models of category generation do not explicitly make this distinction.
-One of the first computational models of category generation, \citeA{jern2013probabilistic}'s hierarchical sampling model, is a Bayesian model that reproduces distributionally similar categories by assuming that the covariance matrix of features of exemplars from generated categories is generated by the same prior that generated the covariance matrix of the known categories.
-An alternative proposed model, PACKER \cite{austerweil2018catgen,conaway2017packer}, is an extension of the classic Generalized Context Model \cite<GCM;>{nosofsky1984choice,nosofsky1986attention}.
-It explicitly incorporates contrast into the similarity function by including a penalty for being similar to exemplars from a given category.
-A contrast parameter allows more different candidate categories to be weighted more heavily than similar candidate categories.
+One of the first computational models of category generation, \cite{jern2013probabilistic}'s hierarchical sampling model, is a Bayesian model that reproduces distributionally similar categories by assuming that the covariance matrix of features of exemplars from generated categories is generated by the same prior that generated the covariance matrix of the known categories.
+An alternative proposed model, PACKER \citep{conaway2017packer}, is an extension of the classic Generalized Context Model \citep[GCM;][]{nosofsky1984choice,nosofsky1986attention}.
+It explicitly incorporates contrast into the similarity function by including a penalty for being similar to exemplars from a known category.
+A contrast parameter allows candidate categories that are more different from the known category to be weighted more heavily than candidate categories similar to the known category.
 Both models are flexible enough to describe how generated categories can be distributionally similar (or different) from an experimenter-defined category.
@@
 
-This issue also extends to the methodology applied in studying category generation.
-In \citeA{ward1994structured}, participants were told to generate aliens that belonged to a different species from a prior group of aliens, without applying any specific label to this new group of aliens.
-Similarly, \citeA{jern2013probabilistic} instructed participants to generate a new, different type of crystal after having observed two different types of crystals.
-In contrast, \citeA{conaway2017packer} prompted participants to generate exemplars from a novel "Beta" category, while avoiding an explicit instruction for participants to create something "different".
-In each of these studies there appears to be implicit assumption that generating an independently identified category is equivalent to generating a category-by-negation.
+This issue also extends to the methodology applied in category generation (as well as most categorization studies.) In
+\cite{ward1994structured}, participants were told to generate aliens that belonged to a different species from a prior
+group of aliens, without applying any specific label to this new group of aliens. Similarly,
+\cite{jern2013probabilistic} instructed participants to generate a new, different type of crystal after having observed
+two different types of crystals. In contrast, \cite{conaway2017packer} prompted participants to generate exemplars from
+a novel ``Beta'' category, while avoiding an explicit instruction for participants to create something ``different''. In
+each of these studies there appears to be an implicit assumption that generating an independently identified category is equivalent to generating a category-by-negation.
 
@@
 In addition, the similarity in distributional statistics should extend not only from the learned category to the first independently identified category to be generated (as previous studies have found), but also between subsequent independently identified categories.
-We test these predictions by adapting and extending a category generation experiment by \citeA{conaway2017packer} and explore the implications of its results.
-
-% In this paper, we investigate how the identity of the generated category affects its structure.
-Specifically, we predict that categories-by-negation should occupy larger areas of the feature space compared to independently identified categories.
-Independently identified categories are predicted to share greater distributional similarities with other independently identified categories.
-We test these predictions by adapting and extending a category generation experiment by \citeA{austerweil2018catgen}.
-
-% [Started reading more about Packing Theory and I don't think we really have a good case about it.
-For one thing, I think it can be argued to support both the Not-Alpha and Beta type groups?
-I've moved Packing theory to the discussion instead.] One hypothesis that can be useful in guiding our intuitions about how generated categories can differ is the packing hypothesis \cite{hidaka2011packing}.
-Their Packing Theory suggests that categories are would inform us that the variation in the range of
-% feature values of 'Not-Alpha' exemplars should be larger than than the variation
-% in the range of feature values of 'Beta' and 'Gamma' exemplars.
-
-% While \citeA{austerweil2018catgen} found that new categories tend to be different from known categories, this does not indicate an equivalence between a newly generated category and a category that is not what was known.
-
-% There might be no difference in the resulting new composition if the musician identifies the new work with a specific name (independently identified category) or that the musician identifies the new work as simply not-jazz (category-by-negation).
-Both are meant to be distinct categories from jazz, but the former is identified as its own category and the latter is identified as a category relative to the previous category.
-
-% If the category is identified by some arbitrary name, that should not offer any more or less information to the observer than stating that the category is not a known category;  consequently, the generated category should not be different in either case.
-[convoluted; revise if there's time] However, this argument assumes that a category-by-negation is labeled as arbitrarily as an independently identified category.
-
-% Labelling the to-be-generated category as "" [Gotta work on the names here.] For instance, suppose a musician has released an album of jazz music.
-For her next album, she does not know what particular genre she wants to release -- she simply knows that it is not jazz.
-The question we are asking is, is there a difference between this and a genre called "goose" (for example)?
-% The current state of affairs is that studies and models don't distinguish between the two.
-
-% The generation of a new category involves at least the concept that the category is novel in some sense, and that it is a category.
-Typically, in laboratory tasks the novelty of the category is communicated to the observer by indicating that this to-be-generated category is \emph{different to} the prior observed category.
-
-% Suppose a band of musicians has released an album of jazz music.
-For their next composition, the band leader suggests that they do something different to jazz.
-The double-bassist agrees, but adds that while they should make a new type of music, they should also give this new genre a new name.
-The band leader is confused -- whether or not their new music is given a name is irrelevant so long as it is different to jazz.
-The double-bassist might argue that the name gives the band 'focus'.
-When creating a new category of objects or exemplars, is there a difference in the resulting category
-
-% When motivated to produce something novel, do For
-% instance, if a musician is creating a new album, is there a difference in the
-% content of the new album when they are trying to compose something that is not
-% one of their previous albums, compared to creating something that is a new genre?
-% [needs work ugh]
+We test these predictions by adapting and extending a category generation experiment by \cite{conaway2017packer} and explore the implications of its results.
 
@@
 It's supposed to be double blind and citing a particular manuscript in preparation all the time may cue reviewers to the identity of the authors...
-Our current experiment closely mirrors the the experimental design of \citeA{austerweil2018catgen} and \citeA{conaway2017packer}, where participants are first trained on a category named 'Alpha' before being tasked to generate exemplars from a new category.
-While the original authors were primarily interested in the measuring the location of generated categories relative to learned categories, our current investigation focuses on analyzing the differences in distributional statistics across different generated categories.
-Consequently, in addition to varying the shape and location of the Alpha category (i.e., across different Alpha conditions where the position of the Alpha exemplars are systematically varied), we include an additional independent variable comprising three different generation conditions: a Not-Alpha condition, where participants generate a new category that is not the learned Alpha category; a Beta-Only condition, where participants generate a new category named 'Beta'; and a Beta-and-Gamma condition, where participants generate a category 'Beta' as well as a category 'Gamma'.
-The resulting 3X3 design is applied in a between-subjects fashion -- participants can be in only one of the nine unique conditions.
+Our current experiment closely mirrors the the experimental design of \cite{conaway2017packer}, where participants are first trained on a category named `Alpha' before being tasked to generate exemplars from a new category.
+While \cite{conaway2017packer} were primarily interested in the measuring the location of generated categories relative to learned categories, our current investigation focuses on analyzing the differences in distributional statistics across different generated categories.
+Consequently, in addition to varying the shape and location of the Alpha category (i.e., across different Alpha conditions where the position of the Alpha exemplars are systematically varied) in three distinct ways, we include an additional independent variable comprising three different generation conditions: a Not-Alpha condition, where participants generate a new category that is not the learned Alpha category; a Beta-Only condition, where participants generate a new category named `Beta'; and a Beta-and-Gamma condition, where participants generate a category `Beta' as well as a category `Gamma'.
+The resulting 3-by-3 design is applied in a between-subjects fashion -- participants can be in only one of the nine unique conditions.
 
-The main advantage of adapting the experiments by \citeA{austerweil2018catgen} \citeA{conaway2017packer} is that the simplicity of their stimuli allow for a straightforward test of the distributional similarities across known and generated categories.
+The main advantage of adapting the experiment by \cite{conaway2017packer} is that the simplicity of their stimuli allow for a straightforward test of the distributional similarities across known and generated categories.
 In addition, the variety of Alpha categories used (i.e., the different shapes of the Alpha categories) also allows us to observe the effect of generated category identity across multiple scenarios.
@@
 counterbalanced across participants. Feature values were evenly-spaced on a
-50X50 grid, giving a near-continuous space from which exemplars can be generated. An example of the feature space is presented in Figure \ref{fig:alphas}a.
+50-by-50 grid, giving a near-continuous space from which exemplars can be generated. An example of the feature space is presented in Figure \ref{fig:alphas}a.
 
-Similar to \citeA{austerweil2018catgen}, our three Alpha conditions can be described as Cluster, Row, and Diagonal.
+The three Alpha conditions are Cluster, Row, and Diagonal.
 In the Cluster condition, Alpha exemplars occupy a small area towards one corner of the feature space.
 In the Row condition, the exemplars are nearly equal along one feature, while equally spread out along the other feature.
-The Diagonal condition is an adaptation of the XOR condition in \citeA{austerweil2018catgen}, except instead of exemplars situated at the corners, they are equally spaced along the diagonal of the feature space in a similar fashion to the diagonal conditions of \citeA{jern2013probabilistic}.
-In addition, exemplar feature values are slightly jittered from their original values in \citeA{austerweil2018catgen} in order to ensure that exemplars are not completely identical along a single feature.
+The Diagonal condition has Alpha exemplars equally spaced along the diagonal of the feature space in a similar fashion to the diagonal conditions of \cite{jern2013probabilistic}.
+In order to ensure that exemplars are not completely identical along any one feature, exemplar feature values are slightly jittered.
 The exact same amount of jitter is applied to all Alpha exemplars within a given Alpha condition.
@@
 
-\begin{figure*}%[H]
+\begin{figure}%[H]
 \begin{center}
-\includegraphics[width=\textwidth]{Figures/conditions.png}
+\includegraphics[width=.5\textwidth]{Figures/conditions.pdf}
 \end{center}
@@
 \label{fig:alphas}
+\end{figure}
+
+\subsubsection{Procedure}
+In the first phase of the experiment (Figure \ref{fig:screens_learn}), participants learned the Alpha category exemplars
+by observing a unique exemplar on each trial. This was repeated over a total of three blocks (four trials per block --
+one corresponding to each unique exemplar,) with the order of exemplar presentation randomized within each block. Prior
+to the presentation of each exemplar, a fixation cross was shown for 1000 ms. Participants were allowed to spend as much
+time as they wanted on each trial and were also shown the full range of possible feature values prior to training.
+
+\begin{figure*}%[H]
+\begin{center}
+    \includegraphics[width=\textwidth]{Figures/screens_learn.png}
+\end{center}
+\caption{Instructions and trials observed by each participant during the category learning phase. The instructions screen is
+shown once, followed by 12 presentations of Alpha exemplars (4 exemplars across 3 blocks.)}
+\label{fig:screens_learn}
 \end{figure*}
 
-\subsubsection{Procedure}
-In the first phase of the experiment, participants learned the Alpha category
-exemplars by observing a unique exemplar on each trial. This
-was repeated over a total of three blocks (four trials per block -- one
-corresponding to each unique exemplar), with the order of exemplar presentation
-randomized within each block. Participants were allowed to spend as much time as
-they wanted on each trial and were also shown the full range of possible feature
-values prior to training.
+The next phase comprised a series of generation trials (Figure \ref{fig:screens_gen}). Depending on their generation
+condition, participants generated either eight exemplars from a category that was `Not-Alpha' (Not-Alpha generation
+condition), eight exemplars from a category called `Beta' (Beta-Only generation condition), or four exemplars from a
+category `Beta' \emph{and} four exemplars from a third category `Gamma' (Beta-and-Gamma generation condition). More
+specifically, participants in the Not-Alpha generation condition were asked to produce ``what [they] think is likely to
+NOT be in the Alpha category'', while participants in the other conditions were asked to produce ``what [they] think is
+likely to be in the Beta [or Gamma] category''. Exemplars were generated on each trial using two on-screen sliding
+scales, with each scale controlling the individual features (color and size) of the generated exemplar. Feature values
+could take any one of 50 evenly-spaced values between the specified boundaries. Previously generated exemplars were not
+allowed to be generated a second time. Participants were shown an on-screen preview of their exemplar on each trial as
+they interacted with the sliders, but could not see previously generated exemplars or exemplars from the Alpha category.
 
-The next phase comprised a series of generation trials.
-Depending on their generation condition, participants were asked to generate either eight exemplars from a category that was 'Not-Alpha' (Not-Alpha generation condition), eight exemplars from a category called 'Beta' (Beta-Only generation condition), or four exemplars from a category 'Beta' \emph{and}
-four exemplars from a third category 'Gamma' (Beta-and-Gamma generation condition).
-Exemplars were generated once per trial using two on-screen sliding scales, with each scale controlling the individual features (color and size) of the generated exemplar.
-Feature values could take any one of 50 evenly-spaced values between the specified boundaries.
-Previously generated exemplars were not allowed to be generated a second time.
-Participants were shown an on-screen preview of their exemplar on each trial as they interacted with the sliders, but could not see previously generated exemplars or exemplars from the Alpha category.
+\begin{figure*}%[H]
+\begin{center}
+    \includegraphics[width=\textwidth]{Figures/screens_gen.png}
+\end{center}
+\caption{Instructions and trials during the generation phase, observed by a participant in the Not-Alpha condition.
+  Participants in the Beta-Only and Beta-Gamma conditions experienced similar trials, with the exception that those in
+  the Beta-Gamma condition were asked to generate 4 Beta exemplars then 4 Gamma exemplars.}
+\label{fig:screens_gen}
+\end{figure*}
+
+% \begin{figure*}%[H]
+% \begin{center}
+%     \includegraphics[width=\textwidth]{Figures/screens_gen.png}
+% \end{center}
+% \caption{Experimental trials observed by a participant in the Not-Alpha condition. Participants in the Beta-Only and
+%   Beta-Gamma conditions experienced similar trials, with the exception that those in the Beta-Gamma condition were asked
+%   to generate 4 Beta exemplars then 4 Gamma exemplars.}
+% \label{fig:profile}
+% \end{figure*}
 
@@
 
-We analyze our data in two stages. First, to provide a coarse overview of the distribution of different patterns of generated categories, we classify the generated categories into
-six different profiles: Positives, where the correlation between the dimensions
-is more than $r$; Negatives, where the correlation between the dimensions is less
-than $-r$; Rows, where the range of values across the x dimension is at least
-five times more than the range across the y dimension; Columns, where the range
-of y dimension values is at least five times more than the range of x dimension
-values; Clusters, where the ranges across both dimensions are less than $a$; and
-Dispersed, where the ranges across both dimensions are more than $a$.
-Next, we compare the generated categories on four key distributional measures: ranges of each feature, the feature correlations, and the area enclosed by the generated exemplars in the feature space (i.e., their convex hull).
-Differences along each of these statistics are performed using Bayesian $t$-tests \cite{rouder2009bayesian}, yielding Bayes factors ($BF_{01}$) which indicate evidence for the null hypothesis when $BF_{01} > 1$, with larger values indicating greater evidence for the null hypothesis.
-$BF_{01} < 1$ indicates evidence for the alternative hypothesis.
-Interpretations of the sizes of Bayes factors are guided by \citeA{jeffreys1961}.
+We analyze our data in two stages. First, to provide a coarse overview of the distribution of different patterns of
+generated categories, we classify the generated categories into six different profiles: Positives, where the correlation
+between the dimensions is more than $r$; Negatives, where the correlation between the dimensions is less than $-r$;
+Rows, where the range of values across the x dimension is at least $d$ times more than the range across the y dimension;
+Columns, where the range of y dimension values is at least $d$ times more than the range of x dimension values;
+Clusters, where the ranges across both dimensions are less than $a$; and Dispersed, where the ranges across both
+dimensions are more than $a$. Next, we compare the generated categories on four key distributional measures: ranges of
+each feature, the feature correlations, and the area enclosed by the generated exemplars in the feature space (i.e.,
+their convex hull). Differences along each of these statistics are performed using Bayesian $t$-tests
+\citep{rouder2009bayesian}, yielding Bayes factors ($BF_{01}$) which indicate evidence for the null hypothesis when
+$BF_{01} > 1$, with larger values indicating greater evidence for the null hypothesis. $BF_{01} < 1$ indicates evidence
+for the alternative hypothesis. Interpretations of the sizes of Bayes factors are guided by \cite{jeffreys1961}.
 
@@
 
-We found that setting $r = .7$ and $a = .25$ was useful in capturing the different profiles of generated categories. A representative sample of each profile is shown in Figure
-\ref{fig:samples} and the frequency plot of the different profiles is presented
-in Figure \ref{fig:profile}.
+We took a subset of the data and tuned each of the profiling parameters such that the profiles of this subset were
+adequately captured. Subsequently, we applied this profiling scheme to the entire dataset. Overall, we found that
+setting $r = .7$, $a = .25$, and $d = 5$ was useful in capturing the different profiles of generated categories. The
+results were robust to moderate variations in the profiling parameters (e.g., setting $.5 < r < .9$, $.1 < a < .4$, and
+$ d > 1 $ returned very similar results.) A representative sample of each profile is shown in Figure \ref{fig:samples}
+and the frequency plot of the different profiles is presented in Figure \ref{fig:profile}.
 
@@
 The Row conditions produced categories that are high on the x dimension but not the y dimension.
-These observations indicate, as was observed in \citeA{austerweil2018catgen}, that the distribution statistics were carried over from the known category to the generated category in these two Alpha conditions.
+These observations indicate that the distributional statistics were carried over from the known category to the generated category in these two Alpha conditions.
 However, the Diagonal conditions did not have similar distributional statistics observed -- instead of a positive correlation, the Diagonal conditions tended to produce large, dispersed categories.
-\citeA{austerweil2018catgen} also observed a similar effect (although they found more evidence for a negative correlation).
+\cite{austerweil2018catgen} also observed a similar effect (although they found more evidence for a negative correlation).
 Overall, in alignment with previous research, at least two out of the three Alpha conditions in our experiment generated categories that share distributional statistics as the known category.
 
-More interestingly, when the data is broken down by the generation conditions (Figure \ref{fig:stats_instr}), we found that compared to the Not-Alpha conditions, there is moderate evidence showing Beta-Only conditions with a lower y dimension range ($t(162) = 3.00, BF_{01} = 0.16$) and moderate to strong evidence that their generated categories are smaller in area ($t(162) = 3.16, BF_{01} = 0.10$).
-There is moderate evidence that the Not-Alpha and Beta conditions share equal range of x dimension values ($t(162) = 1.33, BF_{01} = 4.82$).
+More interestingly, when the data is broken down by the generation conditions (Figure \ref{fig:stats_instr}), we found that compared to the Not-Alpha conditions, there is moderate evidence showing Beta-Only conditions with a lower y dimension range ($t(162) = 3.00$, $BF_{01} = 0.16$) and moderate to strong evidence that their generated categories are smaller in area ($t(162) = 3.16$, $BF_{01} = 0.10$).
+There is moderate evidence that the Not-Alpha and Beta conditions share equal range of x dimension values ($t(162) = 1.33$, $BF_{01} = 4.82$).
 With Beta-and-Gamma conditions, we find greater evidence for smaller and tighter categories compared to the Not-Alpha conditions.
-Specifically, there is very strong evidence that both Beta and Gamma categories from the Beta-and-Gamma condition are smaller in both x ($t(156) = 4.83, BF_{01} = 2.55 \times 10^{-4}$;$t(156) = 4.21, BF_{01} = 2.94 \times 10^{-3}$, respectively) and y ($t(156) = 4.57, BF_{01} = 7.40 \times 10^{-4}$;$t(156) = 7.56, BF_{01} = 4.30 \times 10^{-10}$, respectively) ranges compared to the Not-Alpha conditions.
-Similarly, there is very strong evidence that both categories in the Beta-and-Gamma conditions are smaller in area than the Not-Alpha conditions ($t(156) = 5.70, BF_{01} = 5.41 \times 10^{-6}$;$t(156) = 7.69, BF_{01} = 2.07 \times 10^{-10}$, respectively).
+Specifically, there is very strong evidence that both Beta and Gamma categories from the Beta-and-Gamma condition are smaller in both x ($t(156) = 4.83$, $BF_{01} = 2.55 \times 10^{-4}$; $t(156) = 4.21$, $BF_{01} = 2.94 \times 10^{-3}$, respectively) and y ($t(156) = 4.57$, $BF_{01} = 7.40 \times 10^{-4}$; $t(156) = 7.56$, $BF_{01} = 4.30 \times 10^{-10}$, respectively) ranges compared to the Not-Alpha conditions.
+Similarly, there is very strong evidence that both categories in the Beta-and-Gamma conditions are smaller in area than the Not-Alpha conditions ($t(156) = 5.70$, $BF_{01} = 5.41 \times 10^{-6}$; $t(156) = 7.69$, $BF_{01} = 2.07 \times 10^{-10}$, respectively).
 Overall, when comparing the Not-Alpha conditions to categories from other generation conditions, we consistently find moderate to very strong evidence that Not-Alpha categories are more widely dispersed (in their range values) and also larger (in their area), with the only exception being the comparison of x dimension ranges between the Not-Alpha and Beta-Only conditions.
@@
 When comparing the distributions of the Beta and Gamma categories generated within the Beta-and-Gamma conditions, we find an overall weak-to-moderate evidence for equal distributional statistics.
-Specifically, there was moderate evidence for equal x dimension ranges ($t(150) = 0.56, BF_{01} = 9.47$) and weak evidence for both feature correlations and area size ($t(150) = 1.85, BF_{01} = 2.10$;$t(150) = 1.77, BF_{01} = 2.40$, respectively).
-When measured on their y dimension ranges, we found weak evidence for lower values from Gamma categories compared to the Beta categories ($t(150) = 2.47, BF_{01} = 0.59$).
+Specifically, there was moderate evidence for equal x dimension ranges ($t(150) = 0.56$, $BF_{01} = 9.47$) and weak evidence for both feature correlations and area size ($t(150) = 1.85, BF_{01} = 2.10$; $t(150) = 1.77$, $BF_{01} = 2.40$, respectively).
+When measured on their y dimension ranges, we found weak evidence for lower values from Gamma categories compared to the Beta categories ($t(150) = 2.47$, $BF_{01} = 0.59$).
 
@@
 \section{Discussion}
-At first glance, it seems reasonable to assume that generating a new category X after learning category Y is the same as generating category Not-X.
-An independently identified category should already be a category-by-negation (in that an independently identified category is not what was known).
-Further, if the categories are already identified by arbitrary labels, then it may be easy to assume that identifying the negation of a known category cannot add any additional information in category generation.
+At first glance, it seems reasonable to assume that generating a new category Y after learning category X is the same as generating a new category Not-X.
+An independently identified category should already be a category-by-negation (in that an independently identified category is not what was previously known.) Further, if the categories are already identified by arbitrary labels, then it may be easy to assume that identifying the negation of a known category cannot add any additional information in category generation.
 
@@
 Aside from demonstrating a new effect, this study has continued to show the robustness of the distributional similarities between learned and generated categories.
-In this sense, the results from the different Alpha conditions are similar to those observed in \citeA{austerweil2018catgen}.
+In this sense, the results from the different Alpha conditions are similar to those observed in \cite{austerweil2018catgen}.
 The generated categories from the Cluster condition tend to possess lower x and y ranges, with a correspondingly smaller area, and the generated categories from the Row condition tend to adopt Row-type profiles.
@@
 
-However, one notable difference is that while \citeA{austerweil2018catgen} found evidence of negatively-correlated generated categories in their XOR condition, we found no evidence of negatively-correlated generated categories in our comparable Diagonal condition.
-\citeA{austerweil2018catgen} explained that the presence of negatively-correlated features in this condition was indicated the effect of category contrast -- that is, categories with negatively correlated features are generated because they are particularly different to categories with positively correlated features.
-It is possible that due to the reduced strength of the positive correlation in our study compared to \citeA{austerweil2018catgen} (because of the addition of noise to the feature values), participants were no longer as sensitive to the negative correlations in the experimenter-defined category and therefore started to produce uncorrelated but widely dispersed categories.
+However, one notable difference is that while \cite{austerweil2018catgen} found evidence of negatively-correlated
+generated categories in their XOR condition, we found no evidence of negatively-correlated generated categories in our
+comparable Diagonal condition. \cite{austerweil2018catgen} explained that the presence of negatively-correlated features
+in this condition was indicative of the effect of category contrast -- that is, categories with negatively correlated
+features are generated because they are particularly different to categories with positively correlated features. It is
+possible that due to the reduced strength of the positive correlation in our study compared to
+\cite{austerweil2018catgen} (because of the addition of noise to the feature values), participants were no longer as
+sensitive to the negative correlations in the experimenter-defined category and therefore started to produce
+uncorrelated but widely dispersed categories.
 
-Beyond replicating previous studies, the current study has demonstrated that the consistency in distributional statistics can persist beyond the first generated category.
-However, evidence showing this was ultimately weak.
-One possible reason for this is the relatively small feature space employed in the tasks.
-The generation of a second novel category is necessarily more constrained in the feature space than the generation of the first novel category, possibly resulting in differences in distributional structure.
-By exploring stimuli with less defined boundaries (e.g., orientation), we may expect to see greater consistency in distributional structure over multiple generated categories.
+Beyond replicating previous studies, the current study has demonstrated that the consistency in distributional
+statistics can persist beyond the first generated category. However, evidence showing this was ultimately weak. One
+possible reason for this is the relatively small feature space employed in the tasks. The generation of a second novel
+category is necessarily more constrained in the feature space than the generation of the first novel category, possibly
+contributing to differences in distributional structure. By exploring stimuli features with less defined boundaries
+(e.g., orientation), we may expect to see greater consistency in distributional structure over multiple generated
+categories.
 
@@
 It would be worth investigating how participants might proceed to generate additional categories-by-negation (e.g., by asking observers to generate a category that is Not-Alpha and Not Beta).
-Packing Theory \cite{hidaka2011packing} -- a hypothesis that suggests categories can be neatly 'packed' into the feature space -- may indicate that successive categories-by-negation are generated in a fashion that preferentially occupies spaces between observed categories.
+Packing Theory \citep{hidaka2011packing} -- a hypothesis that suggests categories can be neatly `packed' into the feature space -- may indicate that successive categories-by-negation are generated in a fashion that preferentially occupies spaces between observed categories.
 Further, although none of the current models of category generation can directly account for the effects observed in this study, they may be useful components in a larger category generation framework.
-For instance, future work may consider implementing the hierarchical sampling model from \citeA{jern2013probabilistic} in a framework of overhypotheses \cite{kemp07}, where a prior can be placed over a category identity space, allowing models to behave differently under different regions of generated category identity.
+For instance, future work may consider implementing the hierarchical sampling model from \cite{jern2013probabilistic} in a framework of overhypotheses \citep{kemp07}, where a prior can be placed over a category identity space, allowing models to behave differently under different regions of generated category identity.
 
-Ultimately, the nature of newly generated categories appears to vary depending on the identity they were associated with.
-The extent to which they may differ, and the mechanisms driving these differences represent fascinating areas for future research.
+Ultimately, the nature of newly generated categories appears to vary depending on the identity they were associated
+with. The extent to which they may differ, and the mechanisms driving these differences represent fascinating areas for
+future research.
 
-% \section{Acknowledgments}
+\section{Acknowledgments}
+This work was funded by a Vilas Life Cycle Professorship and the VCRGE at University of Wisconsin-Madison with funding from the WARF
 

### 2019-05-09 Experiments/cogpsych-code/compile-data-5con.py

@@ -44,7 +44,6 @@ for di,databases in enumerate(databases_Set):
         with sqlite3.connect(dbpath) as con:
             data = dict((T, pd.read_sql('SELECT * FROM ' + T, con)) for T in keep_tables)
 
-
         # get the stimuli df, init the alphas
         if num == 0: 
             stimuli = data['stimuli']
@@ -81,8 +80,8 @@ for di,databases in enumerate(databases_Set):
             generation   = data['generation']
         else:
             participants = pd.concat([participants,data['participants']], ignore_index = True)
-            betastats    = pd.concat([betastats,data['betastats']], ignore_index = True)
-            generation   = pd.concat([generation,data['generation']], ignore_index = True)
+            betastats    = pd.concat([betastats,data['betastats']], ignore_index = True,sort=True)
+            generation   = pd.concat([generation,data['generation']], ignore_index = True,sort=True)
 
     # convert experiment num to integer
     experiments['experiment'] = experiments['experiment'].astype(int)

### 2019-05-09 Experiments/xor-cluster-rowABC/analysis/plot.individuals.py

@@ -9,12 +9,14 @@ import Modules.Funcs as funcs
 
 pd.set_option('precision', 2)
 
+makeplots = True #Do plots?
+
 # import data
 con = sqlite3.connect('../data/experiment.db')
 info = pd.read_sql_query("SELECT * from participants", con)
 df = pd.read_sql_query("SELECT * from generation", con)
 alphas = pd.read_sql_query("SELECT * from alphas", con)
-stimuli = pd.read_sql_query("SELECT * from stimuli", con).as_matrix()
+stimuli = pd.read_sql_query("SELECT * from stimuli", con).values
 stats = pd.read_sql_query("SELECT * from betastats", con)
 con.close()
 
@@ -22,18 +24,21 @@ savedir = 'individuals'
 gentypeStr = ['N','B','C'] #not alpha, only beta, beta-gamma
 gentypeStrDisp = ['A\'','B','C'] #not alpha, only beta, beta-gamma
 gentypeCols = [[.3,0,.5],[0,0,.5],[0,.5,0]]
-f, ax= plt.subplots(1,1, figsize=(1.6, 1.6))
+
 
 var_thresh = .25
 cor_thresh = .7
+drat = 5 #dimension ratio for col vs row
 groupcount = {'Positive':0,'Negative':0,'Cluster':0,'Dispersed':0,'Row':0,'Column':0}
 data_arr = []
+
+
 for i, row in info.iterrows():
     pid, condition, gentype = int(row.participant), row.condition, row.gentype
     gentypeStr_p = gentypeStr[gentype]
 
     palphas = alphas[condition]
-    #pbetas_all = df.stimulus[df.participant == pid]
+    pbetas_all = df.stimulus[df.participant == pid]
     if gentype==2:
         pdf = df.loc[df.participant==pid]
         betastr = [gentypeStrDisp[1] if pdf_row.category=='Beta' else gentypeStrDisp[2] for ii,pdf_row in pdf.iterrows() ]
@@ -54,9 +59,9 @@ for i, row in info.iterrows():
         cov  = stats.correlation[stats.participant==pid].values[0]
         group = ''
 
-        if varx > vary*5:
+        if varx > vary*drat:
             group = 'Row'
-        elif vary > varx*5:
+        elif vary > varx*drat:
             group = 'Column'
         elif cov > cor_thresh:
             group = 'Positive'
@@ -68,17 +73,20 @@ for i, row in info.iterrows():
[…]

### 2019-05-09 Modules/Classes/Model.py

@@ -281,7 +281,8 @@ class Exemplar(Model):
         c = None,
         p = 1,
         wrap_ax = None,
-        ax_max = 2):
+                        ax_range = 2,
+                        ax_step = .25):
         """ 
         function to compute summed similarity along rows of 
         X across all items in Y. Resulting array will have one element 
@@ -295,16 +296,18 @@ class Exemplar(Model):
 
         bdlass_ax indicates which axis is boundless (i.e., take the shortest of either the distance or max-distance)
 
-        ax_max is the limits of both axes
+        ax_range is the limits of both axes
+
+        ax_step is the minimum distance of each step on the axis (relevant only when wrapping)
         """
 
         # set weights and c
         if wts is None: wts = self.wts
         if c is None: c = self.specificity
         if p == 1:
-            distance   = Funcs.pdist(np.atleast_2d(X), np.atleast_2d(Y), w = wts, wrap_ax = wrap_ax, ax_max = ax_max)
+            distance   = Funcs.pdist(np.atleast_2d(X), np.atleast_2d(Y), w = wts, wrap_ax = wrap_ax, ax_range = ax_range, ax_step=ax_step)
         else:
-            distance   = Funcs.pdist_gen(np.atleast_2d(X), np.atleast_2d(Y), w = wts, p = p, wrap_ax = wrap_ax, ax_max = ax_max)
+            distance   = Funcs.pdist_gen(np.atleast_2d(X), np.atleast_2d(Y), w = wts, p = p, wrap_ax = wrap_ax, ax_range = ax_range, ax_step = ax_step)
         similarity = np.exp(-float(c) * distance)
         similarity = similarity * float(param)
         return np.sum(similarity, axis = 1)
@@ -335,7 +338,9 @@ class HierSamp(Model):
             ax = int(ax)
             #Keep taking one set of stimuli above and below the perceived space until ratio is large enough
             stim_ax = stimuli[:,ax]
-            stim_diff = np.max(stim_ax) - np.min(stim_ax)
+            stim_unique = np.unique(stim_ax)
+            stim_step = np.abs(stim_unique[0]-stim_unique[1])
+            stim_diff = np.max(stim_ax) - np.min(stim_ax) + stim_step
             stim_up = stimuli.copy()
             stim_dn = stimuli.copy()
             #Note that this isn't complete if we are wrapping more than 1 axis
@@ -350,8 +355,9 @@ class HierSamp(Model):
                 density_ratio = maxden/minden
[…]

### 2019-05-09 Modules/Classes/Simulation.py

@@ -213,7 +213,7 @@ class Trialset(object):
         return self
                 
 
-    def loglike(self, params, model, fixedparams = None, whole_array=False, parmxform = True,seedrng = False, wrap_ax=None):
+    def loglike(self, params, model, fixedparams = None, whole_array=False, parmxform = True,seedrng = False):
         """
         Evaluate a model object's log-likelihood on the
         trial set based on the provided parameters.
@@ -254,12 +254,13 @@ class Trialset(object):
             
             #Iterate over axes wrappings
             for wrap in wraps:
+                #print(wrap)
                 # if it's an assignment task, also compute probabilities for other category (cat0)
                 # ps0 = np.zeros(ps1.shape)
                 if task == 'generate':
                     # compute probabilities of generating exemplar in cat 1
                     ps = model(categories, params, self.stimrange).get_generation_ps(self.stimuli, 1,self.task,seedrng = seedrng,wrap_ax=wrap)
-                    ps_idx = np.where(trial['wrap_ax']==wrap)[0]                    
+                    ps_idx = np.where(trial['wrap_ax']==wrap)[0]
                     #If it's a scalar, take it out of the array
                     if len(ps_idx)==1:
                         ps_idx = ps_idx[0]

### 2019-06-14 Manuscripts/cog-psych/post-revision/post-cogpsych.tex
Massive update with Corner condition writeup. Still rough but probably human readable.

--- 2018-09-19 Manuscripts/cog-psych/revision/cogpsych.tex
@@
 
-Creating and Learning Something Different: \\
+Creating Something Different: \\
 Similarity, Contrast, and Representativeness in Categorization
@@
 computational analyses demonstrate that both models capture different aspects of
-contrast's role in categorization. Our work also serves as a concrete example of
-how well-established categorization models can be applied beyond category learning.
+contrast's role in categorization.
 
@@
 
-% % OLD OLD INTRO:
-% Creativity, innovation, imagination, and the creation of new ideas are among the most fascinating, important and difficult human capabilities to study.
-They are fascinating and important capabilities because they have produced some of the greatest scientific breakthroughs that revolutionized the world.
-Unfortunately, technological and scientific breakthroughs are rare and the result of a great deal of cognitive effort, which makes it difficult to investigate the underlying cognitive representations and processes responsible for them using standard experimental methodology or to formalize them in computational models.
-Although we tend to focus on the most salient products of these processes (e.g., scientific breakthroughs), every person is likely to have generated many novel sentences, thoughts, and/or drawings given the combinatorial explosion of possible statements given a reasonably-sized set of primitive elements and combination rules in language, thought, and perception \citep{goldstone2003}.
-This observation provides a fantastic opportunity: By examining how people create new concepts in a domain amenable to formalizing in computational models, we can investigate (some aspects of) creativity, innovation, and imagination scientifically.
-
-% Towards the aim of providing a formal account of the cognitive mechanisms involved in creativity and innovation, we explore a form of creative cognition that can be studied using standard behavioral methodology and is amenable to formal modeling: category generation.
-In a category generation task, participants are given some background knowledge about a domain (a cover story and/or learning one or more categories in the domain) and then are asked to generate another category in the domain.
-Unlike creativity and innovation, categorization is well studied from a breadth of perspectives, including behavioral, neural, comparative, and computational  \citep{kurtz2015human,mack2013,margolis2015,pothoswills2011}.
-By analyzing a creative task using formal tools from the categorization literature, we can formalize some theories of creativity and test them empirically in a more rigorous manner.
-
-%cut from submission intro:
-
-%Yet, it is also one of the most difficult human capabilities to study.
-We tend to focus on the most salient products of conceptual generation (e.g., scientific breakthroughs), but every person is likely to have generated many novel concepts in ordinary life.
-Indeed, the generation of novel sentences, works of art, and so on may also be considered products of our capacity to create new concepts.
-However, while this capacity is of great interest, it is also highly complex, making it difficult to study in a typical laboratory setting.
-By consequence, we do not have a strong understanding of the representations and processes involved in the generation of new concepts.
-
-%Questions concerning the generation of novel concepts are typically addressed within research on creativity, a tremendously diverse field employing a multitude of experimental and theoretical approaches \citep[see][]{kaufman2010handbook}.
-The processes underlying conceptual generation are commonly investigated using the creative cognition approach \citep{finke1992creative, smith1995creative}, in which the products of creative acts (such as drawings) are analyzed to obtain insights into how they might have been created.
-These approaches have made great advances in our understanding of creativity, but the focus on particular types of creative products such as drawings makes it difficult to formalize the proposed processes and representations within a computational modeling framework.
-As a consequence, this approach does not often employ computational models to provide formal tests of competing theories.
-
-% ------- BEGIN! ------- %
 \begin{flushleft}
@@
 categories of objects that young and old adults learned and have become reliant
-on in the last few decades. In fact, computational researchers have identified
+on in the last few decades. Computational researchers have identified
 generating novel objects and objectives as one of the defining, and most
@@
 embodies the exemplar dissimilarity principle by incorporating a factor that
-repulses members of opposite categories from one another. To foreshadow, we find
-that this notion of contrast predicts human category learning performance, above
-and beyond other category learning models in a novel behavioral experiment. This
-demonstrates that contrast plays a central role in other important
-categorization tasks, and it is not specific to category generation.
+repulses members of opposite categories from one another.
 
@@
 qualitatively and quantitatively analyze the performance of each model in
-capturing human category generation. Afterwards, we find support for category
-contrast predicting the ease at which people learn a pair of categories in a
-novel empirical experiment. We conclude with a discussion of the implications of
-our results for categorization and directions for future
-work.
+capturing human category generation. A third experiment is performed to
+highlight the differences between the two contrast models. We conclude with a
+discussion of the implications of our results for categorization and directions
+for future work.
 
@@
 representativeness -- exemplars are generated such that they resemble the target
-category and thus more distinct from and less similar to other categories. We
+category and are thus more distinct from and less similar to other categories. We
 adopt the formalization given by \cite{tenenbaum01}, where the
@@
 representative of the underlying distribution for a target category. Generally,
-this mechanism of representativeness results in the generation of exemplars that
-are less similar to exemplars from other categories. We see this illustrated in
-Figure \ref{fig:representative-examples}, where the representativeness model
-displays a strong preference for generating exemplars from target category 'B'
-that are further away from the contrast category 'A'.
+with the assumption of unimodal underlying distributions, this mechanism of
+representativeness results in the generation of novel exemplars that are less
+similar to exemplars from the known category. This occurs because the
+representativeness of novel exemplars tend to be highest where the probability
+density of the known category's underlying distribution is the lowest. We see
+this illustrated in Figure \ref{fig:representative-examples}, where the
+representativeness model displays a strong preference for generating exemplars
+from target category 'B' that are further away from the contrast category 'A'.
+Exceptions to this pattern can occur (i.e., the generation of novel exemplars
+similar to known category exemplars) when the known category distribution is
+especially large on all features -- we explore this in more detail in Section
+\ref{section:diffcon} where we highlight an important condition demonstrating
+how PACKER and the representativeness model can make qualitatively different
+predictions.
 
@@
 Bayesian and copy-and-tweak models developed by \cite{jern2013probabilistic}.
-After analyzing the category generation studies, we present third experiment to
-demonstrate the effect of contrast in category learning. This ensures that our
-results are not idiosyncratic to category generation.
+After analyzing the category generation studies, we present a third experiment to
+distinguish between our two conceptualizations of contrast.
 
@@
 
-%Experiment 3 starts here
-\section{Experiment 3}
-
-Experiments 1 and 2 clearly establish the importance of contrast in category
+%Contrast model distinction
+\section{Experiment 3: Differentiating Contrast}
+\label{section:diffcon}
+Experiments 1 and 2 clearly established the importance of contrast in category
 generation. Follow-up model-based analyses illustrated that both contrast models
 account for participant performance better than models that do not take into
-account contrast. We also found more support for contrast as representativeness
-hypothesis over contrast as exemplar dissimilarity hypothesis.
-
-Although category generation is an intriguing aspect of human cognition, it has
-played a relatively limited role in the broader categorization literature. To
-establish that contrast plays a broader role in categorization, we conducted a
-traditional category learning study. In it, participants learned two categories,
-where the categories learned by each participant in Experiment 3 was yoked to
-the `Alpha' shown and `Beta' generated by each participant in Experiment 2. Our goal
-is to investigate whether model performance (in generating categories) is
-consistent with human performance (in learning those categories). Specifically, if contrast plays a broader role in categorization, then `Beta' categories rated as "better" according
-to the contrast models should be easier for participants to learn.
-%\footnote{It is possible to adjust the present models of category generation such that they can make category learning predictions.
-However, such an adjustment makes the strong assumption that the very same processes governing category generation also apply to category learning.
-Taking this approach would fall outside the scope of the current investigation.}
+account contrast. We also found more support for the hypothesis of contrast as
+representativeness versus contrast as exemplar dissimilarity across the
+different Alpha conditions.
+
+While the representativeness model fits better than PACKER, and in spite of
+their fundamentally different structures, both contrast models seem to make
+qualitatively similar predictions. To illustrate this, Figure
+\ref{fig:firstbetas} shows how the modal generating probabilities for the first
+Beta exemplar are similar across both contrast models.\footnote{Model parameters
+  are set to the optimized values from Table \ref{table:global-model-fits}.} The
+highest density regions for all models appear to be near the corners and edges.
+This is particularly the case for the representativeness model, which almost
+exclusively predicts generation at the corners, while PACKER can generate the
+first novel exemplars toward the inner regions of the stimulus space in the XOR
+condition.
+
+\begin{figure}[p]
+    \begin{center} 
+      \includegraphics[width=\textwidth]{figs/firstbetas.pdf}
+      \caption{Contrast model predictions indicating probability of generating
+        the first Beta exemplar for each Alpha condition. Darker shades of
+        blue indicate higher probabilities (normalized within each plot.)}
+      \label{fig:firstbetas}
+    \end{center}
+\end{figure}
+
+These similarities are perhaps unsurprising -- each of the modal generation
+regions are maximally distant from the Alpha exemplars. However, recall that
+unlike PACKER, the representativeness model does not explicitly use a
+distance-based measure in determining exemplar generation probabilities. The
+general pattern of producing exemplars dissimilar from the contrast category is
+the result of the underlying unimodal category structure assumed by the model
+and is not strictly constrained by the distances between exemplars.
+Consequently, it is possible to design a configuration of Alpha exemplars where
+the representativeness model would predict the generation of novel exemplars
+that are \emph{more similar} to the learned exemplars.
+
+In this section, we focus on one particularly useful category type -- the Corner
+category -- to distinguish between the two competing accounts of contrast in
+category generation. Here, each Alpha exemplar is located at the extreme corners
+of the feature space (see Figure \ref{fig:e3-squares}a). We first seek to
+identify what predictions each contrast model generates under this condition,
+followed by an category generation task, and a brief modeling exercise to
+determine which model is a better predictor of the behavioral data.
+
+In the Corner condition, the representativeness model assumes an underlying
+Alpha category distribution that is centred on the feature space, with
+probability densities highest in the centre and lowest in the corners. Since the
+representativeness of novel Beta exemplars generally increase when the
+likelihood of the underlying Alpha category distribution decreases, the regions
+of highest Beta representativeness are consequently located at the corners where
+these Alpha likelihoods are lowest (Figure \ref{fig:e3-squares}b.)
+
+Conversely, PACKER predicts a completely different result. With $\theta_c > 0$,
+the probability of generating a Beta exemplar increases with the dissimilarity
+from all Alpha exemplars. This results in highest Beta generation probabilities
+in the inner regions of the feature space, where the candidate exemplars are
+maximally different from the Alpha exemplars (Figure \ref{fig:e3-squares}c.)
+
+\begin{figure}[p]
+    \begin{center} 
+      \includegraphics[width=.8\textwidth]{figs/e3-squares.pdf}
+      \caption{Alpha exemplars for the Corner condition (a) and the predictions
+        from contrast models PACKER (b) and the representativeness model (c)
+        with square stimuli.}
+      \label{fig:e3-squares}
+    \end{center}
+\end{figure}
+
+[Transition to and justification for this still feels a little shaky]
+
+A large part of the representativeness model's predictive accuracy in
+Experiments 1 and 2 is due to its preference for generating categories located
+at the boundaries of the stimulus features. If at least one of these boundaries
+is removed by utilizing an unbounded feature (i.e., no absolute maximum or
+minimum exists in the stimulus space along this dimension), we should expect to
+see a decrease in the predictive performance of the representativeness model. To
+implement this, in our next experiment we also include a new set of stimuli:
+circles that vary in size (a bounded feature) and orientation (an
+unbounded feature; Figure \ref{fig:e3-circles}a). Before describing the
+details of this experiment, we will briefly explore the contrast model
+predictions with stimuli that comprise one bounded and one unbounded feature.
+
+\begin{figure}[p]
+    \begin{center} 
+      \includegraphics[width=.8\textwidth]{figs/e3-circles.pdf}
+      \caption{Circle stimulus domain (a) and predictions from contrast models
+        PACKER (b) and the representativeness model (c). The unbounded feature
+        (orientation) is represented on the y-axis, while the bounded feature
+        (size) is represented on a x-axis. As with the square stimuli, feature
+        and direction assignment were counterbalanced across participants (in
+        the displayed example, circles are rotated clockwise up the y-axis.) }
+      \label{fig:e3-circles}
+    \end{center}
+\end{figure}
+
+Compared to stimuli with two bounded features, PACKER makes a similar prediction
+in generating the first Beta exemplar when one feature is unbounded, with the
+curious exception that the highest density regions appear compressed along the
+y-axis. Due to the unbounded nature of orientation, the maximum distance from
+the Alpha exemplars is half of the full range of the feature -- for an Alpha
+exemplar oriented at 0 radians, the furthest another exemplar can be situated is
+$\pi$ radians within the full range of orientation is $2\pi$ radians.
+
+For the representativeness model, the single unbounded feature essentially
+prevents the formation of edge effects along that feature. Consequently, this
+results in higher probabilities of the generating the first Beta exemplars along
+the edges of the bounded feature (Figure \ref{fig:e3-circles}c). Since this
+model does not explicitly recognize similarities as distances along the feature
+space (unlike PACKER), it does not experience the compression in density regions
+along the unbounded feature axis. Intuitively, this suggests that the
+representativeness model (which prefers generating exemplars at the edges) will
+be poorer at generating stimuli with an unbounded feature compared to
+stimuli with all features bounded.
+
+% We have identified the conditions under which the contrast models may make
+% differing predictions, however this distinction is not very useful without any
+% empirical support. Data from the next experiment will most certainly support one
+% model or the other
+
+[Maybe work on transition if there's time]
+
 \subsection{Participants, Materials, and Procedure}
 
-Experiment 2 recruited 122 participants who each generated one set of four `Alpha'
-and four `Beta' category exemplars. Experiment 3 recruited the same number of
-participants.
-
-Participants observed four blocks of eight trials. Each trial began with the
-presentation of a fixation cross for 500 ms. This was followed by the
-presentation of one exemplar randomly sampled without replacement from the
-unique category set. Participants were tasked with assigning the presented
-exemplar to either the `Alpha' or `Beta' category with no time limit imposed.
-Feedback was automatically displayed for 2500 ms after each response.
-
-\subsection{Results and Analysis}
-
-Among the 122 originally generated category sets were 102 unique sets (i.e.,
-they contained a unique collection of `Alpha' and `Beta' exemplars). Consequently,
-for our subsequent analyses for this experiment we only use data from the first
-102 participants that were presented with a unique category set.
-
-Overall accuracy of the participants was high, with a mean error rate of $.19$
-($SD = .19$). Aggregated error rates for each block are presented in Figure
-\ref{fig:learningcurve}.
+We recruited 89 participants from Amazon Mechanical Turk for this experiment, 46
+of whom were randomly assigned square stimuli and the remaining 43 assigned
+circle stimuli. Square stimuli were exactly the same as in Experiments 1 and 2.
+Circle stimuli were made up of unfilled circles with a radial line extending
+from its centre to the edge. These stimuli varied along two features -- their
+diameter (or size) ranging from 3.0--5.8 cm, and their orientation ranging from
+0--$2\pi$ radians. The orientation of each circle stimuli was fixed at a
+randomly selected angle for all participants. Similar to Experiments 1 and 2,
+participants first observed a training phase where four Alpha exemplars were
+presented across three blocks followed by the generation phase. All participants
+in this experiment were shown exemplars from the Corner category type.
+
+\subsection{Results and Discussion}
 
@@
     \begin{center}
-    \includegraphics[width=\textwidth/2]{figs/e3-learningcurve.pdf}
-    \caption{Average error rate for each successive block. Green discontinuous
-line represents the overall mean error rate.}
-    \label{fig:learningcurve}
+      \includegraphics[width=\textwidth]{figs/e3-statsboxes.pdf}
+    \caption{Box-plots of the distributional statistics from the categories
+generated in Experiment 3. Boxes depict the median and quartiles of each
+condition, with whiskers placed at 1.5 IQR. All points outside this region are
+marked individually. For circle stimuli, the Y Range represents orientation and
+the X Range represents size.}
+    \label{fig:e3-statsboxes}
     \end{center}
@@
 
-In order to analyze the consistency between model and human performance, we
-optimized each model such that the Pearson correlation between the model's
-negative log-likelihood of generating each unique category set and the
-participant's error in learning that category set is maximized. As with the
-analysis in Section \ref{section:individual-diff}, we simulated the weighting of
-each dimension independently depending on the generated categories.
-
-The greatest correlations were observed when fitting the contrast models.
-Specifically, PACKER correlated with human performance with
-$r = .65, 95\%\,CI\, [.48,.78]$, and the representativeness
-model correlated with human performance with $r = .59, 95\%\,CI\,[.40, .73]$\footnote{Confidence intervals were obtained using a bootstrap
-  method with 1000 bootstrap samples.}.
-Copy-and-tweak yielded a correlation of $r = .45, 95\%\,CI\,[.28, .62]$ and the
-hierarchical Bayesian model yielded a similar correlation of
-$r = .46, 95\%\,CI\, [.30, .62]$. These results are graphically presented in Figure
-\ref{fig:perror_corr}.
-
-\begin{figure}
-    \begin{center}
-    \includegraphics[width=\textwidth]{figs/modelvspptp.pdf}
-    \caption{Scatterplots indicating the correlations between observed
-      participant error and model fit. Blue discontinuous line are best-fit
-      lines. Green discontinuous lines indicate boundaries generated by the 95\%
-      CI around the fitted correlations. }
-    \label{fig:perror_corr}
-    \end{center}
-\end{figure}
-
-To emphasize the strong influence of contrast in maximizing the association
-between category generation and category learning, we computed the correlations
-over wide range of $\theta_{contrast}$ values, with the other parameter values
-of PACKER held constant at their optimized levels. As presented in Figure
-\ref{fig:packer-corr}, the correlation quickly increases with increasing weight
-on contrast, reaching a plateau above values of around 4.0.
-
-\begin{figure}
-    \begin{center}
-    \includegraphics[width=0.8\textwidth]{figs/packer-corr.pdf}
-    \caption{Correlation between PACKER's fit and participant error as a
-function of the $\theta_{contrast}$ parameter. To facilitate comparison,
-PACKER's other parameters ($c$, $\theta_{target}$) were set to the best fitting
-values obtained for copy-and-tweak in Table \ref{table:global-model-fits}. Grey
-dashed line represents the correlation between Copy \& Tweak's fit and
-participant error.}
-    \label{fig:packer-corr}
-    \end{center}
-\end{figure}
+Overall, we found no significant differences between the distributional statistics of each
+generated category between square and circle stimuli (Figure
+\ref{fig:e3-statsboxes}. The only exception is seen in the Y range (representing
+the orientation of circle stimuli,) which is significantly lower for the circle
+stimuli compared to the square stimuli ($t(87) = 4.94$, $p = .001$). This
+provides preliminary support for PACKER, which predicts generating exemplars
+within a shorter range along the unbounded feature compared to a bounded
+feature.
+
+At a rough level (see Table \ref{table:corner_counts}, 35\% of participants who
+observed square stimuli tended to generate their first Beta exemplars along the
+edges of the feature space. In comparison, only 12\% of participants who
+observed circle stimuli generated their first Beta exemplars along the edges of
+the feature space. These proportions more closely match the predictions of
+PACKER (33\% probability of generating first Beta exemplars at the edges for
+squares, and 24\% probability at edges for circles) compared to the
+representativeness model (53\% of the former, 49\% probability of the latter).
+In addition, these results agree with our prediction that the representativeness
+model would be poorer in predicting data from circle stimuli (26\% of
+participants observing circle stimuli) compared to square stimuli (39\%
+participants observing square stimuli). [No proper stats here... feels a little
+dirty. But I suppose there are model fits coming up.]
+
+Quantitative model fits were also obtained by optimizing both contrast models to
+the entire dataset including data from Experiments 1 and 2. Fit statistics and
+parameter values are presented in Table \ref{table:corner-model-fits}. Overall,
+the representativeness model still emerges as the best-fitting model, although
+by a much smaller lead compared to fits to only Experiments 1 and
+2. At an individual level, 67\% of participants in Experiment 3 were better
+fit by PACKER than the representativeness model.
+
+\begin{table} \centering
+  \caption{Frequencies of individual participants best fit by each contrast
+    model  broken down by stimulus type.}
+\label{table:corner_counts}
+\begin{tabular}{ l l l l} \\ Stimulus Type & \textbf{PACKER} &
+                                                               \textbf{Representativeness}& Total \\
+  \hline 
+  Squares & 28 & 18 & 46\\
+  Circles & 32 & 11 & 43\\
+  Total   & 60 & 29 & 89\\
+\end{tabular}
+\end{table}
+\begin{table} \centering
+  \caption{Results of model-fitting to the combined datasets from Experiments 1,
+    2 and 3. Smaller AIC values correspond to better model fits (adjusted
+    for number of parameters.)}
+\label{table:corner-model-fits}
+
+\begin{tabular}{ l l} \\ \textbf{PACKER} & \textbf{Representativeness} \\
+  \hline $AIC = 12052$ & $AIC = 11991$ \\ 
+  $L = -6023$          & $L = -5991$ \\ 
+  $c = 0.60$           & $\kappa = 5.91$  \\ 
+  $\theta_c = 2.78$    & $\nu = 1.00$ \\
+  $\theta_t = 3.17$    & $\lambda = 1.45$ \\
+                       &$\theta = 1.86$ \\
+\end{tabular}
+\end{table}
+
+Ultimately, these results suggest that PACKER's exemplar dissimilarity account
+of contrast, as opposed to the representativeness account, is more appropriate
+for the Corner category type. On almost every measure of analysis in this
+section, PACKER outperformed the representativeness model. Although the
+representativeness model best fit the data from Experiments 1 and 2, it
+struggled when the learned category occupied the extreme corners of the stimulus
+space -- both in cases where the stimuli possess bounded and unbounded features.
+
+[Work on transition time permitting]
 
 \section{General Discussion}
-The sensory impression of every stimulus and event is unique.
-Grouping distinct patterns of sensory information into categories is a fundamental task solved by the mind.
-Most work has focused on how people learn new categories that are provided to them through unlabeled, partially labeled, or fully labeled examples.
-How were these categories first determined?
-Some natural categories are likely to be the result of regularities in the dynamics of our environment.
-But, these are only a subset of the categories that people learn.
-Other categories, such as tools and ideas, were generated by people over time.
-What basic principles underlie how people generate categories?
+The sensory impression of every stimulus and event is unique. Grouping distinct
+patterns of sensory information into categories is a fundamental task solved by
+the mind. Most work has focused on how people learn new categories that are
+provided to them through unlabeled, partially labeled, or fully labeled
+examples. How were these categories first determined? Some natural categories
+are likely to be the result of regularities in the dynamics of our environment.
+But, these are only a subset of the categories that people learn. Other
+categories, such as tools and ideas, were generated by people over time. What
+basic principles underlie how people generate categories?
 
@@
 representativeness model. The exact nature of its superiority over the other
-models is still not yet clear. However, as we explore in a later section, it is
-likely that the representativeness mechanism itself (i.e., independent of the
-hierarchical Bayesian foundation) captures a core aspect of both category
-generation and category learning.
+models is still not yet clear, although we find evidence that it only barely
+surpasses PACKER when considering known categories with extreme exemplars.
+As we explore in a later section, it is likely that the
+representativeness mechanism itself (i.e., independent of the hierarchical
+Bayesian foundation) captures a core aspect of both category generation and
+category learning.
 
@@
 
-Although the focus of this article has been to address the role of contrast in category generation and learning, our findings and approach have relevant implications
+Although the focus of this article has been to address the role of contrast in category generation, our findings and approach have relevant implications
 for research in creative cognition. A central focus of the creative cognition
@@
 contrast in categorization. Specifically, while the representativeness model
-excelled in predicting our category generation data, PACKER was the best-fitting
-model in accounting for participant error in category learning. Given the
-history of their development, this is perhaps not entirely surprising. PACKER
-is implemented as an extension of the GCM, which was originally developed for (and
-has been highly successful in) category learning \citep{nosofsky1986attention,
-  nosofsky1994comparing}. Conversely, the representativeness mechanism was
-conceived as a heuristic describing the relative extent to which an individual
-observation is generated from some unobserved population
-\citep{kahneman1972subjective,abbott16}. This generative nature of the
-representativeness mechanism intuitively lends itself to the prediction of
-category generation data.
+excelled in predicting our overall category generation data, PACKER was the
+best-fitting model in accounting for categories generated from the Corner
+category type, where exemplars are located at the corners of the stimulus
+domain.
+
+% Given the history of their development, this is perhaps not entirely
+% surprising. PACKER is implemented as an extension of the GCM, which was
+% originally developed for (and has been highly successful in) category learning
+% \citep{nosofsky1986attention, nosofsky1994comparing}. Conversely, the
+% representativeness mechanism was conceived as a heuristic describing the
+% relative extent to which an individual observation is generated from some
+% unobserved population \citep{kahneman1972subjective,abbott16}. This generative
+% nature of the representativeness mechanism intuitively lends itself to the
+% prediction of category generation data.
 
@@
 Ultimately, the source of the discrepancy between the performance of the two
-contrast mechanisms in category generation versus category learning is still not
-entirely clear. Our results could be an indication of a fundamental difference
-between the processes involved where exemplar dissimilarity is a larger factor
-in category generation while representativeness exerts more influence in
-category learning. Alternatively, it is possible they are both approximating
-some factor that we still have not fully identified. While a thorough
-investigation is beyond the scope of the current article, the preliminary
-results of this section indicate that further work to disentangle these two
-mechanisms would be promising. In the following section, we explore some other
-limitations of the current work and provide more recommendations for future
-research.
+contrast mechanisms in category generation is still not entirely clear. Our
+results could be an indication of a fundamental difference between the processes
+involved where exemplar dissimilarity is a larger factor under certain category
+configurations while representativeness exerts more influence in others.
+Alternatively, it is possible they are both approximating some factor that we
+still have not fully identified. While a thorough investigation is beyond the
+scope of the current article, the preliminary results of this section indicate
+that further work to disentangle these two mechanisms would be promising. In the
+following section, we explore some other limitations of the current work and
+provide more recommendations for future research.
 
@@
 other conditions such as the nature of the instruction (``Generate a new
-category that is Not Alpha'' vs ``Generate a new category that is Beta'') and
-the boundedness of the features (using features such as hue and orientation that
-do have have strict minima or maxima) can affect the extent to which contrast
-influences categorization. Though our current contrast models can allow the
-influence of contrast to vary, they do not make any predictions regarding the
-specific conditions under which contrast can vary. Consequently,
-addressing this issue would be an immensely promising avenue for future
-research.
+category that is Not Alpha'' vs ``Generate a new category that is Beta'') 
+can affect the extent to which contrast influences categorization. Though our
+current contrast models can allow the influence of contrast to vary, they do not
+make any predictions regarding the specific conditions under which contrast can
+vary. Consequently, addressing this issue would be an immensely promising avenue
+for future research.
 

### 2019-06-14 Experiments/corner/analysis/plot.distance.figs.py

@@ -16,14 +16,14 @@ con = sqlite3.connect('../data/experiment.db')
 info = pd.read_sql_query("SELECT participant, condition, gentype from participants", con)
 generation = pd.read_sql_query("SELECT * from generation", con)
 alphas = pd.read_sql_query("SELECT * from alphas", con)
-stimuli = pd.read_sql_query("SELECT * from stimuli", con).as_matrix()
+stimuli = pd.read_sql_query("SELECT * from stimuli", con).values
 stats = pd.read_sql_query("SELECT * from betastats", con)
 con.close()
 
 stats = pd.merge(stats, info, on = 'participant')
 generation = pd.merge(generation, info, on = 'participant')
 
-gentypeStr = ['N','B','C'] #not alpha, only beta, beta-gamma
+#gentypeStr = ['N','B','C'] #not alpha, only beta, beta-gamma
 
 ngenerations = pd.DataFrame(dict(
     condition = [],
@@ -33,28 +33,28 @@ ngenerations = pd.DataFrame(dict(
 
 
 for c in pd.unique(info.condition):
-    for g in pd.unique(info.gentype):
-        gstr = gentypeStr[g]
-        for i in range(stimuli.shape[0]):
-            count = sum((generation.condition == c) & (generation.stimulus ==i) & (generation.gentype == g))
-            row = dict(condition = c+gstr, stimulus = i, count = count)
-            ngenerations = ngenerations.append(row, ignore_index = True)
+    for i in range(stimuli.shape[0]):
+        count = sum((generation.condition == c) & (generation.stimulus ==i))
+        row = dict(condition = c, stimulus = i, count = count)
+        ngenerations = ngenerations.append(row, ignore_index = True)
 
 
 # fh, ax = plt.subplots(1,2,figsize = (6,2.7))
 fh, ax = plt.subplots(2,1,figsize = (2.7,6))
 
 
-styles = dict(XORN = '-or', XORB = '-og', XORC = '-ob',
-              ClusterN = '-sr', ClusterB = '-sg', ClusterC = '-sb',
-              RowN = '-+r', RowB = '-+g', RowC = '-+b',)
+styles = dict(Corner_S = '-sr', Corner_C = '-og')#dict(XORN = '-or', XORB = '-og', XORC = '-ob',
[…]

### 2019-06-14 Modules/Classes/Model.py

@@ -231,7 +231,7 @@ class Model(object):
         self._update_()
         
 
-    def simulate_generation(self, stimuli, category, nexemplars = None):
+    def simulate_generation(self, stimuli, category, nexemplars = None,wrap_ax = None):
         """
         Simulate the generation of n-exemplars, sourced from stimuli, into a category.
         The resulting category will be added to the model's memory, and the identity 
@@ -249,7 +249,7 @@ class Model(object):
         for i in range(nexemplars):
 
             # compute probabilities, then pick an item
-            ps = self.get_generation_ps(stimuli, category)
+            ps = self.get_generation_ps(stimuli, category, wrap_ax=wrap_ax)
                         #print sum(ps)-1
             num = Funcs.wpick(ps)
             values = np.atleast_2d(stimuli[num,:])

### 2019-06-14 Modules/Funcs/Plot.py

@@ -66,7 +66,7 @@ def plotgradient(h, G, alphas, betas,
      - alphas/betas are the [-1 +1] coordinates of category memebers
      - clim (optional) defines the limits of the colormap.
      - cmap (optional) deinfes the colormap
-     - [alpha/beta]_col: color of alpha and beta markers
+     - [alpha/beta]_col: color of alpha and beta markers     
     """
 
     # generate clims if not provided
@@ -95,12 +95,13 @@ def plotgradient(h, G, alphas, betas,
     for j in range(coords.shape[0]):        
         h.text(coords[j,0],coords[j,1], 'A',color = alpha_col[j], **textsettings)
 
-    coords = gradientspace(betas, G.shape[0])
-    #If number of elements in beta_col doesn't match numel B,
-    #take the first element and expand it to fit B
-    beta_col = checkcolors(beta_col,coords.shape[0])
-    for j in range(coords.shape[0]):
-        h.text(coords[j,0],coords[j,1], 'B', color = beta_col[j], **textsettings)
+    if len(betas)>0:
+        coords = gradientspace(betas, G.shape[0])
+        #If number of elements in beta_col doesn't match numel B,
+        #take the first element and expand it to fit B
+        beta_col = checkcolors(beta_col,coords.shape[0])
+        for j in range(coords.shape[0]):
+            h.text(coords[j,0],coords[j,1], 'B', color = beta_col[j], **textsettings)
     h.set_yticks([])
     h.set_xticks([])
     h.set_aspect('equal', adjustable='box')

### 2019-06-16 Manuscripts/cog-psych/post-revision/post-cogpsych.tex
pass through

@@
 this domain has focused on how the statistical structure of known categories
-generalizes to generated categories, overlooking the extent that the contrast
+generalizes to generated categories, overlooking whether (and if so, how) contrast
 between the known and generated categories is a factor. We report three
@@
 
-Creating and learning new ideas is one of the most fascinating and important
+Creating new ideas is one of the most fascinating and important
 human capabilities. For example, cell phones and smart phones are newly created
@@
 \citep{lake2017,lehman2011,taylor2016}. Yet, the cognitive mechanisms that
-enable people to innovate and learn new innovations are not well understood and
+enable people to innovate are not well understood and
 are understudied by scientists. In part, this is due to the inherent difficulty
 of designing and conducting experiments that test these capabilities. How would
-a scientist devise an experiment that induces a participant to create and learn
-new categories that are as interesting as cell phones and smart phones? As a
-step towards these goals, in this article we investigate the expectations that
-people have when they learn and generate new categories using computational
-models and behavioral experiments.
-
-Generating and learning new concepts are, however, not altogether different from
+a scientist devise an experiment that induces a participant to create 
+new categories that are as interesting as cell phones and smart phones?
+Devising such an experiment is still beyond our capability.
+Instead, we take a step towards this goal, and investigate the fundamental constraints and expectations that people have when they generate new categories.
+
+% we do have tools to study the fundamental constraints on category . As a
+%step towards these goals, in this article we investigate the expectations that
+%people have when they generate new categories using computational
+%models and behavioral experiments.
+
+Creating new concepts are, however, not altogether different from
 the types of behaviors typically studied in cognitive psychology laboratories.
@@
 label. Thus, we can make progress formalizing the processes involved in category
-generation by extending theories of categorization to this case. Then, to ensure
-that these principles reflect basic principles of categorization and are not
-just specific to category generation, we can examine what factors of the
-generated categories affect the ease at which people learn the category.
+generation by extending theories of categorization to this case. This will also serve to test the breadth of the explanatory power of categorization models to tasks beyond typical category learning.
 
@@
 representative of the domain (and thus, not new at all). Contrast from other
-known categories should play a role. Further, scientists in other fields, such
+known categories should play a role. Indeed, scientists in other fields, such
 as marketing \citep{berger16} and sociology \citep{rogers03}, highlight the
@@
 systematic investigation addressing how generated categories {\em differ} from
-what is already known. Although the idea of category contrast is ubiquitous
+what is already known. Although the idea of category contrast is discussed
 throughout the categorization literature, and extends to a variety of other
@@
 strong and systematic effects of category contrast on concept generation, and we
-qualitatively and quantitatively analyze the performance of each model in
+qualitatively and quantitatively analyze the behavioral results and the performance of each model in
 capturing human category generation. A third experiment is performed to
@@
 experiments devised by \citeauthor{jern2013probabilistic} were specifically
-designed to challenge it. Its application is notable as a first step
-toward understanding theories developed in the creative cognition literature
-using well-known formal approaches from the categorization literature.
+designed to challenge it.
 
@@
 
-Prior work on category generation has explored only one factor: distributional
-correspondences between learned and generated categories. As a result, most of
-the computational, theoretical, and empirical efforts have been focused on
-explaining those effects. In this paper, we investigate another important
-factor: category contrast. To generate a novel concept, individuals must produce
-something that is in some capacity {\em different} from what they already know.
-As a consequence, we propose that contrast should be a primary constraint that
-underlies people's expectations about how a set of categories are organized:
-Categories should be different from each other.
-
-Although it is evident that people are {\em capable} of creating new concepts
-and categories, it is not entirely clear how new concepts are systematically
+%Prior work on category generation has explored only one factor: distributional
+%correspondences between learned and generated categories. As a result, most of
+%the computational, theoretical, and empirical efforts have been focused on
+%explaining those effects. In this paper, we investigate another important
+%factor: category contrast. To generate a novel concept, individuals must produce
+%something that is in some capacity {\em different} from what they already know.
+%As a consequence, we propose that contrast should be a primary constraint that
+%underlies people's expectations about how a set of categories are organized:
+%Categories should be different from each other.
+
+Although people are {\em capable} of creating 
+new categories, it is not entirely clear whether (and if so, how) new concepts are systematically
 made different from what is already known. The hierarchical Bayesian model
@@
   Bayesian models. The generative process would need to include a component that
-  presumes contrast.} Note that this does not mean location information and
+  presumes contrast, which is precisely the factor of study in this article.} Note that this does not mean location information and
 contrast cannot be captured using hierarchical Bayesian models. In fact, a model
@@
 much of the theoretical and empirical work in categorization. The focus of our
-work with PACKER concerns the dual constraints of within- and between-class
+work with PACKER concerns how contrast may be captured through the dual constraints of within- and between-class
 similarity; it is not difficult to imagine how such constraints may be
@@
 \subsubsection{Relation Between PACKER and Copy-And-Tweak}
-\label{section:copytweak-packer} The PACKER model is similar to the
-copy-and-tweak model reported by \cite{jern2013probabilistic}: Both models are
-exemplar-based, and both models generate new items according to their similarity
-to known members of the target class. However, PACKER diverges from the
-copy-and-tweak model by including a contrast mechanism, enabling generation
-according to dissimilarity to members of opposing categories. As a consequence,
-copy-and-tweak can be realized as a parameterization of the PACKER model that is
-insensitive to category contrast. Specifically, when $\theta_c = 0$ and
-$\theta_t > 0$ (see Figure \ref{fig:packer-examples}, panel B), $f(x_j) = 0$ for
-$x_j$ belonging to contrast categories; thus, PACKER is not influenced by these
-items, and is mathematically equivalent to a copy-and-tweak approach.
-
-In this paper, we report simulations using this copy-and-tweak model. This model
+%: Both models are
+%exemplar-based, and both models generate new items according to their similarity
+%to known members of the target class. However, PACKER diverges from the
+%copy-and-tweak model by including a contrast mechanism, enabling generation
+%according to dissimilarity to members of opposing categories. As a consequence,
+%copy-and-tweak can be realized as a parameterization of the PACKER model that is
+%insensitive to category contrast. Specifically, when $\theta_c = 0$ and
+%$\theta_t > 0$ (see Figure \ref{fig:packer-examples}, panel B), $f(x_j) = 0$ for
+%$x_j$ belonging to contrast categories; thus, PACKER is not influenced by these
+%items, and is mathematically equivalent to a copy-and-tweak approach.
+
+\label{section:copytweak-packer} The PACKER model is similar to the copy-and-tweak model reported by \cite{jern2013probabilistic}.
+An advantage of formulating copy-and-tweak as a special case parameterization of PACKER. In this paper, we report simulations using this copy-and-tweak model. This model
 fits within the exemplar-based category generation framework defined above, and
@@
 \begin{equation}
-  R(x,h) = \log \dfrac{p(x|h)}{1-p(x|\mathcal{H}^c)} = \log \dfrac{p(x|h)}{\sum_{h'\in H'}{p(x|h')p(h')}} . 
+  R(x,h) = \log \dfrac{p(x|h)}{1-p(x|\mathcal{H}^c)} = \log \dfrac{p(x|h)}{\sum_{h'\in \mathcal{H}^c}{p(x|h')p(h')}} . 
 \label{eq:representativeness}
@@
 probabilities proportional to the multivariate normal likelihood $p(y|h)$, the
-representativeness model generates new exemplars according to their
-representativeness $R(y,h)$. In practice, the probability of generating a
+representativeness model generates new exemplars according to their representativeness $R(y,h)$ as formulated in Equation \ref{eq:representativeness}. In practice, the probability of generating a
 particular candidate $y$ is obtained using an Exponentiated
-\citet{luce1977choice} choice rule:
+choice rule \citep{luce1977choice}:
 \begin{equation}
@@
 
-\subsection{Synopsis and Prognosis}
-
-Research on categorization and category generation has focused on the finding
-that generated categories tend to possess distributional commonalities with
-known categories. However, a fundamental task of categorization is to create
-something {\em new} (i.e., different from what is already known). The manner in
-which generated categories differ from known ones is, nonetheless, poorly
-understood: Existing theories do not make strong predictions about how generated
-concepts should systematically differ from existing ones. Above, we introduced
-two novel approaches to formalizing the roles of similarity and contrast in
-category generation.
-
-In the sections below, we present two experiments demonstrating systematic
-effects of category contrast on category generation inspired by factors
-influencing how the contrast models generate new categories. Our experiments are
-based on \cite{jern2013probabilistic}'s paradigm, which is a straightforward
-translation of the traditional artificial classification paradigm to the task of
-category generation: Participants are first exposed to a single,
-experimenter-defined category, and are then asked to generate members of a new
-category. We then report quantitative analyses of our results comparing the exemplar dissimilarity
-and the representativeness accounts of our results to that of the hierarchical
-Bayesian and copy-and-tweak models developed by \cite{jern2013probabilistic}.
-After analyzing the category generation studies, we present a third experiment to
-distinguish between our two conceptualizations of contrast.
+%\subsection{Synopsis and Prognosis}
+%
+%Research on categorization and category generation has focused on the finding
+%that generated categories tend to possess distributional commonalities with
+%known categories. However, a fundamental task of categorization is to create
+%something {\em new} (i.e., different from what is already known). The manner in
+%which generated categories differ from known ones is, nonetheless, poorly
+%understood: Existing theories do not make strong predictions about how generated
+%concepts should systematically differ from existing ones. Above, we introduced
+%two novel approaches to formalizing the roles of similarity and contrast in
+%category generation.
+%
+%In the sections below, we present two experiments demonstrating systematic
+%effects of category contrast on category generation inspired by factors
+%influencing how the contrast models generate new categories. Our experiments are
+%based on \cite{jern2013probabilistic}'s paradigm, which is a straightforward
+%translation of the traditional artificial classification paradigm to the task of
+%category generation: Participants are first exposed to a single,
+%experimenter-defined category, and are then asked to generate members of a new
+%category. We then report quantitative analyses of our results comparing the exemplar dissimilarity
+%and the representativeness accounts of our results to that of the hierarchical
+%Bayesian and copy-and-tweak models developed by \cite{jern2013probabilistic}.
+%After analyzing the category generation studies, we present a third experiment to
+%distinguish between our two conceptualizations of contrast.
 
@@
 role, exemplars in the generated categories of participants in the XOR condition
-may not be positively correlated -- they may be negatively correlated instead.
+may not be positively correlated -- they may not be correlated, or perhaps even be negatively correlated.
 In this case, contrast and statistical regularities would interact, which would
@@
 properties as previously-known categories.
-Given the increased emphasis of replication within psychology \citep{zwaan2018making}, it is important as it serves as a conceptual replication of \citet{jern2013probabilistic}.
+Given the increased importance of replication within psychology \citep{zwaan2018making}, it is important as it serves as a conceptual replication of \citet{jern2013probabilistic}.
 For each generated category, we
@@
 
-\subsection{Discussion} In Experiment 1 we sought to extend our analysis of the
-\cite{jern2013probabilistic} data by evaluating the influence of category
+\subsection{Discussion} In Experiment 1 we evaluated the influence of category
 contrast on category generation, given qualitatively different types of categories. We found strong evidence for effects of category contrast in each
@@
 non-normality of the data (Shapiro-Wilk normality test $W=0.77, p < .001$ for
-the Middle condition and $W=0.85, p < .001$ for the Bottom condition). Figure
-\ref{fig:e2-yranges} depicts the Y-axis position of the exemplars generated
-within each participant's category.
-Further, the median produced Y-Axis range is much smaller in the Bottom than Middle condition, whereas they are essentially identical in X-Axis range.
-The categories are sorted by overall range,
-then by condition assignment. These data reveal that there were nearly as many
+the Middle condition and $W=0.85, p < .001$ for the Bottom condition).
+Figure \ref{fig:e2-yranges} depicts the Y-Axis range and Y-Axis position of exemplars generated by each participant.
+The categories are sorted by overall range, and then colored by training condition.
+These data reveal that there were nearly as many
 participants who generated categories spanning the entire Y-axis as those who
-generated categories spanning almost none of the Y-axis. The non-normality of
-the Y-axis range distributions thus requires that we use a different approach to
+generated categories spanning almost none of the Y-axis.
+Indeed, the median produced Y-Axis range is much smaller in the Bottom than Middle condition, whereas they are essentially identical in X-Axis range.
+The non-normality of the Y-axis range distributions thus requires that we use a different analytic approach to
 addressing the experiment's main question.
@@
 (e.g., BIC, \citealp{schwarz1978estimating}; $AIC_C$,
-\citealp{hurvich1989regression}). In addition to AIC, table \ref{table:global-model-fits} contains the corresponding log-likelihood ($L$) and the best-fitting
+\citealp{hurvich1989regression}). In addition to AIC, Table \ref{table:global-model-fits} contains the corresponding log-likelihood ($L$) and the best-fitting
 parameter values. These results reveal strong model differentiation: both
@@
 item in the `Beta' category). To ensure this is not driving our results, we
-conducted an identical set of simulations as above, excluding this trial
+conducted an identical set of simulations as above, excluding the first trial
 (leaving 915 responses in the dataset): Again, the representativeness model ($L
@@
 row-like and column-like categories. This covariance matrix specifies the amount
-of variance assumed along each dimension (as well as the dimensional
-correlations) across the domain of categories. The covariance matrix for a newly
+of variance assumed along each dimension (as well as the correlations between dimensions) across the domain of categories. The covariance matrix for a newly
 generated category, $\Sigma_B$, is based on the assumed $\Sigma_0$ as well as
@@
 Experiments 1 and 2 clearly established the importance of contrast in category
-generation. Follow-up model-based analyses illustrated that both contrast models
-account for participant performance better than models that do not take into
+generation. Follow-up model-based analyses illustrated that both contrast models account for participant performance better than models that do not take into
 account contrast. We also found more support for the hypothesis of contrast as
@@
 While the representativeness model fits better than PACKER, and in spite of
-their fundamentally different structures, both contrast models seem to make
-qualitatively similar predictions. To illustrate this, Figure
-\ref{fig:firstbetas} shows how the modal generating probabilities for the first
-Beta exemplar are similar across both contrast models.\footnote{Model parameters
-  are set to the optimized values from Table \ref{table:global-model-fits}.} The
-highest density regions for all models appear to be near the corners and edges.
-This is particularly the case for the representativeness model, which almost
-exclusively predicts generation at the corners, while PACKER can generate the
-first novel exemplars toward the inner regions of the stimulus space in the XOR
-condition.
+their fundamentally different structures, both contrast models make qualitatively similar predictions for Experiments 1 and 2.
+To illustrate this, Figure \ref{fig:firstbetas} shows how the modal generating probabilities for the first Beta exemplar are similar across both contrast models.\footnote{Model parameters are set to the optimized values from Table \ref{table:global-model-fits}.} The highest density regions for both models appear to be near the corners and edges.
+This is particularly the case for the representativeness model, which almost exclusively predicts generation at the corners, while PACKER can generate the first novel exemplars toward the inner regions of the stimulus space in the XOR condition.
 
@@
 regions are maximally distant from the Alpha exemplars. However, recall that
-unlike PACKER, the representativeness model does not explicitly use a
-distance-based measure in determining exemplar generation probabilities. The
-general pattern of producing exemplars dissimilar from the contrast category is
-the result of the underlying unimodal category structure assumed by the model
-and is not strictly constrained by the distances between exemplars.
-Consequently, it is possible to design a configuration of Alpha exemplars where
-the representativeness model would predict the generation of novel exemplars
+unlike PACKER, the representativeness model does not explicitly use a measure sensitive to distance from the location of exemplars in the contrasting category.
+The general pattern of producing exemplars dissimilar from the contrast category is because a unimodal category structure is assumed by the model and it is not strictly constrained by the distances to exemplars of the contrasting categories.
+Consequently, it is possible to design a configuration of Alpha exemplars where the representativeness model would predict the generation of novel exemplars
 that are \emph{more similar} to the learned exemplars.
 
-In this section, we focus on one particularly useful category type -- the Corner
+In this section, we focus on one category type -- the Corner
 category -- to distinguish between the two competing accounts of contrast in
-category generation. Here, each Alpha exemplar is located at the extreme corners
-of the feature space (see Figure \ref{fig:e3-squares}a). We first seek to
-identify what predictions each contrast model generates under this condition,
-followed by an category generation task, and a brief modeling exercise to
-determine which model is a better predictor of the behavioral data.
+category generation.
+Here, each Alpha exemplar is located at the extreme corners of the feature space (see Figure \ref{fig:e3-squares}a).
+We first seek to identify what predictions each contrast model generates under this condition,
+followed by an category generation task, and a model-based analysis of the two contrast models given the behavioral data.
 
 In the Corner condition, the representativeness model assumes an underlying
-Alpha category distribution that is centred on the feature space, with
-probability densities highest in the centre and lowest in the corners. Since the
-representativeness of novel Beta exemplars generally increase when the
+Alpha category distribution that is centered on the feature space, with
+probability densities highest in the center and lowest in the corners. Since the representativeness of novel Beta exemplars generally increase when the
 likelihood of the underlying Alpha category distribution decreases, the regions
@@
 
-[Transition to and justification for this still feels a little shaky]
-
-A large part of the representativeness model's predictive accuracy in
-Experiments 1 and 2 is due to its preference for generating categories located
+Another factor we examine with this category structure is the effect of boundaries. The represenativeness model prefers to generate categories located
 at the boundaries of the stimulus features. If at least one of these boundaries
-is removed by utilizing an unbounded feature (i.e., no absolute maximum or
-minimum exists in the stimulus space along this dimension), we should expect to
-see a decrease in the predictive performance of the representativeness model. To
-implement this, in our next experiment we also include a new set of stimuli:
+is removed by utilizing a toroidal feature (one with no boundaries), such as the rotation of a line.
+The represenativeness model will strongly prefer to generate categories on the non-toroidal or bounded feature as it still has a boundary.
+Experiment 3 manipulates the stimulus space.
+One condition uses the same stimuli as before, squares varying in lightness and size, both of which are bounded features.
+The other condition use a new set of stimuli:
 circles that vary in size (a bounded feature) and orientation (an
 unbounded feature; Figure \ref{fig:e3-circles}a). Before describing the
-details of this experiment, we will briefly explore the contrast model
+details of the experiment, we examine each model's
 predictions with stimuli that comprise one bounded and one unbounded feature.
@@
 
-Compared to stimuli with two bounded features, PACKER makes a similar prediction
-in generating the first Beta exemplar when one feature is unbounded, with the
-curious exception that the highest density regions appear compressed along the
-y-axis. Due to the unbounded nature of orientation, the maximum distance from
+Compared to stimuli with two bounded features, PACKER makes a similar prediction in generating the first Beta exemplar when one feature is unbounded, with the curious exception that the highest density regions appear compressed along the y-axis (for simplicity, we always assign the y-axis as the toroidal feature in the text and figures).
+Due to the unbounded nature of orientation, the maximum distance from
 the Alpha exemplars is half of the full range of the feature -- for an Alpha
-exemplar oriented at 0 radians, the furthest another exemplar can be situated is
-$\pi$ radians within the full range of orientation is $2\pi$ radians.
+exemplar oriented at 0 radians, the furthest another exemplar can be situated is $\pi$ radians within the full range of orientation is $2\pi$ radians.
+
+For the representativeness model, adding an unbounded feature changes the probability distribution for the underlying Bayesian model.
+Whereas it is still Gaussian in one feature, the other feature is toroidal, meaning that it wraps back to $0$ radians once it reaches $2\pi$ (we have conducted a change of variables where $y \rightarrow y \mod 2\pi $, where $\mod$ is the modulus operator).
+As we cross $2\pi$ and reach $0$ again, the remaining probability density also gets wrapped around, meaning that the probability density assigned by the Gaussian distribution on the toroidal dimension at $2\pi$ is actually assigned to $0$ (and $2\pi + \epsilon$ to $\epsilon$ so on).
+This is called the {\em wrapped} Gaussian distribution \citep{mardia1972}.
+Note that the often used distribution for circular dimensions, the Von Mises distribution is an approximation to the wrapped Gaussian distribution.
+See Appendix B for implementational details.
 
 For the representativeness model, the single unbounded feature essentially
-prevents the formation of edge effects along that feature. Consequently, this
-results in higher probabilities of the generating the first Beta exemplars along
-the edges of the bounded feature (Figure \ref{fig:e3-circles}c). Since this
+prevents the formation of edge effects along that feature.
+Consequently, this results in higher probabilities of the generating the first Beta exemplars along the edges of the bounded feature (Figure \ref{fig:e3-circles}c).
+As this
 model does not explicitly recognize similarities as distances along the feature
@@
 along the unbounded feature axis. Intuitively, this suggests that the
-representativeness model (which prefers generating exemplars at the edges) will
-be poorer at generating stimuli with an unbounded feature compared to
-stimuli with all features bounded.
+representativeness model (which prefers generating exemplars at the boundaries) will prefer to generate stimuli on the boundaries of the bounded feature with some amount of variance over unbounded feature.
 
@@
 
-[Maybe work on transition if there's time]
+%[Maybe work on transition if there's time]
 
@@
 
-We recruited 89 participants from Amazon Mechanical Turk for this experiment, 46
-of whom were randomly assigned square stimuli and the remaining 43 assigned
-circle stimuli. Square stimuli were exactly the same as in Experiments 1 and 2.
-Circle stimuli were made up of unfilled circles with a radial line extending
-from its centre to the edge. These stimuli varied along two features -- their
-diameter (or size) ranging from 3.0--5.8 cm, and their orientation ranging from
+We recruited 89 participants from Amazon Mechanical Turk who were randomly assigned to the bounded, square stimuli ($N=46$), or to the toroidal, circle stimuli ($N=43$).
+The square stimuli were exactly the same as in Experiments 1 and 2.
+The circle stimuli were made up of unfilled circles with a radial line extending from its center to its edge.
+These stimuli varied along two features -- their
+diameter ranging from 3.0--5.8 cm, and their orientation ranging from
 0--$2\pi$ radians. The orientation of each circle stimuli was fixed at a
@@
 participants first observed a training phase where four Alpha exemplars were
-presented across three blocks followed by the generation phase. All participants
-in this experiment were shown exemplars from the Corner category type.
+presented across three blocks followed by the generation phase. The training condition was the Corner category for all participants.
 
@@
 
-Overall, we found no significant differences between the distributional statistics of each
-generated category between square and circle stimuli (Figure
-\ref{fig:e3-statsboxes}. The only exception is seen in the Y range (representing
-the orientation of circle stimuli,) which is significantly lower for the circle
-stimuli compared to the square stimuli ($t(87) = 4.94$, $p = .001$). This
+Participants produced circle stimuli with significantly lower Y-range (lower for the unbounded feature) as compared with the square stimuli ($t(87) = 4.94$, $p = .001$).  This
 provides preliminary support for PACKER, which predicts generating exemplars
 within a shorter range along the unbounded feature compared to a bounded
-feature.
-
-At a rough level (see Table \ref{table:corner_counts}, 35\% of participants who
-observed square stimuli tended to generate their first Beta exemplars along the
-edges of the feature space. In comparison, only 12\% of participants who
-observed circle stimuli generated their first Beta exemplars along the edges of
+feature. Otherwise, there were no significant differences between the distributional statistics of each
+generated category between square and circle stimuli (Figure
+\ref{fig:e3-statsboxes})
+
+The contrast models were fit by optimizing their fit to all experiments in this article (Experiments 1 through 3).
+Fit statistics and parameter values are presented in Table \ref{table:corner-model-fits}.
+Overall,
+the representativeness model emerges as the best-fitting model in aggregate, although by a much smaller amount than Experiments 1 and 2. However, analyzing data in aggregate can be misleading.
+
+%by a much smaller lead compared to fits to only Experiments 1 and 2. At an individual level, 67\% of participants in Experiment 3 were better
+%fit by PACKER than the representativeness model.
+
+Table \ref{table:corner_counts} presents frequencies of the best-fitting model for participants in each condition.
+Overall the representativeness model was significantly worse at capturing the results of Experiment 3 (it fit 29 out of the 89 participants better than PACKER; two-tailed Binomial test $p < 0.005$).
+Further, this is mostly due to representativeness model capturing participants in the unbounded circle condition worse than  PACKER (11 of 43 better fit by representativeness; two-tailed Binomial test, $p<0.005$).
+The representativeness model only trended towards fitting participants in the bounded, Square condition better (18 out of 46; two-tailed Binomial test $p = 0.18$).
+Further, 35\% of participants in the bounded condition generated their first Beta exemplars along the
+edges of the feature space, but only 12\% of participants in the toroidal, circle condition generated their first Beta exemplars along the edges of
 the feature space. These proportions more closely match the predictions of
@@
 representativeness model (53\% of the former, 49\% probability of the latter).
-In addition, these results agree with our prediction that the representativeness
-model would be poorer in predicting data from circle stimuli (26\% of
-participants observing circle stimuli) compared to square stimuli (39\%
-participants observing square stimuli). [No proper stats here... feels a little
-dirty. But I suppose there are model fits coming up.]
-
-Quantitative model fits were also obtained by optimizing both contrast models to
-the entire dataset including data from Experiments 1 and 2. Fit statistics and
-parameter values are presented in Table \ref{table:corner-model-fits}. Overall,
-the representativeness model still emerges as the best-fitting model, although
-by a much smaller lead compared to fits to only Experiments 1 and
-2. At an individual level, 67\% of participants in Experiment 3 were better
-fit by PACKER than the representativeness model.
+Thus, the representativeness model's success at capturing Experiments 1 and 2 is in part due to the nature of the stimulus features, as well as the structure of contrasting category.
+%In addition, these results agree with our prediction that the representativeness model would be poorer in predicting data from circle stimuli (26\% of participants observing circle stimuli) compared to square stimuli (39\% participants observing square stimuli).
 
@@
 \label{table:corner_counts}
-\begin{tabular}{ l l l l} \\ Stimulus Type & \textbf{PACKER} &
-                                                               \textbf{Representativeness}& Total \\
+\begin{tabular}{ l l l l l} \\ Stimulus Type & \textbf{PACKER} &
+                                                               \textbf{Representativeness}& Total & Two-tailed Binomial test $p$-value \\
   \hline 
-  Squares & 28 & 18 & 46\\
-  Circles & 32 & 11 & 43\\
-  Total   & 60 & 29 & 89\\
+  Squares & 28 & 18 & 46 & $p < 0.005$\\
+  Circles & 32 & 11 & 43 & $p \approx 0.18$\\
+  Total   & 60 & 29 & 89 & $p < 0.005$\\
 \end{tabular}
@@
 
-Ultimately, these results suggest that PACKER's exemplar dissimilarity account
-of contrast, as opposed to the representativeness account, is more appropriate
-for the Corner category type. On almost every measure of analysis in this
-section, PACKER outperformed the representativeness model. Although the
-representativeness model best fit the data from Experiments 1 and 2, it
-struggled when the learned category occupied the extreme corners of the stimulus
-space -- both in cases where the stimuli possess bounded and unbounded features.
-
-[Work on transition time permitting]
+%Ultimately, these results suggest that PACKER's exemplar dissimilarity account
+%of contrast, as opposed to the representativeness account, is more appropriate
+%for the Corner category type. On almost every measure of analysis in this
+%section, PACKER outperformed the representativeness model. Although the
+%representativeness model best fit the data from Experiments 1 and 2, it
+%struggled when the learned category occupied the extreme corners of the stimulus
+%space -- both in cases where the stimuli possess bounded and unbounded features.
+%
+%[Work on transition time permitting]
 
@@
 
+\subsection{Similarity and Contrast in Cognition}
+We propose that category contrast as a primary constraint in categorization.
+For categories to be useful, they should not all be identical, or in other words, they should be different.
+Thus, a newly generated category should be different than pre-existing categories.
+Beyond its role
+in category generation specifically, category contrast is also of fundamental importance in
+categorization more broadly. All other factors held constant, new categories are
+easier to learn if they are dissimilar to members of other categories, and
+knowledge of highly distinct categories is applied more accurately than that of
+ill-defined categories \citep{ashby1994categorization,imai1965discriminability}.
+Likewise, basic-level categories \citep{rosch1976basic} are thought to be
+abstracted in order to maximize within-category similarity while minimizing
+between-category similarity. Finally, the act of forming category
+representations affects similarity judgments about category members and
+nonmembers, with category members being viewed as more similar to one another
+than members of other categories
+\citep{goldstone1994influences,goldstone2001altering,goldstone1996isolated}.
+
+Beyond the traditional categorization literature, one can find instances of the trade-off between within
+and between-class similarity in linguistic categories over perceptual
+dimensions. For example, \cite{regier2007} showed that the partitioning of color
+categories reflects such a trade-off in a psychological space -- colors are
+partitioned into groups with members that are viewed as highly similar to one
+another yet distinct from other colors. A similar trade-off can be observed in
+phoneme categories. Different exemplars of the same phoneme must be similar to
+one another, while contrasting from other phonemes, such that a listener can
+infer the appropriate phoneme. This pattern has been found and modeled in the
+natural acoustics of American English vowels
+\citep{feldman2013,hillenbrand1995}. As linguistic categories must have been
+created at some point in human history, it is revealing that the constraints of
+emulating distributional structure across categories and having categories
+contrast from one another still bias human category generation today.
+
+The dual forces of within-class similarity and between-class contrast influence
+cognitive functions in a wide variety of domains. The PACKER model is notable in
+that it explicitly interprets this trade-off within the domain of
+categorization, and allows us to begin to understand the relatively understudied
+processes involved in category generation through our more well-developed tools
+for understanding human categorization.
+
+\subsection{Implications for Creative Cognition}
+
+Although the focus of this article has been to address the role of contrast in category generation, our findings and approach have relevant implications
+for research in creative cognition. A central focus of the creative cognition
+approach has been to explain acts of creativity in terms of the mental
+representations and processes that are commonly studied in cognitive psychology
+and cognitive science \citep{finke1992creative,smith1995creative}. However,
+unlike other fields in the study of cognition, creative cognition research
+rarely employs quantitative models to evaluate the explanatory value of such
+representations and processes. Our modeling results provide a concrete example
+of how formal approaches may be used to gain insight into the nature of creative
+cognition.
+
+In addition to demonstrating the utility of formal modeling for studying
+creative cognition, the contrast models here specifically offer an additional
+interpretation of some of the field's most central findings. For example,
+perhaps the most foundational principle from this literature concerns the
+limiting influence of prior knowledge: Individuals create new categories
+composed of features from existing classes, and what is created can be
+influenced drastically through the introduction of cues or examples
+\citep{marsh1999inadvertent,smith1993constraining}. In this paper, we have
+identified another important aspect of the constraining influence of prior
+knowledge: What is generated cannot be the same as what is already known.
+Further, there is systematicity in how generated categories differ from prior
+knowledge. The results of our simulations suggest that this conceptualization of
+difference can be addressed in at least two different ways. Specifically,
+simulations with PACKER demonstrate that the constraining influence of
+difference is concisely explained in terms of a trade-off between
+within-category similarity and between-category dissimilarity. Conversely, the
+representativeness model shows that this influence can also be the result of
+enhancing the representativeness of generated exemplars to their category.
+
+PACKER may offer an additional interpretation of existing accounts of creative
+generation. Most notably, a leading account within the creative cognition
+literature, the Path of Least Resistance \citep{ward1994structured,ward1995s},
+also explains generation in terms of an exemplar-based retrieval process. This
+account was designed to explain the creative generation of natural categories
+(e.g., new species of plants and animals) and as a result relies strongly on the
+hierarchical organization of these categories: Individuals are thought to
+retrieve an example of the higher-level category being generated (e.g.,
+\textit{bird} may be retrieved from the category \textit{animal}), and then
+systematically alter what was retrieved to make something new. As the PACKER
+model does not assume knowledge is hierarchically organized \citep[this is true
+of the exemplar view more broadly, see][]{murphy2016exemplar}, the model may be
+viewed as a formal instantiation of the Path of Least Resistance for application
+in a traditional artificial categorization domain (when there is no established
+hierarchy of categories). PACKER's success in explaining generation within an
+artificial domain motivates future work exploring the nature of category
+contrast within a more naturalistic setting.
+
+The broader study of creativity currently involves a wide breadth of different
+approaches \citep[for a review, see][]{kozbelt2010theories}, such as those based
+on free association \citep{mednick1962associative} and conceptual combination
+\citep{estes2002emergence,murphy1988comprehending}. In addition, recent work in
+the machine learning literature has explored using neural networks to address
+the overall problem of creative generation
+\citep[e.g.,][]{goodfellow2014gan,ho2016gail,kingma2016,chen2016infogan}. In
+contrast to these varied investigations, we reduced our focus by studying a
+highly complex behavior (category generation) as it applies within a
+well-established domain (artificial category learning). We hope that future work
+incorporates and highlights the importance of contrast into theories of
+creativity and adapts the discussed contrast models to work in state-of-the-art creative generation methods.
+
+\subsection{Implications For Categorization}
+
+Categorization research addresses the representations and processes that
+underlie the learning and use of categories. Category learning tasks are
+generally about figuring out which items belong to which category. Once learned,
+categories are generally used to classify new stimuli and to make inferences
+beyond the available information. Our work is fairly unique in that people learn
+a single category through positive examples and then create another category
+that would make sense in the domain. In other words, only one category is
+provided to be learned and the use that is required of that category is to
+generate a contrast category. Such scenarios have direct application to
+real-world situations. For instance, consider being stranded in an unfamiliar
+environment where you eat and get sick from a mushroom with certain features. In
+avoiding such mushrooms and seeking alternative sources of nutrition, it would
+behoove you to imagine what features an edible, safe mushroom would possess (e.g.,
+if the poisonous mushroom had a long stem and gills, perhaps a mushroom with a short stem that did not have gills would be safe to
+consume). In the present studies, we have learned something about the form that
+such expectations are likely to take.
+
+We can think of the category generation task in our studies as asking a person
+to formulate an idea about what set of items in the domain are most
+interestingly {\em not} members of the original category. To meet this
+condition, the items must take some form of coherence that aligns with that of
+the original category and some form of distinctiveness relative to the original
+category. Reflecting the basic level of organization in natural categories, it
+makes sense to generate a set of items that possess strong within-category
+coherence (by importing or systematically transforming the internal structure of
+the original category) as well as strong between-category differentiation (by
+creating maximum contrast with the original category, be it through
+exemplar-based dissimilarity or the maximization of exemplar-category
+representativeness). In this sense one can see the patterns of performance in
+the category generation task as recapitulating the order of semantic
+organization.
+
+The current work also suggests exciting directions for related investigations
+into unsupervised categorization. While the unsupervised categorization
+literature is primarily interested in the generation of categories within a set
+of observed exemplars where no prior categories are learned, our category
+generation studies are focused on the production of novel category exemplars
+themselves. Despite this distinction, both types of categorization research are
+ultimately interested in the question of how categories can be formed. The results
+of our current work indicate that the formation of categories is constrained by
+some measure of contrast. To our knowledge, this explicit investigation of
+category contrast has not yet been attempted in the unsupervised categorization
+literature.
+
+Interestingly, the unidimensional sort bias that is commonly observed in
+unsupervised categorization
+\citep{imai1965discriminability,milton2004influence,ahn1992two}, where
+participants sort unlabelled family resemblance category exemplars by focusing
+on a single feature, was not consistently observed in our tasks. In our
+experiments we only observed this bias in the row condition of Experiment 1,
+whereas in other conditions participants generated categories that typically
+varied along both dimensions. This may be unsurprising at face value, since
+participants in conditions other than the row condition were trained on a prior
+category that varied on both dimensions. However, these results place constraints on how broadly applicable the 
+unidimensional bias is to different types of category construction. Additionally, the
+unidimensional sort bias is observed in category construction tasks in which
+participants develop a category-level organization of a provided domain while
+category generation tasks involve generating a new category relative to one that
+was just acquired (one might say that participants are forging a domain from a
+category rather than categories from a domain). Since the more creative task of
+generating a category from scratch (with respect to an established category)
+does not invoke the unidimensional bias, this suggests that a more creatively
+demanding version of a sort task might also reduce the bias.
+
+\subsection{Exemplar Dissimilarity and Representativeness in Categorization} 
+Although both PACKER and the representativeness model perform better than the
+classical models, they appear to be capturing fundamentally different aspects of
+contrast in categorization. Specifically, the representativeness model
+excelled in predicting categories generated by participants in Experiments 1 and 2.
+Conversely, PACKER was the best-fitting model in accounting for categories generated from the Corner category type, especially when the stimuli had a toroidal feature.
+
+% Given the history of their development, this is perhaps not entirely
+% surprising. PACKER is implemented as an extension of the GCM, which was
+% originally developed for (and has been highly successful in) category learning
+% \citep{nosofsky1986attention, nosofsky1994comparing}. Conversely, the
+% representativeness mechanism was conceived as a heuristic describing the
+% relative extent to which an individual observation is generated from some
+% unobserved population \citep{kahneman1972subjective,abbott16}. This generative
+% nature of the representativeness mechanism intuitively lends itself to the
+% prediction of category generation data.
+
+While the contrast models were born quite naturally out of the corresponding
+classical models, it is worth considering if similar benefits in prediction
+would be seen if the contrast mechanisms were applied to different classical
+models. Although there is no clear way to adapt the exemplar dissimilarity
+mechanism to the current hierarchical Bayesian model, it is a fairly
+straightforward process applying the representativeness mechanism to an exemplar
+model. Specifically, by treating the similarity measure (Equation
+\ref{eq:similarity}) in the current implementation of copy-and-tweak as a
+density estimate for the representativeness mechanism (Equation
+\ref{eq:representativeness}), we gain a new model of representativeness that is
+based on exemplar similarity instead of a multivariate Gaussian
+likelihood.\footnote{An analogous implementation of the representativeness
+  mechanism to PACKER results in both $\theta_c$ and $\theta_t$ adding to
+  constant value that is independent of the similarity to contrast and target
+  exemplars respectively. Consequently, it is formally equivalent to a
+  copy-and-tweak model with representativeness.}
+
+How well does a representativeness model that uses exemplar similarity to
+represent the category’s density over features explain our experimental results?
+We find that this model substantially outperforms its classical counterpart, but
+not to the extent of PACKER or the representativeness model. Specifically,
+fitting this model to the category generation data yielded $L = -4815$ and
+$AIC=9633$ for the entire set of data from Experiments 1 and 2, and $L = -3474$,
+$AIC=6953$ for data excluding the first trials. When used to predict the error
+rates of participants in the category learning task (Experiment 3), the
+performance of the copy-and-tweak with representativeness model correlated with
+participant error with $r=.56, p < .001$. These results suggest that the
+predictive advantages to including the representativeness mechanism is not
+limited to a hierarchical Bayesian model, but can also be found when applied to
+an exemplar model. Although its performance is not as strong as either of the
+fully-developed contrast models, there is a benefit to the representativeness
+mechanism that is independent from any interaction with a hierarchical Bayesian
+framework.
+
+Ultimately, the source of the discrepancy between the performance of the two
+contrast mechanisms in category generation is still not entirely clear. Our
+results could be an indication of a fundamental difference between the processes
+involved where exemplar dissimilarity is a larger factor under certain category
+configurations while representativeness exerts more influence in others.
+Alternatively, it is possible they are both approximating some factor that we
+still have not fully identified. While a thorough investigation is beyond the
+scope of the current article, the preliminary results of this section indicate
+that further work to disentangle these two mechanisms would be promising. In the
+following section, we explore some other limitations of the current work and
+provide more recommendations for future research.
+
+\subsection{Limitations and Future Directions}
+
+Although successful in explaining our results in all three studies (albeit not
+as well as the representativeness model for Experiments 1 and 2), PACKER does
+not provide a full account of what is known about category generation. Most
+notably, in this paper we have not evaluated the model's ability to explain the
+classic finding that generated categories tend to share distributional
+commonalities with previously learned categories
+\citep[see][]{jern2013probabilistic,ward1994structured}. While we successfully
+replicated this effect in Experiment 1, we also found that its influence was
+limited in comparison to the fundamental constraints imposed by category
+contrast. Even within Experiment 1, we found systematic inconsistencies: by
+generating exemplars into unoccupied regions of the space, participants who
+learned an `XOR' category, composed of members that are widely distributed along
+both features and are positively correlated in space, tended to generate
+categories with an opposite (negative) correlation. More generally, PACKER
+inherits the strengths and weaknesses of exemplar models of categorization: It
+provides a simple and flexible model that explains many results, but deviates
+systematically from human performance in some cases.
+
+Nonetheless, these classic effects are a core element of the phenomenology of
+category generation, and PACKER does not include any mechanisms that explain
+them. Instead, through the development and evaluation of the PACKER model, we
+have sought to add new elements into such a phenomenology: The broad and strong
+influence of category contrast, and the interdependence between category
+location and distributional structure. It may be possible to combine the
+hierarchical Bayesian approach proposed by \cite{jern2013probabilistic} with
+PACKER's underlying claims to obtain a ``best of both worlds'' model, capable of
+explaining the role of contrast in category generation, as well as the emulation
+of distributional structure. However, as noted in the introduction, the
+incorporation of category contrast is antithetical to the core principles of a
+traditional, semi-conjugate Bayesian approach. This suggests that category
+generation is a fundamentally different computational-level problem
+\citep[different from those posed by][]{jern2013probabilistic,kemp2014taxonomy}, which is supported by the represenativeness form of their model performing much better.
+
+Characterizing that problem and conducting a rational analysis is an important
+direction for future research. To that aim, we plan to explore the connection
+between exemplar modeling as an Importance Sampling approximation
+\citep{shi10exemplar}, and see what sort of computational-level problem PACKER
+approximates. Once formalized in probabilistic terms, it should also be
+straightforward to incorporate distributional factors into the model. This would
+unite PACKER with the Bayesian representativeness model, and they would differ
+in terms of their assumptions about how people represent category distributions
+(as exemplars or prototypes, respectively). Alternatively, it may be possible to
+integrate the core principles of either model of contrast into other
+categorization models
+\citep[e.g.,][]{love2004sustain,kurtz2007divergent,smith2000thirty}.
+
+The present work focuses on the influence of contrast on categorization,
+primarily by exploring two different conceptualizations of contrast. However,
+throughout this paper we have only explored one type of category generation
+problem, that is the generation of new categories in an
+artificial domain. It is possible that the strength of the influence of contrast on
+category generation decisions varies depending on the nature of the category
+generation problem, for instance, in situations where the boundaries of the
+domain are not clearly defined. To provide some intuition for this, consider the
+following example: an entomologist is asked to draw a new type of insect that
+they have never seen. Clearly, features of the new insect will be constrained by
+the definition of an insect: it will be an arthropod with six legs and a
+relatively hard exoskeleton. However, there are no strict limitations on other
+features such as how large the insect needs to be, or if the insect should have
+a particular type of pattern on its body. In this scenario, it is plausible that
+the entomologist will draw an insect that is fairly similar to insects that
+currently exist (e.g., small in size with a camouflage pattern that matches its
+environment) as opposed to something completely different (e.g., a horse-sized
+beetle with roman numerals on its back), indicating that contrast plays only a
+minor role in the decision making process. In this situation it appears that
+there is a combination of clearly defined constraints on the domain boundary
+that are imposed by the external environment (insects must be arthropods with
+six legs) but also ill-defined constraints that are imposed by the observer
+(insects are probably small and have patterns on their body that suit the
+environment). However, the relation of these contraints (and the interactions
+between them) to the role of contrast is not currently well understood.
+
+The definition of the domain boundary is only one possible factor that could
+influence the effect of contrast on categorization. It is also plausible that
+other conditions such as the nature of the instruction (``Generate a new
+category that is Not Alpha'' vs ``Generate a new category that is Beta'') 
+can affect the extent to which contrast influences categorization. Though our
+current contrast models can allow the influence of contrast to vary, they do not
+make any predictions regarding the specific conditions under which contrast can
+vary. Consequently, addressing this issue would be an immensely promising avenue
+for future research.
+
+\section{Conclusions}
+
+The generation of new concepts and ideas is a highly interesting topic, but it
+is difficult to study in a controlled experimental environment. In this paper,
+we have provided such an examination of category generation as it applies within
+an artificial categorization experiment. Extending the literature on creative
+cognition, our experiments provide a detailed picture of the role of category
+contrast in generation: People seek to create concepts that are distinct from
+what they already know, and the nature of what is created can be influenced by
+what does not yet exist. Our simulations with traditional exemplar models, as
+well as a hierarchical Bayesian model, provide strong support for the claim that
+category contrast is of fundamental importance to categorization. More broadly,
+our results demonstrate that popular explanatory approaches from basic research
+in cognitive science can offer a precise, quantitative account of a behavior as
+complex as that of category generation, and in exploring them, we can increase
+our understanding of categorization as well.
+
+\clearpage
+\section{Acknowledgments} Previous versions of this work were presented at the
+Thirty-Ninth Annual Conference of the Cognitive Science Society and Forty-Ninth
+Annual Meeting of the Society for Mathematical Psychology. Support for this
+research was provided by the Office of the VCRGE at the UW - Madison with
+funding from the WARF. We thank Charles Kemp for feedback on a previous version
+of this manuscript. We also thank him and Alan Jern for providing code and data.
+\end{flushleft}
+
+% references section
+\clearpage \bibliographystyle{apacite} \bibliography{citations.bib}
+\clearpage
+
+\appendix \numberwithin{equation}{section}
+
+\section{The Hierarchical Bayesian Model of Concept Generation}
+\label{ap:hsampling-definition}
+
+\cite{jern2013probabilistic} demonstrated how a hierarchical Bayesian model
+could explain the distributional correspondences between observed and generated
+categories. In their model, exemplars of generated categories were viewed as
+samples from a multivariate Normal distribution over the dimensions of stimulus
+space. The mean of the generated category was independent of the observed
+categories, but the covariance matrix (encoding feature variances and
+correlations) was based on a common prior distribution. Generating a new
+category was thus completed by sampling a new category mean (uniform over
+stimulus space) and covariance matrix from the common prior distribution.
+Because the shared prior distribution's parameters were unobserved, the
+hierarchical Bayesian approach was used to infer its parameters from the
+previous categories (their feature variances and correlations), and then to
+generate the covariance matrix of the new category.
+
+In our implementation of their model\footnote{Note that
+\citet{jern2013probabilistic}'s model is slightly different, as they used a
+semi-conjugate model. Their model acts very similarly to our version.}, each
+category's exemplars are viewed as samples from a multivariate Normal
+distribution with parameters ($\mu, \Sigma$). Category covariance matrices
+(specifying variance and covariance along $k$-dimensions), are assumed to be
+Normal-Inverse-Wishart distributed with parameters: $\nu$ ($>k-1$), $\kappa$
+($>0$), and $\Sigma_D$. $\nu$ and $\kappa$ are treated as free parameters in our
+simulations, and $\Sigma_D$ is the domain-wide covariance matrix from which all
+categories are viewed as samples. Assuming a given $\Sigma_D$, a category
+covariance matrix $\Sigma$ can be computed on the basis of its examples:
+
+\begin{equation} \Sigma = \left[ \Sigma_D \nu + C + \dfrac {\kappa n} {\kappa +
+n} (\bar{x}-\mu)(\bar{x}-\mu)^T \right] (\nu + n)^{-1}
+\label{eq:Sigma_B}
+\end{equation}
+% 
+where $\bar{x}$ and $C$ are the empirical mean and covariance of the category's
+known members, and $n$ is the number of observed members of the category. When
+there are fewer than two known members of the category (and thus no covariance
+to speak of), $\Sigma = \Sigma_D\nu$.
+
+The category mean, $\mu$, can be computed as:
+
+\begin{equation} \mu = \dfrac {\kappa\mu_{0} + n \bar{x}} {\kappa + n}
+    \label{eq:category_mus_jk13}
+\end{equation}
+% 
+where $\mu_{0}$ is the prior mean. In our simulations, $\mu_{0}$ is set to the
+center of the domain. However, when no examples of the target category have been
+observed, generation is assumed to be random. In practice, the model's best fits
+are achieved when the $\kappa$ parameter, which controls the influence of
+$\mu_0$ on $\mu$, is set very close to zero (hence, the influence of $\mu_0$ is
+minimal).
+
+Importantly, the domain-wide covariance matrix $\Sigma_D$ is unobserved and
+needs to be inferred from the observed categories. For conjugacy, if $\Sigma_D$
+is viewed as a sample from an Inverse-Wishart distribution with scale
+$\Sigma_0$, $\Sigma_D$ can be computed as:
+
+ \begin{equation} \Sigma_D = \Sigma_0 + \sum_y{C_y}
+    \label{eq:category_sigmas}
+\end{equation}
+% 
+where $\Sigma_0$ is the prior covariance in the domain. In our simulations,
+$\Sigma_0 = \lambda {\bf I}$, where $\lambda$ is a free parameter controlling
+the expected variance of dimensions (dimensions of the domain covariance matrix
+are expected to be uncorrelated) and ${\bf I}$ is a $k$-by-$k$ identity matrix.
+
+Generated exemplars are drawn from a multivariate Normal distribution specified
+by $(\mu, \Sigma)$. Thus, $p(y)$ is
+\begin{equation} p(y \mid x) = \dfrac {\exp \left \{ \theta \cdot {\rm
+Normal}(y; \mu, \Sigma) \right \} } {\sum_i \exp \left \{ \theta \cdot {\rm
+Normal}(y_i; \mu, \Sigma) \right \} }
+\end{equation}
+% 
+where $\theta$ is a response determinism parameter and ${\rm Normal}(y; \mu,
+\Sigma)$ denotes a multivariate Normal density evaluated at $y$.
+
+\end{document}
+
 \subsection{Finding Contrast Effects in Prior Work}
@@
 not precisely describe the nature of that influence.
-
-\subsection{Similarity and Contrast in Cognition}
-The proposal that category contrast is a primary constraint in categorization is, in some ways, entirely commonsense.
-For categories to be useful, they should not all be identical, or in other words, they should be different.
-Thus, a newly generated category should be different than pre-existing categories.
-Beyond its role
-in category generation specifically, category contrast is also of fundamental importance in
-categorization more broadly. All other factors held constant, new categories are
-easier to learn if they are dissimilar to members of other categories, and
-knowledge of highly distinct categories is applied more accurately than that of
-ill-defined categories \citep{ashby1994categorization,imai1965discriminability}.
-Likewise, basic-level categories \citep{rosch1976basic} are thought to be
-abstracted in order to maximize within-category similarity while minimizing
-between-category similarity. Finally, the act of forming category
-representations affects similarity judgments about category members and
-nonmembers, with category members being viewed as more similar to one another
-than members of other categories
-\citep{goldstone1994influences,goldstone2001altering,goldstone1996isolated}.
-
-Beyond categorization, one can find instances of the trade-off between within
-and between-class similarity in linguistic categories over perceptual
-dimensions. For example, \cite{regier2007} showed that the partitioning of color
-categories reflects such a trade-off in a psychological space -- colors are
-partitioned into groups with members that are viewed as highly similar to one
-another yet distinct from other colors. A similar trade-off can be observed in
-phoneme categories. Different exemplars of the same phoneme must be similar to
-one another, while contrasting from other phonemes, such that a listener can
-infer the appropriate phoneme. This pattern has been found and modeled in the
-natural acoustics of American English vowels
-\citep{feldman2013,hillenbrand1995}. As linguistic categories must have been
-created at some point in human history, it is revealing that the constraints of
-emulating distributional structure across categories and having categories
-contrast from one another still bias human category generation today.
-
-The dual forces of within-class similarity and between-class contrast influence
-cognitive functions in a wide variety of domains. The PACKER model is notable in
-that it explicitly interprets this trade-off within the domain of
-categorization, and allows us to begin to understand the relatively understudied
-processes involved in category generation through our more well-developed tools
-for understanding human categorization.
-
-\subsection{Implications for Creative Cognition}
-
-Although the focus of this article has been to address the role of contrast in category generation, our findings and approach have relevant implications
-for research in creative cognition. A central focus of the creative cognition
-approach has been to explain acts of creativity in terms of the mental
-representations and processes that are commonly studied in cognitive psychology
-and cognitive science \citep{finke1992creative,smith1995creative}. However,
-unlike other fields in the study of cognition, creative cognition research
-rarely employs quantitative models to evaluate the explanatory value of such
-representations and processes. Our modeling results provide a concrete example
-of how formal approaches may be used to gain insight into the nature of creative
-cognition.
-
-In addition to demonstrating the utility of formal modeling for studying
-creative cognition, the contrast models here specifically offer an additional
-interpretation of some of the field's most central findings. For example,
-perhaps the most foundational principle from this literature concerns the
-limiting influence of prior knowledge: Individuals create new categories
-composed of features from existing classes, and what is created can be
-influenced drastically through the introduction of cues or examples
-\citep{marsh1999inadvertent,smith1993constraining}. In this paper, we have
-identified another important aspect of the constraining influence of prior
-knowledge: What is generated cannot be the same as what is already known.
-Further, there is systematicity in how generated categories differ from prior
-knowledge. The results of our simulations suggest that this conceptualization of
-difference can be addressed in at least two different ways. Specifically,
-simulations with PACKER demonstrate that the constraining influence of
-difference is concisely explained in terms of a trade-off between
-within-category similarity and between-category dissimilarity. Conversely, the
-representativeness model shows that this influence can also be the result of
-enhancing the representativeness of generated exemplars to their category.
-
-PACKER may offer an additional interpretation of existing accounts of creative
-generation. Most notably, a leading account within the creative cognition
-literature, the Path of Least Resistance \citep{ward1994structured,ward1995s},
-also explains generation in terms of an exemplar-based retrieval process. This
-account was designed to explain the creative generation of natural categories
-(e.g., new species of plants and animals) and as a result relies strongly on the
-hierarchical organization of these categories: Individuals are thought to
-retrieve an example of the higher-level category being generated (e.g.,
-\textit{bird} may be retrieved from the category \textit{animal}), and then
-systematically alter what was retrieved to make something new. As the PACKER
-model does not assume knowledge is hierarchically organized \citep[this is true
-of the exemplar view more broadly, see][]{murphy2016exemplar}, the model may be
-viewed as a formal instantiation of the Path of Least Resistance for application
-in a traditional artificial categorization domain (when there is no established
-hierarchy of categories). PACKER's success in explaining generation within an
-artificial domain motivates future work exploring the nature of category
-contrast within a more naturalistic setting.
-
-The broader study of creativity currently involves a wide breadth of different
-approaches\citep[for a review see][]{kozbelt2010theories}, such as those based
-on free association \citep{mednick1962associative} and conceptual combination
-\citep{estes2002emergence,murphy1988comprehending}. In addition, recent work in
-the machine learning literature has explored using neural networks to address
-the overall problem of creative generation
-\citep[e.g.,][]{goodfellow2014gan,ho2016gail,kingma2016,chen2016infogan}. In
-contrast to these varied investigations, we reduced our focus by studying a
-highly complex behavior (category generation) as it applies within a
-well-established domain (artificial category learning). We hope that future work
-incorporates and highlights the importance of contrast into theories of
-creativity.
-
-\subsection{Implications For Categorization}
-
-Categorization research addresses the representations and processes that
-underlie the learning and use of categories. Category learning tasks are
-generally about figuring out which items belong to which category. Once learned,
-categories are generally used to classify new stimuli and to make inferences
-beyond the available information. Our work is fairly unique in that people learn
-a single category through positive examples and then create another category
-that would make sense in the domain. In other words, only one category is
-provided to be learned and the use that is required of that category is to
-generate a contrast category. Such scenarios have direct application to
-real-world situations. For instance, consider being stranded in an unfamiliar
-environment where you eat and get sick from a plant with certain features. In
-avoiding such plants and seeking alternative sources of nutrition, it would
-behoove you to imagine what features an edible, safe plant would possess (e.g.,
-if the poisonous plant was bright red, perhaps a non-red plant would be safe to
-consume). In the present studies, we have learned something about the form that
-such expectations are likely to take.
-
-We can think of the category generation task in our studies as asking a person
-to formulate an idea about what set of items in the domain are most
-interestingly {\em not} members of the original category. To meet this
-condition, the items must take some form of coherence that aligns with that of
-the original category and some form of distinctiveness relative to the original
-category. Reflecting the basic level of organization in natural categories, it
-makes sense to generate a set of items that possess strong within-category
-coherence (by importing or systematically transforming the internal structure of
-the original category) as well as strong between-category differentiation (by
-creating maximum contrast with the original category, be it through
-exemplar-based dissimilarity or the maximization of exemplar-category
-representativeness). In this sense one can see the patterns of performance in
-the category generation task as recapitulating the order of semantic
-organization.
-
-The current work also suggests exciting directions for related investigations
-into unsupervised categorization. While the unsupervised categorization
-literature is primarily interested in the generation of categories within a set
-of observed exemplars where no prior categories are learned, our category
-generation studies are focused on the production of novel category exemplars
-themselves. Despite this distinction, both types of categorization research are
-ultimately interested in the question of how categories can be formed. The results
-of our current work indicate that the formation of categories is constrained by
-some measure of contrast. To our knowledge, this explicit investigation of
-category contrast has not yet been attempted in the unsupervised categorization
-literature.
-
-Interestingly, the unidimensional sort bias that is commonly observed in
-unsupervised categorization
-\citep{imai1965discriminability,milton2004influence,ahn1992two}, where
-participants sort unlabelled family resemblance category exemplars by focusing
-on a single feature, was not consistently observed in our tasks. In our
-experiments we only observed this bias in the row condition of Experiment 1,
-whereas in other conditions participants generated categories that typically
-varied along both dimensions. This may be unsurprising at face value, since
-participants in conditions other than the row condition were trained on a prior
-category that varied on both dimensions. However, these results suggest that the
-unidimensional sort bias is not universal to all types of category construction
--- or more specifically, the unidimensional bias may not apply when participants
-need to generate categories with entirely new exemplars. Additionally, the
-unidimensional sort bias is observed in category construction tasks in which
-participants develop a category-level organization of a provided domain while
-category generation tasks involve generating a new category relative to one that
-was just acquired (one might say that participants are forging a domain from a
-category rather than categories from a domain). Since the more creative task of
-generating a category from scratch (with respect to an established category)
-does not invoke the unidimensional bias, this suggests that a more creatively
-demanding version of a sort task might also reduce the bias.
-
-\subsection{Exemplar Dissimilarity and Representativeness in Categorization} 
-Although both PACKER and the representativeness model perform better than the
-classical models, they appear to be capturing fundamentally different aspects of
-contrast in categorization. Specifically, while the representativeness model
-excelled in predicting our overall category generation data, PACKER was the
-best-fitting model in accounting for categories generated from the Corner
-category type, where exemplars are located at the corners of the stimulus
-domain.
-
-% Given the history of their development, this is perhaps not entirely
-% surprising. PACKER is implemented as an extension of the GCM, which was
-% originally developed for (and has been highly successful in) category learning
-% \citep{nosofsky1986attention, nosofsky1994comparing}. Conversely, the
-% representativeness mechanism was conceived as a heuristic describing the
-% relative extent to which an individual observation is generated from some
-% unobserved population \citep{kahneman1972subjective,abbott16}. This generative
-% nature of the representativeness mechanism intuitively lends itself to the
-% prediction of category generation data.
-
-While the contrast models were born quite naturally out of the corresponding
-classical models, it is worth considering if similar benefits in prediction
-would be seen if the contrast mechanisms were applied to different classical
-models. Although there is no clear way to adapt the exemplar dissimilarity
-mechanism to the current hierarchical Bayesian model, it is a fairly
-straightforward process applying the representativeness mechanism to an exemplar
-model. Specifically, by treating the similarity measure (Equation
-\ref{eq:similarity}) in the current implementation of copy-and-tweak as a
-density estimate for the representativeness mechanism (Equation
-\ref{eq:representativeness}), we gain a new model of representativeness that is
-based on exemplar similarity instead of a multivariate Gaussian
-likelihood.\footnote{An analogous implementation of the representativeness
-  mechanism to PACKER results in both $\theta_c$ and $\theta_t$ adding to
-  constant value that is independent of the similarity to contrast and target
-  exemplars respectively. Consequently, it is formally equivalent to a
-  copy-and-tweak model with representativeness.}
-
-How well does a representativeness model that uses exemplar similarity to
-represent the category’s density over features explain our experimental results?
-We find that this model substantially outperforms its classical counterpart, but
-not to the extent of PACKER or the representativeness model. Specifically,
-fitting this model to the category generation data yielded $L = -4815$ and
-$AIC=9633$ for the entire set of data from Experiments 1 and 2, and $L = -3474$,
-$AIC=6953$ for data excluding the first trials. When used to predict the error
-rates of participants in the category learning task (Experiment 3), the
-performance of the copy-and-tweak with representativeness model correlated with
-participant error with $r=.56, p < .001$. These results suggest that the
-predictive advantages to including the representativeness mechanism is not
-limited to a hierarchical Bayesian model, but can also be found when applied to
-an exemplar model. Although its performance is not as strong as either of the
-fully-developed contrast models, there is a benefit to the representativeness
-mechanism that is independent from any interaction with a hierarchical Bayesian
-framework.
-
-Ultimately, the source of the discrepancy between the performance of the two
-contrast mechanisms in category generation is still not entirely clear. Our
-results could be an indication of a fundamental difference between the processes
-involved where exemplar dissimilarity is a larger factor under certain category
-configurations while representativeness exerts more influence in others.
-Alternatively, it is possible they are both approximating some factor that we
-still have not fully identified. While a thorough investigation is beyond the
-scope of the current article, the preliminary results of this section indicate
-that further work to disentangle these two mechanisms would be promising. In the
-following section, we explore some other limitations of the current work and
-provide more recommendations for future research.
-
-\subsection{Limitations and Future Directions}
-
-Although successful in explaining our results in all three studies (albeit not
-as well as the representativeness model for Experiments 1 and 2), PACKER does
-not provide a full account of what is known about category generation. Most
-notably, in this paper we have not evaluated the model's ability to explain the
-classic finding that generated categories tend to share distributional
-commonalities with previously learned categories
-\citep[see][]{jern2013probabilistic,ward1994structured}. While we successfully
-replicated this effect in Experiment 1, we also found that its influence was
-limited in comparison to the fundamental constraints imposed by category
-contrast. Even within Experiment 1, we found systematic inconsistencies: by
-generating exemplars into unoccupied regions of the space, participants who
-learned an `XOR' category, composed of members that are widely distributed along
-both features and are positively correlated in space, tended to generate
-categories with an opposite (negative) correlation. More generally, PACKER
-inherits the strengths and weaknesses of exemplar models of categorization: It
-provides a simple and flexible model that explains many results, but deviates
-systematically from human performance in some cases.
-
-Nonetheless, these classic effects are a core element of the phenomenology of
-category generation, and PACKER does not include any mechanisms that explain
-them. Instead, through the development and evaluation of the PACKER model, we
-have sought to add new elements into such a phenomenology: The broad and strong
-influence of category contrast, and the interdependence between category
-location and distributional structure. It may be possible to combine the
-hierarchical Bayesian approach proposed by \cite{jern2013probabilistic} with
-PACKER's underlying claims to obtain a ``best of both worlds'' model, capable of
-explaining the role of contrast in category generation, as well as the emulation
-of distributional structure. However, as noted in the introduction, the
-incorporation of category contrast is antithetical to the core principles of a
-traditional, semi-conjugate Bayesian approach. This suggests that category
-generation is a fundamentally different computational-level problem
-\citep[different from those posed by][]{jern2013probabilistic,kemp2014taxonomy}.
-
-Characterizing that problem and conducting a rational analysis is an important
-direction for future research. To that aim, we plan to explore the connection
-between exemplar modeling as an Importance Sampling approximation
-\citep{shi10exemplar}, and see what sort of computational-level problem PACKER
-approximates. Once formalized in probabilistic terms, it should also be
-straightforward to incorporate distributional factors into the model. This would
-unite PACKER with the Bayesian representativeness model, and they would differ
-in terms of their assumptions about how people represent category distributions
-(as exemplars or prototypes, respectively). Alternatively, it may be possible to
-integrate the core principles of either model of contrast into other
-categorization models
-\citep[e.g.,][]{love2004sustain,kurtz2007divergent,smith2000thirty}.
-
-The present work focuses on the influence of contrast on categorization,
-primarily by exploring two different conceptualizations of contrast. However,
-throughout this paper we have only explored one type of category generation
-problem, that is the generation of new categories of new categories in an
-artificial domain. It is possible that the strength of the influence of contrast on
-category generation decisions varies depending on the nature of the category
-generation problem, for instance, in situations where the boundaries of the
-domain are not clearly defined. To provide some intuition for this, consider the
-following example: an entomologist is asked to draw a new type of insect that
-they have never seen. Clearly, features of the new insect will be constrained by
-the definition of an insect: it will be an arthropod with six legs and a
-relatively hard exoskeleton. However, there are no strict limitations on other
-features such as how large the insect needs to be, or if the insect should have
-a particular type of pattern on its body. In this scenario, it is plausible that
-the entomologist will draw an insect that is fairly similar to insects that
-currently exist (e.g., small in size with a camouflage pattern that matches its
-environment) as opposed to something completely different (e.g., a horse-sized
-beetle with roman numerals on its back), indicating that contrast plays only a
-minor role in the decision making process. In this situation it appears that
-there is a combination of clearly defined constraints on the domain boundary
-that are imposed by the external environment (insects must be arthropods with
-six legs) but also ill-defined constraints that are imposed by the observer
-(insects are probably small and have patterns on their body that suit the
-environment). However, the relation of these contraints (and the interactions
-between them) to the role of contrast is not currently well understood.
-
-The definition of the domain boundary is only one possible factor that could
-influence the effect of contrast on categorization. It is also plausible that
-other conditions such as the nature of the instruction (``Generate a new
-category that is Not Alpha'' vs ``Generate a new category that is Beta'') 
-can affect the extent to which contrast influences categorization. Though our
-current contrast models can allow the influence of contrast to vary, they do not
-make any predictions regarding the specific conditions under which contrast can
-vary. Consequently, addressing this issue would be an immensely promising avenue
-for future research.
-
-\section{Conclusions}
-
-The generation of new concepts and ideas is a highly interesting topic, but it
-is difficult to study in a controlled experimental environment. In this paper,
-we have provided such an examination of category generation as it applies within
-an artificial categorization experiment. Extending the literature on creative
-cognition, our experiments provide a detailed picture of the role of category
-contrast in generation: People seek to create concepts that are distinct from
-what they already know, and the nature of what is created can be influenced by
-what does not yet exist. Our simulations with traditional exemplar models, as
-well as a hierarchical Bayesian model, provide strong support for the claim that
-category contrast is of fundamental importance to categorization. More broadly,
-our results demonstrate that popular explanatory approaches from basic research
-in cognitive science can offer a precise, quantitative account of a behavior as
-complex as that of category generation, and in exploring them, we can increase
-our understanding of categorization as well.
-
-\clearpage
-\section{Acknowledgments} Previous versions of this work were presented at the
-Thirty-Ninth Annual Conference of the Cognitive Science Society and Forty-Ninth
-Annual Meeting of the Society for Mathematical Psychology. Support for this
-research was provided by the Office of the VCRGE at the UW - Madison with
-funding from the WARF. We thank Charles Kemp for feedback on a previous version
-of this manuscript. We also thank him and Alan Jern for providing code and data.
-\end{flushleft}
-
-% references section
-\clearpage \bibliographystyle{apacite} \bibliography{citations.bib}
-\clearpage
-
-\appendix \numberwithin{equation}{section}
-
-\section{The Hierarchical Bayesian Model of Concept Generation}
-\label{ap:hsampling-definition}
-
-\cite{jern2013probabilistic} demonstrated how a hierarchical Bayesian model
-could explain the distributional correspondences between observed and generated
-categories. In their model, exemplars of generated categories were viewed as
-samples from a multivariate Normal distribution over the dimensions of stimulus
-space. The mean of the generated category was independent of the observed
-categories, but the covariance matrix (encoding feature variances and
-correlations) was based on a common prior distribution. Generating a new
-category was thus completed by sampling a new category mean (uniform over
-stimulus space) and covariance matrix from the common prior distribution.
-Because the shared prior distribution's parameters were unobserved, the
-hierarchical Bayesian approach was used to infer its parameters from the
-previous categories (their feature variances and correlations), and then to
-generate the covariance matrix of the new category.
-
-In our implementation of their model\footnote{Note that
-\citet{jern2013probabilistic}'s model is slightly different, as they used a
-semi-conjugate model. Their model acts very similarly to our version.}, each
-category's exemplars are viewed as samples from a multivariate Normal
-distribution with parameters ($\mu, \Sigma$). Category covariance matrices
-(specifying variance and covariance along $k$-dimensions), are assumed to be
-Normal-Inverse-Wishart distributed with parameters: $\nu$ ($>k-1$), $\kappa$
-($>0$), and $\Sigma_D$. $\nu$ and $\kappa$ are treated as free parameters in our
-simulations, and $\Sigma_D$ is the domain-wide covariance matrix from which all
-categories are viewed as samples. Assuming a given $\Sigma_D$, a category
-covariance matrix $\Sigma$ can be computed on the basis of its examples:
-
-\begin{equation} \Sigma = \left[ \Sigma_D \nu + C + \dfrac {\kappa n} {\kappa +
-n} (\bar{x}-\mu)(\bar{x}-\mu)^T \right] (\nu + n)^{-1}
-\label{eq:Sigma_B}
-\end{equation}
-% 
-where $\bar{x}$ and $C$ are the empirical mean and covariance of the category's
-known members, and $n$ is the number of observed members of the category. When
-there are fewer than two known members of the category (and thus no covariance
-to speak of), $\Sigma = \Sigma_D\nu$.
-
-The category mean, $\mu$, can be computed as:
-
-\begin{equation} \mu = \dfrac {\kappa\mu_{0} + n \bar{x}} {\kappa + n}
-    \label{eq:category_mus_jk13}
-\end{equation}
-% 
-where $\mu_{0}$ is the prior mean. In our simulations, $\mu_{0}$ is set to the
-center of the domain. However, when no examples of the target category have been
-observed, generation is assumed to be random. In practice, the model's best fits
-are achieved when the $\kappa$ parameter, which controls the influence of
-$\mu_0$ on $\mu$, is set very close to zero (hence, the influence of $\mu_0$ is
-minimal).
-
-Importantly, the domain-wide covariance matrix $\Sigma_D$ is unobserved and
-needs to be inferred from the observed categories. For conjugacy, if $\Sigma_D$
-is viewed as a sample from an Inverse-Wishart distribution with scale
-$\Sigma_0$, $\Sigma_D$ can be computed as:
-
- \begin{equation} \Sigma_D = \Sigma_0 + \sum_y{C_y}
-    \label{eq:category_sigmas}
-\end{equation}
-% 
-where $\Sigma_0$ is the prior covariance in the domain. In our simulations,
-$\Sigma_0 = \lambda {\bf I}$, where $\lambda$ is a free parameter controlling
-the expected variance of dimensions (dimensions of the domain covariance matrix
-are expected to be uncorrelated) and ${\bf I}$ is a $k$-by-$k$ identity matrix.
-
-Generated exemplars are drawn from a multivariate Normal distribution specified
-by $(\mu, \Sigma)$. Thus, $p(y)$ is
-\begin{equation} p(y \mid x) = \dfrac {\exp \left \{ \theta \cdot {\rm
-Normal}(y; \mu, \Sigma) \right \} } {\sum_i \exp \left \{ \theta \cdot {\rm
-Normal}(y_i; \mu, \Sigma) \right \} }
-\end{equation}
-% 
-where $\theta$ is a response determinism parameter and ${\rm Normal}(y; \mu,
-\Sigma)$ denotes a multivariate Normal density evaluated at $y$.
-
-\end{document}

### 2019-07-09 Manuscripts/post-cogpsych/post-cogpsych.tex
Modelling for hybrid rep+gcm models finally done. Fixed comments as suggested by Joe, also improved writing in some areas (I think). Ready to send to ken+nolan.

--- 2019-06-16 Manuscripts/cog-psych/post-revision/post-cogpsych.tex
@@
 this domain has focused on how the statistical structure of known categories
-generalizes to generated categories, overlooking whether (and if so, how) contrast
-between the known and generated categories is a factor. We report three
+generalizes to generated categories, overlooking whether (and if so, how)
+contrast between the known and generated categories is a factor. We report three
 experiments demonstrating that contrast between what is known and what is
@@
 
+To our knowledge, no prior literature has yet explicitly discussed any empirical
+effects of contrast in category generation. However, some evidence of contrast
+can be seen in the data from Experiment 3 of cite{jern2013probabilistic}. We
+leave a detailed analysis of this data to the supplemental materials.
+
 \subsection{Contrast as exemplar dissimilarity: The PACKER Model}
@@
 Exceptions to this pattern can occur (i.e., the generation of novel exemplars
-similar to known category exemplars) when the known category distribution is
-especially large on all features -- we explore this in more detail in Section
+similar to known category exemplars) -- we explore this in more detail in Section
 \ref{section:diffcon} where we highlight an important condition demonstrating
@@
 \subsection{Discussion} In Experiment 1 we evaluated the influence of category
-contrast on category generation, given qualitatively different types of categories. We found strong evidence for effects of category contrast in each
+contrast on category generation, given qualitatively different types of
+categories. We found strong evidence for effects of category contrast in each
 condition: Participants were more likely to generate stimuli that are more
@@
 
-For the representativeness model, adding an unbounded feature changes the probability distribution for the underlying Bayesian model.
-Whereas it is still Gaussian in one feature, the other feature is toroidal, meaning that it wraps back to $0$ radians once it reaches $2\pi$ (we have conducted a change of variables where $y \rightarrow y \mod 2\pi $, where $\mod$ is the modulus operator).
-As we cross $2\pi$ and reach $0$ again, the remaining probability density also gets wrapped around, meaning that the probability density assigned by the Gaussian distribution on the toroidal dimension at $2\pi$ is actually assigned to $0$ (and $2\pi + \epsilon$ to $\epsilon$ so on).
-This is called the {\em wrapped} Gaussian distribution \citep{mardia1972}.
-Note that the often used distribution for circular dimensions, the Von Mises distribution is an approximation to the wrapped Gaussian distribution.
-See Appendix B for implementational details.
+For the representativeness model, adding an unbounded feature changes the
+probability distribution for the underlying Bayesian model. Whereas it is still
+Gaussian in one feature, the other feature is toroidal, meaning that it wraps
+back to $0$ radians once it reaches $2\pi$ (we have conducted a change of
+variables where $y \rightarrow y \mod 2\pi $, where $\mod$ is the modulus
+operator). As we cross $2\pi$ and reach $0$ again, the remaining probability
+density also gets wrapped around, meaning that the probability density assigned
+by the Gaussian distribution on the toroidal dimension at $2\pi$ is actually
+assigned to $0$ (and $2\pi + \epsilon$ to $\epsilon$, and so on). This is called the
+{\em wrapped} Gaussian distribution \citep{mardia1972}. Note that the often used
+distribution for circular dimensions, the Von Mises distribution, is an
+approximation to the wrapped Gaussian distribution. See Appendix B for
+implementational details of the representativeness model in toroidal space.
 
@@
 
-We recruited 89 participants from Amazon Mechanical Turk who were randomly assigned to the bounded, square stimuli ($N=46$), or to the toroidal, circle stimuli ($N=43$).
-The square stimuli were exactly the same as in Experiments 1 and 2.
-The circle stimuli were made up of unfilled circles with a radial line extending from its center to its edge.
-These stimuli varied along two features -- their
-diameter ranging from 3.0--5.8 cm, and their orientation ranging from
-0--$2\pi$ radians. The orientation of each circle stimuli was fixed at a
-randomly selected angle for all participants. Similar to Experiments 1 and 2,
-participants first observed a training phase where four Alpha exemplars were
-presented across three blocks followed by the generation phase. The training condition was the Corner category for all participants.
+We recruited 89 participants from Amazon Mechanical Turk who were randomly
+assigned to the bounded, square stimuli ($N=46$), or to the toroidal, circle
+stimuli ($N=43$). The square stimuli were exactly the same as in Experiments 1
+and 2. The circle stimuli were made up of unfilled circles with a radial line
+extending from its center to its edge. These stimuli varied along two features
+-- their diameter ranging from 3.0--5.8 cm, and their orientation ranging from
+$2.587$--$2\pi + 2.587$ radians, with $0$ radians indicating a completely
+horizontal line. The line widths of the circles and radial line were set to 0.05
+cm. Similar to Experiments 1 and 2, participants first observed a training phase
+where four Alpha exemplars were presented across three blocks followed by the
+generation phase. The training condition was the Corner category for all
+participants.
 
@@
 
-The contrast models were fit by optimizing their fit to all experiments in this article (Experiments 1 through 3).
-Fit statistics and parameter values are presented in Table \ref{table:corner-model-fits}.
-Overall,
+The contrast models were fit by optimizing their fit to all experiments in this
+article (Experiments 1 through 3). Fit statistics and parameter values are
+presented in Table \ref{table:corner-model-fits}. Although not our main focus in
+this section, we also provide the statistics for the classical models --
+unsurprisingly, they do not perform as well as the contrast models. Overall,
 the representativeness model emerges as the best-fitting model in aggregate, although by a much smaller amount than Experiments 1 and 2. However, analyzing data in aggregate can be misleading.
@@
 
-\begin{tabular}{ l l} \\ \textbf{PACKER} & \textbf{Representativeness} \\
-  \hline $AIC = 12052$ & $AIC = 11991$ \\ 
-  $L = -6023$          & $L = -5991$ \\ 
-  $c = 0.60$           & $\kappa = 5.91$  \\ 
-  $\theta_c = 2.78$    & $\nu = 1.00$ \\
-  $\theta_t = 3.17$    & $\lambda = 1.45$ \\
-                       &$\theta = 1.86$ \\
+\begin{tabular}{ l l l l} \\ \\ \textbf{PACKER} & \textbf{Copy \& Tweak} &
+\textbf{Representativeness} & \textbf{Hierarchical Bayesian} \\ 
+  \hline $AIC = 12023$ &$AIC = 12779$ &$AIC = 11982$  &$AIC = 12976$\\ 
+  $L = -6008$          &$L = -6388$   &$L = -5987$    &$L = -6484$\\ 
+  $c = 0.51$           &$c = 3.24$    &$\kappa = 6.73$&$\kappa = 0.00$  \\ 
+  $\theta_c = 3.03$    &$\theta_c = 0$ (fixed)  &$\nu = 1.00$ &$\nu = 3.27$\\
+  $\theta_t = 3.46$    &$\theta_t = 2.90$       &$\lambda = 1.90$ &$\lambda = 0.07$\\
+                       &              &$\theta = 2.45$ &$\theta = 3.15$\\
 \end{tabular}
@@
 that both constraints influence generation when explained with an exemplar
-model. Ultimately, our category generation data were best explained by the
-representativeness model. The exact nature of its superiority over the other
-models is still not yet clear, although we find evidence that it only barely
-surpasses PACKER when considering known categories with extreme exemplars.
-As we explore in a later section, it is likely that the
-representativeness mechanism itself (i.e., independent of the hierarchical
-Bayesian foundation) captures a core aspect of both category generation and
-category learning.
+model.
+
+While we found that the representativeness model was found to be a better fit to
+data from Experiments 1 and 2, both contrast models essentially made
+qualitatively similar predictions across all conditions in those experiments. In
+Experiment 3, we collected behavioral data on a category generation condition
+that qualitatively distinguished the two contrast models. Specifically, in this
+condition where exemplars were located at the corners of the stimulus space,
+PACKER predicted the initial generation of centrally-located novel exemplars
+while the representativeness model predicted novel exemplars similarly located
+at the corners. Interestingly, data from this experiment was better fit,
+qualitatively and quantitatively, by PACKER than the representativeness model.
+
+% Ultimately, our category generation data were best explained by the
+% representativeness model. The exact nature of its superiority over the other
+% models is still not yet clear, although we find evidence that it only barely
+% surpasses PACKER when considering known categories with extreme exemplars.
 
@@
 classical models, they appear to be capturing fundamentally different aspects of
-contrast in categorization. Specifically, the representativeness model
-excelled in predicting categories generated by participants in Experiments 1 and 2.
-Conversely, PACKER was the best-fitting model in accounting for categories generated from the Corner category type, especially when the stimuli had a toroidal feature.
+contrast in categorization. Specifically, the representativeness model excelled
+in predicting categories generated by participants in Experiments 1 and 2.
+Conversely, PACKER was the best-fitting model in accounting for categories
+generated from the Corner category type, especially when the stimuli had a
+toroidal feature.
 
@@
 represent the category’s density over features explain our experimental results?
-We find that this model substantially outperforms its classical counterpart, but
-not to the extent of PACKER or the representativeness model. Specifically,
-fitting this model to the category generation data yielded $L = -4815$ and
+We find that for data across all experiments, this model substantially
+outperforms its classical counterpart, but not to the extent of PACKER or the
+representativeness model. Specifically, the model fits yielded $L = -4815$ and
 $AIC=9633$ for the entire set of data from Experiments 1 and 2, and $L = -3474$,
-$AIC=6953$ for data excluding the first trials. When used to predict the error
-rates of participants in the category learning task (Experiment 3), the
-performance of the copy-and-tweak with representativeness model correlated with
-participant error with $r=.56, p < .001$. These results suggest that the
-predictive advantages to including the representativeness mechanism is not
-limited to a hierarchical Bayesian model, but can also be found when applied to
-an exemplar model. Although its performance is not as strong as either of the
-fully-developed contrast models, there is a benefit to the representativeness
-mechanism that is independent from any interaction with a hierarchical Bayesian
-framework.
-
-Ultimately, the source of the discrepancy between the performance of the two
-contrast mechanisms in category generation is still not entirely clear. Our
-results could be an indication of a fundamental difference between the processes
-involved where exemplar dissimilarity is a larger factor under certain category
-configurations while representativeness exerts more influence in others.
-Alternatively, it is possible they are both approximating some factor that we
-still have not fully identified. While a thorough investigation is beyond the
-scope of the current article, the preliminary results of this section indicate
-that further work to disentangle these two mechanisms would be promising. In the
-following section, we explore some other limitations of the current work and
-provide more recommendations for future research.
+$AIC=6953$ for the same data excluding the first trials. When we fit the model
+to data from Experiments 1,2 and 3, we find that $L = -6301$, $AIC = 12606$.
+These results suggest that the predictive advantages to including the
+representativeness mechanism is not limited to a hierarchical Bayesian model,
+but can also be found when applied to an exemplar model. Although its
+performance is not as strong as either of the fully-developed contrast models,
+there is a benefit to the representativeness mechanism that is independent from
+any interaction with a hierarchical Bayesian framework.
+
+Ultimately, it appears that there is no single model that can capture the
+entirety of our empirical data. While the representativeness model, and indeed
+the representativeness mechanism itself, offers a predictive advantage in
+category generation across all out models, PACKER performs better when
+accounting for data from the Corner conditions. While we have established that contrast
+mechanisms are essential in modeling category generation, there is no conclusive
+evidence that one form of contrast is superior to the other across all
+conditions. 
+%Xian: Consider rewriting paragraph if there's time.
 
@@
 
+\newpage
+\section{Representativeness Model in Toroidal Space.}
+\label{ap:wrapgauss}
+
+Assuming a single bounded feature, we can easily represent the probability
+density of a function at any point $x$ on a line as $f(x)$. With a single
+unbounded feature, the points $x$ are mapped onto a unit circle with the
+corresponding $f(x)$ ``wrapped'' around the circle. More formally, we can
+represent the probability density along a wrapped axis $f'(x)$ as:
+\begin{align}
+f'(x) &= ... + f(x-4\pi) + f(x-2\pi) + f(x) + f(x+2\pi) + f(x+4\pi) + ...\\
+     &= \sum_{i=-\infty}^{\infty} f(x + 2\pi i)
+\end{align}
+
+We implement the wrapped axis in the representativeness model when
+calculating the multivariate Gaussian components of the representativeness equation
+$R(x,h)$ (Equation \ref{eq:representativeness} of the main text.) Specifically,
+the probability density of the Gaussian component along a single wrapped axis can be computed as:
+\begin{align}
+p'(x|h) = \sum_{i=-\infty}^{\infty} p(x + 2\pi i|h)
+\end{align}
+where $p(.|h)$ represents the probability density of the Gaussian component $h$
+on a line (i.e., not wrapped around a unit circle.)
+
+In practice, we can approximate the infinite summation by constraining $i$
+within a limited range (e.g., $[-50,50]$). For increased efficiency, we can also
+approximate $p'(x|h)$ by stopping the summation once the ratio of the maximum
+and minimum values of $p(x + 2\pi i|h)$ computed thus far is large enough (e.g.,
+10,000.) For our modeling exercises here we impose both contraints, although the
+first constraint shows some redundancy since we find that $i$ rarely, if ever,
+exceeds a value of 4. Given the unimodal nature of the Gaussian distribution,
+this procedure allows for comparatively efficient approximation of the wrapped
+distribution without needing to compute the infinite sum.
+
 \end{document}
-
-\subsection{Finding Contrast Effects in Prior Work}
-
-Although existing accounts of category generation broadly overlook the role of
-category contrast in determining what is novel versus familiar, it was
-implicitly assumed that learners in previous experiments were {\em successful}
-in creating new categories. Thus, if contrast plays a robust role in category
-genration, its effects should be discernable within the experimental results of
-these studies. To provide a test of the influence of category contrast within
-existing data, we conducted a novel analysis of Experiment 3 from
-\cite{jern2013probabilistic}.\footnote{Although \cite{jern2013probabilistic}
-	reported four experiments, we focus on Experiment 3 because their first two
-	experiments tested generation of items into known categories, and their fourth
-	experiment was identical to Experiment 3 but with a restricted generation
-	space.}
-
-Participants in their experiment were exposed to members of two
-experimenter-defined categories of `crystals' varying in hue, saturation, and
-size. Each category possessed a unique hue, but varied in saturation and size:
-In the `Positive' condition, there was a positive correlation between these
-features (i.e., larger sized crystals were more saturated), and in the
-`Negative' condition this relation was reversed. In the `Neutral' condition,
-there was no correlation between saturation and size. After learning about the
-categories from each condition, participants were asked to generate six
-exemplars belonging to a novel class. As noted above,
-\cite{jern2013probabilistic} found that the generated categories tended to
-follow the distributional properties of the experimenter-defined categories:
-Generated categories were tightly distributed along the hue feature, and
-possessed the same saturation-size correlations as in the learned categories.
-\cite{jern2013probabilistic}, however, did not analyze or discuss how the
-generated categories {\em differed} from the experimenter-defined categories.
-
-Because each experimenter-defined category in the \cite{jern2013probabilistic}
-experiment possessed a distinct hue shared by all members of the category, it is
-sensible that participants might generate a category with a hue distinct from
-the experimenter-provided categories. If category generation were influenced by
-category contrast in this way, the hues of generated categories should be
-systematically different from those of the experimenter-defined categories.
-Unfortunately, stimulus hue was encoded and presented in the
-Hue-Saturation-Value (HSV) color space, which is device-dependent and not
-perceptually normed such that perceived color similarity corresponds to
-proximity in the color space (as opposed to a color space such as CIELAB that is
-device-independent and equidistant sets of points correspond to pairs of colors
-that have the same perceptual similarity; \citealp{wyszecki1967}). Further, they
-did not calibrate their monitor, and so we cannot know the precise colors
-presented to participants. As \cite{jern2013probabilistic} were interested in
-relations between the saturation and length of examples in generated into novel
-categories, these issues do not undercut their analyses and results. However,
-these issues pose a significant challenge to evaluating contrast between the
-experimenter-defined and participant-generated categories along the hue
-dimension. It is plausible that two uncalibrated monitors could display the same
-HSV color and the colors be perceived in different color categories (especially
-for color boundaries that vary over lightness, such as the yellow-brown
-boundary).
-
-Although we cannot know the precise colors that were displayed or perceived, we
-can still analyze their results from a coarse perspective to see whether there
-is preliminary support for contrast. To do so, we binned all possible hues into
-one of eight uniformly-spaced color groups: \{{\em Red}: $0-0.063$, $0.938-1$,
-{\em Yellow}: $0.063-0.188$, {\em Yellow-Green}: $0.188-0.313$, {\em
-	Green-Teal}: $0.313-0.438$, {\em Teal}: $0.438-0.563$, {\em Teal-Blue}:
-$0.563-0.688$, {\em Purple}: $0.688-0.813$, {\em Pink}: $0.813-0.938$\}. In the
-\cite{jern2013probabilistic} experiment, the hue of each experimenter-defined
-category was selected from one of six possible values, each of which falls into
-one of the color groups above (two color groups were not used as a possible hue
-for the experimenter-defined categories). By categorizing the
-participant-generated crystals likewise, we can obtain a broad measure of
-category contrast by determining the proportion of participant-generated
-crystals that fall into the same groups as the experimenter-defined categories:
-If contrast influences the hues of the generated categories, we should observe
-minimal overlap between in the color groupings.
-
-\begin{figure}
-	\begin{center} \inputpgf{figs/}{jk13-huecontrast.pgf}
-		\caption{Analysis of data from \cite{jern2013probabilistic}, Experiment
-			3. Plotted is the number of generated items that share a color group with one of
-			the experimenter-defined classes. The ``Expected'' data follows a Binomial
-			distribution with $p = 2/8 = 1/4$, given there were two experimenter-defined
-			classes, and eight color groups.}
-		\label{fig:jk13-huecontrast}
-	\end{center}
-\end{figure}
-
-These data, shown in Figure \ref{fig:jk13-huecontrast}, reveal a clear pattern:
-The majority of participants in each condition ($n = 22$) generated categories
-possessing entirely distinct hues; with 0/6 exemplars sharing a hue with the
-experimenter-defined categories. These results can be compared to the
-predictions of a Binomial model, which proposes that participants generate hues
-at random. That is, if hue selection is not systematic, the probability that any
-given example will lie in the same color group as an experimenter-defined
-category is given by a Binomial distribution with $p = 2/8=1/4$, as there were
-two experimenter-defined categories and eight possible color groups. Chi-square
-goodness-of-fit tests reveal that the observed distribution in each condition is
-highly inconsistent with the hues being chosen at random, all
-$\chi^2(6,N=22)>200$, $p < .001$. Participants tended to generate items that were
-perceptually distinct from the categories they had learned, and were less likely
-to generate hues possessed by members of the experimenter-defined categories.
-
-It is possible, however, that this data can be explained by a process that does
-not involve contrast. Specifically, since participants were trained on exemplars
-that share the same hue within a category, it is plausible that they generate
-exemplars that all share a common hue. If this were the case, given a $6/8 = .75$
-probability that a hue distinct from the experimenter-defined categories is
-selected, there is a corresponding $.75$ probability (or about 17 out of 22
-participants) that the generated category will have no exemplars that share a
-hue with any  of the experimenter-defined categories. In addition, that there is
-a $2/8 = .25$ probability (about 6 out of 22 participants) that the generated
-category will have all six exemplars that share a hue with any of the
-experimenter-defined categories. This prediction of regularity across categories
-is much closer to the data than the prediction from a Binomial model and does
-not require any mechanism of category contrast.
-
-However, upon analyzing the consistency of hues in the generated categories, it
-is clear that participants do not generally produce categories with entirely
-similar hues across all exemplars. Specifically, only 55\%  of all generated
-categories comprised exemplars with a common hue. Consequently, the assumption
-of regularity across categories may be inappropriate for this data set, leaving
-the contrast-based explanation as the most plausible for our purposes.
-
-Re-analyzing the results from \cite{jern2013probabilistic} provides some
-corroborating support for contrast playing a role in category generation. Taken
-alongside the analyses reported by \cite{jern2013probabilistic}, our analysis
-suggests that generated categories tend to be distinct from {\em and}
-distributionally similar to what is already known. However, it is worth noting
-that our analysis is still limited: The color groups defined above are
-imprecise, and it is not clear that our color grouping is consistent with the
-psychological color boundaries perceived by participants. While we did obtain
-similar results using a variety of alternative groupings, the hue dimension used
-in the \cite{jern2013probabilistic} study does not lend itself straightforwardly
-to the computation of similarities, and thus we cannot be certain of whether our
-coding accurately approximates the psychological space of the stimuli. This
-precludes traditional applications of categorization models to their data as it
-is usually necessary to encode objects in psychological space in order to
-accurately determine the similarity between objects. By consequence, although
-these results likely indicate that contrast exerts {\em some} influence, they do
-not precisely describe the nature of that influence.

### 2019-07-29 Manuscripts/model-descriptions/negative-weighted/four_models.tex
Updates packer manuscript descriptions (long overdue.) Fixing some representativeness weirdness.

%!TEX output_directory = latex_out/

\documentclass[12pt]{article} \usepackage[letterpaper, margin=1in,
headheight=15pt]{geometry} \usepackage{amsmath, listings}

\begin{document}

\section{Generating Exemplars from Categories-by-Negation}

When generating exemplars from categories defined by a label that describes what is \emph{not} already learned, we can
define a simple model where these exemplars are samples drawn with a probability proportional to the inverse (or 1-?)category
distribution of

An interesting property of such a model is that it assumes that categories-by-negation are multi-modal, unlike the
unimodal assumptions currently adopted by current models of category generation.

### 2019-07-29 Experiments/corner/analysis/plot.individuals.py

@@ -7,6 +7,8 @@ import pickle
 execfile('Imports.py')
 import Modules.Funcs as funcs
 
+savefig = False
+
 pd.set_option('precision', 2)
 
 # import data
@@ -22,7 +24,8 @@ savedir = 'individuals'
 gentypeStr = ['N','B','C'] #not alpha, only beta, beta-gamma
 gentypeStrDisp = ['A\'','B','C'] #not alpha, only beta, beta-gamma
 gentypeCols = [[.3,0,.5],[0,0,.5],[0,.5,0]]
-f, ax= plt.subplots(1,1, figsize=(1.6, 1.6))
+if savefig:
+    f, ax= plt.subplots(1,1, figsize=(1.6, 1.6))
 
 var_thresh = .25
 cor_thresh = .7
@@ -72,15 +75,18 @@ for i, row in info.iterrows():
         printstr += gentypeStr[ii+1] + ' ' + group + '\n'
         groupcount[group] += 1
 
-    funcs.plotclasses(ax, stimuli, palphas, pbetas_all, betastr=betastr,betacol = betacol)
+
     data_arr += [[pid, group]] #not quite accurate for gammas?
-    fname = os.path.join(savedir,condition + '-' + gentypeStr_p + '-' + str(pid) + '.png')
-    ax.text(.9,1.25,str(pid))
-    ax.text(-.9,1.0,printstr)
-    f.savefig(fname, bbox_inches='tight', transparent=False)
-    plt.cla()
+    if savefig:
+        funcs.plotclasses(ax, stimuli, palphas, pbetas_all, betastr=betastr,betacol = betacol)
+        fname = os.path.join(savedir,condition + '-' + gentypeStr_p + '-' + str(pid) + '.png')
+        ax.text(.9,1.25,str(pid))
+        ax.text(-.9,1.0,printstr)    
+        f.savefig(fname, bbox_inches='tight', transparent=False)
+        plt.cla()
 
 print(groupcount)
 data = pd.DataFrame(columns=('participant','betagroup'),data = data_arr)
+#print(data)
 with open('individual_group_counts.p','wb') as f:
     pickle.dump(data,f)

### 2019-08-22 Manuscripts/post-cogpsych/post-cogpsych.tex
Fixed last minute typos
Reorganised stimsteps -- now it is intialised when trialsets are initialised.

@@
 effects of contrast in category generation. However, some evidence of contrast
-can be seen in the data from Experiment 3 of cite{jern2013probabilistic}. We
+can be seen in the data from Experiment 3 of \cite{jern2013probabilistic}. We
 leave a detailed analysis of this data to the supplemental materials.
@@
 
-For the representativeness model, the single unbounded feature essentially
-prevents the formation of edge effects along that feature.
+The single unbounded feature prevents ... While the representativeness model
+more closely resembles PACKER under this condition, note that the former still
+more strongly predicts the generation of the first Betas exemplars at the
+bounded edges compared to PACKER (Figure \ref{fig:e3-circles}c.) 
+% For the representativeness model, the single unbounded feature essentially
+% prevents the formation of edge effects along that feature.
 Consequently, this results in higher probabilities of the generating the first Beta exemplars along the edges of the bounded feature (Figure \ref{fig:e3-circles}c).
 As this
-model does not explicitly recognize similarities as distances along the feature
-space (unlike PACKER), it does not experience the compression in density regions
-along the unbounded feature axis. Intuitively, this suggests that the
-representativeness model (which prefers generating exemplars at the boundaries) will prefer to generate stimuli on the boundaries of the bounded feature with some amount of variance over unbounded feature.
+% model does not explicitly recognize similarities as distances along the feature
+% space (unlike PACKER), it does not experience the compression in density regions
+% along the unbounded feature axis. Intuitively, this suggests that the
+% representativeness model (which prefers generating exemplars at the boundaries) will prefer to generate stimuli on the boundaries of the bounded feature with some amount of variance over unbounded feature.
 
@@
 \subsection{Results and Discussion}
+
+%Repeat hypotheses
 
@@
 \newpage
-\section{Representativeness Model in Toroidal Space.}
+\section{Representativeness Model in Toroidal Space}
 \label{ap:wrapgauss}

### 2019-08-22 Modules/Classes/Simulation.py

@@ -21,7 +21,7 @@ class Trialset(object):
         
         # figure out what the stimulus domain is        
         self.stimuli = stimuli
-        self.stimrange = funcs.getrange(stimuli)
+        self.stimrange,self.stimstep = funcs.getrange(stimuli)
                 
         # initialize trials list
         self.Set = [] # compact set
@@ -207,14 +207,17 @@ class Trialset(object):
                 for num, row in rows.groupby('trial'):
                     categories = row.categories.item()[:] #gotta copy the list with [:]                    
                     gen_exemplars = np.array(rows.loc[rows.trial<num, 'stimulus'].values)
-                    gen_cats = rows.loc[rows.trial<num, 'category'].values
-                    for gen_cat in np.unique(gen_cats):
-                        gen_idc = np.array([gi for gi, gc in enumerate(gen_cats) if gc == gen_cat])
-                        categories += [gen_exemplars[gen_idc]]
-                    
-                    #categories = row.categories.item() + [gen_exemplars]
+                    if 'category' in rows.columns:                    
+                        gen_cats = rows.loc[rows.trial<num, 'category'].values
+                        for gen_cat in np.unique(gen_cats):
+                            gen_idc = np.array([gi for gi, gc in enumerate(gen_cats) if gc == gen_cat])
+                            categories += [gen_exemplars[gen_idc]]
+                        add2cat = self.cat2ind(row.category.item())
+                    else:
+                        categories = row.categories.item() + [gen_exemplars]
+                        add2cat = 1
+
                     genstim = row.stimulus.item()
-                    add2cat = self.cat2ind(row.category.item())
                     stimulus = [genstim,add2cat]
                     if not 'wrap_ax' in row.columns:
                         wrap_ax = None
@@ -267,7 +270,16 @@ class Trialset(object):
             params = model.params2dict(params)
             for parmname in fixedparams.keys():
                 params[parmname] = fixedparams[parmname]
-        
+        #Get stimsteps, if it doesn't exist, compute and add as property
+        if hasattr(self,'stimstep'):
+            stimstep = self.stimstep
+        else:
+            ndim = self.stimuli.shape[1]
+            stimstep = []
+            for di in range(ndim):
+                uniques = np.unique(self.stimuli[:,di])
+                st = uniques[1]-uniques[0]
+                stimstep += [st]
         # iterate over trials
         loglike = 0
         ps_list = np.array([])
[…]

### 2019-08-22 Modules/Funcs/Misc.py

@@ -692,14 +692,24 @@ def printProg(i,print_ct = 0, steps = 1, breakline = 40, breakby = 'char'):
 def getrange(stimuli):
     '''
     Get the min and max for each stimulus feature. To be used when initialising models.
+    Also return the steps of each dimension
     '''
     stimrange = []
     for i in range(len(stimuli[0])):
         stimrange += [{'min': stimuli[:,i].min(),
                        'max': stimuli[:,i].max()}]
-    return stimrange
+        
+    ndim = stimuli.shape[1]
+    stimstep = []
+    for di in range(ndim):
+        uniques = np.unique(stimuli[:,di])
+        st = uniques[1]-uniques[0]
+        stimstep += [st]
+
+    return stimrange,stimstep
         
 
+    
 def getModelName(modelname,fetch='short'):
     '''
     Fetches the specified model name. FETCH can be 'short',long','class'.

### 2019-08-23 Manuscripts/post-cogpsych/post-cogpsych.tex
Updated section 7.2 of cog sci manuscript

@@
 
-Another factor we examine with this category structure is the effect of boundaries. The represenativeness model prefers to generate categories located
+Another factor we examine with this category structure is the effect of
+boundaries. The representativeness model prefers to generate categories located
 at the boundaries of the stimulus features. If at least one of these boundaries
-is removed by utilizing a toroidal feature (one with no boundaries), such as the rotation of a line.
-The represenativeness model will strongly prefer to generate categories on the non-toroidal or bounded feature as it still has a boundary.
-Experiment 3 manipulates the stimulus space.
-One condition uses the same stimuli as before, squares varying in lightness and size, both of which are bounded features.
-The other condition use a new set of stimuli:
-circles that vary in size (a bounded feature) and orientation (an
-unbounded feature; Figure \ref{fig:e3-circles}a). Before describing the
-details of the experiment, we examine each model's
-predictions with stimuli that comprise one bounded and one unbounded feature.
+is removed by utilizing a toroidal feature (one with no boundaries), such as the
+rotation of a line.[Note to Xian: Something weird going on here, remember to edit.] The representativeness model will strongly prefer to
+generate categories on the non-toroidal or bounded feature as it still has a
+boundary. Experiment 3 manipulates the stimulus space. One condition uses the
+same stimuli as before, squares varying in lightness and size, both of which are
+bounded features. The other condition use a new set of stimuli: circles that
+vary in size (a bounded feature) and orientation (an unbounded feature; Figure
+\ref{fig:e3-circles}a). Before describing the details of the experiment, we
+examine each model's predictions with stimuli that comprise one bounded and one
+unbounded feature.
 
@@
     \begin{center} 
-      \includegraphics[width=.8\textwidth]{figs/e3-circles.pdf}
+      \includegraphics[width=.8\textwidth]{figs/e3-circles.png}
       \caption{Circle stimulus domain (a) and predictions from contrast models
@@
 
-Compared to stimuli with two bounded features, PACKER makes a similar prediction in generating the first Beta exemplar when one feature is unbounded, with the curious exception that the highest density regions appear compressed along the y-axis (for simplicity, we always assign the y-axis as the toroidal feature in the text and figures).
-Due to the unbounded nature of orientation, the maximum distance from
-the Alpha exemplars is half of the full range of the feature -- for an Alpha
-exemplar oriented at 0 radians, the furthest another exemplar can be situated is $\pi$ radians within the full range of orientation is $2\pi$ radians.
+Compared to stimuli with two bounded features, PACKER makes a similar prediction
+in generating the first Beta exemplar when one feature is unbounded, with the
+curious exception that the highest density regions appear compressed along the
+y-axis (for simplicity, we always assign the y-axis as the toroidal feature in
+the text and figures). Due to the unbounded nature of orientation, the maximum
+distance from the Alpha exemplars is half of the full range of the feature --
+for an Alpha exemplar oriented at 0 radians, the furthest another exemplar can
+be situated is $\pi$ radians within the full range of orientation is $2\pi$
+radians.
 
@@
 
-The single unbounded feature prevents ... While the representativeness model
-more closely resembles PACKER under this condition, note that the former still
-more strongly predicts the generation of the first Betas exemplars at the
-bounded edges compared to PACKER (Figure \ref{fig:e3-circles}c.) 
+The consequence of employing a wrapped Gaussian is presented in Figure
+\ref{fig:e3-circles}c. Edge effects along the unbounded feature (displayed as
+the vertical axis) are no longer present since the edges themselves no longer
+exist. Instead, the probability of generating the first Beta exemplar is
+bimodally distributed at the edges of the bounded feature (displayed as the
+horizontal axis.) Essentially, as with stimuli possessing only bounded features,
+the representativeness model still struggles to generate the first Beta
+exemplars in the center of the feature space.
+
+It may be argued that our expectation that the representativeness model should
+\emph{always} tend to produce first-Betas at the edges may be too strong.
+Specifically, this prediction is based on the model that is optimized to data
+from Experiments 1 and 2, and it is conceivable that the model may make
+drastically different predictions (such as some favoring the initial generation
+of centrally-located Betas more than Betas at the edges) in the current
+experiment when it is actually fit to empirical data from this experiment.
+However, we contend that this prediction of the representativeness model is a
+fundamental property of the model due to its assumption of unimodal category
+distributions. At the very least, the maximum probability of generating a
+centrally located first Beta should almost never be higher than the maximum
+probability of generating a first Beta at an edge.
+
+It is also worth noting that since the representativeness model does not explicitly
+recognize similarities as distances within the feature space (unlike PACKER),
+there is no compression in density regions along the unbounded feature axis.
+This provides us with an additional qualitative marker distinguishing the
+representativeness model from PACKER -- if the range of the entire Beta category
+along the unbounded feature is observed to be shorter than that of the bounded
+feature, this would be stronger evidence for PACKER rather than the
+representativeness model.
+
+% In the following
+% section, we explore -- using both qualitative comparisons and quantitative fits
+% -- whether PACKER or the representativeness model is a better account of
+% behavioral data in this Corner experiment.
+
 % For the representativeness model, the single unbounded feature essentially
@@
 where four Alpha exemplars were presented across three blocks followed by the
-generation phase. The training condition was the Corner category for all
+generation phase, and the features values are divided into 9 discrete steps. The training condition was the Corner category for all
 participants.
@@
 %Repeat hypotheses
+Unlike the previous experiments, Experiment 3 allows us to distinguish between
+PACKER and the representativeness model beyond their quantitative fits by
+identifying specific qualitatively diagnostic markers. Specifically, if PACKER
+is a better account of the data, we would expect a relative decrease in the
+range of the entire Beta category along the unbounded feature compared to the
+bounded feature, as well as a preference for first-generated Beta exemplars away
+from the edges of the feature space. These predictions are correspondingly
+reversed if the representativeness model offers a better account of the data.
+For our analyses, we define the edge to be the 1-tile wide border of each 9-by-9
+large feature space.
+
+Figure \ref{fig:e3-statsboxes} presents the distributional statistics for both
+square and circle stimuli. Participants produced circle stimuli with
+significantly lower Y-range (lower for the unbounded feature) as compared to 
+the square stimuli ($t(87) = 4.94$, $p = .001$), providing early support for PACKER.
+There are no significant differences between the distributional statistics of
+each generated category between square and circle stimuli (Figure
+\ref{fig:e3-statsboxes})
 
@@
 
-Participants produced circle stimuli with significantly lower Y-range (lower for the unbounded feature) as compared with the square stimuli ($t(87) = 4.94$, $p = .001$).  This
-provides preliminary support for PACKER, which predicts generating exemplars
-within a shorter range along the unbounded feature compared to a bounded
-feature. Otherwise, there were no significant differences between the distributional statistics of each
-generated category between square and circle stimuli (Figure
-\ref{fig:e3-statsboxes})
-
-The contrast models were fit by optimizing their fit to all experiments in this
-article (Experiments 1 through 3). Fit statistics and parameter values are
-presented in Table \ref{table:corner-model-fits}. Although not our main focus in
-this section, we also provide the statistics for the classical models --
-unsurprisingly, they do not perform as well as the contrast models. Overall,
-the representativeness model emerges as the best-fitting model in aggregate, although by a much smaller amount than Experiments 1 and 2. However, analyzing data in aggregate can be misleading.
-
-%by a much smaller lead compared to fits to only Experiments 1 and 2. At an individual level, 67\% of participants in Experiment 3 were better
-%fit by PACKER than the representativeness model.
-
-Table \ref{table:corner_counts} presents frequencies of the best-fitting model for participants in each condition.
-Overall the representativeness model was significantly worse at capturing the results of Experiment 3 (it fit 29 out of the 89 participants better than PACKER; two-tailed Binomial test $p < 0.005$).
-Further, this is mostly due to representativeness model capturing participants in the unbounded circle condition worse than  PACKER (11 of 43 better fit by representativeness; two-tailed Binomial test, $p<0.005$).
-The representativeness model only trended towards fitting participants in the bounded, Square condition better (18 out of 46; two-tailed Binomial test $p = 0.18$).
-Further, 35\% of participants in the bounded condition generated their first Beta exemplars along the
-edges of the feature space, but only 12\% of participants in the toroidal, circle condition generated their first Beta exemplars along the edges of
-the feature space. These proportions more closely match the predictions of
-PACKER (33\% probability of generating first Beta exemplars at the edges for
-squares, and 24\% probability at edges for circles) compared to the
-representativeness model (53\% of the former, 49\% probability of the latter).
-Thus, the representativeness model's success at capturing Experiments 1 and 2 is in part due to the nature of the stimulus features, as well as the structure of contrasting category.
+An inspection of the distributions of generated first-Betas (Figure
+\ref{fig:e3-behave}) does not immediately indicate support for either model. The
+square condition reveals first Betas generated at the corners as well as the
+center of the feature space, while the circle condition shows greater
+frequencies at the center. However, we also found that 65\% and 87\% of
+participants generated their first squares and first circles, respectively, away
+from the edge of the feature space.\footnote{Strictly speaking, the circle
+  stimuli do not have edges along its orientation feature, However, for ease of
+  comparison we will continue to refer to these comparable regions between
+  stimuli as edges.} Further, the maximum proportion of first-Betas generated
+off the edges (.13) is higher than that of the maximum proportion in an edge
+(.04) for the circle stimuli, although the maximum proportions are the same for
+squares (.06 for both edge and off-edge.) This preference for generating
+first-Betas away from the edge is in line with our hypothesis
+supporting PACKER.
+
+\begin{figure}
+    \begin{center}
+      \includegraphics[width=.6\textwidth]{figs/e3-behave.pdf}
+    \caption{Heatmaps indicating the relative frequencies of generating the
+      first Beta exemplar in the Corner condition.}
+    \label{fig:e3-behave}
+    \end{center}
+\end{figure}
+
+After fitting the models to data from Experiment 3, there appears to be no major
+qualitative deviations in the distribution of first-Betas for the square stimuli
+(Figures \ref{fig:e3-firstbetas}a and b) from previous expectations (i.e., as
+seen in Figure \ref{fig:e3-squares}), however this is not the case for the
+circle stimuli (Figures \ref{fig:e3-firstbetas}c and d compared to Figure
+\ref{fig:e3-circles})). Specifically, we see greater variance along the
+unbounded feature for PACKER, and most strikingly, that the representativeness
+model is able to expand the distribution of generating first-Betas into more
+centrally located regions \ref{fig:e3-firstbetas} for circle stimuli. However as
+predicted, even with this distribution, the maximum probability of the
+representativeness model generating at an edge (.0438) is higher than generating
+off-edge (.0394), indicating a preference for producing first-Betas at the edge.
+This is in contrast to both PACKER and the empirical data, which both show a
+general preference for centrally-located first-Betas.
+
+% These proportions more closely match the predictions of PACKER
+% (34\% probability of generating first Beta exemplars at the edges for squares,
+% and 24\% probability at edges for circles) compared to the representativeness
+% model (53\% of the former, 49\% probability of the latter). Thus, the
+% representativeness model's success at capturing Experiments 1 and 2 is in part
+% due to the nature of the stimulus features, as well as the structure of
+% contrasting category.
+
+\begin{figure}
+    \begin{center}
+      \includegraphics[width=.6\textwidth]{figs/e3-firstbetas.pdf}
+      \caption{Heatmaps indicating the probabilities of each contrast model
+        generating the first Beta exemplar in the Corner condition for both
+        square and circle stimuli. Models have been fit to data from Experiment 3.}
+    \label{fig:e3-firstbetas}
+    \end{center}
+\end{figure}
+
+The contrast models were fit by optimizing their fit to only Experiment 3. Fit
+statistics and parameter values are presented in Table
+\ref{table:corner-model-fits}. Although not our main focus in this section, we
+also provide the statistics for the classical models -- unsurprisingly, they do
+not perform as well as the contrast models. In line with our qualitative
+observations, PACKER emerges as the best-fitting model in aggregate.
+
+\begin{table} \centering
+  \caption{Results of model-fitting to the combined datasets from Experiments 1,
+    2 and 3. Smaller AIC values correspond to better model fits (adjusted
+    for number of parameters.)}
+\label{table:corner-model-fits}
+
+\begin{tabular}{ l l l l} \\ \\ \textbf{PACKER} & \textbf{Copy \& Tweak} &
+\textbf{Representativeness} & \textbf{Hierarchical Bayesian} \\ 
+  \hline $AIC = 2954$ &$AIC = 2963$ &$AIC = 2971$  &$AIC = 3025$\\ 
+  $L = -1474$          &$L = -1479$   &$L = -1481$    &$L = -1508$\\ 
+  $c = 2.08$           &$c = 3.73$    &$\kappa = 0.00$&$\kappa = 0.29$  \\ 
+  $\theta_c = 0.87$    &$\theta_c = 0$ (fixed)  &$\nu = 4.07 \times 10^20$ &$\nu
+                                                                             = 1.00$\\
+  $\theta_t = 2.18$    &$\theta_t = 2.57$       &$\lambda = 0.14$ &$\lambda = 0.00$\\
+                       &              &$\theta = 0.73$ &$\theta = 7.50$\\
+\end{tabular}
+\end{table}
+
+Table \ref{table:corner_counts} presents frequencies of the best-fitting model
+for participants in each condition. Overall the representativeness model was
+significantly worse at capturing the results of Experiment 3 (it fit 26 out of
+the 89 participants better than PACKER; two-tailed Binomial test $p < 0.001$).
+Further, this is mostly due to the representativeness model capturing participants
+in the unbounded circle condition worse than PACKER (8 of 43 better fit by
+representativeness; two-tailed Binomial test, $p<0.001$). The representativeness
+model only trended towards fitting participants in the bounded, Square condition
+better (18 out of 46; two-tailed Binomial test $p = 0.18$). 
 %In addition, these results agree with our prediction that the representativeness model would be poorer in predicting data from circle stimuli (26\% of participants observing circle stimuli) compared to square stimuli (39\% participants observing square stimuli).
@@
   \hline 
-  Squares & 28 & 18 & 46 & $p < 0.005$\\
-  Circles & 32 & 11 & 43 & $p \approx 0.18$\\
-  Total   & 60 & 29 & 89 & $p < 0.005$\\
+  Squares & 28 & 18 & 46 & $p \approx 0.18$\\
+  Circles & 35 &  8 & 43 & $p < 0.001$\\
+  Total   & 60 & 26 & 89 & $p < 0.001$\\
 \end{tabular}
 \end{table}
-\begin{table} \centering
-  \caption{Results of model-fitting to the combined datasets from Experiments 1,
-    2 and 3. Smaller AIC values correspond to better model fits (adjusted
-    for number of parameters.)}
-\label{table:corner-model-fits}
-
-\begin{tabular}{ l l l l} \\ \\ \textbf{PACKER} & \textbf{Copy \& Tweak} &
-\textbf{Representativeness} & \textbf{Hierarchical Bayesian} \\ 
-  \hline $AIC = 12023$ &$AIC = 12779$ &$AIC = 11982$  &$AIC = 12976$\\ 
-  $L = -6008$          &$L = -6388$   &$L = -5987$    &$L = -6484$\\ 
-  $c = 0.51$           &$c = 3.24$    &$\kappa = 6.73$&$\kappa = 0.00$  \\ 
-  $\theta_c = 3.03$    &$\theta_c = 0$ (fixed)  &$\nu = 1.00$ &$\nu = 3.27$\\
-  $\theta_t = 3.46$    &$\theta_t = 2.90$       &$\lambda = 1.90$ &$\lambda = 0.07$\\
-                       &              &$\theta = 2.45$ &$\theta = 3.15$\\
-\end{tabular}
-\end{table}
-
-%Ultimately, these results suggest that PACKER's exemplar dissimilarity account
-%of contrast, as opposed to the representativeness account, is more appropriate
-%for the Corner category type. On almost every measure of analysis in this
-%section, PACKER outperformed the representativeness model. Although the
-%representativeness model best fit the data from Experiments 1 and 2, it
-%struggled when the learned category occupied the extreme corners of the stimulus
-%space -- both in cases where the stimuli possess bounded and unbounded features.
-%
-%[Work on transition time permitting]
+
+Ultimately, these results suggest that PACKER's exemplar dissimilarity account
+of contrast, as opposed to the representativeness account, offers a better
+explanation of category generation for the Corner category type. On almost every
+measure of analysis in this section, PACKER outperformed the representativeness
+model, indicating that the better fit of the latter model demonstrated in
+Experiments 1 and 2 is not general to all conditions.
 
@@
 generation is a fundamentally different computational-level problem
-\citep[different from those posed by][]{jern2013probabilistic,kemp2014taxonomy}, which is supported by the represenativeness form of their model performing much better.
+\citep[different from those posed by][]{jern2013probabilistic,kemp2014taxonomy}, which is supported by the representativeness form of their model performing much better.
 

### 2019-08-23 Experiments/cogpsych-code/print_gs_results_csv.py

@@ -12,7 +12,7 @@ import numpy as np
 pickledir = 'pickles/'
 writefile = 'globalfits'
 prefix = 'chtc_gs_best_params'
-printdata = ['5con','corner','xcrABC']#'all'
+printdata = ['5con','corner','all']#'all'
 if not printdata is 'all':
     gsfiles = []
     data_names = []

### 2019-08-23 Modules/Classes/ConjugateJK13.py

@@ -203,7 +203,6 @@ class RepresentJK13(HierSamp):
 
         target_is_populated = any(self.assignments == category)                
         mu,Sigma = self.get_musig(stimuli, category)
-                
         # get relative densities
         if np.isnan(Sigma).any() or np.isinf(Sigma).any():
             #target_dist = np.ones(mu.shape) * np.nan
@@ -211,8 +210,8 @@ class RepresentJK13(HierSamp):
         else:
             # #270418 Implementing representational draws
             target_dist_target = multivariate_normal(mean = mu, cov = Sigma)
-            if not self.wrap_ax is None:
-                likelihood_target = self._wrapped_density(target_dist_target,stimuli)
+            if not self.wrap_ax is None:                
+                likelihood_target = self._wrapped_density(target_dist_target,stimuli)                
             else:
                 likelihood_target = target_dist_target.pdf(stimuli)
 

### 2019-09-02 Manuscripts/post-cogpsych/post-cogpsych.tex
Updated draft Exp 3 with correlations.

@@
 items within the same category and those in different categories
-\citep{roschmervis1975,regier2007,shepard1961learning,smith1995creative}.
+\citep{roschmervis1975,regier2007,shepard1961learning,smith1995creative}.\footnote{It
+  is worth noting that sensitivity to feature correlations is not universal and
+  can be dependent on factors such as the salience of correlations and task
+  demands \citep[e.g., see][]{chin2002effect,malt1984correlated,murphy2004big}.}
 Inspired by this work, previous research on the topic of category generation has
@@
 similar {\em statistical regularities} as previously learned categories
-\citep{jern2013probabilistic,ward1994structured}. Although this is an important
-characteristic of generating new categories, it cannot be the only one. Taken to
-the extreme, the best ``new'' category in terms of having the same statistical
-regularities to other categories would be identical to a known category that is
-representative of the domain (and thus, not new at all). Contrast from other
-known categories should play a role. Indeed, scientists in other fields, such
-as marketing \citep{berger16} and sociology \citep{rogers03}, highlight the
-critical importance of contrast in the creation of new ideas.
+\citep{jern2013probabilistic,ward1994structured}. These statistical regularities
+reflect general structural patterns across old and new categories and are not
+restricted to strictly quantitatively similar properties.
+
+Although statistical regularity is an important characteristic of generating new
+categories, it cannot be the only one. Taken to the extreme, the best ``new''
+category in terms of having the same statistical regularities to other
+categories would be identical to a known category that is representative of the
+domain (and thus, not new at all). Contrast from other known categories should
+play a role. Indeed, scientists in other fields, such as marketing
+\citep{berger16} and sociology \citep{rogers03}, highlight the critical
+importance of contrast in the creation of new ideas.
 
@@
 category. To formalize the first technique, we build on the largely successful
-exemplar modeling framework that
+exemplar modeling framework
 \citep{medin1978context,nosofsky1984choice,nosofsky1986attention}. In doing so,
@@
 model can do at capturing contrast is to assume that the new category is placed
-uniformly at random over stimulus space. But, this defeats the purpose of a
+uniformly at random over feature space. But, this defeats the purpose of a
 hierarchy as it is ignored when determining a new category location!\footnote{It
@@
   Representations}). PACKER explains category generation as a balance between
-two fundamental constraints: The category to be generated should not be similar
-to known categories, and exemplars within each category should be similar to one
-another. These ideas are implemented within the well-studied exemplar framework
--- the PACKER model is an extension of the influential Generalized Context Model
-of categorization \citep[GCM;][]{nosofsky1984choice,nosofsky1986attention}.
+two fundamental constraints: The exemplars of the category to be generated
+should not be similar to known categories, and exemplars within each category
+should be similar to one another. These ideas are implemented within the
+well-studied exemplar framework -- the PACKER model is an extension of the
+influential Generalized Context Model of categorization
+\citep[GCM;][]{nosofsky1984choice,nosofsky1986attention}.
 
@@
 similarity calculations, and $c$ ($c>0$) is a specificity parameter controlling
-the spread of exemplar generalization. For simplicity, attention will be
-distributed uniformly in our simulations (unless otherwise noted). The value of
-$r$ depends on the nature of the experimental conditions being simulated: $r=1$
-is appropriate for separable dimensions, whereas $r=2$ is appropriate for
-integral dimensions \citep[e.g.,][]{shepard1964attention,garner1974processing}.
-In our simulations, we set $r=1$ due to the separable nature of the stimulus
-dimensions used in our experiments (see Figure \ref{fig:e1-conditions}).
+the spread of exemplar generalization. For simplicity, in our simulations
+attention will be distributed across each dimension uniformly (unless otherwise
+noted). The value of $r$ depends on the nature of the experimental conditions
+being simulated: $r=1$ is appropriate for separable dimensions, whereas $r=2$ is
+appropriate for integral dimensions
+\citep[e.g.,][]{shepard1964attention,garner1974processing}. In our simulations,
+we set $r=1$ due to the separable nature of the stimulus dimensions used in our
+experiments (see Figure \ref{fig:e1-conditions}).
 
@@
 %determinism.
-It is worth noting that PACKER is only one possible exemplar-based account of
-category generation within our proposed framework. That is, PACKER places
-specific constraints on the possible values of $f(x_j)$, but other
-exemplar-based category generation models with drastically different behavior
-can be formalized in this framework by imposing alternative constraints. For
-example, as will be discussed in more detail below, PACKER is formally
-equivalent to the copy-and-tweak model proposed by \cite{jern2013probabilistic}
-when $\theta_c = 0$ and $\theta_t > 0$. Likewise, when
-$\theta_t = 0$ and $\theta_c > 0$, PACKER can represent a
-contrast-only generation mode, relying exclusively on contrast when generating
-new categories. When $f(x_j) < 0$ for all $x_j$ (regardless of class
-membership), a ``pure-packing'' approach is yielded, generating items in
-unoccupied areas of the domain. Thus, the proposed framework may be used to
-describe a wide variety of qualitatively distinct generation strategies.
 
 \subsubsection{Relation Between PACKER and Copy-And-Tweak}
+\label{section:copytweak-packer}
 %: Both models are
@@
 
-\label{section:copytweak-packer} The PACKER model is similar to the copy-and-tweak model reported by \cite{jern2013probabilistic}.
-An advantage of formulating copy-and-tweak as a special case parameterization of PACKER. In this paper, we report simulations using this copy-and-tweak model. This model
-fits within the exemplar-based category generation framework defined above, and
-is a continuous-dimension adaptation of the model tested by
-\cite{jern2013probabilistic}. By formalizing a model family where PACKER and
-copy-and-tweak are different parameterizations of models within the same
-framework, the comparison between PACKER and copy-and-tweak provides a test of
-the explanatory value of the contrast mechanism based on exemplar dissimilarity:
-The account provided by copy-and-tweak will only equal that of PACKER if the
-contrast mechanism does not offer an advantage (i.e., if $\theta_c > 0$
-significantly improves model fits). Note that the purpose of the article is to
-explore and formally analyze the role of contrast in categorization and thus, we
-leave extending PACKER to incorporate distributional factors for future work.
+It is worth noting that PACKER is only one possible exemplar-based account of
+category generation within our proposed framework. That is, PACKER places
+specific constraints on the possible values of $f(x_j)$, but other
+exemplar-based category generation models with drastically different behavior
+can be formalized in this framework by imposing alternative constraints. For
+example, PACKER is formally equivalent to the copy-and-tweak model proposed by
+\cite{jern2013probabilistic} when $\theta_c = 0$ and $\theta_t > 0$ (in fact, it
+becomes a continuous-dimension adaptation of the original model). Likewise,
+when $\theta_t = 0$ and $\theta_c > 0$, PACKER can represent a contrast-only
+generation mode, relying exclusively on contrast when generating new categories.
+When $f(x_j) < 0$ for all $x_j$ (regardless of class membership), a
+``pure-packing'' approach is yielded, generating items in unoccupied areas of
+the domain. Thus, the proposed framework may be used to describe a wide variety
+of qualitatively distinct generation strategies.
+
+By formalizing a model family where PACKER and copy-and-tweak are different
+parameterizations of models within the same framework, the comparison between
+PACKER and copy-and-tweak provides a test of the explanatory value of the
+contrast mechanism based on exemplar dissimilarity: The account provided by
+copy-and-tweak will only equal that of PACKER if the contrast mechanism does not
+offer an advantage (i.e., if $\theta_c > 0$ significantly improves model fits).
+Note that the purpose of the article is to explore and formally analyze the role
+of contrast in categorization and thus, we leave extending PACKER to incorporate
+distributional factors for future work.
 
@@
 representativeness -- exemplars are generated such that they resemble the target
-category and are thus more distinct from and less similar to other categories. We
-adopt the formalization given by \cite{tenenbaum01}, where the
-representativeness $R(y,h)$ of an item $y$ (or in our case, an exemplar) is the
-relative amount of evidence that is provided by $y$ for a given hypothesis $h$
-in a space of hypotheses $\mathcal{H}$ in contrast to all other hypotheses
+category and are thus more distinct from and less similar to other categories.
+The general idea of representativeness is not new in categorization. For
+instance, the literature on the \emph{graded structure} of categories explores
+how some exemplars are better representations of their categories
+\citep{palmeri2001central,barsalou1985ideals}. In this article we adopt the
+specific formalization of representativeness given by \cite{tenenbaum01}, where
+the representativeness $R(y,h)$ of an item $y$ (or in our case, an exemplar) is
+the relative amount of evidence that is provided by $y$ for a given hypothesis
+$h$ in a space of hypotheses $\mathcal{H}$ in contrast to all other hypotheses
 $\mathcal{H}^c=\mathcal{H} \setminus \{h\}$:
@@
 
-In our simulations, we set the prior mean $\mu_0$ to the center of the stimulus
+In our simulations, we set the prior mean $\mu_0$ to the center of the feature
 space and the prior variance to be isotropic ($\Sigma_0 = \lambda \mathbf{I}$ where
@@
 role, exemplars in the generated categories of participants in the XOR condition
-may not be positively correlated -- they may not be correlated, or perhaps even be negatively correlated.
-In this case, contrast and statistical regularities would interact, which would
-be inconsistent with prevailing theories of category generation
-\citep{jern2013probabilistic}.
+may not be positively correlated -- they may not be correlated, or perhaps even
+be negatively correlated. In this case, contrast and statistical regularities
+would interact, which would be inconsistent with prevailing theories of category
+generation \citep{jern2013probabilistic}.
 
@@
 be generated into a new category. This supports the notion that contrast is a
-fundamental constraint on how categories are related to one another and that statistical
-regularity alone is insufficient.
+fundamental constraint on how categories are related to one another and that
+statistical regularity alone is insufficient.
 
@@
 
-We believe that the failure to replicate dimension correlation emulation is
+We believe that the failure to replicate the emulation of dimension correlation is
 because participants in \cite{jern2013probabilistic} could differentiate the
@@
 differently between conditions due to the differences in the available empty
-stimulus space for generating a new category?
+feature space for generating a new category?
 
@@
 space. Participants assigned to learn the Bottom category should be less likely
-to generate exemplars into the lower regions of the stimulus space (as these
+to generate exemplars into the lower regions of the feature space (as these
 areas possess greater similarity to members of the Bottom category), preferring
@@
 %Contrast model distinction
-\section{Experiment 3: Differentiating Contrast}
+\section{Experiment 3: Contrasting Contrast}
 \label{section:diffcon}
@@
 
-While the representativeness model fits better than PACKER, and in spite of
-their fundamentally different structures, both contrast models make qualitatively similar predictions for Experiments 1 and 2.
-To illustrate this, Figure \ref{fig:firstbetas} shows how the modal generating probabilities for the first Beta exemplar are similar across both contrast models.\footnote{Model parameters are set to the optimized values from Table \ref{table:global-model-fits}.} The highest density regions for both models appear to be near the corners and edges.
-This is particularly the case for the representativeness model, which almost exclusively predicts generation at the corners, while PACKER can generate the first novel exemplars toward the inner regions of the stimulus space in the XOR condition.
+While the representativeness model fits better than PACKER, and in spite of their fundamentally different structures,
+both contrast models make qualitatively similar predictions for Experiments 1 and 2. To illustrate this, Figure
+\ref{fig:firstbetas} shows how the models' generating probabilities for the first Beta exemplar are similar across both
+contrast models.\footnote{Model parameters are set to the optimized values from Table \ref{table:global-model-fits}.}
+Across the five Alpha conditions of Experiments 1 and 2, the first Beta exemplars according to the contrast models were
+generally more probable near the corners and edges. This is particularly the case for the representativeness model,
+which almost exclusively predicts generation at the corners, while PACKER can predict generation in the center of the
+feature space in the XOR condition.
 
@@
         the first Beta exemplar for each Alpha condition. Darker shades of
-        blue indicate higher probabilities (normalized within each plot.)}
+        blue indicate higher probabilities (normalized within each plot).}
       \label{fig:firstbetas}
@@
 
-These similarities are perhaps unsurprising -- each of the modal generation
-regions are maximally distant from the Alpha exemplars. However, recall that
-unlike PACKER, the representativeness model does not explicitly use a measure sensitive to distance from the location of exemplars in the contrasting category.
-The general pattern of producing exemplars dissimilar from the contrast category is because a unimodal category structure is assumed by the model and it is not strictly constrained by the distances to exemplars of the contrasting categories.
-Consequently, it is possible to design a configuration of Alpha exemplars where the representativeness model would predict the generation of novel exemplars
-that are \emph{more similar} to the learned exemplars.
-
-In this section, we focus on one category type -- the Corner
-category -- to distinguish between the two competing accounts of contrast in
-category generation.
-Here, each Alpha exemplar is located at the extreme corners of the feature space (see Figure \ref{fig:e3-squares}a).
-We first seek to identify what predictions each contrast model generates under this condition,
-followed by an category generation task, and a model-based analysis of the two contrast models given the behavioral data.
-
-In the Corner condition, the representativeness model assumes an underlying
-Alpha category distribution that is centered on the feature space, with
-probability densities highest in the center and lowest in the corners. Since the representativeness of novel Beta exemplars generally increase when the
-likelihood of the underlying Alpha category distribution decreases, the regions
-of highest Beta representativeness are consequently located at the corners where
-these Alpha likelihoods are lowest (Figure \ref{fig:e3-squares}b.)
-
-Conversely, PACKER predicts a completely different result. With $\theta_c > 0$,
-the probability of generating a Beta exemplar increases with the dissimilarity
-from all Alpha exemplars. This results in highest Beta generation probabilities
-in the inner regions of the feature space, where the candidate exemplars are
-maximally different from the Alpha exemplars (Figure \ref{fig:e3-squares}c.)
+The effect of contrast for both models is similar in that they both tend to produce exemplars that are distant from the
+Alpha category. However, this is realized in different ways for the two models. The representativeness model does not
+explicitly use a measure sensitive to distance from the location of exemplars in the contrasting category. Instead, it
+is the assumption of unimodal distributions that results in the model producing novel exemplars dissimilar to learned
+exemplars. Consequently, it is possible to design a configuration of Alpha exemplars where the representativeness model
+predicts novel exemplars should be \emph{similar} to the learned exemplars. However, PACKER cannot predict this because
+it assumes that the similarity weight for contrasting exemplars is negative.
+
+In this section, we focus on a different category type -- the Corner category -- that contrasts the two models of
+contrast. Here, each Alpha exemplar is located at the corners of the feature space (see Figure \ref{fig:e3-squares}a).
+
+We first identify the predictions of each model in this condition. Second, we test these contrasting predictions in a
+behavioral experiment. We then conduct a model-based analysis of the two contrast models.
+
+In the Corner condition, the representativeness model estimates the underlying Alpha category distribution to be
+centered on the feature space, with probability densities highest in the center and lowest in the corners. The
+representativeness of novel Beta exemplars generally decreases with the likelihood of the underlying Alpha category
+distribution. Despite the Alpha exemplars being in the corners, the predicted mean will be near the center
+of the Alpha category (the center of the space; See \ref{fig:e3-squares}b). As such, the Alpha likelihood is lowest in
+the corners, and thus, the representativeness of the Beta category will be largest in the corners.
+
+Conversely, PACKER predicts a completely different result. With PACKER, the probability of generating a Beta exemplar
+increases with the dissimilarity from all Alpha exemplars. This results in Betas generated in the center of feature
+space, where the candidate exemplars are as far as possible from the Alpha exemplars (Figure \ref{fig:e3-squares}c).
 
@@
 
-Another factor we examine with this category structure is the effect of
-boundaries. The representativeness model prefers to generate categories located
-at the boundaries of the stimulus features. If at least one of these boundaries
-is removed by utilizing a toroidal feature (one with no boundaries), such as the
-rotation of a line.[Note to Xian: Something weird going on here, remember to edit.] The representativeness model will strongly prefer to
-generate categories on the non-toroidal or bounded feature as it still has a
-boundary. Experiment 3 manipulates the stimulus space. One condition uses the
-same stimuli as before, squares varying in lightness and size, both of which are
-bounded features. The other condition use a new set of stimuli: circles that
-vary in size (a bounded feature) and orientation (an unbounded feature; Figure
-\ref{fig:e3-circles}a). Before describing the details of the experiment, we
-examine each model's predictions with stimuli that comprise one bounded and one
-unbounded feature.
+Another aspect of the contrast in the models' predictions is the role of the boundaries of feature space. The
+representativeness model prefers to generate categories at the boundaries of feature space. PACKER is less sensitive to
+boundaries. Boundaries tend to be where distance to the contrasting category exemplars is maximized. But, as in cases
+like the Corner condition (Figure \ref{fig:e3-squares}a), this is not always the case; Beta exemplars should be as far
+from the Alphas as possible, which is maximized in the center of feature space.
+
+Thus, in this experiment, we also examine how making a boundary toroidal (e.g., line rotation) affects the prediction of
+each model. The representativeness model prefers to generate categories on the feature that still has a boundary.
+PACKER, in contrast, should continue to generate categories that are most distant from the learn category regardless of
+the presence or absence of feature boundaries -- in this condition it would be the center of the feature space.
+Experiment 3 manipulates whether one feature is bounded or toroidal (always displayed as the Y-axis). One condition uses
+the same stimuli as before, squares varying in lightness and size, both of which are bounded features. The other
+condition use a new set of stimuli: circles that vary in size (a bounded feature) and orientation (an unbounded feature;
+Figure \ref{fig:e3-circles}a). Before describing the details of the experiment, we explain in detail each model's
+predictions given the Corner Alphas for a feature space where one feature is bounded and the other is toroidal.
 
@@
     \begin{center} 
-      \includegraphics[width=.8\textwidth]{figs/e3-circles.png}
+      \includegraphics[width=.8\textwidth]{figs/e3-circles.pdf}
       \caption{Circle stimulus domain (a) and predictions from contrast models
@@
         and direction assignment were counterbalanced across participants (in
-        the displayed example, circles are rotated clockwise up the y-axis.) }
+        the displayed example, circles are rotated clockwise up the y-axis). }
       \label{fig:e3-circles}
@@
 
-Compared to stimuli with two bounded features, PACKER makes a similar prediction
-in generating the first Beta exemplar when one feature is unbounded, with the
-curious exception that the highest density regions appear compressed along the
-y-axis (for simplicity, we always assign the y-axis as the toroidal feature in
-the text and figures). Due to the unbounded nature of orientation, the maximum
-distance from the Alpha exemplars is half of the full range of the feature --
-for an Alpha exemplar oriented at 0 radians, the furthest another exemplar can
-be situated is $\pi$ radians within the full range of orientation is $2\pi$
-radians.
-
-For the representativeness model, adding an unbounded feature changes the
-probability distribution for the underlying Bayesian model. Whereas it is still
-Gaussian in one feature, the other feature is toroidal, meaning that it wraps
-back to $0$ radians once it reaches $2\pi$ (we have conducted a change of
-variables where $y \rightarrow y \mod 2\pi $, where $\mod$ is the modulus
-operator). As we cross $2\pi$ and reach $0$ again, the remaining probability
-density also gets wrapped around, meaning that the probability density assigned
-by the Gaussian distribution on the toroidal dimension at $2\pi$ is actually
-assigned to $0$ (and $2\pi + \epsilon$ to $\epsilon$, and so on). This is called the
-{\em wrapped} Gaussian distribution \citep{mardia1972}. Note that the often used
-distribution for circular dimensions, the Von Mises distribution, is an
-approximation to the wrapped Gaussian distribution. See Appendix B for
-implementational details of the representativeness model in toroidal space.
-
-The consequence of employing a wrapped Gaussian is presented in Figure
-\ref{fig:e3-circles}c. Edge effects along the unbounded feature (displayed as
-the vertical axis) are no longer present since the edges themselves no longer
-exist. Instead, the probability of generating the first Beta exemplar is
-bimodally distributed at the edges of the bounded feature (displayed as the
-horizontal axis.) Essentially, as with stimuli possessing only bounded features,
-the representativeness model still struggles to generate the first Beta
-exemplars in the center of the feature space.
-
-It may be argued that our expectation that the representativeness model should
-\emph{always} tend to produce first-Betas at the edges may be too strong.
-Specifically, this prediction is based on the model that is optimized to data
-from Experiments 1 and 2, and it is conceivable that the model may make
-drastically different predictions (such as some favoring the initial generation
-of centrally-located Betas more than Betas at the edges) in the current
-experiment when it is actually fit to empirical data from this experiment.
-However, we contend that this prediction of the representativeness model is a
-fundamental property of the model due to its assumption of unimodal category
-distributions. At the very least, the maximum probability of generating a
-centrally located first Beta should almost never be higher than the maximum
-probability of generating a first Beta at an edge.
-
-It is also worth noting that since the representativeness model does not explicitly
-recognize similarities as distances within the feature space (unlike PACKER),
-there is no compression in density regions along the unbounded feature axis.
-This provides us with an additional qualitative marker distinguishing the
-representativeness model from PACKER -- if the range of the entire Beta category
-along the unbounded feature is observed to be shorter than that of the bounded
-feature, this would be stronger evidence for PACKER rather than the
-representativeness model.
+Compared to stimuli with two bounded features, PACKER makes a similar prediction in generating the first Beta exemplar
+when one feature is unbounded, with the curious exception that the highest density regions appear compressed along the
+toroidal feature. Due to the toroidal nature of orientation, the maximum distance from the Alpha exemplars is half of
+the full range of the feature -- for an Alpha exemplar oriented at 0 radians, the furthest distance to another exemplar
+is $\pi$ radians.
+
+For the representativeness model, adding a toroidal feature changes the probability distribution for the underlying
+Bayesian model. For the toroidal feature, space wraps back to $0$ radians once it reaches $2\pi$ (we have conducted a
+change of variables where $y \rightarrow y$ mod $2\pi$, where mod is the modulus operator). As we cross $2\pi$ and reach
+$0$ again, the remaining probability density also gets wrapped around, meaning that the density at $2\pi$ is added to
+$0$ (and $2\pi + \epsilon$ to $\epsilon$, and so on). This is called the {\em wrapped} Gaussian distribution
+\citep{mardia1972}. Note that the more familiar distribution for circular features, the Von Mises distribution, is an
+approximation to the wrapped Gaussian distribution. See Appendix B for implementational details of the
+representativeness model in a bounded and toroidal feature.
+
+The consequence of employing a wrapped Gaussian is presented in Figure \ref{fig:e3-circles}c. Edge effects along the
+unbounded feature (displayed as the vertical axis) are no longer present since the edges themselves no longer exist.
+Instead, the probability of generating the first Beta exemplar is bimodally distributed at the edges of the bounded
+feature (displayed as the horizontal axis). Thus, the representativeness model and PACKER make orthogonal predictions
+for how a toroidal dimension should affect category generation.
+
+In this experiment, we test human category generation in two cases where PACKER and the representativeness model
+contrast. The representativeness model predicts categories should be distributed in the boundaries of the bounded
+features. PACKER predicts category exemplars should be centered in space and decay exponentially over bounded
+dimension(s).
+
+% It may be argued that our expectation that the representativeness model should
+% \emph{always} tend to produce first-Betas at the edges may be too strong.
+% Specifically, this prediction is based on the model that is optimized to data
+% from Experiments 1 and 2, and it is conceivable that the model may make
+% drastically different predictions (such as some favoring the initial generation
+% of centrally-located Betas more than Betas at the edges) in the current
+% experiment when it is actually fit to empirical data from this experiment.
+% However, we contend that this prediction of the representativeness model is a
+% fundamental property of the model due to its assumption of unimodal category
+% distributions. At the very least, the maximum probability of generating a
+% centrally located first Beta should almost never be higher than the maximum
+% probability of generating a first Beta at an edge.
+
+% It is also worth noting that since the representativeness model does not explicitly
+% recognize similarities as distances within the feature space (unlike PACKER),
+% there is no compression in density regions along the unbounded feature axis.
+% This provides us with an additional qualitative marker distinguishing the
+% representativeness model from PACKER -- if the range of the entire Beta category
+% along the unbounded feature is observed to be shorter than that of the bounded
+% feature, this would be stronger evidence for PACKER rather than the
+% representativeness model.
 
@@
 
-We recruited 89 participants from Amazon Mechanical Turk who were randomly
-assigned to the bounded, square stimuli ($N=46$), or to the toroidal, circle
-stimuli ($N=43$). The square stimuli were exactly the same as in Experiments 1
-and 2. The circle stimuli were made up of unfilled circles with a radial line
-extending from its center to its edge. These stimuli varied along two features
--- their diameter ranging from 3.0--5.8 cm, and their orientation ranging from
-$2.587$--$2\pi + 2.587$ radians, with $0$ radians indicating a completely
-horizontal line. The line widths of the circles and radial line were set to 0.05
-cm. Similar to Experiments 1 and 2, participants first observed a training phase
-where four Alpha exemplars were presented across three blocks followed by the
-generation phase, and the features values are divided into 9 discrete steps. The training condition was the Corner category for all
-participants.
+We recruited 89 participants from Amazon Mechanical Turk who were randomly assigned to the square stimuli ($N=46$), or
+to the circle stimuli ($N=43$). The square stimuli were exactly the same as in Experiments 1 and 2, but with Alpha
+exemplars in the corners of the feature space. The circle stimuli were made up of unfilled circles with a radial line
+extending from its center to its edge. These stimuli varied along two features -- their diameter ranging from 3.0--5.8
+cm, and their orientation ranging from $2.587$--$2\pi + 2.587$ radians, with $0$ radians indicating a completely
+horizontal line. The line widths of the circles and radial line were set to 0.05 cm. Similar to Experiments 1 and 2,
+participants first observed a training phase where four Alpha exemplars were presented across three blocks followed by
+the generation phase, and the features values are divided into 9 discrete steps. The training condition was the Corner
+category for all participants.
 
@@
 %Repeat hypotheses
-Unlike the previous experiments, Experiment 3 allows us to distinguish between
-PACKER and the representativeness model beyond their quantitative fits by
-identifying specific qualitatively diagnostic markers. Specifically, if PACKER
-is a better account of the data, we would expect a relative decrease in the
-range of the entire Beta category along the unbounded feature compared to the
-bounded feature, as well as a preference for first-generated Beta exemplars away
-from the edges of the feature space. These predictions are correspondingly
+Unlike previous experiments, Experiment 3 allows us to distinguish between PACKER and the representativeness model
+beyond their quantitative fits by identifying specific qualitatively diagnostic markers. Specifically, if PACKER is a
+better account of the data, we would expect a relative decrease in the range of the entire Beta category along the
+unbounded feature compared to the bounded feature. In addition, we would expect to see a preference for first-generated
+Beta exemplars away from the boundaries of the feature space -- that is, the probability of generating an exemplar would
+be negatively correlated with its distance from the center for bounded features. These predictions are correspondingly
 reversed if the representativeness model offers a better account of the data.
-For our analyses, we define the edge to be the 1-tile wide border of each 9-by-9
-large feature space.
-
-Figure \ref{fig:e3-statsboxes} presents the distributional statistics for both
-square and circle stimuli. Participants produced circle stimuli with
-significantly lower Y-range (lower for the unbounded feature) as compared to 
-the square stimuli ($t(87) = 4.94$, $p = .001$), providing early support for PACKER.
-There are no significant differences between the distributional statistics of
-each generated category between square and circle stimuli (Figure
-\ref{fig:e3-statsboxes})
+%For our analyses, we define the edge to be the 1-tile wide border of each 9-by-9
+%large feature space.
+
+Figure \ref{fig:e3-statsboxes} presents the distributional statistics for both square and circle stimuli. Participants
+produced circle stimuli with significantly lower Y-range (lower for the unbounded feature) as compared to the square
+stimuli ($t(87) = 4.94$, $p = .001$), which supports PACKER. There are no significant differences between the
+distributional statistics of each generated category between square and circle stimuli (Figure \ref{fig:e3-statsboxes}).
 
@@
 
-An inspection of the distributions of generated first-Betas (Figure
-\ref{fig:e3-behave}) does not immediately indicate support for either model. The
-square condition reveals first Betas generated at the corners as well as the
-center of the feature space, while the circle condition shows greater
-frequencies at the center. However, we also found that 65\% and 87\% of
-participants generated their first squares and first circles, respectively, away
-from the edge of the feature space.\footnote{Strictly speaking, the circle
-  stimuli do not have edges along its orientation feature, However, for ease of
-  comparison we will continue to refer to these comparable regions between
-  stimuli as edges.} Further, the maximum proportion of first-Betas generated
-off the edges (.13) is higher than that of the maximum proportion in an edge
-(.04) for the circle stimuli, although the maximum proportions are the same for
-squares (.06 for both edge and off-edge.) This preference for generating
-first-Betas away from the edge is in line with our hypothesis
-supporting PACKER.
+The distributions of generated first-Betas (Figure \ref{fig:e3-behave}) indicate further qualitative support for PACKER.
+In particular, the probability of generating the first Beta exemplar for circles along the bounded dimension is
+significantly negatively correlated with distance from the center ($r(3) = -.95, p = .01$). Although the correlations
+for squares ($r(3)= -.50, p = .40$ and $r(3)= -.31 p = .61$ for x and y dimensions) were not significant, they were in
+the directions predicted by PACKER. Mean probabilities of of generating first-Betas are presented in Figure
+\ref{fig:e3-distcent}. This analysis was excluded for the toroidal dimension of the circle stimuli because the
+boundaries (and consequently a center) are not defined for this feature.
+
+% An inspection of the distributions of generated first-Betas (Figure
+% \ref{fig:e3-behave}) does not immediately indicate support for either model. The
+% square condition reveals first Betas generated at the corners as well as the
+% center of the feature space, while the circle condition shows greater
+% frequencies at the center. However, we also found that 65\% and 87\% of
+% participants generated their first squares and first circles, respectively, away
+% from the edge of the feature space.\footnote{Strictly speaking, the circle
+%   stimuli do not have edges along its orientation feature, However, for ease of
+%   comparison we will continue to refer to these comparable regions between
+%   stimuli as edges.} Further, the maximum proportion of first-Betas generated
+% off the edges (.13) is higher than that of the maximum proportion in an edge
+% (.04) for the circle stimuli, although the maximum proportions are the same for
+% squares (.06 for both edge and off-edge). This preference for generating
+% first-Betas away from the edge is in line with our hypothesis
+% supporting PACKER.
 
@@
 
-After fitting the models to data from Experiment 3, there appears to be no major
-qualitative deviations in the distribution of first-Betas for the square stimuli
-(Figures \ref{fig:e3-firstbetas}a and b) from previous expectations (i.e., as
-seen in Figure \ref{fig:e3-squares}), however this is not the case for the
-circle stimuli (Figures \ref{fig:e3-firstbetas}c and d compared to Figure
-\ref{fig:e3-circles})). Specifically, we see greater variance along the
-unbounded feature for PACKER, and most strikingly, that the representativeness
-model is able to expand the distribution of generating first-Betas into more
-centrally located regions \ref{fig:e3-firstbetas} for circle stimuli. However as
-predicted, even with this distribution, the maximum probability of the
-representativeness model generating at an edge (.0438) is higher than generating
-off-edge (.0394), indicating a preference for producing first-Betas at the edge.
-This is in contrast to both PACKER and the empirical data, which both show a
-general preference for centrally-located first-Betas.
-
-% These proportions more closely match the predictions of PACKER
-% (34\% probability of generating first Beta exemplars at the edges for squares,
-% and 24\% probability at edges for circles) compared to the representativeness
-% model (53\% of the former, 49\% probability of the latter). Thus, the
-% representativeness model's success at capturing Experiments 1 and 2 is in part
-% due to the nature of the stimulus features, as well as the structure of
-% contrasting category.
+\begin{figure}
+    \begin{center}
+      \includegraphics[width=.6\textwidth]{figs/e3-distcent.pdf}
+      \caption{Empirical mean probabilities of generating first Beta exemplars at each point of distance from the center
+        of the feature space. Plot for the toroidal feature of circle stimuli is not presented because a center is not
+        defined for this feature.}
+    \label{fig:e3-distcent}
+    \end{center}
+\end{figure}
+
+Model generation probabilities of the first Beta exemplars do not qualitatively differ from earlier predictions even
+after fitting to the current dataset. Figures \ref{fig:e3-firstbetas}a-b and \ref{fig:e3-firstbetas}c-d show the model
+predictions for square and circle stimuli respectively. While the predictions for the representativeness model for
+circles appears different from earlier predictions (i.e., in Figure \ref{fig:e3-circles}c), note that our key
+expectation for this model was still met -- specifically, it predicts the highest first-Beta generation probabilities at
+the edges of the bounded features.
+
+% The only exception is that the representativeness model is able to predict decreasing generation probabilities
+% with increasing distance from the center of the feature space along the toroidal feature, but not the bounded feature.
+% As expected, the first Betas tend to be generated at the edges of the bounded feature.
+
+% After fitting the models to data from Experiment 3, there appears to be no major qualitative
+% deviations in the distribution of first-Betas for the square stimuli (Figures \ref{fig:e3-firstbetas}a and b) from
+% previous expectations (i.e., as seen in Figure \ref{fig:e3-squares}), however this is not the case for the circle
+% stimuli (Figures \ref{fig:e3-firstbetas}c and d compared to Figure \ref{fig:e3-circles})). Specifically, we see greater
+% variance along the unbounded feature for PACKER, and most strikingly, that the representativeness model is able to
+% expand the distribution of generating first-Betas into more centrally located regions \ref{fig:e3-firstbetas} for circle
+% stimuli. However as predicted, even with this distribution, the maximum probability of the representativeness model
+% generating at an edge (.0438) is higher than generating off-edge (.0394), indicating a preference for producing
+% first-Betas at the edge. This is in contrast to both PACKER and the empirical data, which both show a general preference
+% for centrally-located first-Betas.
 
@@
 
-The contrast models were fit by optimizing their fit to only Experiment 3. Fit
-statistics and parameter values are presented in Table
-\ref{table:corner-model-fits}. Although not our main focus in this section, we
-also provide the statistics for the classical models -- unsurprisingly, they do
-not perform as well as the contrast models. In line with our qualitative
+The contrast models were fit by optimizing them to only Experiment 3. Fit statistics and parameter values are presented
+in Table \ref{table:corner-model-fits}. Although not our main focus in this section, we also provide the statistics for
+the classical models -- unsurprisingly, they do not perform as well as the contrast models. In line with our qualitative
 observations, PACKER emerges as the best-fitting model in aggregate.
@@
     2 and 3. Smaller AIC values correspond to better model fits (adjusted
-    for number of parameters.)}
+    for number of parameters).}
 \label{table:corner-model-fits}
@@
 structure (feature variance, correlation) and physical instantiation (location
-within the stimulus space) of generated categories: In Experiment 2, we found
+within the feature space) of generated categories: In Experiment 2, we found
 that the unoccupied regions of the domain influenced the distributional
@@
 that qualitatively distinguished the two contrast models. Specifically, in this
-condition where exemplars were located at the corners of the stimulus space,
+condition where exemplars were located at the corners of the feature space,
 PACKER predicted the initial generation of centrally-located novel exemplars
@@
 \subsection{Similarity and Contrast in Cognition}
-We propose that category contrast as a primary constraint in categorization.
-For categories to be useful, they should not all be identical, or in other words, they should be different.
-Thus, a newly generated category should be different than pre-existing categories.
-Beyond its role
-in category generation specifically, category contrast is also of fundamental importance in
+We propose category contrast as a primary constraint in categorization. For
+categories to be useful, they should not all be identical, or in other words,
+they should be different. Thus, a newly generated category should be different
+than pre-existing categories. Beyond its role in category generation
+specifically, category contrast is also of fundamental importance in
 categorization more broadly. All other factors held constant, new categories are
@@
 literature is primarily interested in the generation of categories within a set
-of observed exemplars where no prior categories are learned, our category
-generation studies are focused on the production of novel category exemplars
-themselves. Despite this distinction, both types of categorization research are
-ultimately interested in the question of how categories can be formed. The results
-of our current work indicate that the formation of categories is constrained by
-some measure of contrast. To our knowledge, this explicit investigation of
-category contrast has not yet been attempted in the unsupervised categorization
-literature.
+of observed exemplars where no prior categories are learned
+\citep[e.g.,][]{pothos2011measuring}, our category generation studies are
+focused on the production of novel category exemplars themselves. Despite this
+distinction, both types of categorization research are ultimately interested in
+the question of how categories can be formed. The results of our current work
+indicate that the formation of categories is constrained by some measure of
+contrast. To our knowledge, this explicit investigation of category contrast has
+not yet been attempted in the unsupervised categorization literature.
 
@@
 category was thus completed by sampling a new category mean (uniform over
-stimulus space) and covariance matrix from the common prior distribution.
+feature space) and covariance matrix from the common prior distribution.
 Because the shared prior distribution's parameters were unobserved, the
@@
 calculating the multivariate Gaussian components of the representativeness equation
-$R(x,h)$ (Equation \ref{eq:representativeness} of the main text.) Specifically,
+$R(x,h)$ (Equation \ref{eq:representativeness} of the main text). Specifically,
 the probability density of the Gaussian component along a single wrapped axis can be computed as:
@@
 where $p(.|h)$ represents the probability density of the Gaussian component $h$
-on a line (i.e., not wrapped around a unit circle.)
+on a line (i.e., not wrapped around a unit circle).
 
@@
 and minimum values of $p(x + 2\pi i|h)$ computed thus far is large enough (e.g.,
-10,000.) For our modeling exercises here we impose both contraints, although the
+10,000). For our modeling exercises here we impose both contraints, although the
 first constraint shows some redundancy since we find that $i$ rarely, if ever,

### 2019-09-06 Manuscripts/cogsci-journal/cogsci-journal.tex
Cog Sci manuscript finalized, and ready for submission.

--- 2019-09-02 Manuscripts/post-cogpsych/post-cogpsych.tex
@@
 \vfill
-Joseph L. Austerweil\textsuperscript{1}
-Nolan Conaway\textsuperscript{1}, 
-Shi Xian Liew\textsuperscript{1},
-and Kenneth J. Kurtz\textsuperscript{2}
+Joseph L. Austerweil\textsuperscript{1+}
+Shi Xian Liew\textsuperscript{1+},
+Nolan Conaway\textsuperscript{2}, 
+and Kenneth J. Kurtz\textsuperscript{3}
 \\[\baselineskip]
-\textsuperscript{1}Department of Psychology, University of Wisconsin-Madison, Madison, WI, USA
-\textsuperscript{2}Department of Psychology, Binghamton University, Binghamton, NY, USA
+\textsuperscript{1}Department of Psychology, University of Wisconsin-Madison, Madison, WI, USA\\
+\textsuperscript{2}Shutterstock, New York, NY, USA\\
+\textsuperscript{3}Department of Psychology, Binghamton University, Binghamton, NY, USA
 \\[1in]
@@
 Correspondence concerning this article should be addressed to: 
-Joseph L. Austerweil, 1202 West Johnson Street, Madison, WI 53706.
-E-mail: austerweil@wisc.edu
+Shi Xian Liew, 1202 West Johnson Street, Madison, WI 53706.
+E-mail: shixianliew@gmail.com
+
+\textsuperscript{+}These authors share equal contribution to this work.
 
@@
 inclined to seek out something healthier and less carbohydrate-heavy (e.g., a
-Caesar salad) compared to something similar (a different flavour of instant
+Caesar salad) compared to something similar (a different flavor of instant
 noodles). An actor who is dissatisfied with being typecast as a villain may seek
 out roles as heroes as opposed to playing other antagonistic characters. An
-especially compelling example can be seen in \cite{askin2017} who found that
-new songs that were more different to their peers were generally more popular
-than new songs that were similar to others. It appears that once there is the
+especially compelling example can be seen in \cite{askin2017} who found that new
+songs that were more different to their peers were generally more popular than
+new songs that were similar to others. It appears that once there is the
 motivation to generate something new, there is an associated motivation to
-generate something \emph{different}.
+generate and observe something \emph{different}.
 
@@
 The second hypothesis for how contrast might affect categorization is contrast
-as {\em representativeness}. The representativeness heuristic \citep{kahneman1972subjective}
-states that people make judgments regarding an outcome based on how
-representative it is of the evidence given in the current context. The heuristic
-is a powerful theoretical construct that has been used to capture a range of
-complex patterns of human judgments \citep{kahneman73,tversky74,tversky83},
-especially those that deviate from normative theory (as given by a
-straightforward application of Bayes' rule). For example, the coinflip sequence
-TTHTH is perceived to be more random (due to it being representative of a
-coinflip sequence) than TTTTT (despite both having the same probability of
-having been generated from a fair coin). Although there is a healthy debate as
-to whether perceived randomness {\em really} deviates from normative theory
-\citep{griffiths18,hahn09}, the representativeness heuristic remains one of the
-main explanations of human judgments \citep{reimers18}. To formalize the
-hypothesis that people expect a new category to be representative of the
-opposite of the current category, or contrast as {\em representativeness}, we
-use the Bayesian formulation of representativeness \citep{tenenbaum01}, which
-has been used successfully to capture human performance in color categorization
-\citep{abbott16}, image category learning \citep{abbott11}, and language grammar
-learning \citep{rafferty10}. We do so by extending the state-of-the-art category
-generation model \citep{jern2013probabilistic} and find that human category
-generation is best captured as generating representative, rather than probable,
-samples.
+as {\em representativeness}. The representativeness heuristic
+\citep{kahneman1972subjective} states that people make judgments regarding an
+outcome based on how representative it is of the evidence given in the current
+context. The heuristic is a powerful theoretical construct that has been used to
+capture a range of complex patterns of human judgments
+\citep{kahneman73,tversky74,tversky83}, especially those that deviate from
+normative theory (as given by a straightforward application of Bayes' rule). For
+example, the coinflip sequence TTHTH is perceived to be more random (due to it
+being representative of a coinflip sequence) than TTTTT (despite both having the
+same probability of having been generated from a fair coin). Although there is a
+healthy debate as to whether perceived randomness {\em really} deviates from
+normative theory \citep{griffiths18,hahn09}, the representativeness heuristic
+remains one of the main explanations of human judgments \citep{reimers18}. To
+formalize the hypothesis that people expect a new category to be representative
+of the opposite of the current category, or contrast as {\em
+  representativeness}, we use the Bayesian formulation of representativeness
+\citep{tenenbaum01}, which has been used successfully to capture human
+performance in color categorization \citep{abbott16}, image category learning
+\citep{abbott11}, and language grammar learning \citep{rafferty10}. We do so by
+extending the state-of-the-art category generation model
+\citep{jern2013probabilistic} and find that human category generation can
+also be captured as generating representative, rather than probable, samples.
 
@@
 only features that are not characteristic of the retrieved category (e.g., if
-{\em bird} was retrieved, the presence of {\em wings} would not change, but {\em
-  color} might). This theory incorporates elements of the highly influential
+{\em bird} was retrieved, the presence of {\em wings} would not change, but {\em color} might). This theory incorporates elements of the highly influential
 basic-level categories framework \citep{rosch1975cognitive,rosch1976basic}, as
@@
   is plausible that some hierarchical Bayesian models could be created that
-  generates categories different from each other. However, this model would not
+  generate categories different from each other. However, this model would not
   be a standard application or extension of most pre-existing hierarchical
@@
 But, it requires an additional factor, which in our case is having a different
-goal than predicting the next category.
+goal than predicting exemplars from a new category.
 
@@
 PACKER within an exemplar framework reflects the facts that exemplar models have
-been throughly evaluated, are strongly theoretically motivated, and dominate
+been thoroughly evaluated, are strongly theoretically motivated, and dominate
 much of the theoretical and empirical work in categorization. The focus of our
@@
 \end{equation}
-% 
-%where $\theta$ ($\theta \geq 0$) is a free parameter controlling response
-%determinism.
+
+Note, because $\theta_t$ and $\theta_c$ are unconstrained, their size plays the
+role of the response determinism parameter for the exponentiated Luce choice
+rule.
 
@@
 \label{section:copytweak-packer}
-%: Both models are
-%exemplar-based, and both models generate new items according to their similarity
-%to known members of the target class. However, PACKER diverges from the
-%copy-and-tweak model by including a contrast mechanism, enabling generation
-%according to dissimilarity to members of opposing categories. As a consequence,
-%copy-and-tweak can be realized as a parameterization of the PACKER model that is
-%insensitive to category contrast. Specifically, when $\theta_c = 0$ and
-%$\theta_t > 0$ (see Figure \ref{fig:packer-examples}, panel B), $f(x_j) = 0$ for
-%$x_j$ belonging to contrast categories; thus, PACKER is not influenced by these
-%items, and is mathematically equivalent to a copy-and-tweak approach.
 
@@
 instance, the literature on the \emph{graded structure} of categories explores
-how some exemplars are better representations of their categories
+how some exemplars are better exemplars of their categories
 \citep{palmeri2001central,barsalou1985ideals}. In this article we adopt the
@@
 In Equation \ref{eq:representativeness}, $p(h')$ is the prior probability
-distribution on the hypothesis space that excludes $h$. Note that when there are
-only two hypotheses (e.g. two categories) involved in the domain, this prior
-takes the value of $1$ since there is ever only one alternative hypothesis.
+distribution on the hypothesis space that excludes $h$ (i.e., we effectively set
+$p(h)$ to 0 and then renormalize the priors). Note that when there are only two
+hypotheses (e.g. two categories) involved in the domain, this prior takes the
+value of $1$ since there is ever only one alternative hypothesis.
 
@@
 such that $\theta \geq 0$).
+
+Here, we implement the representativeness mechanism at the level of
+distributions over category exemplars. However, with a hierarchical model, it is
+also possible to apply the representativeness mechanism at other levels, such as
+the level of the distribution of category statistics (i.e., their means and
+covariances). For simplicity, we focus only on one type of representativeness
+model (i.e., where the mechanism is implemented at the level of distribution
+over category exemplars).
 
@@
 (3.0--5.8cm), see Figure \ref{fig:e1-conditions}. The assignment of perceptual
-features (color, size) to axes of the conceptual space (x, y)  and the
+features (color, size) to axes of the conceptual space (X, Y)  and the
 direction of variation along each axis (e.g., $dark \rightarrow light$ or $light
@@
 classic result that generated categories often possess the same distributional
-properties as previously-known categories.
-Given the increased importance of replication within psychology \citep{zwaan2018making}, it is important as it serves as a conceptual replication of \citet{jern2013probabilistic}.
-For each generated category, we
-computed the category range along each axis (X, Y), as well as the correlation
-between features. These data, shown in Figure \ref{fig:e1-statsboxes}, reveal
-broad individual differences: Within each condition, participants generated
-categories spanning the entire X- and Y- axis as well as categories that spanned
-very little along each. Likewise, in each condition participants generated
-categories possessing strongly positive, neutral, and strongly negative
-correlations between the dimensions. Comparing the distributional statistics
-between conditions yields a broad yet, as we will see, misleading replication of
-the classic effect.
+properties as previously-known categories. Given the increased importance of
+replication within psychology \citep{zwaan2018making}, it is important as it
+serves as a conceptual replication of \citet{jern2013probabilistic}. For each
+generated category, we computed the category range along each axis (X, Y), as
+well as the correlation between features. These data, shown in Figure
+\ref{fig:e1-statsboxes}, reveal broad individual differences: Within each
+condition, some participants generated categories spanning the entire X- and Y-
+axis whereas other participants categories that spanned very little along each.
+Likewise, in each condition participants generated categories possessing
+strongly positive, neutral, and strongly negative correlations between the
+dimensions. Comparing the distributional statistics between conditions yields a
+broad yet, as we will see, misleading replication of the classic effect.
 
@@
 
-However, our ability to detect differences in Y-Axis range using a standard
+However, our ability to detect differences in Y-axis range using a standard
 $t$-test between the conditions is, in this case, diminished due to the
@@
 the Middle condition and $W=0.85, p < .001$ for the Bottom condition).
-Figure \ref{fig:e2-yranges} depicts the Y-Axis range and Y-Axis position of exemplars generated by each participant.
+Figure \ref{fig:e2-yranges} depicts the Y-axis range and Y-axis position of exemplars generated by each participant.
 The categories are sorted by overall range, and then colored by training condition.
@@
 generated categories spanning almost none of the Y-axis.
-Indeed, the median produced Y-Axis range is much smaller in the Bottom than Middle condition, whereas they are essentially identical in X-Axis range.
+Indeed, the median produced Y-axis range is much smaller in the Bottom than Middle condition, whereas they are essentially identical in X-axis range.
 The non-normality of the Y-axis range distributions thus requires that we use a different analytic approach to
@@
     \begin{center} \inputpgf{figs/}{e2-yranges.pgf}
-    \caption{Y-Axis range and position of the participant-generated categories
+    \caption{Y-axis range and position of the participant-generated categories
 from Experiment 2. Each line corresponds to a participant's category, with
@@
 
-To obtain a global measure of the quality of each model's account, we fitted the
+To obtain a global measure of the quality of each model's account, we fit the
 parameters of each model to our entire dataset (Experiments 1 and 2 combined),
@@
 model's predictions of the observed responses (1220 responses from 305 total
-participants). We fitted three parameters in the PACKER model ($c$, $\theta_t$,
+participants). We fit three parameters in the PACKER model ($c$, $\theta_t$,
 and $\theta_c$; see Section \ref{section:PACKER-definition}), as well as four in
@@
 \ref{section:representativeness-definition} and Appendix
-\ref{ap:hsampling-definition} respectively). We fitted only two parameters for
+\ref{ap:hsampling-definition} respectively). We fit only two parameters for
 the copy-and-tweak model ($c$, and $\theta_t$), as $\theta_c$ is held constant
@@
 
-While PACKER's advantage may tentatively be attributed to the model's
-sensitivity to category contrast (this will be explored in detail below), the
-advantage shown by copy-and-tweak over the hierarchical Bayesian model may be
-attributed to its exemplar-based representation of category B, as opposed to forcing a
-prototype-based representation as assumed by the hierarchical Bayesian model. As
-observed in Figures \ref{fig:e1-samples} and \ref{fig:e2-samples}, the generated
-categories we observed were often widely distributed, with no items near the
-category prototype. This aspect of the data is inconsistent with the
-multivariate normal distributions (similar to prototypes) used to represent
-categories in the \cite{jern2013probabilistic} model, but can be easily
-accounted for using an exemplar-based approach.
-Interestingly, representativeness using a prototype approach fits better than an exemplar-based approach.
-Understanding these differences more thoroughly is left for future research.
+While PACKER's advantage over copy-and-tweak may tentatively be attributed to
+the model's sensitivity to category contrast (this will be explored in detail
+below), the advantage shown by copy-and-tweak over the hierarchical Bayesian
+model may be attributed to its exemplar-based representation of category B, as
+opposed to forcing a prototype-based representation as assumed by the
+hierarchical Bayesian model. As observed in Figures \ref{fig:e1-samples} and
+\ref{fig:e2-samples}, the generated categories we observed were often widely
+distributed, with no items near the category prototype. This aspect of the data
+is inconsistent with the multivariate normal distributions (similar to
+prototypes) used to represent categories in the \cite{jern2013probabilistic}
+model, but can be easily accounted for using an exemplar-based approach.
+Interestingly, representativeness using a prototype approach fits better than an
+exemplar-based approach.
 
@@
 
-The effect of contrast for both models is similar in that they both tend to produce exemplars that are distant from the
-Alpha category. However, this is realized in different ways for the two models. The representativeness model does not
-explicitly use a measure sensitive to distance from the location of exemplars in the contrasting category. Instead, it
-is the assumption of unimodal distributions that results in the model producing novel exemplars dissimilar to learned
-exemplars. Consequently, it is possible to design a configuration of Alpha exemplars where the representativeness model
-predicts novel exemplars should be \emph{similar} to the learned exemplars. However, PACKER cannot predict this because
-it assumes that the similarity weight for contrasting exemplars is negative.
-
-In this section, we focus on a different category type -- the Corner category -- that contrasts the two models of
-contrast. Here, each Alpha exemplar is located at the corners of the feature space (see Figure \ref{fig:e3-squares}a).
-
-We first identify the predictions of each model in this condition. Second, we test these contrasting predictions in a
-behavioral experiment. We then conduct a model-based analysis of the two contrast models.
+The effect of contrast for both models is similar in that they both tend to
+produce exemplars that are distant from the Alpha category. However, this is
+realized in different ways for the two models. The representativeness model does
+not explicitly use a measure sensitive to distance from the location of
+exemplars in the contrasting category. Instead, it is the assumption of unimodal
+distributions that results in the model producing novel exemplars dissimilar to
+learned exemplars. Consequently, it is possible to design a configuration of
+Alpha exemplars where the representativeness model predicts novel exemplars
+should be \emph{similar} to the learned exemplars. However, PACKER cannot
+predict this because it assumes that the similarity weight for contrasting
+exemplars is negative.
+
+In this section, we focus on a different category type -- the Corner category --
+that contrasts the two models of contrast. Here, each Alpha exemplar is located
+at the corners of the feature space (see Figure \ref{fig:e3-squares}a). We first
+identify the predictions of each model in this condition. Second, we test these
+contrasting predictions in a behavioral experiment. We then conduct a
+model-based analysis of the two contrast models.
 
@@
 
-Thus, in this experiment, we also examine how making a boundary toroidal (e.g., line rotation) affects the prediction of
-each model. The representativeness model prefers to generate categories on the feature that still has a boundary.
-PACKER, in contrast, should continue to generate categories that are most distant from the learn category regardless of
-the presence or absence of feature boundaries -- in this condition it would be the center of the feature space.
-Experiment 3 manipulates whether one feature is bounded or toroidal (always displayed as the Y-axis). One condition uses
-the same stimuli as before, squares varying in lightness and size, both of which are bounded features. The other
-condition use a new set of stimuli: circles that vary in size (a bounded feature) and orientation (an unbounded feature;
-Figure \ref{fig:e3-circles}a). Before describing the details of the experiment, we explain in detail each model's
-predictions given the Corner Alphas for a feature space where one feature is bounded and the other is toroidal.
+Thus, in this experiment, we also examine how making one feature toroidal (e.g.,
+line orientation) affects the prediction of each model. The representativeness
+model prefers to generate categories on the feature that still has a boundary.
+PACKER, in contrast, should continue to generate categories that are most
+distant from the learn category regardless of the presence or absence of feature
+boundaries -- in this condition it would be the center of the bounded feature.
+Experiment 3 manipulates whether one feature is bounded or toroidal (always
+displayed as the Y-axis). One condition uses the same stimuli as before, squares
+varying in lightness and size, both of which are bounded features. The other
+condition use a new set of stimuli: circles that vary in size (a bounded
+feature) and orientation (an unbounded feature; Figure \ref{fig:e3-circles}a).
+Before describing the details of the experiment, we explain in detail each
+model's predictions given the Corner Alphas for a feature space where one
+feature is bounded and the other is toroidal.
 
@@
         PACKER (b) and the representativeness model (c). The unbounded feature
-        (orientation) is represented on the y-axis, while the bounded feature
-        (size) is represented on a x-axis. As with the square stimuli, feature
+        (orientation) is represented on the Y-axis, while the bounded feature
+        (size) is represented on a X-axis. As with the square stimuli, feature
         and direction assignment were counterbalanced across participants (in
-        the displayed example, circles are rotated clockwise up the y-axis). }
+        the displayed example, circles are rotated clockwise up the Y-axis). }
       \label{fig:e3-circles}
@@
 
-Compared to stimuli with two bounded features, PACKER makes a similar prediction in generating the first Beta exemplar
-when one feature is unbounded, with the curious exception that the highest density regions appear compressed along the
-toroidal feature. Due to the toroidal nature of orientation, the maximum distance from the Alpha exemplars is half of
-the full range of the feature -- for an Alpha exemplar oriented at 0 radians, the furthest distance to another exemplar
-is $\pi$ radians.
-
-For the representativeness model, adding a toroidal feature changes the probability distribution for the underlying
-Bayesian model. For the toroidal feature, space wraps back to $0$ radians once it reaches $2\pi$ (we have conducted a
-change of variables where $y \rightarrow y$ mod $2\pi$, where mod is the modulus operator). As we cross $2\pi$ and reach
-$0$ again, the remaining probability density also gets wrapped around, meaning that the density at $2\pi$ is added to
-$0$ (and $2\pi + \epsilon$ to $\epsilon$, and so on). This is called the {\em wrapped} Gaussian distribution
-\citep{mardia1972}. Note that the more familiar distribution for circular features, the Von Mises distribution, is an
-approximation to the wrapped Gaussian distribution. See Appendix B for implementational details of the
-representativeness model in a bounded and toroidal feature.
+Compared to stimuli with two bounded features, PACKER makes a similar prediction
+in generating the first Beta exemplar when one feature is unbounded, with the
+curious exception that the highest density regions appear compressed along the
+toroidal feature. Due to the toroidal nature of orientation, the maximum
+distance from the Alpha exemplars is half of the full range of the feature --
+for an Alpha exemplar oriented at 0 radians, the furthest distance to another
+exemplar is $\pi$ radians.
+
+For the representativeness model, adding a toroidal feature changes the
+probability distribution for the underlying Bayesian model. For the toroidal
+feature, space wraps back to $0$ radians once it reaches $2\pi$ (we have
+conducted a change of variables where $y \rightarrow y$ mod $2\pi$, where mod is
+the modulus operator). As we cross $2\pi$ and reach $0$ again, the remaining
+probability density also gets wrapped around, meaning that the density at $2\pi$
+is added to $0$ (and $2\pi + \epsilon$ to $\epsilon$, and so on). This is called
+the {\em wrapped} Gaussian distribution \citep{mardia1972}. Note that the more
+familiar distribution for circular features, the Von Mises distribution, is an
+approximation to the wrapped Gaussian distribution. See Appendix B for
+implementation details of the representativeness model in a bounded and
+toroidal feature space.
 
@@
 
-% It may be argued that our expectation that the representativeness model should
-% \emph{always} tend to produce first-Betas at the edges may be too strong.
-% Specifically, this prediction is based on the model that is optimized to data
-% from Experiments 1 and 2, and it is conceivable that the model may make
-% drastically different predictions (such as some favoring the initial generation
-% of centrally-located Betas more than Betas at the edges) in the current
-% experiment when it is actually fit to empirical data from this experiment.
-% However, we contend that this prediction of the representativeness model is a
-% fundamental property of the model due to its assumption of unimodal category
-% distributions. At the very least, the maximum probability of generating a
-% centrally located first Beta should almost never be higher than the maximum
-% probability of generating a first Beta at an edge.
-
-% It is also worth noting that since the representativeness model does not explicitly
-% recognize similarities as distances within the feature space (unlike PACKER),
-% there is no compression in density regions along the unbounded feature axis.
-% This provides us with an additional qualitative marker distinguishing the
-% representativeness model from PACKER -- if the range of the entire Beta category
-% along the unbounded feature is observed to be shorter than that of the bounded
-% feature, this would be stronger evidence for PACKER rather than the
-% representativeness model.
-
-% In the following
-% section, we explore -- using both qualitative comparisons and quantitative fits
-% -- whether PACKER or the representativeness model is a better account of
-% behavioral data in this Corner experiment.
-
-% For the representativeness model, the single unbounded feature essentially
-% prevents the formation of edge effects along that feature.
-Consequently, this results in higher probabilities of the generating the first Beta exemplars along the edges of the bounded feature (Figure \ref{fig:e3-circles}c).
-As this
-% model does not explicitly recognize similarities as distances along the feature
-% space (unlike PACKER), it does not experience the compression in density regions
-% along the unbounded feature axis. Intuitively, this suggests that the
-% representativeness model (which prefers generating exemplars at the boundaries) will prefer to generate stimuli on the boundaries of the bounded feature with some amount of variance over unbounded feature.
-
-% We have identified the conditions under which the contrast models may make
-% differing predictions, however this distinction is not very useful without any
-% empirical support. Data from the next experiment will most certainly support one
-% model or the other
-
-%[Maybe work on transition if there's time]
-
 \subsection{Participants, Materials, and Procedure}
 
-We recruited 89 participants from Amazon Mechanical Turk who were randomly assigned to the square stimuli ($N=46$), or
-to the circle stimuli ($N=43$). The square stimuli were exactly the same as in Experiments 1 and 2, but with Alpha
-exemplars in the corners of the feature space. The circle stimuli were made up of unfilled circles with a radial line
-extending from its center to its edge. These stimuli varied along two features -- their diameter ranging from 3.0--5.8
-cm, and their orientation ranging from $2.587$--$2\pi + 2.587$ radians, with $0$ radians indicating a completely
-horizontal line. The line widths of the circles and radial line were set to 0.05 cm. Similar to Experiments 1 and 2,
-participants first observed a training phase where four Alpha exemplars were presented across three blocks followed by
-the generation phase, and the features values are divided into 9 discrete steps. The training condition was the Corner
-category for all participants.
+We recruited 89 participants from Amazon Mechanical Turk who were randomly
+assigned to the square stimuli ($N=46$), or to the circle stimuli ($N=43$). The
+square stimuli were exactly the same as in Experiments 1 and 2, but with Alpha
+exemplars in the corners of the feature space. The circle stimuli were made up
+of unfilled circles with a radial line extending from its center to its edge.
+These stimuli varied along two features: their diameter within the range of
+[3.0, 5.8] cm (inclusive), and their orientation in the range [$2.587$,
+$2\pi + 2.587$] radians (inclusive), with $0$ radians indicating a completely
+horizontal line and orientation increasing in an anti-clockwise direction. The
+line widths of the circles and radial line were set to 0.05 cm. Similar to
+Experiments 1 and 2, participants first observed a training phase where four
+Alpha exemplars were presented across three blocks followed by the generation
+phase, and the features values are divided into 9 discrete steps. The training
+condition was the Corner category for all participants.
 
@@
 
-The distributions of generated first-Betas (Figure \ref{fig:e3-behave}) indicate further qualitative support for PACKER.
-In particular, the probability of generating the first Beta exemplar for circles along the bounded dimension is
-significantly negatively correlated with distance from the center ($r(3) = -.95, p = .01$). Although the correlations
-for squares ($r(3)= -.50, p = .40$ and $r(3)= -.31 p = .61$ for x and y dimensions) were not significant, they were in
-the directions predicted by PACKER. Mean probabilities of of generating first-Betas are presented in Figure
-\ref{fig:e3-distcent}. This analysis was excluded for the toroidal dimension of the circle stimuli because the
-boundaries (and consequently a center) are not defined for this feature.
-
-% An inspection of the distributions of generated first-Betas (Figure
-% \ref{fig:e3-behave}) does not immediately indicate support for either model. The
-% square condition reveals first Betas generated at the corners as well as the
-% center of the feature space, while the circle condition shows greater
-% frequencies at the center. However, we also found that 65\% and 87\% of
-% participants generated their first squares and first circles, respectively, away
-% from the edge of the feature space.\footnote{Strictly speaking, the circle
-%   stimuli do not have edges along its orientation feature, However, for ease of
-%   comparison we will continue to refer to these comparable regions between
-%   stimuli as edges.} Further, the maximum proportion of first-Betas generated
-% off the edges (.13) is higher than that of the maximum proportion in an edge
-% (.04) for the circle stimuli, although the maximum proportions are the same for
-% squares (.06 for both edge and off-edge). This preference for generating
-% first-Betas away from the edge is in line with our hypothesis
-% supporting PACKER.
+The distributions of generated first-Betas (Figure \ref{fig:e3-behave}) indicate
+further qualitative support for PACKER. In particular, the probability of
+generating the first Beta exemplar for circles along the bounded dimension is
+significantly negatively correlated with distance from the center
+($r(3) = -.95, p = .01$). Although the correlations for squares
+($r(3)= -.50, p = .40$ and $r(3)= -.31, p = .61$ for X and Y dimensions
+respectively) were not significant, they were in the directions predicted by
+PACKER. Mean probabilities of generating first-Betas are presented in Figure
+\ref{fig:e3-distcent}. This analysis was excluded for the toroidal dimension of
+the circle stimuli because the boundaries (and consequently a center) are not
+defined for this feature.
 
@@
       \includegraphics[width=.6\textwidth]{figs/e3-distcent.pdf}
-      \caption{Empirical mean probabilities of generating first Beta exemplars at each point of distance from the center
-        of the feature space. Plot for the toroidal feature of circle stimuli is not presented because a center is not
-        defined for this feature.}
+      \caption{Empirical mean probabilities of generating first Beta exemplars
+        at each point of distance from the center of the feature space.
+        Distances are normalized such that 0 represented the center and 1
+        represents the edge. Plot for the toroidal feature of circle stimuli is
+        not presented because a center is not defined for this feature.}
     \label{fig:e3-distcent}
@@
   $c = 2.08$           &$c = 3.73$    &$\kappa = 0.00$&$\kappa = 0.29$  \\ 
-  $\theta_c = 0.87$    &$\theta_c = 0$ (fixed)  &$\nu = 4.07 \times 10^20$ &$\nu
+  $\theta_c = 0.87$    &$\theta_c = 0$ (fixed)  &$\nu = 4.07 \times 10^{20}$ &$\nu
                                                                              = 1.00$\\
@@
 representativeness; two-tailed Binomial test, $p<0.001$). The representativeness
-model only trended towards fitting participants in the bounded, Square condition
-better (18 out of 46; two-tailed Binomial test $p = 0.18$). 
-%In addition, these results agree with our prediction that the representativeness model would be poorer in predicting data from circle stimuli (26\% of participants observing circle stimuli) compared to square stimuli (39\% participants observing square stimuli).
+model fit the bounded Square condition better, but still worse than PACKER (18 out of 46; two-tailed Binomial test $p = 0.18$).
 
@@
 Experiment 3, we collected behavioral data on a category generation condition
-that qualitatively distinguished the two contrast models. Specifically, in this
+that was designed to qualitatively distinguish the two contrast models. Specifically, in this
 condition where exemplars were located at the corners of the feature space,
@@
 while the representativeness model predicted novel exemplars similarly located
-at the corners. Interestingly, data from this experiment was better fit,
+at the corners. Interestingly, data from this experiment were better fit,
 qualitatively and quantitatively, by PACKER than the representativeness model.
-
-% Ultimately, our category generation data were best explained by the
-% representativeness model. The exact nature of its superiority over the other
-% models is still not yet clear, although we find evidence that it only barely
-% surpasses PACKER when considering known categories with extreme exemplars.
 
@@
 beyond the available information. Our work is fairly unique in that people learn
-a single category through positive examples and then create another category
-that would make sense in the domain. In other words, only one category is
-provided to be learned and the use that is required of that category is to
-generate a contrast category. Such scenarios have direct application to
-real-world situations. For instance, consider being stranded in an unfamiliar
-environment where you eat and get sick from a mushroom with certain features. In
-avoiding such mushrooms and seeking alternative sources of nutrition, it would
-behoove you to imagine what features an edible, safe mushroom would possess (e.g.,
-if the poisonous mushroom had a long stem and gills, perhaps a mushroom with a short stem that did not have gills would be safe to
-consume). In the present studies, we have learned something about the form that
-such expectations are likely to take.
+a category through positive examples and then create another category that would
+make sense in the domain. Such scenarios have direct application to real-world
+situations. For instance, consider a rock musician in the 1970s who wants to
+write new types of rock music. Given the diversity existing rock music that the
+musician is aware of, the musician must identify features of this new sub-genre
+that are unique to that sub-genre (e.g., introduce instruments that may not
+typical to rock music at the time, such as accordions or synthesizers). In the
+present studies, we have learned something about the form that such expectations
+on new categories are likely to take.
 
@@
 \citep{imai1965discriminability,milton2004influence,ahn1992two}, where
-participants sort unlabelled family resemblance category exemplars by focusing
+participants sort unlabeled family resemblance category exemplars by focusing
 on a single feature, was not consistently observed in our tasks. In our
@@
 How well does a representativeness model that uses exemplar similarity to
-represent the category’s density over features explain our experimental results?
-We find that for data across all experiments, this model substantially
+represent the category's density over features explain our experimental results?
+We find that for data from experiments 1 and 2, this model substantially
 outperforms its classical counterpart, but not to the extent of PACKER or the
-representativeness model. Specifically, the model fits yielded $L = -4815$ and
-$AIC=9633$ for the entire set of data from Experiments 1 and 2, and $L = -3474$,
-$AIC=6953$ for the same data excluding the first trials. When we fit the model
-to data from Experiments 1,2 and 3, we find that $L = -6301$, $AIC = 12606$.
-These results suggest that the predictive advantages to including the
-representativeness mechanism is not limited to a hierarchical Bayesian model,
-but can also be found when applied to an exemplar model. Although its
-performance is not as strong as either of the fully-developed contrast models,
-there is a benefit to the representativeness mechanism that is independent from
-any interaction with a hierarchical Bayesian framework.
-
-Ultimately, it appears that there is no single model that can capture the
-entirety of our empirical data. While the representativeness model, and indeed
-the representativeness mechanism itself, offers a predictive advantage in
-category generation across all out models, PACKER performs better when
-accounting for data from the Corner conditions. While we have established that contrast
-mechanisms are essential in modeling category generation, there is no conclusive
-evidence that one form of contrast is superior to the other across all
-conditions. 
-%Xian: Consider rewriting paragraph if there's time.
+representativeness model. Specifically, the model fits here yielded $L = -4815$
+and $AIC=9633$ for the entire set of data from Experiments 1 and 2 and
+$L = -3474$, $AIC=6953$ for the same data excluding the first trials (cf. Table
+\ref{table:global-model-fits}). When we fit the model to data from Experiment 3,
+we find that $L = -1486$, $AIC = 2975$, which indicates a poorer fit than its
+classical counterpart (cf. Table \ref{table:corner-model-fits}). Overall, the
+representativeness model performed best when generating categories where
+previous category examples were not at the edges of the feature space.
+Conversely, PACKER performed better when the previous category was at the
+extremes of the features or if a feature is bounded. These results suggest that
+the predictive advantages to including the representativeness mechanism is not
+limited to a hierarchical Bayesian model, but can also be found when applied to
+an exemplar model. Although its performance is not as strong as either of the
+fully-developed contrast models, there can be a benefit to the
+representativeness mechanism that is independent from any interaction with a
+hierarchical Bayesian framework. Ultimately, it appears that there is no single
+model that can capture the entirety of our empirical data.
 
@@
 (insects are probably small and have patterns on their body that suit the
-environment). However, the relation of these contraints (and the interactions
+environment). However, the relation of these constraints (and the interactions
 between them) to the role of contrast is not currently well understood.
@@
 well as a hierarchical Bayesian model, provide strong support for the claim that
-category contrast is of fundamental importance to categorization. More broadly,
-our results demonstrate that popular explanatory approaches from basic research
-in cognitive science can offer a precise, quantitative account of a behavior as
-complex as that of category generation, and in exploring them, we can increase
-our understanding of categorization as well.
+category contrast is of fundamental importance to categorization.
 
@@
 research was provided by the Office of the VCRGE at the UW - Madison with
-funding from the WARF. We thank Charles Kemp for feedback on a previous version
-of this manuscript. We also thank him and Alan Jern for providing code and data.
+funding from the WARF and the Vilas Life Cycle Fellowship. We thank Charles Kemp
+for feedback on a previous version of this manuscript. We also thank him and
+Alan Jern for providing code and data.
 \end{flushleft}
@@
 and minimum values of $p(x + 2\pi i|h)$ computed thus far is large enough (e.g.,
-10,000). For our modeling exercises here we impose both contraints, although the
+10,000). For our modeling exercises here we impose both constraints, although the
 first constraint shows some redundancy since we find that $i$ rarely, if ever,

### 2019-09-06 Experiments/cogpsych-code/validate_data.py

@@ -7,6 +7,9 @@ choices = ['pooled',
            'catassign',
            'xcr',
            'xcrABC',
+           'xcrA',
+           'xcrB',
+           'xcrC',
            'midbot',
            'midbotABC',
            'corner',
@@ -60,6 +63,21 @@ elif dataname == 'xcrABC':
         filename = dataname
         raw_db = 'experiment-xcrABC.db'
         task = "generate"
+elif dataname == 'xcrA':
+        # XOR, Cluster, Row conditions with Not-Alpha condition
+        filename = dataname
+        raw_db = 'experiment-xcrA.db'
+        task = "generate"
+elif dataname == 'xcrB':
+        # XOR, Cluster, Row conditions with Beta-only condition
+        filename = dataname
+        raw_db = 'experiment-xcrB.db'
+        task = "generate"
+elif dataname == 'xcrC':
+        # XOR, Cluster, Row conditions with Beta-Gamma condition
+        filename = dataname
+        raw_db = 'experiment-xcrC.db'
+        task = "generate"        
 elif dataname == 'midbot':
         # experiment 2 only mid bottom conditions
         filename = dataname

### 2019-09-20 Manuscripts/cogsci-journal/cogsci-journal.tex
Added notAvsB subdirectory. Also short writeup on some of the interesting modeling results. Next step, get efficient individual fits working -- treat each ppt as a separate dataset and run 1000 chtc chunks on it.

@@
 Shi Xian Liew\textsuperscript{1+},
-Nolan Conaway\textsuperscript{2}, 
-and Kenneth J. Kurtz\textsuperscript{3}
+Nolan Conaway\textsuperscript{1}, 
+and Kenneth J. Kurtz\textsuperscript{2}
 \\[\baselineskip]
 \textsuperscript{1}Department of Psychology, University of Wisconsin-Madison, Madison, WI, USA\\
-\textsuperscript{2}Shutterstock, New York, NY, USA\\
-\textsuperscript{3}Department of Psychology, Binghamton University, Binghamton, NY, USA
+\textsuperscript{2}Department of Psychology, Binghamton University, Binghamton, NY, USA
 \\[1in]
@@
         at each point of distance from the center of the feature space.
-        Distances are normalized such that 0 represented the center and 1
+        Distances are normalized such that 0 represents the center and 1
         represents the edge. Plot for the toroidal feature of circle stimuli is
@@
 Annual Meeting of the Society for Mathematical Psychology. Support for this
-research was provided by the Office of the VCRGE at the UW - Madison with
+research was provided by the Office of the VCRGE at the UW-Madison with
 funding from the WARF and the Vilas Life Cycle Fellowship. We thank Charles Kemp

### 2019-09-20 Manuscripts/model-descriptions/negative-weighted/four_models.tex
Added notAvsB subdirectory. Also short writeup on some of the interesting modeling results. Next step, get efficient individual fits working -- treat each ppt as a separate dataset and run 1000 chtc chunks on it.

@@
 
-When generating exemplars from categories defined by a label that describes what is \emph{not} already learned, we can
-define a simple model where these exemplars are samples drawn with a probability proportional to the inverse (or 1-?)category
-distribution of
+When generating exemplars from categories defined by a label that describes what
+is \emph{not} already learned, we can imagine a mechanism that focuses on the
+areas of the feature space that is not occupied by learned categories. This can
+be implemented with a simple model where exemplars from these novel categories
+are generated with a probability inversely proportional to the density of all
+other categories. For simplicity, we assume that each individual category has
+its exemplars distributed as a multivariate Gaussian parameterized by a
+$d$-sized vector of empirical means $M$ and a $d$-by-$d$ empirical covariance
+matrix $S$, where $d$ describes the dimensionality (i.e., number of features).
 
-An interesting property of such a model is that it assumes that categories-by-negation are multi-modal, unlike the
-unimodal assumptions currently adopted by current models of category generation.
+More formally, the probability of generating an exemplar $y$ given a set of
+categories $C'$\footnote{This model assumes that the novel category is a negated
+  category with respect to \emph{all} learned categories. It would be
+  appropriate for data from people generating exemplars that are not what they
+  previously learned. However, if participants were instead instructed to
+  generate exemplars that are not from some subset of a larger set of learned
+  categories (e.g., generate not Alphas, having learned both Alpha and Beta
+  category exemplars), this simple model would not be appropriate. For our
+  experiments so far, we only test participants who have learned one category,
+  so there is no problem at this point. For future development, however, we
+  might want to consider negative-space models that can consider a subset of
+  learned categories. This is probably as simple as having $C'$ represent the
+  relevant subset of learned categories.} excluding the novel category can be described as
+proportional to a density function $n(y|C')$ where
+
+\begin{equation}\label{eq:negmech}
+n(y|C') = -\sum_{c \in C'} \textrm{MvN}(y; M_c,S_c)
+\end{equation}
+
+To get the actual choice probabilities, we normalize these values using an
+exponentiated Luce's choice rule:
+
+\begin{equation}\label{eq:negluce}
+p(y|C') = \dfrac
+{\textrm{exp}(\theta \cdot n(y|C'))}
+{\sum_i \textrm{exp}(\theta \cdot n(y|C'))}
+\end{equation}
+where $\theta$ is a response determinism parameter that is freely estimated.
+
+% (An interesting property of such a model is that it assumes that
+% categories-by-negation are multi-modal, unlike the unimodal assumptions
+% currently adopted by current models of category generation.)
+
+Note that with this simple model the probability of generating an exemplar is
+completely dependent on the learned categories -- there is no explicit
+within-category constraints defined by the model.
+
+Although this model is limited in its complexity, it is straightforward to apply
+as an extension to any of the other models of category generation. We describe
+extensions to four models -- PACKER, copy-and-tweak, Jern and Kemp's (2013)
+hierarchical Bayesian model, and the representativeness model -- in the following sections.
+
+\section{Extending PACKER and copy-and-tweak}
+
+We begin by briefly describing the structure of the PACKER and copy-and-tweak
+models (which can be implemented as a special case of PACKER).
+
+These models predict the generation of novel exemplars by measuring the
+candidate exemplar's similarity to other exemplars. Specifically, the similarity
+between exemplars $x_i$ and $x_j$ is computed using Shepard's law:
+
+\begin{equation}
+s\left(x_i,x_j\right) = \exp \left\{ -c \left[\sum_{k}{ w_k \left| x_{ik} - x_{jk} \right|^r }\right]^{1/r} \right\}
+\label{eq:similarity}
+\end{equation}
+where $k$ indicates each feature, $w_k$ is the weight placed on that feature,
+$c$ is a free-to-vary specificity parameer, and $r$ allows us to define the
+distance metric. In most of our simulations, we fix $r$ to 1 (implementing a
+city-block distance metric). Unless otherwise specified, $w_k$ is fixed to be
+equal across all features.
+
+PACKER takes into account the similarity of a candidate exemplar to the existing
+novel category (i.e., the target category) exemplars as well as its
+dissimilarity to contrast category exemplars. This is implemented as an
+aggregated similarity $a$ between candidate $y$ and all other exemplars $x$:
+
+\begin{equation}
+  a(y, x) = \sum_j{f(x_j) s(y, x_j)}
+\end{equation}
+where $f(x_j) = \theta_t$ when $x_j$ is a member of the target category and
+$f(x_j) = \theta_c$ when $x_j$ is a member of a contrast category.
+
+To extend this model with the negative-space mechanism, we linearly combine this
+aggregated similarity with the densities of the contrast categories $n(.|C')$
+(Equation \ref{eq:negmech}):
+
+\begin{equation}\label{eq:negpacker}
+b(y,x) = a(y,x) + \gamma \cdot n(y|C')
+\end{equation}
+where $\gamma$ is a free-to-vary parameter that describes the amount of weight
+placed on the negative-space mechanism.
+
+The probability that a given item $y$ will be generated given the model's memory
+$x$ is computed using relative summed similarity values across all generation
+candidates $y_i$ (i.e., with an exponentiated Luce's choice rule):
+
+\begin{equation}
+p(y) = \dfrac
+{ \exp  \left \{ b \left( y, x \right) \right \} } 
+{ \sum_i{ \exp  \left \{  b \left( y_i, x \right) \right\}  } }
+\label{eq:packer-choice}
+\end{equation}
+
+The negative-space copy-and-tweak model is instantiated in exactly the same way,
+but with $\theta_c$ set to 0, since the model does not account for contrast
+category similarity. In addition, $\theta_t$ and $\gamma$ combine to serve as the
+response determinism parameter that is free to vary. Response determinism for
+PACKER is a combination of $\theta_c$, $\theta_t$, and $\gamma$ parameters.
+
+\section{Extending Hierarchical Bayesian and Representativeness models}
+
+Unlike the exemplar models, the hierarchical Bayesian and Representativeness
+models do not explicitly measure exemplar similarity on the feature space.
+Instead, both of these Bayesian models assume exemplars $x$ within a category
+$c$ are influenced by an underlying multivariate Gaussian distribution
+parameterized by means $\mu_c$ and covariance matrix $\Sigma_c$.
+
+$\mu_c$ and $\Sigma_c$ are assumed to be Normal-Inverse-Wishart distributed. The
+former can be straightforwardly computed with:
+\begin{equation}\label{eq:mu_c}
+  \mu_c = \dfrac
+    {\kappa\mu_{0} + n_c \bar{x}_c}
+    {\kappa + n_c}
+\end{equation}
+where:
+\begin{itemize}
+    \setlength\itemsep{-0.5em}
+    \item $\mu_{0}$ is the prior mean along $d$ dimensions. Here we set it to the middle of the space.
+    \item $\kappa$ is a scalar hyper-parameter, roughly weighting the importance of $\mu_{0}$. $\kappa$ must be greater than zero.
+    \item $n_c$ is the number of observations in $x_c$
+    \item $\bar{x}_c$ is the sample mean along $d$ dimensions
+\end{itemize}
+
+$\Sigma_c$ is influenced by members of the target category $c$ as well as all
+other members through a domain-level covariance matrix $\Sigma_D$. Assuming the
+observed category covariances $S_c$ are Wishart-distributed, we can compute $\Sigma_D$
+with:
+
+\begin{equation}
+  \Sigma_D = \Sigma_0 + \sum_{c}{S_c}
+\end{equation}
+where $\Sigma_0$ is a $d$-by-$d$ prior covariance matrix defined as the
+$d$-dimensional identity matrix $I_d$ weighted by a free parameter $lambda$:
+\begin{equation}
+    \Sigma_0 =  \lambda I_d
+\end{equation}
+
+With this, we can compute $\Sigma_c$ as:
+
+\begin{equation}
+  \Sigma_c = [\Sigma_D\nu + S_c +
+    \dfrac
+    {\kappa n_c}
+    {\kappa + n_c}
+    (\bar{x_c}-\mu_c)(\bar{x_c}-\mu_c)^T
+  ] (\nu + n_c)^{-1}
+  \label{eq:Sigma_c}
+\end{equation}
+where $\nu$ is an additional free parameter weighting the importance of
+$\Sigma_D$.
+
+As with the other models, choice probabilities are computed using the
+exponentiated Luce's choice rule. With some general function $g(\cdot)$:
+
+\begin{equation}\label{eq:hierLuce}
+p(y) = \dfrac
+{ \exp  \{ \theta \cdot g(\cdot) \} } 
+{ \sum_i{ \exp \{ \theta \cdot g(\cdot) \}  } }
+\end{equation}
+
+The specific form of $g(\cdot)$ is where the hierarchical Bayesian model
+diverges from the representativeness model. For the hierarchical Bayesian model
+, $g(\cdot)$ is the density at $y$ under a multivariate Gaussian
+parameterized by $\mu_c$ and $\Sigma_c$:
+
+\begin{equation}\label{eq:hierg}
+g(\cdot) = \textrm{Normal}(y;\mu_c,\Sigma_c)
+\end{equation}
+
+For the representativeness model, $g(\cdot)$ is the representativeness
+$R(y,h)$ of $y$ from the target hypothesis $h$ compared to all other hypotheses
+$h'$. Here, each hypothesis is represented by the probability density of the
+underlying distribution of a specific category. Formally,
+
+\begin{align}
+g(\cdot) & = R(y,h)\\
+      & = \log \dfrac{p(y|h)}{\sum_{h'\in \mathcal{H}^c}{p(y|h')p(h')}} . 
+\label{representativeness}
+\end{align}
+where $h_c$ is a multivariate Gaussian parameterized by $\mu_c$ and $\Sigma_c$
+for a specific category $c$ and $\mathcal{H^c}$ represents the hypothesis set
+from all contrast categories.
+
+Similar to the exemplar models, the extension of both hierarchical Bayesian and
+the representativeness models are performed by linearly combining the
+negative-space densities computed from $n(y|C')$ to the original function prior
+to normalization via the exponentiated Luce choice rule. Formally, we can
+represent the negative-space extensions of both Bayesian models by amending
+Equation \ref{eq:hierLuce} to:
+
+\begin{equation}\label{eq:hierLuceNeg}
+p(y) = \dfrac
+{ \exp  \{ \theta \cdot g(\cdot) + \gamma \cdot n(y|C') \} } 
+{ \sum_i{ \exp \{ \theta \cdot g(\cdot) + \gamma \cdot n(y|C') \}  } }
+\end{equation}
+
+Given the structure of the Bayesian models, we adopt a more principled
+implementation of $n(y|C')$ by specifying $\mu_c$ (from Equation \ref{eq:mu_c})
+and $\Sigma_c$ (from Equation \ref{eq:Sigma_c}) as the means and covariance matrices
+of the multivariate Gaussian, instead of the empirical $M_c$ and $S_c$ values.
+Specifcally, Equation \ref{eq:negmech} becomes:
+
+\begin{equation}\label{eq:hiernegmech}
+n(y|C') = -\sum_{c \in C'} \textrm{MvN}(y; \mu_c,\Sigma_c)
+\end{equation}
+
+\end{document}

### 2019-09-20 Manuscripts/notAvsB/notAvsB.tex
Added notAvsB subdirectory. Also short writeup on some of the interesting modeling results. Next step, get efficient individual fits working -- treat each ppt as a separate dataset and run 1000 chtc chunks on it.

%!TEX output_directory = latex_out/
\documentclass[12pt]{article}
\usepackage[letterpaper, margin=1in, headheight=15pt]{geometry}
\usepackage{amsmath}
\usepackage{setspace}
\usepackage{pgfplots}
\usepackage{fancyhdr}
\usepackage{csquotes}
\usepackage{todonotes}
\usepackage{verbatim}
\usepackage[toc]{appendix}
\usepackage[natbibapa]{apacite}

% set up PGF
\pgfplotsset{compat=1.6}
\newcommand\inputpgf[2]{{
\let\pgfimageWithoutPath\pgfimage
\renewcommand{\pgfimage}[2][]{\pgfimageWithoutPath[##1]{#1/##2}}
\input{#1/#2}
}}

% set up notes-- different backgrounds for Nolan and Joe!
\newcommand\jlanote[1]{\todo[inline, backgroundcolor = lime]{\textbf{JLA}: #1}}
\newcommand\sxlnote[1]{\todo[inline, backgroundcolor = cyan]{\textbf{SXL}: #1}}

% set up header % 
\pagestyle{fancy}
\fancyhf{} % sets both header and footer to nothing
\renewcommand{\headrulewidth}{0pt}
\lhead{RUNNING HEAD: Independently identified categories vs. categories-by-negation}
\fancyhead[R]{\thepage}

\begin{document}

% ------- TITLE PAGE ------- %
% \begin{center}
% \hfill
% \\[1in]

% Some NotAvsB Title

% \vfill
% Shi Xian Liew\textsuperscript{1},
% Kenneth J. Kurtz?,
% Joseph L. Austerweil\textsuperscript{1}
% \\[\baselineskip]
% \textsuperscript{1}Department of Psychology, University of Wisconsin-Madison, Madison, WI, USA
% \\[1in]

% \vfill

% Author Note

% Correspondence concerning this article should be addressed to: 
% Shi Xian Liew, 1202 West Johnson Street, Madison, WI 53706.
% E-mail: shixianliew@gmail.com

% %\textsuperscript{+}These authors share equal contribution to this work.

% \end{center}
% \clearpage

% % ------- ABSTRACT PAGE ------- %
% \doublespacing
% \section*{Abstract}
% Add abstract here

% \setlength\parindent{0.5in}
% {\em Keywords}: categorization; concepts; generation;
% computational modeling; exemplar models; representativeness
% \clearpage

\doublespacing
\begin{flushleft}
\setlength\parindent{0.5in}

\section{Modeling NotA vs B}

This document is a brief presentation of recent modeling work on distinguishing independently-identified categories from
categories-by-negation. Previous work identified a number of descriptive differences between novel
independently-identified categories (i.e., categories that were generated with a label such as 'Beta') compared to
categories-by-negation (i.e., categories generated with respect to \emph{not} being some other learned category, such as
'not Alpha') \citep{liew2019nota}. Specifically, categories-by-negation were found to be larger and spread more widely
than independently-identified categories.

These results are the first indication that categories-by-negation are distinct from independently identified
categories. These are patterns that no current model of category generation can account for -- to our knowledge every
category generation model assumes that all novel categories are independently identified categories. In order to
appropriately capture the data from categories-by-negation, we require a mechanism that takes into account the regions
of the feature space that is in some way \emph{not} where the learned category members are.

One straightforward way to implement this is by placing a distribution over the learned category, and assuming that
every new exemplar from a category-by-negation is sampled with probability inversely proportional to the distribution
over the learned category. For simplicity, we can assume that the exemplars of the learned category are distributed as a
multivariate Gaussian with a mean and covariance directly estimated from the existing learned category members (i.e.,
with the empirical mean and covariance matrix). Formally, the function describing this negative space can be written as:
\begin{equation}\label{eq:negmech}
n(y|C') = -\sum_{c \in C'} \textrm{MvN}(y; M_c,S_c)
\end{equation}
where $y$ is the candidate novel exemplar, $C'$ is the relevant subset of learned categories (for our experiments here
there is only one -- Alpha), MvN is the probability density function for the multivariate Gaussian, and $M_c$ and $S_c$
are the mean and covariance matrix of each learned category respectively.

To obtain the generation probabilities, we can use the exponentiated Luce choice rule over all possible generation
candidates in the feature space: 
\begin{equation}\label{eq:negluce}
p(y|C') = \dfrac
{\textrm{exp}(\theta \cdot n(y|C'))}
{\sum_i \textrm{exp}(\theta \cdot n(y|C'))}
\end{equation}
where $\theta$ is a response determinism parameter that is freely estimated. A tentative name for this model can be the
Simple Negative-Space model (SNS).

\citet{austerweil2019catgen} tested four category generation models in their performance to a basic category generation
task where participants were told to generate a novel 'Beta' category after having observed members of an 'Alpha'
category. These models, PACKER, copy-and-tweak, \citet{jern2013probabilistic}'s hierarchical Bayesian model, and the
representativeness model, can be easily extended to include a negative-space mechanism by linearly combining the
negative-space function in Equation \ref{eq:negmech} with the density estimates of each model. The specific details of
how this is done within each model are presented in the supplemental document [though in the full draft I'll provide
more detail here...], but essentially, $n(y|C')$ is weighted by a free parameter $\gamma$ and then added to each model's
function prior to their normalization via the exponentiated Luce's choice rule. Each of the non-negative-space (positive
space??) models becomes nested within its negative-space variant, but with $\gamma$ set to 0.

Prior to fitting these models, we might have expected that the negative-space models should better predict the data from
the Not-Alpha generation condition compared to the positive-space models. Conversely, the positive-space models should
be better at predicting independently-identified categories, and should therefore be better fit to categories from the
Beta-only generation condition. However, the resulting fits were less straightforward while being more interesting.

%We refer to this new class of models as negative-space (NS) models, and prepend this description to the relevant model
%when referring to the negative-space variant of the model (e.g., negative-space PACKER is NS-PACKER; negative-space
%copy-and-tweak is NS-CT...ugh cumbersome

Models were fit to each generation condition (i.e., Not-Alpha, Beta-Only, and Beta-and-Gamma) separately and the results
are presented in Table \ref{table:lrtest}. The SNS model was the poorest-performing model overall -- however this is not
surprising because the SNS model does not take into account any within-category structure. The SNS model predicts
exemplar generation completely based on the structure of learned categories without any regard for existing exemplars of
the novel category.

\begin{table}
\begin{center} 
  \caption{Fit results and likelihood ratio tests between positive- and negative-space models. PS-LL refers to
    log-likelihood values for positive-space variants of the models. NS-LL refers to log-likelihood values for
    negative-space variants of the models. }
\label{table:lrtest} 
%\vskip 0.12in
\begin{tabular}{llccccc} 
\hline
Condition & &SNS&  Copy-Tweak & Hier. Bayes & PACKER & Represent. \\
\hline
Not-Alpha & PS-LL    & --     & -4818  & -4916   & -4724 & -4767   \\
          & NS-LL    & -5094  & -4801  & -4667   & -4724 & -4677   \\
          & $\chi^2$ & --     & 34.9** & 497.5** & 0     & 177.3** \\
          &          &        &        &         &       &         \\
Beta-Only & PS-LL    & --     & -4644  & -4680   & -4540 & -4575   \\
          & NS-LL    & -5117  & -4643  & -4579   & -4540 & -4575   \\
          & $\chi^2$ & --     &  2.2   & 202.1** & 0     &   0     \\
          &          &        &        &         &       &         \\
Beta-Gamma& PS-LL    & --     & -4348  & -4359   & -4274 & -4333   \\
          & NS-LL    & -4747  & -4344  & -4335   & -4274 & -4306   \\
          & $\chi^2$ & --     &  6.5*  &  46.8** & 0     &  53.7** \\
\hline
*$p<.05$, **$p<.001$
\end{tabular} 
\end{center} 
\end{table}

The best-performing model appears to be PACKER -- this is interesting for a few reasons. First,
\cite{austerweil2019catgen} found that with the same Alpha conditions (i.e., XOR, cluster, and row), the
representativeness model fit better than PACKER. The relative difference in fit for this analysis might be due to the
slight methodological differences between this study and \cite{austerweil2019catgen}, specifically here we ran twice
the number of generated Betas, we had the XOR condition adjusted to an equally-spaced diagonal structure, the feature
space was increased from a 9-by-9 grid to a 50-by-50 grid, and slight jitter was added to the Alpha exemplars.

Second, PACKER is the best-fitting model while also being the only model that has its positive-space variant performing
just as well as the negative-space variant. For this model, adding the negative-space mechanism did not yield any
benefits in prediction.

The models that appeared to benefit the most with a negative-space mechanism are the non-contrast models (5 out of
the 6 likelihood ratio tests were significant, compared to the 2 out of 6 for the contrast models). This suggests that
the negative-space mechanism is also useful as a contrast mechanism -- although the negative-space variants of the
non-contrast models generally do not do as well as their positive-space variants of their contrast counterparts (a
notable exceptiion is the negative-space hierarchical Bayesian model compared to the positive-space representativeness
model). [This is a point I think will want to emphasise in the full paper -- that the negative-space mechanism might act
like a contrast mechanism but it really isn't. Look at how the representativeness model can still benefit from it, especially
in the Not-Alpha condition where it even outperforms PACKER.]

Our initial hypotheses regarding the performance of the negative-space mechanism is most clearly reflected in the
representativeness model. Here, we see that the negative-space mechanism offers a benefit for the Not-Alpha generation
condition, but not the Beta-Only condition. There is also a benefit for the Beta-Gamma condition, suggesting something
about generating multiple (i.e., more than just one) categories that we are not quite capturing yet.

[From here, one of the main things I'm aiming to complete are the fits to individual data. I think it would be really
interesting to look at the advantage of including the negative-space mechanism at the individual level. The other thing
is to consider what it is about PACKER that makes it so good, but yet not quite good enough to beat negative-space
Representativeness in the Not-Alpha condition. ]

\end{flushleft}
% references section
\clearpage \bibliographystyle{apacite} \bibliography{citations.bib}

\end{document}

### 2022-11-23 Manuscripts/cogsci-journal/revision/supp.tex

--- 2018-09-19 Manuscripts/cog-psych/revision/cogpsych.tex
@@
 %!TEX output_directory = latex_out/
+
 \documentclass[12pt]{article}
 \usepackage[letterpaper, margin=1in, headheight=15pt]{geometry}
-\usepackage{amsmath}
+\usepackage{amsmath,fancyvrb}
+\usepackage[natbibapa]{apacite}
+\usepackage{pgfplots}
 \usepackage{setspace}
-\usepackage{pgfplots}
-\usepackage{fancyhdr}
-\usepackage{csquotes}
-\usepackage{todonotes}
-\usepackage{verbatim}
-\usepackage[toc]{appendix}
-\usepackage[natbibapa]{apacite}
 
@@
 
-% set up notes-- different backgrounds for Nolan and Joe!
-\newcommand\nbcnote[1]{\todo[inline, backgroundcolor = yellow]{\textbf{NBC}: #1}}
-\newcommand\jlanote[1]{\todo[inline, backgroundcolor = lime]{\textbf{JLA}: #1}}
-\newcommand\sxlnote[1]{\todo[inline, backgroundcolor = cyan]{\textbf{SXL}: #1}}
-
-% set up header % 
-\pagestyle{fancy}
-\fancyhf{} % sets both header and footer to nothing
-\renewcommand{\headrulewidth}{0pt}
-\lhead{RUNNING HEAD: Similarity, Contrast, and Representativeness in Categorization}
-\fancyhead[R]{\thepage}
-
 \begin{document}
-
-% ------- TITLE PAGE ------- %
-\begin{center}
-\hfill
-\\[1in]
-
-Creating and Learning Something Different: \\
-Similarity, Contrast, and Representativeness in Categorization
-
-\vfill
-Joseph L. Austerweil\textsuperscript{1}
-Nolan Conaway\textsuperscript{1}, 
-Shi Xian Liew\textsuperscript{1},
-and Kenneth J. Kurtz\textsuperscript{2}
-\\[\baselineskip]
-\textsuperscript{1}Department of Psychology, University of Wisconsin-Madison, Madison, WI, USA
-\textsuperscript{2}Department of Psychology, Binghamton University, Binghamton, NY, USA
-\\[1in]
-
-\vfill
-
-Author Note
-
-Correspondence concerning this article should be addressed to: 
-Joseph L. Austerweil, 1202 West Johnson Street, Madison, WI 53706.
-E-mail: austerweil@wisc.edu
-
-\end{center}
-\clearpage
-
-% ------- ABSTRACT PAGE ------- %
 \doublespacing
-\section*{Abstract}
-The ability to generate new concepts and ideas is among the most fascinating
-aspects of human cognition, but we do not have a strong understanding of the
-cognitive processes and representations underlying concept generation. In this
-paper, we study the generation of new categories using the computational and
-behavioral toolkit of traditional artificial category learning. Previous work in
-this domain has focused on how the statistical structure of known categories
-generalizes to generated categories, overlooking the extent that the contrast
-between the known and generated categories is a factor. We report three
-experiments demonstrating that contrast between what is known and what is
-created is of fundamental importance for categorization. We propose two novel
-approaches to modeling category contrast: one focused on exemplar dissimilarity
-and another on the representativeness heuristic. Our experiments and
-computational analyses demonstrate that both models capture different aspects of
-contrast's role in categorization. Our work also serves as a concrete example of
-how well-established categorization models can be applied beyond category learning.
-
-\setlength\parindent{0.5in}
-{\em Keywords}: categorization; concepts; category learning; generation;
-computational modeling; exemplar models; representativeness
-\clearpage
-
-% % OLD OLD INTRO:
-% Creativity, innovation, imagination, and the creation of new ideas are among the most fascinating, important and difficult human capabilities to study.
-They are fascinating and important capabilities because they have produced some of the greatest scientific breakthroughs that revolutionized the world.
-Unfortunately, technological and scientific breakthroughs are rare and the result of a great deal of cognitive effort, which makes it difficult to investigate the underlying cognitive representations and processes responsible for them using standard experimental methodology or to formalize them in computational models.
-Although we tend to focus on the most salient products of these processes (e.g., scientific breakthroughs), every person is likely to have generated many novel sentences, thoughts, and/or drawings given the combinatorial explosion of possible statements given a reasonably-sized set of primitive elements and combination rules in language, thought, and perception \citep{goldstone2003}.
-This observation provides a fantastic opportunity: By examining how people create new concepts in a domain amenable to formalizing in computational models, we can investigate (some aspects of) creativity, innovation, and imagination scientifically.
-
-% Towards the aim of providing a formal account of the cognitive mechanisms involved in creativity and innovation, we explore a form of creative cognition that can be studied using standard behavioral methodology and is amenable to formal modeling: category generation.
-In a category generation task, participants are given some background knowledge about a domain (a cover story and/or learning one or more categories in the domain) and then are asked to generate another category in the domain.
-Unlike creativity and innovation, categorization is well studied from a breadth of perspectives, including behavioral, neural, comparative, and computational  \citep{kurtz2015human,mack2013,margolis2015,pothoswills2011}.
-By analyzing a creative task using formal tools from the categorization literature, we can formalize some theories of creativity and test them empirically in a more rigorous manner.
-
-%cut from submission intro:
-
-%Yet, it is also one of the most difficult human capabilities to study.
-We tend to focus on the most salient products of conceptual generation (e.g., scientific breakthroughs), but every person is likely to have generated many novel concepts in ordinary life.
-Indeed, the generation of novel sentences, works of art, and so on may also be considered products of our capacity to create new concepts.
-However, while this capacity is of great interest, it is also highly complex, making it difficult to study in a typical laboratory setting.
-By consequence, we do not have a strong understanding of the representations and processes involved in the generation of new concepts.
-
-%Questions concerning the generation of novel concepts are typically addressed within research on creativity, a tremendously diverse field employing a multitude of experimental and theoretical approaches \citep[see][]{kaufman2010handbook}.
-The processes underlying conceptual generation are commonly investigated using the creative cognition approach \citep{finke1992creative, smith1995creative}, in which the products of creative acts (such as drawings) are analyzed to obtain insights into how they might have been created.
-These approaches have made great advances in our understanding of creativity, but the focus on particular types of creative products such as drawings makes it difficult to formalize the proposed processes and representations within a computational modeling framework.
-As a consequence, this approach does not often employ computational models to provide formal tests of competing theories.
-
-% ------- BEGIN! ------- %
-\begin{flushleft}
-
-\section{Introduction}
-\setlength\parindent{0.5in}
-
-Creating and learning new ideas is one of the most fascinating and important
-human capabilities. For example, cell phones and smart phones are newly created
-categories of objects that young and old adults learned and have become reliant
-on in the last few decades. In fact, computational researchers have identified
-generating novel objects and objectives as one of the defining, and most
-difficult to formalize, characteristics of intelligent life
-\citep{lake2017,lehman2011,taylor2016}. Yet, the cognitive mechanisms that
-enable people to innovate and learn new innovations are not well understood and
-are understudied by scientists. In part, this is due to the inherent difficulty
-of designing and conducting experiments that test these capabilities. How would
-a scientist devise an experiment that induces a participant to create and learn
-new categories that are as interesting as cell phones and smart phones? As a
-step towards these goals, in this article we investigate the expectations that
-people have when they learn and generate new categories using computational
-models and behavioral experiments.
-
-Generating and learning new concepts are, however, not altogether different from
-the types of behaviors typically studied in cognitive psychology laboratories.
-In particular, generating a member of a novel class can be considered a `special
-case' application of existing category knowledge
-\citep{kemp2014taxonomy,kurtz2015human}. Research in categorization typically
-focuses on what properties of a category affect human learning of an object's
-category given its features \citep{shepard1961learning,kurtz2013human}, or the
-prediction of an object's unobserved features given some of its other features
-and/or its category \citep{markman2003category}. The generation of members of a
-new category consists of inferring {\em all} features for a {\em novel} category
-label. Thus, we can make progress formalizing the processes involved in category
-generation by extending theories of categorization to this case. Then, to ensure
-that these principles reflect basic principles of categorization and are not
-just specific to category generation, we can examine what factors of the
-generated categories affect the ease at which people learn the category.
-
-Previous work in category learning and generation has established that people
-are highly sensitive to the structural properties of categories, such as
-correlations between the features of category members and the relation between
-items within the same category and those in different categories
-\citep{roschmervis1975,regier2007,shepard1961learning,smith1995creative}.
-Inspired by this work, previous research on the topic of category generation has
-explored a similar principle: People tend to create new categories that have
-similar {\em statistical regularities} as previously learned categories
-\citep{jern2013probabilistic,ward1994structured}. Although this is an important
-characteristic of generating new categories, it cannot be the only one. Taken to
-the extreme, the best ``new'' category in terms of having the same statistical
-regularities to other categories would be identical to a known category that is
-representative of the domain (and thus, not new at all). Contrast from other
-known categories should play a role. Further, scientists in other fields, such
-as marketing \citep{berger16} and sociology \citep{rogers03}, highlight the
-critical importance of contrast in the creation of new ideas.
-
-To successfully generate something novel, what is generated must be different
-from what is already known. This fundamental constraint, ``being different'', or
-contrasting from other categories in the relevant domain, is the focus of our
-work. Intuitively, the influence of contrast may be expected in a number of
-real-life scenarios involving the generation of new categories. For instance, a
-college student who has been eating instant noodles for a week may be more
-inclined to seek out something healthier and less carbohydrate-heavy (e.g., a
-Caesar salad) compared to something similar (a different flavour of instant
-noodles). An actor who is dissatisfied with being typecast as a villain may seek
-out roles as heroes as opposed to playing other antagonistic characters. An
-especially compelling example can be seen in \cite{askin2017} who found that
-new songs that were more different to their peers were generally more popular
-than new songs that were similar to others. It appears that once there is the
-motivation to generate something new, there is an associated motivation to
-generate something \emph{different}.
-
-Although implicitly assumed in some work, this constraint has been
-overlooked in previous research: To our knowledge, there has not been any
-systematic investigation addressing how generated categories {\em differ} from
-what is already known. Although the idea of category contrast is ubiquitous
-throughout the categorization literature, and extends to a variety of other
-fields \citep[e.g., color;][]{regier2007}, the idea that a new category should
-be ``different'' is vague, as there are many ways it could be different from a
-previously observed category.
-
-We examine two different definitions of how categories should differ from one
-another: exemplar dissimilarity and representativeness of the alternative
-category. To formalize the first technique, we build on the largely successful
-exemplar modeling framework that
-\citep{medin1978context,nosofsky1984choice,nosofsky1986attention}. In doing so,
-we propose a novel exemplar model of category generation, {\em Producing Alike
-  and Contrasting Knowledge using Exemplar Representations} (PACKER),
-formalizing how new categories should differ from previous categories. It
-embodies the exemplar dissimilarity principle by incorporating a factor that
-repulses members of opposite categories from one another. To foreshadow, we find
-that this notion of contrast predicts human category learning performance, above
-and beyond other category learning models in a novel behavioral experiment. This
-demonstrates that contrast plays a central role in other important
-categorization tasks, and it is not specific to category generation.
-
-The second hypothesis for how contrast might affect categorization is contrast
-as {\em representativeness}. The representativeness heuristic \citep{kahneman1972subjective}
-states that people make judgments regarding an outcome based on how
-representative it is of the evidence given in the current context. The heuristic
-is a powerful theoretical construct that has been used to capture a range of
-complex patterns of human judgments \citep{kahneman73,tversky74,tversky83},
-especially those that deviate from normative theory (as given by a
-straightforward application of Bayes' rule). For example, the coinflip sequence
-TTHTH is perceived to be more random (due to it being representative of a
-coinflip sequence) than TTTTT (despite both having the same probability of
-having been generated from a fair coin). Although there is a healthy debate as
-to whether perceived randomness {\em really} deviates from normative theory
-\citep{griffiths18,hahn09}, the representativeness heuristic remains one of the
-main explanations of human judgments \citep{reimers18}. To formalize the
-hypothesis that people expect a new category to be representative of the
-opposite of the current category, or contrast as {\em representativeness}, we
-use the Bayesian formulation of representativeness \citep{tenenbaum01}, which
-has been used successfully to capture human performance in color categorization
-\citep{abbott16}, image category learning \citep{abbott11}, and language grammar
-learning \citep{rafferty10}. We do so by extending the state-of-the-art category
-generation model \citep{jern2013probabilistic} and find that human category
-generation is best captured as generating representative, rather than probable,
-samples.
-
-The outline of the article is as follows. First we describe previous
-computational formalizations of theories of category generation and empirical
-work investigating them. We then present two hypotheses for how contrast might
-affect categorization and formalize them in computational models. The first is a
-novel exemplar model, which is designed to generate categories that
-systematically differ from existing categories in the domain. The second assumes
-that the goal of category generation is to create representative samples of the
-opposite of the observed categories. We present two experiments demonstrating
-strong and systematic effects of category contrast on concept generation, and we
-qualitatively and quantitatively analyze the performance of each model in
-capturing human category generation. Afterwards, we find support for category
-contrast predicting the ease at which people learn a pair of categories in a
-novel empirical experiment. We conclude with a discussion of the implications of
-our results for categorization and directions for future
-work.
-
-\section{Prior work}
-
-Much of what we know about concept generation and contrast comes from the
-foundational literature on creative cognition. In a series of reports, Ward and
-colleagues
-\citep{ward1995s,ward1994structured,marsh1999inadvertent,ward2002role,smith1993constraining}
-established that category generation is highly constrained by prior knowledge:
-Generated categories tend to consist of features observed in known categories,
-and they tend to exhibit the distributional properties found in known
-categories. In a seminal study, \cite{ward1994structured} asked participants to
-generate new species of alien animals by drawing and describing members of the
-species. People tended to generate species with the same features as on Earth
-(e.g., eyes, legs, wings), and possessing the same feature correlations as on
-Earth (e.g., feathers co-occur with wings). Likewise, aliens drawn from the same
-species tended to share more features with one another compared to members of
-opposite species.
-
-The broader set of observations made by Ward and colleagues provide a great deal
-of insight into the role of prior knowledge in constraining category generation.
-Much of the work from this area
-\citep[e.g.,][]{smith1993constraining,marsh1999inadvertent} focuses on how
-information provided to participants (such as an example of a species generated
-by other participants) can drastically diminish the difference of a new category
-from pre-existing categories. Theoretical accounts of these effects have
-primarily been grounded within the categorization literature. For example, the
-predominant ``Path of Least Resistance'' account
-\citep[see][]{ward1994structured,ward1995s,ward2002role} proposes that, when
-generating a new species of animal, people retrieve from memory a known
-subcategory of animals (e.g., {\em bird}, {\em dog}, {\em horse}), and simply
-change some of the features to make something new. People are thought to change
-only features that are not characteristic of the retrieved category (e.g., if
-{\em bird} was retrieved, the presence of {\em wings} would not change, but {\em
-  color} might). This theory incorporates elements of the highly influential
-basic-level categories framework \citep{rosch1975cognitive,rosch1976basic}, as
-well as the exemplar view \citep{medin1978context,brooks1978nonanalytic}. While
-this work has been incredibly useful in providing a conceptual sketch of
-generation theories, its qualitative nature and the hand-drawn responses used in
-the experiments paradigms precludes the development of formal approaches that
-can be used to test them in a fine-grained manner.
-
-\cite{jern2013probabilistic} recently showed that concept generation could be
-studied in a more controlled manner through the well-developed methods of an
-artificial categorization paradigm \citep[see][for a review]{kurtz2015human}. In
-Experiments 3 and 4 of their article, participants were exposed to members of
-experimenter-defined categories of "crystals" varying in size, hue, and
-saturation. Following a training phase during which the experimenter-defined
-categories were learned, participants were asked to generate novel categories of
-crystals. In a finding mirroring \cite{ward1994structured},
-\cite{jern2013probabilistic} found that participants generated categories with
-the same distributional properties as the experimenter-defined categories. For
-example, after learning categories with a positive correlation between the size
-and saturation features (larger sized crystals were more saturated),
-participants generated novel categories with the same positive correlation. By
-replicating \cite{ward1994structured}, they demonstrated that category generation
-can be studied in a well-known and highly controlled experimental paradigm.
-
-The authors evaluated the predictions of several formal models on their data.
-Most notably, they showed that a hierarchical Bayesian model provided the
-strongest account of their results. Their model views observed examples as
-samples from an underlying category distribution, describing the location of the
-category in the space, as well as how it varies along each feature. In turn,
-each category is viewed as a sample from an underlying {\em domain}
-distribution, specifying distributional commonalities among the observed
-categories. Generated categories are thought to stem from the same domain
-distribution as observed categories, thus the distributional properties of
-observed categories will be preserved within the generated category.
-
-\cite{jern2013probabilistic} additionally tested a ``copy-and-tweak'' model that
-broadly resembles the earlier ``Path of Least Resistance'' account. The core
-proposal is that participants generate new items by copying stored examples from
-memory and tweaking them to generate something new. The copy-and-tweak model
-differs from the Path of Least Resistance account in that it notably omits the
-hierarchical organization of categories, as well as selectivity to which
-features are changed \citep[both of which are factors in the Path of Least
-Resistance account;][]{ward2002role}. Instead, their copy-and-tweak model
-corresponds to a direct exemplar-similarity approach
-\citep[e.g.,][]{nosofsky1984choice,nosofsky1986attention}, generating new items
-according to their similarity to known members of the target category. The
-copy-and-tweak model provided a poor account of their results, as the
-experiments devised by \citeauthor{jern2013probabilistic} were specifically
-designed to challenge it. Its application is notable as a first step
-toward understanding theories developed in the creative cognition literature
-using well-known formal approaches from the categorization literature.
-
-It is worth pointing out that the early observation of regularities in
-distributional properties across categories is not confined to the work of Ward
-and colleagues. Most notably, \citet{thomas98} trained participants to classify
-circles of different sizes, each with a radial line of varying orientation, into
-two categories. During a surprise prediction phase, participants were asked to
-generate a value for a missing feature for a certain category exemplar given a
-value on the other feature. Their results revealed that exemplars tended to
-share the same feature correlations as exemplars from previously learned
-categories. Interestingly, this effect was not consistently observed when the
-given values of a feature fell outside the range of learned values for that
-particular category. For instance, if the learned Category A exemplars were
-generally 8 cm to 10 cm in radius and negatively correlated with the angle of
-the radial line, participants generally produced radial line orientations that
-were also negatively correlated with the exemplar size when the given size was
-between 8 cm and 10 cm. However, when the given size fell outside that range for
-that category (e.g., 14 cm for a Category A exemplar), some participants
-consistently produced radial line orientations that were not negatively
-correlated with the size and that also placed the exemplars in a significantly
-different location in the feature space. These results suggest that when tasked
-to produce exemplars that fall outside what was previously learned, people show
-some tendency to generate exemplars that are different (both distributionally
-and spatially) to what was learned. We explore this idea of contrast
-more deeply in the following section.
-
-\section{And Now for Something (Completely) Different: The Role of Contrast}
-
-Prior work on category generation has explored only one factor: distributional
-correspondences between learned and generated categories. As a result, most of
-the computational, theoretical, and empirical efforts have been focused on
-explaining those effects. In this paper, we investigate another important
-factor: category contrast. To generate a novel concept, individuals must produce
-something that is in some capacity {\em different} from what they already know.
-As a consequence, we propose that contrast should be a primary constraint that
-underlies people's expectations about how a set of categories are organized:
-Categories should be different from each other.
-
-Although it is evident that people are {\em capable} of creating new concepts
-and categories, it is not entirely clear how new concepts are systematically
-made different from what is already known. The hierarchical Bayesian model
-developed by \cite{jern2013probabilistic} assumes that differences between
-observed and generated categories are only due to random variation. The model
-assumes that generated categories are sampled from the same underlying domain
-distribution as observed categories, and will thus share a common distributional
-structure. The model does not make predictions about the {\em location} of the
-category within the domain (the perceptual instantiation of category members).
-Under a strict interpretation of their model, given knowledge of a single
-category within the domain, the most probable new category to be generated is
-located in {\em exactly} the same location and possesses an identical
-distributional structure. This is not an issue with their model specifically,
-but using a broader class of standard hierarchical Bayesian models without any
-additional features \citep[e.g., ][]{griffithsscn07,kemp07}. Many of these
-models assume that at some point of the latent generative process the same
-underlying distribution generates all of the categories and thus, any
-differences between categories are due to {\em noise} and should not be {\em
-  systematic}. If the goal of creating a new category given others in a domain
-is generating from the posterior predictive distribution then the best a
-standard hierarchical Bayesian model without a notion of contrast built into the
-model can do at capturing contrast is to assume that the new category is placed
-uniformly at random over stimulus space. But, this defeats the purpose of a
-hierarchy as it is ignored when determining a new category location!\footnote{It
-  is plausible that some hierarchical Bayesian models could be created that
-  generates categories different from each other. However, this model would not
-  be a standard application or extension of most pre-existing hierarchical
-  Bayesian models. The generative process would need to include a component that
-  presumes contrast.} Note that this does not mean location information and
-contrast cannot be captured using hierarchical Bayesian models. In fact, a model
-instantiating contrast as representativeness is a hierarchical Bayesian model.
-But, it requires an additional factor, which in our case is having a different
-goal than predicting the next category.
-
-The copy-and-tweak model tested by \cite{jern2013probabilistic} also claims
-little about how generated categories should contrast with what is already
-known. In their simulations, the model was only tested on generation after the
-learner had been exposed to members of the target category, and so the model's
-ability to generate a new category from scratch was not evaluated. However, the
-model's generation is based exclusively on similarity to known members of the
-{\em target} category; when there are no members of the target category,
-generation is presumably random.
-
-\subsection{Contrast as exemplar dissimilarity: The PACKER Model}
-
-\label{section:PACKER-definition}
-
-As noted above, the constraint that new concepts should differ from what is
-already known has been largely overlooked in previous work. This is no doubt in
-part due to the vague definition of what it means for a concept to be
-``different'': A generated category may be different from what is already known
-in any number of respects. Towards providing a more precise definition of the
-role of contrast in generation, we formalized contrast in a novel exemplar
-model, PACKER ({\em Producing Alike and Contrasting Knowledge using Exemplar
-  Representations}). PACKER explains category generation as a balance between
-two fundamental constraints: The category to be generated should not be similar
-to known categories, and exemplars within each category should be similar to one
-another. These ideas are implemented within the well-studied exemplar framework
--- the PACKER model is an extension of the influential Generalized Context Model
-of categorization \citep[GCM;][]{nosofsky1984choice,nosofsky1986attention}.
-
-Although, as an exemplar model, one of PACKER's proposals is people represent
-categories in terms of a collection of stored exemplars, we did not pick it
-assuming it is the correct model of human categorization. The choice to develop
-PACKER within an exemplar framework reflects the facts that exemplar models have
-been throughly evaluated, are strongly theoretically motivated, and dominate
-much of the theoretical and empirical work in categorization. The focus of our
-work with PACKER concerns the dual constraints of within- and between-class
-similarity; it is not difficult to imagine how such constraints may be
-instantiated using alternative frameworks (e.g.,
-\citealp{kurtz2007divergent,love2004sustain,smith2000thirty}; for a review of
-categorization models, see \citealp{pothoswills2011}).
-
-Both PACKER and the GCM simulate categorization under the assumption that
-learners represent categories as a collection of exemplars, corresponding to the
-labeled stimuli they have observed. The exemplars are encoded within a
-$k$-dimensional psychological space, and model performance is based on the
-amount of similarity between the item to be categorized and the stored
-exemplars. Similarity between two examples, $s\left(x_i, x_j\right)$, is
-computed as an inverse exponential function of distance
-\citep[following][]{attneave1950,shepard1957stimulus,shepard1987toward}:
-
-\begin{equation} s\left(x_i,x_j\right) = \exp \left\{ -c \left[\sum_{k}{ w_k
-\left| x_{ik} - x_{jk} \right|^r }\right]^{1/r} \right\}
-\label{eq:similarity}
-\end{equation}
-% 
-where $w_k$ is the attention weighting of dimension $k$ ($w_k \geq 0$ and
-$\sum_k{w_k} = 1$), accounting for the relative importance of each dimension in
-similarity calculations, and $c$ ($c>0$) is a specificity parameter controlling
-the spread of exemplar generalization. For simplicity, attention will be
-distributed uniformly in our simulations (unless otherwise noted). The value of
-$r$ depends on the nature of the experimental conditions being simulated: $r=1$
-is appropriate for separable dimensions, whereas $r=2$ is appropriate for
-integral dimensions \citep[e.g.,][]{shepard1964attention,garner1974processing}.
-In our simulations, we set $r=1$ due to the separable nature of the stimulus
-dimensions used in our experiments (see Figure \ref{fig:e1-conditions}).
-
-PACKER (as well as its name) was in part inspired by earlier work from the
-categorization literature \citep{hidaka2011packing,stewart2005}. \citet{hidaka2011packing} argued that
-natural categories ``pack'' the values of features such that different
-categories fill the domain space with distance between one another, while
-keeping items within the same category close together. Inspired by this idea,
-PACKER proposes that generation is constrained by both similarity to members of
-the target category (the category in which a stimulus is being generated) as
-well as similarity to members of other categories: the most desirable generation
-candidates are similar to members of the target category and not similar to
-members of contrast categories. This is achieved by aggregating similarity
-across known exemplars differently according to class membership. The aggregated
-similarity $a(y,x)$ between generation candidate $y$ and stored exemplars $x$ is
-given by:
-
-\begin{equation} a(y, x) = \sum_j{f(x_j) s(y, x_j)}
-\end{equation}
-% 
-where $f(x_j)$ is a function specifying each exemplar's contribution to
-generation. A negative value for $f(x_j)$ produces a `repelling' effect (items
-are less likely to be generated nearby $x_j$), and a positive value produces an
-`attracting' effect (items are more likely to be generated nearby $x_j$). When
-$f(x_j)=0$, the exemplar does not contribute to generation.
-
-PACKER sets $f(x_j)$ depending on exemplar $x_j$'s category membership:
-$f(x_j) = \theta_t$ if $x_j$ is a member of the target category, and
-$f(x_j) = -\theta_c$ if $x_j$ is a member of a contrast category. $\theta_t$ and
-$\theta_c$ are free parameters ($0 \leq \theta_t,\theta_c$) controlling the
-trade-off between within- and between-category similarity. For example, when
-$\theta_t = \theta_c = 0.5$, $f(x_j) = 0.5$ for members of the target category
-and $f(x_j) = -0.5$ for members of other categories; thus, the model is likely
-to generate items that are similar to members of the target category but are not
-similar to members of other categories. In this way, $\theta_t > 0$ with
-$\theta_c = 0$ produces exclusive consideration of target-category members, and
-$\theta_c > 0$ with $\theta_t = 0$ produces exclusive consideration of
-contrast-category members. The combination of $\theta_t$ and $\theta_c$
-parameters thus specifies a wide breadth of possible approaches; by fitting it
-to a dataset, one can describe the relative roles of between-category contrast
-and within-category similarity in generation. See Figure
-\ref{fig:packer-examples} for an illustration of how these parameters control
-the relative influence of within-category similarity and contrast to other
-categories when generating a new exemplar.
-
-\begin{figure}
-	\begin{center} \inputpgf{figs/}{packer-examples.pgf}
-      \caption{PACKER generation of a category `B' example, following exposure
-        to one member of category `A' and one member of category `B'.
-        Predictions are shown for three different parameterizations (differing
-        in values of $\theta_t$ and $\theta_c$): {\em (a)}
-        Predictions based on contrast similarity only. {\em (b)}
-        Predictions based on target similarity only. {\em (c)} Predictions with
-        both constraints considered.}
-		\label{fig:packer-examples}
-	\end{center}
-\end{figure}
-
-The probability that a candidate $y$ will be generated is evaluated using an
-Exponentiated \citet{luce1977choice} choice rule. Candidates with greater values
-of $a(y,x)$ are more likely to be generated than candidates with smaller values:
-% 
-\begin{equation} p(y \mid x) = \dfrac { \exp \left \{ a \left( y, x
-\right) \right \} } { \sum_i{ \exp \left \{ a \left( y_i, x \right)
-\right\} } }
-\label{eq:packer-choice}
-\end{equation}
-% 
-%where $\theta$ ($\theta \geq 0$) is a free parameter controlling response
-%determinism.
-It is worth noting that PACKER is only one possible exemplar-based account of
-category generation within our proposed framework. That is, PACKER places
-specific constraints on the possible values of $f(x_j)$, but other
-exemplar-based category generation models with drastically different behavior
-can be formalized in this framework by imposing alternative constraints. For
-example, as will be discussed in more detail below, PACKER is formally
-equivalent to the copy-and-tweak model proposed by \cite{jern2013probabilistic}
-when $\theta_c = 0$ and $\theta_t > 0$. Likewise, when
-$\theta_t = 0$ and $\theta_c > 0$, PACKER can represent a
-contrast-only generation mode, relying exclusively on contrast when generating
-new categories. When $f(x_j) < 0$ for all $x_j$ (regardless of class
-membership), a ``pure-packing'' approach is yielded, generating items in
-unoccupied areas of the domain. Thus, the proposed framework may be used to
-describe a wide variety of qualitatively distinct generation strategies.
-
-\subsubsection{Relation Between PACKER and Copy-And-Tweak}
-\label{section:copytweak-packer} The PACKER model is similar to the
-copy-and-tweak model reported by \cite{jern2013probabilistic}: Both models are
-exemplar-based, and both models generate new items according to their similarity
-to known members of the target class. However, PACKER diverges from the
-copy-and-tweak model by including a contrast mechanism, enabling generation
-according to dissimilarity to members of opposing categories. As a consequence,
-copy-and-tweak can be realized as a parameterization of the PACKER model that is
-insensitive to category contrast. Specifically, when $\theta_c = 0$ and
-$\theta_t > 0$ (see Figure \ref{fig:packer-examples}, panel B), $f(x_j) = 0$ for
-$x_j$ belonging to contrast categories; thus, PACKER is not influenced by these
-items, and is mathematically equivalent to a copy-and-tweak approach.
-
-In this paper, we report simulations using this copy-and-tweak model. This model
-fits within the exemplar-based category generation framework defined above, and
-is a continuous-dimension adaptation of the model tested by
-\cite{jern2013probabilistic}. By formalizing a model family where PACKER and
-copy-and-tweak are different parameterizations of models within the same
-framework, the comparison between PACKER and copy-and-tweak provides a test of
-the explanatory value of the contrast mechanism based on exemplar dissimilarity:
-The account provided by copy-and-tweak will only equal that of PACKER if the
-contrast mechanism does not offer an advantage (i.e., if $\theta_c > 0$
-significantly improves model fits). Note that the purpose of the article is to
-explore and formally analyze the role of contrast in categorization and thus, we
-leave extending PACKER to incorporate distributional factors for future work.
-
-\subsection{Contrast as representativeness: An extension to Jern and Kemp
-(2013)}
-\label{section:representativeness-definition}
-
-An alternate conceptualization of category contrast is the idea of
-representativeness -- exemplars are generated such that they resemble the target
-category and thus more distinct from and less similar to other categories. We
-adopt the formalization given by \cite{tenenbaum01}, where the
-representativeness $R(y,h)$ of an item $y$ (or in our case, an exemplar) is the
-relative amount of evidence that is provided by $y$ for a given hypothesis $h$
-in a space of hypotheses $\mathcal{H}$ in contrast to all other hypotheses
-$\mathcal{H}^c=\mathcal{H} \setminus \{h\}$:
-\begin{equation}
-  R(x,h) = \log \dfrac{p(x|h)}{1-p(x|\mathcal{H}^c)} = \log \dfrac{p(x|h)}{\sum_{h'\in H'}{p(x|h')p(h')}} . 
-\label{eq:representativeness}
-\end{equation}
-In Equation \ref{eq:representativeness}, $p(h')$ is the prior probability
-distribution on the hypothesis space that excludes $h$. Note that when there are
-only two hypotheses (e.g. two categories) involved in the domain, this prior
-takes the value of $1$ since there is ever only one alternative hypothesis.
-
-The Bayesian formalization of representativeness makes it straightforward to
-extend the existing hierarchical Bayesian model of category generation
-developed by \cite{jern2013probabilistic}.\footnote{We thank Charles Kemp for the suggestion.} Specifically, both models assume that
-exemplars are sampled from a given category distribution ($h$ in
-Equation \ref{eq:representativeness}). Each category distribution is a
-multivariate normal parameterized by a location vector $\mu_k$ and covariance
-matrix $\Sigma_k$ (i.e., $h = \mathcal{N}(\mu_k,\Sigma_k)$). The parameters
-$\mu_k$ and $\Sigma_k$ are assumed to be samples from a prior
-normal-inverse-Wishart (NIW) distribution parameterized by
-$\mu_0,\Sigma_0,\kappa,\nu$. Mathematically,
-\begin{align}
-  \mu_k, \Sigma_k|\mu_0,\Sigma_0,\kappa,\nu &\sim
-                                              \textrm{NIW}(\mu_0,\Sigma_0,\kappa,\nu)
-                                              \label{eq:mvnsample}\\
-  y|C &\sim \mathcal{N}(\mu_k,\Sigma_k) \label{eq:exemplarsample}
-\end{align}
-
-In our simulations, we set the prior mean $\mu_0$ to the center of the stimulus
-space and the prior variance to be isotropic ($\Sigma_0 = \lambda \mathbf{I}$ where
-$\lambda$ is a free parameter and $\mathbf{I}$ is a $d$-by-$d$ identity matrix
-with $d$ representing the number of dimensions or features in the domain). The
-$\kappa$ and $\nu$ parameters are freely estimated within the constraints
-$\kappa > d-1$ and $\nu > 0$.
-
-With the assumption of the NIW prior, the expected location vector $\mu_k$ given exemplars observed from category $k$ is:
-\begin{equation}
-  \mu_k = \dfrac
-    {\kappa\mu_{0} + n_k \bar{x}_k}
-    {\kappa + n_k}
-    \label{eq:category_mus}
-\end{equation}
-where $n_k$ is the number of observed exemplars in category $k$ and $\bar{x}_k$
-is the observed category mean. Note that if there are no observed exemplars in
-the target category (i.e., if the model is generating a completely novel
-category), Equation \ref{eq:category_mus} simplifies to $\mu_k = \mu_0$.
-
-The NIW prior also allows us to infer the category covariance matrix $\Sigma_k$
-by computing the following:
-\begin{equation}
-  \Sigma_k = [\nu \Sigma_D+ C_k +
-    \dfrac
-    {\kappa n_k}
-    {\kappa + n_k}
-    (\bar{x_k}-\mu_k)(\bar{x_k}-\mu_k)^T
-  ] (\nu + n_k)^{-1}
-  \label{eq:Sigma_y}
-\end{equation}
-where $C_k$ is the observed category covariance and $\Sigma_D$ is the domain
-covariance matrix from which $\Sigma_k$ samples are obtained. We can
-infer $\Sigma_D$ from the observed category covariances $C$ and the prior
-covariance $\Sigma_0$. Specifically,
-\begin{equation}
-    \Sigma_D = \Sigma_0 + \sum_{k}{C_k}
-\end{equation}
-
-At this point, both the representativeness model and the hierarchical Bayesian
-model from \cite{jern2013probabilistic} are largely identical. However, both
-models diverge in their computation of exemplar generation probabilities. While
-the original hierarchical Bayesian model produces novel exemplars with
-probabilities proportional to the multivariate normal likelihood $p(y|h)$, the
-representativeness model generates new exemplars according to their
-representativeness $R(y,h)$. In practice, the probability of generating a
-particular candidate $y$ is obtained using an Exponentiated
-\citet{luce1977choice} choice rule:
-\begin{equation}
-  p(y) = \dfrac
-    {\exp( \theta \cdot R(y,h))}
-    {\sum_k \exp(\theta  \cdot R(y_k,h_k))} .
-\label{eq:representative-choice}
-\end{equation}
-where $\theta$ is a freely estimated response determinism parameter (constrained
-such that $\theta \geq 0$).
-
-Despite both the original hierarchical Bayesian model and our representativeness
-model sharing identical hierarchical structures, their distinct response
-processes can yield very different exemplars. Specifically, the hierarchical
-Bayesian model emphasizes the generation of exemplars that maintain
-distributional commonalities across categories. In contrast, the
-representativeness model focuses on the generation of exemplars that are
-representative of the underlying distribution for a target category. Generally,
-this mechanism of representativeness results in the generation of exemplars that
-are less similar to exemplars from other categories. We see this illustrated in
-Figure \ref{fig:representative-examples}, where the representativeness model
-displays a strong preference for generating exemplars from target category 'B'
-that are further away from the contrast category 'A'.
-
-\begin{figure}
-	\begin{center} \inputpgf{figs/}{hbm-examples.pgf}
-		\caption{Generation of a category `B' example, following exposure
-to one member of category `A' and one member of category `B'. Predictions are
-shown for (a) the original hierarchical Bayesian model and (b) the
-representativeness model.}
-		\label{fig:representative-examples}
-	\end{center}
-\end{figure}
-
-\subsection{Synopsis and Prognosis}
-
-Research on categorization and category generation has focused on the finding
-that generated categories tend to possess distributional commonalities with
-known categories. However, a fundamental task of categorization is to create
-something {\em new} (i.e., different from what is already known). The manner in
-which generated categories differ from known ones is, nonetheless, poorly
-understood: Existing theories do not make strong predictions about how generated
-concepts should systematically differ from existing ones. Above, we introduced
-two novel approaches to formalizing the roles of similarity and contrast in
-category generation.
-
-In the sections below, we present two experiments demonstrating systematic
-effects of category contrast on category generation inspired by factors
-influencing how the contrast models generate new categories. Our experiments are
-based on \cite{jern2013probabilistic}'s paradigm, which is a straightforward
-translation of the traditional artificial classification paradigm to the task of
-category generation: Participants are first exposed to a single,
-experimenter-defined category, and are then asked to generate members of a new
-category. We then report quantitative analyses of our results comparing the exemplar dissimilarity
-and the representativeness accounts of our results to that of the hierarchical
-Bayesian and copy-and-tweak models developed by \cite{jern2013probabilistic}.
-After analyzing the category generation studies, we present third experiment to
-demonstrate the effect of contrast in category learning. This ensures that our
-results are not idiosyncratic to category generation.
-
-\section{Experiment 1}
-To begin our investigation, we examined category generation in a well-understood domain using a few disparate category types. We used an
-artificial stimulus design: A two dimensional domain of squares, varying in
-color and size (see Figure \ref{fig:e1-conditions}a). These dimensions
-have been used in numerous classification learning studies
-\cite[e.g.,][]{conaway2016similar,conaway2016generalization,shepard1961learning,nosofsky1994comparing}.
-Unlike those used in the \cite{jern2013probabilistic} experiments, distance on
-these physical dimensions aligns more directly with perceptual similarity,
-allowing us to evaluate the role of  contrast in categorization
-more precisely.
-It also enables more straightforward comparisons to prior work.
-We tested the effects of category contrast after learning one category from a set of qualitatively distinct category structures, as shown in Figure \ref{fig:e1-conditions}.
-
-Figures \ref{fig:e1-conditions}b-d show the values of exemplar dimensions
-belonging to the experimenter-defined categories (`A', or `Alpha') that
-participants were assigned to learn about prior to generating a new category.
-Each participant learned one of the category types during training. In the
-`Cluster' type, category A is a tight cluster of examples in the space.
-Perceptually instantiated, the members of category A might, for example, be
-large and dark in color. In the `Row' type, category A has a row pattern across
-the space, varying along one feature but not the other. Thus, its members might
-all be dark in color but would vary in size. Finally, in the `XOR' type, the
-experimenter-defined category consists of two clusters separated in opposite
-corners of the space, conforming to the exclusive-or logical structure (e.g.,
-members are small and dark or large and light).
-
-It should be noted that in our experiments the assignment of the perceptual
-to conceptual dimensions (e.g., $X \rightarrow Size$, $Y \rightarrow Color$) and the direction of variation along each dimension (e.g., $dark
-\rightarrow light$ or $light \rightarrow dark$) were counterbalanced across
-participants. The category types in Figure \ref{fig:e1-conditions} are plotted
-in conceptual space, rather than perceptual space. Thus, while the
-conceptual organization of the category types remains constant, each category
-type may have a different physical instantiation according to the counterbalance
-assignment. For example, the Cluster type may be large and dark in color, or it
-may be small and light in color, depending on the assignment and direction of
-the dimensions. For this reason, below we will discuss generation within a
-conceptual space, rather than a physically instantiated one.
-
-\begin{figure}
-    \begin{center} \inputpgf{figs/}{e1-conditions.pgf}
-    \caption{Stimulus domain and category types tested in Experiment 1. Stimuli
-are not drawn to scale. Dimension and direction assignment (e.g., large to small or small to large) for color and size were counterbalanced over participants.}
-    \label{fig:e1-conditions}
-    \end{center}
-\end{figure}
-
-After learning about an experimenter-defined category, participants are asked to
-generate examples of a new category. Within this paradigm, an effect of category
-contrast would be realized if participants prefer to generate items in locations
-that are distant (i.e., perceptually dissimilar) from members of category A.
-However, generation is left unconstrained. Critically, participants were not
-asked to generate something different in the prompt. For example, participants
-assigned to the Cluster condition may generate a tightly clustered category in
-the corner opposite of the experimenter-defined category. Alternatively, they
-may generate a tightly clustered category directly overlapping with the
-experimenter-defined category. Further, they may even generate an entirely
-different type of category (e.g., a row category).
-
-Our experimental results also provide a converging test
-that generated categories tend to share distributional properties with known
-categories in the domain \citep{jern2013probabilistic,ward1994structured}. From
-these results, we can predict that, in each condition, participants should
-generate categories that are distributionally similar to the
-experimenter-defined category: In the Cluster condition, generated categories
-should be tightly clustered. In the Row condition, generated categories should
-vary more along the X-axis than the Y-axis. In XOR condition, generated
-categories should be widely distributed across both dimensions, and the two
-dimensions should be positively correlated.
-
-Interestingly, the XOR condition also offers a dissociation between the roles of
-category contrast and the emulation of distributional structure:
-widely-distributed, positively-correlated categories would need to lie along the
-positive diagonal of the space (that is the only place they ``fit''), which is
-already occupied by the experimenter-defined category. Thus, if contrast plays a
-role, exemplars in the generated categories of participants in the XOR condition
-may not be positively correlated -- they may be negatively correlated instead.
-In this case, contrast and statistical regularities would interact, which would
-be inconsistent with prevailing theories of category generation
-\citep{jern2013probabilistic}.
-
-\subsection{Participants and Materials}
-
-183 participants were recruited from Amazon Mechanical Turk. Each participant
-was randomly assigned to one condition: 64 participants were assigned to the
-Cluster condition, 61 were assigned to the Row condition, and 58 were assigned
-to the XOR condition (sample sizes differ due to random assignment). Stimuli
-were squares varying in color (grayscale 9.8\%--90.2\%) and side length
-(3.0--5.8cm), see Figure \ref{fig:e1-conditions}. The assignment of perceptual
-features (color, size) to axes of the conceptual space (x, y)  and the
-direction of variation along each axis (e.g., $dark \rightarrow light$ or $light
-\rightarrow dark$) were counterbalanced across participants.
-
-\subsection{Procedure}
-
-As noted in the introduction of this paper, the task of generating members of a
-new category is well situated as a task in the categorization
-literature: Whereas classification consists of predicting an object's category
-label on the basis of its features, inference consists of predicting an observed
-feature, given a set of observed features and a category label. Generation thus
-consists of predicting {\em all} features of an object, given a novel category
-label. We designed our generation task as an
-extension of the traditional artificial classification learning paradigm. The
-task differs from traditional work in creative cognition primarily through the
-use of an artificial domain, which enables the application of computational
-models. The use of an artificial domain also requires the addition of a training
-phase, during which participants learn about the categories in the domain. As a
-result, unlike most previous studies \citep[e.g.,][]{ward1994structured},
-participants in our studies have no experience with the domain before the start
-of the experiment, and the experimenter-defined categories are not
-hierarchically structured (as are many natural categories).
-
-Participants began the experiment with a short training phase (3 blocks of 4
-trials), where they observed exemplars belonging to the `Alpha' category.
-Participants were instructed to learn as much as they can about the `Alpha'
-category, and that they would answer a series of test questions afterwards. On
-each trial, a single `Alpha' category exemplar was presented, and participants
-were given as much time as they desired to observe it before moving on to the
-next trial. Each block consisted of a single presentation of each of the members
-of the `Alpha' category, in a random order. Participants were shown the range of
-possible colors and sizes prior to training.
-
-Following the training phase, participants were asked to generate four examples
-belonging to another category called `Beta'. As in
-\citet{jern2013probabilistic}, generation was completed using a sliding-scale
-interface. Two scales controlled the values of the two dimensions (color, size)
-for the generated example. An on-screen preview of the example updated whenever
-one of the features was changed. Participants could generate any example along
-an evenly-spaced 9x9 grid (including members of the `Alpha' category), except for
-any previously generated `Beta' exemplars. Neither the members of the `Alpha'
-category nor the previously generated `Beta' examples were visible during
-generation. Prior to beginning the generation phase, participants read the
-following instructions:
-
-\begin{displayquote} As it turns out, there is another category of geometric
-figures called ``Beta''. Instead of showing you examples of the Beta category,
-we would like to know what you think is likely to be in the Beta category.
-
-You will now be given the chance to create examples of any size or color in
-order to show what you expect about the Beta category. You will be asked to
-produce 4 Beta examples - they can be quite similar or quite different to each
-other, depending on what you think makes the most sense for the category.
-
-Each example needs to be unique, but the computer will let you know if you
-accidentally create a repeat.
-\end{displayquote}
-
-\subsection{Results}
-
-We observed a substantial degree of individual differences in our data. In
-Figure \ref{fig:e1-samples} we have plotted sample data from several
-participants, from which it is evident that different participants generated
-qualitatively different category structures. In this section we will focus on
-analyzing the data in aggregate, but in later sections we will explore how these
-individual differences can be explained.
-
-\begin{figure}
-    \begin{center} \inputpgf{figs/}{e1-samples.pgf}
-    \caption{Sample categories generated by participants in Experiment 1.
-Representative samples from common generation profiles are shown.}
-    \label{fig:e1-samples}
-    \end{center}
-\end{figure}
-
-To evaluate the role of contrast, we computed the number of times each stimulus
-was generated, as a function of its average city-block distance from members of
-the experimenter-defined ``Alpha'' category. These data, shown in Figure
-\ref{fig:e1-distanceplots}, reveal a clear pattern: Examples that are more
-distant from members of the experimenter-defined categories are more likely to
-be generated into a new category. This supports the notion that contrast is a
-fundamental constraint on how categories are related to one another and that statistical
-regularity alone is insufficient.
-
-Figure \ref{fig:e1-distanceplots} also depicts, for each participant, the
-average distance of members within the generated category ({\em within-category}
-distance) against the average distance between members of the generated and
-experimenter-defined category ({\em between-category} distance). The narrow
-distribution of between-category distances in the XOR condition reflects the
-widely distributed nature of the experimenter-defined category, reducing the
-possible distances to members of the participant-generated category. These data
-reveal a systematic pattern: The majority of participants generated categories
-with greater between-category distance than within-category distance. That is,
-members of the generated category tended to be more similar to one another than
-to members of the experimenter-defined category. To evaluate this claim
-quantitatively, we conducted t-tests comparing the amount of within- and
-between- class distance in each condition. All conditions possessed greater
-between-category distance: Cluster, $t(63) = 11.43$, $p < .001$; Row, $t(60) =
-13.16$, $p < .001$; and XOR, $t(57) = 3.64$, $p < .001$. These results provide
-further evidence of an effect of category contrast: Participants prefer to
-generate categories that are dissimilar to the learned category but maintain
-some level of internal cohesion.
-
-\begin{figure}
-    \begin{center} \inputpgf{figs/}{e1-distanceplots.pgf}
-    \caption{Experiment 1 results. {\em Left}: Frequency of generation as a
-function of distance from members of the experimenter-defined category. {\em
-Right}: Scatter plot of within-category versus between-category distance in each
-of the participant-generated categories.}
-    \label{fig:e1-distanceplots}
-    \end{center}
-\end{figure}
-
-A secondary goal of this experiment was to examine whether we replicate the
-classic result that generated categories often possess the same distributional
-properties as previously-known categories.
-Given the increased emphasis of replication within psychology \citep{zwaan2018making}, it is important as it serves as a conceptual replication of \citet{jern2013probabilistic}.
-For each generated category, we
-computed the category range along each axis (X, Y), as well as the correlation
-between features. These data, shown in Figure \ref{fig:e1-statsboxes}, reveal
-broad individual differences: Within each condition, participants generated
-categories spanning the entire X- and Y- axis as well as categories that spanned
-very little along each. Likewise, in each condition participants generated
-categories possessing strongly positive, neutral, and strongly negative
-correlations between the dimensions. Comparing the distributional statistics
-between conditions yields a broad yet, as we will see, misleading replication of
-the classic effect.
-
-With respect to ranges along each axis (X, Y), the generated categories from
-each condition tend to reflect the ranges of the experimenter-defined
-categories. The categories generated in the Cluster condition were less widely
-distributed along the X-axis compared to Row, $t(123) = 5.61$, $p < .001$, and
-XOR, $t(120) = 2.68$, $p < .01$. Categories generated in the XOR condition
-were also less widely distributed along the X-axis compared to Row, $t(117) =
-2.56$, $p = .046$. This latter effect was not expected because the
-experimenter-defined categories for XOR and Row had similar X-ranges. However,
-the key finding is that categories from the Cluster condition tended to be more
-tightly distributed along the X-axis.
-
-\begin{figure}
-    \begin{center} \inputpgf{figs/}{e1-statsboxes.pgf}
-    \caption{Box-plots of the distributional statistics from the categories
-generated in Experiment 1. Boxes depict the median and quartiles of each
-condition, with whiskers placed at 1.5 IQR. All points outside this region are
-marked individually.}
-    \label{fig:e1-statsboxes}
-    \end{center}
-\end{figure}
-
-Likewise, categories generated in the Row condition had less Y-axis range
-compared to Cluster, $t(123) = 4.57$, $p < .001$ and XOR, $t(117) = 9.26$, $p <
-.001$, and categories from the Cluster condition had less Y-axis range compared
-to XOR, $t(120) = 3.95$, $p < .001$. As expected, the correlations in the
-Cluster and Row conditions were not systematically positive or negative ($ps >
-.1$). However, the generated categories in the XOR condition tended to possess
-{\em negatively} correlated dimensions, $t(57) = 2.04$, $p = .046$. This
-finding is notable, as it is the opposite of what would be expected based on previous literature \citep{jern2013probabilistic}, assuming
-learners are emulating the distributional structure of the experimenter-defined
-class (which possesses perfectly positively correlated features).
-
-We believe that the failure to replicate dimension correlation emulation is
-because participants in \cite{jern2013probabilistic} could differentiate the
-generated category on a third dimension (hue) to maintain the statistical
-regularities on the other two dimensions. In addition, in our XOR condition the
-stimuli were constrained to the corners of the feature space, whereas in the
-positive diagonal condition of \cite{jern2013probabilistic} the stimuli had no
-such constraint. Although the correlation in the XOR condition is significantly
-negative, it is clear from the box-plot in Figure~\ref{fig:e1-statsboxes} that
-it would be inappropriate to make a strong conclusion (e.g., the median is close
-to zero). However, we can conclude with confidence that there are situations
-where people do not emulate the distributional structure of the given category.
-This indicates that there is more to category generation than the emulation of
-distributional structure of other categories in the domain. Further, as we will
-discuss in more detail in the model-based analysis section, this is expected by
-our proposal that contrast is a fundamental principle in categorization.
-
-\subsection{Discussion} In Experiment 1 we sought to extend our analysis of the
-\cite{jern2013probabilistic} data by evaluating the influence of category
-contrast on category generation, given qualitatively different types of categories. We found strong evidence for effects of category contrast in each
-condition: Participants were more likely to generate stimuli that are more
-distant from (i.e., less similar to) members of a previously-learned category,
-and members of participant-generated categories tended to be more similar to one
-another than to members of previously-learned categories. We also partially
-replicated the classic finding that the distributional structure of generated
-categories reflects that of previously learned categories
-\citep{jern2013probabilistic,ward1994structured}: Members of generated
-categories were more widely distributed along dimensions which were widely
-distributed in the experimenter-defined category.
-
-Notably, however, we also found that participants who learned an XOR category
-(composed of exemplars following a positive diagonal, see Figure
-\ref{fig:e1-conditions}) tended to generate items according to a {\em negative}
-feature correlation -- the opposite of what was present in the previously
-learned category. While this may be difficult to account for under existing
-theoretical approaches (which assume generated categories follow the same
-distributional structure as known categories), it can be concisely explained
-from a category contrast perspective. Specifically, within the XOR condition,
-individuals who seek to generate a category that is perceptually distinct from
-what is already known are left with only the upper-left and bottom-right
-quadrants of the space, as members of the previously-learned XOR category lie in
-the bottom-left and top-right. If examples are generated into both of the
-available quadrants, the generated category will possess a strongly negative
-correlation, opposing that of the experimenter-defined class.
-
-Thus, the core results of Experiment 1 indicate that generated categories can
-systematically differ from what we would expect based on prior work. The
-negative (or null) correlations observed in the XOR condition suggests an
-interesting interaction between contrast with a given category and emulation of
-statistical properties. That is, the constraints on generation imposed by
-category contrast may not simply influence the {\em location} of generated
-categories, but also their distributional structure. In Experiment 2, we test
-this claim more systematically.
-
-\section{Experiment 2}
-
-\begin{figure}
-    \begin{center} \input{figs/e2-conditions.pgf}
-    \caption{Category types tested in Experiment 2.}
-    \label{fig:e2-conditions}
-    \end{center}
-\end{figure}
-
-To test whether category contrast influences the distributional structure of
-generated categories, we sought to identify conditions in which differences in
-the distributional structure of generated categories cannot be explained by the
-distributional structure of the experimenter-defined category. We created two
-new category types (depicted in Figure \ref{fig:e2-conditions}) that possess
-identical distributional structures (both are tight clusters of examples with no
-correlation between features), as they only differ in their Y-axis position: the
-`Bottom' category lies near the bottom of the space, and the `Middle' category
-lies in the center. The distributional equality of these conditions is key to
-the design of the experiment: If the distributional structure of previously
-learned categories were the only influence on the generated categories, we
-should observe no difference in the categories participants generate between
-these two conditions. Will participants distribute their generated category
-differently between conditions due to the differences in the available empty
-stimulus space for generating a new category?
-
-If category contrast influences the distributional structure of the categories
-people generate, then we should observe different types of categories according
-to the shape of the space that is {\em unoccupied} by members of previously
-learned categories. The difference in the Y-axis position between the Bottom and
-Middle conditions produces a considerable change to the shape of the unoccupied
-space. Participants assigned to learn the Bottom category should be less likely
-to generate exemplars into the lower regions of the stimulus space (as these
-areas possess greater similarity to members of the Bottom category), preferring
-instead to distribute exemplars across the upper region of the space. This
-constraint is lifted in the Middle condition, as the Middle category exemplars
-are equidistant to the upper and lower regions of the space. Accordingly,
-participants should be more likely to utilize both of these areas. Thus, if
-category contrast influences the distributional structure of generated
-categories, we should observe more participants in the Middle condition that
-generate examples above {\em and} below the experimenter-defined category.
-
-\subsection{Participants, Materials, and Procedure}
-
-122 participants were recruited from Amazon Mechanical Turk. 61 participants
-were randomly assigned to the Middle and Bottom conditions each. The stimulus
-space and procedure were exactly as in Experiment 1. Participants first
-completed a short training phase, followed by the generation phase. The only
-difference from Experiment 1 was the category types given to participants.
-
-\subsection{Results}
-
-\begin{figure}
-    \begin{center} \inputpgf{figs/}{e2-samples.pgf}
-    \caption{Sample categories generated in Experiment 2. }
-    \label{fig:e2-samples}
-    \end{center}
-\end{figure}
-
-As in Experiment 1, we observed broad differences in the generation approach
-taken by different participants. To characterize the nature of these
-differences, Figure \ref{fig:e2-samples} depicts sample categories generated by
-participants. The data from each condition are organized into four columns based
-on commonly observed patterns of generation: a `Cluster' type of
-tightly-clustered examples, `Row' and `Column' types of exemplars widely
-distributed along the one axis but narrowly along the other, and a `Corners'
-type, wherein participants placed exemplars in disparate corners of the space.
-As before, in this section we focus on analyzing the data in aggregate, but in
-later sections we will focus more specifically on explaining the individual
-differences.
-
-\begin{figure}
-    \begin{center} \inputpgf{figs/}{e2-distanceplots.pgf}
-    \caption{Experiment 2 results. {\em Left}: Frequency of generation as a
-function of distance from members of the experimenter-defined category. {\em
-Right}: Scatter plot of within-category versus between-category distance in each
-of the participant-generated categories.}
-    \label{fig:e2-distanceplots}
-    \end{center}
-\end{figure}
-
-We began our analysis by testing for the broad influence of category contrast on
-generation. As in Experiment 1, we computed the frequency each stimulus was
-generated as a function of its average distance from members of the
-experimenter-defined category, as well as each participant's average within- and
-between- category distance. These data, shown in Figure
-\ref{fig:e2-distanceplots}, yield very similar results. Participants generated
-stimuli that are distant from members of the experimenter-defined category, and
-the categories in each condition tended to possess more between-category than
-within-category distance: Bottom, $t(60) = 5.5$, $p < .001$; Middle, $t(60) =
-2.71$, $p < .01$. We did, however, observe a notable subgroup of participants
-in each condition who generated categories with more within-category than
-between-category distance. Upon manual inspection, many of these individuals
-appear to have assumed a `Corners' strategy, placing exemplars in disparate
-corners of the space, thus producing much more within-category distance, see
-Figure \ref{fig:e2-samples} for examples.
-
-To explore the distributional structure of the generated categories, we computed
-the range of exemplars along each axis (X, Y), as well as the correlation
-between features. These data, shown in Figure \ref{fig:e2-statsboxes}, again
-demonstrate the degree of individual differences observed in our study. In each
-condition, we observed tightly clustered and widely distributed categories along
-each dimension. Although most participants generated uncorrelated categories in
-both conditions, many still produced positively and negatively correlated
-categories.
-
-\begin{figure}
-    \begin{center} \inputpgf{figs/}{e2-statsboxes.pgf}
-    \caption{Box-plots of the distributional statistics from the categories
-generated in Experiment 2. Boxes depict the median and quartiles of each
-condition, with whiskers placed at 1.5 IQR. All points outside this region are
-marked individually.}
-    \label{fig:e2-statsboxes}
-    \end{center}
-\end{figure}
-
-As noted above, if the distributional structure of generated categories is
-influenced by the shape of the space not occupied by members of known
-categories, then participants in the Middle condition would be more likely to
-place exemplars in the upper {\em and} lower regions of the space, as members of
-the experimenter-defined category are equidistant from these regions.
-Participants in the Bottom condition should be less likely to generate category
-members in the bottom regions because members of the experimenter-defined
-category are located there. One way to test these predictions is to analyze the
-Y-axis ranges of the generated categories: If Middle participants utilize the
-upper and lower regions of the space, their categories should vary more along
-the Y-axis. T-Tests comparing the conditions on the distributional statistics,
-however, reveal few between-group differences: the conditions do not differ with
-respect to X-axis range, Y-axis range, or feature correlations ($ps>0.17$).
-
-However, our ability to detect differences in Y-Axis range using a standard
-$t$-test between the conditions is, in this case, diminished due to the
-non-normality of the data (Shapiro-Wilk normality test $W=0.77, p < .001$ for
-the Middle condition and $W=0.85, p < .001$ for the Bottom condition). Figure
-\ref{fig:e2-yranges} depicts the Y-axis position of the exemplars generated
-within each participant's category.
-Further, the median produced Y-Axis range is much smaller in the Bottom than Middle condition, whereas they are essentially identical in X-Axis range.
-The categories are sorted by overall range,
-then by condition assignment. These data reveal that there were nearly as many
-participants who generated categories spanning the entire Y-axis as those who
-generated categories spanning almost none of the Y-axis. The non-normality of
-the Y-axis range distributions thus requires that we use a different approach to
-addressing the experiment's main question.
-
-\begin{figure}
-    \begin{center} \inputpgf{figs/}{e2-yranges.pgf}
-    \caption{Y-Axis range and position of the participant-generated categories
-from Experiment 2. Each line corresponds to a participant's category, with
-notches corresponding to the Y-axis position of exemplars within the category
-(notches may overlap). Participants are sorted by overall range, and then by
-condition. }
-    \label{fig:e2-yranges}
-    \end{center}
-\end{figure}
-
-Because our main prediction concerns the generation of exemplars within the
-upper and lower regions of the domain, we compared the conditions in terms of
-the frequency with which participants generated examples above and below the
-categories. Specifically, we counted the number of participants in each
-condition who placed at least one `Beta' exemplar on the top and bottom `rows'
-of the space (the maximum and minimum possible Y-axis value, respectively). The
-resulting contingencies data are shown in Table \ref{table:e2-subset-table}.
-
-Firstly, it should be noted that nearly every participant utilized the top
-and/or bottom rows: only $10 / 122$ participants generated their category
-entirely within the interior region. Fisher's Exact Tests comparing the
-conditions reveal that more Middle participants generated an exemplar in the
-bottom row, $p < .001$, again demonstrating the role of contrast in guiding
-where exemplars are generated. The conditions did not differ in use of the top
-of the space, $p = .16$, however, more Middle participants placed exemplars in
-the top {\em and} bottom rows, $p = .04$. The latter effect is of particular interest
-here, as it indicates that the shape of the unoccupied space exerts some
-influence on the distributional structure of generated categories: Participants
-in the Middle condition were more likely to generate a category spanning the
-entire Y-axis. Thus, the distributional structure of the generated categories
-can be influenced without any change to the distributional structure of the
-given category. Rather, it can be affected by category contrast alone.
-
-\begin{table}
-\begin{center}
-\caption{Experiment 2 results.}
-\label{table:e2-subset-table} \vskip 0.12in
-\begin{tabular}{ l r r} \textbf{Middle} & Used top row & No top row \\ \hline
-Used bottom row & 28 & 18 \\ No bottom row & 11 & 4 \\ \\ \textbf{Bottom} & Used
-top row & No top row \\ \hline Used bottom row & 16 & 8 \\ No bottom row & 31 &
-6 \\
-\end{tabular}
-\end{center}
-\end{table}
-
-\subsection{Discussion}
-
-In Experiment 2, we replicated the core findings from Experiment 1. Stimuli are
-more likely to be generated if they are distant from exemplars in other
-categories, and most participants generate categories with more between-category
-than within-category distance. However, we additionally found that the {\em
-position} of a previously learned category (rather than its distributional
-structure) influences the types of categories people generate: Participants who
-learned the `Middle' type were more likely to generate categories spanning the
-entire Y-axis of the space. Participants who learned the `Bottom' type were less
-likely to do so as a result of the presence of opposite category exemplars in
-the lower regions of the space.
-
-This finding cannot be explained from the perspective that the distributional
-structure of previously learned categories is the sole determinant of the
-distributional structure of generated categories. However, the observed behavior
-is expected from a category contrast perspective: Participants seeking to
-generate a perceptually distinct category will be more likely to use areas of
-space that are unoccupied by exemplars belonging to previously learned
-categories. In the Middle condition, the upper and lower regions of space are
-equidistant from members of the experimenter-defined category, whereas in the
-Bottom condition, the lower region of the space is closer to members of the
-experimenter-defined category. Thus, while Middle participants may form
-categories around the use of the equally unoccupied areas, the same is not true
-for the Bottom condition.
-
-\section{Model-based Analyses of Experiments 1 and 2}
-
-Experiments 1 and 2 revealed systematic and strong effects of category contrast
-on category generation. In this section, we analyze the performance of the
-different formal models at explaining the experimental results. Specifically, we
-present simulations from the two novel contrast models: PACKER and the
-representativeness model, and compare them to two models that do not incorporate
-contrast: the copy-and-tweak model (discussed in Section
-\ref{section:copytweak-packer}) and an implementation of the hierarchical
-Bayesian model proposed by \cite{jern2013probabilistic}, described in-depth in
-Appendix \ref{ap:hsampling-definition}. For our simulations here, the
-copy-and-tweak model is defined as a variant of PACKER with the $\theta_c$
-parameter constrained to be zero. The comparison of this set of models serves to
-highlight the explanatory role of contrast in categorization: If contrast
-affords little explanatory advantage, then there should be few differences in
-performance between PACKER and copy-and-tweak, or between the representativeness
-and hierarchical Bayesian model. The comparison between these models can also
-emphasize the necessity of contrast and demonstrate that generation cannot be
-explained entirely through the emulation of distributional structure. Each model
-has complementary strengths and weaknesses: Whereas PACKER and copy-and-tweak
-are relatively insensitive to the distributional structure of learned categories
-(relying only on exemplar similarities), the representativeness and hierarchical
-Bayesian model generates categories exclusively on the basis of knowledge of how
-existing classes are distributed.
-
-Our approach in this section is to first broadly evaluate and compare the
-quality of each model's account to our entire dataset (Experiments 1 and 2
-combined), then analyze the ability for each model to explain individual
-differences in each experiment, and lastly we describe the strengths and
-weakness of each model's account of category generation.
-
-\subsection{Parameter-Fitting}
-
-To obtain a global measure of the quality of each model's account, we fitted the
-parameters of each model to our entire dataset (Experiments 1 and 2 combined),
-using a hill-climbing algorithm which maximized the log-likelihood of the
-model's predictions of the observed responses (1220 responses from 305 total
-participants). We fitted three parameters in the PACKER model ($c$, $\theta_t$,
-and $\theta_c$; see Section \ref{section:PACKER-definition}), as well as four in
-the representativeness model and the hierarchical Bayesian model ($\kappa$,
-$\lambda$, $\nu$, and $\theta$; see Section
-\ref{section:representativeness-definition} and Appendix
-\ref{ap:hsampling-definition} respectively). We fitted only two parameters for
-the copy-and-tweak model ($c$, and $\theta_t$), as $\theta_c$ is held constant
-($\theta_c = 0$). Attention ($w$, see Equation \ref{eq:similarity}) in PACKER
-and copy-and-tweak was set uniformly. Parameters were not allowed to vary
-between participants or conditions -- the goal was to obtain the best-fitting
-values to our entire dataset.
-
-\begin{table} \centering
-\caption{Results of model-fitting to the combined datasets from Experiments 1
-and 2. Note that smaller AIC values correspond to better model fits (adjusted
-for number of parameters)}
-\label{table:global-model-fits}
-\begin{tabular}{ l l l l} \\ \textbf{PACKER} & \textbf{Copy \& Tweak} &
-\textbf{Representativeness} & \textbf{Hierarchical Bayesian} \\ \hline $AIC =
-  9069$ & $AIC = 9813$ & $AIC = 8783$ &  $AIC =
-9881$ \\ $L = -4531$ & $L = -4905$ & $L = -4388$ & $L = -4937$ \\ $c = 0.51$ &
-                                                                                $c = 3.22$ & $\kappa = 12.23$ &
-$\kappa < 0.001$ \\ $\theta_c = 3.09$ & $\theta_c = 0$ (fixed) & $\nu = 1.00$ & $\nu = 5.44$ \\
-$\theta_t = 3.47$ & $\theta_t = 3.00$ & $\lambda = 7.04$& $\lambda
-                                                                       = 0.06$
-  \\ & &$\theta = 10.21 $& $\theta = 3.09$
-\\
-\end{tabular}
-\end{table}
-
-Table \ref{table:global-model-fits} contains the model fits. Due to the uneven number of fitted parameters among the models, we
-compare the model fits using the Akaike Information Criterion
-\citep[AIC;][]{akaike1974new}, where smaller values correspond to better fits
-(discounted by model complexity as measured by the number of parameters). The
-same qualitative results were obtained with alternative model comparison metrics
-(e.g., BIC, \citealp{schwarz1978estimating}; $AIC_C$,
-\citealp{hurvich1989regression}). In addition to AIC, table \ref{table:global-model-fits} contains the corresponding log-likelihood ($L$) and the best-fitting
-parameter values. These results reveal strong model differentiation: both
-contrast models (PACKER and the representativeness model) achieved far better
-fits compared to their non-contrast counterparts: copy-and-tweak and the
-hierarchical Bayesian model respectively. Interestingly between the contrast
-models, the hierarchical model (representativeness) outperformed the
-exemplar-based theory (PACKER), whereas between the non-contrast models, the
-reverse is observed. Specifically, here the exemplar-based model
-(copy-and-tweak) performed somewhat better than the hierarchical Bayesian model.
-
-While PACKER's advantage may tentatively be attributed to the model's
-sensitivity to category contrast (this will be explored in detail below), the
-advantage shown by copy-and-tweak over the hierarchical Bayesian model may be
-attributed to its exemplar-based representation of category B, as opposed to forcing a
-prototype-based representation as assumed by the hierarchical Bayesian model. As
-observed in Figures \ref{fig:e1-samples} and \ref{fig:e2-samples}, the generated
-categories we observed were often widely distributed, with no items near the
-category prototype. This aspect of the data is inconsistent with the
-multivariate normal distributions (similar to prototypes) used to represent
-categories in the \cite{jern2013probabilistic} model, but can be easily
-accounted for using an exemplar-based approach.
-Interestingly, representativeness using a prototype approach fits better than an exemplar-based approach.
-Understanding these differences more thoroughly is left for future research.
-
-A key distinction between the contrast and classical models is that only the
-contrast models are capable of making strong predictions about the location of
-new category members when the target class is entirely novel (i.e., no member of
-the category has been observed). Under these circumstances, there are no
-examples to copy, and thus the copy-and-tweak model predicts that items are
-generated at random. Likewise, with no observations on which to condition the
-category distribution, the hierarchical Bayesian model also picks an item at
-random.
-
-Thus, it is possible that the failure of the classical models is simply due to
-their inability to explain each participant's first trial (generating the first
-item in the `Beta' category). To ensure this is not driving our results, we
-conducted an identical set of simulations as above, excluding this trial
-(leaving 915 responses in the dataset): Again, the representativeness model ($L
-= -3286$, $AIC = 6580$) and PACKER ($L = -3377$,
-$AIC = 6759$) achieved better fits than the copy-and-tweak ($L = -3564$,
-$AIC = 7132$) and hierarchical Bayesian ($L = -3597$, $AIC = 7201$) models.
-
-Finally, because copy-and-tweak is nested within PACKER, we can use a likelihood
-ratio test to compare the two models. PACKER explains the aggregate data
-significantly better than copy-and-tweak ($\chi^2(1) = 747, p < .001$ for all
-data and $\chi^2(1) = 375, p < .001$ excluding the first example), providing
-further evidence that category generation is better explained when contrast is
-considered.
-
-Through comparison with the copy-and-tweak model, Figure
-\ref{fig:packer-loglike} more clearly demonstrates the robustness of the
-explanatory gains yielded by PACKER's category contrast mechanism. It displays
-the log-likelihood of the participants' results under PACKER as a function of
-the $\theta_c$ parameter. The model's other parameters ($c$,
-$\theta_t$) were set according to copy-and-tweak's best fits from Table
-\ref{table:global-model-fits}, and thus when $\theta_c=0$, the models are
-equivalent. The figure clearly shows a ``sweet spot'': a convex region in which
-PACKER achieves superior fits as a result of changes to $\theta_c$. The
-best fitting values lie well above the value of 0 assumed by the copy-and-tweak
-model, which demonstrates the robustness of the contrast effect (though note
-PACKER achieves even better fits when its parameters are fitted together, as in
-Table \ref{table:global-model-fits}). In sum, the data are better explained when
-both within-category similarity and category contrast is considered.
-
-\begin{figure}
-    \begin{center}
-      \includegraphics[width=\textwidth]{figs/packer-loglike.pdf}    
-      \caption{PACKER's fit as a function of its prioritization of
-        within-category and between-category similarity (using the $\theta_c$
-        parameter) . To facilitate comparison, PACKER's other parameters ($c$,
-        $\theta_t$) were set to the best fitting values obtained for
-        copy-and-tweak in Table \ref{table:global-model-fits}. The black diamond
-        marker indicates the log-likelihood for the point where
-        $\theta_c=\theta_t$. }
-    \label{fig:packer-loglike}
-    \end{center}
-\end{figure}
-
-\subsection{Individual Differences}
-\label{section:individual-diffs}
-
-As noted in Experiments 1 and 2, we observed a great deal of individual
-differences in the types of categories that participants generated. Within each
-condition, there were a wide variety of category types, such as row and column
-categories (see Figures \ref{fig:e1-samples} and \ref{fig:e2-samples}). The
-simulations reported above serve to evaluate the models while considering the
-entire dataset, but a secondary goal of any formal account should be to provide
-some explanation of how different profiles of performance emerge. Many of the
-individual generation profiles we observed can be described with the models
-simply by tuning the model's parameters in a principled manner. In this section,
-we describe more specifically how the most frequently observed profiles can be
-realized.
-
-By manual inspection, it is evident that the most common profiles of generation
-consist of: (A) a tightly-distributed `cluster' of examples, (B) `row'- and
-`column'-like arrangements (varying widely along one dimension but not the
-other), and (C) a `corners' arrangement with examples placed into disparate
-corners of the space. These four profiles are distinct in terms of the
-distribution of the generated category along each dimension: Whereas the cluster
-profile is tightly distributed along both dimensions, the row and column
-profiles are tightly distributed along just one dimension. Finally, the corners
-profile is widely distributed along both dimensions.
-
-In the framework proposed by PACKER, the cluster and corners profiles arise
-based on different prioritization of within-category similarity versus
-between-category contrast, and the row and column profiles arise based on the
-prioritization of each dimension in the computation of similarity. For example,
-in the cluster profile, there is a high degree of within-category similarity
-along both dimensions, whereas in the corners profile there is minimal
-within-category similarity. Thus, PACKER's proposal is that these individual
-differences arise as a result of different priorities: While the tight cluster
-configuration can be considered PACKER's `default' mode (as it maximizes
-within-category similarity), the corners profile can be produced when
-between-category contrast is put at a higher priority (i.e., $\theta_c > \theta_t$).
-
-Likewise, in the row and column profiles, there is a large degree of
-within-category similarity along one dimension but not the other. These
-differences likely arise due to a differential focus on one dimension over
-another, and thus they can be produced by changes to PACKER's attention weights,
-$w_1$ and $w_2$ (see Equation \ref{eq:similarity}). Traditionally, the attention
-weights in exemplar models are thought to reflect the diagnostic value of each
-dimension towards classifying the known category members
-\citep{nosofsky1984choice,nosofsky1986attention,kruschke1992alcove}, but within
-a generation context the weights specify the importance of within- and
-between-category similarity along each dimension. For example, if all of
-attention is allocated along the X-axis ($w_1=1$ and $w_2=0$), similarity along
-the Y-axis no longer influences performance. As a result, PACKER will create
-categories that are more widely distributed along the Y-axis, as similarity is
-not taken into account along that dimension. As a general principle,
-differentially weighting one dimension will result in the generation of
-categories that are more widely distributed along the ignored dimension,
-conforming to a row- or column-like arrangement. See Figure
-\ref{fig:packer-attention} for a depiction of how attention influences PACKER's
-performance.
-
-\begin{figure}
-    \begin{center} \inputpgf{figs/}{packer-attention-examples.pgf}
-    \caption{PACKER generation of a category `B' example, following exposure to
-one member of category `A' and one member of category `B'. Predictions are shown
-for different attention settings: {\em (a)} Increased weighting of the X-axis.
-{\em (b)} Increased weighting of the Y-axis. {\em (c)} Uniform weighting
-(identical to Figure \ref{fig:packer-examples}).}
-    \label{fig:packer-attention}
-    \end{center}
-\end{figure}
-
-As in PACKER, changes in the parameter settings of the copy-and-tweak model can
-also be used to produce different patterns of generation. Indeed, as
-copy-and-tweak is simply a special case of the PACKER model, the attention
-weights operate exactly as described above to produce row- and column-like
-categories. However, because the model is not influenced by category contrast,
-it is biased toward generating tightly clustered categories, as new items are
-always most likely to be generated near known examples of the target category.
-Thus, the lack of a contrast mechanism prevents the model from explaining why
-some individuals widely distribute their categories to the corners of the space.
-
-For both the hierarchical Bayesian and the representativeness models, the prior
-domain covariance matrix $\Sigma_0$ can be used to explain the generation of
-row-like and column-like categories. This covariance matrix specifies the amount
-of variance assumed along each dimension (as well as the dimensional
-correlations) across the domain of categories. The covariance matrix for a newly
-generated category, $\Sigma_B$, is based on the assumed $\Sigma_0$ as well as
-the distributions of previously learned categories (see Appendix
-\ref{ap:hsampling-definition}). Thus, the importance of each feature can be
-coded into $\Sigma_0$ to alter the dimensional variance of generated categories.
-Because the hierarchical Bayesian model possesses no mechanism to account for
-category contrast, the model is most likely to generate new items that are
-similar to known examples of the target category with no regard for how
-different it is to the contrast category. However, the representativeness model
-predicts that new exemplars should provide more relative evidence to the `Beta'
-category, accounting for the tendency of row-like and column-like profiles
-occupying the edge of the feature space.
-
-While the copy-and-tweak and hierarchical Bayesian models possess mechanisms to
-explain row- and column-like categories, they cannot easily explain why some
-individuals widely distribute their generated categories into disparate corners
-of the space. This, however, reveals a more general limitation: According to the
-copy-and-tweak and hierarchical Bayesian models, the distributional structure of
-generated categories is {\em independent} of their location within the domain.
-For example, although the copy-and-tweak or hierarchical Bayesian models can be
-parameterized to generate row- or column-like categories, there is no mechanism
-in place to ensure that what is generated will be distinct from what is already
-known. In the next subsection, we explore this prediction through an analysis of
-the interdependence between distributional structure and location in category
-generation.
-
-\subsection{Category Location vs. Distributional Structure}
-\label{section:individual-diff}
-As noted above, while all three models make clear claims about the internal
-structure of generated categories, the copy-and-tweak and hierarchical Bayesian
-models do not make any claims about how generated categories should differ from
-what is already known. However, as we observed in the results of Experiment 2,
-the distributional structure of a category is not always independent of its
-location within the domain. To demonstrate this point in more depth, we computed
-the X- and Y- axis ranges of every participant-generated category. Taking the
-difference between these values ($X-Y$) produces a measure of each category's
-orientation in the space: positive difference scores correspond to categories
-with more X-axis range (horizontally aligned, `Row' categories), whereas
-negative difference scores indicate the opposite (vertically aligned, `Column'
-categories). Neutral differences scores indicate there was an equal amount of X-
-and Y-axis range, which can be produced by a number of different category types
-(`Clusters', `Corners', etc; see Figures \ref{fig:e1-samples} and
-\ref{fig:e2-samples}). By plotting, for each possible stimulus, the difference
-scores of categories it was generated within, we can relate the distributional
-structure of generated categories to their location within the domain.
-
-However, because many stimuli were infrequently generated (such items near
-members of the `Alpha' category), we cannot simply compute the empirical average
-of the difference scores, as infrequently generated stimuli would be likely to
-show artificially strong differences. Instead, we used a Bayesian analysis to
-estimate the mean $\mu_x$ on the assumption that the scores $x$ for each
-stimulus are normally distributed with an unknown mean and unknown standard
-deviation. The conjugate Normal-Inverse Gamma distribution provides a
-straightforward method for this estimation:
-
-\begin{equation} \mu_x = \dfrac { \nu_0 \mu_0 + \sum{x} } { \nu_0 + n }
-\label{eq:rangediff-bayes}
-\end{equation} \ where $\mu_0$ is the prior mean, $\nu_0$ is a prior scale
-parameter (controlling the weighting of the $\mu_0$), and $n$ is the number of
-categories in which the stimulus was a member (i.e., the number of scores in
-$x$). The default assumption is that there is an equal amount of range along the
-X- and Y-axes, and so we set $\mu_0 = 0$. Likewise, to give a moderate amount of
-weighting to the prior mean we set $\nu_0 = 1$, though the results are robust to
-a range of values. Within this approach, the resulting aggregation is a
-trade-off between the number of generations and the strength of the range
-difference within each generated category. Infrequently generated stimuli, as
-well as those with mixed positive and negative scores, are given neutral
-difference scores.
-
-\begin{figure}[p]
-    \begin{center} 
-      \includegraphics[width=\textwidth]{figs/range-diff-gradients.pdf}
-      \caption{Behavioral and simulated range difference gradients. Each panel
-        shows, for each stimulus, the dimensional orientation of the categories
-        it was generated into: vertically aligned `columns' (orange) versus
-        horizontally aligned `rows' (purple).}
-      \label{fig:range-diff-gradients}
-    \end{center}
-\end{figure}
-
-The results of our analysis are shown in Figure \ref{fig:range-diff-gradients}
-for the experiment and model results\footnote{Prior to plotting, data were also
-processed using a Gaussian filter with $\sigma = 0.8$.}. The left-most column of
-Figure \ref{fig:range-diff-gradients} displays the effect of category location
-and contrast on the distributional structure of the category generated by
-participants. These data reveal strong and consistent patterns across all the
-conditions we tested in Experiments 1 and 2: Generated categories are more
-tightly distributed along the axis in which they are distinct. For example, in
-the `Cluster' condition, exemplars in the bottom-left of the space are more
-often generated into vertically aligned categories, and exemplars in the
-top-right are more often generated into horizontally aligned categories.
-Similarly, in the `Bottom' and `Middle' conditions, horizontally aligned
-categories are generated above and below the experimenter-defined categories,
-while vertically-aligned categories are generated to the sides. In the `Row'
-condition, most categories are horizontally aligned, and lie along the upper
-areas of the space. There are no strong range difference patterns in the XOR
-condition.
-
-These patterns of performance clearly depict the interdependence between the
-distributional structure and location of generated concepts. Our results can be
-interpreted in terms of local minimization of between-category similarity: By
-distributing the generated category away from members of the
-experimenter-defined category, participants may increase the degree of
-between-category distance without drastically altering the degree of
-within-category similarity.
-
-To explore how well the PACKER, copy-and-tweak, representativeness, and
-hierarchical Bayesian models explain our findings, we conducted simulations
-using an individual-differences approach. As noted in Section
-\ref{section:individual-diffs}, row- and column-like categories can be produced
-by each model through changes to the weighting of each dimension. Given this
-information, we may use the models to simulate each participant's generation
-separately, with the importance of each dimension set according to the relative
-range of the participant's generated category along each dimension.
-
-In the PACKER and copy-and-tweak models, the attention weights, $w$, specify the
-importance of each dimension in the computation of similarity. While there exist
-methods to find the optimal attention weighting scheme given a classification
-\citep[see][]{vanpaemel2012using}, for simplicity we assume that the `Alpha' and
-`Beta' categories are distinct along dimensions that the `Beta' exemplars do
-not vary on. In this case, the weighting for a given participant can be computed
-as:
-
-\begin{equation} w_k = \dfrac {\exp{ \{ -\theta_w \cdot\text{range}(k)} \} } {
-\sum_k {\exp{ \{ -\theta_w \cdot\text{range}(k)} \} } }
-\label{eq:range-weight}
-\end{equation}
-% 
-where $\theta_w$ is a free parameter controlling how differences in range
-correspond to differences in weights (functioning similarly to the $\theta$
-parameter in each of the models), and $\text{range}(k)$ is the range of examples
-generated by the participant along dimension $k$. We used $\theta_w = 1.5$ in
-our simulations, though the results are robust and similar for other $\theta_w$
-values. The resulting $w$ values are thus inversely proportional to the range of
-generated categories along each dimension, with less range corresponding to
-greater weighting.
-
-Unlike the PACKER and copy-and-tweak models, the representativeness and
-hierarchical Bayesian model's dimensional variances correspond to the assumed
-variance of generated categories along each dimension (rather than the inverse
-of the variance). Thus, a different transformation is appropriate for
-incorporating the weights computed in Equation \ref{eq:range-weight}. For the
-hierarchical Bayesian model, we computed the dimensional variances according to:
-$\lambda (1-w_k) 2$, where $\lambda$ is a free parameter specifying the overall
-assumed variance of the domain, and $2$ corresponds to the number of dimensions
-in our experiments\footnote{This calculation applies only in two-dimensional
-  domains, where $w_2 = 1-w_1$.}. Under this approach, evenly distributed
-weights correspond to an assumed variance of $\lambda$. Likewise, larger values
-of $w$, which are produced when the generated category is tightly distributed
-along one dimension, correspond to smaller assumed variances.
-
-Each model was used to simulate each participant's generation independently,
-with the importance of each dimension set according to the participant's
-generated category. The other free parameters within each model were set as in
-Table \ref{table:global-model-fits}. Every participant's generation was
-simulated 2,000 times; given the 305 participants tested across the two
-experiments, each model generated 610,000 categories in total. For comparison
-with our behavioral results, we then computed the range difference gradient
-identically as with the behavioral data. The results are shown in Figure
-\ref{fig:range-diff-gradients}.
-
-As in the more traditional model evaluation analysis described above, the
-contrast models (i.e., PACKER and the representativeness model) provided a much
-closer match to our behavioral results than the copy-and-tweak and hierarchical
-Bayesian models. In all conditions, the contrast models distribute categories
-similarly to the behavioral data: Horizontally-aligned categories tend to be
-placed above and below members of the experimenter-defined category, and
-vertically-aligned categories tend to be placed to the sides. Conversely,
-because the copy-and-tweak and hierarchical Bayesian models are insensitive to
-category contrast, these models do not produce any systematic patterns of
-association between category location and distributional structure. The sole
-exception is within the `Row' condition of Experiment 1, in which the majority
-of participants generated a `Row'-like category, widely distributed along the
-X-axis but not the Y-axis. In these cases, both models are initialized with
-weights that produce Row categories, but because category contrast is not
-considered, categories are uniformly generated across the entire domain, rather
-than concentrated within the upper-regions as observed behaviorally.
-
-%Experiment 3 starts here
-\section{Experiment 3}
-
-Experiments 1 and 2 clearly establish the importance of contrast in category
-generation. Follow-up model-based analyses illustrated that both contrast models
-account for participant performance better than models that do not take into
-account contrast. We also found more support for contrast as representativeness
-hypothesis over contrast as exemplar dissimilarity hypothesis.
-
-Although category generation is an intriguing aspect of human cognition, it has
-played a relatively limited role in the broader categorization literature. To
-establish that contrast plays a broader role in categorization, we conducted a
-traditional category learning study. In it, participants learned two categories,
-where the categories learned by each participant in Experiment 3 was yoked to
-the `Alpha' shown and `Beta' generated by each participant in Experiment 2. Our goal
-is to investigate whether model performance (in generating categories) is
-consistent with human performance (in learning those categories). Specifically, if contrast plays a broader role in categorization, then `Beta' categories rated as "better" according
-to the contrast models should be easier for participants to learn.
-%\footnote{It is possible to adjust the present models of category generation such that they can make category learning predictions.
-However, such an adjustment makes the strong assumption that the very same processes governing category generation also apply to category learning.
-Taking this approach would fall outside the scope of the current investigation.}
-\subsection{Participants, Materials, and Procedure}
-
-Experiment 2 recruited 122 participants who each generated one set of four `Alpha'
-and four `Beta' category exemplars. Experiment 3 recruited the same number of
-participants.
-
-Participants observed four blocks of eight trials. Each trial began with the
-presentation of a fixation cross for 500 ms. This was followed by the
-presentation of one exemplar randomly sampled without replacement from the
-unique category set. Participants were tasked with assigning the presented
-exemplar to either the `Alpha' or `Beta' category with no time limit imposed.
-Feedback was automatically displayed for 2500 ms after each response.
-
-\subsection{Results and Analysis}
-
-Among the 122 originally generated category sets were 102 unique sets (i.e.,
-they contained a unique collection of `Alpha' and `Beta' exemplars). Consequently,
-for our subsequent analyses for this experiment we only use data from the first
-102 participants that were presented with a unique category set.
-
-Overall accuracy of the participants was high, with a mean error rate of $.19$
-($SD = .19$). Aggregated error rates for each block are presented in Figure
-\ref{fig:learningcurve}.
-
-\begin{figure}
-    \begin{center}
-    \includegraphics[width=\textwidth/2]{figs/e3-learningcurve.pdf}
-    \caption{Average error rate for each successive block. Green discontinuous
-line represents the overall mean error rate.}
-    \label{fig:learningcurve}
-    \end{center}
-\end{figure}
-
-In order to analyze the consistency between model and human performance, we
-optimized each model such that the Pearson correlation between the model's
-negative log-likelihood of generating each unique category set and the
-participant's error in learning that category set is maximized. As with the
-analysis in Section \ref{section:individual-diff}, we simulated the weighting of
-each dimension independently depending on the generated categories.
-
-The greatest correlations were observed when fitting the contrast models.
-Specifically, PACKER correlated with human performance with
-$r = .65, 95\%\,CI\, [.48,.78]$, and the representativeness
-model correlated with human performance with $r = .59, 95\%\,CI\,[.40, .73]$\footnote{Confidence intervals were obtained using a bootstrap
-  method with 1000 bootstrap samples.}.
-Copy-and-tweak yielded a correlation of $r = .45, 95\%\,CI\,[.28, .62]$ and the
-hierarchical Bayesian model yielded a similar correlation of
-$r = .46, 95\%\,CI\, [.30, .62]$. These results are graphically presented in Figure
-\ref{fig:perror_corr}.
-
-\begin{figure}
-    \begin{center}
-    \includegraphics[width=\textwidth]{figs/modelvspptp.pdf}
-    \caption{Scatterplots indicating the correlations between observed
-      participant error and model fit. Blue discontinuous line are best-fit
-      lines. Green discontinuous lines indicate boundaries generated by the 95\%
-      CI around the fitted correlations. }
-    \label{fig:perror_corr}
-    \end{center}
-\end{figure}
-
-To emphasize the strong influence of contrast in maximizing the association
-between category generation and category learning, we computed the correlations
-over wide range of $\theta_{contrast}$ values, with the other parameter values
-of PACKER held constant at their optimized levels. As presented in Figure
-\ref{fig:packer-corr}, the correlation quickly increases with increasing weight
-on contrast, reaching a plateau above values of around 4.0.
-
-\begin{figure}
-    \begin{center}
-    \includegraphics[width=0.8\textwidth]{figs/packer-corr.pdf}
-    \caption{Correlation between PACKER's fit and participant error as a
-function of the $\theta_{contrast}$ parameter. To facilitate comparison,
-PACKER's other parameters ($c$, $\theta_{target}$) were set to the best fitting
-values obtained for copy-and-tweak in Table \ref{table:global-model-fits}. Grey
-dashed line represents the correlation between Copy \& Tweak's fit and
-participant error.}
-    \label{fig:packer-corr}
-    \end{center}
-\end{figure}
-
-\section{General Discussion}
-The sensory impression of every stimulus and event is unique.
-Grouping distinct patterns of sensory information into categories is a fundamental task solved by the mind.
-Most work has focused on how people learn new categories that are provided to them through unlabeled, partially labeled, or fully labeled examples.
-How were these categories first determined?
-Some natural categories are likely to be the result of regularities in the dynamics of our environment.
-But, these are only a subset of the categories that people learn.
-Other categories, such as tools and ideas, were generated by people over time.
-What basic principles underlie how people generate categories?
-
-While the bulk of prior research on categorization has focused on the classic
-finding that generated concepts tend to be distributionally similar to known
-concepts, there has been little work addressing the role of contrast in category
-generation: How is it that people are able to create something {\em different}
-from what is already known? We developed two novel models, each incorporating a
-different conceptualization of category contrast. One model is PACKER, which is
-an exemplar-based model that formally specifies the role of contrast in
-generation as exemplar dissimilarity. Specifically, the model proposes that
-categories are represented as exemplars in a multidimensional psychological
-space, and generation is constrained both by within-category and
-between-category similarity: Exemplars belonging to the same category should be
-similar to one another, and exemplars belonging to different categories should
-not be similar to one another. The second model is a novel hierarchical Bayesian
-model with a representativeness mechanism. This model generates exemplars that
-are more representative of (i.e., has greater relative evidence from) the novel
-category compared to the learned category.
-
-We reported two experiments demonstrating systematic effects of category
-contrast in category generation. Members of participant-generated categories
-tended to be highly dissimilar from members of previously-learned categories,
-and were usually more similar to one another than to members of other
-categories. We also observed broad interdependence between the distributional
-structure (feature variance, correlation) and physical instantiation (location
-within the stimulus space) of generated categories: In Experiment 2, we found
-that the unoccupied regions of the domain influenced the distributional
-structure of categories, and in both experiments we observed that participants
-distributed their generated categories to increase contrast with what was
-already known.
-
-We conducted simulations comparing the contrast models' account of our results
-to the classical proposals for category generation: a ``copy-and-tweak'' model
-(realized as a variant of PACKER with no sensitivity to category contrast), and
-a hierarchical Bayesian model designed to explain the classic distributional
-similarity effect. In all simulations, we found that the contrast models
-captured a previously unexplained and unexplored aspect of human category
-generation. In particular, by measuring PACKER's fit as a function of its
-prioritization of within- and between-category similarity, we observed that
-considering either constraint exclusively results in a relatively low-quality
-account. Instead, PACKER's best results were obtained when both constraints are
-considered, indicating that human learners do not generate novel concepts
-exclusively on the basis of within-category similarity or between
-class-contrast. This finding mirrors our behavioral results and demonstrates
-that both constraints influence generation when explained with an exemplar
-model. Ultimately, our category generation data were best explained by the
-representativeness model. The exact nature of its superiority over the other
-models is still not yet clear. However, as we explore in a later section, it is
-likely that the representativeness mechanism itself (i.e., independent of the
-hierarchical Bayesian foundation) captures a core aspect of both category
-generation and category learning.
-
-\subsection{Finding Contrast Effects in Prior Work}
+\section{Finding Contrast Effects in Prior Work}
 
@@
 
-\subsection{Similarity and Contrast in Cognition}
-The proposal that category contrast is a primary constraint in categorization is, in some ways, entirely commonsense.
-For categories to be useful, they should not all be identical, or in other words, they should be different.
-Thus, a newly generated category should be different than pre-existing categories.
-Beyond its role
-in category generation specifically, category contrast is also of fundamental importance in
-categorization more broadly. All other factors held constant, new categories are
-easier to learn if they are dissimilar to members of other categories, and
-knowledge of highly distinct categories is applied more accurately than that of
-ill-defined categories \citep{ashby1994categorization,imai1965discriminability}.
-Likewise, basic-level categories \citep{rosch1976basic} are thought to be
-abstracted in order to maximize within-category similarity while minimizing
-between-category similarity. Finally, the act of forming category
-representations affects similarity judgments about category members and
-nonmembers, with category members being viewed as more similar to one another
-than members of other categories
-\citep{goldstone1994influences,goldstone2001altering,goldstone1996isolated}.
-
-Beyond categorization, one can find instances of the trade-off between within
-and between-class similarity in linguistic categories over perceptual
-dimensions. For example, \cite{regier2007} showed that the partitioning of color
-categories reflects such a trade-off in a psychological space -- colors are
-partitioned into groups with members that are viewed as highly similar to one
-another yet distinct from other colors. A similar trade-off can be observed in
-phoneme categories. Different exemplars of the same phoneme must be similar to
-one another, while contrasting from other phonemes, such that a listener can
-infer the appropriate phoneme. This pattern has been found and modeled in the
-natural acoustics of American English vowels
-\citep{feldman2013,hillenbrand1995}. As linguistic categories must have been
-created at some point in human history, it is revealing that the constraints of
-emulating distributional structure across categories and having categories
-contrast from one another still bias human category generation today.
-
-The dual forces of within-class similarity and between-class contrast influence
-cognitive functions in a wide variety of domains. The PACKER model is notable in
-that it explicitly interprets this trade-off within the domain of
-categorization, and allows us to begin to understand the relatively understudied
-processes involved in category generation through our more well-developed tools
-for understanding human categorization.
-
-\subsection{Implications for Creative Cognition}
-
-Although the focus of this article has been to address the role of contrast in category generation and learning, our findings and approach have relevant implications
-for research in creative cognition. A central focus of the creative cognition
-approach has been to explain acts of creativity in terms of the mental
-representations and processes that are commonly studied in cognitive psychology
-and cognitive science \citep{finke1992creative,smith1995creative}. However,
-unlike other fields in the study of cognition, creative cognition research
-rarely employs quantitative models to evaluate the explanatory value of such
-representations and processes. Our modeling results provide a concrete example
-of how formal approaches may be used to gain insight into the nature of creative
-cognition.
-
-In addition to demonstrating the utility of formal modeling for studying
-creative cognition, the contrast models here specifically offer an additional
-interpretation of some of the field's most central findings. For example,
-perhaps the most foundational principle from this literature concerns the
-limiting influence of prior knowledge: Individuals create new categories
-composed of features from existing classes, and what is created can be
-influenced drastically through the introduction of cues or examples
-\citep{marsh1999inadvertent,smith1993constraining}. In this paper, we have
-identified another important aspect of the constraining influence of prior
-knowledge: What is generated cannot be the same as what is already known.
-Further, there is systematicity in how generated categories differ from prior
-knowledge. The results of our simulations suggest that this conceptualization of
-difference can be addressed in at least two different ways. Specifically,
-simulations with PACKER demonstrate that the constraining influence of
-difference is concisely explained in terms of a trade-off between
-within-category similarity and between-category dissimilarity. Conversely, the
-representativeness model shows that this influence can also be the result of
-enhancing the representativeness of generated exemplars to their category.
-
-PACKER may offer an additional interpretation of existing accounts of creative
-generation. Most notably, a leading account within the creative cognition
-literature, the Path of Least Resistance \citep{ward1994structured,ward1995s},
-also explains generation in terms of an exemplar-based retrieval process. This
-account was designed to explain the creative generation of natural categories
-(e.g., new species of plants and animals) and as a result relies strongly on the
-hierarchical organization of these categories: Individuals are thought to
-retrieve an example of the higher-level category being generated (e.g.,
-\textit{bird} may be retrieved from the category \textit{animal}), and then
-systematically alter what was retrieved to make something new. As the PACKER
-model does not assume knowledge is hierarchically organized \citep[this is true
-of the exemplar view more broadly, see][]{murphy2016exemplar}, the model may be
-viewed as a formal instantiation of the Path of Least Resistance for application
-in a traditional artificial categorization domain (when there is no established
-hierarchy of categories). PACKER's success in explaining generation within an
-artificial domain motivates future work exploring the nature of category
-contrast within a more naturalistic setting.
-
-The broader study of creativity currently involves a wide breadth of different
-approaches\citep[for a review see][]{kozbelt2010theories}, such as those based
-on free association \citep{mednick1962associative} and conceptual combination
-\citep{estes2002emergence,murphy1988comprehending}. In addition, recent work in
-the machine learning literature has explored using neural networks to address
-the overall problem of creative generation
-\citep[e.g.,][]{goodfellow2014gan,ho2016gail,kingma2016,chen2016infogan}. In
-contrast to these varied investigations, we reduced our focus by studying a
-highly complex behavior (category generation) as it applies within a
-well-established domain (artificial category learning). We hope that future work
-incorporates and highlights the importance of contrast into theories of
-creativity.
-
-\subsection{Implications For Categorization}
-
-Categorization research addresses the representations and processes that
-underlie the learning and use of categories. Category learning tasks are
-generally about figuring out which items belong to which category. Once learned,
-categories are generally used to classify new stimuli and to make inferences
-beyond the available information. Our work is fairly unique in that people learn
-a single category through positive examples and then create another category
-that would make sense in the domain. In other words, only one category is
-provided to be learned and the use that is required of that category is to
-generate a contrast category. Such scenarios have direct application to
-real-world situations. For instance, consider being stranded in an unfamiliar
-environment where you eat and get sick from a plant with certain features. In
-avoiding such plants and seeking alternative sources of nutrition, it would
-behoove you to imagine what features an edible, safe plant would possess (e.g.,
-if the poisonous plant was bright red, perhaps a non-red plant would be safe to
-consume). In the present studies, we have learned something about the form that
-such expectations are likely to take.
-
-We can think of the category generation task in our studies as asking a person
-to formulate an idea about what set of items in the domain are most
-interestingly {\em not} members of the original category. To meet this
-condition, the items must take some form of coherence that aligns with that of
-the original category and some form of distinctiveness relative to the original
-category. Reflecting the basic level of organization in natural categories, it
-makes sense to generate a set of items that possess strong within-category
-coherence (by importing or systematically transforming the internal structure of
-the original category) as well as strong between-category differentiation (by
-creating maximum contrast with the original category, be it through
-exemplar-based dissimilarity or the maximization of exemplar-category
-representativeness). In this sense one can see the patterns of performance in
-the category generation task as recapitulating the order of semantic
-organization.
-
-The current work also suggests exciting directions for related investigations
-into unsupervised categorization. While the unsupervised categorization
-literature is primarily interested in the generation of categories within a set
-of observed exemplars where no prior categories are learned, our category
-generation studies are focused on the production of novel category exemplars
-themselves. Despite this distinction, both types of categorization research are
-ultimately interested in the question of how categories can be formed. The results
-of our current work indicate that the formation of categories is constrained by
-some measure of contrast. To our knowledge, this explicit investigation of
-category contrast has not yet been attempted in the unsupervised categorization
-literature.
-
-Interestingly, the unidimensional sort bias that is commonly observed in
-unsupervised categorization
-\citep{imai1965discriminability,milton2004influence,ahn1992two}, where
-participants sort unlabelled family resemblance category exemplars by focusing
-on a single feature, was not consistently observed in our tasks. In our
-experiments we only observed this bias in the row condition of Experiment 1,
-whereas in other conditions participants generated categories that typically
-varied along both dimensions. This may be unsurprising at face value, since
-participants in conditions other than the row condition were trained on a prior
-category that varied on both dimensions. However, these results suggest that the
-unidimensional sort bias is not universal to all types of category construction
--- or more specifically, the unidimensional bias may not apply when participants
-need to generate categories with entirely new exemplars. Additionally, the
-unidimensional sort bias is observed in category construction tasks in which
-participants develop a category-level organization of a provided domain while
-category generation tasks involve generating a new category relative to one that
-was just acquired (one might say that participants are forging a domain from a
-category rather than categories from a domain). Since the more creative task of
-generating a category from scratch (with respect to an established category)
-does not invoke the unidimensional bias, this suggests that a more creatively
-demanding version of a sort task might also reduce the bias.
-
-\subsection{Exemplar Dissimilarity and Representativeness in Categorization} 
-Although both PACKER and the representativeness model perform better than the
-classical models, they appear to be capturing fundamentally different aspects of
-contrast in categorization. Specifically, while the representativeness model
-excelled in predicting our category generation data, PACKER was the best-fitting
-model in accounting for participant error in category learning. Given the
-history of their development, this is perhaps not entirely surprising. PACKER
-is implemented as an extension of the GCM, which was originally developed for (and
-has been highly successful in) category learning \citep{nosofsky1986attention,
-  nosofsky1994comparing}. Conversely, the representativeness mechanism was
-conceived as a heuristic describing the relative extent to which an individual
-observation is generated from some unobserved population
-\citep{kahneman1972subjective,abbott16}. This generative nature of the
-representativeness mechanism intuitively lends itself to the prediction of
-category generation data.
-
-While the contrast models were born quite naturally out of the corresponding
-classical models, it is worth considering if similar benefits in prediction
-would be seen if the contrast mechanisms were applied to different classical
-models. Although there is no clear way to adapt the exemplar dissimilarity
-mechanism to the current hierarchical Bayesian model, it is a fairly
-straightforward process applying the representativeness mechanism to an exemplar
-model. Specifically, by treating the similarity measure (Equation
-\ref{eq:similarity}) in the current implementation of copy-and-tweak as a
-density estimate for the representativeness mechanism (Equation
-\ref{eq:representativeness}), we gain a new model of representativeness that is
-based on exemplar similarity instead of a multivariate Gaussian
-likelihood.\footnote{An analogous implementation of the representativeness
-  mechanism to PACKER results in both $\theta_c$ and $\theta_t$ adding to
-  constant value that is independent of the similarity to contrast and target
-  exemplars respectively. Consequently, it is formally equivalent to a
-  copy-and-tweak model with representativeness.}
-
-How well does a representativeness model that uses exemplar similarity to
-represent the category’s density over features explain our experimental results?
-We find that this model substantially outperforms its classical counterpart, but
-not to the extent of PACKER or the representativeness model. Specifically,
-fitting this model to the category generation data yielded $L = -4815$ and
-$AIC=9633$ for the entire set of data from Experiments 1 and 2, and $L = -3474$,
-$AIC=6953$ for data excluding the first trials. When used to predict the error
-rates of participants in the category learning task (Experiment 3), the
-performance of the copy-and-tweak with representativeness model correlated with
-participant error with $r=.56, p < .001$. These results suggest that the
-predictive advantages to including the representativeness mechanism is not
-limited to a hierarchical Bayesian model, but can also be found when applied to
-an exemplar model. Although its performance is not as strong as either of the
-fully-developed contrast models, there is a benefit to the representativeness
-mechanism that is independent from any interaction with a hierarchical Bayesian
-framework.
-
-Ultimately, the source of the discrepancy between the performance of the two
-contrast mechanisms in category generation versus category learning is still not
-entirely clear. Our results could be an indication of a fundamental difference
-between the processes involved where exemplar dissimilarity is a larger factor
-in category generation while representativeness exerts more influence in
-category learning. Alternatively, it is possible they are both approximating
-some factor that we still have not fully identified. While a thorough
-investigation is beyond the scope of the current article, the preliminary
-results of this section indicate that further work to disentangle these two
-mechanisms would be promising. In the following section, we explore some other
-limitations of the current work and provide more recommendations for future
-research.
-
-\subsection{Limitations and Future Directions}
-
-Although successful in explaining our results in all three studies (albeit not
-as well as the representativeness model for Experiments 1 and 2), PACKER does
-not provide a full account of what is known about category generation. Most
-notably, in this paper we have not evaluated the model's ability to explain the
-classic finding that generated categories tend to share distributional
-commonalities with previously learned categories
-\citep[see][]{jern2013probabilistic,ward1994structured}. While we successfully
-replicated this effect in Experiment 1, we also found that its influence was
-limited in comparison to the fundamental constraints imposed by category
-contrast. Even within Experiment 1, we found systematic inconsistencies: by
-generating exemplars into unoccupied regions of the space, participants who
-learned an `XOR' category, composed of members that are widely distributed along
-both features and are positively correlated in space, tended to generate
-categories with an opposite (negative) correlation. More generally, PACKER
-inherits the strengths and weaknesses of exemplar models of categorization: It
-provides a simple and flexible model that explains many results, but deviates
-systematically from human performance in some cases.
-
-Nonetheless, these classic effects are a core element of the phenomenology of
-category generation, and PACKER does not include any mechanisms that explain
-them. Instead, through the development and evaluation of the PACKER model, we
-have sought to add new elements into such a phenomenology: The broad and strong
-influence of category contrast, and the interdependence between category
-location and distributional structure. It may be possible to combine the
-hierarchical Bayesian approach proposed by \cite{jern2013probabilistic} with
-PACKER's underlying claims to obtain a ``best of both worlds'' model, capable of
-explaining the role of contrast in category generation, as well as the emulation
-of distributional structure. However, as noted in the introduction, the
-incorporation of category contrast is antithetical to the core principles of a
-traditional, semi-conjugate Bayesian approach. This suggests that category
-generation is a fundamentally different computational-level problem
-\citep[different from those posed by][]{jern2013probabilistic,kemp2014taxonomy}.
-
-Characterizing that problem and conducting a rational analysis is an important
-direction for future research. To that aim, we plan to explore the connection
-between exemplar modeling as an Importance Sampling approximation
-\citep{shi10exemplar}, and see what sort of computational-level problem PACKER
-approximates. Once formalized in probabilistic terms, it should also be
-straightforward to incorporate distributional factors into the model. This would
-unite PACKER with the Bayesian representativeness model, and they would differ
-in terms of their assumptions about how people represent category distributions
-(as exemplars or prototypes, respectively). Alternatively, it may be possible to
-integrate the core principles of either model of contrast into other
-categorization models
-\citep[e.g.,][]{love2004sustain,kurtz2007divergent,smith2000thirty}.
-
-The present work focuses on the influence of contrast on categorization,
-primarily by exploring two different conceptualizations of contrast. However,
-throughout this paper we have only explored one type of category generation
-problem, that is the generation of new categories of new categories in an
-artificial domain. It is possible that the strength of the influence of contrast on
-category generation decisions varies depending on the nature of the category
-generation problem, for instance, in situations where the boundaries of the
-domain are not clearly defined. To provide some intuition for this, consider the
-following example: an entomologist is asked to draw a new type of insect that
-they have never seen. Clearly, features of the new insect will be constrained by
-the definition of an insect: it will be an arthropod with six legs and a
-relatively hard exoskeleton. However, there are no strict limitations on other
-features such as how large the insect needs to be, or if the insect should have
-a particular type of pattern on its body. In this scenario, it is plausible that
-the entomologist will draw an insect that is fairly similar to insects that
-currently exist (e.g., small in size with a camouflage pattern that matches its
-environment) as opposed to something completely different (e.g., a horse-sized
-beetle with roman numerals on its back), indicating that contrast plays only a
-minor role in the decision making process. In this situation it appears that
-there is a combination of clearly defined constraints on the domain boundary
-that are imposed by the external environment (insects must be arthropods with
-six legs) but also ill-defined constraints that are imposed by the observer
-(insects are probably small and have patterns on their body that suit the
-environment). However, the relation of these contraints (and the interactions
-between them) to the role of contrast is not currently well understood.
-
-The definition of the domain boundary is only one possible factor that could
-influence the effect of contrast on categorization. It is also plausible that
-other conditions such as the nature of the instruction (``Generate a new
-category that is Not Alpha'' vs ``Generate a new category that is Beta'') and
-the boundedness of the features (using features such as hue and orientation that
-do have have strict minima or maxima) can affect the extent to which contrast
-influences categorization. Though our current contrast models can allow the
-influence of contrast to vary, they do not make any predictions regarding the
-specific conditions under which contrast can vary. Consequently,
-addressing this issue would be an immensely promising avenue for future
-research.
-
-\section{Conclusions}
-
-The generation of new concepts and ideas is a highly interesting topic, but it
-is difficult to study in a controlled experimental environment. In this paper,
-we have provided such an examination of category generation as it applies within
-an artificial categorization experiment. Extending the literature on creative
-cognition, our experiments provide a detailed picture of the role of category
-contrast in generation: People seek to create concepts that are distinct from
-what they already know, and the nature of what is created can be influenced by
-what does not yet exist. Our simulations with traditional exemplar models, as
-well as a hierarchical Bayesian model, provide strong support for the claim that
-category contrast is of fundamental importance to categorization. More broadly,
-our results demonstrate that popular explanatory approaches from basic research
-in cognitive science can offer a precise, quantitative account of a behavior as
-complex as that of category generation, and in exploring them, we can increase
-our understanding of categorization as well.
-
-\clearpage
-\section{Acknowledgments} Previous versions of this work were presented at the
-Thirty-Ninth Annual Conference of the Cognitive Science Society and Forty-Ninth
-Annual Meeting of the Society for Mathematical Psychology. Support for this
-research was provided by the Office of the VCRGE at the UW - Madison with
-funding from the WARF. We thank Charles Kemp for feedback on a previous version
-of this manuscript. We also thank him and Alan Jern for providing code and data.
-\end{flushleft}
-
-% references section
-\clearpage \bibliographystyle{apacite} \bibliography{citations.bib}
-\clearpage
-
-\appendix \numberwithin{equation}{section}
-
-\section{The Hierarchical Bayesian Model of Concept Generation}
-\label{ap:hsampling-definition}
-
-\cite{jern2013probabilistic} demonstrated how a hierarchical Bayesian model
-could explain the distributional correspondences between observed and generated
-categories. In their model, exemplars of generated categories were viewed as
-samples from a multivariate Normal distribution over the dimensions of stimulus
-space. The mean of the generated category was independent of the observed
-categories, but the covariance matrix (encoding feature variances and
-correlations) was based on a common prior distribution. Generating a new
-category was thus completed by sampling a new category mean (uniform over
-stimulus space) and covariance matrix from the common prior distribution.
-Because the shared prior distribution's parameters were unobserved, the
-hierarchical Bayesian approach was used to infer its parameters from the
-previous categories (their feature variances and correlations), and then to
-generate the covariance matrix of the new category.
-
-In our implementation of their model\footnote{Note that
-\citet{jern2013probabilistic}'s model is slightly different, as they used a
-semi-conjugate model. Their model acts very similarly to our version.}, each
-category's exemplars are viewed as samples from a multivariate Normal
-distribution with parameters ($\mu, \Sigma$). Category covariance matrices
-(specifying variance and covariance along $k$-dimensions), are assumed to be
-Normal-Inverse-Wishart distributed with parameters: $\nu$ ($>k-1$), $\kappa$
-($>0$), and $\Sigma_D$. $\nu$ and $\kappa$ are treated as free parameters in our
-simulations, and $\Sigma_D$ is the domain-wide covariance matrix from which all
-categories are viewed as samples. Assuming a given $\Sigma_D$, a category
-covariance matrix $\Sigma$ can be computed on the basis of its examples:
-
-\begin{equation} \Sigma = \left[ \Sigma_D \nu + C + \dfrac {\kappa n} {\kappa +
-n} (\bar{x}-\mu)(\bar{x}-\mu)^T \right] (\nu + n)^{-1}
-\label{eq:Sigma_B}
-\end{equation}
-% 
-where $\bar{x}$ and $C$ are the empirical mean and covariance of the category's
-known members, and $n$ is the number of observed members of the category. When
-there are fewer than two known members of the category (and thus no covariance
-to speak of), $\Sigma = \Sigma_D\nu$.
-
-The category mean, $\mu$, can be computed as:
-
-\begin{equation} \mu = \dfrac {\kappa\mu_{0} + n \bar{x}} {\kappa + n}
-    \label{eq:category_mus_jk13}
-\end{equation}
-% 
-where $\mu_{0}$ is the prior mean. In our simulations, $\mu_{0}$ is set to the
-center of the domain. However, when no examples of the target category have been
-observed, generation is assumed to be random. In practice, the model's best fits
-are achieved when the $\kappa$ parameter, which controls the influence of
-$\mu_0$ on $\mu$, is set very close to zero (hence, the influence of $\mu_0$ is
-minimal).
-
-Importantly, the domain-wide covariance matrix $\Sigma_D$ is unobserved and
-needs to be inferred from the observed categories. For conjugacy, if $\Sigma_D$
-is viewed as a sample from an Inverse-Wishart distribution with scale
-$\Sigma_0$, $\Sigma_D$ can be computed as:
-
- \begin{equation} \Sigma_D = \Sigma_0 + \sum_y{C_y}
-    \label{eq:category_sigmas}
-\end{equation}
-% 
-where $\Sigma_0$ is the prior covariance in the domain. In our simulations,
-$\Sigma_0 = \lambda {\bf I}$, where $\lambda$ is a free parameter controlling
-the expected variance of dimensions (dimensions of the domain covariance matrix
-are expected to be uncorrelated) and ${\bf I}$ is a $k$-by-$k$ identity matrix.
-
-Generated exemplars are drawn from a multivariate Normal distribution specified
-by $(\mu, \Sigma)$. Thus, $p(y)$ is
-\begin{equation} p(y \mid x) = \dfrac {\exp \left \{ \theta \cdot {\rm
-Normal}(y; \mu, \Sigma) \right \} } {\sum_i \exp \left \{ \theta \cdot {\rm
-Normal}(y_i; \mu, \Sigma) \right \} }
-\end{equation}
-% 
-where $\theta$ is a response determinism parameter and ${\rm Normal}(y; \mu,
-\Sigma)$ denotes a multivariate Normal density evaluated at $y$.
+\bibliographystyle{apacite} \bibliography{citations.bib}
 

### 2022-11-25 Manuscripts/model_descriptions/jern-kemp-conjugate/jern-kemp-conjugate.tex
STILL HYPHENS AAHHHHHHHHHHHH

--- 2017-01-30 Manuscripts/model-descriptions/jern-kemp-conjugate/jern-kemp-conjugate.tex
@@
 
-\section*{Updated Version of the Jern \& Kemp (2013) Model}
+\section*{Conjugate Implementation of the Jern \& Kemp (2013) Model}
 
-Working backwards, what is needed for simulation of exemplar generation is the probability of generating a stimulus $y$ given exposure to members of the target category $x_B$:
-\
-\begin{equation}
-p(y | x_B) = ?
-\end{equation}
-\
-where $x_B$ may be empty.
-Jern \& Kemp's model achieves this using a generative process.
 Category members ($x_A$ or $x_B$; more generally written as $x_C$) are assumed to have been generated using an underlying category distribution (specifically, multivariate normal):

### 2022-11-25 Manuscripts/model_descriptions/packer/packer_new.tex
STILL HYPHENS AAHHHHHHHHHHHH

--- 2017-04-04 Manuscripts/model-descriptions/packer/packer.tex
@@
 
-\documentclass[12pt]{article}
-\usepackage[letterpaper, margin=1in, headheight=15pt]{geometry}
-\usepackage{amsmath, listings}
+\documentclass[12pt]{article} \usepackage[letterpaper, margin=1in,
+headheight=15pt]{geometry} \usepackage{amsmath, listings}
 
@@
 
-\section{The PACKER Model: Producing Alike or Contrasting Knowledge with Exemplar Representations}
+\section{The PACKER Model: Producing Alike or Contrasting Knowledge with
+Exemplar Representations}
 
-This model assumes that people represent categories as exemplars in a multidimensional space, and that generation is constrained by both similarity to members of the category being generated \textit{and} dissimilarity to members of other categories.
-The model assumes people generate categories that are dissimilar to known categories and have strong within-class similarity.
+This model assumes that people represent categories as exemplars in a
+multidimensional space, and that generation is constrained by both similarity to
+members of the category being generated \textit{and} dissimilarity to members of
+other categories. The model assumes people generate categories that are
+dissimilar to known categories and have strong within-class similarity.
 
-The similarity between exemplars $x_i$ and $x_j$ is computed using Shepard's law:
+The similarity between exemplars $x_i$ and $x_j$ is computed using Shepard's
+law:
 
-\begin{equation}
-s\left(x_i,x_j\right) = \exp \left\{ -c \left[\sum_{k}{ w_k \left| x_{ik} - x_{jk} \right|^r }\right]^{1/r} \right\}
+\begin{equation} s\left(x_i,x_j\right) = \exp \left\{ -c \left[\sum_{k}{ w_k
+\left| x_{ik} - x_{jk} \right|^r }\right]^{1/r} \right\}
 \label{eq:similarity}
@@
 
-When prompted to make a generation decision, participants are thought to consider both similarity to examples from other categories as well as similarity to examples in the target category.
-More formally, the aggregated similarity $a$ between candidate $y$ and the model's stored exemplars $x$ can be computed as:
+When prompted to make a generation decision, participants are thought to
+consider both similarity to examples from other categories as well as similarity
+to examples in the target category. More formally, the aggregated similarity $a$
+between candidate $y$ and the model's stored exemplars $x$ can be computed as:
 
-\begin{equation}
-  a(y, x) = \sum_j{f(x_j) s(y, x_j)}
+\begin{equation} a(y, x) = \sum_j{f(x_j) s(y, x_j)}
 \end{equation}
 
-Where $f(x_j)$ is a function specifying each stored example's degree of contribution toward generation.
-Although $f(x_j)$ may be set arbitrarily, in PACKER it is set according to class assignment.
-For known members of the target category, $f(x_j) = \gamma$.
-For members of contrast categories, $f(x_j) = -1 + \gamma$.
-$\gamma$ is thus a free parameter ($0 \leq \gamma \leq 1$) controlling the trade-off between within-class similarity and between-class dissimilarity: $\gamma = 1$ produces exclusive consideration of same-category members, and $\gamma = 0$ produces exclusive consideration of opposite-category members.
-When $\gamma = 0.5$, the similarity to contrast categories is effectively subtracted from the similarity to the target category.
+Where $f(x_j)$ is a function specifying each stored example's degree of
+contribution toward generation. Although $f(x_j)$ may be set arbitrarily, in
+PACKER it is set according to class assignment. For known members of the target
+category, $f(x_j) = \theta_t$. For members of contrast categories, $f(x_j) =
+-\theta_c$. $\theta_t$ and $\theta_c$ are free parameters ($0 \leq \theta_c, \theta_t$)
+respectively controlling the within-class similarity and between-class
+dissimilarity: $\theta_t > 0, \theta_c = 0$ produces exclusive consideration of same-category
+members, and $\theta_t = 0, \theta_c > 0$ produces exclusive consideration of opposite-category
+members. When both $\theta_t = \theta_c$, the similarity to contrast categories is
+effectively subtracted from the similarity to the target category.
 
-The probability that a given item $y$ will be generated given the model's memory $x$ is computed using relative summed similarity values across all generation candidates $y_i$:
+The probability that a given item $y$ will be generated given the model's memory
+$x$ is computed using relative summed similarity values across all generation
+candidates $y_i$:
 
-\begin{equation}
-p(y) = \dfrac
-{ \exp  \left \{ \theta \cdot a \left( y, x \right) \right \} } 
-{ \sum_i{ \exp  \left \{ \theta \cdot a \left( y_i, x \right) \right\}  } }
+\begin{equation} p(y) = \dfrac { \exp \left \{ \cdot a \left( y, x
+\right) \right \} } { \sum_i{ \exp \left \{  \cdot a \left( y_i, x \right)
+\right\} } }
 \label{eq:packer-choice}
@@
 
-Where $\theta$ ($\geq 0$) is a free parameter controlling overall response determinism.
+\subsection{An Older Version of PACKER: Didn't make it to any of the revisions, but was briefly considered}
+This model assumes that people represent categories as exemplars in a
+multidimensional space, and that generation is constrained by both similarity to
+members of the category being generated \textit{and} dissimilarity to members of
+other categories. The model assumes people generate categories that are
+dissimilar to known categories and have strong within-class similarity.
+
+The similarity between exemplars $x_i$ and $x_j$ is computed using Shepard's
+law:
+
+\begin{equation} s\left(x_i,x_j\right) = \exp \left\{ -c \left[\sum_{k}{ w_k
+\left| x_{ik} - x_{jk} \right|^r }\right]^{1/r} \right\}
+\label{eq:similarity}
+\end{equation}
+
+When prompted to make a generation decision, participants are thought to
+consider both similarity to examples from other categories as well as similarity
+to examples in the target category. More formally, the aggregated similarity $a$
+between candidate $y$ and the model's stored exemplars $x$ can be computed as:
+
+\begin{equation} a(y, x) = \sum_j{f(x_j) s(y, x_j)}
+\end{equation}
+
+Where $f(x_j)$ is a function specifying each stored example's degree of
+contribution toward generation. Although $f(x_j)$ may be set arbitrarily, in
+PACKER it is set according to class assignment. For known members of the target
+category, $f(x_j) = \gamma$. For members of contrast categories, $f(x_j) =
+\gamma - 1$. $\gamma$ is thus a free parameter ($0 \leq \gamma \leq 1$)
+controlling the trade-off between within-class similarity and between-class
+dissimilarity: $\gamma = 1$ produces exclusive consideration of same-category
+members, and $\gamma = 0$ produces exclusive consideration of opposite-category
+members. When $\gamma = 0.5$, the similarity to contrast categories is
+effectively subtracted from the similarity to the target category.
+
+The probability that a given item $y$ will be generated given the model's memory
+$x$ is computed using relative summed similarity values across all generation
+candidates $y_i$:
+
+\begin{equation} p(y) = \dfrac { \exp \left \{ \theta \cdot a \left( y, x
+\right) \right \} } { \sum_i{ \exp \left \{ \theta \cdot a \left( y_i, x \right)
+\right\} } }
+\label{eq:packer-choice}
+\end{equation}
+
+Where $\theta$ ($\geq 0$) is a free parameter controlling overall response
+determinism.
 

### 2022-11-25 Experiments/xor_cluster_row/analysis/plot.conditions.py

@@ -1,18 +1,21 @@
 import sqlite3, sys
 import numpy as np
 import pandas as pd
+import os
 
 np.set_printoptions(precision = 1)
 
 import matplotlib.pyplot as plt
-my_import_loc = '/home/jausterw/work/generating_categories/Experiments/xor_cluster_row/analysis/Imports.py'
-exec(open(my_import_loc).read())
+#set wd to where this program expects
+os.chdir(sys.path[0])
+
+exec(open('Imports.py').read())
 
 #execfile('Imports.py')
 import Modules.Funcs as funcs
 
 
-con = sqlite3.connect('/home/jausterw/work/generating_categories/Experiments/xor_cluster_row/data/experiment.db')
+con = sqlite3.connect('../data/experiment.db')
 stimuli = pd.read_sql_query("SELECT * from stimuli", con).to_numpy()
 alphas = pd.read_sql_query("SELECT * from alphas", con)
 con.close()

### 2023-01-25 Manuscripts/cbb2023/sn-article.tex
moving to cbb

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                                 %%
%% Please do not use \input{...} to include other tex files.       %%
%% Submit your LaTeX manuscript as one .tex document.              %%
%%                                                                 %%
%% All additional figures and files should be attached             %%
%% separately and not embedded in the \TeX\ document itself.       %%
%%                                                                 %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%\documentclass[referee,sn-basic]{sn-jnl}% referee option is meant for double line spacing

%%=======================================================%%
%% to print line numbers in the margin use lineno option %%
%%=======================================================%%

%%\documentclass[lineno,sn-basic]{sn-jnl}% Basic Springer Nature Reference Style/Chemistry Reference Style

%%======================================================%%
%% to compile with pdflatex/xelatex use pdflatex option %%
%%======================================================%%

%%\documentclass[pdflatex,sn-basic]{sn-jnl}% Basic Springer Nature Reference Style/Chemistry Reference Style

%%\documentclass[sn-basic]{sn-jnl}% Basic Springer Nature Reference Style/Chemistry Reference Style
%%\documentclass[sn-mathphys]{sn-jnl}% Math and Physical Sciences Reference Style
%%\documentclass[sn-aps]{sn-jnl}% American Physical Society (APS) Reference Style
%%\documentclass[sn-vancouver]{sn-jnl}% Vancouver Reference Style
\documentclass[sn-apa]{sn-jnl}% APA Reference Style
%%\documentclass[sn-chicago]{sn-jnl}% Chicago-based Humanities Reference Style
%%\documentclass[sn-standardnature]{sn-jnl}% Standard Nature Portfolio Reference Style
%%\documentclass[default]{sn-jnl}% Default
%%\documentclass[default,iicol]{sn-jnl}% Default with double column layout

%%%% Standard Packages
%%<additional latex packages if required can be included here>
%%%%

%%%%%=============================================================================%%%%
%%%%  Remarks: This template is provided to aid authors with the preparation
%%%%  of original research articles intended for submission to journals published 
%%%%  by Springer Nature. The guidance has been prepared in partnership with 
%%%%  production teams to conform to Springer Nature technical requirements. 
%%%%  Editorial and presentation requirements differ among journal portfolios and 
%%%%  research disciplines. You may find sections in this template are irrelevant 
%%%%  to your work and are empowered to omit any such section if allowed by the 
%%%%  journal you intend to submit to. The submission guidelines and policies 
%%%%  of the journal take precedence. A detailed User Manual is available in the 
%%%%  template package for technical guidance.
%%%%%=============================================================================%%%%

\jyear{2023}%

%% as per the requirement new theorem styles can be included as shown below
\theoremstyle{thmstyleone}%
\newtheorem{theorem}{Theorem}%  meant for continuous numbers
%%\newtheorem{theorem}{Theorem}[section]% meant for sectionwise numbers
%% optional argument [theorem] produces theorem numbering sequence instead of independent numbers for Proposition
\newtheorem{proposition}[theorem]{Proposition}% 
%%\newtheorem{proposition}{Proposition}% to get separate numbers for theorem and proposition etc.

\theoremstyle{thmstyletwo}%
\newtheorem{example}{Example}%
\newtheorem{remark}{Remark}%

\theoremstyle{thmstylethree}%
\newtheorem{definition}{Definition}%

\raggedbottom
%%\unnumbered% uncomment this for unnumbered level heads

%TODO: CONVERT FINAL PGFS TO STANDARD FORMAT
\pgfplotsset{compat=1.6}
\newcommand\inputpgf[2]{{
\let\pgfimageWithoutPath\pgfimage
\renewcommand{\pgfimage}[2][]{\pgfimageWithoutPath[##1]{#1/##2}}
\input{#1/#2}
}}

\begin{document}

\title[Article Title]{Article Title}

%%=============================================================%%
%% Prefix	-> \pfx{Dr}
%% GivenName	-> \fnm{Joergen W.}
%% Particle	-> \spfx{van der} -> surname prefix
%% FamilyName	-> \sur{Ploeg}
%% Suffix	-> \sfx{IV}
%% NatureName	-> \tanm{Poet Laureate} -> Title after name
%% Degrees	-> \dgr{MSc, PhD}
%% \author*[1,2]{\pfx{Dr} \fnm{Joergen W.} \spfx{van der} \sur{Ploeg} \sfx{IV} \tanm{Poet Laureate} 
%%                 \dgr{MSc, PhD}}\email{iauthor@gmail.com}
%%=============================================================%%

\author*[1,2]{\fnm{First} \sur{Author}}\email{iauthor@gmail.com}

\author[2,3]{\fnm{Second} \sur{Author}}\email{iiauthor@gmail.com}
\equalcont{These authors contributed equally to this work.}

\author[1,2]{\fnm{Third} \sur{Author}}\email{iiiauthor@gmail.com}
\equalcont{These authors contributed equally to this work.}

\affil*[1]{\orgdiv{Department}, \orgname{Organization}, \orgaddress{\street{Street}, \city{City}, \postcode{100190}, \state{State}, \country{Country}}}

\affil[2]{\orgdiv{Department}, \orgname{Organization}, \orgaddress{\street{Street}, \city{City}, \postcode{10587}, \state{State}, \country{Country}}}

\affil[3]{\orgdiv{Department}, \orgname{Organization}, \orgaddress{\street{Street}, \city{City}, \postcode{610101}, \state{State}, \country{Country}}}

%%==================================%%
%% sample for unstructured abstract %%
%%==================================%%

\abstract{The abstract serves both as a general introduction to the topic and as a brief, non-technical summary of the main results and their implications. Authors are advised to check the author instructions for the journal they are submitting to for word limits and if structural elements like subheadings, citations, or equations are permitted.}

%%================================%%
%% Sample for structured abstract %%
%%================================%%

% \abstract{\textbf{Purpose:} The abstract serves both as a general introduction to the topic and as a brief, non-technical summary of the main results and their implications. The abstract must not include subheadings (unless expressly permitted in the journal's Instructions to Authors), equations or citations. As a guide the abstract should not exceed 200 words. Most journals do not set a hard limit however authors are advised to check the author instructions for the journal they are submitting to.
% 
% \textbf{Methods:} The abstract serves both as a general introduction to the topic and as a brief, non-technical summary of the main results and their implications. The abstract must not include subheadings (unless expressly permitted in the journal's Instructions to Authors), equations or citations. As a guide the abstract should not exceed 200 words. Most journals do not set a hard limit however authors are advised to check the author instructions for the journal they are submitting to.
% 
% \textbf{Results:} The abstract serves both as a general introduction to the topic and as a brief, non-technical summary of the main results and their implications. The abstract must not include subheadings (unless expressly permitted in the journal's Instructions to Authors), equations or citations. As a guide the abstract should not exceed 200 words. Most journals do not set a hard limit however authors are advised to check the author instructions for the journal they are submitting to.
% 
% \textbf{Conclusion:} The abstract serves both as a general introduction to the topic and as a brief, non-technical summary of the main results and their implications. The abstract must not include subheadings (unless expressly permitted in the journal's Instructions to Authors), equations or citations. As a guide the abstract should not exceed 200 words. Most journals do not set a hard limit however authors are advised to check the author instructions for the journal they are submitting to.}

\keywords{keyword1, Keyword2, Keyword3, Keyword4}

%%\pacs[JEL Classification]{D8, H51}

%%\pacs[MSC Classification]{35A01, 65L10, 65L12, 65L20, 65L70}

\maketitle

\section{Introduction}\label{sec1}

The Introduction section, of referenced text \cite{bib1} expands on the background of the work (some overlap with the Abstract is acceptable). The introduction should not include subheadings.

Springer Nature does not impose a strict layout as standard however authors are advised to check the individual requirements for the journal they are planning to submit to as there may be journal-level preferences. When preparing your text please also be aware that some stylistic choices are not supported in full text XML (publication version), including coloured font. These will not be replicated in the typeset article if it is accepted.

\section{Results}\label{sec2}

Sample body text. Sample body text. Sample body text. Sample body text. Sample body text. Sample body text. Sample body text. Sample body text.

\section{This is an example for first level head---section head}\label{sec3}

\subsection{This is an example for second level head---subsection head}\label{subsec2}

\subsubsection{This is an example for third level head---subsubsection head}\label{subsubsec2}

Sample body text. Sample body text. Sample body text. Sample body text. Sample body text. Sample body text. Sample body text. Sample body text.

\section{Equations}\label{sec4}

Equations in \LaTeX\ can either be inline or on-a-line by itself (``display equations''). For
inline equations use the \verb+$...$+ commands. E.g.: The equation
$H\psi = E \psi$ is written via the command \verb+$H \psi = E \psi$+.

For display equations (with auto generated equation numbers)
one can use the equation or align environments:
\begin{equation}
\|\tilde{X}(k)\|^2 \leq\frac{\sum\limits_{i=1}^{p}\left\|\tilde{Y}_i(k)\right\|^2+\sum\limits_{j=1}^{q}\left\|\tilde{Z}_j(k)\right\|^2 }{p+q}.\label{eq1}
\end{equation}
where,
\begin{align}
D_\mu &=  \partial_\mu - ig \frac{\lambda^a}{2} A^a_\mu \nonumber \\
F^a_{\mu\nu} &= \partial_\mu A^a_\nu - \partial_\nu A^a_\mu + g f^{abc} A^b_\mu A^a_\nu \label{eq2}
\end{align}
Notice the use of \verb+\nonumber+ in the align environment at the end
of each line, except the last, so as not to produce equation numbers on
lines where no equation numbers are required. The \verb+\label{}+ command
should only be used at the last line of an align environment where
\verb+\nonumber+ is not used.
\begin{equation}
Y_\infty = \left( \frac{m}{\textrm{GeV}} \right)^{-3}
    \left[ 1 + \frac{3 \ln(m/\textrm{GeV})}{15}
    + \frac{\ln(c_2/5)}{15} \right]
\end{equation}
The class file also supports the use of \verb+\mathbb{}+, \verb+\mathscr{}+ and
\verb+\mathcal{}+ commands. As such \verb+\mathbb{R}+, \verb+\mathscr{R}+
and \verb+\mathcal{R}+ produces $\mathbb{R}$, $\mathscr{R}$ and $\mathcal{R}$
respectively (refer Subsubsection~\ref{subsubsec2}).

\section{Tables}\label{sec5}

Tables can be inserted via the normal table and tabular environment. To put
footnotes inside tables you should use \verb+\footnotetext[]{...}+ tag.
The footnote appears just below the table itself (refer Tables~\ref{tab1} and \ref{tab2}). 
For the corresponding footnotemark use \verb+\footnotemark[...]+

\begin{table}[h]
\begin{center}
\begin{minipage}{174pt}
\caption{Caption text}\label{tab1}%
\begin{tabular}{@{}llll@{}}
\toprule
Column 1 & Column 2  & Column 3 & Column 4\\
\midrule
row 1    & data 1   & data 2  & data 3  \\
row 2    & data 4   & data 5\footnotemark[1]  & data 6  \\
row 3    & data 7   & data 8  & data 9\footnotemark[2]  \\
\botrule
\end{tabular}
\footnotetext{Source: This is an example of table footnote. This is an example of table footnote.}
\footnotetext[1]{Example for a first table footnote. This is an example of table footnote.}
\footnotetext[2]{Example for a second table footnote. This is an example of table footnote.}
\end{minipage}
\end{center}
\end{table}

\noindent
The input format for the above table is as follows:

%%=============================================%%
%% For presentation purpose, we have included  %%
%% \bigskip command. please ignore this.       %%
%%=============================================%%
\bigskip
\begin{verbatim}
\begin{table}[<placement-specifier>]
\begin{center}
\begin{minipage}{<preferred-table-width>}
\caption{<table-caption>}\label{<table-label>}%
\begin{tabular}{@{}llll@{}}
\toprule
Column 1 & Column 2 & Column 3 & Column 4\\
\midrule
row 1 & data 1 & data 2	 & data 3 \\
row 2 & data 4 & data 5\footnotemark[1] & data 6 \\
row 3 & data 7 & data 8	 & data 9\footnotemark[2]\\
\botrule
\end{tabular}
\footnotetext{Source: This is an example of table footnote. 
This is an example of table footnote.}
\footnotetext[1]{Example for a first table footnote.
This is an example of table footnote.}
\footnotetext[2]{Example for a second table footnote. 
This is an example of table footnote.}
\end{minipage}
\end{center}
\end{table}
\end{verbatim}
\bigskip
%%=============================================%%
%% For presentation purpose, we have included  %%
%% \bigskip command. please ignore this.       %%
%%=============================================%%

\begin{table}[h]
\begin{center}
\begin{minipage}{\textwidth}
\caption{Example of a lengthy table which is set to full textwidth}\label{tab2}
\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}lcccccc@{\extracolsep{\fill}}}
\toprule%
& \multicolumn{3}{@{}c@{}}{Element 1\footnotemark[1]} & \multicolumn{3}{@{}c@{}}{Element 2\footnotemark[2]} \\\cmidrule{2-4}\cmidrule{5-7}%
Project & Energy & $\sigma_{calc}$ & $\sigma_{expt}$ & Energy & $\sigma_{calc}$ & $\sigma_{expt}$ \\
\midrule
Element 3  & 990 A & 1168 & $1547\pm12$ & 780 A & 1166 & $1239\pm100$\\
Element 4  & 500 A & 961  & $922\pm10$  & 900 A & 1268 & $1092\pm40$\\
\botrule
\end{tabular*}
\footnotetext{Note: This is an example of table footnote. This is an example of table footnote this is an example of table footnote this is an example of~table footnote this is an example of table footnote.}
\footnotetext[1]{Example for a first table footnote.}
\footnotetext[2]{Example for a second table footnote.}
\end{minipage}
\end{center}
\end{table}

In case of double column layout, tables which do not fit in single column width should be set to full text width. For this, you need to use \verb+\begin{table*}+ \verb+...+ \verb+\end{table*}+ instead of \verb+\begin{table}+ \verb+...+ \verb+\end{table}+ environment. Lengthy tables which do not fit in textwidth should be set as rotated table. For this, you need to use \verb+\begin{sidewaystable}+ \verb+...+ \verb+\end{sidewaystable}+ instead of \verb+\begin{table*}+ \verb+...+ \verb+\end{table*}+ environment. This environment puts tables rotated to single column width. For tables rotated to double column width, use \verb+\begin{sidewaystable*}+ \verb+...+ \verb+\end{sidewaystable*}+.

\begin{sidewaystable}
\sidewaystablefn%
\begin{center}
\begin{minipage}{\textheight}
\caption{Tables which are too long to fit, should be written using the ``sidewaystable'' environment as shown here}\label{tab3}
\begin{tabular*}{\textheight}{@{\extracolsep{\fill}}lcccccc@{\extracolsep{\fill}}}
\toprule%
& \multicolumn{3}{@{}c@{}}{Element 1\footnotemark[1]}& \multicolumn{3}{@{}c@{}}{Element\footnotemark[2]} \\\cmidrule{2-4}\cmidrule{5-7}%
Projectile & Energy	& $\sigma_{calc}$ & $\sigma_{expt}$ & Energy & $\sigma_{calc}$ & $\sigma_{expt}$ \\
\midrule
Element 3 & 990 A & 1168 & $1547\pm12$ & 780 A & 1166 & $1239\pm100$ \\
Element 4 & 500 A & 961  & $922\pm10$  & 900 A & 1268 & $1092\pm40$ \\
Element 5 & 990 A & 1168 & $1547\pm12$ & 780 A & 1166 & $1239\pm100$ \\
Element 6 & 500 A & 961  & $922\pm10$  & 900 A & 1268 & $1092\pm40$ \\
\botrule
\end{tabular*}
\footnotetext{Note: This is an example of table footnote this is an example of table footnote this is an example of table footnote this is an example of~table footnote this is an example of table footnote.}
\footnotetext[1]{This is an example of table footnote.}
\end{minipage}
\end{center}
\end{sidewaystable}

\section{Figures}\label{sec6}

As per the \LaTeX\ standards you need to use eps images for \LaTeX\ compilation and \verb+pdf/jpg/png+ images for \verb+PDFLaTeX+ compilation. This is one of the major difference between \LaTeX\ and \verb+PDFLaTeX+. Each image should be from a single input .eps/vector image file. Avoid using subfigures. The command for inserting images for \LaTeX\ and \verb+PDFLaTeX+ can be generalized. The package used to insert images in \verb+LaTeX/PDFLaTeX+ is the graphicx package. Figures can be inserted via the normal figure environment as shown in the below example:

%%=============================================%%
%% For presentation purpose, we have included  %%
%% \bigskip command. please ignore this.       %%
%%=============================================%%
\bigskip
\begin{verbatim}
\begin{figure}[<placement-specifier>]
\centering
\includegraphics{<eps-file>}
\caption{<figure-caption>}\label{<figure-label>}
\end{figure}
\end{verbatim}
\bigskip
%%=============================================%%
%% For presentation purpose, we have included  %%
%% \bigskip command. please ignore this.       %%
%%=============================================%%

\begin{figure}[h]%
\centering
\includegraphics[width=0.9\textwidth]{fig.eps}
\caption{This is a widefig. This is an example of long caption this is an example of long caption  this is an example of long caption this is an example of long caption}\label{fig1}
\end{figure}

In case of double column layout, the above format puts figure captions/images to single column width. To get spanned images, we need to provide \verb+\begin{figure*}+ \verb+...+ \verb+\end{figure*}+.

For sample purpose, we have included the width of images in the optional argument of \verb+\includegraphics+ tag. Please ignore this.

\section{Algorithms, Program codes and Listings}\label{sec7}

Packages \verb+algorithm+, \verb+algorithmicx+ and \verb+algpseudocode+ are used for setting algorithms in \LaTeX\ using the format:

%%=============================================%%
%% For presentation purpose, we have included  %%
%% \bigskip command. please ignore this.       %%
%%=============================================%%
\bigskip
\begin{verbatim}
\begin{algorithm}
\caption{<alg-caption>}\label{<alg-label>}
\begin{algorithmic}[1]
. . .
\end{algorithmic}
\end{algorithm}
\end{verbatim}
\bigskip
%%=============================================%%
%% For presentation purpose, we have included  %%
%% \bigskip command. please ignore this.       %%
%%=============================================%%

You may refer above listed package documentations for more details before setting \verb+algorithm+ environment. For program codes, the ``program'' package is required and the command to be used is \verb+\begin{program}+ \verb+...+ \verb+\end{program}+. A fast exponentiation procedure:

\begin{program}
\BEGIN \\ %
  \FOR i:=1 \TO 10 \STEP 1 \DO
     |expt|(2,i); \\ |newline|() \OD %
\rcomment{Comments will be set flush to the right margin}
\WHERE
\PROC |expt|(x,n) \BODY
          z:=1;
          \DO \IF n=0 \THEN \EXIT \FI;
             \DO \IF |odd|(n) \THEN \EXIT \FI;
\COMMENT{This is a comment statement};
                n:=n/2; x:=x*x \OD;
             \{ n>0 \};
             n:=n-1; z:=z*x \OD;
          |print|(z) \ENDPROC
\END
\end{program}

\begin{algorithm}
\caption{Calculate $y = x^n$}\label{algo1}
\begin{algorithmic}[1]
\Require $n \geq 0 \vee x \neq 0$
\Ensure $y = x^n$ 
\State $y \Leftarrow 1$
\If{$n < 0$}\label{algln2}
        \State $X \Leftarrow 1 / x$
        \State $N \Leftarrow -n$
\Else
        \State $X \Leftarrow x$
        \State $N \Leftarrow n$
\EndIf
\While{$N \neq 0$}
        \If{$N$ is even}
            \State $X \Leftarrow X \times X$
            \State $N \Leftarrow N / 2$
        \Else[$N$ is odd]
            \State $y \Leftarrow y \times X$
            \State $N \Leftarrow N - 1$
        \EndIf
\EndWhile
\end{algorithmic}
\end{algorithm}
\bigskip
%%=============================================%%
%% For presentation purpose, we have included  %%
%% \bigskip command. please ignore this.       %%
%%=============================================%%

Similarly, for \verb+listings+, use the \verb+listings+ package. \verb+\begin{lstlisting}+ \verb+...+ \verb+\end{lstlisting}+ is used to set environments similar to \verb+verbatim+ environment. Refer to the \verb+lstlisting+ package documentation for more details.

%%=============================================%%
%% For presentation purpose, we have included  %%
%% \bigskip command. please ignore this.       %%
%%=============================================%%
\bigskip
\begin{minipage}{\hsize}%
\lstset{frame=single,framexleftmargin=-1pt,framexrightmargin=-17pt,framesep=12pt,linewidth=0.98\textwidth,language=pascal}% Set your language (you can change the language for each code-block optionally)
%%% Start your code-block
\begin{lstlisting}
for i:=maxint to 0 do
begin
{ do nothing }
end;
Write('Case insensitive ');
Write('Pascal keywords.');
\end{lstlisting}
\end{minipage}

\section{Cross referencing}\label{sec8}

Environments such as figure, table, equation and align can have a label
declared via the \verb+\label{#label}+ command. For figures and table
environments use the \verb+\label{}+ command inside or just
below the \verb+\caption{}+ command. You can then use the
\verb+\ref{#label}+ command to cross-reference them. As an example, consider
the label declared for Figure~\ref{fig1} which is
\verb+\label{fig1}+. To cross-reference it, use the command 
\verb+Figure \ref{fig1}+, for which it comes up as
``Figure~\ref{fig1}''.

To reference line numbers in an algorithm, consider the label declared for the line number 2 of Algorithm~\ref{algo1} is \verb+\label{algln2}+. To cross-reference it, use the command \verb+\ref{algln2}+ for which it comes up as line~\ref{algln2} of Algorithm~\ref{algo1}.

\subsection{Details on reference citations}\label{subsec7}

Standard \LaTeX\ permits only numerical citations. To support both numerical and author-year citations this template uses \verb+natbib+ \LaTeX\ package. For style guidance please refer to the template user manual.

Here is an example for \verb+\cite{...}+: \cite{bib1}. Another example for \verb+\citep{...}+: \citep{bib2}. For author-year citation mode, \verb+\cite{...}+ prints Jones et al. (1990) and \verb+\citep{...}+ prints (Jones et al., 1990).

All cited bib entries are printed at the end of this article: \cite{bib3}, \cite{bib4}, \cite{bib5}, \cite{bib6}, \cite{bib7}, \cite{bib8}, \cite{bib9}, \cite{bib10}, \cite{bib11} and \cite{bib12}.

\section{Examples for theorem like environments}\label{sec10}

For theorem like environments, we require \verb+amsthm+ package. There are three types of predefined theorem styles exists---\verb+thmstyleone+, \verb+thmstyletwo+ and \verb+thmstylethree+

%%=============================================%%
%% For presentation purpose, we have included  %%
%% \bigskip command. please ignore this.       %%
%%=============================================%%
\bigskip
\begin{tabular}{|l|p{19pc}|}
\hline
\verb+thmstyleone+ & Numbered, theorem head in bold font and theorem text in italic style \\\hline
\verb+thmstyletwo+ & Numbered, theorem head in roman font and theorem text in italic style \\\hline
\verb+thmstylethree+ & Numbered, theorem head in bold font and theorem text in roman style \\\hline
\end{tabular}
\bigskip
%%=============================================%%
%% For presentation purpose, we have included  %%
%% \bigskip command. please ignore this.       %%
%%=============================================%%

For mathematics journals, theorem styles can be included as shown in the following examples:

\begin{theorem}[Theorem subhead]\label{thm1}
Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. 
Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. 
Example theorem text. 
\end{theorem}

Sample body text. Sample body text. Sample body text. Sample body text. Sample body text. Sample body text. Sample body text. Sample body text.

\begin{proposition}
Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text. 
Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text. 
\end{proposition}

Sample body text. Sample body text. Sample body text. Sample body text. Sample body text. Sample body text. Sample body text. Sample body text.

\begin{example}
Phasellus adipiscing semper elit. Proin fermentum massa
ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. 
\end{example}

Sample body text. Sample body text. Sample body text. Sample body text. Sample body text. Sample body text. Sample body text. Sample body text.

\begin{remark}
Phasellus adipiscing semper elit. Proin fermentum massa
ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. 
\end{remark}

Sample body text. Sample body text. Sample body text. Sample body text. Sample body text. Sample body text. Sample body text. Sample body text.

\begin{definition}[Definition sub head]
Example definition text. Example definition text. Example definition text. Example definition text. Example definition text. Example definition text. Example definition text. Example definition text. 
\end{definition}

Additionally a predefined ``proof'' environment is available: \verb+\begin{proof}+ \verb+...+ \verb+\end{proof}+. This prints a ``Proof'' head in italic font style and the ``body text'' in roman font style with an open square at the end of each proof environment.

\begin{proof}
Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. 
\end{proof}

Sample body text. Sample body text. Sample body text. Sample body text. Sample body text. Sample body text. Sample body text. Sample body text.

\begin{proof}[Proof of Theorem~{\upshape\ref{thm1}}]
Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. 
\end{proof}

\noindent
For a quote environment, use \verb+\begin{quote}...\end{quote}+
\begin{quote}
Quoted text example. Aliquam porttitor quam a lacus. Praesent vel arcu ut tortor cursus volutpat. In vitae pede quis diam bibendum placerat. Fusce elementum
convallis neque. Sed dolor orci, scelerisque ac, dapibus nec, ultricies ut, mi. Duis nec dui quis leo sagittis commodo.
\end{quote}

Sample body text. Sample body text. Sample body text. Sample body text. Sample body text (refer Figure~\ref{fig1}). Sample body text. Sample body text. Sample body text (refer Table~\ref{tab3}).

\section{Methods}\label{sec11}

Topical subheadings are allowed. Authors must ensure that their Methods section includes adequate experimental and characterization data necessary for others in the field to reproduce their work. Authors are encouraged to include RIIDs where appropriate.

\textbf{Ethical approval declarations} (only required where applicable) Any article reporting experiment/s carried out on (i)~live vertebrate (or higher invertebrates), (ii)~humans or (iii)~human samples must include an unambiguous statement within the methods section that meets the following requirements:

\begin{enumerate}[1.]
\item Approval: a statement which confirms that all experimental protocols were approved by a named institutional and/or licensing committee. Please identify the approving body in the methods section

\item Accordance: a statement explicitly saying that the methods were carried out in accordance with the relevant guidelines and regulations

\item Informed consent (for experiments involving humans or human tissue samples): include a statement confirming that informed consent was obtained from all participants and/or their legal guardian/s
\end{enumerate}

If your manuscript includes potentially identifying patient/participant information, or if it describes human transplantation research, or if it reports results of a clinical trial then  additional information will be required. Please visit (\url{https://www.nature.com/nature-research/editorial-policies}) for Nature Portfolio journals, (\url{https://www.springer.com/gp/authors-editors/journal-author/journal-author-helpdesk/publishing-ethics/14214}) for Springer Nature journals, or (\url{https://www.biomedcentral.com/getpublished/editorial-policies\#ethics+and+consent}) for BMC.

\section{Discussion}\label{sec12}

Discussions should be brief and focused. In some disciplines use of Discussion or `Conclusion' is interchangeable. It is not mandatory to use both. Some journals prefer a section `Results and Discussion' followed by a section `Conclusion'. Please refer to Journal-level guidance for any specific requirements.

\section{Conclusion}\label{sec13}

Conclusions may be used to restate your hypothesis or research question, restate your major findings, explain the relevance and the added value of your work, highlight any limitations of your study, describe future directions for research and recommendations.

In some disciplines use of Discussion or 'Conclusion' is interchangeable. It is not mandatory to use both. Please refer to Journal-level guidance for any specific requirements.

\backmatter

\bmhead{Supplementary information}

If your article has accompanying supplementary file/s please state so here.

Authors reporting data from electrophoretic gels and blots should supply the full unprocessed scans for key as part of their Supplementary information. This may be requested by the editorial team/s if it is missing.

Please refer to Journal-level guidance for any specific requirements.

\bmhead{Acknowledgments}

Acknowledgments are not compulsory. Where included they should be brief. Grant or contribution numbers may be acknowledged.

Please refer to Journal-level guidance for any specific requirements.

\section*{Declarations}

Some journals require declarations to be submitted in a standardised format. Please check the Instructions for Authors of the journal to which you are submitting to see if you need to complete this section. If yes, your manuscript must contain the following sections under the heading `Declarations':

\begin{itemize}
\item Funding
\item Conflict of interest/Competing interests (check journal-specific guidelines for which heading to use)
\item Ethics approval 
\item Consent to participate
\item Consent for publication
\item Availability of data and materials
\item Code availability 
\item Authors' contributions
\end{itemize}

\noindent
If any of the sections are not relevant to your manuscript, please include the heading and write `Not applicable' for that section.

%%===================================================%%
%% For presentation purpose, we have included        %%
%% \bigskip command. please ignore this.             %%
%%===================================================%%
\bigskip
\begin{flushleft}%
Editorial Policies for:

\bigskip\noindent
Springer journals and proceedings: \url{https://www.springer.com/gp/editorial-policies}

\bigskip\noindent
Nature Portfolio journals: \url{https://www.nature.com/nature-research/editorial-policies}

\bigskip\noindent
\textit{Scientific Reports}: \url{https://www.nature.com/srep/journal-policies/editorial-policies}

\bigskip\noindent
BMC journals: \url{https://www.biomedcentral.com/getpublished/editorial-policies}
\end{flushleft}

\begin{appendices}

\section{Section title of first appendix}\label{secA1}

An appendix contains supplementary information that is not an essential part of the text itself but which may be helpful in providing a more comprehensive understanding of the research problem or it is information that is too cumbersome to be included in the body of the paper.

%%=============================================%%
%% For submissions to Nature Portfolio Journals %%
%% please use the heading ``Extended Data''.   %%
%%=============================================%%

%%=============================================================%%
%% Sample for another appendix section			       %%
%%=============================================================%%

%% \section{Example of another appendix section}\label{secA2}%
%% Appendices may be used for helpful, supporting or essential material that would otherwise 
%% clutter, break up or be distracting to the text. Appendices can consist of sections, figures, 
%% tables and equations etc.

\end{appendices}

%%===========================================================================================%%
%% If you are submitting to one of the Nature Portfolio journals, using the eJP submission   %%
%% system, please include the references within the manuscript file itself. You may do this  %%
%% by copying the reference list from your .bbl file, paste it into the main manuscript .tex %%
%% file, and delete the associated \verb+\bibliography+ commands.                            %%
%%===========================================================================================%%

\bibliography{sn-bibliography}% common bib file
%% if required, the content of .bbl file can be included here once bbl is generated
%%\input sn-article.bbl

%% Default %%
%%\input sn-sample-bib.tex%

\end{document}

### 2023-01-25 Manuscripts/cogsci_2017/paper.tex

--- 2017-04-25 Manuscripts/cogsci-2017/paper.tex
@@
 
-PACKER sets $f(x_j)$ depending on exemplar $j$'s category membership: $f(x_j) = \gamma$ if $x_j$ is a member of the target category, and $f(x_j) = 1 - \gamma$ if $x_j$ is a member of a contrast category.
+PACKER sets $f(x_j)$ depending on exemplar $j$'s category membership: $f(x_j) = \gamma$ if $x_j$ is a member of the target category, and $f(x_j) = \gamma - 1$ if $x_j$ is a member of a contrast category.
 $\gamma$ is thus a free parameter ($0 \leq \gamma \leq 1$) controlling the trade-off between within- and between-category similarity.
@@
 To the best of our knowledge, \citet{jern2013probabilistic} were the first to evaluate computational models of generation.
-Based on their work, we describe two alternative models: a formalization of the \textit{copy-and-tweak} hypothesis \citep{ward1995s}, and the \textit{hierarchical sampling} hypothesis \citep{jern2013probabilistic}.
+Based on their work, we describe two alternative models: a formalization of the \textit{Path of Least Resistance} hypothesis (later termed {\em copy-and-tweak}, see \citealp{jern2013probabilistic}), and the \textit{hierarchical sampling} hypothesis \citep{jern2013probabilistic}.
 
 \subsection{Copy-and-Tweak}
-The \textit{copy-and-tweak} model proposes that generation is a two-part process: First, learners retrieve an observation of the target class from memory, and then they tweak it to make something new.
+The \textit{copy-and-tweak} model, based broadly on the earlier Path of Least Resistance view \citep{ward1994structured,ward1995s}, proposes that participants generate categories by retrieving an observation of the target class from memory, and then tweaking it to make something new.
 \citet{jern2013probabilistic} interpreted this proposal in terms of an exemplar model using the GCM \citep{nosofsky1984choice}.
@@
 In this case,  $f(x_j) = 1$ for members of the target category and $f(x_j) = 0$ for members of other categories; thus, the model considers only target-class similarity, and when no members of the target class are known, the model generates items at random.
-Thus, in our work we provide simulations from the copy-and-tweak account as a variant of PACKER with a fixed $\gamma$ parameter.
 
-Formalizing a model family where PACKER and copy-and-tweak are different parameterizations within the same framework is useful because comparison between the models thus provides a test of the explanatory value of the contrast mechanism: The account provided by copy-and-tweak will only equal that of PACKER if the contrast mechanism does not offer an advantage (i.e., if $\gamma < 1$ significantly improves model fits).
+In our work we provide simulations from the copy-and-tweak account, realized as a variant of PACKER with a fixed $\gamma$ parameter.
+Formalizing a model family where PACKER and copy-and-tweak are different parameterizations within the same framework is useful because comparison between the models provides a test of the explanatory value of the contrast mechanism: The account provided by copy-and-tweak will only equal that of PACKER if the contrast mechanism does not offer an advantage (i.e., if $\gamma < 1$ significantly improves model fits).
 
@@
 Each model's best-fitting parameterization is shown in Table \ref{table:model-fits}.
-Overall, PACKER outperformed copy-and-tweak and the hierarchical sampling model by a considerable margin ($\sim11\%$ improvement in log-likelihood).
+Overall, PACKER outperformed copy-and-tweak and the hierarchical sampling model by a considerable margin ($\sim13\%$ improvement in log-likelihood).
 The parameter settings associated with PACKER's best fit are exactly as expected: a strong preference for items that are similar to members of the target category but are dissimilar to members of the contrast category.

### 2024-02-16 Manuscripts/cbb2023/sn-article.tex
from overleaf

@@
 %%\documentclass[sn-vancouver]{sn-jnl}% Vancouver Reference Style
-\documentclass[sn-apa]{sn-jnl}% APA Reference Style
+\documentclass[pdflatex,sn-apa]{sn-jnl}% APA Reference Style
 %%\documentclass[sn-chicago]{sn-jnl}% Chicago-based Humanities Reference Style
@@
 %%%% Standard Packages
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{csquotes}
 %%<additional latex packages if required can be included here>
@@
 
-%TODO: CONVERT FINAL PGFS TO STANDARD FORMAT
-\pgfplotsset{compat=1.6}
-\newcommand\inputpgf[2]{{
-\let\pgfimageWithoutPath\pgfimage
-\renewcommand{\pgfimage}[2][]{\pgfimageWithoutPath[##1]{#1/##2}}
-\input{#1/#2}
-}}
-
 \begin{document}
 
-\title[Article Title]{Article Title}
+\title[RUNNING HEAD: Similarity, Contrast, and Representativeness in Categorization]{Creating Something Different: Similarity, Contrast, and Representativeness in Categorization}
 
@@
 
-\author*[1,2]{\fnm{First} \sur{Author}}\email{iauthor@gmail.com}
-
-\author[2,3]{\fnm{Second} \sur{Author}}\email{iiauthor@gmail.com}
+\author*[1]{\fnm{Joseph L.} \sur{Austerweil}}\email{austerweil@wisc.edu}
 \equalcont{These authors contributed equally to this work.}
 
-\author[1,2]{\fnm{Third} \sur{Author}}\email{iiiauthor@gmail.com}
+\author[2]{\fnm{Shi Xian} \sur{Liew}}\email{shixianliew@gmail.com}
 \equalcont{These authors contributed equally to this work.}
 
-\affil*[1]{\orgdiv{Department}, \orgname{Organization}, \orgaddress{\street{Street}, \city{City}, \postcode{100190}, \state{State}, \country{Country}}}
-
-\affil[2]{\orgdiv{Department}, \orgname{Organization}, \orgaddress{\street{Street}, \city{City}, \postcode{10587}, \state{State}, \country{Country}}}
-
-\affil[3]{\orgdiv{Department}, \orgname{Organization}, \orgaddress{\street{Street}, \city{City}, \postcode{610101}, \state{State}, \country{Country}}}
+\author[1]{\fnm{Nolan} \sur{Conaway}}\email{nolanbconaway@gmail.com}
+
+\author[3]{\fnm{Kenneth J.} \sur{Kurtz}}\email{kkurtz@binghamton.edu}
+
+\affil*[1]{\orgdiv{Department of Psychology}, \orgname{University of Wisconsin-Madison}, \orgaddress{\street{1202 West Johnson Street}, \city{Madison}, \postcode{53706}, \state{WI}, \country{USA}}}
+
+\affil[2]{\orgdiv{School of Psychology}, \orgname{University of South New Wales}, \orgaddress{\street{Mathews Building}, \city{Sydney}, \postcode{2052}, \state{NSW}, \country{Australia}}}
+
+\affil[3]{\orgdiv{Department of Psychology }\orgname{SUNY Binghamton} \orgaddress{\city{Binghamton}, \postcode{{12042}}, \state{NY} \country{USA}}}
 
@@
 
-\abstract{The abstract serves both as a general introduction to the topic and as a brief, non-technical summary of the main results and their implications.
-Authors are advised to check the author instructions for the journal they are submitting to for word limits and if structural elements like subheadings, citations, or equations are permitted.}
+\abstract{The ability to generate new concepts and ideas is among the most fascinating aspects of human cognition, but we do not have a strong understanding of the
+cognitive processes and representations underlying concept generation. In this
+paper, we study the generation of new categories using the computational and
+behavioral toolkit of traditional artificial category learning. Previous work in
+this domain has focused on how the statistical structure of known categories
+generalizes to generated categories, overlooking whether (and if so, how)
+contrast between the known and generated categories is a factor. We report three
+experiments demonstrating that contrast between what is known and what is
+created is of fundamental importance for categorization. We propose two novel
+approaches to modeling category contrast: one focused on exemplar dissimilarity
+and another on the representativeness heuristic. Our experiments and
+computational analyses demonstrate that both models capture different aspects of
+contrast's role in categorization.}
 
@@
 
-\keywords{keyword1, Keyword2, Keyword3, Keyword4}
+\keywords{categorization, concepts, category learning, generation,
+computational modeling, exemplar models, representativeness}
 
@@
 
+\section{Declarations Section}
+\subsection{Funding}
+The studies were funded by the Wisconsin Alumni Research Foundation through University of Wisconsin-Madison Office of the Vice Chancellor for Graduate Education.
+
+\subsection{Conflict of Interests}
+No authors have any conflicts of interests to declare.
+
+\subsection{Availability of Data,  Materials, and Code }
+All experiment materials, anonymized raw data, and code are available at \url{https://github.com/AusterweilLab/generating_categories}.
+
+\subsection{Authors Contributions}
+The first draft of the manuscript was primarily written by JLA and SXL.
+All authors contributed to writing the manuscript substantially by editing and/or commenting on the manuscript.
+The computational models, and Experiments 1 and 2 were designed by JLA, NC, and KJK.
+Experiments 1 and 2 were implemented and conducted by NC and JLA.
+Experiment 3 was designed by JLA, SXL, and KJK.
+It was implemented and conducted by SXL and JLA.
+Model-based and other analyses were done by NC, SXL, and JLA.
+
+\subsection{Ethical Approval}
+All human subject experiments were originally approved by the Social Behavioral and Education Institutional Review Board (now named the Minimal Risk Research Institutional Review Board) at the University of Wisconsin-Madison (protocol number 2016-1011).
+
+\subsection{Consent to Participate}
+All participants were provided informed consent before participating.
+Given the minimal risk provided by the study, it was determined that an unintentional breach of confidentiality would be the most likely potential risk and so, no identifying information was stored.
+
+\subsection{Consent to Publish}
+Not Applicable.
+
+\newpage
+
 \section{Introduction}\label{sec1}
-
-The Introduction section, of referenced text \cite{bib1} expands on the background of the work (some overlap with the Abstract is acceptable). The introduction should not include subheadings.
-
-Springer Nature does not impose a strict layout as standard however authors are advised to check the individual requirements for the journal they are planning to submit to as there may be journal-level preferences.
-When preparing your text please also be aware that some stylistic choices are not supported in full text XML (publication version), including coloured font.
-These will not be replicated in the typeset article if it is accepted.
-
-\section{Results}\label{sec2}
-
-Sample body text. Sample body text. Sample body text. Sample body text. Sample body text. Sample body text. Sample body text. Sample body text.
-
-\section{This is an example for first level head---section head}\label{sec3}
-
-\subsection{This is an example for second level head---subsection head}\label{subsec2}
-
-\subsubsection{This is an example for third level head---subsubsection head}\label{subsubsec2}
-
-Sample body text. Sample body text. Sample body text. Sample body text. Sample body text. Sample body text. Sample body text. Sample body text.
-
-\section{Equations}\label{sec4}
-
-Equations in \LaTeX\ can either be inline or on-a-line by itself (``display equations''). For
-inline equations use the \verb+$...$+ commands. E.g.: The equation
-$H\psi = E \psi$ is written via the command \verb+$H \psi = E \psi$+.
-
-For display equations (with auto generated equation numbers)
-one can use the equation or align environments:
+Creating new ideas is one of the most fascinating and important
+human capabilities. For example, cell phones and smart phones are newly created
+categories of objects that young and old adults learned and have become reliant
+on in the last few decades. Computational researchers have identified
+generating novel objects and objectives as one of the defining, and most
+difficult to formalize, characteristics of intelligent life
+\citep{lake2017,lehman2011,taylor2016}.
+Recently, generative deep learning methods, such as stable diffusion \cite{rombach2022} and ChatGPT \cite{chatgpt2022}, have captivated researchers and the overall public.
+Their power to generate novel images and text is astonishing.
+They have been reported to win an art contest \cite{roose2022}, and to pass a major graduate medical and business school examinations \cite{kung2022,terwiesch2023}.
+What are the major principles underlying category generation?
+
+Despite recent successes in generation from machine learning methods, the cognitive mechanisms that enable people to innovate are not well understood and
+are understudied by scientists. In part, this is due to the inherent difficulty
+of designing and conducting experiments that test these capabilities. How would
+a scientist devise an experiment that induces a participant to create 
+new categories that are as interesting as cell phones and smart phones?
+Devising such an experiment is still beyond our capability.
+Instead, we take a step towards this goal, and investigate the fundamental constraints and expectations that people have when they generate new categories.
+
+% we do have tools to study the fundamental constraints on category . As a
+%step towards these goals, in this article we investigate the expectations that
+%people have when they generate new categories using computational
+%models and behavioral experiments.
+
+Creating new concepts are, however, not altogether different from
+the types of behaviors typically studied in cognitive psychology laboratories.
+In particular, generating a member of a novel class can be considered a `special
+case' application of existing category knowledge
+\citep{kemp2014taxonomy,kurtz2015human}. Research in categorization typically
+focuses on what properties of a category affect human learning of an object's
+category given its features \citep{shepard1961learning,kurtz2013human}, or the
+prediction of an object's unobserved features given some of its other features
+and/or its category \citep{markman2003category}. The generation of members of a
+new category consists of inferring {\em all} features for a {\em novel} category
+label. Thus, we can make progress formalizing the processes involved in category
+generation by extending theories of categorization to this case. This will also serve to test the breadth of the explanatory power of categorization models to tasks beyond typical category learning.
+
+Previous work in category learning and generation has established that people
+are highly sensitive to the structural properties of categories, such as
+correlations between the features of category members and the relation between
+items within the same category and those in different categories
+\citep{roschmervis1975,regier2007,shepard1961learning,smith1995creative}.\footnote{It
+  is worth noting that sensitivity to feature correlations is not universal and
+  can be dependent on factors such as the salience of correlations and task
+  demands \citep[e.g., see][]{chin2002effect,malt1984correlated,murphy2004big}.}
+Inspired by this work, previous research on the topic of category generation has
+explored a similar principle: People tend to create new categories that have
+similar {\em statistical regularities} as previously learned categories
+\citep{jern2013probabilistic,ward1994structured}. These statistical regularities
+reflect general structural patterns across old and new categories and are not
+restricted to strictly quantitatively similar properties.
+
+Although statistical regularity is an important characteristic of generating new
+categories, it cannot be the only one. Taken to the extreme, the best ``new''
+category in terms of having the same statistical regularities to other
+categories would be identical to a known category that is representative of the
+domain (and thus, not new at all). Contrast from other known categories should
+play a role. Indeed, scientists in other fields, such as marketing
+\citep{berger16} and sociology \citep{rogers03}, highlight the critical
+importance of contrast in the creation of new ideas.
+
+To successfully generate something novel, what is generated must be different
+from what is already known. This fundamental constraint, ``being different'', or
+contrasting from other categories in the relevant domain, is the focus of our
+work. Intuitively, the influence of contrast may be expected in a number of
+real-life scenarios involving the generation of new categories. For instance, a
+college student who has been eating instant noodles for a week may be more
+inclined to seek out something healthier and less carbohydrate-heavy (e.g., a salad) compared to something similar (a different flavor of instant
+noodles). An actor who is dissatisfied with being typecast as a villain may seek
+out roles as heroes as opposed to playing other antagonistic characters. An
+especially compelling example can be seen in \cite{askin2017} who found that new
+songs that were more different to their peers were generally more popular than
+new songs that were similar to others. It appears that once there is the
+motivation to generate something new, there is an associated motivation to
+generate and observe something \emph{different}.
+
+Although implicitly assumed in some work, this constraint has been
+overlooked in previous research: To our knowledge, there has not been any
+systematic investigation addressing how generated categories {\em differ} from
+what is already known. Although the idea of category contrast is discussed
+throughout the categorization literature, and extends to a variety of other
+fields \citep[e.g., color;][]{regier2007}, the idea that a new category should
+be ``different'' is vague, as there are many ways it could be different from a
+previously observed category.
+
+We examine two different definitions of how categories should differ from one
+another: exemplar dissimilarity and representativeness of the alternative
+category. To formalize the first technique, we build on the largely successful
+exemplar modeling framework
+\citep{medin1978context,nosofsky1984choice,nosofsky1986attention}. In doing so,
+we propose a novel exemplar model of category generation, {\em Producing Alike
+  and Contrasting Knowledge using Exemplar Representations} (PACKER),
+formalizing how new categories should differ from previous categories. It
+embodies the exemplar dissimilarity principle by incorporating a factor that
+repulses members of opposite categories from one another.
+
+The second hypothesis for how contrast might affect categorization is contrast
+as {\em representativeness}. The representativeness heuristic
+\citep{kahneman1972subjective} states that people make judgments regarding an
+outcome based on how representative it is of the evidence given in the current
+context. The heuristic is a powerful theoretical construct that has been used to
+capture a range of complex patterns of human judgments
+\citep{kahneman73,tversky74,tversky83}, especially those that deviate from
+normative theory (as given by a straightforward application of Bayes' rule). For
+example, the coinflip sequence TTHTH is perceived to be more random (due to it
+being representative of a coinflip sequence) than TTTTT (despite both having the
+same probability of having been generated from a fair coin). Although there is a
+healthy debate as to whether perceived randomness {\em really} deviates from
+normative theory \citep{griffiths18,hahn09}, the representativeness heuristic
+remains one of the main explanations of human judgments \citep{reimers18}. To
+formalize the hypothesis that people expect a new category to be representative
+of the opposite of the current category, or contrast as {\em
+  representativeness}, we use the Bayesian formulation of representativeness
+\citep{tenenbaum01}, which has been used successfully to capture human
+performance in color categorization \citep{abbott16}, image category learning
+\citep{abbott11}, and language grammar learning \citep{rafferty10}. We do so by
+extending the state-of-the-art category generation model
+\citep{jern2013probabilistic} and find that human category generation can
+also be captured as generating representative, rather than probable, samples.
+
+The outline of the article is as follows. First we describe previous
+computational formalizations of theories of category generation and empirical
+work investigating them. We then present two hypotheses for how contrast might
+affect categorization and formalize them in computational models. The first is a
+novel exemplar model, which is designed to generate categories that
+systematically differ from existing categories in the domain. The second assumes
+that the goal of category generation is to create representative samples of the
+opposite of the observed categories. We present two experiments demonstrating
+strong and systematic effects of category contrast on concept generation, and we
+qualitatively and quantitatively analyze the behavioral results and the performance of each model in
+capturing human category generation. A third experiment is performed to
+highlight the differences between the two contrast models. We conclude with a
+discussion of the implications of our results for categorization and directions
+for future work.
+
+\section{Prior work}
+
+Much of what we know about concept generation and contrast comes from the
+foundational literature on creative cognition. In a series of reports, Ward and
+colleagues
+\citep{ward1995s,ward1994structured,marsh1999inadvertent,ward2002role,smith1993constraining}
+established that category generation is highly constrained by prior knowledge:
+Generated categories tend to consist of features observed in known categories,
+and they tend to exhibit the distributional properties found in known
+categories. In a seminal study, \cite{ward1994structured} asked participants to
+generate new species of alien animals by drawing and describing members of the
+species. People tended to generate species with the same features as on Earth
+(e.g., eyes, legs, wings), and possessing the same feature correlations as on
+Earth (e.g., feathers co-occur with wings). Likewise, aliens drawn from the same
+species tended to share more features with one another compared to members of
+opposite species.
+
+The broader set of observations made by Ward and colleagues provide a great deal
+of insight into the role of prior knowledge in constraining category generation.
+Much of the work from this area
+\citep[e.g.,][]{smith1993constraining,marsh1999inadvertent} focuses on how
+information provided to participants (such as an example of a species generated
+by other participants) can drastically diminish the difference of a new category
+from pre-existing categories. Theoretical accounts of these effects have
+primarily been grounded within the categorization literature. For example, the
+predominant ``Path of Least Resistance'' account
+\citep[see][]{ward1994structured,ward1995s,ward2002role} proposes that, when
+generating a new species of animal, people retrieve from memory a known
+subcategory of animals (e.g., {\em bird}, {\em dog}, {\em horse}), and simply
+change some of the features to make something new. People are thought to change
+only features that are not characteristic of the retrieved category (e.g., if
+{\em bird} was retrieved, the presence of {\em wings} would not change, but {\em color} might). This theory incorporates elements of the highly influential
+basic-level categories framework \citep{rosch1975cognitive,rosch1976basic}, as
+well as the exemplar view \citep{medin1978context,brooks1978nonanalytic}. While
+this work has been incredibly useful in providing a conceptual sketch of
+generation theories, its qualitative nature and the hand-drawn responses used in
+the experiments paradigms precludes the development of formal approaches that
+can be used to test them in a fine-grained manner.
+One other relevant study of note is \cite{lake2015}.
+In one part of their experiments, they had participants generate novel alphabets.
+While their work is fascinating, the model is domain-specific (representing orthography using a grammar over the parameters of B\'{e}zier curves) and the results are difficult to generalize beyond that domain.
+
+\cite{jern2013probabilistic} showed that concept generation could be
+studied in a more controlled manner through the well-developed methods of an
+artificial categorization paradigm \citep[see][for a review]{kurtz2015human}. In
+Experiments 3 and 4 of their article, participants were exposed to members of
+experimenter-defined categories of "crystals" varying in size, hue, and
+saturation. Following a training phase during which the experimenter-defined
+categories were learned, participants were asked to generate novel categories of
+crystals. In a finding mirroring \cite{ward1994structured},
+\cite{jern2013probabilistic} found that participants generated categories with
+the same distributional properties as the experimenter-defined categories. For
+example, after learning categories with a positive correlation between the size
+and saturation features (larger sized crystals were more saturated),
+participants generated novel categories with the same positive correlation. By
+replicating \cite{ward1994structured}, they demonstrated that category generation
+can be studied in a well-known and highly controlled experimental paradigm.
+
+The authors evaluated the predictions of several formal models on their data.
+Most notably, they showed that a hierarchical Bayesian model provided the
+strongest account of their results. Their model views observed examples as
+samples from an underlying category distribution, describing the location of the
+category in the space, as well as how it varies along each feature. In turn,
+each category is viewed as a sample from an underlying {\em domain}
+distribution, specifying distributional commonalities among the observed
+categories. Generated categories are thought to stem from the same domain
+distribution as observed categories, thus the distributional properties of
+observed categories will be preserved within the generated category.
+
+\cite{jern2013probabilistic} additionally tested a ``copy-and-tweak'' model that
+broadly resembles the earlier ``Path of Least Resistance'' account. The core
+proposal is that participants generate new items by copying stored examples from
+memory and tweaking them to generate something new. The copy-and-tweak model
+differs from the Path of Least Resistance account in that it notably omits the
+hierarchical organization of categories, as well as selectivity to which
+features are changed \citep[both of which are factors in the Path of Least
+Resistance account;][]{ward2002role}. Instead, their copy-and-tweak model
+corresponds to a direct exemplar-similarity approach
+\citep[e.g.,][]{nosofsky1984choice,nosofsky1986attention}, generating new items
+according to their similarity to known members of the target category. The
+copy-and-tweak model provided a poor account of the results of \cite{jern2013probabilistic}.
+
+It is worth pointing out that the early observation of regularities in
+distributional properties across categories is not confined to the work of Ward
+and colleagues. Most notably, \citet{thomas98} trained participants to classify
+circles of different sizes, each with a radial line of varying orientation, into
+two categories. During a surprise prediction phase, participants were asked to
+generate a value for a missing feature for a certain category exemplar given a
+value on the other feature. Their results revealed that exemplars tended to
+share the same feature correlations as exemplars from previously learned
+categories. Interestingly, this effect was not consistently observed when the
+given values of a feature fell outside the range of learned values for that
+particular category. For instance, if the learned Category A exemplars were
+generally 8 cm to 10 cm in radius and negatively correlated with the angle of
+the radial line, participants generally produced radial line orientations that
+were also negatively correlated with the exemplar size when the given size was
+between 8 cm and 10 cm. However, when the given size fell outside that range for
+that category (e.g., 14 cm for a Category A exemplar), some participants
+consistently produced radial line orientations that were not negatively
+correlated with the size and that also placed the exemplars in a significantly
+different location in the feature space. These results suggest that when tasked
+to produce exemplars that fall outside what was previously learned, people show
+some tendency to generate exemplars that are different (both distributionally
+and spatially) to what was learned. We explore this idea of contrast
+more deeply in the following section.
+
+\section{And Now for Something (Completely) Different: The Role of Contrast}
+
+%Prior work on category generation has explored only one factor: distributional
+%correspondences between learned and generated categories. As a result, most of
+%the computational, theoretical, and empirical efforts have been focused on
+%explaining those effects. In this paper, we investigate another important
+%factor: category contrast. To generate a novel concept, individuals must produce
+%something that is in some capacity {\em different} from what they already know.
+%As a consequence, we propose that contrast should be a primary constraint that
+%underlies people's expectations about how a set of categories are organized:
+%Categories should be different from each other.
+
+Although people are {\em capable} of creating 
+new categories, it is not entirely clear whether (and if so, how) new concepts are systematically
+made different from what is already known. The hierarchical Bayesian model
+developed by \cite{jern2013probabilistic} assumes that differences between
+observed and generated categories are only due to random variation. The model
+assumes that generated categories are sampled from the same underlying domain
+distribution as observed categories, and will thus share a common distributional
+structure. The model does not make predictions about the {\em location} of the
+category within the domain (the perceptual instantiation of category members).
+Under a strict interpretation of their model, given knowledge of a single
+category within the domain, the most probable new category to be generated is
+located in {\em exactly} the same location and possesses an identical
+distributional structure. This is not an issue with their model specifically,
+but using a broader class of standard hierarchical Bayesian models without any
+additional features \citep[e.g., ][]{griffithsscn07,kemp07}. Many of these
+models assume that at some point of the latent generative process the same
+underlying distribution generates all of the categories and thus, any
+differences between categories are due to {\em noise} and should not be {\em
+  systematic}. If the goal of creating a new category given others in a domain
+is generating from the posterior predictive distribution then the best a
+standard hierarchical Bayesian model without a notion of contrast built into the
+model can do at capturing contrast is to assume that the new category is placed
+uniformly at random over feature space. But, this defeats the purpose of a
+hierarchy as it is ignored when determining a new category location!\footnote{It
+  is plausible that some hierarchical Bayesian models could be created that
+  generate categories different from each other. However, this model would not
+  be a standard application or extension of most pre-existing hierarchical
+  Bayesian models. The generative process would need to include a component that
+  presumes contrast, which is precisely the factor of study in this article.} Note that this does not mean location information and
+contrast cannot be captured using hierarchical Bayesian models \citep[e.g., see][]{sanborn2021}. In fact, a model
+instantiating contrast as representativeness is a hierarchical Bayesian model.
+But, it requires an additional factor, which in our case is having a different
+goal than predicting exemplars from a new category.
+
+The copy-and-tweak model tested by \cite{jern2013probabilistic} also claims
+little about how generated categories should contrast with what is already
+known. In their simulations, the model was only tested on generation after the
+learner had been exposed to members of the target category, and so the model's
+ability to generate a new category from scratch was not evaluated. However, the
+model's generation is based exclusively on similarity to known members of the
+{\em target} category; when there are no members of the target category,
+generation is presumably random.
+
+To our knowledge, no prior literature has yet explicitly discussed any empirical
+effects of contrast in category generation. However, some evidence of contrast
+can be seen in the data from Experiment 3 of \cite{jern2013probabilistic}. We
+leave a detailed analysis of these data to the supplemental materials.
+
+\subsection{Contrast as exemplar dissimilarity: The PACKER Model}
+
+\label{section:PACKER-definition}
+
+As noted above, the constraint that new concepts should differ from what is
+already known has been largely overlooked in previous work. This is no doubt in
+part due to the vague definition of what it means for a concept to be
+``different'': A generated category may be different from what is already known
+in any number of respects. Towards providing a more precise definition of the
+role of contrast in generation, we formalized contrast in a novel exemplar
+model, PACKER ({\em Producing Alike and Contrasting Knowledge using Exemplar
+  Representations}). PACKER explains category generation as a balance between
+two fundamental constraints: The exemplars of the category to be generated
+should not be similar to known categories, and exemplars within each category
+should be similar to one another. These ideas are implemented within the
+well-studied exemplar framework -- the PACKER model is an extension of the
+influential Generalized Context Model of categorization
+\citep[GCM;][]{nosofsky1984choice,nosofsky1986attention}.
+
+Although, as an exemplar model, one of PACKER's proposals is people represent
+categories in terms of a collection of stored exemplars, we did not pick it
+assuming it is the correct model of human categorization. The choice to develop
+PACKER within an exemplar framework reflects the facts that exemplar models have
+been thoroughly evaluated, are strongly theoretically motivated, and dominate
+much of the theoretical and empirical work in categorization. The focus of our
+work with PACKER concerns how contrast may be captured through the dual constraints of within- and between-class
+similarity; it is not difficult to imagine how such constraints may be
+instantiated using alternative frameworks (e.g.,
+\citealp{kurtz2007divergent,love2004sustain,smith2000thirty}; for a review of
+categorization models, see \citealp{pothoswills2011}).
+
+Both PACKER and the GCM simulate categorization under the assumption that
+learners represent categories as a collection of exemplars, corresponding to the
+labeled stimuli they have observed. The exemplars are encoded within a
+$k$-dimensional psychological space, and model performance is based on the
+amount of similarity between the item to be categorized and the stored
+exemplars. Similarity between two examples, $s\left(x_i, x_j\right)$, is
+computed as an inverse exponential function of distance
+\citep[following][]{attneave1950,shepard1957stimulus,shepard1987toward}:
+
 \begin{equation}
-\|\tilde{X}(k)\|^2 \leq\frac{\sum\limits_{i=1}^{p}\left\|\tilde{Y}_i(k)\right\|^2+\sum\limits_{j=1}^{q}\left\|\tilde{Z}_j(k)\right\|^2 }{p+q}.\label{eq1}
+s \left(x_i,x_j \right) = \exp \left\{ -c \left[ \sum_{k}{w_k \lvert x_{ik}-x_{jk} \rvert ^r }  \right]^{1/r}  \right\}
+\label{eq:similarity}
 \end{equation}
-where,
+
+% \begin{equation} 
+% s\left( x_i,x_j\right) = \exp \left\{ -c \left[\sum_{k}{ w_k\left| x_{ik} - x_{jk}\right|^r }\right]^{1/r} \right\}
+% \label{eq:similarity}
+% \end{equation}
+% 
+\noindent where $w_k$ is the attention weighting of dimension $k$ ($w_k \geq 0$ and
+$\sum_k{w_k} = 1$), accounting for the relative importance of each dimension in
+similarity calculations, and $c$ ($c>0$) is a specificity parameter controlling
+the spread of exemplar generalization. For simplicity, in our simulations
+attention will be distributed across each dimension uniformly (unless otherwise
+noted). The value of $r$ depends on the nature of the experimental conditions
+being simulated: $r=1$ is appropriate for separable dimensions, whereas $r=2$ is
+appropriate for integral dimensions
+\citep[e.g.,][]{shepard1964attention,garner1974processing}. In our simulations,
+we set $r=1$ due to the separable nature of the stimulus dimensions used in our
+experiments (see Figure \ref{fig:e1-conditions}).
+
+PACKER (as well as its name) was in part inspired by earlier work from the
+categorization literature \citep{hidaka2011packing,stewart2005}. \citet{hidaka2011packing} argued that
+natural categories ``pack'' the values of features such that different
+categories fill the domain space with distance between one another, while
+keeping items within the same category close together. Inspired by this idea,
+PACKER proposes that generation is constrained by both similarity to members of
+the target category (the category in which a stimulus is being generated) as
+well as similarity to members of other categories: the most desirable generation
+candidates are similar to members of the target category and not similar to
+members of contrast categories. This is achieved by aggregating similarity
+across known exemplars differently according to class membership. The aggregated
+similarity $a(y,x)$ between generation candidate $y$ and stored exemplars $x$ is
+given by:
+
+\begin{equation} a(y, x) = \sum_j{f(x_j) s(y, x_j)}
+\end{equation}
+% 
+where $f(x_j)$ is a function specifying each exemplar's contribution to
+generation. A negative value for $f(x_j)$ produces a `repelling' effect (items
+are less likely to be generated nearby $x_j$), and a positive value produces an
+`attracting' effect (items are more likely to be generated nearby $x_j$). When
+$f(x_j)=0$, the exemplar does not contribute to generation.
+
+PACKER sets $f(x_j)$ depending on exemplar $x_j$'s category membership:
+$f(x_j) = \theta_t$ if $x_j$ is a member of the target category, and
+$f(x_j) = -\theta_c$ if $x_j$ is a member of a contrast category. $\theta_t$ and
+$\theta_c$ are free parameters ($0 \leq \theta_t,\theta_c$) controlling the
+trade-off between within- and between-category similarity. For example, when
+$\theta_t = \theta_c = 0.5$, $f(x_j) = 0.5$ for members of the target category
+and $f(x_j) = -0.5$ for members of other categories; thus, the model is likely
+to generate items that are similar to members of the target category but are not
+similar to members of other categories. In this way, $\theta_t > 0$ with
+$\theta_c = 0$ produces exclusive consideration of target-category members, and
+$\theta_c > 0$ with $\theta_t = 0$ produces exclusive consideration of
+contrast-category members. The combination of $\theta_t$ and $\theta_c$
+parameters thus specifies a wide breadth of possible approaches; by fitting it
+to a dataset, one can describe the relative roles of between-category contrast
+and within-category similarity in generation. See Figure
+\ref{fig:packer-examples} for an illustration of how these parameters control
+the relative influence of within-category similarity and contrast to other
+categories when generating a new exemplar.
+
+\begin{figure}
+	\begin{center} 
+    \includegraphics[width=0.95\textwidth]{fig1.eps}
+      \caption{The effect of weights on PACKER generation of a category `B' example, following exposure
+        to one member of category `A' and one member of category `B'.
+        Predictions are shown for three different parameterizations (differing
+        in values of $\theta_t$ and $\theta_c$): {\em (a)}
+        Predictions based on contrast similarity only. {\em (b)}
+        Predictions based on target similarity only. {\em (c)} Predictions with
+        both constraints considered.}
+		\label{fig:packer-examples}
+	\end{center}
+\end{figure}
+
+The probability that a candidate $y$ will be generated is evaluated using an
+Exponentiated \citet{luce1977choice} choice rule. Candidates with greater values
+of $a(y,x)$ are more likely to be generated than candidates with smaller values:
+% 
+\begin{equation} p(y \mid x) = \dfrac { \exp \left \{ a \left( y, x
+\right) \right \} } { \sum_i{ \exp \left \{ a \left( y_i, x \right)
+\right\} } }
+\label{eq:packer-choice}
+\end{equation}
+
+Note, because $\theta_t$ and $\theta_c$ are unconstrained, their size plays the
+role of the response determinism parameter for the exponentiated Luce choice
+rule.
+
+\subsubsection{Relation Between PACKER and Copy-And-Tweak}
+\label{section:copytweak-packer}
+
+It is worth noting that PACKER is only one possible exemplar-based account of
+category generation within our proposed framework. That is, PACKER places
+specific constraints on the possible values of $f(x_j)$, but other
+exemplar-based category generation models with drastically different behavior
+can be formalized in this framework by imposing alternative constraints. For
+example, PACKER is formally equivalent to the copy-and-tweak model proposed by
+\cite{jern2013probabilistic} when $\theta_c = 0$ and $\theta_t > 0$ (in fact, it
+becomes a continuous-dimension adaptation of the original model). Likewise,
+when $\theta_t = 0$ and $\theta_c > 0$, PACKER can represent a contrast-only
+generation mode, relying exclusively on contrast when generating new categories.
+When $f(x_j) < 0$ for all $x_j$ (regardless of class membership), a
+``pure-packing'' approach is yielded, generating items in unoccupied areas of
+the domain. Thus, the proposed framework may be used to describe a wide variety
+of qualitatively distinct generation strategies.
+
+By formalizing a model family where PACKER and copy-and-tweak are different
+parameterizations of models within the same framework, the comparison between
+PACKER and copy-and-tweak provides a test of the explanatory value of the
+contrast mechanism based on exemplar dissimilarity: The account provided by
+copy-and-tweak will only equal that of PACKER if the contrast mechanism does not
+offer an advantage (i.e., if $\theta_c > 0$ significantly improves model fits).
+Note that the purpose of the article is to explore and formally analyze the role
+of contrast in categorization and thus, we leave extending PACKER to incorporate
+distributional factors for future work.
+
+\subsection{Contrast as representativeness: An extension to Jern and Kemp
+(2013)}
+\label{section:representativeness-definition}
+
+An alternate conceptualization of category contrast is the idea of
+representativeness -- exemplars are generated such that they resemble the target
+category and are thus more distinct from and less similar to other categories.
+The general idea of representativeness is not new in categorization. For
+instance, the literature on the \emph{graded structure} of categories explores
+how some exemplars are better exemplars of their categories
+\citep{palmeri2001central,barsalou1985ideals}. In this article, we adopt the
+specific formalization of representativeness given by \cite{tenenbaum01}, where
+the representativeness $R(y,h)$ of an item $y$ (or in our case, an exemplar) is
+the relative amount of evidence that is provided by $y$ for a given hypothesis
+$h$ in a space of hypotheses $\mathcal{H}$ in contrast to all other hypotheses
+$\mathcal{H}^c=\mathcal{H} \setminus \{h\}$:
+% \begin{equation}
+% R(x,h) = \log \frac
+% \end{equation}
+\begin{equation}
+R(x,h) = \log \frac{p(x \mid h)}{1-p(x \mid \mathcal{H}^c)} = \log \frac{p(x \mid h)}{\sum_{h'\in \mathcal{H}^c}{p(x \mid h')p(h')}} .
+\label{eq:representativeness}
+\end{equation}
+In Equation \ref{eq:representativeness}, $p(h')$ is the prior probability
+distribution on the hypothesis space that excludes $h$ (i.e., we effectively set
+$p(h)$ to 0 and then renormalize the priors). Note that when there are only two
+hypotheses (e.g. two categories) involved in the domain, this prior takes the
+value of $1$ since there is ever only one alternative hypothesis.
+
+The Bayesian formalization of representativeness makes it straightforward to
+extend the existing hierarchical Bayesian model of category generation
+exemplars are sampled from a given category distribution ($h$ in
+Equation \ref{eq:representativeness}). Each category distribution is a
+multivariate normal parameterized by a location vector $\mu_k$ and covariance
+matrix $\Sigma_k$ (i.e., $h = \mathcal{N}(\mu_k,\Sigma_k)$). The parameters
+$\mu_k$ and $\Sigma_k$ are assumed to be samples from a prior
+normal-inverse-Wishart (NIW) distribution parameterized by
+$\mu_0,\Sigma_0,\kappa,\nu$. Mathematically,
 \begin{align}
-D_\mu &=  \partial_\mu - ig \frac{\lambda^a}{2} A^a_\mu \nonumber \\
-F^a_{\mu\nu} &= \partial_\mu A^a_\nu - \partial_\nu A^a_\mu + g f^{abc} A^b_\mu A^a_\nu \label{eq2}
+  \mu_k, \Sigma_k \mid \mu_0,\Sigma_0,\kappa,\nu &\sim
+                                              \textrm{NIW}(\mu_0,\Sigma_0,\kappa,\nu)
+                                              \label{eq:mvnsample}\\
+  y \mid C &\sim \mathcal{N}(\mu_k,\Sigma_k) \label{eq:exemplarsample}
 \end{align}
-Notice the use of \verb+\nonumber+ in the align environment at the end
-of each line, except the last, so as not to produce equation numbers on
-lines where no equation numbers are required. The \verb+\label{}+ command
-should only be used at the last line of an align environment where
-\verb+\nonumber+ is not used.
+
+In our simulations, we set the prior mean $\mu_0$ to the center of the feature
+space and the prior variance to be isotropic ($\Sigma_0 = \lambda \mathbf{I}$ where
+$\lambda$ is a free parameter and $\mathbf{I}$ is a $d$-by-$d$ identity matrix
+with $d$ representing the number of dimensions or features in the domain). The
+$\kappa$ and $\nu$ parameters are freely estimated within the constraints
+$\kappa > d-1$ and $\nu > 0$.
+
+With the assumption of the NIW prior, the expected location vector $\mu_k$ given exemplars observed from category $k$ is:
 \begin{equation}
-Y_\infty = \left( \frac{m}{\textrm{GeV}} \right)^{-3}
-    \left[ 1 + \frac{3 \ln(m/\textrm{GeV})}{15}
-    + \frac{\ln(c_2/5)}{15} \right]
+  \mu_k = \dfrac
+    {\kappa\mu_{0} + n_k \bar{x}_k}
+    {\kappa + n_k}
+    \label{eq:category_mus}
 \end{equation}
-The class file also supports the use of \verb+\mathbb{}+, \verb+\mathscr{}+ and
-\verb+\mathcal{}+ commands. As such \verb+\mathbb{R}+, \verb+\mathscr{R}+
-and \verb+\mathcal{R}+ produces $\mathbb{R}$, $\mathscr{R}$ and $\mathcal{R}$
-respectively (refer Subsubsection~\ref{subsubsec2}).
-
-\section{Tables}\label{sec5}
-
-Tables can be inserted via the normal table and tabular environment. To put
-footnotes inside tables you should use \verb+\footnotetext[]{...}+ tag.
-The footnote appears just below the table itself (refer Tables~\ref{tab1} and \ref{tab2}). 
-For the corresponding footnotemark use \verb+\footnotemark[...]+
-
-\begin{table}[h]
+where $n_k$ is the number of observed exemplars in category $k$ and $\bar{x}_k$
+is the observed category mean. Note that if there are no observed exemplars in
+the target category (i.e., if the model is generating a completely novel
+category), Equation \ref{eq:category_mus} simplifies to $\mu_k = \mu_0$.
+
+The NIW prior also allows us to infer the category covariance matrix $\Sigma_k$
+by computing the following:
+\begin{equation}
+  \Sigma_k = [\nu \Sigma_D+ C_k +
+    \dfrac
+    {\kappa n_k}
+    {\kappa + n_k}
+    (\bar{x_k}-\mu_k)(\bar{x_k}-\mu_k)^T
+  ] (\nu + n_k)^{-1}
+  \label{eq:Sigma_y}
+\end{equation}
+where $C_k$ is the observed category covariance and $\Sigma_D$ is the domain
+covariance matrix from which $\Sigma_k$ samples are obtained. We can
+infer $\Sigma_D$ from the observed category covariances $C$ and the prior
+covariance $\Sigma_0$. Specifically,
+\begin{equation}
+    \Sigma_D = \Sigma_0 + \sum_{k}{C_k}
+\end{equation}
+
+At this point, both the representativeness model and the hierarchical Bayesian
+model from \cite{jern2013probabilistic} are largely identical. However, both
+models diverge in their computation of exemplar generation probabilities. While
+the original hierarchical Bayesian model produces novel exemplars with
+probabilities proportional to the multivariate normal likelihood $p(y \mid h)$, the
+representativeness model generates new exemplars according to their representativeness $R(y,h)$ as formulated in Equation \ref{eq:representativeness}. In practice, the probability of generating a
+particular candidate $y$ is obtained using an Exponentiated
+choice rule \citep{luce1977choice}:
+\begin{equation}
+  p(y) = \dfrac
+    {\exp( \theta \cdot R(y,h))}
+    {\sum_k \exp(\theta  \cdot R(y_k,h_k))} .
+\label{eq:representative-choice}
+\end{equation}
+where $\theta$ is a freely estimated response determinism parameter (constrained
+such that $\theta \geq 0$).
+
+Here, we implement the representativeness mechanism at the level of
+distributions over category exemplars. However, with a hierarchical model, it is
+also possible to apply the representativeness mechanism at other levels, such as
+the level of the distribution of category statistics (i.e., their means and
+covariances). For simplicity, we focus only on one type of representativeness
+model (i.e., where the mechanism is implemented at the level of distribution
+over category exemplars) and formalism \citep[for a related formalization of representativeness, see][]{bordalo21}.
+
+Despite both the original hierarchical Bayesian model and our representativeness
+model sharing identical hierarchical structures, their distinct response
+processes can yield very different exemplars. Specifically, the hierarchical
+Bayesian model emphasizes the generation of exemplars that maintain
+distributional commonalities across categories. In contrast, the
+representativeness model focuses on the generation of exemplars that are
+representative of the underlying distribution for a target category. Generally,
+with the assumption of unimodal underlying distributions, this mechanism of
+representativeness results in the generation of novel exemplars that are less
+similar to exemplars from the known category. This occurs because the
+representativeness of novel exemplars tend to be highest where the probability
+density of the known category's underlying distribution is the lowest. We see
+this illustrated in Figure \ref{fig:representative-examples}, where the
+representativeness model displays a strong preference for generating exemplars
+from target category 'B' that are further away from the contrast category 'A'.
+Exceptions to this pattern can occur (i.e., the generation of novel exemplars
+similar to known category exemplars) -- we explore this in more detail in Section
+\ref{section:diffcon} where we highlight an important condition demonstrating
+how PACKER and the representativeness model can make qualitatively different
+predictions.
+
+\begin{figure}
+	\begin{center} \includegraphics{fig2.eps}
+		\caption{Generation of a category `B' example, following exposure
+to one member of category `A' and one member of category `B'. Predictions are
+shown for (a) the original hierarchical Bayesian model and (b) the
+representativeness model.}
+		\label{fig:representative-examples}
+	\end{center}
+\end{figure}
+
+%\subsection{Synopsis and Prognosis}
+%
+%Research on categorization and category generation has focused on the finding
+%that generated categories tend to possess distributional commonalities with
+%known categories. However, a fundamental task of categorization is to create
+%something {\em new} (i.e., different from what is already known). The manner in
+%which generated categories differ from known ones is, nonetheless, poorly
+%understood: Existing theories do not make strong predictions about how generated
+%concepts should systematically differ from existing ones. Above, we introduced
+%two novel approaches to formalizing the roles of similarity and contrast in
+%category generation.
+%
+%In the sections below, we present two experiments demonstrating systematic
+%effects of category contrast on category generation inspired by factors
+%influencing how the contrast models generate new categories. Our experiments are
+%based on \cite{jern2013probabilistic}'s paradigm, which is a straightforward
+%translation of the traditional artificial classification paradigm to the task of
+%category generation: Participants are first exposed to a single,
+%experimenter-defined category, and are then asked to generate members of a new
+%category. We then report quantitative analyses of our results comparing the exemplar dissimilarity
+%and the representativeness accounts of our results to that of the hierarchical
+%Bayesian and copy-and-tweak models developed by \cite{jern2013probabilistic}.
+%After analyzing the category generation studies, we present a third experiment to
+%distinguish between our two conceptualizations of contrast.
+
+\section{Experiment 1}
+To begin our investigation, we examined category generation in a well-understood domain using a few disparate category types. We used an
+artificial stimulus design: A two dimensional domain of squares, varying in
+color and size (see Figure \ref{fig:e1-conditions}a). These dimensions
+have been used in numerous classification learning studies
+\cite[e.g.,][]{conaway2016similar,conaway2016generalization,shepard1961learning,nosofsky1994comparing}.
+Unlike those used in the \cite{jern2013probabilistic} experiments, distance on
+these physical dimensions aligns more directly with perceptual similarity,
+allowing us to evaluate the role of  contrast in categorization
+more precisely.
+It also enables more straightforward comparisons to prior work.
+We tested the effects of category contrast after learning one category from a set of qualitatively distinct category structures, as shown in Figure \ref{fig:e1-conditions}.
+
+Figures \ref{fig:e1-conditions}b-d show the values of exemplar dimensions
+belonging to the experimenter-defined categories (`A', or `Alpha') that
+participants were assigned to learn about prior to generating a new category.
+Each participant learned one of the category types during training. In the
+`Cluster' type, category A is a tight cluster of examples in the space.
+Perceptually instantiated, the members of category A might, for example, be
+large and dark in color. In the `Row' type, category A has a row pattern across
+the space, varying along one feature but not the other. Thus, its members might
+all be dark in color but would vary in size. Finally, in the `XOR' type, the
+experimenter-defined category consists of two clusters separated in opposite
+corners of the space, conforming to the exclusive-or logical structure (e.g.,
+members are small and dark or large and light).
+
+It should be noted that in our experiments the assignment of the perceptual
+to conceptual dimensions (e.g., $X \rightarrow Size$, $Y \rightarrow Color$) and the direction of variation along each dimension (e.g., $dark
+\rightarrow light$ or $light \rightarrow dark$) were counterbalanced across
+participants. The category types in Figure \ref{fig:e1-conditions} are plotted
+in conceptual space, rather than perceptual space. Thus, while the
+conceptual organization of the category types remains constant, each category
+type may have a different physical instantiation according to the counterbalance
+assignment. For example, the Cluster type may be large and dark in color, or it
+may be small and light in color, depending on the assignment and direction of
+the dimensions. For this reason, below we will discuss generation within a
+conceptual space, rather than a physically instantiated one.
+
+\begin{figure}
+    \begin{center} \includegraphics[width=0.95\textwidth]{figs/e1-conditions.eps}
+    \caption{Stimulus domain and category types tested in Experiment 1. Stimuli
+are not drawn to scale. Dimension and direction assignment (e.g., large to small or small to large) for color and size were counterbalanced over participants.}
+    \label{fig:e1-conditions}
+    \end{center}
+\end{figure}
+
+After learning about an experimenter-defined category, participants are asked to
+generate examples of a new category. Within this paradigm, an effect of category
+contrast would be realized if participants prefer to generate items in locations
+that are distant (i.e., perceptually dissimilar) from members of category A.
+However, generation is left unconstrained. Critically, participants were not
+asked to generate something different in the prompt. For example, participants
+assigned to the Cluster condition may generate a tightly clustered category in
+the corner opposite of the experimenter-defined category. Alternatively, they
+may generate a tightly clustered category directly overlapping with the
+experimenter-defined category. Further, they may even generate an entirely
+different type of category (e.g., a row category).
+
+Our experimental results also provide a converging test
+that generated categories tend to share distributional properties with known
+categories in the domain \citep{jern2013probabilistic,ward1994structured}. From
+these results, we can predict that, in each condition, participants should
+generate categories that are distributionally similar to the
+experimenter-defined category: In the Cluster condition, generated categories
+should be tightly clustered. In the Row condition, generated categories should
+vary more along the X-axis than the Y-axis. In XOR condition, generated
+categories should be widely distributed across both dimensions, and the two
+dimensions should be positively correlated.
+
+Interestingly, the XOR condition also offers a dissociation between the roles of
+category contrast and the emulation of distributional structure:
+widely-distributed, positively-correlated categories would need to lie along the
+positive diagonal of the space (that is the only place they ``fit''), which is
+already occupied by the experimenter-defined category. Thus, if contrast plays a
+role, exemplars in the generated categories of participants in the XOR condition
+may not be positively correlated -- they may not be correlated, or perhaps even
+be negatively correlated. In this case, contrast and statistical regularities
+would interact, which would be inconsistent with prevailing theories of category
+generation \citep{jern2013probabilistic}.
+
+\subsection{Participants and Materials}
+
+183 participants were recruited from Amazon Mechanical Turk. Each participant
+was randomly assigned to one condition: 64 participants were assigned to the
+Cluster condition, 61 were assigned to the Row condition, and 58 were assigned
+to the XOR condition (sample sizes differ due to random assignment). Stimuli
+were squares varying in color (grayscale 9.8\%--90.2\%) and side length
+(3.0--5.8cm), see Figure \ref{fig:e1-conditions}. The assignment of perceptual
+features (color, size) to axes of the conceptual space (X, Y)  and the
+direction of variation along each axis (e.g., $dark \rightarrow light$ or $light
+\rightarrow dark$) were counterbalanced across participants.
+
+\subsection{Procedure}
+
+As noted in the introduction of this paper, the task of generating members of a
+new category is well situated as a task in the categorization
+literature: Whereas classification consists of predicting an object's category
+label on the basis of its features, inference consists of predicting an observed
+feature, given a set of observed features and a category label. Generation thus
+consists of predicting {\em all} features of an object, given a novel category
+label. We designed our generation task as an
+extension of the traditional artificial classification learning paradigm. The
+task differs from traditional work in creative cognition primarily through the
+use of an artificial domain, which enables the application of computational
+models. The use of an artificial domain also requires the addition of a training
+phase, during which participants learn about the categories in the domain. As a
+result, unlike most previous studies \citep[e.g.,][]{ward1994structured},
+participants in our studies have no experience with the domain before the start
+of the experiment, and the experimenter-defined categories are not
+hierarchically structured (as are many natural categories).
+
+Participants began the experiment with a short training phase (3 blocks of 4
+trials), where they observed exemplars belonging to the `Alpha' category.
+Participants were instructed to learn as much as they can about the `Alpha'
+category, and that they would answer a series of test questions afterwards. On
+each trial, a single `Alpha' category exemplar was presented, and participants
+were given as much time as they desired to observe it before moving on to the
+next trial. Each block consisted of a single presentation of each of the members
+of the `Alpha' category, in a random order. Participants were shown the range of
+possible colors and sizes prior to training.
+
+Following the training phase, participants were asked to generate four examples
+belonging to another category called `Beta'. As in
+\citet{jern2013probabilistic}, generation was completed using a sliding-scale
+interface. Two scales controlled the values of the two dimensions (color, size)
+for the generated example. An on-screen preview of the example updated whenever
+one of the features was changed. Participants could generate any example along
+an evenly-spaced 9x9 grid (including members of the `Alpha' category), except for
+any previously generated `Beta' exemplars. Neither the members of the `Alpha'
+category nor the previously generated `Beta' examples were visible during
+generation. Prior to beginning the generation phase, participants read the
+following instructions:
+
+\begin{displayquote} As it turns out, there is another category of geometric
+figures called ``Beta''. Instead of showing you examples of the Beta category,
+we would like to know what you think is likely to be in the Beta category.
+
+You will now be given the chance to create examples of any size or color in
+order to show what you expect about the Beta category. You will be asked to
+produce 4 Beta examples - they can be quite similar or quite different to each
+other, depending on what you think makes the most sense for the category.
+
+Each example needs to be unique, but the computer will let you know if you
+accidentally create a repeat.
+\end{displayquote}
+
+\subsection{Results}
+
+We observed a substantial degree of individual differences in our data. In
+Figure \ref{fig:e1-samples} we have plotted sample data from several
+participants, from which it is evident that different participants generated
+qualitatively different category structures. In this section we will focus on
+analyzing the data in aggregate, but in later sections we will explore how these
+individual differences can be explained.
+
+\begin{figure}
+    \begin{center} \includegraphics[width=0.95\textwidth]{figs/e1-samples.eps}
+    \caption{Sample categories generated by participants in Experiment 1.
+Representative samples from common generation profiles are shown.}
+    \label{fig:e1-samples}
+    \end{center}
+\end{figure}
+
+To evaluate the role of contrast, we computed the number of times each stimulus
+was generated, as a function of its average city-block distance from members of
+the experimenter-defined ``Alpha'' category. These data, shown in Figure
+\ref{fig:e1-distanceplots}, reveal a clear pattern: Examples that are more
+distant from members of the experimenter-defined categories are more likely to
+be generated into a new category. This supports the notion that contrast is a
+fundamental constraint on how categories are related to one another and that
+statistical regularity alone is insufficient.
+
+Figure \ref{fig:e1-distanceplots} also depicts, for each participant, the
+average distance of members within the generated category ({\em within-category}
+distance) against the average distance between members of the generated and
+experimenter-defined category ({\em between-category} distance). The narrow
+distribution of between-category distances in the XOR condition reflects the
+widely distributed nature of the experimenter-defined category, reducing the
+possible distances to members of the participant-generated category. These data
+reveal a systematic pattern: The majority of participants generated categories
+with greater between-category distance than within-category distance. That is,
+members of the generated category tended to be more similar to one another than
+to members of the experimenter-defined category. To evaluate this claim
+quantitatively, we conducted t-tests comparing the amount of within- and
+between- class distance in each condition. All conditions possessed greater
+between-category distance: Cluster, $t(63) = 11.43$, $p < .001$; Row, $t(60) =
+13.16$, $p < .001$; and XOR, $t(57) = 3.64$, $p < .001$. These results provide
+further evidence of an effect of category contrast: Participants prefer to
+generate categories that are dissimilar to the learned category but maintain
+some level of internal cohesion.
+
+\begin{figure}
+    \begin{center} \includegraphics[width=0.95\textwidth]{figs/e1-distanceplots.eps}
+    \caption{Experiment 1 results. {\em Left}: Frequency of generation as a
+function of distance from members of the experimenter-defined category. {\em
+Right}: Scatter plot of within-category versus between-category distance in each
+of the participant-generated categories.}
+    \label{fig:e1-distanceplots}
+    \end{center}
+\end{figure}
+
+A secondary goal of this experiment was to examine whether we replicate the
+classic result that generated categories often possess the same distributional
+properties as previously-known categories. Given the increased importance of
+replication within psychology \citep{zwaan2018making}, it is important as it
+serves as a conceptual replication of \citet{jern2013probabilistic}. For each
+generated category, we computed the category range along each axis (X, Y), as
+well as the correlation between features. These data, shown in Figure
+\ref{fig:e1-statsboxes}, reveal broad individual differences: Within each
+condition, some participants generated categories spanning the entire X- and Y-
+axis whereas other participants categories that spanned very little along each.
+Likewise, in each condition participants generated categories possessing
+strongly positive, neutral, and strongly negative correlations between the
+dimensions. Comparing the distributional statistics between conditions yields a
+broad yet, as we will see, misleading replication of the classic effect.
+
+With respect to ranges along each axis (X, Y), the generated categories from
+each condition tend to reflect the ranges of the experimenter-defined
+categories. The categories generated in the Cluster condition were less widely
+distributed along the X-axis compared to Row, $t(123) = 5.61$, $p < .001$, and
+XOR, $t(120) = 2.68$, $p < .01$. Categories generated in the XOR condition
+were also less widely distributed along the X-axis compared to Row, $t(117) =
+2.56$, $p = .046$. This latter effect was not expected because the
+experimenter-defined categories for XOR and Row had similar X-ranges. However,
+the key finding is that categories from the Cluster condition tended to be more
+tightly distributed along the X-axis.
+
+\begin{figure}
+    \begin{center} \includegraphics[width=0.95\textwidth]{figs/e1-statsboxes.eps}
+    \caption{Box-plots of the distributional statistics from the categories
+generated in Experiment 1. Boxes depict the median and quartiles of each
+condition, with whiskers placed at 1.5 IQR. All points outside this region are
+marked individually.}
+    \label{fig:e1-statsboxes}
+    \end{center}
+\end{figure}
+
+Likewise, categories generated in the Row condition had less Y-axis range
+compared to Cluster, $t(123) = 4.57$, $p < .001$ and XOR, $t(117) = 9.26$, $p <
+.001$, and categories from the Cluster condition had less Y-axis range compared
+to XOR, $t(120) = 3.95$, $p < .001$. As expected, the correlations in the
+Cluster and Row conditions were not systematically positive or negative ($ps >
+.1$). However, the generated categories in the XOR condition tended to possess
+{\em negatively} correlated dimensions, $t(57) = 2.04$, $p = .046$. This
+finding is notable, as it is the opposite of what would be expected based on previous literature \citep{jern2013probabilistic}, assuming
+learners are emulating the distributional structure of the experimenter-defined
+class (which possesses perfectly positively correlated features).
+
+We believe that the failure to replicate the emulation of dimension correlation is
+because participants in \cite{jern2013probabilistic} could differentiate the
+generated category on a third dimension (hue) to maintain the statistical
+regularities on the other two dimensions. In addition, in our XOR condition the
+stimuli were constrained to the corners of the feature space, whereas in the
+positive diagonal condition of \cite{jern2013probabilistic} the stimuli had no
+such constraint. Although the correlation in the XOR condition is significantly
+negative, it is clear from the box-plot in Figure~\ref{fig:e1-statsboxes} that
+it would be inappropriate to make a strong conclusion (e.g., the median is close
+to zero). However, we can conclude with confidence that there are situations
+where people do not emulate the distributional structure of the given category.
+This indicates that there is more to category generation than the emulation of
+distributional structure of other categories in the domain. Further, as we will
+discuss in more detail in the model-based analysis section, this is expected by
+our proposal that contrast is a fundamental principle in categorization.
+
+\subsection{Discussion} In Experiment 1 we evaluated the influence of category
+contrast on category generation, given qualitatively different types of
+categories. We found strong evidence for effects of category contrast in each
+condition: Participants were more likely to generate stimuli that are more
+distant from (i.e., less similar to) members of a previously-learned category,
+and members of participant-generated categories tended to be more similar to one
+another than to members of previously-learned categories. We also partially
+replicated the classic finding that the distributional structure of generated
+categories reflects that of previously learned categories
+\citep{jern2013probabilistic,ward1994structured}: Members of generated
+categories were more widely distributed along dimensions which were widely
+distributed in the experimenter-defined category.
+
+Notably, however, we also found that participants who learned an XOR category
+(composed of exemplars following a positive diagonal, see Figure
+\ref{fig:e1-conditions}) tended to generate items according to a {\em negative}
+feature correlation -- the opposite of what was present in the previously
+learned category. While this may be difficult to account for under existing
+theoretical approaches (which assume generated categories follow the same
+distributional structure as known categories), it can be concisely explained
+from a category contrast perspective. Specifically, within the XOR condition,
+individuals who seek to generate a category that is perceptually distinct from
+what is already known are left with only the upper-left and bottom-right
+quadrants of the space, as members of the previously-learned XOR category lie in
+the bottom-left and top-right. If examples are generated into both of the
+available quadrants, the generated category will possess a strongly negative
+correlation, opposing that of the experimenter-defined class.
+
+Thus, the core results of Experiment 1 indicate that generated categories can
+systematically differ from what we would expect based on prior work. The
+negative (or null) correlations observed in the XOR condition suggests an
+interesting interaction between contrast with a given category and emulation of
+statistical properties. That is, the constraints on generation imposed by
+category contrast may not simply influence the {\em location} of generated
+categories, but also their distributional structure. In Experiment 2, we test
+this claim more systematically.
+
+\section{Experiment 2}
+
+\begin{figure}
+    \begin{center} \includegraphics{figs/e2-conditions.eps}
+    \caption{Category types tested in Experiment 2.}
+    \label{fig:e2-conditions}
+    \end{center}
+\end{figure}
+
+To test whether category contrast influences the distributional structure of
+generated categories, we sought to identify conditions in which differences in
+the distributional structure of generated categories cannot be explained by the
+distributional structure of the experimenter-defined category. We created two
+new category types (depicted in Figure \ref{fig:e2-conditions}) that possess
+identical distributional structures (both are tight clusters of examples with no
+correlation between features), as they only differ in their Y-axis position: the
+`Bottom' category lies near the bottom of the space, and the `Middle' category
+lies in the center. The distributional equality of these conditions is key to
+the design of the experiment: If the distributional structure of previously
+learned categories were the only influence on the generated categories, we
+should observe no difference in the categories participants generate between
+these two conditions. Will participants distribute their generated category
+differently between conditions due to the differences in the available empty
+feature space for generating a new category?
+
+If category contrast influences the distributional structure of the categories
+people generate, then we should observe different types of categories according
+to the shape of the space that is {\em unoccupied} by members of previously
+learned categories. The difference in the Y-axis position between the Bottom and
+Middle conditions produces a considerable change to the shape of the unoccupied
+space. Participants assigned to learn the Bottom category should be less likely
+to generate exemplars into the lower regions of the feature space (as these
+areas possess greater similarity to members of the Bottom category), preferring
+instead to distribute exemplars across the upper region of the space. This
+constraint is lifted in the Middle condition, as the Middle category exemplars
+are equidistant to the upper and lower regions of the space. Accordingly,
+participants should be more likely to utilize both of these areas. Thus, if
+category contrast influences the distributional structure of generated
+categories, we should observe more participants in the Middle condition that
+generate examples above {\em and} below the experimenter-defined category.
+
+\subsection{Participants, Materials, and Procedure}
+
+122 participants were recruited from Amazon Mechanical Turk. 61 participants
+were randomly assigned to the Middle and Bottom conditions each. The stimulus
+space and procedure were exactly as in Experiment 1. Participants first
+completed a short training phase, followed by the generation phase. The only
+difference from Experiment 1 was the category types given to participants.
+
+\subsection{Results}
+
+\begin{figure}
+    \begin{center} \includegraphics[width=0.95\textwidth]{figs/e2-samples.eps}
+    \caption{Sample categories generated in Experiment 2. }
+    \label{fig:e2-samples}
+    \end{center}
+\end{figure}
+
+As in Experiment 1, we observed broad differences in the generation approach
+taken by different participants. To characterize the nature of these
+differences, Figure \ref{fig:e2-samples} depicts sample categories generated by
+participants. The data from each condition are organized into four columns based
+on commonly observed patterns of generation: a `Cluster' type of
+tightly-clustered examples, `Row' and `Column' types of exemplars widely
+distributed along the one axis but narrowly along the other, and a `Corners'
+type, wherein participants placed exemplars in disparate corners of the space.
+As before, in this section we focus on analyzing the data in aggregate, but in
+later sections we will focus more specifically on explaining the individual
+differences.
+
+\begin{figure}
+    \begin{center} \includegraphics[width=0.95\textwidth]{figs/e2-distanceplots.eps}
+    \caption{Experiment 2 results. {\em Left}: Frequency of generation as a
+function of distance from members of the experimenter-defined category. {\em
+Right}: Scatter plot of within-category versus between-category distance in each
+of the participant-generated categories.}
+    \label{fig:e2-distanceplots}
+    \end{center}
+\end{figure}
+
+We began our analysis by testing for the broad influence of category contrast on
+generation. As in Experiment 1, we computed the frequency each stimulus was
+generated as a function of its average distance from members of the
+experimenter-defined category, as well as each participant's average within- and
+between- category distance. These data, shown in Figure
+\ref{fig:e2-distanceplots}, yield very similar results. Participants generated
+stimuli that are distant from members of the experimenter-defined category, and
+the categories in each condition tended to possess more between-category than
+within-category distance: Bottom, $t(60) = 5.5$, $p < .001$; Middle, $t(60) =
+2.71$, $p < .01$. We did, however, observe a notable subgroup of participants
+in each condition who generated categories with more within-category than
+between-category distance. Upon manual inspection, many of these individuals
+appear to have assumed a `Corners' strategy, placing exemplars in disparate
+corners of the space, thus producing much more within-category distance, see
+Figure \ref{fig:e2-samples} for examples.
+
+To explore the distributional structure of the generated categories, we computed
+the range of exemplars along each axis (X, Y), as well as the correlation
+between features. These data, shown in Figure \ref{fig:e2-statsboxes}, again
+demonstrate the degree of individual differences observed in our study. In each
+condition, we observed tightly clustered and widely distributed categories along
+each dimension. Although most participants generated uncorrelated categories in
+both conditions, many still produced positively and negatively correlated
+categories.
+
+\begin{figure}
+    \begin{center} \includegraphics[width=0.95\textwidth]{figs/e2-statsboxes.eps}
+    \caption{Box-plots of the distributional statistics from the categories
+generated in Experiment 2. Boxes depict the median and quartiles of each
+condition, with whiskers placed at 1.5 IQR. All points outside this region are
+marked individually.}
+    \label{fig:e2-statsboxes}
+    \end{center}
+\end{figure}
+
+As noted above, if the distributional structure of generated categories is
+influenced by the shape of the space not occupied by members of known
+categories, then participants in the Middle condition would be more likely to
+place exemplars in the upper {\em and} lower regions of the space, as members of
+the experimenter-defined category are equidistant from these regions.
+Participants in the Bottom condition should be less likely to generate category
+members in the bottom regions because members of the experimenter-defined
+category are located there. One way to test these predictions is to analyze the
+Y-axis ranges of the generated categories: If Middle participants utilize the
+upper and lower regions of the space, their categories should vary more along
+the Y-axis. T-Tests comparing the conditions on the distributional statistics,
+however, reveal few between-group differences: the conditions do not differ with
+respect to X-axis range, Y-axis range, or feature correlations ($ps>0.17$).
+
+However, our ability to detect differences in Y-axis range using a standard
+$t$-test between the conditions is, in this case, diminished due to the
+non-normality of the data (Shapiro-Wilk normality test $W=0.77, p < .001$ for
+the Middle condition and $W=0.85, p < .001$ for the Bottom condition).
+Figure \ref{fig:e2-yranges} depicts the Y-axis range and Y-axis position of exemplars generated by each participant.
+The categories are sorted by overall range, and then colored by training condition.
+These data reveal that there were nearly as many
+participants who generated categories spanning the entire Y-axis as those who
+generated categories spanning almost none of the Y-axis.
+Indeed, the median produced Y-axis range is much smaller in the Bottom than Middle condition, whereas they are essentially identical in X-axis range.
+The non-normality of the Y-axis range distributions thus requires that we use a different analytic approach to
+addressing the experiment's main question.
+
+\begin{figure}
+    \begin{center} \includegraphics[width=0.95\textwidth]{figs/e2-yranges.eps}
+    \caption{Y-axis range and position of the participant-generated categories
+from Experiment 2. Each line corresponds to a participant's category, with
+notches corresponding to the Y-axis position of exemplars within the category
+(notches may overlap). Participants are sorted by overall range, and then by
+condition. }
+    \label{fig:e2-yranges}
+    \end{center}
+\end{figure}
+
+Because our main prediction concerns the generation of exemplars within the
+upper and lower regions of the domain, we compared the conditions in terms of
+the frequency with which participants generated examples above and below the
+categories. Specifically, we counted the number of participants in each
+condition who placed at least one `Beta' exemplar on the top and bottom `rows'
+of the space (the maximum and minimum possible Y-axis value, respectively). The
+resulting contingencies data are shown in Table \ref{table:e2-subset-table}.
+
+Firstly, it should be noted that nearly every participant utilized the top
+and/or bottom rows: only $10 / 122$ participants generated their category
+entirely within the interior region. Fisher's Exact Tests comparing the
+conditions reveal that more Middle participants generated an exemplar in the
+bottom row, $p < .001$, again demonstrating the role of contrast in guiding
+where exemplars are generated. The conditions did not differ in use of the top
+of the space, $p = .16$, however, more Middle participants placed exemplars in
+the top {\em and} bottom rows, $p = .04$. The latter effect is of particular interest
+here, as it indicates that the shape of the unoccupied space exerts some
+influence on the distributional structure of generated categories: Participants
+in the Middle condition were more likely to generate a category spanning the
+entire Y-axis. Thus, the distributional structure of the generated categories
+can be influenced without any change to the distributional structure of the
+given category. Rather, it can be affected by category contrast alone.
+
+\begin{table}
 \begin{center}
-\begin{minipage}{174pt}
-\caption{Caption text}\label{tab1}%
-\begin{tabular}{@{}llll@{}}
-\toprule
-Column 1 & Column 2  & Column 3 & Column 4\\
-\midrule
-row 1    & data 1   & data 2  & data 3  \\
-row 2    & data 4   & data 5\footnotemark[1]  & data 6  \\
-row 3    & data 7   & data 8  & data 9\footnotemark[2]  \\
-\botrule
+\caption{Experiment 2 results.}
+\label{table:e2-subset-table} \vskip 0.12in
+\begin{tabular}{ l r r} \textbf{Middle} & Used top row & No top row \\ \hline
+Used bottom row & 28 & 18 \\ No bottom row & 11 & 4 \\ \\ \textbf{Bottom} & Used
+top row & No top row \\ \hline Used bottom row & 16 & 8 \\ No bottom row & 31 &
+6 \\
 \end{tabular}
-\footnotetext{Source: This is an example of table footnote. This is an example of table footnote.}
-\footnotetext[1]{Example for a first table footnote. This is an example of table footnote.}
-\footnotetext[2]{Example for a second table footnote. This is an example of table footnote.}
-\end{minipage}
 \end{center}
@@
 
-\noindent
-The input format for the above table is as follows:
-
-%%=============================================%%
-%% For presentation purpose, we have included  %%
-%% \bigskip command. please ignore this.       %%
-%%=============================================%%
-\bigskip
-\begin{verbatim}
-\begin{table}[<placement-specifier>]
+\subsection{Discussion}
+
+In Experiment 2, we replicated the core findings from Experiment 1. Stimuli are
+more likely to be generated if they are distant from exemplars in other
+categories, and most participants generate categories with more between-category
+than within-category distance. However, we additionally found that the {\em
+position} of a previously learned category (rather than its distributional
+structure) influences the types of categories people generate: Participants who
+learned the `Middle' type were more likely to generate categories spanning the
+entire Y-axis of the space. Participants who learned the `Bottom' type were less
+likely to do so as a result of the presence of opposite category exemplars in
+the lower regions of the space.
+
+This finding cannot be explained from the perspective that the distributional
+structure of previously learned categories is the sole determinant of the
+distributional structure of generated categories. However, the observed behavior
+is expected from a category contrast perspective: Participants seeking to
+generate a perceptually distinct category will be more likely to use areas of
+space that are unoccupied by exemplars belonging to previously learned
+categories. In the Middle condition, the upper and lower regions of space are
+equidistant from members of the experimenter-defined category, whereas in the
+Bottom condition, the lower region of the space is closer to members of the
+experimenter-defined category. Thus, while Middle participants may form
+categories around the use of the equally unoccupied areas, the same is not true
+for the Bottom condition.
+
+\section{Model-based Analyses of Experiments 1 and 2}
+
+Experiments 1 and 2 revealed systematic and strong effects of category contrast
+on category generation. In this section, we analyze the performance of the
+different formal models at explaining the experimental results. Specifically, we
+present simulations from the two novel contrast models: PACKER and the
+representativeness model, and compare them to two models that do not incorporate
+contrast: the copy-and-tweak model (discussed in Section
+\ref{section:copytweak-packer}) and an implementation of the hierarchical
+Bayesian model proposed by \cite{jern2013probabilistic}, described in-depth in
+Appendix \ref{ap:hsampling-definition}. For our simulations here, the
+copy-and-tweak model is defined as a variant of PACKER with the $\theta_c$
+parameter constrained to be zero. The comparison of this set of models serves to
+highlight the explanatory role of contrast in categorization: If contrast
+affords little explanatory advantage, then there should be few differences in
+performance between PACKER and copy-and-tweak, or between the representativeness
+and hierarchical Bayesian model. The comparison between these models can also
+emphasize the necessity of contrast and demonstrate that generation cannot be
+explained entirely through the emulation of distributional structure. Each model
+has complementary strengths and weaknesses: Whereas PACKER and copy-and-tweak
+are relatively insensitive to the distributional structure of learned categories
+(relying only on exemplar similarities), the representativeness and hierarchical
+Bayesian model generates categories exclusively on the basis of knowledge of how
+existing classes are distributed.
+
+Our approach in this section is to first broadly evaluate and compare the
+quality of each model's account to our entire dataset (Experiments 1 and 2
+combined), then analyze the ability for each model to explain individual
+differences in each experiment, and lastly we describe the strengths and
+weakness of each model's account of category generation.
+
+\subsection{Parameter-Fitting}
+
+To obtain a global measure of the quality of each model's account, we fit the
+parameters of each model to our entire dataset (Experiments 1 and 2 combined),
+using a hill-climbing algorithm which maximized the log-likelihood of the
+model's predictions of the observed responses (1220 responses from 305 total
+participants). We fit three parameters in the PACKER model ($c$, $\theta_t$,
+and $\theta_c$; see Section \ref{section:PACKER-definition}), as well as four in
+the representativeness model and the hierarchical Bayesian model ($\kappa$,
+$\lambda$, $\nu$, and $\theta$; see Section
+\ref{section:representativeness-definition} and Appendix
+\ref{ap:hsampling-definition} respectively). We fit only two parameters for
+the copy-and-tweak model ($c$, and $\theta_t$), as $\theta_c$ is held constant
+($\theta_c = 0$). Attention ($w$, see Equation \ref{eq:similarity}) in PACKER
+and copy-and-tweak was set uniformly. Parameters were not allowed to vary
+between participants or conditions -- the goal was to obtain the best-fitting
+values to our entire dataset.
+
+\begin{table} \centering
+\caption{Results of model-fitting to the combined datasets from Experiments 1
+and 2. Note that smaller AIC values correspond to better model fits (adjusted
+for number of parameters)}
+\label{table:global-model-fits}
+\begin{tabular}{ l l l l} \\ \textbf{PACKER} & \textbf{Copy \& Tweak} &
+\textbf{Representativeness} & \textbf{Hierarchical Bayesian} \\ \hline $AIC =
+  9069$ & $AIC = 9813$ & $AIC = 8783$ &  $AIC =
+9881$ \\ $L = -4531$ & $L = -4905$ & $L = -4388$ & $L = -4937$ \\ $c = 0.51$ &
+                                                                                $c = 3.22$ & $\kappa = 12.23$ &
+$\kappa < 0.001$ \\ $\theta_c = 3.09$ & $\theta_c = 0$ (fixed) & $\nu = 1.00$ & $\nu = 5.44$ \\
+$\theta_t = 3.47$ & $\theta_t = 3.00$ & $\lambda = 7.04$& $\lambda
+                                                                       = 0.06$
+  \\ & &$\theta = 10.21 $& $\theta = 3.09$
+\\
+\end{tabular}
+\end{table}
+
+Table \ref{table:global-model-fits} contains the model fits. Due to the uneven number of fitted parameters among the models, we
+compare the model fits using the Akaike Information Criterion
+\citep[AIC;][]{akaike1974new}, where smaller values correspond to better fits
+(discounted by model complexity as measured by the number of parameters). The
+same qualitative results were obtained with alternative model comparison metrics
+(e.g., BIC, \citealp{schwarz1978estimating}; $AIC_C$,
+\citealp{hurvich1989regression}). In addition to AIC, Table \ref{table:global-model-fits} contains the corresponding log-likelihood ($L$) and the best-fitting
+parameter values. These results reveal strong model differentiation: both
+contrast models (PACKER and the representativeness model) achieved far better
+fits compared to their non-contrast counterparts: copy-and-tweak and the
+hierarchical Bayesian model respectively. Interestingly between the contrast
+models, the hierarchical model (representativeness) outperformed the
+exemplar-based theory (PACKER), whereas between the non-contrast models, the
+reverse is observed. Specifically, here the exemplar-based model
+(copy-and-tweak) performed somewhat better than the hierarchical Bayesian model.
+
+While PACKER's advantage over copy-and-tweak may tentatively be attributed to
+the model's sensitivity to category contrast (this will be explored in detail
+below), the advantage shown by copy-and-tweak over the hierarchical Bayesian
+model may be attributed to its exemplar-based representation of category B, as
+opposed to forcing a prototype-based representation as assumed by the
+hierarchical Bayesian model. As observed in Figures \ref{fig:e1-samples} and
+\ref{fig:e2-samples}, the generated categories we observed were often widely
+distributed, with no items near the category prototype. This aspect of the data
+is inconsistent with the multivariate normal distributions (similar to
+prototypes) used to represent categories in the \cite{jern2013probabilistic}
+model, but can be easily accounted for using an exemplar-based approach.
+Interestingly, representativeness using a prototype approach fits better than an
+exemplar-based approach.
+
+A key distinction between the contrast and classical models is that only the
+contrast models are capable of making strong predictions about the location of
+new category members when the target class is entirely novel (i.e., no member of
+the category has been observed). Under these circumstances, there are no
+examples to copy, and thus the copy-and-tweak model predicts that items are
+generated at random. Likewise, with no observations on which to condition the
+category distribution, the hierarchical Bayesian model also picks an item at
+random.
+
+Thus, it is possible that the failure of the classical models is simply due to
+their inability to explain each participant's first trial (generating the first
+item in the `Beta' category). To ensure this is not driving our results, we
+conducted an identical set of simulations as above, excluding the first trial
+(leaving 915 responses in the dataset): Again, the representativeness model ($L
+= -3286$, $AIC = 6580$) and PACKER ($L = -3377$,
+$AIC = 6759$) achieved better fits than the copy-and-tweak ($L = -3564$,
+$AIC = 7132$) and hierarchical Bayesian ($L = -3597$, $AIC = 7201$) models.
+
+Finally, because copy-and-tweak is nested within PACKER, we can use a likelihood
+ratio test to compare the two models. PACKER explains the aggregate data
+significantly better than copy-and-tweak ($\chi^2(1) = 747, p < .001$ for all
+data and $\chi^2(1) = 375, p < .001$ excluding the first example), providing
+further evidence that category generation is better explained when contrast is
+considered.
+
+Through comparison with the copy-and-tweak model, Figure
+\ref{fig:packer-loglike} more clearly demonstrates the robustness of the
+explanatory gains yielded by PACKER's category contrast mechanism. It displays
+the log-likelihood of the participants' results under PACKER as a function of
+the $\theta_c$ parameter. The model's other parameters ($c$,
+$\theta_t$) were set according to copy-and-tweak's best fits from Table
+\ref{table:global-model-fits}, and thus when $\theta_c=0$, the models are
+equivalent. The figure clearly shows a ``sweet spot'': a convex region in which
+PACKER achieves superior fits as a result of changes to $\theta_c$. The
+best fitting values lie well above the value of 0 assumed by the copy-and-tweak
+model, which demonstrates the robustness of the contrast effect (though note
+PACKER achieves even better fits when its parameters are fitted together, as in
+Table \ref{table:global-model-fits}). In sum, the data are better explained when
+both within-category similarity and category contrast is considered.
+
+\begin{figure}
     \begin{center}
-\begin{minipage}{<preferred-table-width>}
-\caption{<table-caption>}\label{<table-label>}%
-\begin{tabular}{@{}llll@{}}
-\toprule
-Column 1 & Column 2 & Column 3 & Column 4\\
-\midrule
-row 1 & data 1 & data 2	 & data 3 \\
-row 2 & data 4 & data 5\footnotemark[1] & data 6 \\
-row 3 & data 7 & data 8	 & data 9\footnotemark[2]\\
-\botrule
+      \includegraphics[width=\textwidth]{figs/packer-loglike.pdf}    
+      \caption{PACKER's fit as a function of its prioritization of
+        within-category and between-category similarity (using the $\theta_c$
+        parameter) . To facilitate comparison, PACKER's other parameters ($c$,
+        $\theta_t$) were set to the best fitting values obtained for
+        copy-and-tweak in Table \ref{table:global-model-fits}. The black diamond
+        marker indicates the log-likelihood for the point where
+        $\theta_c=\theta_t$. }
+    \label{fig:packer-loglike}
+    \end{center}
+\end{figure}
+
+\subsection{Individual Differences}
+\label{section:individual-diffs}
+
+As noted in Experiments 1 and 2, we observed a great deal of individual
+differences in the types of categories that participants generated. Within each
+condition, there were a wide variety of category types, such as row and column
+categories (see Figures \ref{fig:e1-samples} and \ref{fig:e2-samples}). The
+simulations reported above serve to evaluate the models while considering the
+entire dataset, but a secondary goal of any formal account should be to provide
+some explanation of how different profiles of performance emerge. Many of the
+individual generation profiles we observed can be described with the models
+simply by tuning the model's parameters in a principled manner. In this section,
+we describe more specifically how the most frequently observed profiles can be
+realized.
+
+By manual inspection, it is evident that the most common profiles of generation
+consist of: (A) a tightly-distributed `cluster' of examples, (B) `row'- and
+`column'-like arrangements (varying widely along one dimension but not the
+other), and (C) a `corners' arrangement with examples placed into disparate
+corners of the space. These four profiles are distinct in terms of the
+distribution of the generated category along each dimension: Whereas the cluster
+profile is tightly distributed along both dimensions, the row and column
+profiles are tightly distributed along just one dimension. Finally, the corners
+profile is widely distributed along both dimensions.
+
+In the framework proposed by PACKER, the cluster and corners profiles arise
+based on different prioritization of within-category similarity versus
+between-category contrast, and the row and column profiles arise based on the
+prioritization of each dimension in the computation of similarity. For example,
+in the cluster profile, there is a high degree of within-category similarity
+along both dimensions, whereas in the corners profile there is minimal
+within-category similarity. Thus, PACKER's proposal is that these individual
+differences arise as a result of different priorities: While the tight cluster
+configuration can be considered PACKER's `default' mode (as it maximizes
+within-category similarity), the corners profile can be produced when
+between-category contrast is put at a higher priority (i.e., $\theta_c > \theta_t$).
+
+Likewise, in the row and column profiles, there is a large degree of
+within-category similarity along one dimension but not the other. These
+differences likely arise due to a differential focus on one dimension over
+another, and thus they can be produced by changes to PACKER's attention weights,
+$w_1$ and $w_2$ (see Equation \ref{eq:similarity}). Traditionally, the attention
+weights in exemplar models are thought to reflect the diagnostic value of each
+dimension towards classifying the known category members
+\citep{nosofsky1984choice,nosofsky1986attention,kruschke1992alcove}, but within
+a generation context the weights specify the importance of within- and
+between-category similarity along each dimension. For example, if all of
+attention is allocated along the X-axis ($w_1=1$ and $w_2=0$), similarity along
+the Y-axis no longer influences performance. As a result, PACKER will create
+categories that are more widely distributed along the Y-axis, as similarity is
+not taken into account along that dimension. As a general principle,
+differentially weighting one dimension will result in the generation of
+categories that are more widely distributed along the ignored dimension,
+conforming to a row- or column-like arrangement. See Figure
+\ref{fig:packer-attention} for a depiction of how attention influences PACKER's
+performance.
+
+\begin{figure}
+    \begin{center} \includegraphics[width=0.8\textwidth]{fig13.eps}
+    \caption{PACKER generation of a category `B' example, following exposure to
+one member of category `A' and one member of category `B'. Predictions are shown
+for different attention settings: {\em (a)} Increased weighting of the X-axis.
+{\em (b)} Increased weighting of the Y-axis. {\em (c)} Uniform weighting
+(identical to Figure \ref{fig:packer-examples}).}
+    \label{fig:packer-attention}
+    \end{center}
+\end{figure}
+
+As in PACKER, changes in the parameter settings of the copy-and-tweak model can
+also be used to produce different patterns of generation. Indeed, as
+copy-and-tweak is simply a special case of the PACKER model, the attention
+weights operate exactly as described above to produce row- and column-like
+categories. However, because the model is not influenced by category contrast,
+it is biased toward generating tightly clustered categories, as new items are
+always most likely to be generated near known examples of the target category.
+Thus, the lack of a contrast mechanism prevents the model from explaining why
+some individuals widely distribute their categories to the corners of the space.
+
+For both the hierarchical Bayesian and the representativeness models, the prior
+domain covariance matrix $\Sigma_0$ can be used to explain the generation of
+row-like and column-like categories. This covariance matrix specifies the amount
+of variance assumed along each dimension (as well as the correlations between dimensions) across the domain of categories. The covariance matrix for a newly
+generated category, $\Sigma_B$, is based on the assumed $\Sigma_0$ as well as
+the distributions of previously learned categories (see Appendix
+\ref{ap:hsampling-definition}). Thus, the importance of each feature can be
+coded into $\Sigma_0$ to alter the dimensional variance of generated categories.
+Because the hierarchical Bayesian model possesses no mechanism to account for
+category contrast, the model is most likely to generate new items that are
+similar to known examples of the target category with no regard for how
+different it is to the contrast category. However, the representativeness model
+predicts that new exemplars should provide more relative evidence to the `Beta'
+category, accounting for the tendency of row-like and column-like profiles
+occupying the edge of the feature space.
+
+While the copy-and-tweak and hierarchical Bayesian models possess mechanisms to
+explain row- and column-like categories, they cannot easily explain why some
+individuals widely distribute their generated categories into disparate corners
+of the space. This, however, reveals a more general limitation: According to the
+copy-and-tweak and hierarchical Bayesian models, the distributional structure of
+generated categories is {\em independent} of their location within the domain.
+For example, although the copy-and-tweak or hierarchical Bayesian models can be
+parameterized to generate row- or column-like categories, there is no mechanism
+in place to ensure that what is generated will be distinct from what is already
+known. In the next subsection, we explore this prediction through an analysis of
+the interdependence between distributional structure and location in category
+generation.
+
+\subsection{Category Location vs. Distributional Structure}
+\label{section:individual-diff}
+As noted above, while all three models make clear claims about the internal
+structure of generated categories, the copy-and-tweak and hierarchical Bayesian
+models do not make any claims about how generated categories should differ from
+what is already known. However, as we observed in the results of Experiment 2,
+the distributional structure of a category is not always independent of its
+location within the domain. To demonstrate this point in more depth, we computed
+the X- and Y- axis ranges of every participant-generated category. Taking the
+difference between these values ($X-Y$) produces a measure of each category's
+orientation in the space: positive difference scores correspond to categories
+with more X-axis range (horizontally aligned, `Row' categories), whereas
+negative difference scores indicate the opposite (vertically aligned, `Column'
+categories). Neutral differences scores indicate there was an equal amount of X-
+and Y-axis range, which can be produced by a number of different category types
+(`Clusters', `Corners', etc; see Figures \ref{fig:e1-samples} and
+\ref{fig:e2-samples}). By plotting, for each possible stimulus, the difference
+scores of categories it was generated within, we can relate the distributional
+structure of generated categories to their location within the domain.
+
+However, because many stimuli were infrequently generated (such items near
+members of the `Alpha' category), we cannot simply compute the empirical average
+of the difference scores, as infrequently generated stimuli would be likely to
+show artificially strong differences. Instead, we used a Bayesian analysis to
+estimate the mean $\mu_x$ on the assumption that the scores $x$ for each
+stimulus are normally distributed with an unknown mean and unknown standard
+deviation. The conjugate Normal-Inverse Gamma distribution provides a
+straightforward method for this estimation:
+
+\begin{equation} \mu_x = \dfrac { \nu_0 \mu_0 + \sum{x} } { \nu_0 + n }
+\label{eq:rangediff-bayes}
+\end{equation} \ where $\mu_0$ is the prior mean, $\nu_0$ is a prior scale
+parameter (controlling the weighting of the $\mu_0$), and $n$ is the number of
+categories in which the stimulus was a member (i.e., the number of scores in
+$x$). The default assumption is that there is an equal amount of range along the
+X- and Y-axes, and so we set $\mu_0 = 0$. Likewise, to give a moderate amount of
+weighting to the prior mean we set $\nu_0 = 1$, though the results are robust to
+a range of values. Within this approach, the resulting aggregation is a
+trade-off between the number of generations and the strength of the range
+difference within each generated category. Infrequently generated stimuli, as
+well as those with mixed positive and negative scores, are given neutral
+difference scores.
+
+\begin{figure}[p]
+    \begin{center} 
+      \includegraphics[width=\textwidth]{figs/range-diff-gradients.pdf}
+      \caption{Behavioral and simulated range difference gradients. Each panel
+        shows, for each stimulus, the dimensional orientation of the categories
+        it was generated into: vertically aligned `columns' (orange) versus
+        horizontally aligned `rows' (purple).}
+      \label{fig:range-diff-gradients}
+    \end{center}
+\end{figure}
+
+The results of our analysis are shown in Figure \ref{fig:range-diff-gradients}
+for the experiment and model results\footnote{Prior to plotting, data were also
+processed using a Gaussian filter with $\sigma = 0.8$.}. The left-most column of
+Figure \ref{fig:range-diff-gradients} displays the effect of category location
+and contrast on the distributional structure of the category generated by
+participants. These data reveal strong and consistent patterns across all the
+conditions we tested in Experiments 1 and 2: Generated categories are more
+tightly distributed along the axis in which they are distinct. For example, in
+the `Cluster' condition, exemplars in the bottom-left of the space are more
+often generated into vertically aligned categories, and exemplars in the
+top-right are more often generated into horizontally aligned categories.
+Similarly, in the `Bottom' and `Middle' conditions, horizontally aligned
+categories are generated above and below the experimenter-defined categories,
+while vertically-aligned categories are generated to the sides. In the `Row'
+condition, most categories are horizontally aligned, and lie along the upper
+areas of the space. There are no strong range difference patterns in the XOR
+condition.
+
+These patterns of performance clearly depict the interdependence between the
+distributional structure and location of generated concepts. Our results can be
+interpreted in terms of local minimization of between-category similarity: By
+distributing the generated category away from members of the
+experimenter-defined category, participants may increase the degree of
+between-category distance without drastically altering the degree of
+within-category similarity.
+
+To explore how well the PACKER, copy-and-tweak, representativeness, and
+hierarchical Bayesian models explain our findings, we conducted simulations
+using an individual-differences approach. As noted in Section
+\ref{section:individual-diffs}, row- and column-like categories can be produced
+by each model through changes to the weighting of each dimension. Given this
+information, we may use the models to simulate each participant's generation
+separately, with the importance of each dimension set according to the relative
+range of the participant's generated category along each dimension.
+
+In the PACKER and copy-and-tweak models, the attention weights, $w$, specify the
+importance of each dimension in the computation of similarity. While there exist
+methods to find the optimal attention weighting scheme given a classification
+\citep[see][]{vanpaemel2012using}, for simplicity we assume that the `Alpha' and
+`Beta' categories are distinct along dimensions that the `Beta' exemplars do
+not vary on. In this case, the weighting for a given participant can be computed
+as:
+
+\begin{equation} w_k = \dfrac {\exp{ \{ -\theta_w \cdot\text{range}(k)} \} } {
+\sum_k {\exp{ \{ -\theta_w \cdot\text{range}(k)} \} } }
+\label{eq:range-weight}
+\end{equation}
+% 
+where $\theta_w$ is a free parameter controlling how differences in range
+correspond to differences in weights (functioning similarly to the $\theta$
+parameter in each of the models), and $\text{range}(k)$ is the range of examples
+generated by the participant along dimension $k$. We used $\theta_w = 1.5$ in
+our simulations, though the results are robust and similar for other $\theta_w$
+values. The resulting $w$ values are thus inversely proportional to the range of
+generated categories along each dimension, with less range corresponding to
+greater weighting.
+
+Unlike the PACKER and copy-and-tweak models, the representativeness and
+hierarchical Bayesian model's dimensional variances correspond to the assumed
+variance of generated categories along each dimension (rather than the inverse
+of the variance). Thus, a different transformation is appropriate for
+incorporating the weights computed in Equation \ref{eq:range-weight}. For the
+hierarchical Bayesian model, we computed the dimensional variances according to:
+$\lambda (1-w_k) 2$, where $\lambda$ is a free parameter specifying the overall
+assumed variance of the domain, and $2$ corresponds to the number of dimensions
+in our experiments\footnote{This calculation applies only in two-dimensional
+  domains, where $w_2 = 1-w_1$.}. Under this approach, evenly distributed
+weights correspond to an assumed variance of $\lambda$. Likewise, larger values
+of $w$, which are produced when the generated category is tightly distributed
+along one dimension, correspond to smaller assumed variances.
+
+Each model was used to simulate each participant's generation independently,
+with the importance of each dimension set according to the participant's
+generated category. The other free parameters within each model were set as in
+Table \ref{table:global-model-fits}. Every participant's generation was
+simulated 2,000 times; given the 305 participants tested across the two
+experiments, each model generated 610,000 categories in total. For comparison
+with our behavioral results, we then computed the range difference gradient
+identically as with the behavioral data. The results are shown in Figure
+\ref{fig:range-diff-gradients}.
+
+As in the more traditional model evaluation analysis described above, the
+contrast models (i.e., PACKER and the representativeness model) provided a much
+closer match to our behavioral results than the copy-and-tweak and hierarchical
+Bayesian models. In all conditions, the contrast models distribute categories
+similarly to the behavioral data: Horizontally-aligned categories tend to be
+placed above and below members of the experimenter-defined category, and
+vertically-aligned categories tend to be placed to the sides. Conversely,
+because the copy-and-tweak and hierarchical Bayesian models are insensitive to
+category contrast, these models do not produce any systematic patterns of
+association between category location and distributional structure. The sole
+exception is within the `Row' condition of Experiment 1, in which the majority
+of participants generated a `Row'-like category, widely distributed along the
+X-axis but not the Y-axis. In these cases, both models are initialized with
+weights that produce Row categories, but because category contrast is not
+considered, categories are uniformly generated across the entire domain, rather
+than concentrated within the upper-regions as observed behaviorally.
+
+%Contrast model distinction
+\section{Experiment 3: Contrasting Contrast}
+\label{section:diffcon}
+Experiments 1 and 2 clearly established the importance of contrast in category
+generation. Follow-up model-based analyses illustrated that both contrast models account for participant performance better than models that do not take into
+account contrast. We also found more support for the hypothesis of contrast as
+representativeness versus contrast as exemplar dissimilarity across the
+different Alpha conditions.
+
+While the representativeness model fits better than PACKER, and in spite of their fundamentally different structures,
+both contrast models make qualitatively similar predictions for Experiments 1 and 2. To illustrate this, Figure
+\ref{fig:firstbetas} shows how the models' generating probabilities for the first Beta exemplar are similar across both
+contrast models.\footnote{Model parameters are set to the optimized values from Table \ref{table:global-model-fits}.}
+Across the five Alpha conditions of Experiments 1 and 2, the first Beta exemplars according to the contrast models were
+generally more probable near the corners and edges. This is particularly the case for the representativeness model,
+which almost exclusively predicts generation at the corners, while PACKER can predict generation in the center of the
+feature space in the XOR condition.
+
+\begin{figure}[p]
+    \begin{center} 
+      \includegraphics[width=\textwidth]{figs/firstbetas.pdf}
+      \caption{Contrast model predictions indicating probability of generating
+        the first Beta exemplar for each Alpha condition. Darker shades of
+        blue indicate higher probabilities (normalized within each plot).}
+      \label{fig:firstbetas}
+    \end{center}
+\end{figure}
+
+The effect of contrast for both models is similar in that they both tend to
+produce exemplars that are distant from the Alpha category. However, this is
+realized in different ways for the two models. The representativeness model does
+not explicitly use a measure sensitive to distance from the location of
+exemplars in the contrasting category. Instead, it is the assumption of unimodal
+distributions that results in the model producing novel exemplars dissimilar to
+learned exemplars. Consequently, it is possible to design a configuration of
+Alpha exemplars where the representativeness model predicts novel exemplars
+should be \emph{similar} to the learned exemplars. However, PACKER cannot
+predict this because it assumes that the similarity weight for contrasting
+exemplars is negative.
+
+In this section, we focus on a different category type -- the Corner category --
+that contrasts the two models of contrast. Here, each Alpha exemplar is located
+at the corners of the feature space (see Figure \ref{fig:e3-squares}a). We first
+identify the predictions of each model in this condition. Second, we test these
+contrasting predictions in a behavioral experiment. We then conduct a
+model-based analysis of the two contrast models.
+
+In the Corner condition, the representativeness model estimates the underlying Alpha category distribution to be
+centered on the feature space, with probability densities highest in the center and lowest in the corners. The
+representativeness of novel Beta exemplars generally decreases with the likelihood of the underlying Alpha category
+distribution. Despite the Alpha exemplars being in the corners, the predicted mean will be near the center
+of the Alpha category (the center of the space; See \ref{fig:e3-squares}b). As such, the Alpha likelihood is lowest in
+the corners, and thus, the representativeness of the Beta category will be largest in the corners.
+
+Conversely, PACKER predicts a completely different result. With PACKER, the probability of generating a Beta exemplar
+increases with the dissimilarity from all Alpha exemplars. This results in Betas generated in the center of feature
+space, where the candidate exemplars are as far as possible from the Alpha exemplars (Figure \ref{fig:e3-squares}c).
+
+\begin{figure}[p]
+    \begin{center} 
+      \includegraphics[width=.8\textwidth]{figs/e3-squares.pdf}
+      \caption{Alpha exemplars for the Corner condition (a) and the predictions
+        from contrast models PACKER (b) and the representativeness model (c)
+        with square stimuli.}
+      \label{fig:e3-squares}
+    \end{center}
+\end{figure}
+
+Another aspect of the contrast in the models' predictions is the role of the boundaries of feature space. The
+representativeness model prefers to generate categories at the boundaries of feature space. PACKER is less sensitive to
+boundaries. Boundaries tend to be where distance to the contrasting category exemplars is maximized. But, as in cases
+like the Corner condition (Figure \ref{fig:e3-squares}a), this is not always the case; Beta exemplars should be as far
+from the Alphas as possible, which is maximized in the center of feature space.
+
+Thus, in this experiment, we also examine how making one feature toroidal (e.g.,
+line orientation) affects the prediction of each model. The representativeness
+model prefers to generate categories on the feature that still has a boundary.
+PACKER, in contrast, should continue to generate categories that are most
+distant from the learned category regardless of the presence or absence of feature
+boundaries -- in this condition it would be the center of the bounded feature.
+Experiment 3 manipulates whether one feature is bounded or toroidal (always
+displayed as the Y-axis). One condition uses the same stimuli as before, squares
+varying in lightness and size, both of which are bounded features. The other
+condition use a new set of stimuli: circles that vary in size (a bounded
+feature) and orientation (an unbounded feature; Figure \ref{fig:e3-circles}a).
+Before describing the details of the experiment, we explain in detail each
+model's predictions given the Corner Alphas for a feature space where one
+feature is bounded and the other is toroidal.
+
+\begin{figure}[p]
+    \begin{center} 
+      \includegraphics[width=.8\textwidth]{figs/e3-circles.pdf}
+      \caption{Circle stimulus domain (a) and predictions from contrast models
+        PACKER (b) and the representativeness model (c). The unbounded feature
+        (orientation) is represented on the Y-axis, while the bounded feature
+        (size) is represented on a X-axis. As with the square stimuli, feature
+        and direction assignment were counterbalanced across participants (in
+        the displayed example, circles are rotated clockwise up the Y-axis). }
+      \label{fig:e3-circles}
+    \end{center}
+\end{figure}
+
+Compared to stimuli with two bounded features, PACKER makes a similar prediction
+in generating the first Beta exemplar when one feature is unbounded, with the
+curious exception that the highest density regions appear compressed along the
+toroidal feature. Due to the toroidal nature of orientation, the maximum
+distance from the Alpha exemplars is half of the full range of the feature --
+for an Alpha exemplar oriented at 0 radians, the furthest distance to another
+exemplar is $\pi$ radians.
+
+For the representativeness model, adding a toroidal feature changes the
+probability distribution for the underlying Bayesian model. For the toroidal
+feature, space wraps back to $0$ radians once it reaches $2\pi$ (we have
+conducted a change of variables where $y \rightarrow y$ mod $2\pi$, where mod is
+the modulus operator). As we cross $2\pi$ and reach $0$ again, the remaining
+probability density also gets wrapped around, meaning that the density at $2\pi$
+is added to $0$ (and $2\pi + \epsilon$ to $\epsilon$, and so on). This is called
+the {\em wrapped} Gaussian distribution \citep{mardia1972}. Note that the more
+familiar distribution for circular features, the Von Mises distribution, is an
+approximation to the wrapped Gaussian distribution. See Appendix B for
+implementation details of the representativeness model in a bounded and
+toroidal feature space.
+
+The consequence of employing a wrapped Gaussian is presented in Figure \ref{fig:e3-circles}c. Edge effects along the
+unbounded feature (displayed as the vertical axis) are no longer present since the edges themselves no longer exist.
+Instead, the probability of generating the first Beta exemplar is bimodally distributed at the edges of the bounded
+feature (displayed as the horizontal axis). Thus, the representativeness model and PACKER make orthogonal predictions
+for how a toroidal dimension should affect category generation.
+
+In this experiment, we test human category generation in two cases where PACKER and the representativeness model
+contrast. The representativeness model predicts categories should be distributed in the boundaries of the bounded
+features. PACKER predicts category exemplars should be centered in space and decay exponentially over bounded
+dimension(s).
+
+\subsection{Participants, Materials, and Procedure}
+
+We recruited 89 participants from Amazon Mechanical Turk who were randomly
+assigned to the square stimuli ($N=46$), or to the circle stimuli ($N=43$). The
+square stimuli were exactly the same as in Experiments 1 and 2, but with Alpha
+exemplars in the corners of the feature space. The circle stimuli were made up
+of unfilled circles with a radial line extending from its center to its edge.
+These stimuli varied along two features: their diameter within the range of
+[3.0, 5.8] cm (inclusive), and their orientation in the range [$2.587$,
+$2\pi + 2.587$] radians (inclusive), with $0$ radians indicating a completely
+horizontal line and orientation increasing in an anti-clockwise direction. The
+line widths of the circles and radial line were set to 0.05 cm. Similar to
+Experiments 1 and 2, participants first observed a training phase where four
+Alpha exemplars were presented across three blocks followed by the generation
+phase, and the features values are divided into 9 discrete steps. The training
+condition was the Corner category for all participants.
+
+\subsection{Results and Discussion}
+
+%Repeat hypotheses
+Unlike previous experiments, Experiment 3 allows us to distinguish between PACKER and the representativeness model
+beyond their quantitative fits by identifying specific qualitatively diagnostic markers. Specifically, if PACKER is a
+better account of the data, we would expect a relative decrease in the range of the entire Beta category along the
+unbounded feature compared to the bounded feature. In addition, we would expect to see a preference for first-generated
+Beta exemplars away from the boundaries of the feature space -- that is, the probability of generating an exemplar would
+be negatively correlated with its distance from the center for bounded features. These predictions are correspondingly
+reversed if the representativeness model offers a better account of the data.
+%For our analyses, we define the edge to be the 1-tile wide border of each 9-by-9
+%large feature space.
+
+Figure \ref{fig:e3-statsboxes} presents the distributional statistics for both square and circle stimuli. Participants
+produced circle stimuli with significantly lower Y-range (lower for the unbounded feature) as compared to the square
+stimuli ($t(87) = 4.94$, $p = .001$), which supports PACKER. There are no significant differences between the
+distributional statistics of each generated category between square and circle stimuli (Figure \ref{fig:e3-statsboxes}).
+
+\begin{figure}
+    \begin{center}
+      \includegraphics[width=\textwidth]{figs/e3-statsboxes.pdf}
+    \caption{Box-plots of the distributional statistics from the categories
+generated in Experiment 3. Boxes depict the median and quartiles of each
+condition, with whiskers placed at 1.5 IQR. All points outside this region are
+marked individually. For circle stimuli, the Y Range represents orientation and
+the X Range represents size.}
+    \label{fig:e3-statsboxes}
+    \end{center}
+\end{figure}
+
+The distributions of generated first-Betas (Figure \ref{fig:e3-behave}) indicate
+further qualitative support for PACKER. In particular, the probability of
+generating the first Beta exemplar for circles along the bounded dimension is
+significantly negatively correlated with distance from the center
+($r(3) = -.95, p = .01$). Although the correlations for squares
+($r(3)= -.50, p = .40$ and $r(3)= -.31, p = .61$ for X and Y dimensions
+respectively) were not significant, they were in the directions predicted by
+PACKER. Mean probabilities of generating first-Betas are presented in Figure
+\ref{fig:e3-distcent}. This analysis was excluded for the toroidal dimension of
+the circle stimuli because the boundaries (and consequently a center) are not
+defined for this feature.
+
+\begin{figure}
+    \begin{center}
+      \includegraphics[width=.6\textwidth]{figs/e3-behave.pdf}
+    \caption{Heatmaps indicating the relative frequencies of generating the
+      first Beta exemplar in the Corner condition.}
+    \label{fig:e3-behave}
+    \end{center}
+\end{figure}
+
+\begin{figure}
+    \begin{center}
+      \includegraphics[width=.6\textwidth]{figs/e3-distcent.pdf}
+      \caption{Empirical mean probabilities of generating first Beta exemplars
+        at each point of distance from the center of the feature space.
+        Distances are normalized such that 0 represents the center and 1
+        represents the edge. Plot for the toroidal feature of circle stimuli is
+        not presented because a center is not defined for this feature.}
+    \label{fig:e3-distcent}
+    \end{center}
+\end{figure}
+
+Model generation probabilities of the first Beta exemplars do not qualitatively differ from earlier predictions even
+after fitting to the current dataset. Figures \ref{fig:e3-firstbetas}a-b and \ref{fig:e3-firstbetas}c-d show the model
+predictions for square and circle stimuli respectively. While the predictions for the representativeness model for
+circles appears different from earlier predictions (i.e., in Figure \ref{fig:e3-circles}c), note that our key
+expectation for this model was still met -- specifically, it predicts the highest first-Beta generation probabilities at
+the edges of the bounded features.
+
+% The only exception is that the representativeness model is able to predict decreasing generation probabilities
+% with increasing distance from the center of the feature space along the toroidal feature, but not the bounded feature.
+% As expected, the first Betas tend to be generated at the edges of the bounded feature.
+
+% After fitting the models to data from Experiment 3, there appears to be no major qualitative
+% deviations in the distribution of first-Betas for the square stimuli (Figures \ref{fig:e3-firstbetas}a and b) from
+% previous expectations (i.e., as seen in Figure \ref{fig:e3-squares}), however this is not the case for the circle
+% stimuli (Figures \ref{fig:e3-firstbetas}c and d compared to Figure \ref{fig:e3-circles})). Specifically, we see greater
+% variance along the unbounded feature for PACKER, and most strikingly, that the representativeness model is able to
+% expand the distribution of generating first-Betas into more centrally located regions \ref{fig:e3-firstbetas} for circle
+% stimuli. However as predicted, even with this distribution, the maximum probability of the representativeness model
+% generating at an edge (.0438) is higher than generating off-edge (.0394), indicating a preference for producing
+% first-Betas at the edge. This is in contrast to both PACKER and the empirical data, which both show a general preference
+% for centrally-located first-Betas.
+
+\begin{figure}
+    \begin{center}
+      \includegraphics[width=.6\textwidth]{figs/e3-firstbetas.pdf}
+      \caption{Heatmaps indicating the probabilities of each contrast model
+        generating the first Beta exemplar in the Corner condition for both
+        square and circle stimuli. Models have been fit to data from Experiment 3.}
+    \label{fig:e3-firstbetas}
+    \end{center}
+\end{figure}
+
+The contrast models were fit by optimizing them to only Experiment 3. Fit statistics and parameter values are presented
+in Table \ref{table:corner-model-fits}. Although not our main focus in this section, we also provide the statistics for
+the classical models -- unsurprisingly, they do not perform as well as their corresponding contrast models.
+Copy \& Tweak outperforms the Representativeness model.
+This is likely due to these categories being more challenging to encode with a Gaussian distribution than an exemplar-based model (and its corresponding nonparametric distribution).
+Overall, in line with our qualitative
+observations, PACKER emerges as the best-fitting model in aggregate.
+
+\begin{table} \centering
+  \caption{Results of model-fitting to the combined datasets from Experiments 1,
+    2 and 3. Smaller AIC values correspond to better model fits (adjusted
+    for number of parameters).}
+\label{table:corner-model-fits}
+
+\begin{tabular}{ l l l l} \\ \\ \textbf{PACKER} & \textbf{Copy \& Tweak} &
+\textbf{Representativeness} & \textbf{Hierarchical Bayesian} \\ 
+  \hline $AIC = 2954$ &$AIC = 2963$ &$AIC = 2971$  &$AIC = 3025$\\ 
+  $L = -1474$          &$L = -1479$   &$L = -1481$    &$L = -1508$\\ 
+  $c = 2.08$           &$c = 3.73$    &$\kappa = 0.00$&$\kappa = 0.29$  \\ 
+  $\theta_c = 0.87$    &$\theta_c = 0$ (fixed)  &$\nu = 4.07 \times 10^{20}$ &$\nu
+                                                                             = 1.00$\\
+  $\theta_t = 2.18$    &$\theta_t = 2.57$       &$\lambda = 0.14$ &$\lambda = 0.00$\\
+                       &              &$\theta = 0.73$ &$\theta = 7.50$\\
 \end{tabular}
-\footnotetext{Source: This is an example of table footnote. 
-This is an example of table footnote.}
-\footnotetext[1]{Example for a first table footnote.
-This is an example of table footnote.}
-\footnotetext[2]{Example for a second table footnote. 
-This is an example of table footnote.}
-\end{minipage}
-\end{center}
 \end{table}
-\end{verbatim}
-\bigskip
-%%=============================================%%
-%% For presentation purpose, we have included  %%
-%% \bigskip command. please ignore this.       %%
-%%=============================================%%
-
-\begin{table}[h]
-\begin{center}
-\begin{minipage}{\textwidth}
-\caption{Example of a lengthy table which is set to full textwidth}\label{tab2}
-\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}lcccccc@{\extracolsep{\fill}}}
-\toprule%
-& \multicolumn{3}{@{}c@{}}{Element 1\footnotemark[1]} & \multicolumn{3}{@{}c@{}}{Element 2\footnotemark[2]} \\\cmidrule{2-4}\cmidrule{5-7}%
-Project & Energy & $\sigma_{calc}$ & $\sigma_{expt}$ & Energy & $\sigma_{calc}$ & $\sigma_{expt}$ \\
-\midrule
-Element 3  & 990 A & 1168 & $1547\pm12$ & 780 A & 1166 & $1239\pm100$\\
-Element 4  & 500 A & 961  & $922\pm10$  & 900 A & 1268 & $1092\pm40$\\
-\botrule
-\end{tabular*}
-\footnotetext{Note: This is an example of table footnote.
-This is an example of table footnote this is an example of table footnote this is an example of~table footnote this is an example of table footnote.}
-\footnotetext[1]{Example for a first table footnote.}
-\footnotetext[2]{Example for a second table footnote.}
-\end{minipage}
-\end{center}
+
+Table \ref{table:corner_counts} presents frequencies of the best-fitting model
+for participants in each condition. Overall the representativeness model was
+significantly worse at capturing the results of Experiment 3 (it fit 26 out of
+the 89 participants better than PACKER; two-tailed Binomial test $p < 0.001$).
+Further, this is mostly due to the representativeness model capturing participants
+in the unbounded circle condition worse than PACKER (8 of 43 better fit by
+representativeness; two-tailed Binomial test, $p<0.001$). The representativeness
+model fit the bounded Square condition better, but still worse than PACKER (18 out of 46; two-tailed Binomial test $p = 0.18$).
+
+\begin{table} \centering
+  \caption{Frequencies of individual participants best fit by each contrast
+    model  broken down by stimulus type.}
+\label{table:corner_counts}
+\begin{tabular}{ l l l l l} \\ Stimulus Type & \textbf{PACKER} &
+                                                               \textbf{Representativeness}& Total & Two-tailed Binomial test $p$-value \\
+  \hline 
+  Squares & 28 & 18 & 46 & $p \approx 0.18$\\
+  Circles & 35 &  8 & 43 & $p < 0.001$\\
+  Total   & 60 & 26 & 89 & $p < 0.001$\\
+\end{tabular}
 \end{table}
 
-In case of double column layout, tables which do not fit in single column width should be set to full text width.
-For this, you need to use \verb+\begin{table*}+ \verb+...+ \verb+\end{table*}+ instead of \verb+\begin{table}+ \verb+...+ \verb+\end{table}+ environment.
-Lengthy tables which do not fit in textwidth should be set as rotated table.
-For this, you need to use \verb+\begin{sidewaystable}+ \verb+...+ \verb+\end{sidewaystable}+ instead of \verb+\begin{table*}+ \verb+...+ \verb+\end{table*}+ environment.
-This environment puts tables rotated to single column width.
-For tables rotated to double column width, use \verb+\begin{sidewaystable*}+ \verb+...+ \verb+\end{sidewaystable*}+.
-
-\begin{sidewaystable}
-\sidewaystablefn%
-\begin{center}
-\begin{minipage}{\textheight}
-\caption{Tables which are too long to fit, should be written using the ``sidewaystable'' environment as shown here}\label{tab3}
-\begin{tabular*}{\textheight}{@{\extracolsep{\fill}}lcccccc@{\extracolsep{\fill}}}
-\toprule%
-& \multicolumn{3}{@{}c@{}}{Element 1\footnotemark[1]}& \multicolumn{3}{@{}c@{}}{Element\footnotemark[2]} \\\cmidrule{2-4}\cmidrule{5-7}%
-Projectile & Energy	& $\sigma_{calc}$ & $\sigma_{expt}$ & Energy & $\sigma_{calc}$ & $\sigma_{expt}$ \\
-\midrule
-Element 3 & 990 A & 1168 & $1547\pm12$ & 780 A & 1166 & $1239\pm100$ \\
-Element 4 & 500 A & 961  & $922\pm10$  & 900 A & 1268 & $1092\pm40$ \\
-Element 5 & 990 A & 1168 & $1547\pm12$ & 780 A & 1166 & $1239\pm100$ \\
-Element 6 & 500 A & 961  & $922\pm10$  & 900 A & 1268 & $1092\pm40$ \\
-\botrule
-\end{tabular*}
-\footnotetext{Note: This is an example of table footnote this is an example of table footnote this is an example of table footnote this is an example of~table footnote this is an example of table footnote.}
-\footnotetext[1]{This is an example of table footnote.}
-\end{minipage}
-\end{center}
-\end{sidewaystable}
-
-\section{Figures}\label{sec6}
-
-As per the \LaTeX\ standards you need to use eps images for \LaTeX\ compilation and \verb+pdf/jpg/png+ images for \verb+PDFLaTeX+ compilation.
-This is one of the major difference between \LaTeX\ and \verb+PDFLaTeX+.
-Each image should be from a single input .eps/vector image file.
-Avoid using subfigures.
-The command for inserting images for \LaTeX\ and \verb+PDFLaTeX+ can be generalized.
-The package used to insert images in \verb+LaTeX/PDFLaTeX+ is the graphicx package.
-Figures can be inserted via the normal figure environment as shown in the below example:
-
-%%=============================================%%
-%% For presentation purpose, we have included  %%
-%% \bigskip command. please ignore this.       %%
-%%=============================================%%
-\bigskip
-\begin{verbatim}
-\begin{figure}[<placement-specifier>]
-\centering
-\includegraphics{<eps-file>}
-\caption{<figure-caption>}\label{<figure-label>}
-\end{figure}
-\end{verbatim}
-\bigskip
-%%=============================================%%
-%% For presentation purpose, we have included  %%
-%% \bigskip command. please ignore this.       %%
-%%=============================================%%
-
-\begin{figure}[h]%
-\centering
-\includegraphics[width=0.9\textwidth]{fig.eps}
-\caption{This is a widefig. This is an example of long caption this is an example of long caption  this is an example of long caption this is an example of long caption}\label{fig1}
-\end{figure}
-
-In case of double column layout, the above format puts figure captions/images to single column width. To get spanned images, we need to provide \verb+\begin{figure*}+ \verb+...+ \verb+\end{figure*}+.
-
-For sample purpose, we have included the width of images in the optional argument of \verb+\includegraphics+ tag. Please ignore this.
-
-\section{Algorithms, Program codes and Listings}\label{sec7}
-
-Packages \verb+algorithm+, \verb+algorithmicx+ and \verb+algpseudocode+ are used for setting algorithms in \LaTeX\ using the format:
-
-%%=============================================%%
-%% For presentation purpose, we have included  %%
-%% \bigskip command. please ignore this.       %%
-%%=============================================%%
-\bigskip
-\begin{verbatim}
-\begin{algorithm}
-\caption{<alg-caption>}\label{<alg-label>}
-\begin{algorithmic}[1]
-. . .
-\end{algorithmic}
-\end{algorithm}
-\end{verbatim}
-\bigskip
-%%=============================================%%
-%% For presentation purpose, we have included  %%
-%% \bigskip command. please ignore this.       %%
-%%=============================================%%
-
-You may refer above listed package documentations for more details before setting \verb+algorithm+ environment.
-For program codes, the ``program'' package is required and the command to be used is \verb+\begin{program}+ \verb+...+ \verb+\end{program}+.
-A fast exponentiation procedure:
-
-\begin{program}
-\BEGIN \\ %
-  \FOR i:=1 \TO 10 \STEP 1 \DO
-     |expt|(2,i); \\ |newline|() \OD %
-\rcomment{Comments will be set flush to the right margin}
-\WHERE
-\PROC |expt|(x,n) \BODY
-          z:=1;
-          \DO \IF n=0 \THEN \EXIT \FI;
-             \DO \IF |odd|(n) \THEN \EXIT \FI;
-\COMMENT{This is a comment statement};
-                n:=n/2; x:=x*x \OD;
-             \{ n>0 \};
-             n:=n-1; z:=z*x \OD;
-          |print|(z) \ENDPROC
-\END
-\end{program}
-
-\begin{algorithm}
-\caption{Calculate $y = x^n$}\label{algo1}
-\begin{algorithmic}[1]
-\Require $n \geq 0 \vee x \neq 0$
-\Ensure $y = x^n$ 
-\State $y \Leftarrow 1$
-\If{$n < 0$}\label{algln2}
-        \State $X \Leftarrow 1 / x$
-        \State $N \Leftarrow -n$
-\Else
-        \State $X \Leftarrow x$
-        \State $N \Leftarrow n$
-\EndIf
-\While{$N \neq 0$}
-        \If{$N$ is even}
-            \State $X \Leftarrow X \times X$
-            \State $N \Leftarrow N / 2$
-        \Else[$N$ is odd]
-            \State $y \Leftarrow y \times X$
-            \State $N \Leftarrow N - 1$
-        \EndIf
-\EndWhile
-\end{algorithmic}
-\end{algorithm}
-\bigskip
-%%=============================================%%
-%% For presentation purpose, we have included  %%
-%% \bigskip command. please ignore this.       %%
-%%=============================================%%
-
-Similarly, for \verb+listings+, use the \verb+listings+ package.
-\verb+\begin{lstlisting}+ \verb+...+ \verb+\end{lstlisting}+ is used to set environments similar to \verb+verbatim+ environment.
-Refer to the \verb+lstlisting+ package documentation for more details.
-
-%%=============================================%%
-%% For presentation purpose, we have included  %%
-%% \bigskip command. please ignore this.       %%
-%%=============================================%%
-\bigskip
-\begin{minipage}{\hsize}%
-\lstset{frame=single,framexleftmargin=-1pt,framexrightmargin=-17pt,framesep=12pt,linewidth=0.98\textwidth,language=pascal}% Set your language (you can change the language for each code-block optionally)
-%%% Start your code-block
-\begin{lstlisting}
-for i:=maxint to 0 do
-begin
-{ do nothing }
-end;
-Write('Case insensitive ');
-Write('Pascal keywords.');
-\end{lstlisting}
-\end{minipage}
-
-\section{Cross referencing}\label{sec8}
-
-Environments such as figure, table, equation and align can have a label
-declared via the \verb+\label{#label}+ command. For figures and table
-environments use the \verb+\label{}+ command inside or just
-below the \verb+\caption{}+ command. You can then use the
-\verb+\ref{#label}+ command to cross-reference them. As an example, consider
-the label declared for Figure~\ref{fig1} which is
-\verb+\label{fig1}+. To cross-reference it, use the command 
-\verb+Figure \ref{fig1}+, for which it comes up as
-``Figure~\ref{fig1}''.
-
-To reference line numbers in an algorithm, consider the label declared for the line number 2 of Algorithm~\ref{algo1} is \verb+\label{algln2}+.
-To cross-reference it, use the command \verb+\ref{algln2}+ for which it comes up as line~\ref{algln2} of Algorithm~\ref{algo1}.
-
-\subsection{Details on reference citations}\label{subsec7}
-
-Standard \LaTeX\ permits only numerical citations.
-To support both numerical and author-year citations this template uses \verb+natbib+ \LaTeX\ package.
-For style guidance please refer to the template user manual.
-
-Here is an example for \verb+\cite{...}+: \cite{bib1}.
-Another example for \verb+\citep{...}+: \citep{bib2}.
-For author-year citation mode, \verb+\cite{...}+ prints Jones et al.
-(1990) and \verb+\citep{...}+ prints (Jones et al., 1990).
-
-All cited bib entries are printed at the end of this article: \cite{bib3}, \cite{bib4}, \cite{bib5}, \cite{bib6}, \cite{bib7}, \cite{bib8}, \cite{bib9}, \cite{bib10}, \cite{bib11} and \cite{bib12}.
-
-\section{Examples for theorem like environments}\label{sec10}
-
-For theorem like environments, we require \verb+amsthm+ package. There are three types of predefined theorem styles exists---\verb+thmstyleone+, \verb+thmstyletwo+ and \verb+thmstylethree+
-
-%%=============================================%%
-%% For presentation purpose, we have included  %%
-%% \bigskip command. please ignore this.       %%
-%%=============================================%%
-\bigskip
-\begin{tabular}{|l|p{19pc}|}
-\hline
-\verb+thmstyleone+ & Numbered, theorem head in bold font and theorem text in italic style \\\hline
-\verb+thmstyletwo+ & Numbered, theorem head in roman font and theorem text in italic style \\\hline
-\verb+thmstylethree+ & Numbered, theorem head in bold font and theorem text in roman style \\\hline
-\end{tabular}
-\bigskip
-%%=============================================%%
-%% For presentation purpose, we have included  %%
-%% \bigskip command. please ignore this.       %%
-%%=============================================%%
-
-For mathematics journals, theorem styles can be included as shown in the following examples:
-
-\begin{theorem}[Theorem subhead]\label{thm1}
-Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. 
-Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. 
-Example theorem text. 
-\end{theorem}
-
-Sample body text. Sample body text. Sample body text. Sample body text. Sample body text. Sample body text. Sample body text. Sample body text.
-
-\begin{proposition}
-Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text. 
-Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text. 
-\end{proposition}
-
-Sample body text. Sample body text. Sample body text. Sample body text. Sample body text. Sample body text. Sample body text. Sample body text.
-
-\begin{example}
-Phasellus adipiscing semper elit. Proin fermentum massa
-ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
-at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. 
-\end{example}
-
-Sample body text. Sample body text. Sample body text. Sample body text. Sample body text. Sample body text. Sample body text. Sample body text.
-
-\begin{remark}
-Phasellus adipiscing semper elit. Proin fermentum massa
-ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
-at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. 
-\end{remark}
-
-Sample body text. Sample body text. Sample body text. Sample body text. Sample body text. Sample body text. Sample body text. Sample body text.
-
-\begin{definition}[Definition sub head]
-Example definition text. Example definition text. Example definition text. Example definition text. Example definition text. Example definition text. Example definition text. Example definition text. 
-\end{definition}
-
-Additionally a predefined ``proof'' environment is available: \verb+\begin{proof}+ \verb+...+ \verb+\end{proof}+.
-This prints a ``Proof'' head in italic font style and the ``body text'' in roman font style with an open square at the end of each proof environment.
-
-\begin{proof}
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-\end{proof}
-
-Sample body text. Sample body text. Sample body text. Sample body text. Sample body text. Sample body text. Sample body text. Sample body text.
-
-\begin{proof}[Proof of Theorem~{\upshape\ref{thm1}}]
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-\end{proof}
-
-\noindent
-For a quote environment, use \verb+\begin{quote}...\end{quote}+
-\begin{quote}
-Quoted text example. Aliquam porttitor quam a lacus. Praesent vel arcu ut tortor cursus volutpat. In vitae pede quis diam bibendum placerat. Fusce elementum
-convallis neque. Sed dolor orci, scelerisque ac, dapibus nec, ultricies ut, mi. Duis nec dui quis leo sagittis commodo.
-\end{quote}
-
-Sample body text. Sample body text. Sample body text. Sample body text. Sample body text (refer Figure~\ref{fig1}). Sample body text. Sample body text. Sample body text (refer Table~\ref{tab3}).
-
-\section{Methods}\label{sec11}
-
-Topical subheadings are allowed.
-Authors must ensure that their Methods section includes adequate experimental and characterization data necessary for others in the field to reproduce their work.
-Authors are encouraged to include RIIDs where appropriate.
-
-\textbf{Ethical approval declarations} (only required where applicable) Any article reporting experiment/s carried out on (i)~live vertebrate (or higher invertebrates), (ii)~humans or (iii)~human samples must include an unambiguous statement within the methods section that meets the following requirements:
-
-\begin{enumerate}[1.]
-\item Approval: a statement which confirms that all experimental protocols were approved by a named institutional and/or licensing committee. Please identify the approving body in the methods section
-
-\item Accordance: a statement explicitly saying that the methods were carried out in accordance with the relevant guidelines and regulations
-
-\item Informed consent (for experiments involving humans or human tissue samples): include a statement confirming that informed consent was obtained from all participants and/or their legal guardian/s
-\end{enumerate}
-
-If your manuscript includes potentially identifying patient/participant information, or if it describes human transplantation research, or if it reports results of a clinical trial then  additional information will be required.
-Please visit (\url{https://www.nature.com/nature-research/editorial-policies}) for Nature Portfolio journals, (\url{https://www.springer.com/gp/authors-editors/journal-author/journal-author-helpdesk/publishing-ethics/14214}) for Springer Nature journals, or (\url{https://www.biomedcentral.com/getpublished/editorial-policies\#ethics+and+consent}) for BMC.
-
-\section{Discussion}\label{sec12}
-
-Discussions should be brief and focused.
-In some disciplines use of Discussion or `Conclusion' is interchangeable.
-It is not mandatory to use both.
-Some journals prefer a section `Results and Discussion' followed by a section `Conclusion'.
-Please refer to Journal-level guidance for any specific requirements.
-
-\section{Conclusion}\label{sec13}
-
-Conclusions may be used to restate your hypothesis or research question, restate your major findings, explain the relevance and the added value of your work, highlight any limitations of your study, describe future directions for research and recommendations.
-
-In some disciplines use of Discussion or 'Conclusion' is interchangeable. It is not mandatory to use both. Please refer to Journal-level guidance for any specific requirements.
+Ultimately, these results suggest that PACKER's exemplar dissimilarity account
+of contrast, as opposed to the representativeness account, offers a better
+explanation of category generation for the Corner category type. On almost every
+measure of analysis in this section, PACKER outperformed the representativeness
+model, indicating that the better fit of the latter model demonstrated in
+Experiments 1 and 2 is not general to all conditions.
+
+\section{General Discussion}
+The sensory impression of every stimulus and event is unique. Grouping distinct
+patterns of sensory information into categories is a fundamental task solved by
+the mind. Most work has focused on how people learn new categories that are
+provided to them through unlabeled, partially labeled, or fully labeled
+examples. How were these categories first determined? Some natural categories
+are likely to be the result of regularities in the dynamics of our environment.
+But, these are only a subset of the categories that people learn. Other
+categories, such as tools and ideas, were generated by people over time. What
+basic principles underlie how people generate categories?
+
+While the bulk of prior research on categorization has focused on the classic
+finding that generated concepts tend to be distributionally similar to known
+concepts, there has been little work addressing the role of contrast in category
+generation: How is it that people are able to create something {\em different}
+from what is already known? We developed two novel models, each incorporating a
+different conceptualization of category contrast. One model is PACKER, which is
+an exemplar-based model that formally specifies the role of contrast in
+generation as exemplar dissimilarity. Specifically, the model proposes that
+categories are represented as exemplars in a multidimensional psychological
+space, and generation is constrained both by within-category and
+between-category similarity: Exemplars belonging to the same category should be
+similar to one another, and exemplars belonging to different categories should
+not be similar to one another. The second model is a novel hierarchical Bayesian
+model with a representativeness mechanism. This model generates exemplars that
+are more representative of (i.e., has greater relative evidence from) the novel
+category compared to the learned category.
+
+We reported two experiments demonstrating systematic effects of category
+contrast in category generation. Members of participant-generated categories
+tended to be highly dissimilar from members of previously-learned categories,
+and were usually more similar to one another than to members of other
+categories. We also observed broad interdependence between the distributional
+structure (feature variance, correlation) and physical instantiation (location
+within the feature space) of generated categories: In Experiment 2, we found
+that the unoccupied regions of the domain influenced the distributional
+structure of categories, and in both experiments we observed that participants
+distributed their generated categories to increase contrast with what was
+already known.
+
+We conducted simulations comparing the contrast models' account of our results
+to the classical proposals for category generation: a ``copy-and-tweak'' model
+(realized as a variant of PACKER with no sensitivity to category contrast), and
+a hierarchical Bayesian model designed to explain the classic distributional
+similarity effect. In all simulations, we found that the contrast models
+captured a previously unexplained and unexplored aspect of human category
+generation. In particular, by measuring PACKER's fit as a function of its
+prioritization of within- and between-category similarity, we observed that
+considering either constraint exclusively results in a relatively low-quality
+account. Instead, PACKER's best results were obtained when both constraints are
+considered, indicating that human learners do not generate novel concepts
+exclusively on the basis of within-category similarity or between
+class-contrast. This finding mirrors our behavioral results and demonstrates
+that both constraints influence generation when explained with an exemplar
+model.
+
+While we found that the representativeness model was found to be a better fit to
+data from Experiments 1 and 2, both contrast models essentially made
+qualitatively similar predictions across all conditions in those experiments. In
+Experiment 3, we collected behavioral data on a category generation condition
+that was designed to qualitatively distinguish the two contrast models. Specifically, in this
+condition where exemplars were located at the corners of the feature space,
+PACKER predicted the initial generation of centrally-located novel exemplars
+while the representativeness model predicted novel exemplars similarly located
+at the corners. Interestingly, data from this experiment were better fit,
+qualitatively and quantitatively, by PACKER than the representativeness model.
+
+\subsection{Similarity and Contrast in Cognition}
+We propose category contrast as a primary constraint in categorization. For
+categories to be useful, they should not all be identical, or in other words,
+they should be different. Thus, a newly generated category should be different
+than pre-existing categories. Beyond its role in category generation
+specifically, category contrast is also of fundamental importance in
+categorization more broadly. All other factors held constant, new categories are
+easier to learn if they are dissimilar to members of other categories, and
+knowledge of highly distinct categories is applied more accurately than that of
+ill-defined categories \citep{ashby1994categorization,imai1965discriminability}.
+Likewise, basic-level categories \citep{rosch1976basic} are thought to be
+abstracted in order to maximize within-category similarity while minimizing
+between-category similarity. Finally, the act of forming category
+representations affects similarity judgments about category members and
+nonmembers, with category members being viewed as more similar to one another
+than members of other categories
+\citep{goldstone1994influences,goldstone2001altering,goldstone1996isolated}.
+
+Beyond the traditional categorization literature, one can find instances of the trade-off between within
+and between-class similarity in linguistic categories over perceptual
+dimensions. For example, \cite{regier2007} showed that the partitioning of color
+categories reflects such a trade-off in a psychological space -- colors are
+partitioned into groups with members that are viewed as highly similar to one
+another yet distinct from other colors. A similar trade-off can be observed in
+phoneme categories. Different exemplars of the same phoneme must be similar to
+one another, while contrasting from other phonemes, such that a listener can
+infer the appropriate phoneme. This pattern has been found and modeled in the
+natural acoustics of American English vowels
+\citep{feldman2013,hillenbrand1995}. As linguistic categories must have been
+created at some point in human history, it is revealing that the constraints of
+emulating distributional structure across categories and having categories
+contrast from one another still bias human category generation today.
+
+The dual forces of within-class similarity and between-class contrast influence
+cognitive functions in a wide variety of domains. The PACKER model is notable in
+that it explicitly interprets this trade-off within the domain of
+categorization, and allows us to begin to understand the relatively understudied
+processes involved in category generation through our more well-developed tools
+for understanding human categorization.
+
+\subsection{Implications for Creative Cognition}
+
+Although the focus of this article has been to address the role of contrast in category generation, our findings and approach have relevant implications
+for research in creative cognition. A central focus of the creative cognition
+approach has been to explain acts of creativity in terms of the mental
+representations and processes that are commonly studied in cognitive psychology
+and cognitive science \citep{finke1992creative,smith1995creative}. However,
+unlike other fields in the study of cognition, creative cognition research
+rarely employs quantitative models to evaluate the explanatory value of such
+representations and processes. Our modeling results provide a concrete example
+of how formal approaches may be used to gain insight into the nature of creative
+cognition.
+
+In addition to demonstrating the utility of formal modeling for studying
+creative cognition, the contrast models here specifically offer an additional
+interpretation of some of the field's most central findings. For example,
+perhaps the most foundational principle from this literature concerns the
+limiting influence of prior knowledge: Individuals create new categories
+composed of features from existing classes, and what is created can be
+influenced drastically through the introduction of cues or examples
+\citep{marsh1999inadvertent,smith1993constraining}. In this paper, we have
+identified another important aspect of the constraining influence of prior
+knowledge: What is generated cannot be the same as what is already known.
+Further, there is systematicity in how generated categories differ from prior
+knowledge. The results of our simulations suggest that this conceptualization of
+difference can be addressed in at least two different ways. Specifically,
+simulations with PACKER demonstrate that the constraining influence of
+difference is concisely explained in terms of a trade-off between
+within-category similarity and between-category dissimilarity. Conversely, the
+representativeness model shows that this influence can also be the result of
+enhancing the representativeness of generated exemplars to their category.
+
+PACKER may offer an additional interpretation of existing accounts of creative
+generation. Most notably, a leading account within the creative cognition
+literature, the Path of Least Resistance \citep{ward1994structured,ward1995s},
+also explains generation in terms of an exemplar-based retrieval process. This
+account was designed to explain the creative generation of natural categories
+(e.g., new species of plants and animals) and as a result relies strongly on the
+hierarchical organization of these categories: Individuals are thought to
+retrieve an example of the higher-level category being generated (e.g.,
+\textit{bird} may be retrieved from the category \textit{animal}), and then
+systematically alter what was retrieved to make something new. As the PACKER
+model does not assume knowledge is hierarchically organized \citep[this is true
+of the exemplar view more broadly, see][]{murphy2016exemplar}, the model may be
+viewed as a formal instantiation of the Path of Least Resistance for application
+in a traditional artificial categorization domain (when there is no established
+hierarchy of categories). PACKER's success in explaining generation within an
+artificial domain motivates future work exploring the nature of category
+contrast within a more naturalistic setting.
+
+The broader study of creativity currently involves a wide breadth of different
+approaches \citep[for a review, see][]{kozbelt2010theories}, such as those based
+on free association \citep{mednick1962associative} and conceptual combination
+\citep{estes2002emergence,murphy1988comprehending}. In addition, recent work in
+the machine learning literature has explored using neural networks to address
+the overall problem of creative generation
+\citep[e.g.,][]{goodfellow2014gan,ho2016gail,kingma2016,chen2016infogan}. In
+contrast to these varied investigations, we reduced our focus by studying a
+highly complex behavior (category generation) as it applies within a
+well-established domain (artificial category learning). We hope that future work
+incorporates and highlights the importance of contrast into theories of
+creativity and adapts the discussed contrast models to work in state-of-the-art creative generation methods.
+
+\subsection{Implications For Categorization}
+
+Categorization research addresses the representations and processes that
+underlie the learning and use of categories. Category learning tasks are
+generally about figuring out which items belong to which category. Once learned,
+categories are generally used to classify new stimuli and to make inferences
+beyond the available information. Our work is fairly unique in that people learn
+a category through positive examples and then create another category that would
+make sense in the domain. Such scenarios have direct application to real-world
+situations. For instance, consider a rock musician in the 1970s who wants to
+write new types of rock music. Given the diverse types of existing rock music that the
+musician is aware of, the musician must identify features of this new sub-genre
+that are unique to that sub-genre (e.g., introduce instruments that may not
+typical to rock music at the time, such as accordions or synthesizers). In the
+present studies, we have learned something about the form that such expectations
+on new categories are likely to take.
+
+We can think of the category generation task in our studies as asking a person
+to formulate an idea about what set of items in the domain are most
+interestingly {\em not} members of the original category. To meet this
+condition, the items must take some form of coherence that aligns with that of
+the original category and some form of distinctiveness relative to the original
+category. Reflecting the basic level of organization in natural categories, it
+makes sense to generate a set of items that possess strong within-category
+coherence (by importing or systematically transforming the internal structure of
+the original category) as well as strong between-category differentiation (by
+creating maximum contrast with the original category, be it through
+exemplar-based dissimilarity or the maximization of exemplar-category
+representativeness). In this sense one can see the patterns of performance in
+the category generation task as recapitulating the order of semantic
+organization.
+
+The current work also suggests exciting directions for related investigations
+into unsupervised categorization. While the unsupervised categorization
+literature is primarily interested in the generation of categories within a set
+of observed exemplars where no prior categories are learned
+\citep[e.g.,][]{pothos2011measuring}, our category generation studies are
+focused on the production of novel category exemplars themselves. Despite this
+distinction, both types of categorization research are ultimately interested in
+the question of how categories can be formed. The results of our current work
+indicate that the formation of categories is constrained by some measure of
+contrast. To our knowledge, this explicit investigation of category contrast has
+not yet been attempted in the unsupervised categorization literature.
+
+Interestingly, the unidimensional sort bias that is commonly observed in
+unsupervised categorization
+\citep{imai1965discriminability,milton2004influence,ahn1992two}, where
+participants sort unlabeled family resemblance category exemplars by focusing
+on a single feature, was not consistently observed in our tasks. In our
+experiments we only observed this bias in the row condition of Experiment 1,
+whereas in other conditions participants generated categories that typically
+varied along both dimensions. This may be unsurprising at face value, since
+participants in conditions other than the row condition were trained on a prior
+category that varied on both dimensions. However, these results place constraints on how broadly applicable the 
+unidimensional bias is to different types of category construction. Additionally, the
+unidimensional sort bias is observed in category construction tasks in which
+participants develop a category-level organization of a provided domain while
+category generation tasks involve generating a new category relative to one that
+was just acquired (one might say that participants are forging a domain from a
+category rather than categories from a domain). Since the more creative task of
+generating a category from scratch (with respect to an established category)
+does not invoke the unidimensional bias, this suggests that a more creatively
+demanding version of a sort task might also reduce the bias.
+
+\subsection{Exemplar Dissimilarity and Representativeness in Categorization} 
+Although both PACKER and the representativeness model perform better than the
+classical models, they appear to be capturing fundamentally different aspects of
+contrast in categorization. Specifically, the representativeness model excelled
+in predicting categories generated by participants in Experiments 1 and 2.
+Conversely, PACKER was the best-fitting model in accounting for categories
+generated from the Corner category type, especially when the stimuli had a
+toroidal feature.
+
+% Given the history of their development, this is perhaps not entirely
+% surprising. PACKER is implemented as an extension of the GCM, which was
+% originally developed for (and has been highly successful in) category learning
+% \citep{nosofsky1986attention, nosofsky1994comparing}. Conversely, the
+% representativeness mechanism was conceived as a heuristic describing the
+% relative extent to which an individual observation is generated from some
+% unobserved population \citep{kahneman1972subjective,abbott16}. This generative
+% nature of the representativeness mechanism intuitively lends itself to the
+% prediction of category generation data.
+
+While the contrast models were born quite naturally out of the corresponding
+classical models, it is worth considering if similar benefits in prediction
+would be seen if the contrast mechanisms were applied to different classical
+models. Although there is no clear way to adapt the exemplar dissimilarity
+mechanism to the current hierarchical Bayesian model, it is a fairly
+straightforward process applying the representativeness mechanism to an exemplar
+model. Specifically, by treating the similarity measure (Equation
+\ref{eq:similarity}) in the current implementation of copy-and-tweak as a
+density estimate for the representativeness mechanism (Equation
+\ref{eq:representativeness}), we gain a new model of representativeness that is
+based on exemplar similarity instead of a multivariate Gaussian
+likelihood.\footnote{An analogous implementation of the representativeness
+  mechanism to PACKER results in both $\theta_c$ and $\theta_t$ adding to
+  constant value that is independent of the similarity to contrast and target
+  exemplars respectively. Consequently, it is formally equivalent to a
+  copy-and-tweak model with representativeness.}
+
+How well does a representativeness model that uses exemplar similarity to
+represent the category's density over features explain our experimental results?
+We find that for data from experiments 1 and 2, this model substantially
+outperforms its classical counterpart, but not to the extent of PACKER or the
+representativeness model. Specifically, the model fits here yielded $L = -4815$
+and $AIC=9633$ for the entire set of data from Experiments 1 and 2 and
+$L = -3474$, $AIC=6953$ for the same data excluding the first trials (cf. Table
+\ref{table:global-model-fits}). When we fit the model to data from Experiment 3,
+we find that $L = -1486$, $AIC = 2975$, which indicates a poorer fit than its
+classical counterpart (cf. Table \ref{table:corner-model-fits}). Overall, the
+representativeness model performed best when generating categories where
+previous category examples were not at the edges of the feature space.
+Conversely, PACKER performed better when the previous category was at the
+extremes of the features or if a feature is bounded. These results suggest that
+the predictive advantages to including the representativeness mechanism is not
+limited to a hierarchical Bayesian model, but can also be found when applied to
+an exemplar model. Although its performance is not as strong as either of the
+fully-developed contrast models, there can be a benefit to the
+representativeness mechanism that is independent from any interaction with a
+hierarchical Bayesian framework. Ultimately, it appears that there is no single
+model that can capture the entirety of our empirical data.
+
+\subsection{Limitations and Future Directions}
+
+Although successful in explaining our results in all three studies (albeit not
+as well as the representativeness model for Experiments 1 and 2), PACKER does
+not provide a full account of what is known about category generation. Most
+notably, in this paper we have not evaluated the model's ability to explain the
+classic finding that generated categories tend to share distributional
+commonalities with previously learned categories
+\citep[see][]{jern2013probabilistic,ward1994structured}. While we successfully
+replicated this effect in Experiment 1, we also found that its influence was
+limited in comparison to the fundamental constraints imposed by category
+contrast. Even within Experiment 1, we found systematic inconsistencies: by
+generating exemplars into unoccupied regions of the space, participants who
+learned an `XOR' category, composed of members that are widely distributed along
+both features and are positively correlated in space, tended to generate
+categories with an opposite (negative) correlation. More generally, PACKER
+inherits the strengths and weaknesses of exemplar models of categorization: It
+provides a simple and flexible model that explains many results, but deviates
+systematically from human performance in some cases.
+
+Nonetheless, these classic effects are a core element of the phenomenology of
+category generation, and PACKER does not include any mechanisms that explain
+them. Instead, through the development and evaluation of the PACKER model, we
+have sought to add new elements into such a phenomenology: The broad and strong
+influence of category contrast, and the interdependence between category
+location and distributional structure. It may be possible to combine the
+hierarchical Bayesian approach proposed by \cite{jern2013probabilistic} with
+PACKER's underlying claims to obtain a ``best of both worlds'' model, capable of
+explaining the role of contrast in category generation, as well as the emulation
+of distributional structure. However, as noted in the introduction, the
+incorporation of category contrast is antithetical to the core principles of a
+traditional, semi-conjugate Bayesian approach. This suggests that category
+generation is a fundamentally different computational-level problem
+\citep[different from those posed by][]{jern2013probabilistic,kemp2014taxonomy}, which is supported by the representativeness form of their model performing much better.
+
+Characterizing that problem and conducting a rational analysis is an important
+direction for future research. To that aim, we plan to explore the connection
+between exemplar modeling as an Importance Sampling approximation
+\citep{shi10exemplar}, and see what sort of computational-level problem PACKER
+approximates. Once formalized in probabilistic terms, it should also be
+straightforward to incorporate distributional factors into the model. This would
+unite PACKER with the Bayesian representativeness model, and they would differ
+in terms of their assumptions about how people represent category distributions
+(as exemplars or prototypes, respectively). Alternatively, it may be possible to
+integrate the core principles of either model of contrast into other
+categorization models
+\citep[e.g.,][]{love2004sustain,kurtz2007divergent,smith2000thirty}.
+
+The present work focuses on the influence of contrast on categorization,
+primarily by exploring two different conceptualizations of contrast. However,
+throughout this paper we have only explored one type of category generation
+problem, that is the generation of new categories in an
+artificial domain. It is possible that the strength of the influence of contrast on
+category generation decisions varies depending on the nature of the category
+generation problem, for instance, in situations where the boundaries of the
+domain are not clearly defined. To provide some intuition for this, consider the
+following example: an entomologist is asked to draw a new type of insect that
+they have never seen. Clearly, features of the new insect will be constrained by
+the definition of an insect: it will be an arthropod with six legs and a
+relatively hard exoskeleton. However, there are no strict limitations on other
+features such as how large the insect needs to be, or if the insect should have
+a particular type of pattern on its body. In this scenario, it is plausible that
+the entomologist will draw an insect that is fairly similar to insects that
+currently exist (e.g., small in size with a camouflage pattern that matches its
+environment) as opposed to something completely different (e.g., a horse-sized
+beetle with roman numerals on its back), indicating that contrast plays only a
+minor role in the decision making process. In this situation it appears that
+there is a combination of clearly defined constraints on the domain boundary
+that are imposed by the external environment (insects must be arthropods with
+six legs) but also ill-defined constraints that are imposed by the observer
+(insects are probably small and have patterns on their body that suit the
+environment). However, the relation of these constraints (and the interactions
+between them) to the role of contrast is not currently well understood.
+
+The definition of the domain boundary is only one possible factor that could
+influence the effect of contrast on categorization. It is also plausible that
+other conditions such as the nature of the instruction (``Generate a new
+category that is Not Alpha'' vs ``Generate a new category that is Beta'') 
+can affect the extent to which contrast influences categorization. Though our
+current contrast models can allow the influence of contrast to vary, they do not
+make any predictions regarding the specific conditions under which contrast can
+vary. Consequently, addressing this issue would be an immensely promising avenue
+for future research.
+
+\section{Conclusions}
+
+The generation of new concepts and ideas is a highly interesting topic, but it
+is difficult to study in a controlled experimental environment. In this paper,
+we have provided such an examination of category generation as it applies within
+an artificial categorization experiment. Extending the literature on creative
+cognition, our experiments provide a detailed picture of the role of category
+contrast in generation: People seek to create concepts that are distinct from
+what they already know, and the nature of what is created can be influenced by
+what does not yet exist. Our simulations with traditional exemplar models, as
+well as a hierarchical Bayesian model, provide strong support for the claim that
+category contrast is of fundamental importance to categorization.
 
@@
 
-\bmhead{Supplementary information}
-
-If your article has accompanying supplementary file/s please state so here.
-
-Authors reporting data from electrophoretic gels and blots should supply the full unprocessed scans for key as part of their Supplementary information.
-This may be requested by the editorial team/s if it is missing.
-
-Please refer to Journal-level guidance for any specific requirements.
+All code and data will be publicly available after manuscript acceptance and are available now by request.
 
@@
 
-Acknowledgments are not compulsory. Where included they should be brief. Grant or contribution numbers may be acknowledged.
-
-Please refer to Journal-level guidance for any specific requirements.
-
-\section*{Declarations}
-
-Some journals require declarations to be submitted in a standardised format.
-Please check the Instructions for Authors of the journal to which you are submitting to see if you need to complete this section.
-If yes, your manuscript must contain the following sections under the heading `Declarations':
-
-\begin{itemize}
-\item Funding
-\item Conflict of interest/Competing interests (check journal-specific guidelines for which heading to use)
-\item Ethics approval 
-\item Consent to participate
-\item Consent for publication
-\item Availability of data and materials
-\item Code availability 
-\item Authors' contributions
-\end{itemize}
-
-\noindent
-If any of the sections are not relevant to your manuscript, please include the heading and write `Not applicable' for that section.
-
-%%===================================================%%
-%% For presentation purpose, we have included        %%
-%% \bigskip command. please ignore this.             %%
-%%===================================================%%
-\bigskip
-\begin{flushleft}%
-Editorial Policies for:
-
-\bigskip\noindent
-Springer journals and proceedings: \url{https://www.springer.com/gp/editorial-policies}
-
-\bigskip\noindent
-Nature Portfolio journals: \url{https://www.nature.com/nature-research/editorial-policies}
-
-\bigskip\noindent
-\textit{Scientific Reports}: \url{https://www.nature.com/srep/journal-policies/editorial-policies}
-
-\bigskip\noindent
-BMC journals: \url{https://www.biomedcentral.com/getpublished/editorial-policies}
-\end{flushleft}
+Previous versions of this work were presented at the
+Thirty-Ninth Annual Conference of the Cognitive Science Society and Forty-Ninth
+Annual Meeting of the Society for Mathematical Psychology. Support for this
+research was provided by the Office of the VCRGE at the UW-Madison with
+funding from the WARF and the Vilas Life Cycle Fellowship. We thank Charles Kemp
+for feedback on a previous version of this manuscript. We also thank him and
+Alan Jern for providing code and data.
+
+\bmhead{Conflict of Interest}
+On behalf of all authors, the corresponding author states that there is no conflict of interest.
 
@@
 
-\section{Section title of first appendix}\label{secA1}
-
-An appendix contains supplementary information that is not an essential part of the text itself but which may be helpful in providing a more comprehensive understanding of the research problem or it is information that is too cumbersome to be included in the body of the paper.
-
-%%=============================================%%
-%% For submissions to Nature Portfolio Journals %%
-%% please use the heading ``Extended Data''.   %%
-%%=============================================%%
-
-%%=============================================================%%
-%% Sample for another appendix section			       %%
-%%=============================================================%%
-
-%% \section{Example of another appendix section}\label{secA2}%
-%% Appendices may be used for helpful, supporting or essential material that would otherwise 
-%% clutter, break up or be distracting to the text. Appendices can consist of sections, figures, 
-%% tables and equations etc.
+\section{The Hierarchical Bayesian Model of Concept Generation}
+\label{ap:hsampling-definition}
+
+\cite{jern2013probabilistic} demonstrated how a hierarchical Bayesian model
+could explain the distributional correspondences between observed and generated
+categories. In their model, exemplars of generated categories were viewed as
+samples from a multivariate Normal distribution over the dimensions of stimulus
+space. The mean of the generated category was independent of the observed
+categories, but the covariance matrix (encoding feature variances and
+correlations) was based on a common prior distribution. Generating a new
+category was thus completed by sampling a new category mean (uniform over
+feature space) and covariance matrix from the common prior distribution.
+Because the shared prior distribution's parameters were unobserved, the
+hierarchical Bayesian approach was used to infer its parameters from the
+previous categories (their feature variances and correlations), and then to
+generate the covariance matrix of the new category.
+
+In our implementation of their model\footnote{Note that
+\citet{jern2013probabilistic}'s model is slightly different, as they used a
+semi-conjugate model. Their model acts very similarly to our version.}, each
+category's exemplars are viewed as samples from a multivariate Normal
+distribution with parameters ($\mu, \Sigma$). Category covariance matrices
+(specifying variance and covariance along $k$-dimensions), are assumed to be
+Normal-Inverse-Wishart distributed with parameters: $\nu$ ($>k-1$), $\kappa$
+($>0$), and $\Sigma_D$. $\nu$ and $\kappa$ are treated as free parameters in our
+simulations, and $\Sigma_D$ is the domain-wide covariance matrix from which all
+categories are viewed as samples. Assuming a given $\Sigma_D$, a category
+covariance matrix $\Sigma$ can be computed on the basis of its examples:
+
+\begin{equation} \Sigma = \left[ \Sigma_D \nu + C + \dfrac {\kappa n} {\kappa +
+n} (\bar{x}-\mu)(\bar{x}-\mu)^T \right] (\nu + n)^{-1}
+\label{eq:Sigma_B}
+\end{equation}
+% 
+where $\bar{x}$ and $C$ are the empirical mean and covariance of the category's
+known members, and $n$ is the number of observed members of the category. When
+there are fewer than two known members of the category (and thus no covariance
+to speak of), $\Sigma = \Sigma_D\nu$.
+
+The category mean, $\mu$, can be computed as:
+
+\begin{equation} \mu = \dfrac {\kappa\mu_{0} + n \bar{x}} {\kappa + n}
+    \label{eq:category_mus_jk13}
+\end{equation}
+% 
+where $\mu_{0}$ is the prior mean. In our simulations, $\mu_{0}$ is set to the
+center of the domain. However, when no examples of the target category have been
+observed, generation is assumed to be random. In practice, the model's best fits
+are achieved when the $\kappa$ parameter, which controls the influence of
+$\mu_0$ on $\mu$, is set very close to zero (hence, the influence of $\mu_0$ is
+minimal).
+
+Importantly, the domain-wide covariance matrix $\Sigma_D$ is unobserved and
+needs to be inferred from the observed categories. For conjugacy, if $\Sigma_D$
+is viewed as a sample from an Inverse-Wishart distribution with scale
+$\Sigma_0$, $\Sigma_D$ can be computed as:
+
+ \begin{equation} \Sigma_D = \Sigma_0 + \sum_y{C_y}
+    \label{eq:category_sigmas}
+\end{equation}
+% 
+where $\Sigma_0$ is the prior covariance in the domain. In our simulations,
+$\Sigma_0 = \lambda {\bf I}$, where $\lambda$ is a free parameter controlling
+the expected variance of dimensions (dimensions of the domain covariance matrix
+are expected to be uncorrelated) and ${\bf I}$ is a $k$-by-$k$ identity matrix.
+
+Generated exemplars are drawn from a multivariate Normal distribution specified
+by $(\mu, \Sigma)$. Thus, $p(y)$ is
+\begin{equation} p(y \mid x) = \dfrac {\exp \left \{ \theta \cdot {\rm
+Normal}(y; \mu, \Sigma) \right \} } {\sum_i \exp \left \{ \theta \cdot {\rm
+Normal}(y_i; \mu, \Sigma) \right \} }
+\end{equation}
+% 
+where $\theta$ is a response determinism parameter and ${\rm Normal}(y; \mu,
+\Sigma)$ denotes a multivariate Normal density evaluated at $y$.
+
+\newpage
+\section{Representativeness Model in Toroidal Space}
+\label{ap:wrapgauss}
+
+Assuming a single bounded feature, we can easily represent the probability
+density of a function at any point $x$ on a line as $f(x)$. With a single
+unbounded feature, the points $x$ are mapped onto a unit circle with the
+corresponding $f(x)$ "wrapped" around the circle. More formally, we can
+represent the probability density along a wrapped axis $f'(x)$ as:
+\begin{align}
+f'(x) &= ... + f(x-4\pi) + f(x-2\pi) + f(x) + f(x+2\pi) + f(x+4\pi) + ...\\
+     &= \sum_{i=-\infty}^{\infty} f(x + 2\pi i)
+\end{align}
+
+We implement the wrapped axis in the representativeness model when
+calculating the multivariate Gaussian components of the representativeness equation
+$R(x,h)$ (Equation \ref{eq:representativeness} of the main text). Specifically,
+the probability density of the Gaussian component along a single wrapped axis can be computed as:
+\begin{equation}
+p'(x \mid h) = \sum_{i=-\infty}^{\infty} p(x + 2\pi i \mid h)
+\end{equation}
+where $p(\cdot \mid h)$ represents the probability density of the Gaussian component $h$
+on a line (i.e., not wrapped around a unit circle).
+
+In practice, we can approximate the infinite summation by constraining $i$
+within a limited range (e.g., $[-50,50]$). For increased efficiency, we can also
+approximate $p'(x \mid h)$ by stopping the summation once the ratio of the maximum
+and minimum values of $p(x + 2\pi i  \mid h)$ computed thus far is large enough (e.g.,
+10,000). For our modeling exercises here we impose both constraints, although the
+first constraint shows some redundancy since we find that $i$ rarely, if ever,
+exceeds a value of 4. Given the unimodal nature of the Gaussian distribution,
+this procedure allows for comparatively efficient approximation of the wrapped
+distribution without needing to compute the infinite sum.
 
@@
 
+% references section
+%\clearpage \bibliographystyle{apacite} \bibliography{citations.bib}
+%\clearpage
+
 \bibliography{sn-bibliography}% common bib file

### 2024-02-22 Manuscripts/cbb2023/sn-article.tex
work to get paper compiling as it does in overleaf
getting towards compiling in normal latex environment
Minor progress on revision

@@
 
-\jyear{2023}%
+\jyear{2024}%
 
@@
 
-%%=============================================================%%
-%% Prefix	-> \pfx{Dr}
-%% GivenName	-> \fnm{Joergen W.}
-%% Particle	-> \spfx{van der} -> surname prefix
-%% FamilyName	-> \sur{Ploeg}
-%% Suffix	-> \sfx{IV}
-%% NatureName	-> \tanm{Poet Laureate} -> Title after name
-%% Degrees	-> \dgr{MSc, PhD}
-%% \author*[1,2]{\pfx{Dr} \fnm{Joergen W.} \spfx{van der} \sur{Ploeg} \sfx{IV} \tanm{Poet Laureate} 
-%%                 \dgr{MSc, PhD}}\email{iauthor@gmail.com}
-%%=============================================================%%
-
 \author*[1]{\fnm{Joseph L.} \sur{Austerweil}}\email{austerweil@wisc.edu}
@@
 
-\abstract{The ability to generate new concepts and ideas is among the most fascinating aspects of human cognition, but we do not have a strong understanding of the
-cognitive processes and representations underlying concept generation. In this
-paper, we study the generation of new categories using the computational and
-behavioral toolkit of traditional artificial category learning. Previous work in
+\abstract{The ability to generate new concepts and ideas is among the most fascinating aspects of human cognition,
+but we do not have a strong understanding of the
+cognitive processes and representations underlying concept generation. Previous work in
 this domain has focused on how the statistical structure of known categories
 generalizes to generated categories, overlooking whether (and if so, how)
-contrast between the known and generated categories is a factor. We report three
-experiments demonstrating that contrast between what is known and what is
-created is of fundamental importance for categorization. We propose two novel
+contrast between the known and generated categories is a factor. In this paper, 
+we explore a different factor: contrast from known categories. We propose two novel
 approaches to modeling category contrast: one focused on exemplar dissimilarity
-and another on the representativeness heuristic. Our experiments and
-computational analyses demonstrate that both models capture different aspects of
-contrast's role in categorization.}
+and another based on the representativeness heuristic. Across three behavioral experiments, 
+we find that people generate new categories that contrast from observed categories and 
+distribute exemplars acoss "unoccupied" regions of stimulus space. The model based on 
+the representativeness heuristic captured human category generation better when 
+the known category was well captured by a Gaussian distribution. Conversely, the 
+exemplar-based model captured human generated categories better when the known category
+was not Gaussian distributed. Our results suggest contrast is a fundamental principle used in
+generating exemplars of a new category.}
 
@@
 \citep{abbott11}, and language grammar learning \citep{rafferty10}. We do so by
-extending the state-of-the-art category generation model
+extending a Bayesian category generation model
 \citep{jern2013probabilistic} and find that human category generation can
@@
 strong and systematic effects of category contrast on concept generation, and we
-qualitatively and quantitatively analyze the behavioral results and the performance of each model in
-capturing human category generation. A third experiment is performed to
-highlight the differences between the two contrast models. We conclude with a
+analyze the behavioral results and the performance of each model in
+capturing human category generation at the individual-level. Based on the model-
+based analyses, we develop and conduct a third experiment that distinguishes the two
+computational formulations of contrast. The third experiment found the exemplar-based
+version of contrast to capture human performance better. We conclude with a
 discussion of the implications of our results for categorization and directions
@@
 	\begin{center} 
-    \includegraphics[width=0.95\textwidth]{fig1.eps}
+    \includegraphics[width=0.95\textwidth]{figs/fig1.eps}
       \caption{The effect of weights on PACKER generation of a category `B' example, following exposure
@@
 \begin{figure}
-	\begin{center} \includegraphics{fig2.eps}
+	\begin{center} \includegraphics{figs/fig2.eps}
 		\caption{Generation of a category `B' example, following exposure
@@
 \begin{figure}
-    \begin{center} \includegraphics[width=0.8\textwidth]{fig13.eps}
+    \begin{center} \includegraphics[width=0.8\textwidth]{figs/fig13.eps}
     \caption{PACKER generation of a category `B' example, following exposure to
